### 2018-03-26 02_scripts/01_hyperspectral_indices/01_hs-rename-and-subset.Rmd
add files

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

# Load packages

```{r}
library(sf)
library(raster)
library(magrittr)
library(here)
library(purrr)
library(dplyr)
```

# Read all atmoshperic corrected hyperspectral images

```{r}
map(list.files("/data/patrick/raw/hyperspectral/ATM", 
                    full.names = TRUE, pattern = ".tif$"),
    ~ brick(.x)) -> data_atm

all.plots <- st_read("/home/shares/data/LIFE/mod/plots/location/all.plots.shp") %>% mutate(Name = as.character(Name))
```

# Read plot locations

```{r}
st_read("/data/patrick/mod/plots/location/all.plots.shp") %>% 
  mutate(Name = as.character(Name)) -> plot_locations
```

# Load plots by name and perform renaming and subsetting for each

Names are extracted from the first three digits in the file name
e.g. B101 -> Plot 'Busturi- axpe' (see README for LUT)

1. Change plot shp CRS to HS CRS (faster than changing the CRS of large raster)
2. Crop HS by plot shp
3. Apply UTM Zone 30 CRS
4. Save cropped raster

## Busturi- axpe (Code B101)
```{r}
busturi1 <- raster::brick(files[[1]])

all.plots %>% 
  filter(Name == "a busturi-axpe") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

busturi1 <- raster::crop(busturi1, plot.shp)
busturi1 <- raster::mask(busturi1, plot.shp)
busturi1 <- raster::projectRaster(busturi1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(busturi1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/busturi.tif",
                    overwrite = TRUE)
```

## Laukiz 1, Laukiz 2, Laukiz 3 (Code B103)

```{r}
## Laukiz I
laukiz1 <- raster::brick(files[[2]])

all.plots %>% 
  filter(Name == "Laukiz I") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laukiz1 <- raster::crop(laukiz1, plot.shp)
laukiz1 <- raster::mask(laukiz1, plot.shp)
laukiz1 <- raster::projectRaster(laukiz1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laukiz1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz1.tif",
                    progress = "text", overwrite = TRUE)

## Laukiz II
laukiz2 <- raster::brick(files[[2]])

all.plots %>% 
  filter(Name == "Laukiz II") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laukiz2 <- raster::crop(laukiz2, plot.shp)
laukiz2 <- raster::mask(laukiz2, plot.shp)
laukiz2 <- raster::projectRaster(laukiz2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laukiz2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz2.tif",
                    progress = "text", overwrite = TRUE)

## Laukiz III
laukiz3 <- raster::brick(files[[2]])

all.plots %>% 
  filter(Name == "Laukiz III") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laukiz3 <- raster::crop(laukiz3, plot.shp)
laukiz3 <- raster::mask(laukiz3, plot.shp)
laukiz3 <- raster::projectRaster(laukiz3, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laukiz3, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz3.tif",
                    progress = "text", overwrite = TRUE)

```

## Altube (Code B111 and B211)

```{r}
altube <- raster::brick(files[[3]])

all.plots %>% 
  filter(Name == "a altube") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

altube <- raster::crop(altube, plot.shp)
altube <- raster::mask(altube, plot.shp)
altube <- raster::projectRaster(altube, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(altube, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/altube.tif", 
                    progress = "text", overwrite = TRUE) 
```

## Barazar (Code B113)

```{r}
barazar <- raster::brick(files[[4]])

all.plots %>% 
  filter(Name == "a barazar") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

barazar <- raster::crop(barazar, plot.shp)
barazar <- raster::mask(barazar, plot.shp)
barazar <- raster::projectRaster(barazar, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(barazar, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/barazar.tif",
                    progress = "text", overwrite = TRUE)
```

## Laricio defol. Urkiola (Code B114)

```{r}
Laricio.defol.urkiola <- raster::brick(files[[5]])

all.plots %>% 
  filter(Name == "Laricio defol. Urkiola.") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

Laricio.defol.urkiola <- raster::crop(Laricio.defol.urkiola, plot.shp)
Laricio.defol.urkiola <- raster::mask(Laricio.defol.urkiola, plot.shp)
Laricio.defol.urkiola <- raster::projectRaster(Laricio.defol.urkiola, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(Laricio.defol.urkiola, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio-defol-urkiola.tif",
                    progress = "text", overwrite = TRUE)
```

## Aguinaga pinaster radiata (Code B123)

```{r}
aguinaga.pinaster.radiata <- raster::brick(files[[6]])

all.plots %>% 
  filter(Name == "aguinaga pinaster radiata") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

aguinaga.pinaster.radiata <- raster::crop(aguinaga.pinaster.radiata, plot.shp)
aguinaga.pinaster.radiata <- raster::mask(aguinaga.pinaster.radiata, plot.shp)
aguinaga.pinaster.radiata <- raster::projectRaster(aguinaga.pinaster.radiata, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(aguinaga.pinaster.radiata, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/aguinaga-pinaster-radiata.tif",
                    progress = "text", overwrite = TRUE)
```

## Oiartzun (Code B124)

```{r}
oiartzun <- raster::brick(files[[7]])

all.plots %>% 
  filter(Name == "Oiartzun") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

oiartzun <- raster::crop(oiartzun, plot.shp)
oiartzun <- raster::mask(oiartzun, plot.shp)
oiartzun <- raster::projectRaster(oiartzun, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(oiartzun, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/oiartzun.tif",
                    progress = "text", overwrite = TRUE)
```

## Pinast. Arza (Code B204)

```{r}
pinast.arza <- raster::brick(files[[8]])

all.plots %>% 
  filter(Name == "Pinast. Arza") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

pinast.arza <- raster::crop(pinast.arza, plot.shp)
pinast.arza <- raster::mask(pinast.arza, plot.shp)
pinast.arza <- raster::projectRaster(pinast.arza, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(pinast.arza, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/pinast-arza.tif",
                    progress = "text", overwrite = TRUE) 
```

## Pinas. Santa Koloma (Code B205)

```{r}
pinas.santa.koloma <- raster::brick(files[[9]])

all.plots %>% 
  filter(Name == "Pinas. Santa Koloma") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

pinas.santa.koloma <- raster::crop(pinas.santa.koloma, plot.shp)
pinas.santa.koloma <- raster::mask(pinas.santa.koloma, plot.shp)
pinas.santa.koloma <- raster::projectRaster(pinas.santa.koloma, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(pinas.santa.koloma, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/pinas-santa-koloma.tif",
                    progress = "text", overwrite = TRUE)
```

## Rad. Menagaray 1 and Rad. Mengaray 2 (Code B206)

```{r}
## Rad. Menagaray 1
rad.menagaray.1 <- raster::brick(files[[10]])

all.plots %>% 
  filter(Name == "Rad. Menagaray 1") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

rad.menagaray.1 <- raster::crop(rad.menagaray.1, plot.shp)
rad.menagaray.1 <- raster::mask(rad.menagaray.1, plot.shp)
rad.menagaray.1 <- raster::projectRaster(rad.menagaray.1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(rad.menagaray.1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-menagaray-1.tif",
                    progress = "text", overwrite = TRUE)

## Rad. Menagaray 2
rad.menagaray.2 <- raster::brick(files[[10]])

all.plots %>% 
  filter(Name == "Rad. Menagaray 2") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

rad.menagaray.2 <- raster::crop(rad.menagaray.2, plot.shp)
rad.menagaray.2 <- raster::mask(rad.menagaray.2, plot.shp)
rad.menagaray.2 <- raster::projectRaster(rad.menagaray.2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(rad.menagaray.2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-menagaray-2.tif",
                    progress = "text", overwrite = TRUE) 
```

## Ayala_aiara / Luiando (Code B207)

```{r}
ayala.aiara <- raster::brick(files[[11]])

all.plots %>% 
  filter(Name == "Ayala/Aiara") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

ayala.aiara <- raster::crop(ayala.aiara, plot.shp)
ayala.aiara <- raster::mask(ayala.aiara, plot.shp)
ayala.aiara <- raster::projectRaster(ayala.aiara, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(ayala.aiara, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/ayala-aiara.tif",
                    progress = "text", overwrite = TRUE)
```

## Silvest. Nograro (Code B208)

```{r}
silvest.nograro <- raster::brick(files[[12]])

all.plots %>% 
  filter(Name == "Silvest. Nograro") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

silvest.nograro <- raster::crop(silvest.nograro, plot.shp)
silvest.nograro <- raster::mask(silvest.nograro, plot.shp)
silvest.nograro <- raster::projectRaster(silvest.nograro, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(silvest.nograro, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/silvest-nograro.tif",
                    progress = "text", overwrite = TRUE)
```

## Caranca (Code B209)

```{r}
caranca <- raster::brick(files[[13]])

all.plots %>% 
  filter(Name == "a caranca") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

caranca <- raster::crop(caranca, plot.shp)
caranca <- raster::mask(caranca, plot.shp)
caranca <- raster::projectRaster(caranca, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(caranca, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/caranca.tif",
                    progress = "text", overwrite = TRUE)
```

## Picea Abornicano (Code B210)

```{r}
picea.abornicano <- raster::brick(files[[14]])

all.plots %>% 
  filter(Name == "Picea Abornicano") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

picea.abornicano <- raster::crop(picea.abornicano, plot.shp)
picea.abornicano <- raster::mask(picea.abornicano, plot.shp)
picea.abornicano <- raster::projectRaster(picea.abornicano, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(picea.abornicano, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/picea-abornicano.tif",
                    progress = "text", overwrite = TRUE)
```

## Douglas 1 and Douglas 2 and Rad. Altube (Code B212)

Code 211 bzw. file[[15]] == Altube

```{r}
## douglas 1
douglas.1 <- raster::brick(files[[16]])

all.plots %>% 
  filter(Name == "douglas 1") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.1 <- raster::crop(douglas.1, plot.shp)
douglas.1 <- raster::mask(douglas.1, plot.shp)
douglas.1 <- raster::projectRaster(douglas.1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas.1.tif",
                    progress = "text", overwrite = TRUE)

## douglas 2
douglas.2 <- raster::brick(files[[16]])

all.plots %>% 
  filter(Name == "douglas 2") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.2 <- raster::crop(douglas.2, plot.shp)
douglas.2 <- raster::mask(douglas.2, plot.shp)
douglas.2 <- raster::projectRaster(douglas.2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-2.tif",
                    progress = "text", overwrite = TRUE)

## rad. altube
rad.altube <- raster::brick(files[[16]])

all.plots %>% 
  filter(Name == "Rad. Altube") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

rad.altube <- raster::crop(rad.altube, plot.shp)
rad.altube <- raster::mask(rad.altube, plot.shp)
rad.altube <- raster::projectRaster(rad.altube, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(rad.altube, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-altube.tif",
                    progress = "text", overwrite = TRUE)
```

## Otxandio and Laricio defol. Olaeta (Code B215)

```{r}
## otxandio
otxandio <- raster::brick(files[[17]])

all.plots %>% 
  filter(Name == "a otxandio") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

otxandio <- raster::crop(otxandio, plot.shp)
otxandio <- raster::mask(otxandio, plot.shp)
otxandio <- raster::projectRaster(otxandio, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(otxandio, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/otxandio.tif",
                    progress = "text", overwrite = TRUE)

## Laricio defol. Olaeta
laricio.defol.olaeta <- raster::brick(files[[17]])

all.plots %>% 
  filter(Name == "Laricio defol. Olaeta. pol") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laricio.defol.olaeta <- raster::crop(laricio.defol.olaeta, plot.shp)
laricio.defol.olaeta <- raster::mask(laricio.defol.olaeta, plot.shp)
laricio.defol.olaeta <- raster::projectRaster(laricio.defol.olaeta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laricio.defol.olaeta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio-defol-olaeta.tif",
                    progress = "text", overwrite = TRUE)
```

## Douglas Legutio (Code B317 and B318)
Here B317 is taken
```{r}
douglas.legutio <- raster::brick(files[[18]])

all.plots %>% 
  filter(Name == "Douglas Legutio") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.legutio <- raster::crop(douglas.legutio, plot.shp)
douglas.legutio <- raster::mask(douglas.legutio, plot.shp)
douglas.legutio <- raster::projectRaster(douglas.legutio, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.legutio, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-legutio.tif",
                    progress = "text", overwrite = TRUE)
```
## Laricio Gatzaga (Code B319)

```{r}
laricio.gatzaga <- raster::brick(files[[20]])

all.plots %>% 
  filter(Name == "Laricio defol.Gatzaga") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laricio.gatzaga <- raster::crop(laricio.gatzaga, plot.shp)
laricio.gatzaga <- raster::mask(laricio.gatzaga, plot.shp)
laricio.gatzaga <- raster::projectRaster(laricio.gatzaga, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laricio.gatzaga, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio.gatzaga.tif",
                    progress = "text", overwrite = TRUE)
```
## Azaceta (Code B320)

```{r}
azaceta <- raster::brick(files[[21]])

all.plots %>% 
  filter(Name == "a azaceta") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

azaceta <- raster::crop(azaceta, plot.shp)
azaceta <- raster::mask(azaceta, plot.shp)
azaceta <- raster::projectRaster(azaceta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(azaceta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/azaceta.tif",
                    progress = "text", overwrite = TRUE)
```
## Gipuzkoa and Sequoia Legorreta (Code B321)

```{r}
## gipuzkoa
gipuzkoa <- raster::brick(files[[22]])

all.plots %>% 
  filter(Name == "a gipuzkoa") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

gipuzkoa <- raster::crop(gipuzkoa, plot.shp)
gipuzkoa <- raster::mask(gipuzkoa, plot.shp)
gipuzkoa <- raster::projectRaster(gipuzkoa, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(gipuzkoa, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/gipuzkoa.tif",
                    progress = "text", overwrite = TRUE)

## Sequoia Legorreta
sequoia.legorreta <- raster::brick(files[[22]])

all.plots %>% 
  filter(Name == "Sequoia Legorreta") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

sequoia.legorreta <- raster::crop(sequoia.legorreta, plot.shp)
sequoia.legorreta <- raster::mask(sequoia.legorreta, plot.shp)
sequoia.legorreta <- raster::projectRaster(sequoia.legorreta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(sequoia.legorreta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/sequoia.legorreta.tif",
                    progress = "text", overwrite = TRUE)
```
## Douglas Albiztur (Code B322)

```{r}
douglas.albiztur <- raster::brick(files[[23]])

all.plots %>% 
  filter(Name == "Douglas Albiztur") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.albiztur <- raster::crop(douglas.albiztur, plot.shp)
douglas.albiztur <- raster::mask(douglas.albiztur, plot.shp)
douglas.albiztur <- raster::projectRaster(douglas.albiztur, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.albiztur, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-albiztur.tif",
                    progress = "text", overwrite = TRUE)
```

### 2018-03-26 02_scripts/01_hyperspectral_indices/02_hyperspectral-ndvi-raster.Rmd
add files

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

```{r}
pacman::p_load(raster, hsdar, here, pbapply, parallel)
```

# specify wavelength

```{r}
wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")
```

# read rasters (cropped by plot)

## list files

```{r}
files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
                    pattern = ".tif$", full.names = TRUE)
```

## read in files as "RasterBrick"

```{r}
spectral_data <- lapply(files, function(x) brick(x))
```

# make class "HyperSpecRaster"
```{r}
hyperspecs <- lapply(spectral_data, function(x) hsdar::HyperSpecRaster(x, wavelength))
```

# account for use on local laptop or server
```{r}
if (parallel::detectCores() > 10) {
  cores <- 28
} else (cores <- 1)
```

# calculate NDVI

assign to object to prevent list printing to console ### use 'NDVI2' as formula 
```{r}
pboptions(style = 1, type = "timer")
tmp <- pblapply(hyperspecs, function(x) 
  vegindex(x, "NDVI2", filename = paste0("/home/shares/data/LIFE/mod/hyperspectral/ndvi/", "ndvi2-",
                                         basename(x@file@name)),
           bnames = "NDVI"), cl = cores)
```

### 2018-03-26 02_scripts/01_hyperspectral_indices/03_hyperspectral-ndvi-analysis.Rmd
add files

---
title: "NDVI analysis of hyperspectral data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
# Data & packages

Loads list of 28 cropped (to respective plot) atmospherical corrected hyperspectral images.  
See R script `hs.ndvi.R` for multicore creation.

Move to chunk 'hsdar package' if file exists!
```{r}
pacman::p_load(rgdal, raster, mapview, hsdar, rio, ggplot2, colorRamps)

if (file.exists("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/all-cropped.rds")) {
  files <- as.list(rio::import("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/all-cropped.rds"))
} else {
  files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/", 
                    pattern = ".tif$", full.names = TRUE)

  files1 <- lapply(files, function(x) raster::brick(x))
}
```

Define 'Get values row by row' (vignette `raster::functions`)
* x = multy layer Raster file (applying `brick()`)
* a = value to add to existing value
```{r}
source(here("R/helper.funs/get.values.row.by.row.R"))
```

# Get values from all raster files (bricks!) using `lapply`
```{r}
# files_atm <- lapply(files_atm, function(x) f8(x, 0, progress = "text"))
# saveRDS(files_atm, here("R/tmp/files_atm_with_values.rda")
files_atm <- readRDS("/home/patrick/PhD/R/tmp/files_atm_with_values.rda")
```

Get wavelengths out of README table (http://datascienceplus.com/extracting-tables-from-pdfs-in-r-using-the-tabulizer-package/)

Does not work because wavelength table on page 10 of README is not a table but an image...
```{r}
# devtools::install_github("ropenscilabs/tabulizer")
# devtools::install_github("ropenscilabs/tabulizerjars")

# library(tabulizer)

# tabs <- extract_tables("/home/shares/data/original/hyperspectral/0_README_spanish.pdf", pages = 10)
```

# hsdar package
```{r}
# specify wavelength of each band
wavelength <- c(404.08, 408.5, 412.92, 417.36, 421.81, 426.27, 430.73, 435.20, 
                439.69, 444.18, 448.68, 453.18, 457.69, 462.22, 466.75, 471.29,
                475.83, 480.39, 484.95, 489.52, 494.09, 498.68, 503.26, 507.86,
                512.47, 517.08, 521.70, 526.32, 530.95, 535.58, 540.23, 544.88,
                549.54, 554.20, 558.86, 563.54, 568.22, 572.90, 577.60, 582.29,
                586.99, 591.70, 596.41, 601.13, 605.85, 610.58, 615.31, 620.05,
                624.79, 629.54, 634.29, 639.04, 643.80, 648.56, 653.33, 658.10,
                662.88, 667.66, 672.44, 677.23, 682.02, 686.81, 691.60, 696.40,
                701.21, 706.01, 710.82, 715.64, 720.45, 725.27, 730.09, 734.91,
                739.73, 744.56, 749.39, 754.22, 759.05, 763.89, 768.72, 773.56, 
                778.40, 783.24, 788.08, 792.93, 797.77, 802.62, 807.47, 812.32,
                817.17, 822.02, 826.87, 831.72, 836.57, 841.42, 846.28, 851.13,
                855.98, 860.83, 865.69, 870.54, 875.39, 880.24, 885.09, 889.94,
                894.79, 899.64, 904.49, 909.34, 914.18, 919.03, 923.87, 928.71,
                933.55, 938.39, 943.23, 948.07, 952.90, 957.73, 962.56, 967.39,
                972.22, 977.04, 981.87, 986.68, 991.50, 996.31)
```

Read in raster data either using `rgdal` or `raster` package. 
`raster` seems to be prefered but `rgdal` is a lot faster in reading values

```{r}
## create HyperSpecRaster file
# ra <- hsdar::HyperSpecRaster(files_atm[[1]], wavelength)
# tr <- blockSize(ra)

##  create SpecLib file
# spectral_data1 <- f8(files1[[1]], 0, progress = "text")

# speclib <- hsdar::speclib(spectral_data, wavelength)
```

plot spectral signature
```{r}
# plot(spectral_data, main = "Default Plot")

# plot(spectral_data, FUN = "median") # not working
```

Visualize NDVI
```{r}
mapview::mapview(spectral_data1, zcol = "ndvi", maxpixels = 4664960, na.color = "white")
```

# NDVI analysis

## Read in files

1. List files
2. Read in using `rgdal::readGDAL`
3. Get filenames
4. Rename `band1` to `NDVI` and label SpGrDaFr with plot name
4. Get pixel count of each file
5. Create DF with one NDVI column and a factor column with filenames
```{r}
files.list <- list.files("/home/shares/data/LIFE/mod/hyperspectral/ndvi", 
                    pattern = "ndvi2.*\\.tif", full.names = TRUE)

files.ndvi <- lapply(files.list, function(x) rgdal::readGDAL(x, silent = TRUE))
# rename band name to 'NDVI'
for (i in seq_along(files.ndvi)) {
  colnames(files.ndvi[[i]]@data) <- "NDVI"

}

names.file <- sapply(files.list, function(x) basename(x))
# label SpGrDaFr with plot name
names(files.ndvi) <- names.file

pixel.count <- sapply(files.ndvi, function(x) length(x@data$NDVI))

ndvi.all <- c()
ndvi.all <- unlist(lapply(files.ndvi, function(x) append(ndvi.all, x@data$NDVI)))

plot.names <- c()
for (i in seq_along(names.file)) {
  plot.names <- append(plot.names, rep(basename(names.file[i]), pixel.count[i]))
}

df <- data.frame(NDVI = ndvi.all,
                 plot = as.factor(plot.names))

# remove NA rows
df <- df[complete.cases(df),]
```

## Plot NDVI histograms

```{r}
ndvi.histograms <- ggplot(df, aes(NDVI)) +
  #geom_histogram() + 
  # normalize y-axis of each facet http://stackoverflow.com/questions/16339204/normalizing-faceted-histograms-separately-in-ggplot2
  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
  facet_wrap(~ plot) + 
  ylab("Percentage pixel count") + 
  xlab("NDVI (0.25 to 1)") + 
  scale_x_continuous(limits = c(0.25, 1)) + 
  cowplot::background_grid()
```

## Count pixels per plot (size 1m^2)
```{r}
table(df$plot)
```

## Quantiles

Biased towards low values (non-forest pixels)
```{r}
stats::quantile(df$NDVI)
```

```{r}
boxplot(df$NDVI)
```

```{r}
color <- colorRamps::green2red(2)
rasterVis::levelplot(files.ndvi[[4]])
mapview::mapView(as.raster(files.ndvi[[4]]), zcol = "NDVI", na.color = "blue", color = color)
```

## plot NDVI rasters
```{r}
# convert to raster layer for plotting
files.ndvi.raster <- lapply(files.ndvi, function(x) raster::raster(x))
#set names
for (i in seq_along(files.ndvi.raster)) {
  files.ndvi.raster[[i]]@file@name <- names(files.ndvi)[i]
}

# using mapview
list.plots <- lapply(files.ndvi.raster, function(x) mapview::mapview(x, use.layer.names = TRUE, na.color = "white"))
# using rasterVis::levelplot
list.plots <- lapply(files.ndvi.raster, function(x) rasterVis::levelplot(x,
                                                                         main = x@file@name,
                                                                         margin = FALSE,
                                                                         pretty = TRUE,
                                                                         col.regions = rev(colorRamps::green2red(400)), 
                                                                         at = seq(0.5, 1, len = 101))
)

gridExtra::grid.arrange(list.plots[[1]], list.plots[[2]], list.plots[[3]], 
                                        list.plots[[4]], list.plots[[5]], list.plots[[6]],
                                        list.plots[[7]], list.plots[[8]], list.plots[[9]],
                                        ncol = 3)

gridExtra::grid.arrange(list.plots[[10]], list.plots[[11]], list.plots[[12]],
                                        list.plots[[13]], list.plots[[14]], list.plots[[15]],
                                        list.plots[[16]], list.plots[[17]], list.plots[[18]],
                                        ncol = 3)

gridExtra::grid.arrange(list.plots[[19]], list.plots[[20]], list.plots[[21]],
                                        list.plots[[22]], list.plots[[23]], list.plots[[24]], 
                                        list.plots[[25]], list.plots[[26]], list.plots[[27]], 
                                        list.plots[[28]],
                                        ncol = 3)

```

## save NDVI raster plots

```{r}
png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.1.png", width = 16.51, height = 12.07, units = "in", res = 300)
gridExtra::grid.arrange(list.plots[[1]], list.plots[[2]], list.plots[[3]], 
                                        list.plots[[4]], list.plots[[5]], list.plots[[6]],
                                        list.plots[[7]], list.plots[[8]], list.plots[[9]],
                                        ncol = 3)
dev.off()

png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.2.png", width = 16.51, height = 12.07, units = "in", res = 300)
gridExtra::grid.arrange(list.plots[[10]], list.plots[[11]], list.plots[[12]],
                                        list.plots[[13]], list.plots[[14]], list.plots[[15]],
                                        list.plots[[16]], list.plots[[17]], list.plots[[18]],
                                        ncol = 3)
dev.off()

png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.3.png", width = 16.51, height = 12.07, units = "in", res = 300)
gridExtra::grid.arrange(list.plots[[19]], list.plots[[20]], list.plots[[21]],
                                        list.plots[[22]], list.plots[[23]], list.plots[[24]], 
                                        list.plots[[25]], list.plots[[26]], list.plots[[27]], 
                                        list.plots[[28]],
                                        ncol = 3)
dev.off()
```

## save ndvi histograms
```{r}
png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi.histograms.normalized.png", width = 16.51, height = 12.07, units = "in", res = 300)
ndvi.histograms
dev.off()
```

### 2018-03-26 02_scripts/01_hyperspectral_indices/04_hyperspectral-all-vegindex.Rmd
add files

---
title: "Calculate all possible vegetation indices that the `hsdar` package offers."
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

```{r}
pacman::p_load(raster, hsdar, pbapply, parallel, rasterFunctions, tidyverse)
```

# Specify wavelength

```{r}
wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")
```

# List files

```{r}
files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
                    pattern = ".tif$", full.names = TRUE)
```

# read in files as "RasterBrick"
```{r}
spectral_data <- lapply(files, function(x) brick(x))
```

# make class "HyperSpecRaster"

```{r}
hyperspecs <- lapply(spectral_data, function(x) hsdar::HyperSpecRaster(x, wavelength))
```

# account for use on local laptop or server
```{r}
if (detectCores() > 10) {
  cores <- 28
} else (cores <- 1)
```

# do calculation

Runtime server: ~ 1 min
Runtime MBP: ~ 10 mins

Writing files as `.gri` files because this format preserves layer names. TIFF format does not!!
Layer names are changed using the `bnames` argument of function `rasterFunctions::hsdar.vegIndex()`.

Specify vegetation indices.

```{r}
# subset for test purposes
# hyperspecs <- list(hyperspecs[[17]])

# specify which indices to calculate WITH DERIVATION INDICES
indices <- c("Boochs", "Boochs2","CARI", "Carter", "Carter2", "Carter3", "Carter4", "Carter5",
             "Carter6", "CI", "CI2", "ClAInt", "CRI1", "CRI2", "CRI3", "CRI4",
             "D1", "D2",
             "Datt", "Datt2", "Datt3", "Datt4", "Datt5", "Datt6", "DD", "DDn", "DPI", "DWSI4",
             "EGFR", "EVI", "GDVI_2", "GDVI_3", "GDVI_4", "GI", "Gitelson",
             "Gitelson2", "GMI1", "GMI2", "Green NDVI", "Maccioni", "MCARI",
             "MCARI2", "mND705", "mNDVI", "MPRI", "MSAVI", "mSR",
             "mSR2", "mSR705", "MTCI", "MTVI", "NDVI", "NDVI2", "NDVI3",
             "NPCI", "OSAVI", "PARS", "PRI", "PRI_norm", "PRI*CI2", "PSRI",
             "PSSR", "PSND","SPVI", "PWI", "RDVI", "REP_Li", "SAVI", "SIPI",
             "SPVI","SR", "SR1", "SR2", "SR3", "SR4", "SR5", "SR6", "SR7",
             "SR8", "Sum_Dr1", "Sum_Dr2", "SRPI", "TCARI", "TCARI2", "TGI", "TVI", "Vogelmann",
             "Vogelmann2", "Vogelmann3", "Vogelmann4")
# length(indices)
```

# do veg. indices calculation

```{r }
hyperspecs %>% 
  pblapply(function(x) 
    vegindex(x, indices, filename = 
             gsub(".tif", ".grd", 
                  "/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/",
                    "all-veg-indices-", basename(x@file@name)), na.rm = T,
             bnames = indices), cl = cores) -> rasters

# mapview(tmp[[1]])
```

### 2018-03-26 02_scripts/01_hyperspectral_indices/05_narrow-band-indices.Rmd
add files

---
title: "Calculate all possible narrow band indices"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

**Attention**

This script uses the function `nbi_raster()` which is a custom function and only available from the developtment version of `hsdar` on Github: `devtools::install_github("pat-s/hsdar")`.

```{r}
pacman::p_load(raster, hsdar, pbapply, parallel, tidyverse, here, magrittr)
```

# Load wavelength and rasters

```{r}
wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")

list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
           pattern = ".tif$", full.names = TRUE) %>%
  lapply(function(x) brick(x)) %>% 
  lapply(function(x) HyperSpecRaster(x, wavelength)) -> hyperspecs
```

# account for use on local laptop or server
```{r}
if (detectCores() > 10) {
  cores <- 28
} else (cores <- 1)
```

# do calculation of narrow band indices

Band names will be calculated automatically by function `nbi_raster`

Takes 190 GB RAM on server when being parallalized.

Runtime per plot between 5 - 27 minutes.

Largest file: 24 GB

```{r}
## Calculate NBI
# hyperspecs <- list(hyperspecs[[1]])

hyperspecs %>% 
  pblapply(function(x) 
    nbi_raster(x, filename = gsub(".tif", ".grd", 
                                  "/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/",
                                         "narrow-band-indices-", basename(x@file@name))), cl = cores) 
```

### 2018-03-26 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd
add files

---
title: "Extract indices to points"
output: html_notebook
editor_options: 
chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

```{r}
pacman::p_load(raster, sf, here, magrittr, tidyverse, pbapply, parallel)
```

# Laukiz1

## Load points

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/laukiz1.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> laukiz1_points
```

## Load rasters

```{r}
laukiz1_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz1")
```

```{r}
laukiz1_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz1")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(laukiz1_veg_indices, laukiz1_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Nested parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(laukiz1_nbi, laukiz1_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = 5)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

laukiz1_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> laukiz1_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(laukiz1_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/all-indices-laukiz1.rda")
```

# Laukiz2

## Load points

```{r}
print("Starting Laukiz2")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/laukiz2.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> laukiz2_points
```

## Load rasters

```{r}
laukiz2_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz2")
```

```{r}
laukiz2_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz2")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(laukiz2_veg_indices, laukiz2_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(laukiz2_nbi, laukiz2_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = 5)
  #endCluster()
  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}

### local debugging code
# out_nbi <- raster::extract(laukiz2_nbi, laukiz2_points, buffer = 2, 
#                            fun = mean, df = TRUE)
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

laukiz2_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> laukiz2_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(laukiz2_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/all-indices-laukiz2.rda")
```

# Oiartzun

## Load points

```{r}
print("Starting oiartzun")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/oiartzun.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> oiartzun_points
```

## Load rasters

```{r}
oiartzun_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-oiartzun")
```

```{r}
oiartzun_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-oiartzun")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(oiartzun_veg_indices, oiartzun_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(oiartzun_nbi, oiartzun_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = cores)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

oiartzun_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> oiartzun_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(oiartzun_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/all-indices-oiartzun.rda")
```

# Luiando

## Load points

```{r}
print("Starting luiando")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/luiando/luiando.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> luiando_points
```

## Load rasters

```{r}
luiando_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-ayala-aiara") 
```

```{r}
luiando_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-ayala-aiara")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(luiando_veg_indices, luiando_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(luiando_nbi, luiando_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = cores)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

luiando_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> luiando_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(luiando_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/luiando/all-indices-luiando.rda")
```

### 2018-03-26 02_scripts/02_preprocessing.Rmd
add files

---
title: "Preprocessing of raw data"
author: "Patrick Schratz"
output:
  html_notebook:
    code_folding: none
    fig_width: 10
    highlight: haddock
    number_sections: yes
    toc: yes
---

# Load packages

```{r}
library(purrr)
library(here)
library(readr)
library(tibble)
library(glue)
library(raster)
```

# Read data

Reading in data from .txt files and exporting them as a Geopackage (.gpkg).

## Survey data

```{r}
names = c("laukiz1", "laukiz2", "luiando", "oiartzun")

map(names, 
    ~ as_tibble(read.table(file = glue("/data/patrick/mod/survey_data/2016/{.x}.txt"), 
                           sep = "\t", dec = ",", header = TRUE))) %>% 
  set_names(names) -> data_all
```

## Vegetation indices

```{r}
map(names,
    ~ brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/all-veg-indices-{.x}.gri"))) %>% 
  set_names(names) -> indices_all
```

## Narrow band indices

# Bind indices to response

### 2018-04-04 02_scripts/01_hyperspectral_indices/01_hs-rename-and-subset.Rmd
shiny app 2 by 2

@@
 title: "R Notebook"
-output: html_notebook
-editor_options: 
-  chunk_output_type: inline
+output: 
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
 ---
 
-# Information about this script
-
-This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`
-
-Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).
-
-Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
-It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE,
+                      cache = TRUE)
+```
 
@@
 
-```{r}
+```{r, results='hide'}
 library(sf)
@@
 library(magrittr)
-library(here)
 library(purrr)
@@
 ```{r}
-map(list.files("/data/patrick/raw/hyperspectral/ATM", 
+map(list.data_atm("/data/patrick/raw/hyperspectral/ATM", 
                     full.names = TRUE, pattern = ".tif$"),
     ~ brick(.x)) -> data_atm
-
-all.plots <- st_read("/home/shares/data/LIFE/mod/plots/location/all.plots.shp") %>% mutate(Name = as.character(Name))
 ```
@@
 ```{r}
-st_read("/data/patrick/mod/plots/location/all.plots.shp") %>% 
+st_read("/data/patrick/mod/plots/location/plot_locations.shp") %>% 
   mutate(Name = as.character(Name)) -> plot_locations
@@
 
-## Busturi- axpe (Code B101)
 ```{r}
-busturi1 <- raster::brick(files[[1]])
+process_hyperspec <- function(name_id, name_out, index) {
 
-all.plots %>% 
-  filter(Name == "a busturi-axpe") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
+  image <- data_atm[[index]]
 
-busturi1 <- raster::crop(busturi1, plot.shp)
-busturi1 <- raster::mask(busturi1, plot.shp)
-busturi1 <- raster::projectRaster(busturi1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
+  plot_locations %>% 
+    filter(Name == name_id) %>% 
+    st_transform(25830) %>% 
+    as("Spatial") -> shape_single_plot
 
-raster::writeRaster(busturi1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/busturi.tif",
+  image %>% 
+    crop(shape_single_plot) %>% 
+    mask(shape_single_plot) %>% 
+    projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs") %>% 
+    writeRaster(glue("/data/patrick/mod/hyperspectral/atm-cropped/{name_out}.tif"),
                 overwrite = TRUE)
+  cat(glue("Finished {name_out}."))
+}
 ```
 
-## Laukiz 1, Laukiz 2, Laukiz 3 (Code B103)
+# Create dictionaries to apply over
 
 ```{r}
-## Laukiz I
-laukiz1 <- raster::brick(files[[2]])
-
-all.plots %>% 
-  filter(Name == "Laukiz I") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-laukiz1 <- raster::crop(laukiz1, plot.shp)
-laukiz1 <- raster::mask(laukiz1, plot.shp)
-laukiz1 <- raster::projectRaster(laukiz1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(laukiz1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz1.tif",
-                    progress = "text", overwrite = TRUE)
-
-## Laukiz II
-laukiz2 <- raster::brick(files[[2]])
-
-all.plots %>% 
-  filter(Name == "Laukiz II") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-laukiz2 <- raster::crop(laukiz2, plot.shp)
-laukiz2 <- raster::mask(laukiz2, plot.shp)
-laukiz2 <- raster::projectRaster(laukiz2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(laukiz2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz2.tif",
-                    progress = "text", overwrite = TRUE)
-
-## Laukiz III
-laukiz3 <- raster::brick(files[[2]])
-
-all.plots %>% 
-  filter(Name == "Laukiz III") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-laukiz3 <- raster::crop(laukiz3, plot.shp)
-laukiz3 <- raster::mask(laukiz3, plot.shp)
-laukiz3 <- raster::projectRaster(laukiz3, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(laukiz3, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz3.tif",
-                    progress = "text", overwrite = TRUE)
-
+name_id = c("a busturi-axpe", "Laukiz I", "Laukiz II", "Laukiz III", "a altube", 
+            "a barazar", "Laricio defol. Urkiola.", "aguinaga pinaster radiata",
+            "Oiartzun", "Pinast. Arza", "Pinas. Santa Koloma", "Rad. Menagaray 1",
+            "Rad. Menagaray 2", "Ayala/Aiara", "Silvest. Nograro", "a caranca",
+            "Picea Abornicano", "douglas 1", "douglas 2", "Rad. Altube", 
+            "a otxandio", "Laricio defol. Olaeta. pol", "Douglas Legutio", 
+            "Laricio defol.Gatzaga", "a azaceta", "a gipuzkoa", 
+            "Sequoia Legorreta", "Douglas Albiztur")
+name_out = c("busturi", "laukiz1", "laukiz2", "laukiz3", "altube", "barazar", 
+             "laricio-defol-urkiola", "aguinaga-pinaster-radiata", "oiartzun",
+             "pinast-arza", "pinas-santa-koloma", "rad-menagaray-1",
+             "rad-menagaray-2", "luiando", "silvest-nograro", "caranca",
+             "picea-abornicano", "douglas-1", "douglas-2", "rad-altube",
+             "otxandio", "laricio-defol-olaeta", "douglas-legutio",
+             "laricio-gatzaga", "azaceta", "gipuzkoa", "sequoia-legorreta", 
+             "douglas-albiztur")
+index = c(1, 2, 2, 2, 3, 4, 5, 6, 7, 8, 9, 10, 10, 11, 12, 13, 14, 16, 16, 16,
+          17, 17, 18, 20, 21, 22, 22, 23)
 ```
 
-## Altube (Code B111 and B211)
+Now let's map over all entries in the vectors specified.
 
 ```{r}
-altube <- raster::brick(files[[3]])
-
-all.plots %>% 
-  filter(Name == "a altube") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-altube <- raster::crop(altube, plot.shp)
-altube <- raster::mask(altube, plot.shp)
-altube <- raster::projectRaster(altube, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(altube, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/altube.tif", 
-                    progress = "text", overwrite = TRUE) 
+pwalk(list(name_id, name_out, index), ~ process_hyperspec(..1, ..2, ..3))
 ```
 
-## Barazar (Code B113)
-
 ```{r}
-barazar <- raster::brick(files[[4]])
-
-all.plots %>% 
-  filter(Name == "a barazar") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-barazar <- raster::crop(barazar, plot.shp)
-barazar <- raster::mask(barazar, plot.shp)
-barazar <- raster::projectRaster(barazar, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(barazar, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/barazar.tif",
-                    progress = "text", overwrite = TRUE)
+sessionInfo()
 ```
-
-## Laricio defol. Urkiola (Code B114)
-
-```{r}
-Laricio.defol.urkiola <- raster::brick(files[[5]])
-
-all.plots %>% 
-  filter(Name == "Laricio defol. Urkiola.") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-Laricio.defol.urkiola <- raster::crop(Laricio.defol.urkiola, plot.shp)
-Laricio.defol.urkiola <- raster::mask(Laricio.defol.urkiola, plot.shp)
-Laricio.defol.urkiola <- raster::projectRaster(Laricio.defol.urkiola, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(Laricio.defol.urkiola, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio-defol-urkiola.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Aguinaga pinaster radiata (Code B123)
-
-```{r}
-aguinaga.pinaster.radiata <- raster::brick(files[[6]])
-
-all.plots %>% 
-  filter(Name == "aguinaga pinaster radiata") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-aguinaga.pinaster.radiata <- raster::crop(aguinaga.pinaster.radiata, plot.shp)
-aguinaga.pinaster.radiata <- raster::mask(aguinaga.pinaster.radiata, plot.shp)
-aguinaga.pinaster.radiata <- raster::projectRaster(aguinaga.pinaster.radiata, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(aguinaga.pinaster.radiata, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/aguinaga-pinaster-radiata.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Oiartzun (Code B124)
-
-```{r}
-oiartzun <- raster::brick(files[[7]])
-
-all.plots %>% 
-  filter(Name == "Oiartzun") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-oiartzun <- raster::crop(oiartzun, plot.shp)
-oiartzun <- raster::mask(oiartzun, plot.shp)
-oiartzun <- raster::projectRaster(oiartzun, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(oiartzun, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/oiartzun.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Pinast. Arza (Code B204)
-
-```{r}
-pinast.arza <- raster::brick(files[[8]])
-
-all.plots %>% 
-  filter(Name == "Pinast. Arza") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-pinast.arza <- raster::crop(pinast.arza, plot.shp)
-pinast.arza <- raster::mask(pinast.arza, plot.shp)
-pinast.arza <- raster::projectRaster(pinast.arza, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(pinast.arza, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/pinast-arza.tif",
-                    progress = "text", overwrite = TRUE) 
-```
-
-## Pinas. Santa Koloma (Code B205)
-
-```{r}
-pinas.santa.koloma <- raster::brick(files[[9]])
-
-all.plots %>% 
-  filter(Name == "Pinas. Santa Koloma") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-pinas.santa.koloma <- raster::crop(pinas.santa.koloma, plot.shp)
-pinas.santa.koloma <- raster::mask(pinas.santa.koloma, plot.shp)
-pinas.santa.koloma <- raster::projectRaster(pinas.santa.koloma, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(pinas.santa.koloma, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/pinas-santa-koloma.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Rad. Menagaray 1 and Rad. Mengaray 2 (Code B206)
-
-```{r}
-## Rad. Menagaray 1
-rad.menagaray.1 <- raster::brick(files[[10]])
-
-all.plots %>% 
-  filter(Name == "Rad. Menagaray 1") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-rad.menagaray.1 <- raster::crop(rad.menagaray.1, plot.shp)
-rad.menagaray.1 <- raster::mask(rad.menagaray.1, plot.shp)
-rad.menagaray.1 <- raster::projectRaster(rad.menagaray.1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(rad.menagaray.1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-menagaray-1.tif",
-                    progress = "text", overwrite = TRUE)
-
-## Rad. Menagaray 2
-rad.menagaray.2 <- raster::brick(files[[10]])
-
-all.plots %>% 
-  filter(Name == "Rad. Menagaray 2") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-rad.menagaray.2 <- raster::crop(rad.menagaray.2, plot.shp)
-rad.menagaray.2 <- raster::mask(rad.menagaray.2, plot.shp)
-rad.menagaray.2 <- raster::projectRaster(rad.menagaray.2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(rad.menagaray.2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-menagaray-2.tif",
-                    progress = "text", overwrite = TRUE) 
-```
-
-## Ayala_aiara / Luiando (Code B207)
-
-```{r}
-ayala.aiara <- raster::brick(files[[11]])
-
-all.plots %>% 
-  filter(Name == "Ayala/Aiara") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-ayala.aiara <- raster::crop(ayala.aiara, plot.shp)
-ayala.aiara <- raster::mask(ayala.aiara, plot.shp)
-ayala.aiara <- raster::projectRaster(ayala.aiara, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(ayala.aiara, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/ayala-aiara.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Silvest. Nograro (Code B208)
-
-```{r}
-silvest.nograro <- raster::brick(files[[12]])
-
-all.plots %>% 
-  filter(Name == "Silvest. Nograro") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-silvest.nograro <- raster::crop(silvest.nograro, plot.shp)
-silvest.nograro <- raster::mask(silvest.nograro, plot.shp)
-silvest.nograro <- raster::projectRaster(silvest.nograro, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(silvest.nograro, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/silvest-nograro.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Caranca (Code B209)
-
-```{r}
-caranca <- raster::brick(files[[13]])
-
-all.plots %>% 
-  filter(Name == "a caranca") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-caranca <- raster::crop(caranca, plot.shp)
-caranca <- raster::mask(caranca, plot.shp)
-caranca <- raster::projectRaster(caranca, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(caranca, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/caranca.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Picea Abornicano (Code B210)
-
-```{r}
-picea.abornicano <- raster::brick(files[[14]])
-
-all.plots %>% 
-  filter(Name == "Picea Abornicano") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-picea.abornicano <- raster::crop(picea.abornicano, plot.shp)
-picea.abornicano <- raster::mask(picea.abornicano, plot.shp)
-picea.abornicano <- raster::projectRaster(picea.abornicano, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(picea.abornicano, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/picea-abornicano.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Douglas 1 and Douglas 2 and Rad. Altube (Code B212)
-
-Code 211 bzw. file[[15]] == Altube
-
-```{r}
-## douglas 1
-douglas.1 <- raster::brick(files[[16]])
-
-all.plots %>% 
-  filter(Name == "douglas 1") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-douglas.1 <- raster::crop(douglas.1, plot.shp)
-douglas.1 <- raster::mask(douglas.1, plot.shp)
-douglas.1 <- raster::projectRaster(douglas.1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(douglas.1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas.1.tif",
-                    progress = "text", overwrite = TRUE)
-
-## douglas 2
-douglas.2 <- raster::brick(files[[16]])
-
-all.plots %>% 
-  filter(Name == "douglas 2") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-douglas.2 <- raster::crop(douglas.2, plot.shp)
-douglas.2 <- raster::mask(douglas.2, plot.shp)
-douglas.2 <- raster::projectRaster(douglas.2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(douglas.2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-2.tif",
-                    progress = "text", overwrite = TRUE)
-
-## rad. altube
-rad.altube <- raster::brick(files[[16]])
-
-all.plots %>% 
-  filter(Name == "Rad. Altube") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-rad.altube <- raster::crop(rad.altube, plot.shp)
-rad.altube <- raster::mask(rad.altube, plot.shp)
-rad.altube <- raster::projectRaster(rad.altube, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(rad.altube, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-altube.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Otxandio and Laricio defol. Olaeta (Code B215)
-
-```{r}
-## otxandio
-otxandio <- raster::brick(files[[17]])
-
-all.plots %>% 
-  filter(Name == "a otxandio") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-otxandio <- raster::crop(otxandio, plot.shp)
-otxandio <- raster::mask(otxandio, plot.shp)
-otxandio <- raster::projectRaster(otxandio, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(otxandio, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/otxandio.tif",
-                    progress = "text", overwrite = TRUE)
-
-## Laricio defol. Olaeta
-laricio.defol.olaeta <- raster::brick(files[[17]])
-
-all.plots %>% 
-  filter(Name == "Laricio defol. Olaeta. pol") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-laricio.defol.olaeta <- raster::crop(laricio.defol.olaeta, plot.shp)
-laricio.defol.olaeta <- raster::mask(laricio.defol.olaeta, plot.shp)
-laricio.defol.olaeta <- raster::projectRaster(laricio.defol.olaeta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(laricio.defol.olaeta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio-defol-olaeta.tif",
-                    progress = "text", overwrite = TRUE)
-```
-
-## Douglas Legutio (Code B317 and B318)
-Here B317 is taken
-```{r}
-douglas.legutio <- raster::brick(files[[18]])
-
-all.plots %>% 
-  filter(Name == "Douglas Legutio") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-douglas.legutio <- raster::crop(douglas.legutio, plot.shp)
-douglas.legutio <- raster::mask(douglas.legutio, plot.shp)
-douglas.legutio <- raster::projectRaster(douglas.legutio, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(douglas.legutio, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-legutio.tif",
-                    progress = "text", overwrite = TRUE)
-```
-## Laricio Gatzaga (Code B319)
-
-```{r}
-laricio.gatzaga <- raster::brick(files[[20]])
-
-all.plots %>% 
-  filter(Name == "Laricio defol.Gatzaga") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-laricio.gatzaga <- raster::crop(laricio.gatzaga, plot.shp)
-laricio.gatzaga <- raster::mask(laricio.gatzaga, plot.shp)
-laricio.gatzaga <- raster::projectRaster(laricio.gatzaga, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(laricio.gatzaga, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio.gatzaga.tif",
-                    progress = "text", overwrite = TRUE)
-```
-## Azaceta (Code B320)
-
-```{r}
-azaceta <- raster::brick(files[[21]])
-
-all.plots %>% 
-  filter(Name == "a azaceta") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-azaceta <- raster::crop(azaceta, plot.shp)
-azaceta <- raster::mask(azaceta, plot.shp)
-azaceta <- raster::projectRaster(azaceta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(azaceta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/azaceta.tif",
-                    progress = "text", overwrite = TRUE)
-```
-## Gipuzkoa and Sequoia Legorreta (Code B321)
-
-```{r}
-## gipuzkoa
-gipuzkoa <- raster::brick(files[[22]])
-
-all.plots %>% 
-  filter(Name == "a gipuzkoa") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-gipuzkoa <- raster::crop(gipuzkoa, plot.shp)
-gipuzkoa <- raster::mask(gipuzkoa, plot.shp)
-gipuzkoa <- raster::projectRaster(gipuzkoa, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(gipuzkoa, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/gipuzkoa.tif",
-                    progress = "text", overwrite = TRUE)
-
-## Sequoia Legorreta
-sequoia.legorreta <- raster::brick(files[[22]])
-
-all.plots %>% 
-  filter(Name == "Sequoia Legorreta") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-sequoia.legorreta <- raster::crop(sequoia.legorreta, plot.shp)
-sequoia.legorreta <- raster::mask(sequoia.legorreta, plot.shp)
-sequoia.legorreta <- raster::projectRaster(sequoia.legorreta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(sequoia.legorreta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/sequoia.legorreta.tif",
-                    progress = "text", overwrite = TRUE)
-```
-## Douglas Albiztur (Code B322)
-
-```{r}
-douglas.albiztur <- raster::brick(files[[23]])
-
-all.plots %>% 
-  filter(Name == "Douglas Albiztur") %>% 
-  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
-  as("Spatial") -> plot.shp
-
-douglas.albiztur <- raster::crop(douglas.albiztur, plot.shp)
-douglas.albiztur <- raster::mask(douglas.albiztur, plot.shp)
-douglas.albiztur <- raster::projectRaster(douglas.albiztur, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-
-raster::writeRaster(douglas.albiztur, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-albiztur.tif",
-                    progress = "text", overwrite = TRUE)
-```

### 2018-04-04 02_scripts/01_hyperspectral_indices/04_hyperspectral-all-vegindex.Rmd
shiny app 2 by 2

@@
 title: "Calculate all possible vegetation indices that the `hsdar` package offers."
-output: html_notebook
-editor_options: 
-  chunk_output_type: inline
+output: 
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
 ---
 
-# Information about this script
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE,
+                      cache = TRUE)
+```
 
-This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`
-
-Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).
-
-Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
-It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.
+# Load packages
 
 ```{r}
-pacman::p_load(raster, hsdar, pbapply, parallel, rasterFunctions, tidyverse)
+library(raster)
+library(hsdar)
+library(pbmcapply)
+library(magrittr)
+library(purrr)
+library(glue)
 ```
@@
 ```{r}
-wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")
+wavelength <- readRDS("/data/patrick/mod/hyperspectral/wavelength.Rds")
 ```
 
-# List files
+## list files
 
 ```{r}
-files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
-                    pattern = ".tif$", full.names = TRUE)
+list.files("/data/patrick/mod/hyperspectral/atm-cropped", 
+                    pattern = ".tif$", full.names = TRUE) %>% 
+  map(~ brick(.x)) -> spectral_data
 ```
 
-# read in files as "RasterBrick"
+# Coerce to class "HyperSpecRaster"
+
 ```{r}
-spectral_data <- lapply(files, function(x) brick(x))
+hyperspecs = map(spectral_data, function(x) HyperSpecRaster(x, wavelength))
 ```
 
-# make class "HyperSpecRaster"
+# Specify indices
 
 ```{r}
-hyperspecs <- lapply(spectral_data, function(x) hsdar::HyperSpecRaster(x, wavelength))
-```
-
-# account for use on local laptop or server
-```{r}
-if (detectCores() > 10) {
-  cores <- 28
-} else (cores <- 1)
-```
-
-# do calculation
-
-Runtime server: ~ 1 min
-Runtime MBP: ~ 10 mins
-
-Writing files as `.gri` files because this format preserves layer names. TIFF format does not!!
-Layer names are changed using the `bnames` argument of function `rasterFunctions::hsdar.vegIndex()`.
-
-Specify vegetation indices.
-
-```{r}
-# subset for test purposes
-# hyperspecs <- list(hyperspecs[[17]])
-
 # specify which indices to calculate WITH DERIVATION INDICES
@@
 
-# do veg. indices calculation
+# Calculate vegetation indices
 
-```{r }
-hyperspecs %>% 
-  pblapply(function(x) 
-    vegindex(x, indices, filename = 
-             gsub(".tif", ".grd", 
-                  "/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/",
-                    "all-veg-indices-", basename(x@file@name)), na.rm = T,
-             bnames = indices), cl = cores) -> rasters
+Writing files as `.gri` files because this format preserves layer names. TIFF format does not!!
+Layer names are changed using the `bnames` argument of function `rasterFunctions::hsdar.vegIndex()`.
 
-# mapview(tmp[[1]])
+```{r, cache=FALSE}
+tmp <- pbmclapply(hyperspecs, function(x) 
+  vegindex(x, indices, 
+           filename = 
+             str_replace(glue("/data/patrick/mod/hyperspectral/all-veg-indices/{name}",
+                              name = basename(x@file@name)), ".tif", ".grd"), 
+           na.rm = TRUE, bnames = indices), mc.cores = 28)
 ```

### 2018-04-04 02_scripts/01_hyperspectral_indices/05_narrow-band-indices.Rmd
shiny app 2 by 2

@@
 title: "Calculate all possible narrow band indices"
-output: html_notebook
-editor_options: 
-  chunk_output_type: inline
+output: 
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
 ---
 
-# Information about this script
-
-This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`
-
-Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).
-
-Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
-It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.
-
-**Attention**
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE,
+                      cache = TRUE)
+```
 
@@
 
-```{r}
-pacman::p_load(raster, hsdar, pbapply, parallel, tidyverse, here, magrittr)
+```{r, results='hide'}
+library(raster)
+devtools::install_github("pat-s/hsdar")
+library(hsdar)
+library(pbmcapply)
+library(glue)
+library(magrittr)
+library(purrr)
 ```
 
-# Load wavelength and rasters
+# Specify wavelength
 
 ```{r}
-wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")
-
-list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
-           pattern = ".tif$", full.names = TRUE) %>%
-  lapply(function(x) brick(x)) %>% 
-  lapply(function(x) HyperSpecRaster(x, wavelength)) -> hyperspecs
+wavelength = readRDS("/data/patrick/mod/hyperspectral/wavelength.Rds")
 ```
 
-# account for use on local laptop or server
+# Read rasters (cropped by plot)
+
+## list files
+
 ```{r}
-if (detectCores() > 10) {
-  cores <- 28
-} else (cores <- 1)
+list.files("/data/patrick/mod/hyperspectral/atm-cropped", 
+                    pattern = ".tif$", full.names = TRUE) %>% 
+  map(~ brick(.x)) -> spectral_data
 ```
 
-# do calculation of narrow band indices
+# Coerce to class "HyperSpecRaster"
+
+```{r}
+hyperspecs = map(spectral_data, function(x) HyperSpecRaster(x, wavelength))
+```
+
+# Create narrow band indices
 
@@
 
+```{r, cache=FALSE}
+tmp <- pbmclapply(hyperspecs, function(x)  
+  nbi_raster(x,
+             filename = str_replace(glue("/data/patrick/mod/hyperspectral/",
+                             "narrow-band-indices/narrow-band-indices-{name}",
+                             name = basename(x@file@name)), ".tif", ".grd"), mc.cores = 15)) 
+rm(tmp)
+```
+
 ```{r}
-## Calculate NBI
-# hyperspecs <- list(hyperspecs[[1]])
-
-hyperspecs %>% 
-  pblapply(function(x) 
-    nbi_raster(x, filename = gsub(".tif", ".grd", 
-                                  "/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/",
-                                         "narrow-band-indices-", basename(x@file@name))), cl = cores) 
+sessionInfo()
 ```

### 2018-04-04 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd
shiny app 2 by 2

@@
 title: "Extract indices to points"
-output: html_notebook
-editor_options: 
-chunk_output_type: inline
+output: 
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
 ---
 
-# Information about this script
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE,
+                      cache = TRUE)
+```
 
-This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`
-
-Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).
-
-Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
-It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.
-
-```{r}
-pacman::p_load(raster, sf, here, magrittr, tidyverse, pbapply, parallel)
+```{r, results='hide'}
+library(raster)
+devtools::install_github("pat-s/hsdar")
+library(hsdar)
+library(pbmcapply)
+library(glue)
+library(magrittr)
+library(sf)
+library(purrr)
 ```

### 2018-04-04 02_scripts/02_preprocessing.Rmd
shiny app 2 by 2

@@
 library(raster)
+library(sf)
+library(dplyr)
+library(tidyr)
+library(magrittr)
+library(stringr)
 ```
@@
     ~ as_tibble(read.table(file = glue("/data/patrick/mod/survey_data/2016/{.x}.txt"), 
-                           sep = "\t", dec = ",", header = TRUE))) %>% 
-  set_names(names) -> data_all
+                           sep = "\t", dec = ",", header = TRUE,
+                           stringsAsFactors = FALSE, 
+                           na.strings=c("")))) %>% 
+  set_names(names) %>% 
+  map(~ rename_all(.x, tolower)) -> data_tbl
+```
+
+Laukiz2 has no coordinate info in the txt file. We need to read the shapefile and join them.
+
+### Join Laukiz2 coords
+
+```{r}
+laukiz2_coords = st_read("/data/patrick/raw/survey_data/2016/laukiz2/arboles/Laukiz2_arboles_ETRS89.shp")
+
+data_tbl$laukiz2 = left_join(data_tbl$laukiz2, laukiz2_coords, 
+                             by = c("num.arbol" = "id_arbol"))
+```
+
+Drop observations with no coordinate information.
+
+```{r}
+data_tbl %<>% 
+  map(~ drop_na(.x, c("x", "y")))
+```
+
+Coerce to `sf` objects.
+
+```{r}
+map(data_tbl, ~ st_as_sf(.x, coords = c("x", "y"))) -> data_sf
+```
+
+Select variables related to defoliation.
+
+```{r}
+data_sf %<>% 
+  map(~ select(.x, contains("def")))
+```
+
+Laukiz2 needs some clean up:
+
+"no" -> 0
+"*" -> NA
+Coerce from chr to int.
+Calculate "deftot" which is the mean of all three parts (bottom, mid, top).
+
+`rowSums` will thrown an error due to the "geometry" column so we need to add `drop = TRUE`.
+
+```{r}
+data_sf$laukiz2 %<>% 
+  mutate(x..de.defoliacion.parte.media = 
+           replace(.$x..de.defoliacion.parte.media, 
+                   .$x..de.defoliacion.parte.media == "no", "0")) %>% 
+  mutate(x..defoliacion.en.punta = 
+           replace(.$x..defoliacion.en.punta, 
+                   .$x..defoliacion.en.punta == "no", "0")) %>% 
+  mutate(x..defoliacion.base = 
+           replace(.$x..defoliacion.base,
+                   .$x..defoliacion.base == "no", "0")) %>% 
+  mutate(x..defoliacion.base = 
+           replace(.$x..defoliacion.base,
+                   .$x..defoliacion.base == "*", NA)) %>% 
+  mutate_if(is.character, as.numeric) %>% 
+  mutate(deftot = as.integer(rowMeans(.[, 1:3, drop = TRUE], na.rm = TRUE)))
 ```
@@
     ~ brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/all-veg-indices-{.x}.gri"))) %>% 
-  set_names(names) -> indices_all
+  set_names(names) -> veg_indices_all
 ```
@@
 
+```{r}
+map(names,
+    ~ brick(glue("/data/patrick/mod/hyperspectral/narrow-band-indices/narrow-band-indices-{.x}.gri"))) %>% 
+  set_names(names) -> nbi_indices_all
+```
+
 # Bind indices to response

### 2018-04-04 02_scripts/01_hyperspectral_indices/07_shiny_app/app.R

@@ -1,0 +1,113 @@
+library(shiny)
+library(mapview)
+library(leaflet)
+library(sp)
+
+ui <- pageWithSidebar(
+
+  # App title ----
+  headerPanel("Input"),
+
+  # Sidebar panel for inputs ----
+  sidebarPanel(        # Input: Selector for variable to plot against mpg ----
+                       selectInput("plot", "Plot:",
+                                   c("Aguinaga Pinaster Radiata" = "aguinaga-pinaster-radiata",
+                                   "Altube" = "altube", "Azaceta" = "azaceta",
+                                   "Barazar" = "barazar",
+                                   "Busturi = busturi",
+                                   "Caranca" = "caranca",
+                                   "Douglas 1" = "douglas-1",
+                                   "Douglas 2" = "douglas-2",
+                                   "Douglas Albiztur" = "douglas-albiztur",
+                                   "Douglas Legutio" = "douglas-legutio",
+                                   "Gipuzkoa" = "gipuzkoa",
+                                   "Laricio Defol Olaeta" = "laricio-defol-olaeta",
+                                   "Laricio Defol Urkiola" = "laricio-defol-urkiola",
+                                   "Laricio Gatzaga" = "laricio-gatzaga",
+                                   "Laukiz1" = "laukiz1",
+                                   "Laukiz2" = "laukiz2",
+                                   "Laukiz3" = "laukiz3",
+                                   "Luiando" = "luiando",
+                                   "Oiartzun" = "oiartzun",
+                                   "Otxandio" = "otxandio",
+                                   "Picea Abornicano" = "picea-abornicano",
+                                   "Pinas Santa Koloma" = "pinas-santa-koloma",
+                                   "Pinast Arza" = "pinast-arza",
+                                   "Rad Altube" = "rad-altube",
+                                   "Rad Menagaray 1" = "rad-menagaray-1",
+                                   "Rad Menagaray 2" = "rad-menagaray-2",
+                                   "Sequoia Legorreta" = "sequoia-legorreta",
+                                   "Silvest Nograro" = "silvest-nograro")),
+                       selectInput("index", "Index",
[…]

### 2018-04-04 02_scripts/01_hyperspectral_indices/07_shiny_app/app2.R

@@ -1,0 +1,288 @@
+library(shiny)
+library(mapview)
+library(leaflet)
+library(sp)
+library(shinydashboard)
+
+header <- dashboardHeader(title = "Healthy Forest")
+
+sidebar <- dashboardSidebar(
+  selectInput(
+    "plot", "Plot 1",
+    selected = "laukiz1",
+    c(
+      "Aguinaga Pinaster Radiata" = "aguinaga-pinaster-radiata",
+      "Altube" = "altube", "Azaceta" = "azaceta",
+      "Barazar" = "barazar",
+      "Busturi" = "busturi",
+      "Caranca" = "caranca",
+      "Douglas 1" = "douglas-1",
+      "Douglas 2" = "douglas-2",
+      "Douglas Albiztur" = "douglas-albiztur",
+      "Douglas Legutio" = "douglas-legutio",
+      "Gipuzkoa" = "gipuzkoa",
+      "Laricio Defol Olaeta" = "laricio-defol-olaeta",
+      "Laricio Defol Urkiola" = "laricio-defol-urkiola",
+      "Laricio Gatzaga" = "laricio-gatzaga",
+      "Laukiz1" = "laukiz1",
+      "Laukiz2" = "laukiz2",
+      "Laukiz3" = "laukiz3",
+      "Luiando" = "luiando",
+      "Oiartzun" = "oiartzun",
+      "Otxandio" = "otxandio",
+      "Picea Abornicano" = "picea-abornicano",
+      "Pinas Santa Koloma" = "pinas-santa-koloma",
+      "Pinast Arza" = "pinast-arza",
+      "Rad Altube" = "rad-altube",
+      "Rad Menagaray 1" = "rad-menagaray-1",
+      "Rad Menagaray 2" = "rad-menagaray-2",
+      "Sequoia Legorreta" = "sequoia-legorreta",
+      "Silvest Nograro" = "silvest-nograro"
+    )
+  ),
+  selectInput(
+    "plot2", "Plot 2",
+    selected = "luiando",
+    c(
+      "Aguinaga Pinaster Radiata" = "aguinaga-pinaster-radiata",
+      "Altube" = "altube", "Azaceta" = "azaceta",
+      "Barazar" = "barazar",
+      "Busturi" = "busturi",
+      "Caranca" = "caranca",
+      "Douglas 1" = "douglas-1",
+      "Douglas 2" = "douglas-2",
[…]

### 2018-04-17 01_writing/Schratz_hyperspectral.tex
add first .tex draft, update scripts

%%%% using 'arara' 4.0
% arara: xelatex: {synctex: yes, interaction: nonstopmode}
% arara: bibtex
% arara: xelatex: {synctex: yes, interaction: nonstopmode}
% arara: xelatex: {synctex: yes, interaction: nonstopmode}

% arara: indent: {overwrite: yes}

% arara: clean: { extensions: [aux, bcf, cod, blg, lof, lot, out, toc, log, xml, bak0 ] }

\documentclass[review]{elsarticle}

% Figures Links, mittig und rechts platzieren
\usepackage[export]{adjustbox}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}

% prevents that appendices are moved behind references
\usepackage{placeins}

\usepackage[nolist]{acronym}

\usepackage{longtable}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{float}

% enable linking to subsubsection
\setcounter{secnumdepth}{3}

% various symbols, e.g. \degree
\usepackage{gensymb}

\usepackage[hidelinks]{hyperref}

\usepackage{lineno}
\modulolinenumbers[5]

% set autoref abbr for appendix
\newcommand*{\Appendixautorefname}{appendix}

\journal{Journal "Remote Sensing of environment"}

% line breaks in table cells
\newcommand{\specialcell}[2][l]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

% tilde
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}

%% APA style
\bibliographystyle{model5-names}\biboptions{authoryear}

\begin{document}

\begin{frontmatter}

	\title{title}

	%% Group authors per affiliation:
	\author[FSU]{Patrick Schratz}
	\cortext[mycorrespondingauthor]{Corresponding author}
	\ead{patrick.schratz@uni-jena.de}

	\author[FSU]{Jannes Muenchow}
	\author[NEIKER]{Eugenia Iturritxa}
	%\author[TUDO]{Jakob Richter}
	\author[FSU]{Alexander Brenning}

	\address[FSU]{Department of Geography, GIScience group, Grietgasse 6, 07743, Jena, Germany}
	%\address[NEIKER]{NEIKER, Granja Modelo –Arkaute, Apdo. 46, 01080 Vitoria-Gasteiz, Arab, Spain}
	%\address[TUDO]{Department of Statistics, TU Dortmund University, Germany}

	\begin{abstract}

	\end{abstract}

	\begin{keyword}
		hyperspectral imagery \sep statistical learning \sep spatial cross-validation
	\end{keyword}

\end{frontmatter}

\linenumbers

% längste Abkürzung steht hier!!! in eckigen Klammern
\begin{acronym}[AUROC]

	% geringerer Zeilenabstand
	%\setlength{\itemsep}{-\parsep}
	\acro{ANN}{Artificial Neural Network}
	\acro{AUROC}{Area Under the Receiver Operating Characteristics Curve}
	\acro{BRT}{Boosted Regression Trees}
	\acro{CART}{Classification and Regression Trees}
	\acro{CV}{cross-validation}
	\acro{ENM}{Environmental Niche Modeling}
	\acro{FPR}{False Positive Rate}
	\acro{GAM}{Generalized Additive Model}
	\acro{GBM}{Gradient Boosting Machine}
	\acro{GLM}{Generalized Linear Model}
	\acro{ICGC}{Institut Cartografic i Geologic de Catalunya}
	\acro{IQR}{Interquartile Range}
	\acro{WKNN}{Weighted $k$-nearest neighbor}
	\acro{MARS}{Multivariate Adaptive Regression Splines}
	\acro{MEM}{Maximum Entropy Model}
	\acro{NBI}{Narrow Band Index}
	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
	\acro{PISR}{Potential Incoming Solar Radiation}
	\acro{RBF}{Radial Basis Function}
	\acro{RF}{Random Forest}
	\acro{SDM}{Species Distribution Modeling}
	\acro{SVM}{Support Vector Machines}
	\acro{TPR}{True Positive Rate}
\end{acronym}

\section{Introduction}
\label{sec:intro}

\section{Data and study area}

\cite{Brenning2012}

% \begin{figure} [t!]
% 	\begin{center}
% 		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../../04_figures/01_data/study_area.pdf}}
% 		\caption[Study area]{Spatial distribution of tree observations within the Basque Country, northern Spain, showing infection state by \textit{Diplodia sapinea}.}
% 		\label{fig: study_area}
% 	\end{center}
% \end{figure}

\subsection{Hyperspectral data}

The airborne hyperspectral data was acquired on September 28th and October 5th 2016 around 12 am.
The images were taken by an AISAEAGLE-II sensor from the \ac{ICGC}.
All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by \ac{ICGC}.

Additional information is provided in Table 1:

% parameter limits
\begin{table}[b!]
\centering
\caption[t]{Specifications of hyperspectral data.}
\begingroup\footnotesize
\begin{tabular}{ll}
	\\
	Characteristic         & Value                               \\
	\hline
	Geometric resolution   & 1 m                                 \\
	Radiometric resolution & 12 bit                              \\
	Spectral resolution    & 126 bands (404.08 nm - 996.31 nm)   \\
	Correction:            & Radiometric, geometric, atmospheric
\end{tabular}
\endgroup
\label{tab:hyperparameter_limits}
\end{table}

\section{Methods}

\subsection{Index calculation}

% link to PDF with veg indeces
All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that are offered by the \texttt{hsdar} package have been calculated.
Additionally, all possible \ac{NBI} were calculated from the data using the following formula:

\begin{equation}
	NBI_{i,j} = \frac{B_{i} - B_{j}}{B_{i} + B_{j}}
\end{equation}

\noindent
where $i$ and $j$ are the respective band numbers.

To account for geometric offsets, we calculated every index five times using a buffer from 1 - 5 meter around the centroid of the respective tree.
The mean value of all pixels touched by the buffer was assigned as the final value for each index.
The exact number of contributing pixels of an index cannot be determined as it depends on the location of the tree within the pixel grid.
If a tree is located at the border of a pixel, the same buffer (e.g. 3 m) will include more pixels than if the point is located at the center of a pixel.
Also, if a tree is located at the border of the image data, some directions of the buffer may not contain values.
Missing values were removed from the mean value calculation.
However, the bigger the buffer size, the more pixels are represented by the index value.
In total, 7875 possible \ac{NBI}s have been calculated for each buffer size summing up to 39375 indices.
The same applies to the vegetation indices which sum up to 450 indices leading to a total number of 39825 indices per tree.

\subsection{Ridge regression}

\section*{References}

\bibliography{Biblio_hyperspectral}

\end{document}

### 2018-04-17 02_scripts/03-modeling.Rmd
add first .tex draft, update scripts

---
title: "03-modeling"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(mlr)
library(glue)
library(sf)
library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
```

# Read data

```{r}
data = map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~ readRDS(glue("/data/patrick/mod/tree-per-tree/2016/{.x}/all-indices-{.x}.rda")))
```

# Extract coordinates

```{r}
coords = map(data, ~ as.data.frame(st_coordinates(.x)))
```

# Remove geometry

```{r}
data %<>%
  map(~ st_set_geometry(.x, NULL)) 
```

# Remove columns with NA values

```{r}
test = as.data.table(data[[1]])
setDT(test)[test[, !Reduce(`|`, lapply(.SD, is.na))]]
```

```{r}
task = makeRegrTask(id = "laukiz1", data = data[[1]], 
                    target = "deftot", coordinates = coords[[1]],
                    )
```

```{r}
# predict.type for 'auc' (AUROC)
lrn <- makeLearner("regr.LiblineaRL2L2SVR")
getHyperPars(lrn)
getParamSet(lrn)
```

```{r}
ps <- makeParamSet(makeNumericParam("cost", lower = 0, upper = 100))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 50)
inner <- makeResampleDesc("SpCV", iters = 5)
```

```{r}
tune_wrapper <- makeTuneWrapper(lrn, resampling = inner, par.set = ps,
                               control = ctrl, show.info = TRUE,
                               measures = list(rmse))
```

```{r}
outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 2)
```

```{r}
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
              mc.set.seed = TRUE) # only parallelize the tuning

set.seed(12345)
resa_brt_spatial <- mlr::resample(tune_wrapper, task,
                                  resampling = outer, extract = getTuneResult,
                                  show.info = TRUE, measures = list(rmse))

parallelStop()
```

### 2018-04-26 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd
26/04/18

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE,
-                      cache = TRUE)
+knitr::opts_chunk$set(echo = TRUE)
 ```
+
+# Load packages
 
@@
 library(purrr)
+library(dplyr)
 ```
 
-# Laukiz1
+The points to which the indices are extracted to are cleaned in the file "02_preprocessing.Rmd".
 
-## Load points
+# Create function
+
+See step-by-step example for Laukiz 1 below.
 
 ```{r}
-st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/laukiz1.shp", 
-        quiet = TRUE) %>% 
-  as("Spatial") -> laukiz1_points
-```
+extract_to_plot = function(plot_name, buffer, cores, bf_name) {
+  points = st_read(glue("/data/patrick/mod/tree-per-tree/2016/{plot_name}/{plot_name}.gpkg"),
+          quiet = TRUE)
 
-## Load rasters
+  ras_veg_indices <- brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/{plot_name}.grd"))
 
-```{r}
-laukiz1_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz1")
-```
+  ras_nbi <- brick(glue("/data/patrick/mod/hyperspectral/narrow-band-indices/narrow-band-indices-{plot_name}.grd"))
 
-```{r}
-laukiz1_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz1")
-```
+  # calculate buffered veg index
+  out_veg <- pbmclapply(buffer, function(x) 
+    raster::extract(ras_veg_indices, points, buffer = x, 
+                    fun = mean, df = TRUE,
+                    na.rm = TRUE), mc.cores = cores)
 
-## Extract values to points
+  out_veg %<>% 
+    map2(seq_along(buffer), ~ setNames(.x, glue("bf", .y, "_{name}", 
+                                             name = names(out_veg[[.y]]))))
 
-### Veg indices
+  out_nbi <- pbmclapply(buffer, function(x)
+    raster::extract(ras_nbi, points, buffer = x,
+                    fun = mean, df = TRUE,
+                    na.rm = TRUE) , mc.cores = cores)
 
-Buffer size:
+  out_nbi %<>%
+    map2(seq_along(buffer), ~ setNames(.x, glue("bf", .y, "_{name}",
+                                             name = names(out_nbi[[.y]]))))
 
-* 1m
-* 2m
-* 3m
-* 4m
-* 5m
+  # merge all data frames (buffers)
+  all_veg <- bind_cols(out_veg)
 
-```{r}
-out_veg <- lapply(seq_along(1:5), function(x) 
-  raster::extract(laukiz1_veg_indices, laukiz1_points, buffer = x, 
-                  fun = mean, df = TRUE))
-```
+  all_nbi <- bind_cols(out_nbi)
 
-Label output names according to buffer size
+  points %<>% 
+    bind_cols(all_veg) %>% 
+    bind_cols(all_nbi)
 
-```{r}
-out_veg <- lapply(seq_along(out_veg), function(x) 
-  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
-```
-
-### NBI
-
-This might take a while - nested parallelization here! 
-Only outer parallelization: 1 hour
-
-This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!
-
-NOTE 05 April: Nested parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
-I think we are better of using only outer parallelization here.
-
-```{r}
-if (detectCores() > 45 ) {
-
-  Sys.time <- Sys.time()
-
-  out_nbi <- pblapply(seq_along(1:5), function(x) {
-    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
-    raster::extract(laukiz1_nbi, laukiz1_points, buffer = x, 
-                    fun = mean, df = TRUE) } , cl = 5)
-
-  print(Sys.time() - Sys.time)
-} else {
-  stop("please run this script on a server with more than 45 cores!")
+  # .rda file (data frame)
+  points %>% 
+    saveRDS(glue("/data/patrick/mod/tree-per-tree/2016/{plot_name}/all-indices-{plot_name}_{bf_name}.rda"))
 }
@@
 
-### Bind to points
+# Process
+
+## Buffer: 2
 
 ```{r}
-# merge all data frames
-df_out <- bind_cols(out_veg)
-
-df_out1 <- bind_cols(out_nbi)
-
-laukiz1_points %>% 
-  st_as_sf() %>% 
-  bind_cols(df_out) %>% 
-  bind_cols(df_out1) -> laukiz1_points
+pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"), 
+           function(x) extract_to_plot(x, 2, cores = 1, "bf2"), 
+           mc.cores = 4)
 ```
 
-## Save files
+<!-- ## Buffer: 1 - 10 -->
 
-```{r}
-# .rda file (data frame)
-as.data.frame(laukiz1_points) %>% 
-  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/all-indices-laukiz1.rda")
-```
+<!-- ```{r} -->
+<!-- pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),  -->
+<!--            function(x) extract_to_plot(x, seq(1:10), cores = 10, "bf1-10"),  -->
+<!--            mc.cores = 4) -->
+<!-- ``` -->
 
-# Laukiz2
+<!-- ## Buffer: 1,5,10,15,20 -->
 
-## Load points
+<!-- ```{r} -->
+<!-- pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),  -->
+<!--            function(x) extract_to_plot(x, c(1, 5, 10, 15, 20), cores = 5, "bf1-5-10-15-20"),  -->
+<!--            mc.cores = 4) -->
+<!-- ``` -->
 
-```{r}
-print("Starting Laukiz2")
-```
+<!-- ## Buffer: 1, 10, 20, 30, 40, 50 -->
 
-```{r}
-st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/laukiz2.shp", 
-        quiet = TRUE) %>% 
-  as("Spatial") -> laukiz2_points
-```
-
-## Load rasters
-
-```{r}
-laukiz2_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz2")
-```
-
-```{r}
-laukiz2_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz2")
-```
-
-## Extract values to points
-
-### Veg indices
-
-Buffer size:
-
-* 1m
-* 2m
-* 3m
-* 4m
-* 5m
-
-```{r}
-out_veg <- lapply(seq_along(1:5), function(x) 
-  raster::extract(laukiz2_veg_indices, laukiz2_points, buffer = x, 
-                  fun = mean, df = TRUE))
-```
-
-Label output names according to buffer size
-
-```{r}
-out_veg <- lapply(seq_along(out_veg), function(x) 
-  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
-```
-
-### NBI
-
-This might take a while - nested parallelization here! 
-Only outer parallelization: 1 hour
-
-This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!
-
-NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
-I think we are better of using only outer parallelization here.
-
-```{r}
-if (detectCores() > 45 ) {
-
-  Sys.time <- Sys.time()
-
-  out_nbi <- pblapply(seq_along(1:5), function(x) {
-    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
-    raster::extract(laukiz2_nbi, laukiz2_points, buffer = x, 
-                    fun = mean, df = TRUE) } , cl = 5)
-  #endCluster()
-  print(Sys.time() - Sys.time)
-} else {
-  stop("please run this script on a server with more than 45 cores!")
-}
-
-### local debugging code
-# out_nbi <- raster::extract(laukiz2_nbi, laukiz2_points, buffer = 2, 
-#                            fun = mean, df = TRUE)
-```
-
-### Bind to points
-
-```{r}
-# merge all data frames
-df_out <- bind_cols(out_veg)
-
-df_out1 <- bind_cols(out_nbi)
-
-laukiz2_points %>% 
-  st_as_sf() %>% 
-  bind_cols(df_out) %>% 
-  bind_cols(df_out1) -> laukiz2_points
-```
-
-## Save files
-
-```{r}
-# .rda file (data frame)
-as.data.frame(laukiz2_points) %>% 
-  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/all-indices-laukiz2.rda")
-```
-
-# Oiartzun
-
-## Load points
-
-```{r}
-print("Starting oiartzun")
-```
-
-```{r}
-st_read("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/oiartzun.shp", 
-        quiet = TRUE) %>% 
-  as("Spatial") -> oiartzun_points
-```
-
-## Load rasters
-
-```{r}
-oiartzun_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-oiartzun")
-```
-
-```{r}
-oiartzun_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-oiartzun")
-```
-
-## Extract values to points
-
-### Veg indices
-
-Buffer size:
-
-* 1m
-* 2m
-* 3m
-* 4m
-* 5m
-
-```{r}
-out_veg <- lapply(seq_along(1:5), function(x) 
-  raster::extract(oiartzun_veg_indices, oiartzun_points, buffer = x, 
-                  fun = mean, df = TRUE))
-```
-
-Label output names according to buffer size
-
-```{r}
-out_veg <- lapply(seq_along(out_veg), function(x) 
-  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
-```
-
-### NBI
-
-This might take a while - nested parallelization here! 
-Only outer parallelization: 1 hour
-
-This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!
-
-NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
-I think we are better of using only outer parallelization here.
-
-```{r}
-if (detectCores() > 45 ) {
-
-  Sys.time <- Sys.time()
-
-  out_nbi <- pblapply(seq_along(1:5), function(x) {
-    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
-    raster::extract(oiartzun_nbi, oiartzun_points, buffer = x, 
-                    fun = mean, df = TRUE) } , cl = cores)
-
-  print(Sys.time() - Sys.time)
-} else {
-  stop("please run this script on a server with more than 45 cores!")
-}
-```
-
-### Bind to points
-
-```{r}
-# merge all data frames
-df_out <- bind_cols(out_veg)
-
-df_out1 <- bind_cols(out_nbi)
-
-oiartzun_points %>% 
-  st_as_sf() %>% 
-  bind_cols(df_out) %>% 
-  bind_cols(df_out1) -> oiartzun_points
-```
-
-## Save files
-
-```{r}
-# .rda file (data frame)
-as.data.frame(oiartzun_points) %>% 
-  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/all-indices-oiartzun.rda")
-```
-
-# Luiando
-
-## Load points
-
-```{r}
-print("Starting luiando")
-```
-
-```{r}
-st_read("/home/shares/data/LIFE/mod/tree-per-tree/luiando/luiando.shp", 
-        quiet = TRUE) %>% 
-  as("Spatial") -> luiando_points
-```
-
-## Load rasters
-
-```{r}
-luiando_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-ayala-aiara") 
-```
-
-```{r}
-luiando_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-ayala-aiara")
-```
-
-## Extract values to points
-
-### Veg indices
-
-Buffer size:
-
-* 1m
-* 2m
-* 3m
-* 4m
-* 5m
-
-```{r}
-out_veg <- lapply(seq_along(1:5), function(x) 
-  raster::extract(luiando_veg_indices, luiando_points, buffer = x, 
-                  fun = mean, df = TRUE))
-```
-
-Label output names according to buffer size
-
-```{r}
-out_veg <- lapply(seq_along(out_veg), function(x) 
-  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
-```
-
-### NBI
-
-This might take a while - nested parallelization here! 
-Only outer parallelization: 1 hour
-
-This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!
-
-NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
-I think we are better of using only outer parallelization here.
-
-```{r}
-if (detectCores() > 45 ) {
-
-  Sys.time <- Sys.time()
-
-  out_nbi <- pblapply(seq_along(1:5), function(x) {
-    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
-    raster::extract(luiando_nbi, luiando_points, buffer = x, 
-                    fun = mean, df = TRUE) } , cl = cores)
-
-  print(Sys.time() - Sys.time)
-} else {
-  stop("please run this script on a server with more than 45 cores!")
-}
-```
-
-### Bind to points
-
-```{r}
-# merge all data frames
-df_out <- bind_cols(out_veg)
-
-df_out1 <- bind_cols(out_nbi)
-
-luiando_points %>% 
-  st_as_sf() %>% 
-  bind_cols(df_out) %>% 
-  bind_cols(df_out1) -> luiando_points
-```
-
-## Save files
-
-```{r}
-# .rda file (data frame)
-as.data.frame(luiando_points) %>% 
-  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/luiando/all-indices-luiando.rda")
-```
+<!-- ```{r} -->
+<!-- pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),  -->
+<!--            function(x) extract_to_plot(x, c(1, 10, 20, 30, 40, 50), cores = 6,  -->
+<!--                                        "bf1-10-20-30-40-50"),  -->
+<!--            mc.cores = 2) # due to RAM consumption -->
+<!-- ``` -->

### 2018-04-26 02_scripts/021_preprocessing.Rmd
26/04/18

---
title: "02-preprocessing"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r}
library(purrr)
library(here)
library(readr)
library(tibble)
library(glue)
library(raster)
library(sf)
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
library(fs)
library(pbmcapply)
library(data.table)
```

# Read data

Reading in data from .txt files and exporting them as a Geopackage (.gpkg).

## Survey data

```{r}
names <- c("laukiz1", "laukiz2", "luiando", "oiartzun")

map(
  names,
  ~ as_tibble(read.table(
    file = glue("/data/patrick/mod/survey_data/2016/{.x}.txt"),
    sep = "\t", dec = ",", header = TRUE,
    stringsAsFactors = FALSE,
    na.strings = c("")
  ))
) %>%
  set_names(names) %>%
  map(~ rename_all(.x, tolower)) -> data_tbl
```

Laukiz2 has no coordinate info in the txt file. We need to read the shapefile and join them.

### Join Laukiz2 coords

```{r}
laukiz2_coords <- st_read("/data/patrick/raw/survey_data/2016/laukiz2/arboles/Laukiz2_arboles_ETRS89.shp")

data_tbl$laukiz2 <- left_join(data_tbl$laukiz2, laukiz2_coords,
  by = c("num.arbol" = "id_arbol")
)
```

Drop observations with no coordinate information.

```{r}
data_tbl %<>%
  map(~ drop_na(.x, c("x", "y")))
```

Coerce to `sf` objects.

```{r}
map(data_tbl, ~ st_as_sf(.x, coords = c("x", "y"))) -> data_sf
```

Select variables related to defoliation.

```{r}
data_sf %<>%
  map(~ dplyr::select(.x, contains("def")))
```

Laukiz2 needs some clean up:

"no" -> 0
"*" -> NA
Coerce from chr to int.
Calculate "deftot" which is the mean of all three parts (bottom, mid, top).

`rowSums` will thrown an error due to the "geometry" column so we need to add `drop = TRUE`.

```{r}
data_sf$laukiz2 %<>%
  mutate(
    x..de.defoliacion.parte.media =
      replace(
        .$x..de.defoliacion.parte.media,
        .$x..de.defoliacion.parte.media == "no", "0"
      )
  ) %>%
  mutate(
    x..defoliacion.en.punta =
      replace(
        .$x..defoliacion.en.punta,
        .$x..defoliacion.en.punta == "no", "0"
      )
  ) %>%
  mutate(
    x..defoliacion.base =
      replace(
        .$x..defoliacion.base,
        .$x..defoliacion.base == "no", "0"
      )
  ) %>%
  mutate(
    x..defoliacion.base =
      replace(
        .$x..defoliacion.base,
        .$x..defoliacion.base == "*", NA
      )
  ) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(deftot = as.integer(rowMeans(.[, 1:3, drop = TRUE], na.rm = TRUE))) %>%
  drop_na(deftot) %>%
  select(deftot, geometry)
```

# Write clean data as .gpkg

Writing to cifs mounts is currently not working so we have to copy and delete the files.

```{r}
# write
walk2(data_sf, seq_along(data_sf), ~ write_sf(.x, glue(here("{name}.gpkg"),
  name = names(data_sf[.y])
)))

# copy
walk(names(data_sf), ~ file_copy(glue(here("{.x}.gpkg")),
  glue("/data/patrick/mod/tree-per-tree/2016/{.x}/{.x}.gpkg"),
  overwrite = TRUE
))

# delete
walk(names(data_sf), ~ file_delete(glue(here("{.x}.gpkg"))))
```

### 2018-04-26 02_scripts/022_preprocessing.Rmd
26/04/18

---
title: "02-preprocessing"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r}
library(purrr)
library(tibble)
library(glue)
library(dplyr)
library(tidyr)
library(magrittr)
library(sf)
library(pbmcapply)
library(data.table)
```

# Read points with indices

Read the clean points with all extracted indices for each plot.
Indices are extracted in script `01_hyperspectral/06_extract_indices_*`.

```{r}
data = map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~ readRDS(glue("/data/patrick/mod/tree-per-tree/2016/{.x}/all-indices-{.x}_bf2.rda")))
```

# Remove columns with NA values and "ID" columns

Ref: https://stackoverflow.com/questions/41343900/remove-row-columns-with-any-all-nan-values/41343980#41343980

```{r}
data %<>% 
  map(~ dplyr::select(.x, -contains("_ID"))) %>% 
  map(~ as.data.table(.x)) %>% 
  map(~ Filter(function(x) !any(is.na(x)), .x))
```

# Extract coordinates

The scaling of variables won't work with the `geom` column attached.
So we need to safe the coordinates, remove them and attach them afterwards again.
Coordinates are needed for the spatial partitioning during modeling.

```{r}
coords = map(data, ~ st_as_sf(.x, crs = 32630)) %>% 
  map(~ as.data.frame(st_coordinates(.x)))
coords = saveRDS(coords, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
data %<>% 
  map(~ st_as_sf(.x, crs = 32630)) %>% 
  map(~ st_set_geometry(.x, NULL)) 
```

# Standardize variables

```{r}
# data %<>% 
#   pbmclapply(function(x) mutate_at(x, .funs = scale, 
#                    .vars = vars(-deftot)),
#              mc.cores = 4
#       )
# saveRDS(data, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf1.rda")

# data.table so much faster
data %<>% 
  pbmclapply(function(x) {
    cols = names(x)[2:length(names(x))]
    x[, (cols) := lapply(.SD, scale), .SDcols=cols]
  }, mc.cores = 4
      )
saveRDS(data, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
```

### 2018-04-26 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
26/04/18

---
title: "03-modeling"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(mlr)
library(glue)
library(sf)
library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
library(pbmcapply)
```

# Read data

Variables per plot:

Lauki1: 7594
Laukiz2: 7594
Oiartzun: 7891
Luiando: 6073

Why is there such a huge difference in the number of indices without NAs for the plots?

```{r}
data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
```

# Select plot

```{r}
data = data[[1]]
coords = coords[[1]]
```

# Create task

```{r}
task = makeRegrTask(id = "laukiz1", data = data, 
                    target = "deftot", coordinates = coords)
```

# Modeling

## glmnet

### Ridge

We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
The value of `s` determines which trained model will be chosen.
If `s` != `lambda`, linear interpolation or refitting (this) is applied.

To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.

Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
However, we would do a random partitioning instead of a spatial partitioning here.

Coefficients of ridge regression should not be used for inference as they are biased. 
They can be seen as a variable importance measure. 
The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464

```{r}
# predict.type for 'auc' (AUROC)
lrn_glmnet <- makeLearner("regr.glmnet",
                          alpha = 0,
                          standardize = FALSE,
                          intercept = FALSE)
                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
# getHyperPars(lrn_glmnet)
# getParamSet(lrn_glmnet)
```

To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.

```{r}
glmnet_train = mlr::train(lrn_glmnet, task)
summary(glmnet_train$learner.model$lambda)
```

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 77, upper = 7700))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 200)
inner <- makeResampleDesc("SpCV", iters = 10)
```

#### SpCV

```{r}
tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
                                control = ctrl, show.info = TRUE,
                                measures = list(rmse))
```

```{r}
outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
```

```{r}
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
              mc.set.seed = TRUE) # only parallelize the tuning

set.seed(12345)
resa_glmnet <- mlr::resample(tune_wrapper, task,
                             resampling = outer, extract = getTuneResult,
                             show.info = TRUE, measures = list(rmse))
parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
```

#### TuneParams

```{r}
configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
              mc.set.seed = TRUE) # only parallelize the tuning
set.seed(12345)
params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
                                 par.set = ps_glmnet, control = ctrl, 
                                 measure = list(rmse))
parallelStop()
saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_ridge.rda")
```

#### Tuned results

We use the `s` value from the tuning. 
As we will always use the complete dataset, this should be the best value.

```{r}
lrn_glmnet_tuned <- makeLearner("regr.glmnet",
                                alpha = 0,
                                s = 1810,
                                standardize = FALSE, 
                                intercept = FALSE)

glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
```

Inspect the highest absolute coefficients (positive and negative).
The first entry is the intercept and is therefore left out.

```{r}
tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
  arrange(desc(coef))
```
```{r}
tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
  arrange(coef)
```

```{r}
plot(glmnet_train_tuned$learner.model, "lambda")
```

```{r}
plot(coef(glmnet_train_tuned$learner.model))
```

### Lasso

We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
The value of `s` determines which trained model will be chosen.
If `s` != `lambda`, linear interpolation or refitting is applied.

To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.

Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
However, we would do a random partitioning instead of a spatial partitioning here.

Coefficients of ridge regression should not be used for inference as they are biased. 
They can be seen as a variable importance measure then? The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464

```{r}
lrn_glmnet <- makeLearner("regr.glmnet",
                          alpha = 1,
                          standardize = FALSE,
                          intercept = FALSE)
# getHyperPars(lrn_glmnet)
# getParamSet(lrn_glmnet)
```

To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.

```{r}
glmnet_train = mlr::train(lrn_glmnet, task)
summary(glmnet_train$learner.model$lambda)
```

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 200)
inner <- makeResampleDesc("SpCV", iters = 5)
```

#### SpCV

```{r}
tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
                                control = ctrl, show.info = TRUE,
                                measures = list(rmse))
```

```{r}
outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 2)
```

```{r}
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
              mc.set.seed = TRUE) # only parallelize the tuning

set.seed(12345)
resa_glmnet <- mlr::resample(tune_wrapper, task,
                             resampling = outer, extract = getTuneResult,
                             show.info = TRUE, measures = list(rmse))

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
```

#### TuneParams

```{r}
configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
              mc.set.seed = TRUE) # only parallelize the tuning
set.seed(12345)
params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
                                 par.set = ps_glmnet, control = ctrl, 
                                 measure = list(rmse))
parallelStop()
```

#### Tuned results

We use the `s` value from the tuning. 
As we will always use the complete dataset, this should be the best value.

```{r}
lrn_glmnet_tuned <- makeLearner("regr.glmnet",
                                alpha = 0,
                                s = 1810,
                                standardize = FALSE, 
                                intercept = FALSE)
#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
```

Inspect the highest absolute coefficients (positive and negative).
The first entry is the intercept and is therefore left out.

```{r}
tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
  arrange(desc(coef))
```
```{r}
tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
  arrange(coef)
```

```{r}
plot(glmnet_train_tuned$learner.model, "norm")
```

```{r}
plot(coef(glmnet_train_tuned$learner.model))
```

### Elasticnet

We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
The value of `s` determines which trained model will be chosen.
If `s` != `lambda`, linear interpolation or refitting is applied.

To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.

Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
However, we would do a random partitioning instead of a spatial partitioning here.

Coefficients of ridge regression should not be used for inference as they are biased. 
They can be seen as a variable importance measure then? The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464

```{r}
lrn_glmnet <- makeLearner("regr.glmnet",
                          standardize = FALSE,
                          intercept = FALSE)
#getHyperPars(lrn_glmnet)
#getParamSet(lrn_glmnet)
```

To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.

```{r}
glmnet_train = mlr::train(lrn_glmnet, task)
summary(glmnet_train$learner.model$lambda)
```

For the elasticnet we also tune `alpha`.
As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.

```{r}
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7700),
                          makeNumericParam("alpha", lower = 0, upper = 1))
```

```{r}
ctrl <- makeTuneControlRandom(maxit = 200)
inner <- makeResampleDesc("SpCV", iters = 10)
```

#### SpCV

```{r}
tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
                                control = ctrl, show.info = TRUE,
                                measures = list(rmse))
```

```{r}
outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
```

```{r}
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
              mc.set.seed = TRUE) # only parallelize the tuning

set.seed(12345)
resa_glmnet <- mlr::resample(tune_wrapper, task,
                             resampling = outer, extract = getTuneResult,
                             show.info = TRUE, measures = list(rmse))

parallelStop()
saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
```

#### TuneParams

```{r}
configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
library(parallelMap)
parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
              mc.set.seed = TRUE) # only parallelize the tuning
set.seed(12345)
params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
                                 par.set = ps_glmnet, control = ctrl, 
                                 measure = list(rmse))
parallelStop()
```

#### Tuned results

We use the `s` value from the tuning. 
As we will always use the complete dataset, this should be the best value.

```{r}
lrn_glmnet_tuned <- makeLearner("regr.glmnet",
                                alpha = 0,
                                s = 1810,
                                standardize = FALSE, 
                                intercept = FALSE)
#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
```

Inspect the highest absolute coefficients (positive and negative).
The first entry is the intercept and is therefore left out.

```{r}
tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
  arrange(desc(coef))
```
```{r}
tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
  arrange(coef)
```

```{r}
plot(glmnet_train_tuned$learner.model, "lambda")
```

```{r}
plot(coef(glmnet_train_tuned$learner.model))
```

### 2018-04-26 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
26/04/18

--- 2018-04-26 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
@@
 ```{r}
-data = data[[1]]
-coords = coords[[1]]
+data = data[[2]]
+coords = coords[[2]]
 ```
@@
                           alpha = 0,
-                          standardize = FALSE,
-                          intercept = FALSE)
+                          standardize = FALSE)
                           #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
+getHyperPars(lrn_glmnet)
+getParamSet(lrn_glmnet)
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 77, upper = 7700))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 30, upper = 3300))
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+ctrl <- makeTuneControlRandom(maxit = 500)
+inner <- makeResampleDesc("SpCV", iters = 5)
+```
+
+#### TuneParams
+
+```{r}
+configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
+library(parallelMap)
+parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
+                                 par.set = ps_glmnet, control = ctrl, 
+                                 measure = list(rmse))
+parallelStop()
+saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_ridge.rda")
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_ridge.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.3))
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
+ctrl <- makeTuneControlRandom(maxit = 500)
 inner <- makeResampleDesc("SpCV", iters = 5)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 2)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
 ```
@@
 parallelStop()
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
+                                control = ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+              mc.set.seed = TRUE) # only parallelize the tuning
+
+set.seed(12345)
+resa_glmnet <- mlr::resample(tune_wrapper, task,
+                             resampling = outer, extract = getTuneResult,
+                             show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
 ```
@@
 For the elasticnet we also tune `alpha`.
-As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.
 
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7700),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.3),
                           makeNumericParam("alpha", lower = 0, upper = 1))
@@
 ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+inner <- makeResampleDesc("SpCV", iters = 5)
+```
+
+#### TuneParams
+
+```{r}
+configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
+library(parallelMap)
+parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
+                                 par.set = ps_glmnet, control = ctrl, 
+                                 measure = list(rmse))
+parallelStop()
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
 ```

### 2018-04-26 02_scripts/035_all-plots/03-modeling_all.Rmd
26/04/18

--- 2018-04-26 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
@@
 ```{r}
-data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
-coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
+data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf1.rda")
+coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf1-coords.rda")
 ```
 
-# Select plot
+# Merge into one dataset
+
+Leaving out Luiando (see README for reason).
 
 ```{r}
-data = data[[2]]
-coords = coords[[2]]
+data = as_tibble(rbindlist(data, fill = TRUE))
+data = Filter(function(x) !any(is.na(x)), data)
+coords = as_tibble(rbindlist(coords))
 ```
@@
 The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting (this) is applied.
+If `s` != `lambda`, linear interpolation or refitting is applied.
 
@@
 Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure. 
+They can be seen as a variable importance measure then?
 The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
@@
                              show.info = TRUE, measures = list(rmse))
+
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
 ```
@@
                                 intercept = FALSE)
-
+#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
 glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
@@
 
+#### Variable importance
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40,
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+feature_imp = generateFeatureImportanceData(task, 
+                                            method = "permutation.importance",
+                                            lrn_glmnet_tuned,
+                                            measure = list(rmse),
+                                            nmc = 40)
+parallelStop()
+
+saveRDS(feature_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf1-nmc40.rda")
+
+arrange(feature_imp$res)
+
+melt(feature_imp$res) %>% 
+  arrange(desc(value))
+```
+
 ### Lasso
@@
 ```{r}
+# predict.type for 'auc' (AUROC)
 lrn_glmnet <- makeLearner("regr.glmnet",
                           alpha = 1,
-                          standardize = FALSE,
-                          intercept = FALSE)
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
+                          standardize = FALSE)
+                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
+getHyperPars(lrn_glmnet)
+getParamSet(lrn_glmnet)
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "norm")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Elasticnet
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          standardize = FALSE,
-                          intercept = FALSE)
-#getHyperPars(lrn_glmnet)
-#getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-For the elasticnet we also tune `alpha`.
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.3),
-                          makeNumericParam("alpha", lower = 0, upper = 1))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 5)
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
+saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_lasso.rda")
 ```
@@
 ```
+
+#### Variable importance
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40,
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+feature_imp = generateFeatureImportanceData(task, 
+                                            method = "permutation.importance",
+                                            lrn_glmnet_tuned,
+                                            measure = list(rmse),
+                                            nmc = 40)
+parallelStop()
+
+saveRDS(feature_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf1-nmc40.rda")
+
+arrange(feature_imp$res)
+
+melt(feature_imp$res) %>% 
+  arrange(desc(value))
+```

### 2018-04-26 02_scripts/01_hyperspectral_indices/07_shiny_app/ui.R

@@ -193,7 +193,7 @@ body <- dashboardBody(
           br(),
           "<b>Correction:</b> Radiometric, geometric, atmospheric",
           br(),
-          "<b>Aquisition date:</b> 29 - 30 September 2016"
+          "<b>Acquisition date:</b> 29 - 30 September 2016"
         )), color = "black",
         value = "Metadata information", icon = icon("calendar")
       )

### 2018-05-07 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
07/05/18

@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge_bf2_10folds.rda")
 ```
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 parallelStop()
-saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_ridge.rda")
+saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge.rda")
 ```
@@
                                 alpha = 0,
-                                s = 1810,
+                                s = 7500,
                                 standardize = FALSE, 
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610)))) %>% 
   arrange(desc(coef))
 ```
+
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610)))) %>% 
   arrange(coef)
@@
 ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 5)
+inner <- makeResampleDesc("SpCV", iters = 10)
 ```
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 2)
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
 ```

### 2018-05-07 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
07/05/18

--- 2018-04-26 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
@@
 ```{r}
-task = makeRegrTask(id = "laukiz1", data = data, 
+task = makeRegrTask(id = "laukiz2", data = data, 
                     target = "deftot", coordinates = coords)
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 77, upper = 7700))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 102, upper = 10228))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
 ```
@@
                                  par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
+                                 show.info = TRUE, measure = list(rmse))
 parallelStop()
-saveRDS("/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf1_ridge.rda")
+saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10.2))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7700),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10228),
                           makeNumericParam("alpha", lower = 0, upper = 1))
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
 ```

### 2018-05-07 02_scripts/033-luiando/03-modeling_luiando.Rmd
26/04/18
07/05/18

--- 2018-04-26 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
@@
 ```{r}
-data = data[[1]]
-coords = coords[[1]]
+data = data[[3]]
+coords = coords[[3]]
 ```
@@
 ```{r}
-task = makeRegrTask(id = "laukiz1", data = data, 
+task = makeRegrTask(id = "luiando", data = data, 
                     target = "deftot", coordinates = coords)
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 77, upper = 7700))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 15, upper = 1560))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.56))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_lasso_bf2_10folds.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7700),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1560),
                           makeNumericParam("alpha", lower = 0, upper = 1))
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_elasticnet.rda")
 ```

### 2018-05-07 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
07/05/18

--- 2018-05-07 02_scripts/033-luiando/03-modeling_luiando.Rmd
@@
 Oiartzun: 7891
-Luiando: 6073
+oiartzun: 6073
 
@@
 ```{r}
-data = data[[3]]
-coords = coords[[3]]
+data = data[[4]]
+coords = coords[[4]]
 ```
@@
 ```{r}
-task = makeRegrTask(id = "luiando", data = data, 
+task = makeRegrTask(id = "oiartzun", data = data, 
                     target = "deftot", coordinates = coords)
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 15, upper = 1560))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 55, upper = 5525))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/oiartzun_ridge_bf2_10folds.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.56))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.05, upper = 5.5))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_lasso_bf2_10folds.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/oiartzun_lasso_bf2_10folds.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1560),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.05, upper = 5525),
                           makeNumericParam("alpha", lower = 0, upper = 1))
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_elasticnet.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/oiartzun_elasticnet.rda")
 ```

### 2018-05-07 02_scripts/035_all-plots/03-modeling_all.Rmd
07/05/18

--- 2018-04-26 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
@@
 ```{r, results='hide'}
+devtools::install_github("tidyverse/ggplot2")
+devtools::install_github("pat-s/mlr@blocking_visualization")
+
 library(mlr)
@@
 Laukiz2: 7594
-Oiartzun: 7891
-Luiando: 6073
+Oiartzun: 7841
+Luiando: 7471
 
@@
 
-# Select plot
+# Merge into one dataset
+
+Leaving out Luiando (see README for reason).
 
 ```{r}
-data = data[[1]]
-coords = coords[[1]]
+data = as_tibble(rbindlist(data, fill = TRUE))
+data = Filter(function(x) !any(is.na(x)), data)
+coords = as_tibble(rbindlist(coords))
 ```
@@
 ```{r}
-task = makeRegrTask(id = "laukiz1", data = data, 
-                    target = "deftot", coordinates = coords)
+task = makeRegrTask(id = "all_plots", data = data, 
+                    target = "deftot", coordinates = coords,
+                    blocking = factor(rep(1:4, c(559, 382, 301, 499))))
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 77, upper = 7700))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 35, upper = 3559))
 ```
@@
 ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+inner <- makeResampleDesc("CV", iters = 3)
 ```
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
+outer <- makeResampleDesc("RepCV", folds = 4, reps = 2)
+outer <- makeResampleDesc("Blocking")
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds.rda")
+```
+
+#### Visualize partitions
+
+```{r}
+resa_glmnet= readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds.rda")
+plot = createSpatialResamplingPlots(task, list("Blocking" = resa_glmnet), crs = 32630)
+cowplot::plot_grid(plotlist = plot[["Plots"]], ncol = 4, nrow = 1, labels = plot[["Labels"]])
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.5))
 ```
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 2)
+outer <- makeResampleDesc("Blocking")
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7700),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3559),
                           makeNumericParam("alpha", lower = 0, upper = 1))
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
+outer <- makeResampleDesc("Blocking")
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_elasticnet.rda")
 ```

### 2018-05-09 01_writing/Schratz_hyperspectral.tex
09/05/18

@@
 	\acro{MEM}{Maximum Entropy Model}
-	\acro{NBI}{Narrow Band Index}
+	\acro{NRI}{Normalized Ratio Index}
 	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
@@
 
+\subsection{Ground data}
+% Describe Plots
+
+The four \textit{Pinus radiata} plots Laukiz 1, Laukiz 2, Luiando and Oiartzun are located in the northern part of the Basque Country (\autoref{fig:study_area}).
+Laukiz 1 has the most trees (n = 559) while Laukiz 2 has largest area size.
+All plots besides Luiando are located nearby the coast.
+The data was collected in September 2016.
+
+% Show Map
+
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/data/study_area_hyperspectral.pdf}}
+		\caption{Information about the plot locations, the area of hyperspectral coverage and the number of trees per plot.}
+		\label{fig:study_area}
+	\end{center}
+\end{figure}
+
+% describe the hyperspectral data
+
 \cite{Brenning2012}
@@
 
-The airborne hyperspectral data was acquired on September 28th and October 5th 2016 around 12 am.
+The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around 12 am.
 The images were taken by an AISAEAGLE-II sensor from the \ac{ICGC}.
@@
 
-\subsection{Index calculation}
+\subsection{Derivation of indices}
 
@@
 All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that are offered by the \texttt{hsdar} package have been calculated.
-Additionally, all possible \ac{NBI} were calculated from the data using the following formula:
+Additionally, all possible \ac{NRI} were calculated from the data using the formula:
 
 \begin{equation}
-	NBI_{i,j} = \frac{B_{i} - B_{j}}{B_{i} + B_{j}}
+	NRI_{i,j} = \frac{B_{i} - B_{j}}{B_{i} + B_{j}}
 \end{equation}
@@
 
-To account for geometric offsets, we calculated every index five times using a buffer from 1 - 5 meter around the centroid of the respective tree.
+%To account for geometric offsets, we calculated every index five times using a buffer from 1 - 5 meter around the centroid of the respective tree.
+To account for geometric offsets, we used a buffer of 2 meters around the centroid of the respective tree.
 The mean value of all pixels touched by the buffer was assigned as the final value for each index.
+Missing values were removed from the mean value calculation.
+In total, 7875 \ac{NRI}s have been calculated ($\frac{125*126}{2}$).
+Some indices returned \texttt{NA} values for some observations and were removed from the dataset, leaving a total of 7471 indices that were available for all plots without missing values.
+Note that due to the mass of variables we cannot state which indices in detail have been removed.
+
+\subsection{Penalized regression}
+
+Penalized regression was used to account for the large amount of highly correlated predictor variables.
+The aim was to find the indices that best explain defoliation within the plots.
+
+Due to the amount of highly correlated predictors in the dataset, the independence assumption of the predictors is violated.
+To compensate for that, penalized regression penalizes the coefficients.
+This leads to a substantial decrease in variance and better predictive performance.
+However, it also results in biased coefficients that cannot be used for statistical inference anymore.
+The resulting coefficients can be seen as a measure of variable importance.
+The penalization terms that have been used in this work are explained in the following sections.
+
+\subsubsection{L1 Penalization}
+
+\subsubsection{L2 Penalization}
+
+\subsubsection{Elasticnet}
+
+\subsection{Modeling}
+
+The first step was to decide which penalization term works best for the given data.
+A nested 10-fold spatial \ac{CV} was conducted on every single plot and for the merged dataset, we ran a plot-based \ac{CV} for both performance evaluation and hyperparameter tuning.
+% show plot of spatial partitioning for both?
+This means that three out of the 4 plots formed the training set while the test set consisted of the remaining plot.
+
+Next, we applied the winning method (Ridge) again on the plots and the merged dataset to find the most important coefficients for every plot.
+Hyperparameter tuning of
+
+% Points to mention
+% * All penalization methods are suited for different kinds of dataset structures
+% * Find out which one works best on average and then use this method to create create models for prediction and variable importance
+
+\section{Discussion}
+
+\subsection{Index derivation}
+
 The exact number of contributing pixels of an index cannot be determined as it depends on the location of the tree within the pixel grid.
@@
 Also, if a tree is located at the border of the image data, some directions of the buffer may not contain values.
-Missing values were removed from the mean value calculation.
-However, the bigger the buffer size, the more pixels are represented by the index value.
-In total, 7875 possible \ac{NBI}s have been calculated for each buffer size summing up to 39375 indices.
-The same applies to the vegetation indices which sum up to 450 indices leading to a total number of 39825 indices per tree.
-
-\subsection{Ridge regression}
 

### 2018-05-09 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
09/05/18

@@
 coords = coords[[1]]
+```
+
+Subset to indices shared by all plots.
+
+```{r}
+indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda")
+
+data %<>% 
+  select(indices_shared)
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.7))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10.3))
 ```

### 2018-05-09 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
09/05/18

@@
 coords = coords[[2]]
+```
+
+Subset to indices shared by all plots.
+
+```{r}
+indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda")
+
+data %<>% 
+  select(indices_shared)
 ```
@@
                                 alpha = 0,
-                                s = 1810,
+                                s = 7690,
                                 standardize = FALSE, 
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690)))) %>% 
   arrange(desc(coef))
 ```
+
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690)))) %>% 
   arrange(coef)

### 2018-05-09 02_scripts/033-luiando/03-modeling_luiando.Rmd
09/05/18

@@
 coords = coords[[3]]
+```
+
+Subset to indices shared by all plots.
+
+```{r}
+indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda")
+
+data %<>% 
+  select(indices_shared)
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 15, upper = 1560))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 15, upper = 1558))
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_ridge_bf2_10folds.rda")
 ```
@@
                                 alpha = 0,
-                                s = 1810,
+                                s = 1560,
                                 standardize = FALSE, 
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560)))) %>% 
   arrange(desc(coef))
 ```
+
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560)))) %>% 
   arrange(coef)

### 2018-05-09 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
09/05/18

@@
 coords = coords[[4]]
+```
+
+Subset to indices shared by all plots.
+
+```{r}
+indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda")
+
+data %<>% 
+  select(indices_shared)
 ```
@@
                                 alpha = 0,
-                                s = 1810,
+                                s = 5460,
                                 standardize = FALSE, 
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460)))) %>% 
   arrange(desc(coef))
 ```
+
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460)))) %>% 
   arrange(coef)

### 2018-05-09 02_scripts/035_all-plots/03-modeling_all.Rmd
09/05/18

@@
 devtools::install_github("tidyverse/ggplot2")
-devtools::install_github("pat-s/mlr@blocking_visualization")
+devtools::install_github("mlr-org/mlr@blocking")
 
@@
 coords = as_tibble(rbindlist(coords))
+```
+
+Save the indices shared by all plots so that we can subset the single plots to use the same indices
+
+```{r}
+indices_shared = names(data)
+saveRDS(indices_shared, "/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda")
 ```
@@
 ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("CV", iters = 3)
+inner <- makeResampleDesc("Blocking")
 ```
@@
 ```{r}
-outer <- makeResampleDesc("RepCV", folds = 4, reps = 2)
+# outer <- makeResampleDesc("RepCV", folds = 4, reps = 2)
 outer <- makeResampleDesc("Blocking")
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds_blocking.rda")
 ```
@@
 parallelStop()
-saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge.rda")
+saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge_blocking.rda")
 ```
@@
                                 alpha = 0,
-                                s = 1810,
+                                s = 875,
                                 standardize = FALSE, 
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 875))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 875)))) %>% 
   arrange(desc(coef))
 ```
+
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 875))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 875)))) %>% 
   arrange(coef)
@@
 inner <- makeResampleDesc("SpCV", iters = 10)
+inner <- makeResampleDesc("Blocking")
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds_blocking.rda")
 ```
@@
 ```
+
 ```{r}
@@
 ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+inner <- makeResampleDesc("Blocking")
 ```
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_elasticnet.rda")
+saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_elasticnet_blocking.rda")
 ```

### 2018-05-11 01_writing/Schratz_hyperspectral.tex
11/05/2018

@@
 
-\journal{Journal "Remote Sensing of environment"}
+\journal{Journal "Remote Sensing of Environment"}
 
@@
 	\acro{NRI}{Normalized Ratio Index}
+	\acro{OLS}{Ordinary Least Squares}
 	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
@@
 	\acro{RF}{Random Forest}
+	\acro{RMSE}{Root Mean Square Error}
+	\acro{RR}{Ridge Regression}
+	\acro{RSS}{Residual Sum of Squares}
 	\acro{SDM}{Species Distribution Modeling}
@@
 
-Penalized regression was used to account for the large amount of highly correlated predictor variables.
-The aim was to find the indices that best explain defoliation within the plots.
+The aim of this work was to find the indices that best explain defoliation within the plots.
+We used penalized regression to account for the large amount of highly correlated predictor variables.
+In a standard \ac{OLS} regression one of the assumptions is that the predictor variables should be independent when minimizing the \ac{RSS} \citep{Bare1981, Hastie2001}:
 
-Due to the amount of highly correlated predictors in the dataset, the independence assumption of the predictors is violated.
-To compensate for that, penalized regression penalizes the coefficients.
-This leads to a substantial decrease in variance and better predictive performance.
-However, it also results in biased coefficients that cannot be used for statistical inference anymore.
-The resulting coefficients can be seen as a measure of variable importance.
-The penalization terms that have been used in this work are explained in the following sections.
+\begin{equation}
+	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2}
+\end{equation}
 
-\subsubsection{L1 Penalization}
+where $\beta_{0}$ is the intercept, $\beta_{i}$ the coefficient, $x_{ij}$ the predictor variable and $y_{i}$ the response variable.
 
-\subsubsection{L2 Penalization}
+If this assumption is violated, regression coefficients can be highly biased.
+They can even show the wrong sign and are very sensitive to adding new independent variables or data points to the model.
+These points reduce the robustness and performance of \ac{OLS} regression when dealing with multicollinearity.
+One approach to overcome these limitations is to penalize the coefficients.
+This method leads to a substantial decrease in variance and better predictive performance compared to \ac{OLS} regression.
+However, it also sacrifies the assumption of unbiased coefficients.
+Hence, the resulting coefficients cannot be used for statistical inference but should be interpreted as a measure of variable importance.
 
-\subsubsection{Elasticnet}
+\subsubsection{The ridge penalty}
+
+In \ac{RR} (also called $\ell_{2}$ penalization) the assumption of unbiased coefficients is given up in favor of higher predictive accuracy and reduced variance \citep{Hastie2001}.
+Coefficients are standardized and penalized for their size.
+When minimizing the \ac{RSS}, \ac{RR} adds a penalization term $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ to the equation:
+
+\begin{equation}
+	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p}\beta_{j}^{2}
+\end{equation}
+
+where $\lambda >= 0$ is a tuning parameter responsible for the magnitude of penalization.
+To make the second term (usually referred to as the \textit{shrinkage penalty}) of this equation small, the coefficients $\beta_{j}$ need to become small.
+Unlike to Lasso however, predictors are not removed from the final model and will always be $\beta_{j} >= 0$.
+Hence, $\lambda$ has the effect of shrinking the coefficients when minimizing the \ac{RSS}.
+For $\lambda = 0$, no penalization is done and standard \ac{OLS} applies \citep{James2013}.
+The \textit{shrinkage penalty} is only applied to the coefficients and not to the intercept.
+Also, while \ac{OLS} generates only one set of coefficient estimates, \ac{RR} will create multiple sets for every value of $\lambda$.
+In summary, the advantage of \ac{RR} is based on the bias-variance tradeoff: For $\lambda\to\infty$, the flexibility of the fit is reduced leading to a decrease in variance of the coefficients but also introduces a (substantial) bias.
+For $\lambda = 0$, the variance is high but coefficients are unbiased \citep{James2013}.
+
+\subsubsection{The lasso penalty}
+
+While the Ridge penalty $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ shrinks the coefficients, it does not exclude any predictor from the model by setting the coefficient to zero.
+This is what the Lasso penalty (also called $\ell_{1}$) does: It works similar to a "best subset selection" approach by only keeping the most important variables in the model and reducing the coefficients of unimportant ones down to zero.
+This is done by a different \textit{shrinkage penalty} $\lambda \sum_{j=1}^{p} \mid\beta_{j}\mid$ that is added to the \ac{OLS} term when minimizing the \ac{RSS}:
+
+\begin{equation}
+	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p} \mid\beta_{j}\mid
+\end{equation}
+
+Subsequently, Lasso is performing variable selection and results in a sparse model that has the advantage of a simplied interpretability compared to Ridge models.
+However, coeffcients are still biased due to penalization \citep{Hastie2001, James2013}.
+
+\subsubsection{The elasticnet penalty}
+
+The idea of the \textit{elasticnet penalty} is to combine both penalties Lasso and Ridge and weight these by a parameter $\alpha$ that that needs to be optimized.
+It was introduced by \cite{Zou2005} with the idea to combine the advantages of both penalty terms.
+The \textit{elasticnet penalty} is written as
+
+\begin{equation}
+	\lambda \sum_{j=1}^{p} \alpha\beta_{j}^{2} + (1 - \alpha) \mid\beta_{j}\mid
+\end{equation}
+
+where $\alpha$ weights the contribution of either the Lasso or Ridge penalty.
+For $\alpha = 0.5$ both are weighted equally.
+When $\alpha = 0$ only Ridge is used and for $\alpha = 1$ only Lasso applies \citep{Hastie2001}.
 
@@
 
-The first step was to decide which penalization term works best for the given data.
-A nested 10-fold spatial \ac{CV} was conducted on every single plot and for the merged dataset, we ran a plot-based \ac{CV} for both performance evaluation and hyperparameter tuning.
-% show plot of spatial partitioning for both?
-This means that three out of the 4 plots formed the training set while the test set consisted of the remaining plot.
+\subsubsection{Selecting the best penalty}
 
-Next, we applied the winning method (Ridge) again on the plots and the merged dataset to find the most important coefficients for every plot.
-Hyperparameter tuning of
+As the introduced penalties behave different depending on the characteristics of the dataset, we first conducted a nested 10-fold spatial \ac{CV} to find the best method among those three.
+The comparison was applied to all of the four plots using \ac{RMSE} as the error measure.
+For both the performance evaluation and hyperparameter tuning a spatial sampling using $k$-means clustering was used \citep{sperrorest}.
+
+\subsubsection{Extracting the most important variables}
+
+Next, we applied the winning method (\ac{RR}) on all plots and additionally to the merged dataset of all plots to find the most important coefficients of every setup.
+Hyperparameter tuning of $\lambda$ was done using again a spatial sampling again for the single plots.
+For the merged dataset we used a plot-based sampling approach: One plot serves as the test partition while the remaining plots form the training set.
+This approach has the advantage of a spatially independent test partition but also comes with the limitation of estimating the performance on a single repetition only as the restriction to the plot-level leaves no opportunity to create multiple resampling instances.
+
+\subsubsection{Linking the variables to plot characteristics}
+
+To interpret the outcomes of the models on a plot level, we linked the winning variables of each to attributes of the dataset that describe the underlying plot characteristics.
+For example, a plot which a low tree density might inherit more information from the bare ground in the calculdated indices while a plot with a very high tree density might in contrast contain information from multiple trees.
 
@@
 % * Find out which one works best on average and then use this method to create create models for prediction and variable importance
+
+\section{Results}
+
+\subsection{Comparison of penalties}
+
+% parameter limits
+\begin{table}[b!]
+\centering
+\caption[t]{Predictive performances of lasso, ridge and elasticnet penalties on the single plots and the merged dataset using \ac{RMSE} as the error measure.}
+\begingroup\footnotesize
+\begin{tabular}{llll}
+	\\
+	Plot/Penalty & Lasso  & Ridge  & Elasticnet \\
+	\hline
+	Laukiz 1     & 146.33 & 87.76  & 135.58     \\
+	Laukiz 2     & 146.33 & 87.76  & 135.58     \\
+	Luiando      &        &        &            \\
+	Oiartzun     &        & 106.65 &            \\
+	All Plots    & 56.88  & 55.88  & 56.76      \\
+	\bottomrule
+\end{tabular}
+\endgroup
+\label{tab:penalty_comparison}
+\end{table}
 

### 2018-05-11 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
11/05/2018

@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.05, upper = 5.5))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.5))
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.05, upper = 5525),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 5525),
                           makeNumericParam("alpha", lower = 0, upper = 1))

### 2018-05-14 01_writing/Schratz_hyperspectral.tex
14/05/2018

@@
 
-The idea of the \textit{elasticnet penalty} is to combine both penalties Lasso and Ridge and weight these by a parameter $\alpha$ that that needs to be optimized.
+The \textit{elasticnet penalty} combines both penalties Lasso and Ridge by weighting them with a parameter $\alpha$ that needs to be optimized.
 It was introduced by \cite{Zou2005} with the idea to combine the advantages of both penalty terms.
@@
 As the introduced penalties behave different depending on the characteristics of the dataset, we first conducted a nested 10-fold spatial \ac{CV} to find the best method among those three.
-The comparison was applied to all of the four plots using \ac{RMSE} as the error measure.
+The model comparison was applied to all of the four plots using \ac{RMSE} as the error measure.
 For both the performance evaluation and hyperparameter tuning a spatial sampling using $k$-means clustering was used \citep{sperrorest}.
+For the tuning, 200 random search iterations were used.
+The limits of the tuning space for $\lambda$ were set to the internally calculated (data driven) limits of the R package \texttt{glmnet}.
 
-\subsubsection{Extracting the most important variables}
+\subsubsection{Extraction of the most important variables}
 
-Next, we applied the winning method (\ac{RR}) on all plots and additionally to the merged dataset of all plots to find the most important coefficients of every setup.
-Hyperparameter tuning of $\lambda$ was done using again a spatial sampling again for the single plots.
-For the merged dataset we used a plot-based sampling approach: One plot serves as the test partition while the remaining plots form the training set.
-This approach has the advantage of a spatially independent test partition but also comes with the limitation of estimating the performance on a single repetition only as the restriction to the plot-level leaves no opportunity to create multiple resampling instances.
+Next, we applied the winning method (\ac{RR}) on all plots.
+Hyperparameter tuning of $\lambda$ for the full dataset was again done using a spatial sampling.
+The ten highest coefficients (both positive and negative) were extracted and reported.
 
-\subsubsection{Linking the variables to plot characteristics}
+\subsubsection{Linking variables to plot characteristics}
 
-To interpret the outcomes of the models on a plot level, we linked the winning variables of each to attributes of the dataset that describe the underlying plot characteristics.
-For example, a plot which a low tree density might inherit more information from the bare ground in the calculdated indices while a plot with a very high tree density might in contrast contain information from multiple trees.
+To interpret the outcomes of the models on a plot level, we linked the winning variables of each to dataset attributes that describe the underlying plot characteristics.
+For example, a plot which a low tree density might inherit more information from the bare ground in the calculated indices while a plot with a very high tree density might in contrast contain information from multiple trees.
+Also, the overall level of defoliation for the whole plot might possibly have an effect how well an index is able to describe the situation.
+
+\subsubsection{Creation of a super model with all observations}
+
+After the plot level analysis, we merged all observations into a single dataset and fitted another \ac{RR} model.
+We used a block-level based spatial \ac{CV} for this setup.
+For a total number of $m$ folds ($n$ = number of plots), every plot is once the test set while the training is done on the remaining ones.
+This ensures that the test set is fully spatially independent from the training set.
+Tuning was also performned on the block level within the training set on $n - 1$ folds using 200 random search iterations.
+Due to the fixing of the indices on a plot level, varying those to use multiple repetitions is not possible.
 
@@
 
-\subsection{Comparison of penalties}
+\subsection{Predictive performances}
 
@@
 	\hline
-	Laukiz 1     & 146.33 & 87.76  & 135.58     \\
-	Laukiz 2     & 146.33 & 87.76  & 135.58     \\
-	Luiando      &        &        &            \\
-	Oiartzun     &        & 106.65 &            \\
+	Laukiz 1     & 146.33 & 89.89  & 135.58     \\
+	Laukiz 2     & 172.70 & 30.37  & 37.07      \\
+	Luiando      & 74.85  & 76.02  & 74.77      \\
+	Oiartzun     & 327.93 & 106.65 & 260.38     \\
 	All Plots    & 56.88  & 55.88  & 56.76      \\
@@
 \label{tab:penalty_comparison}
+\end{table}
+
+\begin{table}[b!]
+\centering
+\caption[t]{Predictive performance of \ac{RR} using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.}
+\begingroup\footnotesize
+\begin{tabular}{lll}
+	\\
+	Plot     & Supermodel (Block CV) & Single plot (SpCV) \\
+	\hline
+	Laukiz 1 & 58.95                 & 89.89              \\
+	Laukiz 2 & 27.94                 & 30.37              \\
+	Luiando  & 69.72                 & 76.02              \\
+	Oiartzun & 58.09                 & 106.65             \\
+	\bottomrule
+\end{tabular}
+\endgroup
+\label{tab:supermodel_performance}
 \end{table}
@@
 
+\subsection{Variable importance}
+
+% variable importance
+\begin{table}[b!]
+\centering
+\caption[t]{The ten highest coefficient estimates for every plot and the merged dataset.}
+\begingroup\footnotesize
+\begin{tabular}{lllll}
+	\\
+	Laukiz 1           & Laukiz 2           & Luiando             & Oiartzun            & All Plots          \\
+	\hline
+	b80-b77 (0.0089)   & b89-b84 (0.0093)   & b95-b93 (1.5e-36)   & b23-b18 (-0.0090)   & b78-b77 (0.058)    \\
+	b81-b77 (0.0079)   & b89-b87 (0.0086)   & b109-b106 (1.4e-36) & b23-b19 (-0.0074)   & b115-b113 (0.054)  \\
+	b78-b77 (0.0079)   & b89-b88 (0.0086)   & b92-b95 (1.4e-36)   & b99-b98 (0.0073)    & b82-b77 (0.052)    \\
+	b77-b76 (-0.0078)  & b108-b104 (0.0084) & b114-b6 (-1.2e-36)  & b23-b20 (-0.0070)   & b79-b77 (0.049)    \\
+	b79-b77 (0.0077)   & b89-b85 (0.0083)   & b96-b93 (1.2e-36)   & b10-b8 (-0.0063)    & b80-b77 (0.049)    \\
+	Datt3 (-0.71)      & b92-b84 (0.0083)   & b116-b6 (-1.2e-36)  & b102-b98 (0.0062)   & b81-b77 (0.049)    \\
+	b41-b25 (0.0070)   & b89-b83 (0.0081)   & b114-b5 (-1.2e-36)  & b124-b115 (0.0062)  & b81-b78 (-0.048)   \\
+	b116-b113 (0.0068) & b89-b86 (0.0080)   & b115-b114 (1.2e-36) & b23-b15 (-0.0060)   & b124-b115 (-0.047) \\
+	b82-b77 (0.0068)   & b92-b88 (0.0080)   & b95-b91 (1.1e-36)   & b126-b115 (-0.0060) & b23-b20 (-0.047)   \\
+	b77-b75 (-0.0068)  & b108-b96 (0.0079)  & b114-b8 (-1.1e-36)  & b118-b115 (-0.0059) & b80-b78 (-0.046)   \\
+	\bottomrule
+\end{tabular}
+\endgroup
+\label{tab:variable-importance}
+\end{table}
+
 \section*{References}

### 2018-05-14 02_scripts/022_preprocessing.Rmd
14/05/2018

@@
 #       )
-# saveRDS(data, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf1.rda")
+# saveRDS(data, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
 

### 2018-05-14 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
14/05/2018

@@
 data %<>% 
-  select(indices_shared)
+  dplyr::select(indices_shared)
 ```
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge_bf2_10folds.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge_bf2_10folds.rda")
 ```
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7500))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7500)))) %>% 
   arrange(desc(coef))
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7610)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7500))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7500)))) %>% 
   arrange(coef)
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
 ```

### 2018-05-14 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
14/05/2018

@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
 ```

### 2018-05-14 02_scripts/035_all-plots/03-modeling_all.Rmd
14/05/2018

@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds_blocking.rda")
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds_blocking.rda")
 ```

### 2018-05-14 02_scripts/04_plot_characteristics.Rmd
14/05/2018

---
title: "04-plot characteristics"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
devtools::install_github("tidyverse/ggplot2")
devtools::install_github("mlr-org/mlr@blocking")

library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
library(raster)
library(ggplot2)
library(hrbrthemes)
library(cowplot)
```

# Read data

Variables per plot:

Lauki1: 7594
Laukiz2: 7594
Oiartzun: 7841
Luiando: 7471

Why is there such a huge difference in the number of indices without NAs for the plots?

```{r}
data <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
```

# Merge into one dataset

Leaving out Luiando (see README for reason).

```{r}
data <- as_tibble(rbindlist(data, fill = TRUE)) # combine all plots
data <- Filter(function(x) !any(is.na(x)), data) # filter out columsn containing NA
coords <- as_tibble(rbindlist(coords))
data %<>%
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(559, 382, 301, 499))))
```

# Mean defoliation per plot

```{r}
data %>%
  group_by(plot) %>%
  summarise(mean(deftot))
```

```{r}
data %>%
  group_by(plot) %>%
  ggplot(aes(y = deftot, x = plot)) +
  geom_boxplot() +
  theme_ipsum_rc()
```

# Point density

In Meters.

```{r}
plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")

dist <- map(plots, ~ {
  coords_laukiz1 <- coords %>%
    mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(559, 382, 301, 499)))) %>%
    filter(plot == .x) %>%
    dplyr::select(-plot)

  laukiz1_points <- SpatialPoints(coords = coords_laukiz1, proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

  distance <- mean(as.dist(pointDistance(laukiz1_points, allpairs = TRUE, lonlat = FALSE)))
})
set_names(dist, plots)
```

# Scatterplots

Visualize RMSE vs plot characteristics

```{r}
df <- tibble(
  plot = factor(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")),
  rmse = c(89.89, 30.37, 76.02, 106.65),
  point_dens = c(34.34, 62.51, 33.76, 35.01),
  mean_defol = c(53.22, 17.73, 68.35, 44.63)
) 
```

## RMSE vs point dens

```{r}
rmse_dens <- ggplot(df, aes(x = point_dens, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean point density (m)", y = "RMSE") +
  theme_ipsum_rc() +
  theme(legend.position = "none") +
  theme(plot.margin = unit(c(6,0,6,0), "pt"))
```

## RMSE vs mean defol

```{r}
rmse_defol <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean defoliation (%)", y = "RMSE") +
  theme_ipsum_rc() +
  theme(legend.position = "none") +
  theme(plot.margin = unit(c(6,0,6,0), "pt"))
```

## Plot

### Legend

```{r}
legend <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean defoliation (%)", y = "RMSE", color = "Plot") +
  theme_ipsum_rc()
legend <- get_legend(legend)
```

### Grid

```{r}
plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
```

### 2018-05-15 01_writing/Schratz_hyperspectral.tex
15/05/2018

@@
 % link to PDF with veg indeces
-All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that are offered by the \texttt{hsdar} package have been calculated.
+All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data and offered by the \texttt{hsdar} package have been calculated.
 Additionally, all possible \ac{NRI} were calculated from the data using the formula:
@@
 \begin{equation}
-	NRI_{i,j} = \frac{B_{i} - B_{j}}{B_{i} + B_{j}}
+	NRI_{i,j} = \frac{b_{i} - b_{j}}{b_{i} + b_{j}}
 \end{equation}
@@
 
+% parameter limits
+\begin{table}[b!]
+\centering
+\caption[t]{10-fold 20-times repeated spatial \ac{CV} performances of lasso, ridge and elasticnet penalties on the plot level and the merged dataset using \ac{RMSE} as the error measure.
+Values show the overall mean and standard deviation at the repetition level.}
+\begingroup\footnotesize
+\begin{tabular}{llll}
+	\\
+	Plot/Penalty   & Lasso         & Ridge        & Elasticnet    \\
+	\hline
+	Laukiz 1       & 133.15 (4.66) & 86.71 (2.96) & 121.04 (4.15) \\
+	Laukiz 2       & 91.93 (6.78)  & 29.73 (0.38) & 47.18 (13.70) \\
+	Luiando        & 74.85         & 76.02        & 74.77         \\
+	Oiartzun       & 327.93        & 106.65       & 260.38        \\
+	Merged dataset & 56.88         & 55.88        & 56.76         \\
+	\bottomrule
+\end{tabular}
+\endgroup
+\label{tab:penalty_comparison}
+\end{table}
+
 \subsubsection{Linking variables to plot characteristics}
@@
 Due to the fixing of the indices on a plot level, varying those to use multiple repetitions is not possible.
+The idea behind fitting a supermodel is that this model includes information from all plots rather than just from a single plot.
+This may possibly reduce the predictive error and create a more robust model.
 
@@
 
-\subsection{Predictive performances}
+\subsection{Plot characteristics}
 
-% parameter limits
-\begin{table}[b!]
+Luiando shows the highest defoliation ($\bar{x} = 68.36 \%$) among the plots while Laukiz 2 is the healthiest ($\bar{x} = 17.73 \%$) (\autoref{fig:defol_boxplots}).
+Laukiz 1 and Oiartzun both show a medium defoliation level around 50 \%.
+Oiartzun consists mainly of trees that show either a very high level of defoliation ($70 \% <=$) or none at all.
+Laukiz 1 shows an evenly distributed level of defoliation across the entire plot.
+
+The high defoliation level of Luiando and Oiartzun is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
+Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
+Oiartzun is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
+Laukiz 2 shows a mean tree density of 62.51 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.35 (Laukiz 1), 33.77 (Luiando), 35.02 (Oiartzun)) (\autoref{fig:rmse_defol_dens}).
+
+% defoliation boxplots
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {../03_figures/results/boxplot_defol.pdf}}
+		\caption{Descriptive statistics of the response variable \textit{defoliation}.}
+		\label{fig:defol_boxplots}
+	\end{center}
+\end{figure}
+
+% spectral signatures
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/spectral_signatures.pdf}}
+		\caption{Spectral signatures (mean and standard deviation) of each plot.
+			The colored lines show the most important band for each plot, respectively: Band 80 (773nm, red), band 89 (817nm, blue), band 95 (846nm, orange), band 23 (503nm, green).}
+		\label{fig:spectral_signatures}
+	\end{center}
+\end{figure}
+
+\subsection{Predictive performance}
+
+\begin{table}[t!]
 \centering
-\caption[t]{Predictive performances of lasso, ridge and elasticnet penalties on the single plots and the merged dataset using \ac{RMSE} as the error measure.}
-\begingroup\footnotesize
-\begin{tabular}{llll}
-	\\
-	Plot/Penalty & Lasso  & Ridge  & Elasticnet \\
-	\hline
-	Laukiz 1     & 146.33 & 89.89  & 135.58     \\
-	Laukiz 2     & 172.70 & 30.37  & 37.07      \\
-	Luiando      & 74.85  & 76.02  & 74.77      \\
-	Oiartzun     & 327.93 & 106.65 & 260.38     \\
-	All Plots    & 56.88  & 55.88  & 56.76      \\
-	\bottomrule
-\end{tabular}
-\endgroup
-\label{tab:penalty_comparison}
-\end{table}
-
-\begin{table}[b!]
-\centering
-\caption[t]{Predictive performance of \ac{RR} using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.}
+\caption[t]{Predictive performance of \ac{RR} using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.
+The values for "merged dataset" correspond to the fold for which the respective plot was serving as the test set.
+For "single plot", the values correspond to the mean value of the SpCV at the repetition level (10 folds, 20 repetitions).}
 \begingroup\footnotesize
@@
 	\\
-	Plot     & Supermodel (Block CV) & Single plot (SpCV) \\
+	Plot/Data & Merged dataset (Block CV) & Single plot (SpCV) \\
 	\hline
@@
 \end{table}
+
+\ac{RR} shows the lowest error for three out of four plots (for Luiando \textit{elasticnet} shows a slightly better performance) (\autoref{tab:penalty_comparison}).
+The magnitude of difference for \ac{RR} compared to the other penalties for the plots in which \ac{RR} showed the best performance ranges between XX and XX percent.
+For the merged dataset, all penalties show a similar mean predictive performance that outperform all single plot models besides the Laukiz 2 model.
+
+When comparing the mean predictive performance of the plot level model against the performance of the super model at the plot level (when the respective plot served as the test set), the supermodel also outperforms the Laukiz 2 model (27.94 vs 30.37 RMSE) (\autoref{tab:supermodel_performance}).
+
+The worst performance of the supermodel on the fold level is reported for Luiando (69.72 RMSE) while for the single plot models Oiartzun shows the highest error (106.65 RMSE).
+
+% rmse vs defol and point dens
+\begin{figure} [b!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/mean-defol_mean-point-dens.pdf}}
+		\caption{RMSE vs. mean defoliation and point density}
+		\label{fig:rmse_defol_dens}
+	\end{center}
+\end{figure}
 

### 2018-05-15 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
15/05/2018

@@
 
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_ridge_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
+```
+
 #### TuneParams
@@
 
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
+```
+
 #### TuneParams
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
+```
+
+Calculate repetition mean
+
+```{r}
 resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_elasticnet_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
 ```

### 2018-05-15 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
15/05/2018

@@
 
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
+```
+
 #### TuneParams
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
+```
+
+Calculate repetition mean
+
+```{r}
 resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
+```
+
+Calculate repetition mean
+
+```{r}
 resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
 ```

### 2018-05-15 02_scripts/04_plot_characteristics.Rmd
15/05/2018

@@
 library(cowplot)
+library(here)
+library(ggpubr)
+library(ggsci)
+library(ggrepel)
 ```
@@
 ```{r}
-data %>%
+boxplot_defol = data %>%
   group_by(plot) %>%
-  ggplot(aes(y = deftot, x = plot)) +
-  geom_boxplot() +
-  theme_ipsum_rc()
+  ggboxplot(
+    x = "plot", y = "deftot", color = "plot",
+    add = "jitter", add.params = list(size = "deftot")
+  ) +
+  scale_size(range = c(0.5, 0.5)) +
+  annotate("text",
+    label = expression(bold(atop("n = 559", bar(x)~"= 53.23"))), x = 1, 
+    y = 112, size = 3, colour = "#BC3C29", fontface = 2
+  ) +
+  annotate("text",
+    label = expression(bold(atop("n = 382", bar(x)~"= 17.73"))), x = 2, 
+    y = 112, size = 3, colour = "#0072B5", fontface = 2
+  ) +
+  annotate("text",
+    label = expression(bold(atop("n = 301", bar(x)~"= 68.36"))), x = 3, 
+    y = 112, size = 3, colour = "#E18727", fontface = 2
+  ) +
+  annotate("text",
+    label = expression(bold(atop("n = 499", bar(x)~"= 44.64"))), x = 4, 
+    y = 112, size = 3, colour = "#20854E", fontface = 2
+  ) +
+  scale_color_nejm() +
+  theme(legend.position = "none") +
+  labs(y = "Total defoliation per tree (%)", x = "Plot")
+boxplot_defol
+```
+
+```{r}
+save_plot(here("03_figures/results/boxplot_defol.png"), boxplot_defol)
 ```
@@
 rmse_dens <- ggplot(df, aes(x = point_dens, y = rmse)) +
-  geom_point(aes(color = plot), shape = 19) +
+  geom_point(aes(color = plot), shape = 19, size = 3) +
   labs(x = "Mean point density (m)", y = "RMSE") +
-  theme_ipsum_rc() +
-  theme(legend.position = "none") +
-  theme(plot.margin = unit(c(6,0,6,0), "pt"))
+  theme_ipsum_rc(
+    axis_title_size = 22,
+    axis_title_just = "c"
+  ) +
+  scale_color_nejm() +
+  theme(
+    plot.margin = unit(c(6, 0, 6, 0), "pt"),
+    legend.position = "none",
+    axis.text = element_text(size = 20),
+    legend.text = element_text(size = 20),
+    axis.title.y = element_text(margin = margin(
+      t = 0, r = 15,
+      b = 0, l = 0
+    )),
+    axis.title.x = element_text(margin = margin(
+      t = 15, r = 0,
+      b = 0, l = 0
+    ))
+  )
 ```
@@
 rmse_defol <- ggplot(df, aes(x = mean_defol, y = rmse)) +
-  geom_point(aes(color = plot), shape = 19) +
+  geom_point(aes(color = plot), shape = 19, size = 3) +
   labs(x = "Mean defoliation (%)", y = "RMSE") +
-  theme_ipsum_rc() +
-  theme(legend.position = "none") +
-  theme(plot.margin = unit(c(6,0,6,0), "pt"))
+  theme_ipsum_rc(
+    axis_title_size = 22,
+    axis_title_just = "c"
+  ) +
+  scale_color_nejm() +
+  theme(
+    plot.margin = unit(c(6, 0, 6, 0), "pt"),
+    legend.position = "none",
+    axis.text = element_text(size = 20),
+    legend.text = element_text(size = 20),
+    axis.title.y = element_text(margin = margin(
+      t = 0, r = 15,
+      b = 0, l = 0
+    )),
+    axis.title.x = element_text(margin = margin(
+      t = 15, r = 0,
+      b = 0, l = 0
+    ))
+  )
 ```
@@
 legend <- ggplot(df, aes(x = mean_defol, y = rmse)) +
-  geom_point(aes(color = plot), shape = 19) +
+  geom_point(aes(color = plot), shape = 19, size = 3) +
   labs(x = "Mean defoliation (%)", y = "RMSE", color = "Plot") +
-  theme_ipsum_rc()
+  theme_ipsum_rc(
+    axis_title_size = 16,
+    axis_title_just = "c"
+  ) +
+  scale_color_nejm() +
+  theme(
+    axis.text = element_text(size = 20),
+    legend.text = element_text(size = 20),
+    legend.title = element_text(size = 20),
+    axis.title.y = element_text(margin = margin(
+      t = 0, r = 15,
+      b = 0, l = 0
+    ))
+  )
 legend <- get_legend(legend)
@@
 ```{r}
-plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
+plots_grid <- plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
 ```
+
+```{r}
+save_plot(here("03_figures/results/mean-defol_mean-point-dens.png"), plots_grid, ncol = 3)
+```
+
+# PNG to PDF conversion
+
+Converts all pngs in the specified folder to pdfs. 
+`&&` makes the "convert" call conditional on `cd`.
+`exec` reduces memory waste of shell.
+Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.
+
+```{r png to pdf, echo = FALSE, eval = FALSE}
+# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
+#        parallel convert '{}' '{.}.pdf'")
+
+system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
+       parallel convert '{}' '{.}.pdf'")
+
+# copy to presentation folder
+# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
+#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
+#           overwrite = TRUE)
+```
+
+# Session Info
+
+```{r}
+sessionInfo()
+```

### 2018-05-16 01_writing/Schratz_hyperspectral.tex
16/05/18

@@
 	Laukiz 2       & 91.93 (6.78)  & 29.73 (0.38)  & 47.18 (13.70) \\
-	Luiando        & 74.85         & 76.02        & 74.77         \\
+	Luiando        & 74.83 (0.311) & 75.95 (0.23)  & 74.74 (0.29)  \\
 	Oiartzun       & 327.93        & 106.65        & 260.38        \\
-	Merged dataset & 56.88         & 55.88        & 56.76         \\
+	Merged dataset & 54.80 (17.58) & 53.67 (17.95) & 54.66 (17.67) \\
 	\bottomrule
@@
 
+% single models vs supermodel
 When comparing the mean predictive performance of the plot level model against the performance of the super model at the plot level (when the respective plot served as the test set), the supermodel also outperforms the Laukiz 2 model (27.94 vs 30.37 RMSE) (\autoref{tab:supermodel_performance}).
@@
 The worst performance of the supermodel on the fold level is reported for Luiando (69.72 RMSE) while for the single plot models Oiartzun shows the highest error (106.65 RMSE).
+
+% RMSE vs defol and point dens
+Laukiz 2 showed contrary results compared to all other plots when linking \ac{RMSE} against mean defoliation and mean point density (\autoref{fig:rmse_defol_dens}).
+Laukiz 1, Luiando and Oiartzun show almost identical values in terms of point density (34.35m, 33.77m, 35.01m).
+Due to the high similarity, these should be interpreted as a group and not as single values.
 
@@
 
-\section{Discussion}
+\subsubsection{Variable importance}
 
-\subsection{Index derivation}
-
-The exact number of contributing pixels of an index cannot be determined as it depends on the location of the tree within the pixel grid.
-If a tree is located at the border of a pixel, the same buffer (e.g. 3 m) will include more pixels than if the point is located at the center of a pixel.
-Also, if a tree is located at the border of the image data, some directions of the buffer may not contain values.
-
-\subsection{Variable importance}
+\ac{NRI}s using bands in the wavelength range of 770 nm - 820 nm (band 80 - band 89), which belongs to the infrared region, appear most often among the ten highest coefficient estimates across all plots (\autoref{tab:variable-importance}).
+Only one vegeation index (Datt3) showed up among the most important predictors (Laukiz 1).
+Luiando and Oiartzun also prefered bands with longer (938.39 nm (band 114) - 996.31 nm (band 126)) and shorter wavelengths (480.30 nm (band 18) - 503.26 (band 23)).
+The first range again belongs to the infrared region while the second is within the region of the visible light, transitioning from blue to green.
 
@@
 
+\section{Discussion}
+
+\subsection{Index derivation}
+
+The exact number of contributing pixels to the final index value of an observations cannot be determined as it depends on the location of the tree within the pixel grid.
+If a tree is located at the border of a pixel, a buffer of e.g. three meters will include more pixels than if the point is located at the center of a pixel.
+Also, if a tree is located at the border of the plot, some directions of the buffer will not contain image values.
+
+\subsection{Plot characteristics}
+
+For Laukiz1, Luiando and Oiartzun \ac{RMSE} seems to increase with a higher point density on a first look.
+However, the point densities of these plots are very similar (33.7 m - 35.01 m) and should be interpreted as a group instead of single values.
+Comparing this group to Laukiz 2, which has point density of 62 m, the error decreases with a higher point density.
+Linking the error against the mean defoliation of a plot does not lead to a clear conclusion:
+While Laukiz1, Luiando and Oiartzun give the impression of a decreasing RMSE when defoliation increases, Laukiz 2 clearly shows that a plot with a low defoliation can results in a low RMSE estimate (\autoref{fig:rmse_defol_dens}).
+
+\subsection{Variable importance}
+
 \section*{References}

### 2018-05-16 02_scripts/033-luiando/03-modeling_luiando.Rmd
15/05/2018
16/05/18

@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_ridge_bf2_10folds.rda")
+```
+
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_ridge_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_lasso_bf2_10folds.rda")
+```
+
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_lasso_bf2_10folds.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
 ```
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_elasticnet.rda")
+```
+
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_elasticnet.rda")
+resa_glmnet$measures.test %>% 
+  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
+  group_by(rep) %>% 
+  summarise_at("rmse", mean) %>% 
+  summarise(mean(rmse), sd(rmse))
 ```

### 2018-05-16 02_scripts/035_all-plots/03-modeling_all.Rmd
16/05/18

--- 2018-05-09 02_scripts/035_all-plots/03-modeling_all.Rmd
@@
 
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_ridge_bf2_10folds_blocking.rda")
+resa_glmnet$measures.test %>% 
+  summarise(mean(rmse), sd(rmse))
+```
+
 #### Visualize partitions
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds_blocking.rda")
+```
+
+Calculate repetition mean
+
+```{r}
+resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds_blocking.rda")
+resa_glmnet$measures.test %>% 
+  summarise(mean(rmse), sd(rmse))
 ```

### 2018-05-28 02_scripts/04_plot_characteristics.Rmd
28/05/2018

@@
 
+# Coefficient of variation
+
+```{r}
+data %>%
+  group_by(plot) %>%
+  summarise((sd(deftot) / mean(deftot)) * 100)
+```
+
+#  sd / skewness
+
+```{r}
+data %>%
+  group_by(plot) %>%
+  summarise(((sd(deftot) / mean(deftot)) * 100) / skewness(deftot))
+```
+
+```{r}
+data %>%
+  group_by(plot) %>%
+  summarise(sd(deftot))
+```
+
 ```{r}
@@
   point_dens = c(34.34, 62.51, 33.76, 35.01),
-  mean_defol = c(53.22, 17.73, 68.35, 44.63)
+  mean_defol = c(53.22, 17.73, 68.35, 44.63),
+  cov_defol = c(44.88707	, 115.78180	, 12.17709	, 83.18840),
+  sd_skewness = c(-55.077546, 67.232146, -3.187733	, -401.564010)
 )
@@
 
-## RMSE vs mean defol
+## RMSE vs coef of variation
 
 ```{r}
-rmse_defol <- ggplot(df, aes(x = mean_defol, y = rmse)) +
+rmse_defol <- ggplot(df, aes(x = sd_skewness, y = rmse)) +
   geom_point(aes(color = plot), shape = 19, size = 3) +
-  labs(x = "Mean defoliation (%)", y = "RMSE") +
+  labs(x = "Coefficient of variation (%) / Skewness", y = "RMSE") +
   theme_ipsum_rc(

### 2018-07-11 01_writing/Schratz_hyperspectral.tex
11/07/2018

@@
 \usepackage{float}
+
+\graphicspath{{../03_figures/results/}{./}{../03_figures/data/}}
+
+% enable paragraph as 4th level https://tex.stackexchange.com/questions/60209/how-to-add-an-extra-level-of-sections-with-headings-below-subsubsection
+\usepackage{titlesec}
+\setcounter{secnumdepth}{4}
+\titleformat{\paragraph}
+{\normalfont\small\bfseries}{\theparagraph}{1em}{}
+\titlespacing*{\paragraph}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 
@@
 	\begin{keyword}
-		hyperspectral imagery \sep statistical learning \sep spatial cross-validation
+		hyperspectral imagery \sep penalized regression \sep machine-learning \sep variable importance \sep spatial cross-validation
 	\end{keyword}
@@
 
-\subsection{Ground data}
+\subsection{In-situ data}
 % Describe Plots
@@
 The four \textit{Pinus radiata} plots Laukiz 1, Laukiz 2, Luiando and Oiartzun are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-Laukiz 1 has the most trees (n = 559) while Laukiz 2 has largest area size.
+Oiartzun has the most observations (n = 529) while Laukiz 2 has largest area size.
 All plots besides Luiando are located nearby the coast.
@@
 
-% Show Map
-
 \begin{figure} [t!]
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/data/study_area_hyperspectral.pdf}}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
 		\caption{Information about the plot locations, the area of hyperspectral coverage and the number of trees per plot.}
@@
 \end{figure}
-
-% describe the hyperspectral data
-
-\cite{Brenning2012}
-
-% \begin{figure} [t!]
-% 	\begin{center}
-% 		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../../04_figures/01_data/study_area.pdf}}
-% 		\caption[Study area]{Spatial distribution of tree observations within the Basque Country, northern Spain, showing infection state by \textit{Diplodia sapinea}.}
-% 		\label{fig: study_area}
-% 	\end{center}
-% \end{figure}
 
@@
 
-% link to PDF with veg indeces
+% link to PDF with veg indices
 All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data and offered by the \texttt{hsdar} package have been calculated.
@@
 Missing values were removed from the mean value calculation.
-In total, 7875 \ac{NRI}s have been calculated ($\frac{125*126}{2}$).
-Some indices returned \texttt{NA} values for some observations and were removed from the dataset, leaving a total of 7471 indices that were available for all plots without missing values.
-Note that due to the mass of variables we cannot state which indices in detail have been removed.
+In total, 7875 Normalized Ratio Indices NRI have been calculated ($\frac{125*126}{2}$).
+Some indices returned \texttt{NA} values for some observations and were removed from the dataset, leaving a total of 7471 variables without missing values.
 
-\subsection{Penalized regression}
+\subsection{Exploratory plot characteristics}
 
-The aim of this work was to find the indices that best explain defoliation within the plots.
-We used penalized regression to account for the large amount of highly correlated predictor variables.
-In a standard \ac{OLS} regression one of the assumptions is that the predictor variables should be independent when minimizing the \ac{RSS} \citep{Bare1981, Hastie2001}:
+To get a better impression on the infection state of the four plots, an exploratory data analysis was done.
+The distribution of the response variable \texttt{defoliation} was visualized as well as the spectral signatures of each plot.
 
-\begin{equation}
-	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2}
-\end{equation}
+\subsection{Benchmarking of algorithms}
 
-where $\beta_{0}$ is the intercept, $\beta_{i}$ the coefficient, $x_{ij}$ the predictor variable and $y_{i}$ the response variable.
-
-If this assumption is violated, regression coefficients can be highly biased.
-They can even show the wrong sign and are very sensitive to adding new independent variables or data points to the model.
-These points reduce the robustness and performance of \ac{OLS} regression when dealing with multicollinearity.
-One approach to overcome these limitations is to penalize the coefficients.
-This method leads to a substantial decrease in variance and better predictive performance compared to \ac{OLS} regression.
-However, it also sacrifies the assumption of unbiased coefficients.
-Hence, the resulting coefficients cannot be used for statistical inference but should be interpreted as a measure of variable importance.
+Multiple algorithms were benchmarked on predictive performance to find the best performing one.
+Besides well-known machine-learning algorithms like \ac{RF} and \ac{SVM} we also used \textit{xgboost} due to its promising results in recent machine-learning competitions.
+To expand the algorithm selection even more, we also used penalized L2 (Ridge) regression.
+This algorithm is able to account for the high correlation between the predictors by penalizing the estimated coefficients of the fitted model.
 
@@
 
-\subsubsection{The lasso penalty}
+\subsubsection{Performance estimation}
 
-While the Ridge penalty $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ shrinks the coefficients, it does not exclude any predictor from the model by setting the coefficient to zero.
-This is what the Lasso penalty (also called $\ell_{1}$) does: It works similar to a "best subset selection" approach by only keeping the most important variables in the model and reducing the coefficients of unimportant ones down to zero.
-This is done by a different \textit{shrinkage penalty} $\lambda \sum_{j=1}^{p} \mid\beta_{j}\mid$ that is added to the \ac{OLS} term when minimizing the \ac{RSS}:
+The algorithms were benchmarked using a spatial \ac{CV} on the plot level.
+The dataset was split into four folds, each fold representing one plot.
+Algorithms were trained on three out of four plots and evaluated on the remaining plot.
+Hyperparameter tuning was also performend on the plot level: Each respective training set, consisting of observations from three plots, was again split into three partitions to estimate the best performing hyperparameter set.
 
-\begin{equation}
-	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p} \mid\beta_{j}\mid
-\end{equation}
+% figure visualizing the folds
 
-Subsequently, Lasso is performing variable selection and results in a sparse model that has the advantage of a simplied interpretability compared to Ridge models.
-However, coeffcients are still biased due to penalization \citep{Hastie2001, James2013}.
+\subsection{Variable importance}
 
-\subsubsection{The elasticnet penalty}
+We aimed to find indices that contribute most when predicting defoliation.
+Permutation-based variable importance was applied on the best performing algorithm.
 
-The \textit{elasticnet penalty} combines both penalties Lasso and Ridge by weighting them with a parameter $\alpha$ that needs to be optimized.
-It was introduced by \cite{Zou2005} with the idea to combine the advantages of both penalty terms.
-The \textit{elasticnet penalty} is written as
+% partial dependence plots??
 
-\begin{equation}
-	\lambda \sum_{j=1}^{p} \alpha\beta_{j}^{2} + (1 - \alpha) \mid\beta_{j}\mid
-\end{equation}
+\section{Results}
 
-where $\alpha$ weights the contribution of either the Lasso or Ridge penalty.
-For $\alpha = 0.5$ both are weighted equally.
-When $\alpha = 0$ only Ridge is used and for $\alpha = 1$ only Lasso applies \citep{Hastie2001}.
-
-\subsection{Modeling}
-
-\subsubsection{Selecting the best penalty}
-
-As the introduced penalties behave different depending on the characteristics of the dataset, we first conducted a nested 10-fold spatial \ac{CV} to find the best method among those three.
-The model comparison was applied to all of the four plots using \ac{RMSE} as the error measure.
-For both the performance evaluation and hyperparameter tuning a spatial sampling using $k$-means clustering was used \citep{sperrorest}.
-For the tuning, 200 random search iterations were used.
-The limits of the tuning space for $\lambda$ were set to the internally calculated (data driven) limits of the R package \texttt{glmnet}.
-
-\subsubsection{Extraction of the most important variables}
-
-Next, we applied the winning method (\ac{RR}) on all plots.
-Hyperparameter tuning of $\lambda$ for the full dataset was again done using a spatial sampling.
-The ten highest coefficients (both positive and negative) were extracted and reported.
-
-% parameter limits
+% Performance estimates
 \begin{table}[b!]
 \centering
-\caption[t]{10-fold 20-times repeated spatial \ac{CV} performances of lasso, ridge and elasticnet penalties on the plot level and the merged dataset using \ac{RMSE} as the error measure.
-Values show the overall mean and standard deviation at the repetition level.}
+\caption[t]{Four-fold spatial \ac{CV} performances of RF, RR, SVM and xgboost using \ac{RMSE} as the error measure. 
+Mean and standard deviation are shown.}
 \begingroup\footnotesize
@@
 	\\
-	Plot/Penalty   & Lasso         & Ridge         & Elasticnet    \\
+	RF            & RR            & SVM           & xgboost \\
 	\hline
-	Laukiz 1       & 133.15 (4.66) & 86.71 (2.96)  & 121.04 (4.15) \\
-	Laukiz 2       & 91.93 (6.78)  & 29.73 (0.38)  & 47.18 (13.70) \\
-	Luiando        & 74.83 (0.311) & 75.95 (0.23)  & 74.74 (0.29)  \\
-	Oiartzun       & 327.93        & 106.65        & 260.38        \\
-	Merged dataset & 54.80 (17.58) & 53.67 (17.95) & 54.66 (17.67) \\
+	54.80 (17.58) & 53.67 (17.95) & 54.66 (17.67) &         \\
 	\bottomrule
@@
 
-\subsubsection{Linking variables to plot characteristics}
-
-To interpret the outcomes of the models on a plot level, we linked the winning variables of each to dataset attributes that describe the underlying plot characteristics.
-For example, a plot which a low tree density might inherit more information from the bare ground in the calculated indices while a plot with a very high tree density might in contrast contain information from multiple trees.
-Also, the overall level of defoliation for the whole plot might possibly have an effect how well an index is able to describe the situation.
-
-\subsubsection{Creation of a super model with all observations}
-
-After the plot level analysis, we merged all observations into a single dataset and fitted another \ac{RR} model.
-We used a block-level based spatial \ac{CV} for this setup.
-For a total number of $m$ folds ($n$ = number of plots), every plot is once the test set while the training is done on the remaining ones.
-This ensures that the test set is fully spatially independent from the training set.
-Tuning was also performned on the block level within the training set on $n - 1$ folds using 200 random search iterations.
-Due to the fixing of the indices on a plot level, varying those to use multiple repetitions is not possible.
-The idea behind fitting a supermodel is that this model includes information from all plots rather than just from a single plot.
-This may possibly reduce the predictive error and create a more robust model.
-
-% Points to mention
-% * All penalization methods are suited for different kinds of dataset structures
-% * Find out which one works best on average and then use this method to create create models for prediction and variable importance
-
 \section{Results}
 
-\subsection{Plot characteristics}
-
-Luiando shows the highest defoliation ($\bar{x} = 68.36 \%$) among the plots while Laukiz 2 is the healthiest ($\bar{x} = 17.73 \%$) (\autoref{fig:defol_boxplots}).
-Laukiz 1 and Oiartzun both show a medium defoliation level around 50 \%.
-Oiartzun consists mainly of trees that show either a very high level of defoliation ($70 \% <=$) or none at all.
-Laukiz 1 shows an evenly distributed level of defoliation across the entire plot.
-
-The high defoliation level of Luiando and Oiartzun is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
-Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
-Oiartzun is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
-Laukiz 2 shows a mean tree density of 62.51 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.35 (Laukiz 1), 33.77 (Luiando), 35.02 (Oiartzun)) (\autoref{fig:rmse_defol_dens}).
+\subsection{Exploratory data analysis}
 
@@
 \end{figure}
+
+Luiando shows the highest defoliation ($\bar{x} = 68.36 \%$) among the plots while Laukiz 2 is the healthiest ($\bar{x} = 17.73 \%$) (\autoref{fig:defol_boxplots}).
+Laukiz 1 and Oiartzun both show a medium defoliation level around 50 \%.
+Oiartzun consists mainly of trees that show either a very high level of defoliation ($70 \% <=$) or none at all.
+Laukiz 1 shows an evenly distributed level of defoliation across the entire plot.
+
+The high defoliation level of Luiando and Oiartzun is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
+Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
+Oiartzun is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
+Laukiz 2 shows a mean tree density of 62.51 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.35 (Laukiz 1), 33.77 (Luiando), 35.02 (Oiartzun)) (\autoref{fig:rmse_defol_dens}).
 
@@
 
-% RMSE vs defol and point dens
-Laukiz 2 showed contrary results compared to all other plots when linking \ac{RMSE} against mean defoliation and mean point density (\autoref{fig:rmse_defol_dens}).
-Laukiz 1, Luiando and Oiartzun show almost identical values in terms of point density (34.35m, 33.77m, 35.01m).
-Due to the high similarity, these should be interpreted as a group and not as single values.
+% RMSE vs CV and point dens
+Laukiz 2 showed contrary results compared to all other plots when linking \ac{RMSE} against coefficient of variation and mean point density (\autoref{fig:rmse_defol_dens}).
+Comparing RMSE against $CV/skewness$ shows a $log_{2}(-x)$ relationship.
 
@@
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/mean-defol_mean-point-dens.pdf}}
-		\caption{RMSE vs. mean defoliation and point density}
-		\label{fig:rmse_defol_dens}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/plot-characteristics.pdf}}
+		\caption{RMSE vs. mean point density, coefficient of variation and coefficient of variation / skewness.}
+		\label{fig:plot-characteristics}
 	\end{center}
@@
 
+% band importance
+\begin{figure} [b!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[angle=90, height = 0.95\textheight] {../03_figures/results/important-bands_boxplot.pdf}}
+		\caption{test}
+		\label{fig:important-bands}
+	\end{center}
+\end{figure}
+
 \section{Discussion}
@@
 
-For Laukiz1, Luiando and Oiartzun \ac{RMSE} seems to increase with a higher point density on a first look.
+For Laukiz1, Luiando and Oiartzun \ac{RMSE} seems to increase with a higher point density at a first glance.
 However, the point densities of these plots are very similar (33.7 m - 35.01 m) and should be interpreted as a group instead of single values.
-Comparing this group to Laukiz 2, which has point density of 62 m, the error decreases with a higher point density.
-Linking the error against the mean defoliation of a plot does not lead to a clear conclusion:
-While Laukiz1, Luiando and Oiartzun give the impression of a decreasing RMSE when defoliation increases, Laukiz 2 clearly shows that a plot with a low defoliation can results in a low RMSE estimate (\autoref{fig:rmse_defol_dens}).
+With Laukiz2 being completely off from the other plots in terms of mean point density, no pattern can be extracted from this result.
+Linking RMSE vs coefficient of variation shows the same relationship as linking against mean point density.
+The interesting $log_{2}(-x)$ relationship for RMSE vs.
+coefficient of variation / skewness should be interpreted with caution: The sample size of four plots is not representative to make general statements here.
+This finding should be verified with more observations in future studies.
 

### 2018-07-11 02_scripts/01_hyperspectral_indices/02_hyperspectral-ndvi-raster.Rmd
11/07/2018

@@
 ---
-title: "R Notebook"
-output: html_notebook
-editor_options: 
-  chunk_output_type: inline
+title: "Create NDVI raster files for each plot"
+output: 
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
 ---
 
-# Information about this script
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE,
+                      cache = TRUE)
+```
 
-This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`
+```{r, results='hide'}
+library(raster)
+devtools::install_github("pat-s/hsdar")
+library(hsdar)
+library(pbmcapply)
+library(glue)
+library(magrittr)
+library(purrr)
+```
 
-Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).
-
-Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
-It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.
+# Specify wavelength
 
 ```{r}
-pacman::p_load(raster, hsdar, here, pbapply, parallel)
+wavelength = readRDS("/data/patrick/mod/hyperspectral/wavelength.Rds")
 ```
 
-# specify wavelength
-
-```{r}
-wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")
-```
-
-# read rasters (cropped by plot)
+# Read rasters (cropped by plot)
 
@@
 ```{r}
-files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
-                    pattern = ".tif$", full.names = TRUE)
+list.files("/data/patrick/mod/hyperspectral/atm-cropped", 
+                    pattern = ".tif$", full.names = TRUE) %>% 
+  map(~ brick(.x)) -> spectral_data
 ```
 
-## read in files as "RasterBrick"
+# Coerce to class "HyperSpecRaster"
 
 ```{r}
-spectral_data <- lapply(files, function(x) brick(x))
+hyperspecs = map(spectral_data, function(x) HyperSpecRaster(x, wavelength))
 ```
 
-# make class "HyperSpecRaster"
-```{r}
-hyperspecs <- lapply(spectral_data, function(x) hsdar::HyperSpecRaster(x, wavelength))
+# Calculate NDVI in parallel
+
+* Assign to `tmp` to prevent list printing to console
+* Use 'NDVI2' as formula
+
+```{r, cache=FALSE}
+tmp <- pbmclapply(hyperspecs, function(x) 
+  vegindex(x, "NDVI2", 
+           filename = glue("/data/patrick/mod/hyperspectral/ndvi/ndvi2-{name}",
+                           name = basename(x@file@name)),
+           bnames = "NDVI"), mc.cores = 28)
+rm(tmp)
 ```
-
-# account for use on local laptop or server
-```{r}
-if (parallel::detectCores() > 10) {
-  cores <- 28
-} else (cores <- 1)
-```
-
-# calculate NDVI
-
-assign to object to prevent list printing to console ### use 'NDVI2' as formula 
-```{r}
-pboptions(style = 1, type = "timer")
-tmp <- pblapply(hyperspecs, function(x) 
-  vegindex(x, "NDVI2", filename = paste0("/home/shares/data/LIFE/mod/hyperspectral/ndvi/", "ndvi2-",
-                                         basename(x@file@name)),
-           bnames = "NDVI"), cl = cores)
-```

### 2018-07-11 02_scripts/01_hyperspectral_indices/03_hyperspectral-ndvi-analysis.Rmd
11/07/2018

@@
 title: "NDVI analysis of hyperspectral data"
-output: html_notebook
-editor_options: 
-  chunk_output_type: inline
+output: 
+  bookdown::html_document2:
+    toc: true
+    toc_float: true
 ---
-# Data & packages
 
-Loads list of 28 cropped (to respective plot) atmospherical corrected hyperspectral images.  
-See R script `hs.ndvi.R` for multicore creation.
-
-Move to chunk 'hsdar package' if file exists!
-```{r}
-pacman::p_load(rgdal, raster, mapview, hsdar, rio, ggplot2, colorRamps)
-
-if (file.exists("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/all-cropped.rds")) {
-  files <- as.list(rio::import("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/all-cropped.rds"))
-} else {
-  files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/", 
-                    pattern = ".tif$", full.names = TRUE)
-
-  files1 <- lapply(files, function(x) raster::brick(x))
-}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE,
+                      cache = TRUE)
 ```
 
-Define 'Get values row by row' (vignette `raster::functions`)
-* x = multy layer Raster file (applying `brick()`)
-* a = value to add to existing value
-```{r}
-source(here("R/helper.funs/get.values.row.by.row.R"))
-```
-
-# Get values from all raster files (bricks!) using `lapply`
-```{r}
-# files_atm <- lapply(files_atm, function(x) f8(x, 0, progress = "text"))
-# saveRDS(files_atm, here("R/tmp/files_atm_with_values.rda")
-files_atm <- readRDS("/home/patrick/PhD/R/tmp/files_atm_with_values.rda")
-```
-
-Get wavelengths out of README table (http://datascienceplus.com/extracting-tables-from-pdfs-in-r-using-the-tabulizer-package/)
-
-Does not work because wavelength table on page 10 of README is not a table but an image...
-```{r}
-# devtools::install_github("ropenscilabs/tabulizer")
-# devtools::install_github("ropenscilabs/tabulizerjars")
-
-# library(tabulizer)
-
-# tabs <- extract_tables("/home/shares/data/original/hyperspectral/0_README_spanish.pdf", pages = 10)
-```
-
-# hsdar package
-```{r}
-# specify wavelength of each band
-wavelength <- c(404.08, 408.5, 412.92, 417.36, 421.81, 426.27, 430.73, 435.20, 
-                439.69, 444.18, 448.68, 453.18, 457.69, 462.22, 466.75, 471.29,
-                475.83, 480.39, 484.95, 489.52, 494.09, 498.68, 503.26, 507.86,
-                512.47, 517.08, 521.70, 526.32, 530.95, 535.58, 540.23, 544.88,
-                549.54, 554.20, 558.86, 563.54, 568.22, 572.90, 577.60, 582.29,
-                586.99, 591.70, 596.41, 601.13, 605.85, 610.58, 615.31, 620.05,
-                624.79, 629.54, 634.29, 639.04, 643.80, 648.56, 653.33, 658.10,
-                662.88, 667.66, 672.44, 677.23, 682.02, 686.81, 691.60, 696.40,
-                701.21, 706.01, 710.82, 715.64, 720.45, 725.27, 730.09, 734.91,
-                739.73, 744.56, 749.39, 754.22, 759.05, 763.89, 768.72, 773.56, 
-                778.40, 783.24, 788.08, 792.93, 797.77, 802.62, 807.47, 812.32,
-                817.17, 822.02, 826.87, 831.72, 836.57, 841.42, 846.28, 851.13,
-                855.98, 860.83, 865.69, 870.54, 875.39, 880.24, 885.09, 889.94,
-                894.79, 899.64, 904.49, 909.34, 914.18, 919.03, 923.87, 928.71,
-                933.55, 938.39, 943.23, 948.07, 952.90, 957.73, 962.56, 967.39,
-                972.22, 977.04, 981.87, 986.68, 991.50, 996.31)
-```
-
-Read in raster data either using `rgdal` or `raster` package. 
-`raster` seems to be prefered but `rgdal` is a lot faster in reading values
+# Load packages
 
 ```{r}
-## create HyperSpecRaster file
-# ra <- hsdar::HyperSpecRaster(files_atm[[1]], wavelength)
-# tr <- blockSize(ra)
-
-##  create SpecLib file
-# spectral_data1 <- f8(files1[[1]], 0, progress = "text")
-
-# speclib <- hsdar::speclib(spectral_data, wavelength)
-```
-
-plot spectral signature
-```{r}
-# plot(spectral_data, main = "Default Plot")
-
-# plot(spectral_data, FUN = "median") # not working
-```
-
-Visualize NDVI
-```{r}
-mapview::mapview(spectral_data1, zcol = "ndvi", maxpixels = 4664960, na.color = "white")
+library(purrr)
+library(magrittr)
+library(pbmcapply)
+library(raster)
+library(rasterVis)
+library(viridis)
+library(gridExtra)
+library(tools)
+library(stringr)
+library(tibble)
+library(dplyr)
+library(ggplot2)
 ```
@@
 
-## Read in files
+Note: Names of raster objects will always have a dot in it which cannot be replaced by a dash or a whitespace.
 
-1. List files
-2. Read in using `rgdal::readGDAL`
-3. Get filenames
-4. Rename `band1` to `NDVI` and label SpGrDaFr with plot name
-4. Get pixel count of each file
-5. Create DF with one NDVI column and a factor column with filenames
+That is the reason why we set the names in `levelplot_mod()`.
+
 ```{r}
-files.list <- list.files("/home/shares/data/LIFE/mod/hyperspectral/ndvi", 
-                    pattern = "ndvi2.*\\.tif", full.names = TRUE)
+ndvi_data <- list.files("/data/patrick/mod/hyperspectral/ndvi",
+  pattern = "ndvi2.*\\.tif", full.names = TRUE
+) %>%
+  map(~ raster(.x)) %>%
+  map(~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")) %>%
+  map(~ set_names(.x, gsub(".", " ", str_sub(names(.x), 7), fixed = T)))
 
-files.ndvi <- lapply(files.list, function(x) rgdal::readGDAL(x, silent = TRUE))
-# rename band name to 'NDVI'
-for (i in seq_along(files.ndvi)) {
-  colnames(files.ndvi[[i]]@data) <- "NDVI"
+names_ndvi_data <- map_chr(ndvi_data, names) %>%
+  map_chr(~ str_replace_all(.x, "\\.", " ")) %>%
+  map_chr(toTitleCase)
+```
 
+```{r}
+levelplot_mod <- function(x, names_raster) {
+  levelplot(
+    x,
+    main = names_raster,
+    margin = FALSE,
+    pretty = TRUE, 
+    colorkey = list(
+      width = 0.75, 
+      space = "right",
+      labels = list(
+        at = seq(-1, 1,
+          by = 0.2
+        ),
+        font = 4
+      ), 
+      axis.line = list(col = "black")
+    ),
+    par.strip.text = list(col = "white", cex = 1.2),
+    par.settings = list(
+      axis.line = list(col = "transparent"), 
+      strip.background = list(col = "black")
+    ),
+    col.regions = viridis,
+    at = seq(-1, 1, len = 101)
+  )
 }
+```
 
-names.file <- sapply(files.list, function(x) basename(x))
-# label SpGrDaFr with plot name
-names(files.ndvi) <- names.file
+```{r}
+ndvi_plots <- map2(ndvi_data, names_ndvi_data, ~ levelplot_mod(.x, .y))
+```
 
-pixel.count <- sapply(files.ndvi, function(x) length(x@data$NDVI))
+## Plot 1-9
 
-ndvi.all <- c()
-ndvi.all <- unlist(lapply(files.ndvi, function(x) append(ndvi.all, x@data$NDVI)))
+```{r}
+grid.arrange(grobs = ndvi_plots[c(1:9)], ncol = 3)
+```
 
-plot.names <- c()
-for (i in seq_along(names.file)) {
-  plot.names <- append(plot.names, rep(basename(names.file[i]), pixel.count[i]))
-}
+## Plot 10-19
 
-df <- data.frame(NDVI = ndvi.all,
-                 plot = as.factor(plot.names))
+```{r}
+grid.arrange(grobs = ndvi_plots[c(10:19)], ncol = 3)
+```
 
-# remove NA rows
-df <- df[complete.cases(df),]
+## Plot 20:28
+
+```{r}
+grid.arrange(grobs = ndvi_plots[c(20:28)], ncol = 3)
+```
+
+## Plot demonstration plots
+
+```{r}
+grid.arrange(grobs = ndvi_plots[c(15, 16, 18, 19)], ncol = 2)
 ```
@@
 
+Note: Hernani is missing as we have no hyperspectral coverage for that plot.
+
 ```{r}
-ndvi.histograms <- ggplot(df, aes(NDVI)) +
-  #geom_histogram() + 
-  # normalize y-axis of each facet http://stackoverflow.com/questions/16339204/normalizing-faceted-histograms-separately-in-ggplot2
-  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
-  facet_wrap(~ plot) + 
+plot_values_comb <- map(ndvi_data, ~ as_tibble(.x@data@values)) %>%
+  map2(names_ndvi_data, ~ mutate(.x, Name = .y)) %>%
+  map(na.omit) %>%
+  reduce(full_join, by = c("value", "Name"))
+
+plot_values_comb %<>%
+  mutate(Name = as.factor(Name))
+
+colnames(plot_values_comb) <- c("NDVI", "name")
+```
+
+Normalize y-axis of each facet: http://stackoverflow.com/questions/16339204/normalizing-faceted-histograms-separately-in-ggplot2
+
+```{r}
+ndvi_histograms <- ggplot(plot_values_comb, aes(x = NDVI)) +
+  geom_histogram(aes(y = (..count..) / tapply(..count.., ..PANEL.., sum)[..PANEL..])) +
+  facet_wrap(~ name) +
   ylab("Percentage pixel count") +
-  xlab("NDVI (0.25 to 1)") + 
-  scale_x_continuous(limits = c(0.25, 1)) + 
-  cowplot::background_grid()
+  xlab("NDVI (-1 to 1)") +
+  scale_x_continuous(limits = c(-1, 1)) +
+  hrbrthemes::theme_ipsum_rc()
 ```
@@
 ## Count pixels per plot (size 1m^2)
+
 ```{r}
-table(df$plot)
+plot_values_comb %>%
+  group_by(name) %>%
+  count() %>%
+  arrange(desc(n))
 ```
@@
 Biased towards low values (non-forest pixels)
+
 ```{r}
-stats::quantile(df$NDVI)
+quantile(plot_values_comb$NDVI)
 ```
@@
 ```{r}
-boxplot(df$NDVI)
+boxplot(plot_values_comb$NDVI)
 ```
-
-```{r}
-color <- colorRamps::green2red(2)
-rasterVis::levelplot(files.ndvi[[4]])
-mapview::mapView(as.raster(files.ndvi[[4]]), zcol = "NDVI", na.color = "blue", color = color)
-```
-
-## plot NDVI rasters
-```{r}
-# convert to raster layer for plotting
-files.ndvi.raster <- lapply(files.ndvi, function(x) raster::raster(x))
-#set names
-for (i in seq_along(files.ndvi.raster)) {
-  files.ndvi.raster[[i]]@file@name <- names(files.ndvi)[i]
-}
-
-# using mapview
-list.plots <- lapply(files.ndvi.raster, function(x) mapview::mapview(x, use.layer.names = TRUE, na.color = "white"))
-# using rasterVis::levelplot
-list.plots <- lapply(files.ndvi.raster, function(x) rasterVis::levelplot(x,
-                                                                         main = x@file@name,
-                                                                         margin = FALSE,
-                                                                         pretty = TRUE,
-                                                                         col.regions = rev(colorRamps::green2red(400)), 
-                                                                         at = seq(0.5, 1, len = 101))
-)
-
-gridExtra::grid.arrange(list.plots[[1]], list.plots[[2]], list.plots[[3]], 
-                                        list.plots[[4]], list.plots[[5]], list.plots[[6]],
-                                        list.plots[[7]], list.plots[[8]], list.plots[[9]],
-                                        ncol = 3)
-
-gridExtra::grid.arrange(list.plots[[10]], list.plots[[11]], list.plots[[12]],
-                                        list.plots[[13]], list.plots[[14]], list.plots[[15]],
-                                        list.plots[[16]], list.plots[[17]], list.plots[[18]],
-                                        ncol = 3)
-
-gridExtra::grid.arrange(list.plots[[19]], list.plots[[20]], list.plots[[21]],
-                                        list.plots[[22]], list.plots[[23]], list.plots[[24]], 
-                                        list.plots[[25]], list.plots[[26]], list.plots[[27]], 
-                                        list.plots[[28]],
-                                        ncol = 3)
-
-```
-
-## save NDVI raster plots
-
-```{r}
-png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.1.png", width = 16.51, height = 12.07, units = "in", res = 300)
-gridExtra::grid.arrange(list.plots[[1]], list.plots[[2]], list.plots[[3]], 
-                                        list.plots[[4]], list.plots[[5]], list.plots[[6]],
-                                        list.plots[[7]], list.plots[[8]], list.plots[[9]],
-                                        ncol = 3)
-dev.off()
-
-png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.2.png", width = 16.51, height = 12.07, units = "in", res = 300)
-gridExtra::grid.arrange(list.plots[[10]], list.plots[[11]], list.plots[[12]],
-                                        list.plots[[13]], list.plots[[14]], list.plots[[15]],
-                                        list.plots[[16]], list.plots[[17]], list.plots[[18]],
-                                        ncol = 3)
-dev.off()
-
-png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.3.png", width = 16.51, height = 12.07, units = "in", res = 300)
-gridExtra::grid.arrange(list.plots[[19]], list.plots[[20]], list.plots[[21]],
-                                        list.plots[[22]], list.plots[[23]], list.plots[[24]], 
-                                        list.plots[[25]], list.plots[[26]], list.plots[[27]], 
-                                        list.plots[[28]],
-                                        ncol = 3)
-dev.off()
-```
-
-## save ndvi histograms
-```{r}
-png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi.histograms.normalized.png", width = 16.51, height = 12.07, units = "in", res = 300)
-ndvi.histograms
-dev.off()
-```

### 2018-07-11 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd
14/05/2018
11/07/2018

@@
   out_veg %<>% 
-    map2(seq_along(buffer), ~ setNames(.x, glue("bf", .y, "_{name}", 
+    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}", 
                                              name = names(out_veg[[.y]]))))
@@
   out_nbi %<>%
-    map2(seq_along(buffer), ~ setNames(.x, glue("bf", .y, "_{name}",
+    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
                                              name = names(out_nbi[[.y]]))))

### 2018-07-11 02_scripts/01_hyperspectral_indices/08_extract_bands_to_points.Rmd
11/07/2018

--- 2018-07-11 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd
@@
 library(dplyr)
+library(furrr)
 ```
@@
 
-  ras_veg_indices <- brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/{plot_name}.grd"))
-
-  ras_nbi <- brick(glue("/data/patrick/mod/hyperspectral/narrow-band-indices/narrow-band-indices-{plot_name}.grd"))
+  bands_ras <- brick(glue("/data/patrick/mod/hyperspectral/atm-cropped/{plot_name}.tif"))
 
   # calculate buffered veg index
-  out_veg <- pbmclapply(buffer, function(x) 
-    raster::extract(ras_veg_indices, points, buffer = x, 
+
+  # plan(multiprocess)
+  # out_bands <- future_map(buffer, ~
+  #   raster::extract(bands_ras[[5:126]], points, buffer = .x, 
+  #                   fun = mean, df = TRUE,
+  #                   na.rm = TRUE),.progress = TRUE)
+
+  out_bands <- pbmclapply(buffer, function(x)
+    raster::extract(bands_ras[[5:126]], points, buffer = x,
                     fun = mean, df = TRUE,
@@
 
-  out_veg %<>% 
+  out_bands %<>% 
     map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}", 
-                                             name = names(out_veg[[.y]]))))
-
-  out_nbi <- pbmclapply(buffer, function(x)
-    raster::extract(ras_nbi, points, buffer = x,
-                    fun = mean, df = TRUE,
-                    na.rm = TRUE) , mc.cores = cores)
-
-  out_nbi %<>%
-    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
-                                             name = names(out_nbi[[.y]]))))
+                                             name = names(out_bands[[.y]]))))
 
   # merge all data frames (buffers)
-  all_veg <- bind_cols(out_veg)
-
-  all_nbi <- bind_cols(out_nbi)
+  out_bands <- bind_cols(out_bands)
 
   points %<>% 
-    bind_cols(all_veg) %>% 
-    bind_cols(all_nbi)
+    bind_cols(out_bands)
 
@@
   points %>% 
-    saveRDS(glue("/data/patrick/mod/tree-per-tree/2016/{plot_name}/all-indices-{plot_name}_{bf_name}.rda"))
+    saveRDS(glue("/data/patrick/mod/tree-per-tree/2016/{plot_name}/all-bands-{plot_name}_{bf_name}.rda"))
 }
@@
 ```{r}
+# plan(multiprocess)
+# future_map(c("laukiz1", "laukiz2", "oiartzun", "luiando"), ~
+#            extract_to_plot(.x, 2, "bf2"), .progress = TRUE)
 pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"), 

### 2018-07-11 02_scripts/022_preprocessing.Rmd
11/07/2018

@@
 
+luiando has some NA obs that need to be removed, otherwise the whole column will be removed by `Filter()`
+
 ```{r}
+data[[3]] %<>% 
+  na.omit()
+
 data %<>% 
@@
   map(~ as.data.table(.x)) %>% 
+  #map(~ na.omit(.x)) %>% # some obs in luiando are NA after extraction
+  map(~ dplyr::select(.x, -tree.number, -decoloration, -canker, -diameter, -height)) %>% 
   map(~ Filter(function(x) !any(is.na(x)), .x))

### 2018-07-11 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
11/07/2018

@@
 library(mlr)
+library(mlrMBO)
 library(glue)
@@
 library(pbmcapply)
+library(stringr)
+library(ggpubr)
 ```
@@
 task = makeRegrTask(id = "laukiz1", data = data, 
-                    target = "deftot", coordinates = coords)
+                    target = "defoliation", coordinates = coords)
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 102, upper = 10228))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 78, upper = 7879))
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set =  ps_glmnet))
 inner <- makeResampleDesc("SpCV", iters = 10)
@@
 tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
+                                control = tune.ctrl, show.info = TRUE,
                                 measures = list(rmse))
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+parallelStart(mode = "multicore", level = "mlr.resample", cpus = 20,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
+                                 par.set = ps_glmnet, control = tune.ctrl, 
                                  measure = list(rmse))
@@
 
+#### Important spectral regions
+
+Get the first 100 most important indices.
+We take the absolute value of the coefficient to consider both sides.
+
+We only take the top 100 indices so that the result is not cluttered by mid-ranged indices summping up in the end-result.
+Apparently, we have 56 unique bands within the top 100 indices.
+
+```{r}
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7500))), 
+       coef = abs(as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7500))))) %>% 
+  arrange(desc(coef)) %>% 
+  mutate(coef = coef) -> top_inds_rr_laukiz1
+saveRDS(top_inds_rr_laukiz1, "/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_laukiz1.rda")
+```
+
+Remove the "bf2" information and filter out NRI only.
+
+```{r}
+top_inds_rr_laukiz1 %<>% 
+  mutate(index = str_sub(index, 9, 15)) %>% 
+  filter(str_detect(index, "b"))
+```
+
+Standardize the coefficients
+
+```{r}
+top_inds_rr_laukiz1 %<>% 
+  mutate(coef = as.numeric((coef / min(coef)) * coef))
+```
+
+Extract all band numbers.
+
+```{r}
+top_inds_rr_laukiz1 %>% 
+  dplyr::select(index) %>% 
+  str_match_all("[0-9]+") %>% 
+  flatten() %>% 
+  as.numeric() -> all_bands_rr
+```
+
+```{r}
+hist(all_bands_rr)
+```
+
+Now add a weight to all bands.
+Starting from 100-2 depending on the score.
+why 2? Because we only have 198 values because we removed one vegetation index.
+
+```{r}
+tibble(band = all_bands_rr, weight = rep(top_inds_rr_laukiz1$coef, times = 1, each = 2)) %>% 
+  group_by(band) %>% 
+  summarise(score = sum(weight)) %>% 
+  mutate(band = band) %>% 
+  arrange(desc(score)) -> tibble_most_imp
+```
+
+Classify the bands into spectral regions
+
+1 - 60: visible (404 - 680)
+61 - 71: red edge (680 - 730)
+72 - 126: NIR (730 - 996)
+
+```{r}
+tibble_most_imp %<>% 
+  mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 - 680)", 
+                           (band >= 61 & band <= 71) ~ "Red Edge (680 - 730)",
+                           band >= 72 ~ "NIR (730 - 996)"))) %>% 
+  arrange(desc(score)) %>% 
+  mutate(band = as.factor(band))
+```
+
+centered by the mean of the scaled coefficients.
+
+```{r}
+ggdotchart(tibble_most_imp, x = "band", y = "score",                              # Color by groups
+           sorting = "descending",                       # Sort value in descending order
+           add = "segments",                             # Add segments from y = 0 to dots
+           color = "group",
+           palette = "nejm",
+           group = "score",
+           combine = T,
+           rotate = TRUE,                                # Rotate vertically
+           dot.size = 1,                                 # Large dot size
+           #label = round(tibble_most_imp$sum),                        # Add mpg values as dot labels
+           ggtheme = theme_pubr(base_size = 9, margin = FALSE)) + 
+  labs(y = "Score", color = "Spectral regions")
+```
+
 ### Lasso
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10.3))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7.87))
 ```
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
+parallelStart(mode = "multicore", level = "mlr.resample", cpus = 10,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_lasso_bf2_10folds.rda")
 ```
@@
 lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
+                                alpha = 1,
                                 s = 1810,
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10228),
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.07, upper = 7879),
                           makeNumericParam("alpha", lower = 0, upper = 1))
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+parallelStart(mode = "multicore", level = "mlr.resample", cpus = 10,
               mc.set.seed = TRUE) # only parallelize the tuning

### 2018-07-11 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
11/07/2018

@@
 task = makeRegrTask(id = "laukiz2", data = data, 
-                    target = "deftot", coordinates = coords)
+                    target = "defoliation", coordinates = coords)
 ```
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+parallelStart(mode = "multicore", level = "mlr.resample", cpus = 40,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
+                                 par.set = ps_glmnet, control = tune.ctrl, 
                                  show.info = TRUE, measure = list(rmse))
@@
 
+#### Important spectral regions
+
+Get the first 100 most important indices.
+We take the absolute value of the coefficient to consider both sides.
+
+We only take the top 100 indices so that the result is not cluttered by mid-ranged indices summping up in the end-result.
+
+```{r}
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690))), 
+       coef = abs(as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690))))) %>% 
+  arrange(desc(coef)) %>% 
+  mutate(coef = coef) -> top_inds_rr_laukiz2
+saveRDS(top_inds_rr_laukiz2, "/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_laukiz2.rda")
+```
+
 ### Lasso
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
+parallelStart(mode = "multicore", level = "mlr.resample", cpus = 20,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 ```
+
+### SVM
+
+```{r}
+lrn_ksvm <- makeLearner("regr.ksvm",
+                        kernel = "rbfdot")
+```
+
+```{r}
+ps_svm <- makeParamSet(makeNumericParam("C", lower = -15, upper = 15,
+                                        trafo = function(x) 2 ^ x),
+                       makeNumericParam("sigma", lower = -15, upper = 15,
+                                        trafo = function(x) 2 ^ x))
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_svm))
+inner <- makeResampleDesc("SpCV", iters = 10)
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_ksvm, resampling = inner, par.set = ps_svm,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_svm <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/laukiz2_svm.rda")
+```

### 2018-07-11 02_scripts/033-luiando/03-modeling_luiando.Rmd
11/07/2018

@@
 task = makeRegrTask(id = "luiando", data = data, 
-                    target = "deftot", coordinates = coords)
+                    target = "defoliation", coordinates = coords)
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 15, upper = 1558))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 10, upper = 10000))
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_glmnet))
 inner <- makeResampleDesc("SpCV", iters = 10)
@@
 
+#### Important spectral regions
+
+Get the first 100 most important indices.
+We take the absolute value of the coefficient to consider both sides.
+
+We only take the top 100 indices so that the result is not cluttered by mid-ranged indices summping up in the end-result.
+
+```{r}
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560))), 
+       coef = abs(as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560))))) %>% 
+  arrange(desc(coef)) %>% 
+  mutate(coef = coef * 10000000000000000000000000000000000) -> top_inds_rr_luiando
+saveRDS(top_inds_rr_luiando, "/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_luiando.rda")
+```
+
 ### Lasso
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.56))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 15000))
 ```

### 2018-07-11 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
15/05/2018
11/07/2018

@@
 task = makeRegrTask(id = "oiartzun", data = data, 
-                    target = "deftot", coordinates = coords)
+                    target = "defoliation", coordinates = coords)
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 55, upper = 5525))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 31, upper = 3187))
 ```
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 10)
 ```
@@
 
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
+glmnet_train_tuned_oiartzun = mlr::train(lrn_glmnet_tuned, task)
 ```
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 5460))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 5460)))) %>% 
   arrange(desc(coef))
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 5460))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 5460)))) %>% 
   arrange(coef)
@@
 ```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
+plot(glmnet_train_tuned_oiartzun$learner.model, "lambda")
 ```
@@
 ```{r}
-plot(coef(glmnet_train_tuned$learner.model))
+plot(coef(glmnet_train_tuned_oiartzun$learner.model))
+```
+
+#### Important spectral regions
+
+Get the first 100 most important indices.
+We take the absolute value of the coefficient to consider both sides.
+
+We only take the top 100 indices so that the result is not cluttered by mid-ranged indices summping up in the end-result.
+
+```{r}
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 5460))), 
+       coef = abs(as.vector(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 5460))))) %>% 
+  arrange(desc(coef)) %>% 
+  mutate(coef = coef) -> top_inds_rr_oiartzun
+saveRDS(top_inds_rr_oiartzun, "/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_oiartzun.rda")
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.5))
+ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.18))
 ```
@@
 #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
+glmnet_train_tuned_oiartzun = mlr::train(lrn_glmnet_tuned, task)
 ```
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 1810))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 1810)))) %>% 
   arrange(desc(coef))
@@
 ```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 1810))), 
+       coef = as.vector(as.matrix(coef(glmnet_train_tuned_oiartzun$learner.model, s = 1810)))) %>% 
   arrange(coef)
@@
 ```{r}
-plot(glmnet_train_tuned$learner.model, "norm")
+plot(glmnet_train_tuned_oiartzun$learner.model, "norm")
 ```
@@
 ```{r}
-plot(coef(glmnet_train_tuned$learner.model))
+plot(coef(glmnet_train_tuned_oiartzun$learner.model))
 ```

### 2018-07-11 02_scripts/035_all-plots/03-modeling_all.Rmd
11/07/2018

@@
 ```{r, results='hide'}
-devtools::install_github("tidyverse/ggplot2")
+#devtools::install_github("tidyverse/ggplot2")
 devtools::install_github("mlr-org/mlr@blocking")
@@
 
-Leaving out Luiando (see README for reason).
-
 ```{r}
@@
 task = makeRegrTask(id = "all_plots", data = data, 
-                    target = "deftot", coordinates = coords,
-                    blocking = factor(rep(1:4, c(559, 382, 301, 499))))
+                    target = "defoliation", coordinates = coords,
+                    blocking = factor(rep(1:4, c(479, 451, 291, 529))))
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set =  ps_glmnet))
 inner <- makeResampleDesc("Blocking")
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
+parallelStart(mode = "multicore", level = "mlrMBO.feval", cpus = 40,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 
+#### Important spectral regions
+
+Get the first 100 most important indices.
+We take the absolute value of the coefficient to consider both sides.
+
+We only take the top 100 indices so that the result is not cluttered by mid-ranged indices summping up in the end-result.
+
+```{r}
+tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 875))), 
+       coef = abs(as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 875))))) %>% 
+  arrange(desc(coef)) %>% 
+    mutate(coef = coef) -> top_inds_rr_all_plots
+saveRDS(top_inds_rr_all_plots, "/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_all_plots.rda")
+```
+
 ### Lasso
@@
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
+parallelStart(mode = "multicore", level = "mlr.resample", cpus = 48,
               mc.set.seed = TRUE) # only parallelize the tuning
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_glmnet))
 inner <- makeResampleDesc("Blocking")
@@
 ```
+
+### SVM
+
+```{r}
+lrn_ksvm <- makeLearner("regr.ksvm",
+                        kernel = "rbfdot")
+```
+
+```{r}
+ps_svm <- makeParamSet(makeNumericParam("C", lower = -15, upper = 15,
+                                        trafo = function(x) 2 ^ x),
+                       makeNumericParam("sigma", lower = -15, upper = 15,
+                                        trafo = function(x) 2 ^ x))
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_svm))
+inner <- makeResampleDesc("Blocking")
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_ksvm, resampling = inner, par.set = ps_svm,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("Blocking")
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_svm <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/all_plots_svm_blocking.rda")
+```
+
+### xgboost
+
+```{r}
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error",
+                     eta = 0.1
+                   ))
+```
+
+```{r}
+ps_xgboost = makeParamSet(
+  makeDiscreteParam("nrounds", values = c(10, 100, 1000, 10000)),
+  makeNumericParam("colsample_bytree", lower=0.1, upper=1),
+  makeNumericParam("subsample", lower=0.1, upper=1),
+  makeIntegerParam("max_depth", lower=1, upper=785),
+  makeNumericParam("gamma", lower=0, upper=20)
+)
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+inner <- makeResampleDesc("Blocking")
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("Blocking")
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_svm <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/all_plots_xgboost_blocking.rda")
+```

### 2018-07-11 02_scripts/04_plot_characteristics.Rmd
11/07/2018

@@
 ```{r, results='hide'}
-devtools::install_github("tidyverse/ggplot2")
 devtools::install_github("mlr-org/mlr@blocking")
@@
 data %<>%
-  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(559, 382, 301, 499))))
+  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), 
+                           c(479, 451, 291, 529))))
 ```
@@
   group_by(plot) %>%
-  summarise(mean(deftot))
+  summarise(mean(defoliation))
 ```
@@
   group_by(plot) %>%
-  summarise((sd(deftot) / mean(deftot)) * 100)
+  summarise((sd(defoliation) / mean(defoliation)) * 100)
 ```
@@
   group_by(plot) %>%
-  summarise(((sd(deftot) / mean(deftot)) * 100) / skewness(deftot))
+  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation))
 ```
@@
   group_by(plot) %>%
-  summarise(sd(deftot))
+  summarise(sd(defoliation))
 ```
@@
   ggboxplot(
-    x = "plot", y = "deftot", color = "plot",
-    add = "jitter", add.params = list(size = "deftot")
+    x = "plot", y = "defoliation", color = "plot",
+    add = "jitter", add.params = list(size = "defoliation")
   ) +
@@
   annotate("text",
-    label = expression(bold(atop("n = 559", bar(x)~"= 53.23"))), x = 1, 
+    label = expression(bold(atop("n = 479", bar(x)~"= 55.96"))), x = 1, 
     y = 112, size = 3, colour = "#BC3C29", fontface = 2
@@
   annotate("text",
-    label = expression(bold(atop("n = 382", bar(x)~"= 17.73"))), x = 2, 
+    label = expression(bold(atop("n = 451", bar(x)~"= 13.54"))), x = 2, 
     y = 112, size = 3, colour = "#0072B5", fontface = 2
@@
   annotate("text",
-    label = expression(bold(atop("n = 301", bar(x)~"= 68.36"))), x = 3, 
+    label = expression(bold(atop("n = 291", bar(x)~"= 68.44"))), x = 3, 
     y = 112, size = 3, colour = "#E18727", fontface = 2
@@
   annotate("text",
-    label = expression(bold(atop("n = 499", bar(x)~"= 44.64"))), x = 4, 
+    label = expression(bold(atop("n = 529", bar(x)~"= 69.22"))), x = 4, 
     y = 112, size = 3, colour = "#20854E", fontface = 2
@@
 dist <- map(plots, ~ {
-  coords_laukiz1 <- coords %>%
-    mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(559, 382, 301, 499)))) %>%
+  coords <- coords %>%
+    mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), 
+                             c(479, 451, 291, 529)))) %>%
     filter(plot == .x) %>%
@@
 
-  laukiz1_points <- SpatialPoints(coords = coords_laukiz1, proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
+  points <- SpatialPoints(coords = coords, 
+                          proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
 
-  distance <- mean(as.dist(pointDistance(laukiz1_points, allpairs = TRUE, lonlat = FALSE)))
+  distance <- mean(as.dist(pointDistance(points, allpairs = TRUE, lonlat = FALSE)))
 })
@@
 
+## RMSE vs coef of variation / skewness
+
+```{r}
+rmse_defol_cov_sk <- ggplot(df, aes(x = sd_skewness, y = rmse)) +
+  geom_point(aes(color = plot), shape = 19, size = 3) +
+  labs(x = "CV (%) / Skewness", y = "") +
+  theme_ipsum_rc(
+    axis_title_size = 22,
+    axis_title_just = "c"
+  ) +
+  scale_color_nejm() +
+  theme(
+    plot.margin = unit(c(6, 0, 6, 0), "pt"),
+    legend.position = "none",
+    axis.text = element_text(size = 20),
+    legend.text = element_text(size = 20),
+    axis.title.y = element_text(margin = margin(
+      t = 0, r = 15,
+      b = 0, l = 0
+    )),
+    axis.title.x = element_text(margin = margin(
+      t = 15, r = 0,
+      b = 0, l = 0
+    ))
+  )
+```
+
 ## RMSE vs coef of variation
@@
 ```{r}
-rmse_defol <- ggplot(df, aes(x = sd_skewness, y = rmse)) +
+rmse_defol_cov <- ggplot(df, aes(x = cov_defol, y = rmse)) +
   geom_point(aes(color = plot), shape = 19, size = 3) +
-  labs(x = "Coefficient of variation (%) / Skewness", y = "RMSE") +
+  labs(x = "CV (%)", y = "") +
   theme_ipsum_rc(
@@
 ```{r}
+prow = plot_grid(rmse_dens, rmse_defol_cov, rmse_defol_cov_sk,  align = 'vh', nrow = 1)
+plots_grid <- plot_grid(prow, legend, rel_widths = c(3, .3))
+```
+
+```{r}
 plots_grid <- plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
@@
 ```{r}
-save_plot(here("03_figures/results/mean-defol_mean-point-dens.png"), plots_grid, ncol = 3)
+save_plot(here("03_figures/results/plot-characteristics.png"), plots_grid, ncol = 4)
 ```

### 2018-07-11 02_scripts/06_important_bands.Rmd
11/07/2018

---
title: "06-Important bands"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(ggpubr)
library(dplyr)
library(stringr)
library(purrr)
library(glue)
library(cowplot)
```

Read data

```{r}
ls = map(c("laukiz1", "laukiz2", "luiando", "oiartzun", "all_plots"), 
         ~ readRDS(glue("/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_{name}.rda", name = .x)))
```

#### Important spectral regions

Remove the "bf2" information and filter out NRI only.

```{r}
ls_inds = map(ls, ~ 
                .x %<>% 
                mutate(index = str_sub(index, 9, 17)) %>% 
                filter(str_detect(index, "b"))
)
```

Bring all plots on the same scale (0 -1)

```{r}
ls_inds %<>% 
  map(~ .x %<>% 
        mutate(coef = coef / max(coef)))
```

Extract all band numbers.

```{r}
ls_all_bands_rr =
  map(ls_inds, ~ .x %<>% 
        dplyr::select(index) %>% 
        str_match_all("[0-9]+") %>% 
        flatten() %>% 
        as.numeric())
```

Ok, bands 1-4 did not make it into the data (NA problems).

```{r}
map(ls_all_bands_rr, ~ .x %>% as.factor() %>% levels(.))
```

Create a tibble with the coefficients and the band numbers.

```{r}
ls_tibble_most_imp_all = map2(ls_inds, ls_all_bands_rr, ~
                            tibble(band = .y, weight = rep(.x$coef, times = 1, each = 2))
)
```

Create a tibble with the coefficients and the band numbers.

```{r}
ls_tibble_most_imp_sum = map2(ls_inds, ls_all_bands_rr, ~
                            tibble(band = .y, weight = rep(.x$coef, times = 1, each = 2)) %>% 
                            group_by(band) %>% 
                            summarise(score = sum(weight)) %>% 
                            mutate(band = band) %>% 
                            arrange(desc(score))
)
```

Classify the bands into spectral regions

1 - 60: visible (404 - 680)
61 - 71: red edge (680 - 730)
72 - 126: NIR (730 - 996)

```{r}
ls_tibble_most_imp_sum %<>% 
  map(~ .x %<>% 
        mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 nm - 680 nm)", 
                                           (band >= 61 & band <= 71) ~ "Red Edge (680 nm - 730 nm)",
                                           band >= 72 ~ "NIR (730 nm  - 996 nm)"))) %>% 
        arrange(desc(score))
  ) 
```

```{r}
ls_barplot = map2(ls_tibble_most_imp_sum, c("Laukiz 1", 
                                           "Laukiz 2", 
                                           "Luiando", 
                                           "Oiartzun", 
                                           "All Plots"), ~ 
                       ggbarplot(.x, x = "band", y = "score",                              # Color by groups
                                 add = "segments",                             # Add segments from y = 0 to dots
                                 fill = "group",
                                 color = "group",
                                 palette = "nejm",
                                 group = "band",
                                 title = .y,                              # Rotate vertically
                                 dot.size = 0.5,    
                                 width = 0.5,
                                 ylim = c(0, 50),# Large dot size
                                 ggtheme = theme_pubr(base_size = 4, margin = TRUE,
                                                      legend = "none")) + 
                       labs(y = "Score", color = "Spectral region") + 
                       scale_x_discrete(breaks = seq(0, 120, 20),
                                          labels = seq(0, 120, 20))
)
```

```{r}
legend_bar = ggbarplot(ls_tibble_most_imp_sum[[1]], x = "band", y = "score",                              # Color by groups
                                 add = "segments",                             # Add segments from y = 0 to dots
                                 fill = "group",
                                 palette = "nejm",
                                 group = "band",                            # Rotate vertically                    
                                 ggtheme = theme_pubr(base_size = 5, margin = TRUE,
                                                      legend = "right")) + 
                       labs(y = "Score", fill = "Spectral region")
```

## Barplot

```{r}
legend_plot_bar <- get_legend(legend_bar)
ls_barplot[[6]] = legend_plot_bar
plot_gridded_bar = plot_grid(plotlist = ls_barplot, ncol = 3)
save_plot(plot_gridded_bar,
          filename = here::here("03_figures/results/important-bands_barplot.png"))
```

## Boxplot

Classify the bands into spectral regions

1 - 60: visible (404 - 680)
61 - 71: red edge (680 - 730)
72 - 126: NIR (730 - 996)

```{r}
ls_tibble_most_imp_all %<>% 
  map(~ .x %<>% 
        mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 nm - 680 nm)", 
                                           (band >= 61 & band <= 71) ~ "Red Edge (680 nm - 730 nm)",
                                           band >= 72 ~ "NIR (730 nm  - 996 nm)"))) %>% 
        arrange(desc(weight))
  ) 
```

```{r}
ls_boxplot = map2(ls_tibble_most_imp_all, c("Laukiz 1", 
                                           "Laukiz 2", 
                                           "Luiando", 
                                           "Oiartzun", 
                                           "All Plots"), ~ 
                       ggboxplot(.x, x = "band", y = "weight",                              # Color by groups
                                 #add = "segments",                             # Add segments from y = 0 to dots
                                 #fill = "group",
                                 color = "group",
                                 palette = "nejm",
                                 group = "band",
                                 title = .y,                              # Rotate vertically
                                 width = 0.4,
                                 size = 0.1,
                                 outlier.size = 0.01,
                                 #size = 0.1, # Large dot size
                                 ggtheme = theme_pubr(base_size = 4, margin = TRUE,
                                                      legend = "none")) + 
                       labs(y = "Score", color = "Spectral region") + 
                       scale_x_discrete(breaks = seq(5, 125, 20),
                                          labels = seq(5, 125, 20))
)
```

```{r}
legend_box = ggboxplot(ls_tibble_most_imp_all[[1]], x = "band", y = "weight",                              # Color by groups
                                 #add = "segments",                             # Add segments from y = 0 to dots
                                 color = "group",
                                 palette = "nejm",
                                 group = "band",                            # Rotate vertically                    
                                 ggtheme = theme_pubr(base_size = 5, margin = TRUE,
                                                      legend = "right")) + 
                       labs(y = "Score", color = "Spectral region")
```

```{r}
legend_plot_box <- get_legend(legend_box)
ls_boxplot[[6]] = legend_plot_box
plot_gridded_box = plot_grid(plotlist = ls_boxplot, ncol = 3)
save_plot(plot_gridded_box, base_width = 12,
          filename = here::here("03_figures/results/important-bands_boxplot.png"))
```

## Top 50 indices only

Works only if the data arriving here was subsetted before, i.e. contains only the top XX bands.
Subsetting happens in the respective plot files, e.g. 03-modeling-laukiz1.Rmd.

Classify the bands into spectral regions

1 - 60: visible (404 - 680)
61 - 71: red edge (680 - 730)
72 - 126: NIR (730 - 996)

```{r}
ls_tibble_most_imp %<>% 
  map(~ .x %<>% 
        mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 nm - 680 nm)", 
                                           (band >= 61 & band <= 71) ~ "Red Edge (680 nm - 730 nm)",
                                           band >= 72 ~ "NIR (730 nm  - 996 nm)"))) %>% 
        arrange(desc(score)) %>% 
        mutate(band = as.factor(band))
  )
```

```{r}
ls_dotplots = map2(ls_tibble_most_imp, c("Laukiz 1 (bands = 56)", 
                                         "Laukiz 2 (bands = 41)", 
                                         "Luiando (bands = 61)", 
                                         "Oiartzun (bands = 79)", 
                                         "All Plots (bands = 76)"), ~ 
                     ggdotchart(.x, x = "band", y = "score",                              # Color by groups
                                sorting = "descending",                       # Sort value in descending order
                                add = "segments",                             # Add segments from y = 0 to dots
                                color = "group",
                                palette = "nejm",
                                position = position_dodge(1),
                                group = "band",
                                title = .y,
                                rotate = FALSE,                                # Rotate vertically
                                dot.size = 0.5,                                 # Large dot size
                                ggtheme = theme_pubr(base_size = 4, margin = TRUE,
                                                     legend = "none")) + 
                     labs(y = "Score", color = "Spectral region")
)

# adjust colors for Laukiz2 because "Red Edge" level is missing and hence the colors are different
ls_dotplots[[2]] = ls_dotplots[[2]] + 
  scale_color_manual(values = c("#619CFF", "#F8766D"))
```

```{r}
legend = ggdotchart(ls_tibble_most_imp[[1]], x = "band", y = "score",                              # Color by groups
                    sorting = "descending",                       # Sort value in descending order
                    add = "segments",                             # Add segments from y = 0 to dots
                    color = "group",
                    palette = "nejm",
                    group = "score",
                    rotate = TRUE,                                # Rotate vertically
                    dot.size = 0.5,                                 # Large dot size
                    #label = round(tibble_most_imp$sum),                        # Add mpg values as dot labels
                    ggtheme = theme_pubr(base_size = 3, margin = FALSE,
                                         legend = "right")) + 
  labs(y = "Score", color = "Spectral region")
```

```{r}
legend_plot <- get_legend(legend)
ls_dotplots[[6]] = legend_plot
plot_gridded = plot_grid(plotlist = ls_dotplots, ncol = 3)
save_plot(plot_gridded,
          filename = here::here("03_figures/results/important-bands.png"))
```

## Boxplot

```{r}
as_tibble(data.table::rbindlist(ls_tibble_most_imp)) %>% 
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun", "All Plots"), 
                           c(56, 41, 61, 79, 76)))) %>% 
  mutate(band = as.numeric(as.character(band))) %>% 
  group_by(plot) %>%
  ggboxplot(
    x = "plot", y = "band", color = "plot",
    add = "jitter")
```

```{r}
as_tibble(data.table::rbindlist(ls_tibble_most_imp)) %>% 
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun", "All Plots"), 
                           c(56, 41, 61, 79, 76)))) %>% 
  ggdotchart(x = "band", y = "score",                              # Color by groups
             sorting = "descending",                       # Sort value in descending order
             add = "segments",                             # Add segments from y = 0 to dots
             color = "group",
             palette = "nejm",
             group = "band",
             facet.by = "plot",
             rotate = F,                                # Rotate vertically
             dot.size = 0.5,                                 # Large dot size
             #label = round(tibble_most_imp$sum),                        # Add mpg values as dot labels
             ggtheme = theme_pubr(base_size = 4, margin = FALSE,
                                  legend = "right")) + 
  labs(y = "Score", color = "Spectral region")
```

## Histogram

```{r}
as_tibble(data.table::rbindlist(ls_tibble_most_imp)) %>% 
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun", "All Plots"), 
                           c(56, 41, 61, 79, 76)))) %>% 
  mutate(band = as.numeric(as.character(band))) %>% 
  group_by(plot) %>%
  ggbarplot(
    x = "band", y = "score", color = "plot")
```

# PNG to PDF conversion

Converts all pngs in the specified folder to pdfs. 
`&&` makes the "convert" call conditional on `cd`.
`exec` reduces memory waste of shell.
Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.

```{r png to pdf, echo = FALSE, eval = FALSE}
# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
#        parallel convert '{}' '{.}.pdf'")

system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

# copy to presentation folder
# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
#           overwrite = TRUE)
```

# Session Info

```{r}
sessionInfo()
```

### 2018-07-11 02_scripts/07_benchmark_all_plots.Rmd
11/07/2018

---
title: "03-modeling"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

```{r, results='hide'}
# devtools::install_github("tidyverse/ggplot2")
devtools::install_github("mlr-org/mlr@blocking")

library(mlr)
library(mlrMBO)
library(glue)
library(sf)
library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
library(pbmcapply)
```

# Read data

Variables per plot:

Lauki1: 7594
Laukiz2: 7594
Oiartzun: 7841
Luiando: 7471

Why is there such a huge difference in the number of indices without NAs for the plots?

```{r}
data <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
```

# Merge into one dataset

```{r}
data <- as_tibble(rbindlist(data, fill = TRUE))
data <- Filter(function(x) !any(is.na(x)), data)
coords <- as_tibble(rbindlist(coords))
```

Save the indices shared by all plots so that we can subset the single plots to use the same indices

```{r}
indices_shared <- names(data)
saveRDS(indices_shared, "/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda")
```

# Create task

```{r}
task <- makeRegrTask(
  id = "all_plots", data = data,
  target = "defoliation", coordinates = coords,
  blocking = factor(rep(1:4, c(479, 451, 291, 529)))
)
```

# Benchmark

## Tune strategy

```{r}
ctrl <- makeMBOControl(propose.points = 1L)
ctrl <- setMBOControlTermination(ctrl, iters = 70L)
ctrl <- setMBOControlInfill(ctrl, crit = crit.ei)
ctrl <- setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)

inner <- makeResampleDesc("Blocking")
```

## Run

```{r}
# SVM
ps_svm <- makeParamSet(
  makeNumericParam("C",
                   lower = -15, upper = 15,
                   trafo = function(x) 2^x
  ),
  makeNumericParam("sigma",
                   lower = -15, upper = 15,
                   trafo = function(x) 2^x
  )
)
tune.ctrl <- makeTuneControlMBO(
  mbo.control = ctrl,
  mbo.design = generateDesign(n = 30, par.set = ps_svm)
)
tune_wrapper_svm <- makeTuneWrapper(makeLearner("regr.ksvm", kernel = "rbfdot"),
                                    resampling = inner, par.set = ps_svm,
                                    control = tune.ctrl, show.info = TRUE,
                                    measures = list(rmse)
)

# RIDGE
ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 35, upper = 3559))
tune.ctrl <- makeTuneControlMBO(
  mbo.control = ctrl,
  mbo.design = generateDesign(n = 30, par.set = ps_glmnet)
)
tune_wrapper_glmnet <- makeTuneWrapper(makeLearner("regr.glmnet",
                                                   alpha = 0,
                                                   standardize = FALSE,
                                                   intercept = FALSE),
                                       resampling = inner, par.set = ps_glmnet,
                                       control = tune.ctrl, show.info = TRUE,
                                       measures = list(rmse)
)

# XGBOOST
ps_xgboost <- makeParamSet(
  makeDiscreteParam("nrounds", values = c(10, 100, 1000, 10000)),
  makeNumericParam("colsample_bytree", lower = 0.1, upper = 1),
  makeNumericParam("subsample", lower = 0.1, upper = 1),
  makeIntegerParam("max_depth", lower = 1, upper = 785),
  makeNumericParam("gamma", lower = 0, upper = 20)
)
tune.ctrl <- makeTuneControlMBO(
  mbo.control = ctrl,
  mbo.design = generateDesign(n = 30, par.set = ps_xgboost)
)
tune_wrapper_xgboost <- makeTuneWrapper(makeLearner("regr.xgboost",
                                                    par.vals = list(
                                                      objective = "reg:linear",
                                                      eval_metric = "error",
                                                      eta = 0.1
                                                    )
),
resampling = inner, par.set = ps_xgboost,
control = tune.ctrl, show.info = TRUE,
measures = list(rmse)
)

# RF
ps_rf <- makeParamSet(
  makeIntegerParam("mtry", lower = 1, upper = 7000),
  makeIntegerParam("min.node.size", lower = 1, upper = 10),
  makeNumericParam("sample.fraction", lower = 0.2, upper = 0.9)
)
tune.ctrl <- makeTuneControlMBO(
  mbo.control = ctrl,
  mbo.design = generateDesign(n = 30, par.set = ps_rf)
)
tune_wrapper_rf <- makeTuneWrapper(makeLearner("classif.ranger"),
                                   resampling = inner, par.set = ps_rf,
                                   control = tune.ctrl, show.info = TRUE,
                                   measures = list(rmse)
)

lrns <- list(
  tune_wrapper_svm, tune_wrapper_glmnet,
  tune_wrapper_xgboost, tune_wrapper_rf
)

outer <- makeResampleDesc("Blocking")

library(parallelMap)
parallelStart(
  mode = "multicore", level = "mlrMBO.feval", cpus = 30,
  mc.set.seed = TRUE
) # only parallelize the tuning

set.seed(12345)
bmr <- benchmark(lrns, task, outer, measures = rmse, show.info = TRUE)
parallelStop()
saveRDS(bmr, "/data/patrick/mod/hyperspectral/spcv/all_plots_benchmark.rda")
```

### 2018-07-11 90_funs.R

@@ -1,0 +1,74 @@
+##### my.pairs env
+the.col.hist <- "cyan"
+
+
+panel.hist <- function(x, col.hist=the.col.hist,
+                       cex.cor=NULL,digits=NULL,col=NULL,pch=NULL,cex=NULL,...)
+{
+  dummy <- is.character(cex.cor) | is.numeric(digits) | is.character(cex) | is.character(pch) | is.character(col)
+  usr <- par("usr"); on.exit(par(usr))
+  par(usr = c(usr[1:2], 0, 1.5) )
+  h <- hist(x, plot = FALSE)
+  breaks <- h$breaks; nB <- length(breaks)
+  y <- h$counts; y <- y/max(y)
+  rect(breaks[-nB], 0, breaks[-1], y, col=col.hist, ...)
+}
+
+panel.cor <- function(x, y, digits=2, prefix="", cex.cor=NULL,
+                      col.hist=NULL,
+                      col=NULL,pch=NULL,cex=NULL,...)
+{
+  dummy <- is.character(col.hist) | is.character(cex) | is.character(pch) | is.character(col)
+  usr <- par("usr"); on.exit(par(usr))
+  par(usr = c(0, 1, 0, 1))
+  r <- abs(cor(x, y, use="pairwise.complete.obs"))
+  txt <- format(c(r, 0.123456789), digits=digits)[1]
+  txt <- paste(prefix, txt, sep="")
+  if (is.null(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+  #text(0.5, 0.5, txt, cex = cex.cor * r)
+  text(0.5, 0.5, txt, cex = cex.cor)
+}
+
+my_pairs <- function( x, ... )
+{
+  pairs( x, ...,
+         upper.panel=panel.smooth, lower.panel=panel.cor, diag.panel=panel.hist )
+}
+
+age_imputation <- function(x, y, age, print = TRUE) {
+  wh <- which(is.na(age))
+  N.na <- length(wh)
+  set.seed(1234)
+  for (i in wh) {
+    di <- sqrt((x - x[i])^2 + (y - y[i])^2)
+    wh.nn <- which(di < 50 & !is.na(age))
+    if (length(wh.nn) < 5)
+      wh.nn <- which(di < 100 & !is.na(age))
+    if (length(wh.nn) < 5)
+      wh.nn <- which(di < 200 & !is.na(age))
+    if (length(wh.nn) < 5)
[…]

### 2018-07-13 01_writing/Schratz_hyperspectral.tex
- start intro - update results - update data and sutdy area

@@
 	\begin{keyword}
-		hyperspectral imagery \sep penalized regression \sep machine-learning \sep variable importance \sep spatial cross-validation
+		hyperspectral imagery \sep forest health \sep machine-learning \sep variable importance \sep model comparison
 	\end{keyword}
@@
 	%\setlength{\itemsep}{-\parsep}
+	\acro{AGB}{Above-Ground Biomass}
 	\acro{ANN}{Artificial Neural Network}
@@
 	\acro{MEM}{Maximum Entropy Model}
-	\acro{NRI}{Normalized Ratio Index}
+	\acro{NRI}{Normalized Ratio Indeces}
 	\acro{OLS}{Ordinary Least Squares}
@@
 	\acro{SDM}{Species Distribution Modeling}
+	\acro{SMBO}{Sequential-based Model Optimization}
 	\acro{SVM}{Support Vector Machines}
@@
 
+% Explain how remote sensing is used in forestry (potential to map forest health)
+Remote sensing data is successfully used in forestry to monitor temporal changes across large areas REFERENCES.
+The use of SAR techniques enables scientists to estimate \ac{AGB}.
+Forest health is commonly modelled using optical data from multispectral satellites by applying temporal change detections.
+With a well-trained model, scientists are able to predict the modelled response to a large area, giving valuable information about the environmental condition of this region.
+To model forest health, usually vegetation indices are derived from the band values, i.e. band combinations that are proven to be sensitive to health changes of vegetation.
+However, the pool of possible indices is often limited due to a low spectral resolution of freely available data from multispectral sensors.
+Also, if the spatial resolution of the data is too coarse (e.g. > 5 m), the value of a pixel contains information of multiple trees and possibly even bare-ground information.
+These problems introduce bias into the data which is passed onto the fitted model, leading to non-optimal modeling results.
+
+% Why are we mapping defoliation?
+
+In this study we will use hyperspectral remote sensing data with a spatial resolution of one meter and 126 spectral bands to model defoliation of \textit{Pinus radiata} trees in northern Spain. 
+These trees suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Armillaria} or \textit{Heterobasidion} leading to a spread of cankers or defoliation before the tree dies.
+The fungis are assumed to infect the trees through open wounds, possibly caused by hail damage \citep{Iturritxa2014}.
+The dieback of these trees, which are mainly used as timber, causes high economic damages in the Iberian Peninsula area.
+Remote sensing data in combination with state-of-the-art machine-learning techniques can help to constantly monitor forest health in this region using proxies such as defoliation of the tree crown.
+With this information, an increase in defoliation can serve as an early-warning indicator of a possible infection of a plot.
+
+% Show importance of vegetation indices and NRI
+
+To extract the most information from the available remote sensing data, we not only calculated the most often used vegetation indices (e.g. NDVI) to link against defoliation. 
+The narrow bands of the given data enable the calculation of up to 90 vegetation indices.
+Additionally, all possible combinations of \ac{NRI} were calculated from the data and supplied to the machine-learning algorithms as predictors.
+With their ability to find patterns between the response and its predictors in the data, we hope to increase the predictive accuracy of defoliation based on remote sensing indices.
+
+% Mention the ability of ML algs to handle highly correlated variables
+
+In this study the following objectives are addressed:
+
+\begin{itemize}
+	\item Checking multiple machine-learning algorithms on their ability to model defoliation of \textit{Pinus radiata} trees
+	\item Exploration of the most important indices of the model
+	\item Prediction of defoliation to plots with unknown levels of defoliation
+\end{itemize}
+
 \section{Data and study area}
@@
 
-The four \textit{Pinus radiata} plots Laukiz 1, Laukiz 2, Luiando and Oiartzun are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-Oiartzun has the most observations (n = 529) while Laukiz 2 has largest area size.
-All plots besides Luiando are located nearby the coast.
-The data was collected in September 2016.
+The \textit{Pinus radiata} plots of this study, named \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
+\textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} has the largest area size (1.44 ha).
+All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
+The data collection took place in September 2016.
 
@@
 All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by \ac{ICGC}.
+The first four bands were corrupted, leaving 122 bands with valid information.
 
-Additional information is provided in Table 1:
+Additional data characteristics are provided in Table 1:
 
@@
 
+For all analysis steps we used the open-source statistical programming language R \citep{R_core}.
+The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chenXGBoostScalableTree2016}, \textit{kernlab} \citep{kernlab} (Support Vector Machine), \cite{Vapnik1998}) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
+We used the R package \textit{mlr} for all modeling related steps.
+It provides a standardized interface for a wide variety of statistical and machine-learning models in R simplifying essential modeling tasks such as hyperparameter tuning, model performance evaluation and parallelization \citep{bischlMlrMachineLearning2016}.
+
 \subsection{Derivation of indices}
@@
 % link to PDF with veg indices
-All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data and offered by the \texttt{hsdar} package have been calculated.
+All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that were available in the R package \textit{hsdar} have been calculated.
 Additionally, all possible \ac{NRI} were calculated from the data using the formula:
@@
 %To account for geometric offsets, we calculated every index five times using a buffer from 1 - 5 meter around the centroid of the respective tree.
-To account for geometric offsets, we used a buffer of 2 meters around the centroid of the respective tree.
+To account for geometric offsets, we used a buffer of two meters around the centroid of the respective tree.
 The mean value of all pixels touched by the buffer was assigned as the final value for each index.
@@
 In total, 7875 Normalized Ratio Indices NRI have been calculated ($\frac{125*126}{2}$).
-Some indices returned \texttt{NA} values for some observations and were removed from the dataset, leaving a total of 7471 variables without missing values.
+Due to four corrupted bands and some other numerical problems, few indices returned \texttt{NA} values for some observations.
+These indices were removed from the dataset, leaving a total of 7471 variables without missing values.
 
-\subsection{Exploratory plot characteristics}
+\subsection{Exploratory analysis of plot characteristics}
 
-To get a better impression on the infection state of the four plots, an exploratory data analysis was done.
-The distribution of the response variable \texttt{defoliation} was visualized as well as the spectral signatures of each plot.
+Plot characteristics like age, stand density and defoliation were analysed to show differences among the plots.
+Additionally, the spectral signatures of each plot have been visualized.
 
@@
 Multiple algorithms were benchmarked on predictive performance to find the best performing one.
-Besides well-known machine-learning algorithms like \ac{RF} and \ac{SVM} we also used \textit{xgboost} due to its promising results in recent machine-learning competitions.
-To expand the algorithm selection even more, we also used penalized L2 (Ridge) regression.
-This algorithm is able to account for the high correlation between the predictors by penalizing the estimated coefficients of the fitted model.
+Besides the well-known \ac{SVM} \citep{Vapnik1998} we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
+We also added penalized L2 (Ridge) regression to the algorithm collection due to its ability to handle highly correlated covariates.
+We also want to explain why we did not consider the probably most popular machine-learning algorithm, Random Forest: Due to the high number of variables model fitting times in the range hours for a single model fit were not practicable for this work.
+These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{Probst2018b}.
 
-\subsubsection{The ridge penalty}
+% \subsubsection{The ridge penalty}
 
-In \ac{RR} (also called $\ell_{2}$ penalization) the assumption of unbiased coefficients is given up in favor of higher predictive accuracy and reduced variance \citep{Hastie2001}.
-Coefficients are standardized and penalized for their size.
-When minimizing the \ac{RSS}, \ac{RR} adds a penalization term $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ to the equation:
+% In \ac{RR} (also called $\ell_{2}$ penalization) the assumption of unbiased coefficients is given up in favor of higher predictive accuracy and reduced variance \citep{Hastie2001}.
+% Coefficients are standardized and penalized for their size.
+% When minimizing the \ac{RSS}, \ac{RR} adds a penalization term $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ to the equation:
 
-\begin{equation}
-	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p}\beta_{j}^{2}
-\end{equation}
+% \begin{equation}
+% 	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p}\beta_{j}^{2}
+% \end{equation}
 
-where $\lambda >= 0$ is a tuning parameter responsible for the magnitude of penalization.
-To make the second term (usually referred to as the \textit{shrinkage penalty}) of this equation small, the coefficients $\beta_{j}$ need to become small.
-Unlike to Lasso however, predictors are not removed from the final model and will always be $\beta_{j} >= 0$.
-Hence, $\lambda$ has the effect of shrinking the coefficients when minimizing the \ac{RSS}.
-For $\lambda = 0$, no penalization is done and standard \ac{OLS} applies \citep{James2013}.
-The \textit{shrinkage penalty} is only applied to the coefficients and not to the intercept.
-Also, while \ac{OLS} generates only one set of coefficient estimates, \ac{RR} will create multiple sets for every value of $\lambda$.
-In summary, the advantage of \ac{RR} is based on the bias-variance tradeoff: For $\lambda\to\infty$, the flexibility of the fit is reduced leading to a decrease in variance of the coefficients but also introduces a (substantial) bias.
-For $\lambda = 0$, the variance is high but coefficients are unbiased \citep{James2013}.
+% where $\lambda >= 0$ is a tuning parameter responsible for the magnitude of penalization.
+% To make the second term (usually referred to as the \textit{shrinkage penalty}) of this equation small, the coefficients $\beta_{j}$ need to become small.
+% Unlike to Lasso however, predictors are not removed from the final model and will always be $\beta_{j} >= 0$.
+% Hence, $\lambda$ has the effect of shrinking the coefficients when minimizing the \ac{RSS}.
+% For $\lambda = 0$, no penalization is done and standard \ac{OLS} applies \citep{James2013}.
+% The \textit{shrinkage penalty} is only applied to the coefficients and not to the intercept.
+% Also, while \ac{OLS} generates only one set of coefficient estimates, \ac{RR} will create multiple sets for every value of $\lambda$.
+% In summary, the advantage of \ac{RR} is based on the bias-variance tradeoff: For $\lambda\to\infty$, the flexibility of the fit is reduced leading to a decrease in variance of the coefficients but also introduces a (substantial) bias.
+% For $\lambda = 0$, the variance is high but coefficients are unbiased \citep{James2013}.
 
@@
 
-The algorithms were benchmarked using a spatial \ac{CV} on the plot level.
-The dataset was split into four folds, each fold representing one plot.
-Algorithms were trained on three out of four plots and evaluated on the remaining plot.
-Hyperparameter tuning was also performend on the plot level: Each respective training set, consisting of observations from three plots, was again split into three partitions to estimate the best performing hyperparameter set.
+The algorithms were benchmarked in two ways: 
+(1) Using spatial \ac{CV} for each plot using on the k-means clustering approach of \cite{Brenning2012}. To reduce runtime we used a five-fold five-times repeated CV setup.
+(2) Using spatial \ac{CV} on the plot level with each plot being the test set once. This results in four performance estimates, one for each fold.
+For (1) we only used the best performing algorithm from (2). 
+The reason why the (2) was chosen for algorithm selection is that this model will also be used to spatially predict defoliation in other plots.
 
-% figure visualizing the folds
+% Add figure visualizing both spCV approaches
+
+\subsubsection{Hyperparameter tuning}
+To tune the hyperparameters of the algorithms, we used \ac{SMBO} via the R package \textit{mlrMBO} \citep{mlrMBO}.
+This Bayesian approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting to be evaluated next is proposed based on the setting that performed best.
+This strategy continues until a termination criterion, defined by the user, is reached \citep{Hutter2011, Jones1998}.
+In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting a total budget of 50 evaluated hyperparameter settings per fold.
+The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs such as \textit{random search} or \textit{grid search} \citep{Bergstra2012}.
 
@@
 
-We aimed to find indices that contribute most when predicting defoliation.
-Permutation-based variable importance was applied on the best performing algorithm.
+To find indices that contributed most to model performance, we used permutation-based variable importance on the best performing algorithm.
 
 % partial dependence plots??
-
-\section{Results}
-
-% Performance estimates
-\begin{table}[b!]
-\centering
-\caption[t]{Four-fold spatial \ac{CV} performances of RF, RR, SVM and xgboost using \ac{RMSE} as the error measure. 
-Mean and standard deviation are shown.}
-\begingroup\footnotesize
-\begin{tabular}{llll}
-	\\
-	RF            & RR            & SVM           & xgboost \\
-	\hline
-	54.80 (17.58) & 53.67 (17.95) & 54.66 (17.67) &         \\
-	\bottomrule
-\end{tabular}
-\endgroup
-\label{tab:penalty_comparison}
-\end{table}
 
@@
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {../03_figures/results/boxplot_defol.pdf}}
+		\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {boxplot_defol.pdf}}
 		\caption{Descriptive statistics of the response variable \textit{defoliation}.}
@@
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/spectral_signatures.pdf}}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral_signatures.pdf}}
 		\caption{Spectral signatures (mean and standard deviation) of each plot.
@@
 
-Luiando shows the highest defoliation ($\bar{x} = 68.36 \%$) among the plots while Laukiz 2 is the healthiest ($\bar{x} = 17.73 \%$) (\autoref{fig:defol_boxplots}).
-Laukiz 1 and Oiartzun both show a medium defoliation level around 50 \%.
-Oiartzun consists mainly of trees that show either a very high level of defoliation ($70 \% <=$) or none at all.
-Laukiz 1 shows an evenly distributed level of defoliation across the entire plot.
+\textit{Oiartzun} shows the highest defoliation ($\bar{x} = 69.22 \%$) among the plots while \textit{Laukiz 2} is the healthiest ($\bar{x} = 13.54 \%$) (\autoref{fig:defol_boxplots}).
+All plots besides \textit{Luiando} show an evenly distributed level of defoliation across the entire plot.
 
-The high defoliation level of Luiando and Oiartzun is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
+The high degree of defoliation of \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
 Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
 Oiartzun is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
-Laukiz 2 shows a mean tree density of 62.51 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.35 (Laukiz 1), 33.77 (Luiando), 35.02 (Oiartzun)) (\autoref{fig:rmse_defol_dens}).
+Laukiz 2 shows a mean tree density of 61.59 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.64 (Laukiz 1), 33.01 (Luiando), 34.96 (Oiartzun)) (\autoref{fig:plot-characteristics}).
 
@@
 
+% Performance estimates
+\begin{table}[b!]
+	\centering
+	\caption[t]{Four-fold spatial \ac{CV} performances of RR, SVM and xgboost using \ac{RMSE} as the error measure. 
+	Mean and standard deviation are shown.}
+	\begingroup\footnotesize
+	\begin{tabular}{llll}
+		RR            & SVM           & xgboost \\
+		\hline
+		59.10 (22.71) & 36.23 (15.73) &  &         \\
+		\bottomrule
+	\end{tabular}
+	\endgroup
+	\label{tab:penalty_comparison}
+	\end{table}
+
 \begin{table}[t!]
 \centering
-\caption[t]{Predictive performance of \ac{RR} using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.
-The values for "merged dataset" correspond to the fold for which the respective plot was serving as the test set.
-For "single plot", the values correspond to the mean value of the SpCV at the repetition level (10 folds, 20 repetitions).}
+\caption[t]{Predictive performance of xgboost using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.
+The performance estimates for "All Plots" correspond to the fold for which the respective plot was serving as the test set.
+For "single plot", the values correspond to the mean value of the SpCV at the repetition level (10 folds, 20 repetitions), scored by models fitted on the data of the respective plot only.}
 \begingroup\footnotesize
@@
 	\\
-	Plot/Data & Merged dataset (Block CV) & Single plot (SpCV) \\
+	Plot/Data & All Plots (Block CV) & Single plot (SpCV) \\
 	\hline
@@
 
-% variable importance
-\begin{table}[b!]
-\centering
-\caption[t]{The ten highest coefficient estimates for every plot and the merged dataset.}
-\begingroup\footnotesize
-\begin{tabular}{lllll}
-	\\
-	Laukiz 1           & Laukiz 2           & Luiando             & Oiartzun            & All Plots          \\
-	\hline
-	b80-b77 (0.0089)   & b89-b84 (0.0093)   & b95-b93 (1.5e-36)   & b23-b18 (-0.0090)   & b78-b77 (0.058)    \\
-	b81-b77 (0.0079)   & b89-b87 (0.0086)   & b109-b106 (1.4e-36) & b23-b19 (-0.0074)   & b115-b113 (0.054)  \\
-	b78-b77 (0.0079)   & b89-b88 (0.0086)   & b92-b95 (1.4e-36)   & b99-b98 (0.0073)    & b82-b77 (0.052)    \\
-	b77-b76 (-0.0078)  & b108-b104 (0.0084) & b114-b6 (-1.2e-36)  & b23-b20 (-0.0070)   & b79-b77 (0.049)    \\
-	b79-b77 (0.0077)   & b89-b85 (0.0083)   & b96-b93 (1.2e-36)   & b10-b8 (-0.0063)    & b80-b77 (0.049)    \\
-	Datt3 (-0.71)      & b92-b84 (0.0083)   & b116-b6 (-1.2e-36)  & b102-b98 (0.0062)   & b81-b77 (0.049)    \\
-	b41-b25 (0.0070)   & b89-b83 (0.0081)   & b114-b5 (-1.2e-36)  & b124-b115 (0.0062)  & b81-b78 (-0.048)   \\
-	b116-b113 (0.0068) & b89-b86 (0.0080)   & b115-b114 (1.2e-36) & b23-b15 (-0.0060)   & b124-b115 (-0.047) \\
-	b82-b77 (0.0068)   & b92-b88 (0.0080)   & b95-b91 (1.1e-36)   & b126-b115 (-0.0060) & b23-b20 (-0.047)   \\
-	b77-b75 (-0.0068)  & b108-b96 (0.0079)  & b114-b8 (-1.1e-36)  & b118-b115 (-0.0059) & b80-b78 (-0.046)   \\
-	\bottomrule
-\end{tabular}
-\endgroup
-\label{tab:variable-importance}
-\end{table}
-
 % band importance

### 2018-07-13 02_scripts/031-laukiz1/03-modeling_laukiz1.Rmd
- start intro - update results - update data and sutdy area

@@
 ```
+
+### xgboost
+
+```{r}
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error"
+                   ))
+```
+
+```{r}
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 0, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+inner <- makeResampleDesc("SpCV", iters = 10)
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_rf <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/laukiz1_xgboost.rda")
+```

### 2018-07-13 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
- start intro - update results - update data and sutdy area

@@
 ```
+
+### xgboost
+
+```{r}
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error"
+                   ))
+```
+
+```{r}
+ps_xgboost <- makeParamSet(
+  makeNumericParam("nrounds", lower = 0, upper = 8.64, 
+                   trafo = function (x) round(2^x * 10)),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+inner <- makeResampleDesc("SpCV", iters = 10)
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_rf <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/laukiz2_xgboost.rda")
+```

### 2018-07-13 02_scripts/033-luiando/03-modeling_luiando.Rmd
- start intro - update results - update data and sutdy area

@@
 ```
+
+### xgboost
+
+```{r}
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error"
+                   ))
+```
+
+```{r}
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 0, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+inner <- makeResampleDesc("SpCV", iters = 5)
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 5)
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlr.resample",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_rf <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/luiando_xgboost.rda")
+```

### 2018-07-13 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
- start intro - update results - update data and sutdy area

@@
 ```
+
+### xgboost
+
+```{r}
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error"
+                   ))
+```
+
+```{r}
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 0, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+inner <- makeResampleDesc("SpCV", iters = 10)
+```
+
+#### SpCV
+
+```{r}
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
+                                measures = list(rmse))
+```
+
+```{r}
+outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 5)
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 30, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_rf <- mlr::resample(tune_wrapper, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/oiartzun_xgboost.rda")
+```

### 2018-07-13 02_scripts/035_all-plots/03-modeling_all.Rmd
- start intro - update results - update data and sutdy area

@@
                      objective = "reg:linear",
-                     eval_metric = "error",
-                     eta = 0.1
+                     eval_metric = "error"
                    ))
@@
 ```{r}
-ps_xgboost = makeParamSet(
-  makeDiscreteParam("nrounds", values = c(10, 100, 1000, 10000)),
-  makeNumericParam("colsample_bytree", lower=0.1, upper=1),
-  makeNumericParam("subsample", lower=0.1, upper=1),
-  makeIntegerParam("max_depth", lower=1, upper=785),
-  makeNumericParam("gamma", lower=0, upper=20)
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 0, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
 )
@@
 ctrl = makeMBOControl(propose.points = 1L)
-ctrl = setMBOControlTermination(ctrl, iters = 70L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
 ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
@@
                                mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+# tune.ctrl = makeTuneControlRandom(maxit = 50)
 inner <- makeResampleDesc("Blocking")
@@
 library(parallelMap)
-parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+parallelStart(mode = "multicore", cpus = 4, level = "mlr.resample",
               mc.set.seed = TRUE) # only parallelize the tuning
@@
                           show.info = TRUE, measures = list(rmse))
+# resa_svm <- mlr::resample(lrn_xgboost, task,
+#                           resampling = outer, 
+#                           show.info = TRUE, measures = list(rmse))
 
@@
 ```
+
+### RF
+
+```{r}
+# RF
+ps_rf <- makeParamSet(
+  makeIntegerParam("mtry", lower = 1, upper = 3000),
+  makeIntegerParam("min.node.size", lower = 1, upper = 10),
+  makeNumericParam("sample.fraction", lower = 0.2, upper = 0.9)
+)
+ctrl <- makeMBOControl(propose.points = 1L)
+ctrl <- setMBOControlTermination(ctrl, iters = 20L)
+ctrl <- setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl <- setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl <- makeTuneControlMBO(
+  mbo.control = ctrl,
+  mbo.design = generateDesign(n = 30, par.set = ps_rf)
+)
+inner <- makeResampleDesc("Blocking")
+
+tune_wrapper_rf <- makeTuneWrapper(makeLearner("regr.ranger"),
+                                   resampling = inner, par.set = ps_rf,
+                                   control = tune.ctrl, show.info = TRUE,
+                                   measures = list(rmse)
+)
+```
+
+```{r}
+outer <- makeResampleDesc("Blocking")
+```
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
+              mc.set.seed = TRUE) # only parallelize the tuning
+set.seed(12345)
+resa_svm <- mlr::resample(tune_wrapper_rf, task,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/all_plots_rf_blocking.rda")
+```

### 2018-07-16 01_writing/Schratz_hyperspectral.tex
first draft of introduction (#4)

@@
 	\acro{MEM}{Maximum Entropy Model}
-	\acro{NRI}{Normalized Ratio Indeces}
+	\acro{NRI}{Normalized Ratio Indices}
 	\acro{OLS}{Ordinary Least Squares}
@@
 	\acro{RSS}{Residual Sum of Squares}
+	\acro{SAR}{Synthetic Aperture Radar}
 	\acro{SDM}{Species Distribution Modeling}
@@
 % Explain how remote sensing is used in forestry (potential to map forest health)
-Remote sensing data is successfully used in forestry to monitor temporal changes across large areas REFERENCES.
-The use of SAR techniques enables scientists to estimate \ac{AGB}.
-Forest health is commonly modelled using optical data from multispectral satellites by applying temporal change detections.
-With a well-trained model, scientists are able to predict the modelled response to a large area, giving valuable information about the environmental condition of this region.
-To model forest health, usually vegetation indices are derived from the band values, i.e. band combinations that are proven to be sensitive to health changes of vegetation.
-However, the pool of possible indices is often limited due to a low spectral resolution of freely available data from multispectral sensors.
-Also, if the spatial resolution of the data is too coarse (e.g. > 5 m), the value of a pixel contains information of multiple trees and possibly even bare-ground information.
-These problems introduce bias into the data which is passed onto the fitted model, leading to non-optimal modeling results.
+Data retrieved from remote sensing satellites is successfully used in forestry to monitor temporal changes across large areas \citep{martinezdelcastilloEvaluationForestCover2015,sextonModelPropagationUncertainty2015}.
+The use of \ac{SAR} techniques enables scientists to estimate \ac{AGB} \citep{luSurveyRemoteSensingbased2016, sinhaReviewRadarRemote2015}.
+Forest health is commonly assesed using optical data from multi-/hyperspectral satellites by applying temporal change detections \citep{zhangRemoteSensingSeasonal2016}.
+With the recent success story of machine-learning methods in the field of remote sensing, modeling techniques such as \ac{RF} are frequently used to model relationships of possible triggers to forest health \citep{belgiuRandomForestRemote2016, laryMachineLearningGeosciences2016, michezClassificationRiparianForest2016}.
+
+With a robust model, predictions of the modelled response to large areas is possible, giving valuable information about the condition of this variable in unknown regions.
+To model forest health, usually few variables are extracted based on the spectral signatures of affected and unaffected trees \citep{lelongEvaluationOilPalmFungal2010}. 
+However, spectral (vegetation-)indices have shown the potential to contribute valuable information to increase predictive accuracy of forest pathogens \citep{jiangSatellitederivedVegetationIndices2014, adamczykRededgeVegetationIndices2015}.
+
+However, the amount of possible (vegetation-)indices that could be calculated is often limited due to a low spectral resolution of freely available data from optical multispectral sensors (e.g.
+Sentinel-2).
+Also, there is currently no free data available from hyperspectral sensors that could be used for such studies (after the decommission of the EO-1 Hyperion satellite).
+If the spatial resolution of the data is too coarse (e.g. $> 5 m$), the value of a pixel usually contains information from multiple trees and possibly even bare-ground information.
+This makes the resulting information almost useless to be used for forest health monitoring on a tree level.
 
@@
 
-In this study we will use hyperspectral remote sensing data with a spatial resolution of one meter and 126 spectral bands to model defoliation of \textit{Pinus radiata} trees in northern Spain. 
-These trees suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Armillaria} or \textit{Heterobasidion} leading to a spread of cankers or defoliation before the tree dies.
-The fungis are assumed to infect the trees through open wounds, possibly caused by hail damage \citep{Iturritxa2014}.
-The dieback of these trees, which are mainly used as timber, causes high economic damages in the Iberian Peninsula area.
-Remote sensing data in combination with state-of-the-art machine-learning techniques can help to constantly monitor forest health in this region using proxies such as defoliation of the tree crown.
-With this information, an increase in defoliation can serve as an early-warning indicator of a possible infection of a plot.
+In this study we will use hyperspectral data with a spatial resolution of one meter and 126 spectral bands to model the health status of Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
+The trees in the study area suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum} \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation before the tree dies \citep{mesanzaNativeRhizobacteriaBiocontrol2016, iturritxaBiocontrolFusariumCircinatum2017}.
+In-situ measurements of defoliation on a tree level are used as a proxy to model tree health.
+The fungis are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
+The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{ganleyGlobalClimaticRisk2009}.
+Hyperspectral remote sensing data in combination with state-of-the-art machine-learning techniques is used to help monitoring the health status in this region by early detecting affected trees/plots.
 
@@
 
-To extract the most information from the available remote sensing data, we not only calculated the most often used vegetation indices (e.g. NDVI) to link against defoliation. 
-The narrow bands of the given data enable the calculation of up to 90 vegetation indices.
-Additionally, all possible combinations of \ac{NRI} were calculated from the data and supplied to the machine-learning algorithms as predictors.
-With their ability to find patterns between the response and its predictors in the data, we hope to increase the predictive accuracy of defoliation based on remote sensing indices.
+To extract the most information from the available remote sensing data, we not only calculated the most common vegetation indices like \textit{NDVI} to link against defoliation but all possible ones within the spectral region of the data (400 nm - 1000 nm) that were implemented in the \textit{hsdar} package in R \citep{hsdar}.
+Additionally, all possible combinations of \ac{NRI} were calculated from the data and supplied to a selection of machine-learning algorithms as predictors.
 
@@
 
-In this study the following objectives are addressed:
+Specifically the following objectives are addressed:
 
 \begin{itemize}
-	\item Checking multiple machine-learning algorithms on their ability to model defoliation of \textit{Pinus radiata} trees
-	\item Exploration of the most important indices of the model
-	\item Prediction of defoliation to plots with unknown levels of defoliation
+	\item Comparison of multiple algorithms on their performance to model defoliation of \textit{Pinus radiata} trees using highly-correlated indices
+	\item Exploration of the most important indices of the best performing model
+	\item Prediction of defoliation to \textit{Pinus radiata} plots with an unknown defoliation level
 \end{itemize}

### 2018-07-17 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
clean scripts

--- 2018-05-15 02_scripts/032-laukiz2/03-modeling_laukiz2.Rmd
@@
 Subset to indices shared by all plots.
+Created in "035_all_plots/03_modeling_all.Rmd".
 
@@
 
-## glmnet
-
-### Ridge
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting (this) is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure. 
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
+## xgboost
 
 ```{r}
-# predict.type for 'auc' (AUROC)
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          alpha = 0,
-                          standardize = FALSE,
-                          intercept = FALSE)
-                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error"
+                   ))
 ```
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 102, upper = 10228))
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
 ```
 
-#### SpCV
+```{r}
+inner <- makeResampleDesc("SpCV", iters = 5)
+outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 5)
+```
+
+### SpCV
 
 ```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
                                 measures = list(rmse))
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
-```
-
-```{r}
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
-```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_ridge_bf2_10folds.rda")
-resa_glmnet$measures.test %>% 
-  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
-  group_by(rep) %>% 
-  summarise_at("rmse", mean) %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
+parallelStart(mode = "multicore", cpus = 25, level = "mlr.resample",
               mc.set.seed = TRUE) # only parallelize the tuning
 set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 show.info = TRUE, measure = list(rmse))
-parallelStop()
-saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge.rda")
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 7690,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690)))) %>% 
-  arrange(desc(coef))
-```
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 7690)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Lasso
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          alpha = 1,
-                          standardize = FALSE,
-                          intercept = FALSE)
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10.2))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
+resa_rf <- mlr::resample(tune_wrapper, task,
                           resampling = outer, extract = getTuneResult,
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/laukiz2_xgboost.rda")
 ```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_lasso_bf2_10folds.rda")
-resa_glmnet$measures.test %>% 
-  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
-  group_by(rep) %>% 
-  summarise_at("rmse", mean) %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "norm")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Elasticnet
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          standardize = FALSE,
-                          intercept = FALSE)
-#getHyperPars(lrn_glmnet)
-#getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-For the elasticnet we also tune `alpha`.
-As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.1, upper = 10228),
-                          makeNumericParam("alpha", lower = 0, upper = 1))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
-```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_elasticnet.rda")
-resa_glmnet$measures.test %>% 
-  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
-  group_by(rep) %>% 
-  summarise_at("rmse", mean) %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```

### 2018-07-17 02_scripts/033-luiando/03-modeling_luiando.Rmd
clean scripts

--- 2018-05-16 02_scripts/033-luiando/03-modeling_luiando.Rmd
@@
 # Read data
-
-Variables per plot:
-
-Lauki1: 7594
-Laukiz2: 7594
-Oiartzun: 7891
-Luiando: 6073
-
-Why is there such a huge difference in the number of indices without NAs for the plots?
 
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 15, upper = 1558))
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, 
+                                                           par.set = ps_xgboost))
+```
+```{r}
+inner <- makeResampleDesc("SpCV", iters = 5)
+outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 5)
 ```
 
-#### SpCV
+### SpCV
 
 ```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
                                 measures = list(rmse))
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
-```
-
-```{r}
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_ridge_bf2_10folds.rda")
-```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_ridge_bf2_10folds.rda")
-resa_glmnet$measures.test %>% 
-  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
-  group_by(rep) %>% 
-  summarise_at("rmse", mean) %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
+parallelStart(mode = "multicore", cpus = 25, level = "mlr.resample",
               mc.set.seed = TRUE) # only parallelize the tuning
 set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 show.info = TRUE, measure = list(rmse))
-parallelStop()
-saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge.rda")
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1560,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560)))) %>% 
-  arrange(desc(coef))
-```
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1560)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Lasso
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          alpha = 1,
-                          standardize = FALSE,
-                          intercept = FALSE)
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.56))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
+resa_rf <- mlr::resample(tune_wrapper, task,
                           resampling = outer, extract = getTuneResult,
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_lasso_bf2_10folds.rda")
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/luiando_xgboost.rda")
 ```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_lasso_bf2_10folds.rda")
-resa_glmnet$measures.test %>% 
-  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
-  group_by(rep) %>% 
-  summarise_at("rmse", mean) %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "norm")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Elasticnet
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          standardize = FALSE,
-                          intercept = FALSE)
-#getHyperPars(lrn_glmnet)
-#getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-For the elasticnet we also tune `alpha`.
-As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1560),
-                          makeNumericParam("alpha", lower = 0, upper = 1))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 20)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/luiando_elasticnet.rda")
-```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_elasticnet.rda")
-resa_glmnet$measures.test %>% 
-  dplyr::mutate(rep = rep(1:20, times = 1, each = 10)) %>% 
-  group_by(rep) %>% 
-  summarise_at("rmse", mean) %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```

### 2018-07-17 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
clean scripts

--- 2018-05-11 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
@@
 # Read data
-
-Variables per plot:
-
-Lauki1: 7594
-Laukiz2: 7594
-Oiartzun: 7891
-oiartzun: 6073
-
-Why is there such a huge difference in the number of indices without NAs for the plots?
 
@@
 ```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 55, upper = 5525))
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
+)
 ```
@@
 ```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+```
+
+```{r}
+inner <- makeResampleDesc("SpCV", iters = 5)
+outer <- makeResampleDesc("SpRepCV", folds = 5, reps = 5)
 ```
@@
 ```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
+tune_wrapper <- makeTuneWrapper(lrn_xgboost, resampling = inner, par.set = ps_xgboost,
+                                control = tune.ctrl, show.info = TRUE,
                                 measures = list(rmse))
@@
 ```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
-```
-
-```{r}
 library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/oiartzun_ridge_bf2_10folds.rda")
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
+parallelStart(mode = "multicore", cpus = 25, level = "mlr.resample",
               mc.set.seed = TRUE) # only parallelize the tuning
 set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 show.info = TRUE, measure = list(rmse))
-parallelStop()
-saveRDS(params_tuned_glmnet, "/data/patrick/mod/tmp/tuning_hyperspectral/tuning_bf2_ridge.rda")
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 5460,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460)))) %>% 
-  arrange(desc(coef))
-```
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 5460)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Lasso
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          alpha = 1,
-                          standardize = FALSE,
-                          intercept = FALSE)
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 1.5))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
+resa_rf <- mlr::resample(tune_wrapper, task,
                           resampling = outer, extract = getTuneResult,
@@
 parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/oiartzun_lasso_bf2_10folds.rda")
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/oiartzun_xgboost.rda")
 ```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "norm")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Elasticnet
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          standardize = FALSE,
-                          intercept = FALSE)
-#getHyperPars(lrn_glmnet)
-#getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-For the elasticnet we also tune `alpha`.
-As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.01, upper = 5525),
-                          makeNumericParam("alpha", lower = 0, upper = 1))
-```
-
-```{r}
-ctrl <- makeTuneControlRandom(maxit = 200)
-inner <- makeResampleDesc("SpCV", iters = 10)
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("SpRepCV", folds = 10, reps = 2)
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/oiartzun_elasticnet.rda")
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```

### 2018-07-17 02_scripts/035_all-plots/03-modeling_all.Rmd
clean scripts

@@
 ```{r, results='hide'}
-#devtools::install_github("tidyverse/ggplot2")
 devtools::install_github("mlr-org/mlr@blocking")
@@
 
-Save the indices shared by all plots so that we can subset the single plots to use the same indices
+Save the indices shared by all plots so that we can subset the single plots to use the same indices.
 
@@
 
-## glmnet
-
-### Ridge
+## Ridge
 
@@
 ```{r}
-# predict.type for 'auc' (AUROC)
 lrn_glmnet <- makeLearner("regr.glmnet",
@@
                           intercept = FALSE)
-                          #lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
 ```
@@
 ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
-ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
 tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
-                               mbo.design = generateDesign(n = 30, par.set =  ps_glmnet))
+                               mbo.design = generateDesign(n = 30, 
+                                                           par.set =  ps_glmnet))
+```
+
+```{r}
 inner <- makeResampleDesc("Blocking")
+outer <- makeResampleDesc("Blocking")
 ```
@@
 ```{r}
-# outer <- makeResampleDesc("RepCV", folds = 4, reps = 2)
-outer <- makeResampleDesc("Blocking")
-```
-
-```{r}
 library(parallelMap)
 parallelStart(mode = "multicore", level = "mlrMBO.feval", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
+              mc.set.seed = TRUE)
 
@@
 parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
+              mc.set.seed = TRUE) 
 set.seed(12345)
@@
 
-### Lasso
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          alpha = 1,
-                          standardize = FALSE,
-                          intercept = FALSE)
-# getHyperPars(lrn_glmnet)
-# getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3.5))
-```
-
-```{r}
-ctrl = makeMBOControl(propose.points = 1L)
-ctrl = setMBOControlTermination(ctrl, iters = 70L)
-ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
-ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
-tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
-                               mbo.design = generateDesign(n = 30, par.set =  ps_glmnet))
-inner <- makeResampleDesc("SpCV", iters = 10)
-inner <- makeResampleDesc("Blocking")
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = tune.ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("Blocking")
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.resample", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds_blocking.rda")
-```
-
-Calculate repetition mean
-
-```{r}
-resa_glmnet = readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_lasso_bf2_10folds_blocking.rda")
-resa_glmnet$measures.test %>% 
-  summarise(mean(rmse), sd(rmse))
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = tune.ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "norm")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### Elasticnet
-
-We only need to tune `s`: https://github.com/mlr-org/mlr/issues/1030#issuecomment-233677172
-`glmnet` fits many models during training. However, important is which `s` is chosen for prediction. 
-The value of `s` determines which trained model will be chosen.
-If `s` != `lambda`, linear interpolation or refitting is applied.
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda` values.
-
-Besides tuning with `mlr`, we could use `cv-glmnet()` to find a suitable `lambda` value for a given dataset.
-However, we would do a random partitioning instead of a spatial partitioning here.
-
-Coefficients of ridge regression should not be used for inference as they are biased. 
-They can be seen as a variable importance measure then?
-The highest coefficients are most important for the model (if all variables are on the same scale) https://stats.stackexchange.com/a/154712/101464
-
-```{r}
-lrn_glmnet <- makeLearner("regr.glmnet",
-                          standardize = FALSE,
-                          intercept = FALSE)
-#getHyperPars(lrn_glmnet)
-#getParamSet(lrn_glmnet)
-```
-
-To find a suitable tuning space for `s`, Julia recommends to do a simple `train()` call and check the range of the chosen `lambda`.
-
-```{r}
-glmnet_train = mlr::train(lrn_glmnet, task)
-summary(glmnet_train$learner.model$lambda)
-```
-
-For the elasticnet we also tune `alpha`.
-As we have both penalties, we just use the lower and uppper maxima from both Lasso and Ridge.
-
-```{r}
-ps_glmnet <- makeParamSet(makeNumericParam("s", lower = 0.03, upper = 3559),
-                          makeNumericParam("alpha", lower = 0, upper = 1))
-```
-
-```{r}
-ctrl = makeMBOControl(propose.points = 1L)
-ctrl = setMBOControlTermination(ctrl, iters = 70L)
-ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
-ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
-tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
-                               mbo.design = generateDesign(n = 30, par.set = ps_glmnet))
-inner <- makeResampleDesc("Blocking")
-```
-
-#### SpCV
-
-```{r}
-tune_wrapper <- makeTuneWrapper(lrn_glmnet, resampling = inner, par.set = ps_glmnet,
-                                control = tune.ctrl, show.info = TRUE,
-                                measures = list(rmse))
-```
-
-```{r}
-outer <- makeResampleDesc("Blocking")
-```
-
-```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.resample", cpus = 40,
-              mc.set.seed = TRUE) # only parallelize the tuning
-
-set.seed(12345)
-resa_glmnet <- mlr::resample(tune_wrapper, task,
-                             resampling = outer, extract = getTuneResult,
-                             show.info = TRUE, measures = list(rmse))
-
-parallelStop()
-saveRDS(resa_glmnet, "/data/patrick/mod/hyperspectral/spcv/all_plots_elasticnet_blocking.rda")
-```
-
-#### TuneParams
-
-```{r}
-configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
-library(parallelMap)
-parallelStart(mode = "multicore", level = "mlr.tuneParams", cpus = 48,
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-params_tuned_glmnet = tuneParams(lrn_glmnet, task = task, resampling = inner,
-                                 par.set = ps_glmnet, control = tune.ctrl, 
-                                 measure = list(rmse))
-parallelStop()
-```
-
-#### Tuned results
-
-We use the `s` value from the tuning. 
-As we will always use the complete dataset, this should be the best value.
-
-```{r}
-lrn_glmnet_tuned <- makeLearner("regr.glmnet",
-                                alpha = 0,
-                                s = 1810,
-                                standardize = FALSE, 
-                                intercept = FALSE)
-#lambda = sort(seq(0, 5, length.out = 100), decreasing = T))
-glmnet_train_tuned = mlr::train(lrn_glmnet_tuned, task)
-```
-
-Inspect the highest absolute coefficients (positive and negative).
-The first entry is the intercept and is therefore left out.
-
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(desc(coef))
-```
-```{r}
-tibble(index = rownames(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810))), 
-       coef = as.vector(as.matrix(coef(glmnet_train_tuned$learner.model, s = 1810)))) %>% 
-  arrange(coef)
-```
-
-```{r}
-plot(glmnet_train_tuned$learner.model, "lambda")
-```
-
-```{r}
-plot(coef(glmnet_train_tuned$learner.model))
-```
-
-### SVM
+## SVM
 
@@
 
-### xgboost
+## xgboost
 
@@
 ps_xgboost <- makeParamSet(
-  makeIntegerParam("nrounds", lower = 0, upper = 600),
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
   makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
@@
 ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
-ctrl = setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
 tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
-                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
-# tune.ctrl = makeTuneControlRandom(maxit = 50)
+                               mbo.design = generateDesign(n = 30, 
+                                                           par.set = ps_xgboost))
+```
+
+```{r}
 inner <- makeResampleDesc("Blocking")
+outer <- makeResampleDesc("Blocking")
 ```
@@
 
-```{r}
-outer <- makeResampleDesc("Blocking")
-```
+If we parallelize the tuning using `level = "mlrMBO.feval"` for some unknown reason everything stalls after the first outer fold.
+That's why the parallelization takes place on the resample level.
 
@@
 parallelStart(mode = "multicore", cpus = 4, level = "mlr.resample",
-              mc.set.seed = TRUE) # only parallelize the tuning
+              mc.set.seed = TRUE)
 set.seed(12345)
@@
                           show.info = TRUE, measures = list(rmse))
-# resa_svm <- mlr::resample(lrn_xgboost, task,
-#                           resampling = outer, 
-#                           show.info = TRUE, measures = list(rmse))
-
 parallelStop()
@@
 
-### RF
+#### Tuning
+
+```{r}
+configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
+library(parallelMap)
+parallelStart(mode = "multicore", level = "mlrMBO.feval", cpus = 48,
+              mc.set.seed = TRUE)
+set.seed(12345)
+xgboost_tuned = tuneParams(lrn_xgboost, task = task, resampling = inner,
+                           par.set = ps_xgboost, control = tune.ctrl, 
+                           show.info = TRUE, measure = list(rmse))
+parallelStop()
+saveRDS(xgboost_tuned, "/data/patrick/mod/tmp/tuning_hyperspectral/xgboost_tuned_blocking.rda")
+```
+
+## RF
 

### 2018-07-17 01_writing/Schratz_hyperspectral.tex
update CV table

@@
 All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
-The data collection took place in September 2016.
+In total 1750 observations are available (\textit{Laukiz 1} = 479, \textit{Laukiz 2} = 451, \textit{Luiando} = 291, \textit{Oiartzun} = 529).
+The data was surveyed in September 2016.
 
@@
 The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around 12 am.
-The images were taken by an AISAEAGLE-II sensor from the \ac{ICGC}.
-All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by \ac{ICGC}.
+The images were taken using a AISAEAGLE-II sensor.
+All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by the \ac{ICGC}.
 The first four bands were corrupted, leaving 122 bands with valid information.
-
-Additional data characteristics are provided in Table 1:
+Additional metadata information is available in Table 1:
 
@@
 We also added penalized L2 (Ridge) regression to the algorithm collection due to its ability to handle highly correlated covariates.
-We also want to explain why we did not consider the probably most popular machine-learning algorithm, Random Forest: Due to the high number of variables model fitting times in the range hours for a single model fit were not practicable for this work.
+The probably most popular machine-learning algorithm, Random Forest, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
 These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{Probst2018b}.
@@
 	\hline
-		59.10 (22.71) & 36.23 (15.73) &  &         \\
+	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) \\
 	\bottomrule
@@
 \centering
-\caption[t]{Predictive performance of xgboost using the merged dataset (supermodel) and observations on a plot level only (single plot) with \ac{RMSE} as the error measure.
-The performance estimates for "All Plots" correspond to the fold for which the respective plot was serving as the test set.
-For "single plot", the values correspond to the mean value of the SpCV at the repetition level (10 folds, 20 repetitions), scored by models fitted on the data of the respective plot only.}
+\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations ) and observations from single plots only (Single Plot Observations) with \ac{RMSE} as the error measure.
+The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set.
+Column "single plot", shows the mean performances at the repetition level of a SpCV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
 \begingroup\footnotesize
@@
 	\\
-	Plot/Data & All Plots (Block CV) & Single plot (SpCV) \\
+	Plot/Data & All Observations (Block CV) & Single Plot Observations (SpCV)             \\
 	\hline
-	Laukiz 1  & 58.95                     & 89.89              \\
-	Laukiz 2  & 27.94                     & 30.37              \\
-	Luiando   & 69.72                     & 76.02              \\
-	Oiartzun  & 58.09                     & 106.65             \\ 
+	Laukiz 1  & 22.03                & -                              \\
+	Laukiz 2  & 51.75                & 17.37  (update to rep mean!)   \\
+	Luiando   & 13.20                & -                              \\
+	Oiartzun  & 32.97                & 14.72    (update to rep mean!) \\
 	\bottomrule

### 2018-07-30 02_scripts/031-laukiz1-modeling.Rmd
update scripts add prediction script

--- 2018-07-17 02_scripts/034-oiartzun/03-modeling_oiartzun.Rmd
@@
 library(data.table)
-library(pbmcapply)
+library(stringr)
+library(ggpubr)
 ```
@@
 ```{r}
-data = data[[4]]
-coords = coords[[4]]
+data = data[[1]]
+coords = coords[[1]]
 ```
@@
 ```{r}
-task = makeRegrTask(id = "oiartzun", data = data, 
+task = makeRegrTask(id = "laukiz1", data = data, 
                     target = "defoliation", coordinates = coords)
@@
 tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
-                               mbo.design = generateDesign(n = 30, par.set = ps_xgboost))
+                               mbo.design = generateDesign(n = 30, 
+                                                           par.set = ps_xgboost))
 ```
@@
 
-#### SpCV
+### SpCV
 
@@
 parallelStop()
-saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/oiartzun_xgboost.rda")
+saveRDS(resa_rf, "/data/patrick/mod/hyperspectral/spcv/laukiz1_xgboost.rda")
 ```

### 2018-07-30 02_scripts/035-all-plots-modeling.Rmd
update scripts add prediction script

--- 2018-07-17 02_scripts/035_all-plots/03-modeling_all.Rmd
@@
 library(mlrMBO)
+library(iml)
 library(glue)
@@
 library(pbmcapply)
+library(ggpubr)
 ```
@@
 
+#### Train
+
+```{r}
+xgboost_tuned = readRDS("/data/patrick/mod/tmp/tuning_hyperspectral/xgboost_tuned_blocking.rda")
+
+lrn_xgboost = setHyperPars(makeLearner("regr.xgboost"), 
+                           par.vals = xgboost_tuned$x)
+m = train(lrn_xgboost, task)
+saveRDS(m, "/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
+```
+
+### Variable importance
+
+#### mlr
+
+We can either use the Feat Imp from `xgboost` by calling `getFeatureImportance()` directly on the fitted model or calculate it again using permutation via `generateFeatureImportanceData()`.
+
+```{r}
+# 5h 30 min
+var_imp = generateFeatureImportanceData(task, learner = lrn_xgboost, 
+                                        measure = rmse, nmc = 2)
+saveRDS(var_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf2_ncm2.rda")
+
+var_imp$res = tidyr::gather(var_imp$res, "index", "feat.imp")
+var_imp$res %>% 
+  arrange(desc(feat.imp)) %>% 
+  slice(1:30) %>% 
+  ggdotchart(x = "index", y = "feat.imp",                              # Color by groups
+             sorting = "asc",                       # Sort value in descending order
+             add = "segments",                             # Add segments from y = 0 to dots
+             #color = "index",
+             palette = "nejm",
+             #position = position_dodge(1),
+             group = "index",
+             #title = .y,
+             rotate = T,                                # Rotate vertically
+             dot.size = 0.5,                                 # Large dot size
+             ggtheme = theme_pubr(base_size = 6, margin = F,
+                                  legend = "none")) + 
+  labs(y = "Score", color = "Spectral region")
+```
+
+Visualize top 30 indices
+
+```{r}
+imp.xgb = getFeatureImportance(m) 
+imp.xgb$res = tidyr::gather(imp.xgb$res, "index", "feat.imp")
+
+imp.xgb$res %>% 
+  arrange(desc(feat.imp)) %>% 
+  slice(1:30) %>% 
+  ggdotchart(x = "index", y = "feat.imp",                              # Color by groups
+             sorting = "asc",                       # Sort value in descending order
+             add = "segments",                             # Add segments from y = 0 to dots
+             #color = "index",
+             palette = "nejm",
+             #position = position_dodge(1),
+             group = "index",
+             #title = .y,
+             rotate = T,                                # Rotate vertically
+             dot.size = 0.5,                                 # Large dot size
+             ggtheme = theme_pubr(base_size = 6, margin = F,
+                                  legend = "none")) + 
+  labs(y = "Score", color = "Spectral region")
+```
+```{r}
+ggsave(here::here("03_figures/results/var-imp-xgboost.png"), height = 3,
+       width = 5)
+```
+
+#### iml
+
+```{r}
+X = task$env$data[which(names(task$env$data) != "defoliation")]
+predictor = Predictor$new(m, data = X, y = task$env$data$defoliation)
+imp = FeatureImp$new(predictor, loss = "rmse")
+```
+
+Partial influence
+
+```{r}
+X = task$env$data[which(names(task$env$data) != "defoliation")]
+predictor = Predictor$new(m, data = X, y = task$env$data$defoliation)
+pdp.obj = Partial$new(predictor, feature = "bf2_EVI")
+pdp.obj$plot()
+```
+
 ## RF

### 2018-07-30 02_scripts/04_plot_characteristics.Rmd
update scripts add prediction script

@@
 ```{r}
+mean_defol = data %>%
+  group_by(plot) %>%
+  summarise(mean(defoliation)) %>% 
+  pull(.)
+
 data %>%
@@
 ```{r}
+cov_defol = data %>%
+  group_by(plot) %>%
+  summarise((sd(defoliation) / mean(defoliation)) * 100) %>% 
+  pull(.)
+
 data %>%
@@
 ```{r}
+sd_skewness_defol = data %>%
+  group_by(plot) %>%
+  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation)) %>% pull(.)
+
 data %>%
@@
 ```{r}
+sd_defol = data %>%
+  group_by(plot) %>%
+  summarise(sd(defoliation)) %>% 
+  pull(.)
+
 data %>%
@@
 set_names(dist, plots)
+
+dist_plots = unlist(dist)
 ```
@@
 
+## Load CV results
+
+```{r}
+laukiz1 = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz1_xgboost.rda")
+laukiz2 = readRDS("/data/patrick/mod/hyperspectral/spcv/laukiz2_xgboost.rda")
+luiando = readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_xgboost.rda")
+oiartzun = readRDS("/data/patrick/mod/hyperspectral/spcv/oiartzun_xgboost.rda")
+
+models <- list(laukiz1, laukiz2, luiando, oiartzun)
+
+results_reps <- map(models, ~ .x$measures.test %>% 
+                      dplyr::mutate(rep = rep(1:5, times = 1, each = 5)) %>% 
+                      group_by(rep) %>% 
+                      summarise_at("rmse", mean),
+)
+results_reps <- set_names(results_reps, 
+                          c("laukiz1", "laukiz2", "luiando", "oiartzun"))
+```
+
 ```{r}
@@
   plot = factor(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")),
-  rmse = c(89.89, 30.37, 76.02, 106.65),
-  point_dens = c(34.34, 62.51, 33.76, 35.01),
-  mean_defol = c(53.22, 17.73, 68.35, 44.63),
-  cov_defol = c(44.88707	, 115.78180	, 12.17709	, 83.18840),
-  sd_skewness = c(-55.077546, 67.232146, -3.187733	, -401.564010)
+  rmse = c(mean(results_reps$laukiz1), mean(results_reps$laukiz2), mean(results_reps$luiando), mean(results_reps$oiartzun)),
+  point_dens = dist_plots,
+  mean_defol = mean_defol,
+  cov_defol = cov_defol,
+  sd_skewness = sd_skewness_defol
 )

### 2018-07-30 02_scripts/08_prediction.Rmd
update scripts add prediction script

---
title: "08-prediction"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(purrr)
library(raster)
library(furrr)
```

# Load model

```{r}
mod = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
```

# Load vegetation indices

```{r}
files = list.files("/data/patrick/mod/hyperspectral/all-veg-indices", 
                   pattern = ".grd", full.names = TRUE)

veg_inds = map(files, ~ brick(.))
```

## Format veg inds to sf objects

So that we have a data.frame with coordinates.

```{r}
plan(multicore)
veg_inds %<>% 
  future_map(~ as(., "SpatialPixelsDataFrame")) %>% 
  future_map(~ st_as_sf(.))
```

# Load NBI

```{r}
files_nbi = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
                   pattern = ".grd", full.names = TRUE)

nri = map(files_nbi, ~ brick(.))
```

## Format NRI to sf objects

So that we have a data.frame with coordinates.

```{r}
plan(multicore, workers = 5)
nri %<>% 
  future_map(~ as(., "SpatialPixelsDataFrame"), .progress = TRUE) %>% 
  future_map(~ st_as_sf(.), .progress = TRUE)
```

# Name list elements

```{r}
names = map_chr(files, ~ tools::file_path_sans_ext(basename(.)))

nri %<>% 
  set_names(names)

veg_inds %<>% 
  set_names(names)
```

# Combine NRI and veg inds

### 2018-07-31 01_writing/Schratz_hyperspectral.tex
add title
- add table with most imp variables - results: var imp

@@
 
-	\title{title}
+	\title{Modeling defoliation of Pinus Radiata trees using hyperspectral remote sensing data}
 
@@
 
-%To account for geometric offsets, we calculated every index five times using a buffer from 1 - 5 meter around the centroid of the respective tree.
 To account for geometric offsets, we used a buffer of two meters around the centroid of the respective tree.
@@
 
-To find indices that contributed most to model performance, we used permutation-based variable importance on the best performing algorithm.
+To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
+The score is calculated by taking the contribution of each feature for each tree in the fitted model.
+The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.
 
@@
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {../03_figures/results/plot-characteristics.pdf}}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {plot-characteristics.pdf}}
 		\caption{RMSE vs. mean point density, coefficient of variation and coefficient of variation / skewness.}
@@
 
-\subsubsection{Variable importance}
+\subsection{Variable importance}
 
-\ac{NRI}s using bands in the wavelength range of 770 nm - 820 nm (band 80 - band 89), which belongs to the infrared region, appear most often among the ten highest coefficient estimates across all plots (\autoref{tab:variable-importance}).
-Only one vegeation index (Datt3) showed up among the most important predictors (Laukiz 1).
-Luiando and Oiartzun also prefered bands with longer (938.39 nm (band 114) - 996.31 nm (band 126)) and shorter wavelengths (480.30 nm (band 18) - 503.26 (band 23)).
-The first range again belongs to the infrared region while the second is within the region of the visible light, transitioning from blue to green.
+Variable
 
-% band importance
+The 7 most important features are vegetation indices with \textit{EVI} \citep{hueteComparisonVegetationIndices1997} being the most important one.
+
+\begin{equation}
+	EVI = 2.5*\frac{R_{800}-R_{670}}{R_{800}-(6*R_{670})-(7.5*R_{475})+1)}
+\end{equation}
+
+Vegetation index \textit{GDVI} appears three times among the first seven features.
+This is because the \textit{hsdar} package computes the index four times, with \texttt{n} ranging from 1-4 \citep{wuEstimatingChlorophyllContent2008}:
+
+\begin{equation}
+	GDVI = \frac{R_{800}^n-R_{680}^n}{R_{800}^n+R_{680}^n}
+\end{equation}
+
+The first seven features (\textit{EVI}, \textit{GDVI4}, \textit{D1}, \textit{GDVI3}, \textit{GDVI2}, \textit{mNDVI} and \textit{mSR}) show a substantial difference compared to all following variables.
+
 \begin{figure} [b!]
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[angle=90, height = 0.95\textheight] {../03_figures/results/important-bands_boxplot.pdf}}
-		\caption{test}
-		\label{fig:important-bands}
+		\makebox[\textwidth]{\includegraphics[angle=90, height = 0.95\textheight] {var-imp-xgboost.pdf}}
+		\caption{The 30 most important variables as estimated by the internal variable importance measure of the \textit{xgboost} algorithm. The higher the score, the more important the feature.
+			"bf2" means that a buffer of 2 meter was used to extract the variable information to the tree observation.
+The following pattern describes the index: "NRI" means that a normalized ratio index with the following bands was used.
+Features without "NBI" are vegetation indices, e.g.
+"bf2\_EVI"}.
+%An overview of the vegetation indices can be found \href{https://rdrr.io/cran/hsdar/man/vegindex.html}{here}.}
+		\label{fig:var-imp}
 	\end{center}
 \end{figure}
+
+\begin{table}[t!]
+\centering
+\caption[t]{Formulas of the five most important vegetation indices of the super model. $R$ = Reflectance at wavelength, $D$ = First derivation of reflectance value at wavelength.}
+\begingroup\footnotesize
+\begin{tabular}{llll}
+	\\
+	Acronym & Name                                    & Formula                                                                 & Reference                                   \\
+	\hline
+	EVI     & Enhanced vegetation index               & $2.5*\frac{R_{800}-R_{670}}{R_{800}-(6*R_{670})-(7.5*R_{475})+1)}$        & \cite{hueteComparisonVegetationIndices1997}                                       \\
+	GDVI    & Generalized DVI* & $\frac{R_{800}^n-R_{680}^n}{R_{800}^n+R_{680}^n}$                       & \cite{wuEstimatingChlorophyllContent2008}  \\
+	D1      &         Derivative Index                                & $\frac{D_{730}}{D_{706}}$                                                       & \cite{zarco-tejadaSteadystateChlorophyllFluorescence2003} \\
+	mNDVI   & Normalized DVI*  & $  \frac{R_{800}-R_{680}}{(R_{800}+R_{680}-2*R_{445}}$                       & \cite{simsRelationshipsLeafPigment2002}    \\
+	mSR     & Simple Ratio Index                      & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                   & \cite{simsRelationshipsLeafPigment2002}    \\
+	\bottomrule
+	\multicolumn{2}{l}{\small{* Difference Vegetation Index}}
+\end{tabular}
+\endgroup
+\label{tab:supermodel_performance}
+\end{table}
 
@@
 
-The exact number of contributing pixels to the final index value of an observations cannot be determined as it depends on the location of the tree within the pixel grid.
+The exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
 If a tree is located at the border of a pixel, a buffer of e.g. three meters will include more pixels than if the point is located at the center of a pixel.

### 2018-07-31 01_writing/Schratz_hyperspectral.tex
complete table with performances

@@
 % Performance estimates
-\begin{table}[b!]
+\begin{table}[t!]
 \centering
-\caption[t]{Four-fold spatial \ac{CV} performances of RR, SVM and xgboost using \ac{RMSE} as the error measure.
+\caption[t]{Four-fold block spatial \ac{CV} performances of RR, SVM and xgboost using \ac{RMSE} as the error measure.
 	Mean and standard deviation are shown.}
@@
 
-\begin{table}[t!]
+\begin{table}[b!]
 \centering
-\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations ) and observations from single plots only (Single Plot Observations) with \ac{RMSE} as the error measure.
+\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations) and observations from specific plots only (Single Plot Observations) with \ac{RMSE} as the error measure.
 The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set.
-Column "single plot", shows the mean performances at the repetition level of a SpCV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
+Column "Single Plot Observations", shows the mean performances at the repetition level of a spatial CV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
 \begingroup\footnotesize
@@
 	\hline
-	Laukiz 1  & 22.03                       & -                               \\
-	Laukiz 2  & 51.75                       & 17.37  (update to rep mean!)    \\
-	Luiando   & 13.20                       & -                               \\
-	Oiartzun  & 32.97                       & 14.72    (update to rep mean!)  \\
+	Laukiz 1  & 22.03                       & 19.18                               \\
+	Laukiz 2  & 51.75                       & 17.24      \\
+	Luiando   & 13.20                       & 8.30                             \\
+	Oiartzun  & 32.97                       & 14.40      \\
 	\bottomrule

### 2018-07-31 01_writing/Schratz_hyperspectral.tex
update complete results section

@@
 \endgroup
-\label{tab:penalty_comparison}
+\label{tab:model_comparison}
 \end{table}
@@
 
-\ac{RR} shows the lowest error for three out of four plots (for Luiando \textit{elasticnet} shows a slightly better performance) (\autoref{tab:penalty_comparison}).
-The magnitude of difference for \ac{RR} compared to the other penalties for the plots in which \ac{RR} showed the best performance ranges between XX and XX percent.
-For the merged dataset, all penalties show a similar mean predictive performance that outperform all single plot models besides the Laukiz 2 model.
+\subsubsection{Algorithm benchmarking}
 
-% single models vs supermodel
-When comparing the mean predictive performance of the plot level model against the performance of the super model at the plot level (when the respective plot served as the test set), the supermodel also outperforms the Laukiz 2 model (27.94 vs 30.37 RMSE) (\autoref{tab:supermodel_performance}).
+The \textit{xgboost} algorithm shows the lowest error (33.26 RMSE) when benchmarking the learners on the complete dataset of all plots (\autoref{tab:model_comparison}).
+While the SVM performance is only slightly worse (36.23 RMSE), RR shows a large drop in performance compared to xgboost (59.10 RMSE).
 
-The worst performance of the supermodel on the fold level is reported for Luiando (69.72 RMSE) while for the single plot models Oiartzun shows the highest error (106.65 RMSE).
+\subsubsection{Single models vs. super model}
 
-% RMSE vs CV and point dens
-Laukiz 2 showed contrary results compared to all other plots when linking \ac{RMSE} against coefficient of variation and mean point density (\autoref{fig:rmse_defol_dens}).
-Comparing RMSE against $CV/skewness$ shows a $log_{2}(-x)$ relationship.
+When comparing the mean predictive performance of models fitted at the plot level against the performance of the model that was fitted using all data, the plot-level models show a better performance in all cases (\autoref{tab:supermodel_performance}).
+The highest difference between both models types occurs for \textit{Laukiz2} with a difference of 34.51 RMSE.
 
-% rmse vs defol and point dens
+\subsubsection{RMSE vs. plot characteristics}
+
+\noindent An increase of the error rate was observed with an increases of descriptive plot measures such as mean point density and the coefficient of variation (based on the response \textit{defoliation}) (\autoref{fig:plot-characteristics}).
+
 \begin{figure} [b!]
@@
 		\makebox[\textwidth]{\includegraphics[width=\textwidth] {plot-characteristics.pdf}}
-		\caption{RMSE vs. mean point density, coefficient of variation and coefficient of variation / skewness.}
+		\caption{RMSE vs. mean point density and coefficient of variation (defoliation).}
 		\label{fig:plot-characteristics}
@@
 
-Variable
-
-The 7 most important features are vegetation indices with \textit{EVI} \citep{hueteComparisonVegetationIndices1997} being the most important one.
+\noindent The seven most important features of the super model in this study were vegetation indices with \textit{EVI} \citep{hueteComparisonVegetationIndices1997} being the most important one (\autoref{fig:var-imp}).
 
@@
 
-Vegetation index \textit{GDVI} appears three times among the first seven features.
-This is because the \textit{hsdar} package computes the index four times, with \texttt{n} ranging from 1-4 \citep{wuEstimatingChlorophyllContent2008}:
+\noindent where $R$ = Reflectance at the respective wavelength.
+
+\bigbreak
+
+\noindent Vegetation index \textit{GDVI} appears three times among the first seven most important features (\autoref{fig:var-imp}) with different .
+This is because it was computed four times, with \texttt{n} ranging from 1 - 4 \citep{wuEstimatingChlorophyllContent2008}:
 
@@
 
-The first seven features (\textit{EVI}, \textit{GDVI4}, \textit{D1}, \textit{GDVI3}, \textit{GDVI2}, \textit{mNDVI} and \textit{mSR}) show a substantial difference compared to all following variables.
+\bigbreak
 
-\begin{figure} [b!]
+\noindent The seven most important features (\textit{EVI}, \textit{GDVI4}, \textit{D1}, \textit{GDVI3}, \textit{GDVI2}, \textit{mNDVI} and \textit{mSR}) show a substantial difference compared to all following variables (\autoref{fig:var-imp}).
+
+\begin{figure} [t!]
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[angle=90, height = 0.95\textheight] {var-imp-xgboost.pdf}}
+		\makebox[\textwidth]{\includegraphics[width = 1\textwidth] {var-imp-xgboost.pdf}}
 		\caption{The 30 most important variables as estimated by the internal variable importance measure of the \textit{xgboost} algorithm. The higher the score, the more important the feature.
 			"bf2" means that a buffer of 2 meter was used to extract the variable information to the tree observation.
-The following pattern describes the index: "NRI" means that a normalized ratio index with the following bands was used.
-Features without "NBI" are vegetation indices, e.g.
-"bf2\_EVI"}.
+"NRI" means that a normalized ratio index with the subsequent bands was calculated.
+Features without "NRI" prefix are vegetation indices, e.g.
+"bf2\_EVI".}.
 %An overview of the vegetation indices can be found \href{https://rdrr.io/cran/hsdar/man/vegindex.html}{here}.}
@@
 \endgroup
-\label{tab:supermodel_performance}
+\label{tab:most_imp_vars}
 \end{table}

### 2018-07-31 02_scripts/01_hyperspectral_indices/05_narrow-band-indices.Rmd
NBI -> NRI

@@
 library(purrr)
+library(stringr)
 ```
@@
                              "narrow-band-indices/narrow-band-indices-{name}",
-                             name = basename(x@file@name)), ".tif", ".grd"), mc.cores = 15)) 
+                             name = basename(x@file@name)), ".tif", ".grd"),
+             bnames_prefix = "NRI"), mc.cores = 15)
 rm(tmp)

### 2018-07-31 02_scripts/035-all-plots-modeling.Rmd
scripts

@@
 ```{r}
+m = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
 # 5h 30 min
 var_imp = generateFeatureImportanceData(task, learner = lrn_xgboost, 
-                                        measure = rmse, nmc = 2)
-saveRDS(var_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf2_ncm2.rda")
+                                        measure = rmse, nmc = 50)
+saveRDS(var_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf2_ncm50rda")
 

### 2018-07-31 02_scripts/04_plot_characteristics.Rmd
scripts

@@
 
+all_plots = readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_xgboost_blocking.rda")
+
 models <- list(laukiz1, laukiz2, luiando, oiartzun)
@@
   plot = factor(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")),
-  rmse = c(mean(results_reps$laukiz1), mean(results_reps$laukiz2), mean(results_reps$luiando), mean(results_reps$oiartzun)),
+  rmse = c(mean(results_reps$laukiz1$rmse), mean(results_reps$laukiz2$rmse), mean(results_reps$luiando$rmse), mean(results_reps$oiartzun$rmse)),
   point_dens = dist_plots,
@@
 
+## Differences between models (single vs super)
+
+```{r}
+# laukiz1
+all_plots$measures.test$rmse[3] - mean(results_reps$laukiz1$rmse)
+
+# laukiz2
+all_plots$measures.test$rmse[4] - mean(results_reps$laukiz2$rmse)
+
+# luiando
+all_plots$measures.test$rmse[2] - mean(results_reps$luiando$rmse)
+
+# oiartzun
+all_plots$measures.test$rmse[1] - mean(results_reps$oiartzun$rmse)
+```
+
 ## Plot
@@
 ```{r}
-prow = plot_grid(rmse_dens, rmse_defol_cov, rmse_defol_cov_sk,  align = 'vh', nrow = 1)
-plots_grid <- plot_grid(prow, legend, rel_widths = c(3, .3))
+prow = plot_grid(rmse_dens, rmse_defol_cov,  align = 'vh', nrow = 1)
+plots_grid <- plot_grid(prow, legend, rel_widths = c(2, .5))
 ```
@@
 ```{r}
-plots_grid <- plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
+#plots_grid <- plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
 ```
@@
 ```{r}
-save_plot(here("03_figures/results/plot-characteristics.png"), plots_grid, ncol = 4)
+save_plot(here("03_figures/results/plot-characteristics.png"), plots_grid, ncol = 3)
 ```

### 2018-07-31 02_scripts/08_prediction.Rmd
scripts

@@
 library(furrr)
+library(sf)
+library(pbmcapply)
 ```
@@
 
-# Load NBI
+# Load NRI
 
 ```{r}
-files_nbi = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
+files_nri = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
                    pattern = ".grd", full.names = TRUE)
 
-nri = map(files_nbi, ~ brick(.))
+nri = map(files_nri, ~ brick(.))
 ```
@@
 ```{r}
-plan(multicore, workers = 5)
 nri %<>% 
-  future_map(~ as(., "SpatialPixelsDataFrame"), .progress = TRUE) %>% 
-  future_map(~ st_as_sf(.), .progress = TRUE)
+  pbmclapply(function(x) as(x, "SpatialPixelsDataFrame"), mc.cores = 2) %>% 
+  pbmclapply(function(x) st_as_sf(x), mc.cores = 2)
 ```

### 2018-08-02 01_writing/Schratz_hyperspectral.tex
write disc: error vs plot characteristics - move spectral signatures to appendix

@@
 
-% link to PDF with veg indices
+The idea behind calculating this amount of indices was to use the full information of the high radiometric resolution of the hyperspectral data.
+Besides vegetation indices, that have been shown to be sensitive to changes to the health status of vegetation, we were interested if NRIs of arbitrary band combinations will have a substantial positive effect on the predictive power of the fitted model.
 All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that were available in the R package \textit{hsdar} have been calculated.
@@
 
-\subsection{Exploratory analysis of plot characteristics}
-
-Plot characteristics like age, stand density and defoliation were analysed to show differences among the plots.
-Additionally, the spectral signatures of each plot have been visualized.
-
 \subsection{Benchmarking of algorithms}
@@
 These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{Probst2018b}.
+After the selection of the best model, we checked if the winning algorithm can achieve a similar performance when using only the most important variables compared to using all variables.
+We did not use a robust selection metric here but just selected the \texttt{n} most important variables by inspecting the importance score of the model (\autoref{fig:var-imp}).
 
@@
 
-\subsection{Exploratory data analysis}
+\subsection{Plot characteristics}
 
@@
 		\label{fig:defol_boxplots}
-	\end{center}
-\end{figure}
-
-% spectral signatures
-\begin{figure} [t!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral_signatures.pdf}}
-		\caption{Spectral signatures (mean and standard deviation) of each plot.
-			The colored lines show the most important band for each plot, respectively: Band 80 (773nm, red), band 89 (817nm, blue), band 95 (846nm, orange), band 23 (503nm, green).}
-		\label{fig:spectral_signatures}
 	\end{center}
@@
 
-\begin{table}[b!]
+\begin{table}[t!]
 \centering
@@
 
-\noindent An increase of the error rate was observed with an increases of descriptive plot measures such as mean point density and the coefficient of variation (based on the response \textit{defoliation}) (\autoref{fig:plot-characteristics}).
+\noindent An increase of the error rate was observed with an increase of descriptive plot measures such as mean point density and the coefficient of variation (based on the response \textit{defoliation}) (\autoref{fig:plot-characteristics}).
 
@@
 
-\noindent Vegetation index \textit{GDVI} appears three times among the first seven most important features (\autoref{fig:var-imp}) with different .
+\noindent Vegetation index \textit{GDVI} appears three times among the first seven most important features (\autoref{fig:var-imp}) with different \texttt{n} values.
 This is because it was computed four times, with \texttt{n} ranging from 1 - 4 \citep{wuEstimatingChlorophyllContent2008}:
@@
 
-\begin{figure} [t!]
+\begin{figure} [b!]
 	\begin{center}
@@
 
-\subsection{Index derivation}
+\subsection{Derivation of indices}
+
+The buffer of 2m that we used to generate the index value for each observation can be seen critical.
+When using now buffer at all, the possibility exists that a pixel value is assigned to the tree observation that does not spatially match with the tree observation.
+Using a buffer of more than 2 meters would cause information from other trees to be merged into the pixel value, blurring the actual value of the tree observation.
+Thats why using a buffer of 2m was the best compromise here in our perspective.
 
 The exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-If a tree is located at the border of a pixel, a buffer of e.g. three meters will include more pixels than if the point is located at the center of a pixel.
-Also, if a tree is located at the border of the plot, some directions of the buffer will not contain image values.
+As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels will be used to calculate the final pixel value.
+If a tree observation is located at the border of the plot, some directions of the buffer will not contain image values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
 
-\subsection{Plot characteristics}
+Using a buffer in the first place and the fact that it is unclear how much information from other trees went into the calculation of an index value of a certain observation certainly limits the meaningfullness of the study and has to be considered when making interpretations about the outcome of this study.
 
-For Laukiz1, Luiando and Oiartzun \ac{RMSE} seems to increase with a higher point density at a first glance.
-However, the point densities of these plots are very similar (33.7 m - 35.01 m) and should be interpreted as a group instead of single values.
-With Laukiz2 being completely off from the other plots in terms of mean point density, no pattern can be extracted from this result.
-Linking RMSE vs coefficient of variation shows the same relationship as linking against mean point density.
-The interesting $log_{2}(-x)$ relationship for RMSE vs.
-coefficient of variation / skewness should be interpreted with caution: The sample size of four plots is not representative to make general statements here.
-This finding should be verified with more observations in future studies.
+\subsection{RMSE vs. plot characteristics}
+
+Trying to relate the modeling error to plot characteristics (mean point density, CV) did not show a clear picture: For both comparisons, Laukiz 2 did not follow the pattern that was observed from the other three plots (\autoref{fig:plot-characteristics}) of having an increase in error with an increase in mean point density and CV.
+
+Having an increase in error with an increase in mean point density could be related to the buffer we used in this study. 
+If a tree is surrounded by bare ground, chances are higher that bare ground information will be included in the final index value. 
+By contrast, if the density of trees in a plot is high, the possibility of bare ground information in the calculated index value will be small.
+The fact of having information from another tree in the final pixel value is less problematic than having bare ground information as the latter affects the final index calculation more in a negative way.
+
+Nevertheless, it has to be considered that we did only have four plots here. 
+To make a valid statement about a relationship between modeling error and these plot characteristics, one would need a bigger sample of plots.
 
 \subsection{Variable importance}
+
+\section{Appendix}
+
+\appendix
+% https://tex.stackexchange.com/questions/248704/cross-reference-to-appendix-sections-in-elsarticle-document-class
+\gdef\thesection{\Alph{section}} % corrected redefinition of "\thesection"
+\makeatletter
+\renewcommand\@seccntformat[1]{Appendix \csname the#1\endcsname.\hspace{0.5em}}
+\makeatother
+
+\section{Spectral signatures of each plot}
+
+% spectral signatures
+\begin{figure} [H]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral_signatures.pdf}}
+		\caption{Spectral signatures (mean and standard deviation) of each plot.}
+		\label{fig:spectral_signatures}
+	\end{center}
+\end{figure}
+
+\pagebreak
 

### 2018-08-02 01_writing/Schratz_hyperspectral.tex
disc: single model vs super model

@@
 
-Nevertheless, it has to be considered that we did only have four plots here. 
-To make a valid statement about a relationship between modeling error and these plot characteristics, one would need a bigger sample of plots.
+Nevertheless, it has to be considered that we only had four plots in this work. 
+To make a valid statement about a relationship between modeling error and these plot characteristics, a larger sample size of plots is needed.
+
+\subsection{Predictive Performance}
+
+\subsubsection{Algorithm benchmarking}
+
+The relatively large difference in performance between RR (59.10 RMSE) and the machine-learning models (36.23 and 33.26 RMSE) is surprising. 
+RR has shown promising performance results in other studies when many highly-correlated predictors were involved (REFERECES).
+However, in this study, RR was not able to achieve a sufficient performance score compared to SVM and xgboost even though its hyperparameter $\lambda$ was properly tuned using SMBO.
+
+While xgboost shows a slightly better performance than SVM, the latter has the advantage of only having two hyperparameters that need to be tuned.
+This results in a shorter runtime.
+Nevertheless, xgboost showed the best performance and was subsequently used to fit the models on the plot level and for the prediction.
+
+\subsubsection{Single models vs. super model}
+
+It is expected that models that were trained on the plot level only achieve a better performance than the global model.
+The low performance on Laukiz 2 for the super model is most likely due to the difference of this plot to all others: The fitted model on Laukiz 1, Luiando and Oiartzun is not capable of reaching a good performance on the Laukiz 2 prediction data.
+This is not surprising as Laukiz 2 shows substantially different plot characteristics compared to all others plots in terms of CV, mean point density (\autoref{arg}) and defoliation (\autoref{fig:defol_boxplots}).
+
+The low error for Luiando (8.30 RMSE) for the plot model validates the approach of relating defoliation to vegetation indices and NRI. 
+The overall error of the super model (33.26 RSME) is expected to decrease if more plots are available for training.
+To reach an optimal performance, the fitted model would need to include at least one instance of every plot that shows unique characteristics (i.e.
+here Laukiz 2 is substantially different to the others).
+Possibly more plots showing unique characteristics exists that are not integrated into the fitted model of this work.
 

### 2018-08-02 01_writing/Schratz_hyperspectral.tex
methods: var imp disc: var imp

@@
 The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.
+Using the internal variable importance measure from the xgboost algorithm is quite convienient since it is automatically computed during model fit.
+By contrast to other approaches such as permutation-based variable importance, it is composed out of three parts that contribute to the overall importance score:
+
+\begin{itemize}
+	\item Gain: The relative contribution of the feature to the model 
+	\item Cover metric: How often a feature was selected to be the deciding feature in a tree for a specific observation 
+	\item Frequency How often a feature occurs in all trees of the model 
+\end{itemize}
+
+The \textit{Gain} features is the most important one among the three. 
+All measures sum up to one.
+Downsides of this approach are that due to the contribution of three different parts it is complicated to understand why a specific feature was selected.
+Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
+Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.
 
@@
 \begin{tabular}{llll}
-	RR            & SVM           & xgboost       \\
+	RR            & SVM           & xgboost       & xgboost (7 variables) \\
 	\hline
-	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) \\
+	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) &                       \\
 	\bottomrule
@@
 \begingroup\footnotesize
-\begin{tabular}{lll}
+\begin{tabular}{llll}
 	\\
-	Plot/Data & All Observations (Block CV) & Single Plot Observations (SpCV) \\
+	Plot/Data & All Observations/all variables (Block CV) & All Observations/7 variables (Block CV) & Single Plot Observations (SpCV) \\
 	\hline
-	Laukiz 1  & 22.03                       & 19.18                           \\
-	Laukiz 2  & 51.75                       & 17.24                           \\
-	Luiando   & 13.20                       & 8.30                            \\
-	Oiartzun  & 32.97                       & 14.40                           \\
+	Laukiz 1  & 22.03                                     &                                         & 19.18                           \\
+	Laukiz 2  & 51.75                                     &                                         & 17.24                           \\
+	Luiando   & 13.20                                     &                                         & 8.30                            \\
+	Oiartzun  & 32.97                                     &                                         & 14.40                           \\
 	\bottomrule

### 2018-08-03 01_writing/Schratz_hyperspectral.tex
add xgboost 7 vars only to table and update results

@@
 The \textit{Gain} features is the most important one among the three.
-All measures sum up to one.
-Downsides of this approach are that due to the contribution of three different parts it is complicated to understand why a specific feature was selected.
-Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
-Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.
+All measures sum up to one (REFERENCE!).
 
@@
 	\\
-	Plot/Data & All Observations/all variables (Block CV) & All Observations/7 variables (Block CV) & Single Plot Observations (SpCV) \\
+	Plot/Data & \specialcell{All Observations/ \\{all variables (Block CV)}} & \specialcell{All Observations/         \\ {7 variables (Block CV)}} & \specialcell{Plot level observations/\\{all variables (SpCV)}} \\
 	\hline
-	Laukiz 1  & 22.03                                     &                                         & 19.18                           \\
-	Laukiz 2  & 51.75                                     &                                         & 17.24                           \\
-	Luiando   & 13.20                                     &                                         & 8.30                            \\
-	Oiartzun  & 32.97                                     &                                         & 14.40                           \\
+	Laukiz 1  & 22.03                                                                              &       21.47                         & 19.18 \\
+	Laukiz 2  & 51.75                                                                              &        49.94                        & 17.24 \\
+	Luiando   & 13.20                                                                              &       15.37                        & 8.30  \\
+	Oiartzun  & 32.97                                                                              &    17.62                            & 14.40 \\
 	\bottomrule
@@
 The highest difference between both models types occurs for \textit{Laukiz2} with a difference of 34.51 RMSE.
+
+Using only the seven most important variables (\autoref{fig:var-imp}) shows small increases in performance for Laukiz 1 and Laukiz 2, a small decrease for Luiando and almost a reduction of 50\% of the error for Oiartzun (32.97 vs.
+17.62 RMSE).
 
@@
 
+The best NRI scored rank eight (band 112 and band 62).
+All further places up to rank 30 are NRIs.
+
 \section{Discussion}
@@
 
+An important point that needs to be considered when interpreting the performance results is that we only related defoliation to indices derived from remote sensing data.
+Possible other cofounding factors that could possibly help in predicting defoliation were not considered.
+One example here is tree age: The older a tree the more vunerable it may be to pathogen infections causing defoliation.
+However, such predictors would not be available for a spatial prediction scenario and one of the main goals of this study is to relate defoliation to variables that are available on a larger scale (e.g.
+remote sensing indices).
+
+	[COMPARE ALL VARIABLES VS 7 VARIABLES]
+
 \subsubsection{Single models vs. super model}
@@
 The low performance on Laukiz 2 for the super model is most likely due to the difference of this plot to all others: The fitted model on Laukiz 1, Luiando and Oiartzun is not capable of reaching a good performance on the Laukiz 2 prediction data.
-This is not surprising as Laukiz 2 shows substantially different plot characteristics compared to all others plots in terms of CV, mean point density (\autoref{arg}) and defoliation (\autoref{fig:defol_boxplots}).
+This is not surprising as Laukiz 2 shows substantially different plot characteristics compared to all others plots in terms of the distribution of the response variable \textit{defoliation} (\autoref{fig:defol_boxplots}) and the mean point density of trees (\autoref{fig:plot-characteristics}).
 
@@
 \subsection{Variable importance}
+
+There are some downsides using the internal variable importance approach of xgboost: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
+Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
+Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.
+
+It is not surprising that vegetation indices are most important for the model as they are most sensitive to changes in vegetation health.
+Even though we are not specifically looking at vegetation health but at the level of defoliation, these indices seem to help the model deciding whether a tree is defoliated or not.
+The vegetation indeces could help here in two ways:
+1) Trees that show a high level of defoliation do also reflect their bad health status through the remaining foliation.
+2) Defoliated trees have more influence of bare ground information in their pixel values and will therefore be identified as defoliated by the model.
+
+Even though no NRI made it among the most important variables in this study (stating that the first seven of this study are the most important ones), it is interesting that they occupy all ranks from 8 - 30 (\autoref{fig:var-imp}).
+Our statement saying that the important indices of this study are the first seven can be seen critical as we only based the selection on a visual inspection of the variable importance results (\autoref{fig:var-imp}).
+All indices starting from rank eight do simply have a very small score.
+Nevertheless, one could also make the cut at rank five here.
+However, the break between vegetation indeces and NRI at rank eight also had an influence in setting this subjective break there.
 

### 2018-08-03 01_writing/Schratz_hyperspectral.tex
add discussion about 7 vars model

@@
 
-Using only the seven most important variables (\autoref{fig:var-imp}) shows small increases in performance for Laukiz 1 and Laukiz 2, a small decrease for Luiando and almost a reduction of 50\% of the error for Oiartzun (32.97 vs.
-17.62 RMSE).
+Using only the seven most important variables (\autoref{fig:var-imp}) for the super model shows small increases in performance for Laukiz 1 and Laukiz 2, a small decrease for Luiando and almost a reduction of 50\% of the error for Oiartzun (32.97 vs.
+17.62 RMSE) (\autoref{tab:supermodel_performance}).
 
@@
 
+\subsection{Spatial prediction}
+
 \section{Discussion}
@@
 Possible other cofounding factors that could possibly help in predicting defoliation were not considered.
-One example here is tree age: The older a tree the more vunerable it may be to pathogen infections causing defoliation.
+One example here is tree age: The older a tree the more vulnerable it may be to pathogen infections causing defoliation.
 However, such predictors would not be available for a spatial prediction scenario and one of the main goals of this study is to relate defoliation to variables that are available on a larger scale (e.g.
 remote sensing indices).
-
-	[COMPARE ALL VARIABLES VS 7 VARIABLES]
 
@@
 
+An interesting find is that the supermodel with only seven variables shows a better overall performance than the model with all 7471 variables (\autoref{tab:supermodel_performance}).
+This leads to the conclusion that adding as many variables as possible to a model will not necessarily improve its performance.
+In contrast, too much information can even be problematic for the model as it will have a hard time prioritizing variables.
+However, to find the most important variables in the first place and to check for the performance difference, a model with all variables needs to fitted first.
+This scenario cannot be generalized and different results may occur for other datasets.
+Using a model with only a few predictors does not only simplify prediction tasks but also reduces runtime for hyperparameter tuning and performance estimation.
+
 \subsection{Variable importance}
@@
 It is not surprising that vegetation indices are most important for the model as they are most sensitive to changes in vegetation health.
-Even though we are not specifically looking at vegetation health but at the level of defoliation, these indices seem to help the model deciding whether a tree is defoliated or not.
-The vegetation indeces could help here in two ways:
+Even though we are not directly looking at vegetation health but using the level of defoliation as a proxy, these indices seem to help the model most deciding whether a tree is defoliated or not.
+Vegetation indices can help here in two ways:
 1) Trees that show a high level of defoliation do also reflect their bad health status through the remaining foliation.
-2) Defoliated trees have more influence of bare ground information in their pixel values and will therefore be identified as defoliated by the model.
+2) Defoliated trees have more influence of bare ground information in their pixel values and will therefore be classified as defoliated by the model.
 
-Even though no NRI made it among the most important variables in this study (stating that the first seven of this study are the most important ones), it is interesting that they occupy all ranks from 8 - 30 (\autoref{fig:var-imp}).
+Even though no NRI made it among the most important variables in this study (stating that the first seven of this study are the most important ones), it is interesting that all ranks from 8 - 30 are occupied by NRI (\autoref{fig:var-imp}).
+
 Our statement saying that the important indices of this study are the first seven can be seen critical as we only based the selection on a visual inspection of the variable importance results (\autoref{fig:var-imp}).
-All indices starting from rank eight do simply have a very small score.
-Nevertheless, one could also make the cut at rank five here.
-However, the break between vegetation indeces and NRI at rank eight also had an influence in setting this subjective break there.
+The decision to make a cut between rank seven and eight was based on a combination of two facts: 1) Using only vegetation indices is easier for large scale predictions using satellites like Sentinel-2 (most NRI cannot be used with it) and 2) the drop in the importance score of the variable importance results (\autoref{fig:var-imp}).
+However, based on this statement, we could also have made the cut between rank five and 6 but including the two vegetation indices at rank six and seven will eventually improve the model and does not increase runtime.
+
+\section{Conclusion}
 

### 2018-08-03 01_writing/Schratz_hyperspectral.tex
small corrections in the whole doc

@@
 
-	\title{Modeling defoliation of Pinus Radiata trees using hyperspectral remote sensing data}
+	\title{Modeling defoliation as a proxy for tree health: A case study on Pinus Radiata trees using machine-learning and hyperspectral data}
 
@@
 
-With a robust model, predictions of the modelled response to large areas is possible, giving valuable information about the condition of this variable in unknown regions.
-To model forest health, usually few variables are extracted based on the spectral signatures of affected and unaffected trees \citep{lelongEvaluationOilPalmFungal2010}.
-However, spectral (vegetation-)indices have shown the potential to contribute valuable information to increase predictive accuracy of forest pathogens \citep{jiangSatellitederivedVegetationIndices2014, adamczykRededgeVegetationIndices2015}.
+With a robust model, predictions to large areas can be conducted, providing valuable information about the health condition of forest stands.
+One approach to model forest health is to extract information from spectral signatures of affected and unaffected trees \citep{lelongEvaluationOilPalmFungal2010}.
+Spectral (vegetation-)indices have shown the potential to provide valuable information to increase predictive accuracy of forest pathogens \citep{jiangSatellitederivedVegetationIndices2014, adamczykRededgeVegetationIndices2015}.
 
-However, the amount of possible (vegetation-)indices that could be calculated is often limited due to a low spectral resolution of freely available data from optical multispectral sensors (e.g.
+However, the amount of possible (vegetation-)indices that can be calculated is often limited due to a low spectral resolution of freely available data from optical multispectral sensors (e.g.
 Sentinel-2).
-Also, there is currently no free data available from hyperspectral sensors that could be used for such studies (after the decommission of the EO-1 Hyperion satellite).
+Also, there is currently no freely available data from hyperspectral sensors that could be used for such studies (after the decommission of the EO-1 Hyperion satellite).
 If the spatial resolution of the data is too coarse (e.g. $> 5 m$), the value of a pixel usually contains information from multiple trees and possibly even bare-ground information.
-This makes the resulting information almost useless to be used for forest health monitoring on a tree level.
+This point makes it impossible to use data from sensors like Sentinel-2 to train a model on a tree level.
 
@@
 
-In this study we will use hyperspectral data with a spatial resolution of one meter and 126 spectral bands to model the health status of Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
-The trees in the study area suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum} \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation before the tree dies \citep{mesanzaNativeRhizobacteriaBiocontrol2016, iturritxaBiocontrolFusariumCircinatum2017}.
-In-situ measurements of defoliation on a tree level are used as a proxy to model tree health.
-The fungis are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
+In this study we will use hyperspectral data with a spatial resolution of one meter and 126 spectral bands to model the health status of Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain (\autoref{fig:study_area}).
+The trees in the study area suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum} \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanzaNativeRhizobacteriaBiocontrol2016, iturritxaBiocontrolFusariumCircinatum2017}.
+In-situ measurements of defoliation on a tree level have been collected to serve as the response variable (as a proxy for tree health).
+The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
 The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{ganleyGlobalClimaticRisk2009}.
 Hyperspectral remote sensing data in combination with state-of-the-art machine-learning techniques is used to help monitoring the health status in this region by early detecting affected trees/plots.
+The aim is to spatially predict the fitted model to other plots/the whole Basque Country to create a forest health map.
 
@@
 
-To extract the most information from the available remote sensing data, we not only calculated the most common vegetation indices like \textit{NDVI} to link against defoliation but all possible ones within the spectral region of the data (400 nm - 1000 nm) that were implemented in the \textit{hsdar} package in R \citep{hsdar}.
-Additionally, all possible combinations of \ac{NRI} were calculated from the data and supplied to a selection of machine-learning algorithms as predictors.
+To extract the most information from the available remote sensing data, we not only calculated the most common vegetation indices like \textit{NDVI} to link against defoliation but all possible ones within the spectral region of the data (400 nm - 1000 nm) that are implemented in the \textit{hsdar} package in R \citep{hsdar}.
+Additionally, all possible combinations of \ac{NRI} were calculated from the data and supplied to a selection of different algorithms (\textit{\ac{SVM}}, \textit{\ac{RR}}, \textit{xgboost}) as predictors.
 
@@
 	\item Exploration of the most important indices of the best performing model
-	\item Prediction of defoliation to \textit{Pinus radiata} plots with an unknown defoliation level
+	\item Spatial prediction of defoliation to \textit{Pinus radiata} plots and the whole Basque Country
 \end{itemize}
@@
 
-The \textit{Pinus radiata} plots of this study, named \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
+The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
 \textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} has the largest area size (1.44 ha).
@@
 All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by the \ac{ICGC}.
-The first four bands were corrupted, leaving 122 bands with valid information.
+The first four bands are corrupted, leaving 122 bands with valid information.
 Additional metadata information is available in Table 1:
@@
 % parameter limits
-\begin{table}[b!]
+\begin{table}[t!]
 \centering
@@
 For all analysis steps we used the open-source statistical programming language R \citep{R_core}.
-The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chenXGBoostScalableTree2016}, \textit{kernlab} \citep{kernlab} (Support Vector Machine), \cite{Vapnik1998}) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
+The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chenXGBoostScalableTree2016} (\textit{xgboost}), \textit{kernlab} \citep{kernlab} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
 We used the R package \textit{mlr} for all modeling related steps.
@@
 
-The idea behind calculating this amount of indices was to use the full information of the high radiometric resolution of the hyperspectral data.
-Besides vegetation indices, that have been shown to be sensitive to changes to the health status of vegetation, we were interested if NRIs of arbitrary band combinations will have a substantial positive effect on the predictive power of the fitted model.
-All vegetation indices (90 total) suitable for the wavelength range of the hyperspectral data that were available in the R package \textit{hsdar} have been calculated.
-Additionally, all possible \ac{NRI} were calculated from the data using the formula:
+To use the full information from the hyperspectral data, we not only calculated all possible vegetation indices that are available in the \textit{hsdar} package (90 in total) but also all possible combinations of \ac{NRI}.
+We were interested if NRIs of arbitrary band combinations will have a substantial positive effect on the predictive power of the fitted model.
+The NRI were calculated using the following formula:
 
@@
 
-To account for geometric offsets, we used a buffer of two meters around the centroid of the respective tree.
+\bigbreak
+
+To account for geometric offsets (which were reported with up to 1 m), we used a buffer of two meters around the centroid of the respective tree.
 The mean value of all pixels touched by the buffer was assigned as the final value for each index.
 Missing values were removed from the mean value calculation.
-In total, 7875 Normalized Ratio Indices NRI have been calculated ($\frac{125*126}{2}$).
-Due to four corrupted bands and some other numerical problems, few indices returned \texttt{NA} values for some observations.
+In total, 7875 Normalized Ratio Indices NRI were calculated ($\frac{125*126}{2}$).
+Due to four corrupted bands and other numerical problems, some indices returned \texttt{NA} for specific observations.
 These indices were removed from the dataset, leaving a total of 7471 variables without missing values.
@@
 
-Multiple algorithms were benchmarked on predictive performance to find the best performing one.
-Besides the well-known \ac{SVM} \citep{Vapnik1998} we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
-We also added penalized L2 (Ridge) regression to the algorithm collection due to its ability to handle highly correlated covariates.
-The probably most popular machine-learning algorithm, Random Forest, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
+Three algorithms (\textit{xgboost}, \textit{\ac{SVM}}, \textit{\ac{RR}}) were benchmarked on predictive performance.
+Besides the well-known \ac{SVM} algorithm \citep{Vapnik1998} we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
+We also added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates (REFERENCE).
+The probably most popular machine-learning algorithm, Random Forest \citep{Breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
 These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{Probst2018b}.
 After the selection of the best model, we checked if the winning algorithm can achieve a similar performance when using only the most important variables compared to using all variables.
-We did not use a robust selection metric here but just selected the \texttt{n} most important variables by inspecting the importance score of the model (\autoref{fig:var-imp}).
-
-% \subsubsection{The ridge penalty}
-
-% In \ac{RR} (also called $\ell_{2}$ penalization) the assumption of unbiased coefficients is given up in favor of higher predictive accuracy and reduced variance \citep{Hastie2001}.
-% Coefficients are standardized and penalized for their size.
-% When minimizing the \ac{RSS}, \ac{RR} adds a penalization term $\lambda \sum_{j=1}^{p}\beta_{j}^{2}$ to the equation:
-
-% \begin{equation}
-% 	RSS = \sum_{i=1}^{n} \left(y_{i} - \beta_{0} - \sum_{j=1}^{p} \beta_{j} x_{ij} \right) ^{2} + \lambda \sum_{j=1}^{p}\beta_{j}^{2}
-% \end{equation}
-
-% where $\lambda >= 0$ is a tuning parameter responsible for the magnitude of penalization.
-% To make the second term (usually referred to as the \textit{shrinkage penalty}) of this equation small, the coefficients $\beta_{j}$ need to become small.
-% Unlike to Lasso however, predictors are not removed from the final model and will always be $\beta_{j} >= 0$.
-% Hence, $\lambda$ has the effect of shrinking the coefficients when minimizing the \ac{RSS}.
-% For $\lambda = 0$, no penalization is done and standard \ac{OLS} applies \citep{James2013}.
-% The \textit{shrinkage penalty} is only applied to the coefficients and not to the intercept.
-% Also, while \ac{OLS} generates only one set of coefficient estimates, \ac{RR} will create multiple sets for every value of $\lambda$.
-% In summary, the advantage of \ac{RR} is based on the bias-variance tradeoff: For $\lambda\to\infty$, the flexibility of the fit is reduced leading to a decrease in variance of the coefficients but also introduces a (substantial) bias.
-% For $\lambda = 0$, the variance is high but coefficients are unbiased \citep{James2013}.
+This would simplify the spatial prediction a lot as less indices need to be computed from the data used for prediction.
+Also, simpler models are always preferred over complex ones (REFERENCE).
 
@@
 
-The algorithms were benchmarked in two ways:
-(1) Using spatial \ac{CV} for each plot using on the k-means clustering approach of \cite{Brenning2012}. To reduce runtime we used a five-fold five-times repeated CV setup.
-(2) Using spatial \ac{CV} on the plot level with each plot being the test set once. This results in four performance estimates, one for each fold.
-For (1) we only used the best performing algorithm from (2).
-The reason why the (2) was chosen for algorithm selection is that this model will also be used to spatially predict defoliation in other plots.
+\noindent The algorithms were benchmarked in two ways:
+1) Using spatial block \ac{CV} on the plot level with each plot serving as the test set once.
+This approach results in four performance estimates, one for each fold.
+2) Using five-fold five-times repeated spatial \ac{CV} for each plot based on the k-means clustering approach of \cite{sperrorest}.
+As the best algorithm of 1) will be used for the spatial prediction, we first conducted the model selection on this setup and only applied the winning algorithm on 2).
 
@@
 \subsubsection{Hyperparameter tuning}
-To tune the hyperparameters of the algorithms, we used \ac{SMBO} via the R package \textit{mlrMBO} \citep{mlrMBO}.
+
+\noindent To tune the hyperparameters of the algorithms, we used \ac{SMBO} via the R package \textit{mlrMBO} \citep{mlrMBO}.
 This Bayesian approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
-After these \textit{n} tries have been evaluated, a new hyperparameter setting to be evaluated next is proposed based on the setting that performed best.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting which is going to be evaluated next is proposed based on a surrogate model.
 This strategy continues until a termination criterion, defined by the user, is reached \citep{Hutter2011, Jones1998}.
-In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting a total budget of 50 evaluated hyperparameter settings per fold.
+In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting in a total budget of 50 evaluated hyperparameter settings per fold.
 The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs such as \textit{random search} or \textit{grid search} \citep{Bergstra2012}.
@@
 
-To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
+\noindent To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
 The score is calculated by taking the contribution of each feature for each tree in the fitted model.
 The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.
-Using the internal variable importance measure from the xgboost algorithm is quite convienient since it is automatically computed during model fit.
-By contrast to other approaches such as permutation-based variable importance, it is composed out of three parts that contribute to the overall importance score:
+The variable importance measure is automatically computed during model fit.
+In contrast to other approaches such as permutation-based ones, the \textit{xgboost} score is composed out of three parts that contribute to the overall importance \citep{chenXGBoostScalableTree2016}:
 
@@
 
-The \textit{Gain} features is the most important one among the three.
-All measures sum up to one (REFERENCE!).
+\noindent The \textit{Gain} features is the most important one among the three.
+All measures sum up to one \citep{chenXGBoostScalableTree2016}.
 
@@
 
-\textit{Oiartzun} shows the highest defoliation ($\bar{x} = 69.22 \%$) among the plots while \textit{Laukiz 2} is the healthiest ($\bar{x} = 13.54 \%$) (\autoref{fig:defol_boxplots}).
+\noindent \textit{Oiartzun} shows the highest defoliation ($\bar{x} = 69.22 \%$) among the plots while \textit{Laukiz 2} is the healthiest ($\bar{x} = 13.54 \%$) (\autoref{fig:defol_boxplots}).
 All plots besides \textit{Luiando} show an evenly distributed level of defoliation across the entire plot.
 
-The high degree of defoliation of \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
+\noindent The high overall degree of defoliation of \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
 Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
-Oiartzun is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
-Laukiz 2 shows a mean tree density of 61.59 m \autoref{fig:rmse_defol_dens}) while all other plots are more dense (34.64 (Laukiz 1), 33.01 (Luiando), 34.96 (Oiartzun)) (\autoref{fig:plot-characteristics}).
+\textit{Oiartzun} is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
+\textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots are more dense (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
 
@@
 \centering
-\caption[t]{Predictive performance of \textit{xgboost} using all observations (All Observations) and observations from specific plots only (Single Plot Observations) with \ac{RMSE} as the error measure.
+\caption[t]{Predictive performance of \textit{xgboost} using all observations and all variables (All Observations/all variables), all observations and the seven most important variables only (All Observations/7 variables) and observations from specific plots only (Plot level observation/all variables) with \ac{RMSE} as the error measure.
 The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set.
-Column "Single Plot Observations", shows the mean performances at the repetition level of a spatial CV (5 folds, 5 repetitions), scored by using data of the respective plot only.}
+Column "Plot level observation", shows the mean RMSE estimates at the repetition level of a five-fold five-times repeated spatial CV, scored by using data of the respective plot only.}
 \begingroup\footnotesize
@@
 
-The \textit{xgboost} algorithm shows the lowest error (33.26 RMSE) when benchmarking the learners on the complete dataset of all plots (\autoref{tab:model_comparison}).
-While the SVM performance is only slightly worse (36.23 RMSE), RR shows a large drop in performance compared to xgboost (59.10 RMSE).
+\noindent The \textit{xgboost} algorithm showed the lowest error (33.26 RMSE) when benchmarking the learners on the complete dataset of all plots (\autoref{tab:model_comparison}).
+While the \textit{SVM} performance was only slightly worse (36.23 RMSE), \textit{RR} showed a substantially worse performance than \textit{xgboost} (59.10 RMSE).
 
@@
 
-When comparing the mean predictive performance of models fitted at the plot level against the performance of the model that was fitted using all data, the plot-level models show a better performance in all cases (\autoref{tab:supermodel_performance}).
-The highest difference between both models types occurs for \textit{Laukiz2} with a difference of 34.51 RMSE.
+\noindent When comparing the mean predictive performance of models fitted at the plot level against the performance of the model that was fitted using all data (super model), the plot-level models showed a better performance in all cases (\autoref{tab:supermodel_performance}).
+The highest difference between both model types occurred for plot \textit{Laukiz2} with a difference of 34.51 RMSE.
 
-Using only the seven most important variables (\autoref{fig:var-imp}) for the super model shows small increases in performance for Laukiz 1 and Laukiz 2, a small decrease for Luiando and almost a reduction of 50\% of the error for Oiartzun (32.97 vs.
+Using only the seven most important variables (\autoref{fig:var-imp}) for the super model showed small increases in performance for \textit{Laukiz 1} and \textit{Laukiz 2}, a small decrease for \textit{Luiando} and almost a reduction of 50\% of the error for \textit{Oiartzun} (32.97 vs.
 17.62 RMSE) (\autoref{tab:supermodel_performance}).
@@
 
-\noindent An increase of the error rate was observed with an increase of descriptive plot measures such as mean point density and the coefficient of variation (based on the response \textit{defoliation}) (\autoref{fig:plot-characteristics}).
+\noindent An increase of the error rate was observed with an increase of descriptive plot measures such as mean point density and the coefficient of variation (based on the response variable \textit{defoliation}) (\autoref{fig:plot-characteristics}).
 
-\begin{figure} [b!]
+\begin{figure} [t!]
 	\begin{center}
@@
 
-\noindent The seven most important features of the super model in this study were vegetation indices with \textit{EVI} \citep{hueteComparisonVegetationIndices1997} being the most important one (\autoref{fig:var-imp}).
+\noindent The seven most important features of the super model in this study were vegetation indices with the \textit{EVI} \citep{hueteComparisonVegetationIndices1997} being the most important one (\autoref{fig:var-imp}).
 
@@
 
-\noindent The seven most important features (\textit{EVI}, \textit{GDVI4}, \textit{D1}, \textit{GDVI3}, \textit{GDVI2}, \textit{mNDVI} and \textit{mSR}) show a substantial difference compared to all following variables (\autoref{fig:var-imp}).
+\noindent The seven most important features (\textit{EVI}, \textit{GDVI4}, \textit{D1}, \textit{GDVI3}, \textit{GDVI2}, \textit{mNDVI} and \textit{mSR}) show a substantial difference in the importance score compared to all following variables (\autoref{fig:var-imp}).
 
@@
 		\caption{The 30 most important variables as estimated by the internal variable importance measure of the \textit{xgboost} algorithm. The higher the score, the more important the feature.
-			"bf2" means that a buffer of 2 meter was used to extract the variable information to the tree observation.
+			"bf2" notes that a buffer of 2 meter was used to extract the variable information to the tree observation.
 "NRI" means that a normalized ratio index with the subsequent bands was calculated.
@@
 
-The best NRI scored rank eight (band 112 and band 62).
+\noindent The best NRI scored rank eight (band 112 and band 62).
 All further places up to rank 30 are NRIs.
@@
 
-The buffer of 2m that we used to generate the index value for each observation can be seen critical.
-When using now buffer at all, the possibility exists that a pixel value is assigned to the tree observation that does not spatially match with the tree observation.
-Using a buffer of more than 2 meters would cause information from other trees to be merged into the pixel value, blurring the actual value of the tree observation.
-Thats why using a buffer of 2m was the best compromise here in our perspective.
+\noindent The buffer of 2 m that we used to generate the index value for each observation can be seen critical.
+When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation that does not spatially match.
+Using a buffer of more than 2 meters would increase the probability of merging information from other trees into the pixel value, blurring the actual value of the tree observation.
+Thats why in our view using a buffer of 2 m was the best compromise here.
 
 The exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels will be used to calculate the final pixel value.
-If a tree observation is located at the border of the plot, some directions of the buffer will not contain image values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
+As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels are touched by the circle that will then be used to calculate the final pixel value.
+If a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
 
-Using a buffer in the first place and the fact that it is unclear how much information from other trees went into the calculation of an index value of a certain observation certainly limits the meaningfullness of the study and has to be considered when making interpretations about the outcome of this study.
+Using a buffer in the first place and the fact that it is unclear how much information from other trees went into the calculation of an index value of a certain observation introduced a bias of an unknown magnitude into the data.
+This has to be considered when making interpretations about the outcome of this study.
 
@@
 
-Trying to relate the modeling error to plot characteristics (mean point density, CV) did not show a clear picture: For both comparisons, Laukiz 2 did not follow the pattern that was observed from the other three plots (\autoref{fig:plot-characteristics}) of having an increase in error with an increase in mean point density and CV.
+\noindent Relating the modeling error to plot characteristics (mean point density, coefficient of variation) did not show a clear picture: For both comparisons, \textit{Laukiz 2} did not follow the pattern that was observed from the other three plots (\autoref{fig:plot-characteristics}) of having an increase in error with an increase in mean point density and coefficient of variation.
 
-Having an increase in error with an increase in mean point density could be related to the buffer we used in this study.
-If a tree is surrounded by bare ground, chances are higher that bare ground information will be included in the final index value.
-By contrast, if the density of trees in a plot is high, the possibility of bare ground information in the calculated index value will be small.
-The fact of having information from another tree in the final pixel value is less problematic than having bare ground information as the latter affects the final index calculation more in a negative way.
-
-Nevertheless, it has to be considered that we only had four plots in this work.
-To make a valid statement about a relationship between modeling error and these plot characteristics, a larger sample size of plots is needed.
+It needs to be considered that we only looked at four plots in this work.
+To make a robust statement about a possible relationship between modeling error and plot characteristics, a larger sample size of plots is needed.
 
@@
 
-The relatively large difference in performance between RR (59.10 RMSE) and the machine-learning models (36.23 and 33.26 RMSE) is surprising.
-RR has shown promising performance results in other studies when many highly-correlated predictors were involved (REFERECES).
-However, in this study, RR was not able to achieve a sufficient performance score compared to SVM and xgboost even though its hyperparameter $\lambda$ was properly tuned using SMBO.
+\noindent The relatively large difference in performance between \textit{RR} (59.10 RMSE) and the machine-learning models (36.23 and 33.26 RMSE) is remarkable.
+\textit{RR} has shown promising performance results in other studies when many highly-correlated predictors were involved (REFERECES).
+However, in this study, \textit{RR} was not able to achieve a sufficient performance score compared to \textit{SVM} and \textit{xgboost} even though its hyperparameter $\lambda$ was properly tuned using SMBO.
 
-While xgboost shows a slightly better performance than SVM, the latter has the advantage of only having two hyperparameters that need to be tuned.
+While \textit{xgboost} shows a slightly better performance than \textit{SVM}, the latter has the advantage of only having two hyperparameters that need to be tuned.
 This results in a shorter runtime.
-Nevertheless, xgboost showed the best performance and was subsequently used to fit the models on the plot level and for the prediction.
+Nevertheless, \textit{xgboost} showed the best performance and was subsequently selected to fit the models on the plot level and for the spatial prediction.
 

### 2018-08-05 02_scripts/035-all-plots-modeling.Rmd
finish prediction, create plots

@@
 
-## RF
+## xgboost 7 vars only
 
 ```{r}
-# RF
-ps_rf <- makeParamSet(
-  makeIntegerParam("mtry", lower = 1, upper = 3000),
-  makeIntegerParam("min.node.size", lower = 1, upper = 10),
-  makeNumericParam("sample.fraction", lower = 0.2, upper = 0.9)
-)
-ctrl <- makeMBOControl(propose.points = 1L)
-ctrl <- setMBOControlTermination(ctrl, iters = 20L)
-ctrl <- setMBOControlInfill(ctrl, crit = crit.ei)
-ctrl <- setMBOControlMultiPoint(ctrl, method = "cl", cl.lie = min)
-tune.ctrl <- makeTuneControlMBO(
-  mbo.control = ctrl,
-  mbo.design = generateDesign(n = 30, par.set = ps_rf)
-)
-inner <- makeResampleDesc("Blocking")
+data_7 = data %>% 
+  select(defoliation, bf2_EVI, bf2_GDVI_4, bf2_GDVI_3, bf2_GDVI_2, bf2_D1, bf2_mNDVI, bf2_mSR)
 
-tune_wrapper_rf <- makeTuneWrapper(makeLearner("regr.ranger"),
-                                   resampling = inner, par.set = ps_rf,
-                                   control = tune.ctrl, show.info = TRUE,
-                                   measures = list(rmse)
-)
+task_7 = makeRegrTask(id = "all_plots", data = data_7, 
+                    target = "defoliation", coordinates = coords,
+                    blocking = factor(rep(1:4, c(479, 451, 291, 529))))
+
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 4, level = "mlr.resample",
+              mc.set.seed = TRUE)
+set.seed(12345)
+resa_svm <- mlr::resample(tune_wrapper, task_7,
+                          resampling = outer, extract = getTuneResult,
+                          show.info = TRUE, measures = list(rmse))
+parallelStop()
+saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/all_plots_xgboost_blocking_7vars.rda")
 ```
 
+#### Tuning
+
 ```{r}
-outer <- makeResampleDesc("Blocking")
+configureMlr(on.learner.error = "warn", on.error.dump = TRUE)
+library(parallelMap)
+parallelStart(mode = "multicore", level = "mlrMBO.feval", cpus = 48,
+              mc.set.seed = TRUE)
+set.seed(12345)
+xgboost_tuned = tuneParams(lrn_xgboost, task = task_7, resampling = inner,
+                           par.set = ps_xgboost, control = tune.ctrl, 
+                           show.info = TRUE, measure = list(rmse))
+parallelStop()
+saveRDS(xgboost_tuned, "/data/patrick/mod/tmp/tuning_hyperspectral/xgboost_tuned_blocking_7vars.rda")
 ```
 
+#### Train
+
 ```{r}
-library(parallelMap)
-parallelStart(mode = "multicore", cpus = 40, level = "mlrMBO.feval",
-              mc.set.seed = TRUE) # only parallelize the tuning
-set.seed(12345)
-resa_svm <- mlr::resample(tune_wrapper_rf, task,
-                          resampling = outer, extract = getTuneResult,
-                          show.info = TRUE, measures = list(rmse))
+xgboost_tuned = readRDS("/data/patrick/mod/tmp/tuning_hyperspectral/xgboost_tuned_blocking.rda")
 
-parallelStop()
-saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/all_plots_rf_blocking.rda")
+lrn_xgboost = setHyperPars(makeLearner("regr.xgboost"), 
+                           par.vals = xgboost_tuned$x)
+m = train(lrn_xgboost, task)
+saveRDS(m, "/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned_7vars.rda")
 ```

### 2018-08-05 02_scripts/08_prediction.Rmd
finish prediction, create plots

@@
 library(pbmcapply)
+library(magrittr)
+library(dplyr)
+library(data.table)
+library(rasterVis)
+library(viridis)
+library(ggpubr)
 ```
@@
 ```{r}
-mod = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
+mod = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned_7vars.rda")
 ```
@@
 
+Prefix all vars with "bf2" so that they match with the names of the fitted model
+
+```{r}
+veg_inds %<>% 
+  map(~ select_all(.x, .funs = funs(paste0("bf2_", .))))
+```
+
+Use only the indices used for training
+
+```{r}
+veg_inds %<>% 
+  map(~ dplyr::select(.x, bf2_EVI, bf2_GDVI_4, bf2_GDVI_3, bf2_GDVI_2, 
+                      bf2_mNDVI, bf2_mSR, bf2_D1))
+```
+
 # Load NRI
 
+We only use the 7 most important variables. No NRI among them.
+
 ```{r}
-files_nri = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
-                   pattern = ".grd", full.names = TRUE)
-
-nri = map(files_nri, ~ brick(.))
+# files_nri = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
+#                    pattern = ".grd", full.names = TRUE)
+# 
+# nri = map(files_nri, ~ brick(.))
 ```
@@
 ```{r}
-nri %<>% 
-  pbmclapply(function(x) as(x, "SpatialPixelsDataFrame"), mc.cores = 2) %>% 
-  pbmclapply(function(x) st_as_sf(x), mc.cores = 2)
+# if (file.exists("/data/patrick/mod/hyperspectral/prediction/nri_basque.rda")) {
+#   nri = readRDS("/data/patrick/mod/hyperspectral/prediction/nri_basque.rda")
+# } else {
+#   nri %<>% 
+#     pbmclapply(function(x) as(x, "SpatialPixelsDataFrame"), mc.cores = 2) %>% 
+#     pbmclapply(function(x) st_as_sf(x), mc.cores = 2)
+# }
 ```
@@
 
-nri %<>% 
-  set_names(names)
+# nri %<>% 
+#   set_names(names)
 
@@
 
-# Combine NRI and veg inds
+<!-- # Combine NRI and veg inds -->
+
+<!-- ```{r} -->
+<!-- if (file.exists("/data/patrick/mod/hyperspectral/prediction/prediction_df.rda")) { -->
+<!--   inds_combined = readRDS("/data/patrick/mod/hyperspectral/prediction/prediction_df.rda") -->
+<!-- } else { -->
+<!--   #options(future.globals.maxSize = 55000 * 1024^2) -->
+<!--   #plan(multicore, workers = 2) -->
+<!--   inds_combined = map2(veg_inds, nri, ~ st_join(.x, .y)) -->
+<!--   saveRDS(inds_combined, "/data/patrick/mod/hyperspectral/prediction/prediction_df.rda") -->
+<!-- } -->
+<!-- ``` -->
+
+<!-- Prefix all vars with "bf2" so that they match with the names of the fitted model -->
+
+<!-- ```{r} -->
+<!-- inds_combined %<>%  -->
+<!--   map(~ select_all(.x, .funs = funs(paste0("bf2_", .)))) -->
+<!-- ``` -->
+
+<!-- Use only the indices used for training -->
+
+<!-- ```{r} -->
+<!-- indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda") -->
+
+<!-- inds_combined %<>%  -->
+<!--   map(~ dplyr::select(.x, indices_shared)) -->
+<!-- ``` -->
+
+Extract coordinates of all plots
+
+```{r}
+pred_df_x = map(veg_inds, ~ as.integer(st_coordinates(.x)[, 1]))
+pred_df_y = map(veg_inds, ~ as.integer(st_coordinates(.x)[, 2]))
+
+coords2 = map2(pred_df_x, pred_df_y, ~ data.frame(.x, .y))
+coords2 = map(coords2, ~ dplyr::rename(.x, x = .x, y = .y))
+```
+
+# Predict
+
+Make prediction data.frame
+
+```{r}
+veg_inds %<>% 
+  map(~ st_set_geometry(.x, NULL))
+
+saveRDS(veg_inds, "/data/patrick/mod/modeling/paper1_2018/prediction_df.rda")
+```
+
+# Standardize variables
+
+```{r}
+veg_inds %<>% 
+  map(~ as.data.table(.x)) %>% 
+  pbmclapply(function(x) {
+    cols = names(x)[2:length(names(x))]
+    x[, (cols) := lapply(.SD, scale), .SDcols=cols]
+  }, mc.cores = 28
+      ) %>% 
+  map(~ as(.x, "data.frame"))
+```
+
+```{r}
+#plan(multicore, workers = 28)
+
+# correct column order of features: https://github.com/dmlc/xgboost/issues/1809
+veg_inds %<>% 
+  map(~ .x[mod$learner.model$feature_names])
+
+pred = map(veg_inds, ~ predict(mod$learner.model, newdata = as.matrix(.x)))
+```
+
+# Combine prediction results with coords
+
+```{r}
+all = map2(coords2, pred, ~ cbind(.x, .y)) %>% 
+  map(~ rename(.x, defoliation = .y))
+```
+
+# Transfer to SPDF and export raster
+
+```{r}
+all_spdf = map(all, ~ SpatialPixelsDataFrame(points = .x[, c("x", "y")], 
+                                             data = as.data.frame(.x[, "defoliation"]), 
+                                             proj4string = CRS("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"))) %>% 
+  map(~ set_names(.x, "defoliation"))
+
+# create raster objects
+raster_list <- map(all_spdf, ~ raster(.x))
+
+# write out names as there are some probs doing it dynamically in walk
+names = map_chr(files, ~ tools::file_path_sans_ext(basename(.x)))
+
+# WHY DOES this work??
+iwalk(raster_list, ~ writeRaster(.x, glue::glue("/data/patrick/mod/hyperspectral/prediction/defol-pred-{.y}.tif"),
+                          overwrite = TRUE)
+)
+```
+
+# Plot demonstration plots
+
+```{r}
+# lattice.options(
+#   layout.heights=list(bottom.padding=list(x=0), top.padding=list(x=0)),
+#   layout.widths=list(left.padding=list(x=0), right.padding=list(x=0))
+# )
+
+raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
+
+name = c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
+
+maps = imap(raster_list_wgs84[c(15, 16, 18, 19)], ~ levelplot( 
+  .x[["defoliation"]], 
+  main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
+  margin = FALSE,                             
+  between = list(x = 0.0, y = 0.0), 
+  at = seq(30, 80, length.out = 100),
+  colorkey = list(
+    width = 0.75,                       
+    space = 'right',  
+    labels = list(#at = c(0, 25, 50, 75, 100),
+                font = 4),      
+    axis.line = list(col = 'black')
+  ),
+  par.strip.text = list(col = "white", cex = 1.2),    
+  col.regions = viridis)    
+)
+```
+
+```{r}
+png("../03_figures/results/predictions_demon_plots.png", units = "in", width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps, nrow = 2, ncol = 2)
+dev.off()
+
+```
+
+## Histograms
+
+```{r}
+df_all= map(all_spdf[c(15, 16, 18, 19)], ~ as(.x, "data.frame"))
+
+hists_all = imap(df_all, ~ gghistogram(.x, x = "defoliation", 
+                                       title = tools::toTitleCase(.y)))
+
+patchwork::wrap_plots(hists_all)
+```

### 2018-08-05 02_scripts/09_vis_sp_partitions.Rmd
finish prediction, create plots

---
title: "03-study area"
output:
  html_notebook:
    code_folding: none
    fig_width: 10
    highlight: haddock
    number_sections: yes
    toc: yes
---

# Packages

```{r}
devtools::install_github("mlr-org/mlr@fix-create-resamp-plots")
library(mlr)
library(tibble)
library(data.table)
library(cowplot)
library(ggrepel)
library(magrittr)
```

# Visualize partitions

Here we visualize the partitioning of the first repetition (5 folds) in the performance estimation level of the spatial and non-spatial cross-validation setting.

## Blocking

```{r fig.height=3, fig.width=13}
data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")

data = as_tibble(rbindlist(data, fill = TRUE))
data = Filter(function(x) !any(is.na(x)), data)
coords = as_tibble(rbindlist(coords))
data$plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(479, 451, 291, 529)))
data = cbind(data, coords)

data %<>% 
  #dplyr::slice(c(200, 700, 1100, 1400)) %>% 
  dplyr::select(defoliation, plot, X, Y)

#rownames(data) = c("200", "700", "1100", "1400")

task = makeRegrTask(id = "all_plots", data = data, 
                    target = "defoliation", coordinates = data[, c("X", "Y")])

resamp_blocking <- readRDS("/data/patrick/mod/hyperspectral/spcv/all_plots_svm_blocking.rda")

plist <- createSpatialResamplingPlots(task, list("Block CV" = resamp_blocking),
                                      crs = 32630, repetitions = 1, point.size = 1.5,
                                      y.axis.breaks = c(43.35, 43.1), 
                                      x.axis.breaks = c(-2.8, -2),
                                      axis.text.size = 18)

# We need to subset the data with geom_label_repel to only plot 4 repels
# We need to make sure to select one instance of every plot
plist[["Plots"]] = purrr::map(plist[["Plots"]], ~ .x + 
                         geom_label_repel(aes(x = X, y = Y, label = plot), 
                                          min.segment.length = unit(1, "lines"),
                                          data = subset(.x$data, rownames(.x$data) %in% 
                                                          c(200, 700, 1100, 1400))) +
                           labs(y = "", x = "") +
                           #theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
                           #hrbrthemes::theme_ipsum_rc(plot_margin = margin(0, 0, 0, 0))
                       )

plots = plot_grid(plotlist = plist[["Plots"]], labels = plist[["Labels"]],
                  nrow = 2, ncol = 2)

plots = patchwork::wrap_plots(plist[["Plots"]])

save_plot("../03_figures/data/blocking_folds.png", 
          plots, ncol = 2, nrow = 2)
```

## sp sp

```{r}
data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")

data = data[[4]]
coords = coords[[4]]

data %<>% 
  dplyr::select(defoliation)

task = makeRegrTask(id = "oiartzun", data = data, 
                    target = "defoliation", coordinates = coords)

resamp_sp <- readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_xgboost.rda")

plist <- createSpatialResamplingPlots(task, list("SpCV" = resamp_sp),
                                      crs = 32630, repetitions = 1, point.size = 1.5,
                                      #y.axis.breaks = c(42.8, 43.2), 
                                      #x.axis.breaks = c(-2, -3),
                                      axis.text.size = 18)

plot_grid(plot)
```

# for presentation

```{r}
sp_pres <- plot_grid(plist[["Plots"]][[1]], plist[["Plots"]][[2]], plist[["Plots"]][[3]], 
                plist[["Plots"]][[4]], plist[["Plots"]][[5]], ncol = 5, nrow = 1, 
                labels = plist[["Labels"]][1:5], label_size = 18)

nsp_pres <- plot_grid(plist[["Plots"]][[6]], plist[["Plots"]][[7]], plist[["Plots"]][[8]], 
                 plist[["Plots"]][[9]], plist[["Plots"]][[10]],
                 ncol = 5, nrow = 1, labels = plist[["Labels"]][6:10],
                 label_size = 18)

final_plot_pres <- plot_grid(sp_pres, nsp_pres, ncol = 1)

save_plot("../04_figures/01_data/spcv_nspcv_folds_pres.png", 
          final_plot_pres, ncol = 5, nrow = 2)
```

# PNG to PDF conversion

Converts all pngs in the specified folder to pdfs. 
`&&` makes the "convert" call conditional on `cd`.
`exec` reduces memory waste of shell.
Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.

```{r png to pdf, echo = FALSE, eval = FALSE}
library(furrr)
library(magick)
imgs = list.files("/home/patrick/PhD/papers/01_model_comparison/04_figures/01_data", 
                  pattern = ".png", full.names = TRUE)

plan(multiprocess, workers = length(imgs))
future_walk(imgs, ~ image_write(image_read(.x), 
                                path = stringr::str_replace(.x, ".png", ".pdf"), 
                                format = "pdf"))

system("cd /home/patrick/PhD/papers/01_model_comparison/04_figures/01_data && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

system("cd /home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/ && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

system("cp /home/patrick/PhD/papers/01_model_comparison/04_figures/01_data/study_area.png /home/patrick/PhD/presentations/paper1/figs/study_area.png")
system("cp /home/patrick/PhD/papers/01_model_comparison/04_figures/01_data/spcv_nspcv_folds_pres.png /home/patrick/PhD/presentations/paper1/figs/spcv_nspcv_folds_pres.png")
```

### 2018-08-06 01_writing/Schratz_hyperspectral.tex
add spatial prediction on a plot level results:: spatial prediction discussion:: spatial prediction

@@
 
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width = 1\textwidth] {predictions_demon_plots.pdf}}
+		\caption{Spatially predicted defoliation (in \%) from \textit{xgboost} of \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
+		\label{fig:pred_demon_plots}
+	\end{center}
+\end{figure}
+
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width = 0.6\textwidth] {predictions_demon_plots_hists.pdf}}
+		\caption{Histograms of predicted defoliation (in \%) from \textit{xgboost} of \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
+		\label{fig:pred_demon_plots_hists}
+	\end{center}
+\end{figure}
+
+\noindent The plots with a higher mean defoliation (\textit{Luiando} and \textit{Oiartzun}) showed a good visual separability compared to the healthier plots \textit{Laukiz 1} and \textit{Laukiz 2} (\autoref{fig:pred_demon_plots}).
+All predicted values ranged between 30 \% and 80 \% with \textit{Luiando} showing the smallest variance (\autoref{fig:pred_demon_plots_hists}).
+
+The high error of the super model for \textit{Laukiz 2} (49.94 RMSE) is also visible in the respective histogram as most predictions range between 40 \% - 50 \% (\autoref{fig:pred_demon_plots_hists}) while in fact most trees of \textit{Laukiz 2} show an actual defoliation of around 20 \% (\autoref{fig:defol_boxplots}).
+
+For the less defoliated plots \textit{Laukiz 1} and \textit{Laukiz 2} a subtle level of separation between trees and bare ground is visible (\autoref{fig:pred_demon_plots}).
+This is not the case for the other two higher defoliated plots for which bare ground and defoliated predictions are mainly in the same value range (around 60 \% - 80 \%).
+
+[ADD FIGURE WITH SENTINEL2 PREDICTION]
+
 \section{Discussion}
@@
 An important point that needs to be considered when interpreting the performance results is that we only related defoliation to indices derived from remote sensing data.
-Possible other cofounding factors that could possibly help in predicting defoliation were not considered.
+Possible other local variables that could help in predicting defoliation were not considered.
 One example here is tree age: The older a tree the more vulnerable it may be to pathogen infections causing defoliation.
@@
 
-There are some downsides using the internal variable importance approach of xgboost: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
+\noindent There are some downsides using the internal variable importance approach of xgboost: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
 Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
@@
 However, based on this statement, we could also have made the cut between rank five and 6 but including the two vegetation indices at rank six and seven will eventually improve the model and does not increase runtime.
+
+\subsection{Spatial prediction}
+
+\noindent The spatial predictions showed that the model tries to avoid making extreme predictions since most values ranged between 30 \% and 80 \%. 
+This behavior is mainly triggered by the conentration of the model on the distribution of defoliation in the \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} training plots.
+This focus then causes the high prediction error for \textit{Laukiz 2} for which a lot of defoliation values actually range between 0 \% and 20 \%.
+The fitted model would need more training samples for plots that show a similar defoliation as \textit{Laukiz 2} to reach a better performance for similar plots.
 

### 2018-08-07 01_writing/Schratz_hyperspectral.tex
write conclusion

@@
 This approach results in four performance estimates, one for each fold.
-2) Using five-fold five-times repeated spatial \ac{CV} for each plot based on the k-means clustering approach of \cite{sperrorest}.
+2) Using five-fold five-times repeated spatial \ac{CV} for each plot based on the k-means clustering approach of \cite{sperrorest} (\autoref{fig:spcv}).
 As the best algorithm of 1) will be used for the spatial prediction, we first conducted the model selection on this setup and only applied the winning algorithm on 2).
 
-% Add figure visualizing both spCV approaches
+% defoliation boxplots
+\begin{figure} [b!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=1\textwidth] {spcv_folds_demon_plots.pdf}}
+		\caption{Fold 1 of the spatial partitioning using k-means clustering for \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
+		\label{fig:spcv}
+	\end{center}
+\end{figure}
 
@@
 	\hline
-	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) &                       \\
+	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) & 29.59   (16.09)              \\
 	\bottomrule
@@
 \noindent The spatial predictions showed that the model tries to avoid making extreme predictions since most values ranged between 30 \% and 80 \%.
-This behavior is mainly triggered by the conentration of the model on the distribution of defoliation in the \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} training plots.
+This behavior is mainly triggered by the concentration of the model on the distribution of defoliation in the \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} training plots.
 This focus then causes the high prediction error for \textit{Laukiz 2} for which a lot of defoliation values actually range between 0 \% and 20 \%.
@@
 \section{Conclusion}
+
+\noindent In this work we used various indices derived from hyperspectral remote sensing data to estimate defoliation in northern Spain.
+In the algorithm comparison \textit{xgboost} showed the best performance among the tested ones.
+Even though \ac{RR} is able to handle highly correlated data, it was far from achieving an acceptable performance in this work.
+
+The fitted models on the plot level showed a promising performance with RMSE values between 8 and 20.
+The performance of the model containing all data was acceptable (RMSE 29.59) but can be improved by adding more observations from other plots in future studies.
+
+The spatial prediction showed a satisfying result making it possible to distinguish highly defoliated plots from plots with low defoliation easily.
+The most important indices for the fitted model were widely known vegetations indices such as EVI, GDVI or NDVI.
+It would be interesting to link the indices of this work to other indicators of forest/vegetation health in future studies and analyze their importance and the model performances.
+In summary, defoliation can be successfully used as a response variable to model forest health using remote sensing indices.
 

### 2018-08-07 02_scripts/08_prediction.Rmd
update scripts

@@
 
+Plot title only
+
 ```{r}
@@
   .x[["defoliation"]], 
-  main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
+  #main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
+  main = glue::glue("{tools::toTitleCase(.y)}"), 
   margin = FALSE,                             
@@
 
+Descriptive title
+
+```{r}
+# lattice.options(
+#   layout.heights=list(bottom.padding=list(x=0), top.padding=list(x=0)),
+#   layout.widths=list(left.padding=list(x=0), right.padding=list(x=0))
+# )
+
+raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
+
+name = c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
+
+maps = imap(raster_list_wgs84[c(15, 16, 18, 19)], ~ levelplot( 
+  .x[["defoliation"]], 
+  main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
+  #main = glue::glue("{tools::toTitleCase(.y)}"), 
+  margin = FALSE,                             
+  between = list(x = 0.0, y = 0.0), 
+  at = seq(30, 80, length.out = 100),
+  colorkey = list(
+    width = 0.75,                       
+    space = 'right',  
+    labels = list(#at = c(0, 25, 50, 75, 100),
+                font = 4),      
+    axis.line = list(col = 'black')
+  ),
+  par.strip.text = list(col = "white", cex = 1.2),    
+  col.regions = viridis)    
+)
+```
+
+```{r}
+png("../03_figures/results/predictions_demon_plots_descriptive_title.png", units = "in", width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps, nrow = 2, ncol = 2)
+dev.off()
+
+```
+
 ## Histograms
@@
 
-patchwork::wrap_plots(hists_all)
+hists_grid = patchwork::wrap_plots(hists_all)
+
+cowplot::save_plot(hists_grid, filename = 
+                   "../03_figures/results/predictions_demon_plots_hists.png", 
+                   ncol = 2, nrow = 2,
+                   )
 ```
+
+# PNG to PDF conversion
+
+Converts all pngs in the specified folder to pdfs. 
+`&&` makes the "convert" call conditional on `cd`.
+`exec` reduces memory waste of shell.
+Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.
+
+```{r png to pdf, echo = FALSE, eval = FALSE}
+# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
+#        parallel convert '{}' '{.}.pdf'")
+
+system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
+       parallel convert '{}' '{.}.pdf'")
+
+# copy to presentation folder
+# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
+#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
+#           overwrite = TRUE)
+```
+
+# Session Info
+
+```{r}
+sessionInfo()
+```

### 2018-08-07 02_scripts/09_vis_sp_partitions.Rmd
update scripts

@@
 library(magrittr)
+library(purrr)
 ```
@@
                                                           c(200, 700, 1100, 1400))) +
-                           labs(y = "", x = "") +
+                           labs(y = "", x = "") 
                            #theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
@@
 
+Visualize partitions for all four plots and then take only fold 1 of each.
+
 ```{r}
+
+all = imap(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~ {
   data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
@@
 
-data = data[[4]]
-coords = coords[[4]]
+  data = data[[.y]]
+  coords = coords[[.y]]
 
@@
 
-task = makeRegrTask(id = "oiartzun", data = data, 
+  task = makeRegrTask(id = "data", data = data, 
                       target = "defoliation", coordinates = coords)
 
-resamp_sp <- readRDS("/data/patrick/mod/hyperspectral/spcv/luiando_xgboost.rda")
+  resamp_sp <- readRDS(glue::glue("/data/patrick/mod/hyperspectral/spcv/{.x}_xgboost.rda"))
 
@@
                                         crs = 32630, repetitions = 1, point.size = 1.5,
-                                      #y.axis.breaks = c(42.8, 43.2), 
-                                      #x.axis.breaks = c(-2, -3),
+                                        y.axis.breaks = c(43.3541, 43.353, 43.2648, 43.0955), 
+                                        x.axis.breaks = c(-2.93, -2.9293, -2.99577,
+                                                          -1.845),
                                         axis.text.size = 18)
+})
 
-plot_grid(plot)
+plots = plot_grid(all[[1]][["Plots"]][[1]],
+                  all[[2]][["Plots"]][[1]],
+                  all[[3]][["Plots"]][[1]],
+                  all[[4]][["Plots"]][[1]],
+                  labels = c("Fold 1 (Laukiz 1)", "Fold 1 (Laukiz 2)",
+                             "Fold 3 (Luiando)", "Fold 4 (Oiartzun)"),
+                  nrow = 1, ncol = 4)
+
+save_plot(plots, filename = "../03_figures/data/spcv_folds_demon_plots.png",
+          ncol = 4, nrow = 1)
 ```
@@
 library(magick)
-imgs = list.files("/home/patrick/PhD/papers/01_model_comparison/04_figures/01_data", 
+imgs = list.files("/home/patrick/PhD/papers/03_hyperspectral/03_figures/data", 
                   pattern = ".png", full.names = TRUE)

### 2018-08-08 01_writing/Schratz_hyperspectral.tex
add disc:comparison to other studies make plot names itlaic rearrange figures

@@
 \usepackage{amsmath}
+
+\usepackage{enumerate}
 
@@
 
-	\title{Modeling defoliation as a proxy for tree health: A case study on Pinus Radiata trees using machine-learning and hyperspectral data}
+	\title{Modeling defoliation as a proxy for tree health: A case study using machine-learning and hyperspectral remote sensing data}
 
@@
 	\acro{IQR}{Interquartile Range}
-	\acro{WKNN}{Weighted $k$-nearest neighbor}
 	\acro{MARS}{Multivariate Adaptive Regression Splines}
 	\acro{MEM}{Maximum Entropy Model}
+	\acro{NDII}{Normalized Difference Infrared Index}
+	\acro{NIR}{Near-Infrared}
 	\acro{NRI}{Normalized Ratio Indices}
@@
 
-For all analysis steps we used the open-source statistical programming language R \citep{R_core}.
+\noindent For all analysis steps we used the open-source statistical programming language R \citep{R_core}.
 The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chenXGBoostScalableTree2016} (\textit{xgboost}), \textit{kernlab} \citep{kernlab} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
@@
 It provides a standardized interface for a wide variety of statistical and machine-learning models in R simplifying essential modeling tasks such as hyperparameter tuning, model performance evaluation and parallelization \citep{bischlMlrMachineLearning2016}.
+We provide the complete code and required data as a Mendeley dataset to make this work fully reproducible \citep{}.
 
@@
 
-To use the full information from the hyperspectral data, we not only calculated all possible vegetation indices that are available in the \textit{hsdar} package (90 in total) but also all possible combinations of \ac{NRI}.
-We were interested if NRIs of arbitrary band combinations will have a substantial positive effect on the predictive power of the fitted model.
-The NRI were calculated using the following formula:
+\noindent To use the full information from the hyperspectral data, we calculated all possible vegetation indices that are available in the \textit{hsdar} package (90 in total) and all possible \ac{NRI} combinations.
+We were interested if NRIs of arbitrary band combinations will have a substantial effect on the predictive power of the fitted model.
+The NRIs were calculated using the following formula:
 
@@
 
-To account for geometric offsets (which were reported with up to 1 m), we used a buffer of two meters around the centroid of the respective tree.
-The mean value of all pixels touched by the buffer was assigned as the final value for each index.
-Missing values were removed from the mean value calculation.
-In total, 7875 Normalized Ratio Indices NRI were calculated ($\frac{125*126}{2}$).
-Due to four corrupted bands and other numerical problems, some indices returned \texttt{NA} for specific observations.
-These indices were removed from the dataset, leaving a total of 7471 variables without missing values.
+\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), we used a buffer of two meters around the centroid of the respective tree.
+The mean value of all pixels touched by the buffer was assigned as the final value of each index.
+In total, $\frac{125*126}{2} = 7875$ NRIs were calculated .
+Due to four corrupted bands and numerical problems for some band combinations, some indices returned \texttt{NA} for specific observations.
+We removed all indices from the dataset that showed one or more \texttt{NA} values (across all plots) since we valued a single observation more than having an additional NRI as a predictor variable.
+In total, 7471 indices had no \texttt{NA} values and were subsequently used as predictors.
 
@@
 
-Three algorithms (\textit{xgboost}, \textit{\ac{SVM}}, \textit{\ac{RR}}) were benchmarked on predictive performance.
-Besides the well-known \ac{SVM} algorithm \citep{Vapnik1998} we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
-We also added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates (REFERENCE).
-The probably most popular machine-learning algorithm, Random Forest \citep{Breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
+\noindent Three algorithms (\textit{xgboost}, \textit{\ac{SVM}}, \textit{\ac{RR}}) were benchmarked on their predictive performance.
+Besides the well-known \ac{SVM} algorithm \citep{Vapnik1998}, we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
+We also added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerlRidgeRegressionBiased1970}.
+One of the most popular machine-learning algorithm, Random Forest \citep{Breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
 These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{Probst2018b}.
 After the selection of the best model, we checked if the winning algorithm can achieve a similar performance when using only the most important variables compared to using all variables.
-This would simplify the spatial prediction a lot as less indices need to be computed from the data used for prediction.
-Also, simpler models are always preferred over complex ones (REFERENCE).
+A successful feature selection simplifies the spatial prediction task because the prediction dataset needs to consist of less variables.
+Furthermore, model complexity and fitting times are reduced.
 
@@
 \noindent The algorithms were benchmarked in two ways:
-1) Using spatial block \ac{CV} on the plot level with each plot serving as the test set once.
-This approach results in four performance estimates, one for each fold.
-2) Using five-fold five-times repeated spatial \ac{CV} for each plot based on the k-means clustering approach of \cite{sperrorest} (\autoref{fig:spcv}).
-As the best algorithm of 1) will be used for the spatial prediction, we first conducted the model selection on this setup and only applied the winning algorithm on 2).
 
-% defoliation boxplots
-\begin{figure} [b!]
+\begin{enumerate}
+	\item Using spatial block \ac{CV} on the plot level with each plot serving as the test set once.
+	      Subsequently, four performance estimates were retrieved, one for each fold.
+	\item Using five-fold five-time repeated spatial \ac{CV} within each plot based on the k-means clustering approach of \cite{sperrorest} (\autoref{fig:spcv}).
+\end{enumerate}
+
+\noindent As we used the best algorithm of 1) for the spatial prediction, we first conducted the model selection on this setup and only applied the winning algorithm on 2).
+
+\begin{figure} [t!]
 	\begin{center}
@@
 This Bayesian approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
-After these \textit{n} tries have been evaluated, a new hyperparameter setting which is going to be evaluated next is proposed based on a surrogate model.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
+The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
+Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
 This strategy continues until a termination criterion, defined by the user, is reached \citep{Hutter2011, Jones1998}.
 In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting in a total budget of 50 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs such as \textit{random search} or \textit{grid search} \citep{Bergstra2012}.
+The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs, such as random search or grid search \citep{Bergstra2012}.
 
@@
 
-\noindent The high overall degree of defoliation of \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
-Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to Laukiz 1 and Laukiz 2.
+\noindent The high degree of defoliation for \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
+Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to \textit{Laukiz 1} and \textit{Laukiz 2}.
 \textit{Oiartzun} is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
-\textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots are more dense (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
+
+\textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots have a higher density (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
 
@@
 \centering
-\caption[t]{Four-fold block spatial \ac{CV} performances of RR, SVM and xgboost using \ac{RMSE} as the error measure.
+\caption[t]{Spatial block \ac{CV} performances of \textit{RR}, \textit{SVM} and \textit{xgboost} using \ac{RMSE} as the error measure.
 	Mean and standard deviation are shown.}
@@
 
-\begin{table}[t!]
+\begin{table}[b!]
 \centering
 \caption[t]{Predictive performance of \textit{xgboost} using all observations and all variables (All Observations/all variables), all observations and the seven most important variables only (All Observations/7 variables) and observations from specific plots only (Plot level observation/all variables) with \ac{RMSE} as the error measure.
-The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set.
-Column "Plot level observation", shows the mean RMSE estimates at the repetition level of a five-fold five-times repeated spatial CV, scored by using data of the respective plot only.}
+The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set (block CV).
+Column "Plot level observations", shows the mean RMSE estimates at the repetition level of a five-fold five-time repeated spatial CV, scored by using data of the respective plot only.}
 \begingroup\footnotesize
@@
 
-\noindent When comparing the mean predictive performance of models fitted at the plot level against the performance of the model that was fitted using all data (super model), the plot-level models showed a better performance in all cases (\autoref{tab:supermodel_performance}).
-The highest difference between both model types occurred for plot \textit{Laukiz2} with a difference of 34.51 RMSE.
+\noindent Comparing the mean predictive performance of models fitted at the plot level against the performance of the model that was fitted using all data (super model), the plot-level models showed a better performance in all cases (\autoref{tab:supermodel_performance}).
+The highest difference between both datasets occurred for plot \textit{Laukiz 2} with a difference of 34.51 RMSE.
 
@@
 
-\noindent The seven most important features (\textit{EVI}, \textit{GDVI4}, \textit{D1}, \textit{GDVI3}, \textit{GDVI2}, \textit{mNDVI} and \textit{mSR}) show a substantial difference in the importance score compared to all following variables (\autoref{fig:var-imp}).
+\noindent The seven most important features (\textit{EVI}, \textit{GDVI\_4}, \textit{D1}, \textit{GDVI\_3}, \textit{GDVI\_2}, \textit{mNDVI} and \textit{mSR}) showed a substantial difference in the importance score compared to all following variables (\autoref{fig:var-imp}).
 
@@
 \noindent The best NRI scored rank eight (band 112 and band 62).
-All further places up to rank 30 are NRIs.
+All further places up to rank 30 were occupied by NRIs.
 
@@
 \noindent The plots with a higher mean defoliation (\textit{Luiando} and \textit{Oiartzun}) showed a good visual separability compared to the healthier plots \textit{Laukiz 1} and \textit{Laukiz 2} (\autoref{fig:pred_demon_plots}).
-All predicted values ranged between 30 \% and 80 \% with \textit{Luiando} showing the smallest variance (\autoref{fig:pred_demon_plots_hists}).
+All predicted values ranged between 30 \% and 80 \% defoliation with \textit{Luiando} showing the smallest variance (\autoref{fig:pred_demon_plots_hists}).
 
-The high error of the super model for \textit{Laukiz 2} (49.94 RMSE) is also visible in the respective histogram as most predictions range between 40 \% - 50 \% (\autoref{fig:pred_demon_plots_hists}) while in fact most trees of \textit{Laukiz 2} show an actual defoliation of around 20 \% (\autoref{fig:defol_boxplots}).
+The high error of the super model for \textit{Laukiz 2} (49.94 RMSE) is also visible in the respective histogram as most predictions range between a defoliation of 40 \% - 50 \% (\autoref{fig:pred_demon_plots_hists}) while in fact most trees of \textit{Laukiz 2} show an actual defoliation of around 20 \% (\autoref{fig:defol_boxplots}).
 
@@
 \noindent The buffer of 2 m that we used to generate the index value for each observation can be seen critical.
-When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation that does not spatially match.
+When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation that does not spatially match (due to the geometric offset of 1 m in the hyperspectral data).
 Using a buffer of more than 2 meters would increase the probability of merging information from other trees into the pixel value, blurring the actual value of the tree observation.
@@
 
-The exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels are touched by the circle that will then be used to calculate the final pixel value.
+another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
+As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels are touched by the buffer.
 If a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
 
-Using a buffer in the first place and the fact that it is unclear how much information from other trees went into the calculation of an index value of a certain observation introduced a bias of an unknown magnitude into the data.
+All these points introduced a bias of an unknown magnitude into the data.
 This has to be considered when making interpretations about the outcome of this study.
@@
 
-It is expected that models that were trained on the plot level only achieve a better performance than the global model.
-The low performance on Laukiz 2 for the super model is most likely due to the difference of this plot to all others: The fitted model on Laukiz 1, Luiando and Oiartzun is not capable of reaching a good performance on the Laukiz 2 prediction data.
-This is not surprising as Laukiz 2 shows substantially different plot characteristics compared to all others plots in terms of the distribution of the response variable \textit{defoliation} (\autoref{fig:defol_boxplots}) and the mean point density of trees (\autoref{fig:plot-characteristics}).
+\noindent It is expected that models that were trained using observations from the respective plot only only achieve a better performance than the super model which was trained on observations from multiple plots (\autoref{tab:supermodel_performance}).
+The low performance on \textit{Laukiz 2} for the super model is most likely due to the difference of this plot to all others: The fitted model on \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} is not capable of reaching a good performance when \textit{Laukiz 2} is the evaluation dataset.
+This is not surprising as \textit{Laukiz 2} shows substantially different plot characteristics compared to all others plots in terms of the distribution of the response variable \textit{defoliation} (\autoref{fig:defol_boxplots}) and the mean point density of trees (\autoref{fig:plot-characteristics}).
 
-The low error for Luiando (8.30 RMSE) for the plot model validates the approach of relating defoliation to vegetation indices and NRI.
-The overall error of the super model (33.26 RSME) is expected to decrease if more plots are available for training.
-To reach an optimal performance, the fitted model would need to include at least one instance of every plot that shows unique characteristics (i.e.
-here Laukiz 2 is substantially different to the others).
-Possibly more plots showing unique characteristics exists that are not integrated into the fitted model of this work.
+The low error for Luiando (8.30 RMSE) for the plot model validates the approach of relating defoliation to vegetation indices and NRIs.
+The overall error of the super model (33.26 RSME) is expected to decrease if more plots would be available for training.
+However, even if the fitted model would include at least one instance of every plot showing unique characteristics (e.g.
+here \textit{Laukiz 2} is substantially different to the others), the overall error would not become smaller than the error achieved when using data from the respective plot only.
 
-An interesting find is that the supermodel with only seven variables shows a better overall performance than the model with all 7471 variables (\autoref{tab:supermodel_performance}).
-This leads to the conclusion that adding as many variables as possible to a model will not necessarily improve its performance.
-In contrast, too much information can even be problematic for the model as it will have a hard time prioritizing variables.
+An interesting find is that the model with only seven variables shows a better overall performance than the model with all 7471 variables (\autoref{tab:supermodel_performance}).
+This leads to the conclusion that adding as many variables as possible to a model will not necessarily improve its performance but instead add noise to the model.
+Too much information can be problematic for a model as it will have a hard time distinguishing between noise and important information in the variables \citep{liFeatureSelectionData2017, guyonIntroductionVariableFeature2003}.
 However, to find the most important variables in the first place and to check for the performance difference, a model with all variables needs to fitted first.
-This scenario cannot be generalized and different results may occur for other datasets.
-Using a model with only a few predictors does not only simplify prediction tasks but also reduces runtime for hyperparameter tuning and performance estimation.
+Using a model with only a few predictors does not only simplify prediction tasks but also reduces runtime for hyperparameter tuning and performance estimation \citep{liuComputationalMethodsFeature2007}.
 
@@
 
-\noindent There are some downsides using the internal variable importance approach of xgboost: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
+\noindent There are some downsides using the internal variable importance approach of \textit{xgboost}: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
 Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
-Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal xgboost approach was sufficient for this work.
+Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal \textit{xgboost} approach was sufficient for this work.
 
-It is not surprising that vegetation indices are most important for the model as they are most sensitive to changes in vegetation health.
-Even though we are not directly looking at vegetation health but using the level of defoliation as a proxy, these indices seem to help the model most deciding whether a tree is defoliated or not.
-Vegetation indices can help here in two ways:
-1) Trees that show a high level of defoliation do also reflect their bad health status through the remaining foliation.
-2) Defoliated trees have more influence of bare ground information in their pixel values and will therefore be classified as defoliated by the model.
+It is expected that vegetation indices are most important for the model as these are most sensitive to changes in vegetation health \citep{croftApplicabilityEmpiricalVegetation2014}.
+Even though we are not directly looking at vegetation health but using the level of defoliation as a proxy for tree health, vegetation indices were most important for the fitted model of this study (\autoref{fig:var-imp}).
 
-Even though no NRI made it among the most important variables in this study (stating that the first seven of this study are the most important ones), it is interesting that all ranks from 8 - 30 are occupied by NRI (\autoref{fig:var-imp}).
+\noindent Vegetation indices can help assessing defoliation in two ways:
 
-Our statement saying that the important indices of this study are the first seven can be seen critical as we only based the selection on a visual inspection of the variable importance results (\autoref{fig:var-imp}).
-The decision to make a cut between rank seven and eight was based on a combination of two facts: 1) Using only vegetation indices is easier for large scale predictions using satellites like Sentinel-2 (most NRI cannot be used with it) and 2) the drop in the importance score of the variable importance results (\autoref{fig:var-imp}).
-However, based on this statement, we could also have made the cut between rank five and 6 but including the two vegetation indices at rank six and seven will eventually improve the model and does not increase runtime.
+\begin{itemize}
+	\item Trees that show a high level of defoliation do also reflect their bad health status through the remaining foliation.
+	\item Defoliated trees have more influence of bare ground information in their pixel values and will therefore be classified as defoliated by the model.
+\end{itemize}
+
+\noindent Even though no NRI made it among the most important variables in this study (\autoref{fig:var-imp}) (stating that the first seven of this study are the most important ones), it is notable that all ranks from 8 - 30 are occupied by NRI (\autoref{fig:var-imp}).
+However, their relative importance was very small compared to the first seven ranks.
+
+Restricting the important indices of this study to the first seven ranks can be seen critical as we only based the selection on a visual inspection of the variable importance results (\autoref{fig:var-imp}).
+The decision to make a cut between rank seven and eight was based on a combination of two facts:
+
+\begin{itemize}
+	\item Using only vegetation indices is easier for large scale predictions using satellites like Sentinel-2 (most NRIs cannot be used with it because the spectral bands do not exist).
+	\item The drop in the importance score of the variable importance results (\autoref{fig:var-imp}).
+\end{itemize}
+
+\noindent However, based on these two points, we could also have made the cut between rank five and 6 but including the two vegetation indices at rank six and seven will eventually improve the model and does not increase runtime.
 
@@
 
-\noindent The spatial predictions showed that the model tries to avoid making extreme predictions since most values ranged between 30 \% and 80 \%.
-This behavior is mainly triggered by the concentration of the model on the distribution of defoliation in the \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} training plots.
-This focus then causes the high prediction error for \textit{Laukiz 2} for which a lot of defoliation values actually range between 0 \% and 20 \%.
-The fitted model would need more training samples for plots that show a similar defoliation as \textit{Laukiz 2} to reach a better performance for similar plots.
+\noindent The spatial predictions showed that the model tries to avoid making extreme predictions since most values ranged between a defoliation of 30 \% and 80 \%.
+This behavior is mainly triggered by overfitting on the observations of the \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} training plots.
+The overfit then causes a high prediction error for \textit{Laukiz 2} for which a lot of defoliation values actually range between 0 \% and 20 \%.
+The fitted model would need more training samples of plots with an unusual defoliation distribution to become more robust and achieve a better overall performance.
+
+\subsection{Comparison to other studies}
+
+\noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengarajanModelingForestDefoliation2016}.
+This corresponds with the finding of this study that the most important variables are located in the NIR region.
+
+\cite{goodbodyDigitalAerialPhotogrammetry2018} used NDVI and structural measures in as inputs for a partial least squares analysis to model defoliation caused by the spruce budworm.
+Results showed that metrics from spectral features were most important.
+Incorporating spectral metrics could be a possible enhancement for future studies.
+
+\cite{townsendGeneralLandsatModel2012} used Landsat data to model defoliation caused by insect herbivores.
+They found that the \ac{NDII} ($\frac{Band 4 - Band 5}{Band 4 + Band 5}$) and the moisture stress index ($\frac{Band 5}{Band 4}$) gave better results than using NDVI.
+Overall, they used 10 vegetation indices derived from Landsat data.
+
+MODIS data was used by \cite{debeursEstimatingEffectGypsy2008} to model defoliation caused by the gypsy moth using vegetation indices such as NDVI, EVI, NDWI and NDII.
+
+All of these examples validate the approach of using vegetation indices to model defoliation.
+Even though the spatial resolution of the data in these studies varied between hundreds of meters (MODIS) \cite{debeursEstimatingEffectGypsy2008} and centimeters \cite{goodbodyDigitalAerialPhotogrammetry2018}, high resolution data is preferred to fit accurate models.
+Also, the importance of certain indices (e.g. NDVI, EVI) will vary based on the data and resolution.
+The finding of this work that the vegetation indices GDVI and EVI are most important for the fitted model could not be verified by other studies.
+However, the presented ones did often only use a small subset of the vegetation indices that were used in this study.
+
+We could not find a study that used recent machine-learning techniques in combination with a high amount of variables to model defoliation.
+This fact highlights the importance of this work and will hopefully encourage scientists to use machine-learning techniques, feature selection methods and a wider selection of vegetation indices when assessing defoliation in the future.
 
@@
 
-\noindent In this work we used various indices derived from hyperspectral remote sensing data to estimate defoliation in northern Spain.
+\noindent In this work we used various indices derived from hyperspectral remote sensing data to estimate defoliationas a proxy for tree health in northern Spain.
 In the algorithm comparison \textit{xgboost} showed the best performance among the tested ones.
@@
 
-The fitted models on the plot level showed a promising performance with RMSE values between 8 and 20.
-The performance of the model containing all data was acceptable (RMSE 29.59) but can be improved by adding more observations from other plots in future studies.
+The fitted models on the plot level showed a promising performance with RMSE values between 8 and 20 which validated the approach of relating defoliation to remote sensing indices.
+The performance of the model containing all data (four plots) was acceptable (RMSE 29.59) but can be improved by adding more observations from other plots in future studies.
 
@@
 The most important indices for the fitted model were widely known vegetations indices such as EVI, GDVI or NDVI.
-It would be interesting to link the indices of this work to other indicators of forest/vegetation health in future studies and analyze their importance and the model performances.
-In summary, defoliation can be successfully used as a response variable to model forest health using remote sensing indices.
+In future studies it would be interesting to link the indices of this work to other indicators of forest/vegetation health and analyze their importance and the model performances.
 

### 2018-08-09 01_writing/Schratz_hyperspectral.tex
- improve intro

@@
 	\acro{NIR}{Near-Infrared}
-	\acro{NRI}{Normalized Ratio Indices}
+	\acro{NRI}{Normalized Ratio Index}
 	\acro{OLS}{Ordinary Least Squares}
@@
 	\acro{SMBO}{Sequential-based Model Optimization}
-	\acro{SVM}{Support Vector Machines}
+	\acro{SVM}{Support Vector Machine}
 	\acro{TPR}{True Positive Rate}
@@
 % Explain how remote sensing is used in forestry (potential to map forest health)
-Data retrieved from remote sensing satellites is successfully used in forestry to monitor temporal changes across large areas \citep{martinezdelcastilloEvaluationForestCover2015,sextonModelPropagationUncertainty2015}.
+\noindent Data retrieved from remote sensing satellites is successfully used in forestry to monitor temporal changes across large areas \citep{martinezdelcastilloEvaluationForestCover2015,sextonModelPropagationUncertainty2015}.
 The use of \ac{SAR} techniques enables scientists to estimate \ac{AGB} \citep{luSurveyRemoteSensingbased2016, sinhaReviewRadarRemote2015}.
-Forest health is commonly assesed using optical data from multi-/hyperspectral satellites by applying temporal change detections \citep{zhangRemoteSensingSeasonal2016}.
+Forest health is commonly assesed using optical data from multi-/hyperspectral satellites by applying either temporal change detections \citep{zhangRemoteSensingSeasonal2016} or by using vegetation indices to monitor the current state \citep{townsendGeneralLandsatModel2012}.
 With the recent success story of machine-learning methods in the field of remote sensing, modeling techniques such as \ac{RF} are frequently used to model relationships of possible triggers to forest health \citep{belgiuRandomForestRemote2016, laryMachineLearningGeosciences2016, michezClassificationRiparianForest2016}.
 
-With a robust model, predictions to large areas can be conducted, providing valuable information about the health condition of forest stands.
-One approach to model forest health is to extract information from spectral signatures of affected and unaffected trees \citep{lelongEvaluationOilPalmFungal2010}.
-Spectral (vegetation-)indices have shown the potential to provide valuable information to increase predictive accuracy of forest pathogens \citep{jiangSatellitederivedVegetationIndices2014, adamczykRededgeVegetationIndices2015}.
+With a robust model, predictions to large areas can be made, providing valuable information about the health condition of forest stands.
+One approach to model forest health is to extract information from spectral information of affected and unaffected trees \citep{lelongEvaluationOilPalmFungal2010}.
+Vegetation indices have shown the potential to provide valuable information to increase the predictive accuracy of forest pathogens \citep{jiangSatellitederivedVegetationIndices2014, adamczykRededgeVegetationIndices2015}.
 
@@
 Sentinel-2).
-Also, there is currently no freely available data from hyperspectral sensors that could be used for such studies (after the decommission of the EO-1 Hyperion satellite).
-If the spatial resolution of the data is too coarse (e.g. $> 5 m$), the value of a pixel usually contains information from multiple trees and possibly even bare-ground information.
-This point makes it impossible to use data from sensors like Sentinel-2 to train a model on a tree level.
+Also, there is currently no freely available data from hyperspectral sensors that could be used for such studies (after the decommission of the EO-1 Hyperion satellite in January 2017).
+To assess forest health on a tree scale, the spatial resolution of the data should be below 5 m as otherwise the value of a pixel may contain information from multiple trees and possibly even bare-ground information.
+With this limitation, data from sateillites with a coarser resolution (e.g. Sentinel-2) can be used for prediction tasks but not for training purposes of algorithms on a tree level.
 
@@
 In this study we will use hyperspectral data with a spatial resolution of one meter and 126 spectral bands to model the health status of Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain (\autoref{fig:study_area}).
-The trees in the study area suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum} \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanzaNativeRhizobacteriaBiocontrol2016, iturritxaBiocontrolFusariumCircinatum2017}.
-In-situ measurements of defoliation on a tree level have been collected to serve as the response variable (as a proxy for tree health).
+The trees in the study area suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanzaNativeRhizobacteriaBiocontrol2016, iturritxaBiocontrolFusariumCircinatum2017}.
+In-situ measurements of defoliation on a tree level (as a proxy for tree health) have been collected to serve as the response variable.
 The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
 The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{ganleyGlobalClimaticRisk2009}.
-Hyperspectral remote sensing data in combination with state-of-the-art machine-learning techniques is used to help monitoring the health status in this region by early detecting affected trees/plots.
-The aim is to spatially predict the fitted model to other plots/the whole Basque Country to create a forest health map.
 
-% Show importance of vegetation indices and NRI
+We used state-of-the-art machine-learning techniques in combination with high resolution remote sensing information to model forest health on a tree level.
+Furthermore, we aimed to reduce the amount of predictors using feature selection to simplify the spatial prediction of the fitted model.
 
-To extract the most information from the available remote sensing data, we not only calculated the most common vegetation indices like \textit{NDVI} to link against defoliation but all possible ones within the spectral region of the data (400 nm - 1000 nm) that are implemented in the \textit{hsdar} package in R \citep{hsdar}.
-Additionally, all possible combinations of \ac{NRI} were calculated from the data and supplied to a selection of different algorithms (\textit{\ac{SVM}}, \textit{\ac{RR}}, \textit{xgboost}) as predictors.
-
-% Mention the ability of ML algs to handle highly correlated variables
-
-Specifically the following objectives are addressed:
+\noindent Specifically the following objectives were addressed:
 
 \begin{itemize}
-	\item Comparison of multiple algorithms on their performance to model defoliation of \textit{Pinus radiata} trees using highly-correlated indices
-	\item Exploration of the most important indices of the best performing model
-	\item Spatial prediction of defoliation to \textit{Pinus radiata} plots and the whole Basque Country
+	\item Performance comparison of multiple algorithms modeling defoliation of \textit{Pinus radiata} trees using highly-correlated features
+	\item Exploration of the most important variables
+	\item Spatial prediction of \textit{Pinus radiata} defoliation
 \end{itemize}
@@
 \subsection{In-situ data}
-% Describe Plots
 
-The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
+\noindent The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
 \textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} has the largest area size (1.44 ha).
@@
 
-\begin{figure} [b!]
+\begin{figure} [t!]
 	\begin{center}
@@
 
-The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around 12 am.
+\noindent The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around 12 am.
 The images were taken using a AISAEAGLE-II sensor.
@@
 % parameter limits
-\begin{table}[t!]
+\begin{table}[b!]
 \centering
@@
 It provides a standardized interface for a wide variety of statistical and machine-learning models in R simplifying essential modeling tasks such as hyperparameter tuning, model performance evaluation and parallelization \citep{bischlMlrMachineLearning2016}.
-We provide the complete code and required data as a Mendeley dataset to make this work fully reproducible \citep{}.
+We provide the complete code and required data as a Mendeley dataset to make this work fully reproducible [STILL TO COME].
 
@@
 
-\noindent Three algorithms (\textit{xgboost}, \textit{\ac{SVM}}, \textit{\ac{RR}}) were benchmarked on their predictive performance.
+\noindent Three algorithms (\textit{xgboost}, \textit{\ac{SVM}} and \textit{\ac{RR}}) were benchmarked on their predictive performance.
 Besides the well-known \ac{SVM} algorithm \citep{Vapnik1998}, we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
-We also added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerlRidgeRegressionBiased1970}.
+We added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerlRidgeRegressionBiased1970}.
 One of the most popular machine-learning algorithm, Random Forest \citep{Breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
@@
 \begin{enumerate}
-	\item Using spatial block \ac{CV} on the plot level with each plot serving as the test set once.
-	      Subsequently, four performance estimates were retrieved, one for each fold.
+	\item Using spatial block \ac{CV} on the plot level with each plot serving as the test set once. Four performance estimates were retrieved, one for each fold.
 	\item Using five-fold five-time repeated spatial \ac{CV} within each plot based on the k-means clustering approach of \cite{sperrorest} (\autoref{fig:spcv}).
@@
 \subsubsection{Hyperparameter tuning}
+\label{meth:smbo}
 
@@
 This strategy continues until a termination criterion, defined by the user, is reached \citep{Hutter2011, Jones1998}.
+
 In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting in a total budget of 50 evaluated hyperparameter settings per fold.
@@
 \noindent To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
-The score is calculated by taking the contribution of each feature for each tree in the fitted model.
+This score is calculated by taking the contribution of each feature for each tree in the fitted model.
 The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.
@@
 
-\textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots have a higher density (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
+\noindent \textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots have a higher density (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
 
@@
 This is not the case for the other two higher defoliated plots for which bare ground and defoliated predictions are mainly in the same value range (around 60 \% - 80 \%).
-
-	[ADD FIGURE WITH SENTINEL2 PREDICTION]
 
@@
 \noindent The relatively large difference in performance between \textit{RR} (59.10 RMSE) and the machine-learning models (36.23 and 33.26 RMSE) is remarkable.
-\textit{RR} has shown promising performance results in other studies when many highly-correlated predictors were involved (REFERECES).
-However, in this study, \textit{RR} was not able to achieve a sufficient performance score compared to \textit{SVM} and \textit{xgboost} even though its hyperparameter $\lambda$ was properly tuned using SMBO.
+\textit{RR} has shown promising results in other studies when many highly-correlated predictors were involved \citep{hernandezUsingRidgeRegression2015, imaniRidgeRegressionbasedFeature2015}.
+However, in this study, \textit{RR} was not able to achieve a sufficient performance score compared to \textit{SVM} and \textit{xgboost} even though its hyperparameter $\lambda$ was properly tuned using SMBO (\autoref{meth:smbo}).
 
-While \textit{xgboost} shows a slightly better performance than \textit{SVM}, the latter has the advantage of only having two hyperparameters that need to be tuned.
+While \textit{xgboost} showed a slightly better performance than \textit{SVM}, the latter has the advantage of only having two hyperparameters that need to be tuned.
 This results in a shorter runtime.

### 2018-11-28 02_scripts/version-1/021_preprocessing.Rmd
save state

---
title: "02-preprocessing"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r}
library(purrr)
library(here)
library(tibble)
library(raster)
library(glue)
library(sf)
library(dplyr)
library(magrittr)
```

<!-- # Read data -->

<!-- Reading in data from .txt files and exporting them as a Geopackage (.gpkg). -->

<!-- ## Survey data -->

<!-- ```{r} -->
<!-- names <- c("laukiz1", "laukiz2", "luiando", "oiartzun") -->

<!-- map( -->
<!--   names, -->
<!--   ~ as_tibble(readr::read_csv( -->
<!--     file = glue("/data/patrick/mod/survey_data/2016/{.x}.csv" -->
<!--                 ), -->
<!--   )) -->
<!-- ) %>% -->
<!--   set_names(names) %>% -->
<!--   map(~ rename_all(.x, tolower)) -> data_tbl -->
<!-- ``` -->

<!-- ## Clean Laukiz 1 & Oiartzun coords -->

<!-- ```{r} -->
<!-- data_tbl$laukiz1 %<>% -->
<!--   select(-x, -y) %>%  -->
<!--   rename(x = lon, y = punto) -->

<!-- data_tbl$oiartzun %<>% -->
<!--   select(-x, -y) %>%  -->
<!--   rename(x = lon, y = punto) -->
<!-- ``` -->

<!-- Laukiz2 has no coordinate info in the txt file. We need to read the shapefile and join them. -->

<!-- ### Join Laukiz2 coords -->

<!-- ```{r} -->
<!-- laukiz2_coords <- st_read("/data/patrick/raw/survey_data/2016/laukiz2/arboles/Laukiz2_arboles_ETRS89.shp") -->

<!-- data_tbl$laukiz2 <- left_join(data_tbl$laukiz2, laukiz2_coords, -->
<!--   by = c("num arbol" = "id_arbol") -->
<!-- ) -->
<!-- ``` -->

<!-- Drop observations with no coordinate information. -->

<!-- ```{r} -->
<!-- data_tbl %<>% -->
<!--   map(~ drop_na(.x, c("x", "y"))) -->
<!-- ``` -->

<!-- Coerce to `sf` objects. -->

<!-- ```{r} -->
<!-- map(data_tbl, ~ st_as_sf(.x, coords = c("x", "y"))) -> data_sf -->
<!-- ``` -->

<!-- Select variables related to defoliation. -->

<!-- ```{r} -->
<!-- data_sf %<>% -->
<!--   map(~ dplyr::select(.x, contains("def"))) -->
<!-- ``` -->

<!-- Laukiz2 needs some clean up: -->

<!-- "no" -> 0 -->
<!-- "*" -> NA -->
<!-- Coerce from chr to int. -->
<!-- Calculate "deftot" which is the mean of all three parts (bottom, mid, top). -->

<!-- `rowSums` will thrown an error due to the "geometry" column so we need to add `drop = TRUE`. -->

<!-- ```{r} -->
<!-- data_sf$laukiz2 %<>% -->
<!--   mutate( -->
<!--     `% de defoliacion parte media` = -->
<!--       replace( -->
<!--         .$`% de defoliacion parte media`, -->
<!--         .$`% de defoliacion parte media` == "no", "0" -->
<!--       ) -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     `% defoliacion en punta` = -->
<!--       replace( -->
<!--         .$`% defoliacion en punta`, -->
<!--         .$`% defoliacion en punta` == "no", "0" -->
<!--       ) -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     `% defoliacion base` = -->
<!--       replace( -->
<!--         .$`% defoliacion base`, -->
<!--         .$`% defoliacion base` == "no", "0" -->
<!--       ) -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     `% defoliacion base` = -->
<!--       replace( -->
<!--         `% defoliacion base`, -->
<!--         `% defoliacion base` == "*", NA -->
<!--       ) -->
<!--   ) %>% -->
<!--   mutate_if(is.character, as.numeric) %>% -->
<!--   mutate(deftot = as.integer(rowMeans(.[, 1:3, drop = TRUE], na.rm = TRUE))) %>% -->
<!--   drop_na(deftot) %>% -->
<!--   dplyr::select(deftot, geometry) -->
<!-- ``` -->

<!-- # Write clean data as .gpkg -->

<!-- The error messages can be ignored, writing works: https://github.com/r-spatial/sf/issues/628#issuecomment-383363473 -->

<!-- ```{r} -->
<!-- walk2(data_sf, seq_along(data_sf), ~ st_write(.x, glue("/data/patrick/mod/tree-per-tree/2016/{name}/{name}.gpkg", -->
<!--   name = names(data_sf[.y])), delete_dsn = TRUE)) -->
<!-- ``` -->

# New cleaned data From Eugenia June 2018

```{r}
data_tbl = map(c("laukiz1", "laukiz2", "luiando", "oiartzun", "hernani"), 
               ~ readr::read_csv2(glue("/data/patrick/raw/survey_data/2016/{.x}/{.x}.csv")) %>% 
                 #na.omit() %>% 
                 rename_all(tolower) %>% 
                 rename(height = heigth) %>% 
               filter_all(any_vars(!is.na(.)))) %>% 
  set_names(c("laukiz1", "laukiz2", "luiando", "oiartzun", "hernani"))

data_tbl$laukiz2 %<>% 
  rename(canker = cankers) 
```

## Join coordinates

### Hernani

```{r}
hernani_coords1 <- st_read("/data/patrick/raw/survey_data/2016/hernani/hernani/arboles/arboles_hernani.shp") %>% 
  select(id) 
hernani_coords2 = as.data.frame(st_coordinates(hernani_coords1)) %>% 
  bind_cols(hernani_coords1) %>% 
  st_as_sf() %>% 
  st_set_geometry(NULL) %>% 
  rename_all(tolower)

data_tbl$hernani <- left_join(data_tbl$hernani, hernani_coords2,
  by = c("tree number" = "id")
)
```

### Laukiz 2

```{r}
laukiz2_coords <- st_read("/data/patrick/raw/survey_data/2016/laukiz2/arboles/Laukiz2_arboles_ETRS89.shp") %>% 
  st_set_geometry(NULL)

data_tbl$laukiz2 <- left_join(data_tbl$laukiz2, laukiz2_coords,
  by = c("tree number" = "id_arbol")
)
```

### Laukiz 1

```{r}
laukiz1_coords <- st_read("/data/patrick/raw/survey_data/2016/laukiz1/escaner/arboles_laukiz1_CT.shp") %>% 
  select(num)

laukiz1_coords2 = as.data.frame(st_coordinates(laukiz1_coords)) %>% 
  bind_cols(laukiz1_coords) %>% 
  st_as_sf() %>% 
  st_set_geometry(NULL) %>% 
  rename_all(tolower)

data_tbl$laukiz1 <- left_join(data_tbl$laukiz1, laukiz1_coords2,
  by = c("tree number" = "num")
)
```

### Luiando

```{r}
luiando_coords <- st_read("/data/patrick/raw/survey_data/2016/luiando/puntos_escaner/luiando_diam_CT_chapa.shp") %>% 
  select(id)

luiando_coords2 = as.data.frame(st_coordinates(luiando_coords)) %>% 
  bind_cols(luiando_coords) %>% 
  st_as_sf() %>% 
  st_set_geometry(NULL) %>% 
  rename_all(tolower)

data_tbl$luiando <- left_join(data_tbl$luiando, luiando_coords2,
  by = c("tree number" = "id")
)
```

### Oiartzun

```{r}
oiartzun_coords <- st_read("/data/patrick/raw/survey_data/2016/oiartzun/ARBOLES_antiguo/oiartzun_etrs89.shp") %>% 
  select(x, y, num) %>% 
  mutate(num = as.integer(as.character(num))) %>% 
  st_set_geometry(NULL)

data_tbl$oiartzun <- left_join(data_tbl$oiartzun, oiartzun_coords,
  by = c("tree number" = "num")
)
```

## Clean up

Ensure that no value > 100 exist.

```{r}
data_tbl %<>% 
  map(~ .x %>% 
        mutate(decoloration = if_else(decoloration > 100, 100, 
                                      as.numeric(decoloration))) %>% 
        mutate(defoliation = if_else(defoliation > 100, 100, 
                                      as.numeric(defoliation))))
```

NA means that there is no defoliation -> 0

```{r}
data_tbl %<>% 
  map(~ .x %>% 
        mutate(decoloration = if_else(is.na(decoloration), 0, 
                                      as.numeric(decoloration))) %>% 
        mutate(defoliation = if_else(is.na(defoliation), 0, 
                                     as.numeric(defoliation))) %>% 
        mutate(canker = if_else(is.na(canker), 0, 
                                as.numeric(canker))))
```

## st_as_sf() and save

```{r}
data_sf = map(data_tbl, ~ tidyr::drop_na(.x, c("x", "y")) %>% 
                st_as_sf(coords = c("x", "y"), crs = 32630))
# hernani is not needed
data_sf$hernani = NULL
```

```{r}
walk2(data_sf, seq_along(data_sf), 
      ~ st_write(.x, glue("/data/patrick/mod/tree-per-tree/2016/{name}/{name}.gpkg",
                          name = names(data_sf[.y])), delete_dsn = TRUE))
```

### 2018-11-28 02_scripts/version-1/035-all-plots-modeling.Rmd
save state

--- 2018-08-05 02_scripts/035-all-plots-modeling.Rmd
@@
 ```{r}
-m = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
-# 5h 30 min
+# 5h 30 min with 2 nmc
 var_imp = generateFeatureImportanceData(task, learner = lrn_xgboost, 
                                         measure = rmse, nmc = 50)
-saveRDS(var_imp, "/data/patrick/mod/hyperspectral/variable-importance/bf2_ncm50rda")
+var_imp = readRDS("/data/patrick/mod/hyperspectral/variable-importance/bf2_ncm50rda")
 
@@
 ```{r}
+m = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
 imp.xgb = getFeatureImportance(m) 
@@
 ```{r}
+model_xgboost_defol = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
+library("doParallel")
+cl = makePSOCKcluster(40)
+# Registers cluster
+registerDoParallel(cl)
+
 X = task$env$data[which(names(task$env$data) != "defoliation")]
-predictor = Predictor$new(m, data = X, y = task$env$data$defoliation)
-imp = FeatureImp$new(predictor, loss = "rmse")
+
+# correct column order of features: https://github.com/dmlc/xgboost/issues/1809
+X = X[model_xgboost_defol$learner.model$feature_names]
+
+predictor = Predictor$new(model_xgboost_defol, data = X, y = task$env$data$defoliation)
+imp = FeatureImp$new(predictor, loss = "rmse", parallel = TRUE, n.repetitions = 100)
+stopCluster(cl)
+saveRDS(imp, "/data/patrick/mod/hyperspectral/variable-importance/feature-imp-iml.rda")
+
+imp = readRDS("/data/patrick/mod/hyperspectral/variable-importance/feature-imp-iml.rda")
 ```
@@
 ```
+
+## xgboost feat sel and tune
+
+```{r}
+lrn_xgboost <- makeLearner("regr.xgboost",
+                   par.vals = list(
+                     objective = "reg:linear",
+                     eval_metric = "error"
+                   ))
+```
+
+```{r}
+lrn_xgboost = makeFilterWrapper(learner = lrn_xgboost,
+                                imp.learner = c("regr.xgboost"))
+```
+
+```{r}
+ps_xgboost_mrmr <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20),
+  makeDiscreteParam("fw.method", "mrmr"),
+  makeNumericParam("fw.perc", lower = 0, upper = 1)
+)
+
+ps_xgboost_permut <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20),
+  makeDiscreteParam("fw.method", "permutation.importance"),
+  makeNumericParam("fw.perc", lower = 0, upper = 1)
+)
+```
+
+```{r}
+ctrl = makeMBOControl(propose.points = 1L)
+ctrl = setMBOControlTermination(ctrl, iters = 20L)
+ctrl = setMBOControlInfill(ctrl, crit = crit.ei)
+tune.ctrl = makeTuneControlMBO(mbo.control = ctrl,
+                               mbo.design = generateDesign(n = 30, 
+                                                           par.set = ps_xgboost))
+```
+
+```{r}
+inner <- makeResampleDesc("Blocking")
+outer <- makeResampleDesc("Blocking")
+```
+
+### tune wrapper
+
+```{r}
+tune_wrapper_mrmr <- makeTuneWrapper(lrn_xgboost, resampling = inner, 
+                                     par.set = ps_xgboost_mrmr,
+                                     control = tune.ctrl, show.info = TRUE,
+                                     measures = list(rmse))
+tune_wrapper_permut <- makeTuneWrapper(lrn_xgboost, resampling = inner, 
+                                       par.set = ps_xgboost_permut,
+                                       control = tune.ctrl, show.info = TRUE,
+                                       measures = list(rmse))
+```
+
+### Resample
+
+```{r}
+library(parallelMap)
+parallelStart(mode = "multicore", cpus = 30, level = "mlrMBO.feval",
+              mc.set.seed = TRUE)
+# parallelStart(mode = "multicore", cpus = 4, level = "mlr.resample",
+#               mc.set.seed = TRUE)
+set.seed(12345)
+resa_svm <- mlr::resample(tune_wrapper, task, models = FALSE,
+                          resampling = outer, 
+                          extract = function(x) list(getFeatSelResult, getTuneResult),
+                          show.info = TRUE, measures = list(rmse))
+
+benchm <- benchmark(tune_wrapper, task, models = FALSE,
+                          resampling = outer, 
+                          extract = function(x) list(getFeatSelResult, getTuneResult),
+                          show.info = TRUE, measures = list(rmse))
+
+parallelStop()
+saveRDS(resa_svm, "/data/patrick/mod/hyperspectral/spcv/all_plots_xgboost_fs_tune.rda")
+```

### 2018-11-28 02_scripts/version-1/07_benchmark_all_plots.Rmd
save state

--- 2018-07-11 02_scripts/07_benchmark_all_plots.Rmd
@@
 ps_xgboost <- makeParamSet(
-  makeDiscreteParam("nrounds", values = c(10, 100, 1000, 10000)),
-  makeNumericParam("colsample_bytree", lower = 0.1, upper = 1),
-  makeNumericParam("subsample", lower = 0.1, upper = 1),
-  makeIntegerParam("max_depth", lower = 1, upper = 785),
-  makeNumericParam("gamma", lower = 0, upper = 20)
+  makeNumericParam("nrounds", lower = 0, upper = 8.64, 
+                   trafo = function (x) round(2^x * 10)),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20)
 )
@@
                                                       objective = "reg:linear",
-                                                      eval_metric = "error",
-                                                      eta = 0.1
+                                                      eval_metric = "error"
                                                     )
@@
 )
-tune_wrapper_rf <- makeTuneWrapper(makeLearner("classif.ranger"),
+tune_wrapper_rf <- makeTuneWrapper(makeLearner("regr.ranger"),
                                    resampling = inner, par.set = ps_rf,

### 2018-11-28 02_scripts/version-1/08_prediction.Rmd
save state

--- 2018-08-07 02_scripts/08_prediction.Rmd
@@
 ```{r}
-# lattice.options(
-#   layout.heights=list(bottom.padding=list(x=0), top.padding=list(x=0)),
-#   layout.widths=list(left.padding=list(x=0), right.padding=list(x=0))
-# )
-
 raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
@@
 dev.off()
-
 ```
@@
 
+# Plot all plots
+
+Descriptive title
+
+```{r}
+raster_list = map(list.files("/data/patrick/mod/hyperspectral/prediction/", 
+                             pattern = "defol-pred*", full.names = TRUE), 
+                  ~ raster(.x)
+                  )
+raster_list_wgs84 = map(raster_list, 
+                        ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")) %>% 
+  map(~ set_names(.x, "defoliation"))
+
+name = c("Aguinaga", "Altube", "Azaceta", "Barazar", "busturi", "Caranca",
+         "Douglas 1", "Douglas 2", "Douglas Albiztur", "Douglas Legutio",
+         "Gipuzkoa", "Laricio Defol Olaeta", "Laricio Defol Urkiola",
+         "Laricio Gatzaga", "Laukiz 1", "Laukiz 2", "Laukiz 3", "Luiando", 
+         "Oiartzun", "Oxtandio", "Picea Abornicano", "Pinas Santa Koloma",
+         "Pinast Arza", "Rad Altube", "Rad Menagaray", "Sequoia Legorreta", 
+         "Silvest Nograro")
+
+maps = imap(raster_list_wgs84, ~ levelplot( 
+  .x[["defoliation"]], 
+  main = glue::glue("{tools::toTitleCase(name[.y])}: Defoliation [%]"), 
+  #main = glue::glue("{tools::toTitleCase(.y)}"), 
+  margin = FALSE,                             
+  between = list(x = 0.0, y = 0.0), 
+  at = seq(30, 82, length.out = 100), # min/max from map(raster_list_wgs84, ~ summary(.x))
+  colorkey = list(
+    width = 0.75,                       
+    space = 'right',  
+    labels = list(#at = c(0, 25, 50, 75, 100),
+                font = 4),      
+    axis.line = list(col = 'black')
+  ),
+  par.strip.text = list(col = "white", cex = 1.2),    
+  col.regions = viridis)    
+)
+```
+
+```{r}
+png("../03_figures/results/predictions_plots-1-4.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[1:4], nrow = 2, ncol = 2)
+dev.off()
+
+png("../03_figures/results/predictions_plots-5-8.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[5:8], nrow = 2, ncol = 2)
+dev.off()
+
+png("../03_figures/results/predictions_plots-9-12.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[9:12], nrow = 2, ncol = 2)
+dev.off()
+
+png("../03_figures/results/predictions_plots-12-16-.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[13:16], nrow = 2, ncol = 2)
+dev.off()
+
+png("../03_figures/results/predictions_plots-17-20.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[17:20], nrow = 2, ncol = 2)
+dev.off()
+
+png("../03_figures/results/predictions_plots-21-24.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[21:24], nrow = 2, ncol = 2)
+dev.off()
+
+png("../03_figures/results/predictions_plots-25-28.png", units = "in", 
+    width = 13, height = 9.21, res = 300)
+gridExtra::grid.arrange(grobs = maps[25:28], nrow = 2, ncol = 2)
+dev.off()
+```
+
 # PNG to PDF conversion

### 2018-11-28 02_scripts/feature-set-HS-bands/HS-filter-borda.R

@@ -1,0 +1,114 @@
+remotes::install_github("pat-s/mlr@fs-pat-all")
+library(mlr)
+library(mlrMBO)
+library(magrittr)
+
+# Task ----------------
+
+# Read data
+
+data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
+coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
+
+# Merge into one dataset
+
+data = tibble::as_tibble(data.table::rbindlist(data, fill = TRUE))
+data = Filter(function(x) !any(is.na(x)), data)
+coords = tibble::as_tibble(data.table::rbindlist(coords))
+
+task = makeRegrTask(id = "defoliation-all-plots", data = data,
+                    target = "defoliation", coordinates = coords,
+                    blocking = factor(rep(1:4, c(479, 451, 291, 529))))
+
+# Learners ---------------------
+
+lrn_xgboost <- makeLearner("regr.xgboost",
+                           par.vals = list(
+                             objective = "reg:linear",
+                             eval_metric = "error"
+                           )) %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+lrn_ksvm <- makeLearner("regr.ksvm",
+                        kernel = "rbfdot") %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+# Resampling ----------------
+
+inner <- makeResampleDesc("CV", fixed = TRUE)
+outer <- makeResampleDesc("CV", fixed = TRUE)
+
+# Param sets ----------------
+
+ps_xgboost <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20),
+  makeDiscreteVectorParam("fw.basal.methods", len = 8,
+                          values = c("carscore", "cforest.importance", "FSelector_chi.squared", "FSelectorRcpp_gain.ratio",
[…]

### 2018-11-28 02_scripts/feature-set-VI/VI-filter-borda.R

@@ -1,0 +1,236 @@
+remotes::install_github(c("pat-s/mlr@fs-pat-all", "pat-s/ParamHelpers@distinct"))
+library(mlr)
+library(mlrMBO)
+library(magrittr)
+
+# Task ----------------
+
+task = readRDS("/data/patrick/mod/paper3/task/task-VI-[paper3].rda")
+
+# Learners ---------------------
+
+lrn_xgboost <- makeLearner("regr.xgboost",
+                           par.vals = list(
+                             objective = "reg:linear",
+                             eval_metric = "error"
+                           )) %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+lrn_ksvm <- makeLearner("regr.ksvm",
+                        kernel = "rbfdot") %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+lrn_rf <- makeLearner(
+  "regr.ranger") %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+lrn_ridge <- makeLearner("regr.glmnet", id = "ridge",
+                         alpha = 0,
+                         standardize = FALSE,
+                         intercept = FALSE) %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+ridge_train = train(makeLearner("regr.glmnet",
+                                alpha = 0, standardize = FALSE),
+                    task)
+lambda_ridge_min = min(ridge_train$learner.model$lambda)
+lambda_ridge_max = max(ridge_train$learner.model$lambda)
+
+lrn_lasso <- makeLearner("regr.glmnet", id = "lasso",
+                         alpha = 1,
+                         standardize = FALSE,
+                         intercept = FALSE) %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
+lasso_train = train(makeLearner("regr.glmnet", id = "elnet",
+                                alpha = 1, standardize = FALSE),
+                    task)
+lambda_lasso_min = min(lasso_train$learner.model$lambda)
+lambda_lasso_max = max(lasso_train$learner.model$lambda)
+
+lrn_elnet <- makeLearner("regr.glmnet",
+                         standardize = FALSE,
+                         intercept = FALSE) %>%
+  makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
+
[…]

### 2018-11-28 02_scripts/version-1/035-all-plots-modeling/xgboost-chi-squared.R

@@ -1,0 +1,106 @@
+devtools::install_github("mlr-org/mlr@dd983c382bbd50d820ce2dceb809cb46e4bc5d7b")
+
+library(mlr)
+library(mlrMBO)
+library(tibble)
+library(purrr)
+library(magrittr)
+library(data.table)
+
+# Read data
+
+data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
+coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
+
+# Merge into one dataset
+
+data = as_tibble(rbindlist(data, fill = TRUE))
+data = Filter(function(x) !any(is.na(x)), data)
+coords = as_tibble(rbindlist(coords))
+
+# Task ----------------
+
+task = makeRegrTask(id = "defoliation-all-plots", data = data,
+                    target = "defoliation", coordinates = coords,
+                    blocking = factor(rep(1:4, c(479, 451, 291, 529))))
+
+# Resampling ----------------
+
+inner <- makeResampleDesc("CV", fixed = TRUE)
+outer <- makeResampleDesc("CV", fixed = TRUE)
+
+# Learners ----------------
+
+lrn_xgboost <- makeLearner("regr.xgboost",
+                           par.vals = list(
+                             objective = "reg:linear",
+                             eval_metric = "error"
+                           )) %>%
+  makeFilterWrapper()
+
+
+lrn_ksvm <- makeLearner("regr.ksvm",
+                        kernel = "rbfdot") %>%
+  makeFilterWrapper()
+
+# Param sets ----------------
+
+ps_xgboost_chi.squared <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20),
+  makeDiscreteParam("fw.method", "chi.squared"),
+  makeNumericParam("fw.perc", lower = 0, upper = 1)
+)
+
+
[…]

### 2018-11-28 02_scripts/version-1/035-all-plots-modeling/xgboost-mrmr.R

@@ -1,0 +1,106 @@
+devtools::install_github("mlr-org/mlr@featsel_praznik")
+
+library(mlr)
+library(mlrMBO)
+library(tibble)
+library(purrr)
+library(magrittr)
+library(data.table)
+
+# Read data
+
+data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
+coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
+
+# Merge into one dataset
+
+data = as_tibble(rbindlist(data, fill = TRUE))
+data = Filter(function(x) !any(is.na(x)), data)
+coords = as_tibble(rbindlist(coords))
+
+# Task ----------------
+
+task = makeRegrTask(id = "defoliation-all-plots", data = data,
+                    target = "defoliation", coordinates = coords,
+                    blocking = factor(rep(1:4, c(479, 451, 291, 529))))
+
+# Resampling ----------------
+
+inner <- makeResampleDesc("CV", fixed = TRUE)
+outer <- makeResampleDesc("CV", fixed = TRUE)
+
+# Learners ----------------
+
+lrn_xgboost <- makeLearner("regr.xgboost",
+                           par.vals = list(
+                             objective = "reg:linear",
+                             eval_metric = "error"
+                           )) %>%
+  makeFilterWrapper()
+
+
+lrn_ksvm <- makeLearner("regr.ksvm",
+                        kernel = "rbfdot") %>%
+  makeFilterWrapper()
+
+# Param sets ----------------
+
+ps_xgboost_mrmr <- makeParamSet(
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20),
+  makeDiscreteParam("fw.method", "mrmr"),
+  makeNumericParam("fw.perc", lower = 0, upper = 1)
+)
+
+
+ps_svm_mrmr <- makeParamSet(
+  makeNumericParam("C", lower = -15, upper = 15,
+                   trafo = function(x) 2 ^ x),
[…]

### 2019-02-01 README.Rmd
restructure

---
output: github_document
---

[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/pat-s/paper_hyperspectral/commits/master)  
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/)   
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)  
<!-- [![Travis-CI Build Status](https://travis-ci.org/pat-s/paper_hyperspectral.png?branch=master)](https://travis-ci.org/pat-s/paper_hyperspectral)  -->

### Compendium URL

### Authors

**Patrick Schratz** (patrick.schratz@gmail.com) [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0003-0748-6624)    
Jannes Muenchow [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-7834-4717)   
Bernd Bischl [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6002-6980)   
Alexander Brenning [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6640-679X)

### Contents

This repository contains the research compendium of our work on comparing algorithms across multiple feature sets and filtering methods (including ensemble filter methods).

- Using machine-learning algorithms to model defoliation of _Pinus Radiata_ trees.

- keywords

    - high-dimensional data handling
    - feature selection
    - machine-learning
    - ecology

- Compare filtering methods (ensemble filter methods) across various algorithms and datasets

- Predict defoliation to all available plots (24) and the whole Basque Country (at 200 m resolution)

### How to use

#### Read the code, access the data

See the [`analysis`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis) directory on GitHub for the source code that generated the figures and statistical results contained in the manuscript.
See the [`data`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis/data) directory for instructions how to access the raw data discussed in the manuscript.

#### Install the R package

[![Build Status](https://travis-ci.org/pat-s/paper_hyperspectral.svg?branch=master)](https://travis-ci.org/pat-s/paper_hyperspectral)

This repository is organized as an R package, providing functions and raw data to reproduce and extend the analysis reported in the publication. 
Note that this package has been written explicitly for this project and may not be suitable for more general use.

This project is setup with a [drake workflow](https://github.com/ropensci/drake), ensuring reproducibility.
Intermediate targets/objects will be stored in a hidden `.drake` directory.

The R library of this project is managed by [packrat](https://rstudio.github.io/packrat/).
This makes sure that the exact same package versions are used when recreating the project.
When calling `packrat::restore()`, all required packages will be installed with their specific version.

Please note that this project was built with R version 3.5.2 on a Debian 9 operating system.
The packrat packages from this project **are not compatible with R versions prior version 3.5.0.**
For reproducibility, it is recommended to replicate the analysis using the included Dockerfile.
Instructions can be found [ħere](https://github.com/pat-s/paper_hyperspectral#docker).
(In general, it should be possible to reproduce the analysis on any other operating system.)

To clone the project, a working installation of `git` is required.
Open a terminal in the directory of your choice and execute:

```sh
git clone git@github.com:pat-s/paper_hyperspectral.git
```

Then start R in this directory and run

```{r README-1, eval = FALSE}
packrat::restore()
source("analysis/drake.R")
make(plan, keep_going = TRUE, console_log_file = stdout()) 
# use more cores with make(plan, jobs = <number of cores>)
```

## Runtime

Predicted total runtime (based on all target runtimes ever built and stored in the internal `config` file)

```{r README-2 }
source("analysis/drake.R")
predict_runtime(config, from_scratch = TRUE)
```

Acceleration when parallelizing the `make()` call

```{r README-3, warning=FALSE, message=FALSE}
time <- c()
for (jobs in 1:10){
time[jobs] <- predict_runtime(
drake_config(),
jobs = jobs,
from_scratch = TRUE,
known_times = build_times(targets_only = TRUE)$elapsed
)
}

library(ggplot2)
ggplot(data.frame(time = time / 3600, jobs = ordered(1:10), group = 1)) +
geom_line(aes(x = jobs, y = time, group = group)) +
scale_y_continuous(breaks = 0:10 * 4, limits = c(0, 29)) +
ggpubr::theme_pubr() +
xlab("jobs argument of make()") +
ylab("Predicted runtime of make() (hours)")
```

### Licenses

Text: CC-BY-4.0 http://creativecommons.org/licenses/by/4.0/

Code: MIT http://opensource.org/licenses/MIT year: 2019, copyright holder: Patrick Schratz

Data: CC0 http://creativecommons.org/publicdomain/zero/1.0/

### Notes and resources

* The [issues tracker](https://github.com/pat-s/paper_hyperspectral/issues) is the place to report problems or ask questions

* See the repository [history](https://github.com/pat-s/paper_hyperspectral/commits/master) for a fine-grained view of progress and changes.

* The organisation of this compendium is based on the work of [Carl Boettiger](http://www.carlboettiger.info/) and [Ben Marwick](https://github.com/benmarwick).

## Literature

- What to do first (Feature selection or tuning): https://datascience.stackexchange.com/a/19778/24475
- Hastie, T., Friedman, J., & Tibshirani, R. (2001). The Elements of Statistical Learning. Springer New York. https://doi.org/10.1007/978-0-387-21606-5 (Kapitel 16)
- FSel algs overview: http://featureselection.asu.edu/algorithms.php
- De Jay, N., Papillon-Cavanagh, S., Olsen, C., El-Hachem, N., Bontempi, G., & Haibe-Kains, B. (2013). mRMRe: an R package for parallelized mRMR ensemble feature selection. Bioinformatics, 29(18), 2365–2368. https://doi.org/10/f48vxc
- Drotár, P., Gazda, J., & Smékal, Z. (2015). An experimental comparison of feature selection methods on two-class biomedical datasets. Computers in Biology and Medicine, 66, 1–10. https://doi.org/10/f72ttc
- Drotár, P., Šimoňák, S., Pietriková, E., Chovanec, M., Chovancová, E., Ádám, N., … Biňas, M. (2017). Comparison of Filter Techniques for Two-Step Feature Selection. COMPUTING AND INFORMATICS, 36(3), 597–617. https://doi.org/10/gbntp7
- Drotár, P., Gazda, M., & Gazda, J. (2017). Heterogeneous ensemble feature selection based on weighted Borda count. In 2017 9th International Conference on Information Technology and Electrical Engineering (ICITEE) (pp. 1–4). https://doi.org/10.1109/ICITEED.2017.8250495
- Nie, F., Huang, H., Cai, X., & Ding, C. H. (2010). Efficient and Robust Feature Selection via Joint \mathscrl2,1-Norms Minimization. In J. D. Lafferty, C. K. I. Williams, J. Shawe-Taylor, R. S. Zemel, & A. Culotta (Eds.), Advances in Neural Information Processing Systems 23 (pp. 1813–1821). Curran Associates, Inc. Retrieved from http://papers.nips.cc/paper/3988-efficient-and-robust-feature-selection-via-joint-l21-norms-minimization.pdf

### 2019-02-01 R/fun-data-preprocessing.R

@@ -1,0 +1,74 @@
+remove_na = function(data) {
+  # Remove columns with NA values and "ID" columns
+  # Ref: https://stackoverflow.com/questions/41343900/remove-row-columns-with-any-all-nan-values/41343980#41343980
+  # luiando has some NA obs that need to be removed, otherwise the whole column will be removed by `Filter()`
+  # for the nri_vi stack 3 more bands are removed than for the "bands" only
+  data[[3]] %<>%
+    na.omit()
+
+  data %<>%
+    map(~ dplyr::select(data, -contains("_ID"))) %>%
+    map(~ as.data.table(data)) %>%
+    #map(~ na.omit(.x)) %>% # some obs in luiando are NA after extraction
+    map(~ dplyr::select(data, -tree.number, -decoloration, -canker, -diameter, -height)) %>%
+    map(~ Filter(function(x) !any(is.na(x)), data))
+
+  data %<>%
+    map(~ st_as_sf(data, crs = 32630)) %>%
+    map(~ st_set_geometry(data, NULL))
+
+  return(data)
+}
+
+extract_coords = function(data) {
+
+  # Extract coords ----------------------------------------------------------
+
+  # Extract coordinates
+  # The scaling of variables won't work with the `geom` column attached.
+  # So we need to safe the coordinates, remove them and attach them afterwards again.
+  # Coordinates are needed for the spatial partitioning during modeling.
+
+  coords = map(data, ~ st_as_sf(data, crs = 32630)) %>%
+    map(~ as.data.frame(st_coordinates(data)))
+
+  return(coords)
+}
+
+standardize = function(data) {
+  data %<>%
+    pbmclapply(function(x) {
+      cols = names(x)[2:length(names(x))]
+      x[, (cols) := lapply(.SD, scale), .SDcols=cols]
+    }, mc.cores = 4
+    )
+
+  return(data)
+}
+
+mutate_defol = function(data) {
+
+  data %<>%
+    mutate(defoliation = case_when(.data$defoliation == 0 ~ 0.001,
+                                   TRUE ~ as.numeric(defoliation)))
+}
+
+log_response = function(data, response) {
+
+  data[[response]] = log(data[[response]])
+  return(data)
+}
+
+split_in_feature_sets = function(data, set) {
+
+  if (set == "nri") {
+    data_split = data[["data_vi_nri"]] %>%
+      select(matches("nri|defol"))
+  } else if (set == "vi") {
[…]

### 2019-02-01 R/fun-download.R

@@ -1,0 +1,35 @@
+download_trees = function(path) {
+  if (!file.exists("data/gpkg/oiartzun.gpkg")) {
+    curl_download(path,
+                  destfile = glue(tempdir(), "/trees.zip"), quiet = FALSE)
+    unzip(glue(tempdir(), "/trees.zip"), exdir = "data/gpkg")
+  }
+
+  files = map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~
+        st_read(glue("data/gpkg/{.x}.gpkg"), quiet = TRUE))
+  return(files)
+}
+
+download_locations = function(path) {
+  if (!file.exists("data/gpkg/all-plots.gpkg")) {
+    curl_download(path,
+                  destfile = "data/gpkg/all-plots.gpkg", quiet = FALSE)
+  }
+
+  files = st_read("data/gpkg/all-plots.gpkg") %>%
+    mutate(Name = as.character(ignore(Name)))
+  return(files)
+}
+
+download_hyperspectral = function(path) {
+  if (!file.exists("data/hyperspectral/B101_P1N_A090F03_ATM_S.tif")) {
+    curl_download(path,
+                  destfile = glue(tempdir(), "/hs.zip"), quiet = FALSE)
+      unzip(glue(tempdir(), "/hs.zip"), exdir = "data/hyperspectral/")
+  }
+
+  files = map(list.files("data/hyperspectral",
+                         full.names = TRUE, pattern = ".tif$"),
+              ~ stack(.x))
+  return(files)
+}

### 2019-02-01 R/fun-hyperspectral.R

@@ -1,0 +1,116 @@
+process_hyperspec = function(data, id, name, index, plots, name_out) {
+
+  plan(future.callr::callr, workers = 28)
+
+  # for later
+  dir_create("data/hyperspectral/vi/")
+  dir_create("data/hyperspectral/ndvi/")
+  dir_create("data/hyperspectral/nri/")
+
+  out = future_pmap(list(id, index, name_out), ~
+                      process_hyperspec_helper(data = data, id = ..1,
+                                               index = ..2, name_out = ..3,
+                                               plots = plots))
+
+  out %<>% set_names(name_out)
+  return(out)
+
+}
+
+process_hyperspec_helper <- function(data, id, name, index, plots, name_out) {
+
+  image <- data[[index]]
+
+  # project
+  plots %>%
+    dplyr::filter(Name == ignore(id)) %>%
+    st_transform(25830) %>%
+    as("Spatial") -> shape_single_plot
+
+  image_masked = image %>%
+    raster::crop(shape_single_plot) %>%
+    raster::mask(shape_single_plot) %>%
+    raster::projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
+  cat(glue("Finished {name_out}."))
+
+  return(image_masked)
+}
+
+extract_indices_to_plot = function(plot_name, buffer, bf_name, tree_data,
+                                   veg_indices, nri_indices) {
+
+  # calculate buffered veg index
+  out_veg <- future_map(buffer, function(x)
+    raster::extract(veg_indices, tree_data, buffer = x,
+                    fun = mean, df = TRUE,
+                    na.rm = TRUE), mc.cores = cores)
+
+  out_veg %<>%
+    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
+                                                name = names(out_veg[[.y]]))))
+
+  out_nbi <- future_map(buffer, function(y)
+    raster::extract(nri_indices, tree_data, buffer = y,
+                    fun = mean, df = TRUE,
+                    na.rm = TRUE))
+
+  out_nbi %<>%
+    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
+                                                name = names(out_nbi[[.y]]))))
+
+  # merge all data frames (buffers)
+  all_veg <- bind_cols(out_veg)
+
+  all_nbi <- bind_cols(out_nbi)
+
+  points %<>%
[…]

### 2019-02-01 analysis/download.R

@@ -1,0 +1,125 @@
+# Raw data -----------------------------------------------------------
+
+data_hs_raw = download_hyperspectral("https://ndownloader.figshare.com/articles/7570958/versions/1")
+
+plot_locations = download_locations("https://ndownloader.figshare.com/files/14090993")
+
+name_id = c("a busturi-axpe", "Laukiz I", "Laukiz II", "Laukiz III", "a altube",
+            "a barazar", "Laricio defol. Urkiola.", "aguinaga pinaster radiata",
+            "Oiartzun", "Pinast. Arza", "Pinas. Santa Koloma", "Rad. Menagaray 1",
+            "Rad. Menagaray 2", "Ayala/Aiara", "Silvest. Nograro", "a caranca",
+            "Picea Abornicano", "douglas 1", "douglas 2", "Rad. Altube",
+            "a otxandio", "Laricio defol. Olaeta. pol", "Douglas Legutio",
+            "Laricio defol.Gatzaga", "a azaceta", "a gipuzkoa",
+            "Sequoia Legorreta", "Douglas Albiztur")
+name_out = c("busturi", "laukiz1", "laukiz2", "laukiz3", "altube", "barazar",
+             "laricio-defol-urkiola", "aguinaga-pinaster-radiata", "oiartzun",
+             "pinast-arza", "pinas-santa-koloma", "rad-menagaray-1",
+             "rad-menagaray-2", "luiando", "silvest-nograro", "caranca",
+             "picea-abornicano", "douglas-1", "douglas-2", "rad-altube",
+             "otxandio", "laricio-defol-olaeta", "douglas-legutio",
+             "laricio-gatzaga", "azaceta", "gipuzkoa", "sequoia-legorreta",
+             "douglas-albiztur")
+index = c(1, 2, 2, 2, 3, 4, 5, 6, 7, 8, 9, 10, 10, 11, 12, 13, 14, 16, 16, 16,
+          17, 17, 18, 20, 21, 22, 22, 23)
+
+wavelength = c(404.08, 408.5, 412.92, 417.36, 421.81, 426.27, 430.73, 435.2,
+               439.69, 444.18, 448.68, 453.18, 457.69, 462.22, 466.75, 471.29,
+               475.83, 480.39, 484.95, 489.52, 494.09, 498.68, 503.26, 507.86,
+               512.47, 517.08, 521.7, 526.32, 530.95, 535.58, 540.23, 544.88,
[…]

### 2019-02-01 analysis/drake.R

@@ -1,0 +1,48 @@
+needs::needs(drake, hsdar, sf, raster, dplyr, purrr, glue, R.utils, pbmcapply,
+             magrittr, stringr, ggplot2, furrr, data.table, mlrCPO, curl, stringr,
+             fs)
+
+# Plans -----------------------------------------------------------
+
+data_plan = code_to_plan("analysis/data_preprocessing.R")
+download_plan = code_to_plan("analysis/download.R")
+hyperspectral_plan = code_to_plan("analysis/hyperspectral-processing.R")
+# learners = code_to_plan("code/learner/learner.R")
+# resampling = code_to_plan("code/mlr-settings/resampling.R")
+# param_set = code_to_plan("code/mlr-settings/param-set.R")
+# tune_ctrl = code_to_plan("code/mlr-settings/tune_ctrl_mbo.R")
+# tuning_wrapper = code_to_plan("code/mlr-settings/tuning.R")
+# sourceDirectory("code/benchmark/")
+# sourceDirectory("code/prediction/")
+# sourceDirectory("code/reports/")
+#
+sourceDirectory("R")
+
+#  grouping for visualization -----------------------------------------------------------------------
+
+data_plan$stage = "data"
+download_plan$stage = "data"
+hyperspectral_plan$stage = "data"
+# resampling$stage = "mlr_settings"
+# param_set$stage = "mlr_settings"
+# tune_ctrl$stage = "mlr_settings"
+# tuning_wrapper$stage = "learner"
+# benchmark_plan$stage = "benchmark"
+# prediction$stage = "prediction"
+# reports$stage = "reports"
+
+# # Combine all -------------------------------------------------------------
+#
+plan = bind_plans(data_plan, download_plan, hyperspectral_plan)
+
+# set cores of targets
+
+plan %<>%
+  mutate(resources = case_when(target == "data_hs_preprocessed" ~
+                                 list(list(cores = 20, gpus = 0)),
+                               target == "coords" ~
+                                 list(list(cores = 2, gpus = 0)),
+                               TRUE ~ list(list(cores = 1, gpus = 0)))
+  )
+
+config = drake_config(plan)

### 2019-02-01 analysis/hyperspectral-processing.R

@@ -1,0 +1,34 @@
+# crop rasters
+data_hs_preprocessed = process_hyperspec(data = data_hs_raw, plots = plot_locations,
+                                         name_out = name_out, index = index,
+                                         id = name_id)
+
+# create hyperspec rasters from cropped rasters
+#hyperspecs = map(data_hs_preprocessed, ~ HyperSpecRaster(.x, wavelength))
+
+# calculate various indices for all plots
+ndvi_rasters = map(data_hs_preprocessed, ~ HyperSpecRaster(.x, wavelength)) %>%
+  future_imap(~ vegindex(.x, "NDVI2",
+                         filename = glue("data/hyperspectral/ndvi/ndvi2-{.y}"),
+                         bnames = "NDVI"))
+
+veg_indices = map(data_hs_preprocessed, ~ HyperSpecRaster(.x, wavelength)) %>%
+  calc_veg_indices(indices)
+
+nri_indices = map(data_hs_preprocessed, ~ HyperSpecRaster(.x, wavelength)) %>%
+  calc_nri_indices(indices)
+
+# extract all indices to tree data
+trees_with_indices = future_map(c("laukiz1", "laukiz2", "oiartzun", "luiando"), ~
+                                  extract_indices_to_plot(.x, buffer = 2,
+                                                          bf_name = "bf2",
+                                                          tree_data = tree_per_tree,
+                                                          veg_indices = veg_indices,
+                                                          nri_indices = nri_indices))
+
+# extract all indices to tree data
+trees_with_bands = future_map(c("laukiz1", "laukiz2", "oiartzun", "luiando"), ~
+                                extract_bands_to_plot(.x, buffer = 2,
+                                                      bf_name = "bf2",
+                                                      tree_data = tree_per_tree,
+                                                      hyperspectral_bands = data_hs_preprocessed))

### 2019-02-01 code/data/data_preprocessing.R

@@ -1,53 +1,0 @@
-data_ls = list("vi_nri" = trees_with_indices, "bands" = trees_with_bands)
-
-#plan(ignore(multiprocess))
-# data_clean = plan(ignore(multiprocess), workers = 2); future_imap(data_ls, ~ remove_na(.x))
-#
-#plan(ignore(multiprocess))
-
-coords = future_imap(data_ls, ~ extract_coords(.x))
-
-#plan(ignore(multiprocess))
-#data_standardized = plan(ignore(multiprocess), workers = 2); future_imap(data_ls, ~ standardize(.x))
-
-# Merge into one datase
-data_vi_nri_clean = as_tibble(rbindlist(data_clean[[1]], fill = TRUE)) %>%
-  Filter(function(x) !any(is.na(x)), data_clean[[1]])
-
-coords_vi_nri = as_tibble(rbindlist(coords[[1]]))
-
-data_bands_clean = as_tibble(rbindlist(data_clean[[1]], fill = TRUE)) %>%
-  Filter(function(x) !any(is.na(x)), data_clean[[1]])
-
-coords_bands = as_tibble(rbindlist(coords[[1]]))
-
-# modify defoliation value (we get errors if defol is exactly zero)
-data_mod = map(list(data_vi_nri, data_bands_clean), ~ mutate_defol(.x))
-
-# split into feature sets -------------------------------------------------
-
-data_nri = split_in_feature_sets(data_mod, "nri")
-
-data_vi = split_in_feature_sets(data_mod, "vi")
-
-data_bands = split_in_feature_sets(data_mod, "bands")
-
-# log transform response variable -----------------------------------------
-
-data_nri_log = log_response(data_nri, "defoliation")
-data_vi_log = log_response(data_vi, "defoliation")
-data_bands_log = log_response(data_bands, "defoliation")
-
-# create tasks -----------------------------------------
-
-task_vi = makeRegrTask(id = "defoliation-all-plots-VI", data = data_vi_log,
-                       target = "defoliation", coordinates = coords_vi_nri,
-                       blocking = factor(rep(1:4, c(479, 451, 291, 529))))
-
-task_nri = makeRegrTask(id = "defoliation-all-plots-NRI", data = data_nri_log,
-                        target = "defoliation", coordinates = coords_vi_nri,
-                        blocking = factor(rep(1:4, c(479, 451, 291, 529))))
-
-task_bands = makeRegrTask(id = "defoliation-all-plots-bands", data = data_bands_log,
[…]

### 2019-02-01 code/data/hyperspectral-processing.R

@@ -1,47 +1,0 @@
-# crop rasters
-data_hs_preprocessed = future_pmap(list(name_id, index, name_out), ~
-                                      process_hyperspec(data = data_hs_raw, id = ..1,
-                                                        index = ..2, name_out = ..3,
-                                                        plots = plot_locations))
-
-# create hyperspec rasters from cropped rasters
-hyperspecs = map(data_hs_preprocessed, function(x) HyperSpecRaster(x, wavelength))
-
-# calculate various indices for all plots
-ndvi_rasters = pbmclapply(hyperspecs, function(x)
-  vegindex(x, "NDVI2",
-           filename = glue("data/hyperspectral/ndvi/ndvi2-{name}",
-                           name = basename(x@file@name)),
-           bnames = "NDVI"), mc.cores = 28)
-
-veg_indices <- pbmclapply(hyperspecs, function(x)
-  vegindex(x, indices,
-           filename =
-             str_replace(glue("data/hyperspectral/all-veg-indices/{name}",
-                              name = basename(x@file@name)), ".tif", ".grd"),
-           na.rm = TRUE, bnames = indices), mc.cores = 28)
-
-nri_indices <- pbmclapply(hyperspecs, function(x)
-  nbi_raster(x,
-             filename = str_replace(glue("data/hyperspectral/",
-                                         "narrow-band-indices/narrow-band-indices-{name}",
-                                         name = basename(x@file@name)), ".tif", ".grd"),
-             bnames_prefix = "NRI"), mc.cores = 15)
-
-# extract all indices to tree data
-trees_with_indices = pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),
-                                function(x) extract_indices_to_plot(x, buffer = 2,
-                                                                    cores = 1,
-                                                                    bf_name = "bf2",
-                                                                    tree_data = tree_per_tree,
-                                                                    veg_indices = veg_indices,
-                                                                    nri_indices = nri_indices),
-                                mc.cores = 4)
-
-# extract all indices to tree data
-trees_with_bands = pbmclapply(c("laukiz1", "laukiz2", "oiartzun", "luiando"),
-                              function(x) extract_bands_to_plot(x, buffer = 2,
-                                                                cores = 1, bf_name = "bf2",
-                                                                tree_data = tree_per_tree,
-                                                                hyperspectral_bands = data_hs_preprocessed),
-                              mc.cores = 4)

### 2019-02-01 code/drake.R

@@ -1,47 +1,0 @@
-needs::needs(drake, hsdar, sf, raster, dplyr, purrr, glue, R.utils, pbmcapply,
-             magrittr, stringr, ggplot2, furrr, data.table, mlrCPO)
-
-# Plans -----------------------------------------------------------
-
-data_plan = code_to_plan("code/data/data_preprocessing.R")
-download_plan = code_to_plan("code/data/download.R")
-hyperspectral_plan = code_to_plan("code/data/hyperspectral-processing.R")
-# learners = code_to_plan("code/learner/learner.R")
-# resampling = code_to_plan("code/mlr-settings/resampling.R")
-# param_set = code_to_plan("code/mlr-settings/param-set.R")
-# tune_ctrl = code_to_plan("code/mlr-settings/tune_ctrl_mbo.R")
-# tuning_wrapper = code_to_plan("code/mlr-settings/tuning.R")
-# sourceDirectory("code/benchmark/")
-# sourceDirectory("code/prediction/")
-# sourceDirectory("code/reports/")
-#
-sourceDirectory("code/functions/")
-
-#  grouping for visualization -----------------------------------------------------------------------
-
-data_plan$stage = "data"
-download_plan$stage = "data"
-hyperspectral_plan$stage = "data"
-# resampling$stage = "mlr_settings"
-# param_set$stage = "mlr_settings"
-# tune_ctrl$stage = "mlr_settings"
-# tuning_wrapper$stage = "learner"
-# benchmark_plan$stage = "benchmark"
-# prediction$stage = "prediction"
-# reports$stage = "reports"
-
-# # Combine all -------------------------------------------------------------
-#
-plan = bind_plans(data_plan, download_plan, hyperspectral_plan)
-
-# set cores of targets
-
-plan %<>%
-  mutate(resources = case_when(target == "data_hs_preprocessed" ~
-                                 list(list(cores = 20, gpus = 0)),
-                               target == "coords" ~
-                                 list(list(cores = 2, gpus = 0)),
-                               TRUE ~ list(list(cores = 1, gpus = 0)))
-  )
-
-config = drake_config(plan)

### 2019-02-01 code/functions/data/data-preprocessing.R

@@ -1,78 +1,0 @@
-remove_na = function(data) {
-  # Remove columns with NA values and "ID" columns
-  # Ref: https://stackoverflow.com/questions/41343900/remove-row-columns-with-any-all-nan-values/41343980#41343980
-  # luiando has some NA obs that need to be removed, otherwise the whole column will be removed by `Filter()`
-  # for the nri_vi stack 3 more bands are removed than for the "bands" only
-  data[[3]] %<>%
-    na.omit()
-
-  data %<>%
-    map(~ dplyr::select(data, -contains("_ID"))) %>%
-    map(~ as.data.table(data)) %>%
-    #map(~ na.omit(.x)) %>% # some obs in luiando are NA after extraction
-    map(~ dplyr::select(data, -tree.number, -decoloration, -canker, -diameter, -height)) %>%
-    map(~ Filter(function(x) !any(is.na(x)), data))
-
-  data %<>%
-    map(~ st_as_sf(data, crs = 32630)) %>%
-    map(~ st_set_geometry(data, NULL))
-
-  return(data)
-}
-
-# future_init = function(workers) {
-#   do.call(future::plan, list(strategy = multiprocess, ... = workers))
-# }
-
-extract_coords = function(data, future_setting) {
-
-  # Extract coords ----------------------------------------------------------
-
-  # Extract coordinates
-  # The scaling of variables won't work with the `geom` column attached.
-  # So we need to safe the coordinates, remove them and attach them afterwards again.
-  # Coordinates are needed for the spatial partitioning during modeling.
-
-  coords = map(data, ~ st_as_sf(data, crs = 32630)) %>%
-    map(~ as.data.frame(st_coordinates(data)))
-
-  return(coords)
-}
-
-standardize = function(data) {
-  data %<>%
-    pbmclapply(function(x) {
-      cols = names(x)[2:length(names(x))]
-      x[, (cols) := lapply(.SD, scale), .SDcols=cols]
-    }, mc.cores = 4
-    )
-
-  return(data)
-}
-
-mutate_defol = function(data) {
-
-  data %<>%
-    mutate(defoliation = case_when(.data$defoliation == 0 ~ 0.001,
-                                   TRUE ~ as.numeric(defoliation)))
-}
-
-log_response = function(data, response) {
-
-  data[[response]] = log(data[[response]])
-  return(data)
-}
-
-split_in_feature_sets = function(data, set) {
-
-  if (set == "nri") {
[…]

### 2019-02-01 code/functions/data/hyperspectral.R

@@ -1,80 +1,0 @@
-process_hyperspec <- function(data, id, name, index, plots, name_out, future_setup) {
-
-  image <- data[[index]]
-
-  # project
-  plots %>%
-    dplyr::filter(Name == ignore(id)) %>%
-    st_transform(25830) %>%
-    as("Spatial") -> shape_single_plot
-
-  image_masked = image %>%
-    raster::crop(shape_single_plot) %>%
-    raster::mask(shape_single_plot) %>%
-    raster::projectRaster(crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
-  cat(glue("Finished {name_out}."))
-
-  return(image_masked)
-}
-
-extract_indices_to_plot = function(plot_name, buffer, cores, bf_name, tree_data,
-                                   veg_indices, nri_indices) {
-
-  # calculate buffered veg index
-  out_veg <- pbmclapply(buffer, function(x)
-    raster::extract(veg_indices, tree_data, buffer = x,
-                    fun = mean, df = TRUE,
-                    na.rm = TRUE), mc.cores = cores)
-
-  out_veg %<>%
-    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
-                                                name = names(out_veg[[.y]]))))
-
-  out_nbi <- pbmclapply(buffer, function(x)
-    raster::extract(nri_indices, tree_data, buffer = x,
-                    fun = mean, df = TRUE,
-                    na.rm = TRUE) , mc.cores = cores)
-
-  out_nbi %<>%
-    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
-                                                name = names(out_nbi[[.y]]))))
-
-  # merge all data frames (buffers)
-  all_veg <- bind_cols(out_veg)
-
-  all_nbi <- bind_cols(out_nbi)
-
-  points %<>%
-    bind_cols(all_veg) %>%
-    bind_cols(all_nbi)
-
-  return(points)
-}
-
-extract_bands_to_plot = function(plot_name, buffer, cores, bf_name, tree_data, hyperspectral_bands) {
-
-  # calculate buffered veg index
-
-  # plan(multiprocess)
-  # out_bands <- future_map(buffer, ~
-  #   raster::extract(bands_ras[[5:126]], points, buffer = .x,
-  #                   fun = mean, df = TRUE,
-  #                   na.rm = TRUE),.progress = TRUE)
-
[…]

### 2019-02-01 drake-viz.R

@@ -1,58 +1,0 @@
-source("code/drake.R")
-
-# Visualization -----------------------------------------------------------
-
-### no data_plan grouping
-vis_drake_graph(config, group = "stage", clusters = c("task", "learner",
-                                                      "mlr_settings",
-                                                      "benchmark_plan",
-                                                      "prediction"),
-                targets_only = TRUE, show_output_files = FALSE)
-
-sankey_drake_graph(config, group = "stage", clusters = c("task", "learner",
-                                                                 "mlr_settings",
-                                                                 "benchmark_plan",
-                                                                 "prediction"),
-                   targets_only = TRUE, show_output_files = FALSE)
-
-drake_ggraph(config, group = "stage", clusters = c("task", "learner",
-                                                           "mlr_settings",
-                                                           "benchmark_plan",
-                                                           "prediction"),
-             targets_only = TRUE, show_output_files = FALSE) +
-  ggpubr::theme_pubr()
-
-### no learner and mlr_settings grouping
-vis_drake_graph(config, group = "stage", clusters = c("task", "benchmark_plan",
-                                                              "prediction"),
-                targets_only = TRUE, show_output_files = FALSE,
-                selfcontained = TRUE)
-### no benchmark_plan grouping
-vis_drake_graph(config, group = "stage", clusters = c("task", "data_plan",
-                                                              "learner", "mlr_settings"),
-                targets_only = TRUE, show_output_files = FALSE,
-                selfcontained = TRUE)
-
-#
-#
-# ### no benchmark_plan grouping
-#
-# vis_drake_graph(config, group = "stage", clusters = c("data_plan", "task", "learner",
-#                                                       "mlr_settings"),
-#                 targets_only = TRUE, show_output_files = FALSE,
-#                 navigationButtons = FALSE, file = "drake.png")
-vis_drake_graph(config, group = "stage", clusters = c("data", "task", "learner",
-                                                      "mlr_settings",
-                                                      "prediction"),
-                targets_only = TRUE, show_output_files = FALSE)
-sankey_drake_graph(config, group = "stage", clusters = c("data_plan", "task", "learner",
-                                                                 "mlr_settings"),
-                   targets_only = TRUE, show_output_files = FALSE)
-
-drake_ggraph(config, group = "stage", clusters = c("data_plan", "task", "learner",
-                                                           "mlr_settings"),
[…]

### 2019-03-09 README.Rmd
update

@@
 
-### Compendium URL
+**title**
 
-### Authors
+# Authors
 
@@
 
-### Contents
+# Contents
 
@@
 
-### How to use
+# How to use
 
-#### Read the code, access the data
+## Read the code, access the data
 
@@
 
-#### Install the R package
+## Install the R package
 
@@
 packrat::restore()
-source("analysis/drake.R")
-make(plan, keep_going = TRUE, console_log_file = stdout()) 
-# use more cores with make(plan, jobs = <number of cores>)
+r_make()
 ```
 
-## Runtime
+# Runtime
 
-Predicted total runtime (based on all target runtimes ever built and stored in the internal `config` file)
+Predicted total runtime.
+The time is based on all targets that have been created so far in the project.
+Estimated time will differ depending on the CPU speed and possible availability of a HPC.
 
 ```{r README-2 }
-source("analysis/drake.R")
-predict_runtime(config, from_scratch = TRUE)
+r_predict_runtime()
 ```
 
-Acceleration when parallelizing the `make()` call
-
-```{r README-3, warning=FALSE, message=FALSE}
-time <- c()
-for (jobs in 1:10){
-time[jobs] <- predict_runtime(
-drake_config(),
-jobs = jobs,
-from_scratch = TRUE,
-known_times = build_times(targets_only = TRUE)$elapsed
-)
-}
-
-library(ggplot2)
-ggplot(data.frame(time = time / 3600, jobs = ordered(1:10), group = 1)) +
-geom_line(aes(x = jobs, y = time, group = group)) +
-scale_y_continuous(breaks = 0:10 * 4, limits = c(0, 29)) +
-ggpubr::theme_pubr() +
-xlab("jobs argument of make()") +
-ylab("Predicted runtime of make() (hours)")
-```
-
-### Licenses
+# Licenses
 
@@
 
-### Notes and resources
+# Notes and resources
 
@@
 
-## Literature
+# Literature
 

### 2019-03-09 R/fun-download.R

@@ -1,22 +1,22 @@
 download_trees = function(path) {
-  if (!file.exists("data/gpkg/oiartzun.gpkg")) {
+  if (!file.exists("data/gpkg/oiartzun.shp")) {
     curl_download(path,
                   destfile = glue(tempdir(), "/trees.zip"), quiet = FALSE)
     unzip(glue(tempdir(), "/trees.zip"), exdir = "data/gpkg")
   }
 
   files = map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~
-        st_read(glue("data/gpkg/{.x}.gpkg"), quiet = TRUE))
+        st_read(glue("data/gpkg/{.x}.shp"), quiet = TRUE))
   return(files)
 }
 
 download_locations = function(path) {
-  if (!file.exists("data/gpkg/all-plots.gpkg")) {
+  if (!file.exists("data/gpkg/all-plots.shp")) {
     curl_download(path,
-                  destfile = "data/gpkg/all-plots.gpkg", quiet = FALSE)
+                  destfile = "data/gpkg/all-plots.shp", quiet = FALSE)
   }
 
-  files = st_read("data/gpkg/all-plots.gpkg") %>%
+  files = st_read("data/gpkg/all-plots.shp") %>%
     mutate(Name = as.character(ignore(Name)))
   return(files)
 }
@@ -30,6 +30,6 @@ download_hyperspectral = function(path) {
 
   files = map(list.files("data/hyperspectral",
                          full.names = TRUE, pattern = ".tif$"),
-              ~ stack(.x))
+              ~ brick(.x))
   return(files)
 }

### 2019-03-09 R/fun-hyperspectral.R

@@ -1,6 +1,7 @@
 process_hyperspec = function(data, id, name, index, plots, name_out) {
 
-  plan(future.callr::callr, workers = 28)
+  #plan(future.callr::callr, workers = 28)
+  plan("multisession", workers = ignore(28))
 
   # for later
   dir_create("data/hyperspectral/vi/")
@@ -12,7 +13,22 @@ process_hyperspec = function(data, id, name, index, plots, name_out) {
                                                index = ..2, name_out = ..3,
                                                plots = plots))
 
+  cat("Finished processing.")
+
   out %<>% set_names(name_out)
+
+  dir_create("data/raster/hs-preprocessed/")
+
+  # write to disk as otherwise it will be stored as a tmp file that is not accessible later on
+  iwalk(out, ~ writeRaster(.x, glue("data/raster/hs-preprocessed/hs-preprocessed-{.y}"),
+                                  overwrite = TRUE))
+
+  cat("Finished writing.")
+
+  out = map(list.files("data/raster/hs-preprocessed/", pattern = ".grd", full.names = TRUE), ~ brick(.x))
+
+  out %<>% set_names(sort(name_out))
+
   return(out)
 
 }
@@ -40,37 +56,39 @@ extract_indices_to_plot = function(plot_name, buffer, bf_name, tree_data,
                                    veg_indices, nri_indices) {
 
   # calculate buffered veg index
-  out_veg <- future_map(buffer, function(x)
-    raster::extract(veg_indices, tree_data, buffer = x,
-                    fun = mean, df = TRUE,
-                    na.rm = TRUE), mc.cores = cores)
-
-  out_veg %<>%
-    map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
-                                                name = names(out_veg[[.y]]))))
-
-  out_nbi <- future_map(buffer, function(y)
-    raster::extract(nri_indices, tree_data, buffer = y,
+  veg_out <- map(buffer, function(x)
+    raster::extract(veg_indices[[plot_name]], tree_data[[plot_name]], buffer = x,
                     fun = mean, df = TRUE,
                     na.rm = TRUE))
 
-  out_nbi %<>%
+  veg_out %<>%
     map2(seq_along(buffer), ~ setNames(.x, glue("bf{buffer}_{name}",
-                                                name = names(out_nbi[[.y]]))))
[…]

### 2019-03-09 analysis/download.R

@@ -58,7 +58,8 @@ indices <- c("Boochs", "Boochs2","CARI", "Carter", "Carter2", "Carter3", "Carter
 #       plot_locations)
 
 # load tree-per-tree data
-tree_per_tree = download_trees("https://ndownloader.figshare.com/articles/7588730/versions/1")
+tree_per_tree = download_trees("https://ndownloader.figshare.com/articles/7588730/versions/1") %>%
+  set_names(c("laukiz1", "laukiz2", "luiando", "oiartzun"))
 
 # heterobasidion_data = preprocessing_custom("https://data.mendeley.com/datasets/kmy95t22fy/2/files/41deb463-d378-46da-a25e-26044ce253fe/heterobasidion-armillaria.gpkg",
 #                                            study_area = data_basque, drop_vars = "armillaria",

### 2019-03-09 analysis/drake.R

@@ -1,48 +1,0 @@
-needs::needs(drake, hsdar, sf, raster, dplyr, purrr, glue, R.utils, pbmcapply,
-             magrittr, stringr, ggplot2, furrr, data.table, mlrCPO, curl, stringr,
-             fs)
-
-# Plans -----------------------------------------------------------
-
-data_plan = code_to_plan("analysis/data_preprocessing.R")
-download_plan = code_to_plan("analysis/download.R")
-hyperspectral_plan = code_to_plan("analysis/hyperspectral-processing.R")
-# learners = code_to_plan("code/learner/learner.R")
-# resampling = code_to_plan("code/mlr-settings/resampling.R")
-# param_set = code_to_plan("code/mlr-settings/param-set.R")
-# tune_ctrl = code_to_plan("code/mlr-settings/tune_ctrl_mbo.R")
-# tuning_wrapper = code_to_plan("code/mlr-settings/tuning.R")
-# sourceDirectory("code/benchmark/")
-# sourceDirectory("code/prediction/")
-# sourceDirectory("code/reports/")
-#
-sourceDirectory("R")
-
-#  grouping for visualization -----------------------------------------------------------------------
-
-data_plan$stage = "data"
-download_plan$stage = "data"
-hyperspectral_plan$stage = "data"
-# resampling$stage = "mlr_settings"
-# param_set$stage = "mlr_settings"
-# tune_ctrl$stage = "mlr_settings"
-# tuning_wrapper$stage = "learner"
-# benchmark_plan$stage = "benchmark"
-# prediction$stage = "prediction"
-# reports$stage = "reports"
-
-# # Combine all -------------------------------------------------------------
-#
-plan = bind_plans(data_plan, download_plan, hyperspectral_plan)
-
-# set cores of targets
-
-plan %<>%
-  mutate(resources = case_when(target == "data_hs_preprocessed" ~
-                                 list(list(cores = 20, gpus = 0)),
-                               target == "coords" ~
-                                 list(list(cores = 2, gpus = 0)),
-                               TRUE ~ list(list(cores = 1, gpus = 0)))
-  )
-
-config = drake_config(plan)

### 2019-03-09 exec-drake.R

@@ -1,0 +1,22 @@
+# make project
+
+r_make()
+
+# visualize
+
+r_vis_drake_graph(group = "stage", clusters = c("task", "learner",
+                                                "mlr_settings",
+                                                "benchmark_plan",
+                                                "prediction"),
+                  targets_only = TRUE, show_output_files = FALSE)
+
+r_vis_drake_graph()
+
+# see outdated
+
+r_outdated()
+
+# Misc
+r_predict_runtime()
+
+r_predict_workers()

### 2019-04-10 analysis/rmd/defoliation-maps.Rmd
add analysis/rmd

---
title: "Creation of prediction maps"
output: html_document
author: "Patrick Schratz, FSU Jena"
---

```{r setup}
knitr::opts_knit$set(base.dir = 'analysis/rmd/')
knitr::opts_chunk$set(fig.retina=3, fig.align = 'center',
                      out.width = "100%")
loadd(defoliation_raster_2017, defoliation_raster_2018,
      defoliation_df_2017, defoliation_df_2018)
```

# Defoliation evaluation {.tabset .tabset-fade}

## Maps

```{r defoliation-map-2017, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
create_prediction_map(list(defoliation_raster_2017), "xgboost")
```

```{r defoliation-map-2018, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
create_prediction_map(list(defoliation_raster_2018), "xgboost")
```

## Histograms

```{r}
ggplot(defoliation_df_2017, aes(x = defoliation)) +
  geom_histogram() +
  labs(x = "Defoliation (%)") +
  theme_pubr()
```

```{r}
ggplot(defoliation_df_2018, aes(x = defoliation)) +
  geom_histogram() +
  labs(x = "Defoliation (%)") +
  theme_pubr()
```

### 2019-04-12 analysis/rmd/cloud-cover.Rmd
add cloud-cover.Rmd to reports

---
title: "Caclulcate cloud cover of scenes"
output: html_document
author: "Marc Becker"
---

```{r setup, include=FALSE}
knitr::opts_knit$set(base.dir = 'analysis/rmd/')
knitr::opts_chunk$set(fig.retina=3, fig.align = 'center', fig.width = 6.93, fig.height = 6.13,
                      out.width = "100%")

loadd(records, aoi, cloud_mosaic)
```

```{r}
records %<>% split(records$beginposition)

# Unique dates
date_scenes =  
  records %>%
  map(~ dplyr::select(., beginposition)) %>%
  map(~ unique(.)) %>%
  map_chr(~ str_sub(., 1, 10))

# Scenes per mosaic/ date
count_scenes =
  records %>%
  map_dbl(~ nrow(.))

# Row number of mosaic
row_scenes =
  records %>%
  map(~ dplyr::select(., relativeorbitnumber)) %>%
  map(~ pull(.)) %>%
  map_chr(~ unique(.))

# Calculate area of aoi
aoi_area = st_area(aoi) %>%
  sum()

# Calculate cloud cover
cloud_scenes =
  cloud_mosaic %>%
  map(~ st_read(., quiet=TRUE)) %>%
  map(~ st_set_crs(., "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")) %>%
  map(~ st_intersection(st_buffer(., 0), aoi)) %>%
  map(possibly(~ st_union(.), NA)) %>%
  map(possibly(~ st_area(.), NA)) %>%
  map_dbl(possibly(~ sum(.), NA)) %>%
  map_dbl(possibly(~ ./aoi_area, NA))

ext_scenes =
  records %>%
  map(~ dplyr::select(., footprint)) %>%
  map_depth(2, ~ st_as_sfc(.)) %>%
  map_depth(2, ~ st_set_crs(., "+proj=longlat +datum=WGS84 +no_defs ")) %>%
  map_depth(2, ~ st_transform(., "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")) %>%
  set_names(date_scenes)
```

```{r}
tibble(date = date_scenes, row = row_scenes, count = count_scenes, clouds = cloud_scenes) %>%
  saveRDS("scenes.rds")
```

2018-04-19 O% Check
2018-06-23 0% Check
2018-07-23 0% Check
2018-08-12 0% Check
2018-08-27 0% Check
2018-09-01 0% Check
2018-09-11 0% Check
2018-09-26 2% Check

```{r}
mapView(ext_scenes[[56]]) + mapView(aoi)
```

### 2019-04-12 analysis/rmd/timeline.Rmd
add timeline.Rmd

---
title: "Timeline"
output: html_document
author: "Patrick Schratz, FSU Jena"
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(base.dir = 'analysis/rmd/')
knitr::opts_chunk$set(fig.retina=3, fig.align = 'center',
                      out.width = "100%")
```

```{r}
library(vistime)
library(tidyverse)
library(lubridate)

d = read.csv(stringsAsFactors = FALSE, text = "event,start,duration,group
structure,0,2,orga
polish repo/code,10,3,orga
select journal,9,1.2,orga
drake pipeline,0,3,coding
finish benchmarks,1,10,coding
figures and tables,10,4,coding
literature,5,3,writing
review #1,12,1.2,writing
optimize,13,1.2,writing
review #2,14,1.2,writing
optimize,15,1.2,writing
first draft,9,3.1,writing
submit,16,1,writing"
)

start_date = as_date(lubridate::ymd("2019-04-08", tz = "GMT"))
d$start = start_date + d$start * 7
d$end = d$start + d$duration * 7
vistime(d, linewidth = 20)
```

### 2019-04-24 analysis/rmd/cloud-cover.Rmd
style

@@
 ---
-title: "Caclulcate cloud cover of scenes"
+title: "Calculate cloud cover of scenes"
 output: html_document
@@
 knitr::opts_knit$set(base.dir = 'analysis/rmd/')
-knitr::opts_chunk$set(fig.retina=3, fig.align = 'center', fig.width = 6.93, fig.height = 6.13,
-                      out.width = "100%")
+knitr::opts_chunk$set(fig.retina = 3, fig.align = 'center', fig.width = 6.93, 
+                      fig.height = 6.13, out.width = "100%")
 

### 2019-04-24 exec-drake.R

@@ -4,11 +4,15 @@ r_make()
 
 # visualize
 
-r_vis_drake_graph(group = "stage", clusters = c("task", "learner",
-                                                "mlr_settings",
-                                                "benchmark_plan",
-                                                "prediction"),
-                  targets_only = TRUE, show_output_files = FALSE)
+r_vis_drake_graph(
+  group = "stage", clusters = c(
+    "task", "learner",
+    "mlr_settings",
+    "benchmark_plan",
+    "prediction"
+  ),
+  targets_only = TRUE, show_output_files = FALSE
+)
 
 r_vis_drake_graph()
 
@@ -21,8 +25,11 @@ r_predict_runtime()
 
 r_predict_workers()
 
-make(plan, verbose = 2, targets = "train_xgboost", lazy_load = "promise",
-     console_log_file = "log/drake.log", cache_log_file = "log/cache3.log",
-     caching = "worker", template = list(log_file = "log/worker%a.log", n_cpus= 16, memory = 120000),
-     prework = quote(future::plan(future::multisession, workers = 3)),
-     garbage_collection = TRUE, jobs = 1, parallelism = "clustermq")
+make(plan,
+  verbose = 2, targets = "data_list_nri_vi_bands", lazy_load = "promise",
+  console_log_file = "log/drake.log", cache_log_file = "log/cache3.log",
+  caching = "worker",
+  template = list(log_file = "log/worker%a.log", n_cpus = 16, memory = 120000),
+  prework = quote(future::plan(future::multisession, workers = 3)),
+  garbage_collection = TRUE, jobs = 1, parallelism = "clustermq"
+)

### 2019-05-03 README.Rmd
README
explain sub-projects in README

@@
 
-**title**
+**Modeling defoliation as a proxy for tree health: Comparison of feature-selection methods across multiple feature sets derived from hyperspectral data**
 
@@
 # Contents
+
+## Paper
 
@@
 - Predict defoliation to all available plots (24) and the whole Basque Country (at 200 m resolution)
+
+The following directories belong to this project
+
+- `analysis/01-download.R`
+- `analysis/02-hyperspectral-processing.R`
+- `analysis/04-data-processing.R`
+- `analysis/05-modeling/`
+
+## Other Content
+
+In addition, it contains the workflow for a sub-project: Predicting defoliation to the Basque Country (for the years 2017 and 2018) using Sentinel-2 data.
+
+The following "targets" are important for this sub-project:
+
+- `defoliation_raster` (in `analysis/03-sentinel-processing`)
 
@@
 r_make()
-```
-
-# Runtime
-
-Predicted total runtime.
-The time is based on all targets that have been created so far in the project.
-Estimated time will differ depending on the CPU speed and possible availability of a HPC.
-
-```{r README-2 }
-r_predict_runtime()
 ```

### 2019-05-03 analysis/rmd/defoliation-maps.Rmd
defoliation-maps: VI derivations working

@@
 loadd(defoliation_raster_2017, defoliation_raster_2018,
-      defoliation_df_2017, defoliation_df_2018)
+      defoliation_df)
 ```

### 2019-05-03 R/04-fun-sentinel-processing.R

@@ -324,7 +324,7 @@ mask_mosaic <- function(image_mosaic, cloud_mosaic, aoi, forest_mask) {
     ))
 
   return(image_mosaic %>%
-    map(~ str_glue("data/sentinel/image_mask/", .))
+    map(~ str_glue("data/sentinel/image_mask/", basename(.)))
   )
 
 }
@@ -345,9 +345,9 @@ mask_mosaic <- function(image_mosaic, cloud_mosaic, aoi, forest_mask) {
 calculate_vi <- function(image) {
 
   if (grepl("2017", image[1])) {
-    year = 2017
+    year = "2017"
   } else if (grepl("2018", image[1])) {
-    year = 2018
+    year = "2018"
   }
 
   # EVI

### 2019-05-14 analysis/rmd/defoliation-maps.Rmd
defoliaton maps ready

@@
 ---
-title: "Creation of prediction maps"
-output: html_document
-author: "Patrick Schratz, FSU Jena"
+title: "Prediction maps of defoliation at trees"
+output: 
+  html_document: 
+    css: style.css
+    theme: readable
+    number_sections: yes
+    toc: yes
+    toc_float: true
+    toc_collapsed: true
+    df_print: tibble
+    includes:
+      in_header: header.html
+author: "Patrick Schratz, Marc Becker, Friedrich-Schiller-University Jena"
 ---
 
-```{r setup}
-knitr::opts_knit$set(base.dir = 'analysis/rmd/')
-knitr::opts_chunk$set(fig.retina=3, fig.align = 'center',
-                      out.width = "100%")
-loadd(defoliation_raster_2017, defoliation_raster_2018,
-      defoliation_df)
+```{r setup, include=FALSE}
+knitr::opts_knit$set(base.dir = "analysis/rmd/")
+knitr::opts_chunk$set(
+  fig.retina = 3,
+  fig.align = "center",
+  out.width = "100%",
+  echo = FALSE
+)
+
+# load drake objects
+loadd(defoliation_raster,
+      defoliation_raster_relative,
+      defoliation_df,
+      defoliation_df_relative,
+
+      tune_xgboost
+)
 ```
+
+# Predicted defoliation (absolute)
+
+Predicted defoliation by XGBOOST on 20 m spatial resolution in the Basque Country.
+
+With an RMSE of 36 % (for a range of 0 - 100), the model focuses its prediction to the mean of the response range (0 - 100) to achieve acceptable performance results on the test set.
+
+The following vegetation indices have been used, selected by the internal variable importance rating of XGBOOST:
+
+- EVI
+- GDVI4
+- GDVI3
+- GDVI2
+- D1
+- mNDVI
+- mSR
+
+# Optimization path of hyperparameter tuning
+
+The following shows the optimization path (RMSE) of the hyperparameter tuning for the XGBOOST model using 20 MBO iterations (with a starting design of 30).
+
+```{r}
+getOptPathY(tune_xgboost$opt.path)
+```
+
+Some more details on the optimization path are presented below:
+
+```{r}
+plotOptPath(tune_xgboost$opt.path)
+```
+
+# Predicted Defoliation (relative)
+
+Since these absolute values do not reflect the truth, we decided to create a relative Index from the predicted absolute values.
+
+This was done by calling
+
+```{r , eval=FALSE}
+scale(data$defoliation, center = FALSE,
+      scale = max(data$defoliation, na.rm = TRUE)/100)
+```
+
+which re-scales the data from 0 - 100. 
+The code was adapted from [this](https://stackoverflow.com/a/19462405/4185785) Stackoverflow answer.
 
@@
 
-## Maps
+## Absolute defoliation {.tabset .tabset-fade}
+
+### Maps
+
+#### 2017
 
 ```{r defoliation-map-2017, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
-create_prediction_map(list(defoliation_raster_2017), "xgboost")
+plot = create_defoliation_map(defoliation_raster[[1]],
+                              algorithm = "xgboost",
+                              limits = c(35, 60), title = "Defoliation of trees [%]"
+)
+suppressWarnings(suppressMessages(print(plot)))
 ```
 
+#### 2018
+
 ```{r defoliation-map-2018, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
-create_prediction_map(list(defoliation_raster_2018), "xgboost")
+plot = create_defoliation_map(defoliation_raster[[2]],
+                              algorithm = "xgboost",
+                              limits = c(35, 60), title = "Defoliation of trees [%]"
+)
+suppressWarnings(suppressMessages(print(plot)))
 ```
 
-## Histograms
+### Histograms
 
 ```{r}
-ggplot(defoliation_df_2017, aes(x = defoliation)) +
+suppressWarnings(suppressMessages(ggplot(defoliation_df[[1]], aes(x = defoliation)) +
   geom_histogram() +
@@
   theme_pubr()
+))
 ```
@@
 ```{r}
-ggplot(defoliation_df_2018, aes(x = defoliation)) +
+suppressWarnings(suppressMessages(ggplot(defoliation_df[[2]], aes(x = defoliation)) +
   geom_histogram() +
@@
   theme_pubr()
+))
 ```
+
+## Relative defoliation {.tabset .tabset-fade}
+
+### Maps
+
+#### 2017
+
+```{r defoliation-map-relative-2017, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
+plot = create_defoliation_map(defoliation_raster_relative[[1]],
+                              algorithm = "xgboost",
+                              limits = c(65, 100), title = "Relative defoliation of trees"
+)
+suppressWarnings(suppressMessages(print(plot)))
+```
+
+#### 2018
+
+```{r defoliation-map-relative-2018, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
+plot = create_defoliation_map(defoliation_raster_relative[[2]],
+                              algorithm = "xgboost",
+                              limits = c(65, 100), title = "Relative defoliation of trees"
+)
+suppressWarnings(suppressMessages(print(plot)))
+```
+
+### Histograms
+
+```{r}
+suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[1]], aes(x = defoliation)) +
+                   geom_histogram() +
+                   labs(x = "Relative defoliation of trees") +
+                   theme_pubr()
+))
+```
+
+```{r}
+suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[2]], aes(x = defoliation)) +
+                   geom_histogram() +
+                   labs(x = "Relative defoliation of trees") +
+                   theme_pubr()
+))
+```

### 2019-05-14 R/04-fun-sentinel-processing.R

@@ -324,9 +324,41 @@ mask_mosaic <- function(image_mosaic, cloud_mosaic, aoi, forest_mask) {
     ))
 
   return(image_mosaic %>%
-    map(~ str_glue("data/sentinel/image_mask/", basename(.)))
-  )
+    map(~ str_glue("data/sentinel/image_mask/", basename(.))))
+}
 
+#' @title mask_vi
+#' @description Masks the VIs
+#'
+#' @param image_vi (`character`)\cr File name of raster brick/stack.
+#' @return masked VIs (`brick`)
+#' @export
+mask_vi <- function(image_vi) {
+
+  # Get raster
+  veg_inds <- image_vi %>%
+    as.list() %>%
+    flatten() %>%
+    map(~ raster(.))
+
+  # Mask with each other
+  for (ind in veg_inds) {
+    veg_inds %<>%
+      map(~ raster::mask(., ind))
+  }
+
+  return(veg_inds)
+
+  # browser()
+  # # Get names
+  # ind_names <-
+  #   image_vi %>%
+  #   as.list() %>% flatten() %>%
+  #   map_chr(~ tools::file_path_sans_ext(basename(.)))
+  #
+  # # Return for drake
+  # veg_inds %>%
+  #   set_names(ind_names)
 }
 
 #' @title calculate_vi
@@ -343,11 +375,10 @@ mask_mosaic <- function(image_mosaic, cloud_mosaic, aoi, forest_mask) {
 #' @return calculated VIs (file name)
 #' @export
 calculate_vi <- function(image) {
-
   if (grepl("2017", image[1])) {
-    year = "2017"
+    year <- "2017"
   } else if (grepl("2018", image[1])) {
-    year = "2018"
+    year <- "2018"
   }
 
   # EVI
@@ -408,39 +439,11 @@ calculate_vi <- function(image) {
 
 
   return(list(
-    list.files("data/sentinel/image_vi", full.names = TRUE,
-                         pattern = year)
+    list.files("data/sentinel/image_vi",
+      full.names = TRUE,
+      pattern = year
     )
-  )
-}
-
-#' @title mask_vi
-#' @description Masks the VIs
-#'
-#' @param image_vi (`character`)\cr File name of raster brick/stack.
-#' @return masked VIs (`brick`)
-#' @export
-mask_vi <- function(image_vi) {
-
-  # Get raster
-  veg_inds <-
-    image_vi %>%
-    map(~ raster(.))
[…]

### 2019-05-21 analysis/rmd/defoliation-maps.Rmd
upd defoliation-maps.Rmd

@@
 ```{r}
-getOptPathY(tune_xgboost$opt.path)
+suppressMessages(getOptPathY(tune_xgboost$opt.path))
 ```
@@
 ```{r}
-plotOptPath(tune_xgboost$opt.path)
+suppressMessages(plotOptPath(tune_xgboost$opt.path))
 ```
@@
 
-```{r , eval=FALSE}
+```{r , eval=FALSE, echo = TRUE}
 scale(data$defoliation, center = FALSE,
@@
 plot = create_defoliation_map(defoliation_raster[[1]],
-                              algorithm = "xgboost",
+                              algorithm = "XGBOOST",
                               limits = c(35, 60), title = "Defoliation of trees [%]"
@@
 plot = create_defoliation_map(defoliation_raster[[2]],
-                              algorithm = "xgboost",
+                              algorithm = "XGBOOST",
                               limits = c(35, 60), title = "Defoliation of trees [%]"
@@
 
+2017
+
 ```{r}
-suppressWarnings(suppressMessages(ggplot(defoliation_df[[1]], aes(x = defoliation)) +
+hist = suppressWarnings(suppressMessages(ggplot(defoliation_df[[1]], aes(x = defoliation)) +
   geom_histogram() +
@@
 ))
+
+suppressMessages(print(hist))
 ```
 
+2018
+
 ```{r}
-suppressWarnings(suppressMessages(ggplot(defoliation_df[[2]], aes(x = defoliation)) +
+hist = suppressWarnings(suppressMessages(ggplot(defoliation_df[[2]], aes(x = defoliation)) +
   geom_histogram() +
@@
 ))
+suppressMessages(print(hist))
 ```
@@
 plot = create_defoliation_map(defoliation_raster_relative[[1]],
-                              algorithm = "xgboost",
+                              algorithm = "XGBOOST",
                               limits = c(65, 100), title = "Relative defoliation of trees"
@@
 plot = create_defoliation_map(defoliation_raster_relative[[2]],
-                              algorithm = "xgboost",
+                              algorithm = "XGBOOST",
                               limits = c(65, 100), title = "Relative defoliation of trees"
@@
 
+2017
+
 ```{r}
-suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[1]], aes(x = defoliation)) +
+hist = suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[1]], aes(x = defoliation)) +
                    geom_histogram() +
@@
 ))
+suppressMessages(print(hist))
 ```
 
+2018
+
 ```{r}
-suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[2]], aes(x = defoliation)) +
+hist = suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[2]], aes(x = defoliation)) +
                    geom_histogram() +
@@
 ))
+suppressMessages(print(hist))
 ```
+
+```{r echo = FALSE}
+rmarkdown::render(here("analysis/rmd/defoliation-maps.md"), 
+                  output_file = here("analysis/rmd/defoliation-maps.html"))
+```
+
+```{r echo = FALSE}
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/analysis/rmd/defoliation-maps.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/report-defoliation.html")
+```

### 2019-05-22 analysis/about.Rmd
first version

--- 2019-05-03 README.Rmd
@@
 ---
-output: github_document
+title: "About"
+output:
+  workflowr::wflow_html:
+    toc: false
+editor_options:
+  chunk_output_type: console
 ---
-
-[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/pat-s/paper_hyperspectral/commits/master)  
-[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/)   
-[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)  
-<!-- [![Travis-CI Build Status](https://travis-ci.org/pat-s/paper_hyperspectral.png?branch=master)](https://travis-ci.org/pat-s/paper_hyperspectral)  -->
 

### 2019-05-22 analysis/defoliation-maps.Rmd
first version

--- 2019-05-21 analysis/rmd/defoliation-maps.Rmd
@@
 title: "Prediction maps of defoliation at trees"
-output: 
-  html_document: 
-    css: style.css
-    theme: readable
-    number_sections: yes
-    toc: yes
-    toc_float: true
-    toc_collapsed: true
-    df_print: tibble
-    includes:
-      in_header: header.html
+output: workflowr::wflow_html
+editor_options:
+  chunk_output_type: console
 author: "Patrick Schratz, Marc Becker, Friedrich-Schiller-University Jena"
@@
 ```{r setup, include=FALSE}
-knitr::opts_knit$set(base.dir = "analysis/rmd/")
+#knitr::opts_knit$set(base.dir = "analysis/rmd/")
 knitr::opts_chunk$set(
@@
 )
+source(here::here("code/99-packages.R"))
+sourceDirectory("R")
 
@@
 
-```{r defoliation-map-2017, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
+```{r defoliation-map-2017, fig.height = 5.5, fig.width = 8.5}
 plot = create_defoliation_map(defoliation_raster[[1]],
@@
 
-```{r defoliation-map-2018, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
+```{r defoliation-map-2018, fig.height = 5.5, fig.width = 8.5}
 plot = create_defoliation_map(defoliation_raster[[2]],
@@
 
-```{r defoliation-map-relative-2017, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
+```{r defoliation-map-relative-2017, fig.height = 5.5, fig.width = 8.5}
 plot = create_defoliation_map(defoliation_raster_relative[[1]],
@@
 
-```{r defoliation-map-relative-2018, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
+```{r defoliation-map-relative-2018, fig.height = 5.5, fig.width = 8.5}
 plot = create_defoliation_map(defoliation_raster_relative[[2]],
@@
 ```{r echo = FALSE}
-rmarkdown::render(here("analysis/rmd/defoliation-maps.md"), 
-                  output_file = here("analysis/rmd/defoliation-maps.html"))
+# rmarkdown::render(here("analysis/rmd/defoliation-maps.md"), 
+#                   output_file = here("analysis/rmd/defoliation-maps.html"))
 ```
@@
 ```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/analysis/rmd/defoliation-maps.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/report-defoliation.html")
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/analysis/defoliation-maps.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/report-defoliation.html")
 ```

### 2019-05-22 analysis/license.Rmd
first version

---
title: "License"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

# MIT License

Copyright (c) 2019 Patrick Schratz

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.

### 2019-05-22 analysis/index.Rmd
update index

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Welcome to my research website.

Click on this [link](defoliation-maps.html) to see my results.

### 2019-05-22 analysis/index.Rmd
restructure for workflowr

@@
 
-Welcome to my research website.
+[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/pat-s/paper_hyperspectral/commits/master)  
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/)   
+[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
 
-Click on this [link](defoliation-maps.html) to see my results.
+[Project Description](about.html)
+
+# Results
+
+[Report: Defoliation](defoliation-maps.html).
+
+# 
+Powered by [drake](https://github.com/ropensci/drake) + [workflowr](https://github.com/jdblischak/workflowr) and [mlr](https://github.com/mlr-org/mlr)

### 2019-05-22 _drake.R

@@ -1,30 +1,30 @@
-source("analysis/99-packages.R")
+source("code/99-packages.R")
 
 # Plans -----------------------------------------------------------
 
-download_plan = code_to_plan("analysis/01-download.R")
-hyperspectral_plan = code_to_plan("analysis/02-hyperspectral-processing.R")
-sentinel_plan = code_to_plan("analysis/03-sentinel-processing.R")
-data_plan = code_to_plan("analysis/04-data-preprocessing.R")
+download_plan = code_to_plan("code/01-download.R")
+hyperspectral_plan = code_to_plan("code/02-hyperspectral-processing.R")
+sentinel_plan = code_to_plan("code/03-sentinel-processing.R")
+data_plan = code_to_plan("code/04-data-preprocessing.R")
 
 # project
-learners_plan = code_to_plan("analysis/05-modeling/project/learner.R")
-resampling_plan = code_to_plan("analysis/05-modeling/project/resamp.R")
-param_set_plan = code_to_plan("analysis/05-modeling/project/param-set.R")
-tune_ctrl_plan = code_to_plan("analysis/05-modeling/project/tune-ctrl.R")
-tuning_plan = code_to_plan("analysis/05-modeling/project/tune.R")
-train_plan = code_to_plan("analysis/05-modeling/project/train.R")
-task_plan = code_to_plan("analysis/05-modeling/project/task.R")
+learners_plan = code_to_plan("code/05-modeling/project/learner.R")
+resampling_plan = code_to_plan("code/05-modeling/project/resamp.R")
+param_set_plan = code_to_plan("code/05-modeling/project/param-set.R")
+tune_ctrl_plan = code_to_plan("code/05-modeling/project/tune-ctrl.R")
+tuning_plan = code_to_plan("code/05-modeling/project/tune.R")
+train_plan = code_to_plan("code/05-modeling/project/train.R")
+task_plan = code_to_plan("code/05-modeling/project/task.R")
 
 # paper
-learners_paper_plan = code_to_plan("analysis/05-modeling/paper/learner.R")
-resampling_paper_plan = code_to_plan("analysis/05-modeling/paper/resampling.R")
-param_set_paper_plan = code_to_plan("analysis/05-modeling/paper/param-sets.R")
-tune_ctrl_paper_plan = code_to_plan("analysis/05-modeling/paper/tune-ctrl.R")
-tuning_paper_plan = code_to_plan("analysis/05-modeling/paper/tune-wrapper.R")
[…]

### 2019-05-22 analysis/07-reports.R

@@ -1,14 +1,0 @@
-reports_plan = drake_plan(
-
-  defoliation_maps = knit(
-    knitr_in("analysis/rmd/defoliation-maps.Rmd"),
-    output = file_out("analysis/rmd/defoliation-maps.md"), # Important: must be the output file
-    quiet = TRUE),
-
-  cloud_cover = knit(
-    knitr_in("analysis/rmd/cloud-cover.Rmd"),
-    output = file_out("analysis/rmd/cloud-cover.md"), # Important: must be the output file
-    quiet = TRUE)
-
-)
-

### 2019-05-22 analysis/report-defoliation.Rmd
wflow_publish("analysis/report-defoliation.Rmd")

--- 2019-05-22 analysis/defoliation-maps.Rmd
@@
 title: "Prediction maps of defoliation at trees"
-output: workflowr::wflow_html
+output: 
+  workflowr::wflow_html:
+    css: style.css
+    theme: readable
+    number_sections: yes
+    toc: yes
+    toc_float: true
+    toc_collapsed: true
+    df_print: tibble
+    includes:
+      in_header: header.html
 editor_options:
@@
 ```{r setup, include=FALSE}
-#knitr::opts_knit$set(base.dir = "analysis/rmd/")
 knitr::opts_chunk$set(
@@
 ```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/analysis/defoliation-maps.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/report-defoliation.html")
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/analysis/report-defoliation.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/"
+       )
 ```
+
+```{r echo = FALSE}
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super \
+  /home/patrick/papers/2019-feature-selection/docs/figure/defoliation-maps.Rmd/* \
+  -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/figure/report-defoliation.Rmd/"
+)
+```

### 2019-05-23 analysis/index.Rmd
wflow_publish(all = T)

@@
 
-[Report: Defoliation](defoliation-maps.html).
+- [Report: Defoliation](defoliation-maps.html)
 
 # 
-Powered by [drake](https://github.com/ropensci/drake) + [workflowr](https://github.com/jdblischak/workflowr) and [mlr](https://github.com/mlr-org/mlr)
+Powered by [drake](https://github.com/ropensci/drake) + [workflowr](https://github.com/jdblischak/workflowr) + [mlr](https://github.com/mlr-org/mlr)

### 2019-05-23 analysis/report-defoliation.Rmd
wflow_publish(all = T)

@@
 title: "Prediction maps of defoliation at trees"
-output: 
-  workflowr::wflow_html:
-    css: style.css
-    theme: readable
-    number_sections: yes
-    toc: yes
-    toc_float: true
-    toc_collapsed: true
-    df_print: tibble
-    includes:
-      in_header: header.html
+output: workflowr::wflow_html
 editor_options:

### 2019-05-23 inst/rsync-jupiter.sh

@@ -1,2 +1,13 @@
-rsync -rlptDvzog --chown=www-data:www-data --fake-super analysis/paper/submission/3/appendices/prediction-map* -e ssh
-                   │patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-pathogen-infection/
+########
+# LIFE
+########
+
+# `defoliation-report` --------------- copy defoliation report to LIFE
+rsync -rlptDvzog --chown=www-data:www-data --fake-super \
+  /home/patrick/papers/2019-feature-selection/analysis/rmd/defoliation-maps.html \
+  -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/report-defoliation.html
+
+# `defoliation-maps` --------------- copy defoliation report to LIFE
+rsync -rlptDvzog --chown=www-data:www-data --fake-super \
+  /home/patrick/papers/2019-feature-selection/analysis/figures/defoliation-map* \
+  -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/

### 2019-05-23 analysis/report-defoliation.Rmd
wflow_publish(all = T)

@@
 ```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/analysis/report-defoliation.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/"
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/docs/report-defoliation.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/"
        )

### 2019-05-23 analysis/report-defoliation.Rmd
wflow_publish("analysis/report-defoliation.Rmd")

@@
 ```{r echo = FALSE}
-# rmarkdown::render(here("analysis/rmd/defoliation-maps.md"), 
-#                   output_file = here("analysis/rmd/defoliation-maps.html"))
-```
-
-```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/docs/report-defoliation.html -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/"
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super \
+  /home/patrick/papers/2019-feature-selection/docs/figure/defoliation-maps.Rmd/* \
+  -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
 )
 ```
-
-```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super \
-  /home/patrick/papers/2019-feature-selection/docs/figure/defoliation-maps.Rmd/* \
-  -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/figure/report-defoliation.Rmd/"
-)
-```

### 2019-05-24 analysis/report-defoliation.Rmd
wflow_publish("analysis/report-defoliation.Rmd")

@@
 title: "Prediction maps of defoliation at trees"
-output: workflowr::wflow_html
+output: 
+  workflowr::wflow_html:
+    includes:
+      in_header: header.html
 editor_options:

### 2019-05-24 analysis/report-defoliation.Rmd
update scripts

@@
 ```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super \
-  /home/patrick/papers/2019-feature-selection/docs/figure/defoliation-maps.Rmd/* \
-  -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/* -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
 )

### 2019-05-24 code/07-reports.R

@@ -1,16 +1,7 @@
 reports_plan = drake_plan(
 
-  # defoliation_maps = knit(
-  #   knitr_in("analysis/rmd/defoliation-maps.Rmd"),
-  #   output = file_out("analysis/rmd/defoliation-maps.md"),
-  #   quiet = TRUE),
+  defoliation_maps_wfr = wflow_publish(knitr_in("analysis/report-defoliation.Rmd"), view = FALSE),
 
-  defoliation_maps_wfr = wflow_build(knitr_in("analysis/defoliation-maps.Rmd"), view = FALSE)#,
-
-  # cloud_cover = knit(
-  #   knitr_in("analysis/rmd/cloud-cover.Rmd"),
-  #   output = file_out("analysis/rmd/cloud-cover.md"),
-  #   quiet = TRUE)
-
+  strings_in_dots = "literals"
 )
 

### 2019-05-24 analysis/report-defoliation.Rmd
wflow_publish(knitr_in("analysis/report-defoliation.Rmd"), view = FALSE)

@@
 ```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/* -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
+system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
 )

### 2019-05-27 analysis/index.Rmd
analysis/timeline.Rmd

--- 2019-05-22 analysis/index.Rmd
@@
 
-Welcome to my research website.
+[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/pat-s/2019-feature-selection/commits/master) 
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2582969.svg)](https://doi.org/10.5281/zenodo.2582969)
 
-Click on this [link](defoliation-maps.html) to see my results.
+This [workflowr]() site contains all information and results about this research project.
+
+Project details can be found in the [About](about.html) page.
+
+**Result shortcuts**
+
+1. [Prediction maps of defoliation at trees](report-defoliation.html)
+
+Powered by [drake](https://github.com/ropensci/drake) + [workflowr](https://github.com/jdblischak/workflowr) + [mlr](https://github.com/mlr-org/mlr).

### 2019-05-27 analysis/about.Rmd
wflow_publish("analysis/about.Rmd")

@@
 
-- `analysis/01-download.R`
-- `analysis/02-hyperspectral-processing.R`
-- `analysis/04-data-processing.R`
-- `analysis/05-modeling/`
+- `code/01-download.R`
+- `code/02-hyperspectral-processing.R`
+- `code/04-data-processing.R`
+- `code/05-modeling/`
+- `code/06-benchmark-matrix/`
+- `code/07-reports/`
 
@@
 
-In addition, it contains the workflow for a sub-project: Predicting defoliation to the Basque Country (for the years 2017 and 2018) using Sentinel-2 data.
+In addition, this repo contains the workflow for an analysis related to the [LIFE 14 ENV/ES/000179 LIFE HEALTHY FOREST](http://www.lifehealthyforest.com/) project: Predicting defoliation at trees for the Basque Country (for the years 2017 and 2018) using Sentinel-2 data.
 
-The following "targets" are important for this sub-project:
-
-- `defoliation_raster` (in `analysis/03-sentinel-processing`)
+Target `defoliation_maps_wfr` builds are targets necessary for the final results [report](https://2019-feature-selection.pjs-web.de/report-defoliation.html).
 
@@
 
-See the [`analysis`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis) directory on GitHub for the source code that generated the figures and statistical results contained in the manuscript.
+See the [`code`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis) directory on GitHub for the source code that generated the figures and statistical results contained in the manuscript.
 See the [`data`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis/data) directory for instructions how to access the raw data discussed in the manuscript.
@@
 ```
-
-# Licenses
-
-Text: CC-BY-4.0 http://creativecommons.org/licenses/by/4.0/
-
-Code: MIT http://opensource.org/licenses/MIT year: 2019, copyright holder: Patrick Schratz
-
-Data: CC0 http://creativecommons.org/publicdomain/zero/1.0/
 

### 2019-05-29 analysis/report-defoliation.Rmd
minor

--- 2019-05-24 analysis/report-defoliation.Rmd
@@
 ```{r echo = FALSE}
-system("rsync -rlptDvzog --chown=www-data:www-data --fake-super /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/* -e ssh patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
+system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
 )

### 2019-05-29 code/07-reports.R

@@ -1,7 +1,5 @@
 reports_plan = drake_plan(
 
-  defoliation_maps_wfr = wflow_publish(knitr_in("analysis/report-defoliation.Rmd"), view = FALSE),
-
-  strings_in_dots = "literals"
+  defoliation_maps_wfr = wflow_publish(knitr_in("analysis/report-defoliation.Rmd"), view = FALSE)
 )
 

### 2019-06-01 analysis/paper/Schratz_hyperspectral.tex
restructure
major refactoring of manuscript

--- 2018-08-09 01_writing/Schratz_hyperspectral.tex
@@
 \usepackage{float}
+
+% https://tex.stackexchange.com/questions/165115/getting-not-defining-perthousnad-and-not-defining-micro-when-compiling-beamer
+\usepackage{textcomp}
 
@@
 	\author[NEIKER]{Eugenia Iturritxa}
-	%\author[TUDO]{Jakob Richter}
+	\author[LMU]{Bernd Bischl}
 	\author[FSU]{Alexander Brenning}
@@
 	\address[FSU]{Department of Geography, GIScience group, Grietgasse 6, 07743, Jena, Germany}
-	%\address[NEIKER]{NEIKER, Granja Modelo –Arkaute, Apdo. 46, 01080 Vitoria-Gasteiz, Arab, Spain}
-	%\address[TUDO]{Department of Statistics, TU Dortmund University, Germany}
+	\address[NEIKER]{NEIKER, Granja Modelo –Arkaute, Apdo. 46, 01080 Vitoria-Gasteiz, Arab, Spain}
+	\address[LMU]{Department of Statistics, Chair for computational Statistics, Ludwig-Maximilian University Munich, Germany}
 
@@
 	\begin{keyword}
-		hyperspectral imagery \sep forest health \sep machine-learning \sep variable importance \sep model comparison
+		hyperspectral imagery \sep forest health modeling \sep machine-learning \sep feature-selection  \sep model comparison
 	\end{keyword}
@@
 % Explain how remote sensing is used in forestry (potential to map forest health)
-\noindent Data retrieved from remote sensing satellites is successfully used in forestry to monitor temporal changes across large areas \citep{martinezdelcastilloEvaluationForestCover2015,sextonModelPropagationUncertainty2015}.
-The use of \ac{SAR} techniques enables scientists to estimate \ac{AGB} \citep{luSurveyRemoteSensingbased2016, sinhaReviewRadarRemote2015}.
-Forest health is commonly assesed using optical data from multi-/hyperspectral satellites by applying either temporal change detections \citep{zhangRemoteSensingSeasonal2016} or by using vegetation indices to monitor the current state \citep{townsendGeneralLandsatModel2012}.
-With the recent success story of machine-learning methods in the field of remote sensing, modeling techniques such as \ac{RF} are frequently used to model relationships of possible triggers to forest health \citep{belgiuRandomForestRemote2016, laryMachineLearningGeosciences2016, michezClassificationRiparianForest2016}.
+% Link remote sensing and machine-learning
+The use of machine-learning algorithms for analyzing remote sensing data has seen a huge increase in the last decade.
+Ever since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased enormously (REF).
+At the same time, the implementation of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
+Scientists can nowadays relively easily feed large amounts of data into complex algorithms in a semi-automated way.
+These facts combined with a high temporal and spatial coverage of input data enable the possibility to analyze temporal changes of the environment on a large scale.
 
-With a robust model, predictions to large areas can be made, providing valuable information about the health condition of forest stands.
-One approach to model forest health is to extract information from spectral information of affected and unaffected trees \citep{lelongEvaluationOilPalmFungal2010}.
-Vegetation indices have shown the potential to provide valuable information to increase the predictive accuracy of forest pathogens \citep{jiangSatellitederivedVegetationIndices2014, adamczykRededgeVegetationIndices2015}.
+% link to forest health analysis and show exemplary studies
+The monitoring of forest ecosystems is one exemplary field that profits from these new possibilities.
+Specifically the monitoring of forest health got simplified with the ability of analyzing large scale data in efficient algorithms that are able to predict to unknown areas.
+For this task, usually optical data from multi-/hyperspectral satellites is used.
+Various indices are calculated from the data with the aim of carrying out temporal change detections \citep{zhang2016} or analyses about the forest health status for a specific date \citep{townsend2012}.
+% TODO: check if refs match context
+Vegetation indices have shown the potential to provide valuable information for analyzing triggers to forest health \citep{jiang2014, adamczyk2015}.
+More and more studies focus on using machine-learning algorithms such as \ac{RF} \citep{belgiu2016, lary2016, michez2016}.
+Overall, the idea is to fit a robust model which can be used to make predictions to (large) unknown areas, providing valuable information about the health condition of forest stands without the need for ground-truth data in these regions.
 
-However, the amount of possible (vegetation-)indices that can be calculated is often limited due to a low spectral resolution of freely available data from optical multispectral sensors (e.g.
-Sentinel-2).
-Also, there is currently no freely available data from hyperspectral sensors that could be used for such studies (after the decommission of the EO-1 Hyperion satellite in January 2017).
-To assess forest health on a tree scale, the spatial resolution of the data should be below 5 m as otherwise the value of a pixel may contain information from multiple trees and possibly even bare-ground information.
-With this limitation, data from sateillites with a coarser resolution (e.g. Sentinel-2) can be used for prediction tasks but not for training purposes of algorithms on a tree level.
-
-% Why are we mapping defoliation?
-
-In this study we will use hyperspectral data with a spatial resolution of one meter and 126 spectral bands to model the health status of Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain (\autoref{fig:study_area}).
-The trees in the study area suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanzaNativeRhizobacteriaBiocontrol2016, iturritxaBiocontrolFusariumCircinatum2017}.
-In-situ measurements of defoliation on a tree level (as a proxy for tree health) have been collected to serve as the response variable.
-The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
-The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{ganleyGlobalClimaticRisk2009}.
-
-We used state-of-the-art machine-learning techniques in combination with high resolution remote sensing information to model forest health on a tree level.
-Furthermore, we aimed to reduce the amount of predictors using feature selection to simplify the spatial prediction of the fitted model.
-
-\noindent Specifically the following objectives were addressed:
+% now mention why this study is important
+However, we have the impression that the amount of (vegetation-)indices used in forest health studies could be higher, possibly leading to more valuable results.
+Possible reasons include the low spectral resolution of optical multispectral sensors, the lack of freely available hyperspectral data or the lack of knowledge about existing indices.
+% TODO: Ref for high-dim
+Another more applied issue is the fact of high-dimensionality which is introduced by the calculation of hundreds/thousands of indices.
+% TODO: Ref model interpr
+Even though ML algorithms are capable of handling possibly highly-correlated input variables, model runtime increases substantially and the interpretation of results becomes more complicated.
+Also, in recent years the awareness of conducting proper tuning of model hyperparameters has increased in applied modelling fields.
+In this study we addressed these challenges by focusing on the following research questions:
 
 \begin{itemize}
-	\item Performance comparison of multiple algorithms modeling defoliation of \textit{Pinus radiata} trees using highly-correlated features
-	\item Exploration of the most important variables
-	\item Spatial prediction of \textit{Pinus radiata} defoliation
+	\item How to handle and analyze high-dimensional data including model optimization while accounting for spatial characteristics?
+	\item Does the calculation of arbitrary indices add any value to the analysis or is the band information from the sensor already sufficient?
+	\item How can interpretability of the model be maintained while at the same time using all possible input data?
 \end{itemize}
 
+\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of ensemble filters.
+To model the numeric response variable defoliation at trees model-based optimization was applied for hyperparameter tuning and a spatial block \ac{CV} was used to account for spatial autocorrelation in the data.
+
 \section{Data and study area}
+\noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain (\autoref{fig:study_area}).
+The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanza2016, iturritxa2017}.
+In-situ measurements of defoliation at trees (as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\%.
+The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
+The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{Ganley2009}.
 
@@
 \noindent The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-\textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} has the largest area size (1.44 ha).
+\textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} features the largest area size (1.44 ha).
 All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
@@
 
-\begin{figure} [t!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
-		\caption{Information about the plot locations, the area of hyperspectral coverage and the number of trees per plot.}
-		\label{fig:study_area}
-	\end{center}
-\end{figure}
+%\begin{figure} [t!]
+%\begin{center}
+%\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
+%\caption{Information about the plot locations, the area of hyperspectral coverage and the number of trees per plot.}
+%\label{fig:study_area}
+%\end{center}
+%\end{figure}
 
@@
 The first four bands are corrupted, leaving 122 bands with valid information.
-Additional metadata information is available in Table 1:
+Additional metadata information is available in \autoref{tab:hyperparameter_limits}.
 
@@
 
-\noindent For all analysis steps we used the open-source statistical programming language R \citep{R_core}.
-The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chenXGBoostScalableTree2016} (\textit{xgboost}), \textit{kernlab} \citep{kernlab} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
-We used the R package \textit{mlr} for all modeling related steps.
-It provides a standardized interface for a wide variety of statistical and machine-learning models in R simplifying essential modeling tasks such as hyperparameter tuning, model performance evaluation and parallelization \citep{bischlMlrMachineLearning2016}.
-We provide the complete code and required data as a Mendeley dataset to make this work fully reproducible [STILL TO COME].
-
 \subsection{Derivation of indices}
-
-\noindent To use the full information from the hyperspectral data, we calculated all possible vegetation indices that are available in the \textit{hsdar} package (90 in total) and all possible \ac{NRI} combinations.
-We were interested if NRIs of arbitrary band combinations will have a substantial effect on the predictive power of the fitted model.
-The NRIs were calculated using the following formula:
+% TODO: Mention different feature sets
+\noindent To use the full information from the hyperspectral data, we calculated all possible vegetation indices supported by the \textit{hsdar} package (90 in total) as well as all possible \ac{NRI} combinations.
+NRIs were calculated using the following formula:
 
@@
 
+\subsection{Feature selection methods}
+
+% TODO: List filter methods
+% TODO: Explain the idea of ensemble filters
+
 \subsection{Benchmarking of algorithms}
@@
 \noindent Three algorithms (\textit{xgboost}, \textit{\ac{SVM}} and \textit{\ac{RR}}) were benchmarked on their predictive performance.
-Besides the well-known \ac{SVM} algorithm \citep{Vapnik1998}, we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chenXGBoostScalableTree2016}.
-We added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerlRidgeRegressionBiased1970}.
-One of the most popular machine-learning algorithm, Random Forest \citep{Breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
-These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{Probst2018b}.
+Besides the well-known \ac{SVM} algorithm \citep{vapnik1998}, we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chen2016}.
+We added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerl1970}.
+One of the most popular machine-learning algorithm, Random Forest \citep{breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
+These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{probst2018}.
 After the selection of the best model, we checked if the winning algorithm can achieve a similar performance when using only the most important variables compared to using all variables.
@@
 
-\begin{figure} [t!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=1\textwidth] {spcv_folds_demon_plots.pdf}}
-		\caption{Fold 1 of the spatial partitioning using k-means clustering for \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
-		\label{fig:spcv}
-	\end{center}
-\end{figure}
+%\begin{figure} [t!]
+%\begin{center}
+%\makebox[\textwidth]{\includegraphics[width=1\textwidth] {spcv_folds_demon_plots.pdf}}
+%\caption{Fold 1 of the spatial partitioning using k-means clustering for \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
+%\label{fig:spcv}
+%\end{center}
+%\end{figure}
 
@@
 
-\noindent To tune the hyperparameters of the algorithms, we used \ac{SMBO} via the R package \textit{mlrMBO} \citep{mlrMBO}.
+\noindent To tune the hyperparameters of the algorithms, we used \ac{SMBO} via the R package \textit{mlrMBO} \citep{mlrmbo}.
 This Bayesian approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
@@
 Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
-This strategy continues until a termination criterion, defined by the user, is reached \citep{Hutter2011, Jones1998}.
+This strategy continues until a termination criterion, defined by the user, is reached \citep{hutter2011, jones1998}.
 
 In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting in a total budget of 50 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs, such as random search or grid search \citep{Bergstra2012}.
+The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs, such as random search or grid search \citep{bergstra2012}.
 
-\subsection{Variable importance}
-
-\noindent To find indices that contributed most to model performance, we used the internal variable importance measure of the \textit{xgboost} algorithm.
-This score is calculated by taking the contribution of each feature for each tree in the fitted model.
-The higher the score of a variable, the more important it is for the fitted model when making predictions \citep{chenXGBoostScalableTree2016}.
-The variable importance measure is automatically computed during model fit.
-In contrast to other approaches such as permutation-based ones, the \textit{xgboost} score is composed out of three parts that contribute to the overall importance \citep{chenXGBoostScalableTree2016}:
-
-\begin{itemize}
-	\item Gain: The relative contribution of the feature to the model
-	\item Cover metric: How often a feature was selected to be the deciding feature in a tree for a specific observation
-	\item Frequency How often a feature occurs in all trees of the model
-\end{itemize}
-
-\noindent The \textit{Gain} features is the most important one among the three.
-All measures sum up to one \citep{chenXGBoostScalableTree2016}.
-
-% partial dependence plots??
+\noindent The complete study was done using the open-source statistical programming language R \citep{rcoreteam2018}.
+The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chen2016} (\textit{xgboost}), \textit{e1071} \citep{e1071} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
+The filter methods of the following packages were used: \textit{praznik}, \textit{fselectorrcpp}.
+The R package \textit{mlr} \citep{mlr} was used for all modeling related steps and \textit{drake} for setting up the workflow of the project.
+This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
 
@@
 
-% defoliation boxplots
-\begin{figure} [t!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {boxplot_defol.pdf}}
-		\caption{Descriptive statistics of the response variable \textit{defoliation}.}
-		\label{fig:defol_boxplots}
-	\end{center}
-\end{figure}
+%% defoliation boxplots
+%\begin{figure} [t!]
+%\begin{center}
+%\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {boxplot_defol.pdf}}
+%\caption{Descriptive statistics of the response variable \textit{defoliation}.}
+%\label{fig:defol_boxplots}
+%\end{center}
+%\end{figure}
 
-\noindent \textit{Oiartzun} shows the highest defoliation ($\bar{x} = 69.22 \%$) among the plots while \textit{Laukiz 2} is the healthiest ($\bar{x} = 13.54 \%$) (\autoref{fig:defol_boxplots}).
-All plots besides \textit{Luiando} show an evenly distributed level of defoliation across the entire plot.
+%\noindent \textit{Oiartzun} shows the highest defoliation ($\bar{x} = 69.22 \%$) among the plots while \textit{Laukiz 2} is the healthiest ($\bar{x} = 13.54 \%$) (\autoref{fig:defol_boxplots}).
+%All plots besides \textit{Luiando} show an evenly distributed level of defoliation across the entire plot.
 
-\noindent The high degree of defoliation for \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
-Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to \textit{Laukiz 1} and \textit{Laukiz 2}.
-\textit{Oiartzun} is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
+%\noindent The high degree of defoliation for \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
+%Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to \textit{Laukiz 1} and \textit{Laukiz 2}.
+%\textit{Oiartzun} is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
 
-\noindent \textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots have a higher density (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
+%\noindent \textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots have a higher density (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
 
@@
 
-% Performance estimates
-\begin{table}[t!]
-\centering
-\caption[t]{Spatial block \ac{CV} performances of \textit{RR}, \textit{SVM} and \textit{xgboost} using \ac{RMSE} as the error measure.
-	Mean and standard deviation are shown.}
-\begingroup\footnotesize
-\begin{tabular}{llll}
-	RR            & SVM           & xgboost       & xgboost (7 variables) \\
-	\hline
-	59.10 (22.71) & 36.23 (15.73) & 33.26 (16.61) & 29.59   (16.09)       \\
-	\bottomrule
-\end{tabular}
-\endgroup
-\label{tab:model_comparison}
-\end{table}
-
-\begin{table}[b!]
-\centering
-\caption[t]{Predictive performance of \textit{xgboost} using all observations and all variables (All Observations/all variables), all observations and the seven most important variables only (All Observations/7 variables) and observations from specific plots only (Plot level observation/all variables) with \ac{RMSE} as the error measure.
-The performance estimates for "All Observations" correspond to the fold for which the respective plot was serving as the test set (block CV).
-Column "Plot level observations", shows the mean RMSE estimates at the repetition level of a five-fold five-time repeated spatial CV, scored by using data of the respective plot only.}
-\begingroup\footnotesize
-\begin{tabular}{llll}
-	\\
-	Plot/Data & \specialcell{All Observations/ \\{all variables (Block CV)}} & \specialcell{All Observations/         \\ {7 variables (Block CV)}} & \specialcell{Plot level observations/\\{all variables (SpCV)}} \\
-	\hline
-	Laukiz 1  & 22.03                                                                              & 21.47                          & 19.18 \\
-	Laukiz 2  & 51.75                                                                              & 49.94                          & 17.24 \\
-	Luiando   & 13.20                                                                              & 15.37                          & 8.30  \\
-	Oiartzun  & 32.97                                                                              & 17.62                          & 14.40 \\
-	\bottomrule
-\end{tabular}
-\endgroup
-\label{tab:supermodel_performance}
-\end{table}
-
 \subsubsection{Algorithm benchmarking}
-
-\noindent The \textit{xgboost} algorithm showed the lowest error (33.26 RMSE) when benchmarking the learners on the complete dataset of all plots (\autoref{tab:model_comparison}).
-While the \textit{SVM} performance was only slightly worse (36.23 RMSE), \textit{RR} showed a substantially worse performance than \textit{xgboost} (59.10 RMSE).
-
-\subsubsection{Single models vs. super model}
-
-\noindent Comparing the mean predictive performance of models fitted at the plot level against the performance of the model that was fitted using all data (super model), the plot-level models showed a better performance in all cases (\autoref{tab:supermodel_performance}).
-The highest difference between both datasets occurred for plot \textit{Laukiz 2} with a difference of 34.51 RMSE.
-
-Using only the seven most important variables (\autoref{fig:var-imp}) for the super model showed small increases in performance for \textit{Laukiz 1} and \textit{Laukiz 2}, a small decrease for \textit{Luiando} and almost a reduction of 50\% of the error for \textit{Oiartzun} (32.97 vs.
-17.62 RMSE) (\autoref{tab:supermodel_performance}).
-
-\subsubsection{RMSE vs. plot characteristics}
-
-\noindent An increase of the error rate was observed with an increase of descriptive plot measures such as mean point density and the coefficient of variation (based on the response variable \textit{defoliation}) (\autoref{fig:plot-characteristics}).
-
-\begin{figure} [t!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {plot-characteristics.pdf}}
-		\caption{RMSE vs. mean point density and coefficient of variation (defoliation).}
-		\label{fig:plot-characteristics}
-	\end{center}
-\end{figure}
-
-\subsection{Variable importance}
-
-\noindent The seven most important features of the super model in this study were vegetation indices with the \textit{EVI} \citep{hueteComparisonVegetationIndices1997} being the most important one (\autoref{fig:var-imp}).
-
-\begin{equation}
-	EVI = 2.5*\frac{R_{800}-R_{670}}{R_{800}-(6*R_{670})-(7.5*R_{475})+1)}
-\end{equation}
-
-\noindent where $R$ = Reflectance at the respective wavelength.
-
-\bigbreak
-
-\noindent Vegetation index \textit{GDVI} appears three times among the first seven most important features (\autoref{fig:var-imp}) with different \texttt{n} values.
-This is because it was computed four times, with \texttt{n} ranging from 1 - 4 \citep{wuEstimatingChlorophyllContent2008}:
-
-\begin{equation}
-	GDVI = \frac{R_{800}^n-R_{680}^n}{R_{800}^n+R_{680}^n}
-\end{equation}
-
-\bigbreak
-
-\noindent The seven most important features (\textit{EVI}, \textit{GDVI\_4}, \textit{D1}, \textit{GDVI\_3}, \textit{GDVI\_2}, \textit{mNDVI} and \textit{mSR}) showed a substantial difference in the importance score compared to all following variables (\autoref{fig:var-imp}).
-
-\begin{figure} [b!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width = 1\textwidth] {var-imp-xgboost.pdf}}
-		\caption{The 30 most important variables as estimated by the internal variable importance measure of the \textit{xgboost} algorithm. The higher the score, the more important the feature.
-			"bf2" notes that a buffer of 2 meter was used to extract the variable information to the tree observation.
-"NRI" means that a normalized ratio index with the subsequent bands was calculated.
-Features without "NRI" prefix are vegetation indices, e.g.
-"bf2\_EVI".}.
-%An overview of the vegetation indices can be found \href{https://rdrr.io/cran/hsdar/man/vegindex.html}{here}.}
-		\label{fig:var-imp}
-	\end{center}
-\end{figure}
-
-\begin{table}[t!]
-\centering
-\caption[t]{Formulas of the five most important vegetation indices of the super model. $R$ = Reflectance at wavelength, $D$ = First derivation of reflectance value at wavelength.}
-\begingroup\footnotesize
-\begin{tabular}{llll}
-	\\
-	Acronym & Name                      & Formula                                                            & Reference                                                 \\
-	\hline
-	EVI     & Enhanced vegetation index & $2.5*\frac{R_{800}-R_{670}}{R_{800}-(6*R_{670})-(7.5*R_{475})+1)}$ & \cite{hueteComparisonVegetationIndices1997}               \\
-	GDVI    & Generalized DVI*          & $\frac{R_{800}^n-R_{680}^n}{R_{800}^n+R_{680}^n}$                  & \cite{wuEstimatingChlorophyllContent2008}                 \\
-	D1      & Derivative Index          & $\frac{D_{730}}{D_{706}}$                                          & \cite{zarco-tejadaSteadystateChlorophyllFluorescence2003} \\
-	mNDVI   & Normalized DVI*           & $  \frac{R_{800}-R_{680}}{(R_{800}+R_{680}-2*R_{445}}$             & \cite{simsRelationshipsLeafPigment2002}                   \\
-	mSR     & Simple Ratio Index        & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                          & \cite{simsRelationshipsLeafPigment2002}                   \\
-	\bottomrule
-	\multicolumn{2}{l}{\small{* Difference Vegetation Index}}
-\end{tabular}
-\endgroup
-\label{tab:most_imp_vars}
-\end{table}
-
-\noindent The best NRI scored rank eight (band 112 and band 62).
-All further places up to rank 30 were occupied by NRIs.
-
-\subsection{Spatial prediction}
-
-\begin{figure} [b!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width = 1\textwidth] {predictions_demon_plots.pdf}}
-		\caption{Spatially predicted defoliation (in \%) from \textit{xgboost} of \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
-		\label{fig:pred_demon_plots}
-	\end{center}
-\end{figure}
-
-\begin{figure} [t!]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width = 0.6\textwidth] {predictions_demon_plots_hists.pdf}}
-		\caption{Histograms of predicted defoliation (in \%) from \textit{xgboost} of \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
-		\label{fig:pred_demon_plots_hists}
-	\end{center}
-\end{figure}
-
-\noindent The plots with a higher mean defoliation (\textit{Luiando} and \textit{Oiartzun}) showed a good visual separability compared to the healthier plots \textit{Laukiz 1} and \textit{Laukiz 2} (\autoref{fig:pred_demon_plots}).
-All predicted values ranged between 30 \% and 80 \% defoliation with \textit{Luiando} showing the smallest variance (\autoref{fig:pred_demon_plots_hists}).
-
-The high error of the super model for \textit{Laukiz 2} (49.94 RMSE) is also visible in the respective histogram as most predictions range between a defoliation of 40 \% - 50 \% (\autoref{fig:pred_demon_plots_hists}) while in fact most trees of \textit{Laukiz 2} show an actual defoliation of around 20 \% (\autoref{fig:defol_boxplots}).
-
-For the less defoliated plots \textit{Laukiz 1} and \textit{Laukiz 2} a subtle level of separation between trees and bare ground is visible (\autoref{fig:pred_demon_plots}).
-This is not the case for the other two higher defoliated plots for which bare ground and defoliated predictions are mainly in the same value range (around 60 \% - 80 \%).
 
@@
 
-\noindent The relatively large difference in performance between \textit{RR} (59.10 RMSE) and the machine-learning models (36.23 and 33.26 RMSE) is remarkable.
-\textit{RR} has shown promising results in other studies when many highly-correlated predictors were involved \citep{hernandezUsingRidgeRegression2015, imaniRidgeRegressionbasedFeature2015}.
-However, in this study, \textit{RR} was not able to achieve a sufficient performance score compared to \textit{SVM} and \textit{xgboost} even though its hyperparameter $\lambda$ was properly tuned using SMBO (\autoref{meth:smbo}).
-
-While \textit{xgboost} showed a slightly better performance than \textit{SVM}, the latter has the advantage of only having two hyperparameters that need to be tuned.
-This results in a shorter runtime.
-Nevertheless, \textit{xgboost} showed the best performance and was subsequently selected to fit the models on the plot level and for the spatial prediction.
-
-An important point that needs to be considered when interpreting the performance results is that we only related defoliation to indices derived from remote sensing data.
-Possible other local variables that could help in predicting defoliation were not considered.
-One example here is tree age: The older a tree the more vulnerable it may be to pathogen infections causing defoliation.
-However, such predictors would not be available for a spatial prediction scenario and one of the main goals of this study is to relate defoliation to variables that are available on a larger scale (e.g.
-remote sensing indices).
-
-\subsubsection{Single models vs. super model}
-
-\noindent It is expected that models that were trained using observations from the respective plot only only achieve a better performance than the super model which was trained on observations from multiple plots (\autoref{tab:supermodel_performance}).
-The low performance on \textit{Laukiz 2} for the super model is most likely due to the difference of this plot to all others: The fitted model on \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} is not capable of reaching a good performance when \textit{Laukiz 2} is the evaluation dataset.
-This is not surprising as \textit{Laukiz 2} shows substantially different plot characteristics compared to all others plots in terms of the distribution of the response variable \textit{defoliation} (\autoref{fig:defol_boxplots}) and the mean point density of trees (\autoref{fig:plot-characteristics}).
-
-The low error for Luiando (8.30 RMSE) for the plot model validates the approach of relating defoliation to vegetation indices and NRIs.
-The overall error of the super model (33.26 RSME) is expected to decrease if more plots would be available for training.
-However, even if the fitted model would include at least one instance of every plot showing unique characteristics (e.g.
-here \textit{Laukiz 2} is substantially different to the others), the overall error would not become smaller than the error achieved when using data from the respective plot only.
-
-An interesting find is that the model with only seven variables shows a better overall performance than the model with all 7471 variables (\autoref{tab:supermodel_performance}).
-This leads to the conclusion that adding as many variables as possible to a model will not necessarily improve its performance but instead add noise to the model.
-Too much information can be problematic for a model as it will have a hard time distinguishing between noise and important information in the variables \citep{liFeatureSelectionData2017, guyonIntroductionVariableFeature2003}.
-However, to find the most important variables in the first place and to check for the performance difference, a model with all variables needs to fitted first.
-Using a model with only a few predictors does not only simplify prediction tasks but also reduces runtime for hyperparameter tuning and performance estimation \citep{liuComputationalMethodsFeature2007}.
-
 \subsection{Variable importance}
-
-\noindent There are some downsides using the internal variable importance approach of \textit{xgboost}: Due to the contribution of three different parts to the overall importance score it is complicated to understand why a specific feature was selected.
-Furthermore the importance calculation approach is only valid for this algorithm and cannot be compared to others.
-Nevertheless, as we only relied on the variable importance for this specific algorithm, using the internal \textit{xgboost} approach was sufficient for this work.
-
-It is expected that vegetation indices are most important for the model as these are most sensitive to changes in vegetation health \citep{croftApplicabilityEmpiricalVegetation2014}.
-Even though we are not directly looking at vegetation health but using the level of defoliation as a proxy for tree health, vegetation indices were most important for the fitted model of this study (\autoref{fig:var-imp}).
-
-\noindent Vegetation indices can help assessing defoliation in two ways:
-
-\begin{itemize}
-	\item Trees that show a high level of defoliation do also reflect their bad health status through the remaining foliation.
-	\item Defoliated trees have more influence of bare ground information in their pixel values and will therefore be classified as defoliated by the model.
-\end{itemize}
-
-\noindent Even though no NRI made it among the most important variables in this study (\autoref{fig:var-imp}) (stating that the first seven of this study are the most important ones), it is notable that all ranks from 8 - 30 are occupied by NRI (\autoref{fig:var-imp}).
-However, their relative importance was very small compared to the first seven ranks.
-
-Restricting the important indices of this study to the first seven ranks can be seen critical as we only based the selection on a visual inspection of the variable importance results (\autoref{fig:var-imp}).
-The decision to make a cut between rank seven and eight was based on a combination of two facts:
-
-\begin{itemize}
-	\item Using only vegetation indices is easier for large scale predictions using satellites like Sentinel-2 (most NRIs cannot be used with it because the spectral bands do not exist).
-	\item The drop in the importance score of the variable importance results (\autoref{fig:var-imp}).
-\end{itemize}
-
-\noindent However, based on these two points, we could also have made the cut between rank five and 6 but including the two vegetation indices at rank six and seven will eventually improve the model and does not increase runtime.
 
@@
 
-\noindent The spatial predictions showed that the model tries to avoid making extreme predictions since most values ranged between a defoliation of 30 \% and 80 \%.
-This behavior is mainly triggered by overfitting on the observations of the \textit{Laukiz 1}, \textit{Luiando} and \textit{Oiartzun} training plots.
-The overfit then causes a high prediction error for \textit{Laukiz 2} for which a lot of defoliation values actually range between 0 \% and 20 \%.
-The fitted model would need more training samples of plots with an unusual defoliation distribution to become more robust and achieve a better overall performance.
-
 \subsection{Comparison to other studies}
 
-\noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengarajanModelingForestDefoliation2016}.
+\noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengara2016}.
 This corresponds with the finding of this study that the most important variables are located in the NIR region.
 
-\cite{goodbodyDigitalAerialPhotogrammetry2018} used NDVI and structural measures in as inputs for a partial least squares analysis to model defoliation caused by the spruce budworm.
+\cite{goodbody2018} used NDVI and structural measures in as inputs for a partial least squares analysis to model defoliation caused by the spruce budworm.
 Results showed that metrics from spectral features were most important.
@@
 
-\cite{townsendGeneralLandsatModel2012} used Landsat data to model defoliation caused by insect herbivores.
+\cite{townsend2012} used Landsat data to model defoliation caused by insect herbivores.
 They found that the \ac{NDII} ($\frac{Band 4 - Band 5}{Band 4 + Band 5}$) and the moisture stress index ($\frac{Band 5}{Band 4}$) gave better results than using NDVI.
@@
 
-MODIS data was used by \cite{debeursEstimatingEffectGypsy2008} to model defoliation caused by the gypsy moth using vegetation indices such as NDVI, EVI, NDWI and NDII.
+% TODO: Find missing reference
+%MODIS data was used by \cite{debeursEstimatingEffectGypsy2008} to model defoliation caused by the gypsy moth using vegetation indices such as NDVI, EVI, NDWI and NDII.
 
-All of these examples validate the approach of using vegetation indices to model defoliation.
-Even though the spatial resolution of the data in these studies varied between hundreds of meters (MODIS) \cite{debeursEstimatingEffectGypsy2008} and centimeters \cite{goodbodyDigitalAerialPhotogrammetry2018}, high resolution data is preferred to fit accurate models.
-Also, the importance of certain indices (e.g. NDVI, EVI) will vary based on the data and resolution.
-The finding of this work that the vegetation indices GDVI and EVI are most important for the fitted model could not be verified by other studies.
-However, the presented ones did often only use a small subset of the vegetation indices that were used in this study.
+%All of these examples validate the approach of using vegetation indices to model defoliation.
+%Even though the spatial resolution of the data in these studies varied between hundreds of meters (MODIS) \cite{debeursEstimatingEffectGypsy2008} and centimeters \cite{goodbody2018}, high resolution data is preferred to fit accurate models.
+%Also, the importance of certain indices (e.g. NDVI, EVI) will vary based on the data and resolution.
+%The finding of this work that the vegetation indices GDVI and EVI are most important for the fitted model could not be verified by other studies.
+%However, the presented ones did often only use a small subset of the vegetation indices that were used in this study.
 
-We could not find a study that used recent machine-learning techniques in combination with a high amount of variables to model defoliation.
-This fact highlights the importance of this work and will hopefully encourage scientists to use machine-learning techniques, feature selection methods and a wider selection of vegetation indices when assessing defoliation in the future.
+%We could not find a study that used recent machine-learning techniques in combination with a high amount of variables to model defoliation.
+%This fact highlights the importance of this work and will hopefully encourage scientists to use machine-learning techniques, feature selection methods and a wider selection of vegetation indices when assessing defoliation in the future.
 
@@
 % spectral signatures
-\begin{figure} [H]
-	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral_signatures.pdf}}
-		\caption{Spectral signatures (mean and standard deviation) of each plot.}
-		\label{fig:spectral_signatures}
-	\end{center}
-\end{figure}
+%\begin{figure} [H]
+%\begin{center}
+%\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral_signatures.pdf}}
+%\caption{Spectral signatures (mean and standard deviation) of each plot.}
+%\label{fig:spectral_signatures}
+%\end{center}
+%\end{figure}
 

### 2019-06-03 analysis/paper/Schratz_hyperspectral.tex
paper: section: Dealing with high-dimensionality

@@
 In-situ measurements of defoliation at trees (as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\%.
-The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{Iturritxa2014}.
-The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{Ganley2009}.
+The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{iturritxa2014}.
+The dieback of these trees, which are mainly used as timber, causes high economic damages \citep{ganley2009}.
 
@@
 % TODO: Mention different feature sets
-\noindent To use the full information from the hyperspectral data, we calculated all possible vegetation indices supported by the \textit{hsdar} package (90 in total) as well as all possible \ac{NRI} combinations.
-NRIs were calculated using the following formula:
+\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the \textit{hsdar} package (90 in total) as well as all possible \ac{NRI} combinations were calculated.
+The following formula was used for the NRI calculation:
 
@@
 
-\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), we used a buffer of two meters around the centroid of the respective tree.
-The mean value of all pixels touched by the buffer was assigned as the final value of each index.
+\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), a buffer of two meters around the centroid of the respective tree was used.
+The mean of all pixels touched by this buffer was calculated and assigned to the observations.
 In total, $\frac{125*126}{2} = 7875$ NRIs were calculated .
-Due to four corrupted bands and numerical problems for some band combinations, some indices returned \texttt{NA} for specific observations.
-We removed all indices from the dataset that showed one or more \texttt{NA} values (across all plots) since we valued a single observation more than having an additional NRI as a predictor variable.
-In total, 7471 indices had no \texttt{NA} values and were subsequently used as predictors.
+Due to four corrupted bands and numerical calculation errors for some band combinations, a total of 7471 indices was available for each observation.
+If an index returned \texttt{NA} for one plot, the index was removed from all plots.
+This decision was based on the fact that due to the mass of available indices and the low number of observations, the latter was valued higher.
+
+\subsection{Benchmarking design}
+
+The benchmarking matrix of this study consists of the following algorithms:
+
+\begin{itemize}
+	\item  Extreme Gradient Boosting (\ac{XGBOOST})
+	\item  Random Forest (\ac{RF})
+	\item  Penalized Regression (both L1 and L2)
+	\item  Support Vector Machine (\ac{SVM})
+\end{itemize}
+
+\ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
+Extreme Gradient Boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
+Penalized Regression is a statistical modeling technique that is capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model.
+The penalties are also known as LASSO (L1) and RIDGE (L2).
+The former does not allow the removal of variables from the model (penalization to zero) while the latter does.
+Both penalties can also be combined.
+The combined approach is called "elastic net".
+
+Three feature sets were used in this study:
+
+\begin{itemize}
+	\item The raw hyperspectal band information (HR, $n = 122$)
+	\item Vegetation Indices (\ac{VI}, $n = 90$)
+	\item Normalized Ratio Indices (\ac{NRI}, $n = 7471$)
+\end{itemize}
+
+An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
+Maximum Bayesian Optimization \citep{mlrmbo} was used for parameter optimization.
+This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
+The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
+Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
+This strategy continues until a termination criterion, defined by the user, is reached \citep{hutter2011, jones1998}.
+An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
+The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum
+This applies when being compared to methods that do not use information from previous runs, such as random search or grid search \citep{bergstra2012}.
+
+A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible.
+Each tree plot served as one fold within the resampling setup, resulting in four folds total.
+For the inner level, $p - 1$ folds were used (with $p$ being the number of plots).
+
+\subsection{Dealing with high-dimensionality}
+% TODO talk about filter vs wrapper vs PCA
+The case of high-dimensionality in machine-learning is a research field on its own and numerous approaches were developed in recent years to optimize this case.
+% TODO talk more about the general problem and solution, maybe 1-2 sentences
+
+So called "Filter methods" are one approach.
+The basic concept relies on the ranking of all features of the dataset using certain heuristics.
+Various filter methods exist with different capabilities what feature types are supported (numeric, nominal).
+After the coveriates have been ranked by a feature method, a decision needs to be made which features to keep and which to discard.
+This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
+Essentially, the goal is to find the best threshold at which the model achieves a good performance while requiring the least amount of variables.
+
+Another approach is to use algorithms that are normally used for hyperparameter optimization such as \textit{Random Search} or a \textit{Genetic Algorithm}.
+Here, first a subset of features is chosen based on the selected algorithm.
+Now the model is fitted on the data and the performance is evaluated.
+The hyperparameter tuning can only start after the feature-selection optimization has ended.
+This method is very expensive since two optimization stages need to be run.
+This approach is also referred to as "wrapper method" and was not considered in this work due to the expensive runtime.
+
+A third commonly used method is the "Principal Component Analysis".
+Here, the main components of the feature space are extracted and combined.
+Usually the first two components are taken since these contain the major information of the coveriates.
+By using the automatically estimated explained variance of the main components, the model can rely on a few features containing the condensed information of most variables of the data.
+This enables the fitting of very cheap models with a minimal loss of predictor information.
+
+% TODO mention that filter run before hyperpar tuning and are cached
+% TODO mention that the % value of the remaining features is tuned as a normal hyperpar
 
@@
 
-\noindent Three algorithms (\textit{xgboost}, \textit{\ac{SVM}} and \textit{\ac{RR}}) were benchmarked on their predictive performance.
-Besides the well-known \ac{SVM} algorithm \citep{vapnik1998}, we also used \textit{xgboost} which is ensemble method relying on the idea of tree boosting that gained a lot of attention in recent years \citep{chen2016}.
 We added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerl1970}.
-One of the most popular machine-learning algorithm, Random Forest \citep{breiman2001}, was not considered for this study: Due to the high number of variables, model fitting times in the range hours for a single model fit were not practicable for this work.
 These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{probst2018}.
-After the selection of the best model, we checked if the winning algorithm can achieve a similar performance when using only the most important variables compared to using all variables.
 A successful feature selection simplifies the spatial prediction task because the prediction dataset needs to consist of less variables.
@@
 
-\bibliography{Biblio_hyperspectral}
+\bibliography{Biblio}
 

### 2019-06-05 analysis/paper/Schratz_hyperspectral.tex
rewrite methodology

@@
 
+\subsubsection{Feature sets}
+
 Three feature sets were used in this study:
@@
 \end{itemize}
+
+\subsubsection{Hyperparameter Optimization}
 
@@
 
+\subsubsection{Spatial resampling}
+
 A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible.
@@
 \subsection{Dealing with high-dimensionality}
-% TODO talk about filter vs wrapper vs PCA
-The case of high-dimensionality in machine-learning is a research field on its own and numerous approaches were developed in recent years to optimize this case.
-% TODO talk more about the general problem and solution, maybe 1-2 sentences
+% Define "high-dimensionality"
+The usage of the term "high-dimensionality" in the machine-learning field practically means to deal with dataset that has a high amount of covariates.
+There is no clear threshold at which number the term can be applied.
+There has been an informal agreement to speak of "high-dimensonialty" if the number of variables exceeds the number of observations. REF
 
+The case of a high-dimensional dataset comes with several challenges for both model fitting and evaluation.
+There is an increase in time for fitting a model to the data.
+Possible noise maybe introduced into the model by highly-correlated variables.
+The interpretation and prediction of models containing hundreds of covariates becomes more complicated, up to to the point at which interpretation or prediction is not possible anymore.
+Especially in the last decade the amount of data increased a lot and also the environmental modeling field profited from this as well.
+Numerous approaches were developed to deal with this situation and the field of "feature selection" can be safely seen as a standalone research field.
+
+\subsubsection{Filter methods}
+
+% Filter methods
 So called "Filter methods" are one approach.
-The basic concept relies on the ranking of all features of the dataset using certain heuristics.
-Various filter methods exist with different capabilities what feature types are supported (numeric, nominal).
-After the coveriates have been ranked by a feature method, a decision needs to be made which features to keep and which to discard.
+The basic concept relies on the ranking of all features of the dataset using heuristics of the respective algorithm.
+Some are tailored towards specific types of variables (numeric or nominal data).
+After the covariates have been ranked, a decision needs to be made what features to keep and which to discard.
 This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
-Essentially, the goal is to find the best threshold at which the model achieves a good performance while requiring the least amount of variables.
+Essentially, the number of covariates in the model is treated as a hyperparameter of the respective model.
+The goal is to find the threshold (usually in \% of covariates) at which the model achieves a good performance while requiring the least amount of variables.
+In well-implemented software solutions the calculations of the ranking is only done once and then cached.
+In subsequent tuning iterations the results are taken from the cache instead of being calculated again.
 
-Another approach is to use algorithms that are normally used for hyperparameter optimization such as \textit{Random Search} or a \textit{Genetic Algorithm}.
-Here, first a subset of features is chosen based on the selected algorithm.
+% Ensembe filter methods
+Besides the idea of using single filter methods to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models.
+This aligns with the recent rise of the "ensemble" approach which was first used to combine the predictions of multiple models.
+In this work we also used one ensemble filter method based on the "Borda" aggregation.
+Here, the ranks of all single filters are summed up leading to a final ranking which consists of the ranks achieved by all filter algorithms.
+
+Most filter methods are correlation or entropy based.
+A list of all filter methods used in this work along with their respective formulas can be found in Appendix X.
+% TODO: Link appendix
+% TODO: List filter methods
+
+\subsubsection{Other approaches}
+
+% Wrapper approach
+Another approach to reduce the number of variables is to apply algorithms that are normally used for hyperparameter optimization such as \textit{Random Search} or a \textit{Genetic Algorithm}.
+Here, first a (random) subset of features is chosen based on the selected algorithm.
+In comparison to "filters", no ranking is applied here.
 Now the model is fitted on the data and the performance is evaluated.
-The hyperparameter tuning can only start after the feature-selection optimization has ended.
-This method is very expensive since two optimization stages need to be run.
-This approach is also referred to as "wrapper method" and was not considered in this work due to the expensive runtime.
+A major downside of this approach is that hyperparameter tuning can only start after the feature-selection optimization has ended.
+Hence, this method is very expensive because two optimization stages need to be run.
+In science this approach is also referred to as "wrapper method".
+It was not considered in this work due to the expensive runtime.
 
+% PCA
 A third commonly used method is the "Principal Component Analysis".
 Here, the main components of the feature space are extracted and combined.
-Usually the first two components are taken since these contain the major information of the coveriates.
-By using the automatically estimated explained variance of the main components, the model can rely on a few features containing the condensed information of most variables of the data.
-This enables the fitting of very cheap models with a minimal loss of predictor information.
-
-% TODO mention that filter run before hyperpar tuning and are cached
-% TODO mention that the % value of the remaining features is tuned as a normal hyperpar
-
-\subsection{Feature selection methods}
-
-% TODO: List filter methods
-% TODO: Explain the idea of ensemble filters
-
-\subsection{Benchmarking of algorithms}
-
-We added penalized L2 (Ridge) regression to the portfolio due to its ability to handle highly correlated covariates \citep{hoerl1970}.
-These high fitting times are caused by hyperparameter \texttt{mtry} which scales with the number of variables \citep{probst2018}.
-A successful feature selection simplifies the spatial prediction task because the prediction dataset needs to consist of less variables.
-Furthermore, model complexity and fitting times are reduced.
-
-\subsubsection{Performance estimation}
-
-\noindent The algorithms were benchmarked in two ways:
-
-\begin{enumerate}
-	\item Using spatial block \ac{CV} on the plot level with each plot serving as the test set once. Four performance estimates were retrieved, one for each fold.
-	\item Using five-fold five-time repeated spatial \ac{CV} within each plot based on the k-means clustering approach of \cite{sperrorest} (\autoref{fig:spcv}).
-\end{enumerate}
-
-\noindent As we used the best algorithm of 1) for the spatial prediction, we first conducted the model selection on this setup and only applied the winning algorithm on 2).
-
-%\begin{figure} [t!]
-%\begin{center}
-%\makebox[\textwidth]{\includegraphics[width=1\textwidth] {spcv_folds_demon_plots.pdf}}
-%\caption{Fold 1 of the spatial partitioning using k-means clustering for \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
-%\label{fig:spcv}
-%\end{center}
-%\end{figure}
-
-\subsubsection{Hyperparameter tuning}
-\label{meth:smbo}
-
-\noindent To tune the hyperparameters of the algorithms, we used \ac{SMBO} via the R package \textit{mlrMBO} \citep{mlrmbo}.
-This Bayesian approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
-After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
-The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
-Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
-This strategy continues until a termination criterion, defined by the user, is reached \citep{hutter2011, jones1998}.
-
-In this work we used an initial design of 30 randomly composed hyperparameter settings and a termination criterion of 20 iterations, resulting in a total budget of 50 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is that it substantially reduces the tuning budget which is needed to find a setting close to the global minimum compared to methods that do not use information from previous runs, such as random search or grid search \citep{bergstra2012}.
+Usually the first two components are taken since these contain the major information of the covariates.
+By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the condensed information of most variables of the data.
+This enables a very cheap model fitting with a minimal loss of predictor information.
+However, the resulting features used in the model do not relate anymore to the raw input variables since they are a combination of multiple covariates.
+Interpretation is not possible anymore.
 
@@
 The filter methods of the following packages were used: \textit{praznik}, \textit{fselectorrcpp}.
-The R package \textit{mlr} \citep{mlr} was used for all modeling related steps and \textit{drake} for setting up the workflow of the project.
+The R package \textit{mlr} \citep{mlr} was used for all modeling related steps.
+\textit{drake} was used for structuring the work and ensuring reproducbility.
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).

### 2019-06-09 analysis/paper/Schratz_hyperspectral.tex
restructure ToC
add refs to introduction

@@
 % Link remote sensing and machine-learning
-The use of machine-learning algorithms for analyzing remote sensing data has seen a huge increase in the last decade.
-Ever since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased enormously (REF).
-At the same time, the implementation of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
-Scientists can nowadays relively easily feed large amounts of data into complex algorithms in a semi-automated way.
-These facts combined with a high temporal and spatial coverage of input data enable the possibility to analyze temporal changes of the environment on a large scale.
+The use of machine-learning algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
+Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
+At the same time, the implementation and usability of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
+Scientists can nowadays relatively easily feed large amounts of information into complex models in a semi-automated way.
+These points enable the possibility to analyze (temporal) changes of the environment on a large scale \citep{ma2015}.
 
@@
 The monitoring of forest ecosystems is one exemplary field that profits from these new possibilities.
-Specifically the monitoring of forest health got simplified with the ability of analyzing large scale data in efficient algorithms that are able to predict to unknown areas.
+Specifically the monitoring of forest health profited from the ability of analyzing large scale data with efficient algorithms able to predict to unknown areas \citep{hawrylo2018, mascaro2014}.
 For this task, usually optical data from multi-/hyperspectral satellites is used.
-Various indices are calculated from the data with the aim of carrying out temporal change detections \citep{zhang2016} or analyses about the forest health status for a specific date \citep{townsend2012}.
-% TODO: check if refs match context
+Various indices are calculated from the data with the aim of carrying out temporal change detections \citep{zhang2016} or conduct analyses about the current health status \citep{townsend2012}.
 Vegetation indices have shown the potential to provide valuable information for analyzing triggers to forest health \citep{jiang2014, adamczyk2015}.
-More and more studies focus on using machine-learning algorithms such as \ac{RF} \citep{belgiu2016, lary2016, michez2016}.
+Most studies focus on applying the Random Forest algorithm \ac{RF} \citep{belgiu2016, lary2016, michez2016}.
+Others are only used in few studies such as Support Vector Machines \citep{clark2018} or deep-learning algorithms such as Artifical Neural Networks \citep{ingram2005, rocha2018}.
 Overall, the idea is to fit a robust model which can be used to make predictions to (large) unknown areas, providing valuable information about the health condition of forest stands without the need for ground-truth data in these regions.
@@
 
+\subsection{Exploratory data analysis}
+% TODO: Explain why this is necessary here (differentes between plots)
+
 \subsection{Derivation of indices}
@@
 \subsubsection{Feature sets}
+% TODO: talk about the reasons behind different features sets: environmental implications
 
@@
 
-\subsubsection{Algorithm benchmarking}
-
 \section{Discussion}
@@
 
-\subsection{RMSE vs. plot characteristics}
+\subsection{Performance vs. plot characteristics}
 
@@
 
-\subsection{Variable importance}
-
-\subsection{Spatial prediction}
+\subsection{Feature selection methods}
 
 \subsection{Comparison to other studies}
+% TODO: Mention that (probably) most environmental studies only use PCA
+% TODO: find studies comparing filter methods
+% TODO: Find studies tackling the case of defoliation
 
@@
 
-\section{Conclusion}
+\section{Outlook and conclusion}
 

### 2019-06-09 analysis/paper/Schratz_hyperspectral.tex
more refs and rewrite of intro

@@
 
-% arara: clean: { extensions: [aux, bcf, cod, blg, lof, lot, out, toc, log, xml, bak0 ] }
+% arara: clean: { extensions: [aux, bcf, cod, blg, lof, lot, out, toc, xml, bak0 ] }
 
@@
 	\acro{MEM}{Maximum Entropy Model}
+	\acro{ML}{Machine-Learning}
 	\acro{NDII}{Normalized Difference Infrared Index}
@@
 % Link remote sensing and machine-learning
-The use of machine-learning algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
+The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
 Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
@@
 % now mention why this study is important
-However, we have the impression that the amount of (vegetation-)indices used in forest health studies could be higher, possibly leading to more valuable results.
-Possible reasons include the low spectral resolution of optical multispectral sensors, the lack of freely available hyperspectral data or the lack of knowledge about existing indices.
-% TODO: Ref for high-dim
-Another more applied issue is the fact of high-dimensionality which is introduced by the calculation of hundreds/thousands of indices.
-% TODO: Ref model interpr
-Even though ML algorithms are capable of handling possibly highly-correlated input variables, model runtime increases substantially and the interpretation of results becomes more complicated.
+In forest healh studies, it is valuable to have a lot input variables.
+When searching for possible triggers to forest health within forests, it is unclear upfront which information from the data might reveal important relationships during model fitting.
+If hyperspectral data is available, lots of indices can be derived.
+One issue that comes with this approach is the case of "high-dimensionality" \citep{trunk1979, xu2016}.
+Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increases substantially and the interpretation of results becomes more complicated.
 Also, in recent years the awareness of conducting proper tuning of model hyperparameters has increased in applied modelling fields.
-In this study we addressed these challenges by focusing on the following research questions:
+The reseach questions of this study are formulated around the handling of high-dimensionality in environmental modeling:
 
@@
 
-\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of ensemble filters.
-To model the numeric response variable defoliation at trees model-based optimization was applied for hyperparameter tuning and a spatial block \ac{CV} was used to account for spatial autocorrelation in the data.
+\noindent State-of-the-art machine-learning techniques were compared using three feature sets with different feature-selection approaches including a relatively novel approach of ensemble filters.
+To model the numeric response variable defoliation at trees model-based optimization was applied for hyperparameter tuning.
+Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.
 

### 2019-06-10 analysis/paper/Schratz_hyperspectral.tex
more refs for methods section

@@
 
-\subsection{Dealing with high-dimensionality}
-% Define "high-dimensionality"
-The usage of the term "high-dimensionality" in the machine-learning field practically means to deal with dataset that has a high amount of covariates.
-There is no clear threshold at which number the term can be applied.
-There has been an informal agreement to speak of "high-dimensonialty" if the number of variables exceeds the number of observations. REF
+\subsection{Dimension reduction}
 
-The case of a high-dimensional dataset comes with several challenges for both model fitting and evaluation.
-There is an increase in time for fitting a model to the data.
-Possible noise maybe introduced into the model by highly-correlated variables.
-The interpretation and prediction of models containing hundreds of covariates becomes more complicated, up to to the point at which interpretation or prediction is not possible anymore.
-Especially in the last decade the amount of data increased a lot and also the environmental modeling field profited from this as well.
-Numerous approaches were developed to deal with this situation and the field of "feature selection" can be safely seen as a standalone research field.
+It is important to distinguish the term "dimension reduction" from "high-dimensionality".
+The former applies to the general methodology of reducing dimensions of features.
+In the modelling field this means to find the best subset of covariates which provides the most predictive power to the model.
+The latter in contrast is restricted to the case when a dataset has $p > n$ where $p$ is the number of covariates and $n$ the number of observations \citep{Hastie2001}.
+Hence, there is no absolute value at which the term can be applied, both for $n$ and $p$.
+Hence for this study the word "high-dimensionality" only applies to the experiments containing the "NRI" feature set (7470 features for 1759 observations).
+
+The case of a feature-rich dataset comes with several challenges for both model fitting and evaluation.
+
+\begin{itemize}
+  \item Fitting times of the model increase
+  \item Noise is possibly introduced into the model by highly-correlated variables \citep{johnstoneiainm2009}
+  % I think we can phrase this better
+  \item The interpretation and prediction of models containing hundreds of covariates becomes more complicated, up to to the point at which interpretation or prediction is not possible anymore \citep{johnstoneiainm2009}
+\end{itemize}
+
+Especially in the last decade the amount of environmental data increased a lot, paramount by the increased data coverage of satellites.
+The need of applying automated feature-selection methods emerges in more and more studies.
+In the following sub-sections different approaches for selecting features are introduced.
 
@@
 So called "Filter methods" are one approach.
-The basic concept relies on the ranking of all features of the dataset using heuristics of the respective algorithm.
-Some are tailored towards specific types of variables (numeric or nominal data).
-After the covariates have been ranked, a decision needs to be made what features to keep and which to discard.
+The concept of "Filter methods" originates on the idea of ranking features using certain heuristics of an algorithm \citep{chandrashekar2014}.
+Some filter methods are restricted towards specific types of variables (numeric or nominal data).
+After the covariates have been ranked, a decision needs to be made what features to keep and which to discard \citep{drotar2015}.
 This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
@@
 The goal is to find the threshold (usually in \% of covariates) at which the model achieves a good performance while requiring the least amount of variables.
-In well-implemented software solutions the calculations of the ranking is only done once and then cached.
-In subsequent tuning iterations the results are taken from the cache instead of being calculated again.
+In well-implemented software solutions the calculations of the ranking is only done once and then cached \citep{mlr}.
+Hence in subsequent tuning iterations the results are loaded from the cache instead of getting calculated again.
 
 % Ensembe filter methods
-Besides the idea of using single filter methods to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models.
-This aligns with the recent rise of the "ensemble" approach which was first used to combine the predictions of multiple models.
-In this work we also used one ensemble filter method based on the "Borda" aggregation.
+Besides the idea of using single filter methods to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \citep{abeel2010, drotar2017a}.
+This aligns with the recent rise of the "ensemble" approach in model fitting which uses stacking to combine the predictions of multiple models \citep{polikar2012, feurer2015}.
+In this work we also used one ensemble filter method based on the "Borda" aggregation logic \citep{drotar2017a}.
 Here, the ranks of all single filters are summed up leading to a final ranking which consists of the ranks achieved by all filter algorithms.
@@
 
+% I grouped these two since we focus on filter methods in this study - or is this already a bit biased towards filter methods?
 \subsubsection{Other approaches}

### 2019-06-15 analysis/eda.Rmd
add EDA and spectral signatures report

---
title: "Exploratory data analysis"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

source(here::here("code/99-packages.R"))
sourceDirectory("R")

# load drake objects
loadd(vi_data, nri_data, bands_data, coords_vi_nri_clean)
library("DataExplorer")
library("dplyr")
library("ggsci")
library("ggplot2")
```

# Datasets {.tabset .tabset-fade}

## VI

### Overview

```{r eda-1}
intro <- introduce(vi_data)
knitr::kable(intro)
```

### Histograms

```{r eda-2}
plot_histogram(vi_data)
```

### PCA

```{r eda-3}
plot_prcomp(vi_data)
```

### Corr

```{r eda-4}
plot_correlation(vi_data)
```

## NRI

### Overview

```{r eda-5}
intro <- introduce(nri_data)
knitr::kable(intro)
```

### Histograms

No histograms for NRI -> too many features.

### PCA

```{r eda-6}
plot_prcomp(nri_data)
```

### Corr

```{r eda-7}
plot_correlation(nri_data)
```

## HR

### Overview

```{r eda-8}
intro <- introduce(bands_data)
knitr::kable(intro)
```

### Histograms

```{r eda-9}
plot_histogram(bands_data)
```

### PCA

```{r eda-10}
plot_prcomp(bands_data)
```

### Corr

```{r eda-11}
plot_correlation(bands_data)
```

# Custom plots

```{r eda-12}
vi_data_plot <- vi_data %>%
  mutate(plot = factor(rep(
    c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
    c(479, 451, 300, 529)
  )))
```

# Mean defoliation per plot

```{r eda-13}
mean_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(mean(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(mean(defoliation))
```

# Coefficient of variation

```{r eda-14}
cov_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise((sd(defoliation) / mean(defoliation)) * 100) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise((sd(defoliation) / mean(defoliation)) * 100)
```

# sd / skewness

```{r eda-15}
sd_skewness_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation))
```

```{r eda-16}
sd_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(sd(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(sd(defoliation))
```

```{r eda-17}
boxplot_defol <- vi_data_plot %>%
  group_by(plot) %>%
  ggboxplot(
    x = "plot", y = "defoliation", color = "plot",
    add = "jitter", add.params = list(size = "defoliation")
  ) +
  scale_size(range = c(0.5, 0.5)) +
  annotate("text",
    label = expression(bold(atop("n = 479", bar(x) ~ "= 55.96"))), x = 1,
    y = 112, size = 3, colour = "#BC3C29", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 451", bar(x) ~ "= 13.54"))), x = 2,
    y = 112, size = 3, colour = "#0072B5", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 300", bar(x) ~ "= 68.44"))), x = 3,
    y = 112, size = 3, colour = "#E18727", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 529", bar(x) ~ "= 69.22"))), x = 4,
    y = 112, size = 3, colour = "#20854E", fontface = 2
  ) +
  scale_color_nejm() +
  theme(legend.position = "none") +
  labs(y = "Total defoliation per tree (%)", x = "Plot")
boxplot_defol
```

# Point density

In Meters.

```{r eda-18}
plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
dist <- map(plots, ~ {
  coords <- coords_vi_nri_clean %>%
    mutate(plot = factor(rep(
      c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
      c(479, 451, 300, 529)
    ))) %>%
    filter(plot == .x) %>%
    dplyr::select(-plot)
  points <- SpatialPoints(
    coords = coords,
    proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
  )
  distance <- mean(as.dist(pointDistance(points, 
                                         allpairs = TRUE, 
                                         lonlat = FALSE)))
})
set_names(dist, plots)
dist_plots <- unlist(dist)
```

### 2019-06-15 code/07-reports.R

@@ -1,5 +1,9 @@
 reports_plan = drake_plan(
+  
+  eda_wfr = wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE),
 
-  defoliation_maps_wfr = wflow_publish(knitr_in("analysis/report-defoliation.Rmd"), view = FALSE)
+  defoliation_maps_wfr = wflow_publish(knitr_in("analysis/report-defoliation.Rmd"), view = FALSE),
+  
+  spectral_signatures_wfr = wflow_publish(knitr_in("analysis/spectral-signatures.Rmd"), view = FALSE),
 )
 

### 2019-06-15 analysis/eda.Rmd
update reports

@@
 
-```{r eda-17}
+```{r defoliation-distribution-plot, dev = c("png", "pdf")}
 boxplot_defol <- vi_data_plot %>%
@@
 ```
+
+```{r echo = FALSE}
+file_move(list.files(here("docs/figure/eda.Rmd/"), 
+                     pattern = "defoliation-distribution-plot", 
+                     full.names = TRUE),
+          here("code/98-paper/journal/"))
+```

### 2019-06-15 code/98-paper/journal/manuscript.tex
add manuscript.tex
upd bib
fix latex warnings, update bib

--- 2019-06-10 analysis/paper/Schratz_hyperspectral.tex
@@
 
-% arara: clean: { extensions: [aux, bcf, cod, blg, lof, lot, out, toc, xml, bak0 ] }
+% arara: clean: { extensions: [bcf, cod, blg, lof, lot, out, toc, xml, bak0 ] }
 
@@
 \graphicspath{{../03_figures/results/}{./}{../03_figures/data/}}
-
-% enable paragraph as 4th level https://tex.stackexchange.com/questions/60209/how-to-add-an-extra-level-of-sections-with-headings-below-subsubsection
-\usepackage{titlesec}
-\setcounter{secnumdepth}{4}
-\titleformat{\paragraph}
-{\normalfont\small\bfseries}{\theparagraph}{1em}{}
-\titlespacing*{\paragraph}
-{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 
@@
 	\acro{TPR}{True Positive Rate}
+	\acro{VI}{vegetation index}
+	\acro{XGBOOST}{extreme gradient boosting}
 \end{acronym}
@@
 
+%\begin{figure} [t!]
+%\begin{center}
+%\makebox[\textwidth]{\includegraphics[width=\textwidth] {defoliation-distribution-plot.pdf}}
+%\caption{Distribution of defoliation at trees for plots Laukiz 1, Laukiz 2, Luiando and Oiartzun.}
+%\label{fig:defol-distr}
+%\end{center}
+%\end{figure}
+
 \subsection{In-situ data}
@@
 
-%\begin{figure} [t!]
-%\begin{center}
-%\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
-%\caption{Information about the plot locations, the area of hyperspectral coverage and the number of trees per plot.}
-%\label{fig:study_area}
-%\end{center}
-%\end{figure}
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
+		\caption{Information about plot location and size, and spatial distribution of trees.}
+		\label{fig:study_area}
+	\end{center}
+\end{figure}
 
@@
 	\item Fitting times of the model increase
-  \item Noise is possibly introduced into the model by highly-correlated variables \citep{johnstoneiainm2009}
+	\item Noise is possibly introduced into the model by highly-correlated variables \citep{johnstoneiainm.2009}
 	      % I think we can phrase this better
-  \item The interpretation and prediction of models containing hundreds of covariates becomes more complicated, up to to the point at which interpretation or prediction is not possible anymore \citep{johnstoneiainm2009}
+	\item The interpretation and prediction of models containing hundreds of covariates becomes more complicated, up to to the point at which interpretation or prediction is not possible anymore \citep{johnstoneiainm.2009}
 \end{itemize}
@@
 
-\noindent Relating the modeling error to plot characteristics (mean point density, coefficient of variation) did not show a clear picture: For both comparisons, \textit{Laukiz 2} did not follow the pattern that was observed from the other three plots (\autoref{fig:plot-characteristics}) of having an increase in error with an increase in mean point density and coefficient of variation.
-
-It needs to be considered that we only looked at four plots in this work.
-To make a robust statement about a possible relationship between modeling error and plot characteristics, a larger sample size of plots is needed.
-
 \subsection{Predictive Performance}
@@
 
-\noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengara2016}.
+\noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengarajan2016}.
 This corresponds with the finding of this study that the most important variables are located in the NIR region.

### 2019-06-15 code/98-paper/journal/manuscript.tex
update manuscript and biblio

@@
 \section{Data and study area}
-\noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain (\autoref{fig:study_area}).
+\noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
 The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanza2016, iturritxa2017}.
-In-situ measurements of defoliation at trees (as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\%.
+In-situ measurements of defoliation at trees (as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}).
 The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{iturritxa2014}.
@@
 
-%\begin{figure} [t!]
-%\begin{center}
-%\makebox[\textwidth]{\includegraphics[width=\textwidth] {defoliation-distribution-plot.pdf}}
-%\caption{Distribution of defoliation at trees for plots Laukiz 1, Laukiz 2, Luiando and Oiartzun.}
-%\label{fig:defol-distr}
-%\end{center}
-%\end{figure}
+\begin{figure} [t!]
+\begin{center}
+\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {defoliation-distribution-plot-1.pdf}}
+\caption{Distribution of defoliation at trees for plots Laukiz 1, Laukiz 2, Luiando and Oiartzun.}
+\label{fig:defol-distr}
+\end{center}
+\end{figure}
 
@@
 All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
-In total 1750 observations are available (\textit{Laukiz 1} = 479, \textit{Laukiz 2} = 451, \textit{Luiando} = 291, \textit{Oiartzun} = 529).
+In total 1759 observations are available (\textit{Laukiz 1} = 479, \textit{Laukiz 2} = 451, \textit{Luiando} = 300, \textit{Oiartzun} = 529).
 The data was surveyed in September 2016.
@@
 		\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
-		\caption{Information about plot location and size, and spatial distribution of trees.}
+		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}
 		\label{fig:study_area}
@@
 
-\subsection{Exploratory data analysis}
-% TODO: Explain why this is necessary here (differentes between plots)
-
 \subsection{Derivation of indices}
 % TODO: Mention different feature sets
-\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the \textit{hsdar} package (90 in total) as well as all possible \ac{NRI} combinations were calculated.
+\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the R package \textit{hsdar} (90 in total) as well as all possible \ac{NRI} combinations were calculated.
 The following formula was used for the NRI calculation:
@@
 In total, $\frac{125*126}{2} = 7875$ NRIs were calculated .
-Due to four corrupted bands and numerical calculation errors for some band combinations, a total of 7471 indices was available for each observation.
-If an index returned \texttt{NA} for one plot, the index was removed from all plots.
-This decision was based on the fact that due to the mass of available indices and the low number of observations, the latter was valued higher.
+Due to four corrupted bands and numerical calculation errors for some band combinations, a total of 7471 indices were available for each observation.
+All indices which contained missing values for at least one plot were removed from the dataset.
+Observations were given more importance than indices due to the mass of available \ac{NRI} indices.
 
@@
 
-The benchmarking matrix of this study consists of the following algorithms:
+\noindent The benchmarking matrix of this study consists of the following algorithms:
 
 \begin{itemize}
-	\item  Extreme Gradient Boosting (\ac{XGBOOST})
-	\item  Random Forest (\ac{RF})
+	\item  Extreme Gradient Boosting (XGBOOST)
+	\item  Random Forest (RF)
 	\item  Penalized Regression (both L1 and L2)
-	\item  Support Vector Machine (\ac{SVM})
+	\item  Support Vector Machine (SVM)
 \end{itemize}
 
-\ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
+\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
 Extreme Gradient Boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
-Penalized Regression is a statistical modeling technique that is capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model.
-The penalties are also known as LASSO (L1) and RIDGE (L2).
+Penalized Regression is a statistical modeling technique that is capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \citep{hastie2001}.
+The penalties are also known as "lasso" (L1) and "ridge" (L2).
 The former does not allow the removal of variables from the model (penalization to zero) while the latter does.
 Both penalties can also be combined.
-The combined approach is called "elastic net".
+The combined approach is called "elastic net" but was not used in this work.
 
@@
 
-Three feature sets were used in this study:
+\noindent Three feature sets were used in this study with each representing a different way of feature-engineering:
 
 \begin{itemize}
-	\item The raw hyperspectal band information (HR, $n = 122$)
-	\item Vegetation Indices (\ac{VI}, $n = 90$)
-	\item Normalized Ratio Indices (\ac{NRI}, $n = 7471$)
+  \item The raw hyperspectal band information (HR): No feature engineering
+  \item Vegetation Indices (\ac{VI}: Expert-based feature engineering
+  \item Normalized Ratio Indices (\ac{NRI}): Automated feature-engineering
 \end{itemize}
@@
 
-An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
+\noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
 Maximum Bayesian Optimization \citep{mlrmbo} was used for parameter optimization.
@@
 
-A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible.
-Each tree plot served as one fold within the resampling setup, resulting in four folds total.
-For the inner level, $p - 1$ folds were used (with $p$ being the number of plots).
+\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible.
+Each plot served as one fold within the resampling setup, resulting in four folds total.
+For the inner level (hyperparameter tuning), $p - 1$ folds were used (with $p$ being the number of plots). REF SCHRATZ
 
@@
 
-It is important to distinguish the term "dimension reduction" from "high-dimensionality".
-The former applies to the general methodology of reducing dimensions of features.
-In the modelling field this means to find the best subset of covariates which provides the most predictive power to the model.
-The latter in contrast is restricted to the case when a dataset has $p > n$ where $p$ is the number of covariates and $n$ the number of observations \citep{Hastie2001}.
-Hence, there is no absolute value at which the term can be applied, both for $n$ and $p$.
-Hence for this study the word "high-dimensionality" only applies to the experiments containing the "NRI" feature set (7470 features for 1759 observations).
+\noindent It is important to distinguish the term "dimension reduction" from "high-dimensionality".
+The former refers to the general idea of reducing features from a dataset \citep{vandermaaten2007}.
+In modeling this means finding the best subset of covariates which provide the most predictive power to the model.
+The latter term in contrast is the definition of a dataset attribute:
+When $p > n,$ where $p$ is the number of covariates and $n$ the number of observations \citep{hastie2001}, the dataset qualifies as "high-dimensional".
+There is no absolute value at which the term can be applied, both for $n$ and $p$.
+Hence for this study the word "high-dimensionality" only applies to the experiments using the "NRI" feature set (7470 features for 1759 observations).
 
@@
 	      % I think we can phrase this better
-	\item The interpretation and prediction of models containing hundreds of covariates becomes more complicated, up to to the point at which interpretation or prediction is not possible anymore \citep{johnstoneiainm.2009}
+	\item Interpretation and prediction becomes more complex, up to the point at which it is not possible anymore \citep{johnstoneiainm.2009}
 \end{itemize}
 
-Especially in the last decade the amount of environmental data increased a lot, paramount by the increased data coverage of satellites.
-The need of applying automated feature-selection methods emerges in more and more studies.
-In the following sub-sections different approaches for selecting features are introduced.
+\noindent In the last decade the amount of environmental data increased a lot, paramount by the increased data coverage of satellites.
+With this the need of applying automated feature-selection methods emerged.
+In the following sections the currently most used feature selection approaches are briefly introduced.
 
@@
 % Filter methods
-So called "Filter methods" are one approach.
-The concept of "Filter methods" originates on the idea of ranking features using certain heuristics of an algorithm \citep{chandrashekar2014}.
+\noindent The concept of "Filter methods" originates on the idea of ranking features using certain heuristics of an algorithm \citep{chandrashekar2014}.
 Some filter methods are restricted towards specific types of variables (numeric or nominal data).
@@
 This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
-Essentially, the number of covariates in the model is treated as a hyperparameter of the respective model.
-The goal is to find the threshold (usually in \% of covariates) at which the model achieves a good performance while requiring the least amount of variables.
-In well-implemented software solutions the calculations of the ranking is only done once and then cached \citep{mlr}.
-Hence in subsequent tuning iterations the results are loaded from the cache instead of getting calculated again.
+Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
+The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
+In well-implemented software solutions the filter calculation is only done once and then cached, saving valuable resources \citep{mlr}.
 
 % Ensembe filter methods
-Besides the idea of using single filter methods to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \citep{abeel2010, drotar2017a}.
-This aligns with the recent rise of the "ensemble" approach in model fitting which uses stacking to combine the predictions of multiple models \citep{polikar2012, feurer2015}.
-In this work we also used one ensemble filter method based on the "Borda" aggregation logic \citep{drotar2017a}.
-Here, the ranks of all single filters are summed up leading to a final ranking which consists of the ranks achieved by all filter algorithms.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \citep{abeel2010, drotar2017a}.
+This aligns with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models \citep{polikar2012, feurer2015}.
+In this work the "Borda" ensemble filter was applied \citep{drotar2017a}.
+Here, the ranking is based on the sum of all single filters.
 
@@
 % Wrapper approach
-Another approach to reduce the number of variables is to apply algorithms that are normally used for hyperparameter optimization such as \textit{Random Search} or a \textit{Genetic Algorithm}.
-Here, first a (random) subset of features is chosen based on the selected algorithm.
-In comparison to "filters", no ranking is applied here.
+\noindent Another approach to reduce the number of variables is to apply algorithms that are used for hyperparameter optimization such as "Random Search" or a "Generic Simulated Annealing".
+First a (random) subset of features is chosen based on the selected algorithm.
+In comparison to "filters", no ranking is applied in this step.
 Now the model is fitted on the data and the performance is evaluated.
-A major downside of this approach is that hyperparameter tuning can only start after the feature-selection optimization has ended.
+A disadvantage of this approach is that hyperparameter tuning can only start after the feature-selection optimization has ended.
 Hence, this method is very expensive because two optimization stages need to be run.
-In science this approach is also referred to as "wrapper method".
+In science this approach is also referred to as "wrapper method" \citep{chandrashekar2014, kohavi1997}.
 It was not considered in this work due to the expensive runtime.
@@
 % PCA
-A third commonly used method is the "Principal Component Analysis".
+A third commonly used method is the "Principal Component Analysis" \citep{pearson1901, jolliffe2016}.
 Here, the main components of the feature space are extracted and combined.
 Usually the first two components are taken since these contain the major information of the covariates.
-By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the condensed information of most variables of the data.
+By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
 This enables a very cheap model fitting with a minimal loss of predictor information.
@@
 The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chen2016} (\textit{xgboost}), \textit{e1071} \citep{e1071} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
-The filter methods of the following packages were used: \textit{praznik}, \textit{fselectorrcpp}.
+The filter methods of the following packages were used: \textit{praznik \citep{praznik}}, \textit{FSelectorRcpp} \citep{fselectorrcpp}.
 The R package \textit{mlr} \citep{mlr} was used for all modeling related steps.
@@
 \section{Results}
-
-\subsection{Plot characteristics}
-
-%% defoliation boxplots
-%\begin{figure} [t!]
-%\begin{center}
-%\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {boxplot_defol.pdf}}
-%\caption{Descriptive statistics of the response variable \textit{defoliation}.}
-%\label{fig:defol_boxplots}
-%\end{center}
-%\end{figure}
-
-%\noindent \textit{Oiartzun} shows the highest defoliation ($\bar{x} = 69.22 \%$) among the plots while \textit{Laukiz 2} is the healthiest ($\bar{x} = 13.54 \%$) (\autoref{fig:defol_boxplots}).
-%All plots besides \textit{Luiando} show an evenly distributed level of defoliation across the entire plot.
-
-%\noindent The high degree of defoliation for \textit{Luiando} and \textit{Oiartzun} is also visible in the spectral signatures of the plots (\autoref{fig:spectral_signatures}).
-%Both plots show lower mean reflectance values around the wavelength range 800 nm - 1000 nm compared to \textit{Laukiz 1} and \textit{Laukiz 2}.
-%\textit{Oiartzun} is almost completely missing the reflectance drop at around 815 nm that is visible for all other plots but instead shows a higher magnitude for the reflectance increase at around 920 nm.
-
-%\noindent \textit{Laukiz 2} shows a mean tree density of 61.59 m (\autoref{fig:plot-characteristics}) while all other plots have a higher density (34.64 m (Laukiz 1), 33.01 m (Luiando), 34.96 m (Oiartzun)) (\autoref{fig:plot-characteristics}).
 
@@
 
-\noindent The buffer of 2 m that we used to generate the index value for each observation can be seen critical.
-When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation that does not spatially match (due to the geometric offset of 1 m in the hyperspectral data).
+\noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
+When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation which does not spatially match with the hyperspectral (due to the geometric offset of 1 m).
 Using a buffer of more than 2 meters would increase the probability of merging information from other trees into the pixel value, blurring the actual value of the tree observation.
@@
 
-another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
+Another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
 As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels are touched by the buffer.

### 2019-06-16 code/98-paper/journal/manuscript.tex
Updates from Overleaf

@@
 
+Most other studies analyzing defoliation operated on the plot rather than the tree level due to spatial resolution of the satellite products that served as the input data (\citep{townsend2012, debeurs2008a}.
+Studies focusing on the tree-level used ground-level methods such as \ac{ALS} or \ac{LiDAR} \citep{meng2018, kalin2019}.
+
 \noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengarajan2016}.

### 2019-06-16 code/98-paper/journal/manuscript.tex
manuscript

--- 2019-06-15 code/98-paper/journal/manuscript.tex
@@
 
-% arara: indent: {overwrite: yes}
+%% arara: indent: {overwrite: yes}
 
@@
 	\acro{AGB}{Above-Ground Biomass}
+	\acro{ALS}{airborn laser scanning}
 	\acro{ANN}{Artificial Neural Network}
@@
 	\acro{CART}{Classification and Regression Trees}
+	\acro{CNN}{convolutional neural networks}
 	\acro{CV}{cross-validation}
+	\acro{DAP}{digital aerial photogrammetry}
 	\acro{ENM}{Environmental Niche Modeling}
 	\acro{FPR}{False Positive Rate}
+	\acro{FFS}{forward feature-selection}
+	\acro{FS}{feature-selection}
 	\acro{GAM}{Generalized Additive Model}
@@
 	\acro{NRI}{Normalized Ratio Index}
-	\acro{OLS}{Ordinary Least Squares}
+	\acro{OLS}{ordinary least squares}
+	\acro{LiDAR}{light detection and ranging}
 	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
 	\acro{PISR}{Potential Incoming Solar Radiation}
+	\acro{PLS}{partial least-squares}
 	\acro{RBF}{Radial Basis Function}
@@
 
-% Ensembe filter methods
+% Ensemble filter methods
 Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \citep{abeel2010, drotar2017a}.
@@
 The R package \textit{mlr} \citep{mlr} was used for all modeling related steps.
-\textit{drake} was used for structuring the work and ensuring reproducbility.
+\textit{drake} was used for structuring the work and ensuring reproducibility.
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
@@
 Using a buffer of more than 2 meters would increase the probability of merging information from other trees into the pixel value, blurring the actual value of the tree observation.
-Thats why in our view using a buffer of 2 m was the best compromise here.
+That is why in our view using a buffer of 2 m was the best compromise here.
 
@@
 
-\subsection{Comparison to other studies}
-% TODO: Mention that (probably) most environmental studies only use PCA
-% TODO: find studies comparing filter methods
-% TODO: Find studies tackling the case of defoliation
+\subsection{Comparison to other studies}A
+% NOTE: no environmental studies use filter methods, some use FFS
+% NOTE: only few studies analyze defoliation
 
-\noindent Other studies analyzing defoliation found that reflectance differences between defoliated and 30 \% defoliated trees of up to 10\% exist in the \ac{NIR} region \citep{rengarajan2016}.
-This corresponds with the finding of this study that the most important variables are located in the NIR region.
+\noindent Most other studies analyzing defoliation operated on the plot rather than the tree level.
+This originated simply due to spatial resolution of the satellite products that served as the input data \citep{townsend2012, debeurs2008a, rengarajan2016}.
 
-\cite{goodbody2018} used NDVI and structural measures in as inputs for a partial least squares analysis to model defoliation caused by the spruce budworm.
-Results showed that metrics from spectral features were most important.
-Incorporating spectral metrics could be a possible enhancement for future studies.
+Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR} \citep{meng2018, kalin2019}.
+\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
+Both study designs are substantially different compared to the setup of this work.
+In addition, no spatial \ac{CV} or \ac{FS} was used.
+\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
+Study results indicated that spectral metrics were found to be most helpful for the model.
+Incorporating such metrics (both spectral and structural) could be a possible enhancement for future works.
 
-\cite{townsend2012} used Landsat data to model defoliation caused by insect herbivores.
-They found that the \ac{NDII} ($\frac{Band 4 - Band 5}{Band 4 + Band 5}$) and the moisture stress index ($\frac{Band 5}{Band 4}$) gave better results than using NDVI.
-Overall, they used 10 vegetation indices derived from Landsat data.
+\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
+\cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
+A grid-search was used for hyperparameter tuning and a \ac{FFS} for variable selection.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach \citep{meyer2018} on a Random Forest classifier.
 
-% TODO: Find missing reference
-%MODIS data was used by \cite{debeursEstimatingEffectGypsy2008} to model defoliation caused by the gypsy moth using vegetation indices such as NDVI, EVI, NDWI and NDII.
-
-%All of these examples validate the approach of using vegetation indices to model defoliation.
-%Even though the spatial resolution of the data in these studies varied between hundreds of meters (MODIS) \cite{debeursEstimatingEffectGypsy2008} and centimeters \cite{goodbody2018}, high resolution data is preferred to fit accurate models.
-%Also, the importance of certain indices (e.g. NDVI, EVI) will vary based on the data and resolution.
-%The finding of this work that the vegetation indices GDVI and EVI are most important for the fitted model could not be verified by other studies.
-%However, the presented ones did often only use a small subset of the vegetation indices that were used in this study.
-
-%We could not find a study that used recent machine-learning techniques in combination with a high amount of variables to model defoliation.
-%This fact highlights the importance of this work and will hopefully encourage scientists to use machine-learning techniques, feature selection methods and a wider selection of vegetation indices when assessing defoliation in the future.
+In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
+Most studies used only one algorithm (usually Random Forest) without strong arguments why this particular one has been selected.
+In our view the number of possible derivable input features is often higher than the number of variables used in the end, missing out potential helpful information from the data.
+These findings underline the importance of this work in the field of forest health modeling.
 
@@
 
-\noindent In this work we used various indices derived from hyperspectral remote sensing data to estimate defoliationas a proxy for tree health in northern Spain.
+\noindent In this work we used various indices derived from hyperspectral remote sensing data to estimate defoliation as a proxy for tree health in northern Spain.
 In the algorithm comparison \textit{xgboost} showed the best performance among the tested ones.
@@
 The spatial prediction showed a satisfying result making it possible to distinguish highly defoliated plots from plots with low defoliation easily.
-The most important indices for the fitted model were widely known vegetations indices such as EVI, GDVI or NDVI.
+The most important indices for the fitted model were widely known vegetation indices such as EVI, GDVI or NDVI.
 In future studies it would be interesting to link the indices of this work to other indicators of forest/vegetation health and analyze their importance and the model performances.

### 2019-06-16 code/98-paper/journal/manuscript.tex
manuscript

@@
 Also, in recent years the awareness of conducting proper tuning of model hyperparameters has increased in applied modelling fields.
-The reseach questions of this study are formulated around the handling of high-dimensionality in environmental modeling:
+The reseach questions of this study are formulated around the handling of feature-rich datasets in environmental modeling:
 
 \begin{itemize}
-	\item How to handle and analyze high-dimensional data including model optimization while accounting for spatial characteristics?
-	\item Does the calculation of arbitrary indices add any value to the analysis or is the band information from the sensor already sufficient?
-	\item How can interpretability of the model be maintained while at the same time using all possible input data?
-\end{itemize}
+	\item Do different environmental feature sets show a substantial different in performance when predicting defoliation?
+    \item How do feature-selection methods perform amongst each other and do they improve the predictive performance?
+    \item Is is possible to map defolation at a tree level using high-resolution hyperspectral data at a sufficient error rate?
+    \item Which features were most important to the model to detect defoliation at trees?\end{itemize}
 

### 2019-06-16 code/98-paper/journal/manuscript.tex
clean discussion

@@
 
-\noindent In this work we used various indices derived from hyperspectral remote sensing data to estimate defoliation as a proxy for tree health in northern Spain.
-In the algorithm comparison \textit{xgboost} showed the best performance among the tested ones.
-Even though \ac{RR} is able to handle highly correlated data, it was far from achieving an acceptable performance in this work.
-
-The fitted models on the plot level showed a promising performance with RMSE values between 8 and 20 which validated the approach of relating defoliation to remote sensing indices.
-The performance of the model containing all data (four plots) was acceptable (RMSE 29.59) but can be improved by adding more observations from other plots in future studies.
-
-The spatial prediction showed a satisfying result making it possible to distinguish highly defoliated plots from plots with low defoliation easily.
-The most important indices for the fitted model were widely known vegetation indices such as EVI, GDVI or NDVI.
-In future studies it would be interesting to link the indices of this work to other indicators of forest/vegetation health and analyze their importance and the model performances.
-
 \section{Appendix}

### 2019-06-16 code/98-paper/journal/manuscript.tex
rewrite introduction

@@
 % Link remote sensing and machine-learning
-The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
+\noindent The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \citep{lary2016}.
 Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
@@
 % link to forest health analysis and show exemplary studies
-The monitoring of forest ecosystems is one exemplary field that profits from these new possibilities.
-Specifically the monitoring of forest health profited from the ability of analyzing large scale data with efficient algorithms able to predict to unknown areas \citep{hawrylo2018, mascaro2014}.
-For this task, usually optical data from multi-/hyperspectral satellites is used.
-Various indices are calculated from the data with the aim of carrying out temporal change detections \citep{zhang2016} or conduct analyses about the current health status \citep{townsend2012}.
-Vegetation indices have shown the potential to provide valuable information for analyzing triggers to forest health \citep{jiang2014, adamczyk2015}.
-Most studies focus on applying the Random Forest algorithm \ac{RF} \citep{belgiu2016, lary2016, michez2016}.
-Others are only used in few studies such as Support Vector Machines \citep{clark2018} or deep-learning algorithms such as Artifical Neural Networks \citep{ingram2005, rocha2018}.
-Overall, the idea is to fit a robust model which can be used to make predictions to (large) unknown areas, providing valuable information about the health condition of forest stands without the need for ground-truth data in these regions.
+Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \citep{mascaro2014, urban2018}.
+The advantages of conducting analysis at large scale make these points a promising toolset also for forest health analysis.
+This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \citep{hawrylo2018}.
+Different approaches exist including temporal change detections \citep{zhang2016} or analysis about the current health status of forests on a stand level \citep{townsend2012}.
+Defoliation at trees is an information which is able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \citep{townsend2012, goodbody2018}.
+Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios \citep{jiang2014, adamczyk2015}.
+Since both triggers and tree types are highly variable across regions, it cannot be known apriori which indices are most important for describing the actual relationship.
+This emphasizes the need to extract as much information as possible from the available input data to generate features that can possibly explain the modeled relationship.
+Besides \ac{VI} a less known type of indices that can be calculated from remote-sensing products are \ac{NRI}.
+\ac{NRI} are arbitrary combinations of band information from remote-sensing data.
+Especially when working with hyperspectral data, \ac{NRI} provide the possibility to calculate hundreds of indices that can be used during modeling.
 
-% now mention why this study is important
-In forest healh studies, it is valuable to have a lot input variables.
-When searching for possible triggers to forest health within forests, it is unclear upfront which information from the data might reveal important relationships during model fitting.
-If hyperspectral data is available, lots of indices can be derived.
-One issue that comes with this approach is the case of "high-dimensionality" \citep{trunk1979, xu2016}.
-Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increases substantially and the interpretation of results becomes more complicated.
-Also, in recent years the awareness of conducting proper tuning of model hyperparameters has increased in applied modelling fields.
-The reseach questions of this study are formulated around the handling of feature-rich datasets in environmental modeling:
+% now that we introduced the importance of remote sensing indices for forest health,
+
+Despite its popularity in environmental modeling, there are no studies so far that used machine-learning algorithms in combination with remote-sensing data to analyze defoliation on a tree level.
+This study aims at closing this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
+To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms is used.
+The approach of generating many features comes along with the introduction of "high-dimensionality" into the analysis \citep{trunk1979, xu2016}.
+Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increase substantially and the interpretation of results becomes more complicated.
+Besides modeling defoliation at trees this work also aims at providing an exemplary analysis of how to deal with spatial feature-rich data derived from a hyperspectral remote-sensing source.
+
+The reseach questions of this study are the following:
 
 \begin{itemize}
-	\item Do different environmental feature sets show a substantial different in performance when predicting defoliation?
-    \item How do feature-selection methods perform amongst each other and do they improve the predictive performance?
-    \item Is is possible to map defolation at a tree level using high-resolution hyperspectral data at a sufficient error rate?
+	\item Do different environmental feature sets show substantial differences in performance when predicting defoliation at trees?
+    \item How are different feature-selection methods performing and do they improve the predictive performance of the final model?
+    \item Is it possible to map defoliation at a tree level using high-resolution hyperspectral data at a sufficient error rate?
     \item Which features were most important to the model to detect defoliation at trees?\end{itemize}
@@
 \noindent State-of-the-art machine-learning techniques were compared using three feature sets with different feature-selection approaches including a relatively novel approach of ensemble filters.
-To model the numeric response variable defoliation at trees model-based optimization was applied for hyperparameter tuning.
+Model-based optimization was applied for hyperparameter tuning of the chosen algorithms.
 Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.

### 2019-06-18 analysis/eda.Rmd
move figure
fix git

@@
 intro <- introduce(vi_data)
-knitr::kable(intro)
+
+intro_df <- data.frame(
+  "Name" = c("Rows", "Columns",
+             "Discrete columns", "Continuous columns", "All missing columns",
+             "Missing observations", "Complete Rows",
+             "Total observations"),
+  "Value" = c(
+    format(intro[["rows"]], big.mark = ","),
+    format(intro[["columns"]], big.mark = ","),
+    format(intro[["discrete_columns"]], big.mark = ","),
+    format(intro[["continuous_columns"]], big.mark = ","),
+    format(intro[["all_missing_columns"]], big.mark = ","),
+    format(intro[["total_missing_values"]], big.mark = ","),
+    format(intro[["complete_rows"]], big.mark = ","),
+    format(intro[["total_observations"]], big.mark = ",")
+  )
+)
+kable(intro_df)
 ```
@@
 intro <- introduce(nri_data)
-knitr::kable(intro)
+intro_df <- data.frame(
+  "Name" = c("Rows", "Columns",
+             "Discrete columns", "Continuous columns", "All missing columns",
+             "Missing observations", "Complete Rows",
+             "Total observations",
+  "Value" = c(
+    format(intro[["rows"]], big.mark = ","),
+    format(intro[["columns"]], big.mark = ","),
+    format(intro[["discrete_columns"]], big.mark = ","),
+    format(intro[["continuous_columns"]], big.mark = ","),
+    format(intro[["all_missing_columns"]], big.mark = ","),
+    format(intro[["total_missing_values"]], big.mark = ","),
+    format(intro[["complete_rows"]], big.mark = ","),
+    format(intro[["total_observations"]], big.mark = ",")
+  )
+)
+kable(intro_df)
 ```
@@
 intro <- introduce(bands_data)
-knitr::kable(intro)
+intro_df <- data.frame(
+  "Name" = c("Rows", "Columns",
+             "Discrete columns", "Continuous columns", "All missing columns",
+             "Missing observations", "Complete Rows",
+             "Total observations",
+  "Value" = c(
+    format(intro[["rows"]], big.mark = ","),
+    format(intro[["columns"]], big.mark = ","),
+    format(intro[["discrete_columns"]], big.mark = ","),
+    format(intro[["continuous_columns"]], big.mark = ","),
+    format(intro[["all_missing_columns"]], big.mark = ","),
+    format(intro[["total_missing_values"]], big.mark = ","),
+    format(intro[["complete_rows"]], big.mark = ","),
+    format(intro[["total_observations"]], big.mark = ",")
+  )
+)
+kable(intro_df)
 ```
@@
 ```{r echo = FALSE}
-file_move(list.files(here("docs/figure/eda.Rmd/"), 
-                     pattern = "defoliation-distribution-plot", 
-                     full.names = TRUE),
-          here("code/98-paper/journal/"))
+file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
+          "code/98-paper/journal/")
 ```

### 2019-06-18 analysis/test/cloud-cover.Rmd
fix git

--- 2019-04-24 analysis/rmd/cloud-cover.Rmd
@@
                       fig.height = 6.13, out.width = "100%")
+source(here::here("code/99-packages.R"))
 
@@
 ```{r}
-records %<>% split(records$beginposition)
+records %<>% split(records[[2]]$beginposition)
 

### 2019-06-18 code/05-modeling/paper/learner.R

@@ -1,7 +1,7 @@
 # Random Forest -----------------------------------------------------------
 
 lrn_rf <- makeLearner(
-  "regr.ranger"
+  "regr.ranger.pow"
 )
 
 # XGBOOST -----------------------------------------------------------------
@@ -17,8 +17,9 @@ lrn_xgboost <- makeLearner(
 # SVM ---------------------------------------------------------------------
 
 lrn_svm <- makeLearner(
-  "regr.svm",
-  scale = FALSE
+  "regr.ksvm",
+  scaled = FALSE,
+  fit = FALSE
 )
 
 # LASSO ---------------------------------------------------------------------

### 2019-06-18 code/05-modeling/paper/param-sets.R

@@ -5,7 +5,7 @@ ps_xgboost_filter = makeParamSet(
   makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
   makeNumericParam("subsample", lower = 0.25, upper = 1),
   makeIntegerParam("max_depth", lower = 1, upper = 10),
-  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("sigma", lower = 0, upper = 10),
   makeNumericParam("eta", lower = 0.001, upper = 0.6),
   makeNumericParam("min_child_weight", lower = 0, upper = 20),
   makeNumericParam("fw.perc", lower = 0, upper = 1)
@@ -16,67 +16,66 @@ ps_xgboost = makeParamSet(
   makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
   makeNumericParam("subsample", lower = 0.25, upper = 1),
   makeIntegerParam("max_depth", lower = 1, upper = 10),
-  makeNumericParam("gamma", lower = 0, upper = 10),
+  makeNumericParam("sigma", lower = 0, upper = 10),
   makeNumericParam("eta", lower = 0.001, upper = 0.6),
   makeNumericParam("min_child_weight", lower = 0, upper = 20)
 )
 
 ps_xgboost_pca =  makeParamSet(
-    makeIntegerParam("nrounds", lower = 10, upper = 600),
-    makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
-    makeNumericParam("subsample", lower = 0.25, upper = 1),
-    makeIntegerParam("max_depth", lower = 1, upper = 10),
-    makeNumericParam("gamma", lower = 0, upper = 10),
-    makeNumericParam("eta", lower = 0.001, upper = 0.6),
-    makeNumericParam("min_child_weight", lower = 0, upper = 20),
-    makeIntegerParam("ppc.pcaComp", lower = 1, upper = 10)
-  )
+  makeIntegerParam("nrounds", lower = 10, upper = 600),
+  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
+  makeNumericParam("subsample", lower = 0.25, upper = 1),
+  makeIntegerParam("max_depth", lower = 1, upper = 10),
+  makeNumericParam("sigma", lower = 0, upper = 10),
+  makeNumericParam("eta", lower = 0.001, upper = 0.6),
+  makeNumericParam("min_child_weight", lower = 0, upper = 20),
[…]

### 2019-06-18 code/06-benchmark-matrix.R

@@ -40,29 +40,29 @@ bm_plan = drake_plan(bm = target(
                 xgboost_no_filter,
                 xgboost_pca,
 
-                ridge_borda,
-                ridge_info.gain,
-                ridge_gain.ratio,
-                ridge_variance,
-                ridge_rank.cor,
-                ridge_linear.cor,
-                ridge_mrmr,
-                ridge_cmim,
-                ridge_carscore,
+                # ridge_borda,
+                # ridge_info.gain,
+                # ridge_gain.ratio,
+                # ridge_variance,
+                # ridge_rank.cor,
+                # ridge_linear.cor,
+                # ridge_mrmr,
+                # ridge_cmim,
+                # ridge_carscore,
                 ridge_no_filter,
-                ridge_pca,
+                #ridge_pca,
 
-                lasso_borda,
-                lasso_info.gain,
-                lasso_gain.ratio,
-                lasso_variance,
-                lasso_rank.cor,
-                lasso_linear.cor,
-                lasso_mrmr,
-                lasso_cmim,
-                lasso_carscore,
-                lasso_no_filter,
-                lasso_pca
+                # lasso_borda,
+                # lasso_info.gain,
+                # lasso_gain.ratio,
+                # lasso_variance,
+                # lasso_rank.cor,
+                # lasso_linear.cor,
+                # lasso_mrmr,
+                # lasso_cmim,
+                # lasso_carscore,
+                lasso_no_filter#,
+                #lasso_pca
 
                 ),
     # task_name = c("vi", "nri", "hr"), # set pretty names for benchmark matrix

### 2019-06-18 code/99-packages.R

@@ -31,3 +31,4 @@ suppressPackageStartupMessages(library(here))
 suppressPackageStartupMessages(library(workflowr))
 suppressPackageStartupMessages(library(praznik))
 suppressPackageStartupMessages(library(mRMRe))
+suppressPackageStartupMessages(library(kernlab))

### 2019-06-19 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE)

--- 2019-06-15 analysis/eda.Rmd
@@
 # load drake objects
-loadd(vi_data, nri_data, bands_data, coords_vi_nri_clean)
+loadd(vi_data, nri_data, bands_data, coords_vi_nri_clean,
+
+      fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
+      fv_nri_rank, fv_nri_cor,
+      fv_nri_mrmr,
+      fv_nri_cmim, fv_nri_var,
+
+      fv_vi_car, fv_vi_info.gain, fv_vi_gain.ratio,
+      fv_vi_rank, fv_vi_cor,
+      fv_vi_mrmr,
+      fv_vi_cmim, fv_vi_var,
+
+      fv_hr_car, fv_hr_info.gain, fv_hr_gain.ratio,
+      fv_hr_rank, fv_hr_cor,
+      fv_hr_mrmr,
+      fv_hr_cmim, fv_hr_var
+      )
 library("DataExplorer")
@@
 
-```{r eda-1}
+```{r eda-1 }
 intro <- introduce(vi_data)
-knitr::kable(intro)
+intro_df <- data.frame(
+  "Name" = c("Rows", "Columns",
+             "Discrete columns", "Continuous columns", "All missing columns",
+             "Missing observations", "Complete Rows",
+             "Total observations"),
+  "Value" = c(
+    format(intro[["rows"]], big.mark = ","),
+    format(intro[["columns"]], big.mark = ","),
+    format(intro[["discrete_columns"]], big.mark = ","),
+    format(intro[["continuous_columns"]], big.mark = ","),
+    format(intro[["all_missing_columns"]], big.mark = ","),
+    format(intro[["total_missing_values"]], big.mark = ","),
+    format(intro[["complete_rows"]], big.mark = ","),
+    format(intro[["total_observations"]], big.mark = ",")
+  )
+)
+knitr::kable(intro_df)
 ```
@@
 
-```{r eda-2}
+```{r eda-2 }
 plot_histogram(vi_data)
@@
 
-```{r eda-3}
+```{r eda-3 }
 plot_prcomp(vi_data)
@@
 
-```{r eda-4}
+```{r eda-4 }
 plot_correlation(vi_data)
@@
 
-```{r eda-5}
+```{r eda-5 }
 intro <- introduce(nri_data)
-knitr::kable(intro)
+intro_df <- data.frame(
+  "Name" = c("Rows", "Columns",
+             "Discrete columns", "Continuous columns", "All missing columns",
+             "Missing observations", "Complete Rows",
+             "Total observations"),
+  "Value" = c(
+    format(intro[["rows"]], big.mark = ","),
+    format(intro[["columns"]], big.mark = ","),
+    format(intro[["discrete_columns"]], big.mark = ","),
+    format(intro[["continuous_columns"]], big.mark = ","),
+    format(intro[["all_missing_columns"]], big.mark = ","),
+    format(intro[["total_missing_values"]], big.mark = ","),
+    format(intro[["complete_rows"]], big.mark = ","),
+    format(intro[["total_observations"]], big.mark = ",")
+  )
+)
+kable(intro_df)
 ```
@@
 
-```{r eda-6}
+```{r eda-6 }
 plot_prcomp(nri_data)
@@
 
-```{r eda-7}
+```{r eda-7 }
 plot_correlation(nri_data)
@@
 
-```{r eda-8}
+```{r eda-8 }
 intro <- introduce(bands_data)
-knitr::kable(intro)
+intro_df <- data.frame(
+  "Name" = c("Rows", "Columns",
+             "Discrete columns", "Continuous columns", "All missing columns",
+             "Missing observations", "Complete Rows",
+             "Total observations"),
+  "Value" = c(
+    format(intro[["rows"]], big.mark = ","),
+    format(intro[["columns"]], big.mark = ","),
+    format(intro[["discrete_columns"]], big.mark = ","),
+    format(intro[["continuous_columns"]], big.mark = ","),
+    format(intro[["all_missing_columns"]], big.mark = ","),
+    format(intro[["total_missing_values"]], big.mark = ","),
+    format(intro[["complete_rows"]], big.mark = ","),
+    format(intro[["total_observations"]], big.mark = ",")
+  )
+)
+kable(intro_df)
 ```
@@
 
-```{r eda-9}
+```{r eda-9 }
 plot_histogram(bands_data)
@@
 
-```{r eda-10}
+```{r eda-10 }
 plot_prcomp(bands_data)
@@
 
-```{r eda-11}
+```{r eda-11 }
 plot_correlation(bands_data)
@@
 
-```{r eda-12}
+```{r eda-12 }
 vi_data_plot <- vi_data %>%
@@
 
-```{r eda-13}
+```{r eda-13 }
 mean_defol <- vi_data_plot %>%
@@
 
-```{r eda-14}
+```{r eda-14 }
 cov_defol <- vi_data_plot %>%
@@
 
-```{r eda-15}
+```{r eda-15 }
 sd_skewness_defol <- vi_data_plot %>%
@@
 
-```{r eda-16}
+```{r eda-16 }
 sd_defol <- vi_data_plot %>%
@@
 
-```{r eda-17}
+```{r eda-17, dev = c("png", "pdf")}
 boxplot_defol <- vi_data_plot %>%
@@
 
-```{r eda-18}
+```{r eda-18 }
 plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
@@
 ```
+
+```{r eda-19, echo = FALSE}
+file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
+          "code/98-paper/journal/")
+```
+
+# Correlation of filter methods {.tabset .tabset-fade}
+
+## VI
+
+```{r eda-20}
+# combine filter values
+
+fv_all = dplyr::bind_rows(fv_vi_car$data, 
+                          fv_vi_info.gain$data, 
+                          fv_vi_gain.ratio$data,
+                          fv_vi_rank$data, 
+                          fv_vi_cor$data, 
+                          fv_vi_mrmr, 
+                          fv_vi_cmim$data, 
+                          fv_vi_var$data) %>%
+  dplyr::select(value, method) %>%
+  mutate(id=rep(1:7381, 7)) %>% 
+  tidyr::spread(method, value) %>% 
+  dplyr::select(-id)
+
+corr = cor(fv_all)
+p.mat = cor_pmat(corr)
+
+ggcorrplot(corr,
+           #p.mat = p.mat,
+           hc.order = TRUE,
+           type = "lower",
+           lab = TRUE)
+```
+
+## NRI
+
+```{r eda-21}
+# combine filter values
+
+fv_all = dplyr::bind_rows(fv_nri_car$data, 
+                          fv_nri_info.gain$data, 
+                          fv_nri_gain.ratio$data,
+                          fv_nri_rank$data, 
+                          fv_nri_cor$data, 
+                          fv_nri_mrmr, 
+                          fv_nri_cmim$data, 
+                          fv_nri_var$data) %>%
+  dplyr::select(value, method) %>%
+  mutate(id=rep(1:7381, 7)) %>% 
+  tidyr::spread(method, value) %>% 
+  dplyr::select(-id)
+
+corr = cor(fv_all)
+p.mat = cor_pmat(corr)
+
+ggcorrplot(corr,
+           #p.mat = p.mat,
+           hc.order = TRUE,
+           type = "lower",
+           lab = TRUE)
+```
+
+## HR
+
+```{r eda-22}
+# combine filter values
+
+fv_all = dplyr::bind_rows(fv_hr_car$data, 
+                          fv_hr_info.gain$data, 
+                          fv_hr_gain.ratio$data,
+                          fv_hr_rank$data, 
+                          fv_hr_cor$data, 
+                          fv_hr_mrmr, 
+                          fv_hr_cmim$data, 
+                          fv_hr_var$data) %>%
+  dplyr::select(value, method) %>%
+  mutate(id=rep(1:7381, 7)) %>% 
+  tidyr::spread(method, value) %>% 
+  dplyr::select(-id)
+
+corr = cor(fv_all)
+p.mat = cor_pmat(corr)
+
+ggcorrplot(corr,
+           #p.mat = p.mat,
+           hc.order = TRUE,
+           type = "lower",
+           lab = TRUE)
+```

### 2019-06-21 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE)

@@
 
-```{r eda-17, dev = c("png", "pdf")}
+```{r defoliation-distribution-plot, dev = c("png", "pdf")}
 boxplot_defol <- vi_data_plot %>%
@@
 ```{r eda-19, echo = FALSE}
-file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
-          "code/98-paper/journal/")
+# file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
+#           "code/98-paper/journal/")
 ```
@@
                           fv_vi_cor$data, 
-                          fv_vi_mrmr, 
+                          fv_vi_mrmr$data, 
                           fv_vi_cmim$data, 
@@
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:7381, 7)) %>% 
+  mutate(id=rep(1:89, 8)) %>% 
   tidyr::spread(method, value) %>% 
+  mutate_if(is.numeric, list(~na_if(., -Inf))) %>% # Inf in mrmr
+  na.omit() %>% # NA in cmim
   dplyr::select(-id)
@@
 corr = cor(fv_all)
-p.mat = cor_pmat(corr)
+#p.mat = cor_pmat(corr)
 
@@
            #p.mat = p.mat,
-           hc.order = TRUE,
+           hc.order = F,
            type = "lower",
@@
                           fv_nri_cor$data, 
-                          fv_nri_mrmr, 
+                          fv_nri_mrmr$data, 
                           fv_nri_cmim$data, 
@@
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:7381, 7)) %>% 
+  mutate(id=rep(1:7381, 8)) %>% 
   tidyr::spread(method, value) %>% 
@@
                           fv_hr_cor$data, 
-                          fv_hr_mrmr, 
+                          fv_hr_mrmr$data, 
                           fv_hr_cmim$data, 
@@
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:7381, 7)) %>% 
+  mutate(id=rep(1:122, 8)) %>% 
   tidyr::spread(method, value) %>% 

### 2019-06-21 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE)

@@
       fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
+      fv_nri_symuncert, fv_nri_relief,
       fv_nri_rank, fv_nri_cor,
@@
       fv_vi_car, fv_vi_info.gain, fv_vi_gain.ratio,
+      fv_vi_symuncert, fv_vi_relief,
       fv_vi_rank, fv_vi_cor,
@@
       fv_hr_car, fv_hr_info.gain, fv_hr_gain.ratio,
+      fv_hr_symuncert, fv_hr_relief,
       fv_hr_rank, fv_hr_cor,
@@
                           fv_vi_gain.ratio$data,
+                          fv_vi_symuncert$data,
+                          fv_vi_relief$data,
                           fv_vi_rank$data, 
@@
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:89, 8)) %>% 
+  mutate(id=rep(1:89, 10)) %>% 
   tidyr::spread(method, value) %>% 
@@
                           fv_nri_gain.ratio$data,
+                          fv_nri_symuncert$data,
+                          fv_nri_relief$data,
                           fv_nri_rank$data, 
@@
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:7381, 8)) %>% 
+  mutate(id=rep(1:7381, 10)) %>% 
   tidyr::spread(method, value) %>% 
@@
            #p.mat = p.mat,
-           hc.order = TRUE,
+           hc.order = FALSE,
            type = "lower",
@@
                           fv_hr_gain.ratio$data,
+                          fv_hr_symuncert$data,
+                          fv_hr_relief$data,
                           fv_hr_rank$data, 
@@
                           fv_hr_cmim$data,
-                          fv_hr_var$data) %>%
+                          fv_hr_var$data
+) %>%
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:122, 8)) %>% 
+  mutate(id=rep(1:122, 10)) %>% 
   tidyr::spread(method, value) %>% 

### 2019-06-23 code/98-paper/journal/manuscript.tex
change title
add list of simple filters

@@
 
-%% arara: indent: {overwrite: yes}
+% arara: indent: {overwrite: yes}
 
@@
 
-	\title{Modeling defoliation as a proxy for tree health: A case study using machine-learning and hyperspectral remote sensing data}
-
+	\title{Hyperspectral tree defoliation modeling with a focus on feature-selection and model interpretation}
 	%% Group authors per affiliation:
@@
 
+\paragraph{Ensemble filter methods}
+
 % Ensemble filter methods
 Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \citep{abeel2010, drotar2017a}.
-This aligns with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models \citep{polikar2012, feurer2015}.
+This approach is referred to as "ensemble filter" \citep{dietterich2000}.
+Ensemble filters align with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhence predictive performance \citep{polikar2012, feurer2015}.
 In this work the "Borda" ensemble filter was applied \citep{drotar2017a}.
-Here, the ranking is based on the sum of all single filters.
+The final order is the sum of all single filters ranking.
 
-Most filter methods are correlation or entropy based.
-A list of all filter methods used in this work along with their respective formulas can be found in Appendix X.
-% TODO: Link appendix
-% TODO: List filter methods
+\paragraph{Ensuring a fair weighting in the ensemble}
+Filter methods can be grouped into classes: Correlation based, entropy based, linear and non-linear methods.
+It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased.
+This was taken care of by comparing the linear correlations of all generated filter values for each method against each other.
+In case filters showed high correlations with each other, these were not included into the ensemble filter.
+The goal was to create an ensemble filter showing low to medium correlations with each other.
+
+\paragraph{Description of used filter methods}
+
+Filter methods can be classified as follows:
+
+\begin{itemize}
+	\item univariate/multivariate (scoring based on a single variable / multiple variables)
+	\item linear/non-linear (calculation of linear/non-linear interaction terms)
+	\item entropy/correlation (scoring based on derivations of entropy or correlation based approaches)
+\end{itemize}
+
+% filter methods
+\begin{table}[b!]
+\centering
+\caption[t]{List of filter methods used in this work}
+\begingroup\footnotesize
+\begin{tabular}{lll}
+	\\
+	Name                                         & Group                             & Reference          \\
+	\hline
+	Linear correlation (Pearson)                 & univariate, linear, correlation   & \cite{pearson1901} \\
+	Information gain                             & univariate, non-linear, entropy   & \cite{quinlan1986} \\
+	Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy & \cite{zhao2013}    \\
+	Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
+	Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
+	Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
+\end{tabular}
+\endgroup
+\label{tab:filter-methods}
+\end{table}
 
 % I grouped these two since we focus on filter methods in this study - or is this already a bit biased towards filter methods?
-\subsubsection{Other approaches}
+\subsubsection{Wrapper methods and PCA}
 

### 2019-06-23 code/98-paper/journal/manuscript.tex
update wrapper methods

@@
 	\acro{PISR}{Potential Incoming Solar Radiation}
+	\acro{PCA}{principal component analysis}
 	\acro{PLS}{partial least-squares}
@@
 
+\noindent Other approaches to assess feature importance are so called "wrapper methods" and the \ac{PCA} \citep{das2001, jolliffe2016}.
+A short introduction to both is given below.
+
 % Wrapper approach
-\noindent Another approach to reduce the number of variables is to apply algorithms that are used for hyperparameter optimization such as "Random Search" or a "Generic Simulated Annealing".
+\noindent "Wrapper methods" \citep{chandrashekar2014, kohavi1997} apply algorithms that are used for hyperparameter optimization such as "Random Search" or "Generic Simulated Annealing".
 First a (random) subset of features is chosen based on the selected algorithm.
@@
 Now the model is fitted on the data and the performance is evaluated.
+This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
 A disadvantage of this approach is that hyperparameter tuning can only start after the feature-selection optimization has ended.
-Hence, this method is very expensive because two optimization stages need to be run.
-In science this approach is also referred to as "wrapper method" \citep{chandrashekar2014, kohavi1997}.
-It was not considered in this work due to the expensive runtime.
+Hence, this method is very expensive because two optimization stages need to be run in sequential order.
+"Wrapper methods" were not considered in this work due to the expensive runtime.
 
 % PCA
-A third commonly used method is the "Principal Component Analysis" \citep{pearson1901, jolliffe2016}.
+A method that is very distinct from filters and wrappers is the "Principal Component Analysis" \citep{pearson1901, jolliffe2016}.
 Here, the main components of the feature space are extracted and combined.

### 2019-06-23 code/98-paper/journal/manuscript.tex
restructure methods, list benchmark matrix

@@
 
-\subsection{Benchmarking design}
-
-\noindent The benchmarking matrix of this study consists of the following algorithms:
-
-\begin{itemize}
-	\item  Extreme Gradient Boosting (XGBOOST)
-	\item  Random Forest (RF)
-	\item  Penalized Regression (both L1 and L2)
-	\item  Support Vector Machine (SVM)
-\end{itemize}
-
-\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
-Extreme Gradient Boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
-Penalized Regression is a statistical modeling technique that is capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \citep{hastie2001}.
-The penalties are also known as "lasso" (L1) and "ridge" (L2).
-The former does not allow the removal of variables from the model (penalization to zero) while the latter does.
-Both penalties can also be combined.
-The combined approach is called "elastic net" but was not used in this work.
-
-\subsubsection{Feature sets}
-% TODO: talk about the reasons behind different features sets: environmental implications
-
-\noindent Three feature sets were used in this study with each representing a different way of feature-engineering:
-
-\begin{itemize}
-	\item The raw hyperspectal band information (HR): No feature engineering
-	\item Vegetation Indices (\ac{VI}: Expert-based feature engineering
-	\item Normalized Ratio Indices (\ac{NRI}): Automated feature-engineering
-\end{itemize}
-
-\subsubsection{Hyperparameter Optimization}
-
-\noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
-Maximum Bayesian Optimization \citep{mlrmbo} was used for parameter optimization.
-This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
-After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
-The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
-Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
-This strategy continues until a termination criterion, defined by the user, is reached \citep{hutter2011, jones1998}.
-An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum
-This applies when being compared to methods that do not use information from previous runs, such as random search or grid search \citep{bergstra2012}.
-
-\subsubsection{Spatial resampling}
-
-\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible.
-Each plot served as one fold within the resampling setup, resulting in four folds total.
-For the inner level (hyperparameter tuning), $p - 1$ folds were used (with $p$ being the number of plots). REF SCHRATZ
-
 \subsection{Dimension reduction}
@@
 
+\subsection{Benchmarking design}
+
+\subsubsection{Algorithms}
+
+\noindent The benchmarking matrix of this study consists of the following algorithms:
+
+\begin{itemize}
+	\item  Extreme Gradient Boosting (XGBOOST)
+	\item  Random Forest (RF)
+	\item  Penalized Regression (both L1 and L2)
+	\item  Support Vector Machine (SVM)
+\end{itemize}
+
+\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
+Extreme Gradient Boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
+Penalized Regression is a statistical modeling technique that is capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \citep{hastie2001}.
+The penalties are also known as "lasso" (L1) and "ridge" (L2).
+The former does not allow the removal of variables from the model (penalization to zero) while the latter does.
+Both penalties can also be combined.
+The combined approach is called "elastic net" but was not used in this work.
+
+\subsubsection{Feature sets}
+% TODO: talk about the reasons behind different features sets: environmental implications
+
+\noindent Three feature sets were used in this study with each representing a different way of feature-engineering:
+
+\begin{itemize}
+	\item The raw hyperspectal band information (HR: No feature engineering)
+	\item Vegetation Indices (\ac{VI}: Expert-based feature engineering)
+	\item Normalized Ratio Indices (\ac{NRI}: Automated feature-engineering)
+\end{itemize}
+
+\subsubsection{Hyperparameter Optimization}
+
+\noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
+Maximum Bayesian Optimization \citep{mlrmbo} was used for parameter optimization.
+This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
+The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
+Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
+This strategy continues until a termination criterion, defined by the user, is reached \citep{hutter2011, jones1998}.
+An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
+The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum
+This applies when being compared to methods that do not use information from previous runs, such as random search or grid search \citep{bergstra2012}.
+
+For the filter methods, the percentage of features was optimized as a hyperparameter.
+The number of main components was tuned for the \ac{PCA} based benchmarking experiments.
+
+\subsubsection{Spatial resampling}
+
+\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible \citep{schratz2019, sperrorest}.
+Each plot served as one fold within the resampling setup, resulting in four folds total.
+For the inner level (hyperparameter tuning), $p - 1$ folds were used (with $p$ being the number of plots).
+
+\subsubsection{Summary}
+
+\noindent In total the benchmarking matrix consisted of 120 experiments (3 feature sets, 5 algorithms, 8 feature-selection methods).
+
+\subsection{Research compendium}
+
 \noindent The complete study was done using the open-source statistical programming language R \citep{rcoreteam2018}.
-The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chen2016} (\textit{xgboost}), \textit{e1071} \citep{e1071} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
+The algorithm implementations of the following packages have been used: \textit{xgboost} \citep{chen2016} (\textit{xgboost}), \textit{kernlab} \citep{kernlab} (Support Vector Machine) and \textit{glmnet} \citep{glmnet} (Ridge Regression).
 The filter methods of the following packages were used: \textit{praznik \citep{praznik}}, \textit{FSelectorRcpp} \citep{fselectorrcpp}.

### 2019-06-24 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE)

@@
 # load drake objects
-loadd(vi_data, nri_data, bands_data, coords_vi_nri_clean,
+loadd(
+  vi_data, nri_data, bands_data, coords_vi_nri_clean,
 
@@
 intro_df <- data.frame(
-  "Name" = c("Rows", "Columns",
+  "Name" = c(
+    "Rows", "Columns",
     "Discrete columns", "Continuous columns", "All missing columns",
     "Missing observations", "Complete Rows",
-             "Total observations"),
+    "Total observations"
+  ),
   "Value" = c(
@@
     allpairs = TRUE,
-                                         lonlat = FALSE)))
+    lonlat = FALSE
+  )))
 })
@@
 
-fv_all = dplyr::bind_rows(fv_vi_car$data, 
+fv_all <- dplyr::bind_rows(
+  fv_vi_car$data,
   fv_vi_info.gain$data,
@@
   fv_vi_cmim$data,
-                          fv_vi_var$data) %>%
+  fv_vi_var$data
+) %>%
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:89, 10)) %>% 
+  mutate(id = rep(1:89, 10)) %>%
   tidyr::spread(method, value) %>%
-  mutate_if(is.numeric, list(~na_if(., -Inf))) %>% # Inf in mrmr
+  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
   na.omit() %>% # NA in cmim
@@
 
-corr = cor(fv_all)
-#p.mat = cor_pmat(corr)
+corr <- cor(fv_all, method = "spearman")
+# p.mat = cor_pmat(corr)
 
 ggcorrplot(corr,
-           #p.mat = p.mat,
+  # p.mat = p.mat,
   hc.order = F,
   type = "lower",
-           lab = TRUE)
+  lab = TRUE
+)
 ```
@@
 
-fv_all = dplyr::bind_rows(fv_nri_car$data, 
+fv_all <- dplyr::bind_rows(
+  fv_nri_car$data,
   fv_nri_info.gain$data,
@@
   fv_nri_cmim$data,
-                          fv_nri_var$data) %>%
+  fv_nri_var$data
+) %>%
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:7381, 10)) %>% 
+  mutate(id = rep(1:7381, 10)) %>%
   tidyr::spread(method, value) %>%
@@
 
-corr = cor(fv_all)
-p.mat = cor_pmat(corr)
+corr <- cor(fv_all, method = "spearman")
+p.mat <- cor_pmat(corr)
 
 ggcorrplot(corr,
-           #p.mat = p.mat,
+  # p.mat = p.mat,
   hc.order = FALSE,
   type = "lower",
-           lab = TRUE)
+  lab = TRUE
+)
 ```
@@
 
-fv_all = dplyr::bind_rows(fv_hr_car$data, 
+fv_all <- dplyr::bind_rows(
+  fv_hr_car$data,
   fv_hr_info.gain$data,
@@
   dplyr::select(value, method) %>%
-  mutate(id=rep(1:122, 10)) %>% 
+  mutate(id = rep(1:122, 10)) %>%
   tidyr::spread(method, value) %>%
@@
 
-corr = cor(fv_all)
-p.mat = cor_pmat(corr)
+corr <- cor(fv_all, method = "spearman")
+p.mat <- cor_pmat(corr)
 
 ggcorrplot(corr,
-           #p.mat = p.mat,
+  # p.mat = p.mat,
   hc.order = F,
   type = "lower",
-           lab = TRUE)
+  lab = TRUE
+)
 ```

### 2019-06-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE)

@@
   fv_hr_mrmr,
-  fv_hr_cmim, fv_hr_var
+  fv_hr_cmim, fv_hr_var,
+
+  fv_hr_info.gain5, fv_hr_info.gain10,
+  fv_hr_info.gain15, fv_hr_info.gain20,
+  fv_hr_info.gain25, fv_hr_info.gain30,
+
+  fv_vi_info.gain5, fv_vi_info.gain10,
+  fv_vi_info.gain15, fv_vi_info.gain20,
+  fv_vi_info.gain25, fv_vi_info.gain30,
+
+  fv_nri_info.gain5, fv_nri_info.gain10,
+  fv_nri_info.gain15, fv_nri_info.gain20,
+  fv_nri_info.gain25, fv_nri_info.gain30
 )
@@
 
+### Filter methods amongst each other
+
 ```{r eda-20}
@@
 # p.mat = cor_pmat(corr)
+
+ggcorrplot(corr,
+           # p.mat = p.mat,
+           hc.order = F,
+           type = "lower",
+           lab = TRUE
+)
+```
+
+### Number of bins of `FSelectorRcpp::information.gain()`
+
+```{r}
+fv_all <- dplyr::bind_rows(
+  fv_vi_info.gain5$data, fv_vi_info.gain10$data,
+  fv_vi_info.gain15$data, fv_vi_info.gain20$data,
+  fv_vi_info.gain25$data, fv_vi_info.gain30$data
+)  %>%
+  dplyr::select(value, method) %>%
+  mutate(id = rep(1:89, 10)) %>%
+  tidyr::spread(method, value) %>%
+  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
+  na.omit() %>% 
+  dplyr::select(-id)
+
+corr <- cor(fv_all, method = "spearman")
 

### 2019-06-26 analysis/filter-correlation.Rmd
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE)

--- 2019-06-26 analysis/eda.Rmd
@@
 ---
-title: "Exploratory data analysis"
+title: "Correlation of filter methods"
 output: 
@@
 loadd(
-  vi_data, nri_data, bands_data, coords_vi_nri_clean,
-
   fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
@@
 )
-library("DataExplorer")
-library("dplyr")
-library("ggsci")
-library("ggplot2")
-```
-
-# Datasets {.tabset .tabset-fade}
-
-## VI
-
-### Overview
-
-```{r eda-1 }
-intro <- introduce(vi_data)
-intro_df <- data.frame(
-  "Name" = c(
-    "Rows", "Columns",
-    "Discrete columns", "Continuous columns", "All missing columns",
-    "Missing observations", "Complete Rows",
-    "Total observations"
-  ),
-  "Value" = c(
-    format(intro[["rows"]], big.mark = ","),
-    format(intro[["columns"]], big.mark = ","),
-    format(intro[["discrete_columns"]], big.mark = ","),
-    format(intro[["continuous_columns"]], big.mark = ","),
-    format(intro[["all_missing_columns"]], big.mark = ","),
-    format(intro[["total_missing_values"]], big.mark = ","),
-    format(intro[["complete_rows"]], big.mark = ","),
-    format(intro[["total_observations"]], big.mark = ",")
-  )
-)
-knitr::kable(intro_df)
-```
-
-### Histograms
-
-```{r eda-2 }
-plot_histogram(vi_data)
-```
-
-### PCA
-
-```{r eda-3 }
-plot_prcomp(vi_data)
-```
-
-### Corr
-
-```{r eda-4 }
-plot_correlation(vi_data)
-```
-
-## NRI
-
-### Overview
-
-```{r eda-5 }
-intro <- introduce(nri_data)
-intro_df <- data.frame(
-  "Name" = c(
-    "Rows", "Columns",
-    "Discrete columns", "Continuous columns", "All missing columns",
-    "Missing observations", "Complete Rows",
-    "Total observations"
-  ),
-  "Value" = c(
-    format(intro[["rows"]], big.mark = ","),
-    format(intro[["columns"]], big.mark = ","),
-    format(intro[["discrete_columns"]], big.mark = ","),
-    format(intro[["continuous_columns"]], big.mark = ","),
-    format(intro[["all_missing_columns"]], big.mark = ","),
-    format(intro[["total_missing_values"]], big.mark = ","),
-    format(intro[["complete_rows"]], big.mark = ","),
-    format(intro[["total_observations"]], big.mark = ",")
-  )
-)
-kable(intro_df)
-```
-
-### Histograms
-
-No histograms for NRI -> too many features.
-
-### PCA
-
-```{r eda-6 }
-plot_prcomp(nri_data)
-```
-
-### Corr
-
-```{r eda-7 }
-plot_correlation(nri_data)
-```
-
-## HR
-
-### Overview
-
-```{r eda-8 }
-intro <- introduce(bands_data)
-intro_df <- data.frame(
-  "Name" = c(
-    "Rows", "Columns",
-    "Discrete columns", "Continuous columns", "All missing columns",
-    "Missing observations", "Complete Rows",
-    "Total observations"
-  ),
-  "Value" = c(
-    format(intro[["rows"]], big.mark = ","),
-    format(intro[["columns"]], big.mark = ","),
-    format(intro[["discrete_columns"]], big.mark = ","),
-    format(intro[["continuous_columns"]], big.mark = ","),
-    format(intro[["all_missing_columns"]], big.mark = ","),
-    format(intro[["total_missing_values"]], big.mark = ","),
-    format(intro[["complete_rows"]], big.mark = ","),
-    format(intro[["total_observations"]], big.mark = ",")
-  )
-)
-kable(intro_df)
-```
-
-### Histograms
-
-```{r eda-9 }
-plot_histogram(bands_data)
-```
-
-### PCA
-
-```{r eda-10 }
-plot_prcomp(bands_data)
-```
-
-### Corr
-
-```{r eda-11 }
-plot_correlation(bands_data)
-```
-
-# Custom plots
-
-```{r eda-12 }
-vi_data_plot <- vi_data %>%
-  mutate(plot = factor(rep(
-    c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-    c(479, 451, 300, 529)
-  )))
-```
-
-# Mean defoliation per plot
-
-```{r eda-13 }
-mean_defol <- vi_data_plot %>%
-  group_by(plot) %>%
-  summarise(mean(defoliation)) %>%
-  pull(.)
-vi_data_plot %>%
-  group_by(plot) %>%
-  summarise(mean(defoliation))
-```
-
-# Coefficient of variation
-
-```{r eda-14 }
-cov_defol <- vi_data_plot %>%
-  group_by(plot) %>%
-  summarise((sd(defoliation) / mean(defoliation)) * 100) %>%
-  pull(.)
-vi_data_plot %>%
-  group_by(plot) %>%
-  summarise((sd(defoliation) / mean(defoliation)) * 100)
-```
-
-# sd / skewness
-
-```{r eda-15 }
-sd_skewness_defol <- vi_data_plot %>%
-  group_by(plot) %>%
-  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation)) %>%
-  pull(.)
-vi_data_plot %>%
-  group_by(plot) %>%
-  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation))
-```
-
-```{r eda-16 }
-sd_defol <- vi_data_plot %>%
-  group_by(plot) %>%
-  summarise(sd(defoliation)) %>%
-  pull(.)
-vi_data_plot %>%
-  group_by(plot) %>%
-  summarise(sd(defoliation))
-```
-
-```{r defoliation-distribution-plot, dev = c("png", "pdf")}
-boxplot_defol <- vi_data_plot %>%
-  group_by(plot) %>%
-  ggboxplot(
-    x = "plot", y = "defoliation", color = "plot",
-    add = "jitter", add.params = list(size = "defoliation")
-  ) +
-  scale_size(range = c(0.5, 0.5)) +
-  annotate("text",
-           label = expression(bold(atop("n = 479", bar(x) ~ "= 55.96"))), x = 1,
-           y = 112, size = 3, colour = "#BC3C29", fontface = 2
-  ) +
-  annotate("text",
-           label = expression(bold(atop("n = 451", bar(x) ~ "= 13.54"))), x = 2,
-           y = 112, size = 3, colour = "#0072B5", fontface = 2
-  ) +
-  annotate("text",
-           label = expression(bold(atop("n = 300", bar(x) ~ "= 68.44"))), x = 3,
-           y = 112, size = 3, colour = "#E18727", fontface = 2
-  ) +
-  annotate("text",
-           label = expression(bold(atop("n = 529", bar(x) ~ "= 69.22"))), x = 4,
-           y = 112, size = 3, colour = "#20854E", fontface = 2
-  ) +
-  scale_color_nejm() +
-  theme(legend.position = "none") +
-  labs(y = "Total defoliation per tree (%)", x = "Plot")
-boxplot_defol
-```
-
-# Point density
-
-In Meters.
-
-```{r eda-18 }
-plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
-dist <- map(plots, ~ {
-  coords <- coords_vi_nri_clean %>%
-    mutate(plot = factor(rep(
-      c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-      c(479, 451, 300, 529)
-    ))) %>%
-    filter(plot == .x) %>%
-    dplyr::select(-plot)
-  points <- SpatialPoints(
-    coords = coords,
-    proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
-  )
-  distance <- mean(as.dist(pointDistance(points,
-                                         allpairs = TRUE,
-                                         lonlat = FALSE
-  )))
-})
-set_names(dist, plots)
-dist_plots <- unlist(dist)
-```
-
-```{r eda-19, echo = FALSE}
-# file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
-#           "code/98-paper/journal/")
 ```
@@
 # Correlation of filter methods {.tabset .tabset-fade}
+
+Spearman's rank correlation is used since we compare rankings against each other.
 
@@
 ```{r}
+# recode levels so that they are unique
+fv_all_list = map2(list(fv_vi_info.gain5$data, fv_vi_info.gain10$data,
+  fv_vi_info.gain15$data, fv_vi_info.gain20$data,
+  fv_vi_info.gain25$data, fv_vi_info.gain30$data),
+  list(5, 10, 15, 20, 25, 30), ~
+    dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}"))))
+
 fv_all <- dplyr::bind_rows(
-  fv_vi_info.gain5$data, fv_vi_info.gain10$data,
-  fv_vi_info.gain15$data, fv_vi_info.gain20$data,
-  fv_vi_info.gain25$data, fv_vi_info.gain30$data
+  fv_all_list
 )  %>%
   dplyr::select(value, method) %>%
-  mutate(id = rep(1:89, 10)) %>%
+  mutate(id = rep(1:89, 6)) %>%
   tidyr::spread(method, value) %>%

### 2019-06-27 analysis/eda.Rmd
add new report

@@
 library("ggplot2")
+library("rgdal")
 ```

### 2019-06-27 code/07-reports.R

@@ -3,6 +3,8 @@ reports_plan_paper = drake_plan(
   eda_wfr = wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE),
 
   spectral_signatures_wfr = wflow_publish(knitr_in("analysis/spectral-signatures.Rmd"), view = FALSE),
+
+  filter_correlations_wfr = wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE)
 )
 
 reports_plan_project = drake_plan(

### 2019-06-27 analysis/filter-correlation.Rmd
add new report
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE)

@@
 )
+library("rgdal")
 ```
@@
 
-Spearman's rank correlation is used since we compare rankings against each other.
+Spearman's rank correlation is used because we compare rankings against each other.
 
@@
 
-```{r eda-20}
+```{r eda-20, warning=FALSE}
 # combine filter values

### 2019-06-27 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE)

@@
 ```
-
-# Correlation of filter methods {.tabset .tabset-fade}
-
-## VI
-
-### Filter methods amongst each other
-
-```{r eda-20}
-# combine filter values
-
-fv_all <- dplyr::bind_rows(
-  fv_vi_car$data,
-  fv_vi_info.gain$data,
-  fv_vi_gain.ratio$data,
-  fv_vi_symuncert$data,
-  fv_vi_relief$data,
-  fv_vi_rank$data,
-  fv_vi_cor$data,
-  fv_vi_mrmr$data,
-  fv_vi_cmim$data,
-  fv_vi_var$data
-) %>%
-  dplyr::select(value, method) %>%
-  mutate(id = rep(1:89, 10)) %>%
-  tidyr::spread(method, value) %>%
-  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
-  na.omit() %>% # NA in cmim
-  dplyr::select(-id)
-
-corr <- cor(fv_all, method = "spearman")
-# p.mat = cor_pmat(corr)
-
-ggcorrplot(corr,
-           # p.mat = p.mat,
-           hc.order = F,
-           type = "lower",
-           lab = TRUE
-)
-```
-
-### Number of bins of `FSelectorRcpp::information.gain()`
-
-```{r}
-fv_all <- dplyr::bind_rows(
-  fv_vi_info.gain5$data, fv_vi_info.gain10$data,
-  fv_vi_info.gain15$data, fv_vi_info.gain20$data,
-  fv_vi_info.gain25$data, fv_vi_info.gain30$data
-)  %>%
-  dplyr::select(value, method) %>%
-  mutate(id = rep(1:89, 10)) %>%
-  tidyr::spread(method, value) %>%
-  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
-  na.omit() %>% 
-  dplyr::select(-id)
-
-corr <- cor(fv_all, method = "spearman")
-
-ggcorrplot(corr,
-           # p.mat = p.mat,
-           hc.order = F,
-           type = "lower",
-           lab = TRUE
-)
-```
-
-## NRI
-
-```{r eda-21}
-# combine filter values
-
-fv_all <- dplyr::bind_rows(
-  fv_nri_car$data,
-  fv_nri_info.gain$data,
-  fv_nri_gain.ratio$data,
-  fv_nri_symuncert$data,
-  fv_nri_relief$data,
-  fv_nri_rank$data,
-  fv_nri_cor$data,
-  fv_nri_mrmr$data,
-  fv_nri_cmim$data,
-  fv_nri_var$data
-) %>%
-  dplyr::select(value, method) %>%
-  mutate(id = rep(1:7381, 10)) %>%
-  tidyr::spread(method, value) %>%
-  dplyr::select(-id)
-
-corr <- cor(fv_all, method = "spearman")
-p.mat <- cor_pmat(corr)
-
-ggcorrplot(corr,
-           # p.mat = p.mat,
-           hc.order = FALSE,
-           type = "lower",
-           lab = TRUE
-)
-```
-
-## HR
-
-```{r eda-22}
-# combine filter values
-
-fv_all <- dplyr::bind_rows(
-  fv_hr_car$data,
-  fv_hr_info.gain$data,
-  fv_hr_gain.ratio$data,
-  fv_hr_symuncert$data,
-  fv_hr_relief$data,
-  fv_hr_rank$data,
-  fv_hr_cor$data,
-  fv_hr_mrmr$data,
-  fv_hr_cmim$data,
-  fv_hr_var$data
-) %>%
-  dplyr::select(value, method) %>%
-  mutate(id = rep(1:122, 10)) %>%
-  tidyr::spread(method, value) %>%
-  dplyr::select(-id)
-
-corr <- cor(fv_all, method = "spearman")
-p.mat <- cor_pmat(corr)
-
-ggcorrplot(corr,
-           # p.mat = p.mat,
-           hc.order = F,
-           type = "lower",
-           lab = TRUE
-)
-```

### 2019-06-28 code/98-paper/journal/manuscript.tex
add todos
update methods sections

@@
 
+% TODO: Mention the use of mtry.pow
+% TODO: Explain the binning in the FSelectorfilters
+% TODO: Explain that we used Spearman cor in the corrplot
+
 \subsection{Derivation of indices}
@@
 It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased.
-This was taken care of by comparing the linear correlations of all generated filter values for each method against each other.
+This was taken care of by calculating the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
 In case filters showed high correlations with each other, these were not included into the ensemble filter.
-The goal was to create an ensemble filter showing low to medium correlations with each other.
+This ensures that the ensembe filter is not biased towards a certain group of methods yielding highly similar rankings.
 
@@
 % I grouped these two since we focus on filter methods in this study - or is this already a bit biased towards filter methods?
+
+\noindent Filter method "information gain" is by creation only defined for class responses:
+
+\begin{equation}
+	H(Class) + H(Attribute) - H(Class, Attribute)
+\end{equation}
+
+where $H$ is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
+To be able to use this method with a numeric response, the variable is discretized into equal bins and treated as a class variable.
+While the number of bins can be treated as a hyperparameter of the filter method, the default of the R package implementation of \texttt{$n_{bin}$ = 5} was used after rank correlations of > 0.9 for different bin sizes have been explored.
+
 \subsubsection{Wrapper methods and PCA}
@@
 Interpretation is not possible anymore.
+
+\subsubsection{Calculation of the model's feature importance}
+
+\noindent A permutation-based feature importance estimation was done for the best algorithm of each feature set and filter method, respectively.
+This method caluclates the loss of predictive performance for each feature by permuting it in a random manner.
+The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
+
+\subsubsection{Linking feature importance to wavelength regions}
+\noindent For ecological interpretation purposes we linked the ten most important indices of the winning models for each feature set to the spectral regions of the hyperspectral data.
+For feature set HR and NRI a direct linking to the respective bands of the hyperspectral sensor was done.
+For the vegetation indices all bands covered by the spectral range of calculated vegetation indices were counted and summed up.
 
@@
 
-For the filter methods, the percentage of features was optimized as a hyperparameter.
-The number of main components was tuned for the \ac{PCA} based benchmarking experiments.
+For the filter methods, the percentage of features was added to the models as a hyperparameter.
+The number of main components was tuned for the instances which used \ac{PCA} instead of filtering.
+Random Forest hyperparameter \texttt{$m_{try}$} was transformed from taking absolute values to a relative parameter depending on the number of tasks: $p^{m_{try}}$, where $p$ is the number of features.
+This was necessary to ensure that \texttt{$m_{try}$} was not chosen out of bounds during tuning:
+After the filtering, only a subset of features of the task is used for optimizing the models hyperparameters.
+The size of this subset is always different since the number of features is also optimized.
+Hence, the value of  \texttt{$m_{try}$} needs to be created flexible based on the number features.
 

### 2019-06-28 analysis/eval-performance.Rmd
add eval-performance report

---
title: "Evaluation of performances"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

source(here::here("code/99-packages.R"))
sourceDirectory("R")

# load drake objects
loadd(
  bm_aggregated
)
library("mlr")
library("magrittr")
```

```{r eval-performance-1 }
bm_aggregated %>%
  getBMRAggrPerformances(as.df = TRUE) %>%
  arrange(task.id, desc(rmse.test.mean))
```

### Boxplot comparison

```{r performance-boxplot, dev = c("png", "pdf")}
plt <- plotBMRBoxplots(bm_aggregated,
  measure = brier, pretty.names = FALSE,
  # order.lrn = c(
  #   "classif.gam.tuned", 
  #   "classif.kknn.tuned",
  #   "classif.ranger.tuned",
  #   "classif.svm.tuned", 
  #   "classif.gbm.tuned",
  #   "classif.xgboost.tuned"
  # )
) +
  aes(color = learner.id) +
  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
  scale_colour_nejm() +
  labs(y = "Brier score", x = "") +
  theme_pubr() +
  theme(strip.text.x = element_text(size = 8),
        axis.text.x = element_text(size = 7)
        )

#levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt
```

### Aggregated performances

```{r performance-aggregated, dev = c("png", "pdf")}
plt2 <- plotBMRSummary(bm_aggregated, pretty.names = FALSE) +
  scale_colour_nejm() +
  theme_pubr()

#levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")

plt2
```

### 2019-06-28 analysis/filter-correlation.Rmd
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE)
export filter-correlation images

@@
 
-```{r eda-20, warning=FALSE}
+```{r correlation-filter-vi, warning=FALSE, dev = c("png", "pdf")}
 # combine filter values
@@
 
-```{r}
+```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
 # recode levels so that they are unique
@@
 
-```{r eda-21}
+```{r filter-correlation-3, warning=FALSE}
 # combine filter values
@@
 
-```{r eda-22}
+```{r filter-correlation-4, warning=FALSE}
 # combine filter values
@@
 ```
+
+```{r filter-correlation-5, echo = FALSE}
+file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
+          "code/98-paper/journal/")
+file_move("docs/figure/filter-correlation.Rmd/correlation-filter-vi-1.pdf",
+          "code/98-paper/journal/")
+```

### 2019-06-28 R/06-mlr-paper.R
misc

@@ -9,7 +9,7 @@ benchmark_custom_no_models <- function(learner, task) {
 
   parallelStart(
     mode = "multicore",
-    #level = "mlr.resample",
+    level = "mlr.resample",
     cpus = 4
   )
   set.seed(12345, kind = "L'Ecuyer-CMRG")

### 2019-06-28 code/07-reports.R

@@ -4,7 +4,9 @@ reports_plan_paper = drake_plan(
 
   spectral_signatures_wfr = wflow_publish(knitr_in("analysis/spectral-signatures.Rmd"), view = FALSE),
 
-  filter_correlations_wfr = wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE)
+  filter_correlations_wfr = wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE),
+
+  eval_performance = wflow_publish(knitr_in("analysis/eval_performance.Rmd"), view = FALSE)
 )
 
 reports_plan_project = drake_plan(

### 2019-06-28 code/98-paper/journal/manuscript.tex
more methods

@@
 
-% TODO: Mention the use of mtry.pow
-% TODO: Explain the binning in the FSelectorfilters
 % TODO: Explain that we used Spearman cor in the corrplot
@@
 
-\subsubsection{Calculation of the model's feature importance}
-
-\noindent A permutation-based feature importance estimation was done for the best algorithm of each feature set and filter method, respectively.
-This method caluclates the loss of predictive performance for each feature by permuting it in a random manner.
-The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
-
-\subsubsection{Linking feature importance to wavelength regions}
-\noindent For ecological interpretation purposes we linked the ten most important indices of the winning models for each feature set to the spectral regions of the hyperspectral data.
-For feature set HR and NRI a direct linking to the respective bands of the hyperspectral sensor was done.
-For the vegetation indices all bands covered by the spectral range of calculated vegetation indices were counted and summed up.
-
 \subsection{Benchmarking design}
@@
 
-\subsubsection{Summary}
+In total the benchmarking matrix consisted of 120 experiments (3 feature sets, 5 algorithms, 8 feature-selection methods).
 
-\noindent  In total the benchmarking matrix consisted of 120 experiments (3 feature sets, 5 algorithms, 8 feature-selection methods).
+\subsection{Calculation of the model's feature importance}
+\noindent A permutation-based feature importance estimation was done for the best algorithm of each feature set and filter method, respectively.
+This method caluclates the loss of predictive performance for each feature by permuting it in a random manner.
+The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
+
+\subsection{Linking feature importance to wavelength regions}
+\noindent For ecological interpretation purposes we linked the ten most important indices of the winning models for each feature set to the spectral regions of the hyperspectral data.
+For feature set HR and NRI a direct linking to the respective bands of the hyperspectral sensor was done.
+For the vegetation indices all bands covered by the spectral range of calculated vegetation indices were counted and summed up.
 

### 2019-06-30 code/98-paper/journal/manuscript.tex
small optimizations

@@
 
-% TODO: Explain that we used Spearman cor in the corrplot
-
 \subsection{Derivation of indices}
@@
 
-\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), a buffer of two meters around the centroid of the respective tree was used.
-The mean of all pixels touched by this buffer was calculated and assigned to the observations.
+\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), a buffer of two meters around the centroid of the respective tree was drawn.
+The value assigned to a tree observations was the mean of all pixels which touched the respective buffer around the tree.
 In total, $\frac{125*126}{2} = 7875$ NRIs were calculated .
-Due to four corrupted bands and numerical calculation errors for some band combinations, a total of 7471 indices were available for each observation.
-All indices which contained missing values for at least one plot were removed from the dataset.
-Observations were given more importance than indices due to the mass of available \ac{NRI} indices.
+Due to four corrupted bands of the sensor a total of 7471 indices were available for each observation.
 
@@
 The former refers to the general idea of reducing features from a dataset \citep{vandermaaten2007}.
-In modeling this means finding the best subset of covariates which provide the most predictive power to the model.
-The latter term in contrast is the definition of a dataset attribute:
-When $p > n,$ where $p$ is the number of covariates and $n$ the number of observations \citep{hastie2001}, the dataset qualifies as "high-dimensional".
+In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
+"High-dimensionality" in contrast is defined as a dataset attribute and applies when $p > n,$ where $p$ is the number of covariates and $n$ the number of observations \citep{hastie2001}.
 There is no absolute value at which the term can be applied, both for $n$ and $p$.
@@
 \begin{itemize}
-	\item Fitting times of the model increase
-	\item Noise is possibly introduced into the model by highly-correlated variables \citep{johnstoneiainm.2009}
-	      % I think we can phrase this better
-	\item Interpretation and prediction becomes more complex, up to the point at which it is not possible anymore \citep{johnstoneiainm.2009}
+  \item Model fitting times increase
+  \item Noise is possibly introduced into models by highly-correlated variables \citep{johnstoneiainm.2009}
+  \item Model interpretation and prediction tasks become more complicated \citep{johnstoneiainm.2009}
 \end{itemize}
 
-\noindent In the last decade the amount of environmental data increased a lot, paramount by the increased data coverage of satellites.
-With this the need of applying automated feature-selection methods emerged.
-In the following sections the currently most used feature selection approaches are briefly introduced.
+\noindent In the following sections a brief overview about sub-categories of feature-selection approaches is given.
+Due to the focus of this study on the use of filter methods, all other appraoches were grouped into a single section.
 
@@
 \paragraph{Ensuring a fair weighting in the ensemble}
+
 Filter methods can be grouped into classes: Correlation based, entropy based, linear and non-linear methods.
@@
 
-% I grouped these two since we focus on filter methods in this study - or is this already a bit biased towards filter methods?
-
 \noindent Filter method "information gain" is by creation only defined for class responses:
@@
 While the number of bins can be treated as a hyperparameter of the filter method, the default of the R package implementation of \texttt{$n_{bin}$ = 5} was used after rank correlations of > 0.9 for different bin sizes have been explored.
+% TODO: Add reference to appendix, put plot into appendix
 
@@
 
-\noindent Other approaches to assess feature importance are so called "wrapper methods" and the \ac{PCA} \citep{das2001, jolliffe2016}.
+\noindent Other approaches to assess feature importance are "wrapper methods" and the \ac{PCA} \citep{das2001, jolliffe2016}.
 A short introduction to both is given below.
@@
 First a (random) subset of features is chosen based on the selected algorithm.
-In comparison to "filters", no ranking is applied in this step.
+In comparison to "filters", no ranking is done in this step.
 Now the model is fitted on the data and the performance is evaluated.
 This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
-A disadvantage of this approach is that hyperparameter tuning can only start after the feature-selection optimization has ended.
-Hence, this method is very expensive because two optimization stages need to be run in sequential order.
-"Wrapper methods" were not considered in this work due to the expensive runtime.
+A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature-selection optimization finished.
+Hence, the "wrapper approach" is an expensive optimization method because two stages need to be run in sequential order.
+Due to their extensive runtimes, "wrapper methods" were not considered in this work.
 
 % PCA
-A method that is very distinct from filters and wrappers is the "Principal Component Analysis" \citep{pearson1901, jolliffe2016}.
+A method with a completely different approach compared to filters and wrappers is the "Principal Component Analysis" \citep{pearson1901, jolliffe2016}.
 Here, the main components of the feature space are extracted and combined.
-Usually the first two components are taken since these contain the major information of the covariates.
+Most often the first two extracted main components are used since these contain the major information of the covariates.
 By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
-This enables a very cheap model fitting with a minimal loss of predictor information.
-However, the resulting features used in the model do not relate anymore to the raw input variables since they are a combination of multiple covariates.
-Interpretation is not possible anymore.
+This enables cheap model fitting with balanced loss of predictor information.
+The disadvantage of this methodology is the lack of interpretatbility because the main components cannot be related back to the original covariates.
 
@@
 \noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
-Extreme Gradient Boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
-Penalized Regression is a statistical modeling technique that is capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \citep{hastie2001}.
-The penalties are also known as "lasso" (L1) and "ridge" (L2).
-The former does not allow the removal of variables from the model (penalization to zero) while the latter does.
+Extreme gradient boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
+Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \citep{hastie2001}.
+Common penalties are "lasso" (L1) and "ridge" (L2).
+The former does not allow the full removal of variables from the model (penalization to zero) while the latter does.
 Both penalties can also be combined.
@@
 
-\subsection{Calculation of the model's feature importance}
+\subsection{Calculation of model feature importance}
 \noindent A permutation-based feature importance estimation was done for the best algorithm of each feature set and filter method, respectively.
@@
 \subsection{Performance vs. plot characteristics}
+% TODO: Really needed?
 
@@
 
-\subsubsection{Algorithm benchmarking}
+\subsubsection{Algorithm differences}
+
+\subsubsection{Feature set differences}
 
@@
 
-\subsection{Comparison to other studies}A
+\subsubsection{Variable importance vs. filter methods}
+
+\subsubsection{Ecological Interpretation}
+
+% TODO: talk about the link to RS spectra
+
+\subsection{Comparison to other studies}
 % NOTE: no environmental studies use filter methods, some use FFS
@@
 Most studies used only one algorithm (usually Random Forest) without strong arguments why this particular one has been selected.
-In our view the number of possible derivable input features is often higher than the number of variables used in the end, missing out potential helpful information from the data.
+In our view the number of possible derivable input features is often higher than the actual number of variables used in the end, missing out on potentially helpful information from the raw data.
 These findings underline the importance of this work in the field of forest health modeling.

### 2019-06-30 code/98-paper/journal/manuscript.tex
change acronyms to lower case

@@
 	%\setlength{\itemsep}{-\parsep}
-	\acro{AGB}{Above-Ground Biomass}
+	\acro{AGB}{above-ground biomass}
 	\acro{ALS}{airborn laser scanning}
-	\acro{ANN}{Artificial Neural Network}
-	\acro{AUROC}{Area Under the Receiver Operating Characteristics Curve}
-	\acro{BRT}{Boosted Regression Trees}
-	\acro{CART}{Classification and Regression Trees}
+	\acro{ANN}{artificial neural network}
+	\acro{AUROC}{area under the receiver operating characteristics curve}
+	\acro{BRT}{boosted regression trees}
+	\acro{CART}{classification and regression trees}
 	\acro{CNN}{convolutional neural networks}
@@
 	\acro{DAP}{digital aerial photogrammetry}
-	\acro{ENM}{Environmental Niche Modeling}
-	\acro{FPR}{False Positive Rate}
+	\acro{ENM}{environmental niche modeling}
+	\acro{FPR}{false positive rate}
 	\acro{FFS}{forward feature-selection}
 	\acro{FS}{feature-selection}
-	\acro{GAM}{Generalized Additive Model}
-	\acro{GBM}{Gradient Boosting Machine}
-	\acro{GLM}{Generalized Linear Model}
-	\acro{ICGC}{Institut Cartografic i Geologic de Catalunya}
-	\acro{IQR}{Interquartile Range}
-	\acro{MARS}{Multivariate Adaptive Regression Splines}
-	\acro{MEM}{Maximum Entropy Model}
-	\acro{ML}{Machine-Learning}
-	\acro{NDII}{Normalized Difference Infrared Index}
-	\acro{NIR}{Near-Infrared}
-	\acro{NRI}{Normalized Ratio Index}
+	\acro{GAM}{generalized additive model}
+	\acro{GBM}{gradient boosting machine}
+	\acro{GLM}{generalized linear model}
+	\acro{ICGC}{institut cartografic i geologic de catalunya}
+	\acro{IQR}{interquartile range}
+	\acro{MARS}{multivariate adaptive regression splines}
+	\acro{MEM}{maximum entropy model}
+	\acro{ML}{machine-learning}
+	\acro{NDII}{normalized difference infrared index}
+	\acro{NIR}{near-infrared}
+	\acro{NRI}{normalized ratio index}
 	\acro{OLS}{ordinary least squares}
 	\acro{LiDAR}{light detection and ranging}
-	\acro{LOWESS}{Locally Weighted Scatter Plot Smoothing}
-	\acro{PISR}{Potential Incoming Solar Radiation}
+	\acro{LOWESS}{locally weighted scatter plot smoothing}
+	\acro{PISR}{potential incoming solar radiation}
 	\acro{PCA}{principal component analysis}
 	\acro{PLS}{partial least-squares}
-	\acro{RBF}{Radial Basis Function}
-	\acro{RF}{Random Forest}
-	\acro{RMSE}{Root Mean Square Error}
-	\acro{RR}{Ridge Regression}
-	\acro{RSS}{Residual Sum of Squares}
-	\acro{SAR}{Synthetic Aperture Radar}
-	\acro{SDM}{Species Distribution Modeling}
-	\acro{SMBO}{Sequential-based Model Optimization}
-	\acro{SVM}{Support Vector Machine}
-	\acro{TPR}{True Positive Rate}
+	\acro{RBF}{radial basis function}
+	\acro{RF}{random forest}
+	\acro{RMSE}{root mean square error}
+	\acro{RR}{ridge regression}
+	\acro{RSS}{residual sum of squares}
+	\acro{SAR}{synthetic aperture radar}
+	\acro{SDM}{species distribution modeling}
+	\acro{SMBO}{sequential-based model optimization}
+	\acro{SVM}{support vector machine}
+	\acro{TPR}{true positive rate}
 	\acro{VI}{vegetation index}

### 2019-06-30 code/98-paper/journal/manuscript.tex
optimize introduction and data

@@
 
-	\title{Hyperspectral tree defoliation modeling with a focus on feature-selection and model interpretation}
+	\title{Supporting ecological decision making using feature-selection and variable importance}
 	%% Group authors per affiliation:
@@
 At the same time, the implementation and usability of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
-Scientists can nowadays relatively easily feed large amounts of information into complex models in a semi-automated way.
-These points enable the possibility to analyze (temporal) changes of the environment on a large scale \citep{ma2015}.
+Scientists can nowadays relatively easily feed large amounts of (environmental) information into complex models and analyze the data in a semi-automated way.
+This simplifies the analysis of (temporal) changes of the environment on a large scale \citep{ma2015}.
 
@@
 Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \citep{mascaro2014, urban2018}.
-The advantages of conducting analysis at large scale make these points a promising toolset also for forest health analysis.
+The ability of predicting to large unknown areas qualifies these tools as a promising toolset for forest health analysis.
 This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \citep{hawrylo2018}.
-Different approaches exist including temporal change detections \citep{zhang2016} or analysis about the current health status of forests on a stand level \citep{townsend2012}.
-Defoliation at trees is an information which is able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \citep{townsend2012, goodbody2018}.
+
+Approaches for analyzing forest health include temporal change detections \citep{zhang2016} or describing the current health status of forests on a stand level \citep{townsend2012}.
+In such studies, defoliation at trees is a variable able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \citep{townsend2012, goodbody2018}.
+
 Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios \citep{jiang2014, adamczyk2015}.
@@
 This emphasizes the need to extract as much information as possible from the available input data to generate features that can possibly explain the modeled relationship.
-Besides \ac{VI} a less known type of indices that can be calculated from remote-sensing products are \ac{NRI}.
-\ac{NRI} are arbitrary combinations of band information from remote-sensing data.
-Especially when working with hyperspectral data, \ac{NRI} provide the possibility to calculate hundreds of indices that can be used during modeling.
+Besides \ac{VI} a less known index type which can be calculated from remote-sensing products is the \ac{NRI}.
+\ac{NRI}s are arbitrary combinations of band information from remote-sensing data.
+Especially when working with hyperspectral data, hundreds of \ac{NRI}s can be derived and used during modeling.
 
@@
 This study aims at closing this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
-To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms is used.
-The approach of generating many features comes along with the introduction of "high-dimensionality" into the analysis \citep{trunk1979, xu2016}.
+To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms was used.
+The approach of generating many features comes along with the possible introduction of "high-dimensionality" into the analysis \citep{trunk1979, xu2016}.
+
 Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increase substantially and the interpretation of results becomes more complicated.
-Besides modeling defoliation at trees this work also aims at providing an exemplary analysis of how to deal with spatial feature-rich data derived from a hyperspectral remote-sensing source.
+Besides modeling defoliation at trees and interpreting the model results from an ecological point of view, this work also aims at providing an exemplary analysis of how to deal with spatial feature-rich datasets.
 
@@
 	\item Do different environmental feature sets show substantial differences in performance when predicting defoliation at trees?
-	\item How are different feature-selection methods performing and do they improve the predictive performance of the final model?
-	\item Is it possible to map defoliation at a tree level using high-resolution hyperspectral data at a sufficient error rate?
-	\item Which features were most important to the model to detect defoliation at trees?\end{itemize}
+    \item How are various feature-selection methods influencing the predictive performance of the models?
+	\item Which features are most important and how can these be interpreted in an ecological context?
+\end{itemize}
 
-\noindent State-of-the-art machine-learning techniques were compared using three feature sets with different feature-selection approaches including a relatively novel approach of ensemble filters.
-Model-based optimization was applied for hyperparameter tuning of the chosen algorithms.
+\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of "ensemble filters".
+Model-based optimization was applied for hyperparameter tuning of the chosen algorithms which included an optimization of the feature set.
 Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.
@@
 \section{Data and study area}
+% TODO: Add picture of affected trees
+
 \noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
 The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \citep{mesanza2016, iturritxa2017}.
-In-situ measurements of defoliation at trees (as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}).
+In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}).
 The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \citep{iturritxa2014}.
@@
 		\makebox[\textwidth]{\includegraphics[width=0.7\textwidth] {defoliation-distribution-plot-1.pdf}}
-		\caption{Distribution of defoliation at trees for plots Laukiz 1, Laukiz 2, Luiando and Oiartzun.}
+        \caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
 		\label{fig:defol-distr}
@@
 \subsection{Derivation of indices}
-% TODO: Mention different feature sets
-\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the R package \textit{hsdar} (90 in total) as well as all possible \ac{NRI} combinations were calculated.
+\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the R package \textit{hsdar} (89 in total) as well as all possible \ac{NRI} combinations were calculated.
 The following formula was used for the NRI calculation:
@@
 
-% spectral signatures
-%\begin{figure} [H]
-%\begin{center}
-%\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral_signatures.pdf}}
-%\caption{Spectral signatures (mean and standard deviation) of each plot.}
-%\label{fig:spectral_signatures}
-%\end{center}
-%\end{figure}
+\begin{figure} [H]
+\begin{center}
+\makebox[\textwidth]{\includegraphics[width=\textwidth] {spectral-signatures-1.pdf}}
+\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}
+\label{fig:spectral-signatures}
+\end{center}
+\end{figure}
 
 \pagebreak
+
+\section{Correlation among filter methods}
+
+\begin{figure} [H]
+\begin{center}
+\makebox[\textwidth]{\includegraphics[width=\textwidth] {filter-correlation-nri-1.pdf}}
+\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}
+\label{fig:correlation-filters}
+\end{center}
+\end{figure}
+
+\section{Effect of different \texttt{$n_{bins}$} values on filter "information gain"}
+
+\begin{figure} [H]
+\begin{center}
+\makebox[\textwidth]{\includegraphics[width=\textwidth] {correlation-nbins-1.pdf}}
+\caption{Spearman correlation of filter information gain using different \texttt{$n_{bins}$} values for discretization of the numeric response.}
+\label{fig:correlation-nbins}
+\end{center}
+\end{figure}
 

### 2019-06-30 analysis/filter-correlation.Rmd
workflowr::wflow_publish("analysis/filter-correlation.Rmd")

@@
 
-source(here::here("code/99-packages.R"))
-sourceDirectory("R")
+#source(here::here("code/99-packages.R"))
+library("dplyr")
+library("tidyr")
+library("ggcorrplot")
+library("purrr")
+library("glue")
+library("fs")
+R.utils::sourceDirectory("R")
 
 # load drake objects
-loadd(
+drake::loadd(
   fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
@@
 )
-library("rgdal")
 ```
@@
 
-Spearman's rank correlation is used because we compare rankings against each other.
+Spearman's rank correlation is used because rankings are compared.
 
@@
 
-```{r correlation-filter-vi, warning=FALSE, dev = c("png", "pdf")}
+```{r filter-correlation-1, warning=FALSE}
 # combine filter values
@@
 
-```{r filter-correlation-3, warning=FALSE}
+```{r filter-correlation-nri, dev = c("png", "pdf"), warning=FALSE}
 # combine filter values
@@
 ```{r filter-correlation-5, echo = FALSE}
-file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
-          "code/98-paper/journal/")
-file_move("docs/figure/filter-correlation.Rmd/correlation-filter-vi-1.pdf",
-          "code/98-paper/journal/")
+# file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
+#           "code/98-paper/journal/")
+# file_move("docs/figure/filter-correlation.Rmd/correlation-filter-nri-1.pdf",
+#           "code/98-paper/journal/")
 ```

### 2019-06-30 analysis/feature-importance.Rmd
add feature importance report

---
title: "Feature importance analysis"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

library("dplyr")
library("tidyr")
library("ggpubr")
library("purrr")
library("glue")
library("fs")
R.utils::sourceDirectory("R")

# load drake objects
drake::loadd(
)

library(dplyr)
library(hsdar)
```

```{r }
# create datasets of most imp features for each feature set
data_vi = data.frame(class = 
                    c("VI", "VI", "VI"),
                  rank = 
                    c(1, 3, 3),
                  wavelength = 
                    c(404, 404, 788),
                  feature = c("FeatureA", "FeatureB", "FeatureC")
)

data_nri = data.frame(class = 
                    c("NRI", "NRI", "NRI"),
                  rank = 
                    c(1, 3, 3),
                  wavelength = 
                    c(4050, 470, 600),
                  feature = c("FeatureA", "FeatureB", "FeatureC")
)

# create spectral signature of vegetation 
# PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`
# reflectance is scaled to 0-10 to be able to plot it in the same plot as the filter rankings -> the y-axes need to be the same
# PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm to 1000 nm
# because we order from 1 - 10 with 1 being the best, we have to reverse the scaling of the reflectance values
spectra_sim = hsdar::PROSAIL()
spectra_df = data.frame(reflectance = as.vector(spectra_sim@spectra@spectra_ma),
                wavelength = seq(400, 2500, 1)) %>% 
  dplyr::filter(wavelength < 1000) %>% 
  # scale the reflectance to fit into the range of the y-axis for the filter ranking (the 10 - is to reverse the scale)
  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10))

# to be able to plot the sec y axis on the plot, we need to supply the scaled R object -> we need to save the object separately
reflectance = as.vector(spectra_sim@spectra@spectra_ma)
reflectance_scaled = scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10)

# here we bind the simulated data with the filter rankings
# Important: Append the feature set class to the simulated data so when merged all together the facetting works (only works if we have class levels for all entries)
# This is also the reason why we cannot create an initial data.frame containing the results of all feature sets
data_vi_merged = bind_rows(spectra_df, data_vi) %>% 
  mutate(class = "VI")

data_nri_merged = bind_rows(spectra_df, data_nri) %>% 
  mutate(class = "NRI")

data_all = bind_rows(data_vi_merged, data_nri_merged)

ggscatter(data_all, x = "wavelength", y = "rank", color = "class", 
          #add = "segments",
          palette = "nejm",
          size = 3,
          facet.by = "class"
) +
  geom_segment(aes(x = wavelength, y = 10, xend = wavelength, yend = rank), color = "grey") + 
  #scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
  scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE)/-1 ),
                                                                                             labels = c(10, 0.75, 0.55, 0.25, 0),
                                                                                             name = "Reflectance [%]")) +
  labs(y = "Importance", x = "Wavelength [nm]", color = "Feature set") +
  geom_line(aes(x = wavelength, y = reflectance), linetype = "dotted") +
  ggrepel::geom_label_repel(label = data_all$feature, nudge_x = 0.5, nudge_y = 0.5)
```

### 2019-07-01 analysis/eda.Rmd
misc
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 library("ggplot2")
-library("rgdal")
 ```
@@
 ```{r eda-19, echo = FALSE}
-# file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
-#           "code/98-paper/journal/")
+file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
+          "code/98-paper/journal/")
 ```

### 2019-07-01 analysis/eda.Rmd
update reports

--- 2019-06-27 analysis/eda.Rmd
@@
 
-source(here::here("code/99-packages.R"))
-sourceDirectory("R")
+R.utils::sourceDirectory("R")
 
 # load drake objects
-loadd(
+drake::loadd(
   vi_data, nri_data, bands_data, coords_vi_nri_clean,
@@
 library("ggsci")
+library("ggpubr")
 library("ggplot2")
-library("rgdal")
+library("knitr")
 ```

### 2019-07-01 analysis/filter-correlation.Rmd
update reports

@@
 
-```{r filter-correlation-nri, dev = c("png", "pdf"), warning=FALSE}
+```{r correlation-filter-nri, dev = c("png", "pdf"), warning=FALSE}
 # combine filter values
@@
 ```{r filter-correlation-5, echo = FALSE}
+system("sleep 5")
 # file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",

### 2019-07-02 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 library("knitr")
+library("purrr")
+library("sp")
+library("raster")
 ```

### 2019-07-10 analysis/eval-performance.Rmd
update reports
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-source(here::here("code/99-packages.R"))
-sourceDirectory("R")
+R.utils::sourceDirectory("R")
 
 # load drake objects
-loadd(
+drake::loadd(
   bm_aggregated
@@
 library("magrittr")
-```
-
-```{r eval-performance-1 }
-bm_aggregated %>%
-  getBMRAggrPerformances(as.df = TRUE) %>%
-  arrange(task.id, desc(rmse.test.mean))
-```
-
-### Boxplot comparison
-
-```{r performance-boxplot, dev = c("png", "pdf")}
-plt <- plotBMRBoxplots(bm_aggregated,
-  measure = brier, pretty.names = FALSE,
-  # order.lrn = c(
-  #   "classif.gam.tuned", 
-  #   "classif.kknn.tuned",
-  #   "classif.ranger.tuned",
-  #   "classif.svm.tuned", 
-  #   "classif.gbm.tuned",
-  #   "classif.xgboost.tuned"
-  # )
-) +
-  aes(color = learner.id) +
-  geom_boxplot(width = 0.6, outlier.size = 0.3, outlier.shape = 19) +
-  scale_colour_nejm() +
-  labs(y = "Brier score", x = "") +
-  theme_pubr() +
-  theme(strip.text.x = element_text(size = 8),
-        axis.text.x = element_text(size = 7)
-        )
-
-#levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")
-
-plt
+library("dplyr")
+library("ggplot2")
+library("ggrepel")
+library("ggsci")
+library("ggpubr")
 ```
@@
 
-```{r performance-aggregated, dev = c("png", "pdf")}
-plt2 <- plotBMRSummary(bm_aggregated, pretty.names = FALSE) +
-  scale_colour_nejm() +
+#### (Table) All leaner/filter/task combinations ordered by performance.
+
+```{r eval-performance-1}
+getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  mutate(task.id = recode_factor(task.id,
+    `defoliation-all-plots-HR` = "HR",
+    `defoliation-all-plots-VI` = "VI",
+    `defoliation-all-plots-NRI` = "NRI"
+  )) %>%
+  tidyr::separate(learner.id, c("learner_group", "filter"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  group_by(learner_group, task.id, filter) %>%
+  summarise(best = min(rmse.test.rmse)) %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  arrange(best) %>%
+  knitr::kable()
+```
+
+#### (Table) Best learner/filter/task combination
+
+```{r eval-performance-2}
+getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  tidyr::separate(learner.id, c("learner_group", "filter"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  group_by(learner_group) %>%
+  slice(which.min(rmse.test.rmse)) %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  arrange(rmse.test.rmse) %>%
+  select(-6) %>%
+  knitr::kable()
+```
+
+#### (Plot) Best learner/filter combs for all tasks
+
+```{r eval-performance-3}
+results_aggr <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  mutate(task.id = recode_factor(task.id,
+    `defoliation-all-plots-HR` = "HR",
+    `defoliation-all-plots-VI` = "VI",
+    `defoliation-all-plots-NRI` = "NRI"
+  )) %>%
+  tidyr::separate(learner.id, c("learner_group"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  group_by(learner_group, task.id) %>%
+  summarise(best = min(rmse.test.rmse))
+
+results_aggr %>%
+  ggplot(aes(x = best, y = task.id)) +
+  geom_point(aes(color = learner_group, size = 1.7)) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", color = "Learner") +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(1.8, 4)) +
+  geom_label_repel(
+    label = round(results_aggr$best, 3),
+    size = 3, nudge_y = 0.3
+  ) +
   theme_pubr()
-
-#levels(plt$data$learner.id) <- c("GAM", "KNN", "RF", "SVM", "BRT", "XGBOOST")
-
-plt2
 ```

### 2019-07-10 _drake.R
add feat imp plan

@@ -28,6 +28,7 @@ pca_paper_plan = code_to_plan("code/05-modeling/paper/pca-wrapper.R")
 tuning_paper_plan = code_to_plan("code/05-modeling/paper/tune-wrapper.R")
 eda_paper_plan = code_to_plan("code/05-modeling/paper/eda.R")
 aggregate_paper_plan = code_to_plan("code/061-aggregate.R")
+feature_imp_plan = code_to_plan("code/05-modeling/paper/feature-importance.R")
 source("code/06-benchmark-matrix.R")
 
 source("code/07-reports.R")
@@ -63,7 +64,7 @@ plan_project = bind_plans(data_plan, download_plan, hyperspectral_plan, learners
 plan_paper = bind_plans(data_plan, download_plan, hyperspectral_plan, learners_paper_plan,
                         resampling_paper_plan, param_set_paper_plan, tune_ctrl_paper_plan,
                         filter_paper_plan, tuning_paper_plan, bm_plan, reports_plan_paper,
-                        pca_paper_plan, eda_paper_plan, aggregate_paper_plan
+                        pca_paper_plan, eda_paper_plan, aggregate_paper_plan, feature_imp_plan
 )
 
 options(clustermq.scheduler = "slurm",
@@ -79,21 +80,25 @@ plan_paper %<>% mutate(stage = as.factor(stage))
 #              verbose = 2, targets = "defoliation_maps_wfr", lazy_load = "promise",
 #              console_log_file = "log/drake.log", cache_log_file = "log/cache3.log",
 #              caching = "worker",
-#              template = list(log_file = "log/worker%a.log", n_cpus = 3, memory = 50000),
+#              template = list(log_file = "log/worker%a.log", n_cpus = 15, memory = 80000),
 #              prework = quote(future::plan(future::multisession, workers = 3)),
-#              garbage_collection = TRUE, jobs = 1, parallelism = "clustermq"
+#              garbage_collection = TRUE, jobs = 3, parallelism = "clustermq"
 # )
 
 # paper -------------------------------------------------------------------
 
 drake_config(plan_paper,
-             #target = c("bm_nri_task_rf_borda", "bm_nri_task_svm_mrmr"),
+             # target = c(
+             # #"eda_wfr"
+             # #"filter_correlations_wfr"
+             # "spectral_signatures_wfr"
+             # ),
              verbose = 2, lazy_load = "promise",
[…]

### 2019-07-10 code/05-modeling/paper/feature-importance.R

@@ -1,0 +1,9 @@
+fi_permut_vi = generateFilterValuesData(vi_task, method = "permutation.importance",
+                                        imp.learner = "regr.ksvm", nmc = 100,
+                                        measure = rmse)
+fi_permut_nri = generateFilterValuesData(nri_task, method = "permutation.importance",
+                                         imp.learner = "regr.ksvm", nmc = 10,
+                                         measure = rmse)
+fi_permut_hr = generateFilterValuesData(hr_task, method = "permutation.importance",
+                                        imp.learner = "regr.ksvm", nmc = 100,
+                                        measure = rmse)

### 2019-08-05 code/98-paper/journal/manuscript.tex
save state

@@
 	\item Do different environmental feature sets show substantial differences in performance when predicting defoliation at trees?
+    \item Does the combination of feature sets have an substantial effect on predictive performance?
     \item How are various feature-selection methods influencing the predictive performance of the models?
-	\item Which features are most important and how can these be interpreted in an ecological context?
+	\item Which features are most important for the models and how can these be interpreted in an ecological context?
 \end{itemize}
@@
 	\begin{center}
-		\makebox[\textwidth]{\includegraphics[width=\textwidth] {study_area_hyperspectral.pdf}}
+		\makebox[\textwidth]{\includegraphics[width=\textwidth] {study-area-hyperspectral.pdf}}
 		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}

### 2019-08-12 code/98-paper/ieee/ieee-manuscript.tex
start transfering manuscript to IEEE template

%%%% using 'arara' 4.0
% arara: pdflatex: {synctex: yes, interaction: nonstopmode}
% arara: bibtex
% arara: latexindent
% arara: pdflatex: {synctex: yes, interaction: nonstopmode}
% arara: pdflatex: {synctex: yes, interaction: nonstopmode}

%% bare_jrnl.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% see http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[letterpaper, journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  \graphicspath{{pdf/}{jpg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and
% Axel Sommerfeldt. This package may be useful when used in conjunction with
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage[nolist]{acronym}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Patrick~Schratz,~\IEEEmembership{Member,~IEEE,}
        Jannes~Muenchow,~\IEEEmembership{Member,~IEEE,}
        Eugenia Iturritxa,~\IEEEmembership{Member,~IEEE,}
        Bernd~Bischl,~\IEEEmembership{Member,~IEEE,}
        and Alexander~Brenning,~\IEEEmembership{Member,~IEEE}
\thanks{P. Schratz, J. Muenchow and A. Brenning are with the Department
of Geography, GIScience group, Friedrich-Schiller-University of Jena, Germany.}% <-this % stops a space
\thanks{B. Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
\thanks{E. Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Arab, Spain.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
}

% note the % following the last \IEEEmembership and also \thanks -
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
%
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{IEEE Transactions on Geoscience and Remote Sensing}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
%
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
hyperspectral imagery, forest health modeling, machine-learning, feature-selection, model comparison
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\renewcommand{\IEEEiedlistdecl}{\IEEEsetlabelwidth{S

ONET}}
\begin{acronym}

	% geringerer Zeilenabstand

	%\setlength{\itemsep}{-\parsep}

	\acro{AGB}{above-ground biomass}
	\acro{ALS}{airborn laser scanning}
	\acro{ANN}{artificial neural network}
	\acro{AUROC}{area under the receiver operating characteristics curve}
	\acro{BRT}{boosted regression trees}
	\acro{CART}{classification and regression trees}
	\acro{CNN}{convolutional neural networks}
	\acro{CV}{cross-validation}
	\acro{DAP}{digital aerial photogrammetry}
	\acro{ENM}{environmental niche modeling}
	\acro{FPR}{false positive rate}
	\acro{FFS}{forward feature-selection}
	\acro{FS}{feature-selection}
	\acro{GAM}{generalized additive model}
	\acro{GBM}{gradient boosting machine}
	\acro{GLM}{generalized linear model}
	\acro{ICGC}{institut cartografic i geologic de catalunya}
	\acro{IQR}{interquartile range}
	\acro{MARS}{multivariate adaptive regression splines}
	\acro{MEM}{maximum entropy model}
	\acro{ML}{machine-learning}
	\acro{NDII}{normalized difference infrared index}
	\acro{NIR}{near-infrared}
	\acro{NRI}{normalized ratio index}
	\acro{OLS}{ordinary least squares}
	\acro{LiDAR}{light detection and ranging}
	\acro{LOWESS}{locally weighted scatter plot smoothing}
	\acro{PISR}{potential incoming solar radiation}
	\acro{PCA}{principal component analysis}
	\acro{PLS}{partial least-squares}
	\acro{RBF}{radial basis function}
	\acro{RF}{random forest}
	\acro{RMSE}{root mean square error}
	\acro{RR}{ridge regression}
	\acro{RSS}{residual sum of squares}
	\acro{SAR}{synthetic aperture radar}
	\acro{SDM}{species distribution modeling}
	\acro{SMBO}{sequential-based model optimization}
	\acro{SVM}{support vector machine}
	\acro{TPR}{true positive rate}
	\acro{VI}{vegetation index}
	\acro{XGBOOST}{extreme gradient boosting}
\end{acronym}
\renewcommand{\IEEEiedlistdecl}{\relax}% reset back

\section{Introduction}
% Explain how remote sensing is used in forestry (potential to map forest health)

% Link remote sensing and machine-learning

\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
At the same time, the implementation and usability of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
Scientists can nowadays relatively easily feed large amounts of (environmental) information into complex models and analyze the data in a semi-automated way.
This simplifies the analysis of (temporal) changes of the environment on a large scale \cite{ma2015}.

% link to forest health analysis and show exemplary studies

Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
The ability of predicting to large unknown areas qualifies these tools as a promising toolset for forest health analysis.
This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \cite{hawrylo2018}.

Approaches for analyzing forest health include temporal change detections \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
In such studies, defoliation at trees is a variable able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \cite{townsend2012, goodbody2018}.

Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios \cite{jiang2014, adamczyk2015}.
Since both triggers and tree types are highly variable across regions, it cannot be known apriori which indices are most important for describing the actual relationship.
This emphasizes the need to extract as much information as possible from the available input data to generate features that can possibly explain the modeled relationship.
Besides \ac{VI} a less known index type which can be calculated from remote-sensing products is the \ac{NRI}.
\ac{NRI}s are arbitrary combinations of band information from remote-sensing data.
Especially when working with hyperspectral data, hundreds of \ac{NRI}s can be derived and used during modeling.

% now that we introduced the importance of remote sensing indices for forest health,

Despite its popularity in environmental modeling, there are no studies so far that used machine-learning algorithms in combination with remote-sensing data to analyze defoliation on a tree level.
This study aims at closing this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms was used.
The approach of generating many features comes along with the possible introduction of "high-dimensionality" into the analysis \cite{trunk1979, xu2016}.

Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increase substantially and the interpretation of results becomes more complicated.
Besides modeling defoliation at trees and interpreting the model results from an ecological point of view, this work also aims at providing an exemplary analysis of how to deal with spatial feature-rich datasets.

The reseach questions of this study are the following:

\begin{itemize}

	\item Do different environmental feature sets show substantial differences in performance when predicting defoliation at trees?

    \item Does the combination of feature sets have an substantial effect on predictive performance?

    \item How are various feature-selection methods influencing the predictive performance of the models?

	\item Which features are most important for the models and how can these be interpreted in an ecological context?

\end{itemize}

\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of "ensemble filters".
Model-based optimization was applied for hyperparameter tuning of the chosen algorithms which included an optimization of the feature set.
Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Data and study area}

\noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}).
The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
The dieback of these trees, which are mainly used as timber, causes high economic damages \cite{ganley2009}.

\begin{figure} [t!]
\centering
	\begin{center}
		\includegraphics[width=0.5\textwidth] {defoliation-distribution-plot-1.pdf}
        \caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
		\label{fig:defol-distr}
	\end{center}
\end{figure}

\subsection{In-situ data}

\noindent The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
\textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} features the largest area size (1.44 ha).
All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
In total 1759 observations are available (\textit{Laukiz 1} = 479, \textit{Laukiz 2} = 451, \textit{Luiando} = 300, \textit{Oiartzun} = 529).
The data was surveyed in September 2016.

\begin{figure} [t!]
	\begin{center}
\centering
		\includegraphics[width=0.5\textwidth] {study-area-hyperspectral.pdf}
		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}
		\label{fig:study_area}
	\end{center}
\end{figure}

\begin{figure} [t!]
	\begin{center}
		\includegraphics[width=0.5\textwidth] {defol-grid-3000px.jpg}
        \caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
		\label{fig:defol-trees}
	\end{center}
\end{figure}

\subsection{Hyperspectral data}

\noindent The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around 12 am.
The images were taken using a AISAEAGLE-II sensor.
All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by the \ac{ICGC}.
The first four bands are corrupted, leaving 122 bands with valid information.
Additional metadata information is available in \autoref{tab:hyperparameter_limits}.

% parameter limits

\begin{table}[t]
\centering
\caption[t]{Specifications of hyperspectral data.}
\begingroup
\begin{tabular}{ll}
	\\
	Characteristic         & Value                               \\
	\hline
	Geometric resolution   & 1 m                                 \\
	Radiometric resolution & 12 bit                              \\
	Spectral resolution    & 126 bands (404.08 nm - 996.31 nm)   \\
	Correction:            & Radiometric, geometric, atmospheric
\end{tabular}
\endgroup
\label{tab:hyperparameter_limits}
\end{table}

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices

\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,Biblio}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%}
%
%\end{thebibliography}

% biography section
%
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{pschratz}}]{Patrick Schratz}
%\end{IEEEbiography}

\begin{IEEEbiography}
    [{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{pschratz}}]{Patrick Schratz}
Biography text here.
\end{IEEEbiography}

% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Patrick Schratz}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{Jannes Muenchow}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Alexander Brenning}
Biography text here.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Bernd Bischl}
Biography text here.
\end{IEEEbiographynophoto}

\begin{IEEEbiographynophoto}{Eugenia Iturritxa}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2019-08-13 code/98-paper/ieee/ieee-manuscript.tex

@@
 % arara: bibtex
-% arara: latexindent
+%% arara: latexindent
 % arara: pdflatex: {synctex: yes, interaction: nonstopmode}
@@
 
-\renewcommand{\IEEEiedlistdecl}{\IEEEsetlabelwidth{S
-
-ONET}}
+\renewcommand{\IEEEiedlistdecl}{\IEEEsetlabelwidth{SONET}}
 \begin{acronym}
@@
 
-\section{Conclusion}
-The conclusion goes here.
+\section{Methods}
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
+\subsection{Derivation of indices}
+\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the R package \textit{hsdar} (89 in total) as well as all possible \ac{NRI} combinations were calculated.
+The following formula was used for the NRI calculation:
 
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
+\begin{equation}
+	NRI_{i,j} = \frac{b_{i} - b_{j}}{b_{i} + b_{j}}
+\end{equation}
+
+\noindent
+where $i$ and $j$ are the respective band numbers.
+
+\bigbreak
+
+\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), a buffer of two meters around the centroid of the respective tree was drawn.
+The value assigned to a tree observations was the mean of all pixels which touched the respective buffer around the tree.
+In total, $\frac{125*126}{2} = 7875$ NRIs were calculated .
+Due to four corrupted bands of the sensor a total of 7471 indices were available for each observation.
+
+\subsection{Dimension reduction}
+
+\noindent It is important to distinguish the term "dimension reduction" from "high-dimensionality".
+The former refers to the general idea of reducing features from a dataset \cite{vandermaaten2007}.
+In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
+"High-dimensionality" in contrast is defined as a dataset attribute and applies when $p > n,$ where $p$ is the number of covariates and $n$ the number of observations \cite{hastie2001}.
+There is no absolute value at which the term can be applied, both for $n$ and $p$.
+Hence for this study the word "high-dimensionality" only applies to the experiments using the "NRI" feature set (7470 features for 1759 observations).
+
+The case of a feature-rich dataset comes with several challenges for both model fitting and evaluation.
+
+\begin{itemize}
+  \item Model fitting times increase
+  \item Noise is possibly introduced into models by highly-correlated variables \cite{johnstoneiainm.2009}
+  \item Model interpretation and prediction tasks become more complicated \cite{johnstoneiainm.2009}
+\end{itemize}
+
+\noindent In the following sections a brief overview about sub-categories of feature-selection approaches is given.
+Due to the focus of this study on the use of filter methods, all other appraoches were grouped into a single section.
+
+\subsubsection{Filter methods}
+
+% Filter methods
+\noindent The concept of "Filter methods" originates on the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
+Some filter methods are restricted towards specific types of variables (numeric or nominal data).
+After the covariates have been ranked, a decision needs to be made what features to keep and which to discard \cite{drotar2015}.
+This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
+Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
+The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
+In well-implemented software solutions the filter calculation is only done once and then cached, saving valuable resources \cite{mlr}.
+
+\paragraph{Ensemble filter methods}
+
+% Ensemble filter methods
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
+This approach is referred to as "ensemble filter" \cite{dietterich2000}.
+Ensemble filters align with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhence predictive performance \cite{polikar2012, feurer2015}.
+In this work the "Borda" ensemble filter was applied \cite{drotar2017a}.
+The final order is the sum of all single filters ranking.
+
+\paragraph{Ensuring a fair weighting in the ensemble}
+
+Filter methods can be grouped into classes: Correlation based, entropy based, linear and non-linear methods.
+It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased.
+This was taken care of by calculating the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
+In case filters showed high correlations with each other, these were not included into the ensemble filter.
+This ensures that the ensembe filter is not biased towards a certain group of methods yielding highly similar rankings.
+
+\paragraph{Description of used filter methods}
+
+Filter methods can be classified as follows:
+
+\begin{itemize}
+	\item univariate/multivariate (scoring based on a single variable / multiple variables)
+	\item linear/non-linear (calculation of linear/non-linear interaction terms)
+	\item entropy/correlation (scoring based on derivations of entropy or correlation based approaches)
+\end{itemize}
+
+% filter methods
+\begin{table}[b!]
+\centering
+\caption[t]{List of filter methods used in this work}
+\begingroup\footnotesize
+\begin{tabular}{lll}
+	\\
+	Name                                         & Group                             & Reference          \\
+	\hline
+	Linear correlation (Pearson)                 & univariate, linear, correlation   & \cite{pearson1901} \\
+	Information gain                             & univariate, non-linear, entropy   & \cite{quinlan1986} \\
+	Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy & \cite{zhao2013}    \\
+	Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
+	Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
+	Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
+\end{tabular}
+\endgroup
+\label{tab:filter-methods}
+\end{table}
+
+\noindent Filter method "information gain" is only defined for nominal response variables:
+
+\begin{equation}
+	H(Class) + H(Attribute) - H(Class, Attribute)
+\end{equation}
+
+where $H$ is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
+To be able to use this method with a numeric response, the variable is discretized into equal bins and treated as a class variable.
+While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{$n_{bin}$ = 10} after rank correlations of > 0.9 for different bin sizes were observed in a side analysis.
+
+\subsubsection{Wrapper methods and PCA}
+
+\noindent Other approaches to assess feature importance are "wrapper methods" and the \ac{PCA} \cite{das2001, jolliffe2016}.
+A short introduction to both is given below.
+
+% Wrapper approach
+\noindent "Wrapper methods" \cite{chandrashekar2014, kohavi1997} apply algorithms that are used for hyperparameter optimization such as "Random Search" or "Generic Simulated Annealing".
+First a (random) subset of features is chosen based on the selected algorithm.
+In comparison to "filters", no ranking is done in this step.
+Now the model is fitted on the data and the performance is evaluated.
+This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
+A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature-selection optimization finished.
+Hence, the "wrapper approach" is an expensive optimization method because two stages need to be run in sequential order.
+Due to their extensive runtimes, "wrapper methods" were not considered in this work.
+
+% PCA
+A method with a completely different approach compared to filters and wrappers is the "Principal Component Analysis" \cite{pearson1901, jolliffe2016}.
+Here, the main components of the feature space are extracted and combined.
+Most often the first two extracted main components are used since these contain the major information of the covariates.
+By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
+This enables cheap model fitting with balanced loss of predictor information.
+The disadvantage of this methodology is the lack of interpretatbility because the main components cannot be related back to the original covariates.
+
+\subsection{Benchmarking design}
+
+\subsubsection{Algorithms}
+
+\noindent The benchmarking matrix of this study consists of the following algorithms:
+
+\begin{itemize}
+	\item  Extreme Gradient Boosting (XGBOOST)
+	\item  Random Forest (RF)
+	\item  Penalized Regression (both L1 and L2)
+	\item  Support Vector Machine (SVM)
+\end{itemize}
+
+\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
+Extreme gradient boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
+Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \cite{hastie2001}.
+Common penalties are "lasso" (L1) and "ridge" (L2).
+The former does not allow the full removal of variables from the model (penalization to zero) while the latter does.
+Both penalties can also be combined.
+The combined approach is called "elastic net" but was not used in this work.
+
+\subsubsection{Feature sets}
+% TODO: talk about the reasons behind different features sets: environmental implications
+
+\noindent Three feature sets were used in this study with each representing a different way of feature-engineering:
+
+\begin{itemize}
+	\item The raw hyperspectal band information (HR: No feature engineering)
+	\item Vegetation Indices (\ac{VI}: Expert-based feature engineering)
+	\item Normalized Ratio Indices (\ac{NRI}: Automated feature-engineering)
+\end{itemize}
+
+\subsubsection{Hyperparameter Optimization}
+
+\noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
+Maximum Bayesian Optimization \cite{mlrmbo} was used for parameter optimization.
+This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
+The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
+Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
+This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.
+An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
+The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum
+This applies when being compared to methods that do not use information from previous runs, such as random search or grid search \cite{bergstra2012}.
+
+For the filter methods, the percentage of features was added to the models as a hyperparameter.
+The number of main components was tuned for the instances which used \ac{PCA} instead of filtering.
+Random Forest hyperparameter \texttt{$m_{try}$} was transformed from taking absolute values to a relative parameter depending on the number of tasks: $p^{m_{try}}$, where $p$ is the number of features.
+This was necessary to ensure that \texttt{$m_{try}$} was not chosen out of bounds during tuning:
+After the filtering, only a subset of features of the task is used for optimizing the models hyperparameters.
+The size of this subset is always different since the number of features is also optimized.
+Hence, the value of  \texttt{$m_{try}$} needs to be created flexible based on the number features.
+
+\subsubsection{Spatial resampling}
+
+\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
+Each plot served as one fold within the resampling setup, resulting in four folds total.
+For the inner level (hyperparameter tuning), $p - 1$ folds were used (with $p$ being the number of plots).
+
+In total the benchmarking matrix consisted of 120 experiments (3 feature sets, 5 algorithms, 8 feature-selection methods).
+
+\subsection{Calculation of model feature importance}
+\noindent A permutation-based feature importance estimation was done for the best algorithm of each feature set and filter method, respectively.
+This method caluclates the loss of predictive performance for each feature by permuting it in a random manner.
+The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
+
+\subsection{Linking feature importance to wavelength regions}
+\noindent For ecological interpretation purposes we linked the ten most important indices of the winning models for each feature set to the spectral regions of the hyperspectral data.
+For feature set HR and NRI a direct linking to the respective bands of the hyperspectral sensor was done.
+For the vegetation indices all bands covered by the spectral range of calculated vegetation indices were counted and summed up.
+
+\subsection{Research compendium}
+
+\noindent The complete study was done using the open-source statistical programming language R \cite{rcoreteam2018}.
+The algorithm implementations of the following packages have been used: \textit{xgboost} \cite{chen2016} (\textit{xgboost}), \textit{kernlab} \cite{kernlab} (Support Vector Machine) and \textit{glmnet} \cite{glmnet} (Ridge Regression).
+The filter methods of the following packages were used: \textit{praznik \cite{praznik}}, \textit{FSelectorRcpp} \cite{fselectorrcpp}.
+The R package \textit{mlr} \cite{mlr} was used for all modeling related steps.
+\textit{drake} was used for structuring the work and ensuring reproducibility.
+This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
+
+\section{Results}
+
+\subsection{Predictive performance}
+
+\section{Discussion}
+
+\subsection{Derivation of indices}
+
+\noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
+When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation which does not spatially match with the hyperspectral (due to the geometric offset of 1 m).
+Using a buffer of more than 2 meters would increase the probability of merging information from other trees into the pixel value, blurring the actual value of the tree observation.
+That is why in our view using a buffer of 2 m was the best compromise here.
+
+Another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
+As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels are touched by the buffer.
+If a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
+
+All these points introduced a bias of an unknown magnitude into the data.
+This has to be considered when making interpretations about the outcome of this study.
+
+\subsection{Performance vs. plot characteristics}
+% TODO: Really needed?
+
+\subsection{Predictive Performance}
+
+\subsubsection{Algorithm differences}
+
+\subsubsection{Feature set differences}
+
+\subsection{Feature selection methods}
+
+\subsubsection{Variable importance vs. filter methods}
+
+\subsubsection{Ecological Interpretation}
+
+% TODO: talk about the link to RS spectra
+
+\subsection{Comparison to other studies}
+% NOTE: no environmental studies use filter methods, some use FFS
+% NOTE: only few studies analyze defoliation
+
+\noindent Most other studies analyzing defoliation operated on the plot rather than the tree level.
+This originated simply due to spatial resolution of the satellite products that served as the input data \cite{townsend2012, debeurs2008a, rengarajan2016}.
+
+Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
+\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
+Both study designs are substantially different compared to the setup of this work.
+In addition, no spatial \ac{CV} or \ac{FS} was used.
+\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
+Study results indicated that spectral metrics were found to be most helpful for the model.
+Incorporating such metrics (both spectral and structural) could be a possible enhancement for future works.
+
+\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
+\cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
+A grid-search was used for hyperparameter tuning and a \ac{FFS} for variable selection.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach \cite{meyer2018} on a Random Forest classifier.
+
+In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
+Most studies used only one algorithm (usually Random Forest) without strong arguments why this particular one has been selected.
+In our view the number of possible derivable input features is often higher than the actual number of variables used in the end, missing out on potentially helpful information from the raw data.
+These findings underline the importance of this work in the field of forest health modeling.
+
+\section{Outlook and conclusion}
 
@@
 
-\section{Proof of the First Zonklar Equation}
-Appendix one text goes here.
+\section{Spectral signatures of each plot}
 
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
+\begin{figure} [h]
+\begin{center}
+\includegraphics[width=0.5\textwidth] {spectral-signatures-1.pdf}
+\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}
+\label{fig:spectral-signatures}
+\end{center}
+\end{figure}
 
-% use section* for acknowledgment
-\section*{Acknowledgment}
+\pagebreak
 
-The authors would like to thank...
+\section{Correlation among filter methods}
 
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
+\begin{figure} [h]
+\begin{center}
+\includegraphics[width=0.5\textwidth] {correlation-filter-nri-1.pdf}
+\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}
+\label{fig:correlation-filters}
+\end{center}
+\end{figure}
 
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
+\section{Effect of different \texttt{$n_{bins}$} values on filter "information gain"}
 
-% references section
+\begin{figure} [h]
+\begin{center}
+\includegraphics[width=0.5\textwidth] {correlation-nbins-1.pdf}
+\caption{Spearman correlation of filter information gain using different \texttt{$n_{bins}$} values for discretization of the numeric response.}
+\label{fig:correlation-nbins}
+\end{center}
+\end{figure}
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
 \bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
 \bibliography{IEEEabrv,Biblio}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{
-%
-%\bibitem{IEEEhowto:kopka}
-%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-%}
-%
-%\end{thebibliography}
 
-% biography section
-%
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{pschratz}}]{Patrick Schratz}
-%\end{IEEEbiography}
+\section*{References}
 
-\begin{IEEEbiography}
-    [{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{pschratz}}]{Patrick Schratz}
-Biography text here.
-\end{IEEEbiography}
-
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}{Patrick Schratz}
-%Biography text here.
-%\end{IEEEbiography}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Jannes Muenchow}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiographynophoto}{Alexander Brenning}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-\begin{IEEEbiographynophoto}{Bernd Bischl}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-\begin{IEEEbiographynophoto}{Eugenia Iturritxa}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
 \end{document}

### 2019-08-19 analysis/eval-performance.Rmd
add new tasks to performance eval report

@@
     `defoliation-all-plots-VI` = "VI",
-    `defoliation-all-plots-NRI` = "NRI"
+    `defoliation-all-plots-NRI` = "NRI",
+    `defoliation-all-plots-HR-NRI` = "HR-NRI",
+    `defoliation-all-plots-HR-VI` = "HR-VI",
+    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+
   )) %>%
@@
     `defoliation-all-plots-VI` = "VI",
-    `defoliation-all-plots-NRI` = "NRI"
+    `defoliation-all-plots-NRI` = "NRI",
+    `defoliation-all-plots-HR-NRI` = "HR-NRI",
+    `defoliation-all-plots-HR-VI` = "HR-VI",
+    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
   )) %>%

### 2019-08-31 code/04-data-preprocessing.R
add response-normality Rmd

@@ -1,16 +1,18 @@
-# read all datasets as list --------------------------------------------------------------------
+# read all datasets as list ----------------------------------------------------
 
 data_list_nri_vi_bands <- list("vi_nri" = trees_with_indices, "bands" = trees_with_bands)
 
-# clean datasets --------------------------------------------------------------------
+# clean datasets ---------------------------------------------------------------
 
 data_clean_single_plots <- map(data_list_nri_vi_bands, ~ clean_single_plots(.x))
 
-# extract coordinates of cleaned data --------------------------------------------------------------------
+# extract coordinates of cleaned data ------------------------------------------
+
 
 data_clean_single_plots_coords <- map(data_list_nri_vi_bands, ~ extract_coords(.x))
 
-# Merge the dataframes of vi and nri --------------------------------------------------------------------
+# Merge the dataframes of vi and nri -------------------------------------------
+
 # (stored in separate lists for each plot) into one dataset.
 # Possible variables with NAs are removed again, but this is
 # actually happening in `clean_single_plots()`.
@@ -23,7 +25,8 @@ data_bands_clean <- as_tibble(rbindlist(data_clean_single_plots[[2]], fill = TRU
   Filter(function(x) !any(is.na(x)), .)
 coords_bands_clean <- as_tibble(rbindlist(data_clean_single_plots_coords[[2]]))
 
-# modify defoliation value (we get errors if defol is exactly zero) --------------------------------------------------------------------
+# modify defoliation value (we get errors if defol is exactly zero) ------------
+
 
 data_trim_defoliation <- map(
   list(data_vi_nri_clean, data_bands_clean),
@@ -31,7 +34,7 @@ data_trim_defoliation <- map(
 ) %>%
   set_names(c("data_vi_nri", "data_bands"))
 
-# split into feature sets --------------------------------------------------------------------
+# split into feature sets ------------------------------------------------------
 
[…]

### 2019-08-31 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-```{r eval-performance-3}
+```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
 results_aggr <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
@@
   )) %>%
-  tidyr::separate(learner.id, c("learner_group"),
+  tidyr::separate(learner.id, c("learner_group", "filter"),
     remove = FALSE,
@@
   ) %>%
+  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
+  mutate(learner_group = recode_factor(learner_group, `XG` = "XGBOOST")) %>% 
   group_by(learner_group, task.id) %>%
-  summarise(best = min(rmse.test.rmse))
+  # get the best performance per learner and task
+  # this is better than summarise() because we can keep additional columns
+  # in constrast to summariuse which only keeps the grouping columns and the
+  # summarised one
+  slice(which.min(rmse.test.rmse))
 
 results_aggr %>%
-  ggplot(aes(x = best, y = task.id)) +
-  geom_point(aes(color = learner_group, size = 1.7)) +
-  scale_color_nejm() +
-  labs(x = "RMSE", y = "Task", color = "Learner") +
+  ggplot(aes(x = rmse.test.rmse, y = task.id)) +
+  geom_point(aes(color = learner_group), size = 2.5) +
+  scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
+  labs(x = "RMSE log(x)", y = "Task", color = "Learner") +
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(1.8, 4)) +
+  scale_x_continuous(limits = c(30, 70)) +
+  # geom_label_repel(
+  #   label = round(results_aggr$rmse.test.rmse, 3),
+  #   size = 3, nudge_y = 0.3
+  # ) +
   geom_label_repel(
-    label = round(results_aggr$best, 3),
-    size = 3, nudge_y = 0.3
+    label = paste(results_aggr$filter,",",round(results_aggr$rmse.test.rmse, 3)),
+    size = 2.8
   ) +

### 2019-08-31 analysis/eval-performance.Rmd
add filter vs no filter comparison plot

@@
 
-### Aggregated performances
+## (Table) All leaner/filter/task combinations ordered by performance.
 
-#### (Table) All leaner/filter/task combinations ordered by performance.
+Taking into account all tasks, learners and filters: Who won?
 
-```{r eval-performance-1}
+```{r eval-performance-1, warning=FALSE}
 getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
@@
 
-#### (Table) Best learner/filter/task combination
+## (Table) Best learner/filter/task combination
 
-```{r eval-performance-2}
+Learners: On which task and using which filter did every learner score their best result on?
+
+```{r eval-performance-2, warning=FALSE}
 getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
@@
 
-#### (Plot) Best learner/filter combs for all tasks
+## (Plot) Best learner/filter combs for all tasks
 
@@
   scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
-  labs(x = "RMSE log(x)", y = "Task", color = "Learner") +
+  labs(x = "RMSE", y = "Task", color = "Learner") +
   guides(size = FALSE) +
@@
 ```
+
+## (Plot) Best filter combination of each learner vs. no filter per task
+
+Showing the final effect of applying feature selection to a learner for each task.
+
+```{r filter-effect, warning=FALSE, dev = c("png", "pdf")}
+# we rbind two filtered data.frames
+# doign the filtering in one step is complicated here
+
+best_filter = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  mutate(task.id = recode_factor(task.id,
+    `defoliation-all-plots-HR` = "HR",
+    `defoliation-all-plots-VI` = "VI",
+    `defoliation-all-plots-NRI` = "NRI",
+    `defoliation-all-plots-HR-NRI` = "HR-NRI",
+    `defoliation-all-plots-HR-VI` = "HR-VI",
+    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+
+  )) %>%
+  tidyr::separate(learner.id, c("learner_group", "filter"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter != "No Filter")  %>%
+  group_by(learner_group, task.id) %>% 
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF") %>%
+  slice(which.min(rmse.test.rmse))
+
+best_no_filter = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  mutate(task.id = recode_factor(task.id,
+    `defoliation-all-plots-HR` = "HR",
+    `defoliation-all-plots-VI` = "VI",
+    `defoliation-all-plots-NRI` = "NRI",
+    `defoliation-all-plots-HR-NRI` = "HR-NRI",
+    `defoliation-all-plots-HR-VI` = "HR-VI",
+    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+
+  )) %>%
+  tidyr::separate(learner.id, c("learner_group", "filter"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter == "No Filter")  %>%
+  group_by(learner_group, task.id) %>% 
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF") %>%
+  slice(which.min(rmse.test.rmse))
+
+comb = bind_rows(best_filter, best_no_filter)
+
+comb %>% 
+  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  geom_jitter(aes(color = filter), size = 2.5, width=0, height = 0.2) +
+  facet_wrap(~ task.id) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", color = "Filter") +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(30, 45)) +
+  # geom_label_repel(
+  #   label = round(results_aggr$rmse.test.rmse, 3),
+  #   size = 3, nudge_y = 0.3
+  # ) +
+  # geom_label_repel(
+  #   label = paste(comb$task.id),
+  #   size = 2.8
+  # ) +
+  theme_pubr() +
+  theme(panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"))
+```

### 2019-08-31 code/98-paper/ieee/ieee-manuscript.tex
start writing results

@@
 
-\subsection{Predictive performance}
-
+\subsection{Predictive performance - general}
+
+% - results across tasks show no substantial difference
+% - RMSE around 34
+% MBO tuned penalized methods are substantially worse
+% Carscore, Info und Borda were at least once among the best combination for a learner
+% RF Car is the overall winner
+% Ridge-CV hit and miss (hit = NRI, miss = VI)
+% all MBO penalized methods have no difference across tasks
+
+\noindent No practical differences in RMSE were observed among the feature sets (\autoref{fig:perf-result}).
+SVM showed the lowest error in four out of six tasks.
+A noticable difference between SVM and all other learners was visible in all feature sets without NRI variables (difference around 3-4 RMSE).
+
+In all tasks with NRI variables, the difference was either relatively small (0.2, HR-NRI-VI) or another learner achieved a better result (NRI, Ridge-CV).
+Overall, the response variable defoliation at trees could be modeled with an error of 34 \%.
+Learner RF in combination with a Carscore filter scored the absolute lowest error on the NRI task with a RMSE of 3.396.
+
+L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV (50 vs 58 RMSE).
+Regression with Ridge penalty showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was too high to be visualized (HR-NRI-VI; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI).
+All penalized methods which were tuned using MBO show exactly the same error across all tasks.
+
+\begin{figure} [t!]
+\centering
+	\begin{center}
+		\includegraphics[width=0.5\textwidth] {performance-results-1.pdf}
+        \caption{Scored predictive performance (RMSE) of models across tasks.
+Prefix "CV" denotes that the learner was optimized using internal 10-fold CV while prefix "MBO" means that Bayesian optimization was used for hyperparameter optimization.
+The abbreviations on the y-axis refer to the combinations of feature sets on which each model was scored on.
+Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = "Carscore", Info = "Information Gain", Borda = "Borda").
+The second value in the attached label of each point shows the scored RMSE value of the respective setting.}
+		\label{fig:perf-result}
+	\end{center}
+\end{figure}
 \section{Discussion}

### 2019-08-31 code/98-paper/ieee/ieee-manuscript.tex
optimize results - performance

--- 2019-08-13 code/98-paper/ieee/ieee-manuscript.tex
@@
 
+% - RMSE around 34
+% - MBO tuned penalized methods are substantially worse
+% - Carscore, Info und Borda were at least once among the best combination for a learner
+\noindent No practical differences in RMSE were observed among the feature sets (\autoref{fig:perf-result}).
+SVM showed the lowest error in four out of six tasks.
+Overall, the response variable "defoliation at trees" could be modeled with an error of 34 \%.
+
+% tasks
+% - no substantial difference between feature sets
+% - combining feature sets makes no difference
+% - expert based feature sets show no better result than data-driven ones
+% - penalized methods seem to have problems with VI feature set
+No practical difference was observed between feature sets.
+The combination of feature sets showed no increase in performance.
+All learners scored evenly across all tasks.
+Feature sets containing expert-based variables (VI) showed no substantial difference to data-driven tasks (NRI).
+
+% learner
+% - RF Car is the overall winner
+% - Ridge-CV hit and miss (hit = NRI, miss = VI)
+% - all MBO penalized methods have no difference across tasks
+Random Forest combined with the "Carscore" filter showed the best absolute performance (RMSE of 3.396).
+A noticable difference between SVM and all other learners was visible for all feature sets without NRI variables (difference in RMSE of 3-4).
+L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
+Regression with Ridge penalty showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI).
+All penalized methods tuned using MBO showed exactly the same error across all tasks.
+
+% filters
+% - influence varies based on algorithm
+% - no substantial improvement on these datasets compared to no filter
+% - no substantial descrese in performance
+% - no advantage of ensemble borda filter compared to others
+Differences between filter methods were neglibly small in this study.
+Ensemble filters showed no substantial difference to simple filter methods.
+However, there was also no change to worse when using filter methods compared to using none.
+Comparing absolute values, filters "Carscore", "Information gain" and the "Borda" ensemble filter scored the best results across all learners and tasks.
+
+\begin{figure} [t!]
+\centering
+	\begin{center}
+		\includegraphics[width=0.5\textwidth] {performance-results-1.pdf}
+        \caption{Scored predictive performance (RMSE) of models across tasks.
+Prefix "CV" denotes that the learner was optimized using internal 10-fold CV while prefix "MBO" means that Bayesian optimization was used for hyperparameter optimization.
+The abbreviations on the y-axis refer to the combinations of feature sets on which each model was scored on.
+Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = "Carscore", Info = "Information Gain", Borda = "Borda").
+The second value in the attached label of each point shows the scored RMSE value of the respective setting.}
+		\label{fig:perf-result}
+	\end{center}
+\end{figure}
 \section{Discussion}

### 2019-08-31 code/98-paper/ieee/ieee-manuscript.tex
update methods section with feature imp

@@
 \noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
-Maximum Bayesian Optimization \cite{mlrmbo} was used for parameter optimization.
+Maximum Bayesian Optimization \cite{mlrmbo} (MBO) was used for parameter optimization.
 This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
@@
 This applies when being compared to methods that do not use information from previous runs, such as random search or grid search \cite{bergstra2012}.
+In addition to MBO, the penalization term of Lasso and Ridge was optimized using the internal non-spatial CV approach due to promising results in other applications.
 
@@
 \subsection{Calculation of model feature importance}
-\noindent A permutation-based feature importance estimation was done for the best algorithm of each feature set and filter method, respectively.
-This method caluclates the loss of predictive performance for each feature by permuting it in a random manner.
+\noindent A permutation-based feature importance estimation was conducted for the best algorithm of each feature set and filter method, respectively.
+This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
 The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
+To balance both runtime and variance reduction, we opted for 25 iterations of each feature permutation.
 
@@
 \end{figure}
+
+\subsection{Permutation-based variable importance}
 \section{Discussion}

### 2019-08-31 code/98-paper/ieee/ieee-manuscript.tex
update derivation of indices

@@
 
+% FIXME: add table of first 20(?) ordered performance results across all tasks
+
+% plot performance results
 \begin{figure} [t!]
@@
 \noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
-When using no buffer at all, the possibility is high that a pixel value gets assigned to the tree observation which does not spatially match with the hyperspectral (due to the geometric offset of 1 m).
-Using a buffer of more than 2 meters would increase the probability of merging information from other trees into the pixel value, blurring the actual value of the tree observation.
-That is why in our view using a buffer of 2 m was the best compromise here.
+When using no buffer at all, the possibility is high that a pixel value gets assigned to a tree which does not spatially match with the actual recorded hyperspectral reflectance due to the geometric offset of 1 m.
+Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
+In our view using a buffer of 2 m is a good compromise between both worlds.
 
 Another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-As the buffer is a circle, it depends on the exact location of a tree observation within a pixel how much surrounding pixels are touched by the buffer.
+As the buffer is circular, determining the total number of contributing pixels depends on the exact location of a tree within that pixel as this sets the base how many pixels in each direction will be touched by the buffer.
 If a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
 
-All these points introduced a bias of an unknown magnitude into the data.
-This has to be considered when making interpretations about the outcome of this study.
+The magniute of bias introduced by these facts cannot be quantified.
+However, it has to be considered when making interpretations about the outcome of this study.
 

### 2019-08-31 code/98-paper/ieee/ieee-manuscript.tex
add jose as author, - add notes to discussion section

@@
         Eugenia Iturritxa,~\IEEEmembership{Member,~IEEE,}
+        José Cortés, ~\IEEEmembership{Member,~IEEE,}
         Bernd~Bischl,~\IEEEmembership{Member,~IEEE,}
         and Alexander~Brenning,~\IEEEmembership{Member,~IEEE}
-\thanks{P. Schratz, J. Muenchow and A. Brenning are with the Department
+\thanks{P. Schratz, J. Muenchow, J. Cortés and A. Brenning are with the Department
 of Geography, GIScience group, Friedrich-Schiller-University of Jena, Germany.}% <-this % stops a space
@@
 
-The magniute of bias introduced by these facts cannot be quantified.
+The magnitude of bias introduced by these facts cannot be quantified.
 However, it has to be considered when making interpretations about the outcome of this study.
@@
 
+% - Ensemble FS show no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
+% - Effect varies across learnes -> suggest to use FS even on feature sets with p < 100?
+% - When calculating multiple filters, consider to only use ensemble filter?
+% - PCA shows similar performance but probably requires tuning of main components -> check runtime!
+
 \subsubsection{Variable importance vs. filter methods}
+
+WIP
 

### 2019-08-31 code/98-paper/ieee/performance-best-per-learner.tex

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Sat Aug 31 23:13:02 2019
\begin{table}[ht]
\centering
\begin{tabular}{rllllr}
  \hline
 & Task & Learner ID & Learner Group & Filter & RMSE \\ 
  \hline
1 & defoliation-all-plots-NRI & RF Car & RF & Car & 33.96 \\ 
  2 & defoliation-all-plots-NRI & Ridge-CV & Ridge-CV & No Filter & 33.98 \\ 
  3 & defoliation-all-plots-HR-VI & SVM Borda & SVM & Borda & 34.25 \\ 
  4 & defoliation-all-plots-HR & Lasso-CV & Lasso-CV & No Filter & 50.45 \\ 
  5 & defoliation-all-plots-HR-VI & Ridge-MBO & Ridge-MBO & No Filter & 58.56 \\ 
  6 & defoliation-all-plots-HR & Lasso-MBO & Lasso-MBO & No Filter & 58.56 \\ 
   \hline
\end{tabular}
\end{table}

### 2019-09-01 code/98-paper/ieee/performance-best-per-learner.tex
optimize table best perf per learner

@@
 % latex table generated in R 3.5.2 by xtable 1.8-3 package
-% Sat Aug 31 23:13:02 2019
-\begin{table}[ht]
+% Sun Sep  1 10:28:46 2019
+\begin{table}[h!]
 \centering
-\begin{tabular}{rllllr}
+\caption{Best performance of each learner across any task and filter method.} 
+\label{tab:best-learner-perf}
+\begin{tabular}{rlllr}
   \hline
- & Task & Learner ID & Learner Group & Filter & RMSE \\ 
+ & Task & Learner Group & Filter & RMSE \\ 
   \hline
-1 & defoliation-all-plots-NRI & RF Car & RF & Car & 33.96 \\ 
-  2 & defoliation-all-plots-NRI & Ridge-CV & Ridge-CV & No Filter & 33.98 \\ 
-  3 & defoliation-all-plots-HR-VI & SVM Borda & SVM & Borda & 34.25 \\ 
-  4 & defoliation-all-plots-HR & Lasso-CV & Lasso-CV & No Filter & 50.45 \\ 
-  5 & defoliation-all-plots-HR-VI & Ridge-MBO & Ridge-MBO & No Filter & 58.56 \\ 
-  6 & defoliation-all-plots-HR & Lasso-MBO & Lasso-MBO & No Filter & 58.56 \\ 
+1 & defoliation-all-plots-NRI & RF & Car & 33.96 \\ 
+  2 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 33.98 \\ 
+  3 & defoliation-all-plots-HR-VI & SVM & Borda & 34.25 \\ 
+  4 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
+  5 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.56 \\ 
+  6 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
    \hline

### 2019-09-01 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

--- 2019-08-31 analysis/eval-performance.Rmd
@@
 library("ggpubr")
+library("ggbeeswarm")
+library("flextable")
+library("xtable")
 ```
 
-### Aggregated performances
+## (Table) All leaner/filter/task combinations ordered by performance.
 
-#### (Table) All leaner/filter/task combinations ordered by performance.
+Taking into account all tasks, learners and filters: Who won?
 
-```{r eval-performance-1}
-getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+```{r eval-performance-1, warning=FALSE}
+table1 <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
   mutate(task.id = recode_factor(task.id,
@@
     `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
-
   )) %>%
@@
   summarise(best = min(rmse.test.rmse)) %>%
-  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  arrange(best) %>%
-  knitr::kable()
+  dplyr::rename(
+    "Learner Group" = learner_group,
+    "Task" = task.id,
+    "Filter" = filter,
+    "RMSE" = best
+  ) %>%
+  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
+  arrange(RMSE)
+
+# save as latex table
+table1 %>%
+  ungroup() %>%
+  arrange(RMSE) %>%
+  slice(1:15) %>%
+  xtable(
+    type = "latex",
+    caption = "Top 15 results for any task/learner/filter combination, sorted by performance.",
+    label = "tab:perf-top-15"
+  ) %>%
+  print(
+    file = "code/98-paper/ieee/performance-top-20.tex",
+    include.rownames = TRUE,
+    latex.environments = c("center"),
+    table.placement = "h!",
+    caption.placement = "top"
+  )
+
+table1 %>%
+  flextable()
 ```
 
-#### (Table) Best learner/filter/task combination
+## (Table) Best learner/filter/task combination
 
-```{r eval-performance-2}
-getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+Learners: On which task and using which filter did every learner score their best result on?
+
+```{r eval-performance-2, warning=FALSE}
+table2 <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
   tidyr::separate(learner.id, c("learner_group", "filter"),
@@
   arrange(rmse.test.rmse) %>%
+  dplyr::rename(
+    "Learner Group" = learner_group,
+    "Learner ID" = learner.id,
+    "Task" = task.id,
+    "Filter" = filter,
+    "RMSE" = rmse.test.rmse
+  ) %>%
   select(-6) %>%
-  knitr::kable()
+  select(-`Learner ID`)
+
+# save as latex table
+table2 %>%
+  xtable(
+    type = "latex",
+    caption = "Best performance of each learner across any task and filter method.",
+    label = "tab:best-learner-perf"
+  ) %>%
+  print(
+    file = "code/98-paper/ieee/performance-best-per-learner.tex",
+    include.rownames = TRUE,
+    latex.environments = c("center"),
+    table.placement = "h!",
+    caption.placement = "top"
+  )
+
+table2 %>%
+  flextable()
 ```
 
-#### (Plot) Best learner/filter combs for all tasks
+## (Plot) Best learner/filter combs for all tasks
 
@@
   ggplot(aes(x = rmse.test.rmse, y = task.id)) +
-  geom_point(aes(color = learner_group), size = 2.5) +
+  #geom_jitter(aes(color = learner_group), size = 2, width = 0, height = 0.3) +
+  # geom_dotplot(aes(fill = learner_group), binaxis="y", 
+  #                        stackdir="up") +
+  geom_beeswarm(groupOnX=FALSE, aes(color = learner_group), size = 2.5) +
   scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
-  labs(x = "RMSE log(x)", y = "Task", color = "Learner") +
+  labs(x = "RMSE", y = "Task", color = "Learner") +
   guides(size = FALSE) +
   scale_x_continuous(limits = c(30, 70)) +
-  # geom_label_repel(
-  #   label = round(results_aggr$rmse.test.rmse, 3),
-  #   size = 3, nudge_y = 0.3
-  # ) +
   geom_label_repel(
-    label = paste(results_aggr$filter,",",round(results_aggr$rmse.test.rmse, 3)),
-    size = 2.8
+    # subset data to remove out of bounds values
+    data = results_aggr[results_aggr$rmse.test.rmse < 100, ],
+    # from ggbeeswarm, avoid overlapping of points by labels
+    position = position_quasirandom(),
+    aes(label = paste(filter, ",", round(rmse.test.rmse, 2))),
+    size = 2.8,
+    min.segment.length = 0.1,
+    seed = 123,
+    point.padding = 0.5
   ) +
-  theme_pubr()
+  theme_pubr() +
+  theme(
+    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+    axis.title.y = element_blank()
+  )
 ```
+
+## (Plot) Best filter combination of each learner vs. no filter per task
+
+Showing the final effect of applying feature selection to a learner for each task.
+
+```{r filter-effect, warning=FALSE, dev = c("png", "pdf")}
+# we rbind two filtered data.frames
+# doing the filtering in one step is (at least) complicated here
+
+best_filter <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  mutate(task.id = recode_factor(task.id,
+    `defoliation-all-plots-HR` = "HR",
+    `defoliation-all-plots-VI` = "VI",
+    `defoliation-all-plots-NRI` = "NRI",
+    `defoliation-all-plots-HR-NRI` = "HR-NRI",
+    `defoliation-all-plots-HR-VI` = "HR-VI",
+    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+  )) %>%
+  tidyr::separate(learner.id, c("learner_group", "filter"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter != "No Filter") %>%
+  group_by(learner_group, task.id) %>%
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
+  slice(which.min(rmse.test.rmse))
+
+best_no_filter <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+  mutate(task.id = recode_factor(task.id,
+    `defoliation-all-plots-HR` = "HR",
+    `defoliation-all-plots-VI` = "VI",
+    `defoliation-all-plots-NRI` = "NRI",
+    `defoliation-all-plots-HR-NRI` = "HR-NRI",
+    `defoliation-all-plots-HR-VI` = "HR-VI",
+    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+  )) %>%
+  tidyr::separate(learner.id, c("learner_group", "filter"),
+    remove = FALSE,
+    sep = " "
+  ) %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter == "No Filter") %>%
+  group_by(learner_group, task.id) %>%
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
+  slice(which.min(rmse.test.rmse))
+
+comb <- bind_rows(best_filter, best_no_filter)
+
+comb %>%
+  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  geom_beeswarm(groupOnX=FALSE, aes(color = filter), size = 2.5) +
+  facet_wrap(~task.id) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", color = "Filter") +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(30, 45)) +
+  theme_pubr() +
+  theme(
+    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+    axis.title.y = element_blank()
+  )
+```

### 2019-09-01 analysis/eval-performance.Rmd
style files using tidyverse style

@@
 
-Taking into account all tasks, learners and filters: Who won?
+Overall leaderboard across all settings, sorted descending by performance.
 
@@
 Learners: On which task and using which filter did every learner score their best result on?
+
+*-CV: L2 penalized regression using the internal 10-fold CV tuning of the `glmnet` package
+*-MBO: L2 penalized regression using using MBO for hyperparameter optimization.
 
@@
   ggplot(aes(x = rmse.test.rmse, y = task.id)) +
-  #geom_jitter(aes(color = learner_group), size = 2, width = 0, height = 0.3) +
+  # geom_jitter(aes(color = learner_group), size = 2, width = 0, height = 0.3) +
   # geom_dotplot(aes(fill = learner_group), binaxis="y",
   #                        stackdir="up") +
-  geom_beeswarm(groupOnX=FALSE, aes(color = learner_group), size = 2.5) +
+  geom_beeswarm(groupOnX = FALSE, aes(color = learner_group), size = 2.5) +
   scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
@@
 Showing the final effect of applying feature selection to a learner for each task.
+The more left a certain filter appears for a given task compared to the purple dot (No Filter), the higher the effectivity of applying feature selection for that given learner on the given task.
 
@@
   ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
-  geom_beeswarm(groupOnX=FALSE, aes(color = filter), size = 2.5) +
+  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2.5) +
   facet_wrap(~task.id) +

### 2019-09-01 analysis/report-defoliation.Rmd
add new report
style files using tidyverse style

@@
 # load drake objects
-loadd(defoliation_raster,
+loadd(
+  defoliation_raster,
   defoliation_raster_relative,
@@
 ```{r , eval=FALSE, echo = TRUE}
-scale(data$defoliation, center = FALSE,
-      scale = max(data$defoliation, na.rm = TRUE)/100)
+scale(data$defoliation,
+  center = FALSE,
+  scale = max(data$defoliation, na.rm = TRUE) / 100
+)
 ```
@@
 ```{r defoliation-map-2017, fig.height = 5.5, fig.width = 8.5}
-plot = create_defoliation_map(defoliation_raster[[1]],
+plot <- create_defoliation_map(defoliation_raster[[1]],
   algorithm = "XGBOOST",
@@
 ```{r defoliation-map-2018, fig.height = 5.5, fig.width = 8.5}
-plot = create_defoliation_map(defoliation_raster[[2]],
+plot <- create_defoliation_map(defoliation_raster[[2]],
   algorithm = "XGBOOST",
@@
 ```{r}
-hist = suppressWarnings(suppressMessages(ggplot(defoliation_df[[1]], aes(x = defoliation)) +
+hist <- suppressWarnings(suppressMessages(ggplot(defoliation_df[[1]], aes(x = defoliation)) +
   geom_histogram() +
   labs(x = "Defoliation (%)") +
-  theme_pubr()
-))
+  theme_pubr()))
 
@@
 ```{r}
-hist = suppressWarnings(suppressMessages(ggplot(defoliation_df[[2]], aes(x = defoliation)) +
+hist <- suppressWarnings(suppressMessages(ggplot(defoliation_df[[2]], aes(x = defoliation)) +
   geom_histogram() +
   labs(x = "Defoliation (%)") +
-  theme_pubr()
-))
+  theme_pubr()))
 suppressMessages(print(hist))
@@
 ```{r defoliation-map-relative-2017, fig.height = 5.5, fig.width = 8.5}
-plot = create_defoliation_map(defoliation_raster_relative[[1]],
+plot <- create_defoliation_map(defoliation_raster_relative[[1]],
   algorithm = "XGBOOST",
@@
 ```{r defoliation-map-relative-2018, fig.height = 5.5, fig.width = 8.5}
-plot = create_defoliation_map(defoliation_raster_relative[[2]],
+plot <- create_defoliation_map(defoliation_raster_relative[[2]],
   algorithm = "XGBOOST",
@@
 ```{r}
-hist = suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[1]], aes(x = defoliation)) +
+hist <- suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[1]], aes(x = defoliation)) +
   geom_histogram() +
   labs(x = "Relative defoliation of trees") +
-                   theme_pubr()
-))
+  theme_pubr()))
 suppressMessages(print(hist))
@@
 ```{r}
-hist = suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[2]], aes(x = defoliation)) +
+hist <- suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[2]], aes(x = defoliation)) +
   geom_histogram() +
   labs(x = "Relative defoliation of trees") +
-                   theme_pubr()
-))
+  theme_pubr()))
 suppressMessages(print(hist))
@@
 ```{r echo = FALSE}
-system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/"
-)
+system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/")
 ```

### 2019-09-01 R/01-fun-download.R

@@ -4,15 +4,16 @@
 #'
 #' @template url
 #' @name download
-download_trees = function(url) {
+download_trees <- function(url) {
   if (!file.exists("data/gpkg/oiartzun.shp")) {
     curl_download(url,
-                  destfile = glue(tempdir(), "/trees.zip"), quiet = FALSE)
+      destfile = glue(tempdir(), "/trees.zip"), quiet = FALSE
+    )
     unzip(glue(tempdir(), "/trees.zip"), exdir = "data/gpkg")
   }
 
-  files = map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~
-        st_read(glue("data/gpkg/{.x}.gpkg"), quiet = TRUE))
+  files <- map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~
+  st_read(glue("data/gpkg/{.x}.gpkg"), quiet = TRUE))
   return(files)
 }
 
@@ -20,14 +21,14 @@ download_trees = function(url) {
 #' `download_locations()`: Download plot location vector data
 #' @inheritParams download_trees
 #' @rdname download
-download_locations = function(url) {
-
+download_locations <- function(url) {
   if (!file.exists("data/gpkg/plot-locations.gpkg")) {
     curl_download(url,
-                  destfile = "data/gpkg/plot-locations.gpkg", quiet = FALSE)
+      destfile = "data/gpkg/plot-locations.gpkg", quiet = FALSE
+    )
   }
 
-  files = st_read("data/gpkg/plot-locations.gpkg") %>%
+  files <- st_read("data/gpkg/plot-locations.gpkg") %>%
     mutate(Name = as.character(ignore(Name)))
   return(files)
 }
@@ -36,16 +37,20 @@ download_locations = function(url) {
 #' `download_hyperspectral()`: Download hyperspectral data
 #' @inheritParams download_trees
 #' @rdname download
-download_hyperspectral = function(url) {
+download_hyperspectral <- function(url) {
   if (!file.exists("data/hyperspectral/B101_P1N_A090F03_ATM_S.tif")) {
     curl_download(url,
-                  destfile = glue(tempdir(), "/hs.zip"), quiet = FALSE)
-      unzip(glue(tempdir(), "/hs.zip"), exdir = "data/hyperspectral/")
+      destfile = glue(tempdir(), "/hs.zip"), quiet = FALSE
+    )
[…]

### 2019-09-01 R/02-fun-data-preprocessing.R

@@ -11,8 +11,7 @@
 #' @name data_preprocessing
 #' @export
 
-clean_single_plots = function(data) {
-
+clean_single_plots <- function(data) {
   data %<>%
     map(~ as_tibble(.x)) %<>%
     map(~ dplyr::select(.x, -contains("_ID"))) %>%
@@ -33,9 +32,8 @@ clean_single_plots = function(data) {
 #' @param data (`list`)\cr List containing multiple `data.frames`.
 #' @return List of `data.frames` with X/Y information
 #' @export
-extract_coords = function(data) {
-
-  coords = map(data, ~ st_as_sf(.x, crs = 32630)) %>%
+extract_coords <- function(data) {
+  coords <- map(data, ~ st_as_sf(.x, crs = 32630)) %>%
     map(~ as.data.frame(st_coordinates(.x)))
 
   return(coords)
@@ -52,13 +50,12 @@ extract_coords = function(data) {
 #' @details Standardization applies from `2:length(names(data))`, so it is
 #'   expected that the response is sorted first in the data.
 #' @export
-standardize = function(data, cores) {
+standardize <- function(data, cores) {
   data %<>%
     pbmclapply(function(x) {
-      cols = names(x)[2:length(names(x))]
-      x[, (cols) := lapply(.SD, scale), .SDcols=cols]
-    }, mc.cores = cores
-    )
+      cols <- names(x)[2:length(names(x))]
+      x[, (cols) := lapply(.SD, scale), .SDcols = cols]
+    }, mc.cores = cores)
 
   return(data)
 }
@@ -70,12 +67,12 @@ standardize = function(data, cores) {
 #' @param data (`data.frame`)\cr data.frame.
 #' @return `data.frame`
 #' @export
-mutate_defol = function(data) {
-
+mutate_defol <- function(data) {
   data %<>%
-    dplyr::mutate(defoliation = case_when(.data$defoliation == 0 ~ 0.001,
-                                          TRUE ~ as.numeric(.data$defoliation))
-    )
+    dplyr::mutate(defoliation = case_when(
+      .data$defoliation == 0 ~ 0.001,
+      TRUE ~ as.numeric(.data$defoliation)
+    ))
 }
 
 #' @title log_response
@@ -85,9 +82,8 @@ mutate_defol = function(data) {
 #' @param response (`character`)\cr Name of response.
 #' @return `data.frame`
 #' @export
-log_response = function(data, response) {
-
-  data[[response]] = log(data[[response]])
[…]

### 2019-09-01 R/03-fun-hyperspectral.R

@@ -140,10 +140,14 @@ extract_bands_to_plot <- function(plot_name, buffer, tree_data,
     ))
 
   out_bands %<>%
-    map2(seq_along(buffer), ~ setNames(.x,
-                                       str_replace(glue("bf{buffer}_{name}",
-                                                        name = names(out_bands[[.y]]))
-                                                   , "B.*_S.", "B")))
+    map2(seq_along(buffer), ~ setNames(
+      .x,
+      str_replace(
+        glue("bf{buffer}_{name}",
+          name = names(out_bands[[.y]])
+        ), "B.*_S.", "B"
+      )
+    ))
 
   # merge all data frames (buffers)
   out_bands <- bind_cols(out_bands)
@@ -168,7 +172,7 @@ calc_veg_indices <- function(hyperspecs, indices) {
   veg_y <- future_lapply(seq_along(hyperspecs), FUN = function(ii)
     vegindex(hyperspecs[[ii]], indices,
       filename =
-          glue("data/hyperspectral/vi/{names(hyperspecs)[[ii]]}.grd"),
+        glue("data/hyperspectral/vi/{names(hyperspecs)[[ii]]}.grd"),
       bnames = indices, na.rm = TRUE
     )) %>%
     set_names(names(hyperspecs))

### 2019-09-01 R/04-fun-sentinel-processing.R

@@ -530,7 +530,7 @@ prediction_raster <- function(data, year, relative = NULL) {
   )
 
   if (!is.null(relative)) {
-    year = glue("{year}-relative")
+    year <- glue("{year}-relative")
   }
   raster(all_spdf) %>%
     writeRaster(glue("data/sentinel/image_defoliation/defoliation-{year}.tif"),
@@ -588,9 +588,10 @@ create_defoliation_map <- function(data, algorithm, limits, title) {
 #' @return (`character`)
 #' @export
 scale_defoliation <- function(data) {
-
-  data$defoliation = scale(data$defoliation, center = FALSE,
-                           scale = max(data$defoliation, na.rm = TRUE)/100)
+  data$defoliation <- scale(data$defoliation,
+    center = FALSE,
+    scale = max(data$defoliation, na.rm = TRUE) / 100
+  )
 
   return(data)
 }

### 2019-09-01 R/fun-visualization.R

@@ -3,24 +3,23 @@
 #' @param model_name Algorithm name
 #' @param benchmark_object mlr benchmark object containing performance
 #' @param resampling String with the resampling description that should appear on the map
-create_prediction_map = function(prediction_raster, model_name, benchmark_object = NULL,
-                                 resampling = NULL) {
-
+create_prediction_map <- function(prediction_raster, model_name, benchmark_object = NULL,
+                                  resampling = NULL) {
   if (!is.null(resampling)) {
     if (resampling == "spatial/spatial") {
-      resampling_file = "sp_sp"
+      resampling_file <- "sp_sp"
     } else if (resampling == "spatial/no tuning") {
-      resampling_file = "sp_non"
+      resampling_file <- "sp_non"
     } else if (resampling == "spatial/non-spatial") {
-      resampling_file = "sp_nsp"
+      resampling_file <- "sp_nsp"
     } else if (resampling == "non-spatial/non-spatial") {
-      resampling_file = "nsp_nsp"
+      resampling_file <- "nsp_nsp"
     } else if (resampling == "non-spatial/no tuning") {
-      resampling_file = "nsp_non"
+      resampling_file <- "nsp_non"
     }
   }
 
-  out_maps = imap(prediction_raster, ~ {
+  out_maps <- imap(prediction_raster, ~ {
 
     # score = getBMRAggrPerformances(benchmark_object, as.df = TRUE) %>%
     #   dplyr::filter(task.id == .y) %>%
@@ -30,18 +29,21 @@ create_prediction_map = function(prediction_raster, model_name, benchmark_object
     ggplot() +
       annotation_map_tile(zoomin = -1, type = "cartolight") +
       layer_spatial(.x, aes(fill = stat(band1))) +
-      #scale_alpha_continuous(na.value = 0) +
+      # scale_alpha_continuous(na.value = 0) +
       scale_fill_viridis_c(na.value = NA, name = "Defoliation of trees (in %)", limits = c(35, 65)) +
       # spatial-aware automagic scale bar
       annotation_scale(location = "tl") +
       # spatial-aware automagic north arrow
       annotation_north_arrow(location = "br", which_north = "true") +
-      theme_pubr(legend = "right")  +
-      theme(legend.key.size = unit(2,"line"),
-            plot.margin = margin(1.5, 0, 1, 0)) +
-      labs(caption = glue("Algorithm: {toupper(model_name)},",
[…]

### 2019-09-01 code/05-modeling/paper/param-sets.R

@@ -1,6 +1,6 @@
 # XGBOOST -----------------------------------------------------------------
 
-ps_xgboost_filter = makeParamSet(
+ps_xgboost_filter <- makeParamSet(
   makeIntegerParam("nrounds", lower = 10, upper = 600),
   makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
   makeNumericParam("subsample", lower = 0.25, upper = 1),
@@ -11,7 +11,7 @@ ps_xgboost_filter = makeParamSet(
   makeNumericParam("fw.perc", lower = 0, upper = 1)
 )
 
-ps_xgboost = makeParamSet(
+ps_xgboost <- makeParamSet(
   makeIntegerParam("nrounds", lower = 10, upper = 600),
   makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
   makeNumericParam("subsample", lower = 0.25, upper = 1),
@@ -21,7 +21,7 @@ ps_xgboost = makeParamSet(
   makeNumericParam("min_child_weight", lower = 0, upper = 20)
 )
 
-ps_xgboost_pca =  makeParamSet(
+ps_xgboost_pca <- makeParamSet(
   makeIntegerParam("nrounds", lower = 10, upper = 600),
   makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
   makeNumericParam("subsample", lower = 0.25, upper = 1),
@@ -34,45 +34,57 @@ ps_xgboost_pca =  makeParamSet(
 
 # SVM ---------------------------------------------------------------------
 
-ps_svm_filter = makeParamSet(
-  makeNumericParam("C", lower = -10, upper = 10,
-                   trafo = function(x) 2 ^ x),
-  makeNumericParam("sigma", lower = -5, upper = 5,
-                   trafo = function(x) 2 ^ x),
+ps_svm_filter <- makeParamSet(
+  makeNumericParam("C",
+    lower = -10, upper = 10,
+    trafo = function(x) 2^x
+  ),
+  makeNumericParam("sigma",
+    lower = -5, upper = 5,
+    trafo = function(x) 2^x
+  ),
   makeNumericParam("fw.perc", lower = 0, upper = 1)
 )
 
-ps_svm = makeParamSet(
-  makeNumericParam("C", lower = -10, upper = 10,
-                   trafo = function(x) 2 ^ x),
-  makeNumericParam("sigma", lower = -5, upper = 5,
[…]

### 2019-09-01 code/05-modeling/paper/tune-wrapper.R

@@ -1,223 +1,295 @@
 # SVM ---------------------------------------------------------------------
 
-svm_borda = makeTuneWrapper(filter_wrapper_svm_borda, resampling = inner,
-                            par.set = ps_svm_filter,
-                            control = tune.ctrl_svm, show.info = TRUE,
-                            measures = list(rmse)) %>%
+svm_borda <- makeTuneWrapper(filter_wrapper_svm_borda,
+  resampling = inner,
+  par.set = ps_svm_filter,
+  control = tune.ctrl_svm, show.info = TRUE,
+  measures = list(rmse)
+) %>%
   magrittr::inset("id", "SVM Borda")
 
-svm_info.gain = makeTuneWrapper(filter_wrapper_svm_info.gain, resampling = inner,
-                                par.set = ps_svm_filter,
-                                control = tune.ctrl_svm, show.info = TRUE,
-                                measures = list(rmse)) %>%
+svm_info.gain <- makeTuneWrapper(filter_wrapper_svm_info.gain,
+  resampling = inner,
+  par.set = ps_svm_filter,
+  control = tune.ctrl_svm, show.info = TRUE,
+  measures = list(rmse)
+) %>%
   magrittr::inset("id", "SVM Info Gain")
 
-svm_gain.ratio = makeTuneWrapper(filter_wrapper_svm_gain.ratio, resampling = inner,
-                                 par.set = ps_svm_filter,
-                                 control = tune.ctrl_svm, show.info = TRUE,
-                                 measures = list(rmse)) %>%
+svm_gain.ratio <- makeTuneWrapper(filter_wrapper_svm_gain.ratio,
+  resampling = inner,
+  par.set = ps_svm_filter,
+  control = tune.ctrl_svm, show.info = TRUE,
+  measures = list(rmse)
+) %>%
   magrittr::inset("id", "SVM Gain Ratio")
 
-svm_relief = makeTuneWrapper(filter_wrapper_svm_relief, resampling = inner,
-                             par.set = ps_svm_filter,
-                             control = tune.ctrl_svm, show.info = TRUE,
-                             measures = list(rmse)) %>%
+svm_relief <- makeTuneWrapper(filter_wrapper_svm_relief,
+  resampling = inner,
+  par.set = ps_svm_filter,
+  control = tune.ctrl_svm, show.info = TRUE,
+  measures = list(rmse)
+) %>%
   magrittr::inset("id", "SVM Relief")
 
[…]

### 2019-09-01 code/06-benchmark-matrix.R

@@ -1,87 +1,89 @@
 # see https://ropenscilabs.github.io/drake-manual/plans.html#large-plans
 
-bm_plan = drake_plan(
+bm_plan <- drake_plan(
   bm = drake::target(
     benchmark_custom_no_models(learner = learner, task = task),
     transform = cross(
-      task = c("vi" = vi_task, "nri" = nri_task, "hr" = hr_task,
-               "hr_nri_vi" = hr_nri_vi_task, "hr_nri" = hr_nri_task,
-               "hr_vi" = hr_vi_task),
+      task = c(
+        "vi" = vi_task, "nri" = nri_task, "hr" = hr_task,
+        "hr_nri_vi" = hr_nri_vi_task, "hr_nri" = hr_nri_task,
+        "hr_vi" = hr_vi_task
+      ),
 
-               ## only running with the non-transformed variable
+      ## only running with the non-transformed variable
 
-               # "vi_boxcox" = vi_task_boxcox, "nri_boxcox" = nri_task_boxcox,
-               # "hr_boxcox" = hr_task_boxcox, "hr_nri_boxcox" = hr_nri_task_boxcox,
-               # "hr_vi_boxcox" = hr_vi_task_boxcox, "hr_nri_vi_boxcox" = hr_nri_vi_task_boxcox,
-               #
-               # "vi_log" = vi_task_log, "nri_log" = nri_task_log,
-               # "hr_log" = hr_task_log, "hr_nri_log" = hr_nri_task_log,
-               # "hr_vi_log" = hr_vi_task_log, "hr_nri_vi_log" = hr_nri_vi_task_log),
+      # "vi_boxcox" = vi_task_boxcox, "nri_boxcox" = nri_task_boxcox,
+      # "hr_boxcox" = hr_task_boxcox, "hr_nri_boxcox" = hr_nri_task_boxcox,
+      # "hr_vi_boxcox" = hr_vi_task_boxcox, "hr_nri_vi_boxcox" = hr_nri_vi_task_boxcox,
+      #
+      # "vi_log" = vi_task_log, "nri_log" = nri_task_log,
+      # "hr_log" = hr_task_log, "hr_nri_log" = hr_nri_task_log,
+      # "hr_vi_log" = hr_vi_task_log, "hr_nri_vi_log" = hr_nri_vi_task_log),
 
       learner = c(
-                  svm_info.gain,
-                  svm_relief,
-                  svm_linear.cor,
-                  svm_mrmr,
-                  svm_cmim,
[…]

### 2019-09-01 code/move-figures.R

@@ -1,9 +1,24 @@
-
-file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
-          "code/98-paper/journal/")
-file_move("docs/figure/filter-correlation.Rmd/correlation-filter-nri-1.pdf",
-          "code/98-paper/journal/")
-file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
-          "code/98-paper/journal/")
-file_move(here("docs/figure/spectral-signatures.Rmd/spectral-signatures-1.pdf"),
-          here("code/98-paper/journal/"))
+file_move(
+  "docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
+  "code/98-paper/ieee/pdf"
+)
+file_move(
+  "docs/figure/filter-correlation.Rmd/correlation-filter-nri-1.pdf",
+  "code/98-paper/ieee/pdf"
+)
+file_move(
+  "docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
+  "code/98-paper/ieee/pdf"
+)
+file_move(
+  here("docs/figure/spectral-signatures.Rmd/spectral-signatures-1.pdf"),
+  here("code/98-paper/ieee/pdf")
+)
+file_move(
+  here("docs/figure/eval-performance.Rmd/performance-results-1.pdf"),
+  here("code/98-paper/ieee/pdf")
+)
+file_move(
+  here("docs/figure/eval-performance.Rmd/filter-effect-1.pdf"),
+  here("code/98-paper/ieee/pdf")
+)

### 2019-09-01 analysis/eval-performance.Rmd

@@
     table.placement = "h!",
-    caption.placement = "top"
+    caption.placement = "top",
+    timestamp = NULL
   )

### 2019-09-01 analysis/feature-importance.Rmd
add feat imp plan

@@
 # load drake objects
-drake::loadd(
-)
+drake::loadd(fi_permut_vi, fi_permut_nri, fi_permut_hr)
 
@@
 # create datasets of most imp features for each feature set
-data_vi = data.frame(class = 
+data_vi <- data.frame(
+  class =
     c("VI", "VI", "VI"),
@@
 
-data_nri = data.frame(class = 
+data_nri <- data.frame(
+  class =
     c("NRI", "NRI", "NRI"),
@@
 # because we order from 1 - 10 with 1 being the best, we have to reverse the scaling of the reflectance values
-spectra_sim = hsdar::PROSAIL()
-spectra_df = data.frame(reflectance = as.vector(spectra_sim@spectra@spectra_ma),
-                wavelength = seq(400, 2500, 1)) %>% 
+spectra_sim <- hsdar::PROSAIL()
+spectra_df <- data.frame(
+  reflectance = as.vector(spectra_sim@spectra@spectra_ma),
+  wavelength = seq(400, 2500, 1)
+) %>%
   dplyr::filter(wavelength < 1000) %>%
   # scale the reflectance to fit into the range of the y-axis for the filter ranking (the 10 - is to reverse the scale)
-  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10))
+  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE) / 10))
 
 # to be able to plot the sec y axis on the plot, we need to supply the scaled R object -> we need to save the object separately
-reflectance = as.vector(spectra_sim@spectra@spectra_ma)
-reflectance_scaled = scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10)
+reflectance <- as.vector(spectra_sim@spectra@spectra_ma)
+reflectance_scaled <- scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE) / 10)
 
@@
 # This is also the reason why we cannot create an initial data.frame containing the results of all feature sets
-data_vi_merged = bind_rows(spectra_df, data_vi) %>% 
+data_vi_merged <- bind_rows(spectra_df, data_vi) %>%
   mutate(class = "VI")
 
-data_nri_merged = bind_rows(spectra_df, data_nri) %>% 
+data_nri_merged <- bind_rows(spectra_df, data_nri) %>%
   mutate(class = "NRI")
 
-data_all = bind_rows(data_vi_merged, data_nri_merged)
+data_all <- bind_rows(data_vi_merged, data_nri_merged)
 
-ggscatter(data_all, x = "wavelength", y = "rank", color = "class", 
-          #add = "segments",
+ggscatter(data_all,
+  x = "wavelength", y = "rank", color = "class",
+  # add = "segments",
   palette = "nejm",
@@
   geom_segment(aes(x = wavelength, y = 10, xend = wavelength, yend = rank), color = "grey") +
-  #scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
+  # scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
   scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
-  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE)/-1 ),
+  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE) / -1),
     labels = c(10, 0.75, 0.55, 0.25, 0),
-                                                                                             name = "Reflectance [%]")) +
+    name = "Reflectance [%]"
+  )) +
   labs(y = "Importance", x = "Wavelength [nm]", color = "Feature set") +

### 2019-09-01 analysis/filter-correlation.Rmd

@@
 
-#source(here::here("code/99-packages.R"))
+# source(here::here("code/99-packages.R"))
 library("dplyr")
@@
 # recode levels so that they are unique
-fv_all_list = map2(list(fv_vi_info.gain5$data, fv_vi_info.gain10$data,
+fv_all_list <- map2(
+  list(
+    fv_vi_info.gain5$data, fv_vi_info.gain10$data,
     fv_vi_info.gain15$data, fv_vi_info.gain20$data,
-  fv_vi_info.gain25$data, fv_vi_info.gain30$data),
+    fv_vi_info.gain25$data, fv_vi_info.gain30$data
+  ),
   list(5, 10, 15, 20, 25, 30), ~
-    dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}"))))
+  dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}")))
+)
 

### 2019-09-01 code/98-paper/ieee/performance-best-per-learner.tex

@@
 % latex table generated in R 3.5.2 by xtable 1.8-3 package
-% Sun Sep  1 10:28:46 2019
+% Sun Sep  1 15:08:58 2019
 \begin{table}[h!]
@@
   \hline
-1 & defoliation-all-plots-NRI & RF & Car & 33.96 \\ 
-  2 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 33.98 \\ 
-  3 & defoliation-all-plots-HR-VI & SVM & Borda & 34.25 \\ 
-  4 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
-  5 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.56 \\ 
-  6 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
+1 & defoliation-all-plots-HR-NRI-VI & XGBoost & No Filter & 33.95 \\ 
+  2 & defoliation-all-plots-NRI & RF & Car & 33.96 \\ 
+  3 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 33.98 \\ 
+  4 & defoliation-all-plots-HR-VI & SVM & Borda & 34.25 \\ 
+  5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
+  6 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.56 \\ 
+  7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
    \hline

### 2019-09-02 analysis/eval-performance.Rmd
wflow_publish("analysis/eval-performance.Rmd")

@@
 
-## (Table) All leaner/filter/task combinations ordered by performance.
-
-Overall leaderboard across all settings, sorted descending by performance.
-
-```{r eval-performance-1, warning=FALSE}
-table1 <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+```{r}
+df_simple = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
   mutate(task.id = recode_factor(task.id,
@@
     sep = " "
-  ) %>%
+  )
+```
+
+## (Table) All leaner/filter/task combinations ordered by performance.
+
+Overall leaderboard across all settings, sorted descending by performance.
+
+```{r eval-performance-1, warning=FALSE}
+table1 <- df_simple %>%
   group_by(learner_group, task.id, filter) %>%
@@
 ```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
-results_aggr <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
-  mutate(task.id = recode_factor(task.id,
-    `defoliation-all-plots-HR` = "HR",
-    `defoliation-all-plots-VI` = "VI",
-    `defoliation-all-plots-NRI` = "NRI",
-    `defoliation-all-plots-HR-NRI` = "HR-NRI",
-    `defoliation-all-plots-HR-VI` = "HR-VI",
-    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
-  )) %>%
-  tidyr::separate(learner.id, c("learner_group", "filter"),
-    remove = FALSE,
-    sep = " "
-  ) %>%
+results_aggr <- df_simple %>%
   mutate(filter = replace(filter, is.na(filter), "NF")) %>%
@@
   #                        stackdir="up") +
-  geom_beeswarm(groupOnX = FALSE, aes(color = learner_group), size = 2.5) +
+  geom_beeswarm(groupOnX = FALSE, aes(color = learner_group), size = 3) +
   scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
@@
     panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
-    axis.title.y = element_blank()
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12),
+    axis.text.y = element_text(angle = 45),
+    plot.margin = unit(c(6, 6, 6, 0), "pt")
   )
@@
 
-## (Plot) Best filter combination of each learner vs. no filter per task
+## (Plot) Best filter combination of each learner vs. no filter per task vs. Borda
 
@@
 
-best_filter <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
-  mutate(task.id = recode_factor(task.id,
-    `defoliation-all-plots-HR` = "HR",
-    `defoliation-all-plots-VI` = "VI",
-    `defoliation-all-plots-NRI` = "NRI",
-    `defoliation-all-plots-HR-NRI` = "HR-NRI",
-    `defoliation-all-plots-HR-VI` = "HR-VI",
-    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
-  )) %>%
-  tidyr::separate(learner.id, c("learner_group", "filter"),
-    remove = FALSE,
-    sep = " "
-  ) %>%
+best_filter <- df_simple %>%
   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
@@
   ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
-  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2.5) +
+  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 3) +
   facet_wrap(~task.id) +
@@
     panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
-    axis.title.y = element_blank()
+    axis.title.y = element_blank(),
+    axis.text.y = element_text(angle = 45),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12),
+    plot.margin = unit(c(6, 6, 6, 0), "pt")
   )
 ```
+
+## (Plot) Best filter combination of each learner vs. Borda filter
+
+Showing the final effect of the ensemble Borda filter vs the best scoring simple filter.
+
+```{r filter-effect-ensemble, warning=FALSE}
+# we rbind two filtered data.frames
+# doing the filtering in one step is (at least) complicated here
+
+best_filter <- df_simple %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter != "No Filter") %>%
+  group_by(learner_group, task.id) %>%
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
+  slice(which.min(rmse.test.rmse))
+
+borda <- df_simple %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter == "Borda") %>%
+  group_by(learner_group, task.id) %>%
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
+  slice(which.min(rmse.test.rmse))
+
+comb_borda <- bind_rows(best_filter, borda)
+
+comb_borda %>%
+  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 3) +
+  facet_wrap(~task.id) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", color = "Filter") +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(30, 45)) +
+  theme_pubr() +
+  theme(
+    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+    axis.title.y = element_blank(),
+    axis.text.y = element_text(angle = 45),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12),
+    plot.margin = unit(c(6, 6, 6, 0), "pt")
+  )
+```
+
+## (Plot) Performances of all filter methods
+
+```{r filter-perf-all, warning=FALSE, dev = c("png", "pdf")}
+
+perf_all_filters <- df_simple %>%
+  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+  dplyr::filter(filter != "No Filter") %>%
+  group_by(learner_group, task.id) %>%
+  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost")
+
+perf_all_filters %>%
+  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2) +
+  facet_wrap(~task.id) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", color = "Filter") +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(30, 45)) +
+  theme_pubr() +
+  theme(
+    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+    axis.title.y = element_blank(),
+    axis.text.y = element_text(angle = 45),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12),
+    plot.margin = unit(c(6, 6, 6, 0), "pt")
+  )
+```

### 2019-09-02 code/98-paper/ieee/ieee-manuscript.tex
save state

@@
 
+\usepackage{booktabs}
 \begin{document}
@@
 % - all MBO penalized methods have no difference across tasks
-Random Forest combined with the "Carscore" filter showed the best absolute performance (RMSE of 3.396).
+Random Forest combined with the "Carscore" filter showed the best absolute performance (RMSE of 3.396) (\autoref{tab:perf-top-15}).
 A noticable difference between SVM and all other learners was visible for all feature sets without NRI variables (difference in RMSE of 3-4).
+Considering a grouped performance by learners across all tasks and filters, SVM showed the most robust results (\autoref{tab:perf-top-15}).
 L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with Ridge penalty showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI).
+Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI) (\autoref{tab:perf-top-15}).
 All penalized methods tuned using MBO showed exactly the same error across all tasks.
@@
 
-% FIXME: add table of first 20(?) ordered performance results across all tasks
+\input{performance-top-20}
+
+\input{performance-best-per-learner.tex}
 

### 2019-09-02 code/98-paper/ieee/ieee-manuscript.tex
update results: performance

@@
 % arara: bibtex
-%% arara: latexindent
+% arara: latexindent
 % arara: pdflatex: {synctex: yes, interaction: nonstopmode}
@@
 % filters
-% - influence varies based on algorithm
-% - no substantial improvement on these datasets compared to no filter
-% - no substantial descrese in performance
+% - influence varies based on algorithm and task (SVM very small influence, RF medium, XGB medium)
+% - no substantial descrease in performance when using filters
 % - no advantage of ensemble borda filter compared to others
-Differences between filter methods were neglibly small in this study.
-Ensemble filters showed no substantial difference to simple filter methods.
-However, there was also no change to worse when using filter methods compared to using none.
-Comparing absolute values, filters "Carscore", "Information gain" and the "Borda" ensemble filter scored the best results across all learners and tasks.
-
+Effects of filter methods on performance differed between algorithms:
+With the expection of one or two methods, SVM showed a constant performance of all filters across tasks (\autoref{fig:filter-perf-all}).
+RF showed a normal distribution for most tasks with one or two different methods taking the lead on each task.
+For RF, filter "Carscore" showed the best performance in three out of six tasks while this was achieved for XGBoost by filter "CMIM".
+XGBoost showed a diverse distribution with less clustering of filters as SVM and RF (referring to RMSE).
+
+% filter vs no filter
+When comparing the usage of filters vs using no filter at all, there was only one instance (XGBoost on the HR-NRI task) when the model with no filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects}).
+In general, the effect of filtering vs.
+no filtering differed without a clear pattern across tasks: For example, SVM and RF showed a substantial increase for the NRI task when filtering the data while there was almost no effect for XGBoost (\autoref{fig:filter-effects}).
+For task HR, there was a medium effect of applying filtering for RF, almost none for SVM and a large one for XGBoost.
+
+% latex table top 20 absolute performances
 \input{performance-top-20}
 
+% latex table best performance per learner
 \input{performance-best-per-learner.tex}
@@
 
+% plot no filter vs best filter for each model and task
+\begin{figure} [t!]
+\centering
+	\begin{center}
+		\includegraphics[width=0.5\textwidth] {filter-effect-1.pdf}
+        \caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}
+		\label{fig:filter-effects}
+	\end{center}
+\end{figure}
+\subsection{Permutation-based variable importance}
+
+% plot no filter vs best filter for each model and task
+\begin{figure} [t!]
+\centering
+	\begin{center}
+		\includegraphics[width=0.5\textwidth] {filter-perf-all-1.pdf}
+        \caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}
+		\label{fig:filter-perf-all}
+	\end{center}
+\end{figure}
+
 \subsection{Permutation-based variable importance}

### 2019-09-03 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
     latex.environments = c("center"),
-    table.placement = "h!",
+    table.placement = "ht!",
     caption.placement = "top",
@@
     latex.environments = c("center"),
-    table.placement = "h!",
+    table.placement = "ht!",
+    scalebox = 0.90,
     caption.placement = "top",

### 2019-09-12 code/98-paper/ieee/ieee-manuscript.tex

@@
 
-\documentclass[letterpaper, journal]{IEEEtran}
+\documentclass[letterpaper, peerreview]{IEEEtran}
 %
@@
 \usepackage{booktabs}
+\usepackage{adjustbox}
+
 \begin{document}
@@
 \title{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
+\IEEEpeerreviewmaketitle{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
 %
@@
 %
+
+% suppress "underfull hbox" warnings
+\hbadness=99999
 
@@
 % page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
+ \ifCLASSOPTIONpeerreview
+ %\begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+ \fi
+
 % For peerreview papers, this IEEEtran command inserts a page break and
@@
 
-\subsection{Subsection Heading Here}
-Subsection text here.
-
 % needed in second column of first page if using \IEEEpubid
 %\IEEEpubidadjcol
-
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
 
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth] {defoliation-distribution-plot-1.pdf}
+		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
         \caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
@@
 \centering
-		\includegraphics[width=0.5\textwidth] {study-area-hyperspectral.pdf}
+		\includegraphics[width=0.48\textwidth] {study-area-hyperspectral.pdf}
 		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth] {defol-grid-3000px.jpg}
+		\includegraphics[width=0.48\textwidth] {defol-grid-3000px.jpg}
         \caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
@@
 \begingroup\footnotesize
+\begin{adjustbox}{width={0.48\textwidth},totalheight={\textheight},keepaspectratio}
 \begin{tabular}{lll}
@@
 \end{tabular}
+\end{adjustbox}
 \endgroup
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth] {performance-results-1.pdf}
+		\includegraphics[width=0.48\textwidth] {performance-results-1.pdf}
         \caption{Scored predictive performance (RMSE) of models across tasks.
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth] {filter-effect-1.pdf}
+		\includegraphics[width=0.48\textwidth] {filter-effect-1.pdf}
         \caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth] {filter-perf-all-1.pdf}
+		\includegraphics[width=0.48\textwidth] {filter-perf-all-1.pdf}
         \caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}
@@
 When using no buffer at all, the possibility is high that a pixel value gets assigned to a tree which does not spatially match with the actual recorded hyperspectral reflectance due to the geometric offset of 1 m.
+
 Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
@@
 
-\begin{figure} [h]
+\begin{figure} [ht]
 \begin{center}
-\includegraphics[width=0.5\textwidth] {spectral-signatures-1.pdf}
+\includegraphics[width=0.48\textwidth] {spectral-signatures-1.pdf}
 \caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}
@@
 
-\begin{figure} [h]
+\begin{figure} [ht]
 \begin{center}
-\includegraphics[width=0.5\textwidth] {correlation-filter-nri-1.pdf}
+\includegraphics[width=0.48\textwidth] {correlation-filter-nri-1.pdf}
 \caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}
@@
 
-\section{Effect of different \texttt{$n_{bins}$} values on filter "information gain"}
-
-\begin{figure} [h]
+\section{Effect of different \texorpdfstring{$n_{bins}$}{nbins} values on filter "information gain"}
+
+\begin{figure} [ht]
 \begin{center}
-\includegraphics[width=0.5\textwidth] {correlation-nbins-1.pdf}
+\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
 \caption{Spearman correlation of filter information gain using different \texttt{$n_{bins}$} values for discretization of the numeric response.}

### 2019-09-18 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-```{r}
+```{r, warning=FALSE}
 df_simple = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%

### 2019-09-19 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
+saveRDS(table1, "code/98-paper/presentation/table-perf.rda")
+
 table1 %>%
@@
   )
+
+saveRDS(table2, "code/98-paper/presentation/table-best-learner-per-task.rda")
 
@@
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 45)) +
+  scale_x_continuous(limits = c(30, 47)) +
   theme_pubr() +

### 2019-09-24 code/98-paper/ieee/performance-best-per-learner.tex
remove timestamps
update tables

--- 2019-09-01 code/98-paper/ieee/performance-best-per-learner.tex
@@
 % latex table generated in R 3.5.2 by xtable 1.8-3 package
-% Sun Sep  1 10:28:46 2019
-\begin{table}[h!]
+\begin{table}[ht!]
 \centering
@@
 \label{tab:best-learner-perf}
+\scalebox{0.9}{
 \begin{tabular}{rlllr}
@@
   \hline
-1 & defoliation-all-plots-NRI & RF & Car & 33.96 \\ 
-  2 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 33.98 \\ 
-  3 & defoliation-all-plots-HR-VI & SVM & Borda & 34.25 \\ 
-  4 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
-  5 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.56 \\ 
-  6 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
+1 & defoliation-all-plots-HR-NRI & XGBoost & Pearson & 33.84 \\ 
+  2 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
+  3 & defoliation-all-plots-HR-NRI & SVM & Pearson & 34.33 \\ 
+  4 & defoliation-all-plots-HR-NRI & RF & Car & 34.34 \\ 
+  5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
+  6 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.55 \\ 
+  7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
    \hline
 \end{tabular}
+}
 \end{table}

### 2019-10-02 code/98-paper/ieee/performance-best-per-learner.tex
upd learner table

@@
   \hline
-1 & defoliation-all-plots-HR-NRI & XGBoost & Pearson & 33.84 \\ 
-  2 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
-  3 & defoliation-all-plots-HR-NRI & SVM & Pearson & 34.33 \\ 
-  4 & defoliation-all-plots-HR-NRI & RF & Car & 34.34 \\ 
+1 & defoliation-all-plots-HR-NRI-VI & XGBoost & Borda & 32.37 \\ 
+  2 & defoliation-all-plots-NRI & RF & Car & 33.50 \\ 
+  3 & defoliation-all-plots-HR-NRI & SVM & Car & 33.95 \\ 
+  4 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
   5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
-  6 & defoliation-all-plots-HR-VI & Ridge-MBO & No Filter & 58.55 \\ 
+  6 & defoliation-all-plots-VI & Ridge-MBO & No Filter & 58.55 \\ 
   7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 

### 2019-10-31 code/98-paper/ieee/performance-best-per-learner.tex
update tables

@@
   \hline
-1 & defoliation-all-plots-HR-NRI-VI & XGBoost & Borda & 32.37 \\ 
-  2 & defoliation-all-plots-NRI & RF & Car & 33.50 \\ 
-  3 & defoliation-all-plots-HR-NRI & SVM & Car & 33.95 \\ 
-  4 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
+1 & defoliation-all-plots-NRI & RF & Car & 33.52 \\ 
+  2 & defoliation-all-plots-HR-NRI-VI & XGBoost & Borda & 33.91 \\ 
+  3 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
+  4 & defoliation-all-plots-HR-VI & SVM & Borda & 34.35 \\ 
   5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
-  6 & defoliation-all-plots-VI & Ridge-MBO & No Filter & 58.55 \\ 
+  6 & defoliation-all-plots-HR & Ridge-MBO & No Filter & 58.28 \\ 
   7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 

### 2019-11-01 code/98-paper/ieee/performance-best-per-learner.tex
save learner results for watching changes

--- 2019-10-02 code/98-paper/ieee/performance-best-per-learner.tex
@@
   \hline
-1 & defoliation-all-plots-HR-NRI-VI & XGBoost & Borda & 32.37 \\ 
-  2 & defoliation-all-plots-NRI & RF & Car & 33.50 \\ 
-  3 & defoliation-all-plots-HR-NRI & SVM & Car & 33.95 \\ 
-  4 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
+1 & defoliation-all-plots-NRI & RF & Car & 33.52 \\ 
+  2 & defoliation-all-plots-NRI & XGBoost & Borda & 33.91 \\ 
+  3 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
+  4 & defoliation-all-plots-HR-VI & SVM & Borda & 34.35 \\ 
   5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
-  6 & defoliation-all-plots-VI & Ridge-MBO & No Filter & 58.55 \\ 
+  6 & defoliation-all-plots-HR & Ridge-MBO & No Filter & 58.28 \\ 
   7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 

### 2019-12-09 analysis/eval-performance.Rmd
add scatterplots to vis BM perf

@@
 
-```{r, warning=FALSE}
-df_simple = getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+```{r}
+df_perf = getBMRPerformances(bm_aggregated, as.df = T) %>%
   mutate(task.id = recode_factor(task.id,
@@
 
+Aggregate performances and add standard error column.
+
+```{r aggr-perf}
+df_perf %<>% 
+  group_by(task.id, learner.id, filter) %>% 
+  mutate(rmse_aggr = round(mean(rmse), 3)) %>% 
+  mutate(se = round(sd(rmse), 3)) %>% 
+  select(-rmse, iter) %>% 
+  ungroup()
+```
+
 ## (Table) All leaner/filter/task combinations ordered by performance.
@@
 ```{r eval-performance-1, warning=FALSE}
-table1 <- df_simple %>%
-  group_by(learner_group, task.id, filter) %>%
-  summarise(best = min(rmse.test.rmse)) %>%
+table1 <- df_perf %>%
+  group_by(learner.id, task.id, filter) %>%
+  slice(which.min(rmse_aggr)) %>%
   dplyr::rename(
-    "Learner Group" = learner_group,
+    "Model" = learner_group,
+    "Learner ID" = learner.id,
     "Task" = task.id,
     "Filter" = filter,
-    "RMSE" = best
+    "RMSE" = rmse_aggr,
+    "Standard Error" = se,
   ) %>%
+  ungroup() %>% 
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
+  select(-iter, -`Learner ID`, -rsq, -expvar) %>% 
   arrange(RMSE)
@@
 table1 %>%
-  flextable()
+  flextable() %>% 
+  autofit()
 ```
@@
 
-*-CV: L2 penalized regression using the internal 10-fold CV tuning of the `glmnet` package
-*-MBO: L2 penalized regression using using MBO for hyperparameter optimization.
+*CV: L2 penalized regression using the internal 10-fold CV tuning of the `glmnet` package
+
+*MBO: L2 penalized regression using using MBO for hyperparameter optimization.
 
@@
 ```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
-results_aggr <- df_simple %>%
+results_aggr <- df_perf %>%
   mutate(filter = replace(filter, is.na(filter), "NF")) %>%
@@
 
-```{r filter-effect, warning=FALSE, dev = c("png", "pdf")}
+```{r filter-effect-best-vs-no-filter, warning=FALSE, dev = c("png", "pdf")}
 # we rbind two filtered data.frames
@@
 
-best_filter <- df_simple %>%
+best_filter <- df_perf %>%
   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
@@
 
-borda <- df_simple %>%
+borda <- df_perf %>%
   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
@@
 ```{r filter-perf-all, warning=FALSE, dev = c("png", "pdf")}
-
-perf_all_filters <- df_simple %>%
+perf_all_filters <- df_perf %>%
   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
@@
 ```
+
+## (Plot) Scatterplots of filter methods vs. no filter for each learner and task
+
+Showing the final effect of applying feature selection to a learner for each task.
+All filters are summarized into a boxplot whereas using "no filter" appears as a standalone observation.
+
+```{r filter-effect-all-vs-no-filter, warning=FALSE, dev = c("png", "pdf")}
+results_aggr1 <- df_perf %>%
+  filter(learner_group != "Ridge-MBO", learner_group != "Ridge-CV") %>%
+  filter(learner_group != "Lasso-MBO", learner_group != "Lasso-CV") %>%
+  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
+  mutate(learner_group = as.factor(learner_group)) %>%
+  mutate(learner_group = recode(learner_group, `XGBoost` = "XG")) %>%
+  group_by(learner_group, task.id, filter) %>% 
+  # we actually took the mean already in chunk 'aggr-perf'. This is only to get
+  # summarise() working
+  summarise(perf = mean(rmse_aggr))
+
+results_aggr1 %>%
+  ggplot(aes(x = perf, y = learner_group)) +
+  geom_beeswarm(
+    data = results_aggr1[results_aggr1$filter != "NF", ], size = 1.1, shape = 3,
+    groupOnX = FALSE, aes(color = "Filter")
+  ) +
+  geom_point(data = results_aggr1[results_aggr1$filter == "NF", ], 
+             size = 2, shape = 19, aes(color = "No Filter")) +
+  facet_wrap(~task.id) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", colour = NULL) +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(30, 50)) +
+  theme_pubr() +
+  theme(
+    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12),
+  )
+```
+
+## (Plot) Scatterplots of filter methods vs. Borda for each learner and task
+
+Showing the final effect of applying feature selection to a learner for each task.
+All filters are summarized into a boxplot whereas using "no filter" appears as a standalone observation.
+
+```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf")}
+results_aggr2 <- df_perf %>%
+  na.omit() %>% 
+  filter(learner_group != "Ridge-MBO", learner_group != "Ridge-CV") %>%
+  filter(learner_group != "Lasso-MBO", learner_group != "Lasso-CV") %>%
+  mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%
+  group_by(learner_group, task.id, filter) %>% 
+  # we actually took the mean already in chunk 'aggr-perf'. This is only to get
+  # summarise() working
+  summarise(perf = mean(rmse_aggr))
+
+results_aggr2 %>%
+  ggplot(aes(x = perf, y = learner_group)) +
+  geom_beeswarm(
+    data = results_aggr2[results_aggr2$filter != "Borda", ],
+    shape = 3, size = 1.1, aes(color = "Simple Filter"),
+    groupOnX = FALSE
+  ) +
+  geom_point(data = results_aggr2[results_aggr2$filter == "Borda", ], 
+             shape = 19, size = 2, aes(color = "Borda Filter")) +
+  facet_wrap(~task.id) +
+  scale_color_nejm() +
+  labs(x = "RMSE", y = "Task", colour = NULL) +
+  guides(size = FALSE) +
+  scale_x_continuous(limits = c(30, 48)) +
+  theme_pubr() +
+  theme(
+    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12),
+  )
+```

### 2019-12-09 code/98-paper/ieee/performance-best-per-learner.tex
upd table bes-per-learner

@@
-% latex table generated in R 3.5.2 by xtable 1.8-3 package
+% latex table generated in R 3.6.1 by xtable 1.8-3 package
 \begin{table}[ht!]
@@
 \scalebox{0.9}{
-\begin{tabular}{rlllr}
+\begin{tabular}{rlllrr}
   \hline
- & Task & Learner Group & Filter & RMSE \\ 
+ & Task & Model & Filter & RMSE & Standard Error \\ 
   \hline
-1 & defoliation-all-plots-NRI & RF & Car & 33.52 \\ 
-  2 & defoliation-all-plots-NRI & XGBoost & Borda & 33.91 \\ 
-  3 & defoliation-all-plots-NRI & Ridge-CV & No Filter & 34.05 \\ 
-  4 & defoliation-all-plots-HR-VI & SVM & Borda & 34.35 \\ 
-  5 & defoliation-all-plots-HR & Lasso-CV & No Filter & 50.45 \\ 
-  6 & defoliation-all-plots-HR & Ridge-MBO & No Filter & 58.28 \\ 
-  7 & defoliation-all-plots-HR & Lasso-MBO & No Filter & 58.56 \\ 
+1 & HR & XGBoost & CMIM & 30.83 & 13.90 \\ 
+  2 & HR & RF & MRMR & 31.44 & 15.08 \\ 
+  3 & HR-NRI-VI & SVM & Relief & 31.58 & 15.60 \\ 
+  4 & NRI & Ridge-CV & No Filter & 33.23 & 8.22 \\ 
+  5 & HR & Lasso-CV & No Filter & 47.22 & 20.52 \\ 
+  6 & HR & Lasso-MBO & No Filter & 55.11 & 22.84 \\ 
+  7 & HR & Ridge-MBO & No Filter & 55.11 & 22.84 \\ 
    \hline

### 2019-12-09 analysis/eval-performance.Rmd
Standard Error -> SE

@@
     "RMSE" = rmse_aggr,
-    "Standard Error" = se,
+    "SE" = se,
   ) %>%
@@
 ```{r eval-performance-2, warning=FALSE}
-table2 <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
-  tidyr::separate(learner.id, c("learner_group", "filter"),
-    remove = FALSE,
-    sep = " "
-  ) %>%
+table2 <- df_perf %>% 
   group_by(learner_group) %>%
-  slice(which.min(rmse.test.rmse)) %>%
+  slice(which.min(rmse_aggr)) %>%
   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  arrange(rmse.test.rmse) %>%
+  arrange(rmse_aggr) %>%
   dplyr::rename(
-    "Learner Group" = learner_group,
+    "Model" = learner_group,
     "Learner ID" = learner.id,
@@
     "Filter" = filter,
-    "RMSE" = rmse.test.rmse
+    "RMSE" = rmse_aggr,
+    "SE" = se,
   ) %>%
-  select(-6) %>%
+  select(-rsq, -expvar, -iter) %>%
   select(-`Learner ID`)
@@
 table2 %>%
-  flextable()
+  flextable() %>% 
+  autofit()
 ```

### 2019-12-10 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
   # summarised one
-  slice(which.min(rmse.test.rmse))
+  slice(which.min(rmse_aggr))
 
 results_aggr %>%
-  ggplot(aes(x = rmse.test.rmse, y = task.id)) +
+  ggplot(aes(x = rmse_aggr, y = task.id)) +
   # geom_jitter(aes(color = learner_group), size = 2, width = 0, height = 0.3) +
@@
     # subset data to remove out of bounds values
-    data = results_aggr[results_aggr$rmse.test.rmse < 100, ],
+    data = results_aggr[results_aggr$rmse_aggr < 100, ],
     # from ggbeeswarm, avoid overlapping of points by labels
     position = position_quasirandom(),
-    aes(label = paste(filter, ",", round(rmse.test.rmse, 2))),
+    aes(label = paste(filter, ",", round(rmse_aggr, 2))),
     size = 2.8,
@@
   dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-  slice(which.min(rmse.test.rmse))
+  slice(which.min(rmse_aggr))
 
-best_no_filter <- getBMRAggrPerformances(bm_aggregated, as.df = TRUE) %>%
+best_no_filter <- df_perf %>%
   mutate(task.id = recode_factor(task.id,
@@
   dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-  slice(which.min(rmse.test.rmse))
+  slice(which.min(rmse_aggr))
 
@@
 comb %>%
-  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  ggplot(aes(x = rmse_aggr, y = learner_group)) +
   geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 3) +
@@
 comb_borda %>%
-  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  ggplot(aes(x = rmse_aggr, y = learner_group)) +
   geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 3) +
@@
 perf_all_filters %>%
-  ggplot(aes(x = rmse.test.rmse, y = learner_group)) +
+  ggplot(aes(x = rmse_aggr, y = learner_group)) +
   geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2) +

### 2019-12-10 code/98-paper/ieee/performance-best-per-learner.tex
Standard Error -> SE
update perf tables

@@
   \hline
- & Task & Model & Filter & RMSE & Standard Error \\ 
+ & Task & Model & Filter & RMSE & SE \\ 
   \hline
-1 & HR & XGBoost & CMIM & 30.83 & 13.90 \\ 
+1 & HR & XGBoost & Borda & 30.56 & 13.37 \\ 
   2 & HR & RF & MRMR & 31.44 & 15.08 \\ 

### 2019-12-21 code/98-paper/ieee/ieee-manuscript.tex
update bib and clean .tex

@@
 
-% Note that keywords are not normally used for peerreview papers.
+% Note that keywords are not normally used for peer-review papers.
 \begin{IEEEkeywords}
@@
 
-% For peerreview papers, this IEEEtran command inserts a page break and
+% For peer-review papers, this IEEEtran command inserts a page break and
 % creates the second title. It will be ignored for other modes.
@@
 Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios \cite{jiang2014, adamczyk2015}.
-Since both triggers and tree types are highly variable across regions, it cannot be known apriori which indices are most important for describing the actual relationship.
+Since both triggers and tree types are highly variable across regions, it cannot be known a-priori which indices are most important for describing the actual relationship.
 This emphasizes the need to extract as much information as possible from the available input data to generate features that can possibly explain the modeled relationship.
@@
 
-The reseach questions of this study are the following:
+The research questions of this study are the following:
 
@@
 
-\begin{figure} [t!]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {defol-grid-3000px.jpg}
-        \caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
-		\label{fig:defol-trees}
-	\end{center}
-\end{figure}
+%\begin{figure} [t!]
+%	\begin{center}
+%		\includegraphics[width=0.48\textwidth] {defol-grid-3000px.jpg}
+%		\caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
+%		\label{fig:defol-trees}
+%	\end{center}
+%\end{figure}
 
@@
 \noindent In the following sections a brief overview about sub-categories of feature-selection approaches is given.
-Due to the focus of this study on the use of filter methods, all other appraoches were grouped into a single section.
+Due to the focus of this study on the use of filter methods, all other approaches were grouped into a single section.
 
@@
 This approach is referred to as "ensemble filter" \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhence predictive performance \cite{polikar2012, feurer2015}.
+Ensemble filters align with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015}.
 In this work the "Borda" ensemble filter was applied \cite{drotar2017a}.
@@
 In case filters showed high correlations with each other, these were not included into the ensemble filter.
-This ensures that the ensembe filter is not biased towards a certain group of methods yielding highly similar rankings.
+This ensures that the ensemble filter is not biased towards a certain group of methods yielding highly similar rankings.
 
@@
 This enables cheap model fitting with balanced loss of predictor information.
-The disadvantage of this methodology is the lack of interpretatbility because the main components cannot be related back to the original covariates.
+The disadvantage of this methodology is the lack of interpretability because the main components cannot be related back to the original covariates.
 
@@
 \begin{itemize}
-	\item The raw hyperspectal band information (HR: No feature engineering)
+	\item The raw hyperspectral band information (HR: No feature engineering)
 	\item Vegetation Indices (\ac{VI}: Expert-based feature engineering)
@@
 Random Forest combined with the "Carscore" filter showed the best absolute performance (RMSE of 3.396) (\autoref{tab:perf-top-15}).
-A noticable difference between SVM and all other learners was visible for all feature sets without NRI variables (difference in RMSE of 3-4).
+A noticeable difference between SVM and all other learners was visible for all feature sets without NRI variables (difference in RMSE of 3-4).
 Considering a grouped performance by learners across all tasks and filters, SVM showed the most robust results (\autoref{tab:perf-top-15}).
@@
 % - influence varies based on algorithm and task (SVM very small influence, RF medium, XGB medium)
-% - no substantial descrease in performance when using filters
+% - no substantial decrease in performance when using filters
 % - no advantage of ensemble borda filter compared to others
 Effects of filter methods on performance differed between algorithms:
-With the expection of one or two methods, SVM showed a constant performance of all filters across tasks (\autoref{fig:filter-perf-all}).
+With the exception of one or two methods, SVM showed a constant performance of all filters across tasks (\autoref{fig:filter-perf-all}).
 RF showed a normal distribution for most tasks with one or two different methods taking the lead on each task.
@@
 % - Ensemble FS show no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
-% - Effect varies across learnes -> suggest to use FS even on feature sets with p < 100?
+% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
 % - When calculating multiple filters, consider to only use ensemble filter?
@@
 \noindent Most other studies analyzing defoliation operated on the plot rather than the tree level.
-This originated simply due to spatial resolution of the satellite products that served as the input data \cite{townsend2012, debeurs2008a, rengarajan2016}.
+This originated simply due to spatial resolution of the satellite products that served as the input data \cite{townsend2012, debeurs2008, rengarajan2016}.
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,Biblio}
+\bibliography{Biblio}
 

### 2020-01-10 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 ```{r}
-df_perf = getBMRPerformances(bm_aggregated, as.df = T) %>%
+df_perf <- getBMRPerformances(bm_aggregated, as.df = T) %>%
   mutate(task.id = recode_factor(task.id,
@@
   mutate(learner_group = recode_factor(learner_group, `XG` = "XGBOOST")) %>%
-  group_by(learner_group, task.id) %>%
-  # get the best performance per learner and task
-  # this is better than summarise() because we can keep additional columns
-  # in constrast to summariuse which only keeps the grouping columns and the
+  group_by(learner_group, filter, task.id) %>%
+  ### get the best performance per learner and task
+  # this group_by() & slice() approach is better than summarise() because we can
+  # keep additional columns
+  # in constrast to summarise which only keeps the grouping columns and the
   # summarised one
+  slice(which.min(rmse_aggr)) %>% # this groups the CV iters
+  ungroup() %>%
+  group_by(task.id, learner_group) %>%
   slice(which.min(rmse_aggr))
@@
     position = position_quasirandom(),
-    aes(label = paste(filter, ",", round(rmse_aggr, 2))),
+    aes(label = paste0(filter, ",", round(rmse_aggr, 2))),
     size = 2.8,
@@
   facet_wrap(~task.id) +
-  scale_color_nejm() +
+  # scale_color_nejm() +
   labs(x = "RMSE", y = "Task", color = "Filter") +
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 47)) +
+  scale_x_continuous(limits = c(30, 50)) +
   theme_pubr() +
@@
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 45)) +
+  scale_x_continuous(limits = c(30, 50)) +
   theme_pubr() +
@@
 ```{r filter-perf-all, warning=FALSE, dev = c("png", "pdf")}
-perf_all_filters <- df_perf %>%
-  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  dplyr::filter(filter != "No Filter") %>%
-  group_by(learner_group, task.id) %>%
-  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost")
-
-perf_all_filters %>%
-  ggplot(aes(x = rmse_aggr, y = learner_group)) +
-  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2) +
-  facet_wrap(~task.id) +
-  scale_color_nejm() +
-  labs(x = "RMSE", y = "Task", color = "Filter") +
-  guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 45)) +
-  theme_pubr() +
-  theme(
-    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
-    axis.title.y = element_blank(),
-    axis.text.y = element_text(angle = 45),
-    legend.text = element_text(size = 12),
-    legend.title = element_text(size = 12),
-    plot.margin = unit(c(6, 6, 6, 0), "pt")
-  )
+# perf_all_filters <- df_perf %>%
+#   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
+#   dplyr::filter(filter != "No Filter") %>%
+#   group_by(learner_group, task.id) %>%
+#   dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
+#   slice(which.min(rmse_aggr))
+#   summarise(perf = mean(rmse_aggr), filter = as.character(filter))
+#
+# perf_all_filters %>%
+#   ggplot(aes(x = rmse_aggr, y = learner_group)) +
+#   geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2) +
+#   facet_wrap(~task.id) +
+#   scale_color_nejm() +
+#   labs(x = "RMSE", y = "Task", color = "Filter") +
+#   guides(size = FALSE) +
+#   scale_x_continuous(limits = c(30, 45)) +
+#   theme_pubr() +
+#   theme(
+#     panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
+#     axis.title.y = element_blank(),
+#     axis.text.y = element_text(angle = 45),
+#     legend.text = element_text(size = 12),
+#     legend.title = element_text(size = 12),
+#     plot.margin = unit(c(6, 6, 6, 0), "pt")
+#   )
 ```
@@
 Showing the final effect of applying feature selection to a learner for each task.
-All filters are summarized into a boxplot whereas using "no filter" appears as a standalone observation.
+All filters are colored in the same way whereas using "no filter" appears in a different color.
 
@@
   ) +
-  geom_point(data = results_aggr1[results_aggr1$filter == "NF", ], 
-             size = 2, shape = 19, aes(color = "No Filter")) +
+  geom_point(
+    data = results_aggr1[results_aggr1$filter == "NF", ],
+    size = 2, shape = 19, aes(color = "No Filter")
+  ) +
   facet_wrap(~task.id) +
@@
 Showing the final effect of applying feature selection to a learner for each task.
-All filters are summarized into a boxplot whereas using "no filter" appears as a standalone observation.
+All filters are summarized into a a single color whereas the "Borda" filter appears in its own color.
 
@@
   ) +
-  geom_point(data = results_aggr2[results_aggr2$filter == "Borda", ], 
-             shape = 19, size = 2, aes(color = "Borda Filter")) +
+  geom_point(
+    data = results_aggr2[results_aggr2$filter == "Borda", ],
+    shape = 19, size = 2, aes(color = "Borda Filter")
+  ) +
   facet_wrap(~task.id) +

### 2020-01-15 analysis/filter-correlation.Rmd
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE, verbose = TRUE)

@@
 ) %>%
-  dplyr::select(value, method) %>%
+  dplyr::select(value, filter) %>%
   mutate(id = rep(1:89, 10)) %>%
-  tidyr::spread(method, value) %>%
+  tidyr::spread(filter, value) %>%
   mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
@@
   list(5, 10, 15, 20, 25, 30), ~
-  dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}")))
+  dplyr::mutate(.x, filter = recode(.x$filter, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}")))
 )
@@
 ) %>%
-  dplyr::select(value, method) %>%
+  dplyr::select(value, filter) %>%
   mutate(id = rep(1:89, 6)) %>%
-  tidyr::spread(method, value) %>%
+  tidyr::spread(filter, value) %>%
   mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
@@
 ) %>%
-  dplyr::select(value, method) %>%
+  dplyr::select(value, filter) %>%
   mutate(id = rep(1:7381, 10)) %>%
-  tidyr::spread(method, value) %>%
+  tidyr::spread(filter, value) %>%
   dplyr::select(-id)
@@
 ) %>%
-  dplyr::select(value, method) %>%
+  dplyr::select(value, filter) %>%
   mutate(id = rep(1:122, 10)) %>%
-  tidyr::spread(method, value) %>%
+  tidyr::spread(filter, value) %>%
   dplyr::select(-id)

### 2020-02-09 docs/00-manuscripts/ieee/ieee-manuscript.tex
move paper to docs/manuscript

--- 2019-12-21 code/98-paper/ieee/ieee-manuscript.tex
@@
 % V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
+%\cite{} output to follow that of the IEEE. Loading the cite package will
 % result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
+% 'compressed/ranged'. e.g., [1], [9], [2], [7], [5], [6] without using
 % cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
+%\cite will automatically add leading space, if needed. Use cite.sty's
 % noadjust option (cite.sty V3.8 and later) if you want to turn this off
@@
 %
-\ifCLASSINFOpdf
+\ifCLASSINFOpdf{}
   \usepackage[pdftex]{graphicx}
@@
 % http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
+% Another good source of documentation is 'Using Imported Graphics in
+% LaTeX2e' by Keith Reckdahl which can be found at:
 % http://www.ctan.org/pkg/epslatex
@@
 % not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
+% which can result in 'jaggedy'/blurry rendering of lines and letters as
 % well as large increases in file sizes.
@@
 % in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
+% and invoke subfig.sty's 'caption=false' package option (available since
 % subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
@@
 % the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
+% as the top. (e.g., '\begin{figure*}[!b]' is not normally possible in
 % LaTeX2e). It also provides a command:
@@
 % spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
+% prevent endfloat from 'marking' where the figures would have been placed
 % in the text. The two hack lines of code above are a slight modification of
@@
 
-% suppress "underfull hbox" warnings
+% suppress 'underfull hbox' warnings
 \hbadness=99999
@@
 	Eugenia Iturritxa,~\IEEEmembership{Member,~IEEE,}
-        José Cortés, ~\IEEEmembership{Member,~IEEE,}
+	José Cortés,~\IEEEmembership{Member,~IEEE,}
 	Bernd~Bischl,~\IEEEmembership{Member,~IEEE,}
@@
 % a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
+% line to be shifted left slightly. This is one of those 'LaTeX things'. For
+% instance, '\textbf{A} \textbf{B}' will typeset as 'A B' not 'AB'. To get
+% 'AB' then you have to do: '\textbf{A}\textbf{B}'
 % \thanks is no different in this regard, so shield the last } of each \thanks
@@
 % page as needed:
- \ifCLASSOPTIONpeerreview
+\ifCLASSOPTIONpeerreview{}
 %\begin{center} \bfseries EDICS Category: 3-BBND \end{center}
@@
 % creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
+\IEEEpeerreviewmaketitle{}
 
@@
 
-\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
+\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade~\cite{lary2016}.
 Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
@@
 Scientists can nowadays relatively easily feed large amounts of (environmental) information into complex models and analyze the data in a semi-automated way.
-This simplifies the analysis of (temporal) changes of the environment on a large scale \cite{ma2015}.
+This simplifies the analysis of (temporal) changes of the environment on a large scale\cite{ma2015}.
 
@@
 
-Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
+Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping\cite{mascaro2014, urban2018}.
 The ability of predicting to large unknown areas qualifies these tools as a promising toolset for forest health analysis.
-This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \cite{hawrylo2018}.
-
-Approaches for analyzing forest health include temporal change detections \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
-In such studies, defoliation at trees is a variable able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \cite{townsend2012, goodbody2018}.
-
-Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios \cite{jiang2014, adamczyk2015}.
+This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees\cite{hawrylo2018}.
+
+Approaches for analyzing forest health include temporal change detection\cite{zhang2016} or describing the current health status of forests on a stand level\cite{townsend2012}.
+In such studies, defoliation at trees is a variable able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers\cite{townsend2012, goodbody2018}.
+
+Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios\cite{jiang2014, adamczyk2015}.
 Since both triggers and tree types are highly variable across regions, it cannot be known a-priori which indices are most important for describing the actual relationship.
@@
 To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms was used.
-The approach of generating many features comes along with the possible introduction of "high-dimensionality" into the analysis \cite{trunk1979, xu2016}.
+The approach of generating many features comes along with the possible
+introduction of high-dimensionality into the analysis\cite{trunk1979,
+	xu2016}.
 
@@
 
-\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of "ensemble filters".
+\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of ensemble filters.
 Model-based optimization was applied for hyperparameter tuning of the chosen algorithms which included an optimization of the feature set.
@@
 %
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
+% Reminder: the 'draftcls' or 'draftclsnofoot', not 'draft', class
 % option should be used if it is desired that the figures are to be
@@
 % - or typically anywhere on the first page for that matter. Also,
-% in-text middle ("here") positioning is typically not used, but it
+% in-text middle ('here') positioning is typically not used, but it
 % is allowed and encouraged for Computer Society conferences (but
@@
 \noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
-The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
-In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}).
-The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
-The dieback of these trees, which are mainly used as timber, causes high economic damages \cite{ganley2009}.
+The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation\cite{mesanza2016, iturritxa2017}.
+In-situ measurements of defoliation at trees (serving as a proxy for tree
+health) were collected to serve as the response variable \textit{defoliation}
+spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}). % chktex 8
+The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage\cite{iturritxa2014}.
+The dieback of these trees, which are mainly used as timber, causes high economic damages\cite{ganley2009}.
 
@@
 		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
-        \caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}
-		\label{fig:defol-distr}
+		\caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}\label{fig:defol-distr}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {study-area-hyperspectral.pdf}
-		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}
-		\label{fig:study_area}
+		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}\label{fig:study_area}
 	\end{center}
@@
 	Characteristic         & Value                               \\
-	\hline
+	\toprule
 	Geometric resolution   & 1 m                                 \\
 	Radiometric resolution & 12 bit                              \\
-	Spectral resolution    & 126 bands (404.08 nm - 996.31 nm)   \\
+	Spectral resolution    & 126 bands (404.08 nm --- 996.31 nm)   \\
 	Correction:            & Radiometric, geometric, atmospheric
 \end{tabular}
-\endgroup
-\label{tab:hyperparameter_limits}
+\endgroup\label{tab:hyperparameter_limits}
 \end{table}
@@
 \noindent
-where $i$ and $j$ are the respective band numbers.
-
-\bigbreak
+where \(i\) and \(j\) are the respective band numbers.
+
+\bigbreak{}
 
@@
 The value assigned to a tree observations was the mean of all pixels which touched the respective buffer around the tree.
-In total, $\frac{125*126}{2} = 7875$ NRIs were calculated .
+In total, \(\frac{125*126}{2} = 7875\) NRIs were calculated.
 Due to four corrupted bands of the sensor a total of 7471 indices were available for each observation.
@@
 
-\noindent It is important to distinguish the term "dimension reduction" from "high-dimensionality".
-The former refers to the general idea of reducing features from a dataset \cite{vandermaaten2007}.
+\noindent It is important to distinguish the term 'dimension reduction' from 'high-dimensionality'.
+The former refers to the general idea of reducing features from a dataset\cite{vandermaaten2007}.
 In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
-"High-dimensionality" in contrast is defined as a dataset attribute and applies when $p > n,$ where $p$ is the number of covariates and $n$ the number of observations \cite{hastie2001}.
-There is no absolute value at which the term can be applied, both for $n$ and $p$.
-Hence for this study the word "high-dimensionality" only applies to the experiments using the "NRI" feature set (7470 features for 1759 observations).
+'High-dimensionality' in contrast is defined as a dataset attribute and applies
+when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of
+observations\cite{hastie2001}.
+There is no absolute value at which the term can be applied, both for \(n\) and \(p\).
+Hence for this study the word ‘high-dimensionality‘ only applies to the
+experiments using the ‘NRI‘ feature set (7470 features for 1759 observations).
 
@@
 	\item Model fitting times increase
-  \item Noise is possibly introduced into models by highly-correlated variables \cite{johnstoneiainm.2009}
-  \item Model interpretation and prediction tasks become more complicated \cite{johnstoneiainm.2009}
+	\item Noise is possibly introduced into models by highly-correlated variables\cite{johnstoneiainm.2009}
+	\item Model interpretation and prediction tasks become more complicated\cite{johnstoneiainm.2009}
 \end{itemize}
@@
 % Filter methods
-\noindent The concept of "Filter methods" originates on the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
+\noindent The concept of 'Filter methods' originates on the idea of ranking features using certain heuristics of an algorithm\cite{chandrashekar2014}.
 Some filter methods are restricted towards specific types of variables (numeric or nominal data).
-After the covariates have been ranked, a decision needs to be made what features to keep and which to discard \cite{drotar2015}.
+After the covariates have been ranked, a decision needs to be made what features to keep and which to discard\cite{drotar2015}.
 This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
@@
 The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
-In well-implemented software solutions the filter calculation is only done once and then cached, saving valuable resources \cite{mlr}.
+In well-implemented software solutions the filter calculation is only done once
+and then cached, saving valuable resources\cite{mlr}.
 
@@
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as "minimum", "mean" or "sum" are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
-This approach is referred to as "ensemble filter" \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the "ensemble" approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015}.
-In this work the "Borda" ensemble filter was applied \cite{drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables,
+studies showed that combining several filters using statistical operations such
+as 'minimum', 'mean' or 'sum' are able to enhance the predictive performance of
+the resulting models\cite{abeel2010, drotar2017a}.
+This approach is referred to as 'ensemble filter'\cite{dietterich2000}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance\cite{polikar2012, feurer2015}.
+In this work the 'Borda' ensemble filter was applied\cite{drotar2017a}.
 The final order is the sum of all single filters ranking.
@@
 		Name                                         & Group                             & Reference          \\
-	\hline
-	Linear correlation (Pearson)                 & univariate, linear, correlation   & \cite{pearson1901} \\
-	Information gain                             & univariate, non-linear, entropy   & \cite{quinlan1986} \\
-	Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy & \cite{zhao2013}    \\
-	Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
-	Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
-	Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
+		\toprule
+		Linear correlation (Pearson)                 & univariate, linear, correlation   &\cite{pearson1901} \\
+		Information gain                             & univariate, non-linear, entropy   &\cite{quinlan1986} \\
+		Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy &\cite{zhao2013}    \\
+		Carscore                                     & multivariate, linear, correlation &\cite{zuber2011}   \\
+		Relief                                       & multivariate, linear, entropy,    &\cite{kira1992}    \\
+		Conditional minimal information maximization & multivariate, linear, entropy     &\cite{fleuret2004}
 	\end{tabular}
 \end{adjustbox}
-\endgroup
-\label{tab:filter-methods}
+\endgroup\label{tab:filter-methods}
 \end{table}
 
-\noindent Filter method "information gain" is only defined for nominal response variables:
+\noindent Filter method 'information gain' is only defined for nominal response variables:
 
@@
 
-where $H$ is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
+where \(H\) is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
 To be able to use this method with a numeric response, the variable is discretized into equal bins and treated as a class variable.
-While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{$n_{bin}$ = 10} after rank correlations of > 0.9 for different bin sizes were observed in a side analysis.
+While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of > 0.9 for different bin sizes were observed in a side analysis.
 
@@
 
-\noindent Other approaches to assess feature importance are "wrapper methods" and the \ac{PCA} \cite{das2001, jolliffe2016}.
+\noindent Other approaches to assess feature importance are 'wrapper methods' and the \ac{PCA}\cite{das2001, jolliffe2016}.
 A short introduction to both is given below.
@@
 % Wrapper approach
-\noindent "Wrapper methods" \cite{chandrashekar2014, kohavi1997} apply algorithms that are used for hyperparameter optimization such as "Random Search" or "Generic Simulated Annealing".
+\noindent 'Wrapper methods'\cite{chandrashekar2014, kohavi1997} apply algorithms that are used for hyperparameter optimization such as 'Random Search' or 'Generic Simulated Annealing'.
 First a (random) subset of features is chosen based on the selected algorithm.
-In comparison to "filters", no ranking is done in this step.
+In comparison to 'filters', no ranking is done in this step.
 Now the model is fitted on the data and the performance is evaluated.
@@
 A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature-selection optimization finished.
-Hence, the "wrapper approach" is an expensive optimization method because two stages need to be run in sequential order.
-Due to their extensive runtimes, "wrapper methods" were not considered in this work.
+Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.
+Due to their extensive runtimes, 'wrapper methods' were not considered in this work.
 
 % PCA
-A method with a completely different approach compared to filters and wrappers is the "Principal Component Analysis" \cite{pearson1901, jolliffe2016}.
+A method with a completely different approach compared to filters and wrappers is the 'Principal Component Analysis'\cite{pearson1901, jolliffe2016}.
 Here, the main components of the feature space are extracted and combined.
@@
 Extreme gradient boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
-Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model \cite{hastie2001}.
-Common penalties are "lasso" (L1) and "ridge" (L2).
+Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
+Common penalties are 'lasso' (L1) and 'ridge' (L2).
 The former does not allow the full removal of variables from the model (penalization to zero) while the latter does.
 Both penalties can also be combined.
-The combined approach is called "elastic net" but was not used in this work.
+The combined approach is called 'elastic net' but was not used in this work.
 
@@
 \begin{itemize}
-	\item The raw hyperspectral band information (HR: No feature engineering)
+	\item The raw hyperspectral band information (HR: No feature engineering) %chktex 13
 	\item Vegetation Indices (\ac{VI}: Expert-based feature engineering)
@@
 \noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
-Maximum Bayesian Optimization \cite{mlrmbo} (MBO) was used for parameter optimization.
+Maximum Bayesian Optimization\cite{mlrmbo} (MBO) was used for parameter optimization.
 This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
@@
 Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
-This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.
+This strategy continues until a termination criterion, defined by the user, is reached\cite{hutter2011, jones1998}.
 An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
 The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum
-This applies when being compared to methods that do not use information from previous runs, such as random search or grid search \cite{bergstra2012}.
+This applies when being compared to methods that do not use information from previous runs, such as random search or grid search\cite{bergstra2012}.
 In addition to MBO, the penalization term of Lasso and Ridge was optimized using the internal non-spatial CV approach due to promising results in other applications.
@@
 The number of main components was tuned for the instances which used \ac{PCA} instead of filtering.
-Random Forest hyperparameter \texttt{$m_{try}$} was transformed from taking absolute values to a relative parameter depending on the number of tasks: $p^{m_{try}}$, where $p$ is the number of features.
-This was necessary to ensure that \texttt{$m_{try}$} was not chosen out of bounds during tuning:
+Random Forest hyperparameter \texttt{\(m_{try}\)} was transformed from taking absolute values to a relative parameter depending on the number of tasks: \(p^{m_{try}}\), where \(p\) is the number of features.
+This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen out of bounds during tuning:
 After the filtering, only a subset of features of the task is used for optimizing the models hyperparameters.
 The size of this subset is always different since the number of features is also optimized.
-Hence, the value of  \texttt{$m_{try}$} needs to be created flexible based on the number features.
+Hence, the value of  \texttt{\(m_{try}\)} needs to be created flexible based on the number features.
 
@@
 
-\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
+\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible\cite{schratz2019, sperrorest}.
 Each plot served as one fold within the resampling setup, resulting in four folds total.
-For the inner level (hyperparameter tuning), $p - 1$ folds were used (with $p$ being the number of plots).
+For the inner level (hyperparameter tuning), \(p - 1\) folds were used (with \(p\) being the number of plots).
 
@@
 
-\noindent The complete study was done using the open-source statistical programming language R \cite{rcoreteam2018}.
-The algorithm implementations of the following packages have been used: \textit{xgboost} \cite{chen2016} (\textit{xgboost}), \textit{kernlab} \cite{kernlab} (Support Vector Machine) and \textit{glmnet} \cite{glmnet} (Ridge Regression).
-The filter methods of the following packages were used: \textit{praznik \cite{praznik}}, \textit{FSelectorRcpp} \cite{fselectorrcpp}.
-The R package \textit{mlr} \cite{mlr} was used for all modeling related steps.
+\noindent The complete study was done using the open-source statistical programming language R\cite{rcoreteam2018}.
+The algorithm implementations of the following packages have been used: \textit{xgboost}\cite{chen2016} (\textit{xgboost}), \textit{kernlab}\cite{kernlab} (Support Vector Machine) and \textit{glmnet}\cite{glmnet} (Ridge Regression).
+The filter methods of the following packages were used: \textit{praznik\cite{praznik}}, \textit{FSelectorRcpp}\cite{fselectorrcpp}.
+The R package \textit{mlr}\cite{mlr} was used for all modeling related steps.
 \textit{drake} was used for structuring the work and ensuring reproducibility.
@@
 SVM showed the lowest error in four out of six tasks.
-Overall, the response variable "defoliation at trees" could be modeled with an error of 34 \%.
+Overall, the response variable 'defoliation at trees' could be modeled with an error of 34 \%.
 
@@
 % - all MBO penalized methods have no difference across tasks
-Random Forest combined with the "Carscore" filter showed the best absolute performance (RMSE of 3.396) (\autoref{tab:perf-top-15}).
-A noticeable difference between SVM and all other learners was visible for all feature sets without NRI variables (difference in RMSE of 3-4).
+Random Forest combined with the 'Carscore' filter showed the best absolute performance (RMSE of 3.396) (\autoref{tab:perf-top-15}).
+A noticeable difference between SVM and all other learners was visible for all
+feature sets without NRI variables (difference in RMSE of 3 --- 4).
 Considering a grouped performance by learners across all tasks and filters, SVM showed the most robust results (\autoref{tab:perf-top-15}).
 L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI) (\autoref{tab:perf-top-15}).
+Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI\@; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI) (\autoref{tab:perf-top-15}).
 All penalized methods tuned using MBO showed exactly the same error across all tasks.
@@
 RF showed a normal distribution for most tasks with one or two different methods taking the lead on each task.
-For RF, filter "Carscore" showed the best performance in three out of six tasks while this was achieved for XGBoost by filter "CMIM".
+For RF, filter 'Carscore' showed the best performance in three out of six tasks while this was achieved for XGBoost by filter 'CMIM'.
 XGBoost showed a diverse distribution with less clustering of filters as SVM and RF (referring to RMSE).
@@
 When comparing the usage of filters vs using no filter at all, there was only one instance (XGBoost on the HR-NRI task) when the model with no filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects}).
-In general, the effect of filtering vs.
-no filtering differed without a clear pattern across tasks: For example, SVM and RF showed a substantial increase for the NRI task when filtering the data while there was almost no effect for XGBoost (\autoref{fig:filter-effects}).
+In general, the effect of filtering vs.\ no filtering differed without a clear pattern across tasks: For example, SVM and RF showed a substantial increase for the NRI task when filtering the data while there was almost no effect for XGBoost (\autoref{fig:filter-effects}).
 For task HR, there was a medium effect of applying filtering for RF, almost none for SVM and a large one for XGBoost.
@@
 		\caption{Scored predictive performance (RMSE) of models across tasks.
-Prefix "CV" denotes that the learner was optimized using internal 10-fold CV while prefix "MBO" means that Bayesian optimization was used for hyperparameter optimization.
+Prefix 'CV' denotes that the learner was optimized using internal 10-fold CV while prefix 'MBO' means that Bayesian optimization was used for hyperparameter optimization.
 The abbreviations on the y-axis refer to the combinations of feature sets on which each model was scored on.
-Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = "Carscore", Info = "Information Gain", Borda = "Borda").
-The second value in the attached label of each point shows the scored RMSE value of the respective setting.}
-		\label{fig:perf-result}
+Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = 'Carscore', Info = 'Information Gain', Borda = 'Borda').
+The second value in the attached label of each point shows the scored RMSE value of the respective setting.}\label{fig:perf-result}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {filter-effect-1.pdf}
-        \caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}
-		\label{fig:filter-effects}
+		\caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}\label{fig:filter-effects}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {filter-perf-all-1.pdf}
-        \caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}
-		\label{fig:filter-perf-all}
+		\caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}\label{fig:filter-perf-all}
 	\end{center}
@@
 
-\subsection{Performance vs. plot characteristics}
+\subsection{Performance vs.\ plot characteristics}
 % TODO: Really needed?
@@
 
-\subsubsection{Variable importance vs. filter methods}
+\subsubsection{Variable importance vs.\ filter methods}
 
@@
 \noindent Most other studies analyzing defoliation operated on the plot rather than the tree level.
-This originated simply due to spatial resolution of the satellite products that served as the input data \cite{townsend2012, debeurs2008, rengarajan2016}.
-
-Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
-\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
+This originated simply due to spatial resolution of the satellite products that served as the input data\cite{townsend2012, debeurs2008, rengarajan2016}.
+
+Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
+\~cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
 Both study designs are substantially different compared to the setup of this work.
 In addition, no spatial \ac{CV} or \ac{FS} was used.
-\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
+\~cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
 Study results indicated that spectral metrics were found to be most helpful for the model.
@@
 
-\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
-\cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
+\~cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
+\~cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
 A grid-search was used for hyperparameter tuning and a \ac{FFS} for variable selection.
-\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach \cite{meyer2018} on a Random Forest classifier.
+\~cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach\cite{meyer2018} on a Random Forest classifier.
 
@@
 
-\appendices
+\appendices{}
 
@@
 		\includegraphics[width=0.48\textwidth] {spectral-signatures-1.pdf}
-\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}
-\label{fig:spectral-signatures}
+		\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}\label{fig:spectral-signatures}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {correlation-filter-nri-1.pdf}
-\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}
-\label{fig:correlation-filters}
+		\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}\label{fig:correlation-filters}
 	\end{center}
@@
 
-\section{Effect of different \texorpdfstring{$n_{bins}$}{nbins} values on filter "information gain"}
+\section{Effect of different \texorpdfstring{\(n_{bins}\)}{nbins} values on filter 'information gain'}
 
@@
 		\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
-\caption{Spearman correlation of filter information gain using different \texttt{$n_{bins}$} values for discretization of the numeric response.}
-\label{fig:correlation-nbins}
+		\caption{Spearman correlation of filter information gain using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.}\label{fig:correlation-nbins}
 	\end{center}

### 2020-02-09 docs/00-manuscripts/ieee/performance-best-per-learner.tex
move paper to docs/manuscript

--- 2019-12-10 code/98-paper/ieee/performance-best-per-learner.tex
@@
   \hline
-1 & HR & XGBoost & Borda & 30.56 & 13.37 \\ 
-  2 & HR & RF & MRMR & 31.44 & 15.08 \\ 
-  3 & HR-NRI-VI & SVM & Relief & 31.58 & 15.60 \\ 
-  4 & NRI & Ridge-CV & No Filter & 33.23 & 8.22 \\ 
+1 & HR & RF & Relief & 30.66 & 13.57 \\ 
+  2 & HR & XGBoost & Borda & 30.75 & 13.75 \\ 
+  3 & HR-NRI & SVM & Borda & 31.70 & 15.62 \\ 
+  4 & NRI & Ridge-CV & No Filter & 33.31 & 8.17 \\ 
   5 & HR & Lasso-CV & No Filter & 47.22 & 20.52 \\ 

### 2020-02-09 docs/00-manuscripts/journal/manuscript.tex
move paper to docs/manuscript

--- 2019-08-05 code/98-paper/journal/manuscript.tex
@@
 
+\begin{figure} [t!]
+	\begin{center}
+		\makebox[\textwidth]{\includegraphics[width=0.5\textwidth] {defol-grid-3000px.jpg}}
+        \caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
+		\label{fig:defol-trees}
+	\end{center}
+\end{figure}
+
 \subsection{Hyperspectral data}
@@
 
-\noindent Filter method "information gain" is by creation only defined for class responses:
+\noindent Filter method "information gain" is only defined for nominal response variables:
 
@@
 To be able to use this method with a numeric response, the variable is discretized into equal bins and treated as a class variable.
-While the number of bins can be treated as a hyperparameter of the filter method, the default of the R package implementation of \texttt{$n_{bin}$ = 5} was used after rank correlations of > 0.9 for different bin sizes have been explored.
-% TODO: Add reference to appendix, put plot into appendix
+While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{$n_{bin}$ = 10} after rank correlations of > 0.9 for different bin sizes were observed in a side analysis.
 

### 2020-02-09 code/05-modeling/paper/eda.R

@@ -1,88 +1,0 @@
-fv_nri_car <- generateFilterValuesData(nri_task, "carscore")
-fv_nri_info.gain <- generateFilterValuesData(nri_task, "FSelectorRcpp_information.gain")
-fv_nri_gain.ratio <- generateFilterValuesData(nri_task, "FSelectorRcpp_gain.ratio")
-fv_nri_symuncert <- generateFilterValuesData(nri_task, "FSelectorRcpp_symmetrical.uncertainty")
-fv_nri_relief <- generateFilterValuesData(nri_task, "FSelector_relief")
-fv_nri_rank <- generateFilterValuesData(nri_task, "rank.correlation")
-fv_nri_cor <- generateFilterValuesData(nri_task, "linear.correlation")
-fv_nri_mrmr <- generateFilterValuesData(nri_task, "praznik_MRMR")
-fv_nri_praznik_mrmr <- generateFilterValuesData(nri_task, "praznik_MRMR")
-fv_nri_cmim <- generateFilterValuesData(nri_task, "praznik_CMIM")
-fv_nri_var <- generateFilterValuesData(nri_task, "variance")
-
-fv_vi_car <- generateFilterValuesData(vi_task, "carscore")
-fv_vi_info.gain <- generateFilterValuesData(vi_task, "FSelectorRcpp_information.gain")
-fv_vi_gain.ratio <- generateFilterValuesData(vi_task, "FSelectorRcpp_gain.ratio")
-fv_vi_symuncert <- generateFilterValuesData(vi_task, "FSelectorRcpp_symmetrical.uncertainty")
-fv_vi_relief <- generateFilterValuesData(vi_task, "FSelector_relief")
-fv_vi_rank <- generateFilterValuesData(vi_task, "rank.correlation")
-fv_vi_cor <- generateFilterValuesData(vi_task, "linear.correlation")
-fv_vi_mrmr <- generateFilterValuesData(vi_task, "praznik_MRMR")
-fv_vi_cmim <- generateFilterValuesData(vi_task, "praznik_CMIM")
-fv_vi_var <- generateFilterValuesData(vi_task, "variance")
-
-fv_hr_car <- generateFilterValuesData(hr_task, "carscore")
-fv_hr_info.gain <- generateFilterValuesData(hr_task, "FSelectorRcpp_information.gain")
[…]

### 2020-02-09 code/05-modeling/paper/feature-importance.R

@@ -1,42 +1,0 @@
-fi_permut_vi <- feature_imp_parallel(vi_task,
-  learner = "regr.ksvm", nmc = 25, cpus = 25,
-  measure = list(
-    setAggregation(rmse, test.mean), setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean)
-  )
-)
-fi_permut_nri <- feature_imp_parallel(nri_task,
-  learner = "regr.xgboost", nmc = 25, cpus = 25,
-  measure = list(
-    setAggregation(rmse, test.mean), setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean)
-  )
-)
-fi_permut_hr <- feature_imp_parallel(hr_task,
-  learner = "regr.ksvm", nmc = 25, cpus = 25,
-  measure = list(
-    setAggregation(rmse, test.mean), setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean)
-  )
-)
-fi_permut_hr_vi <- feature_imp_parallel(hr_vi_task,
-  learner = "regr.ksvm", nmc = 25, cpus = 25,
-  measure = list(
-    setAggregation(rmse, test.mean), setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean)
-  )
-)
-fi_permut_hr_nri <- feature_imp_parallel(hr_nri_task,
-  learner = "regr.xgboost", nmc = 25, cpus = 25,
-  measure = list(
-    setAggregation(rmse, test.mean), setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean)
-  )
-)
-fi_permut_hr_nri_vi <- feature_imp_parallel(hr_nri_vi_task,
-  learner = "regr.xgboost", nmc = 25, cpus = 25,
-  measure = list(
-    setAggregation(rmse, test.mean), setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean)
-  )
-)

### 2020-02-09 code/05-modeling/paper/filter-wrapper.R

@@ -1,119 +1,0 @@
-# Random Forest -----------------------------------------------------------
-
-filter_wrapper_rf_borda <- makeFilterWrapper(lrn_rf,
-  fw.method = "E-Borda", cache = TRUE,
-  fw.base.methods = c(
-    "FSelectorRcpp_information.gain",
-    "linear.correlation",
-    "praznik_MRMR", "praznik_CMIM",
-    "carscore",
-    "FSelector_relief"
-  ),
-  more.args = list("FSelectorRcpp_information.gain" = list(equal = TRUE, nbins = 10)) # FSelectorRcpp
-)
-
-filter_wrapper_rf_gain.ratio <- makeFilterWrapper(lrn_rf, fw.method = "FSelectorRcpp_gain.ratio", cache = TRUE, equal = TRUE, nbins = 10)
-filter_wrapper_rf_info.gain <- makeFilterWrapper(lrn_rf, fw.method = "FSelectorRcpp_information.gain", cache = TRUE, equal = TRUE, nbins = 10)
-filter_wrapper_rf_relief <- makeFilterWrapper(lrn_rf, fw.method = "FSelector_relief", cache = TRUE)
-filter_wrapper_rf_variance <- makeFilterWrapper(lrn_rf, fw.method = "variance", cache = TRUE)
-filter_wrapper_rf_rank.cor <- makeFilterWrapper(lrn_rf, fw.method = "rank.correlation", cache = TRUE)
-filter_wrapper_rf_linear.cor <- makeFilterWrapper(lrn_rf, fw.method = "linear.correlation", cache = TRUE)
-filter_wrapper_rf_mrmr <- makeFilterWrapper(lrn_rf, fw.method = "praznik_MRMR", cache = TRUE)
-filter_wrapper_rf_cmim <- makeFilterWrapper(lrn_rf, fw.method = "praznik_CMIM", cache = TRUE)
-filter_wrapper_rf_carscore <- makeFilterWrapper(lrn_rf, fw.method = "carscore", cache = TRUE)
-
-# XGBOOST -----------------------------------------------------------------
-
-filter_wrapper_xgboost_borda <- makeFilterWrapper(lrn_xgboost,
-  fw.method = "E-Borda", cache = TRUE,
-  fw.base.methods = c(
-    "FSelectorRcpp_information.gain",
-    "linear.correlation",
-    "praznik_MRMR", "praznik_CMIM",
-    "carscore",
[…]

### 2020-02-09 code/05-modeling/paper/learner.R

@@ -1,59 +1,0 @@
-# Random Forest -----------------------------------------------------------
-
-lrn_rf <- makeLearner(
-  "regr.ranger.pow"
-)
-
-# XGBOOST -----------------------------------------------------------------
-
-lrn_xgboost <- makeLearner(
-  "regr.xgboost",
-  par.vals = list(
-    objective = "reg:linear",
-    eval_metric = "error"
-  )
-)
-
-# SVM ---------------------------------------------------------------------
-
-lrn_svm <- makeLearner(
-  "regr.ksvm",
-  scaled = FALSE,
-  fit = FALSE
-)
-
-# LASSO ---------------------------------------------------------------------
-
-lrn_lasso <- makeLearner("regr.glmnet",
-  id = "Lasso-MBO",
-  alpha = 1,
-  standardize = FALSE,
-  intercept = FALSE
-)
-
-lrn_lassocv <- makeLearner("regr.cvglmnet",
-  id = "Lasso-CV",
-  alpha = 1,
-  standardize = FALSE,
-  intercept = FALSE
-)
-
-# RIDGE ---------------------------------------------------------------------
-
-lrn_ridge <- makeLearner("regr.glmnet",
-  id = "Ridge-MBO",
-  alpha = 0,
-  standardize = FALSE,
-  intercept = FALSE
-)
-
-lrn_ridgecv <- makeLearner("regr.cvglmnet",
-  id = "Ridge-CV",
-  alpha = 0,
-  standardize = FALSE,
-  intercept = FALSE
-)
-
-# LM ---------------------------------------------------------------------
-
-lrn_lm <- makeLearner("regr.glm", id = "lm")

### 2020-02-09 code/05-modeling/paper/param-sets.R

@@ -1,176 +1,0 @@
-# XGBOOST -----------------------------------------------------------------
-
-ps_xgboost_filter <- makeParamSet(
-  makeIntegerParam("nrounds", lower = 10, upper = 600),
-  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
-  makeNumericParam("subsample", lower = 0.25, upper = 1),
-  makeIntegerParam("max_depth", lower = 1, upper = 10),
-  makeNumericParam("gamma", lower = 0, upper = 10),
-  makeNumericParam("eta", lower = 0.001, upper = 0.6),
-  makeNumericParam("min_child_weight", lower = 0, upper = 20),
-  makeNumericParam("fw.perc", lower = 0, upper = 1)
-)
-
-ps_xgboost <- makeParamSet(
-  makeIntegerParam("nrounds", lower = 10, upper = 600),
-  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
-  makeNumericParam("subsample", lower = 0.25, upper = 1),
-  makeIntegerParam("max_depth", lower = 1, upper = 10),
-  makeNumericParam("gamma", lower = 0, upper = 10),
-  makeNumericParam("eta", lower = 0.001, upper = 0.6),
-  makeNumericParam("min_child_weight", lower = 0, upper = 20)
-)
-
-ps_xgboost_pca <- makeParamSet(
-  makeIntegerParam("nrounds", lower = 10, upper = 600),
-  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
-  makeNumericParam("subsample", lower = 0.25, upper = 1),
-  makeIntegerParam("max_depth", lower = 1, upper = 10),
-  makeNumericParam("gamma", lower = 0, upper = 10),
-  makeNumericParam("eta", lower = 0.001, upper = 0.6),
-  makeNumericParam("min_child_weight", lower = 0, upper = 20),
-  makeIntegerParam("ppc.pcaComp", lower = 1, upper = 10)
-)
-
-# SVM ---------------------------------------------------------------------
-
-ps_svm_filter <- makeParamSet(
-  makeNumericParam("C",
-    lower = -10, upper = 10,
-    trafo = function(x) 2^x
-  ),
-  makeNumericParam("sigma",
-    lower = -5, upper = 5,
-    trafo = function(x) 2^x
-  ),
-  makeNumericParam("fw.perc", lower = 0, upper = 1)
-)
-
[…]

### 2020-02-09 code/05-modeling/paper/pca-wrapper.R

@@ -1,9 +1,0 @@
-pca_wrapper_svm <- makePreprocWrapperCaret(lrn_svm, ppc.pca = TRUE)
-
-pca_wrapper_xgboost <- makePreprocWrapperCaret(lrn_xgboost, ppc.pca = TRUE)
-
-pca_wrapper_rf <- makePreprocWrapperCaret(lrn_rf, ppc.pca = TRUE)
-
-pca_wrapper_ridge <- makePreprocWrapperCaret(lrn_ridge, ppc.pca = TRUE)
-
-pca_wrapper_lasso <- makePreprocWrapperCaret(lrn_lasso, ppc.pca = TRUE)

### 2020-02-09 code/05-modeling/paper/tune-ctrl.R

@@ -1,25 +1,0 @@
-tune.ctrl_xgboost <- makeTuneControlRandom(maxit = 100L)
-tune.ctrl_svm <- makeTuneControlRandom(maxit = 100L)
-tune.ctrl_rf <- makeTuneControlRandom(maxit = 100L)
-tune.ctrl_ridge <- makeTuneControlRandom(maxit = 100L)
-tune.ctrl_lasso <- makeTuneControlRandom(maxit = 100L)
-
-# mbo tuning, filter -----------------------------------------------------------
-
-tune.ctrl_xgboost_mbo_filter <- tune_ctrl_mbo_30n_70it(ps_xgboost_filter)
-tune.ctrl_svm_mbo_filter <- tune_ctrl_mbo_30n_70it(ps_svm_filter)
-tune.ctrl_rf_mbo_filter <- tune_ctrl_mbo_30n_70it(ps_rf_filter)
-
-# mbo tuning, pca --------------------------------------------------------------
-
-tune.ctrl_xgboost_mbo_pca <- tune_ctrl_mbo_30n_70it(ps_xgboost_pca)
-tune.ctrl_svm_mbo_pca <- tune_ctrl_mbo_30n_70it(ps_svm_pca)
-tune.ctrl_rf_mbo_pca <- tune_ctrl_mbo_30n_70it(ps_rf_pca)
-
-# mbo tuning, no filter --------------------------------------------------------
-
-tune.ctrl_xgboost_mbo <- tune_ctrl_mbo_30n_70it(ps_xgboost)
-tune.ctrl_svm_mbo <- tune_ctrl_mbo_30n_70it(ps_svm)
-tune.ctrl_rf_mbo <- tune_ctrl_mbo_30n_70it(ps_rf)
-tune.ctrl_ridge_mbo <- tune_ctrl_mbo_30n_70it(ps_ridge)
-tune.ctrl_lasso_mbo <- tune_ctrl_mbo_30n_70it(ps_lasso)

### 2020-02-09 code/05-modeling/paper/tune-wrapper-mbo.R

@@ -1,649 +1,0 @@
-# SVM ---------------------------------------------------------------------
-
-svm_borda_mbo <- makeTuneWrapper(filter_wrapper_svm_borda,
-  resampling = inner,
-  par.set = ps_svm_filter,
-  control = tune.ctrl_svm_mbo_filter, show.info = TRUE,
-  measures = list(
-    setAggregation(rmse, test.mean),
-    setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean),
-    timetrain, timeboth
-  )
-) %>%
-  magrittr::inset("id", "SVM Borda")
-
-svm_info.gain_mbo <- makeTuneWrapper(filter_wrapper_svm_info.gain,
-  resampling = inner,
-  par.set = ps_svm_filter,
-  control = tune.ctrl_svm_mbo_filter, show.info = TRUE,
-  measures = list(
-    setAggregation(rmse, test.mean),
-    setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean),
-    timetrain, timeboth
-  )
-) %>%
-  magrittr::inset("id", "SVM Info Gain")
-
-svm_gain.ratio_mbo <- makeTuneWrapper(filter_wrapper_svm_gain.ratio,
-  resampling = inner,
-  par.set = ps_svm_filter,
-  control = tune.ctrl_svm_mbo_filter, show.info = TRUE,
-  measures = list(
-    setAggregation(rmse, test.mean),
-    setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean),
-    timetrain, timeboth
-  )
-) %>%
-  magrittr::inset("id", "SVM Gain Ratio")
-
-svm_relief_mbo <- makeTuneWrapper(filter_wrapper_svm_relief,
-  resampling = inner,
-  par.set = ps_svm_filter,
-  control = tune.ctrl_svm_mbo_filter, show.info = TRUE,
-  measures = list(
-    setAggregation(rmse, test.mean),
-    setAggregation(rsq, test.mean),
-    setAggregation(expvar, test.mean),
-    timetrain, timeboth
-  )
-) %>%
-  magrittr::inset("id", "SVM Relief")
-
-svm_variance_mbo <- makeTuneWrapper(filter_wrapper_svm_variance,
-  resampling = inner,
-  par.set = ps_svm_filter,
-  control = tune.ctrl_svm_mbo_filter, show.info = TRUE,
[…]

### 2020-02-09 code/05-modeling/project/param-set.R

@@ -1,9 +1,0 @@
-ps_xgboost <- makeParamSet(
-  makeIntegerParam("nrounds", lower = 10, upper = 600),
-  makeNumericParam("colsample_bytree", lower = 0.3, upper = 0.7),
-  makeNumericParam("subsample", lower = 0.25, upper = 1),
-  makeIntegerParam("max_depth", lower = 1, upper = 10),
-  makeNumericParam("gamma", lower = 0, upper = 10),
-  makeNumericParam("eta", lower = 0.001, upper = 0.6),
-  makeNumericParam("min_child_weight", lower = 0, upper = 20)
-)

### 2020-02-09 code/05-modeling/project/task.R

@@ -1,22 +1,0 @@
-# filter 7 most important variables according to xgboost variable imp Mon Mar 11 22:57:56 2019 ------------------------------
-
-data_7_most_imp = nri_vi_data %>%
-  subset(select = c(defoliation, bf2_EVI, bf2_GDVI_4, bf2_GDVI_3, bf2_GDVI_2, bf2_D1, bf2_mNDVI, bf2_mSR))
-  #dplyr::select(defoliation, bf2_EVI, bf2_GDVI_4, bf2_GDVI_3, bf2_GDVI_2, bf2_D1, bf2_mNDVI, bf2_mSR) %>%
-  #as_tibble()
-
-# trim outliers of variables Mon Mar 11 22:59:05 2019 ------------------------------
-
-data_7_most_imp %<>%
-  mutate(bf2_EVI = case_when(bf2_EVI < 0 ~ 0, bf2_EVI > 1.5 ~ 1.5, TRUE ~ as.numeric(bf2_EVI))) %>%
-  mutate(bf2_mSR = case_when(bf2_mSR < -3 ~ -3, bf2_mSR > 4 ~ 4, TRUE ~ as.numeric(bf2_mSR))) %>%
-  mutate(bf2_D1 = case_when(bf2_D1 > 2.5 ~ 2.5, TRUE ~ as.numeric(bf2_D1))) %>%
-  mutate(bf2_GDVI_4 = case_when(bf2_GDVI_4 < -4 ~ -4, TRUE ~ as.numeric(bf2_GDVI_4))) %>%
-  mutate(bf2_GDVI_3 = case_when(bf2_GDVI_3 < -4 ~ -4, TRUE ~ as.numeric(bf2_GDVI_3))) %>%
-  mutate(bf2_GDVI_2 = case_when(bf2_GDVI_2 < -4 ~ -4, TRUE ~ as.numeric(bf2_GDVI_2))) %>%
-  mutate(bf2_mNDVI = case_when(bf2_mNDVI < -3 ~ -3, TRUE ~ as.numeric(bf2_mNDVI)))
-
-task_7_most_imp = makeRegrTask(id = "all_plots", data = data_7_most_imp,
-                               target = "defoliation", coordinates = coords_vi_nri_clean,
-                               blocking = factor(rep(1:4, c(479, 451, 300, 529))))
-

### 2020-02-09 code/05-modeling/project/train.R

@@ -1,2 +1,0 @@
-train_xgboost = train_wrapper("regr.xgboost", tune_object = tune_xgboost,
-                              task = task_7_most_imp)

### 2020-02-09 code/05-modeling/project/tune.R

@@ -1,6 +1,0 @@
-tune_xgboost = tune_wrapper(learner = lrn_xgboost,
-                            tune.control = tune_ctrl_xgboost, show.info = TRUE,
-                            par.set = ps_xgboost, resampling = resampl_inner,
-                            measure = list(rmse), task = task_7_most_imp
-                            )
-

### 2020-02-09 code/98-paper/ieee/jpg/create-img-grid.R

@@ -1,31 +1,0 @@
-# using imagemagick system library
-
-# paper ------------------------------------------------------------------------
-
-system2("montage", args = c("./code/98-paper/ieee/jpg/10-defol.jpg",
-                            "./code/98-paper/ieee/jpg/20-defol.jpg",
-                            "./code/98-paper/ieee/jpg/40-defol.jpg",
-                            "./code/98-paper/ieee/jpg/60-70-defol.jpg",
-                            "-geometry +20+20",
-                            "-tile 2x2",
-                            "./code/98-paper/ieee/jpg/defol-grid-2000px.jpg"))
-
-# resize
-system2("convert", args = c("./code/98-paper/ieee/jpg/defol-grid-2000px.jpg",
-                           "-resize 2000x2000",
-                           "./code/98-paper/ieee/jpg/defol-grid-2000px.jpg"))
-
-# presentation -----------------------------------------------------------------
-
-system2("montage", args = c("./code/98-paper/ieee/jpg/10-defol.jpg",
-                            "./code/98-paper/ieee/jpg/20-defol.jpg",
-                            "./code/98-paper/ieee/jpg/40-defol.jpg",
-                            "./code/98-paper/ieee/jpg/60-70-defol.jpg",
-                            "-geometry +20+20",
-                            "-tile 4x1",
-                            "./code/98-paper/ieee/jpg/defol-grid-pres-1500px.jpg"))
-
-# resize
-system2("convert", args = c("./code/98-paper/ieee/jpg/defol-grid-pres-1500px.jpg",
-                            "-resize 2000x1500",
-                            "./code/98-paper/ieee/jpg/defol-grid-pres-1500px.jpg"))

### 2020-02-13 docs/00-manuscripts/ieee/ieee-manuscript.tex
rewrite intro

@@
   % declare the path(s) where your graphic files are
-  \graphicspath{{pdf/}{jpg/}}
+  \graphicspath{{./pdf/}{./jpg/}}
   % and their extensions so you won't have to specify these with
@@
 	\acro{MEM}{maximum entropy model}
-	\acro{ML}{machine-learning}
+	\acro{ML}{machine learning}
 	\acro{NDII}{normalized difference infrared index}
@@
 \IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade~\cite{lary2016}.
-Especially since the launch of the first Sentinel satellite in 2014 the amount of freely available imagery data has increased greatly.
-At the same time, the implementation and usability of algorithms (both statistical and machine-learning) has been greatly simplified in the open-source community.
-Scientists can nowadays relatively easily feed large amounts of (environmental) information into complex models and analyze the data in a semi-automated way.
-This simplifies the analysis of (temporal) changes of the environment on a large scale\cite{ma2015}.
+This goes in line with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
+At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from open-source efforts.
+Scientists can nowadays relatively easily process large amounts of (environmental) information using various learning algorithms.
+This makes it possible to extend the matrix of possible options in a semi-automated way, possibly stumbling across unexpected findings of process settings that would have never been tested otherwise~\cite{ma2015}.
 
@@
 
-Machine-learning methods in combination with remote-sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping\cite{mascaro2014, urban2018}.
-The ability of predicting to large unknown areas qualifies these tools as a promising toolset for forest health analysis.
-This research field aims to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees\cite{hawrylo2018}.
-
-Approaches for analyzing forest health include temporal change detection\cite{zhang2016} or describing the current health status of forests on a stand level\cite{townsend2012}.
-In such studies, defoliation at trees is a variable able to serve as a proxy for forest health by describing the impact of biotic and abiotic pest triggers\cite{townsend2012, goodbody2018}.
-
-Vegetation indices have shown the potential to provide valuable information when analyzing such scenarios\cite{jiang2014, adamczyk2015}.
-Since both triggers and tree types are highly variable across regions, it cannot be known a-priori which indices are most important for describing the actual relationship.
-This emphasizes the need to extract as much information as possible from the available input data to generate features that can possibly explain the modeled relationship.
-Besides \ac{VI} a less known index type which can be calculated from remote-sensing products is the \ac{NRI}.
-\ac{NRI}s are arbitrary combinations of band information from remote-sensing data.
-Especially when working with hyperspectral data, hundreds of \ac{NRI}s can be derived and used during modeling.
+Machine learning methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping\cite{mascaro2014, urban2018}.
+The ability of predicting to large unknown areas qualifies these tools as a promising toolset for such tasks.
+One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees\cite{hawrylo2018}.
+
+Other approaches for analyzing forest health include temporal change detection\cite{zhang2016} or describing the current health status of forests on a stand level\cite{townsend2012}.
+In such studies, the defoliation of trees serves as a proxy for forest health by describing the impact of biotic and abiotic pest triggers\cite{townsend2012, goodbody2018}.
+
+Vegetation indices have shown the potential to provide valuable information when analyzing forest health\cite{jiang2014, adamczyk2015}.
+It is usually unclear which vegetation indices or spectral bands are most sensitive in reflecting changes of environmental features related to forest health.
+This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help understanding the modeled relationship.
+Besides vegetation indices a less known index type which can be derived from spectral information are normalized ration indices (NRI).
+In contrast to \ac{VI}, \ac{NRI}s are not based on an expert-based formulas following environmental heuristics but follow a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
+Especially when working with hyperspectral data, hundreds of \ac{NRI} features can be derived this way.
 

### 2020-02-14 analysis/eda.Rmd
style files using tidyverse style
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

--- 2019-07-01 analysis/eda.Rmd
@@
 drake::loadd(
-  vi_data, nri_data, bands_data, coords_vi_nri_clean,
+  vi_data_corrected_buffer2, nri_data_corrected_buffer2,
+  bands_data_corrected_buffer2, coords_vi_nri_clean_corrected,
 
-  fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
-  fv_nri_symuncert, fv_nri_relief,
-  fv_nri_rank, fv_nri_cor,
-  fv_nri_mrmr,
-  fv_nri_cmim, fv_nri_var,
+  filter_values,
+  filter_info_gain_nbins,
 
-  fv_vi_car, fv_vi_info.gain, fv_vi_gain.ratio,
-  fv_vi_symuncert, fv_vi_relief,
-  fv_vi_rank, fv_vi_cor,
-  fv_vi_mrmr,
-  fv_vi_cmim, fv_vi_var,
-
-  fv_hr_car, fv_hr_info.gain, fv_hr_gain.ratio,
-  fv_hr_symuncert, fv_hr_relief,
-  fv_hr_rank, fv_hr_cor,
-  fv_hr_mrmr,
-  fv_hr_cmim, fv_hr_var,
-
-  fv_hr_info.gain5, fv_hr_info.gain10,
-  fv_hr_info.gain15, fv_hr_info.gain20,
-  fv_hr_info.gain25, fv_hr_info.gain30,
-
-  fv_vi_info.gain5, fv_vi_info.gain10,
-  fv_vi_info.gain15, fv_vi_info.gain20,
-  fv_vi_info.gain25, fv_vi_info.gain30,
-
-  fv_nri_info.gain5, fv_nri_info.gain10,
-  fv_nri_info.gain15, fv_nri_info.gain20,
-  fv_nri_info.gain25, fv_nri_info.gain30
+  trees_with_bands_corrected_buffer2,
+  trees_with_bands_corrected_no_buffer
 )
@@
 ```{r eda-1 }
-intro <- introduce(vi_data)
+intro <- introduce(vi_data_corrected_buffer2)
 intro_df <- data.frame(
@@
 ```{r eda-2 }
-plot_histogram(vi_data)
+plot_histogram(vi_data_corrected_buffer2)
 ```
@@
 ```{r eda-3 }
-plot_prcomp(vi_data)
+plot_prcomp(vi_data_corrected_buffer2)
 ```
@@
 ```{r eda-4 }
-plot_correlation(vi_data)
+plot_correlation(vi_data_corrected_buffer2)
 ```
@@
 ```{r eda-5 }
-intro <- introduce(nri_data)
+intro <- introduce(nri_data_corrected_buffer2)
 intro_df <- data.frame(
@@
 ```{r eda-6 }
-plot_prcomp(nri_data)
+plot_prcomp(nri_data_corrected_buffer2)
 ```
@@
 ```{r eda-7 }
-plot_correlation(nri_data)
+plot_correlation(nri_data_corrected_buffer2)
 ```
@@
 ```{r eda-8 }
-intro <- introduce(bands_data)
+intro <- introduce(bands_data_corrected_buffer2)
 intro_df <- data.frame(
@@
 ```{r eda-9 }
-plot_histogram(bands_data)
+plot_histogram(bands_data_corrected_buffer2)
 ```
@@
 ```{r eda-10 }
-plot_prcomp(bands_data)
+plot_prcomp(bands_data_corrected_buffer2)
 ```
@@
 ```{r eda-11 }
-plot_correlation(bands_data)
+plot_correlation(bands_data_corrected_buffer2)
 ```
@@
 ```{r eda-12 }
-vi_data_plot <- vi_data %>%
+vi_data_plot <- vi_data_corrected_buffer2 %>%
   mutate(plot = factor(rep(
     c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-    c(479, 451, 300, 529)
+    c(559, 451, 301, 497)
   )))
@@
 dist <- map(plots, ~ {
-  coords <- coords_vi_nri_clean %>%
+  coords <- coords_vi_nri_clean_corrected %>%
     mutate(plot = factor(rep(
       c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-      c(479, 451, 300, 529)
+      c(559, 451, 301, 497)
     ))) %>%
@@
 ```
+
+# Effects of different buffer sizes when extracting values to trees
+
+The following buffer sizes were used to extract the remote sensing information (1 m spatial resolution) to the trees (points):
+
+- 0 m
+- 0.5 m
+- 1 m
+- 1.5 m
+- 2 m
+
+For this EDA only the direct reflectance values are used.
+It is assumed that the VI and NRI datasets will be affected in the same way by the different buffer sizes.
+
+```{r}
+compare_buffer <- function(trees_with_buffer, trees_without_buffer, plot_name) {
+  if (plot_name == "laukiz1") {
+    targets_sub_ids <- 1
+  } else if (plot_name == "laukiz2") {
+    targets_sub_ids <- 2
+  } else if (plot_name == "luiando") {
+    targets_sub_ids <- 3
+  } else if (plot_name == "oiartzun") {
+    targets_sub_ids <- 4
+  }
+
+  foo_buffer <- trees_with_buffer[seq(targets_sub_ids, 12, 4)]
+
+  foo_buffer <- foo_buffer %>%
+    map(~ sf::st_set_geometry(., NULL)) %>%
+    map(~ dplyr::summarise_at(., vars(starts_with("B")), mean, na.rm = TRUE)) %>%
+    map(~ tidyr::pivot_longer(., everything(),
+      names_to = "band", values_to = "reflectance"
+    )) %>%
+    dplyr::bind_rows(.id = "Buffer size")
+
+  foo_no_buffer <- trees_without_buffer[[targets_sub_ids]]
+
+  foo_no_buffer <- foo_no_buffer %>%
+    sf::st_set_geometry(NULL) %>%
+    dplyr::summarise_at(vars(starts_with("B")), mean) %>%
+    tidyr::pivot_longer(everything(),
+      names_to = "band", values_to = "reflectance"
+    ) %>%
+    dplyr::bind_rows(.id = "Buffer size") %>%
+    dplyr::mutate(`Buffer size` = "No Buffer")
+
+  all <- bind_rows(foo_buffer, foo_no_buffer) %>%
+    dplyr::mutate(band = as.numeric(stringr::str_replace(band, "B", "")))
+
+  plot <- ggplot(all, aes(x = band, y = reflectance)) +
+    geom_line(aes(color = `Buffer size`), size = 0.6) +
+    labs(title = tools::toTitleCase(plot_name)) +
+    ggsci::scale_color_nejm() +
+    ggpubr::theme_pubr(legend = "right")
+
+  return(plot)
+}
+```
+
+```{r}
+map(
+  c("laukiz1", "laukiz2", "luiando", "oiartzun"),
+  ~ compare_buffer(
+    trees_with_bands_corrected_buffer2,
+    trees_with_bands_corrected_no_buffer,
+    .x
+  )
+) %>%
+  patchwork::wrap_plots() +
+  patchwork::guide_area() +
+  patchwork::plot_layout(guides = "collect", ncol = 3, nrow = 2) +
+  patchwork::plot_annotation(
+    title = "Comparing the effect of different buffer sizes when extracting reflectance values to points obsverations (trees)",
+    subtitle = "Mean values of each hyperspectral band for each plot "
+  )
+```

### 2020-02-15 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-```{r}
+```{r buffer-size-comp}
 compare_buffer <- function(trees_with_buffer, trees_without_buffer, plot_name) {

### 2020-02-18 docs/00-manuscripts/ieee/ieee-manuscript.tex
rephrase intro

@@
 Vegetation indices have shown the potential to provide valuable information when analyzing forest health\cite{jiang2014, adamczyk2015}.
-It is usually unclear which vegetation indices or spectral bands are most sensitive in reflecting changes of environmental features related to forest health.
+Most vegetation indices were developed with the aim of being sensitive to changes of specific wavelength regions, serving as a proxy for underlying plant processes. 
+However, often enough indices developed for different purposes than the one to be analyzed can help explaining complex relationships.
 This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help understanding the modeled relationship.
-Besides vegetation indices a less known index type which can be derived from spectral information are normalized ration indices (NRI).
-In contrast to \ac{VI}, \ac{NRI}s are not based on an expert-based formulas following environmental heuristics but follow a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
+A less known index type which can be derived from spectral information is called normalized ratio index (NRI).
+In contrast to \ac{VI}, \ac{NRI}s are not based on an expert-based formulas following environmental heuristics but use a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
 Especially when working with hyperspectral data, hundreds of \ac{NRI} features can be derived this way.
@@
 
-Despite its popularity in environmental modeling, there are no studies so far that used machine-learning algorithms in combination with remote-sensing data to analyze defoliation on a tree level.
-This study aims at closing this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
-To cope with the state-of-the-art methodology in modeling, a methodology which combines feature-selection and hyperparameter tuning across multiple algorithms was used.
-The approach of generating many features comes along with the possible
-introduction of high-dimensionality into the analysis\cite{trunk1979,
-	xu2016}.
+Despite its popularity in environmental modeling, there are no studies so far which used machine learning algorithms in combination with remote sensing data to analyze defoliation on a tree level.
+This study aims to close this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
+We make use of the latest state-of-the-art methodology in machine learning by combining feature-selection and hyperparameter tuning across multiple algorithms.
+Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high-dimensionality in environmental modeling \cite{trunk1979, xu2016}.
 
 Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increase substantially and the interpretation of results becomes more complicated.
-Besides modeling defoliation at trees and interpreting the model results from an ecological point of view, this work also aims at providing an exemplary analysis of how to deal with spatial feature-rich datasets.
+At the same time, one has to deal with the presence of spatial autocorrelation in the data.
+The dataset used in this study comes with a spatial grouping at the plot level which needs to be accounted for.
 
@@
 
-	\item Do different environmental feature sets show substantial differences in performance when predicting defoliation at trees?
+	\item Do different environmental feature sets show differences in performance when modeling defoliation at trees?
 
@@
 
-	\item How are various feature-selection methods influencing the predictive performance of the models?
+	\item How are feature-selection methods influencing the predictive performance of the models?
 
@@
 
-\noindent State-of-the-art machine-learning techniques were compared using three feature sets across different feature-selection approaches including a relatively novel approach of ensemble filters.
-Model-based optimization was applied for hyperparameter tuning of the chosen algorithms which included an optimization of the feature set.
-Spatial block \ac{CV} was used to account for spatial autocorrelation in the data.
+% FIXME: Rephrase
+\noindent \ac{RF}, \ac{SVM} and extreme gradient boosting were compared using six feature sets across different feature selection methods, including a new promising approach of ensemble filters.
+Bayesian Optimization, also known as model-based optimization, was used for hyperparameter tuning.
+The feature set size was optimized at the same time.
+By this, we avoided multiple optimization layers which may result in a biased performance estimation or long model fitting times.
+A spatial \ac{CV} approach on the plot level was used to account for spatial autocorrelation in the data.
 

### 2020-02-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize filter methods section

@@
 \usepackage{amsmath}
+\newcommand*\mean[1]{\bar{#1}}
 % A popular package from the American Mathematical Society that provides
@@
 
+\usepackage[utf8]{inputenc}
+% enables custom font styles which are not available in IEEEtrans by default
+% https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran
+
+\renewcommand{\sfdefault}{cmss}
+\renewcommand{\rmdefault}{cmr}
+\renewcommand{\ttdefault}{cmt}
+
 \begin{document}
@@
 	\acro{GLM}{generalized linear model}
-	\acro{ICGC}{institut cartografic i geologic de catalunya}
+	\acro{ICGC}{Institut Cartografic I Geologic de Catalunya}
 	\acro{IQR}{interquartile range}
@@
 \noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
-The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum} leading to a spread of cankers or defoliation\cite{mesanza2016, iturritxa2017}.
-In-situ measurements of defoliation at trees (serving as a proxy for tree
-health) were collected to serve as the response variable \textit{defoliation}
-spanning a range from 0\% - 100\% (\autoref{fig:defol-distr}). % chktex 8
+The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation~\cite{mesanza2016, iturritxa2017}.
+In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0\% - 100\% (\autoref{fig:defol-distr}).
+% chktex 8
 The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage\cite{iturritxa2014}.
@@
 		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
-		\caption{Response variable defoliation at trees visualized as violin plots showing the variance of in-situ measurements for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.}\label{fig:defol-distr}
+		\caption{Response variable \textit{defoliation} at trees for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.
+\texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation.}\label{fig:defol-distr}
 	\end{center}
@@
 \noindent The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-\textit{Oiartzun} has the most observations (n = 529) while \textit{Laukiz 2} features the largest area size (1.44 ha).
+\textit{Oiartzun} has the most observations (n = 559) while \textit{Laukiz 2} shows the largest area size (1.44 ha).
 All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
-In total 1759 observations are available (\textit{Laukiz 1} = 479, \textit{Laukiz 2} = 451, \textit{Luiando} = 300, \textit{Oiartzun} = 529).
+In total 1808 observations are available (\textit{Laukiz 1} = 559, \textit{Laukiz 2} = 451, \textit{Luiando} = 301, \textit{Oiartzun} = 497).
 The data was surveyed in September 2016.
@@
 \subsection{Derivation of indices}
-\noindent To use the full information from the hyperspectral data, all possible vegetation indices supported by the R package \textit{hsdar} (89 in total) as well as all possible \ac{NRI} combinations were calculated.
+\noindent To use the full potential from the hyperspectral data, all possible vegetation indices supported by the R package $\left\{hsdar\right\}$ (89 in total) as well as all possible \ac{NRI} combinations were calculated.
 The following formula was used for the NRI calculation:
@@
 
-\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}), a buffer of two meters around the centroid of the respective tree was drawn.
-The value assigned to a tree observations was the mean of all pixels which touched the respective buffer around the tree.
+\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used during extraction of the reflectance values.
+Subsequently, the value assigned to a tree observations was the mean of all pixels which were touched by the buffer drawn around each tree.
+A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
+This is how the \texttt{extract()} function of the R package $\left\{raster\right\}$ handles buffer extraction.
 In total, \(\frac{125*126}{2} = 7875\) NRIs were calculated.
@@
 
-\noindent It is important to distinguish the term 'dimension reduction' from 'high-dimensionality'.
+\noindent We want to stress the differences between 'dimension reduction' and 'high-dimensionality'.
 The former refers to the general idea of reducing features from a dataset\cite{vandermaaten2007}.
 In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
-'High-dimensionality' in contrast is defined as a dataset attribute and applies
-when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of
-observations\cite{hastie2001}.
-There is no absolute value at which the term can be applied, both for \(n\) and \(p\).
+'High-dimensionality' in contrast is a dataset attribute and applies when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of observations\cite{hastie2001}.
+There is no absolute value at which the term applies, both for \(n\) and \(p\).
 Hence for this study the word ‘high-dimensionality‘ only applies to the
-experiments using the ‘NRI‘ feature set (7470 features for 1759 observations).
+experiments using the ‘NRI‘ feature set (> 7000 features for 1808 observations).
 
@@
 \begin{itemize}
-	\item Model fitting times increase
-	\item Noise is possibly introduced into models by highly-correlated variables\cite{johnstoneiainm.2009}
-	\item Model interpretation and prediction tasks become more complicated\cite{johnstoneiainm.2009}
+	\item Model fitting times increase.
+	\item Noise is possibly introduced into models by highly-correlated variables\cite{johnstoneiainm.2009}.
+	\item Model interpretation and prediction tasks become more challenging\cite{johnstoneiainm.2009}.
 \end{itemize}
 
-\noindent In the following sections a brief overview about sub-categories of feature-selection approaches is given.
-Due to the focus of this study on the use of filter methods, all other approaches were grouped into a single section.
+\noindent In the following sections a brief overview about sub-categories of feature selection approaches is given.
+Due to the focus of this study on the use of filter methods, which are a sub-group of feature selection methods, other approaches were grouped into a single section.
 
@@
 % Filter methods
-\noindent The concept of 'Filter methods' originates on the idea of ranking features using certain heuristics of an algorithm\cite{chandrashekar2014}.
-Some filter methods are restricted towards specific types of variables (numeric or nominal data).
-After the covariates have been ranked, a decision needs to be made what features to keep and which to discard\cite{drotar2015}.
-This step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
+\noindent The concept of filters originates on the idea of ranking features using certain heuristics of an algorithm\cite{chandrashekar2014}.
+Some filter methods are restricted towards specific types of variables (numeric or nominal).
+Filters do only rank features, they do not choose which covariates to drop or keep\cite{drotar2015}.
+The selection of features to keep for the model fitting step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
 Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
 The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
-In well-implemented software solutions the filter calculation is only done once
-and then cached, saving valuable resources\cite{mlr}.
+In well-implemented software solutions the filter calculation is only done once and then cached, saving comptutational resources\cite{mlr}.
 
@@
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables,
-studies showed that combining several filters using statistical operations such
-as 'minimum', 'mean' or 'sum' are able to enhance the predictive performance of
-the resulting models\cite{abeel2010, drotar2017a}.
-This approach is referred to as 'ensemble filter'\cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in machine-learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance\cite{polikar2012, feurer2015}.
-In this work the 'Borda' ensemble filter was applied\cite{drotar2017a}.
-The final order is the sum of all single filters ranking.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum', 'mean' or 'sum' are able to enhance the predictive performance of the resulting models\cite{abeel2010, drotar2017a}.
+This approach is referred to as 'ensemble filter'~\cite{dietterich2000}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance~\cite{polikar2012, feurer2015}.
+In this work the 'Borda' ensemble filter was applied~\cite{drotar2017a}.
+For this filter, the final order is the sum of all single filters.
 
@@
 Filter methods can be grouped into classes: Correlation based, entropy based, linear and non-linear methods.
-It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased.
-This was taken care of by calculating the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
-In case filters showed high correlations with each other, these were not included into the ensemble filter.
-This ensures that the ensemble filter is not biased towards a certain group of methods yielding highly similar rankings.
+It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased towards these.
+In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
+In case pairs of filters with a correlation above 0.9 were discovered, only one of these was included into the ensemble filter.
+By this we ensured that the ensemble filter is not biased towards a certain group of methods yielding highly similar rankings.
 
@@
 
-\noindent Filter method 'information gain' is only defined for nominal response variables:
+\noindent Filter 'Information Gain' is only defined for nominal response variables:
 
@@
 where \(H\) is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
-To be able to use this method with a numeric response, the variable is discretized into equal bins and treated as a class variable.
-While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of > 0.9 for different bin sizes were observed in a side analysis.
+To be able to use this method with a numeric response (defoliation in our case), the variable is discretized into equal bins and treated as a class variable.
+While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of $> 0.9$ for different bin sizes were observed in a side analysis.
 

### 2020-02-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize feature selection section

@@
 
-\noindent Other approaches to assess feature importance are 'wrapper methods' and the \ac{PCA}\cite{das2001, jolliffe2016}.
-A short introduction to both is given below.
+\noindent Other approaches to assess feature importance are so called 'wrapper methods' and the \ac{PCA}\cite{das2001, jolliffe2016}.
 
 % Wrapper approach
-\noindent 'Wrapper methods'\cite{chandrashekar2014, kohavi1997} apply algorithms that are used for hyperparameter optimization such as 'Random Search' or 'Generic Simulated Annealing'.
+\noindent Wrappers \cite{chandrashekar2014, kohavi1997} apply algorithms that are also used for hyperparameter optimization such as 'Random Search' or 'Generic Simulated Annealing'.
 First a (random) subset of features is chosen based on the selected algorithm.
-In comparison to 'filters', no ranking is done in this step.
+In comparison to filters, no ranking is done in this step.
 Now the model is fitted on the data and the performance is evaluated.
@@
 Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.
-Due to their extensive runtimes, 'wrapper methods' were not considered in this work.
+Due to their extensive runtimes, wrappers were not considered in this work.
 
 % PCA
-A method with a completely different approach compared to filters and wrappers is the 'Principal Component Analysis'\cite{pearson1901, jolliffe2016}.
+A method with a completely different approach compared to filters and wrappers is the \ac{PCA}\cite{pearson1901, jolliffe2016}.
 Here, the main components of the feature space are extracted and combined.

### 2020-02-19 analysis/filter-correlation.Rmd

@@
 drake::loadd(
-  fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
-  fv_nri_symuncert, fv_nri_relief,
-  fv_nri_rank, fv_nri_cor,
-  fv_nri_mrmr,
-  fv_nri_cmim, fv_nri_var,
+  filter_values,
+  filter_values_fselectorrcpp,
 
-  fv_vi_car, fv_vi_info.gain, fv_vi_gain.ratio,
-  fv_vi_symuncert, fv_vi_relief,
-  fv_vi_rank, fv_vi_cor,
-  fv_vi_mrmr,
-  fv_vi_cmim, fv_vi_var,
-
-  fv_hr_car, fv_hr_info.gain, fv_hr_gain.ratio,
-  fv_hr_symuncert, fv_hr_relief,
-  fv_hr_rank, fv_hr_cor,
-  fv_hr_mrmr,
-  fv_hr_cmim, fv_hr_var,
-
-  fv_hr_info.gain5, fv_hr_info.gain10,
-  fv_hr_info.gain15, fv_hr_info.gain20,
-  fv_hr_info.gain25, fv_hr_info.gain30,
-
-  fv_vi_info.gain5, fv_vi_info.gain10,
-  fv_vi_info.gain15, fv_vi_info.gain20,
-  fv_vi_info.gain25, fv_vi_info.gain30,
-
-  fv_nri_info.gain5, fv_nri_info.gain10,
-  fv_nri_info.gain15, fv_nri_info.gain20,
-  fv_nri_info.gain25, fv_nri_info.gain30
+  filter_info_gain_nbins
 )
@@
 
+The idea behind is was to analyze the correlation between filter rankings.
+We only wanted to included filters which have a somewhat unique ranking.
+Otherwise, when creating ensemble filters, certain filters would implicitly be weighted more than others.
+
+Takeaway:
+
+- Only use one of "information gain", "gain ratio", "sym uncert"
+
+- Either use Spearman or Pearson correlation
+
 ```{r filter-correlation-1, warning=FALSE}
-# combine filter values
-
-fv_all <- dplyr::bind_rows(
-  fv_vi_car$data,
-  fv_vi_info.gain$data,
-  fv_vi_gain.ratio$data,
-  fv_vi_symuncert$data,
-  fv_vi_relief$data,
-  fv_vi_rank$data,
-  fv_vi_cor$data,
-  fv_vi_mrmr$data,
-  fv_vi_cmim$data,
-  fv_vi_var$data
-) %>%
+fv_all_vi <- map(filter_values[8:14], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:89, 10)) %>%
+  mutate(id = rep(1:89, 7)) %>%
   tidyr::spread(filter, value) %>%
-  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
-  na.omit() %>% # NA in cmim
   dplyr::select(-id)
 
-corr <- cor(fv_all, method = "spearman")
-# p.mat = cor_pmat(corr)
-
-ggcorrplot(corr,
-  # p.mat = p.mat,
-  hc.order = F,
-  type = "lower",
-  lab = TRUE
-)
-```
-
-### Number of bins of `FSelectorRcpp::information.gain()`
-
-```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
-# recode levels so that they are unique
-fv_all_list <- map2(
-  list(
-    fv_vi_info.gain5$data, fv_vi_info.gain10$data,
-    fv_vi_info.gain15$data, fv_vi_info.gain20$data,
-    fv_vi_info.gain25$data, fv_vi_info.gain30$data
-  ),
-  list(5, 10, 15, 20, 25, 30), ~
-  dplyr::mutate(.x, filter = recode(.x$filter, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}")))
-)
-
-fv_all <- dplyr::bind_rows(
-  fv_all_list
-) %>%
+fv_all_vi_fselecrcpp <- map(filter_values_fselectorrcpp[4:6], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:89, 6)) %>%
+  mutate(id = rep(1:89, 3)) %>%
   tidyr::spread(filter, value) %>%
-  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
-  na.omit() %>%
   dplyr::select(-id)
 
-corr <- cor(fv_all, method = "spearman")
+fv_all_vi_merged <- bind_cols(fv_all_vi, fv_all_vi_fselecrcpp)
 
-ggcorrplot(corr,
+corr_vi <- cor(fv_all_vi_merged, method = "spearman")
+
+ggcorrplot(corr_vi,
   # p.mat = p.mat,
@@
 ```{r correlation-filter-nri, dev = c("png", "pdf"), warning=FALSE}
-# combine filter values
-
-fv_all <- dplyr::bind_rows(
-  fv_nri_car$data,
-  fv_nri_info.gain$data,
-  fv_nri_gain.ratio$data,
-  fv_nri_symuncert$data,
-  fv_nri_relief$data,
-  fv_nri_rank$data,
-  fv_nri_cor$data,
-  fv_nri_mrmr$data,
-  fv_nri_cmim$data,
-  fv_nri_var$data
-) %>%
+fv_all_nri <- map(filter_values[15:21], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:7381, 10)) %>%
+  mutate(id = rep(1:7381, 7)) %>%
   tidyr::spread(filter, value) %>%
@@
 
-corr <- cor(fv_all, method = "spearman")
-p.mat <- cor_pmat(corr)
+fv_all_nri_fselecrcpp <- map(filter_values_fselectorrcpp[7:9], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
+  dplyr::select(value, filter) %>%
+  mutate(id = rep(1:7381, 3)) %>%
+  tidyr::spread(filter, value) %>%
+  dplyr::select(-id)
 
-ggcorrplot(corr,
+fv_all_nri_merged <- bind_cols(fv_all_nri, fv_all_nri_fselecrcpp)
+
+corr_nri <- cor(fv_all_nri_merged, method = "spearman")
+
+ggcorrplot(corr_nri,
   # p.mat = p.mat,
-  hc.order = FALSE,
+  hc.order = F,
   type = "lower",
@@
 ```{r filter-correlation-4, warning=FALSE}
-# combine filter values
-
-fv_all <- dplyr::bind_rows(
-  fv_hr_car$data,
-  fv_hr_info.gain$data,
-  fv_hr_gain.ratio$data,
-  fv_hr_symuncert$data,
-  fv_hr_relief$data,
-  fv_hr_rank$data,
-  fv_hr_cor$data,
-  fv_hr_mrmr$data,
-  fv_hr_cmim$data,
-  fv_hr_var$data
-) %>%
+fv_all_hr <- map(filter_values[1:7], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:122, 10)) %>%
+  mutate(id = rep(1:122, 7)) %>%
   tidyr::spread(filter, value) %>%
@@
 
-corr <- cor(fv_all, method = "spearman")
-p.mat <- cor_pmat(corr)
+fv_all_hr_fselecrcpp <- map(filter_values_fselectorrcpp[1:3], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
+  dplyr::select(value, filter) %>%
+  mutate(id = rep(1:122, 3)) %>%
+  tidyr::spread(filter, value) %>%
+  dplyr::select(-id)
 
-ggcorrplot(corr,
+fv_all_hr_merged <- bind_cols(fv_all_hr, fv_all_hr_fselecrcpp)
+
+corr_hr <- cor(fv_all_hr_merged, method = "spearman")
+
+ggcorrplot(corr_hr,
+  # p.mat = p.mat,
+  hc.order = F,
+  type = "lower",
+  lab = TRUE
+)
+```
+
+## Number of bins of `FSelectorRcpp::information.gain()`
+
+Analyzing the effect of a different `nbins` value on the filter values of filter "Information Gain".
+
+- Lower correlation / highest difference: `nbins = 5` vs. `nbins = 30`
+
+-> We decided to use with `nbins = 10` in the analysis.
+
+The hidden default of nbins when setting `equal = TRUE` in `FSelectorRcpp::information_gain()` is 5.
+
+```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
+# recode levels so that they are unique
+fv_info_gain <- map(filter_info_gain_nbins[1:6], ~ .x[["data"]]) %>%
+  dplyr::bind_rows() %>%
+  dplyr::select(value, filter) %>%
+  mutate(id = rep(seq(5, 30, 5), each = 122)) %>%
+  tidyr::unite(filter, c("filter", "id")) %>%
+  mutate(id = rep(seq(1, length(.data$value) / 6), 6)) %>%
+  tidyr::pivot_wider(
+    names_from = filter,
+    values_from = value
+  ) %>%
+  dplyr::select(-id)
+
+corr_inf_gain <- cor(fv_info_gain, method = "spearman")
+
+ggcorrplot(corr_inf_gain,
   # p.mat = p.mat,

### 2020-02-23 docs/00-manuscripts/ieee/ieee-manuscript.tex

@@
 
+\usepackage[detect-all]{siunitx}
+
 \begin{document}
@@
 	\acro{MARS}{multivariate adaptive regression splines}
+	\acro{MBO}{model-based optimization}
 	\acro{MEM}{maximum entropy model}
@@
 		\toprule
-		Linear correlation (Pearson)                 & univariate, linear, correlation   &\cite{pearson1901} \\
-		Information gain                             & univariate, non-linear, entropy   &\cite{quinlan1986} \\
-		Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy &\cite{zhao2013}    \\
-		Carscore                                     & multivariate, linear, correlation &\cite{zuber2011}   \\
-		Relief                                       & multivariate, linear, entropy,    &\cite{kira1992}    \\
-		Conditional minimal information maximization & multivariate, linear, entropy     &\cite{fleuret2004}
+		Linear correlation (Pearson)                 & univariate, linear, correlation   & \cite{pearson1901} \\
+		Information gain                             & univariate, non-linear, entropy   & \cite{quinlan1986} \\
+		Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy & \cite{zhao2013}    \\
+		Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
+		Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
+		Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
 	\end{tabular}
@@
 	\item  Random Forest (RF)
-	\item  Penalized Regression (both L1 and L2)
+	\item  Penalized Regression (both L1 (Lasso) and L2 (Ridge))
 	\item  Support Vector Machine (SVM)
@@
 \noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
-Extreme gradient boosting (commonly referred to as \ac{XGBOOST}) showed promising results in benchmark competitions in recent years.
+\ac{XGBOOST} (commonly abbreviated as XGBOOST) showed promising results in benchmark competitions in recent years.
 Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
@@
 
-\noindent Three feature sets were used in this study with each representing a different way of feature-engineering:
+\noindent Three feature sets were used in this study with each representing a different way of feature engineering:
 
 \begin{itemize}
-	\item The raw hyperspectral band information (HR: No feature engineering) %chktex 13
-	\item Vegetation Indices (\ac{VI}: Expert-based feature engineering)
-	\item Normalized Ratio Indices (\ac{NRI}: Automated feature-engineering)
+	\item The raw hyperspectral band information (HR): No feature engineering) %chktex 13
+	\item Vegetation Indices (\ac{VI}): Expert-based feature engineering)
+	\item Normalized Ratio Indices (\ac{NRI}): Data-driven feature engineering)
 \end{itemize}
 
+The idea of splitting the features into different sets originated from the question whether feature engineered indices from reflectance values have a positive effect on model performance.
+Benchmarking all models on these distinct groups of features makes it easier to draw conclusions on their impact when keeping all other variables such as model type, tuning strategy and partioning method constant.
+However, rather than only looking at these three groups we decided to also take their combinations into account for the overall comparison:
+
+\begin{itemize}
+	\item HR + VI %chktex 13
+	\item HR + NRI
+	\item HR + VI + NRI
+\end{itemize}
+
 \subsubsection{Hyperparameter Optimization}
@@
 \noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
-Maximum Bayesian Optimization\cite{mlrmbo} (MBO) was used for parameter optimization.
+Model-based Optimization\cite{mlrmbo} (MBO) was used for hyperparameter optimization.
 This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
-After these \textit{n} tries have been evaluated, a new hyperparameter setting, which is going to be evaluated next, is proposed based on a fitted regression model.
-The regression model estimates the performance of the machine-learning method for unknown hyperparameter settings.
-Using these estimates, a new promising hyperparameter setting is proposed to be evaluated next.
+After these \textit{n} tries have been evaluated, a new hyperparameter setting, going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
 This strategy continues until a termination criterion, defined by the user, is reached\cite{hutter2011, jones1998}.
+
 An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum
-This applies when being compared to methods that do not use information from previous runs, such as random search or grid search\cite{bergstra2012}.
-In addition to MBO, the penalization term of Lasso and Ridge was optimized using the internal non-spatial CV approach due to promising results in other applications.
+The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum.
+\ac{MBO} shines when being compared to methods that do not use information from previous runs, such as random search or grid search\cite{bergstra2012}.
+For the Lasso and Ridge models however the internal non-spatial CV approach due to promising results in previous applications.
 
 For the filter methods, the percentage of features was added to the models as a hyperparameter.
-The number of main components was tuned for the instances which used \ac{PCA} instead of filtering.
-Random Forest hyperparameter \texttt{\(m_{try}\)} was transformed from taking absolute values to a relative parameter depending on the number of tasks: \(p^{m_{try}}\), where \(p\) is the number of features.
-This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen out of bounds during tuning:
+For \ac{PCA}, the number of main components was tuned instead.
+Random Forest hyperparameter \texttt{\(m_{try}\)} was modified into power transformed version \(p^{m_{try}}\) (ranging from 1 to $p$) to work on relative values of the respective datasets feature count.
+Here \(p\) is the number of features of the dataset.
+This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen out of bounds during tuning with respect to the datasets feature count.
 After the filtering, only a subset of features of the task is used for optimizing the models hyperparameters.
 The size of this subset is always different since the number of features is also optimized.
-Hence, the value of  \texttt{\(m_{try}\)} needs to be created flexible based on the number features.
+Hence, the value of \texttt{\(m_{try}\)} needs to be created dynamically based on the datasets feature count.
 
@@
 
-\noindent A spatial block nested cross-validation was chosen to reduce the influence of spatial autocorrelation as much as possible\cite{schratz2019, sperrorest}.
+\noindent A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible\cite{schratz2019, sperrorest}.
 Each plot served as one fold within the resampling setup, resulting in four folds total.
@@
 
-In total the benchmarking matrix consisted of 120 experiments (3 feature sets, 5 algorithms, 8 feature-selection methods).
+In total the benchmarking matrix consisted of 156 experiments (6 feature sets * 3 ML algorithms * 8 feature-selection methods + 2 * 6 models of the penalized algorithms).
 

### 2020-02-24 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-R.utils::sourceDirectory("R")
+source(here::here("code/99-packages.R"))
+load_packages()
 
 # load drake objects
-drake::loadd(
-  bm_aggregated
+loadd(
+  bm_aggregated_new
 )
-library("mlr")
-library("magrittr")
-library("dplyr")
-library("ggplot2")
+library("xtable")
+library("flextable")
+library("ggbeeswarm")
+library("ggsci")
 library("ggrepel")
-library("ggsci")
 library("ggpubr")
-library("ggbeeswarm")
-library("flextable")
-library("xtable")
 ```
@@
 ```{r}
-df_perf <- getBMRPerformances(bm_aggregated, as.df = T) %>%
+df_perf <- getBMRPerformances(bm_aggregated_new, as.df = T) %>%
   mutate(task.id = recode_factor(task.id,
-    `defoliation-all-plots-HR` = "HR",
-    `defoliation-all-plots-VI` = "VI",
-    `defoliation-all-plots-NRI` = "NRI",
-    `defoliation-all-plots-HR-NRI` = "HR-NRI",
-    `defoliation-all-plots-HR-VI` = "HR-VI",
-    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+    `hr` = "HR",
+    `vi` = "VI",
+    `nri` = "NRI",
+    `nri_vi` = "NRI_VI",
+    `hr_nri` = "HR-NRI",
+    `hr_vi` = "HR-VI",
+    `hr_nri_vi` = "HR-NRI-VI",
   )) %>%
@@
     remove = FALSE,
-    sep = " "
+    sep = " MBO "
   )
@@
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
-  select(-iter, -`Learner ID`, -rsq, -expvar) %>%
+  select(-iter, -`Learner ID`) %>%
   arrange(RMSE)
@@
   print(
-    file = "code/98-paper/ieee/performance-top-20.tex",
+    file = "docs/00-manuscripts/ieee/performance-top-20.tex",
     include.rownames = TRUE,
@@
 
-saveRDS(table1, "code/98-paper/presentation/table-perf.rda")
+saveRDS(table1, "docs/00-manuscripts/presentation/table-perf.rda")
 
@@
   ) %>%
-  select(-rsq, -expvar, -iter) %>%
+  select(-iter) %>%
   select(-`Learner ID`)
@@
   print(
-    file = "code/98-paper/ieee/performance-best-per-learner.tex",
+    file = "docs/00-manuscripts/ieee/performance-best-per-learner.tex",
     include.rownames = TRUE,
@@
 
-saveRDS(table2, "code/98-paper/presentation/table-best-learner-per-task.rda")
+saveRDS(table2, "docs/00-manuscripts/ieee/presentation/table-best-learner-per-task.rda")
 
@@
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 70)) +
+  scale_x_continuous(limits = c(27, 47)) +
   geom_label_repel(
@@
 
+**DEPRECATED**: Please see the scatterplots below for a better visulization
+
 Showing the final effect of applying feature selection to a learner for each task.
@@
   mutate(task.id = recode_factor(task.id,
-    `defoliation-all-plots-HR` = "HR",
-    `defoliation-all-plots-VI` = "VI",
-    `defoliation-all-plots-NRI` = "NRI",
-    `defoliation-all-plots-HR-NRI` = "HR-NRI",
-    `defoliation-all-plots-HR-VI` = "HR-VI",
-    `defoliation-all-plots-HR-NRI-VI` = "HR-NRI-VI",
+    `hr` = "HR",
+    `vi` = "VI",
+    `nri` = "NRI",
+    `nri_vi` = "NRI_VI",
+    `hr_nri` = "HR-NRI",
+    `hr_vi` = "HR-VI",
+    `hr_nri_vi` = "HR-NRI-VI",
   )) %>%
@@
     remove = FALSE,
-    sep = " "
+    sep = " MBO "
   ) %>%
@@
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 50)) +
+  scale_x_continuous(limits = c(27, 47)) +
   theme_pubr() +
@@
 results_aggr1 <- df_perf %>%
-  filter(learner_group != "Ridge-MBO", learner_group != "Ridge-CV") %>%
-  filter(learner_group != "Lasso-MBO", learner_group != "Lasso-CV") %>%
-  mutate(filter = replace(filter, is.na(filter), "NF")) %>%
+  filter(learner_group != "Ridge-CV") %>%
+  filter(learner_group != "Lasso-CV") %>%
+  # mutate(filter = replace(filter, "No Filter", "NF")) %>%
   mutate(learner_group = as.factor(learner_group)) %>%
   mutate(learner_group = recode(learner_group, `XGBoost` = "XG")) %>%
+  mutate(filter = recode(filter, `No Filter` = "NF")) %>%
+  mutate(learner_group = fct_rev(learner_group)) %>%
   group_by(learner_group, task.id, filter) %>%
@@
   facet_wrap(~task.id) +
-  scale_color_nejm() +
+  scale_color_nejm(guide = guide_legend(override.aes = list(shape = c(3, 19)))) +
   labs(x = "RMSE", y = "Task", colour = NULL) +
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 50)) +
+  scale_x_continuous(limits = c(27, 51)) +
   theme_pubr() +
@@
   na.omit() %>%
-  filter(learner_group != "Ridge-MBO", learner_group != "Ridge-CV") %>%
-  filter(learner_group != "Lasso-MBO", learner_group != "Lasso-CV") %>%
+  filter(learner_group != "Ridge-CV") %>%
+  filter(learner_group != "Lasso-CV") %>%
   mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%
@@
   # summarise() working
-  summarise(perf = mean(rmse_aggr))
+  summarise(perf = mean(rmse_aggr)) %>% 
+  ungroup()
 
 results_aggr2 %>%
-  ggplot(aes(x = perf, y = learner_group)) +
+  ggplot(aes(x = perf, y = fct_reorder(learner_group, -perf))) +
   geom_beeswarm(
     data = results_aggr2[results_aggr2$filter != "Borda", ],
-    shape = 3, size = 1.1, aes(color = "Simple Filter"),
+    shape = 3, size = 1.1, aes(color = "Filter"),
     groupOnX = FALSE
@@
   facet_wrap(~task.id) +
-  scale_color_nejm() +
+  scale_color_manual(guide = guide_legend(override.aes = list(shape = c(19, 3))),
+                     values = c("Filter" = "#BC8C29FF",
+                     "Borda Filter" = "#0072B5FF")) +
   labs(x = "RMSE", y = "Task", colour = NULL) +
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(30, 48)) +
+  scale_x_continuous(limits = c(27, 51)) +
   theme_pubr() +

### 2020-02-24 analysis/about.Rmd
add jose to authors

@@
 **Patrick Schratz** (patrick.schratz@gmail.com) [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0003-0748-6624)
+
 Jannes Muenchow [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-7834-4717)
+
+Eugenia Iturritxa [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0002-0577-3315)
+
+José Cortés
+
 Bernd Bischl [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6002-6980)
+
 Alexander Brenning [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6640-679X)

### 2020-02-24 analysis/eda.Rmd
do everything for buffer of 2m and no buffer

@@
 
+library("drake")
+library("magrittr")
 R.utils::sourceDirectory("R")
+R.utils::sourceDirectory("code")
 
 # load drake objects
-drake::loadd(
+loadd(
   vi_data_corrected_buffer2, nri_data_corrected_buffer2,
@@
 library("raster")
+library("fs")
 ```
@@
   annotate("text",
-    label = expression(bold(atop("n = 479", bar(x) ~ "= 55.96"))), x = 1,
+    label = expression(bold(atop("n = 559", bar(x) ~ "= 57.23"))), x = 1,
     y = 112, size = 3, colour = "#BC3C29", fontface = 2
@@
   annotate("text",
-    label = expression(bold(atop("n = 300", bar(x) ~ "= 68.44"))), x = 3,
+    label = expression(bold(atop("n = 301", bar(x) ~ "= 68.36"))), x = 3,
     y = 112, size = 3, colour = "#E18727", fontface = 2
@@
   annotate("text",
-    label = expression(bold(atop("n = 529", bar(x) ~ "= 69.22"))), x = 4,
+    label = expression(bold(atop("n = 407", bar(x) ~ "= 69.07"))), x = 4,
     y = 112, size = 3, colour = "#20854E", fontface = 2
@@
 ```
+
+```{r echo=FALSE}
+copy_figures()
+```

### 2020-02-24 code/042-data-preprocessing-no-buffer.R

@@ -1,0 +1,90 @@
+data_preprocessing_plan_no_buffer <- drake_plan(
+
+  # read all datasets as list ----------------------------------------------------
+    data_list_nri_vi_bands_corrected_no_buffer = target(
+    list(
+      "vi_nri" = trees_with_indices_corrected_no_buffer,
+      "bands" = trees_with_bands_corrected_no_buffer
+    )
+  ),
+
+  # clean datasets ---------------------------------------------------------------
+
+  data_clean_single_plots_corrected_no_buffer = target(
+    map(
+      data_list_nri_vi_bands_corrected_no_buffer,
+      ~ clean_single_plots(
+        .x,
+        # marcos tree data col names differ among plots, hence the warnings of
+        # somce columns not being available
+        c(
+          "diameter", "height", "zona", "tree.numbe", "DECOLORATI", "CANKER",
+          "TREE_NUMBE", "heigth", "CANKERS", "tree_numbe", "geom", "ID"
+        ),
+        remove_coords = TRUE
+      )
+    )
+  ),
+
+  # Merge the dataframes of vi and nri -------------------------------------------
+
+  # (stored in separate lists for each plot) into one dataset.
+  # Possible variables with NAs are removed again, but this is
+  # actually happening in `clean_single_plots()`.
+
+  data_vi_nri_clean_corrected_no_buffer = target(
+    as_tibble(rbindlist(data_clean_single_plots_corrected_no_buffer[[1]], fill = TRUE)) %>%
+      Filter(function(x) !any(is.na(x)), .)
+  ),
+
+  data_bands_clean_corrected_no_buffer = target(
+    as_tibble(rbindlist(data_clean_single_plots_corrected_no_buffer[[2]], fill = TRUE)) %>%
+      Filter(function(x) !any(is.na(x)), .)
+  ),
+
+  # modify defoliation value (we get errors if defol is exactly zero) ------------
+
+  data_trim_defoliation_corrected_no_buffer = target(
+    map(
+      list(data_vi_nri_clean_corrected_no_buffer,
+           data_bands_clean_corrected_no_buffer),
+      ~ mutate_defol(.x)
+    ) %>%
+      set_names(c("data_vi_nri", "data_bands"))
+  ),
+
+  # split into feature sets ------------------------------------------------------
+
+  nri_data_corrected_no_buffer = target(
[…]

### 2020-02-24 code/05-tasks.R

@@ -1,254 +1,0 @@
-tasks_plan <- drake_plan(
-
-  data_old = list(
-    bands_data,
-    vi_data,
-    nri_data,
-    hr_nri_data,
-    nri_vi_data,
-    hr_nri_vi_data
-  ),
-
-  coordinates_list_corrected = list(
-    coords_bands_clean_corrected,
-    coords_vi_nri_clean_corrected,
-    coords_vi_nri_clean_corrected,
-    coords_vi_nri_clean_corrected,
-    coords_vi_nri_clean_corrected,
-    coords_vi_nri_clean_corrected
-  ),
-
-  coordinates_list = list(
-    coords_bands_clean,
-    coords_vi_nri_clean,
-    coords_vi_nri_clean,
-    coords_vi_nri_clean,
-    coords_vi_nri_clean,
-    coords_vi_nri_clean
-  ),
-
-  data_corrected =
-    list(
-      bands_data_corrected,
-      vi_data_corrected,
-      nri_data_corrected,
-      hr_nri_data_corrected,
-      nri_vi_data_corrected,
-      hr_nri_vi_data_corrected
-    ),
-
-  # Non-log transformed tasks ----------------------------------------------------
-  task_old = target(
-    makeRegrTask(task_names,
-                 data_old[[1]],
-      target = "defoliation",
-      coordinates = coordinates_list[[1]],
-      blocking = factor(rep(1:4, c(479, 451, 300, 529))),
-    ),
-    transform = map(data_old,
-      task_names = c(
-        "hr",
-        "vi",
-        "nri",
-        "hr_nri",
-        "nri_vi",
-        "hr_nri_vi"
-      ),
-      coordinates_list
-    )
-  ),
-
-  task_new = target(
-    makeRegrTask(
-      id = task_names,
-      data = data_corrected[[1]],
-      target = "defoliation",
-      coordinates = coordinates_list_corrected[[1]],
-      blocking = factor(rep(1:4, c(559, 451, 301, 497))),
-    ),
-    transform = map(data_corrected,
-      task_names = c(
-        "hr",
-        "vi",
-        "nri",
-        "hr_nri",
-        "nri_vi",
-        "hr_nri_vi"
-      ),
-      coordinates_list_corrected
-    )
-  )
-)
-
[…]

### 2020-02-24 code/051-tasks-buffer2.R

@@ -1,0 +1,78 @@
+tasks_plan_buffer2 <- drake_plan(
+  data_old = list(
+    bands_data,
+    vi_data,
+    nri_data,
+    hr_nri_data,
+    nri_vi_data,
+    hr_nri_vi_data
+  ),
+
+  coordinates_list_corrected = list(
+    coords_bands_clean_corrected,
+    coords_vi_nri_clean_corrected,
+    coords_vi_nri_clean_corrected,
+    coords_vi_nri_clean_corrected,
+    coords_vi_nri_clean_corrected,
+    coords_vi_nri_clean_corrected
+  ),
+
+  coordinates_list = list(
+    coords_bands_clean,
+    coords_vi_nri_clean,
+    coords_vi_nri_clean,
+    coords_vi_nri_clean,
+    coords_vi_nri_clean,
+    coords_vi_nri_clean
+  ),
+
+  data_corrected_buffer2 =
+    list(
+      bands_data_corrected_buffer2,
+      vi_data_corrected_buffer2,
+      nri_data_corrected_buffer2,
+      hr_nri_data_corrected_buffer2,
+      nri_vi_data_corrected_buffer2,
+      hr_nri_vi_data_corrected_buffer2
+    ),
+
+  # Non-log transformed tasks ----------------------------------------------------
+  task_old_buffer2 = target(
+    makeRegrTask(
+      id = task_names_buffer2,
+      data = data_old[[1]],
+      target = "defoliation",
+      coordinates = coordinates_list[[1]],
+      blocking = factor(rep(1:4, c(559, 451, 301, 497))),
+    ),
+    dynamic = map(
+      data_old,
+      task_names_buffer2,
+      coordinates_list
+    )
+  ),
+
+  task_names_buffer2 = c(
+    "hr_buffer2",
+    "vi_buffer2",
+    "nri_buffer2",
+    "hr_nri_buffer2",
+    "nri_vi_buffer2",
+    "hr_nri_vi_buffer2"
+  ),
+
+  task_new_buffer2 = target(
+    makeRegrTask(
+      id = task_names_buffer2,
+      data = data_corrected_buffer2[[1]],
+      target = "defoliation",
+      coordinates = coordinates_list_corrected[[1]],
+      blocking = factor(rep(1:4, c(559, 451, 301, 497))),
+    ),
+    dynamic = map(
+      data_corrected_buffer2,
+      task_names_buffer2,
+      coordinates_list_corrected
+    )
+  )
+)

### 2020-02-24 code/052-tasks-no-buffer.R

@@ -1,0 +1,41 @@
+tasks_plan_no_buffer <- drake_plan(
+  data_corrected_no_buffer =
+    list(
+      bands_data_corrected_no_buffer,
+      vi_data_corrected_no_buffer
+    ),
+
+  # Without a buffer we get many NAs for some obs, especially for NRI tasks.
+  # This would result in dropping many features as observations are more important.
+  # Number of features with no NAs without using a buffer: 101 69 4951 5051 5019 5119
+  # Number of features with no NAs with a buffer of 2m:    124 90 7382 7505 7471 7594
+  task_names_no_buffer = c(
+    "hr_no_buffer",
+    "vi_no_buffer" # ,
+    # "nri_no_buffer",
+    # "hr_nri_no_buffer",
+    # "nri_vi_no_buffer",
+    # "hr_nri_vi_no_buffer"
+  ),
+
+  coordinates_list_corrected_no_buffer = list(
+    coordinates_list_corrected[[1]],
+    coordinates_list_corrected[[2]]
+  ),
+
+  # Non-log transformed tasks ----------------------------------------------------
+  task_new_no_buffer = target(
+    makeRegrTask(
+      id = task_names_no_buffer,
+      data = data_corrected_no_buffer[[1]],
+      target = "defoliation",
+      coordinates = coordinates_list_corrected_no_buffer[[1]],
+      blocking = factor(rep(1:4, c(559, 451, 301, 497))),
+    ),
+    dynamic = map(
+      data_corrected_no_buffer,
+      task_names_no_buffer,
+      coordinates_list_corrected_no_buffer
+    )
+  )
+)

### 2020-02-24 code/07-benchmark-matrix.R

@@ -1,72 +1,0 @@
-# see https://ropenscilabs.github.io/drake-manual/plans.html#large-plans
-
-benchmark_plan <- drake_plan(
-
-  tasks_bm_old = list(
-    task_old_hr,
-    task_old_vi,
-    task_old_nri,
-    task_old_hr_nri,
-    task_old_nri_vi,
-    task_old_hr_nri_vi
-  ),
-
-  tasks_bm_new = list(
-    task_new_hr,
-    task_new_vi,
-    task_new_nri,
-    task_new_hr_nri,
-    task_new_nri_vi,
-    task_new_hr_nri_vi
-  ),
-
-  learners_penalized = list(
-    lrn_lasso,
-    lrn_ridge
-  ),
-
-  benchmark_no_models_new = target(
-    benchmark(
-      learners = tune_wrappers_mbo,
-      tasks = tasks_bm_new[[1]],
-      models = FALSE,
-      keep.pred = TRUE,
-      resamplings = makeResampleDesc("CV", fixed = TRUE),
-      show.info = TRUE,
-      measures = list(
-        setAggregation(rmse, test.mean)
-      )
-    ),
-    dynamic = cross(tune_wrappers_mbo, tasks_bm_new)
-  ),
-
-  benchmark_no_models_new_penalized = target(
-    benchmark(
-      learners = learners_penalized[[1]],
-      tasks = tasks_bm_new[[1]],
-      models = FALSE,
-      keep.pred = TRUE,
-      resamplings = makeResampleDesc("CV", fixed = TRUE),
-      show.info = TRUE,
-      measures = list(
-        setAggregation(rmse, test.mean)
-      )
-    ),
-    dynamic = cross(learners_penalized, tasks_bm_new)
-  ),
-
-  benchmark_no_models_old = target(
-    benchmark(
-      learners = tune_wrappers_mbo,
-      tasks = tasks_bm[[1]],
-      models = FALSE,
-      keep.pred = TRUE,
-      resamplings = makeResampleDesc("CV", fixed = TRUE),
-      show.info = TRUE,
-      measures = list(
-        setAggregation(rmse, test.mean)
-      )
-    ),
-    dynamic = cross(tune_wrappers_mbo, tasks_bm_old)
-  ),
-)

### 2020-02-24 code/071-benchmark-matrix-buffer2.R

@@ -1,0 +1,90 @@
+# see https://ropenscilabs.github.io/drake-manual/plans.html#large-plans
+
+benchmark_plan_buffer2 <- drake_plan(
+
+  learners_penalized = list(
+    lrn_lasso,
+    lrn_ridge
+  ),
+
+  learners_keep_models = list(
+    tune_wrappers_mbo %>%
+      keep(~ "SVM MBO MRMR" %in% .x$id)
+  ),
+
+  benchmark_no_models_new_buffer2 = target(
+    benchmark(
+      learners = tune_wrappers_mbo,
+      tasks = task_new_buffer2,
+      models = FALSE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(tune_wrappers_mbo, task_new_buffer2)
+  ),
+
+  benchmark_no_models_new_penalized_buffer2 = target(
+    benchmark(
+      learners = learners_penalized,
+      tasks = task_new_buffer2,
+      models = FALSE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(learners_penalized, task_new_buffer2)
+  ),
+
+  benchmark_models_new_penalized_buffer2 = target(
+    benchmark(
+      learners = learners_penalized,
+      tasks = task_new_buffer2,
+      models = TRUE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(learners_penalized, task_new_buffer2)
+  ),
+
+  # used in response-normality.Rmd
+  benchmark_models_new_buffer2 = target(
+    benchmark(
+      learners = learners_keep_models,
+      tasks = task_new_buffer2,
+      models = TRUE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
[…]

### 2020-02-24 code/072-benchmark-matrix-no-buffer.R

@@ -1,0 +1,65 @@
+# see https://ropenscilabs.github.io/drake-manual/plans.html#large-plans
+
+benchmark_plan_no_buffer <- drake_plan(
+
+  benchmark_no_models_new_no_buffer = target(
+    benchmark(
+      learners = tune_wrappers_mbo,
+      tasks = task_new_no_buffer,
+      models = FALSE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(tune_wrappers_mbo, task_new_no_buffer)
+  ),
+
+  benchmark_no_models_new_penalized_no_buffer = target(
+    benchmark(
+      learners = learners_penalized[[1]],
+      tasks = task_new_no_buffer[[1]],
+      models = FALSE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(learners_penalized, task_new_no_buffer)
+  ),
+
+  benchmark_models_new_penalized_no_buffer = target(
+    benchmark(
+      learners = learners_penalized[[1]],
+      tasks = task_new_no_buffer[[1]],
+      models = TRUE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(learners_penalized, task_new_no_buffer)
+  ),
+
+  # used in response-normality.Rmd
+  benchmark_models_new_no_buffer = target(
+    benchmark(
+      learners = learners_keep_models[[1]],
+      tasks = task_new_no_buffer[[1]],
+      models = TRUE,
+      keep.pred = TRUE,
+      resamplings = makeResampleDesc("CV", fixed = TRUE),
+      show.info = TRUE,
+      measures = list(
+        setAggregation(rmse, test.mean)
+      )
+    ),
+    dynamic = cross(learners_keep_models, task_new_no_buffer)
+  )
+)

### 2020-02-24 code/08-aggregate.R

@@ -1,8 +1,0 @@
-bm_aggregated_plan = drake_plan(
-
-  bm_aggregated_new = target(
-    mergeBenchmarkResults(benchmark_no_models_new,
-                          benchmark_no_models_new_penalized)
-  )
-)
-

### 2020-02-24 code/081-aggregate_buffer2.R

@@ -1,0 +1,17 @@
+bm_aggregated_plan_buffer2 <- drake_plan(
+  bm_aggregated_temp1_buffer2 = target(
+    mergeBenchmarkResults(
+      benchmark_no_models_new_buffer2
+    )
+  ),
+  bm_aggregated_temp2_buffer2 = target(
+    mergeBenchmarkResults(
+      benchmark_no_models_new_penalized_buffer2
+    )
+  ),
+  bm_aggregated_new_buffer2 = target(
+    mergeBenchmarkResults(list(
+      bm_aggregated_temp1_buffer2, bm_aggregated_temp2_buffer2
+    ))
+  )
+)

### 2020-02-24 analysis/eval-performance.Rmd
add prefixes to plots and tables

@@
 loadd(
-  bm_aggregated_new
+  bm_aggregated_new_buffer2
 )
@@
 ```{r}
-df_perf <- getBMRPerformances(bm_aggregated_new, as.df = T) %>%
+df_perf <- getBMRPerformances(bm_aggregated_new_buffer2, as.df = T) %>%
   mutate(task.id = recode_factor(task.id,
-    `hr` = "HR",
-    `vi` = "VI",
-    `nri` = "NRI",
-    `nri_vi` = "NRI_VI",
-    `hr_nri` = "HR-NRI",
-    `hr_vi` = "HR-VI",
-    `hr_nri_vi` = "HR-NRI-VI",
+    `hr_buffer2` = "HR",
+    `vi_buffer2` = "VI",
+    `nri_buffer2` = "NRI",
+    `nri_vi_buffer2` = "NRI_VI",
+    `hr_nri_buffer2` = "HR-NRI",
+    `hr_vi_buffer2` = "HR-VI",
+    `hr_nri_vi_buffer2` = "HR-NRI-VI",
   )) %>%
@@
 
-## (Table) All leaner/filter/task combinations ordered by performance.
+## (Table) T1 All leaner/filter/task combinations ordered by performance.
 
@@
 
-## (Table) Best learner/filter/task combination
+## (Table) T2 Best learner/filter/task combination
 
@@
 
-saveRDS(table2, "docs/00-manuscripts/ieee/presentation/table-best-learner-per-task.rda")
+saveRDS(table2, "docs/00-manuscripts/presentation/table-best-learner-per-task.rda")
 
@@
 
-## (Plot) Best learner/filter combs for all tasks
+## (Plot) P1 Best learner/filter combs for all tasks
 
@@
 
-## (Plot) Best filter combination of each learner vs. no filter per task vs. Borda
+## (Plot) P2 Best filter combination of each learner vs. no filter per task vs. Borda
 
@@
 
-## (Plot) Best filter combination of each learner vs. Borda filter
+## (Plot) P3 Best filter combination of each learner vs. Borda filter
 
@@
 
-## (Plot) Performances of all filter methods
+## (Plot) P4 Performances of all filter methods
 
@@
 
-## (Plot) Scatterplots of filter methods vs. no filter for each learner and task
+## (Plot) P5 Scatterplots of filter methods vs. no filter for each learner and task
 
@@
 
-## (Plot) Scatterplots of filter methods vs. Borda for each learner and task
+## (Plot) P6 Scatterplots of filter methods vs. Borda for each learner and task
 
@@
   facet_wrap(~task.id) +
-  scale_color_manual(guide = guide_legend(override.aes = list(shape = c(19, 3))),
-                     values = c("Filter" = "#BC8C29FF",
-                     "Borda Filter" = "#0072B5FF")) +
+  scale_color_manual(
+    guide = guide_legend(override.aes = list(shape = c(19, 3))),
+    values = c(
+      "Filter" = "#BC3C29FF",
+      "Borda Filter" = "#0072B5FF"
+    )
+  ) +
   labs(x = "RMSE", y = "Task", colour = NULL) +
@@
 ```
+
+```{r echo = FALSE}
+copy_figures()
+```

### 2020-02-25 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-source(here::here("code/99-packages.R"))
-load_packages()
+R.utils::sourceDirectory("R")
+library("drake")
 
@@
 library("ggpubr")
+library("here")
+library("mlr")
+library("dplyr")
+library("forcats")
+```
+
+Last update:
+
+```{r}
+date()
 ```
@@
   print(
-    file = "docs/00-manuscripts/ieee/performance-top-20.tex",
+    file = here("docs/00-manuscripts/ieee/performance-top-20.tex"),
     include.rownames = TRUE,
@@
 
-saveRDS(table1, "docs/00-manuscripts/presentation/table-perf.rda")
+saveRDS(table1, here("docs/00-manuscripts/presentation/table-perf.rda"))
 
@@
   print(
-    file = "docs/00-manuscripts/ieee/performance-best-per-learner.tex",
+    file = here("docs/00-manuscripts/ieee/performance-best-per-learner.tex"),
     include.rownames = TRUE,
@@
 
-saveRDS(table2, "docs/00-manuscripts/presentation/table-best-learner-per-task.rda")
+saveRDS(table2, here("docs/00-manuscripts/presentation/table-best-learner-per-task.rda"))
 
@@
 ```
-
-```{r echo = FALSE}
-copy_figures()
-```

### 2020-02-28 analysis/feature-importance.Rmd
save feature-importance state

--- 2019-06-30 analysis/feature-importance.Rmd
@@
   fig.align = "center",
-  fig.width = 6.93,
-  fig.height = 6.13,
+  fig.width = 8.5,
+  fig.asp = 0.66,
   out.width = "100%",
@@
 
+library("drake")
+library("hsdar")
 library("dplyr")
-library("tidyr")
+library("ggplot2")
 library("ggpubr")
-library("purrr")
-library("glue")
-library("fs")
-R.utils::sourceDirectory("R")
+library("ggpmisc")
+library("patchwork")
 
 # load drake objects
-drake::loadd(
+loadd(
+  fi_permut_hr_buffer2,
+  fi_permut_vi_buffer2
+)
+```
+
+Sort the permutation feature importance results
+
+```{r feature-importance-1 }
+fi_ranked_hr <- fi_permut_hr_buffer2$res %>%
+  tibble::rownames_to_column("measure") %>%
+  tidyr::pivot_longer(
+    cols = starts_with("B"),
+    values_to = "importance", names_to = "feature"
+  ) %>%
+  mutate(wavelength = seq(420, 995, 4.75)) %>%
+  mutate(numeric_id = seq(5, 126, 1)) %>%
+  arrange(desc(importance)) %>%
+  mutate(rank = row_number()) %>%
+  select(-measure)
+fi_ranked_hr
+
+fi_ranked_vi <- fi_permut_vi_buffer2$res %>%
+  tibble::rownames_to_column("measure") %>%
+  tidyr::pivot_longer(
+    cols = starts_with("B"),
+    values_to = "importance", names_to = "feature"
+  ) %>%
+  # mutate(wavelength = seq(420, 995, 4.75)) %>%
+  arrange(desc(importance)) %>%
+  mutate(rank = row_number()) %>%
+  select(-measure)
+fi_ranked_vi
+```
+
+# Create a virtual spectral signature of vegetation using PROSAIL.
+
+PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`.
+Reflectance is scaled to 0-10 to be able to plot it in the same plot as the feature importance rankings -> the axis limits for the y and z axis needs to match.
+
+PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm - 1000 nm.
+Because we order from 1 - 10 with 1 being the best rank, we have to reverse the scaling of the reflectance values.
+
+```{r feature-importance-2 }
+spectra_sim <- hsdar::PROSAIL()
+spectra_df <- data.frame(
+  reflectance = as.vector(spectra_sim@spectra@spectra_ma),
+  wavelength = seq(400, 2500, 1)
+) %>%
+  dplyr::filter(wavelength < 1000) %>%
+  # scale the reflectance to fit into the range of the y-axis for the filter
+  # ranking (the 10 - is to reverse the scale)
+  dplyr::mutate(reflectance = 10 - scale(reflectance,
+    center = FALSE,
+    scale = max(reflectance, na.rm = TRUE) / 10
+  ))
+
+# to be able to plot the sec y axis on the plot, we need to supply the scaled R
+# object -> we need to save the object separately
+reflectance <- as.vector(spectra_sim@spectra@spectra_ma)
+reflectance_scaled <- scale(reflectance,
+  center = FALSE,
+  scale = max(reflectance, na.rm = TRUE) / 10
+)
+```
+
+Next we bind the simulated data with the feature importance rankings.
+To join both data.frames we need to round the reflectance centers of the bands to integers to match with the reflectance values created by PROSAIL.
+
+```{r feature-importance-3 }
+fi_ranked_hr$wavelength <- round(fi_ranked_hr$wavelength)
+
+data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
+  mutate(class = "HR")
+
+# data_vi_merged <- left_join(spectra_df, fi_ranked_vi, by = c("wavelength")) %>%
+#   mutate(class = "vi") %>%
+#   mutate(reflectance = as.numeric(.data$reflectance))
+```
+
+```{r feature-importance-4, echo=FALSE}
+plot_reflectance_imp_rank <- function(data) {
+  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["rank"]])) +
+    geom_line(aes(x = wavelength, y = reflectance, color = "Spectral Signature of Vegetation"),
+      linetype = "solid",
+      size = 0.3
+    ) +
+    geom_point(size = 2, aes(color = "Band"), show.legend = T) +
+    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
+    scale_y_reverse(
+      expand = c(0, 0),
+      limits = c(10L, 0),
+      # breaks = scales::pretty_breaks(),
+      breaks = seq(1, 10, 1),
+      sec.axis = sec_axis(~ scale(-.,
+        center = FALSE,
+        scale = max(., na.rm = TRUE) / -1
+      ),
+      labels = c(1.0, 0.75, 0.55, 0.25, 0),
+      name = "Reflectance [%]"
+      )
+    ) +
+    guides(color = guide_legend(
+      title = NULL,
+      override.aes = list(
+        linetype = c("blank", "solid", "solid"),
+        shape = c(19, NA, NA)
+      )
+    )) +
+    labs(
+      title = "Permutation-based Variable Importance Results",
+      subtitle = paste0(
+        "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
+        "Labels show the hyperspectral band number of dataset 'HR'."
+      ),
+      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
+      y = "Importance Rank", x = "Wavelength [nm]"
+    ) +
+    ggrepel::geom_label_repel(
+      label = data$feature,
+      nudge_x = 0.5, nudge_y = 0.5,
+      label.size = 0.15
+    ) +
+    ggpubr::theme_pubclean()
+}
+
+plot_reflectance_imp_absolute <- function(data, x_identifier, class) {
+  pl <- ggplot(data, aes(x = .data[[x_identifier]], y = .data[["importance"]])) +
+    labs(y = "Importance", x = "Band number") +
+    geom_segment(aes(
+      x = .data[[x_identifier]], y = 0,
+      xend = .data[[x_identifier]], yend = .data[["importance"]]
+    ),
+    color = "grey", show.legend = FALSE
+    ) +
+    geom_point(size = 1, color = "black", show.legend = T) +
+    labs(
+      title = glue::glue("Permutation-based Variable Importance for Dataset '{class}'"),
+      subtitle = "Absolute importance values by band",
+      caption = "Learner: SVM; 100 Monte-Carlo iterations"
+    ) +
+    ggpubr::theme_pubclean()
+
+  if (is.character(data[[x_identifier]])) {
+    pl
+  } else {
+    pl + scale_x_continuous(breaks = seq(5, 125, 5))
+    pl
+  }
+}
+```
+
+# Plots by dataset
+
+## HR
+
+### P1 Absolute permutation based Var Imp
+
+```{r fi-permut-hr-abs, dev = c("png", "pdf")}
+p1 <- fi_ranked_hr %>%
+  plot_reflectance_imp_absolute("numeric_id", class = "HR")
+p1
+```
+
+### P2 Top ten features along the spectral curve
+
+```{r feature-importance-6}
+table_inds <- tibble::tribble(
+  ~Rank, ~Name, ~Formula,
+
+  "1", "Vogelmann2", "(R_{734}-R_{747})/(R_{715}+R_{726})",
+  "2", "Vogelmann4", "(R_{734}-R_{747})/(R_{715}+R_{720})",
+  "3", "Vogelmann", "R_{740}/R_{720}",
+  "4", "NPCI", "(R_{680}-R_{430})/(R_{680}+R_{430})",
+  "5", "Vogelmann3", "D_{715}/D_{705}",
+  # "D2"        , "$D_{705}/D_{722}$",
+  # "Datt3"     , "$D_{754}/D_{704}$",
+  # "PWI"       , "$R_{900}/R_{970}$",
+  # "SR7"       , "$R_{440}/R_{690}$",
+  # "SRPI"      , "$R_{430}/R_{680}$"
+)
+```
+
+```{r fi-permut-vi-hr-ranked, dev = c("png", "pdf"), warning=FALSE}
+table_inset <- tibble(x = 0.4, y = 1.15, tb = list(table_inds))
+
+p2 <- data_hr_merged %>%
+  plot_reflectance_imp_rank() +
+  # Vogelmann 2 (P1)
+  geom_segment(aes(
+    x = 715, xend = 747,
+    y = 1, yend = 1,
+    color = "Range of VegIndex"
+  ), size = 0.5) +
+
+  # Vogelmann 4 (P2)
+  geom_segment(aes(
+    x = 720, xend = 747,
+    y = 2, yend = 2
+  ), color = "green", size = 0.5) +
+
+  # Vogelmann  (P3)
+  geom_segment(aes(
+    x = 720, xend = 740,
+    y = 3, yend = 3
+  ), color = "green", size = 0.5) +
+
+  # NPCI (P4)
+  geom_segment(aes(
+    x = 680, xend = 690,
+    y = 4, yend = 4
+  ), color = "green", size = 0.5) +
+  geom_segment(aes(
+    x = 430, xend = 440,
+    y = 4, yend = 4
+  ), color = "green", size = 0.5) +
+
+  # Vogelmann 3 (P5)
+  geom_segment(aes(
+    x = 705, xend = 715,
+    y = 5, yend = 5,
+  ), color = "green", size = 0.5) +
+
+  # D2 (P6)
+  geom_segment(aes(
+    x = 705, xend = 722,
+    y = 6, yend = 6
+  ), color = "green", size = 0.5) +
+
+  # Datt3 (P7)
+  geom_segment(aes(
+    x = 704, xend = 709,
+    y = 7, yend = 7
+  ), color = "green", size = 0.5) +
+  geom_segment(aes(
+    x = 754, xend = 759,
+    y = 7, yend = 7
+  ), color = "green", size = 0.5) +
+
+  # PWI (P8)
+  geom_segment(aes(
+    x = 900, xend = 910,
+    y = 8, yend = 8
+  ), color = "green", size = 0.5) +
+  geom_segment(aes(
+    x = 970, xend = 980,
+    y = 8, yend = 8
+  ), color = "green", size = 0.5) +
+
+  # SR7 (P9)
+  geom_segment(aes(
+    x = 440, xend = 450,
+    y = 9, yend = 9
+  ), color = "green", size = 0.5) +
+  geom_segment(aes(
+    x = 690, xend = 700,
+    y = 9, yend = 9
+  ), color = "green", size = 0.5) +
+
+  # SRPI (P10)
+  geom_segment(aes(
+    x = 430, xend = 440,
+    y = 10, yend = 10
+  ), color = "green", size = 0.5) +
+  geom_segment(aes(
+    x = 680, xend = 690,
+    y = 10, yend = 10
+  ), color = "green", size = 0.5) +
+  # annotate(geom = "table", x = 395, y = 0,
+  #          label = list(table_inds), size = 2.4,
+  #        vjust = 1, hjust = 0)
+  ggpmisc::geom_table_npc(
+    data = table_inset, aes(npcx = x, npcy = y, label = tb),
+    hjust = 1, vjust = 1, size = 2.35
+  ) +
+  scale_color_manual(
+    values = c(
+      "Band" = "black", "Spectral Signature of Vegetation" = "black",
+      "Range of VegIndex" = "green"
+    )
   )
 
-library(dplyr)
-library(hsdar)
+p2
 ```
 
-```{r }
-# create datasets of most imp features for each feature set
-data_vi = data.frame(class = 
-                    c("VI", "VI", "VI"),
-                  rank = 
-                    c(1, 3, 3),
-                  wavelength = 
-                    c(404, 404, 788),
-                  feature = c("FeatureA", "FeatureB", "FeatureC")
-)
+Combined
 
-data_nri = data.frame(class = 
-                    c("NRI", "NRI", "NRI"),
-                  rank = 
-                    c(1, 3, 3),
-                  wavelength = 
-                    c(4050, 470, 600),
-                  feature = c("FeatureA", "FeatureB", "FeatureC")
-)
+```{r feature-importance-8, fig.width=9.5, fig.asp=1.1}
+p1 / p2
+```
 
-# create spectral signature of vegetation 
-# PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`
-# reflectance is scaled to 0-10 to be able to plot it in the same plot as the filter rankings -> the y-axes need to be the same
-# PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm to 1000 nm
-# because we order from 1 - 10 with 1 being the best, we have to reverse the scaling of the reflectance values
-spectra_sim = hsdar::PROSAIL()
-spectra_df = data.frame(reflectance = as.vector(spectra_sim@spectra@spectra_ma),
-                wavelength = seq(400, 2500, 1)) %>% 
-  dplyr::filter(wavelength < 1000) %>% 
-  # scale the reflectance to fit into the range of the y-axis for the filter ranking (the 10 - is to reverse the scale)
-  dplyr::mutate(reflectance = 10 - scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10))
+## VI
 
-# to be able to plot the sec y axis on the plot, we need to supply the scaled R object -> we need to save the object separately
-reflectance = as.vector(spectra_sim@spectra@spectra_ma)
-reflectance_scaled = scale(reflectance, center = FALSE, scale = max(reflectance, na.rm = TRUE)/10)
+### P3 Absolute permutation based Var Imp
 
-# here we bind the simulated data with the filter rankings
-# Important: Append the feature set class to the simulated data so when merged all together the facetting works (only works if we have class levels for all entries)
-# This is also the reason why we cannot create an initial data.frame containing the results of all feature sets
-data_vi_merged = bind_rows(spectra_df, data_vi) %>% 
-  mutate(class = "VI")
+```{r fi-permut-vi-abs, dev = c("png", "pdf")}
+p3 <- fi_ranked_vi %>%
+  plot_reflectance_imp_absolute("feature", class = "VI") +
+  ggpubr::rotate_x_text()
+p3
+```
 
-data_nri_merged = bind_rows(spectra_df, data_nri) %>% 
-  mutate(class = "NRI")
+Vogelmann2 $(R_{734}-R_{747})/(R_{715}+R_{726})$ Vogelmann et al. (1993)
+Vogelmann4 $(R_{734}-R_{747})/(R_{715}+R_{720})$ Vogelmann et al. (1993)
+Vogelmann3 $D_{715}/D_{705}$ Vogelmann et al. (1993)
+Vogelmann  $R_{740}/R_{720}$ Vogelmann et al. (1993)
+NPCI       $(R_{680}-R_{430})/(R_{680}+R_{430})$
+D2         $D_{705}/D_{722}$
+Datt3      $D_{754}/D_{704}$
+PWI        $R_{900}/R_{970}$
+SR7        $R_{440}/R_{690}$
+SRPI       $R_{430}/R_{680}$
 
-data_all = bind_rows(data_vi_merged, data_nri_merged)
+Dxxx: First derivation of reflectance values at wavelength 'xxx'.
+Rxxx: Reflectance at wavelength 'xxx'.
 
-ggscatter(data_all, x = "wavelength", y = "rank", color = "class", 
-          #add = "segments",
-          palette = "nejm",
-          size = 3,
-          facet.by = "class"
-) +
-  geom_segment(aes(x = wavelength, y = 10, xend = wavelength, yend = rank), color = "grey") + 
-  #scale_y_continuous(limits = c(0L, 10L), breaks = scales::pretty_breaks()) +
-  scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
-  scale_y_reverse(limits = c(10L, 0L), breaks = scales::pretty_breaks(), sec.axis = sec_axis(~ scale(-., center = FALSE, scale = max(., na.rm = TRUE)/-1 ),
-                                                                                             labels = c(10, 0.75, 0.55, 0.25, 0),
-                                                                                             name = "Reflectance [%]")) +
-  labs(y = "Importance", x = "Wavelength [nm]", color = "Feature set") +
-  geom_line(aes(x = wavelength, y = reflectance), linetype = "dotted") +
-  ggrepel::geom_label_repel(label = data_all$feature, nudge_x = 0.5, nudge_y = 0.5)
-```
+Reference: `?hsdar::vegindex()`

### 2020-03-03 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE, verbose = TRUE)

@@
   fi_permut_hr_buffer2,
-  fi_permut_vi_buffer2
+  fi_permut_vi_buffer2,
+
+  df_wavelengths_from_indices
 )
@@
 
-Sort the permutation feature importance results
+Preview the ordered feature importance results for datasets "HR" and "VI".
 
@@
   mutate(rank = row_number()) %>%
-  select(-measure)
+  select(-measure) %>%
+  mutate(feature = stringr::str_replace(feature, "bf2_", ""))
 fi_ranked_vi
@@
   dplyr::filter(wavelength < 1000) %>%
-  # scale the reflectance to fit into the range of the y-axis for the filter
-  # ranking (the 10 - is to reverse the scale)
-  dplyr::mutate(reflectance = 10 - scale(reflectance,
+  # scale the reflectance to [0, 2] to play nicely with the y-axis later (mean dec in rmse)
+  dplyr::mutate(reflectance = scale(reflectance,
     center = FALSE,
-    scale = max(reflectance, na.rm = TRUE) / 10
+    scale = max(reflectance, na.rm = TRUE) / 2
   ))
-
-# to be able to plot the sec y axis on the plot, we need to supply the scaled R
-# object -> we need to save the object separately
-reflectance <- as.vector(spectra_sim@spectra@spectra_ma)
-reflectance_scaled <- scale(reflectance,
-  center = FALSE,
-  scale = max(reflectance, na.rm = TRUE) / 10
-)
 ```
@@
 ```{r feature-importance-3 }
+# round the wavelengths of the HR dataset to match with the simulated ones
 fi_ranked_hr$wavelength <- round(fi_ranked_hr$wavelength)
@@
 data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
-  mutate(class = "HR")
-
-# data_vi_merged <- left_join(spectra_df, fi_ranked_vi, by = c("wavelength")) %>%
-#   mutate(class = "vi") %>%
-#   mutate(reflectance = as.numeric(.data$reflectance))
+  mutate(class = "HR") %>%
+  mutate(reflectance = as.numeric(.data$reflectance))
 ```
 
-```{r feature-importance-4, echo=FALSE}
-plot_reflectance_imp_rank <- function(data) {
-  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["rank"]])) +
-    geom_line(aes(x = wavelength, y = reflectance, color = "Spectral Signature of Vegetation"),
+To label only a subset of the data, a custom data.frame is created.
+
+```{r feature-importance-4 }
+df_wavelengths_from_indices_imp <- df_wavelengths_from_indices %>%
+  left_join(fi_ranked_vi, by = c("class" = "feature")) %>%
+  na.omit()
+
+df_wavelengths_from_indices_label <- df_wavelengths_from_indices_imp %>%
+  arrange(rank) %>%
+  group_by(class, rank) %>%
+  dplyr::slice(1) %>%
+  ungroup() %>%
+  arrange(rank) %>%
+  slice(1:10)
+
+df_wavelengths_from_hr <- data_hr_merged %>%
+  arrange(rank) %>%
+  group_by(class, rank) %>%
+  dplyr::slice(1) %>%
+  ungroup() %>%
+  arrange(rank) %>%
+  slice(1:10) %>%
+  select(-class) %>%
+  rename(class = feature)
+
+df_label_all <- df_wavelengths_from_indices_label %>%
+  bind_rows(df_wavelengths_from_hr)
+```
+
+```{r feature-importance-5, echo=FALSE}
+plot_reflectance_imp <- function(data) {
+  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
+    geom_line(aes(
+      x = .data[["wavelength"]], y = reflectance,
+      color = "Spectral Signature of Vegetation"
+    ),
     linetype = "solid",
-      size = 0.3
+    size = 0.6
     ) +
-    geom_point(size = 2, aes(color = "Band"), show.legend = T) +
+    # HR: plot only rank 11:last
+    geom_point(
+      data = data_hr_merged[which(!(data_hr_merged$rank %in% c(1:10))), ],
+      aes(color = "HR: 11:122"),
+      size = 1.8,
+      shape = 5
+    ) +
+    # HR: plot only rank 1:10
+    geom_point(
+      data = data_hr_merged[which(data_hr_merged$rank %in% c(1:10)), ],
+      aes(color = "HR: 1:10"),
+      size = 1.8,
+    ) +
+
+    # VI: rank 1:10
+    geom_point(
+      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class),
+      size = 2, shape = 4, color = "darkgreen"
+    ) +
+    # VI: rank 1:10
+    geom_line(
+      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
+      size = 0.2, linetype = "dashed"
+    ) +
+
+    # VI: rank 11:89
+    geom_point(
+      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class),
+      color = "grey",
+      shape = 5, size = 2
+    ) +
+    # # VI: rank 11:89
+    geom_line(
+      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
+      size = 0.2, linetype = "dashed"
+    ) +
+
+    scale_color_manual(values = c(
+      "HR: 11:122" = "grey",
+      "HR: 1:10" = "black",
+      "Spectral Signature of Vegetation" = "black",
+      "VI: 1:10" = "darkgreen",
+      "VI: 11:89" = "grey"
+    )) +
     scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
-    scale_y_reverse(
-      expand = c(0, 0),
-      limits = c(10L, 0),
-      # breaks = scales::pretty_breaks(),
-      breaks = seq(1, 10, 1),
+    scale_y_continuous(
       sec.axis = sec_axis(~ scale(-.,
         center = FALSE,
-        scale = max(., na.rm = TRUE) / -1
+        scale = max(., na.rm = TRUE)
       ),
       labels = c(1.0, 0.75, 0.55, 0.25, 0),
-      name = "Reflectance [%]"
+      name = "Scaled Reflectance [0, 1]"
       )
@@
       override.aes = list(
-        linetype = c("blank", "solid", "solid"),
-        shape = c(19, NA, NA)
+        linetype = c("blank", "blank", "solid", "dashed", "dashed"),
+        shape = c(19, 5, NA, 4, 4),
+        color = c("black", "grey", "black", "darkgreen", "grey")
       )
@@
     labs(
-      title = "Permutation-based Variable Importance Results",
+      title = "Spatial Variable Importance",
       subtitle = paste0(
         "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
-        "Labels show the hyperspectral band number of dataset 'HR'."
+        "Labels show the hyperspectral band number for dataset 'HR' and the index name for dataset 'VI'."
       ),
       caption = "Learner: SVM; 100 Monte-Carlo Iterations",
-      y = "Importance Rank", x = "Wavelength [nm]"
+      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
     ) +
     ggrepel::geom_label_repel(
-      label = data$feature,
-      nudge_x = 0.5, nudge_y = 0.5,
-      label.size = 0.15
+      # only label the best 10 features
+      data = df_label_all,
+      label = df_label_all$class,
+      nudge_x = 0.1, nudge_y = 0.1,
+      size = 3
     ) +
@@
 
+# P1 Main plot
+
+```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE}
+p1 <- data_hr_merged %>%
+  plot_reflectance_imp()
+p1
+```
+
 # Plots by dataset
@@
 
-### P1 Absolute permutation based Var Imp
+### P2 Absolute permutation based Var Imp
 
 ```{r fi-permut-hr-abs, dev = c("png", "pdf")}
-p1 <- fi_ranked_hr %>%
+p2 <- fi_ranked_hr %>%
   plot_reflectance_imp_absolute("numeric_id", class = "HR")
-p1
-```
-
-### P2 Top ten features along the spectral curve
-
-```{r feature-importance-6}
-table_inds <- tibble::tribble(
-  ~Rank, ~Name, ~Formula,
-
-  "1", "Vogelmann2", "(R_{734}-R_{747})/(R_{715}+R_{726})",
-  "2", "Vogelmann4", "(R_{734}-R_{747})/(R_{715}+R_{720})",
-  "3", "Vogelmann", "R_{740}/R_{720}",
-  "4", "NPCI", "(R_{680}-R_{430})/(R_{680}+R_{430})",
-  "5", "Vogelmann3", "D_{715}/D_{705}",
-  # "D2"        , "$D_{705}/D_{722}$",
-  # "Datt3"     , "$D_{754}/D_{704}$",
-  # "PWI"       , "$R_{900}/R_{970}$",
-  # "SR7"       , "$R_{440}/R_{690}$",
-  # "SRPI"      , "$R_{430}/R_{680}$"
-)
-```
-
-```{r fi-permut-vi-hr-ranked, dev = c("png", "pdf"), warning=FALSE}
-table_inset <- tibble(x = 0.4, y = 1.15, tb = list(table_inds))
-
-p2 <- data_hr_merged %>%
-  plot_reflectance_imp_rank() +
-  # Vogelmann 2 (P1)
-  geom_segment(aes(
-    x = 715, xend = 747,
-    y = 1, yend = 1,
-    color = "Range of VegIndex"
-  ), size = 0.5) +
-
-  # Vogelmann 4 (P2)
-  geom_segment(aes(
-    x = 720, xend = 747,
-    y = 2, yend = 2
-  ), color = "green", size = 0.5) +
-
-  # Vogelmann  (P3)
-  geom_segment(aes(
-    x = 720, xend = 740,
-    y = 3, yend = 3
-  ), color = "green", size = 0.5) +
-
-  # NPCI (P4)
-  geom_segment(aes(
-    x = 680, xend = 690,
-    y = 4, yend = 4
-  ), color = "green", size = 0.5) +
-  geom_segment(aes(
-    x = 430, xend = 440,
-    y = 4, yend = 4
-  ), color = "green", size = 0.5) +
-
-  # Vogelmann 3 (P5)
-  geom_segment(aes(
-    x = 705, xend = 715,
-    y = 5, yend = 5,
-  ), color = "green", size = 0.5) +
-
-  # D2 (P6)
-  geom_segment(aes(
-    x = 705, xend = 722,
-    y = 6, yend = 6
-  ), color = "green", size = 0.5) +
-
-  # Datt3 (P7)
-  geom_segment(aes(
-    x = 704, xend = 709,
-    y = 7, yend = 7
-  ), color = "green", size = 0.5) +
-  geom_segment(aes(
-    x = 754, xend = 759,
-    y = 7, yend = 7
-  ), color = "green", size = 0.5) +
-
-  # PWI (P8)
-  geom_segment(aes(
-    x = 900, xend = 910,
-    y = 8, yend = 8
-  ), color = "green", size = 0.5) +
-  geom_segment(aes(
-    x = 970, xend = 980,
-    y = 8, yend = 8
-  ), color = "green", size = 0.5) +
-
-  # SR7 (P9)
-  geom_segment(aes(
-    x = 440, xend = 450,
-    y = 9, yend = 9
-  ), color = "green", size = 0.5) +
-  geom_segment(aes(
-    x = 690, xend = 700,
-    y = 9, yend = 9
-  ), color = "green", size = 0.5) +
-
-  # SRPI (P10)
-  geom_segment(aes(
-    x = 430, xend = 440,
-    y = 10, yend = 10
-  ), color = "green", size = 0.5) +
-  geom_segment(aes(
-    x = 680, xend = 690,
-    y = 10, yend = 10
-  ), color = "green", size = 0.5) +
-  # annotate(geom = "table", x = 395, y = 0,
-  #          label = list(table_inds), size = 2.4,
-  #        vjust = 1, hjust = 0)
-  ggpmisc::geom_table_npc(
-    data = table_inset, aes(npcx = x, npcy = y, label = tb),
-    hjust = 1, vjust = 1, size = 2.35
-  ) +
-  scale_color_manual(
-    values = c(
-      "Band" = "black", "Spectral Signature of Vegetation" = "black",
-      "Range of VegIndex" = "green"
-    )
-  )
-
 p2
-```
-
-Combined
-
-```{r feature-importance-8, fig.width=9.5, fig.asp=1.1}
-p1 / p2
 ```
@@
 Reference: `?hsdar::vegindex()`
+
+## Combined
+
+```{r fi-hr-vi-abs-comb, fig.width=9.5, fig.asp=1.1}
+p2 / p3
+```

### 2020-03-03 docs/00-manuscripts/ieee/performance-best-per-learner.tex
update learner result tables

@@
-% latex table generated in R 3.6.1 by xtable 1.8-3 package
+% latex table generated in R 3.6.3 by xtable 1.8-3 package
 \begin{table}[ht!]
@@
   \hline
-1 & HR & RF & Relief & 30.66 & 13.57 \\ 
-  2 & HR & XGBoost & Borda & 30.75 & 13.75 \\ 
-  3 & HR-NRI & SVM & Borda & 31.70 & 15.62 \\ 
-  4 & NRI & Ridge-CV & No Filter & 33.31 & 8.17 \\ 
-  5 & HR & Lasso-CV & No Filter & 47.22 & 20.52 \\ 
-  6 & HR & Lasso-MBO & No Filter & 55.11 & 22.84 \\ 
-  7 & HR & Ridge-MBO & No Filter & 55.11 & 22.84 \\ 
+1 & HR-NRI & SVM & Car & 27.98 & 19.19 \\ 
+  2 & HR-NRI & RF & Car & 31.08 & 17.09 \\ 
+  3 & NRI\_VI & XGBOOST & Relief & 32.10 & 17.73 \\ 
+  4 & NRI & Ridge-CV & No Filter & 32.97 & 7.24 \\ 
+  5 & HR & Lasso-CV & No Filter & 45.96 & 18.35 \\ 
    \hline

### 2020-03-03 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-```{r}
+```{r, warning = FALSE}
 df_perf <- getBMRPerformances(bm_aggregated_new_buffer2, as.df = T) %>%
@@
 
-Aggregate performances and add standard error column.
-
 ```{r aggr-perf}
+# Aggregate performances and add standard error column.
 df_perf %<>%
@@
 
-## (Plot) P2 Best filter combination of each learner vs. no filter per task vs. Borda
-
-**DEPRECATED**: Please see the scatterplots below for a better visulization
-
-Showing the final effect of applying feature selection to a learner for each task.
-The more left a certain filter appears for a given task compared to the purple dot (No Filter), the higher the effectivity of applying feature selection for that given learner on the given task.
-
-```{r filter-effect-best-vs-no-filter, warning=FALSE, dev = c("png", "pdf")}
-# we rbind two filtered data.frames
-# doing the filtering in one step is (at least) complicated here
-
-best_filter <- df_perf %>%
-  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  dplyr::filter(filter != "No Filter") %>%
-  group_by(learner_group, task.id) %>%
-  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-  slice(which.min(rmse_aggr))
-
-best_no_filter <- df_perf %>%
-  mutate(task.id = recode_factor(task.id,
-    `hr` = "HR",
-    `vi` = "VI",
-    `nri` = "NRI",
-    `nri_vi` = "NRI_VI",
-    `hr_nri` = "HR-NRI",
-    `hr_vi` = "HR-VI",
-    `hr_nri_vi` = "HR-NRI-VI",
-  )) %>%
-  tidyr::separate(learner.id, c("learner_group", "filter"),
-    remove = FALSE,
-    sep = " MBO "
-  ) %>%
-  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  dplyr::filter(filter == "No Filter") %>%
-  group_by(learner_group, task.id) %>%
-  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-  slice(which.min(rmse_aggr))
-
-comb <- bind_rows(best_filter, best_no_filter)
-
-comb %>%
-  ggplot(aes(x = rmse_aggr, y = learner_group)) +
-  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 3) +
-  facet_wrap(~task.id) +
-  # scale_color_nejm() +
-  labs(x = "RMSE", y = "Task", color = "Filter") +
-  guides(size = FALSE) +
-  scale_x_continuous(limits = c(27, 47)) +
-  theme_pubr() +
-  theme(
-    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
-    axis.title.y = element_blank(),
-    axis.text.y = element_text(angle = 45),
-    legend.text = element_text(size = 12),
-    legend.title = element_text(size = 12),
-    plot.margin = unit(c(6, 6, 6, 0), "pt")
-  )
-```
-
-## (Plot) P3 Best filter combination of each learner vs. Borda filter
-
-Showing the final effect of the ensemble Borda filter vs the best scoring simple filter.
-
-```{r filter-effect-ensemble, warning=FALSE}
-# we rbind two filtered data.frames
-# doing the filtering in one step is (at least) complicated here
-
-best_filter <- df_perf %>%
-  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  dplyr::filter(filter != "No Filter") %>%
-  group_by(learner_group, task.id) %>%
-  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-  slice(which.min(rmse_aggr))
-
-borda <- df_perf %>%
-  mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-  dplyr::filter(filter == "Borda") %>%
-  group_by(learner_group, task.id) %>%
-  dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-  slice(which.min(rmse_aggr))
-
-comb_borda <- bind_rows(best_filter, borda)
-
-comb_borda %>%
-  ggplot(aes(x = rmse_aggr, y = learner_group)) +
-  geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 3) +
-  facet_wrap(~task.id) +
-  scale_color_nejm() +
-  labs(x = "RMSE", y = "Task", color = "Filter") +
-  guides(size = FALSE) +
-  scale_x_continuous(limits = c(27, 47)) +
-  theme_pubr() +
-  theme(
-    panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
-    axis.title.y = element_blank(),
-    axis.text.y = element_text(angle = 45),
-    legend.text = element_text(size = 12),
-    legend.title = element_text(size = 12),
-    plot.margin = unit(c(6, 6, 6, 0), "pt")
-  )
-```
-
-## (Plot) P4 Performances of all filter methods
-
-```{r filter-perf-all, warning=FALSE, dev = c("png", "pdf")}
-# perf_all_filters <- df_perf %>%
-#   mutate(filter = replace(filter, is.na(filter), "No Filter")) %>%
-#   dplyr::filter(filter != "No Filter") %>%
-#   group_by(learner_group, task.id) %>%
-#   dplyr::filter(learner_group == "SVM" | learner_group == "RF" | learner_group == "XGBoost") %>%
-#   slice(which.min(rmse_aggr))
-#   summarise(perf = mean(rmse_aggr), filter = as.character(filter))
-#
-# perf_all_filters %>%
-#   ggplot(aes(x = rmse_aggr, y = learner_group)) +
-#   geom_beeswarm(groupOnX = FALSE, aes(color = filter), size = 2) +
-#   facet_wrap(~task.id) +
-#   scale_color_nejm() +
-#   labs(x = "RMSE", y = "Task", color = "Filter") +
-#   guides(size = FALSE) +
-#   scale_x_continuous(limits = c(30, 45)) +
-#   theme_pubr() +
-#   theme(
-#     panel.grid.major.y = element_line(size = 0.1, linetype = "dashed"),
-#     axis.title.y = element_blank(),
-#     axis.text.y = element_text(angle = 45),
-#     legend.text = element_text(size = 12),
-#     legend.title = element_text(size = 12),
-#     plot.margin = unit(c(6, 6, 6, 0), "pt")
-#   )
-```
-
-## (Plot) P5 Scatterplots of filter methods vs. no filter for each learner and task
+## (Plot) P2 Scatterplots of filter methods vs. no filter for each learner and task
 
@@
   # summarise() working
-  summarise(perf = mean(rmse_aggr))
+  summarise(perf = mean(rmse_aggr)) %>% 
+  ungroup() %>% 
+  # we need to reverse the order on porpuse here so that ggplot reverses it
+  # again later
+  mutate(learner_group = fct_relevel(learner_group, "XGBOOST", "SVM", "RF"))
 
@@
 
-## (Plot) P6 Scatterplots of filter methods vs. Borda for each learner and task
+## (Plot) P3 Scatterplots of filter methods vs. Borda for each learner and task
 
@@
   summarise(perf = mean(rmse_aggr)) %>%
-  ungroup()
+  ungroup() %>% 
+  # we need to reverse the order on porpuse here so that ggplot reverses it
+  # again later
+  mutate(learner_group = fct_relevel(learner_group, "XGBOOST", "SVM", "RF"))
 
 results_aggr2 %>%
-  ggplot(aes(x = perf, y = fct_reorder(learner_group, -perf))) +
+  ggplot(aes(x = perf, y = learner_group)) +
   geom_beeswarm(

### 2020-03-04 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-  df_wavelengths_from_indices
+  df_wavelengths_from_indices,
+  spec_sigs
 )
@@
 data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
+  left_join(spec_sigs, by = c("wavelength")) %>% 
   mutate(class = "HR") %>%
-  mutate(reflectance = as.numeric(.data$reflectance))
+  mutate(reflectance = as.numeric(reflectance))
 ```
@@
 ```{r feature-importance-5, echo=FALSE}
-plot_reflectance_imp <- function(data) {
+plot_reflectance_imp_hr <- function(data) {
   ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
-    geom_line(aes(
-      x = .data[["wavelength"]], y = reflectance,
-      color = "Spectral Signature of Vegetation"
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$laukiz1), ],
+      aes(
+      x = wavelength, y = laukiz1,
+      color = "Spectral Signature of Laukiz1"
     ),
+    na.rm = TRUE,
     linetype = "solid",
@@
     ) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$laukiz2), ],
+      aes(
+      x = wavelength, y = laukiz2,
+      color = "Spectral Signature of Laukiz2"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$luiando), ],
+      aes(
+      x = wavelength, y = luiando,
+      color = "Spectral Signature of Luiando"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$oiartzun), ],
+      aes(
+      x = wavelength, y = oiartzun,
+      color = "Spectral Signature of Oiartzun"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+
     # HR: plot only rank 11:last
@@
       size = 1.8,
-      shape = 5
+      shape = 1,
+      na.rm = TRUE
     ) +
@@
       aes(color = "HR: 1:10"), 
+      shape = 20,
       size = 1.8,
-    ) +
-
-    # VI: rank 1:10
-    geom_point(
-      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
-      aes(x = wavelength, y = importance, group = class),
-      size = 2, shape = 4, color = "darkgreen"
-    ) +
-    # VI: rank 1:10
-    geom_line(
-      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
-      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
-      size = 0.2, linetype = "dashed"
-    ) +
-
-    # VI: rank 11:89
-    geom_point(
-      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
-      aes(x = wavelength, y = importance, group = class),
-      color = "grey",
-      shape = 5, size = 2
-    ) +
-    # # VI: rank 11:89
-    geom_line(
-      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
-      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
-      size = 0.2, linetype = "dashed"
+      na.rm = TRUE
     ) +
@@
       "HR: 1:10" = "black",
-      "Spectral Signature of Vegetation" = "black",
-      "VI: 1:10" = "darkgreen",
-      "VI: 11:89" = "grey"
+      "Spectral Signature of Laukiz1" = "#BC3C29",
+      "Spectral Signature of Laukiz2" = "#0072B5",
+      "Spectral Signature of Luiando" = "#E18727",
+      "Spectral Signature of Oiartzun" = "#20854E"
     )) +
@@
       override.aes = list(
-        linetype = c("blank", "blank", "solid", "dashed", "dashed"),
-        shape = c(19, 5, NA, 4, 4),
-        color = c("black", "grey", "black", "darkgreen", "grey")
+        linetype = c("blank", "blank", "solid", "solid", "solid" ,"solid"),
+        shape = c(20, 1, NA, NA, NA, NA),
+        color = c("black", "grey", "#BC3C29", "#0072B5", "#E18727", "#20854E")
       )
@@
     labs(
-      title = "Spatial Variable Importance",
+      title = "Permutation-based Variable Importance for dataset 'HR'",
       subtitle = paste0(
-        "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
-        "Labels show the hyperspectral band number for dataset 'HR' and the index name for dataset 'VI'."
+        "The ten most important features are labeled by their band number."
       ),
@@
       # only label the best 10 features
-      data = df_label_all,
-      label = df_label_all$class,
+      data = df_wavelengths_from_hr,
+      label = df_wavelengths_from_hr$class,
       nudge_x = 0.1, nudge_y = 0.1,
-      size = 3
+      size = 3,
+      na.rm = TRUE
+    ) +
+    ggpubr::theme_pubclean()
+}
+
+plot_reflectance_imp_vi <- function(data) {
+  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$laukiz1), ],
+      aes(
+      x = wavelength, y = laukiz1,
+      color = "Spectral Signature of Laukiz1"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$laukiz2), ],
+      aes(
+      x = wavelength, y = laukiz2,
+      color = "Spectral Signature of Laukiz2"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$luiando), ],
+      aes(
+      x = wavelength, y = luiando,
+      color = "Spectral Signature of Luiando"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+    geom_line(
+      data = data_hr_merged[!is.na(data_hr_merged$oiartzun), ],
+      aes(
+      x = wavelength, y = oiartzun,
+      color = "Spectral Signature of Oiartzun"
+    ),
+    na.rm = TRUE,
+    linetype = "solid",
+    size = 0.6
+    ) +
+
+    # VI: rank 1:10
+    geom_point(
+      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class),
+      size = 1.4, shape = 20
+    ) +
+    # VI: rank 1:10
+    geom_line(
+      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
+      size = 0.2, linetype = "dashed"
+    ) +
+
+    # VI: rank 11:89
+    geom_point(
+      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class),
+      color = "grey",
+      shape = 1, size = 2
+    ) +
+    # VI: rank 11:89
+    geom_line(
+      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
+      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
+      size = 0.2, linetype = "dashed"
+    ) +
+
+    scale_color_manual(values = c(
+      "VI: 11:89" = "grey",
+      "VI: 1:10" = "black",
+      "Spectral Signature of Laukiz1" = "#BC3C29",
+      "Spectral Signature of Laukiz2" = "#0072B5",
+      "Spectral Signature of Luiando" = "#E18727",
+      "Spectral Signature of Oiartzun" = "#20854E"
+    )) +
+    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
+    scale_y_continuous(
+      sec.axis = sec_axis(~ scale(-.,
+        center = FALSE,
+        scale = max(., na.rm = TRUE)
+      ),
+      labels = c(1.0, 0.75, 0.55, 0.25, 0),
+      name = "Scaled Reflectance [0, 1]"
+      )
+    ) +
+    guides(color = guide_legend(
+      title = NULL,
+      override.aes = list(
+        linetype = c("solid", "solid", "solid" ,"solid","dashed", "dashed"),
+        shape = c(NA, NA, NA, NA, 20, 1),
+        color = c("#BC3C29", "#0072B5", "#E18727", "#20854E", "black", "grey")
+      )
+    )) +
+    labs(
+      title = "Permutation-based Variable Importance for dataset 'VI'",
+      subtitle = paste0(
+        "The ten most important features are labeled by their index name."
+      ),
+      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
+      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
+    ) +
+    ggrepel::geom_label_repel(
+      # only label the best 10 features
+      data = df_wavelengths_from_indices_label,
+      label = df_wavelengths_from_indices_label$class,
+      nudge_x = 0.1, nudge_y = 0.1,
+      size = 3,
+      na.rm = TRUE
     ) +
@@
 
-```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE}
-p1 <- data_hr_merged %>%
-  plot_reflectance_imp()
-p1
+```{r fi-permut-hr, warning=FALSE, fig.show='hide'}
+p11 <- data_hr_merged %>%
+  plot_reflectance_imp_hr()
+p11
+```
+
+```{r fi-permut-vi, warning=FALSE, fig.show='hide'}
+p12 <- data_hr_merged %>%
+  plot_reflectance_imp_vi()
+p12
+```
+
+```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE, fig.asp=1.1}
+p11 / p12
 ```

### 2020-03-05 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE, verbose = TRUE)

@@
       label = df_wavelengths_from_indices_label$class,
-      nudge_x = 0.1, nudge_y = 0.1,
       size = 3,
@@
 
-```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE, fig.asp=1.1}
+```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE, fig.asp=1.2}
 p11 / p12
@@
 Vogelmann2 $(R_{734}-R_{747})/(R_{715}+R_{726})$ Vogelmann et al. (1993)
+
 Vogelmann4 $(R_{734}-R_{747})/(R_{715}+R_{720})$ Vogelmann et al. (1993)
+
 Vogelmann3 $D_{715}/D_{705}$ Vogelmann et al. (1993)
+
 Vogelmann  $R_{740}/R_{720}$ Vogelmann et al. (1993)
+
 NPCI       $(R_{680}-R_{430})/(R_{680}+R_{430})$
+
 D2         $D_{705}/D_{722}$
+
 Datt3      $D_{754}/D_{704}$
+
 PWI        $R_{900}/R_{970}$
+
 SR7        $R_{440}/R_{690}$
+
 SRPI       $R_{430}/R_{680}$

### 2020-03-05 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
     `nri_buffer2` = "NRI",
-    `nri_vi_buffer2` = "NRI_VI",
+    `nri_vi_buffer2` = "NRI-VI",
     `hr_nri_buffer2` = "HR-NRI",
@@
   arrange(RMSE) %>%
-  slice(1:15) %>%
+  slice(1:10) %>%
   xtable(
     type = "latex",
-    caption = "Top 15 results for any task/learner/filter combination, sorted by performance.",
-    label = "tab:perf-top-15"
+    caption = "Top 10 results for any task/learner/filter combination, sorted by performance.",
+    label = "tab:perf-top-10"
   ) %>%
   print(
-    file = here("docs/00-manuscripts/ieee/performance-top-20.tex"),
+    file = here("docs/00-manuscripts/ieee/performance-top-10.tex"),
     include.rownames = TRUE,
@@
   geom_beeswarm(
-    data = results_aggr1[results_aggr1$filter != "NF", ], size = 1.1, shape = 3,
+    data = results_aggr1[results_aggr1$filter != "NF", ], size = 2.2, shape = 3,
     groupOnX = FALSE, aes(color = "Filter")
@@
     data = results_aggr1[results_aggr1$filter == "NF", ],
-    size = 2, shape = 19, aes(color = "No Filter")
+    size = 2.2, shape = 19, aes(color = "No Filter")
   ) +
@@
     data = results_aggr2[results_aggr2$filter != "Borda", ],
-    shape = 3, size = 1.1, aes(color = "Filter"),
+    shape = 3, size = 2.2, aes(color = "Filter"),
     groupOnX = FALSE
@@
     data = results_aggr2[results_aggr2$filter == "Borda", ],
-    shape = 19, size = 2, aes(color = "Borda Filter")
+    shape = 19, size = 2.2, aes(color = "Borda Filter")
   ) +

### 2020-03-05 docs/00-manuscripts/ieee/performance-best-per-learner.tex
eval-performance: increase point size, fix typo

--- 2020-02-09 docs/00-manuscripts/ieee/performance-best-per-learner.tex
@@
   \hline
-1 & HR & RF & Relief & 30.66 & 13.57 \\ 
-  2 & HR & XGBoost & Borda & 30.75 & 13.75 \\ 
-  3 & HR-NRI & SVM & Borda & 31.70 & 15.62 \\ 
-  4 & NRI & Ridge-CV & No Filter & 33.31 & 8.17 \\ 
-  5 & HR & Lasso-CV & No Filter & 47.22 & 20.52 \\ 
-  6 & HR & Lasso-MBO & No Filter & 55.11 & 22.84 \\ 
-  7 & HR & Ridge-MBO & No Filter & 55.11 & 22.84 \\ 
+1 & HR-NRI & SVM & Car & 27.98 & 19.19 \\ 
+  2 & HR-NRI & RF & Car & 31.08 & 17.09 \\ 
+  3 & NRI-VI & XGBOOST & Relief & 32.10 & 17.73 \\ 
+  4 & NRI & Ridge-CV & No Filter & 32.97 & 7.24 \\ 
+  5 & HR & Lasso-CV & No Filter & 45.96 & 18.35 \\ 
    \hline

### 2020-03-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
finish results chapter

@@
 	\acro{AGB}{above-ground biomass}
+	\acro{ALE}{accumulated local effects}
 	\acro{ALS}{airborn laser scanning}
@@
 	\acro{PCA}{principal component analysis}
+	\acro{PDP}{partial dependence plots}
 	\acro{PLS}{partial least-squares}
@@
 The idea of splitting the features into different sets originated from the question whether feature engineered indices from reflectance values have a positive effect on model performance.
-Benchmarking all models on these distinct groups of features makes it easier to draw conclusions on their impact when keeping all other variables such as model type, tuning strategy and partioning method constant.
+Benchmarking all models on these distinct groups of features makes it easier to draw conclusions on their impact when keeping all other variables such as model type, tuning strategy and partitioning method constant.
 However, rather than only looking at these three groups we decided to also take their combinations into account for the overall comparison:
@@
 \noindent A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible\cite{schratz2019, sperrorest}.
-Each plot served as one fold within the resampling setup, resulting in four folds total.
-For the inner level (hyperparameter tuning), \(p - 1\) folds were used (with \(p\) being the number of plots).
-
-In total the benchmarking matrix consisted of 156 experiments (6 feature sets * 3 ML algorithms * 8 feature-selection methods + 2 * 6 models of the penalized algorithms).
-
-\subsection{Calculation of model feature importance}
-\noindent A permutation-based feature importance estimation was conducted for the best algorithm of each feature set and filter method, respectively.
+Each plot served as one fold within the cross-validation setting, resulting in four iterations total.
+For the inner level (hyperparameter tuning), \(y - 1\) folds were used (with \(y\) being the number of plots).
+
+In total the benchmarking matrix consisted of 156 experiments (6 feature sets * 3 ML algorithms * 8 feature-selection methods + 2 * 6 L1 and L2 models).
+
+\subsection{Feature importance estimation}
+
+% FIXME Talk about the bias of permutation FI in highly correlated settings
+\noindent Estimating feature importance for highly-correlated datasets is a complicated task.
+The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
+Methods like \ac{PDP} do not produce reliable estimates in such situations because they create unrealistic situations between covariates.
+Due to the noisiness of the dataset, estimates from model-agnostic approaches such as permutation-based feature importance should also be taken with care.
+
+However, as of today there are no unbiased approaches for estimating feature importance for (highly-) correlated datasets.
+Nevertheless, recent methods like \ac{ALE} plots aim to tackle this problem by utilizing marginal distributions of features over a small window to focus on the effect of a single feature.
+However,
+
+In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset.
+With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while keeping modest on over-interpreting the results.
+We used the algorithm which showed the best performance for the HR task to calculate the feature importance.
+
 This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
 The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
-To balance both runtime and variance reduction, we opted for 25 iterations of each feature permutation.
+Permuting every feature $y$ times comes with some computational cost.
+To balance the effort/gain ratio and to be able to relate the results to wavelength regions, we decided to estimate feature importance only for the dataset with the raw reflectance values (HR).
 
@@
 
-% - RMSE around 34
+% - RMSE around 27
 % - MBO tuned penalized methods are substantially worse
-% - Carscore, Info und Borda were at least once among the best combination for a learner
-\noindent No practical differences in RMSE were observed among the feature sets (\autoref{fig:perf-result}).
-SVM showed the lowest error in four out of six tasks.
-Overall, the response variable 'defoliation at trees' could be modeled with an error of 34 \%.
+\noindent Overall, the response variable 'defoliation at trees' could be modeled with an error of 28 \%.
+SVM showed no differences in RMSE across feature sets whereas other learners like RF differed up to seven percentage points (HR-NRI vs. VI) (\autoref{fig:perf-result}).
+Ridge and Lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
+SVM showed the best performance of all learners with an absolute difference up around three percentage points to the next model (RF) (\autoref{tab:best-learner-perf}).
 
@@
 % - no substantial difference between feature sets
-% - combining feature sets makes no difference
+% - combining feature sets has only small impact
 % - expert based feature sets show no better result than data-driven ones
 % - penalized methods seem to have problems with VI feature set
-No practical difference was observed between feature sets.
-The combination of feature sets showed no increase in performance.
-All learners scored evenly across all tasks.
-Feature sets containing expert-based variables (VI) showed no substantial difference to data-driven tasks (NRI).
+The combination of feature sets showed small increases in performance for some learners.
+RF and XGBoost scored slightly better on the combined datasets HR-NRI and NRI-VI, respectively (\autoref{fig:perf-result}).
+Datasets containing derived features (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
 
 % learner
-% - RF Car is the overall winner
+% - SVM Car is the overall winner
 % - Ridge-CV hit and miss (hit = NRI, miss = VI)
 % - all MBO penalized methods have no difference across tasks
-Random Forest combined with the 'Carscore' filter showed the best absolute performance (RMSE of 3.396) (\autoref{tab:perf-top-15}).
-A noticeable difference between SVM and all other learners was visible for all
-feature sets without NRI variables (difference in RMSE of 3 --- 4).
-Considering a grouped performance by learners across all tasks and filters, SVM showed the most robust results (\autoref{tab:perf-top-15}).
-L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In three out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI\@; HR-VI, VI), for HR-NRI and HR the error was medium (around 40 RMSE), and for task NRI Ridge regression achieved almost the same score as the best performing model (3.39 RMSE, NRI) (\autoref{tab:perf-top-15}).
-All penalized methods tuned using MBO showed exactly the same error across all tasks.
+SVM combined with the 'Carscore' filter achieved the best performance (RMSE of 27.98) (\autoref{tab:perf-top-10}).
+%L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
+Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI), for HR-NRI and HR the error was around 40 RMSE, and for task NRI the Ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
+%All penalized methods tuned using MBO showed exactly the same error across all tasks.
 
 % filters
-% - influence varies based on algorithm and task (SVM very small influence, RF medium, XGB medium)
+% - influence varies based on algorithm and task (SVM no influence, RF medium, XGB medium)
 % - no substantial decrease in performance when using filters
 % - no advantage of ensemble borda filter compared to others
-Effects of filter methods on performance differed between algorithms:
-With the exception of one or two methods, SVM showed a constant performance of all filters across tasks (\autoref{fig:filter-perf-all}).
-RF showed a normal distribution for most tasks with one or two different methods taking the lead on each task.
-For RF, filter 'Carscore' showed the best performance in three out of six tasks while this was achieved for XGBoost by filter 'CMIM'.
-XGBoost showed a diverse distribution with less clustering of filters as SVM and RF (referring to RMSE).
-
-% filter vs no filter
-When comparing the usage of filters vs using no filter at all, there was only one instance (XGBoost on the HR-NRI task) when the model with no filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects}).
-In general, the effect of filtering vs.\ no filtering differed without a clear pattern across tasks: For example, SVM and RF showed a substantial increase for the NRI task when filtering the data while there was almost no effect for XGBoost (\autoref{fig:filter-effects}).
-For task HR, there was a medium effect of applying filtering for RF, almost none for SVM and a large one for XGBoost.
+Effects of filter methods on performance differed greatly between algorithms:
+SVM showed no variation in performance across filters (\autoref{fig:filter-effects-no-filter}).
+Using filters for RF showed a substantial increase in performance for all tasks besides VI for which the overall difference of all filters was also smallest (\autoref{fig:filter-effects-no-filter}).
+XGBoost showed a high dependency on filtering the data: In 4 out of 6 tasks using no filter would results in the worst or second worst performance.
+However, in contrast for dataset NRI using no filter results in the best performance.
+XGBoost shows the highest overall differences between filters for a single task: For dataset HR, the range is up to 14 percentage points (Carscore vs.
+no filter)(\autoref{fig:filter-effects-no-filter}).
+
+% filter vs. no filter
+% - Substantial effect for XGBoost and RF
+% - SVM unaffected
+When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) when the model without filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
+For SVM filters and no filter come in at the same performance for tasks VI and NRI even though \autoref{fig:filter-effects-no-filter} lists "No Filter" as the best option.
+
+% Borda vs. other filters
+% - most often among the first best filters for each task
+% - never the best
+The Borda filter did not achieve the best performance for any learner across any task \autoref{fig:filter-effects-borda}.
+For RF and XGBoost it ranked within the first 50\% when looking at all filters used for a task.
+For XGBoost on the VI task the Borda filter scored the second worst performance.
 
 % latex table top 20 absolute performances
-\input{performance-top-20}
+\input{performance-top-10}
 
@@
 
-% plot no filter vs best filter for each model and task
+% plot no filter vs all other filters for each model and task
 \begin{figure} [t!]
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {filter-effect-1.pdf}
-		\caption{Model performances in RMSE when using no filter method compared to the best scoring filter method for each learner across all tasks.}\label{fig:filter-effects}
+		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-no-filter-1.pdf}
+		\caption{Model performances in RMSE when using no filter method compared to all other filters across all tasks.}\label{fig:filter-effects-no-filter}
 	\end{center}
 \end{figure}
-\subsection{Permutation-based variable importance}
-
-% plot no filter vs best filter for each model and task
+
+% plot Borda vs all other filters for each model and task
 \begin{figure} [t!]
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {filter-perf-all-1.pdf}
-		\caption{Model performances in RMSE of all used filter methods and PCA for each learner across all tasks.}\label{fig:filter-perf-all}
+		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-borda-filter-1.pdf}
+		\caption{Model performances in RMSE when using the Borda filter method compared to all other filters for each learner across all tasks.}\label{fig:filter-effects-borda}
 	\end{center}
@@
 \subsection{Permutation-based variable importance}
+
+% Variable Importance
+% - VI: Vogelmann indices are best
+% - All: Clustering around the red edge (700 - 750 nm)
+% - Overall highest: 1.7 RMSE decrease - not that much
+
+% FIXME: explain in methods why only 2 datasets are shown
+The most important features of each dataset showed an average decrease in RMSE of 1.57 (HR, B69) and 1.79 (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm.
+For feature set HR four features in the infrared region (920 nm  - 1000 nm) were identified to be important by causing a mean RMSE decrease of around one percentage point.
+Overall, most features show a small importance only with average decreases in RMSE below 0.5.
+
+% permutation based var imp for datasets HR and VI
+\begin{figure} [t!]
+	\centering
+	\begin{center}
+		\includegraphics[width=0.48\textwidth] {fi-permut-vi-hr-1.pdf}
+		\caption{Variable importance for feature sets HR and VI: Mean decrease in RMSE for one-hundred feature permutations using the SVM learner.
+The wavelength range on the x-axis matches the range of the hyperspectral sensor (400 nm - 1000 nm).
+For each dataset, the ten most important features were highlighted as black dots and labeled by name.
+Grey dots represent features from importance rank 11 to last.
+The spectral signature (mean) of each plot was added as a reference on a normalized reflectance scale [0, 1] (secondary y-axis).
+VI features were decomposed into their individual formula parts; all instances being connected via dashed lines.
+Each VI feature is composed out of at least two instances.}\label{fig:fi-permut-vi-hr}
+	\end{center}
+\end{figure}
+
 \section{Discussion}
 
-\subsection{Derivation of indices}
+\subsection{Derivation of Indices}
 
@@
 
-\section{Spectral signatures of each plot}
-
-\begin{figure} [ht]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {spectral-signatures-1.pdf}
-		\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}\label{fig:spectral-signatures}
-	\end{center}
-\end{figure}
+% \section{Spectral signatures of each plot}
+
+% \begin{figure} [ht]
+% 	\begin{center}
+% 		\includegraphics[width=0.48\textwidth] {spectral-signatures-1.pdf}
+% 		\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}\label{fig:spectral-signatures}
+% 	\end{center}
+% \end{figure}
 

### 2020-03-05 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
     aes(label = paste0(filter, ",", round(rmse_aggr, 2))),
-    size = 2.8,
+    size = 4,
     min.segment.length = 0.1,
@@
   ) +
-  theme_pubr() +
+  theme_pubr(base_size = 14) +
   theme(
@@
     axis.title.y = element_blank(),
-    legend.text = element_text(size = 12),
-    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 14),
+    legend.title = element_text(size = 14),
     axis.text.y = element_text(angle = 45),
@@
   scale_x_continuous(limits = c(27, 51)) +
-  theme_pubr() +
+  theme_pubr(base_size = 14) +
   theme(
@@
     axis.title.y = element_blank(),
-    legend.text = element_text(size = 12),
-    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 14),
+    legend.title = element_text(size = 14),
   )

### 2020-03-05 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE, verbose = TRUE)

@@
       label = df_wavelengths_from_hr$class,
-      nudge_x = 0.1, nudge_y = 0.1,
-      size = 3,
+      size = 4,
       na.rm = TRUE
     ) +
-    ggpubr::theme_pubclean()
+    ggpubr::theme_pubclean(base_size = 14)
 }
@@
       label = df_wavelengths_from_indices_label$class,
-      size = 3,
+      size = 4,
       na.rm = TRUE
     ) +
-    ggpubr::theme_pubclean()
+    ggpubr::theme_pubclean(base_size = 14)
 }

### 2020-03-05 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE, verbose = TRUE)

@@
   ungroup()
+```
+
+Fold performances of "SVM MBO No Filter" on the HR Task
+
+- Fold 1: Luiando
+- Fold 2: Laukiz1
+- Fold 3: Laukiz2
+- Fold 4: Oiartzun
+
+```{r}
+df <- mlr::getBMRPerformances(bm_aggregated_new_buffer2,
+  "hr_buffer2", "SVM MBO No Filter",
+  as.df = TRUE
+)
+df_tab <- df %>%
+  dplyr::select(iter, rmse) %>%
+  dplyr::rename(Plot = iter, RMSE = rmse) %>%
+  dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::mutate(Plot = forcats::fct_recode(Plot,
+    Laukiz1 = "2", Laukiz2 = "3",
+    Luiando = "1", Oiartzun = "4"
+  )) %>% 
+  dplyr::mutate(RMSE = round(RMSE, 2))
+
+df_tab %>%
+  xtable::xtable(
+    type = "latex",
+    caption = "Single fold performances for learner SVM on the HR dataset without using a filter.",
+    label = "tab:svm-single-fold-perf"
+  ) %>%
+  print(
+    file = here::here("docs/00-manuscripts/ieee/performance-svm-single-plot.tex"),
+    include.rownames = TRUE,
+    latex.environments = c("center"),
+    table.placement = "ht!",
+    caption.placement = "top",
+    timestamp = NULL
+  )
+
+saveRDS(df_tab, here("docs/00-manuscripts/presentation/table-svm-single-plot.rda"))
+
+flextable::flextable(df_tab)
 ```

### 2020-03-05 docs/00-manuscripts/ieee/performance-svm-single-plot.tex
add performance single plot table

% latex table generated in R 3.6.1 by xtable 1.8-3 package
\begin{table}[ht!]
\centering
\caption{Single fold performances for learner SVM on the HR dataset without using a filter.} 
\label{tab:svm-single-fold-perf}
\begin{tabular}{rlr}
  \hline
 & Plot & RMSE \\ 
  \hline
1 & Luiando & 9.00 \\ 
  2 & Laukiz1 & 21.17 \\ 
  3 & Laukiz2 & 54.26 \\ 
  4 & Oiartzun & 28.05 \\ 
   \hline
\end{tabular}
\end{table}

### 2020-03-08 analysis/inspect-glmnet-models.Rmd
wflow_publish(knitr_in("analysis/inspect-glmnet-models.Rmd"), view = FALSE, verbose = TRUE)

---
title: "Inspecting {glmnet} performances"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = TRUE
)

R.utils::sourceDirectory("R")
library("drake")
library("mlr")
library("glmnet")

# load drake objects
loadd(
  benchmark_models_new_penalized_buffer2,
  task_new_buffer2
)

```

Last update:

```{r}
date()
```

## Inspect fitted models during CV

Inspect Ridge regression on VI task in detail because the error is enourmus.

First extract the models.

```{r}
models = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["models"]]
```

Then look at the fold performances

```{r}
benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["measures.test"]][["rmse"]]
```

We see a high error on Fold 1 (= Laukiz 2).
The others are also quite high but not "out of bounds".

Because this models used the internal optimization of the lambda sequence (`cv.glmnet`), let's look at the value which was chosen for prediction (parameter `s` which defaults to `s="lambda.1se"`):

```{r}
purrr::map_dbl(models, ~ .x[["learner.model"]][["lambda.1se"]])
```

It seems that the `lambda.1se` value for Fold 1 is way smaller than for the other 3 folds.
However, all values seem to be quite high.

Let's look at the full lambda sequence

```{r}
purrr::map_int(models, ~ length(.x[["learner.model"]][["lambda"]]))
```

Interestingly, the lambda length of fold 1 is not 100 (default) but only 5.

## Train/predict manually for Fold 1 via {glmnet}

To inspect further, let's refit a {glmnet} model directly on the training data of Fold 1 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:

```{r}
train_inds_fold1 = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["pred"]][["instance"]][["train.inds"]][[1]]

obs_train_f1 = as.matrix(task_new_buffer2[[2]]$env$data[train_inds_fold1, getTaskFeatureNames(task_new_buffer2[[2]])])
target_f1 = getTaskTargets(task_new_buffer2[[2]])[train_inds_fold1]
```

Fit `cv.glmnet`

```{r}
set.seed(1)
modf1 = glmnet::cv.glmnet(obs_train_f1, target_f1, nfolds = 5)

modf1$lambda.1se
```

Ok, a value of `0.85` is **very** different to what happened during the CV (4.211054e+08).

Predict on Laukiz 2 now.

```{r}
pred_inds_fold1 = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["pred"]][["instance"]][["test.inds"]][[1]]

obs_pred_f1 = as.matrix(task_new_buffer2[[2]]$env$data[pred_inds_fold1, getTaskFeatureNames(task_new_buffer2[[2]])])

pred = predict(modf1, newx = obs_pred_f1)
```

Calculate the error

```{r}
truth = task_new_buffer2[[2]]$env$data[pred_inds_fold1, "defoliation"]

mlr:::measureRMSE(truth, pred)
```

Ok, RMSE of 52. Still not great but at least not out of bounds.

## Train/predict manually for Fold 1 via {mlr}

Now let's fit the same subset (= Fold 1) via mlr

```{r}
lrn = makeLearner("regr.cvglmnet", nfolds = 5)
task_f1 = subsetTask(task_new_buffer2[[2]], train_inds_fold1)

set.seed(1)
mod = train(lrn, task_f1)
```

Check lambda sequence and `lambda.1se`:

```{r}
mod$learner.model$lambda
```

```{r}
mod$learner.model$lambda.1se
```

Check for equality between {mlr} and {glmnet} directly

```{r}
all.equal(modf1$lambda.1se, mod$learner.model$lambda.1se)
```

Ok, so {mlr} works correctly when fitting the model for a single fold. 
Hence, somethings goes wrong during CV?

### 2020-03-08 analysis/index.Rmd
fix git
workflowr::wflow_publish("analysis/index.Rmd")

@@
 
-[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/pat-s/2019-feature-selection/commits/master) 
-[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.6.0-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2582969.svg)](https://doi.org/10.5281/zenodo.2582969)
@@
 Powered by [drake](https://github.com/ropensci/drake) + [workflowr](https://github.com/jdblischak/workflowr) + [mlr](https://github.com/mlr-org/mlr).
+
+# Acknowledgements
+
+This work has been supported by the the European Union "Healthy Forest" Project LIFE14 ENV/ES/000179.

### 2020-03-11 analysis/inspect-glmnet-models.Rmd
workflowr::wflow_publish("analysis/inspect-glmnet-models.Rmd")

@@
 library("glmnet")
+library("ggplot2")
+library("magrittr")
+
+options(crayon.enabled = TRUE, pillar.bold = TRUE, scipen = 999)
+fansi::set_knit_hooks(knitr::knit_hooks)
 
@@
 loadd(
-  benchmark_models_new_penalized_buffer2,
+  benchmark_models_new_penalized_mbo_buffer2,
   task_new_buffer2
 )
-
 ```
@@
 
-```{r}
+```{r, echo=FALSE, results='asis'}
 date()
 ```
+
+## General glmnet notes
+
+- {glmnet} does its own internal 10-fold CV optimization when using `cv.glmnet()`.
+  It iterates over `lambda` and chooses the most robust values for prediction via parameter `s` in `predict.glmnet()`.
+Supplying a custom lambda sequence does not make much sense since the internal heuristics are quite good (if one wants to use non-spatial optimization).
+- To conduct a spatial optimization, one needs to use `glmnet()` directly. 
+  This implementation does not do an internal optimization for `lambda`  and hence `s` can/needs to be tuned directly by the user.
+Because it is hard to come up with good tuning ranges in this case, one can fit a `cv.glmnet()` on the data and use the borders of the estimated `lambda` as upper and lower borders of the tuning space.
 
@@
 ```{r}
-models = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["models"]]
+models <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["models"]]
 ```
@@
 ```{r}
-benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["measures.test"]][["rmse"]]
+benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["measures.test"]][["rmse"]]
 ```
 
-We see a high error on Fold 1 (= Laukiz 2).
+We see a high error on Fold 2 (= Laukiz 2).
 The others are also quite high but not "out of bounds".
@@
 
-It seems that the `lambda.1se` value for Fold 1 is way smaller than for the other 3 folds.
+It seems that the `lambda.1se` value for Fold 4 is way higher than for the other 3 folds.
 However, all values seem to be quite high.
@@
 ```{r}
-train_inds_fold1 = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["pred"]][["instance"]][["train.inds"]][[1]]
+train_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["train.inds"]][[4]]
 
-obs_train_f1 = as.matrix(task_new_buffer2[[2]]$env$data[train_inds_fold1, getTaskFeatureNames(task_new_buffer2[[2]])])
-target_f1 = getTaskTargets(task_new_buffer2[[2]])[train_inds_fold1]
+obs_train_f4 <- as.matrix(task_new_buffer2[[2]]$env$data[train_inds_fold4, getTaskFeatureNames(task_new_buffer2[[2]])])
+target_f4 <- getTaskTargets(task_new_buffer2[[2]])[train_inds_fold4]
 ```
@@
 set.seed(1)
-modf1 = glmnet::cv.glmnet(obs_train_f1, target_f1, nfolds = 5)
+modf4 <- glmnet::cv.glmnet(obs_train_f4, target_f4, alpha = 0)
 
-modf1$lambda.1se
+modf4$lambda.1se
 ```
@@
 ```{r}
-pred_inds_fold1 = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["pred"]][["instance"]][["test.inds"]][[1]]
+pred_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["test.inds"]][[4]]
 
-obs_pred_f1 = as.matrix(task_new_buffer2[[2]]$env$data[pred_inds_fold1, getTaskFeatureNames(task_new_buffer2[[2]])])
+obs_pred_f4 <- as.matrix(task_new_buffer2[[2]]$env$data[pred_inds_fold4, getTaskFeatureNames(task_new_buffer2[[2]])])
 
-pred = predict(modf1, newx = obs_pred_f1)
+pred <- predict(modf4, newx = obs_pred_f4, s = modf4$lambda.1se)
 ```
@@
 ```{r}
-truth = task_new_buffer2[[2]]$env$data[pred_inds_fold1, "defoliation"]
+truth <- task_new_buffer2[[2]]$env$data[pred_inds_fold4, "defoliation"]
 
@@
 
-Ok, RMSE of 52. Still not great but at least not out of bounds.
-
-## Train/predict manually for Fold 1 via {mlr}
-
-Now let's fit the same subset (= Fold 1) via mlr
+Ok, RMSE of 97073324139.
+This is most likely because of a few. observations which were predicted completely out of bounds.
 
 ```{r}
-lrn = makeLearner("regr.cvglmnet", nfolds = 5)
-task_f1 = subsetTask(task_new_buffer2[[2]], train_inds_fold1)
+qplot(pred, geom = "histogram")
+```
+
+Ok, its one observation (row id  = 737).
+
+Let's have a look at the predictor values for this observation.
+
+```{r}
+summary(obs_train_f4[737, ])
+```
+
+Ok, how does this compare to summaries of other observations?
+
+```{r}
+lapply(seq(500:510), function(x) summary(obs_train_f4[x, ]))
+```
+
+We have some higher values for obs 737 but nothing which stands out.
+
+Let's look at the model coefficients and Partial Dependence Plots (PDP):
+
+```{r}
+coef(modf4)
+```
+
+Feature "bf2_PRI_norm" has a quite high value.
+
+Let's figure out which are the ten most important features and create PDPs for these:
+
+```{r}
+top_ten_abs <- coef(modf4) %>%
+  as.matrix() %>%
+  as.data.frame() %>%
+  dplyr::rename(coef = `1`) %>%
+  dplyr::mutate(feature = rownames(coef(modf4))) %>%
+  dplyr::slice(-1) %>%
+  dplyr::mutate(coef_abs = abs(coef)) %>%
+  dplyr::arrange(desc(coef_abs)) %>%
+  dplyr::slice(1:10) %>%
+  dplyr::pull(feature)
+```
+
+### Partial Dependence Plots
+
+For PDP we use a model trained with {mlr} and check for equality first.
+
+```{r}
+lrn <- makeLearner("regr.cvglmnet", alpha = 0)
+task_f4 <- subsetTask(task_new_buffer2[[2]], train_inds_fold4)
 
 set.seed(1)
-mod = train(lrn, task_f1)
+mod_mlr <- train(lrn, task_f4)
 ```
@@
 ```{r}
-mod$learner.model$lambda
+mod_mlr$learner.model$lambda
 ```
@@
 ```{r}
-mod$learner.model$lambda.1se
+mod_mlr$learner.model$lambda.1se
 ```
@@
 ```{r}
-all.equal(modf1$lambda.1se, mod$learner.model$lambda.1se)
+all.equal(modf4$lambda.1se, mod_mlr$learner.model$lambda.1se)
 ```
 
-Ok, so {mlr} works correctly when fitting the model for a single fold. 
-Hence, somethings goes wrong during CV?
+```{r}
+pdp <- generatePartialDependenceData(mod_mlr, task_f4, features = top_ten_abs)
+```
+
+```{r}
+plotPartialDependence(pdp)
+```
+
+Individual PDP
+
+```{r}
+pdp_ind <- generatePartialDependenceData(mod_mlr, task_f4,
+  features = top_ten_abs,
+  individual = TRUE
+)
+plotPartialDependence(pdp_ind)
+```
+
+## Compare arguments of the `do.call` call in the benchmark with the manual one
+
+`args` were saved during a debug call of `mlr::benchmark()`
+
+```{r}
+args <- readRDS("args-bm-laukiz2.rda")
+target_bm <- args$y
+features_bm <- args$x
+
+train_inds_bm <- as.numeric(rownames(features_bm))
+
+all.equal(sort(target_bm), sort(target_f4))
+
+all.equal(sort(train_inds_bm), sort(train_inds_fold4))
+```

### 2020-03-11 analysis/inspect-glmnet-models.Rmd
workflowr::wflow_publish("analysis/inspect-glmnet-models.Rmd")

@@
 library("magrittr")
+library("plotmo")
 
@@
   Supplying a custom lambda sequence does not make much sense since the internal heuristics are quite good (if one wants to use non-spatial optimization).
+  See this [stats.stackexchange question](https://stats.stackexchange.com/a/415248/101464) for how lambda defaults are estimated.
 - To conduct a spatial optimization, one needs to use `glmnet()` directly. 
@@
 
-## Train/predict manually for Fold 1 via {glmnet}
+## Train/predict via {glmnet} manually
 
@@
 
+```{r}
+plotres(modf4)
+```
+
+```{r}
+plot_glmnet(modf4$glmnet.fit)
+```
+
+```{r}
+plotmo(modf4$glmnet.fit)
+```
+
 Let's figure out which are the ten most important features and create PDPs for these:
@@
 
+Let's look at the x values for observation 737:
+
+```{r}
+obs_train_f4[737, top_ten_abs]
+```
+
+Looks ok.
+
 ## Compare arguments of the `do.call` call in the benchmark with the manual one

### 2020-03-11 analysis/inspect-glmnet-models.Rmd
workflowr::wflow_publish("analysis/inspect-glmnet-models.Rmd")

@@
 ```{r}
-models <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["models"]]
+models <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["models"]]
 ```
@@
 
-We see a high error on Fold 2 (= Laukiz 2).
+We see a high error on Fold 4 (= Laukiz 2).
 The others are also quite high but not "out of bounds".
-
-Because this models used the internal optimization of the lambda sequence (`cv.glmnet`), let's look at the value which was chosen for prediction (parameter `s` which defaults to `s="lambda.1se"`):
-
-```{r}
-purrr::map_dbl(models, ~ .x[["learner.model"]][["lambda.1se"]])
-```
-
-It seems that the `lambda.1se` value for Fold 4 is way higher than for the other 3 folds.
-However, all values seem to be quite high.
 
@@
 
-Interestingly, the lambda length of fold 1 is not 100 (default) but only 5.
+Interestingly, the lambda length of model 1 is not 100 (default) but only 5.
 
@@
 
-To inspect further, let's refit a {glmnet} model directly on the training data of Fold 1 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:
+To inspect further, let's refit a {glmnet} model directly on the training data of Fold 4 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:
 
@@
 ```
-
-Ok, a value of `0.85` is **very** different to what happened during the CV (4.211054e+08).
 
@@
 Ok, how does this compare to summaries of other observations?
+NB: Obs. 500 - 510 were chosen randomly.
+The purpose here is to see if something within the predictors for specific observations looks abnormal.
 
@@
 
-Feature "bf2_PRI_norm" has a quite high value.
+Feature "bf2_PRI_norm" has a quite high value (-95).
 
@@
   dplyr::pull(feature)
+
+top_ten_abs
 ```
@@
 
-Looks ok.
-
-## Compare arguments of the `do.call` call in the benchmark with the manual one
-
-`args` were saved during a debug call of `mlr::benchmark()`
-
-```{r}
-args <- readRDS("args-bm-laukiz2.rda")
-target_bm <- args$y
-features_bm <- args$x
-
-train_inds_bm <- as.numeric(rownames(features_bm))
-
-all.equal(sort(target_bm), sort(target_f4))
-
-all.equal(sort(train_inds_bm), sort(train_inds_fold4))
-```
+Looks ok - they are all within a normal range with respectv to the PDP estimates.

### 2020-03-16 docs/00-manuscripts/ieee/ieee-manuscript.tex
add conclusion

@@
 Hence for this study the word ‘high-dimensionality‘ only applies to the
-experiments using the ‘NRI‘ feature set (> 7000 features for 1808 observations).
+experiments using the ‘NRI‘ feature set ($>$ 7000 features for 1808 observations).
 
@@
 
+Even though the feature selection task was handed over to the filter methods in this study, we ensured to not include features with a pair-wise correlation of 1.
+Having such can cause undesired effects during model fitting and feature importance calculation.
+To account for such, all pair-wise correlations between features were calculated.
+For pairs which exceeded the threshold of 0.9999999999 the one with the largest mean absolute correlation was removed.
+This process was repeated $p$ times, each time calculating a new correlation matrix.
+
+This preprocessing step resulted in the following final number of predictors: HR (122), VI (86), NRI ().
+
 \subsubsection{Hyperparameter Optimization}
@@
 SVM showed the best performance of all learners with an absolute difference up around three percentage points to the next model (RF) (\autoref{tab:best-learner-perf}).
+A high intra-fold variance was observed: Predicting on Luiando resulted in an RMSE value of 9.0 for learner SVM (without filter) but up to 54.26 RMSE when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).
 
@@
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI), for HR-NRI and HR the error was around 40 RMSE, and for task NRI the Ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
-%All penalized methods tuned using MBO showed exactly the same error across all tasks.
+Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI).
+For HR-NRI and HR the error was around 40 RMSE, and for task NRI the Ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
+In all settings for which Ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
+This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.
+% FIXME: All penalized methods tuned using MBO showed exactly the same error across all tasks.
 
@@
 \input{performance-best-per-learner.tex}
+
+% latex table top 20 absolute performances
+\input{performance-svm-single-plot.tex}
 
@@
 
-\subsection{Permutation-based variable importance}
+\subsection{Variable importance}
+
+\subsubsection{Permutation-based Variable Importance}
 
@@
 For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm.
-For feature set HR four features in the infrared region (920 nm  - 1000 nm) were identified to be important by causing a mean RMSE decrease of around one percentage point.
+For feature set HR four features in the infrared region (920 nm  - 1000 nm) were identified by the model to be important due to causing a mean RMSE decrease of around one percentage point.
 Overall, most features show a small importance only with average decreases in RMSE below 0.5.
@@
 
-\noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
-When using no buffer at all, the possibility is high that a pixel value gets assigned to a tree which does not spatially match with the actual recorded hyperspectral reflectance due to the geometric offset of 1 m.
-
+\noindent The decision to use a buffer of 2 m to generate the index value for each observation has to be seen critical.
+Due to the reported geometric offset of up to 1 m for the hyperspectral data, the possibility of assigning a wrong pixel value to a tree observation would be high.
 Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
 In our view using a buffer of 2 m is a good compromise between both worlds.
+In addition, a side-analysis showed that some trees would be assigned missing values when relying on a single pixel.
+This then would result in the removal of that particular tree which we wanted to avoid.
+Even though we provide no results showing the influence of multiple buffer values on the extraction, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
 
 Another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-As the buffer is circular, determining the total number of contributing pixels depends on the exact location of a tree within that pixel as this sets the base how many pixels in each direction will be touched by the buffer.
-If a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
-
-The magnitude of bias introduced by these facts cannot be quantified.
-However, it has to be considered when making interpretations about the outcome of this study.
+According to the extract function of the \texttt{raster} package, a pixel is included if the buffer of the point observation hits the centroid of the respective pixel, not just any part of it.
+As the buffer is circular, determining the total number of contributing pixels for each tree depends on the exact location of a tree within the pixel grid.
+Next, if a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
+The magnitude of uncertainty introduced by these facts cannot be quantified.
+However, we assume these to be of minor importance for the subsequent model fitting step.
 
 \subsection{Performance vs.\ plot characteristics}
-% TODO: Really needed?
+\label{subsec:perf-plot-char}
+
+\noindent The high differences in RMSE observed for the fold performances (\autoref{tab:svm-single-fold-perf}) are a result of model overfitting.
+An RMSE of 54.26 indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
+Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
+The model overfitted on learning relationships related to medium-high defoliation values but missed out on low defoliation values.
+In cases when Laukiz2 is in the training set, the model is able to half the error and even reduce it down to a value of 9 RMSE in the case of Luiando.
 
 \subsection{Predictive Performance}
+The best aggregated performance of this study (SVM + Carscore filter, 27.98 RMSE) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be 19.
+However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
+With the clear presence of model overfitting in this study it can be postulated that more training data representing all health stages of a plot is needed.
+A model can only make robust predictions if it has learned relationships across the whole range of the response.
+Hence, care should be taken when making statements about the overall performance of the fitted models of this study due to their high variance when predicting.
+However, when inspecting the fold level performances, one can say that the model does a decent job predicting defoliation ranging from 50\% to 100\% but fails for 0\% - 50\%.
+This applies to all learners of this study.
 
 \subsubsection{Algorithm differences}
+An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
+% FIXME: Add refs
+In the environmental community we have rarely seen a study in the last years in which SVM outperforms other models that clearly.
+The absolute difference of around three percentage points is also worth noting: We could not find a study in which SVM came out as the best model in such clear way.
+All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).
+
+We were a bit surprised about the performances of the penalized methods Lasso and Ridge.
+The former was not able to achieve an error below 45.96 RMSE across any dataset while Ridge came at least somewhat close to RF and XGBoost for some feature sets.
+However, there were also three instances (VI, NRI-VI, HR-NRI-VI) for which the error of the fitted Ridge model was out of bounds.
+We cannot preclude possible mistakes on our side when optimizing the penalization terms of the penalized learners.
+
+% FIXME: Talk about L1 and L2 models
 
 \subsubsection{Feature set differences}
+One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
+With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial changes in performance.
+RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
+We have no explanation for this specific result, especially because the SVM does not show this difference.
+% FIXME: Talk about L1 and L2 models -> there we have quite some differences!
 
@@
 
-% - Ensemble FS show no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
+% - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
 % - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
@@
 \subsection{Comparison to other studies}
-% NOTE: no environmental studies use filter methods, some use FFS
-% NOTE: only few studies analyze defoliation
+% - no environmental studies use filter methods, some use FFS
+% - only few studies analyze defoliation
+% - most FS is done in bioinformatics
 
 \noindent Most other studies analyzing defoliation operated on the plot rather than the tree level.
-This originated simply due to spatial resolution of the satellite products that served as the input data\cite{townsend2012, debeurs2008, rengarajan2016}.
+This is due to the low spatial resolution of used satellite products which served as the input data, making a tree-level study infeasible \cite{townsend2012, debeurs2008, rengarajan2016}.
 
 Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
-\~cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
+\cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
 Both study designs are substantially different compared to the setup of this work.
 In addition, no spatial \ac{CV} or \ac{FS} was used.
-\~cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
+\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
 Study results indicated that spectral metrics were found to be most helpful for the model.
@@
 
-\~cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
-\~cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
-A grid-search was used for hyperparameter tuning and a \ac{FFS} for variable selection.
-\~cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach\cite{meyer2018} on a Random Forest classifier.
+\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable (woody cover).
+\cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees for eucalyptus forests.
+A grid-search was used for hyperparameter tuning and \ac{FFS} for variable selection.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a Random Forest classifier.
 
 In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
-Most studies used only one algorithm (usually Random Forest) without strong arguments why this particular one has been selected.
-In our view the number of possible derivable input features is often higher than the actual number of variables used in the end, missing out on potentially helpful information from the raw data.
-These findings underline the importance of this work in the field of forest health modeling.
+Most studies used only one algorithm (usually Random Forest) without (strong) arguments why this particular one has been selected or why only one model was used.
+This is not surprising: Most environmental/ecological datasets are not high-dimensional.
+In contrast, the number of predictors if often monadic and issues related to correlations are solved manually instead of relying on an automated approach.
+The bioinformatics field faces high-dimensional feature sets more often.
+Hence more studies using (filter-based) feature-selection approaches can be found for this field \cite{guo2019, radovic2017}.
+If a field only rarely faces high-dimensional dataset issues the motivation and expertise of using advanced methods to solve such are rather low.
+We hope that this work can give some guidance and serve as a starting point for tackling high-dimensional problems in environmental modeling.
 
 \section{Outlook and conclusion}
+This study analyzed defoliation at trees in north Spain by using hyperspectral data as input for machine learning models making heavy use of filter-based feature selection methods.
+It was shown that substantial differences in performance can occur depending on which feature selection methods and machine learning algorithms are combined.
+SVM showed the most robust behavior across all highly-correlated datasets and was able to predict the response variable of this study substantially better than other methods.
+
+Filter methods showed their ability to improve predictive performance for datasets with many features.
+Ensemble filter methods did not show a substantial improvement over legacy filter methods.
+
+The potential of predicting defoliation for single trees with the given toolset of this study was rather limited, seen on the average error of 27\%. 
+However, with better training data covering more variety of the response variable, performance could be highly improved in future studies.
 

### 2020-03-16 docs/00-manuscripts/ieee/ieee-manuscript.tex
update outlook structure

@@
 
-	\item Does the combination of feature sets have an substantial effect on predictive performance?
+	\item Does combining feature sets have an substantial effect on predictive performance?
 
@@
 \noindent We want to stress the differences between 'dimension reduction' and 'high-dimensionality'.
-The former refers to the general idea of reducing features from a dataset\cite{vandermaaten2007}.
+The former refers to the general idea of reducing features from a dataset \cite{vandermaaten2007}.
 In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
-'High-dimensionality' in contrast is a dataset attribute and applies when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of observations\cite{hastie2001}.
+'High-dimensionality' in contrast is a dataset attribute and applies when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of observations \cite{hastie2001}.
 There is no absolute value at which the term applies, both for \(n\) and \(p\).
-Hence for this study the word ‘high-dimensionality‘ only applies to the
-experiments using the ‘NRI‘ feature set ($>$ 7000 features for 1808 observations).
+Hence for this study the word ‘high-dimensionality‘ only applies to the experiments using the ‘NRI‘ feature set ($>$ 7000 features for 1808 observations).
 
@@
 % Filter methods
-\noindent The concept of filters originates on the idea of ranking features using certain heuristics of an algorithm\cite{chandrashekar2014}.
+\noindent The concept of filters originates on the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
 Some filter methods are restricted towards specific types of variables (numeric or nominal).
@@
 The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
-In well-implemented software solutions the filter calculation is only done once and then cached, saving comptutational resources\cite{mlr}.
+In well-implemented software solutions the filter calculation is only done once and then cached, saving computational resources \cite{mlr}.
 
@@
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum', 'mean' or 'sum' are able to enhance the predictive performance of the resulting models\cite{abeel2010, drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum', 'mean' or 'sum' are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
 This approach is referred to as 'ensemble filter'~\cite{dietterich2000}.
@@
 
-This preprocessing step resulted in the following final number of predictors: HR (122), VI (86), NRI ().
+This preprocessing step resulted in the following final number of predictors: HR (122), VI (86), NRI (7467).
 
@@
 \noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
-Model-based Optimization\cite{mlrmbo} (MBO) was used for hyperparameter optimization.
+Model-based Optimization \cite{mlrmbo} (MBO) was used for hyperparameter optimization.
 This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
@@
 
+% RQ3: How are feature-selection methods influencing the predictive performance of the models?
 Filter methods showed their ability to improve predictive performance for datasets with many features.
@@
 
+% RQ2: Does combining feature sets have an substantial effect on predictive performance?
+% RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
+Creating feature sets composed out of derivated features did not help.
+In contrast, dataset VI showed a substantial worse performance than the dataset with the original features (HR).
+Combining feature sets did not show a substantial improvement on predictive performance.
+
+% RQ4: Which features are most important for the models and how can these be interpreted in an ecological context?
+Features along the red-edge wavelength region were most important for models to achieve good predictions.
+With respect to dedicated vegetation indices, the "Vogelmann" index with all of its versions was seen as the most important index for the tested SVM algorithm.
+
+% What about the general potential of analyzing defol with hyperspectral data?
 The potential of predicting defoliation for single trees with the given toolset of this study was rather limited, seen on the average error of 27\%. 

### 2020-03-17 analysis/feature-importance.Rmd
workflowr::wflow_publish("analysis/feature-importance.Rmd")

@@
 library("patchwork")
+library("iml")
 
@@
   fi_permut_vi_buffer2,
+
+  fi_ale_hr_buffer2,
 
@@
 ```
+
+# ALE plots
+
+ALE plots via package {iml}
+
+```{r}
+fi_ale_hr_buffer2$plot(ncols = 2)
+```

### 2020-03-22 analysis/feature-importance.Rmd
workflowr::wflow_publish("analysis/feature-importance.Rmd")
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   fi_ale_hr_buffer2,
+  fi_ale_vi_buffer2,
+  fi_ale_hr_buffer2_gs20,
+  fi_ale_vi_buffer2_gs20,
 
@@
       override.aes = list(
-        linetype = c("blank", "blank", "solid", "solid", "solid" ,"solid"),
+        linetype = c("blank", "blank", "solid", "solid", "solid", "solid"),
         shape = c(20, 1, NA, NA, NA, NA),
@@
       override.aes = list(
-        linetype = c("solid", "solid", "solid" ,"solid","dashed", "dashed"),
+        linetype = c("solid", "solid", "solid", "solid", "dashed", "dashed"),
         shape = c(NA, NA, NA, NA, 20, 1),
@@
 
-# Plots by dataset
+# Plots by dataset {.tabset .tabset-fade}
 
@@
 
-# ALE plots
+# ALE plots {.tabset .tabset-fade}
 
@@
 
+## P2 HR
+
+### Grid size: 100
+
+Top ten HR features from permutation Vimp
+
+```{r fi-hr-ale, dev = c("png", "pdf")}
+fi_ale_hr_buffer2$plot(
+  features = df_wavelengths_from_hr$class,
+  ncol = 2
+)
+```
+
+### Grid size: 20
+
+Top ten HR features from permutation Vimp
+
 ```{r}
-fi_ale_hr_buffer2$plot(ncols = 2)
+fi_ale_hr_buffer2_gs20$plot(
+  features = df_wavelengths_from_hr$class,
+  ncol = 2
+)
 ```
+
+## P3 VI
+
+### Grid size: 100
+
+Top ten VI features from permutation Vimp
+
+```{r fi-vi-ale, dev = c("png", "pdf")}
+fi_ale_vi_buffer2$plot(
+  features = paste0("bf2_", df_wavelengths_from_indices_label$class),
+  ncol = 2
+)
+```
+
+### Grid size: 20
+
+Top ten HR features from permutation Vimp
+
+```{r}
+fi_ale_vi_buffer2_gs20$plot(
+  features = paste0("bf2_", df_wavelengths_from_indices_label$class),
+  ncol = 2
+)
+```

### 2020-03-22 analysis/eda.Rmd
remove copy_figures() from eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
     label = expression(bold(atop("n = 559", bar(x) ~ "= 57.23"))), x = 1,
-    y = 112, size = 3, colour = "#BC3C29", fontface = 2
+    y = 112, size = 4, colour = "#BC3C29", fontface = 2
   ) +
@@
     label = expression(bold(atop("n = 451", bar(x) ~ "= 13.54"))), x = 2,
-    y = 112, size = 3, colour = "#0072B5", fontface = 2
+    y = 112, size = 4, colour = "#0072B5", fontface = 2
   ) +
@@
     label = expression(bold(atop("n = 301", bar(x) ~ "= 68.36"))), x = 3,
-    y = 112, size = 3, colour = "#E18727", fontface = 2
+    y = 112, size = 4, colour = "#E18727", fontface = 2
   ) +
   annotate("text",
-    label = expression(bold(atop("n = 407", bar(x) ~ "= 69.07"))), x = 4,
-    y = 112, size = 3, colour = "#20854E", fontface = 2
+    label = expression(bold(atop("n = 497", bar(x) ~ "= 69.07"))), x = 4,
+    y = 112, size = 4, colour = "#20854E", fontface = 2
   ) +
   scale_color_nejm() +
+  ggpubr::theme_pubr(base_size = 14) +
   theme(legend.position = "none") +
@@
 ```
-
-```{r echo=FALSE}
-copy_figures()
-```

### 2020-03-23 docs/00-manuscripts/ieee/ieee-manuscript.tex

--- 2020-03-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 
-\noindent The decision to use a buffer of 2 m to generate the index value for each observation is questionable.
-When using no buffer at all, the possibility is high that a pixel value gets assigned to a tree which does not spatially match with the actual recorded hyperspectral reflectance due to the geometric offset of 1 m.
-
+\noindent The decision to use a buffer of 2 m to generate the index value for each observation has to be seen critical.
+Due to the reported geometric offset of up to 1 m for the hyperspectral data, the possibility of assigning a wrong pixel value to a tree observation would be high.
 Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
 In our view using a buffer of 2 m is a good compromise between both worlds.
+In addition, a side-analysis showed that some trees would be assigned missing values when relying on a single pixel.
+This then would result in the removal of that particular tree which we wanted to avoid.
+Even though we provide no results showing the influence of multiple buffer values on the extraction, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
 
 Another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-As the buffer is circular, determining the total number of contributing pixels depends on the exact location of a tree within that pixel as this sets the base how many pixels in each direction will be touched by the buffer.
-If a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
-
-The magnitude of bias introduced by these facts cannot be quantified.
-However, it has to be considered when making interpretations about the outcome of this study.
+According to the extract function of the \texttt{raster} package, a pixel is included if the buffer of the point observation hits the centroid of the respective pixel, not just any part of it.
+As the buffer is circular, determining the total number of contributing pixels for each tree depends on the exact location of a tree within the pixel grid.
+Next, if a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
+The magnitude of uncertainty introduced by these facts cannot be quantified.
+However, we assume these to be of minor importance for the subsequent model fitting step.
 
 \subsection{Performance vs.\ plot characteristics}
-% TODO: Really needed?
+\label{subsec:perf-plot-char}
+
+\noindent The high differences in RMSE observed for the fold performances (\autoref{tab:svm-single-fold-perf}) are a result of model overfitting.
+An RMSE of 54.26 indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
+Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
+The model overfitted on learning relationships related to medium-high defoliation values but missed out on low defoliation values.
+In cases when Laukiz2 is in the training set, the model is able to half the error and even reduce it down to a value of 9 RMSE in the case of Luiando.
 
 \subsection{Predictive Performance}
+The best aggregated performance of this study (27.98 RMSE) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+When leaving out Laukiz2 during aggregation, the mean RMSE would be 19.
+Taking into account the overfitting, more training data representing the health stages of a plot is needed.
+A model can only make robust predictions if it has learned relationships which span the whole range of the response variable.
+Hence, the aggregated mean RMSE of 27 here should not be used to make statements about the trained model of this study.
+In contrast, by inspecting the fold level performances, one can say that the model does a decent job predicting defoliation ranging from 50\% to 100\% but fails for 0\% - 50\%.
+This applies to all learners of this study.
 
@@
 Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
-\~cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
+\cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
 Both study designs are substantially different compared to the setup of this work.
 In addition, no spatial \ac{CV} or \ac{FS} was used.
-\~cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
+\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
 Study results indicated that spectral metrics were found to be most helpful for the model.
@@
 
-\~cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
-\~cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
+\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable.
+\cite{shendryk2016} used machine-learning models with \ac{ALS} data to study dieback of trees for eucalyptus forests.
 A grid-search was used for hyperparameter tuning and a \ac{FFS} for variable selection.
-\~cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach\cite{meyer2018} on a Random Forest classifier.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and a relatively novel spatial \ac{FS} approach\cite{meyer2018} on a Random Forest classifier.
 

### 2020-03-23 analysis/about.Rmd

@@
 
-José Cortés
+José Cortés [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0003-2567-8689)
 
@@
 - What to do first (Feature selection or tuning): https://datascience.stackexchange.com/a/19778/24475
+- {glmnet}: How to select lambda and alpha: https://stats.stackexchange.com/questions/357190/glmnet-how-to-select-lambda-and-alpha/415248#415248
 - Hastie, T., Friedman, J., & Tibshirani, R. (2001). The Elements of Statistical Learning. Springer New York. https://doi.org/10.1007/978-0-387-21606-5 (Kapitel 16)

### 2020-03-23 analysis/inspect-glmnet-models.Rmd

@@
   benchmark_models_new_penalized_mbo_buffer2,
-  task_new_buffer2
+  task_new_buffer2,
+  task_new_buffer2_reduced_cor
 )
@@
 qplot(pred, geom = "histogram")
+```
+
+```{r}
+pred[which(pred > 100), , drop = FALSE]
 ```
@@
 Looks ok - they are all within a normal range with respectv to the PDP estimates.
+
+## Train/predict via {glmnet} manually on dataset with filtered feature correlation
+
+To inspect further, let's refit a {glmnet} model directly on the training data of Fold 4 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:
+
+```{r}
+train_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["train.inds"]][[4]]
+
+obs_train_f4 <- as.matrix(task_new_buffer2_reduced_cor[[2]]$env$data[train_inds_fold4, getTaskFeatureNames(task_new_buffer2_reduced_cor[[2]])])
+target_f4 <- getTaskTargets(task_new_buffer2_reduced_cor[[2]])[train_inds_fold4]
+```
+
+Fit `cv.glmnet`
+
+```{r}
+set.seed(1)
+modf4 <- glmnet::cv.glmnet(obs_train_f4, target_f4, alpha = 0)
+
+modf4$lambda.1se
+```
+
+Predict on Laukiz 2 now.
+
+```{r}
+pred_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["test.inds"]][[4]]
+
+obs_pred_f4 <- as.matrix(task_new_buffer2_reduced_cor[[2]]$env$data[pred_inds_fold4, getTaskFeatureNames(task_new_buffer2_reduced_cor[[2]])])
+
+pred <- predict(modf4, newx = obs_pred_f4, s = modf4$lambda.1se)
+```
+
+Calculate the error
+
+```{r}
+truth <- task_new_buffer2_reduced_cor[[2]]$env$data[pred_inds_fold4, "defoliation"]
+
+mlr:::measureRMSE(truth, pred)
+```

### 2020-03-23 docs/00-manuscripts/ieee/ieee-manuscript.tex
rephrase introducton part

--- 2020-03-16 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
-%%%% using 'arara' 4.0
-% arara: pdflatex: {synctex: yes, interaction: nonstopmode}
-% arara: bibtex
-% arara: latexindent
-% arara: pdflatex: {synctex: yes, interaction: nonstopmode}
-% arara: pdflatex: {synctex: yes, interaction: nonstopmode}
-
-%% bare_jrnl.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% see http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
 \documentclass[letterpaper, peerreview]{IEEEtran}
@@
 % \documentclass[journal]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
 
@@
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at:
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is 'Using Imported Graphics in
-% LaTeX2e' by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in 'jaggedy'/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
 
@@
 % many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
 
 % *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
 
@@
 % *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
+\ifCLASSOPTIONcompsoc
+ \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+\else
+ \usepackage[caption=false,font=footnotesize]{subfig}
+\fi
 % subfig.sty, written by Steven Douglas Cochran, is the modern replacement
@@
 % on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
 
@@
 
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and
-% Axel Sommerfeldt. This package may be useful when used in conjunction with
-% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
-% submissions have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from 'marking' where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.4.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/pkg/endfloat
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a
-% page by themselves.
-
 % *** PDF, URL AND HYPERLINK PACKAGES ***
@@
 % systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
 
@@
 
-% FIXME: Rephrase
-\noindent \ac{RF}, \ac{SVM} and extreme gradient boosting were compared using six feature sets across different feature selection methods, including a new promising approach of ensemble filters.
-Bayesian Optimization, also known as model-based optimization, was used for hyperparameter tuning.
-The feature set size was optimized at the same time.
-By this, we avoided multiple optimization layers which may result in a biased performance estimation or long model fitting times.
+\noindent \ac{RF}, \ac{SVM} and extreme gradient boosting were fitted on six feature sets using different feature selection methods.
+Bayesian Optimization, also known as model-based optimization, was used for hyperparameter tuning and optimization of the number of features.
+By merging all optimization steps into a single step, multiple optimization stages, which may introduce bias cause long model fitting times, were avoided.
 A spatial \ac{CV} approach on the plot level was used to account for spatial autocorrelation in the data.
@@
 After these \textit{n} tries have been evaluated, a new hyperparameter setting, going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
-This strategy continues until a termination criterion, defined by the user, is reached\cite{hutter2011, jones1998}.
+This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.
 
@@
 The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum.
-\ac{MBO} shines when being compared to methods that do not use information from previous runs, such as random search or grid search\cite{bergstra2012}.
-For the Lasso and Ridge models however the internal non-spatial CV approach due to promising results in previous applications.
+\ac{MBO} shines when being compared to methods that do not use information from previous runs, such as random search or grid search \cite{bergstra2012}.
 
@@
 
-\noindent A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible\cite{schratz2019, sperrorest}.
+\noindent A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
 Each plot served as one fold within the cross-validation setting, resulting in four iterations total.
@@
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
-Methods like \ac{PDP} do not produce reliable estimates in such situations because they create unrealistic situations between covariates.
+Methods like \ac{PDP} do not produce reliable estimates in such scenarios because unrealistic situations between covariates are created \cite{molnar2019}.
+% FIXME: Add ref
 Due to the noisiness of the dataset, estimates from model-agnostic approaches such as permutation-based feature importance should also be taken with care.
+This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
+The more important the feature is for the model, the higher the variance in the measure, reflecting the importance of the specific feature for the current model.
+Permuting every feature $y$ times comes with some computational cost.
+To balance the effort/gain ratio and cope with limited visual space in plots, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.
 
 However, as of today there are no unbiased approaches for estimating feature importance for (highly-) correlated datasets.
-Nevertheless, recent methods like \ac{ALE} plots aim to tackle this problem by utilizing marginal distributions of features over a small window to focus on the effect of a single feature.
-However,
+Nevertheless, recent methods like \ac{ALE} plots aim to tackle this problem by utilizing marginal distributions of features over a small window to focus on the effect of a single feature \cite{apley2019, molnar2019}.
+However, ALE plots need visual inspection and as the number of features increases, making objective comparisons becomes more complicated.
 
@@
 With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while keeping modest on over-interpreting the results.
-We used the algorithm which showed the best performance for the HR task to calculate the feature importance.
-
-This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
-The more important the feature is for the model, the higher the loss in the used error measure will be, reflecting the importance of the specific feature for the current model.
-Permuting every feature $y$ times comes with some computational cost.
-To balance the effort/gain ratio and to be able to relate the results to wavelength regions, we decided to estimate feature importance only for the dataset with the raw reflectance values (HR).
+We used the algorithm which showed the best performance for the HR (SVM) task to calculate the feature importance.
 
 \subsection{Linking feature importance to wavelength regions}
+
 \noindent For ecological interpretation purposes we linked the ten most important indices of the winning models for each feature set to the spectral regions of the hyperspectral data.
@@
 
+% FIXME: Update used packages
 \noindent The complete study was done using the open-source statistical programming language R\cite{rcoreteam2018}.
@@
 Datasets containing derived features (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
+All learners besides SVM show a substantially worse performance on the VI dataset (around five percentage points).
 
@@
 % FIXME: All penalized methods tuned using MBO showed exactly the same error across all tasks.
+% FIXME: Mention the bad perf of RIDGE on VI
 
@@
 
-% latex table top 20 absolute performances
+% latex table top 10 absolute performances
 \input{performance-top-10}
+
+% latex table worst 10 absolute performances
+%\input{performance-worst-10}
 
@@
 
-% FIXME: explain in methods why only 2 datasets are shown
 The most important features of each dataset showed an average decrease in RMSE of 1.57 (HR, B69) and 1.79 (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
-For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm.
-For feature set HR four features in the infrared region (920 nm  - 1000 nm) were identified by the model to be important due to causing a mean RMSE decrease of around one percentage point.
-Overall, most features show a small importance only with average decreases in RMSE below 0.5.
+For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm (the so called "red-edge").
+For feature set HR four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around one percentage point).
+Overall, most features show only a small importance with average decreases in RMSE below 0.5.
 
@@
 
+\subsubsection{ALE plots}
+
+In addition to permutation-based variable importance, \ac{ALE} plots were used to analyze the importance of features.
+ALE plots
+
 \section{Discussion}
@@
 
-% \section{Spectral signatures of each plot}
-
-% \begin{figure} [ht]
-% 	\begin{center}
-% 		\includegraphics[width=0.48\textwidth] {spectral-signatures-1.pdf}
-% 		\caption{Spectral signatures (mean and standard deviation) of Laukiz1, Laukiz2, Luiando and Oiartzun.}\label{fig:spectral-signatures}
-% 	\end{center}
-% \end{figure}
+\section{SVM ALE plots for task HR}
+
+\begin{figure} [ht]
+	\begin{center}
+		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
+		\caption{ALE plots on dataset HR of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance.}\label{fig:fi-hr-ale}
+	\end{center}
+\end{figure}
+
+\pagebreak
+
+\section{SVM ALE plots for task VI}
+
+\begin{figure} [ht]
+	\begin{center}
+		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
+		\caption{ALE plots on dataset VI of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance.}\label{fig:fi-vi-ale}
+	\end{center}
+\end{figure}
 

### 2020-03-23 docs/00-manuscripts/ieee/ieee-manuscript.tex
update used packages and biblio

@@
 
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the 'draftcls' or 'draftclsnofoot', not 'draft', class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex,
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that the IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command,
-% and the \label for the overall figure must come after \caption.
-% \hfil is used as a separator to get equal spacing.
-% Watch out that the combined width of all the subfigures on a
-% line do not exceed the text width or a line break will occur.
-%
-%\begin{figure*}[!t]
-%\centering
-%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
-%\label{fig_second_case}}
-%\caption{Simulation results for the network.}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat[]), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
-% labels, the optional argument to \subfloat must be present. If a
-% subcaption is not desired, just leave its contents blank,
-% e.g., \subfloat[].
-
-% An example of a floating table. Note that, for IEEE style tables, the
-% \caption command should come BEFORE the table and, given that table
-% captions serve much like titles, are usually capitalized except for words
-% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
-% and up, which are usually not capitalized unless they are the first or
-% last word of the caption. Table text will default to \footnotesize as
-% the IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that the IEEE does not put floats in the very first column
-% - or typically anywhere on the first page for that matter. Also,
-% in-text middle ('here') positioning is typically not used, but it
-% is allowed and encouraged for Computer Society conferences (but
-% not Computer Society journals). Most IEEE journals/conferences use
-% top floats exclusively.
-% Note that, LaTeX2e, unlike IEEE journals/conferences, places
-% footnotes above bottom floats. This can be corrected via the
-% \fnbelowfloat command of the stfloats package.
-
 \section{Data and study area}
@@
 % chktex 8
-The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage\cite{iturritxa2014}.
-The dieback of these trees, which are mainly used as timber, causes high economic damages\cite{ganley2009}.
+The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
+The dieback of these trees, which are mainly used as timber, causes high economic damages \cite{ganley2009}.
 
@@
 
-% FIXME: Update used packages
-\noindent The complete study was done using the open-source statistical programming language R\cite{rcoreteam2018}.
+\noindent The complete study was done using the open-source statistical programming language R \cite{rcoreteam2019}.
+All R packages used in this study can be found in linked repositories.
+Due to space limitations we will only mention the packages of the used algorithms and filter methods.
+
 The algorithm implementations of the following packages have been used: \textit{xgboost}\cite{chen2016} (\textit{xgboost}), \textit{kernlab}\cite{kernlab} (Support Vector Machine) and \textit{glmnet}\cite{glmnet} (Ridge Regression).
-The filter methods of the following packages were used: \textit{praznik\cite{praznik}}, \textit{FSelectorRcpp}\cite{fselectorrcpp}.
-The R package \textit{mlr}\cite{mlr} was used for all modeling related steps.
-\textit{drake} was used for structuring the work and ensuring reproducibility.
+The filter implementations of the following packages have been used: \textit{praznik\cite{praznik}}, \textit{FSelectorRcpp}\cite{fselectorrcpp}.
+The R package \textit{mlr} \cite{mlr} was used for all modeling related steps.
+\textit{drake} \cite{drake} was used for structuring the work and reproducibility.
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
+The code base is available on GitHub (\url{https://github.com/pat-s/2019-feature-selection}).
 

### 2020-03-23 docs/00-manuscripts/ieee/ieee-manuscript.tex
write disc: feature selection

@@
 This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.
-% FIXME: All penalized methods tuned using MBO showed exactly the same error across all tasks.
-% FIXME: Mention the bad perf of RIDGE on VI
+
+FIXME: Mention: All penalized methods tuned using MBO showed exactly the same error across all tasks.
+FIXME: Mention the bad perf of RIDGE on VI
 
@@
 An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
-% FIXME: Add refs
-In the environmental community we have rarely seen a study in the last years in which SVM outperforms other models that clearly.
-The absolute difference of around three percentage points is also worth noting: We could not find a study in which SVM came out as the best model in such clear way.
+
+FIXME: Add refs
+There were few environmental modeling studies only in recent years in which SVM outperformed other models that clearly.
+The absolute difference of around three percentage points is also worth noting:
+We could not find a study in which SVM came out as the best model in such clear way.
 All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).
@@
 
-% FIXME: Talk about L1 and L2 models
+FIXME: Talk about L1 and L2 model performance
 
@@
 We have no explanation for this specific result, especially because the SVM does not show this difference.
-% FIXME: Talk about L1 and L2 models -> there we have quite some differences!
+
+FIXME: Talk about L1 and L2 models -> there we have quite some differences!
 
@@
 
+% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
+The usefulness of filters in this study was apparent across all feature sets, even when $p$ was rather small (HR).
+Even though the effect of feature selection methods varied across algorithms, small to substantial improvement in predictive performance were achieved in almost all cases.
+No optimal algorithm/filter/dataset setting was observed in this study, leading to the conclusion that one needs to try out various combinations of such to find the best performing one.
+
 % - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
-% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
-% - When calculating multiple filters, consider to only use ensemble filter?
+We have no explanation why the Borda ensemble method used in this study did not score better than the average performance of other filter methods.
+Ensemble methods shine in complicated scenarios when simple filter methods face problems due to their focus on certain areas of features.
+We assume that the datasets used in this study were not complex enough to bring up these problems for simple features.
+Therefore it makes sense that, due to its averaging nature, the Borda filter scored average results across all instances.
+However, this cannot be known upfront and we believe that an ensemble filter should be part of every filter portfolio when doing filter-based feature selection.
+
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
+The PCA approach did neither show the best or worst result in the lineup.
+It shines in terms of runtime since it reduces the amount of features highly.
+However, it also removes the option for a post-analysis of feature importance by squashing predictors into principal components and users need to decided upfront what they value more.
+Since filters are only calculated once due to caching, the runtime advantage might in fact be neglibile in practice.
 
@@
 		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
-		\caption{ALE plots on dataset HR of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance.}\label{fig:fi-hr-ale}
+		\caption{ALE plots on dataset HR of SVM learner.
+Subset showing the ten most important features according to the permutation-based variable importance.
+The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-hr-ale}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
-		\caption{ALE plots on dataset VI of SVM learner. Subset showing the ten most important features according to the permutation-based variable importance.}\label{fig:fi-vi-ale}
+		\caption{ALE plots on dataset VI of SVM learner.
+Subset showing the ten most important features according to the permutation-based variable importance.
+The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-vi-ale}
 	\end{center}

### 2020-03-23 docs/00-manuscripts/ieee/ieee-manuscript.tex
upd appendices spacing, upd table width

@@
 		\\
-		Name                                         & Group                             & Reference          \\
+		Name                                         & Group                             & Ref.               \\
 		\toprule
@@
 We used the algorithm which showed the best performance for the HR (SVM) task to calculate the feature importance.
+ALE plots were not discussed in greater detail and are available as supplementarty material.
 
@@
 
-\subsubsection{ALE plots}
-
-In addition to permutation-based variable importance, \ac{ALE} plots were used to analyze the importance of features.
-ALE plots
-
 \section{Discussion}
@@
 
-\subsubsection{Variable importance vs.\ filter methods}
-
-WIP
-
-\subsubsection{Ecological Interpretation}
-
-% TODO: talk about the link to RS spectra
+\subsubsection{Linking feature importance to spectral characteristics}
+
+Not surpringly the most important features for both HR and VI datasets were identified around the red egde of the spectra, specifically in the range of 700 nm to 750 nm.
+
+FIXME: Add ref
+This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and its opposite.
+It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
+Looking at the spectral curves of the plots, we can observe quite some variance in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
 
@@
 
-\pagebreak
-
 \section{SVM ALE plots for task VI}
@@
 
+\section{Correlation among filter methods}
+
 \pagebreak
-
-\section{Correlation among filter methods}
 
@@
 
+\pagebreak
+
 \bibliographystyle{IEEEtran}

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: update keywords

@@
 \begin{IEEEkeywords}
-	hyperspectral imagery, forest health modeling, machine-learning, feature-selection, model comparison
+	hyperspectral imagery, forest health monitoring, machine learning, feature selection, model comparison, imaging spectroscopy
 \end{IEEEkeywords}
@@
 We used the algorithm which showed the best performance for the HR (SVM) task to calculate the feature importance.
-ALE plots were not discussed in greater detail and are available as supplementarty material.
+ALE plots were not discussed in greater detail and are available as supplementary material.
 
@@
 However, it also removes the option for a post-analysis of feature importance by squashing predictors into principal components and users need to decided upfront what they value more.
-Since filters are only calculated once due to caching, the runtime advantage might in fact be neglibile in practice.
+Since filters are only calculated once due to caching, the runtime advantage might in fact be negligible in practice.
 
@@
 
-Not surpringly the most important features for both HR and VI datasets were identified around the red egde of the spectra, specifically in the range of 700 nm to 750 nm.
+Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 700 nm to 750 nm.
 
@@
 % RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
+FIXME: "derivated" is not a real word?
 Creating feature sets composed out of derivated features did not help.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: feature-selection - feature selection

@@
 	\acro{FPR}{false positive rate}
-	\acro{FFS}{forward feature-selection}
-	\acro{FS}{feature-selection}
+	\acro{FFS}{forward feature selection}
+	\acro{FS}{feature selection}
 	\acro{GAM}{generalized additive model}
@@
 This study aims to close this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
-We make use of the latest state-of-the-art methodology in machine learning by combining feature-selection and hyperparameter tuning across multiple algorithms.
+We make use of the latest state-of-the-art methodology in machine learning by combining feature selection and hyperparameter tuning across multiple algorithms.
 Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high-dimensionality in environmental modeling \cite{trunk1979, xu2016}.
@@
 This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
-A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature-selection optimization finished.
+A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature selection optimization finished.
 Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: change title

@@
 % Do not put math or special symbols in the title.
-\title{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
-\IEEEpeerreviewmaketitle{Supporting Ecological Decision Making Using Feature-Selection and Variable Importance}
+\title{Evaluating the Supportive Effect of Feature Selection Methods on Machine Learning Models in Monitoring Forest Health Using Hyperspectral Remote Sensing Data}
+\IEEEpeerreviewmaketitle{Evaluating the Supportive Effect of Feature Selection Methods on Machine Learning Models in Monitoring Forest Health Using Hyperspectral Remote Sensing Data}
 %

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 This goes in line with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
-At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from open-source efforts.
+At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from the open-source community.
 Scientists can nowadays relatively easily process large amounts of (environmental) information using various learning algorithms.
@@
 
-Machine learning methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping\cite{mascaro2014, urban2018}.
+Machine learning methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
 The ability of predicting to large unknown areas qualifies these tools as a promising toolset for such tasks.
-One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees\cite{hawrylo2018}.
-
-Other approaches for analyzing forest health include temporal change detection\cite{zhang2016} or describing the current health status of forests on a stand level\cite{townsend2012}.
-In such studies, the defoliation of trees serves as a proxy for forest health by describing the impact of biotic and abiotic pest triggers\cite{townsend2012, goodbody2018}.
-
-Vegetation indices have shown the potential to provide valuable information when analyzing forest health\cite{jiang2014, adamczyk2015}.
+One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \cite{hawrylo2018}.
+
+Other approaches for analyzing forest health include temporal change detection \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
+In such studies, the defoliation of trees serves as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \cite{townsend2012, goodbody2018}.
+
+Vegetation indices have shown the potential to provide valuable information when analyzing forest health \cite{jiang2014, adamczyk2015}.
 Most vegetation indices were developed with the aim of being sensitive to changes of specific wavelength regions, serving as a proxy for underlying plant processes. 
-However, often enough indices developed for different purposes than the one to be analyzed can help explaining complex relationships.
-This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help understanding the modeled relationship.
+However, often enough indices developed for different purposes than the one to be analyzed can help to explain complex relationships.
+This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship.
 A less known index type which can be derived from spectral information is called normalized ratio index (NRI).

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: optimize VI vs NRI intro

@@
 This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship.
-A less known index type which can be derived from spectral information is called normalized ratio index (NRI).
-In contrast to \ac{VI}, \ac{NRI}s are not based on an expert-based formulas following environmental heuristics but use a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
-Especially when working with hyperspectral data, hundreds of \ac{NRI} features can be derived this way.
+A less known index type which can be derived from spectral information is the normalized ratio index (NRI).
+In contrast to most \ac{VI} indices, the \ac{NRI} does not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
+Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
 

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: Rephrase intro why FS is important, add pena ref

@@
 This study aims to close this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
-We make use of the latest state-of-the-art methodology in machine learning by combining feature selection and hyperparameter tuning across multiple algorithms.
+Method-wise it makes use of machine-learning methods in combination with feature selection and hyperparameter tuning.
 Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high-dimensionality in environmental modeling \cite{trunk1979, xu2016}.
 
-Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, the fitting time of models increase substantially and the interpretation of results becomes more complicated.
-At the same time, one has to deal with the presence of spatial autocorrelation in the data.
-The dataset used in this study comes with a spatial grouping at the plot level which needs to be accounted for.
+Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
+At the same time, it is necessary to account for the presence of spatial autocorrelation in the data.
+The dataset used in this study comes with a spatial grouping at the plot level which needs to be accounted for \cite{sperrorest, pena2015}.
 

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: rephrase research questions
AB-review: Remove unncessary paragraph in intro

@@
 
-	\item Does combining feature sets have an substantial effect on predictive performance?
+	\item Can predictive performance be substantially improved by combining feature sets?
 
@@
 \end{itemize}
-
-\noindent \ac{RF}, \ac{SVM} and extreme gradient boosting were fitted on six feature sets using different feature selection methods.
-Bayesian Optimization, also known as model-based optimization, was used for hyperparameter tuning and optimization of the number of features.
-By merging all optimization steps into a single step, multiple optimization stages, which may introduce bias cause long model fitting times, were avoided.
-A spatial \ac{CV} approach on the plot level was used to account for spatial autocorrelation in the data.
 

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: style

@@
 The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation~\cite{mesanza2016, iturritxa2017}.
-In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0\% - 100\% (\autoref{fig:defol-distr}).
-% chktex 8
+In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
 The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-\noindent The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around 12 am.
+\noindent The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around noon.
 The images were taken using a AISAEAGLE-II sensor.
-All preprocessing steps (geometric, radiometric, atmospheric) have been conducted by the \ac{ICGC}.
+All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.
 The first four bands are corrupted, leaving 122 bands with valid information.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: opt method: index derivation

@@
 
-\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used during extraction of the reflectance values.
-Subsequently, the value assigned to a tree observations was the mean of all pixels which were touched by the buffer drawn around each tree.
+\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used for extraction of the reflectance values.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
-This is how the \texttt{extract()} function of the R package $\left\{raster\right\}$ handles buffer extraction.
-In total, \(\frac{125*126}{2} = 7875\) NRIs were calculated.
-Due to four corrupted bands of the sensor a total of 7471 indices were available for each observation.
+In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated due to the first four bands of the sensor being corrupted.
 

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: grammar

@@
 	\item Noise is possibly introduced into models by highly-correlated variables\cite{johnstoneiainm.2009}.
-	\item Model interpretation and prediction tasks become more challenging\cite{johnstoneiainm.2009}.
+	\item Model interpretation and prediction become more challenging\cite{johnstoneiainm.2009}.
 \end{itemize}
 
-\noindent In the following sections a brief overview about sub-categories of feature selection approaches is given.
+\noindent In the following sections a brief overview of feature selection sub-categories is given.
 Due to the focus of this study on the use of filter methods, which are a sub-group of feature selection methods, other approaches were grouped into a single section.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: remove statement about caching

@@
 % Filter methods
-\noindent The concept of filters originates on the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
+\noindent The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
 Some filter methods are restricted towards specific types of variables (numeric or nominal).
@@
 The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
-In well-implemented software solutions the filter calculation is only done once and then cached, saving computational resources \cite{mlr}.
 

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
style

@@
 	\item Model fitting times increase.
-	\item Noise is possibly introduced into models by highly-correlated variables\cite{johnstoneiainm.2009}.
-	\item Model interpretation and prediction become more challenging\cite{johnstoneiainm.2009}.
+	\item Noise is possibly introduced into models by highly-correlated variables \cite{johnstoneiainm.2009}.
+	\item Model interpretation and prediction become more challenging \cite{johnstoneiainm.2009}.
 \end{itemize}
@@
 Some filter methods are restricted towards specific types of variables (numeric or nominal).
-Filters do only rank features, they do not choose which covariates to drop or keep\cite{drotar2015}.
+Filters do only rank features, they do not choose which covariates to drop or keep \cite{drotar2015}.
 The selection of features to keep for the model fitting step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
@@
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum', 'mean' or 'sum' are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
-This approach is referred to as 'ensemble filter'~\cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance~\cite{polikar2012, feurer2015}.
-In this work the 'Borda' ensemble filter was applied~\cite{drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
+This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015}.
+In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
 For this filter, the final order is the sum of all single filters.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
style
AB-review: style

@@
 
-Filter methods can be grouped into classes: Correlation based, entropy based, linear and non-linear methods.
+Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
 It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased towards these.
@@
 	\item linear/non-linear (calculation of linear/non-linear interaction terms)
-	\item entropy/correlation (scoring based on derivations of entropy or correlation based approaches)
+	\item entropy/correlation (scoring based on derivations of entropy or correlation-based approaches)
 \end{itemize}

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: remove link between tuning and FS

@@
 % Wrapper approach
-\noindent Wrappers \cite{chandrashekar2014, kohavi1997} apply algorithms that are also used for hyperparameter optimization such as 'Random Search' or 'Generic Simulated Annealing'.
+\noindent Wrappers \cite{chandrashekar2014, kohavi1997} apply search algorithms such as 'random search' or 'generic simulated annealing'.
 First a (random) subset of features is chosen based on the selected algorithm.

### 2020-04-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: style
AB-review: address PCA content remarks

@@
 Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.
-Due to their extensive runtimes, wrappers were not considered in this work.
+Due to their high computational cost, wrappers were not considered in this work.
 
@@
 Here, the main components of the feature space are extracted and combined.
-Most often the first two extracted main components are used since these contain the major information of the covariates.
+The number of principal components needed to explain a large portion of the variance is dynamic and depends on the dataset.
+In practice, this number often ranges between two and five.
 By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
 This enables cheap model fitting with balanced loss of predictor information.
-The disadvantage of this methodology is the lack of interpretability because the main components cannot be related back to the original covariates.
+A disadvantage is the reduced interpretability due to the main components being a squashed version of the original covariates.
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: style, add pena ref

@@
 Common penalties are 'lasso' (L1) and 'ridge' (L2).
-The former does not allow the full removal of variables from the model (penalization to zero) while the latter does.
+The latter does not allow the full removal of variables from the model (penalization to zero) while the former does.
 Both penalties can also be combined.
@@
 
-The idea of splitting the features into different sets originated from the question whether feature engineered indices from reflectance values have a positive effect on model performance.
+The idea of splitting the features into different sets originated from the question whether feature engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated previously in a spectrotemporal setting \cite{pena2017}.
 Benchmarking all models on these distinct groups of features makes it easier to draw conclusions on their impact when keeping all other variables such as model type, tuning strategy and partitioning method constant.
@@
 
-Even though the feature selection task was handed over to the filter methods in this study, we ensured to not include features with a pair-wise correlation of 1.
+Even though the feature selection task was handed over to the filter methods in this study, we ensured to not include features with a pairwise correlation of 1.
 Having such can cause undesired effects during model fitting and feature importance calculation.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-Even though the feature selection task was handed over to the filter methods in this study, we ensured to not include features with a pairwise correlation of 1.
+Even though the feature selection step was passed onto the filter methods in this study, we ensured to not include features with a pairwise correlation of 1.
 Having such can cause undesired effects during model fitting and feature importance calculation.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-This preprocessing step resulted in the following final number of predictors: HR (122), VI (86), NRI (7467).
+This preprocessing step reduced the feature sets to 122 HR features, 86 VIs and 7467 NRIs.
 
@@
 
-\noindent An exhaustive hyperparameter tuning was applied during nested spatial \ac{CV} for all algorithms.
-Model-based Optimization \cite{mlrmbo} (MBO) was used for hyperparameter optimization.
+\noindent Hyperparameters were tuned using model-based optimization within a nested spatial CV \cite{mlrmbo, schratz2019}.
 This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-\noindent Hyperparameters were tuned using model-based optimization within a nested spatial CV \cite{mlrmbo, schratz2019}.
+\noindent Hyperparameters were tuned using \ac{MBO} within a nested spatial CV \cite{mlrmbo, schratz2019}.
 This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
@@
 The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum.
-\ac{MBO} shines when being compared to methods that do not use information from previous runs, such as random search or grid search \cite{bergstra2012}.
+\ac{MBO} may outperform methods that do not use information from previous iterations, such as random search or grid search \cite{bergstra2012}.
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-For the filter methods, the percentage of features was added to the models as a hyperparameter.
-For \ac{PCA}, the number of main components was tuned instead.
-Random Forest hyperparameter \texttt{\(m_{try}\)} was modified into power transformed version \(p^{m_{try}}\) (ranging from 1 to $p$) to work on relative values of the respective datasets feature count.
+For the filter methods, the percentage of dataset's features was added to the models as a hyperparameter.
+For \ac{PCA}, the number of principal components was tuned instead.
+Random Forest hyperparameter \texttt{\(m_{try}\)} was modified into power transformed \(p^{m_{try}}\) (ranging from 1 to $p$) to work on relative values of the respective datasets feature count.
 Here \(p\) is the number of features of the dataset.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 For \ac{PCA}, the number of principal components was tuned instead.
-Random Forest hyperparameter \texttt{\(m_{try}\)} was modified into power transformed \(p^{m_{try}}\) (ranging from 1 to $p$) to work on relative values of the respective datasets feature count.
+Random Forest hyperparameter \texttt{\(m_{try}\)} was modified into power transformed \(p^{m_{try}}\) (ranging from 1 to $p$) to work on relative values of the respective dataset's feature count.
 Here \(p\) is the number of features of the dataset.
-This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen out of bounds during tuning with respect to the datasets feature count.
-After the filtering, only a subset of features of the task is used for optimizing the models hyperparameters.
-The size of this subset is always different since the number of features is also optimized.
-Hence, the value of \texttt{\(m_{try}\)} needs to be created dynamically based on the datasets feature count.
+This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen out of bounds during tuning with respect to the dataset's feature count'
+After the filtering, only a subset of features of the task is used for optimizing the model's hyperparameters.
+The size of this subset is always different since the number of features is also optimized'
+Hence, the value of \texttt{\(m_{try}\)} needs to be created dynamically based on the dataset's feature count.
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 \noindent A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
-Each plot served as one fold within the cross-validation setting, resulting in four iterations total.
-For the inner level (hyperparameter tuning), \(y - 1\) folds were used (with \(y\) being the number of plots).
+The \ac{RMSE} was chosen as the error measure.
+Each plot served as one fold within the cross-validation setting, resulting in four iterations in total.
+For the inner level (hyperparameter tuning), \(k - 1\) folds were used where \(k\) is the number of plots.
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-In total the benchmarking matrix consisted of 156 experiments (6 feature sets * 3 ML algorithms * 8 feature-selection methods + 2 * 6 L1 and L2 models).
+In total the benchmarking matrix consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods + 2 $\times$ 6 L1 and L2 models).
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
-Methods like \ac{PDP} do not produce reliable estimates in such scenarios because unrealistic situations between covariates are created \cite{molnar2019}.
+Methods like \ac{PDP} do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
 % FIXME: Add ref

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
-The more important the feature is for the model, the higher the variance in the measure, reflecting the importance of the specific feature for the current model.
-Permuting every feature $y$ times comes with some computational cost.
-To balance the effort/gain ratio and cope with limited visual space in plots, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.
+The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
+Fitting many models on the dataset containing the permuted feature comes with some computational cost.
+For increased efficiency, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 
-However, as of today there are no unbiased approaches for estimating feature importance for (highly-) correlated datasets.
+However, as of today there are no completely satisfactory approaches for estimating feature importance for (highly) correlated datasets.
 Nevertheless, recent methods like \ac{ALE} plots aim to tackle this problem by utilizing marginal distributions of features over a small window to focus on the effect of a single feature \cite{apley2019, molnar2019}.
@@
 In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset.
-With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while keeping modest on over-interpreting the results.
-We used the algorithm which showed the best performance for the HR (SVM) task to calculate the feature importance.
+With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
+We used the best-performing algorithm from the HR task (i.e. SVM) to calculate the feature importance.
 ALE plots were not discussed in greater detail and are available as supplementary material.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 % - MBO tuned penalized methods are substantially worse
-\noindent Overall, the response variable 'defoliation at trees' could be modeled with an error of 28 \%.
-SVM showed no differences in RMSE across feature sets whereas other learners like RF differed up to seven percentage points (HR-NRI vs. VI) (\autoref{fig:perf-result}).
+\noindent Overall, the response variable 'tree defoliation' could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
+SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, Lasso and Ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
 Ridge and Lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 Ridge and Lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
-SVM showed the best performance of all learners with an absolute difference up around three percentage points to the next model (RF) (\autoref{tab:best-learner-perf}).
-A high intra-fold variance was observed: Predicting on Luiando resulted in an RMSE value of 9.0 for learner SVM (without filter) but up to 54.26 RMSE when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).
+SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
+Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE value of 9.0 for learner SVM (without filter) but up to 54.26 RMSE when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
-Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE value of 9.0 for learner SVM (without filter) but up to 54.26 RMSE when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).
+Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 for learner SVM (without filter) but up to 54.26 p.p.
+when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: clarify performance scores with respect to feature sets

@@
 The combination of feature sets showed small increases in performance for some learners.
-RF and XGBoost scored slightly better on the combined datasets HR-NRI and NRI-VI, respectively (\autoref{fig:perf-result}).
-Datasets containing derived features (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
+RF and XGBoost scored slightly better on the combined datasets HR-NRI and NRI-VI, respectively, compared to their standalone variants (HR, NRI, VI) (\autoref{fig:perf-result}).
+Datasets containing derived features only (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
 All learners besides SVM show a substantially worse performance on the VI dataset (around five percentage points).

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: clarify VI datasets performances
AB-review: lasso and ridge all lower case

@@
 	\item  Random Forest (RF)
-	\item  Penalized Regression (both L1 (Lasso) and L2 (Ridge))
+	\item  Penalized Regression (both L1 (lasso) and L2 (ridge))
 	\item  Support Vector Machine (SVM)
@@
 
-The algorithm implementations of the following packages have been used: \textit{xgboost}\cite{chen2016} (\textit{xgboost}), \textit{kernlab}\cite{kernlab} (Support Vector Machine) and \textit{glmnet}\cite{glmnet} (Ridge Regression).
+The algorithm implementations of the following packages have been used: \textit{xgboost}\cite{chen2016} (\textit{xgboost}), \textit{kernlab}\cite{kernlab} (Support Vector Machine) and \textit{glmnet}\cite{glmnet} (ridge regression).
 The filter implementations of the following packages have been used: \textit{praznik\cite{praznik}}, \textit{FSelectorRcpp}\cite{fselectorrcpp}.
@@
 \noindent Overall, the response variable 'tree defoliation' could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
-SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, Lasso and Ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
-Ridge and Lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
+SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
+Ridge and lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
 SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
@@
 Datasets containing derived features only (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
-All learners besides SVM show a substantially worse performance on the VI dataset (around five percentage points).
+All learners besides SVM show a substantially worse performance when using only the VI dataset (around five percentage points worse than their respective best performance).
 
@@
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with Ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI).
-For HR-NRI and HR the error was around 40 RMSE, and for task NRI the Ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
-In all settings for which Ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
+Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI).
+For HR-NRI and HR the error was around 40 RMSE, and for task NRI the ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
+In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
 This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.
@@
 
-We were a bit surprised about the performances of the penalized methods Lasso and Ridge.
-The former was not able to achieve an error below 45.96 RMSE across any dataset while Ridge came at least somewhat close to RF and XGBoost for some feature sets.
-However, there were also three instances (VI, NRI-VI, HR-NRI-VI) for which the error of the fitted Ridge model was out of bounds.
+We were a bit surprised about the performances of the penalized methods lasso and ridge.
+The former was not able to achieve an error below 45.96 RMSE across any dataset while ridge came at least somewhat close to RF and XGBoost for some feature sets.
+However, there were also three instances (VI, NRI-VI, HR-NRI-VI) for which the error of the fitted ridge model was out of bounds.
 We cannot preclude possible mistakes on our side when optimizing the penalization terms of the penalized learners.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording
AB-review: add more explanation on the prediction outlier of RR

@@
 % - all MBO penalized methods have no difference across tasks
-SVM combined with the 'Carscore' filter achieved the best performance (RMSE of 27.98) (\autoref{tab:perf-top-10}).
+SVM combined with the 'Carscore' filter achieved the best overall performance (RMSE of 27.98 p.p.) (\autoref{tab:perf-top-10}).
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
 Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI).
-For HR-NRI and HR the error was around 40 RMSE, and for task NRI the ridge learner achieved almost the same score as the best performing model (32.9 RMSE) (\autoref{tab:perf-top-10}).
+For HR-NRI and HR RMSE was around 40 p.p., and for task NRI the ridge learner achieved almost the same score as the best performing model (32.9 p.p) (\autoref{tab:perf-top-10}).
 In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
+The influential observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
 This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 SVM showed no variation in performance across filters (\autoref{fig:filter-effects-no-filter}).
-Using filters for RF showed a substantial increase in performance for all tasks besides VI for which the overall difference of all filters was also smallest (\autoref{fig:filter-effects-no-filter}).
+Using filters for RF showed a substantial increase in performance for all tasks with the exception of VI, for which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
 XGBoost showed a high dependency on filtering the data: In 4 out of 6 tasks using no filter would results in the worst or second worst performance.

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: clarify SVM filter vs no filter performance

@@
 However, in contrast for dataset NRI using no filter results in the best performance.
-XGBoost shows the highest overall differences between filters for a single task: For dataset HR, the range is up to 14 percentage points (Carscore vs.
-no filter)(\autoref{fig:filter-effects-no-filter}).
+XGBoost shows the highest overall differences between filters for a single task:
+For feature set HR, the range is up to 14 percentage points (Carscore vs. no filter)(\autoref{fig:filter-effects-no-filter}).
 
@@
 When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) when the model without filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
-For SVM filters and no filter come in at the same performance for tasks VI and NRI even though \autoref{fig:filter-effects-no-filter} lists "No Filter" as the best option.
+For SVM, all filters and 'no filter' achieved the same performance on tasks VI and NRI even though \autoref{fig:fig:perf-result} lists "No Filter" as the best option.
 

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: wording

@@
 When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) when the model without filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
-For SVM, all filters and 'no filter' achieved the same performance on tasks VI and NRI even though \autoref{fig:fig:perf-result} lists "No Filter" as the best option.
+For SVM, all filters and 'no filter' achieved the same performance on tasks VI and NRI even though \autoref{fig:perf-result} lists "No Filter" as the best option.
 
@@
 
-The most important features of each dataset showed an average decrease in RMSE of 1.57 (HR, B69) and 1.79 (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+The most important features of each dataset showed an average decrease in RMSE of 1.57 p.p. (HR, B69) and 1.79 p.p (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
 For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm (the so called "red-edge").

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: Figure 3 caption

@@
 		\includegraphics[width=0.48\textwidth] {performance-results-1.pdf}
-		\caption{Scored predictive performance (RMSE) of models across tasks.
-Prefix 'CV' denotes that the learner was optimized using internal 10-fold CV while prefix 'MBO' means that Bayesian optimization was used for hyperparameter optimization.
-The abbreviations on the y-axis refer to the combinations of feature sets on which each model was scored on.
-Labels attached to each point in space show which filter method was used for scoring features during the feature selection process (NF = no filter, Car = 'Carscore', Info = 'Information Gain', Borda = 'Borda').
-The second value in the attached label of each point shows the scored RMSE value of the respective setting.}\label{fig:perf-result}
+		\caption{Predictive performance (RMSE) of models across tasks.
+Suffix 'CV' denotes that the learner was optimized using internal 10-fold CV while prefix 'MBO' means that Bayesian optimization was used for hyperparameter optimization.
+Abbreviations on the vertical axis refer to the combinations of feature sets on which each model was scored on.
+Labels represent the feature selection method (NF = no filter, Car = 'Carscore', Info = 'Information Gain', Borda = 'Borda').
+The second value of each label shows the RMSE value of the respective setting.}\label{fig:perf-result}
 	\end{center}

### 2020-04-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: Caption Figure 4

@@
 		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-borda-filter-1.pdf}
-		\caption{Model performances in RMSE when using the Borda filter method compared to all other filters for each learner across all tasks.}\label{fig:filter-effects-borda}
+		\caption{Predictive performances in RMSE when using the Borda filter method compared to all other filters for each learner across all tasks.}\label{fig:filter-effects-borda}
 	\end{center}

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
AB-review: disc: rephrase
AB review: wording

@@
 Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
-In our view using a buffer of 2 m is a good compromise between both worlds.
+In our view using a buffer of 2 m is a good compromise between including information of too many surrounding trees and not accounting for the geometric offset.
 In addition, a side-analysis showed that some trees would be assigned missing values when relying on a single pixel.
@@
 
-Another critical point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
+Another relevant point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
 According to the extract function of the \texttt{raster} package, a pixel is included if the buffer of the point observation hits the centroid of the respective pixel, not just any part of it.

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: rephrase disc: performance vs. plot charateristics

@@
 % permutation based var imp for datasets HR and VI
-\begin{figure} [t!]
+\begin{figure*} [t!]
 	\centering
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {fi-permut-vi-hr-1.pdf}
+		\includegraphics[width=0.98\textwidth] {fi-permut-vi-hr-1.pdf}
 		\caption{Variable importance for feature sets HR and VI: Mean decrease in RMSE for one-hundred feature permutations using the SVM learner.
@@
 	\end{center}
-\end{figure}
+\end{figure*}
 
@@
 Another relevant point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-According to the extract function of the \texttt{raster} package, a pixel is included if the buffer of the point observation hits the centroid of the respective pixel, not just any part of it.
+According to the extract function of the \texttt{raster} package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
 As the buffer is circular, determining the total number of contributing pixels for each tree depends on the exact location of a tree within the pixel grid.
-Next, if a tree observation is located at the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using less pixels than if the tree observation is located in the middle of the plot.
+Next, if a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using fewer pixels than if the tree observation is located in the middle of the plot.
 The magnitude of uncertainty introduced by these facts cannot be quantified.
@@
 
-\noindent The high differences in RMSE observed for the fold performances (\autoref{tab:svm-single-fold-perf}) are a result of model overfitting.
-An RMSE of 54.26 indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
+The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
+An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
 Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
-The model overfitted on learning relationships related to medium-high defoliation values but missed out on low defoliation values.
-In cases when Laukiz2 is in the training set, the model is able to half the error and even reduce it down to a value of 9 RMSE in the case of Luiando.
+In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
+This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
+In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: error measure wording

@@
 \subsection{Predictive Performance}
-The best aggregated performance of this study (SVM + Carscore filter, 27.98 RMSE) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+The best aggregated performance of this study (SVM + Carscore filter, RMSE 27.98 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
 Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be 19.

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: error measure adjustments
A.B. review: rephrase

@@
 The best aggregated performance of this study (SVM + Carscore filter, RMSE 27.98 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
-Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be 19.
+Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be 19 p.p..
 However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
@@
 A model can only make robust predictions if it has learned relationships across the whole range of the response.
-Hence, care should be taken when making statements about the overall performance of the fitted models of this study due to their high variance when predicting.
+Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by a suboptimal distributed training data.
 However, when inspecting the fold level performances, one can say that the model does a decent job predicting defoliation ranging from 50\% to 100\% but fails for 0\% - 50\%.

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: calm down wording for svm win

@@
 
-FIXME: Add refs
-There were few environmental modeling studies only in recent years in which SVM outperformed other models that clearly.
-The absolute difference of around three percentage points is also worth noting:
-We could not find a study in which SVM came out as the best model in such clear way.
+FIXME: Add refs when SVM was clearly worse then other algs.
+In other (environmental) studies in recent years, SVM showed most often less good performances and was not able to outperform competing algorithms such as RF.
+The difference to the second best performing model of around three percentage points is also worth noting:
+We are not aware of a study in which SVM ranked as the best model that clearly.
 All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: rephrase disc: algor diffs

@@
 We were a bit surprised about the performances of the penalized methods lasso and ridge.
-The former was not able to achieve an error below 45.96 RMSE across any dataset while ridge came at least somewhat close to RF and XGBoost for some feature sets.
-However, there were also three instances (VI, NRI-VI, HR-NRI-VI) for which the error of the fitted ridge model was out of bounds.
-We cannot preclude possible mistakes on our side when optimizing the penalization terms of the penalized learners.
+The former was not able to achieve an RMSE below 46.0 p.p. across any feature set while ridge came at least somewhat close to RF and XGBoost for some feature sets.
+However, there were also three settings (VI, NRI-VI, HR-NRI-VI) for which the RMSE of the ridge model was out of bounds.
+That said, we cannot preclude possible mistakes on our side when optimizing the penalization terms of the penalized learners.
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: optimize wording

@@
 The former was not able to achieve an RMSE below 46.0 p.p. across any feature set while ridge came at least somewhat close to RF and XGBoost for some feature sets.
-However, there were also three settings (VI, NRI-VI, HR-NRI-VI) for which the RMSE of the ridge model was out of bounds.
-That said, we cannot preclude possible mistakes on our side when optimizing the penalization terms of the penalized learners.
+However, there were also three settings (VI, NRI-VI, HR-NRI-VI) for which the RMSE of the ridge model was unreasonably large.
+Careful inspection of the hyperparameter optimization of these learners did not reveal any implementation mistakes.
+However, one observation turned out to be influential even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
+We therefore left it in the model, but we acknowledge that these learners seemed particularly sensitive to the input data.
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: rephrase feature set differences

@@
 One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
-With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial changes in performance.
+With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial and consistent changes in performance.
 RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
-We have no explanation for this specific result, especially because the SVM does not show this difference.
+We have no explanation for this specific result, especially because the best performing SVM model does not reflect this difference.
+In addition, all learners showed a slightly better average performance on feature sets which included NRIs (e.g. NRI-VI, HR-NRI, HR-NRI-VI) than on those which did not (i.e. VI and HR only).
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: rephrase disc FS

@@
 % - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
-The usefulness of filters in this study was apparent across all feature sets, even when $p$ was rather small (HR).
+The usefulness of filters in this study was apparent across all feature sets, even for the smallest feature set, HR.
 Even though the effect of feature selection methods varied across algorithms, small to substantial improvement in predictive performance were achieved in almost all cases.
@@
 We have no explanation why the Borda ensemble method used in this study did not score better than the average performance of other filter methods.
-Ensemble methods shine in complicated scenarios when simple filter methods face problems due to their focus on certain areas of features.
-We assume that the datasets used in this study were not complex enough to bring up these problems for simple features.
+FIXME: Add ref when and why Borda filters are good.
+Ensemble methods have been proven to outperform simple filters in complicated scenarios when the latter face problems due to their focus on certain feature characteristics.
+We suggest that the datasets used in this study were not complex enough to bring up these problems for simple features.
 Therefore it makes sense that, due to its averaging nature, the Borda filter scored average results across all instances.

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: clarify runtime adv. of PCA vs filters

@@
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
-The PCA approach did neither show the best or worst result in the lineup.
-It shines in terms of runtime since it reduces the amount of features highly.
-However, it also removes the option for a post-analysis of feature importance by squashing predictors into principal components and users need to decided upfront what they value more.
-Since filters are only calculated once due to caching, the runtime advantage might in fact be negligible in practice.
+The PCA approach did neither show the best or worst result when being compared to other feature-selection approaches.
+It is able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
+However, it also removes the possibility to conduct a post-analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
+Since filter scores only need to be calculated once in a benchmark setting, the runtime advantage might in fact be negligible in practice.
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: disc: wording

@@
 
-FIXME: Add ref
-This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and its opposite.
+FIXME: Add ref.
+This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and their respective opposites.
 It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, we can observe quite some variance in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
+Looking at the spectral curves of the plots, we can observe quite some dynamic in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
 
@@
 Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
-\cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information out of ground-level RGB photos using \ac{CNN}.
+\cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
 Both study designs are substantially different compared to the setup of this work.
@@
 \cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees for eucalyptus forests.
-A grid-search was used for hyperparameter tuning and \ac{FFS} for variable selection.
-\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a Random Forest classifier.
+A grid search was used for hyperparameter tuning and \ac{FFS} for variable selection.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a random forest classifier.
 
 In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
-Most studies used only one algorithm (usually Random Forest) without (strong) arguments why this particular one has been selected or why only one model was used.
+Most studies used only one algorithm without (strong) arguments why this particular one has been selected or why only one model was used.
 This is not surprising: Most environmental/ecological datasets are not high-dimensional.
-In contrast, the number of predictors if often monadic and issues related to correlations are solved manually instead of relying on an automated approach.
+In contrast, the number of predictors is often less than ten and issues related to correlations are solved manually instead of relying on an automated approach.
 The bioinformatics field faces high-dimensional feature sets more often.
@@
 If a field only rarely faces high-dimensional dataset issues the motivation and expertise of using advanced methods to solve such are rather low.
-We hope that this work can give some guidance and serve as a starting point for tackling high-dimensional problems in environmental modeling.
+We hope that this work can give some guidance and serve as a starting point for tackling high-dimensional problems in environmental remote sensing.
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: conclusion

@@
 \section{Outlook and conclusion}
-This study analyzed defoliation at trees in north Spain by using hyperspectral data as input for machine learning models making heavy use of filter-based feature selection methods.
+
+This study analyzed defoliation at trees in north Spain by using hyperspectral data as input for machine learning models making heavy use of filter-based feature-selection methods.
 It was shown that substantial differences in performance can occur depending on which feature selection methods and machine learning algorithms are combined.
@@
 % RQ3: How are feature-selection methods influencing the predictive performance of the models?
-Filter methods showed their ability to improve predictive performance for datasets with many features.
-Ensemble filter methods did not show a substantial improvement over legacy filter methods.
+Filter methods are able to improve predictive performance for datasets with many features in certain instances.
+Their effectiveness depends on the chosen algorithm and dataset characteristics.
+Ensemble filter methods did not show a substantial improvement over individual filter methods.
 
@@
 % RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
-FIXME: "derivated" is not a real word?
-Creating feature sets composed out of derivated features did not help.
+The creation of derived feature sets did not contribute to an improvement in model performance.
 In contrast, dataset VI showed a substantial worse performance than the dataset with the original features (HR).

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: Conclusion wording

@@
 The creation of derived feature sets did not contribute to an improvement in model performance.
-In contrast, dataset VI showed a substantial worse performance than the dataset with the original features (HR).
+In contrast, dataset VI showed a substantially poorer performance than the dataset with the original features (HR).
 Combining feature sets did not show a substantial improvement on predictive performance.
@@
 % What about the general potential of analyzing defol with hyperspectral data?
-The potential of predicting defoliation for single trees with the given toolset of this study was rather limited, seen on the average error of 27\%. 
-However, with better training data covering more variety of the response variable, performance could be highly improved in future studies.
+The potential of predicting defoliation for single trees as conducted in this study was rather limited with an average RMSE of 27 p.p..
+We suggest that training data covering a wider range of defoliation values in a larger number of forest plantations would further improve the proposed models and result in more generalizable monitoring algorithms.
 

### 2020-04-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
A.B. review: wording for appendix figures

@@
 		\includegraphics[width=0.48\textwidth] {correlation-filter-nri-1.pdf}
-		\caption{Spearman correlation of filter rankings between various filter methods. Results of the NRI feature set are shown.}\label{fig:correlation-filters}
+		\caption{Spearman correlations of NRI feature rankings obtained with different filters.}\label{fig:correlation-filters}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
-		\caption{Spearman correlation of filter information gain using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.}\label{fig:correlation-nbins}
+		\caption{Spearman correlations of rankings obtained with the information gain filter using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.}\label{fig:correlation-nbins}
 	\end{center}

### 2020-04-08 docs/00-manuscripts/ieee/ieee-manuscript.tex
update title

@@
 % Do not put math or special symbols in the title.
-\title{Evaluating the Supportive Effect of Feature Selection Methods on Machine Learning Models in Monitoring Forest Health Using Hyperspectral Remote Sensing Data}
-\IEEEpeerreviewmaketitle{Evaluating the Supportive Effect of Feature Selection Methods on Machine Learning Models in Monitoring Forest Health Using Hyperspectral Remote Sensing Data}
+\title{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
+\IEEEpeerreviewmaketitle{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
 %

### 2020-04-12 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 # load drake objects
-loadd(
-  bm_aggregated_new_buffer2
-)
+# loadd(
+#   bm_aggregated_new_buffer2
+# )
+bm_aggregated_new_buffer2 = readRDS(here::here("temporary-bmr.rda"))
 library("xtable")
@@
     type = "latex",
-    caption = "Single fold performances for learner SVM on the HR dataset without using a filter.",
+    caption = "Test fold performances for learner SVM on the HR dataset without using a filter.
+For each row, the model was trained on observations from fold all other plots and tested on the observations of the given plot.",
     label = "tab:svm-single-fold-perf"
@@
 
-Overall leaderboard across all settings, sorted descending by performance.
+Overall leaderboard across all settings, sorted ascending by performance.
 
@@
 
-saveRDS(table1, here("docs/00-manuscripts/presentation/table-perf.rda"))
+saveRDS(table1, here("docs/00-manuscripts/presentation/table-perf-top-10.rda"))
 
@@
 
+## (Table) T3 All leaner/filter/task combinations ordered descending by performance.
+
+Overall leaderboard across all settings, sorted descending by performance.
+
+```{r eval-performance-3, warning=FALSE}
+table3 <- df_perf %>%
+  group_by(learner.id, task.id, filter) %>%
+  slice(which.min(rmse_aggr)) %>%
+  dplyr::rename(
+    "Model" = learner_group,
+    "Learner ID" = learner.id,
+    "Task" = task.id,
+    "Filter" = filter,
+    "RMSE" = rmse_aggr,
+    "SE" = se,
+  ) %>%
+  ungroup() %>%
+  mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
+  select(-iter, -`Learner ID`) %>%
+  arrange(RMSE)
+
+# save as latex table
+table3 %>%
+  ungroup() %>%
+  arrange(desc(RMSE)) %>%
+  slice(1:10) %>%
+  xtable(
+    type = "latex",
+    caption = "Worst 10 results for any task/learner/filter combination, sorted by performance.",
+    label = "tab:perf-worst-10"
+  ) %>%
+  print(
+    file = here("docs/00-manuscripts/ieee/performance-worst-10.tex"),
+    include.rownames = TRUE,
+    latex.environments = c("center"),
+    table.placement = "ht!",
+    caption.placement = "top",
+    timestamp = NULL
+  )
+
+saveRDS(table3, here("docs/00-manuscripts/presentation/table-perf-worst-10.rda"))
+
+table3 %>%
+  ungroup() %>%
+  arrange(desc(RMSE)) %>%
+  slice(1:10) %>% 
+  flextable() %>%
+  autofit()
+```
+
 ## (Plot) P1 Best learner/filter combs for all tasks
@@
     axis.title.y = element_blank(),
-    legend.text = element_text(size = 14),
+    legend.text = element_text(size = 13),
     legend.title = element_text(size = 14),

### 2020-04-18 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 # load drake objects
-# loadd(
-#   bm_aggregated_new_buffer2
-# )
-bm_aggregated_new_buffer2 = readRDS(here::here("temporary-bmr.rda"))
+loadd(
+  bm_aggregated_new_buffer2
+)
+
 library("xtable")
@@
   mutate(filter = replace(filter, is.na(filter), "NF")) %>%
-  mutate(learner_group = recode_factor(learner_group, `XG` = "XGBOOST")) %>%
+  mutate(learner_group = recode_factor(learner_group, `XG` = "XGBoost")) %>%
   group_by(learner_group, filter, task.id) %>%
@@
     legend.text = element_text(size = 13),
-    legend.title = element_text(size = 14),
+    legend.title = element_text(size = 13),
     axis.text.y = element_text(angle = 45),
@@
   # again later
-  mutate(learner_group = fct_relevel(learner_group, "XGBOOST", "SVM", "RF"))
+  mutate(learner_group = fct_relevel(learner_group, "XGBoost", "SVM", "RF"))
 

### 2020-04-18 docs/00-manuscripts/ieee/ieee-manuscript.tex
correct permutation execution

@@
 The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
-Fitting many models on the dataset containing the permuted feature comes with some computational cost.
+Predicting and calculating the error measure multiple times per feature comes with some computational cost.
 For increased efficiency, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.
@@
 In our view using a buffer of 2 m is a good compromise between including information of too many surrounding trees and not accounting for the geometric offset.
-In addition, a side-analysis showed that some trees would be assigned missing values when relying on a single pixel.
-This then would result in the removal of that particular tree which we wanted to avoid.
 Even though we provide no results showing the influence of multiple buffer values on the extraction, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.

### 2020-04-18 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   select(-iter, -`Learner ID`) %>%
+  mutate(RMSE = round(RMSE, 3)) %>% 
   arrange(RMSE)

### 2020-04-18 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
     sep = " MBO "
-  )
+  ) %>%
+  mutate(learner_group = recode_factor(learner_group,
+    `XGBOOST` = "XGBoost"
+  ))
 ```
@@
 results_aggr1 <- df_perf %>%
-  filter(learner_group != "Ridge-CV") %>%
-  filter(learner_group != "Lasso-CV") %>%
+  filter(learner_group != "Ridge-MBO") %>%
+  filter(learner_group != "Lasso-MBO") %>%
   # mutate(filter = replace(filter, "No Filter", "NF")) %>%
@@
   na.omit() %>%
-  filter(learner_group != "Ridge-CV") %>%
-  filter(learner_group != "Lasso-CV") %>%
+  filter(learner_group != "Ridge-MBO") %>%
+  filter(learner_group != "Lasso-MBO") %>%
   mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%

### 2020-04-18 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   ) %>%
-  mutate(wavelength = seq(420, 995, 4.75)) %>%
-  mutate(numeric_id = seq(5, 126, 1)) %>%
-  arrange(desc(importance)) %>%
-  mutate(rank = row_number()) %>%
-  select(-measure)
+  dplyr::mutate(wavelength = seq(420, 995, 4.75)) %>%
+  dplyr::mutate(numeric_id = seq(5, 126, 1)) %>%
+  dplyr::arrange(desc(importance)) %>%
+  dplyr::mutate(rank = row_number()) %>%
+  dplyr::select(-measure)
 fi_ranked_hr
@@
   # mutate(wavelength = seq(420, 995, 4.75)) %>%
-  arrange(desc(importance)) %>%
-  mutate(rank = row_number()) %>%
-  select(-measure) %>%
-  mutate(feature = stringr::str_replace(feature, "bf2_", ""))
+  dplyr::arrange(desc(importance)) %>%
+  dplyr::mutate(rank = row_number()) %>%
+  dplyr::select(-measure) %>%
+  dplyr::mutate(feature = stringr::str_replace(feature, "bf2_", ""))
 fi_ranked_vi
@@
   dplyr::slice(1) %>%
-  ungroup() %>%
-  arrange(rank) %>%
-  slice(1:10)
+  dplyr::ungroup() %>%
+  dplyr::arrange(rank) %>%
+  dplyr::slice(1:10)
 
@@
   dplyr::slice(1) %>%
-  ungroup() %>%
-  arrange(rank) %>%
-  slice(1:10) %>%
-  select(-class) %>%
-  rename(class = feature)
+  dplyr::ungroup() %>%
+  dplyr::arrange(rank) %>%
+  dplyr::slice(1:10) %>%
+  dplyr::select(-class) %>%
+  dplyr::rename(class = feature)
 
@@
   ncol = 2
+) +
+  theme_pubr(base_size = 14) +
+  theme(
+    legend.text = element_text(size = 13),
+    legend.title = element_text(size = 13)
   )

### 2020-04-18 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   echo = FALSE
+)
+
+options(
+scipen = 999
 )
@@
 ```{r, warning = FALSE}
-df_perf <- getBMRPerformances(bm_aggregated_new_buffer2, as.df = T) %>%
+df_perf <- getBMRPerformances(bm_aggregated_new_buffer2, as.df = TRUE) %>%
   mutate(task.id = recode_factor(task.id,
@@
   guides(size = FALSE) +
-  scale_x_continuous(limits = c(27, 47)) +
+  scale_x_continuous(limits = c(27, 40), breaks = seq(28, 40, 2)) +
   geom_label_repel(

### 2020-04-18 docs/00-manuscripts/ieee/ieee-manuscript.tex

@@
-\documentclass[letterpaper, peerreview]{IEEEtran}
+\documentclass[letterpaper, peerreview, draftcls]{IEEEtran}
 %
@@
 
-\author{Patrick~Schratz,~\IEEEmembership{Member,~IEEE,}
-	Jannes~Muenchow,~\IEEEmembership{Member,~IEEE,}
-	Eugenia Iturritxa,~\IEEEmembership{Member,~IEEE,}
-	José Cortés,~\IEEEmembership{Member,~IEEE,}
-	Bernd~Bischl,~\IEEEmembership{Member,~IEEE,}
-	and Alexander~Brenning,~\IEEEmembership{Member,~IEEE}
-	\thanks{P. Schratz, J. Muenchow, J. Cortés and A. Brenning are with the Department
+\author{Patrick~Schratz,
+	Jannes~Muenchow,
+	Eugenia~Iturritxa,
+	José~Cortés,
+	Bernd~Bischl,
+	and Alexander~Brenning
+	\thanks{P.Schratz, J.Muenchow, J.Cortés and A.Brenning are with the Department
 		of Geography, GIScience group, Friedrich-Schiller-University of Jena, Germany.}% <-this % stops a space
-	\thanks{B. Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
-	\thanks{E. Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Arab, Spain.}% <-this % stops a space
+	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
+	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Arab, Spain.}% <-this % stops a space
 	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}

### 2020-04-18 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   echo = FALSE
+)
+
+options(
+  scipen = 999
 )
@@
       data = data_hr_merged[which(!(data_hr_merged$rank %in% c(1:10))), ],
-      aes(color = "HR: 11:122"),
+      aes(color = "HR: Rank 11:122"),
       size = 1.8,
@@
       data = data_hr_merged[which(data_hr_merged$rank %in% c(1:10)), ],
-      aes(color = "HR: 1:10"),
+      aes(color = "HR: Rank 1:10"),
       shape = 20,
@@
     scale_color_manual(values = c(
-      "HR: 11:122" = "grey",
-      "HR: 1:10" = "black",
+      "HR: Rank 11:122" = "grey",
+      "HR: Rank 1:10" = "black",
       "Spectral Signature of Laukiz1" = "#BC3C29",
@@
       data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
-      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
+      aes(x = wavelength, y = importance, group = class, color = "VI: Rank 1:10"),
       size = 0.2, linetype = "dashed"
@@
       data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
-      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
+      aes(x = wavelength, y = importance, group = class, color = "VI: Rank 11:89"),
       size = 0.2, linetype = "dashed"
@@
     scale_color_manual(values = c(
-      "VI: 11:89" = "grey",
-      "VI: 1:10" = "black",
+      "VI: Rank 11:89" = "grey",
+      "VI: Rank 1:10" = "black",
       "Spectral Signature of Laukiz1" = "#BC3C29",

### 2020-04-18 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   ncol = 2
-) +
-  theme_pubr(base_size = 14) +
+) *
+  theme_pubr(base_size = 12) *
   theme(

### 2020-04-18 analysis/feature-importance.Rmd

@@
   ncol = 2
-) *
-  theme_pubr(base_size = 12) *
+) &
+  theme_pubr(base_size = 12) &
   theme(

### 2020-04-18 docs/00-manuscripts/ieee/performance-best-per-learner.tex

@@
   \hline
-1 & HR-NRI & SVM & Car & 27.98 & 19.19 \\ 
-  2 & HR-NRI & RF & Car & 31.08 & 17.09 \\ 
-  3 & NRI-VI & XGBOOST & Relief & 32.10 & 17.73 \\ 
-  4 & NRI & Ridge-CV & No Filter & 32.97 & 7.24 \\ 
-  5 & HR & Lasso-CV & No Filter & 45.96 & 18.35 \\ 
+1 & NRI & SVM & Info Gain & 27.99 & 19.15 \\ 
+  2 & NRI & RF & Car & 30.77 & 16.86 \\ 
+  3 & VI & Lasso-MBO & No Filter & 31.01 & 14.71 \\ 
+  4 & HR-NRI-VI & XGBoost & Borda & 31.05 & 17.01 \\ 
+  5 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
    \hline

### 2020-04-18 docs/00-manuscripts/ieee/performance-svm-single-plot.tex

@@
 \centering
-\caption{Single fold performances for learner SVM on the HR dataset without using a filter.} 
+\caption{Test fold performances for learner SVM on the HR dataset without using a filter.
+For each row, the model was trained on observations from fold all other plots and tested on the observations of the given plot.}
 \label{tab:svm-single-fold-perf}
@@
   \hline
-1 & Luiando & 9.00 \\ 
-  2 & Laukiz1 & 21.17 \\ 
-  3 & Laukiz2 & 54.26 \\ 
-  4 & Oiartzun & 28.05 \\ 
+1 & Luiando & 21.17 \\ 
+  2 & Laukiz1 & 28.05 \\ 
+  3 & Laukiz2 & 9.00 \\ 
+  4 & Oiartzun & 54.26 \\ 
    \hline

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
     legend.text = element_text(size = 13),
-    legend.title = element_text(size = 13)
+    legend.title = element_text(size = 13),
+    axis.title.y = element_text(size = 10),
+    axis.text.y = element_text(size = 10)
   )

### 2020-04-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
wording: rephrase side analysis

@@
 % guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
+% single column figure to Gbe placed prior to an earlier double column
 % figure.
@@
 To be able to use this method with a numeric response (defoliation in our case), the variable is discretized into equal bins and treated as a class variable.
-While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of $> 0.9$ for different bin sizes were observed in a side analysis.
+While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of $> 0.9$ for different bin sizes were observed during data exploration.
 
@@
 % permutation based var imp for datasets HR and VI
-\begin{figure*} [t!]
+\begin{figure*} [ht!]
 	\centering
 	\begin{center}
-		\includegraphics[width=0.98\textwidth] {fi-permut-vi-hr-1.pdf}
+		\includegraphics[width=0.9\textwidth] {fi-permut-vi-hr-1.pdf}
 		\caption{Variable importance for feature sets HR and VI: Mean decrease in RMSE for one-hundred feature permutations using the SVM learner.
@@
 
-\begin{figure} [ht]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
-		\caption{ALE plots on dataset VI of SVM learner.
+% \begin{figure} [ht]
+% 	\begin{center}
+% 		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
+% 		\caption{ALE plots on dataset VI of SVM learner.
 Subset showing the ten most important features according to the permutation-based variable importance.
 The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-vi-ale}
-	\end{center}
-\end{figure}
+% 	\end{center}
+% \end{figure}
 

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-18 analysis/feature-importance.Rmd
@@
 ) &
-  theme_pubr(base_size = 12) &
+  theme_pubr(base_size = 14) &
   theme(
     legend.text = element_text(size = 13),
-    legend.title = element_text(size = 13)
+    legend.title = element_text(size = 13),
+    axis.title.y = element_text(size = 8),
+    axis.text.y = element_text(size = 9)
   )

### 2020-04-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
upd caption of ALE plots

@@
 According to the extract function of the \texttt{raster} package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
-As the buffer is circular, determining the total number of contributing pixels for each tree depends on the exact location of a tree within the pixel grid.
-Next, if a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated using fewer pixels than if the tree observation is located in the middle of the plot.
-The magnitude of uncertainty introduced by these facts cannot be quantified.
-However, we assume these to be of minor importance for the subsequent model fitting step.
+As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of that tree within the pixel grid.
+Next, if a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation would be located in the middle of the plot.
+
+The hsdar R package was used for the calculation of vegetation indices \cite{hsdar}.
+All indices that could be calculated with the given spectral range of the data were used.
+Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on this preselection.
+
+The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
+Such limitations and uncertainties apply to most applied studies and cannot be completely avoided.
 
@@
 		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
-		\caption{ALE plots on dataset HR of SVM learner.
-Subset showing the ten most important features according to the permutation-based variable importance.
+		\caption{ALE plots of SVM on dataset HR.
+The ten most important features from the permutation-based variable importance estimation were used.
 The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-hr-ale}

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 
-# Create a virtual spectral signature of vegetation using PROSAIL.
+# Create a virtual Spectral Signature (mean) of vegetation using PROSAIL.
 
-PROSAIL is a algorithm simulating spectral signatures of vegetation, see `?hsdar::PROSAIL`.
+PROSAIL is a algorithm simulating Spectral Signature (mean)s of vegetation, see `?hsdar::PROSAIL`.
 Reflectance is scaled to 0-10 to be able to plot it in the same plot as the feature importance rankings -> the axis limits for the y and z axis needs to match.
 
-PROSAIL returns a spectral signature from 400 nm to 2500 nm -> we take the values only and subset to 400 nm - 1000 nm.
+PROSAIL returns a Spectral Signature (mean) from 400 nm to 2500 nm -> we take the values only and subset to 400 nm - 1000 nm.
 Because we order from 1 - 10 with 1 being the best rank, we have to reverse the scaling of the reflectance values.
@@
         x = wavelength, y = laukiz1,
-        color = "Spectral Signature of Laukiz1"
+        color = "Spectral Signature (mean) of Laukiz1"
       ),
@@
         x = wavelength, y = laukiz2,
-        color = "Spectral Signature of Laukiz2"
+        color = "Spectral Signature (mean) of Laukiz2"
       ),
@@
         x = wavelength, y = luiando,
-        color = "Spectral Signature of Luiando"
+        color = "Spectral Signature (mean) of Luiando"
       ),
@@
         x = wavelength, y = oiartzun,
-        color = "Spectral Signature of Oiartzun"
+        color = "Spectral Signature (mean) of Oiartzun"
       ),
@@
       shape = 20,
-      size = 1.8,
+      size = 2.2,
       na.rm = TRUE
@@
       "HR: Rank 1:10" = "black",
-      "Spectral Signature of Laukiz1" = "#BC3C29",
-      "Spectral Signature of Laukiz2" = "#0072B5",
-      "Spectral Signature of Luiando" = "#E18727",
-      "Spectral Signature of Oiartzun" = "#20854E"
+      "Spectral Signature (mean) of Laukiz1" = "#BC3C29",
+      "Spectral Signature (mean) of Laukiz2" = "#0072B5",
+      "Spectral Signature (mean) of Luiando" = "#E18727",
+      "Spectral Signature (mean) of Oiartzun" = "#20854E"
     )) +
@@
       caption = "Learner: SVM; 100 Monte-Carlo Iterations",
-      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
+      y = "Mean Decrease in RMSE", x = "Wavelength [nm]"
     ) +
@@
       aes(x = wavelength, y = importance, group = class),
-      size = 1.4, shape = 20
+      size = 2.2, shape = 20
     ) +
@@
       "VI: Rank 1:10" = "black",
-      "Spectral Signature of Laukiz1" = "#BC3C29",
-      "Spectral Signature of Laukiz2" = "#0072B5",
-      "Spectral Signature of Luiando" = "#E18727",
-      "Spectral Signature of Oiartzun" = "#20854E"
+      "Spectral Signature (mean) of Laukiz1" = "#BC3C29",
+      "Spectral Signature (mean) of Laukiz2" = "#0072B5",
+      "Spectral Signature (mean) of Luiando" = "#E18727",
+      "Spectral Signature (mean) of Oiartzun" = "#20854E"
     )) +

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   ncol = 2
+) & scale_y_continuous(
+  breaks = c(0.0002, 0, -0.002),
+  labels = c(0.0002, 0, -0.002)
 ) &

### 2020-04-19 docs/00-manuscripts/ieee/ieee-manuscript.tex

@@
 Features along the red-edge wavelength region were most important for models to achieve good predictions.
-With respect to dedicated vegetation indices, the "Vogelmann" index with all of its versions was seen as the most important index for the tested SVM algorithm.
+With respect to dedicated vegetation indices, the "Vogelmann" index with all of its derivatives was seen as the most important index for the best performing SVM model.
+This matches well with the actual purpose of these indices: 
+These were invented to detect defoliation on sugar marple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
 

### 2020-04-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
Disc: restructure and reformat sections, rename "derivation of indices" to a more general term

@@
 
-\subsection{Derivation of Indices}
-
-\noindent The decision to use a buffer of 2 m to generate the index value for each observation has to be seen critical.
-Due to the reported geometric offset of up to 1 m for the hyperspectral data, the possibility of assigning a wrong pixel value to a tree observation would be high.
-Using a buffer of more than 2 meters increases the probability of including information from other trees into the pixel value, blurring the actual value of the tree observation.
-In our view using a buffer of 2 m is a good compromise between including information of too many surrounding trees and not accounting for the geometric offset.
-Even though we provide no results showing the influence of multiple buffer values on the extraction, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
-
-Another relevant point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-According to the extract function of the \texttt{raster} package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
-As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of that tree within the pixel grid.
-Next, if a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation would be located in the middle of the plot.
-
-The hsdar R package was used for the calculation of vegetation indices \cite{hsdar}.
-All indices that could be calculated with the given spectral range of the data were used.
-Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on this preselection.
-
-The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
-Such limitations and uncertainties apply to most applied studies and cannot be completely avoided.
-
-\subsection{Performance vs.\ plot characteristics}
-\label{subsec:perf-plot-char}
-
-The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
-An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
-Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
-In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
-This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
-In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
-
 \subsection{Predictive Performance}
@@
 
-\subsubsection{Algorithm differences}
+\subsubsection{Model differences}
 An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
@@
 
-FIXME: Talk about L1 and L2 models -> there we have quite some differences!
+\subsection{Performance vs.\ plot characteristics}
+\label{subsec:perf-plot-char}
+
+The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
+An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
+Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
+In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
+This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
+In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
 
@@
 
-\subsubsection{Linking feature importance to spectral characteristics}
+\subsection{Quality of the data}
+
+\noindent The decision to use a buffer of 2 m for the extraction of an observation's reflectance value was a complex decision.
+Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
+By using a buffer of 2 meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
+However, the difference between single pixel values might often be higher than the difference between a smoothed value composed out of a buffer.
+In our view using a buffer of 2 m was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
+Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
+
+Another relevant point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
+According to the extract function of the \texttt{raster} package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
+As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of that tree within the pixel grid.
+Next, if a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation would be located in the middle of the plot.
+
+The hsdar R package was used for the calculation of vegetation indices \cite{hsdar}.
+All indices that could be calculated with the given spectral range of the data were used.
+Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on this preselection.
+
+The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
+Such limitations and uncertainties apply to most applied studies and cannot be completely avoided.
+
+FIXME: Talk about L1 and L2 models -> there we have quite some differences!
+
+\subsection{Linking feature importance to spectral characteristics}
 
@@
 This matches well with the actual purpose of these indices: 
-These were invented to detect defoliation on sugar marple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
+These were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
 

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-19 analysis/feature-importance.Rmd
@@
   ncol = 2
+) & scale_y_continuous(
+  n.breaks = 3
 ) &
+  labs(y = "ALE") &
   theme_pubr(base_size = 14) &

### 2020-04-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
talk about the bad performance of the VI feature set

@@
 
-Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 700 nm to 750 nm.
-
-FIXME: Add ref.
-This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and their respective opposites.
+Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 680 nm to 750 nm.
+
+This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and their respective opposites \cite{horler1983}.
 It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
 Looking at the spectral curves of the plots, we can observe quite some dynamic in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
+
+A possible explanation for the less good performances of the models scored on the VI dataset compared to all other feature sets could be the lack of features in the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
+The majority of the VI covers the range 550 nm - 800 nm.
+Only one index (PWI) covers information in the range beyond 900 nm.
+The spectral signatures of the plots show some variation around 940 nm, especially plot Oiartzun, which could be of potential help for the models when distinguishing between high and low defoliation.
+In addition, four our out of ten of the most important features in the HR dataset (B113, B115, B124 and B126) are located in this range (930 nm - 1000 nm), confirming the importance of this spectral range.
 

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-19 analysis/feature-importance.Rmd
@@
   ncol = 2
+) & scale_y_continuous(
+  n.breaks = 3,
+  minor_breaks = NULL
 ) &
+  labs(y = "ALE") &
   theme_pubr(base_size = 14) &

### 2020-04-19 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-19 analysis/feature-importance.Rmd
@@
 ) & scale_y_continuous(
-  n.breaks = 3
+  n.breaks = 2
 ) &

### 2020-04-19 docs/00-manuscripts/ieee/ieee-manuscript.tex
disc: add section about hyperspectral RS studies

@@
 Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
-\cite{meng2018} used \ac{OLS} regression methods while\cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
+\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
 Both study designs are substantially different compared to the setup of this work.
@@
 Incorporating such metrics (both spectral and structural) could be a possible enhancement for future works.
+
+% write about hyperspectral remote sensing
+The field of (hyperspectral) remote sensing has a strong focus on using random forest for modeling in recent years \cite{belgiu2016}.
+In high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
+To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016}, PCA and adjusted feature selection \cite{rochac2016}.
+We could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
+Potential reasons could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
+Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for upcoming modeling studies using hyperspectral remote sensing data.
+
+% RS studies with more/other algs than RF
+Studies which include more than one model usually operate in a low-dimensional setting 
+\cite{xu2019}.
 

### 2020-04-20 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 ```{r fi-hr-ale, dev = c("png", "pdf")}
-fi_ale_hr_buffer2$plot(
+ale_p1 = fi_ale_hr_buffer2$plot(
   features = df_wavelengths_from_hr$class,
@@
   )
+
+# Custom y axis breaks for subplots
+ale_p1[[2]] = ale_p1[[2]] + scale_y_continuous(
+  n.breaks = 3
+)
+
+ale_p1[[3]] = ale_p1[[3]] + scale_y_continuous(
+  n.breaks = 4
+)
+
+ale_p1[[7]] = ale_p1[[7]] + scale_y_continuous(
+  n.breaks = 4
+)
+
+ale_p1[[8]] = ale_p1[[8]] + scale_y_continuous(
+  n.breaks = 4
+)
+
+ale_p1[[9]] = ale_p1[[9]] + scale_y_continuous(
+  n.breaks = 4
+)
+
+ale_p1[[10]] = ale_p1[[10]] + scale_y_continuous(
+  n.breaks = 4
+)
+
 ```
@@
 ```{r}
-fi_ale_hr_buffer2_gs20$plot(
+ale_p2 = fi_ale_hr_buffer2_gs20$plot(
   features = df_wavelengths_from_hr$class,
@@
   )
+
+# Custom y axis breaks for subplots
+ale_p2[[1]] = ale_p2[[1]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p2[[2]] = ale_p2[[2]] + scale_y_continuous(
+  n.breaks = 5
+)
+ale_p2[[3]] = ale_p2[[3]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p2[[4]] = ale_p2[[4]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p2[[6]] = ale_p2[[6]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p2[[7]] = ale_p2[[7]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p2[[10]] = ale_p2[[10]] + scale_y_continuous(
+  n.breaks = 3
+)
 ```
@@
 ```{r fi-vi-ale, dev = c("png", "pdf")}
-fi_ale_vi_buffer2$plot(
+ale_p3 = fi_ale_vi_buffer2$plot(
   features = paste0("bf2_", df_wavelengths_from_indices_label$class),
@@
   )
+
+# Custom y axis breaks for subplots
+ale_p3[[1]] = ale_p3[[1]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p3[[2]] = ale_p3[[2]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p3[[3]] = ale_p3[[3]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p3[[4]] = ale_p3[[4]] + scale_y_continuous(
+  n.breaks = 4
+)
 ```
@@
 ```{r}
-fi_ale_vi_buffer2_gs20$plot(
+ale_p4 = fi_ale_vi_buffer2_gs20$plot(
   features = paste0("bf2_", df_wavelengths_from_indices_label$class),
@@
   )
+
+# Custom y axis breaks for subplots
+ale_p4[[2]] = ale_p3[[2]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p4[[3]] = ale_p3[[3]] + scale_y_continuous(
+  n.breaks = 4
+)
+ale_p4[[4]] = ale_p3[[4]] + scale_y_continuous(
+  n.breaks = 4
+)
 ```

### 2020-04-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-18 analysis/eval-performance.Rmd
@@
 loadd(
-  bm_aggregated_new_buffer2
+  bm_aggregated_new_buffer2,
+
+  # for T4
+  benchmark_tune_results_hr_nri_vi
 )
@@
   dplyr::mutate(Plot = forcats::fct_recode(Plot,
-    Laukiz1 = "2", Laukiz2 = "3",
-    Luiando = "1", Oiartzun = "4"
+    Laukiz1 = "1", Laukiz2 = "4",
+    Luiando = "3", Oiartzun = "2"
   )) %>%
@@
 ```
+
+## (Table) T4 Number of features selected during tuning
+
+The model/task combinations which were selected relate to the best performance of the respective algorithm on the HR-NRI-VI task in the overall benchmark.
+
+```{r eval-performance-5}
+bmr_inspect_tune <- mergeBenchmarkResults(benchmark_tune_results_hr_nri_vi)
+
+df_tbl <- getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
+  dplyr::rename(Learner = learner.id) %>%
+  dplyr::rename(Plot = iter) %>%
+  dplyr::rename(RMSE = rmse.test.mean) %>%
+  dplyr::rename("Features (%)" = fw.perc) %>%
+  dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::mutate(Plot = forcats::fct_recode(Plot,
+    Oiartzun = "2", Luiando = "3",
+    Laukiz1 = "1", Laukiz2 = "4"
+  )) %>%
+  dplyr::mutate(Learner = forcats::fct_recode(Learner,
+    "RF Car" = "RF MBO Car", "XGBoost Borda" = "XGBOOST MBO Borda",
+    "SVM Relief" = "SVM MBO Relief"
+  )) %>%
+  dplyr::mutate(Learner = as.character(Learner)) %>%
+  dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::group_by(Learner) %>%
+  dplyr::select(Learner, Plot, "Features (%)", RMSE)
+
+rle.lengths <- rle(df_tbl[[1]])$lengths
+first <- !duplicated(df_tbl[[1]])
+df_tbl[[1]][!first] <- ""
+
+# define appearance of \multirow
+df_tbl[[1]][first] <-
+  paste0("\\midrule\\multirow{", rle.lengths, "}{*}{\\textbf{", df_tbl[[1]][first], "}}")
+df_tbl %>%
+  xtable::xtable(
+    type = "latex",
+    caption = "Percentage of selected features of each fold during optimization for selected learners",
+    label = "tab:tune-perc-sel-features"
+  ) %>%
+  print(
+    file = here("docs/00-manuscripts/ieee/tune-perc-sel-features.tex"),
+    include.rownames = TRUE,
+    latex.environments = c("center"),
+    table.placement = "ht!",
+    caption.placement = "top",
+    timestamp = NULL
+  )
+```
+
+Aggregated mean and standard deviation:
+
+```{r}
+getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
+  dplyr::rename(Learner = learner.id) %>%
+  dplyr::rename(Plot = iter) %>%
+  dplyr::rename(RMSE = rmse.test.mean) %>%
+  dplyr::rename("Features (%)" = fw.perc) %>%
+  dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::mutate(Plot = forcats::fct_recode(Plot,
+    Oiartzun = "2", Luiando = "3",
+    Laukiz1 = "1", Laukiz2 = "4"
+  )) %>%
+  dplyr::mutate(Learner = forcats::fct_recode(Learner,
+    "RF Car" = "RF MBO Car", "XGBoost Borda" = "XGBOOST MBO Borda",
+    "SVM Relief" = "SVM MBO Relief"
+  )) %>%
+  dplyr::mutate(Learner = as.character(Learner)) %>%
+  dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::group_by(Learner) %>%
+  dplyr::summarise(
+    "Mean (Features (%))" = mean(`Features (%)`),
+    "SD (Features (%))" = sd(`Features (%)`)
+  )
+```

### 2020-04-29 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
add table listing percentage of selected features per plot

% latex table generated in R 3.6.1 by xtable 1.8-3 package
\begin{table}[ht!]
\centering
\caption{Percentage of selected features of each fold during optimization for selected learners} 
\label{tab:tune-perc-sel-features}
\begin{tabular}{rllrr}
  \hline
 & Learner & Plot & Features (\%) & RMSE \\ 
  \hline
1 & $\backslash$midrule$\backslash$multirow\{4\}\{*\}\{$\backslash$textbf\{RF Car\}\} & Laukiz1 & 0.00 & 27.48 \\ 
  2 &  & Oiartzun & 0.00 & 37.20 \\ 
  3 &  & Luiando & 0.00 & 36.98 \\ 
  4 &  & Laukiz2 & 0.03 & 15.39 \\ 
  5 & $\backslash$midrule$\backslash$multirow\{4\}\{*\}\{$\backslash$textbf\{XGBoost Borda\}\} & Laukiz1 & 0.01 & 15.00 \\ 
  6 &  & Oiartzun & 0.68 & 36.15 \\ 
  7 &  & Luiando & 1.00 & 38.19 \\ 
  8 &  & Laukiz2 & 1.00 & 29.62 \\ 
  9 & $\backslash$midrule$\backslash$multirow\{4\}\{*\}\{$\backslash$textbf\{SVM Relief\}\} & Laukiz1 & 0.00 & 35.50 \\ 
  10 &  & Oiartzun & 0.71 & 37.48 \\ 
  11 &  & Luiando & 0.71 & 14.89 \\ 
  12 &  & Laukiz2 & 0.95 & 37.19 \\ 
   \hline
\end{tabular}
\end{table}

### 2020-04-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 library("forcats")
+```
+
+```{r, comment="", results="asis", echo=FALSE}
+old.hooks <- fansi::set_knit_hooks(knitr::knit_hooks)
 ```
@@
   )
+
+df_tbl
 ```

### 2020-04-30 analysis/feature-importance.Rmd

@@
 )
-
+ale_p1
 ```
@@
 )
+ale_p2
 ```
@@
 )
+ale_p3
 ```
@@
 )
+ale_p4
 ```

### 2020-04-30 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update perc-sel-table

@@
 \begin{tabular}{rllrr}
-  \hline
- & Learner & Plot & Features (\%) & RMSE \\ 
-  \hline
-1 & $\backslash$midrule$\backslash$multirow\{4\}\{*\}\{$\backslash$textbf\{RF Car\}\} & Laukiz1 & 0.00 & 27.48 \\ 
+  \toprule
+ & Learner & Plot & Features (%) & RMSE \\ 
+  \midrule
+1 & \midrule\multirow{4}{*}{\textbf{RF Car}} & Laukiz1 & 0.00 & 27.48 \\ 
   2 &  & Oiartzun & 0.00 & 37.20 \\ 
@@
   4 &  & Laukiz2 & 0.03 & 15.39 \\ 
-  5 & $\backslash$midrule$\backslash$multirow\{4\}\{*\}\{$\backslash$textbf\{XGBoost Borda\}\} & Laukiz1 & 0.01 & 15.00 \\ 
+  5 & \midrule\multirow{4}{*}{\textbf{XGBoost Borda}} & Laukiz1 & 0.01 & 15.00 \\ 
   6 &  & Oiartzun & 0.68 & 36.15 \\ 
@@
   8 &  & Laukiz2 & 1.00 & 29.62 \\ 
-  9 & $\backslash$midrule$\backslash$multirow\{4\}\{*\}\{$\backslash$textbf\{SVM Relief\}\} & Laukiz1 & 0.00 & 35.50 \\ 
+  9 & \midrule\multirow{4}{*}{\textbf{SVM Relief}} & Laukiz1 & 0.00 & 35.50 \\ 
   10 &  & Oiartzun & 0.71 & 37.48 \\ 
@@
   12 &  & Laukiz2 & 0.95 & 37.19 \\ 
-   \hline
+   \bottomrule
 \end{tabular}

### 2020-04-30 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-29 analysis/eval-performance.Rmd
@@
   dplyr::rename(RMSE = rmse.test.mean) %>%
-  dplyr::rename("Features (%)" = fw.perc) %>%
+  dplyr::rename("Features (\\%)" = fw.perc) %>%
   dplyr::mutate(Plot = as.character(Plot)) %>%
@@
   dplyr::group_by(Learner) %>%
-  dplyr::select(Learner, Plot, "Features (%)", RMSE)
+  dplyr::select(Learner, Plot, "Features (\\%)", RMSE)
 
@@
   paste0("\\midrule\\multirow{", rle.lengths, "}{*}{\\textbf{", df_tbl[[1]][first], "}}")
-df_tbl %>%
+table4 <- df_tbl %>%
   xtable::xtable(
@@
     caption = "Percentage of selected features of each fold during optimization for selected learners",
-    label = "tab:tune-perc-sel-features"
-  ) %>%
+    label = "tab:tune-perc-sel-features",
+    digits = c(0, 0, 0, 5, 2)
+  )
+
+table4 %>%
   print(
     file = here("docs/00-manuscripts/ieee/tune-perc-sel-features.tex"),
-    include.rownames = TRUE,
+    include.rownames = FALSE,
     latex.environments = c("center"),
@@
     caption.placement = "top",
-    timestamp = NULL
+    timestamp = NULL,
+    booktabs = TRUE,
+    # important to treat content of multirow as latex content
+    sanitize.text.function = force
   )
+
+saveRDS(table4, here("docs/00-manuscripts/presentation/tune-perc-sel-feature.rda"))
+
+table4 %>%
+  flextable() %>%
+  autofit()
 ```
@@
     "SD (Features (%))" = sd(`Features (%)`)
-  )
+  ) %>% 
+  flextable() %>% 
+  autofit()
 ```

### 2020-04-30 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update perc-sel-table

@@
 \label{tab:tune-perc-sel-features}
-\begin{tabular}{rllrr}
+\begin{tabular}{llrr}
   \toprule
- & Learner & Plot & Features (%) & RMSE \\ 
+Learner & Plot & Features (\%) & RMSE \\ 
   \midrule
-1 & \midrule\multirow{4}{*}{\textbf{RF Car}} & Laukiz1 & 0.00 & 27.48 \\ 
-  2 &  & Oiartzun & 0.00 & 37.20 \\ 
-  3 &  & Luiando & 0.00 & 36.98 \\ 
-  4 &  & Laukiz2 & 0.03 & 15.39 \\ 
-  5 & \midrule\multirow{4}{*}{\textbf{XGBoost Borda}} & Laukiz1 & 0.01 & 15.00 \\ 
-  6 &  & Oiartzun & 0.68 & 36.15 \\ 
-  7 &  & Luiando & 1.00 & 38.19 \\ 
-  8 &  & Laukiz2 & 1.00 & 29.62 \\ 
-  9 & \midrule\multirow{4}{*}{\textbf{SVM Relief}} & Laukiz1 & 0.00 & 35.50 \\ 
-  10 &  & Oiartzun & 0.71 & 37.48 \\ 
-  11 &  & Luiando & 0.71 & 14.89 \\ 
-  12 &  & Laukiz2 & 0.95 & 37.19 \\ 
+\midrule\multirow{4}{*}{\textbf{RF Car}} & Laukiz1 & 0.00443 & 27.48 \\ 
+   & Oiartzun & 0.00004 & 37.20 \\ 
+   & Luiando & 0.00002 & 36.98 \\ 
+   & Laukiz2 & 0.03037 & 15.39 \\ 
+  \midrule\multirow{4}{*}{\textbf{XGBoost Borda}} & Laukiz1 & 0.00604 & 15.00 \\ 
+   & Oiartzun & 0.68227 & 36.15 \\ 
+   & Luiando & 0.99513 & 38.19 \\ 
+   & Laukiz2 & 0.99976 & 29.62 \\ 
+  \midrule\multirow{4}{*}{\textbf{SVM Relief}} & Laukiz1 & 0.00017 & 35.50 \\ 
+   & Oiartzun & 0.71146 & 37.48 \\ 
+   & Luiando & 0.70672 & 14.89 \\ 
+   & Laukiz2 & 0.94696 & 37.19 \\ 
    \bottomrule

### 2020-04-30 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 
+# save to file
 table4 %>%
@@
 table4 %>%
-  flextable() %>%
-  autofit()
+  print(
+    include.rownames = FALSE,
+    latex.environments = c("center"),
+    table.placement = "ht!",
+    caption.placement = "top",
+    timestamp = NULL,
+    booktabs = TRUE,
+    # important to treat content of multirow as latex content
+    sanitize.text.function = force
+  )
 ```

### 2020-04-30 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 table4 %>%
-  print(
-    include.rownames = FALSE,
-    latex.environments = c("center"),
-    table.placement = "ht!",
-    caption.placement = "top",
-    timestamp = NULL,
-    booktabs = TRUE,
-    # important to treat content of multirow as latex content
-    sanitize.text.function = force
-  )
+  flextable() %>% 
+  autofit() %>% 
+  fontsize(size = 13)
 ```
@@
   flextable() %>% 
-  autofit()
+  autofit() %>% 
+  fontsize(size = 13)
 ```

### 2020-04-30 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
-  select(-iter, -`Learner ID`) %>%
+  select(-iter, -`Learner ID`, -SE) %>%
   mutate(RMSE = round(RMSE, 3)) %>%
@@
     type = "latex",
-    caption = "Worst 10 results for any task/learner/filter combination, sorted by performance.",
+    caption = "Worst 10 results for any task/learner/filter combination, sorted descreasing by RMSE",
     label = "tab:perf-worst-10"
@@
     type = "latex",
-    caption = "Percentage of selected features of each fold during optimization for selected learners",
+    caption = "Optimized number of features for selected learner-filter settings across folds.",
     label = "tab:tune-perc-sel-features",

### 2020-04-30 docs/00-manuscripts/ieee/ieee-manuscript.tex
upd results and discussion for lasso scores

@@
 \usepackage{booktabs}
+\usepackage{multirow}
 \usepackage{adjustbox}
@@
 SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
-Ridge and lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
+Ridge faced major issues in four tasks due to one observation that was predicted off the scale.
 SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
@@
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In two out of six tasks (all including VI variables) the error was out of bounds (HR-NRI-VI, HR-VI, VI).
-For HR-NRI and HR RMSE was around 40 p.p., and for task NRI the ridge learner achieved almost the same score as the best performing model (32.9 p.p) (\autoref{tab:perf-top-10}).
+Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was out of bounds (\autoref{tab:perf-worst-10}).
+For NRI and HR RMSE was 31.16 and 35.45 p.p., respectively.
 In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
-The influential observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
-This then caused the error estimates of these folds and the average estimate across all folds to be out of bounds.
-
-FIXME: Mention: All penalized methods tuned using MBO showed exactly the same error across all tasks.
-FIXME: Mention the bad perf of RIDGE on VI
+The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
+This one outlier caused the error estimates of these folds and the average estimate across all folds to be off the scale.
 
@@
 
+% Selected percentage of features across plots
+High differences were observed between the amount of features selected during tuning for the subsequent fitting process.
+Most features were selected during optimization for plot Laukiz2 and least for Laukiz1 (\autoref{tab:tune-perc-sel-features}).
+RF used only one feature for plots Luiando and Oiartzun (7675 $\times$ 0.00004) while for Laukiz1 34 and for Laukiz2 230 features were seen as optimal by the optimizer.
+In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features) more than two-third of all available features.
+
 % latex table top 10 absolute performances
@@
 % latex table worst 10 absolute performances
-%\input{performance-worst-10}
+\input{performance-worst-10}
 
@@
 \input{performance-svm-single-plot.tex}
+
+% latex table percentage of selected features during tuning
+\input{tune-perc-sel-features.tex}
 
@@
 
-We were a bit surprised about the performances of the penalized methods lasso and ridge.
-The former was not able to achieve an RMSE below 46.0 p.p. across any feature set while ridge came at least somewhat close to RF and XGBoost for some feature sets.
-However, there were also three settings (VI, NRI-VI, HR-NRI-VI) for which the RMSE of the ridge model was unreasonably large.
-Careful inspection of the hyperparameter optimization of these learners did not reveal any implementation mistakes.
-However, one observation turned out to be influential even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
+Penalized methods showed promising performances, especially when taking runtime into account.
+When removing features with a correlation of 1, lasso is able to score performances around 31 p.p. and shows somewhat surprisingly the best performance on the VI task.
+The same general conclusion applies to ridge if one discards the problematic results on four tasks.
+A careful inspection of the hyperparameter optimization procedure of these models did not reveal any implementation mistakes.
+One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
 We therefore left it in the model, but we acknowledge that these learners seemed particularly sensitive to the input data.
-
-FIXME: Talk about L1 and L2 model performance
 
@@
 In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
+
+The high differences of selected features per plot plot during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
+While SVM and XGBoost require a substantial portion of features in most cases to achieve robust predictions, RF is able to achieve the best results with a relatively low amount of features.
+This fact leads to reduced computational runtime, especially if model parameters are dependent on the amount of features (e.g. \texttt{\(m_{try}\)}).
+Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, at least for RF in this study.
+However, the results of (\autoref{tab:tune-perc-sel-features}) should be taken with care as they rely on single model-filter combinations and are subject to random variation.
+More in-depth research is needed to investigate the effect of filters on other criteria (such as runtime) than performance, leading to a multi-criteria optimization problem.
 
@@
 In high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
-To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016}, PCA and adjusted feature selection \cite{rochac2016}.
+To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
 We could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
@@
 % RS studies with more/other algs than RF	
-Studies which include more than one model usually operate in a low-dimensional setting 
-\cite{xu2019}.
+Remote sensing studies comparing multiple models often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
 

### 2020-04-30 docs/00-manuscripts/ieee/ieee-manuscript.tex
don't try to compare SVM performance to other studies

@@
 An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
-
-FIXME: Add refs when SVM was clearly worse then other algs.
-In other (environmental) studies in recent years, SVM showed most often less good performances and was not able to outperform competing algorithms such as RF.
-The difference to the second best performing model of around three percentage points is also worth noting:
-We are not aware of a study in which SVM ranked as the best model that clearly.
-All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).
+These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
+Comparing these differences (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
 
@@
 \label{subsec:perf-plot-char}
-
 The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).

### 2020-04-30 docs/00-manuscripts/ieee/ieee-manuscript.tex
rewrite disc about ensemble filter perf

@@
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
@@
 
-% - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
-We have no explanation why the Borda ensemble method used in this study did not score better than the average performance of other filter methods.
-FIXME: Add ref when and why Borda filters are good.
-Ensemble methods have been proven to outperform simple filters in complicated scenarios when the latter face problems due to their focus on certain feature characteristics.
-We suggest that the datasets used in this study were not complex enough to bring up these problems for simple features.
-Therefore it makes sense that, due to its averaging nature, the Borda filter scored average results across all instances.
-However, this cannot be known upfront and we believe that an ensemble filter should be part of every filter portfolio when doing filter-based feature selection.
+% - Ensemble FS shows no advantage to simple methods -> always one simple filter which is better. Ensemble filters are good across many different datasets
+The models which used the Borda ensemble method in this study did not score better on average than models which used a single filter or no filter at all.
+Ensemble methods have higher stability and robustness than single ones and have shown promising results in \cite{drotar2017a}.
+Hence, their main advantage are stable performances across datasets with varying characteristics.
+Certain single filter methods might yield better model performances for certain datasets but fail on others.
+The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBOOST on task HR-NRI-VI) a single filter outperformed the ensemble filter.
+However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
+Therefore no general statement can yet be made if the use of ensemble filters leads to a substantial better performance than applying one of the various single filter methods.
+More case studies applying ensemble filter methods are needed.
+We believe that ensemble filters can be a promising option in a machine-learning feature-selection portfolio.
 

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
mention bias in permutation FI for correlated features

@@
 
-% FIXME Talk about the bias of permutation FI in highly correlated settings
 \noindent Estimating feature importance for highly-correlated datasets is a complicated task.
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
-Methods like \ac{PDP} do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
-% FIXME: Add ref
-Due to the noisiness of the dataset, estimates from model-agnostic approaches such as permutation-based feature importance should also be taken with care.
-This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
+Methods like \ac{PDP} or permutation-based approaches do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
+Permutation feature importance calculates the loss of predictive performance for each feature by permuting it in a random manner.
 The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
@@
 The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
-Such limitations and uncertainties apply to most applied studies and cannot be completely avoided.
-
-FIXME: Talk about L1 and L2 models -> there we have quite some differences!
+Such limitations and uncertainties apply to many environmental studies and cannot be completely avoided.
 

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
rephrase relation to other studies

@@
 \cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
-Both study designs are substantially different compared to the setup of this work.
-In addition, no spatial \ac{CV} or \ac{FS} was used.
+However, both did not use spatial \ac{CV} or \ac{FS}.
 \cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
@@
 To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
-We could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
+However, we could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
 Potential reasons could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
@@
 
-% RS studies with more/other algs than RF	
+% RS studies with more/other algorithms than RF	
 Remote sensing studies comparing multiple models often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
 
-\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable (woody cover).
+\cite{shendryk2016, ludwig2019} are more similar in their methodology but focus on a different response variable (woody cover).
 \cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees for eucalyptus forests.
@@
 In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
-Most studies used only one algorithm without (strong) arguments why this particular one has been selected or why only one model was used.
-This is not surprising: Most environmental/ecological datasets are not high-dimensional.
-In contrast, the number of predictors is often less than ten and issues related to correlations are solved manually instead of relying on an automated approach.
-The bioinformatics field faces high-dimensional feature sets more often.
+Also, most environmental/ecological datasets are not high-dimensional.
+In fact, the number of predictors is often less than ten and issues related to correlations are solved manually instead of relying on an automated approach.
+The bioinformatics field faces high-dimensional datasets more often.
 Hence more studies using (filter-based) feature-selection approaches can be found for this field \cite{guo2019, radovic2017}.
-If a field only rarely faces high-dimensional dataset issues the motivation and expertise of using advanced methods to solve such are rather low.
+However bioinformatics differs conceptually in many ways from environmental modeling and we do not want to dig deeper just for the sake of finding any similar studies.
+If a field only rarely faces high-dimensional dataset issues (like environmental modeling) the motivation and expertise of using advanced methods to solve such are rather low.
 We hope that this work can give some guidance and serve as a starting point for tackling high-dimensional problems in environmental remote sensing.

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
mention explicitly that all feature sets depend on the 2 m buffered values

@@
 \noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used for extraction of the reflectance values.
+These averaged reflectance values were used as the base to derive all feature sets.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
@@
 	\item The raw hyperspectral band information (HR): No feature engineering) %chktex 13
-	\item Vegetation Indices (\ac{VI}): Expert-based feature engineering)
-	\item Normalized Ratio Indices (\ac{NRI}): Data-driven feature engineering)
+	\item Vegetation Indices (\ac{VIs}): Expert-based feature engineering)
+	\item Normalized Ratio Indices (\ac{NRIs}): Data-driven feature engineering)
 \end{itemize}
@@
 
-\noindent The decision to use a buffer of 2 m for the extraction of an observation's reflectance value was a complex decision.
+\noindent The decision to use a buffer of two meters for the extraction of an observation's reflectance value was a complex decision.
 Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
-By using a buffer of 2 meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
+By using a buffer of two meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
 However, the difference between single pixel values might often be higher than the difference between a smoothed value composed out of a buffer.
-In our view using a buffer of 2 m was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
+In our view using a buffer of two meters was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
 Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove SAC from intro, mention FS importance

@@
 
+% mention why FS is used / important
 Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
-At the same time, it is necessary to account for the presence of spatial autocorrelation in the data.
-The dataset used in this study comes with a spatial grouping at the plot level which needs to be accounted for \cite{sperrorest, pena2015}.
+Feature selection approaches can help to address this issue, reducing possible in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
 
@@
 
-\subsection{Dimension reduction}
-
-\noindent We want to stress the differences between 'dimension reduction' and 'high-dimensionality'.
-The former refers to the general idea of reducing features from a dataset \cite{vandermaaten2007}.
-In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
-'High-dimensionality' in contrast is a dataset attribute and applies when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of observations \cite{hastie2001}.
-There is no absolute value at which the term applies, both for \(n\) and \(p\).
-Hence for this study the word ‘high-dimensionality‘ only applies to the experiments using the ‘NRI‘ feature set ($>$ 7000 features for 1808 observations).
+\subsection{Feature selection}
 
@@
 
-\noindent In the following sections a brief overview of feature selection sub-categories is given.
-Due to the focus of this study on the use of filter methods, which are a sub-group of feature selection methods, other approaches were grouped into a single section.
+To reduce the dimensions of a dataset, conceptually different approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or approaches like \ac{PCA} \cite{das2001, jolliffe2016}.
+Since the main focus of this study is on investigating the effect of filter methods on predictive performance, only this sub-group will be introduced in more detail in the following.
 
@@
 % Filter methods
-\noindent The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
+The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
 Some filter methods are restricted towards specific types of variables (numeric or nominal).
@@
 The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
-
-\paragraph{Ensemble filter methods}
 
@@
 
-\paragraph{Ensuring a fair weighting in the ensemble}
-
+% Ensuring a fair weighting in the ensemble
 Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
@@
 
-\paragraph{Description of used filter methods}
+\subsubsection{Description of used filter methods}
 
@@
 
-\subsubsection{Wrapper methods and PCA}
-
-\noindent Other approaches to assess feature importance are so called 'wrapper methods' and the \ac{PCA}\cite{das2001, jolliffe2016}.
-
-% Wrapper approach
-\noindent Wrappers \cite{chandrashekar2014, kohavi1997} apply search algorithms such as 'random search' or 'generic simulated annealing'.
-First a (random) subset of features is chosen based on the selected algorithm.
-In comparison to filters, no ranking is done in this step.
-Now the model is fitted on the data and the performance is evaluated.
-This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
-A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature selection optimization finished.
-Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.
-Due to their high computational cost, wrappers were not considered in this work.
-
-% PCA
-A method with a completely different approach compared to filters and wrappers is the \ac{PCA}\cite{pearson1901, jolliffe2016}.
-Here, the main components of the feature space are extracted and combined.
-The number of principal components needed to explain a large portion of the variance is dynamic and depends on the dataset.
-In practice, this number often ranges between two and five.
-By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
-This enables cheap model fitting with balanced loss of predictor information.
-A disadvantage is the reduced interpretability due to the main components being a squashed version of the original covariates.
-
 \subsection{Benchmarking design}

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
harmonize xgboost names

@@
 	\acro{VI}{vegetation index}
-	\acro{XGBOOST}{extreme gradient boosting}
+	\acro{XGBoost}{extreme gradient boosting}
 \end{acronym}
@@
 \begin{itemize}
-	\item  Extreme Gradient Boosting (XGBOOST)
+	\item  Extreme Gradient Boosting (XGBoost)
 	\item  Random Forest (RF)
@@
 \noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
-\ac{XGBOOST} (commonly abbreviated as XGBOOST) showed promising results in benchmark competitions in recent years.
+\ac{XGBoost} (commonly abbreviated as XGBoost) showed promising results in benchmark competitions in recent years.
 Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
@@
 Certain single filter methods might yield better model performances for certain datasets but fail on others.
-The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBOOST on task HR-NRI-VI) a single filter outperformed the ensemble filter.
+The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBoost on task HR-NRI-VI) a single filter outperformed the ensemble filter.
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
rephrase disc: feature set differences

@@
 One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
-With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial and consistent changes in performance.
-RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
-We have no explanation for this specific result, especially because the best performing SVM model does not reflect this difference.
-In addition, all learners showed a slightly better average performance on feature sets which included NRIs (e.g. NRI-VI, HR-NRI, HR-NRI-VI) than on those which did not (i.e. VI and HR only).
+With respect to \autoref{fig:perf-result}, no overall trend was found for all models that related to specific feature sets.
+The performance of RF and XGBoost on the VI feature set was about six percentage points lower than on others.
+One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
+In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e. NRI-VI, HR-NRI, HR-NRI-VI).
 

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
rewrite conclusion

@@
 \section{Outlook and conclusion}
-
-This study analyzed defoliation at trees in north Spain by using hyperspectral data as input for machine learning models making heavy use of filter-based feature-selection methods.
-It was shown that substantial differences in performance can occur depending on which feature selection methods and machine learning algorithms are combined.
+This study analyzed defoliation at trees in northern Spain by using hyperspectral data as input for machine-learning models which incorporated hyperparameter tuning and filter-based feature selection.
+Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
 SVM showed the most robust behavior across all highly-correlated datasets and was able to predict the response variable of this study substantially better than other methods.
@@
 % RQ3: How are feature-selection methods influencing the predictive performance of the models?
-Filter methods are able to improve predictive performance for datasets with many features in certain instances.
-Their effectiveness depends on the chosen algorithm and dataset characteristics.
-Ensemble filter methods did not show a substantial improvement over individual filter methods.
+Filter methods were able to improve the predictive performance on datasets in some instances.
+Their effectiveness depends on the algorithm and the dataset characteristics.
+Ensemble filter methods did not show a substantial improvement over individual filter methods in this study.
 
@@
 % RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
-The creation of derived feature sets did not contribute to an improvement in model performance.
-In contrast, dataset VI showed a substantially poorer performance than the dataset with the original features (HR).
-Combining feature sets did not show a substantial improvement on predictive performance.
+The creation of derived feature sets was able to improve predictive performance in most cases.
+In contrast, datasets which covered only a fraction of the available spectral range (i.e. dataset VI) showed a worse performance than the ones which covered the full range (400 nm - 1000 nm; HR, NRI).
+Combining feature sets helped most models in improving their predictive performance.
 
@@
 These were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
+However, feature importance for highly-correlated features is challenging.
+Results might be biased and should be taken with care.
 
 % What about the general potential of analyzing defol with hyperspectral data?
-The potential of predicting defoliation for single trees as conducted in this study was rather limited with an average RMSE of 27 p.p..
-We suggest that training data covering a wider range of defoliation values in a larger number of forest plantations would further improve the proposed models and result in more generalizable monitoring algorithms.
+The potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 27 percentage points scored by the best performing model.
+More training data covering a wider range of defoliation values in a larger number of forest plantations would probably help to build better models that can achieve more robust predictions.
 

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
rewrite disc: effect of filters

@@
 % - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
-The usefulness of filters in this study was apparent across all feature sets, even for the smallest feature set, HR.
-Even though the effect of feature selection methods varied across algorithms, small to substantial improvement in predictive performance were achieved in almost all cases.
-No optimal algorithm/filter/dataset setting was observed in this study, leading to the conclusion that one needs to try out various combinations of such to find the best performing one.
+% - Filters can also have a negative impact
+% - Filters are interesting for reducing runtime and improve interpretability, not only for performance
+The usefulness of filters in this study varied.
+While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all \autoref{fig:filter-effects-no-filter}.
+Since these negative cases were not caused by a specific filter method, one should include multiple filters in a modeling design if filters are going to be used.
+While filters can improve the performance of models, they might be of more interesting in other aspects than performance: Reducing variables can reduce computational efforts in high-dimensional scenarios and might simplify interpretability of models.
+Here, filters are a lot cheaper than wrapper methods due to their ability of being treated as a hyperparameter during the optimization stage.
 
@@
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
-The PCA approach did neither show the best or worst result when being compared to other feature-selection approaches.
-It is able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
-However, it also removes the possibility to conduct a post-analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
+PCA used as a filter in this work did neither show the best nor worst result compared to other feature-selection approaches.
+It was able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
+However, it also removed the possibility to conduct a post-hoc analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
 Since filter scores only need to be calculated once in a benchmark setting, the runtime advantage might in fact be negligible in practice.

### 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
clarify that meng2018 used FS
add interpreation of ale plots

@@
 
-In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset.
+In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset, with a focus on the former.
 With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
 We used the best-performing algorithm from the HR task (i.e. SVM) to calculate the feature importance.
-ALE plots were not discussed in greater detail and are available as supplementary material.
+The ALE plots show a small relative change compared to their respective mean effects for all chosen features ($< \mid 0.001 \mid$) (\autoref{fig:fi-hr-ale}).
+Most ALE curves show a high dynamic across the value range of each variable, within their respective effect range.
 
@@
 
-\noindent The decision to use a buffer of two meters for the extraction of an observation's reflectance value was a complex decision.
+The decision to use a buffer of two meters for the extraction of an observation's reflectance value was a complex decision.
 Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
@@
 It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, we can observe quite some dynamic in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
+Looking at the spectral curves of the plots, one can observe some apparent reflectance differences in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
 
@@
 
+% ALE plots
+The ALE plots are hard to interpret due to their semi-meaningful absolute values which differ for each band.
+For example, a reflectance value of 100 has a very different meaning in B67 than in B124 because for the latter reflectance values of vegetation are naturally higher.
+Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
+However, no baseline reflectance value for every feature exists since the absolute reflectance values depend on the sensor characteristics.
+Hence, it is hard to make reasonable interpretations of harsh drops/jumps within curves of certain features, e.g. the one at value 240 for feature B115.
+ALE plots can be a powerful tool for interpreting feature importances of correlated features but might be limited in interpretability if the absolute feature values are only semi-meaningful.
+
 \subsection{Comparison to other studies}
@@
 \cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
-However, both did not use spatial \ac{CV} or \ac{FS}.
+However, both did not use spatial \ac{CV} and \cite{kalin2019} no \ac{FS}.
 \cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
@@
 
-\section{SVM ALE plots for task VI}
+%\section{SVM ALE plots for task VI}
 
@@
 
+\pagebreak
 \section{Correlation among filter methods}
-
-\pagebreak
 

### 2020-05-03 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize wording of complete manuscript

@@
 \usepackage{adjustbox}
+
+\usepackage{csquotes}
 
@@
 
-\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade~\cite{lary2016}.
+\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade~ cite{lary2016}.
 This goes in line with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
@@
 Scientists can nowadays relatively easily process large amounts of (environmental) information using various learning algorithms.
-This makes it possible to extend the matrix of possible options in a semi-automated way, possibly stumbling across unexpected findings of process settings that would have never been tested otherwise~\cite{ma2015}.
+This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would have never been tested otherwise \cite{ma2015}.
 
@@
 This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship.
-A less known index type which can be derived from spectral information is the normalized ratio index (NRI).
-In contrast to most \ac{VI} indices, the \ac{NRI} does not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
+A less known index type which can be derived from spectral information is the \ac{NRI}.
+In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
 Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
@@
 
-Despite its popularity in environmental modeling, there are no studies so far which used machine learning algorithms in combination with remote sensing data to analyze defoliation on a tree level.
+Despite its popularity in environmental modeling, there are no studies so far which used machine-learning algorithms in combination with remote sensing data to analyze defoliation on a tree level.
 This study aims to close this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
-Method-wise it makes use of machine-learning methods in combination with feature selection and hyperparameter tuning.
+The methodology of this study uses of machine-learning methods in combination with feature selection and hyperparameter tuning.
 Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high-dimensionality in environmental modeling \cite{trunk1979, xu2016}.
@@
 
-	\item Do different environmental feature sets show differences in performance when modeling defoliation at trees?
+	\item Do different (environmental) feature sets show differences in performance when modeling defoliation at trees?
 
@@
 
-	\item Which features are most important for the models and how can these be interpreted in an ecological context?
+	\item Which features are most important and how can these be interpreted in an environmental context?
 
@@
 
-\noindent Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
-The trees in the study area plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation~\cite{mesanza2016, iturritxa2017}.
+Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
+The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation~\cite{mesanza2016, iturritxa2017}.
 In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
-The fungi are assumed to infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
+It is assumed that the fungi infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
 The dieback of these trees, which are mainly used as timber, causes high economic damages \cite{ganley2009}.
@@
 
-\noindent The \textit{Pinus radiata} plots of this study, namely \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-\textit{Oiartzun} has the most observations (n = 559) while \textit{Laukiz 2} shows the largest area size (1.44 ha).
-All plots besides \textit{Luiando} are located nearby the coast (\autoref{fig:study_area}).
-In total 1808 observations are available (\textit{Laukiz 1} = 559, \textit{Laukiz 2} = 451, \textit{Luiando} = 301, \textit{Oiartzun} = 497).
+The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
+Oiartzun has the most observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
+All plots besides Luiando are located nearby the coast (\autoref{fig:study_area}).
+In total 1808 observations are available Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
 The data was surveyed in September 2016.
@@
 
-\noindent The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around noon.
-The images were taken using a AISAEAGLE-II sensor.
+The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around noon.
+The images were taken by an AISAEAGLE-II sensor.
 All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.
-The first four bands are corrupted, leaving 122 bands with valid information.
+The first four bands were corrupted, leaving 122 bands with valid information.
 Additional metadata information is available in \autoref{tab:hyperparameter_limits}.
@@
 \subsection{Derivation of indices}
-\noindent To use the full potential from the hyperspectral data, all possible vegetation indices supported by the R package $\left\{hsdar\right\}$ (89 in total) as well as all possible \ac{NRI} combinations were calculated.
+\noindent To use the hyperspectral data's full potential, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
 The following formula was used for the NRI calculation:
@@
 
-\noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used for extraction of the reflectance values.
-These averaged reflectance values were used as the base to derive all feature sets.
+\noindent To account for geometric offsets within the hyperspectral data, which were reported with up to 1 m from \ac{ICGC}, a buffer of two meters around the centroid of each tree was used when extracting the reflectance values.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
+All of those pixels formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
 In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated due to the first four bands of the sensor being corrupted.
@@
 
-To reduce the dimensions of a dataset, conceptually different approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or approaches like \ac{PCA} \cite{das2001, jolliffe2016}.
-Since the main focus of this study is on investigating the effect of filter methods on predictive performance, only this sub-group will be introduced in more detail in the following.
+To reduce the feature space of a dataset, conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or \ac{PCA} \cite{das2001, jolliffe2016}.
+Since this study focuses mainly on filter methods, only this sub-group will be introduced in greater detail in the following subsections.
 
@@
 The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
-Some filter methods are restricted towards specific types of variables (numeric or nominal).
-Filters do only rank features, they do not choose which covariates to drop or keep \cite{drotar2015}.
-The selection of features to keep for the model fitting step is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
+Some filter methods can only deal with specific types of variables (numeric or nominal).
+Filters only rank features, they do not decide which covariates to drop or keep \cite{drotar2015}.
+The selection which features to keep for model fitting is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
 Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
-The goal is to optimize the number of features (using the ranked covariates) at which the model achieves the best performance.
+The goal is to optimize the number of features, after ranking was done, to the point at which the model achieves the best performance.
 
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' are able to enhance the predictive performance of the resulting models \cite{abeel2010, drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' are able to enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
-For this filter, the final order is the sum of all single filters.
+For this filter, the final feature order is the sum of all single filters ranks.
 
@@
 Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
-It is important to not give certain classes too much weight in the ensemble as otherwise the final result will be biased towards these.
+Care needs to be taken to not weight certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
 In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
-In case pairs of filters with a correlation above 0.9 were discovered, only one of these was included into the ensemble filter.
-By this we ensured that the ensemble filter is not biased towards a certain group of methods yielding highly similar rankings.
+If filter pairs showed a correlation of 0.9 or higher, only one of the two was included into the ensemble filter, selected at random.
+By this it was ensured that the ensemble filter composition was not biased towards a certain group of filter methods.
 
@@
 \begin{itemize}
-	\item univariate/multivariate (scoring based on a single variable / multiple variables)
-	\item linear/non-linear (calculation of linear/non-linear interaction terms)
-	\item entropy/correlation (scoring based on derivations of entropy or correlation-based approaches)
+	\item Univariate/multivariate (scoring based on a single variable / multiple variables).
+	\item Linear/non-linear (calculation of linear/non-linear interaction terms).
+	\item Entropy/correlation (scoring based on derivations of entropy or correlation-based approaches).
 \end{itemize}
@@
 
-\noindent Filter 'Information Gain' is only defined for nominal response variables:
+The filter 'Information Gain' is only defined for nominal response variables:
 
@@
 where \(H\) is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
-To be able to use this method with a numeric response (defoliation in our case), the variable is discretized into equal bins and treated as a class variable.
-While the number of bins can be treated as a hyperparameter of the filter method, we decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of $> 0.9$ for different bin sizes were observed during data exploration.
+In order to use this method with a numeric response (percentage defoliation at trees), the variable was discretized into equal bins and treated as a class variable.
+While the number of bins could be treated as a hyperparameter of the filter method, it was decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of $> 0.9$ for different bin sizes were observed during data exploration.
 
@@
 
-\noindent The benchmarking matrix of this study consists of the following algorithms:
+The following learners were used in this work:
 
@@
 
-\noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
-\ac{XGBoost} (commonly abbreviated as XGBoost) showed promising results in benchmark competitions in recent years.
-Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
+Random forest and {SVM} are well established algorithms widely used in (environmental) modeling.
+Extreme Gradient Boosting (commonly abbreviated as XGBoost) showed promising results in benchmark studies in recent years.
+Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by penalizing the coefficients of the model \cite{hastie2001}.
 Common penalties are 'lasso' (L1) and 'ridge' (L2).
-The latter does not allow the full removal of variables from the model (penalization to zero) while the former does.
-Both penalties can also be combined.
-The combined approach is called 'elastic net' but was not used in this work.
+Ridge does not remove variables from the model (penalization to zero) but just shrinks them to effectively zero, keeping them in the model.
+The combination of both penalities is called 'elastic net' but was not used in this work.
 
 \subsubsection{Feature sets}
-% TODO: talk about the reasons behind different features sets: environmental implications
-
-\noindent Three feature sets were used in this study with each representing a different way of feature engineering:
+
+Three feature sets were used in this study with each representing a different way of feature engineering:
 
 \begin{itemize}
-	\item The raw hyperspectral band information (HR): No feature engineering) %chktex 13
-	\item Vegetation Indices (\ac{VIs}): Expert-based feature engineering)
-	\item Normalized Ratio Indices (\ac{NRIs}): Data-driven feature engineering)
+	\item The raw hyperspectral band information (HR): no feature engineering) %chktex 13
+	\item Vegetation Indices (\ac{VI}s): expert-based feature engineering)
+	\item Normalized Ratio Indices (\ac{NRI}s): data-driven feature engineering)
 \end{itemize}
 
-The idea of splitting the features into different sets originated from the question whether feature engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated previously in a spectrotemporal setting \cite{pena2017}.
-Benchmarking all models on these distinct groups of features makes it easier to draw conclusions on their impact when keeping all other variables such as model type, tuning strategy and partitioning method constant.
-However, rather than only looking at these three groups we decided to also take their combinations into account for the overall comparison:
+The idea of splitting the features into different sets originated from the question whether feature-engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated previously in a spectro-temporal setting \cite{pena2017}.
+Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
+However, rather than only looking at these three groups also combinations of such were taken into account:
 
@@
 
-Even though the feature selection step was passed onto the filter methods in this study, we ensured to not include features with a pairwise correlation of 1.
-Having such can cause undesired effects during model fitting and feature importance calculation.
-To account for such, all pair-wise correlations between features were calculated.
-For pairs which exceeded the threshold of 0.9999999999 the one with the largest mean absolute correlation was removed.
-This process was repeated $p$ times, each time calculating a new correlation matrix.
-
-This preprocessing step reduced the feature sets to 122 HR features, 86 VIs and 7467 NRIs.
+Even though the feature-selection step should be solely left to the filter methods in this study,it was ensured a priori to account for features with a pairwise correlation of 1.
+Having such features within the data can cause undesired effects during model fitting and feature importance calculation.
+Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of 0.9999999999, the feature with the largest mean absolute correlation was removed from the dataset.
+This process was repeated $p$ times, each time calculating a fresh correlation matrix.
+
+This preprocessing step reduced the amount of covariates to 122 (HR), 86 (VI) and 7467 (NRI).
 
@@
 
-\noindent Hyperparameters were tuned using \ac{MBO} within a nested spatial CV \cite{mlrmbo, schratz2019}.
-This approach first composes \textit{n} randomly chosen hyperparameter settings out of a user defined search space.
-After these \textit{n} tries have been evaluated, a new hyperparameter setting, going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
+Hyperparameters were tuned using \ac{MBO} within a nested spatial \ac{CV} \cite{mlrmbo, schratz2019}.
+In MBO first \textit{n} randomly chosen hyperparameter settings out of a user defined search space are composed.
+After these \textit{n} settings have been evaluated, one new setting, which going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
 This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.
 
-An initial design of 30 randomly composed hyperparameter settings and a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is the substantial reduction of the tuning budget required to find a setting which is close to the global minimum.
+In this work, an initial design of 30 randomly composed hyperparameter settings in combination with a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
+The advantage of this tuning approach is the substantial reduction of the tuning budget which is required to find a setting close to the global optimization minimum.
 \ac{MBO} may outperform methods that do not use information from previous iterations, such as random search or grid search \cite{bergstra2012}.
 
-For the filter methods, the percentage of dataset's features was added to the models as a hyperparameter.
+To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage.
 For \ac{PCA}, the number of principal components was tuned instead.
-Random Forest hyperparameter \texttt{\(m_{try}\)} was modified into power transformed \(p^{m_{try}}\) (ranging from 1 to $p$) to work on relative values of the respective dataset's feature count.
-Here \(p\) is the number of features of the dataset.
-This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen out of bounds during tuning with respect to the dataset's feature count'
-After the filtering, only a subset of features of the task is used for optimizing the model's hyperparameters.
-The size of this subset is always different since the number of features is also optimized'
-Hence, the value of \texttt{\(m_{try}\)} needs to be created dynamically based on the dataset's feature count.
+The RF hyperparameter \texttt{\(m_{try}\)} was modified into power transformed \(p^{m_{try}}\) (ranging from 1 to $p$, with \(p\) being the number of features of the dataset) to work on relative values of the respective dataset's feature count.
+This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features left after optimizing the feature percentage during tuning.
 
@@
 Each plot served as one fold within the cross-validation setting, resulting in four iterations in total.
-For the inner level (hyperparameter tuning), \(k - 1\) folds were used where \(k\) is the number of plots.
-
-In total the benchmarking matrix consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods + 2 $\times$ 6 L1 and L2 models).
+For the inner level (hyperparameter tuning), \(k - 1\) folds were used with \(k\) being the number of plots.
+
+In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods + 2 $\times$ 6 L1 and L2 models).
 
@@
 \noindent Estimating feature importance for highly-correlated datasets is a complicated task.
-The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
+The correlation between covariates makes it challenging to calculate an unbiased estimate for single features \cite{molnar2019}.
 Methods like \ac{PDP} or permutation-based approaches do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
-Permutation feature importance calculates the loss of predictive performance for each feature by permuting it in a random manner.
+Permutation feature importance calculates the loss of predictive performance for each feature by permuting it in a random manner during prediction.
 The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
@@
 
-However, as of today there are no completely satisfactory approaches for estimating feature importance for (highly) correlated datasets.
+However, as of today there are no unbiased approaches for estimating feature importance for (highly) correlated datasets.
 Nevertheless, recent methods like \ac{ALE} plots aim to tackle this problem by utilizing marginal distributions of features over a small window to focus on the effect of a single feature \cite{apley2019, molnar2019}.
-However, ALE plots need visual inspection and as the number of features increases, making objective comparisons becomes more complicated.
-
-In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset, with a focus on the former.
+However, ALE plots need visual inspection and as the number of features increases, making objective comparisons/conclusions becomes challenging.
+
+In this work we applied permutation-based feature importance and \ac{ALE} plots (HR only) to estimate feature importance.
 With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
-We used the best-performing algorithm from the HR task (i.e. SVM) to calculate the feature importance.
-The ALE plots show a small relative change compared to their respective mean effects for all chosen features ($< \mid 0.001 \mid$) (\autoref{fig:fi-hr-ale}).
-Most ALE curves show a high dynamic across the value range of each variable, within their respective effect range.
+We used the best-performing algorithm on the HR task (i.e. SVM) to calculate the feature importance.
 
@@
 
-\noindent For ecological interpretation purposes we linked the ten most important indices of the winning models for each feature set to the spectral regions of the hyperspectral data.
-For feature set HR and NRI a direct linking to the respective bands of the hyperspectral sensor was done.
-For the vegetation indices all bands covered by the spectral range of calculated vegetation indices were counted and summed up.
+For environmental interpretation purposes the ten most important indices of the best performing models of feature sets HR and VI were linked to the spectral regions of the hyperspectral data.
+The aim was to visualize the most important features along the spectral curve of the plots to better understand which spectral regions were most important for the model
 
@@
 
-\noindent The complete study was done using the open-source statistical programming language R \cite{rcoreteam2019}.
-All R packages used in this study can be found in linked repositories.
-Due to space limitations we will only mention the packages of the used algorithms and filter methods.
-
-The algorithm implementations of the following packages have been used: \textit{xgboost}\cite{chen2016} (\textit{xgboost}), \textit{kernlab}\cite{kernlab} (Support Vector Machine) and \textit{glmnet}\cite{glmnet} (ridge regression).
-The filter implementations of the following packages have been used: \textit{praznik\cite{praznik}}, \textit{FSelectorRcpp}\cite{fselectorrcpp}.
-The R package \textit{mlr} \cite{mlr} was used for all modeling related steps.
-\textit{drake} \cite{drake} was used for structuring the work and reproducibility.
+All tasks of this study were conducted using the open-source statistical programming language R \cite{rcoreteam2019}.
+A complete list of all R packages used in this study can be found in linked repositories.
+Due to space limitations only the selected packages with high impact on this work will be explicitly cited.
+
+The algorithm implementations of the following packages have been used: xgboost \cite{chen2016} (\textit{Extreme Gradient Boosting}), kernlab \cite{kernlab} (Support Vector Machine) and glmnet \cite{glmnet} (penalized regression).
+The filter implementations of the following packages have been used: praznik \cite{praznik}, FSelectorRcpp \cite{fselectorrcpp}.
+Package mlr \cite{mlr} was used for all modeling related steps.
+drake \cite{drake} was used for structuring the work and reproducibility.
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
@@
 % - MBO tuned penalized methods are substantially worse
-\noindent Overall, the response variable 'tree defoliation' could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
+Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
 SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
-Ridge faced major issues in four tasks due to one observation that was predicted off the scale.
-SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
+Ridge faced major issues in four tasks due to one observation which was predicted off the response scale (i.e. $> 100$).
+SVM showed the best performance overall with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
 Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 for learner SVM (without filter) but up to 54.26 p.p.
@@
 
-% tasks
+% feature sets
 % - no substantial difference between feature sets
@@
 Datasets containing derived features only (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
-All learners besides SVM show a substantially worse performance when using only the VI dataset (around five percentage points worse than their respective best performance).
+All learners besides SVM showed a substantially worse performance on the VI dataset compared to all others (around five percentage points worse than their respective best performance).
 
@@
 % - all MBO penalized methods have no difference across tasks
-SVM combined with the 'Carscore' filter achieved the best overall performance (RMSE of 27.98 p.p.) (\autoref{tab:perf-top-10}).
+SVM combined with the \enquote{Carscore} filter achieved the best overall performance (RMSE of 27.98 p.p.) (\autoref{tab:perf-top-10}).
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was out of bounds (\autoref{tab:perf-worst-10}).
-For NRI and HR RMSE was 31.16 and 35.45 p.p., respectively.
-In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high value (in the millions).
+Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was enourmous (\autoref{tab:perf-worst-10}).
+For NRI and HR RMSE was 31.16 p.p. and 35.45 p.p., respectively.
+In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high defoliation value (in the millions).
 The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
-This one outlier caused the error estimates of these folds and the average estimate across all folds to be off the scale.
+This one outlier caused the error estimates of these folds and the average estimate across all folds to be offscale.
 
@@
 Using filters for RF showed a substantial increase in performance for all tasks with the exception of VI, for which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
-XGBoost showed a high dependency on filtering the data: In 4 out of 6 tasks using no filter would results in the worst or second worst performance.
-However, in contrast for dataset NRI using no filter results in the best performance.
+XGBoost showed a high dependency on filtering the data: In 4 out of 6 tasks using no filter resulted in the worst or second worst performance.
+In contrast, using no filter on dataset NRI resulted in the best performance.
 XGBoost shows the highest overall differences between filters for a single task:
-For feature set HR, the range is up to 14 percentage points (Carscore vs. no filter)(\autoref{fig:filter-effects-no-filter}).
+for feature set HR, the range is up to 14 percentage points (\enquote{Carscore} vs. \enquote{no filter})(\autoref{fig:filter-effects-no-filter}).
 
@@
 % - SVM unaffected
-When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) when the model without filtering showed a slighly better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
-For SVM, all filters and 'no filter' achieved the same performance on tasks VI and NRI even though \autoref{fig:perf-result} lists "No Filter" as the best option.
+When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) when a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
+For SVM, all filters and \enquote{no filter} achieved the same performance on tasks VI and NRI even though \autoref{fig:perf-result} lists "No Filter" as the best option.
 
@@
 The Borda filter did not achieve the best performance for any learner across any task \autoref{fig:filter-effects-borda}.
-For RF and XGBoost it ranked within the first 50\% when looking at all filters used for a task.
-For XGBoost on the VI task the Borda filter scored the second worst performance.
+For RF and XGBoost it most often ranked within the first 50\% with respect to all filters of a specific task.
+For XGBoost on the VI task, the Borda filter scored the second worst performance.
 
@@
 Most features were selected during optimization for plot Laukiz2 and least for Laukiz1 (\autoref{tab:tune-perc-sel-features}).
-RF used only one feature for plots Luiando and Oiartzun (7675 $\times$ 0.00004) while for Laukiz1 34 and for Laukiz2 230 features were seen as optimal by the optimizer.
+RF used only one feature for plots Luiando and Oiartzun (0.00004 \% of 7675) while for Laukiz1 34 and for Laukiz2 230 features were seen as optimal during optimization.
 In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features) more than two-third of all available features.
@@
 
-The most important features of each dataset showed an average decrease in RMSE of 1.57 p.p. (HR, B69) and 1.79 p.p (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
-For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm (the so called "red-edge").
-For feature set HR four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around one percentage point).
-Overall, most features show only a small importance with average decreases in RMSE below 0.5.
+The most important features for datasets HR and VI showed an average decrease in RMSE of 1.57 p.p. (HR, B69) and 1.79 p.p (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
+For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around one percentage point).
+Overall, most features showed only a small importance with average decreases in RMSE below 0.5 percentage points.
 
@@
 
+\subsubsection{ALE Plots}
+
+The ALE plots show a small relative change compared to their respective mean effects for all chosen features ($< \mid 0.001 \mid$) (\autoref{fig:fi-hr-ale}).
+Most ALE curves show a high dynamic across the value range of each variable, within their respective effect range.
+
 \section{Discussion}
@@
 \subsection{Predictive Performance}
-The best aggregated performance of this study (SVM + Carscore filter, RMSE 27.98 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
-Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be 19 p.p..
+
+The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE 27.99 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 percentage points.
 However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
-With the clear presence of model overfitting in this study it can be postulated that more training data representing all health stages of a plot is needed.
+Due to the apparent presence of model overfitting in this study it can be postulated that more training data representing all health stages of a plot is needed.
 A model can only make robust predictions if it has learned relationships across the whole range of the response.
 Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by a suboptimal distributed training data.
-However, when inspecting the fold level performances, one can say that the model does a decent job predicting defoliation ranging from 50\% to 100\% but fails for 0\% - 50\%.
-This applies to all learners of this study.
+However, when inspecting the fold level performances, it can be concluded that the model did a decent job predicting defoliation ranging from 50\% to 100\% but failed for 0\% - 50\%.
+This applied to all learners of this study.
 
 \subsubsection{Model differences}
+
 An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
 These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
-Comparing these differences (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
+However, comparing this finding (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
 
 Penalized methods showed promising performances, especially when taking runtime into account.
-When removing features with a correlation of 1, lasso is able to score performances around 31 p.p. and shows somewhat surprisingly the best performance on the VI task.
+When removing features with a correlation of nearly 1, lasso is able to score performances around 31 p.p. and shows, somewhat surprisingly, the best performance on the VI task.
 The same general conclusion applies to ridge if one discards the problematic results on four tasks.
-A careful inspection of the hyperparameter optimization procedure of these models did not reveal any implementation mistakes.
+Regarding the problematic predictions of the rigde learner, a careful inspection of the hyperparameter optimization procedure of the fitted models did not reveal any apparent implementation mistakes.
 One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
-We therefore left it in the model, but we acknowledge that these learners seemed particularly sensitive to the input data.
+It was therefore left in the model, but it should be acknowledged that these learners seemed particularly sensitive to the input data.
 
 \subsubsection{Feature set differences}
+
 One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
-With respect to \autoref{fig:perf-result}, no overall trend was found for all models that related to specific feature sets.
+With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.
 The performance of RF and XGBoost on the VI feature set was about six percentage points lower than on others.
@@
 \label{subsec:perf-plot-char}
-The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
+
+The large differences in RMSE obtained on different test folds can be attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
 An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
 Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
-In this specific case, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
-This caused overfitting on the medium-high values, degrading predictive performance in other scenarios.
+In the prediction scenario for Laukiz2, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
+This caused overfitting on the medium-high values, degrading the model's predictive performance in other scenarios.
 In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
@@
 The high differences of selected features per plot plot during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
-While SVM and XGBoost require a substantial portion of features in most cases to achieve robust predictions, RF is able to achieve the best results with a relatively low amount of features.
+While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF is able to achieve the best results with a relatively low amount of features.
 This fact leads to reduced computational runtime, especially if model parameters are dependent on the amount of features (e.g. \texttt{\(m_{try}\)}).
-Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, at least for RF in this study.
-However, the results of (\autoref{tab:tune-perc-sel-features}) should be taken with care as they rely on single model-filter combinations and are subject to random variation.
-More in-depth research is needed to investigate the effect of filters on other criteria (such as runtime) than performance, leading to a multi-criteria optimization problem.
+Hence, regardless of the potential advantage of using filters for increased predictive performance, it should be noted that these can have a strong positive effect on runtime, at least for RF in this study.
+
+Ultimatevily, the results of (\autoref{tab:tune-perc-sel-features}) should be taken with care as they rely on single model-filter combinations and are subject to random variation.
+More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criteria optimization problem.
 
@@
 % - Filters are interesting for reducing runtime and improve interpretability, not only for performance
-The usefulness of filters in this study varied.
+The usefulness of filters with respect to predictive performance in this study varied.
 While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all \autoref{fig:filter-effects-no-filter}.
-Since these negative cases were not caused by a specific filter method, one should include multiple filters in a modeling design if filters are going to be used.
-While filters can improve the performance of models, they might be of more interesting in other aspects than performance: Reducing variables can reduce computational efforts in high-dimensional scenarios and might simplify interpretability of models.
+Since these negative cases were not caused by a specific filter method, it is recommended to test multiple filters in a study if filters are going to be used.
+While filters can improve the performance of models, they might be of more interesting in other aspects than performance: reducing variables can reduce computational efforts in high-dimensional scenarios and might enhance the interpretability of models.
 Here, filters are a lot cheaper than wrapper methods due to their ability of being treated as a hyperparameter during the optimization stage.
@@
 Hence, their main advantage are stable performances across datasets with varying characteristics.
-Certain single filter methods might yield better model performances for certain datasets but fail on others.
+Single filter methods might yield better model performances on certain datasets but fail on others.
 The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBoost on task HR-NRI-VI) a single filter outperformed the ensemble filter.
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
-Therefore no general statement can yet be made if the use of ensemble filters leads to a substantial better performance than applying one of the various single filter methods.
-More case studies applying ensemble filter methods are needed.
-We believe that ensemble filters can be a promising option in a machine-learning feature-selection portfolio.
+Therefore no general statement can be made from the results of this study if the use of ensemble filters leads to a substantial better performance than applying one of the various single filter methods.
+For this, more case studies applying ensemble filter methods are needed.
+In any case, ensemble filters can be a promising addition to a machine-learning feature-selection portfolio.
 
@@
 
-The decision to use a buffer of two meters for the extraction of an observation's reflectance value was a complex decision.
-Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
+The decision to use a buffer of two meters for the extraction of reflectance values was a complex process.
+Due to the reported geometric offset of up to 1 m within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different observation was reasonably high.
 By using a buffer of two meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
-However, the difference between single pixel values might often be higher than the difference between a smoothed value composed out of a buffer.
-In our view using a buffer of two meters was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
+However, when using no buffer at all, the difference between single, neighboring pixels might even be higher compared to a smoothed value composed out of a buffer.
+It was concluded that using a buffer of two meters is a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
 Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
 
-Another relevant point is that the exact number of contributing pixels to the final index value of an observation cannot be determined as it depends on the location of the tree within the pixel grid.
-According to the extract function of the \texttt{raster} package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
-As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of that tree within the pixel grid.
-Next, if a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation would be located in the middle of the plot.
-
-The hsdar R package was used for the calculation of vegetation indices \cite{hsdar}.
+Another point worth discussing is that the exact number of contributing pixels to the final index value of an observation cannot be exactly determined: it depends on the location of the tree within the pixel grid.
+According to the extract function of the raster package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
+As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of a tree within the pixel grid.
+If a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation would be located in the middle of the plot.
+
+The R package hsdar was used for the calculation of vegetation indices \cite{hsdar}.
 All indices that could be calculated with the given spectral range of the data were used.
-Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on this preselection.
-
-The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
-Such limitations and uncertainties apply to many environmental studies and cannot be completely avoided.
+Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on the preselection of indices offered by the package.
+
+Overall, the magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
+Such limitations and uncertainties apply to most environmental studies and cannot be completely avoided.
 
@@
 
-This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and their respective opposites \cite{horler1983}.
-It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, one can observe some apparent reflectance differences in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
-
-A possible explanation for the less good performances of the models scored on the VI dataset compared to all other feature sets could be the lack of features in the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
-The majority of the VI covers the range 550 nm - 800 nm.
+This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and its respective counterpart \cite{horler1983}.
+However, four out of ten of the most important features of dataset HR are located between 920 nm and 1000 nm.
+Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area - especially for plot Oiartzun - which might explain why these features were considered as important by the model.
+
+A possible explanation for the less good performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
+The majority of VI features covers the range between 550 nm - 800 nm.
 Only one index (PWI) covers information in the range beyond 900 nm.
-The spectral signatures of the plots show some variation around 940 nm, especially plot Oiartzun, which could be of potential help for the models when distinguishing between high and low defoliation.
-In addition, four our out of ten of the most important features in the HR dataset (B113, B115, B124 and B126) are located in this range (930 nm - 1000 nm), confirming the importance of this spectral range.
 
@@
 The ALE plots are hard to interpret due to their semi-meaningful absolute values which differ for each band.
-For example, a reflectance value of 100 has a very different meaning in B67 than in B124 because for the latter reflectance values of vegetation are naturally higher.
+For example, a reflectance value of 100 has a very different meaning in B67 than in B124 because for the latter, reflectance values of vegetation are naturally higher.
 Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
@@
 
-\noindent Most other studies analyzing defoliation operated on the plot rather than the tree level.
+Most other studies analyzing defoliation operated on the plot rather than the tree level.
 This is due to the low spatial resolution of used satellite products which served as the input data, making a tree-level study infeasible \cite{townsend2012, debeurs2008, rengarajan2016}.
@@
 The field of (hyperspectral) remote sensing has a strong focus on using random forest for modeling in recent years \cite{belgiu2016}.
-In high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
+However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
 To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
 However, we could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
-Potential reasons could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
+Potential reasons for this gap could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
 Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for upcoming modeling studies using hyperspectral remote sensing data.
@@
 % RS studies with more/other algorithms than RF	
-Remote sensing studies comparing multiple models often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
+When looking for remote sensing studies which compare multiple models, it turned out that such often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
 
@@
 
-In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
-Also, most environmental/ecological datasets are not high-dimensional.
-In fact, the number of predictors is often less than ten and issues related to correlations are solved manually instead of relying on an automated approach.
+In summary, no studies which used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to forest health were found.
+This might relate to the fact that most environmental/ecological datasets are not high-dimensional.
+In fact, the number of predictors is often less than ten and issues related to correlations are often solved manually instead of relying on an automated approach.
+
 The bioinformatics field faces high-dimensional datasets more often.
 Hence more studies using (filter-based) feature-selection approaches can be found for this field \cite{guo2019, radovic2017}.
-However bioinformatics differs conceptually in many ways from environmental modeling and we do not want to dig deeper just for the sake of finding any similar studies.
+Yet bioinformatics differs conceptually in many ways from environmental modeling and digging deeper just for the sake of finding any similar studies is not a good idea.
 If a field only rarely faces high-dimensional dataset issues (like environmental modeling) the motivation and expertise of using advanced methods to solve such are rather low.
-We hope that this work can give some guidance and serve as a starting point for tackling high-dimensional problems in environmental remote sensing.
+We hope that this work and its methodology raises awarness about the application of filter methods to tackle high-dimensional problems in the environmental modeling field.
 
 \section{Outlook and conclusion}
-This study analyzed defoliation at trees in northern Spain by using hyperspectral data as input for machine-learning models which incorporated hyperparameter tuning and filter-based feature selection.
+
+This study analyzed defoliation at trees in northern Spain by using hyperspectral data as input for machine-learning models which used hyperparameter tuning and filter-based feature selection.
 Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
@@
 % RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
-The creation of derived feature sets was able to improve predictive performance in most cases.
-In contrast, datasets which covered only a fraction of the available spectral range (i.e. dataset VI) showed a worse performance than the ones which covered the full range (400 nm - 1000 nm; HR, NRI).
-Combining feature sets helped most models in improving their predictive performance.
+The addition of derived feature sets was in most cases able to improve predictive performance.
+In contrast, feature sets which focused on only a small fraction of the available spectral range (i.e.
+dataset VI) showed a worse performance than the ones which covered wider range (400 nm - 1000 nm; HR, NRI).
+NRIs can be seen as a valuable addition for optimizing predictive performance in vegetation related studies.
 
 % RQ4: Which features are most important for the models and how can these be interpreted in an ecological context?
-Features along the red-edge wavelength region were most important for models to achieve good predictions.
+Features along the red edge wavelength region were most important for models during prediction.
 With respect to dedicated vegetation indices, the "Vogelmann" index with all of its derivatives was seen as the most important index for the best performing SVM model.
@@
 These were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
-However, feature importance for highly-correlated features is challenging.
+However, feature importance for highly-correlated features remains a challenging task.
 Results might be biased and should be taken with care.
@@
 The potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 27 percentage points scored by the best performing model.
-More training data covering a wider range of defoliation values in a larger number of forest plantations would probably help to build better models that can achieve more robust predictions.
+More training data covering a wider range of defoliation values in a larger number of forest plantations is needed to train better models which can create more robust predictions.
 

### 2020-05-03 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
     type = "latex",
-    caption = "Optimized number of features for selected learner-filter settings across folds.",
+    caption = "Optimized number of features for selected learner-filter settings across folds for task HR-NRI-VI.",
     label = "tab:tune-perc-sel-features",

### 2020-05-03 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 
-```{r fi-hr-ale, dev = c("png", "pdf")}
+```{r fi-hr-ale, dev = c("png", "pdf"), warning=FALSE}
 ale_p1 = fi_ale_hr_buffer2$plot(
@@
 ale_p1[[10]] = ale_p1[[10]] + scale_y_continuous(
-  n.breaks = 4
+  n.breaks = 5
 )
@@
 
-```{r}
+```{r, warning=FALSE}
 ale_p2 = fi_ale_hr_buffer2_gs20$plot(
@@
 
-```{r fi-vi-ale, dev = c("png", "pdf")}
+```{r fi-vi-ale, dev = c("png", "pdf"), warning=FALSE}
 ale_p3 = fi_ale_vi_buffer2$plot(
@@
 
-```{r}
+```{r, warning=FALSE}
 ale_p4 = fi_ale_vi_buffer2_gs20$plot(

### 2020-05-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
add abstract

@@
 \begin{abstract}
-	The abstract goes here.
+	Analyzing highly-correlated feature-rich datasets from hyperspectral remote sensing data is a challenging task.
+	This study compared multiple machine-learning and statistical learning methods using filter-based feature-selection methods on a hyperspectral dataset.
+	In the case study, defoliation at trees (\%) in northern Spain was modeled as a function of band reflectance.
+	Multiple feature sets were compared using derived features (expert-based and data-driven).
+	Among the selected methods, support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by about three percentage points.
+	The combination of feature sets showed small increases in predictive performance while no substantial differences between single feature sets were observed.
+	Filter methods helped in certain task-learner-filter settings to achieve better predictive performances while ensemble filters did not have a substantial impact on performance.
+	Depending on the dataset-learner combination, filter methods can help to improve predictive performance and lower computational runtime.
+	Permutation-based feature importance estimated features around the red edge to be most important for the models.
+	However, features in the near-infrared region (800 nm - 1000 nm) were also important to achieve acceptable performances.
+	The dataset used in this work was suboptimal and more training data is needed for more generalizable conclusions.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-	hyperspectral imagery, forest health monitoring, machine learning, feature selection, model comparison, imaging spectroscopy
+	hyperspectral imagery, forest health monitoring, machine learning, feature selection, model comparison, filter, imaging spectroscopy
 \end{IEEEkeywords}
@@
 		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
-		\caption{Response variable \textit{defoliation} at trees for plots \textit{Laukiz 1}, \textit{Laukiz 2}, \textit{Luiando} and \textit{Oiartzun}.
+		\caption{Response variable \textit{defoliation} at trees for plots Laukiz1, Laukiz2, Luiando and Oiartzun.
 \texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation.}\label{fig:defol-distr}

### 2020-05-06 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-04-30 analysis/eval-performance.Rmd
@@
     type = "latex",
-    caption = "Top 10 results for any task/learner/filter combination, sorted by performance.",
+    caption = "Best ten results for any task/learner/filter combination, sorted ascending by RMSE",
     label = "tab:perf-top-10"
@@
     type = "latex",
-    caption = "Best performance of each learner across any task and filter method.",
+    caption = "Best performance of each learner across any task and filter method, sorted ascending by RMSE",
     label = "tab:best-learner-perf"
@@
     type = "latex",
-    caption = "Worst 10 results for any task/learner/filter combination, sorted descreasing by RMSE",
+    caption = "Lower ten results for any task/learner/filter combination, sorted decreasing by RMSE",
     label = "tab:perf-worst-10"
@@
   dplyr::mutate(Learner = forcats::fct_recode(Learner,
-    "RF Car" = "RF MBO Car", "XGBoost Borda" = "XGBOOST MBO Borda",
+    "RF Car" = "RF MBO Car", "XGB Borda" = "XGBOOST MBO Borda",
     "SVM Relief" = "SVM MBO Relief"
@@
     type = "latex",
-    caption = "Optimized number of features for selected learner-filter settings across folds.",
+    caption = "Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE",
     label = "tab:tune-perc-sel-features",

### 2020-05-06 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update tables
update tables and figures

@@
 \centering
-\caption{Percentage of selected features of each fold during optimization for selected learners} 
+\caption{Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE} 
 \label{tab:tune-perc-sel-features}
@@
    & Laukiz2 & 0.03037 & 15.39 \\ 
-  \midrule\multirow{4}{*}{\textbf{XGBoost Borda}} & Laukiz1 & 0.00604 & 15.00 \\ 
+  \midrule\multirow{4}{*}{\textbf{XGB Borda}} & Laukiz1 & 0.00604 & 15.00 \\ 
    & Oiartzun & 0.68227 & 36.15 \\ 

### 2020-05-06 docs/00-manuscripts/ieee/ieee-manuscript.tex

--- 2020-05-01 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 	\acro{VI}{vegetation index}
-	\acro{XGBoost}{extreme gradient boosting}
+	\acro{XGBOOST}{extreme gradient boosting}
 \end{acronym}
@@
 
-% mention why FS is used / important
 Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
-Feature selection approaches can help to address this issue, reducing possible in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
+At the same time, it is necessary to account for the presence of spatial autocorrelation in the data.
+The dataset used in this study comes with a spatial grouping at the plot level which needs to be accounted for \cite{sperrorest, pena2015}.
 
@@
 \noindent To account for geometric offsets (which were reported with up to 1 m from \ac{ICGC}) within the hyperspectral data, a buffer of two meters around the centroid of each tree was used for extraction of the reflectance values.
-These averaged reflectance values were used as the base to derive all feature sets.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
@@
 
-\subsection{Feature selection}
+\subsection{Dimension reduction}
+
+\noindent We want to stress the differences between 'dimension reduction' and 'high-dimensionality'.
+The former refers to the general idea of reducing features from a dataset \cite{vandermaaten2007}.
+In modeling this means finding the best subset of covariates which provide the most predictive power to the model or extracting the main components of the features using a \ac{PCA}.
+'High-dimensionality' in contrast is a dataset attribute and applies when \(p > n\), where \(p\) is the number of covariates and \(n\) the number of observations \cite{hastie2001}.
+There is no absolute value at which the term applies, both for \(n\) and \(p\).
+Hence for this study the word ‘high-dimensionality‘ only applies to the experiments using the ‘NRI‘ feature set ($>$ 7000 features for 1808 observations).
 
@@
 
-To reduce the dimensions of a dataset, conceptually different approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or approaches like \ac{PCA} \cite{das2001, jolliffe2016}.
-Since the main focus of this study is on investigating the effect of filter methods on predictive performance, only this sub-group will be introduced in more detail in the following.
+\noindent In the following sections a brief overview of feature selection sub-categories is given.
+Due to the focus of this study on the use of filter methods, which are a sub-group of feature selection methods, other approaches were grouped into a single section.
 
@@
 % Filter methods
-The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
+\noindent The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
 Some filter methods are restricted towards specific types of variables (numeric or nominal).
@@
 
+\paragraph{Ensemble filter methods}
+
 % Ensemble filter methods
@@
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015}.
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
@@
 
-% Ensuring a fair weighting in the ensemble
+\paragraph{Ensuring a fair weighting in the ensemble}
+
 Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
@@
 
-\subsubsection{Description of used filter methods}
+\paragraph{Description of used filter methods}
 
@@
 
+\subsubsection{Wrapper methods and PCA}
+
+\noindent Other approaches to assess feature importance are so called 'wrapper methods' and the \ac{PCA}\cite{das2001, jolliffe2016}.
+
+% Wrapper approach
+\noindent Wrappers \cite{chandrashekar2014, kohavi1997} apply search algorithms such as 'random search' or 'generic simulated annealing'.
+First a (random) subset of features is chosen based on the selected algorithm.
+In comparison to filters, no ranking is done in this step.
+Now the model is fitted on the data and the performance is evaluated.
+This is done multiple times, depending on the defined stopping criteria set by the user (performance, runtime, evaluations).
+A disadvantage of this approach is that hyperparameter tuning can only be applied after the feature selection optimization finished.
+Hence, the 'wrapper approach' is an expensive optimization method because two stages need to be run in sequential order.
+Due to their high computational cost, wrappers were not considered in this work.
+
+% PCA
+A method with a completely different approach compared to filters and wrappers is the \ac{PCA}\cite{pearson1901, jolliffe2016}.
+Here, the main components of the feature space are extracted and combined.
+The number of principal components needed to explain a large portion of the variance is dynamic and depends on the dataset.
+In practice, this number often ranges between two and five.
+By using the (automatically estimated) explained variance of the main components, the model can rely on a few features containing the majority of information available in the data.
+This enables cheap model fitting with balanced loss of predictor information.
+A disadvantage is the reduced interpretability due to the main components being a squashed version of the original covariates.
+
 \subsection{Benchmarking design}
@@
 \begin{itemize}
-	\item  Extreme Gradient Boosting (XGBoost)
+	\item  Extreme Gradient Boosting (XGBOOST)
 	\item  Random Forest (RF)
@@
 \noindent \ac{RF} and {SVM} are well established algorithms that are widely used in environmental modeling.
-\ac{XGBoost} (commonly abbreviated as XGBoost) showed promising results in benchmark competitions in recent years.
+\ac{XGBOOST} (commonly abbreviated as XGBOOST) showed promising results in benchmark competitions in recent years.
 Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by applying a penalization term which shrinks the coefficients of the model\cite{hastie2001}.
@@
 	\item The raw hyperspectral band information (HR): No feature engineering) %chktex 13
-	\item Vegetation Indices (\ac{VIs}): Expert-based feature engineering)
-	\item Normalized Ratio Indices (\ac{NRIs}): Data-driven feature engineering)
+	\item Vegetation Indices (\ac{VI}): Expert-based feature engineering)
+	\item Normalized Ratio Indices (\ac{NRI}): Data-driven feature engineering)
 \end{itemize}
@@
 
+% FIXME Talk about the bias of permutation FI in highly correlated settings
 \noindent Estimating feature importance for highly-correlated datasets is a complicated task.
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features.
-Methods like \ac{PDP} or permutation-based approaches do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
-Permutation feature importance calculates the loss of predictive performance for each feature by permuting it in a random manner.
+Methods like \ac{PDP} do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
+% FIXME: Add ref
+Due to the noisiness of the dataset, estimates from model-agnostic approaches such as permutation-based feature importance should also be taken with care.
+This method calculates the loss of predictive performance for each feature by permuting it in a random manner.
 The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
@@
 
-In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset, with a focus on the former.
+In this work we applied permutation-based feature importance and \ac{ALE} plots to calculate feature importance for the HR dataset.
 With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
 We used the best-performing algorithm from the HR task (i.e. SVM) to calculate the feature importance.
-The ALE plots show a small relative change compared to their respective mean effects for all chosen features ($< \mid 0.001 \mid$) (\autoref{fig:fi-hr-ale}).
-Most ALE curves show a high dynamic across the value range of each variable, within their respective effect range.
+ALE plots were not discussed in greater detail and are available as supplementary material.
 
@@
 SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
-Ridge faced major issues in four tasks due to one observation that was predicted off the scale.
+Ridge and lasso faced major issues up to the point of not achieving an error below 50 for datasets HR-NRI-VI and NRI-VI.
 SVM showed the best performance of all learners with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
@@
 The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
-This one outlier caused the error estimates of these folds and the average estimate across all folds to be off the scale.
+This one outlier caused the error estimates of these folds and the average estimate across all folds to be out of bounds.
 
@@
 An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
-These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
-Comparing these differences (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
-
-Penalized methods showed promising performances, especially when taking runtime into account.
-When removing features with a correlation of 1, lasso is able to score performances around 31 p.p. and shows somewhat surprisingly the best performance on the VI task.
-The same general conclusion applies to ridge if one discards the problematic results on four tasks.
-A careful inspection of the hyperparameter optimization procedure of these models did not reveal any implementation mistakes.
-One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
+
+FIXME: Add refs when SVM was clearly worse then other algs.
+In other (environmental) studies in recent years, SVM showed most often less good performances and was not able to outperform competing algorithms such as RF.
+The difference to the second best performing model of around three percentage points is also worth noting:
+We are not aware of a study in which SVM ranked as the best model that clearly.
+All of these statements are based on the assumption that a proper hyperparameter tuning was conducted for all models (which we also claim for ourselves in this study).
+
+We were a bit surprised about the performances of the penalized methods lasso and ridge.
+The former was not able to achieve an RMSE below 46.0 p.p. across any feature set while ridge came at least somewhat close to RF and XGBoost for some feature sets.
+However, there were also three settings (VI, NRI-VI, HR-NRI-VI) for which the RMSE of the ridge model was unreasonably large.
+Careful inspection of the hyperparameter optimization of these learners did not reveal any implementation mistakes.
+However, one observation turned out to be influential even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
 We therefore left it in the model, but we acknowledge that these learners seemed particularly sensitive to the input data.
+
+FIXME: Talk about L1 and L2 model performance
 
@@
 One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
-With respect to \autoref{fig:perf-result}, no overall trend was found for all models that related to specific feature sets.
-The performance of RF and XGBoost on the VI feature set was about six percentage points lower than on others.
-One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
-In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e. NRI-VI, HR-NRI, HR-NRI-VI).
+With respect to \autoref{fig:perf-result} one can say that overall the different feature sets did not result in substantial and consistent changes in performance.
+RF and XGBoost showed a somewhat surprising decrease in performance of about six percentage points on task VI compared to all others.
+We have no explanation for this specific result, especially because the best performing SVM model does not reflect this difference.
+In addition, all learners showed a slightly better average performance on feature sets which included NRIs (e.g. NRI-VI, HR-NRI, HR-NRI-VI) than on those which did not (i.e. VI and HR only).
 
@@
 \label{subsec:perf-plot-char}
+
 The large differences in RMSE obtained on different test folds are attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
@@
 % - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
-% - Filters can also have a negative impact
-% - Filters are interesting for reducing runtime and improve interpretability, not only for performance
-The usefulness of filters in this study varied.
-While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all \autoref{fig:filter-effects-no-filter}.
-Since these negative cases were not caused by a specific filter method, one should include multiple filters in a modeling design if filters are going to be used.
-While filters can improve the performance of models, they might be of more interesting in other aspects than performance: Reducing variables can reduce computational efforts in high-dimensional scenarios and might simplify interpretability of models.
-Here, filters are a lot cheaper than wrapper methods due to their ability of being treated as a hyperparameter during the optimization stage.
-
-% - Ensemble FS shows no advantage to simple methods -> always one simple filter which is better. Ensemble filters are good across many different datasets
-The models which used the Borda ensemble method in this study did not score better on average than models which used a single filter or no filter at all.
-Ensemble methods have higher stability and robustness than single ones and have shown promising results in \cite{drotar2017a}.
-Hence, their main advantage are stable performances across datasets with varying characteristics.
-Certain single filter methods might yield better model performances for certain datasets but fail on others.
-The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBoost on task HR-NRI-VI) a single filter outperformed the ensemble filter.
-However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
-Therefore no general statement can yet be made if the use of ensemble filters leads to a substantial better performance than applying one of the various single filter methods.
-More case studies applying ensemble filter methods are needed.
-We believe that ensemble filters can be a promising option in a machine-learning feature-selection portfolio.
+The usefulness of filters in this study was apparent across all feature sets, even for the smallest feature set, HR.
+Even though the effect of feature selection methods varied across algorithms, small to substantial improvement in predictive performance were achieved in almost all cases.
+No optimal algorithm/filter/dataset setting was observed in this study, leading to the conclusion that one needs to try out various combinations of such to find the best performing one.
+
+% - Ensemble FS shows no advantage to simple methods -> maybe to less features, features too correlated? check diff between effect on NRI and VI
+We have no explanation why the Borda ensemble method used in this study did not score better than the average performance of other filter methods.
+FIXME: Add ref when and why Borda filters are good.
+Ensemble methods have been proven to outperform simple filters in complicated scenarios when the latter face problems due to their focus on certain feature characteristics.
+We suggest that the datasets used in this study were not complex enough to bring up these problems for simple features.
+Therefore it makes sense that, due to its averaging nature, the Borda filter scored average results across all instances.
+However, this cannot be known upfront and we believe that an ensemble filter should be part of every filter portfolio when doing filter-based feature selection.
 
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
-PCA used as a filter in this work did neither show the best nor worst result compared to other feature-selection approaches.
-It was able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
-However, it also removed the possibility to conduct a post-hoc analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
+The PCA approach did neither show the best or worst result when being compared to other feature-selection approaches.
+It is able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
+However, it also removes the possibility to conduct a post-analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
 Since filter scores only need to be calculated once in a benchmark setting, the runtime advantage might in fact be negligible in practice.
@@
 
-The decision to use a buffer of two meters for the extraction of an observation's reflectance value was a complex decision.
+\noindent The decision to use a buffer of 2 m for the extraction of an observation's reflectance value was a complex decision.
 Due to the reported geometric offset of up to 1 m within the hyperspectral data, the possibility of assigning a value to an observation which would actually refer to a different observation was reasonably high.
-By using a buffer of two meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
+By using a buffer of 2 meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
 However, the difference between single pixel values might often be higher than the difference between a smoothed value composed out of a buffer.
-In our view using a buffer of two meters was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
+In our view using a buffer of 2 m was a good compromise between including information of too many surrounding trees and not accounting for the geometric offset at all.
 Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
@@
 The magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
-Such limitations and uncertainties apply to many environmental studies and cannot be completely avoided.
+Such limitations and uncertainties apply to most applied studies and cannot be completely avoided.
+
+FIXME: Talk about L1 and L2 models -> there we have quite some differences!
 
@@
 It is also worth mentioning that four out of ten of the most important features of dataset HR lie between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, one can observe some apparent reflectance differences in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
+Looking at the spectral curves of the plots, we can observe quite some dynamic in this area, especially for Oiartzun, which might explain why these features were considered as important by the model.
 
@@
 
-% ALE plots
-The ALE plots are hard to interpret due to their semi-meaningful absolute values which differ for each band.
-For example, a reflectance value of 100 has a very different meaning in B67 than in B124 because for the latter reflectance values of vegetation are naturally higher.
-Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
-However, no baseline reflectance value for every feature exists since the absolute reflectance values depend on the sensor characteristics.
-Hence, it is hard to make reasonable interpretations of harsh drops/jumps within curves of certain features, e.g. the one at value 240 for feature B115.
-ALE plots can be a powerful tool for interpreting feature importances of correlated features but might be limited in interpretability if the absolute feature values are only semi-meaningful.
-
 \subsection{Comparison to other studies}
@@
 \cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
-However, both did not use spatial \ac{CV} and \cite{kalin2019} no \ac{FS}.
+Both study designs are substantially different compared to the setup of this work.
+In addition, no spatial \ac{CV} or \ac{FS} was used.
 \cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
@@
 To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
-However, we could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
+We could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
 Potential reasons could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
@@
 
-% RS studies with more/other algorithms than RF	
+% RS studies with more/other algs than RF	
 Remote sensing studies comparing multiple models often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
 
-\cite{shendryk2016, ludwig2019} are more similar in their methodology but focus on a different response variable (woody cover).
+\cite{shendryk2016, ludwig2019} are studies which are more similar in their methodology but focus on a different response variable (woody cover).
 \cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees for eucalyptus forests.
@@
 In summary, we could not find studies using filter methods for \ac{FS} or \ac{NRI} indices in their work with a relation to forest health.
-Also, most environmental/ecological datasets are not high-dimensional.
-In fact, the number of predictors is often less than ten and issues related to correlations are solved manually instead of relying on an automated approach.
-The bioinformatics field faces high-dimensional datasets more often.
+Most studies used only one algorithm without (strong) arguments why this particular one has been selected or why only one model was used.
+This is not surprising: Most environmental/ecological datasets are not high-dimensional.
+In contrast, the number of predictors is often less than ten and issues related to correlations are solved manually instead of relying on an automated approach.
+The bioinformatics field faces high-dimensional feature sets more often.
 Hence more studies using (filter-based) feature-selection approaches can be found for this field \cite{guo2019, radovic2017}.
-However bioinformatics differs conceptually in many ways from environmental modeling and we do not want to dig deeper just for the sake of finding any similar studies.
-If a field only rarely faces high-dimensional dataset issues (like environmental modeling) the motivation and expertise of using advanced methods to solve such are rather low.
+If a field only rarely faces high-dimensional dataset issues the motivation and expertise of using advanced methods to solve such are rather low.
 We hope that this work can give some guidance and serve as a starting point for tackling high-dimensional problems in environmental remote sensing.
@@
 \section{Outlook and conclusion}
-This study analyzed defoliation at trees in northern Spain by using hyperspectral data as input for machine-learning models which incorporated hyperparameter tuning and filter-based feature selection.
-Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
+
+This study analyzed defoliation at trees in north Spain by using hyperspectral data as input for machine learning models making heavy use of filter-based feature-selection methods.
+It was shown that substantial differences in performance can occur depending on which feature selection methods and machine learning algorithms are combined.
 SVM showed the most robust behavior across all highly-correlated datasets and was able to predict the response variable of this study substantially better than other methods.
@@
 % RQ3: How are feature-selection methods influencing the predictive performance of the models?
-Filter methods were able to improve the predictive performance on datasets in some instances.
-Their effectiveness depends on the algorithm and the dataset characteristics.
-Ensemble filter methods did not show a substantial improvement over individual filter methods in this study.
+Filter methods are able to improve predictive performance for datasets with many features in certain instances.
+Their effectiveness depends on the chosen algorithm and dataset characteristics.
+Ensemble filter methods did not show a substantial improvement over individual filter methods.
 
@@
 % RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
-The creation of derived feature sets was able to improve predictive performance in most cases.
-In contrast, datasets which covered only a fraction of the available spectral range (i.e. dataset VI) showed a worse performance than the ones which covered the full range (400 nm - 1000 nm; HR, NRI).
-Combining feature sets helped most models in improving their predictive performance.
+The creation of derived feature sets did not contribute to an improvement in model performance.
+In contrast, dataset VI showed a substantially poorer performance than the dataset with the original features (HR).
+Combining feature sets did not show a substantial improvement on predictive performance.
 
@@
 These were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
-However, feature importance for highly-correlated features is challenging.
-Results might be biased and should be taken with care.
 
 % What about the general potential of analyzing defol with hyperspectral data?
-The potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 27 percentage points scored by the best performing model.
-More training data covering a wider range of defoliation values in a larger number of forest plantations would probably help to build better models that can achieve more robust predictions.
+The potential of predicting defoliation for single trees as conducted in this study was rather limited with an average RMSE of 27 p.p..
+We suggest that training data covering a wider range of defoliation values in a larger number of forest plantations would further improve the proposed models and result in more generalizable monitoring algorithms.
 
@@
 
-%\section{SVM ALE plots for task VI}
+\section{SVM ALE plots for task VI}
 
@@
 
+\section{Correlation among filter methods}
+
 \pagebreak
-\section{Correlation among filter methods}
 

### 2020-05-06 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
       aes(color = "HR: Rank 11:122"),
-      size = 1.8,
+      size = 2.5,
       shape = 1,
@@
       shape = 20,
-      size = 2.2,
+      size = 3,
       na.rm = TRUE
@@
       aes(x = wavelength, y = importance, group = class),
-      size = 2.2, shape = 20
+      size = 3, shape = 20
     ) +
@@
       color = "grey",
-      shape = 1, size = 2
+      shape = 1, size = 2.5
     ) +
@@
 ale_p1[[10]] = ale_p1[[10]] + scale_y_continuous(
-  n.breaks = 5
+  n.breaks = 4
 )

### 2020-05-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
update bernds affil

--- 2020-05-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 		of Geography, GIScience group, Friedrich-Schiller-University of Jena, Germany.}% <-this % stops a space
-	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
+	\thanks{B.Bischl is head of the statistical learning and data science group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
 	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Arab, Spain.}% <-this % stops a space

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize filter table wording related to interaction terms

--- 2020-05-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 \usepackage{adjustbox}
+
+% line breaks in table cells
+\newcommand{\specialcell}[2][l]{%
+  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
 
@@
 The methodology of this study uses of machine-learning methods in combination with feature selection and hyperparameter tuning.
+In addition, feature importance and feature effects are evaluated.
 Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high-dimensionality in environmental modeling \cite{trunk1979, xu2016}.
@@
 
-To reduce the feature space of a dataset, conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or \ac{PCA} \cite{das2001, jolliffe2016}.
-Since this study focuses mainly on filter methods, only this sub-group will be introduced in greater detail in the following subsections.
+To reduce the feature space of a dataset, conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or \ac{PCA} \cite{das2001, guyon2003, jolliffe2016}.
+In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
+Due to the focus on filter methods, only this sub-group of feature selection methods will be introduced in greater detail in the following subsections.
 
@@
 % Filter methods
-The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{chandrashekar2014}.
+The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{guyon2003}.
 Some filter methods can only deal with specific types of variables (numeric or nominal).
@@
 
-Filter methods can be classified as follows:
+Filter methods can be classified as follows (\autoref{tab:filter-methods}):
 
@@
 	\item Univariate/multivariate (scoring based on a single variable / multiple variables).
-	\item Linear/non-linear (calculation of linear/non-linear interaction terms).
+	\item Linear/non-linear (usage of linear/non-linear calculations).
 	\item Entropy/correlation (scoring based on derivations of entropy or correlation-based approaches).
@@
 	\item  Penalized Regression (both L1 (lasso) and L2 (ridge))
-	\item  Support Vector Machine (SVM)
+	\item  Support Vector Machine (SVM, RBF Kernel)
 \end{itemize}
@@
 
-\subsection{Feature importance estimation}
+\subsection{Feature importance and feature effects}
 
@@
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was enourmous (\autoref{tab:perf-worst-10}).
+Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was enormous (\autoref{tab:perf-worst-10}).
 For NRI and HR RMSE was 31.16 p.p. and 35.45 p.p., respectively.
@@
 The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
-This one outlier caused the error estimates of these folds and the average estimate across all folds to be offscale.
+This one outlier caused the error estimates of these folds and the average estimate across all folds to be off scale.
 
@@
 The same general conclusion applies to ridge if one discards the problematic results on four tasks.
-Regarding the problematic predictions of the rigde learner, a careful inspection of the hyperparameter optimization procedure of the fitted models did not reveal any apparent implementation mistakes.
+Regarding the problematic predictions of the ridge learner, a careful inspection of the hyperparameter optimization procedure of the fitted models did not reveal any apparent implementation mistakes.
 One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
@@
 
-Ultimatevily, the results of (\autoref{tab:tune-perc-sel-features}) should be taken with care as they rely on single model-filter combinations and are subject to random variation.
+Ultimately, the results of (\autoref{tab:tune-perc-sel-features}) should be taken with care as they rely on single model-filter combinations and are subject to random variation.
 More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criteria optimization problem.
@@
 
+\section{Hyperparameter tuning ranges}
+
+% parameter limits
+\begin{table}[b!]
+\centering
+\caption[t]{Hyperparameter ranges and types for each model.
+	Hyperparameter notations from the respective R packages were used.}
+\begingroup\footnotesize
+\begin{tabular}{llllrrr}
+	\\
+	\specialcell{Model                                          \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
+	\toprule
+	\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$m_{try}.power$}    & int & 0         & 0.5      & -       \\
+	 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
+	 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
+	\midrule
+	\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
+	 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
+	\midrule
+	\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
+	 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
+	 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
+	 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
+	 & \texttt{gamma}              & int & 0        & 10      & 0   \\
+	 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
+	 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
+	\bottomrule
+\end{tabular}
+\endgroup
+\label{tab:hyperparameter_limits}
+\end{table}
+
+\pagebreak
+
 \bibliographystyle{IEEEtran}

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
lar

@@
 Having such features within the data can cause undesired effects during model fitting and feature importance calculation.
-Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of 0.9999999999, the feature with the largest mean absolute correlation was removed from the dataset.
+Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of 0.9999999999, the feature with the largest mean absolute correlation across all variables was removed from the dataset.
 This process was repeated $p$ times, each time calculating a fresh correlation matrix.

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
sell RC a bit better

@@
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
-The code base is available on GitHub (\url{https://github.com/pat-s/2019-feature-selection}).
+Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://github.com/pat-s/2019-feature-selection}),listing more side-analyses that were carried out through the creation of this study.
 

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
mention the limitation of nrounds = 600

@@
 
+The upper range of 600 used for hyperparameter {\texttt{nrounds}} for model XGBoost in this study could a potential limiting factor with respect to the performance of the datasets including NRI variables (\autoref{tab:hyperparameter_limits}).
+This setting was a compromise between runtime and tuning space extension aimed to work well for most feature sets.
+It could be that better predictive performances could have been achieved with a {\texttt{nrounds}} upper limit close to the number of features of the NRI datasets.
+
 \subsubsection{Feature set differences}

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
clarify scope of the study in intro

@@
 
-The research questions of this study are the following:
+% clarify the focus of this study
+This study shows how high-dimensional datasets can be handled effectively with machine-learning methods while still being conduct inference on the fitted models.
+The predictive power of non-linear methods and their ability to handle highly-correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
+However, this study clearly focuses on investigating the effects of filter methods and feature set types on predictive performance rather than on interpreting feature effects.
+
+That said, the research questions of this study are the following:
 

### 2020-05-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 df_perf %<>%
-  group_by(task.id, learner.id, filter) %>%
-  mutate(rmse_aggr = round(mean(rmse), 3)) %>%
-  mutate(se = round(sd(rmse), 3)) %>%
-  select(-rmse, iter) %>%
-  ungroup()
+  dplyr::group_by(task.id, learner.id, filter) %>%
+  dplyr::mutate(rmse_aggr = round(mean(rmse), 3)) %>%
+  dplyr::mutate(se = round(sd(rmse), 3)) %>%
+  dplyr::select(-rmse, iter) %>%
+  dplyr::ungroup()
 ```
@@
   dplyr::mutate(Learner = forcats::fct_recode(Learner,
-    "RF Car" = "RF MBO Car", "XGB Borda" = "XGBOOST MBO Borda",
-    "SVM Relief" = "SVM MBO Relief"
+    "RF \\\\ Car" = "RF MBO Car", "XGB \\\\ Borda" = "XGBOOST MBO Borda",
+    "SVM \\\\ Relief" = "SVM MBO Relief"
   )) %>%
@@
   dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::mutate("\\#" = case_when(
+    Plot == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
+    Plot == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
+    Plot == "Luiando" ~ ceiling(`Features (\\%)` * 301),
+    Plot == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
+  )) %>% 
   dplyr::group_by(Learner) %>%
-  dplyr::select(Learner, Plot, "Features (\\%)", RMSE)
+  dplyr::select(Learner, Plot, "Features (\\%)", "\\#", RMSE)
 
@@
 df_tbl[[1]][first] <-
-  paste0("\\midrule\\multirow{", rle.lengths, "}{*}{\\textbf{", df_tbl[[1]][first], "}}")
+  paste0("\\midrule\\multirow{", rle.lengths, "}{*}{\\specialcell{", df_tbl[[1]][first], "}}")
+# remove redundant midrule from first entry
+df_tbl[[1]][first][1] = gsub("\\\\midrule", "", df_tbl[[1]][first][1])
+
 table4 <- df_tbl %>%
@@
     label = "tab:tune-perc-sel-features",
-    digits = c(0, 0, 0, 5, 2)
+    digits = c(0, 0, 0, 5, 0, 2)
   )

### 2020-05-29 docs/00-manuscripts/ieee/tune-perc-sel-features.tex

@@
 \label{tab:tune-perc-sel-features}
-\begin{tabular}{llrr}
+\begin{tabular}{llrrr}
   \toprule
-Learner & Plot & Features (\%) & RMSE \\ 
+Learner & Plot & Features (\%) & \# & RMSE \\ 
   \midrule
-\midrule\multirow{4}{*}{\textbf{RF Car}} & Laukiz1 & 0.00443 & 27.48 \\ 
-   & Oiartzun & 0.00004 & 37.20 \\ 
-   & Luiando & 0.00002 & 36.98 \\ 
-   & Laukiz2 & 0.03037 & 15.39 \\ 
-  \midrule\multirow{4}{*}{\textbf{XGB Borda}} & Laukiz1 & 0.00604 & 15.00 \\ 
-   & Oiartzun & 0.68227 & 36.15 \\ 
-   & Luiando & 0.99513 & 38.19 \\ 
-   & Laukiz2 & 0.99976 & 29.62 \\ 
-  \midrule\multirow{4}{*}{\textbf{SVM Relief}} & Laukiz1 & 0.00017 & 35.50 \\ 
-   & Oiartzun & 0.71146 & 37.48 \\ 
-   & Luiando & 0.70672 & 14.89 \\ 
-   & Laukiz2 & 0.94696 & 37.19 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.00443 & 3 & 27.48 \\ 
+   & Oiartzun & 0.00004 & 1 & 37.20 \\ 
+   & Luiando & 0.00002 & 1 & 36.98 \\ 
+   & Laukiz2 & 0.03037 & 14 & 15.39 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.00604 & 4 & 15.00 \\ 
+   & Oiartzun & 0.68227 & 340 & 36.15 \\ 
+   & Luiando & 0.99513 & 300 & 38.19 \\ 
+   & Laukiz2 & 0.99976 & 451 & 29.62 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00017 & 1 & 35.50 \\ 
+   & Oiartzun & 0.71146 & 354 & 37.48 \\ 
+   & Luiando & 0.70672 & 213 & 14.89 \\ 
+   & Laukiz2 & 0.94696 & 428 & 37.19 \\ 
    \bottomrule

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
update selected features table with absolute number of features

@@
 \begin{abstract}
-	Analyzing highly-correlated feature-rich datasets from hyperspectral remote sensing data is a challenging task.
-	This study compared multiple machine-learning and statistical learning methods using filter-based feature-selection methods on a hyperspectral dataset.
-	In the case study, defoliation at trees (\%) in northern Spain was modeled as a function of band reflectance.
-	Multiple feature sets were compared using derived features (expert-based and data-driven).
-	Among the selected methods, support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by about three percentage points.
-	The combination of feature sets showed small increases in predictive performance while no substantial differences between single feature sets were observed.
-	Filter methods helped in certain task-learner-filter settings to achieve better predictive performances while ensemble filters did not have a substantial impact on performance.
-	Depending on the dataset-learner combination, filter methods can help to improve predictive performance and lower computational runtime.
+	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
+	The effect of filter-based feature-selection methods on predictive performance was compared.
+	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
+	Feature wise, defoliation at trees (\%) was modeled as a function of band reflectance.
+	Variable importance was assed using permutation-based feature importance.
+
+	With respect to the used models, support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
+	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between single feature sets were observed.
+	Filter methods helped in certain dataset-learner-filter settings to achieve better predictive performances while ensemble filters did not have a substantial impact on performance.
+	For certain dataset-learner combinations, filter methods helped to improve predictive performance and lower computational runtime.
+
 	Permutation-based feature importance estimated features around the red edge to be most important for the models.
-	However, features in the near-infrared region (800 nm - 1000 nm) were also important to achieve acceptable performances.
-	The dataset used in this work was suboptimal and more training data is needed for more generalizable conclusions.
+	However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the best performances.
+
+	The dataset used in this work was suboptimal with respect to the achieved preditive performances and more training data is needed for more generalizable conclusions.
+	Filter methods have potential to help in high-dimensional situations while still keeping the ability to interpret feature effects of fitted models, which is an essential constrain in environmental modeling studies.
+
 \end{abstract}
@@
 \begin{IEEEkeywords}
-	hyperspectral imagery, forest health monitoring, machine learning, feature selection, model comparison, filter, imaging spectroscopy
+	hyperspectral imagery, forest health monitoring, machine learning, feature selection, feature effects, model comparison, filter, imaging spectroscopy
 \end{IEEEkeywords}

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
revewrite methods: feature importance calc

@@
 
-\noindent A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
+A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
 The \ac{RMSE} was chosen as the error measure.
@@
 
-\noindent Estimating feature importance for highly-correlated datasets is a complicated task.
+Estimating feature importance for datasets with highly-correlated features is a complicated task.
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features \cite{molnar2019}.
-Methods like \ac{PDP} or permutation-based approaches do not produce reliable estimates in such scenarios because unrealistic combinations between covariates are created \cite{molnar2019}.
-Permutation feature importance calculates the loss of predictive performance for each feature by permuting it in a random manner during prediction.
-The more important the feature is for the model, the higher the mean loss in the measure, reflecting the importance of the specific feature for the current model.
-Predicting and calculating the error measure multiple times per feature comes with some computational cost.
-For increased efficiency, we estimated the feature importance only for datasets HR and VI using the best performing learner of the benchmark.
-
-However, as of today there are no unbiased approaches for estimating feature importance for (highly) correlated datasets.
-Nevertheless, recent methods like \ac{ALE} plots aim to tackle this problem by utilizing marginal distributions of features over a small window to focus on the effect of a single feature \cite{apley2019, molnar2019}.
-However, ALE plots need visual inspection and as the number of features increases, making objective comparisons/conclusions becomes challenging.
-
-In this work we applied permutation-based feature importance and \ac{ALE} plots (HR only) to estimate feature importance.
-With the limitations of both methods in mind we aimed to get a general overview of the feature importance of the hyperspectral bands while trying to avoid the over-interpretation of results.
-We used the best-performing algorithm on the HR task (i.e. SVM) to calculate the feature importance.
+Methods like \ac{PDP} or permutation-based approaches may produce unreliable estimates in such scenarios because unrealistic situations between covariates are created \cite{molnar2019}.
+Estimating feature importance for machine learning methods is a complicated procedure for which many different approaches, model-specific and agnostic, exist \cite{friedman2001, hastie2001, greenwell2018}.
+The development of robust methods which enable an unbiased estimation of feature importance for highly-correlated variables are subject to current research.
+
+In this work permutation-based feature importance and \ac{ALE} plots (HR only) were calculated to estimate feature importance / effects \cite{apley2019, molnar2019}.
+With the limitations of both methods in mind when applied to correlated features, the aim was to get a general overview of the feature importance of the hyperspectral bands while trying to avoid an over-interpretation of results.
+The best-performing algorithm on the HR task (i.e. SVM) was used for the feature importance calculation.
 

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove noindent

@@
 \subsection{Derivation of indices}
-\noindent To use the hyperspectral data's full potential, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
+
+To use the hyperspectral data's full potential, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
 The following formula was used for the NRI calculation:
@@
 
-\noindent To account for geometric offsets within the hyperspectral data, which were reported with up to 1 m from \ac{ICGC}, a buffer of two meters around the centroid of each tree was used when extracting the reflectance values.
+To account for geometric offsets within the hyperspectral data, which were reported with up to 1 m from \ac{ICGC}, a buffer of two meters around the centroid of each tree was used when extracting the reflectance values.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove some "we" sentences

@@
 It was concluded that using a buffer of two meters is a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
-Even though no results showing the influence of multiple buffer values on the extraction were provided, we hypothesize that the relationships between features would not change substantially, leading to the almost identical model results.
+Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to the almost identical model results.
 
@@
 To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
-However, we could not find a study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
+However, no study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
 Potential reasons for this gap could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
add bommert and binder refs

@@
 
-To reduce the feature space of a dataset, conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or \ac{PCA} \cite{das2001, guyon2003, jolliffe2016}.
+To reduce the feature space of a dataset, conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or \ac{PCA} \cite{bommert2020, das2001, guyon2003, jolliffe2016}.
 In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
@@
 
-Hyperparameters were tuned using \ac{MBO} within a nested spatial \ac{CV} \cite{mlrmbo, schratz2019}.
+Hyperparameters were tuned using \ac{MBO} within a nested spatial \ac{CV} \cite{mlrmbo, binder2020, schratz2019}.
 In MBO first \textit{n} randomly chosen hyperparameter settings out of a user defined search space are composed.
@@
 
-To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage.
+To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage (\cite{binder2020}).
 For \ac{PCA}, the number of principal components was tuned instead.

### 2020-05-29 docs/00-manuscripts/ieee/ieee-manuscript.tex
spellcheck

@@
 \fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's 'caption=false' package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
 
 % *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to Gbe placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., '\begin{figure*}[!b]' is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
 
@@
 \usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
 
@@
 \usepackage[utf8]{inputenc}
-% enables custom font styles which are not available in IEEEtrans by default
+% enables custom font styles which are not available in IEEEtran by default
 % https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran
@@
 % author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
+% note positions of commas and non-breaking spaces ( ~ ) LaTeX will not break
 % a structure at a ~ so this keeps an author's name from being broken across
@@
 	Feature wise, defoliation at trees (\%) was modeled as a function of band reflectance.
-	Variable importance was assed using permutation-based feature importance.
+	Variable importance was assessed using permutation-based feature importance.
 
@@
 
-	The dataset used in this work was suboptimal with respect to the achieved preditive performances and more training data is needed for more generalizable conclusions.
+	The dataset used in this work was suboptimal with respect to the achieved predictive performances and more training data is needed for more generalizable conclusions.
 	Filter methods have potential to help in high-dimensional situations while still keeping the ability to interpret feature effects of fitted models, which is an essential constrain in environmental modeling studies.
@@
 \end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-\ifCLASSOPTIONpeerreview{}
-%\begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-\fi
 
@@
 
-\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade~ cite{lary2016}.
+\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
 This goes in line with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
@@
 Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
-The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation~\cite{mesanza2016, iturritxa2017}.
+The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
 In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
@@
 
-To use the hyperspectral data's full potential, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
+To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
 The following formula was used for the NRI calculation:
@@
 Ridge does not remove variables from the model (penalization to zero) but just shrinks them to effectively zero, keeping them in the model.
-The combination of both penalities is called 'elastic net' but was not used in this work.
+The combination of both penalties is called 'elastic net' but was not used in this work.
 
@@
 All indices that could be calculated with the given spectral range of the data were used.
-Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on the preselection of indices offered by the package.
+Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on the pre-selection of indices offered by the package.
 
@@
 Hence, it is hard to make reasonable interpretations of harsh drops/jumps within curves of certain features, e.g. the one at value 240 for feature B115.
-ALE plots can be a powerful tool for interpreting feature importances of correlated features but might be limited in interpretability if the absolute feature values are only semi-meaningful.
+ALE plots can be a powerful tool for interpreting feature importance of correlated features but might be limited in interpretability if the absolute feature values are only semi-meaningful.
 
@@
 If a field only rarely faces high-dimensional dataset issues (like environmental modeling) the motivation and expertise of using advanced methods to solve such are rather low.
-We hope that this work and its methodology raises awarness about the application of filter methods to tackle high-dimensional problems in the environmental modeling field.
+We hope that this work and its methodology raises awareness about the application of filter methods to tackle high-dimensional problems in the environmental modeling field.
 

### 2020-06-03 docs/00-manuscripts/ieee/ieee-manuscript.tex
grammar check by marie

@@
 	The dataset used in this work was suboptimal with respect to the achieved predictive performances and more training data is needed for more generalizable conclusions.
-	Filter methods have potential to help in high-dimensional situations while still keeping the ability to interpret feature effects of fitted models, which is an essential constrain in environmental modeling studies.
+	Filter methods show potential to be helpful in high-dimensional situations while still keeping the ability to interpret feature effects of fitted models, which is an essential constrain in environmental modeling studies.
 
@@
 Scientists can nowadays relatively easily process large amounts of (environmental) information using various learning algorithms.
-This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would have never been tested otherwise \cite{ma2015}.
+This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would never have been tested otherwise \cite{ma2015}.
 
@@
 
-Machine learning methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
-The ability of predicting to large unknown areas qualifies these tools as a promising toolset for such tasks.
+Machine-learning methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
+The ability of predicting into unknown space qualifies these tools as a promising toolset for such tasks.
 One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \cite{hawrylo2018}.
@@
 This study aims to close this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
-The methodology of this study uses of machine-learning methods in combination with feature selection and hyperparameter tuning.
+The methodology of this study uses machine-learning methods in combination with feature selection and hyperparameter tuning.
 In addition, feature importance and feature effects are evaluated.
@@
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
-For this filter, the final feature order is the sum of all single filters ranks.
+It's final feature order is determined by the sum of all single filters ranks.
 
@@
 Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
-Care needs to be taken to not weight certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
+Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
 In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
@@
 In order to use this method with a numeric response (percentage defoliation at trees), the variable was discretized into equal bins and treated as a class variable.
-While the number of bins could be treated as a hyperparameter of the filter method, it was decided to use \texttt{\(n_{bin}\) = 10} after rank correlations of $> 0.9$ for different bin sizes were observed during data exploration.
+After feature rank correlations of $> 0.9$ betwforeen different bin sizes were observed in a side analysis, \texttt{\(n_{bin}\) = 10} was found to be a reasonable setting to go with.
 
@@
 Random forest and {SVM} are well established algorithms widely used in (environmental) modeling.
-Extreme Gradient Boosting (commonly abbreviated as XGBoost) showed promising results in benchmark studies in recent years.
+Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmark studies in recent years.
 Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by penalizing the coefficients of the model \cite{hastie2001}.
@@
 In MBO first \textit{n} randomly chosen hyperparameter settings out of a user defined search space are composed.
-After these \textit{n} settings have been evaluated, one new setting, which going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
+After these \textit{n} settings have been evaluated, one new setting, which is going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
 This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.
@@
 For environmental interpretation purposes the ten most important indices of the best performing models of feature sets HR and VI were linked to the spectral regions of the hyperspectral data.
-The aim was to visualize the most important features along the spectral curve of the plots to better understand which spectral regions were most important for the model
+The aim was to visualize the most important features along the spectral curve of the plots to better understand which spectral regions were most important for the model.
 
@@
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
-Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://github.com/pat-s/2019-feature-selection}),listing more side-analyses that were carried out through the creation of this study.
+Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://github.com/pat-s/2019-feature-selection}),listing more side-analyses that were carried out during the creation of this study.
 
@@
 Ridge faced major issues in four tasks due to one observation which was predicted off the response scale (i.e. $> 100$).
-SVM showed the best performance overall with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
+SVM showed the best overall performance with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
 Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 for learner SVM (without filter) but up to 54.26 p.p.
@@
 % - never the best
-The Borda filter did not achieve the best performance for any learner across any task \autoref{fig:filter-effects-borda}.
+The Borda filter did not achieve the best performance for any learner across any task (\autoref{fig:filter-effects-borda}).
 For RF and XGBoost it most often ranked within the first 50\% with respect to all filters of a specific task.
@@
 A model can only make robust predictions if it has learned relationships across the whole range of the response.
-Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by a suboptimal distributed training data.
+Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by suboptimal distributed training data.
 However, when inspecting the fold level performances, it can be concluded that the model did a decent job predicting defoliation ranging from 50\% to 100\% but failed for 0\% - 50\%.
@@
 
-The upper range of 600 used for hyperparameter {\texttt{nrounds}} for model XGBoost in this study could a potential limiting factor with respect to the performance of the datasets including NRI variables (\autoref{tab:hyperparameter_limits}).
+The upper range of 600 used for hyperparameter {\texttt{nrounds}} for model XGBoost in this study could be a potential limiting factor with respect to the performance of the datasets including NRI variables (\autoref{tab:hyperparameter_limits}).
 This setting was a compromise between runtime and tuning space extension aimed to work well for most feature sets.
@@
 The usefulness of filters with respect to predictive performance in this study varied.
-While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all \autoref{fig:filter-effects-no-filter}.
+While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all (\autoref{fig:filter-effects-no-filter}).
 Since these negative cases were not caused by a specific filter method, it is recommended to test multiple filters in a study if filters are going to be used.
@@
 It was concluded that using a buffer of two meters is a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
-Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to the almost identical model results.
+Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
 
@@
 As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of a tree within the pixel grid.
-If a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation would be located in the middle of the plot.
+If a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation is located in the middle of the plot.
 
@@
 However, four out of ten of the most important features of dataset HR are located between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area - especially for plot Oiartzun - which might explain why these features were considered as important by the model.
-
-A possible explanation for the less good performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
+Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area - especially for plot Oiartzun - which might explain why these features were considered important by the model.
+
+A possible explanation for the worse performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
 The majority of VI features covers the range between 550 nm - 800 nm.
@@
 % write about hyperspectral remote sensing
-The field of (hyperspectral) remote sensing has a strong focus on using random forest for modeling in recent years \cite{belgiu2016}.
-However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computational expensive.
+The field of (hyperspectral) remote sensing has a strong focus on using random forests for modeling in recent years \cite{belgiu2016}.
+However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
 To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.

### 2020-06-21 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   dplyr::rename(Plot = iter, RMSE = rmse) %>%
-  dplyr::mutate(Plot = as.character(Plot)) %>%
-  dplyr::mutate(Plot = forcats::fct_recode(Plot,
+  dplyr::mutate(`Test Plot` = as.character(Plot)) %>%
+  dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
     Laukiz1 = "1", Laukiz2 = "4",
@@
     caption = "Test fold performances for learner SVM on the HR dataset without using a filter.
-For each row, the model was trained on observations from fold all other plots and tested on the observations of the given plot.",
+For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot.",
     label = "tab:svm-single-fold-perf"
@@
     type = "latex",
-    caption = "Best ten results for any task/learner/filter combination, sorted ascending by RMSE",
+    caption = "Best ten results among all combinations, sorted in ascending order of RMSE",
     label = "tab:perf-top-10"
@@
     type = "latex",
-    caption = "Lower ten results for any task/learner/filter combination, sorted decreasing by RMSE",
+    caption = "Worst ten results among all combinations, sorted in decreasing order of RMSE",
     label = "tab:perf-worst-10"
@@
   dplyr::rename("Features (\\%)" = fw.perc) %>%
-  dplyr::mutate(Plot = as.character(Plot)) %>%
-  dplyr::mutate(Plot = forcats::fct_recode(Plot,
+  dplyr::mutate(`Test Plot` = as.character(Plot)) %>%
+  dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
     Oiartzun = "2", Luiando = "3",
@@
   dplyr::mutate(Learner = as.character(Learner)) %>%
-  dplyr::mutate(Plot = as.character(Plot)) %>%
+  dplyr::mutate(`Test Plot` = as.character(`Test Plot`)) %>%
   dplyr::mutate("\\#" = case_when(
-    Plot == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
-    Plot == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
-    Plot == "Luiando" ~ ceiling(`Features (\\%)` * 301),
-    Plot == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
+    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
+    `Test Plot` == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
+    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 301),
+    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
   )) %>% 
   dplyr::group_by(Learner) %>%
-  dplyr::select(Learner, Plot, "Features (\\%)", "\\#", RMSE)
+  dplyr::select(Learner, `Test Plot`, "Features (\\%)", "\\#", RMSE)
 

### 2020-06-21 docs/00-manuscripts/ieee/ieee-manuscript.tex
address v4 review from Alex

@@
 	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
-	Feature wise, defoliation at trees (\%) was modeled as a function of band reflectance.
-	Variable importance was assessed using permutation-based feature importance.
-
-	With respect to the used models, support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
-	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between single feature sets were observed.
-	Filter methods helped in certain dataset-learner-filter settings to achieve better predictive performances while ensemble filters did not have a substantial impact on performance.
-	For certain dataset-learner combinations, filter methods helped to improve predictive performance and lower computational runtime.
+	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
+
+	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
+	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
+	For some combinations of learners and feature sets, filter methods achieved better predictive performances than the unfiltered feature sets, while ensemble filters did not have a substantial impact on performance.
 
@@
 
-	The dataset used in this work was suboptimal with respect to the achieved predictive performances and more training data is needed for more generalizable conclusions.
-	Filter methods show potential to be helpful in high-dimensional situations while still keeping the ability to interpret feature effects of fitted models, which is an essential constrain in environmental modeling studies.
+	More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
+	Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.
 
@@
 	\acro{ALE}{accumulated local effects}
-	\acro{ALS}{airborn laser scanning}
+	\acro{ALS}{airborne laser scanning}
 	\acro{ANN}{artificial neural network}
@@
 	\acro{GLM}{generalized linear model}
-	\acro{ICGC}{Institut Cartografic I Geologic de Catalunya}
+	\acro{ICGC}{Institut Cartografic i Geologic de Catalunya}
 	\acro{IQR}{interquartile range}
@@
 Scientists can nowadays relatively easily process large amounts of (environmental) information using various learning algorithms.
-This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would never have been tested otherwise \cite{ma2015}.
+This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would never have been explored otherwise \cite{ma2015}.
 
@@
 The ability of predicting into unknown space qualifies these tools as a promising toolset for such tasks.
-One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing defoliation at trees \cite{hawrylo2018}.
-
-Other approaches for analyzing forest health include temporal change detection \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
+One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing tree defoliation \cite{hawrylo2018}.
+
+Other approaches for analyzing forest health include change detection \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
 In such studies, the defoliation of trees serves as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \cite{townsend2012, goodbody2018}.
@@
 A less known index type which can be derived from spectral information is the \ac{NRI}.
-In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (arbitrary) combinations of spectral bands.
+In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands.
 Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
@@
 
-Despite its popularity in environmental modeling, there are no studies so far which used machine-learning algorithms in combination with remote sensing data to analyze defoliation on a tree level.
-This study aims to close this gap by analyzing defoliation at trees in northern Spain using airborne hyperspectral data.
+Despite their popularity in environmental modeling, there are no studies so far which used machine-learning algorithms in combination with remote sensing data to analyze defoliation at the tree level.
+This study aims to close this gap by analyzing tree defoliation in northern Spain using airborne hyperspectral data.
 The methodology of this study uses machine-learning methods in combination with feature selection and hyperparameter tuning.
 In addition, feature importance and feature effects are evaluated.
-Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high-dimensionality in environmental modeling \cite{trunk1979, xu2016}.
+Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high dimensionality in environmental modeling \cite{trunk1979, xu2016}.
 
 % mention why FS is used / important
-Even though \ac{ML} algorithms are capable of handling highly-correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
-Feature selection approaches can help to address this issue, reducing possible in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
+Even though \ac{ML} algorithms are capable of handling highly correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
+Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
 
 % clarify the focus of this study
-This study shows how high-dimensional datasets can be handled effectively with machine-learning methods while still being conduct inference on the fitted models.
-The predictive power of non-linear methods and their ability to handle highly-correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
+This study shows how high-dimensional datasets can be handled effectively with machine-learning methods while still being able to interpret the fitted models.
+The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
 However, this study clearly focuses on investigating the effects of filter methods and feature set types on predictive performance rather than on interpreting feature effects.
 
-That said, the research questions of this study are the following:
+Considering these opportunities and challenges, the research questions of this study are the following:
 
@@
 
-	\item Do different (environmental) feature sets show differences in performance when modeling defoliation at trees?
+	\item Do different (environmental) feature sets show differences in performance when modeling defoliation of trees?
 
@@
 
-	\item How are feature-selection methods influencing the predictive performance of the models?
+	\item How do different feature selection methods influence the predictive performance of the models?
 
@@
 The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
-In-situ measurements of defoliation at trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
+In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
 It is assumed that the fungi infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
@@
 Oiartzun has the most observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
-All plots besides Luiando are located nearby the coast (\autoref{fig:study_area}).
+All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
 In total 1808 observations are available Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
-The data was surveyed in September 2016.
+Field surveys were conducted in September 2016.
 
@@
 The first four bands were corrupted, leaving 122 bands with valid information.
-Additional metadata information is available in \autoref{tab:hyperparameter_limits}.
+Additional metadata information is available in \autoref{tab:hyperparameter_metadata}.
 
@@
 \end{tabular}
-\endgroup\label{tab:hyperparameter_limits}
+\endgroup\label{tab:hyperparameter_metadata}
 \end{table}
@@
 All of those pixels formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
-In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated due to the first four bands of the sensor being corrupted.
+In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated.
 
@@
 
-The case of a feature-rich dataset comes with several challenges for both model fitting and evaluation.
+High-dimensional, feature-rich datasets come with with several challenges for both model fitting and evaluation.
 
@@
 	\item Model fitting times increase.
-	\item Noise is possibly introduced into models by highly-correlated variables \cite{johnstoneiainm.2009}.
+	\item Noise is possibly introduced into models by highly correlated variables \cite{johnstoneiainm.2009}.
 	\item Model interpretation and prediction become more challenging \cite{johnstoneiainm.2009}.
@@
 % Filter methods
-The concept of filters originates from the idea of ranking features using certain heuristics of an algorithm \cite{guyon2003}.
+The concept of filters originates from the idea of ranking features following a score calculated by an algorithm \cite{guyon2003}.
 Some filter methods can only deal with specific types of variables (numeric or nominal).
@@
 The selection which features to keep for model fitting is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
-Essentially, the number of covariates in the model is treated as a hyperparameter of the model.
-The goal is to optimize the number of features, after ranking was done, to the point at which the model achieves the best performance.
+Essentially, the number of covariates in the model is treated as a additional hyperparameter of the model.
+The goal is to optimize the number of ranked features to the point at which the model achieves the best performance.
 
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' are able to enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' is able to enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
@@
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
-It's final feature order is determined by the sum of all single filters ranks.
+Its final feature order is determined by the sum of all single filters ranks.
 
@@
 where \(H\) is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
-In order to use this method with a numeric response (percentage defoliation at trees), the variable was discretized into equal bins and treated as a class variable.
-After feature rank correlations of $> 0.9$ betwforeen different bin sizes were observed in a side analysis, \texttt{\(n_{bin}\) = 10} was found to be a reasonable setting to go with.
+In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins and treated as a class variable.
+After feature rank correlations of $> 0.9$ between different bin sizes were observed in a side analysis, \texttt{\(n_{bin}\) = 10} was found to be a reasonable setting to go with.
 
@@
 
-Random forest and {SVM} are well established algorithms widely used in (environmental) modeling.
+Random forest and {SVM} are well established algorithms widely used in (environmental) remote sensing.
 Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmark studies in recent years.
@@
 Ridge does not remove variables from the model (penalization to zero) but just shrinks them to effectively zero, keeping them in the model.
-The combination of both penalties is called 'elastic net' but was not used in this work.
 
@@
 
-Three feature sets were used in this study with each representing a different way of feature engineering:
+Three feature sets were used in this study, each representing a different approach to feature engineering:
 
@@
 
-Even though the feature-selection step should be solely left to the filter methods in this study,it was ensured a priori to account for features with a pairwise correlation of 1.
+Even though the feature-selection step should be solely left to the filter methods in this study, it was ensured a priori to account for features with a pairwise correlation of 1.
 Having such features within the data can cause undesired effects during model fitting and feature importance calculation.
-Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of 0.9999999999, the feature with the largest mean absolute correlation across all variables was removed from the dataset.
-This process was repeated $p$ times, each time calculating a fresh correlation matrix.
-
-This preprocessing step reduced the amount of covariates to 122 (HR), 86 (VI) and 7467 (NRI).
+Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of $1 - 10^{-10}$, the feature with the largest mean absolute correlation across all variables was removed from the dataset.
+
+This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).
 
@@
 In this work, an initial design of 30 randomly composed hyperparameter settings in combination with a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
-The advantage of this tuning approach is the substantial reduction of the tuning budget which is required to find a setting close to the global optimization minimum.
+The advantage of this tuning approach is a substantial reduction of the tuning budget that is required to find a setting close to the global optimization minimum.
 \ac{MBO} may outperform methods that do not use information from previous iterations, such as random search or grid search \cite{bergstra2012}.
@@
 For \ac{PCA}, the number of principal components was tuned instead.
-The RF hyperparameter \texttt{\(m_{try}\)} was modified into power transformed \(p^{m_{try}}\) (ranging from 1 to $p$, with \(p\) being the number of features of the dataset) to work on relative values of the respective dataset's feature count.
+The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_try = p^{x_try}$.
+The exponent $x_try$, (ranging from 1 to $p_sel$, where $p_sel$ is the number of features selected with the corresponding filter settings), was optimized between 0 (one randomly selected feature) and 0.5 ($\sqrt{p_sel}$ randomly selected candidate features).
 This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features left after optimizing the feature percentage during tuning.
@@
 
-In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods + 2 $\times$ 6 L1 and L2 models).
+In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
 
@@
 
-Estimating feature importance for datasets with highly-correlated features is a complicated task.
+Estimating feature importance for datasets with highly correlated features is a complicated task for which many different approaches, model-specific and agnostic, exist \cite{friedman2001, hastie2001, greenwell2018}..
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features \cite{molnar2019}.
 Methods like \ac{PDP} or permutation-based approaches may produce unreliable estimates in such scenarios because unrealistic situations between covariates are created \cite{molnar2019}.
-Estimating feature importance for machine learning methods is a complicated procedure for which many different approaches, model-specific and agnostic, exist \cite{friedman2001, hastie2001, greenwell2018}.
-The development of robust methods which enable an unbiased estimation of feature importance for highly-correlated variables are subject to current research.
+The development of robust methods which enable an unbiased estimation of feature importance for highly correlated variables are subject to current research.
 
@@
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
-Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://github.com/pat-s/2019-feature-selection}),listing more side-analyses that were carried out during the creation of this study.
+Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://github.com/pat-s/2019-feature-selection}), listing more side-analyses that were carried out during the creation of this study.
 
@@
 % - MBO tuned penalized methods are substantially worse
-Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 \% percentage points (p.p.).
+Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.).
 SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
 Ridge faced major issues in four tasks due to one observation which was predicted off the response scale (i.e. $> 100$).
-SVM showed the best overall performance with a difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
-Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 for learner SVM (without filter) but up to 54.26 p.p.
+SVM showed the best overall performance with a mean difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
+Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 p.p.
+for learner SVM (without filter) but up to 54.3 p.p.
 when testing on Laukiz2 (\autoref{tab:svm-single-fold-perf}).
@@
 % - SVM unaffected
-When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) when a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
+When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) where a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
 For SVM, all filters and \enquote{no filter} achieved the same performance on tasks VI and NRI even though \autoref{fig:perf-result} lists "No Filter" as the best option.
@@
 % Selected percentage of features across plots
-High differences were observed between the amount of features selected during tuning for the subsequent fitting process.
+% FIXME: Lösung einfügen nach Klärung mit Alex
+Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process.
 Most features were selected during optimization for plot Laukiz2 and least for Laukiz1 (\autoref{tab:tune-perc-sel-features}).
 RF used only one feature for plots Luiando and Oiartzun (0.00004 \% of 7675) while for Laukiz1 34 and for Laukiz2 230 features were seen as optimal during optimization.
-In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features) more than two-third of all available features.
+In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features) more than two-thirds of all available features.
 
@@
 		\caption{Predictive performance (RMSE) of models across tasks.
-Suffix 'CV' denotes that the learner was optimized using internal 10-fold CV while prefix 'MBO' means that Bayesian optimization was used for hyperparameter optimization.
+Suffix 'CV' denotes that the learner was optimized using internal 10-fold CV while prefix 'MBO' means that model-based optimization was used for hyperparameter optimization.
 Abbreviations on the vertical axis refer to the combinations of feature sets on which each model was scored on.
@@
 For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
-For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around one percentage point).
-Overall, most features showed only a small importance with average decreases in RMSE below 0.5 percentage points.
+For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around 1 percentage point).
+Overall, most features showed only a small importance with average decreases in RMSE below 0.5 p.p..
 
@@
 
-The ALE plots show a small relative change compared to their respective mean effects for all chosen features ($< \mid 0.001 \mid$) (\autoref{fig:fi-hr-ale}).
-Most ALE curves show a high dynamic across the value range of each variable, within their respective effect range.
+Most ALE plots show a small absolute change in ALE compared to their respective mean effects (= 0) for all chosen features (less than $\pm 0.001$) (\autoref{fig:fi-hr-ale}).
+Also, a medium-to-high relative dynamic over the respective reflectance value range of each variable was observed.
 
@@
 However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
-Due to the apparent presence of model overfitting in this study it can be postulated that more training data representing all health stages of a plot is needed.
+Due to the apparent presence of model overfitting in this study it can be postulated that more training data representing a greater variety of situations is needed.
 A model can only make robust predictions if it has learned relationships across the whole range of the response.
-Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by suboptimal distributed training data.
-However, when inspecting the fold level performances, it can be concluded that the model did a decent job predicting defoliation ranging from 50\% to 100\% but failed for 0\% - 50\%.
+Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
+However, when inspecting the fold level performances, it can be concluded that the model performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
 This applied to all learners of this study.
@@
 
-An interesting find is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
+An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
 These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
-However, comparing this finding (both relatively and absolute) to other studies would not be fair since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
+However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
 
@@
 One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
-It was therefore left in the model, but it should be acknowledged that these learners seemed particularly sensitive to the input data.
-
-The upper range of 600 used for hyperparameter {\texttt{nrounds}} for model XGBoost in this study could be a potential limiting factor with respect to the performance of the datasets including NRI variables (\autoref{tab:hyperparameter_limits}).
-This setting was a compromise between runtime and tuning space extension aimed to work well for most feature sets.
-It could be that better predictive performances could have been achieved with a {\texttt{nrounds}} upper limit close to the number of features of the NRI datasets.
+It was therefore left in the model, but it should be acknowledged that this learners seemed particularly sensitive to the input data.
+
+A potential limiting factor in this study could be the upper limit of 600 iterations used for the XGBoost algorithm (hyperparameter \texttt{nrounds}), especially for feature sets including NRIs (\autoref{tab:hyperparameter_limits}).
+This setting was a compromise between runtime and tuning space extension with the goal to work well for most feature sets.
+It may be recommendable to increase this upper limit to a value closer to the number of features in the dataset in order to be able to exploit the full spectrum of this hyperparameter.
 
@@
 The large differences in RMSE obtained on different test folds can be attributed to model overfitting (\autoref{tab:svm-single-fold-perf}).
-An RMSE of 54.26 p.p. indicates a complete failure of the model in the prediction stage for this plot (Laukiz2).
+An RMSE of 54.26 p.p. reveals the model's inability to predict tree defoliation on this plot (Laukiz2).
 Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (\autoref{fig:defol-distr}).
-In the prediction scenario for Laukiz2, the model was trained on data containing mostly medium-high defoliation values and only few low ones.
-This caused overfitting on the medium-high values, degrading the model's predictive performance in other scenarios.
-In cases when Laukiz2 was in the training set, the overall aggregated RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
-
-The high differences of selected features per plot plot during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
-While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF is able to achieve the best results with a relatively low amount of features.
-This fact leads to reduced computational runtime, especially if model parameters are dependent on the amount of features (e.g. \texttt{\(m_{try}\)}).
+In the prediction scenario for Laukiz2, the model was trained on data containing mostly medium-to-high defoliation values and only few low ones.
+This caused overfitting on the medium-to-high values, degrading the model's predictive performance in other scenarios.
+When Laukiz2 was in the training set, the overall mean RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).
+
+The large differences of selected features per fold during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
+While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF was able to achieve the best results with a relatively small amount of features.
+Realizing early that few features are needed during tuning to reach adequate performances can reduce the overall computational runtime substantially, especially when iterating on features like \texttt{\(m_{try}\)} which are dependent on the number of features.
 Hence, regardless of the potential advantage of using filters for increased predictive performance, it should be noted that these can have a strong positive effect on runtime, at least for RF in this study.
 
-Ultimately, the results of (\autoref{tab:tune-perc-sel-features}) should be taken with care as they rely on single model-filter combinations and are subject to random variation.
+Ultimately, the results of \autoref{tab:tune-perc-sel-features} should be taken with care as they rely on single model-filter combinations and are subject to random variation.
 More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criteria optimization problem.
@@
 The usefulness of filters with respect to predictive performance in this study varied.
-While some filters were able to enhance model performances (up to 5 p.p. for RF and XGBoost), some caused a worse performance than using no filter at all (\autoref{fig:filter-effects-no-filter}).
+While the performance of some models (up to 5 p.p.
+for RF and XGBoost) was improved by specific filters, some models achieved a poorer performance with filters than without them (\autoref{fig:filter-effects-no-filter}).
 Since these negative cases were not caused by a specific filter method, it is recommended to test multiple filters in a study if filters are going to be used.
-While filters can improve the performance of models, they might be of more interesting in other aspects than performance: reducing variables can reduce computational efforts in high-dimensional scenarios and might enhance the interpretability of models.
-Here, filters are a lot cheaper than wrapper methods due to their ability of being treated as a hyperparameter during the optimization stage.
+While filters can improve the performance of models, they might be more interesting in other aspects than performance: reducing variables can reduce computational efforts in high-dimensional scenarios and might enhance the interpretability of models.
+Filters are a lot cheaper to compute than wrapper methods and the final feature subset selection can be added as an additional hyperparameter to the model optimization stage.
 
@@
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
-Therefore no general statement can be made from the results of this study if the use of ensemble filters leads to a substantial better performance than applying one of the various single filter methods.
-For this, more case studies applying ensemble filter methods are needed.
-In any case, ensemble filters can be a promising addition to a machine-learning feature-selection portfolio.
+In summary, Borda performs no better than a randomly selected filter method in this study.
+More case studies applying ensemble filter methods are needed to verify this finding.
+Nevertheless, ensemble filters can be a promising addition to a machine-learning feature-selection portfolio.
 
@@
 PCA used as a filter in this work did neither show the best nor worst result compared to other feature-selection approaches.
-It was able to reduce model fitting times substantially due to its nature of reducing the number of predictors to a minimum.
+It was able to reduce model fitting times substantially.
 However, it also removed the possibility to conduct a post-hoc analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
@@
 
-\subsection{Quality of the data}
-
+\subsection{Linking feature importance to spectral characteristics}
+
+Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 680 nm to 750 nm.
+
+This area has the highest ability to distinguish between reflectances related to a high density / high foliage density und thus the health status of vegetation and its respective counterpart \cite{horler1983}.
+However, four out of ten of the most important features of dataset HR are located between 920 nm and 1000 nm.
+Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area - especially for plot Oiartzun - which might explain why these features were considered important by the model.
+
+A possible explanation for the worse performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
+The majority of VI features covers the range between 550 nm - 800 nm.
+Only one index (PWI) covers information in the range beyond 900 nm.
+
+% ALE plots
+The ALE plots are hard to interpret because their values are interpreted on a relative scale and differ for each band.
+When linking band reflectance values against ALE values of a specific band, a value of 100 has a very different meaning in B67 than in B124 because for the latter, reflectance values of vegetation are naturally higher.
+Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
+However, no baseline reflectance value for every feature exists since the absolute reflectance values depend on the sensor characteristics.
+ALE plots can be a powerful tool for interpreting feature importance of correlated features but might be limited in their interpretability because only relative comparisons can be conducted.
+
+\subsection{Data quality}
+
+Environmental datasets always come with some constraints that can have potential influence on the modeling process and its outcome.
+The following paragraph discusses these.
 The decision to use a buffer of two meters for the extraction of reflectance values was a complex process.
@@
 
-Another point worth discussing is that the exact number of contributing pixels to the final index value of an observation cannot be exactly determined: it depends on the location of the tree within the pixel grid.
+Another point worth discussing is that the exact number of contributing pixels to the final index value of an observation cannot be determined precisely: it depends on the location of the tree within the pixel grid.
 According to the extract function of the raster package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
@@
 Such limitations and uncertainties apply to most environmental studies and cannot be completely avoided.
-
-\subsection{Linking feature importance to spectral characteristics}
-
-Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 680 nm to 750 nm.
-
-This area has the highest ability to distinguish between reflectances related to a high density / high healthiness of vegetation and its respective counterpart \cite{horler1983}.
-However, four out of ten of the most important features of dataset HR are located between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area - especially for plot Oiartzun - which might explain why these features were considered important by the model.
-
-A possible explanation for the worse performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
-The majority of VI features covers the range between 550 nm - 800 nm.
-Only one index (PWI) covers information in the range beyond 900 nm.
-
-% ALE plots
-The ALE plots are hard to interpret due to their semi-meaningful absolute values which differ for each band.
-For example, a reflectance value of 100 has a very different meaning in B67 than in B124 because for the latter, reflectance values of vegetation are naturally higher.
-Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
-However, no baseline reflectance value for every feature exists since the absolute reflectance values depend on the sensor characteristics.
-Hence, it is hard to make reasonable interpretations of harsh drops/jumps within curves of certain features, e.g. the one at value 240 for feature B115.
-ALE plots can be a powerful tool for interpreting feature importance of correlated features but might be limited in interpretability if the absolute feature values are only semi-meaningful.
 
@@
 
-Most other studies analyzing defoliation operated on the plot rather than the tree level.
-This is due to the low spatial resolution of used satellite products which served as the input data, making a tree-level study infeasible \cite{townsend2012, debeurs2008, rengarajan2016}.
+While most defoliation studies operate on the plot level using coarser-resolution multispectral satellite data \cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
+Among these, \cite{meng2018, kalin2019} used ground-level methods such as \ac{ALS} or \ac{LiDAR}.
 
@@
 \cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
-Study results indicated that spectral metrics were found to be most helpful for the model.
-Incorporating such metrics (both spectral and structural) could be a possible enhancement for future works.
+Study results indicated that spectral features were found to be most helpful for the model.
+Incorporating such (both spectral and structural) could be a possible enhancement for future works.
 
@@
 However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
-To account for this and the high-dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
-However, no study which made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
-Potential reasons for this gap could be an easier access of applying wrappers methods and a higher general awareness of such compared to filter methods.
-Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for upcoming modeling studies using hyperspectral remote sensing data.
+To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
+However, no study that made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
+Potential reasons for this gap could be an easier access to wrappers methods and a higher general awareness of such compared to filter methods.
+Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.
 
 % RS studies with more/other algorithms than RF
-When looking for remote sensing studies which compare multiple models, it turned out that such often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
+When looking for remote sensing studies that compare multiple models, it turned out that these often operate in a low-dimensional predictor space \cite{xu2019} or use wrapper methods explicitly \cite{georganos2018}.
 
@@
 \cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a random forest classifier.
-
-In summary, no studies which used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to forest health were found.
-This might relate to the fact that most environmental/ecological datasets are not high-dimensional.
-In fact, the number of predictors is often less than ten and issues related to correlations are often solved manually instead of relying on an automated approach.
-
-The bioinformatics field faces high-dimensional datasets more often.
+\cite{zandler2015} shows a similar setup: they used hyperspectral vegetation indices, a nested CV approach for performance estimation and estimated variable importance targeting woody biomass as the response.
+In the results, lasso showed the best performance among the chosen methods.
+However, the authors did not optimize the hyperparameters of RF which makes a fair comparison problematic since the other models used perform internal optimization.
+The discussion section of \cite{zandler2015} lists additional studies that made use of shrinkage models for high dimensional remote sensing modeling.
+
+In summary, no studies which used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to tree health were found.
+This might relate to the fact that most environmental datasets are not high-dimensional.
+In fact, many studies use fewer than ten features and issues related to correlations are often solved manually instead of relying on an automated approach.
+This can be subjective and may limit the reproducibility of results.
+
+Other fields (e.g. bioinformatics) face high-dimensional datasets more often.
 Hence more studies using (filter-based) feature-selection approaches can be found for this field \cite{guo2019, radovic2017}.
-Yet bioinformatics differs conceptually in many ways from environmental modeling and digging deeper just for the sake of finding any similar studies is not a good idea.
-If a field only rarely faces high-dimensional dataset issues (like environmental modeling) the motivation and expertise of using advanced methods to solve such are rather low.
+Yet bioinformatics differs conceptually in many ways from environmental modeling and therefore no greater focus was put into comparing studies of this field.
+The availability of high dimensional feature sets will increase in the future due to higher temporal and spectral resolutions of sensors.
+In addition, a high spatial resolution comes with the possibility to calculate many textural features.
+Hence, the ability to deal with high dimensional datasets becomes more important and unbiased robust approaches are needed.
 We hope that this work and its methodology raises awareness about the application of filter methods to tackle high-dimensional problems in the environmental modeling field.
 
-\section{Outlook and conclusion}
-
-This study analyzed defoliation at trees in northern Spain by using hyperspectral data as input for machine-learning models which used hyperparameter tuning and filter-based feature selection.
+\section{Conclusion}
+
+This study analyzed defoliation of trees in northern Spain by using hyperspectral data as input for machine-learning models which used hyperparameter tuning and filter-based feature selection.
 Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
@@
 % RQ3: How are feature-selection methods influencing the predictive performance of the models?
-Filter methods were able to improve the predictive performance on datasets in some instances.
+Filter methods were able to improve the predictive performance on datasets in some instances, although there was no clear and systematic pattern.
 Their effectiveness depends on the algorithm and the dataset characteristics.
@@
 dataset VI) showed a worse performance than the ones which covered wider range (400 nm - 1000 nm; HR, NRI).
-NRIs can be seen as a valuable addition for optimizing predictive performance in vegetation related studies.
+NRIs can be seen as a valuable addition for optimizing predictive performance in remote sensing of vegetation.
 
@@
 Features along the red edge wavelength region were most important for models during prediction.
-With respect to dedicated vegetation indices, the "Vogelmann" index with all of its derivatives was seen as the most important index for the best performing SVM model.
+With respect to dedicated vegetation indices, all versions of the Vogelmann index were seen as the most important index for the best performing SVM model.
 This matches well with the actual purpose of these indices:
-These were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) in 1988 \cite{vogelmann1993}.
-However, feature importance for highly-correlated features remains a challenging task.
-Results might be biased and should be taken with care.
+These were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel}) \cite{vogelmann1993}.
+However, assessing the feature importance for highly correlated features remains a challenging task.
+Results might be biased and should be taken with care to avoid overgeneralizing from individual studies.
 
 % What about the general potential of analyzing defol with hyperspectral data?
-The potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 27 percentage points scored by the best performing model.
+Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 27 percentage points scored by the best performing model.
 More training data covering a wider range of defoliation values in a larger number of forest plantations is needed to train better models which can create more robust predictions.
@@
 	\toprule
-	\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$m_{try}.power$}    & int & 0         & 0.5      & -       \\
+	\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
 	 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
@@
 
+\pagebreak
+
+% The photo area is 1 in wide and 1.25 in long. The IEEE recommends that author photo images should be of 220 dpi (dots per inch) resolution and in gray scale with 8 bits/sample.
+\begin{IEEEbiography}
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/pschratz.jpeg}}]{Patrick Schratz}
+%	[]{Patrick Schratz}
+	is a Data Scientist currently working in Zurich, CH.
+    His area of expertise is applied machine learning, more specifically the field of environmental modeling.
+    He is also a PhD candidate at the GIScience group, Department of Geography at University of Jena where he acts as a researcher in the environmental modeling field.
+    At cynkra GmbH, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.
+\end{IEEEbiography}
+\begin{IEEEbiography}
+	[]{Jannes Muenchow}
+	Biography text here.
+\end{IEEEbiography}
+\begin{IEEEbiography}
+	[]{Eugenia Iturritxa}
+	Biography text here.
+\end{IEEEbiography}
+\begin{IEEEbiography}
+	[]{José Cortés}
+	Biography text here.
+\end{IEEEbiography}
+\begin{IEEEbiography}
+	[]{Bernd Bischl}
+	Biography text here.
+\end{IEEEbiography}
+\begin{IEEEbiography}
+	[]{Alexander Brenning}
+	Biography text here.
+\end{IEEEbiography}
+
 \end{document}

### 2020-07-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
add bio and pictures

@@
 %	[]{Patrick Schratz}
-	is a Data Scientist currently working in Zurich, CH.
+	is a Data Scientist working in Zurich, CH.
     His area of expertise is applied machine learning, more specifically the field of environmental modeling.
-    He is also a PhD candidate at the GIScience group, Department of Geography at University of Jena where he acts as a researcher in the environmental modeling field.
-    At cynkra GmbH, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.
+    He is a PhD candidate at the GIScience group, Department of Geography at University of Jena where he acts as a researcher in the environmental modeling field.
+    At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.
 \end{IEEEbiography}
@@
 \begin{IEEEbiography}
-	[]{Eugenia Iturritxa}
-	Biography text here.
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/eiturritxa.jpg}}]{Eugenia Iturritxa}
+	received the Ecology and Ph.D degree in Plant Protection from the Basque country University in 2001.
+	Since 1999 her main research has focused on forest health, on the study of diverse species of native and introduced pathogenic fungi in forests and forest plantations.
+	Her research includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.
 \end{IEEEbiography}
@@
 	[]{José Cortés}
-	Biography text here.
+	received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
+	Currently he is a PhD student at Friedrich-Schiller-University, Jena, Germany.
+	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC, a joint program with the Max Planck Institute for Biogeochemistry.
+	His research focuses on spatio-temporal trend detection in environmental data.
 \end{IEEEbiography}
 \begin{IEEEbiography}
-	[]{Bernd Bischl}
-	Biography text here.
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/bbischl.jpg}}]{Bernd Bischl}
+	obtained his Ph.D from Dortmund Technical University in 2013.
+	He is a professor for "Statistical Learning and Data Science" at the Department of Statistics at the Ludwig-Maximilians-University Munich and a director of the "Munich Center for Machine Learning".
+	His research interests include AutoML, model selection, interpretable ML and XAI.
 \end{IEEEbiography}
 \begin{IEEEbiography}
-	[]{Alexander Brenning}
-	Biography text here.
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/abrenning.jpg}}]{Alexander Brenning}
+	(Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
+	He served as an assistant professor and tenured associate professor in geomatics at the University from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
+	Dr. Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.
+	He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.
 \end{IEEEbiography}

### 2020-07-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
add jannes pic and bio

@@
 \begin{IEEEbiography}
-	[]{Jannes Muenchow}
-	Biography text here.
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/jmuenchow.jpg}}]{Jannes Muenchow}
+	is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
+	He has earned his PhD at the Friedrich-Alexander University Erlangen-Nuremberg (Germany) in 2013.
+	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia in 2016 (Friedrich Schiller University Jena).
+	Currently, his research focuses on developing open source tools for ecology, geomorphology and qualitative GIS.
+	He is a co-author of the book "Geocomputation with R".
 \end{IEEEbiography}

### 2020-07-21 docs/00-manuscripts/ieee/ieee-manuscript.tex
more small details after reviews

@@
 For \ac{PCA}, the number of principal components was tuned instead.
-The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_try = p^{x_try}$.
-The exponent $x_try$, (ranging from 1 to $p_sel$, where $p_sel$ is the number of features selected with the corresponding filter settings), was optimized between 0 (one randomly selected feature) and 0.5 ($\sqrt{p_sel}$ randomly selected candidate features).
+%The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_{try} = p^x_{try}$.
+The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_\textrm{try} = p_\textrm{sel}^t$ as a function of the number of selected features, $p_\textrm{sel}$.
+It was thus tuned on a logarithmic scale by varying $t$ between 0 (i.e. $m_\textrm{try} = 1$) and 0.5 (i.e. $m_\textrm{try}=\sqrt{p_\textrm{sel}}$).
 This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features left after optimizing the feature percentage during tuning.
@@
 Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process.
-Most features were selected during optimization for plot Laukiz2 and least for Laukiz1 (\autoref{tab:tune-perc-sel-features}).
-RF used only one feature for plots Luiando and Oiartzun (0.00004 \% of 7675) while for Laukiz1 34 and for Laukiz2 230 features were seen as optimal during optimization.
-In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features) more than two-thirds of all available features.
+Most features were selected during optimization when training without Laukiz2 and least when Laukiz1 served as the test set (\autoref{tab:tune-perc-sel-features}).
+RF used only one feature when Luiando and Oiartzun served as the test data (0.00004 \% of 7675) while 34 and 230 features were seen as optimal during optimization when Laukiz1 and Laukiz2 were the test data, respectively.
+In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features; test set) more than two-thirds of all available features.
 
@@
 While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF was able to achieve the best results with a relatively small amount of features.
-Realizing early that few features are needed during tuning to reach adequate performances can reduce the overall computational runtime substantially, especially when iterating on features like \texttt{\(m_{try}\)} which are dependent on the number of features.
+Realizing early that few features are needed during tuning to reach adequate performances can reduce the overall computational runtime substantially, especially when iterating over parameters such as $m_\textrm{try}$ whose optimum (and range) depends on the number of features.
 Hence, regardless of the potential advantage of using filters for increased predictive performance, it should be noted that these can have a strong positive effect on runtime, at least for RF in this study.
@@
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
-PCA used as a filter in this work did neither show the best nor worst result compared to other feature-selection approaches.
+PCA, acting as a filter method in this work, did neither show the best nor worst result compared to other feature-selection approaches.
 It was able to reduce model fitting times substantially.
-However, it also removed the possibility to conduct a post-hoc analysis of feature importance by combining predictors into principal components, forcing users to decide upfront what they value more.
-Since filter scores only need to be calculated once in a benchmark setting, the runtime advantage might in fact be negligible in practice.
+Depending on the use case, PCA can be an interesting option to reduce dimensionality while keeping runtime low.
+However, information about the total number of features used by the model is lost when applying this technique.
+Since filter scores only need to be calculated once for a given dataset in a benchmark setting, the runtime advantage of a PCA vs. filter methods might in fact be negligible in practice.
 

### 2020-08-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
add joses photo
update latin names of fungi after eugenias review

@@
 The ability of predicting into unknown space qualifies these tools as a promising toolset for such tasks.
-One aspect of this research field is to enhance the understanding of biotic and abiotic triggers, for example by analyzing tree defoliation \cite{hawrylo2018}.
+One aspect of this research field is to enhance the understanding of biotic and abiotic stress triggers, for example by analyzing tree defoliation \cite{hawrylo2018}.
 
@@
 
-Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata}) plantations in northern Spain.
-The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea}, \textit{Fusarium circinatum}, \textit{Armillaria mellea} or \textit{Heterobasidion annosum}, leading to a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
+Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata D. Don}) plantations in northern Spain.
+The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P.
+Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.} and \textit{Dothisthroma septosporum (Dorogin) M.
+Morelet} causing a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
+The last two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
 In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
@@
 \begin{IEEEbiography}
-	[]{José Cortés}
+	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/jcortes.png}}]{José Cortés}
 	received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.

### 2020-08-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
incorportate Joses comments

@@
 \begin{equation}
-	NRI_{i,j} = \frac{b_{i} - b_{j}}{b_{i} + b_{j}}
+	NRI_{i,j} = \frac{band_{i} - band_{j}}{band_{i} + band_{j}}
 \end{equation}
@@
 \noindent
-where \(i\) and \(j\) are the respective band numbers.
+with \(i\) and \(j\) being the respective band numbers.
 
@@
 
-Estimating feature importance for datasets with highly correlated features is a complicated task for which many different approaches, model-specific and agnostic, exist \cite{friedman2001, hastie2001, greenwell2018}..
+Estimating feature importance for datasets with highly correlated features is a complicated task for which many different approaches, model-specific and agnostic, exist \cite{friedman2001, hastie2001, greenwell2018}.
 The correlation between covariates makes it challenging to calculate an unbiased estimate for single features \cite{molnar2019}.
@@
 However, no study that made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
-Potential reasons for this gap could be an easier access to wrappers methods and a higher general awareness of such compared to filter methods.
+Potential reasons for this gap could be an easier access to wrapper methods and a higher general awareness of such compared to filter methods.
 Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.
@@
 
-\section*{References}
-
 \pagebreak
@@
 	Currently he is a PhD student at Friedrich-Schiller-University, Jena, Germany.
-	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC, a joint program with the Max Planck Institute for Biogeochemistry.
+	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
 	His research focuses on spatio-temporal trend detection in environmental data.

### 2020-08-05 analysis/about.Rmd
update about.Rmd

--- 2020-02-24 analysis/about.Rmd
@@
 
-    - high-dimensional data handling
+  - hyperspectral imagery
+  - forest health monitoring
+  - machine learning
   - feature selection
-    - machine-learning
-    - ecology
+  - feature effects
+  - model comparison
+  - filter
+  - imaging spectroscopy
 
@@
 
-[![Build Status](https://travis-ci.org/pat-s/paper_hyperspectral.svg?branch=master)](https://travis-ci.org/pat-s/paper_hyperspectral)
-
 This repository is organized as an R package, providing functions and raw data to reproduce and extend the analysis reported in the publication.
@@
 
-The R library of this project is managed by [packrat](https://rstudio.github.io/packrat/).
+The R library of this project is managed by [renv](https://rstudio.github.io/renv/index.html).
 This makes sure that the exact same package versions are used when recreating the project.
-When calling `packrat::restore()`, all required packages will be installed with their specific version.
+When calling `renv::restore()`, all required packages will be installed with their specific version.
 
-Please note that this project was built with R version 3.5.2 on a Debian 9 operating system.
-The packrat packages from this project **are not compatible with R versions prior version 3.5.0.**
-For reproducibility, it is recommended to replicate the analysis using the included Dockerfile.
-Instructions can be found [ħere](https://github.com/pat-s/paper_hyperspectral#docker).
-(In general, it should be possible to reproduce the analysis on any other operating system.)
+Please note that this project was built with R version 3.6.1 on a CentOS 7 operating system.
+Some packages from this project **are not compatible with R versions prior version 3.6.0.**
 
@@
 ```sh
-git clone git@github.com:pat-s/paper_hyperspectral.git
+git clone https://github.com/pat-s/2019-feature-selection.git
 ```
@@
 ```{r README-1, eval = FALSE}
-packrat::restore()
+renv::restore()
 r_make()
@@
 
-* The [issues tracker](https://github.com/pat-s/paper_hyperspectral/issues) is the place to report problems or ask questions
-
-* See the repository [history](https://github.com/pat-s/paper_hyperspectral/commits/master) for a fine-grained view of progress and changes.
-
 * The organisation of this compendium is based on the work of [Carl Boettiger](http://www.carlboettiger.info/) and [Ben Marwick](https://github.com/benmarwick).
-
-# Literature
-
-- What to do first (Feature selection or tuning): https://datascience.stackexchange.com/a/19778/24475
-- Hastie, T., Friedman, J., & Tibshirani, R. (2001). The Elements of Statistical Learning. Springer New York. https://doi.org/10.1007/978-0-387-21606-5 (Kapitel 16)
-- FSel algs overview: http://featureselection.asu.edu/algorithms.php
-- De Jay, N., Papillon-Cavanagh, S., Olsen, C., El-Hachem, N., Bontempi, G., & Haibe-Kains, B.
-(2013).
-mRMRe: an R package for parallelized mRMR ensemble feature selection.
-Bioinformatics, 29(18), 2365–2368.
-https://doi.org/10/f48vxc
-- Drotár, P., Gazda, J., & Smékal, Z.
-(2015).
-An experimental comparison of feature selection methods on two-class biomedical datasets.
-Computers in Biology and Medicine, 66, 1–10.
-https://doi.org/10/f72ttc
-- Drotár, P., Šimoňák, S., Pietriková, E., Chovanec, M., Chovancová, E., Ádám, N., … Biňas, M.
-(2017).
-Comparison of Filter Techniques for Two-Step Feature Selection.
-COMPUTING AND INFORMATICS, 36(3), 597–617.
-https://doi.org/10/gbntp7
-- Drotár, P., Gazda, M., & Gazda, J.
-(2017).
-Heterogeneous ensemble feature selection based on weighted Borda count.
-In 2017 9th International Conference on Information Technology and Electrical Engineering (ICITEE) (pp.
-1–4).
-https://doi.org/10.1109/ICITEED.2017.8250495
-- Nie, F., Huang, H., Cai, X., & Ding, C.
-H.
-(2010).
-Efficient and Robust Feature Selection via Joint \mathscrl2,1-Norms Minimization.
-In J.
-D.
-Lafferty, C.
-K.
-I.
-Williams, J.
-Shawe-Taylor, R.
-S.
-Zemel, & A.
-Culotta (Eds.), Advances in Neural Information Processing Systems 23 (pp.
-1813–1821).
-Curran Associates, Inc.
-Retrieved from http://papers.nips.cc/paper/3988-efficient-and-robust-feature-selection-via-joint-l21-norms-minimization.pdf

### 2020-08-05 analysis/about.Rmd
workflowr::wflow_publish("analysis/about.Rmd")

@@
 
-- Using machine-learning algorithms to model defoliation of _Pinus Radiata_ trees.
-
 - keywords
@@
   - imaging spectroscopy
+
+- Using machine-learning algorithms to model defoliation of _Pinus Radiata_ trees.
 
@@
 
-Target `defoliation_maps_wfr` builds are targets necessary for the final results [report](https://2019-feature-selection.pjs-web.de/report-defoliation.html).
+Target `defoliation_maps_wfr` builds subsequent argets which are necessary for the final results [report](https://pat-s.github.io/2019-feature-selection/report-defoliation.html).
 
@@
 See the [`code`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis) directory on GitHub for the source code that generated the figures and statistical results contained in the manuscript.
-See the [`data`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis/data) directory for instructions how to access the raw data discussed in the manuscript.
+See the [`data`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis/data) directory for instructions how to access the raw data used in the manuscript.
 
@@
 This repository is organized as an R package, providing functions and raw data to reproduce and extend the analysis reported in the publication.
-Note that this package has been written explicitly for this project and may not be suitable for more general use.
+Note that this package has been written explicitly for this project and is not suited a for more general use.
 
@@
 
-* The organisation of this compendium is based on the work of [Carl Boettiger](http://www.carlboettiger.info/) and [Ben Marwick](https://github.com/benmarwick).
+* The organisation of this compendium was inspired by the works of [Carl Boettiger](http://www.carlboettiger.info/) and [Ben Marwick](https://github.com/benmarwick).

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
add abbreviation comment
replace goes in line with coincide

@@
 	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
-
+% JM: I wouldn't indicate abbreviations in the abstract if you don't use them anymore within the abstract
 	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
@@
 \IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
-This goes in line with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
+Naturally, this coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
 At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from the open-source community.

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
replace relatively easily
add missing parenthesis

@@
 At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from the open-source community.
-Scientists can nowadays relatively easily process large amounts of (environmental) information using various learning algorithms.
+Scientists can nowadays process large amounts of (environmental) information with relative ease using various learning algorithms.
 This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would never have been explored otherwise \cite{ma2015}.
@@
 All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
-In total 1808 observations are available Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
+In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
 Field surveys were conducted in September 2016.

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
slightly rewrite flight campaign sentence
delete redundant with

@@
 
-The airborne hyperspectral data was acquired during two flight campaigns on September 28th and October 5th 2016, both around noon.
+The airborne hyperspectral data was acquired during two flight campaigns which took at noon on September 28th and October 5th 2016.
 The images were taken by an AISAEAGLE-II sensor.
@@
 
-High-dimensional, feature-rich datasets come with with several challenges for both model fitting and evaluation.
+High-dimensional, feature-rich datasets come with several challenges for both model fitting and evaluation.
 

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
I guess this is meant

@@
 In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
-Due to the focus on filter methods, only this sub-group of feature selection methods will be introduced in greater detail in the following subsections.
+ Due to the focus on filter methods in this manuscript, only this sub-group of feature selection methods will be introduced in greater detail in the following subsections.
 

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
replace is able to by can

@@
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' is able to enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
add Borda comment
add four
delete redundant words

@@
 Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
+% JM: I have no idea what the Borda ensemble filter is. Maybe add one more sentence on this specific filter
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
@@
 % Ensuring a fair weighting in the ensemble
-Filter methods can be grouped into classes: Correlation-based, entropy-based, linear and non-linear methods.
+Filter methods can be grouped into four classes: Correlation-based, entropy-based, linear and non-linear methods.
 Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
@@
 If filter pairs showed a correlation of 0.9 or higher, only one of the two was included into the ensemble filter, selected at random.
-By this it was ensured that the ensemble filter composition was not biased towards a certain group of filter methods.
+This ensured that the ensemble filter composition was not biased towards a certain group of filter methods.
 

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
add comment for Pat
slightly rewrite spectro-temporal setting sentence

@@
 % Ensuring a fair weighting in the ensemble
+% JM: reading further on, it might be that there are only three classes, please refine, Patrick!
 Filter methods can be grouped into four classes: Correlation-based, entropy-based, linear and non-linear methods.
@@
 
-The idea of splitting the features into different sets originated from the question whether feature-engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated previously in a spectro-temporal setting \cite{pena2017}.
+The idea of splitting the features into different sets originated from the question whether feature-engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated by \cite{pena2017} within a spectro-temporal setting.
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
add comment
delete superfluous s

@@
 The idea of splitting the features into different sets originated from the question whether feature-engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated by \cite{pena2017} within a spectro-temporal setting.
+% JM: just a little reminder to avoid sentences where the main verb appears long after the sentence has started
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
@@
 The same general conclusion applies to ridge if one discards the problematic results on four tasks.
-Regarding the problematic predictions of the ridge learner, a careful inspection of the hyperparameter optimization procedure of the fitted models did not reveal any apparent implementation mistakes.
+Regarding the problematic predictions of the ridge learner, a careful inspection of the hyperparameter optimization procedure of the fitted models ensured again that everything was correctly implemented.
 One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
-It was therefore left in the model, but it should be acknowledged that this learners seemed particularly sensitive to the input data.
+It was therefore left in the model, but it should be acknowledged that this learner seemed particularly sensitive to the input data.
 
@@
 This setting was a compromise between runtime and tuning space extension with the goal to work well for most feature sets.
-It may be recommendable to increase this upper limit to a value closer to the number of features in the dataset in order to be able to exploit the full spectrum of this hyperparameter.
+It may be recommendable to increase this upper limit to a value closer to the number of features in the dataset in order to be able to exploit the full range of this hyperparameter.
 
@@
 
-One objective which this study aimed to answer was whether expert-based or data-driven feature engineering has a positive influence on model performance.
+One objective of ths study was whether expert-based or data-driven feature engineering has a positive influence on model performance.
 With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.

### 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
rewrite using positive language

--- 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 The same general conclusion applies to ridge if one discards the problematic results on four tasks.
-Regarding the problematic predictions of the ridge learner, a careful inspection of the hyperparameter optimization procedure of the fitted models did not reveal any apparent implementation mistakes.
+Regarding the problematic predictions of the ridge learner, a careful inspection of the hyperparameter optimization procedure of the fitted models ensured again that everything was correctly implemented.
 One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.

### 2020-08-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
simplify
Update docs/00-manuscripts/ieee/ieee-manuscript.tex

--- 2020-08-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 % JM: reading further on, it might be that there are only three classes, please refine, Patrick!
-Filter methods can be grouped into four classes: Correlation-based, entropy-based, linear and non-linear methods.
+Filter methods can be grouped into groups which are formed out of three binary classes: multivariate or univariate feature use, correlation or entropy-based importance weighting and linear and non-linear filter methodology.
 Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the final ranking result will be biased.

### 2020-08-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
Apply suggestions from code review

@@
 
-The airborne hyperspectral data was acquired during two flight campaigns which took at noon on September 28th and October 5th 2016.
+The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5th 2016.
 The images were taken by an AISAEAGLE-II sensor.
@@
 
-The idea of splitting the features into different sets originated from the question whether feature-engineered indices from reflectance values have a positive effect on model performance, as for example demonstrated by \cite{pena2017} within a spectro-temporal setting.
+The idea of splitting the features into different sets originated from the question whether feature-engineered indices derived from reflectance values have a positive effect on model performance.
+\cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
 % JM: just a little reminder to avoid sentences where the main verb appears long after the sentence has started
@@
 This setting was a compromise between runtime and tuning space extension with the goal to work well for most feature sets.
-It may be recommendable to increase this upper limit to a value closer to the number of features in the dataset in order to be able to exploit the full range of this hyperparameter.
+It may be recommendable to increase this upper limit to a value closer to the number of features in the dataset in order to be able to exploit the full potential of this hyperparameter.
 
@@
 
-One objective of ths study was whether expert-based or data-driven feature engineering has a positive influence on model performance.
+One objective of this study was whether expert-based or data-driven feature engineering has a positive influence on model performance.
 With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.

### 2020-08-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove trailing komma in table

@@
 \centering
-\caption[t]{List of filter methods used in this work}
+\caption{List of filter methods used in this work}
+\label{tab:filter-methods}
 \begingroup\footnotesize
@@
 		Carscore                                     & multivariate, linear, correlation & \cite{zuber2011}   \\
-		Relief                                       & multivariate, linear, entropy,    & \cite{kira1992}    \\
+		Relief                                       & multivariate, linear, entropy    & \cite{kira1992}    \\
 		Conditional minimal information maximization & multivariate, linear, entropy     & \cite{fleuret2004}
@@
 \end{adjustbox}
-\endgroup\label{tab:filter-methods}
+\endgroup
 \end{table}

### 2020-08-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove jannes inline comments
fix table width of tuning ranges table using tabularx
update keywords

@@
 
+\usepackage{tabularx}
 \usepackage{booktabs}
@@
 	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
-% JM: I wouldn't indicate abbreviations in the abstract if you don't use them anymore within the abstract
 	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
@@
 \begin{IEEEkeywords}
-	hyperspectral imagery, forest health monitoring, machine learning, feature selection, feature effects, model comparison, filter, imaging spectroscopy
+	hyperspectral imagery, forest health monitoring, machine learning, feature selection, feature effects, model comparison
 \end{IEEEkeywords}
@@
 Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
-% JM: I have no idea what the Borda ensemble filter is. Maybe add one more sentence on this specific filter
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
@@
 % Ensuring a fair weighting in the ensemble
-% JM: reading further on, it might be that there are only three classes, please refine, Patrick!
 Filter methods can be grouped into groups which are formed out of three binary classes: multivariate or univariate feature use, correlation or entropy-based importance weighting and linear and non-linear filter methodology.
@@
 \cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
-% JM: just a little reminder to avoid sentences where the main verb appears long after the sentence has started
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
@@
 \begingroup\footnotesize
-\begin{tabular}{llllrrr}
+\begin{tabularx}{.5\textwidth}{llXXrXX}
 	\\
@@
 	\bottomrule
-\end{tabular}
+\end{tabularx}
 \endgroup

### 2020-12-10 docs/00-manuscripts/ieee/ieee-manuscript.tex
add funding info
add table with all vegindex references

@@
 
+\usepackage{xtab}
+
 \usepackage{csquotes}
@@
 	%\setlength{\itemsep}{-\parsep}
-
 	\acro{AGB}{above-ground biomass}
@@
 
-\begin{figure} [t!]
-	\begin{center}
-		\centering
-		\includegraphics[width=0.48\textwidth] {study-area-hyperspectral.pdf}
-		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}\label{fig:study_area}
-	\end{center}
-\end{figure}
+% \begin{figure} [t!]
+% 	\begin{center}
+% 		\centering
+% 		\includegraphics[width=0.48\textwidth] {study-area-hyperspectral.pdf}
+% 		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}\label{fig:study_area}
+% 	\end{center}
+% \end{figure}
 
@@
 
+\section{Acknowledgments}
+This work was funded by the EU LIFE Healthy Forest project (LIFE14 ENV/ES/000179) and the German Scholars Organization/Carl Zeiss Foundation.
+
 \appendices{}
@@
 		Hyperparameter notations from the respective R packages were used.}
-\begingroup\footnotesize
+	\begingroup\scriptsize
 	\begin{tabularx}{.5\textwidth}{llXXrXX}
@@
 \end{table}
+
+\begingroup\scriptsize
+\setlength\tabcolsep{4pt}  % default value: 6pt
+\begin{xtabular*}{0.3\textwidth}{lll}
+	\bfseries{Name} & \bfseries{Formula}                                                                                  & \bfseries{Reference*}      \\
+	Boochs          & $D_{703}$                                                                                           & \cite{boochs1990}       \\
+	Boochs2         & $D_{720}$                                                                                           & \cite{boochs1990}       \\
+	CAI             & $0.5 \times (R_{2000} + R_{2200}) -R_{2100}$                                                        & \cite{nagler2003}       \\
+	\midrule
+	CARI            & $a = (R_{700}-R_{550}) / 150$                                                                       & \cite{walthall1994}          \\
+	& $b = R_{550}-(a\times 550)$                                                                                         &                                  \\
+	& $\frac{R_{700}\times | (a\times 670+R_{670}+b)}{R_{670}\times(a^2+1)| ^{0.5}}$                                      &                                  \\
+	\midrule
+	Carter          & $R_{695}/R_{420}$                                                                                   & \cite{carter1994}              \\
+	Carter2         & $R_{695}/R_{760}$                                                                                   & \cite{carter1994}              \\
+	Carter3         & $R_{605}/R_{760}$                                                                                   & \cite{carter1994}              \\
+	Carter4         & $R_{710}/R_{760}$                                                                                   & \cite{carter1994}              \\
+	Carter5         & $R_{695}/R_{670}$                                                                                   & \cite{carter1994}              \\
+	Carter6         & $R_{550}$                                                                                           & \cite{carter1994}              \\
+	CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                   & \cite{zarco-tejada2003a} \\
+	CI2             & $R_{760}/R_{700}-1$                                                                                 & \cite{gitelson2003}     \\
+	ClAInt          & $\int_{600 nm}^{735 nm} R$                                                                          & \cite{oppelt2004}   \\
+	CRI1            & $1/R_{515}-1/R_{550}$                                                                               & \cite{gitelson2003}     \\
+	CRI2            & $1/R_{515}-1/R_{770}$                                                                               & \cite{gitelson2003}     \\
+	CRI3            & $1/R_{515}-1/R_{550}\times R_{770}$                                                                 & \cite{gitelson2003}     \\
+	CRI4            & $1/R_{515}-1/R_{700}\times R_{770}$                                                                 & \cite{gitelson2003}     \\
+	D1              & $D_{730}/D_{706}$                                                                                   & \cite{zarco-tejada2003a} \\
+	D2              & $D_{705}/D_{722}$                                                                                   & \cite{zarco-tejada2003a} \\
+	Datt            & $(R_{850}-R_{710})/(R_{850}-R_{680})$                                                               & \cite{datt1999}               \\
+	Datt2           & $R_{850}/R_{710}$                                                                                   & \cite{datt1999}               \\
+	Datt3           & $D_{754}/D_{704}$                                                                                   & \cite{datt1999}               \\
+	Datt4           & $R_{672}/(R_{550} \times R_{708})$                                                                  & \cite{datt1998}                \\
+	Datt5           & $R_{672}/R_{550}$                                                                                   & \cite{datt1998}                \\
+	Datt6           & $(R_{860})/(R_{550}\times R_{708})$                                                                 & \cite{datt1998}                \\
+	Datt7           & $(R_{860} - R_{2218})/(R_{860} - R_{1928})$                                                         & \cite{datt1999a}               \\
+	Datt8           & $(R_{860} - R_{1788})/(R_{860} - R_{1928})$                                                         & \cite{datt1999a}               \\
+	DD              & $(R_{749}-R_{720})-(R_{701}-R_{672})$                                                               & \cite{maire2004}     \\
+	DDn             & $2\times (R_{710}-R_{660}-R_{760})$                                                                 & \cite{lemaire2008}     \\
+	DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                       & \cite{zarco-tejada2003a} \\
+	DWSI1           & $R_{80}/R_{1660}$                                                                                   & \cite{apan2004}         \\
+	DWSI2           & $R_{1660}/R_{550}$                                                                                  & \cite{apan2004}         \\
+	DWSI3           & $R_{1660}/R_{680}$                                                                                  & \cite{apan2004}         \\
+	DWSI4           & $R_{550}/R_{680}$                                                                                   & \cite{apan2004}         \\
+	DWSI5           & $(R_{800} + R_{550})/(R_{1660} + R_{680})$                                                          & \cite{apan2004}         \\
+	EGFN            & $\frac{(\max(D_{650:750})-\max(D_{500:550}))}{(\max(D_{650:750})+\max(D_{500:550}))}$               & \cite{penuelas1994}     \\
+	EGFR            & $\max(D_{650:750})/\max(D_{500:550})$                                                               & \cite{penuelas1994}     \\
+	EVI             & $\frac{2.5\times ((R_{800}-R_{670}) }{ (R_{800}-(6\times R_{670})-(7.5\times R_{475})+1)}$          & \cite{huete1997a}        \\
+	GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$**                                                   & \cite{wu2014}                  \\
+	GI              & $R_{554}/R_{677}$                                                                                   & \cite{smith1995}        \\
+	Gitelson        & $1/R_{700}$                                                                                         & \cite{gitelson1999}     \\
+	Gitelson2       & $(R_{750}-R_{800}/R_{695}-R_{740})-1$                                                               & \cite{gitelson2003}   \\
+	GMI1            & $R_{750}/R_{550}$                                                                                   & \cite{gitelson2003}     \\
+	GMI2            & $R_{750}/R_{700}$                                                                                   & \cite{gitelson2003}     \\
+	Green NDVI      & $\frac{R_{800}-R_{550}}{R_{800}+R_{550}}$                                                           & \cite{gitelson1996}     \\
+	LWVI\_1         & $\frac{(R_{1094}-R_{983})}{(R_{1094}+R_{983})}$                                                     & \cite{galvao2005}       \\
+	LWVI\_2         & $\frac{R_{1094}-R_{1205}}{R_{1094}+R_{1205}}$                                                       & \cite{galvao2005}       \\
+	Maccioni        & $\frac{R_{780}-R_{710})}{R_{780}-R_{680}}$                                                          & \cite{maccioni2001}     \\
+	\midrule
+	MCARI           & \parbox{5.5cm}{$((R_{700}-R_{670})-0.2\times (R_{700}-R_{550})) \times (R_{700}/R_{670})$}          & \cite{daughtry2000}     \\
+	\midrule
+	MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}         & \cite{wu2008a}           \\
+	\midrule
+	mND705          & $\frac{(R_{750}-R_{705})}{R_{750}+R_{705}-2\times R_{445}}$ & \cite{sims2002a} \\
+	mNDVI           &  $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$ & \cite{sims2002a} \\
+	MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                           & \cite{hernandez-clemente2011}\\
+	MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                & \cite{qi1994}\\
+	MSI             & $\frac{R_{1600}}{R_{817}}$                                                                          & \cite{huntjr1989}\\
+	mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                           & \cite{sims2002a} \\
+	mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                              & \cite{chen1996}\\
+	mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                           & \cite{sims2002a} \\
+	MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                           & \cite{dash2007}\\
+	\midrule
+	MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}            & \cite{haboudane2002}\\
+	\midrule
+	NDLI           & $\frac{log(1/R_{1754}) - log(1/R_{1680})}{log(1/R_{1754}) + log(1/R_{1680})}$                        & \cite{serrano2002} \\
+	NDNI           & $\frac{log(1/R_{1510}) - log(1/R_{1680})}{log(1/R_{1510}) + log(1/R_{1680})}$                        & \cite{serrano2002} \\
+	NDVI           & $\frac{R_{800}-R_{680}}{R_{800}+R_{680}}$                                                            & \cite{tucker1979} \\
+	NDVI2          & $\frac{R_{750}-R_{705}}{R_{750}+R_{705}}$                                                            & \cite{gitelson1994} \\
+	NDVI3          & $\frac{R_{682}-R_{553}}{R_{682}+R_{553}}$                                                            & \cite{guanter2005} \\
+	NDWI           & $\frac{R_{860}-R_{1240}}{R_{860}+R_{1240}}$                                                          & \cite{gao1996} \\
+	NPCI           & $\frac{R_{680}-R_{430}}{R_{680}+R_{430}}$                                                            & \cite{penuelas1994} \\
+	OSAVI          & $\frac{(1+0.16) \times (R_{800}-R_{670})}{R_{800}+R_{670}+0.16 }$                                    & \cite{rondeaux1996} \\
+	OSAVI2         & $\frac{(1+0.16)\times (R_{750}-R_{705})}{R_{750}+R_{705}+0.16) }$                                    & \cite{wu2008a} \\
+	PARS           & $\frac{R_{746}}{R_{513}}$                                                                            & \cite{chappelle1992} \\
+	PRI            & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            & \cite{gamon1992}\\
+	PRI\_norm      & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                & \cite{zarco-tejada2013a}\\
+	PRI*CI2        & $PRI*CI2$                                                                                            & \cite{garrity2011} \\
+	PSRI           & $\frac{R_{678}-R_{500}}{R_{750}}$                                                                    & \cite{merzlyak1999} \\
+	PSSR           & $\frac{R_{800}}{R_{635}}$                                                                            & \cite{blackburn1998} \\
+	PSND           & $\frac{R_{800}-R_{470}}{R_{800}-R_{470})}$                                                           & \cite{blackburn1998} \\
+	PWI            & $\frac{R_{900}}{R_{970}}$                                                                            & \cite{penuelas1997} \\
+	RDVI           & $\frac{R_{800}-R_{670}}{ \sqrt{R_{800}+R_{670}}}$                                                    & \cite{roujean1995} \\
+	\midrule
+	REP\_LE        & \parbox{3.8cm}{Red-edge position through linear extrapolation}                                       & \cite{cho2006} \\
+	\midrule
+	REP\_Li        & $R_{re}=\frac{R_{670}+R_{780})}{2}$                                                                  & \cite{guyot1988} \\
+	& $\frac{700 + 40 \times ((R_{re} -R_{700})}{(R_{740}-R_{700}))}$ & \\
+	\midrule
+	SAVI           & $\frac{(1+L)\times (R_{800}-R_{670})}{(R_{800}+R_{670}+L)}$                                          & \cite{huete1988} \\
+	SIPI           & $\frac{R_{800}-R_{445}}{R_{800}-R_{680}}$                                                            & \cite{pen-uelas1995} \\
+	\midrule
+	SPVI           & \parbox{3.8cm}{$0.4 \times 3.7 \times (R_{800}-R_{670})-1.2 \times ((R_{530}-R_{670})^2)^{0.5}$}     & \cite{vincini2006} \\
+	\midrule
+	SR             & $\frac{R_{800}}{R_{680}}$                                                                            & \cite{jordan1969} \\
+	SR1            & $\frac{R_{750}}{R_{700}}$                                                                            & \cite{gitelson1997}\\
+	SR2            & $\frac{R_{752}}{R_{690}}$                                                                            & \cite{gitelson1997}\\
+	SR3            & $\frac{R_{750}}{R_{550}}$                                                                            & \cite{gitelson1997}\\
+	SR4            & $\frac{R_{700}}{R_{670}}$                                                                            & \cite{mcmurtrey1994} \\
+	SR5            & $\frac{R_{675}}{R_{700}}$                                                                            & \cite{chappelle1992} \\
+	SR6            & $\frac{R_{750}}{R_{710}}$                                                                            & \cite{zarco-tejada1999} \\
+	SR7            & $\frac{R_{440}}{R_{690}}$                                                                            & \cite{lichtenthaler1996} \\
+	SR8            & $\frac{R_{515}}{R_{550}}$                                                                            & \cite{hernandez-clemente2012}\\
+	SRPI           & $\frac{R_{430}}{R_{680}}$                                                                            & \cite{pen-uelas1995}\\
+	SRWI           & $\frac{R_{850}}{R_{1240}}$                                                                           & \cite{zarco-tejada2003a}\\
+	Sum\_Dr1       & $\sum_{i=626}^{795} D1_i$                                                                            & \cite{elvidge1995} \\
+	Sum\_Dr2       & $\sum_{i=680}^{780} D1_i$                                                                            & \cite{filella1994} \\
+	SWIR FI        & $\frac{R_{2133}^2}{R_{2225} \times R_{2209}^3}$                                                      & \cite{levin2007} \\
+	\midrule
+	SWIR LI        & \parbox{3.8cm}{$3.87  \times (R_{2210} - R_{2090}) - 27.51 \times (R_{2280} - R_{2090}) - 0.2$}      & \cite{lobell2001} \\
+	\midrule
+	SWIR SI        & \parbox{3.8cm}{$-41.59 \times (R_{2210} - R_{2090}) + 1.24 \times (R_{2280} - R_{2090}) + 0.64 $}    & \cite{lobell2001} \\
+	\midrule
+	SWIR VI        & \parbox{3.8cm}{$37.72  \times (R_{2210} - R_{2090}) + 6.27 \times (R_{2280} - R_{2090}) + 0.57$}     & \cite{lobell2001} \\
+	\midrule
+	TCARI          & \parbox{3.8cm}{$3*((R_{700}-R_{670})-0.2\times R_{700}-R_{550})\times (R_{700}/R_{670}))$}           & \cite{haboudane2002} \\
+	\midrule
+	TCARI/OSAVI    & TCARI/OSAVI                                                                                          & \cite{haboudane2002} \\
+	\midrule
+	TCARI2         & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} & \cite{wu2008a} \\
+	\midrule
+	TCARI2/OSAVI2  & TCARI2/OSAVI2                                                                                        & \cite{wu2008a} \\
+	\midrule
+	TGI            & \parbox{3.8cm}{$-0.5 (190 (R_{670} - R_{550} ) - 120 (R_{670} - R_{480}))$}                          & \cite{hunt2013} \\
+	\midrule
+	TVI            &\parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}               & \cite{broge2001} \\
+	\midrule
+	Vogelmann      & $\frac{R_{740}}{R_{720}}$                                                                            & \cite{vogelmann1993} \\
+	Vogelmann2     & $\frac{R_{734}-R_{747}}{R_{715}+R_{726})}$                                                           & \cite{vogelmann1993} \\
+	Vogelmann3     & $\frac{D_{715}}{D_{705}}$                                                                            & \cite{vogelmann1993} \\
+	Vogelmann4     & $\frac{R_{734}-R_{747}}{R_{715}+R_{720}}$                                                            & \cite{vogelmann1993} \\
+\end{xtabular*}
+\endgroup
 

### 2020-12-13 docs/00-manuscripts/ieee/ieee-manuscript.tex
discuss the lack of SWIR coverage

@@
 	\acro{NRI}{normalized ratio index}
+	\acro{NDMI}{normalized difference moisture index}
 	\acro{OLS}{ordinary least squares}
@@
 
-% \begin{figure} [t!]
-% 	\begin{center}
-% 		\centering
-% 		\includegraphics[width=0.48\textwidth] {study-area-hyperspectral.pdf}
-% 		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}\label{fig:study_area}
-% 	\end{center}
-% \end{figure}
+\begin{figure*} [ht!]
+	\begin{center}
+		\centering
+		\includegraphics[width=\textwidth] {study-area-hyperspectral.pdf}
+		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}\label{fig:study_area}
+	\end{center}
+\end{figure*}
 
@@
 % Selected percentage of features across plots
-% FIXME: Lösung einfügen nach Klärung mit Alex
 Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process.
@@
 
-The R package hsdar was used for the calculation of vegetation indices \cite{hsdar}.
-All indices that could be calculated with the given spectral range of the data were used.
-Even though this selection included a large number of available indices, some possibly relevant indices might have been missed by relying on the pre-selection of indices offered by the package.
+The available hyperspectral data covered a wavelength between 400 nm and 1000 nm.
+Hence, the wavelength range of the shortwave infrared (SWIR) region is not covered in this study.
+Given that this wavelength range is often used in forest health studies \cite{hais2019}, e.g. when calculating the \ac{NDMI} index \cite{gao1996}, this marks a clear limitation of the dataset at hand.
+
+The R package hsdar was used for the calculation of vegetation indices \cite{lehnert2016}.
+All indices that could be calculated with the given spectral range of the data (400 nm - 1000 nm) were used.
+This means even though \autoref{tab:vegindices} lists all available indices of the package, not all listed indices were used in this study.
+Even though this selection included a large number of indices, some possibly helpful indices might have been missed due to the restriction of the hyperspectral data.
 
@@
 \end{xtabular*}
+\label{tab:vegindices}
 \endgroup

### 2020-12-13 docs/00-manuscripts/ieee/ieee-manuscript.tex
mention the lack of multi-temporal ini-situ data

@@
 
+The dataset consists of in-situ data collected within September 2016 matched against remote sensing data acquired at the end of September 2016.
+A multi-temporal dataset consisting of in-situ data from different phenology stages would possibly improve the achieved model performances.
+However, this would also require matching hyperspectral data of these additional timestamps.
+
 The R package hsdar was used for the calculation of vegetation indices \cite{lehnert2016}.

### 2020-12-25 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 
+VI = Vegetation Indices
+NRI = Normalized Ratio Indices
+HR = Hyperspectral reflectances (image bands)
+
 # Datasets {.tabset .tabset-fade}
@@
 ```{r eda-3 }
-plot_prcomp(vi_data_corrected_buffer2)
+vi_data_corrected_buffer2_no_defol = vi_data_corrected_buffer2
+vi_data_corrected_buffer2_no_defol$defoliation <- NULL
+pca <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol,
+  variance_cap = 0.98,
+  ggtheme = ggpubr::theme_pubclean()
+)
+pca[[1]]
 ```
@@
 ```{r eda-5 }
+# remove response
+nri_data_corrected_buffer2$defoliation <- NULL
+
 intro <- introduce(nri_data_corrected_buffer2)
@@
 ```{r eda-6 }
-plot_prcomp(nri_data_corrected_buffer2)
+pca <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2,
+  variance_cap = 0.98,
+  ggtheme = ggpubr::theme_pubclean()
+)
+pca[[1]]
 ```
@@
 ```{r eda-8 }
+# remove response
+bands_data_corrected_buffer2$defoliation <- NULL
+
 intro <- introduce(bands_data_corrected_buffer2)
@@
 
-```{r eda-10 }
-plot_prcomp(bands_data_corrected_buffer2)
+```{r eda-10 pca-bands, dev = c("png", "pdf")}
+pca <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2,
+  variance_cap = 0.98,
+  ggtheme = ggpubr::theme_pubclean()
+)
+pca[[1]]
 ```

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 
-VI = Vegetation Indices
-NRI = Normalized Ratio Indices
-HR = Hyperspectral reflectances (image bands)
+- VI = Vegetation Indices
+- NRI = Normalized Ratio Indices
+- HR = Hyperspectral reflectances (image bands)
 
@@
 
-```{r eda-10 pca-bands, dev = c("png", "pdf")}
+```{r pca-bands, dev = c("png", "pdf")}
 pca <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2,
@@
 #           "code/98-paper/journal/")
+# file_move("docs/figure/eda.Rmd/pca-bands-1.pdf",
+          # "code/98-paper/journal/")
 ```
@@
 
-```{r}
+```{r buffer-size-comp-plot}
 map(

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

--- 2020-03-22 analysis/eda.Rmd
@@
 
+- VI = Vegetation Indices
+- NRI = Normalized Ratio Indices
+- HR = Hyperspectral reflectances (image bands)
+
 # Datasets {.tabset .tabset-fade}
@@
 ```{r eda-3 }
-plot_prcomp(vi_data_corrected_buffer2)
+# remove response
+vi_data_corrected_buffer2_no_defol_no_response <- vi_data_corrected_buffer2
+vi_data_corrected_buffer2_no_defol_no_response$defoliation <- NULL
+
+pca_vi <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol_no_response,
+  variance_cap = 0.90,
+  ggtheme = ggpubr::theme_pubclean(),
+  geom_label_args = list(size = 3),
+  parallel = TRUE
+)
+pca_vi[[1]]
 ```
@@
 ```{r eda-6 }
-plot_prcomp(nri_data_corrected_buffer2)
+# remove response
+nri_data_corrected_buffer2_no_response <- nri_data_corrected_buffer2
+nri_data_corrected_buffer2_no_response$defoliation <- NULL
+
+pca_nri <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2_no_response,
+  variance_cap = 0.9,
+  ggtheme = ggpubr::theme_pubclean(),
+  geom_label_args = list(size = 3),
+  parallel = TRUE
+)
+pca_nri[[1]]
 ```
@@
 
-```{r eda-7 }
-plot_correlation(nri_data_corrected_buffer2)
-```
+No correlation plot for NRI -> too many features.
 
@@
 ```{r eda-10 }
-plot_prcomp(bands_data_corrected_buffer2)
+# remove response
+bands_data_corrected_buffer2_no_response <- bands_data_corrected_buffer2
+bands_data_corrected_buffer2_no_response$defoliation <- NULL
+
+pca_hr <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2_no_response,
+  variance_cap = 0.90,
+  geom_label_args = list(size = 3),
+  ggtheme = ggpubr::theme_pubclean()
+)
+pca_hr[[1]]
 ```
@@
 #           "code/98-paper/journal/")
+# file_move("docs/figure/eda.Rmd/pca-bands-1.pdf",
+# "code/98-paper/journal/")
 ```
@@
 ```
+
+# PCA feature set comparison
+
+Comparing the "proportion of variance explained" (POV) for all feature sets.
+
+```{r pca-pov-comparison, dev = c("png", "pdf")}
+pca_vi[[1]] / pca_nri[[1]] / pca_hr[[1]] *
+  labs(title = NULL) *
+  theme(
+    axis.title.y = element_text(size = 8),
+    axis.title.x = element_text(size = 8)
+  ) +
+  patchwork::plot_annotation(
+    tag_levels = list(c("VI", "NRI", "HR")),
+    title = "Proportin of variance explained by principal components",
+    subtitle = "Labels indicate cumulative % of explained variance"
+  )
+```

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 )
-pca_vi[[1]]
+pca_vi = pca_vi[[1]] +
+  labs(title = "Target POV: 90%")
+pca_vi
 ```
@@
 )
-pca_nri[[1]]
+pca_nri = pca_nri[[1]] + 
+  labs(title = "Target POV: 90%")
+pca_nri
 ```
@@
 pca_hr <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2_no_response,
-  variance_cap = 0.90,
+  variance_cap = 0.98,
   geom_label_args = list(size = 3),
@@
 )
-pca_hr[[1]]
+pca_hr = pca_hr[[1]] + 
+  labs(title = "Target POV: 90%")
+pca_hr
 ```
@@
 ```{r pca-pov-comparison, dev = c("png", "pdf")}
-pca_vi[[1]] / pca_nri[[1]] / pca_hr[[1]] *
-  labs(title = NULL) *
+pca_vi / pca_nri / pca_hr *
   theme(
+    plot.title = element_text(size = 8),
     axis.title.y = element_text(size = 8),
@@
     tag_levels = list(c("VI", "NRI", "HR")),
-    title = "Proportin of variance explained by principal components",
+    title = "Proportion of variance explained by principal components",
     subtitle = "Labels indicate cumulative % of explained variance"

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 1),
   parallel = TRUE
@@
   variance_cap = 0.98,
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 1),
   ggtheme = ggpubr::theme_pubclean()

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 1),
+  geom_label_args = list(size = 2),
   parallel = TRUE
@@
   variance_cap = 0.98,
-  geom_label_args = list(size = 1),
+  geom_label_args = list(size = 2),
   ggtheme = ggpubr::theme_pubclean()

### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

--- 2020-12-26 analysis/eda.Rmd
@@
   theme(
-    plot.title = element_text(size = 8),
+    plot.title = element_text(size = 10),
     axis.title.y = element_text(size = 8),
-    axis.title.x = element_text(size = 8)
+    axis.title.x = element_text(size = 10)
   ) +

### 2020-12-26 docs/00-manuscripts/ieee/ieee-manuscript.tex
add PCA pov analysis

@@
 \documentclass[letterpaper, peerreview, draftcls]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[journal]{../sty/IEEEtran}
 
@@
 	\acro{PISR}{potential incoming solar radiation}
+    \acro{PC}{principal component}
 	\acro{PCA}{principal component analysis}
@@
 	\acro{PLS}{partial least-squares}
+	\acro{POV}{proportion of variance explained}
 	\acro{RBF}{radial basis function}
@@
 %\end{figure}
+
+\subsection{Correlation among predictors}
+
+A \ac{PCA} was done to analyse the correlation within the variables of the different feature sets (FIXME figref).
 
@@
 \section{Results}
+
+\subsection{Correlation analysis}
+
+Inspecting the results of the \ac{PCA} for all feature sets, HR showed a high correlation among variables with a \ac{POV} of 84\% in the first \ac{PC} and 14\% in the second \ac{PC}, summing up to a cumulative \ac{POV} of 98\% (\autoref{fig:pca-pov}).
+
+Feature sets VI (51\% PC1, 25\% PC2, 7\% PC3) and NRI (51\% PC1, 25\% PC2, 8\% PC3) showed a more diverse distribution across their respective \ac{PC}s.
+
+\begin{figure} [t!]
+	\centering
+	\begin{center}
+		\includegraphics[width=0.48\textwidth] {pca-pov-comparison-1.pdf}
+		\caption{Proportion of variance explained by a Principal Component Analysis for all feature sets.
+		To show at least 2 PC for every feature set, the target POV was set to 98\% for VI and NRI and to 90\% for HR.}
+		\label{fig:pca-pov}
+	\end{center}
+\end{figure}
 

### 2020-12-26 docs/00-manuscripts/ieee/ieee-manuscript.tex
add more information how the data surveyed
Explain why NDVI was chosen as the base for NRI calculation

@@
 Field surveys were conducted in September 2016.
+Defoliation was measured via visual inspection using 5\% intervals.
+No estimation error was recorded.
 
@@
 To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
-The following formula was used for the NRI calculation:
+The following NDVI-based formula was used for the NRI calculation:
 
@@
 \bigbreak{}
+
+The choice of NDVI serving as the base formula for NRI calculation was determined by the implementation of the \texttt{nri} function in the \texttt{hsdar} R package.
 
@@
 This applied to all learners of this study.
+The overall performance achieved in this study can be classified as "fair".
 
@@
 % parameter limits
-\begin{table}[b!]
+\begin{table}[h!]
 	\centering
@@
 \end{table}
+
+\section{Vegetation index formulas}
 

### 2020-12-27 docs/00-manuscripts/ieee/ieee-manuscript.tex
Add geometric, radiometric and atmospheric correction information to appendix F

@@
 
+\section{Radiometric, Geometric and Atmospheric correction of hyperspectral data}
+
+The following information was provided externally by the company which was in charge for the image acquisition and preprocessing of the data.
+
+\textbf{Radiometric Correction}
+
+\begin{quote}
+The conversion of DNs to spectral radiance was made by using a software designed for the instrument.
+Images are originally scaled in 12 bits but they were radiometrically calibrated to 16 bits, reserving the highest value: 65535, for null values.
+The procedure was applied on the 23 previously selected images.
+Thereafter, the geometric and atmospheric corrections were further applied on them.
+\end{quote}
+
+\textbf{Geometric Correction}
+
+\begin{quote}
+The aim of this procedure was to reduce the positional errors of the images.
+The cartography of reference was GeoEuskadi, UTM, zone 30, datum ETRS-89.
+The position was accomplished by coupling an Applanix POS AV 410 system to the sensor, which integrates GPS and IMU systems.
+The system provides geographic coordinates of the terrain and relative coordinates of the aircraft (attitude) at each scanned line.
+Additionally a DSM from GeoEuskadi with a spatial resolution of 1 m was used.
+The ortorectified hyperspectral images were compared to orthoimages 1:5000 of GeoEuskadi and the RMSE was calculated, which was below the GSD in the across and along track directions
+\end{quote}
+
+\textbf{Atmospheric Correction}
+
+\begin{quote}
+The radiance measured by and instrument depends on the illumination geometry and the reflective properties of the observed surface.
+Radiation may be absorbed or scattered (Rayleigh and Mie scattering). Scattering is responsible of the adjacency effect, i.e., radiation coming from neighbors areas to the target pixel.
+The MODTRAN code was used to modelling the effect of the atmosphere on the radiation.
+To represent the aerosols of the study area it was used the rural model, and the optical thickness was estimated on pixels with a high vegetation cover.
+Columnar water vapor was estimated by a linear regression ratio where the spectral radiance of each pixel at the band of the maximum water absorption (~906 nm) is compared to its theoretical value in absence of absorption.
+Nonetheless, this technique is unreliable when a spectral resolution as the one required here is used.
+To solve that, a manual selection of the water vapor parameter was made according to the smoothness observed on the reflectance peak at 960 nm.
+The mid-latitude summer atmosphere model was also used.
+The output of this procedure was reflectance from the target pixel scaled between 0 and 10,000.
+\end{quote}
+
+\textbf{Acquisition days}
+
+\begin{quote}
+The image acquisitions were originally attempted during one day (29) but due to the variable meteorological conditions some stands had to be imaged the next day.
+\end{quote}
+
+\pagebreak
+
 \bibliographystyle{IEEEtran}

### 2021-02-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
improve responses to reviewer2 and add changes to manuscript

@@
-\documentclass[letterpaper, peerreview, draftcls]{IEEEtran}
+% \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
+\documentclass[peerreview]{IEEEtran}
+
+\usepackage[switch]{lineno}
 
@@
 \begin{document}
+\linenumbers
 %
@@
 	\acro{ENM}{environmental niche modeling}
+	\acro{FFS}{forward feature selection}
 	\acro{FPR}{false positive rate}
-	\acro{FFS}{forward feature selection}
 	\acro{FS}{feature selection}
@@
 	\acro{IQR}{interquartile range}
+	\acro{LiDAR}{light detection and ranging}
+	\acro{LOWESS}{locally weighted scatter plot smoothing}
 	\acro{MARS}{multivariate adaptive regression splines}
@@
 	\acro{NDII}{normalized difference infrared index}
+	\acro{NDMI}{normalized difference moisture index}
 	\acro{NIR}{near-infrared}
 	\acro{NRI}{normalized ratio index}
-	\acro{NDMI}{normalized difference moisture index}
 	\acro{OLS}{ordinary least squares}
-	\acro{LiDAR}{light detection and ranging}
-	\acro{LOWESS}{locally weighted scatter plot smoothing}
-	\acro{PISR}{potential incoming solar radiation}
-    \acro{PC}{principal component}
+	\acro{OMNBR}{optimized multiple narrow-band reflectance}
 	\acro{PCA}{principal component analysis}
 	\acro{PDP}{partial dependence plots}
+	\acro{PISR}{potential incoming solar radiation}
 	\acro{PLS}{partial least-squares}
@@
 	\acro{XGBoost}{extreme gradient boosting}
+    \acro{PC}{principal component}
 \end{acronym}
@@
 
-A \ac{PCA} was done to analyse the correlation within the variables of the different feature sets (FIXME figref).
+A \ac{PCA} was done to analyse the correlation within the variables of the different feature sets (\autoref{fig:pca-pov}).
 
@@
 
-To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
-The following NDVI-based formula was used for the NRI calculation:
+To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated.
+
+NRIs follow the \ac{OMNBR} concept of data-driven information extraction from narrow-band indices of hyperspectral data \cite{thenkabail2000,thenkabail2018}.
+In principal any index can be used for this data-driven concept.
+Due to the popularity of the NDVI index in vegetation applications and the availability of this index in the \texttt{nri()} function of the \texttt{hsdar} R package, the following formula was applied to every available band combination:
 
@@
 \bigbreak{}
-
-The choice of NDVI serving as the base formula for NRI calculation was determined by the implementation of the \texttt{nri} function in the \texttt{hsdar} R package.
 
@@
 
-where \(H\) is the conditional entropy of the response variable (class) or the feature (attribute), respectively.
+where \(H\) is the conditional entropy of the response variable (class or Y) or the feature (attribute or X), respectively.
+$H(X)$ is Shannon's Entropy \cite{shannon1948} for a variable X and $H(X, Y)$ is a joint Shannon's Entropy for a variable X with a condition to Y.
+$H(X)$ itself is defined as
+
+\begin{equation}
+H(X) = - \sum_{i=1}^{n} P(x_i)\log_bP(x_i)
+\end{equation}
+
+where $b$ is the base of the logarithm used (with $b$ commonly being set to 2).
+
 In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins and treated as a class variable.
@@
 
+In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
+The selected hyperparameter settings are shown in \autoref{tab:hyperparameter_limits}.
+In addition the code which generated the matrix is available in the research compendium of this study (\url{10.5281/zenodo.2635403}).
+
 \subsubsection{Feature sets}
@@
 \ac{MBO} may outperform methods that do not use information from previous iterations, such as random search or grid search \cite{bergstra2012}.
-
-To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage (\cite{binder2020}).
+Tuning ranges are shown in \autoref{tab:hyperparameter_limits}.
+
+To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage \cite{binder2020}.
 For \ac{PCA}, the number of principal components was tuned instead.
@@
 For the inner level (hyperparameter tuning), \(k - 1\) folds were used with \(k\) being the number of plots.
-
-In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
 

### 2021-02-07 analysis/report-defoliation.Rmd

@@
 source(here::here("code/99-packages.R"))
+load_packages()
 sourceDirectory("R")
@@
 
-  tune_xgboost
+  tune_xgboost_plan_project
 )
-library("rgdal")
 ```
@@
 ```{r}
-suppressMessages(getOptPathY(tune_xgboost$opt.path))
+suppressMessages(getOptPathY(tune_xgboost_plan_project$opt.path))
 ```
@@
 ```{r}
-suppressMessages(plotOptPath(tune_xgboost$opt.path))
+suppressMessages(plotOptPath(tune_xgboost_plan_project$opt.path))
 ```
@@
 ```{r}
-hist <- suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[1]], aes(x = defoliation)) +
+hist <- suppressWarnings(suppressMessages(ggplot(
+  defoliation_df_relative[[1]],
+  aes(x = defoliation)
+) +
   geom_histogram() +
@@
 ```{r}
-hist <- suppressWarnings(suppressMessages(ggplot(defoliation_df_relative[[2]], aes(x = defoliation)) +
+hist <- suppressWarnings(suppressMessages(ggplot(
+  defoliation_df_relative[[2]],
+  aes(x = defoliation)
+) +
   geom_histogram() +
@@
 ```{r echo = FALSE}
-system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/")
+#system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/")
 ```

### 2021-02-09 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 3, nudge_y = 0.05),
   parallel = TRUE
@@
   labs(title = "Target POV: 90%")
-pca_vi
+pca_vi[[1]]
 ```
@@
     coords = coords,
+    # EPSG: 25830
     proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

### 2021-02-10 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3, nudge_y = 0.05),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   parallel = TRUE
@@
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   parallel = TRUE
@@
   variance_cap = 0.98,
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   ggtheme = ggpubr::theme_pubclean()

### 2021-02-11 docs/00-manuscripts/ieee/ieee-manuscript.tex
update rw4 replies

@@
 
-Inspecting the results of the \ac{PCA} for all feature sets, HR showed a high correlation among variables with a \ac{POV} of 84\% in the first \ac{PC} and 14\% in the second \ac{PC}, summing up to a cumulative \ac{POV} of 98\% (\autoref{fig:pca-pov}).
-
-Feature sets VI (51\% PC1, 25\% PC2, 7\% PC3) and NRI (51\% PC1, 25\% PC2, 8\% PC3) showed a more diverse distribution across their respective \ac{PC}s.
+The PCA-based correlation analysis in \autoref{fig:pca-pov} shows that, depending on the feature set, between two (HR) and five (VI) \ac{PC} are needed to explain 90\% of variance.
+The less PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
+Hence, a high intra-feature correlation can be observed for feature set HR which reaches 98\% with only two PCs.
+Feature sets NRI and VI are more diverse and require four (NRI) and five (VI) PCs to explain 90\% of the variance in the data, respectively.
 
@@
 
+In addition, data quality issues (\autoref{subsec:data-quality}) might have an influence on model performances.
+These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g.
+soil reflectance) and the possible geometrical offset of trees in the field.
+
 \subsubsection{Model differences}
@@
 
-\subsection{Data quality}
+\subsection{Data Quality}
+\label{subsec:data-quality}
 
@@
 The following paragraph discusses these.
-The decision to use a buffer of two meters for the extraction of reflectance values was a complex process.
+
+Finding a suitable approach to extract the remote sensing information from each tree was a complex process.
 Due to the reported geometric offset of up to 1 m within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different observation was reasonably high.
-By using a buffer of two meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
-However, when using no buffer at all, the difference between single, neighboring pixels might even be higher compared to a smoothed value composed out of a buffer.
-It was concluded that using a buffer of two meters is a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
+Finally it was decided to apply a buffer around each recorded tree location and not to solely rely on the reported location during value extraction.
+By using a buffer of multiple meters, the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
+However, when using no buffer at all, the difference between single, neighboring pixels might be quite high and involves the risk of including non-tree data.
+It was concluded that using a buffer of two meters can be a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
 Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
+Besides using a buffer to extract the hyperspectral information, a segmentation could be considered.
+However, this method would require more effort for no clear added value in our view and move the focus of this manuscript more to data preprocessing and away from the focus on filter-based feature selection methods.
 
@@
 		\toprule
-		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
+		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_{try}$}    & dbl & 0         & 0.5      & -       \\
 		 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\

### 2021-02-16 analysis/test/cloud-cover.Rmd
new raster CRS notation

@@
   map(~ st_read(., quiet=TRUE)) %>%
-  map(~ st_set_crs(., "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")) %>%
+  map(~ st_set_crs(., 25830)) %>%
   map(~ st_intersection(st_buffer(., 0), aoi)) %>%
@@
   map_depth(2, ~ st_as_sfc(.)) %>%
-  map_depth(2, ~ st_set_crs(., "+proj=longlat +datum=WGS84 +no_defs ")) %>%
-  map_depth(2, ~ st_transform(., "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")) %>%
+  map_depth(2, ~ st_set_crs(., 4326)) %>%
+  map_depth(2, ~ st_transform(., 32630)) %>%
   set_names(date_scenes)

### 2021-02-16 R/01-fun-download.R

@@ -69,7 +69,7 @@ download_hyperspectral <- function(url) {
     list.files("data/hyperspectral",
       full.names = TRUE, pattern = ".tif$"
     ),
-    ~ brick(.x)
+    ~ brick(.x, crs = 25830)
   )
   return(files)
 }

### 2021-02-16 R/04-fun-sentinel-processing.R

@@ -252,7 +252,7 @@ mosaic_clouds <- function(records) {
     st_polygon(list(cbind(c(0, 1, 1, 0, 0), c(0, 0, 1, 1, 0)))) %>%
     st_sfc() %>%
     st_sf(tibble(name = "Dummy")) %>%
-    st_set_crs("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
+    st_set_crs(4326)
 
   # Build cloud mask mosaic
   vec_mosaic_cloud_mask <-
@@ -265,7 +265,7 @@ mosaic_clouds <- function(records) {
     map(~ purrr::discard(., function(x) all(is.na(x)))) %>%
     map(~ do.call(rbind, .)) %>%
     map_if(~ is.null(.), ~vec_dummy_cloud_mask) %>%
-    map(possibly(~ st_set_crs(., "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"), NA))
+    map(possibly(~ st_set_crs(., 4326), NA))
 
   # Set cloud mask filename
   file_mosaic_cloud_mask <-
@@ -308,7 +308,7 @@ mask_mosaic <- function(image_mosaic,
     lapply(FUN = sf::st_read, quiet = TRUE) %>%
     lapply(
       FUN = sf::st_set_crs,
-      value = "+proj=utm +zone=30 +datum=WGS84 +units=m +noread_defs"
+      value = 4326
     )
 
   # Mask cloud with aoi
@@ -627,6 +627,7 @@ prediction_raster <- function(data, year, relative = NULL) {
   all_spdf <- SpatialPixelsDataFrame(
     points = data[, c("x", "y")],
     data = as.data.frame(data[, "defoliation"]),
+    # EPSG: 4326
     proj4string = CRS("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")
   )
 

### 2021-02-19 analysis/about.Rmd
workflowr::wflow_publish("analysis/about.Rmd")

@@
 
-## Read the code, access the data
+## Reading the code, accessing the data
 
@@
 
-## Install the R package
+## Installing the R package
 
@@
 
-Please note that this project was built with R version 3.6.1 on a CentOS 7 operating system.
+Please note that this project was built with R version 4.0.3 on a CentOS 7 operating system.
 Some packages from this project **are not compatible with R versions prior version 3.6.0.**
@@
 
+## Creating targets with {drake}
+
+Calling `r_make()` will create targets specified in `drake_config(targets = <target>)` in `_drake.R` with the additional drake settings specified.
+
+Out of the 400+ targets in this project, the following targets are important:
+
+-   `bm_aggregated_new_buffer2`: Aggregated benchmark results of all models using a 2 meter buffer for hyperspectral data extraction.
+-   `eda_wfr`: Creates the [report which shows Exploratory Data Analysis (EDA) plots and tables](https://pat-s.github.io/2019-feature-selection/eda.html).
+-   `eval_performance_wfr`: Creates the [report which evaluates the model performances](https://pat-s.github.io/2019-feature-selection/eval-performance.html).
+-   `spectral_signatures_wfr`: Creates the [report which inspects the spectral signatures of the hyperspectral data](https://pat-s.github.io/2019-feature-selection/spectral-signatures.html).
+-   `feature_importance_wfr`: Creates the [report which inspects the feature importance of variables](https://pat-s.github.io/2019-feature-selection/feature-importance.html).
+-   `filter_correlations_wfr`: Creates the [report which inspects correlations among filter methods](https://pat-s.github.io/2019-feature-selection/feature-importance.html).
+
+Note that most reports require some/all fitted models.
+Creating these (e.g. target `benchmark_no_models_new_buffer2`) is a costly process and takes several days on a HPC and way longer on a single machine.
+
 # Notes and resources

### 2021-02-19 analysis/index.Rmd
workflowr::wflow_publish("analysis/index.Rmd")

@@
 
-[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.6.0-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-4.0.3-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2582969.svg)](https://doi.org/10.5281/zenodo.2582969)

### 2021-02-20 analysis/index.Rmd
bump to R 4.0.4

@@
 
-[![minimal R version](https://img.shields.io/badge/R%3E%3D-4.0.3-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-4.0.4-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2582969.svg)](https://doi.org/10.5281/zenodo.2582969)

### 2021-02-24 docs/00-manuscripts/ieee/performance-best-per-learner.tex
line-endings
update workflowr site for submission
update latex performance tables

--- 2020-03-05 docs/00-manuscripts/ieee/performance-best-per-learner.tex
@@
-% latex table generated in R 3.6.1 by xtable 1.8-3 package
+% latex table generated in R 4.0.4 by xtable 1.8-4 package
 \begin{table}[ht!]
 \centering
-\caption{Best performance of each learner across any task and filter method.} 
+\caption{Best performance of each learner across any task and filter method, sorted ascending by RMSE} 
 \label{tab:best-learner-perf}
 \scalebox{0.9}{
-\begin{tabular}{rlllrr}
+\begin{tabular}{rlllrrr}
   \hline
- & Task & Model & Filter & RMSE & SE \\ 
+ & Task & Model & Filter & rsq & RMSE & SE \\ 
   \hline
-1 & HR-NRI & SVM & Car & 27.98 & 19.19 \\ 
-  2 & HR-NRI & RF & Car & 31.08 & 17.09 \\ 
-  3 & NRI-VI & XGBOOST & Relief & 32.10 & 17.73 \\ 
-  4 & NRI & Ridge-CV & No Filter & 32.97 & 7.24 \\ 
-  5 & HR & Lasso-CV & No Filter & 45.96 & 18.35 \\ 
+1 & NRI & SVM & Info Gain & -0.07 & 28.03 & 19.13 \\ 
+  2 & HR-NRI & XGBoost & Car & -0.30 & 31.02 & 17.63 \\ 
+  3 & VI & Lasso-MBO & No Filter & -5.97 & 31.11 & 14.91 \\ 
+  4 & NRI & Ridge-MBO & No Filter & -0.18 & 31.16 & 15.03 \\ 
+  5 & NRI & RF & Car & -5.97 & 32.14 & 15.97 \\ 
    \hline

### 2021-02-24 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update tables and figures
update latex performance tables

--- 2020-05-06 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
@@
-% latex table generated in R 3.6.1 by xtable 1.8-3 package
+% latex table generated in R 4.0.3 by xtable 1.8-4 package
 \begin{table}[ht!]
@@
 \label{tab:tune-perc-sel-features}
-\begin{tabular}{llrr}
+\begin{tabular}{llrrr}
   \toprule
-Learner & Plot & Features (\%) & RMSE \\ 
+Learner & Test Plot & Features (\%) & \# & RMSE \\ 
   \midrule
-\midrule\multirow{4}{*}{\textbf{RF Car}} & Laukiz1 & 0.00443 & 27.48 \\ 
-   & Oiartzun & 0.00004 & 37.20 \\ 
-   & Luiando & 0.00002 & 36.98 \\ 
-   & Laukiz2 & 0.03037 & 15.39 \\ 
-  \midrule\multirow{4}{*}{\textbf{XGB Borda}} & Laukiz1 & 0.00604 & 15.00 \\ 
-   & Oiartzun & 0.68227 & 36.15 \\ 
-   & Luiando & 0.99513 & 38.19 \\ 
-   & Laukiz2 & 0.99976 & 29.62 \\ 
-  \midrule\multirow{4}{*}{\textbf{SVM Relief}} & Laukiz1 & 0.00017 & 35.50 \\ 
-   & Oiartzun & 0.71146 & 37.48 \\ 
-   & Luiando & 0.70672 & 14.89 \\ 
-   & Laukiz2 & 0.94696 & 37.19 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.00485 & 3 & 27.77 \\ 
+   & Oiartzun & 0.00002 & 1 & 37.20 \\ 
+   & Luiando & 0.02875 & 9 & 15.57 \\ 
+   & Laukiz2 & 0.04573 & 21 & 38.63 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.50710 & 284 & 38.19 \\ 
+   & Oiartzun & 0.73142 & 364 & 29.85 \\ 
+   & Luiando & 0.94800 & 286 & 36.91 \\ 
+   & Laukiz2 & 0.23877 & 108 & 15.33 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.19391 & 109 & 14.85 \\ 
+   & Oiartzun & 0.00025 & 1 & 35.31 \\ 
+   & Luiando & 0.04491 & 14 & 36.25 \\ 
+   & Laukiz2 & 0.00015 & 1 & 37.48 \\ 
    \bottomrule

### 2021-03-03 docs/00-manuscripts/ieee/ieee-manuscript.tex
Update ieee-manuscript.tex

@@
 
-To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated.
+To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI}s were calculated.
 
 NRIs follow the \ac{OMNBR} concept of data-driven information extraction from narrow-band indices of hyperspectral data \cite{thenkabail2000,thenkabail2018}.
-In principal any index can be used for this data-driven concept.
-Due to the popularity of the NDVI index in vegetation applications and the availability of this index in the \texttt{nri()} function of the \texttt{hsdar} R package, the following formula was applied to every available band combination:
+While various index formulations such as band ratios or normalized ratios are available, indices involving the same bands are strongly correlated.
+Since the widely-used NDVI index belongs to the family of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following NDI formula to combine all pairs of reflectances:
 
@@
 \noindent
-with \(i\) and \(j\) being the respective band numbers.
+where \(i\) and \(j\) are the respective band numbers.
 

### 2021-03-04 docs/00-manuscripts/ieee/ieee-manuscript.tex
Update ieee-manuscript.tex

@@
 Field surveys were conducted in September 2016.
-Defoliation was measured via visual inspection using 5\% intervals.
-No estimation error was recorded.
+Defoliation was assessed visually in the field by experienced forest pathologists using an established field protocol.
+Values were recorded at three levels of the tree and combined into an overall percentage. No estimation error was recorded.
 
@@
 These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g.
-soil reflectance) and the possible geometrical offset of trees in the field.
+soil reflectance) and the possible positional offset of measured GPS coordinates of trees.
 
@@
 
-The following information was provided externally by the company which was in charge for the image acquisition and preprocessing of the data.
+The following information was provided by the Institut Carogràfic i Geològic de Catalunya, which was in charge of image acquisition and data preprocessing.
+
+\textbf{Sensor Characteristics}
+
+\begin{quote}
+The AisaEAGLE II sensor was used for airborne image acquisition with a field of view of 37.7°.
+Its spectral resolution is 2.4 nm in the range from 400 to 1000 nm.
+\end{quote}
 
@@
 \begin{quote}
-The radiance measured by and instrument depends on the illumination geometry and the reflective properties of the observed surface.
+The radiance measured by an instrument depends on the illumination geometry and the reflective properties of the observed surface.
 Radiation may be absorbed or scattered (Rayleigh and Mie scattering). Scattering is responsible of the adjacency effect, i.e., radiation coming from neighbors areas to the target pixel.
 The MODTRAN code was used to modelling the effect of the atmosphere on the radiation.
-To represent the aerosols of the study area it was used the rural model, and the optical thickness was estimated on pixels with a high vegetation cover.
+To represent the aerosols of the study area the rural model was used, and the optical thickness was estimated on pixels with a high vegetation cover.
 Columnar water vapor was estimated by a linear regression ratio where the spectral radiance of each pixel at the band of the maximum water absorption (~906 nm) is compared to its theoretical value in absence of absorption.
 Nonetheless, this technique is unreliable when a spectral resolution as the one required here is used.
-To solve that, a manual selection of the water vapor parameter was made according to the smoothness observed on the reflectance peak at 960 nm.
+To resolve this, the water vapor parameter was selected manually according to the smoothness observed on the reflectance peak at 960 nm.
 The mid-latitude summer atmosphere model was also used.

### 2021-03-07 analysis/eval-performance.Rmd
fix plot order in eval-performance

@@
     remove = FALSE,
-    sep = " MBO "
+    sep = " MBO ", 
+    fill = "right",
   ) %>%
@@
   dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
-    Laukiz1 = "1", Laukiz2 = "4",
-    Luiando = "3", Oiartzun = "2"
+    Laukiz1 = "2", Laukiz2 = "3",
+    Luiando = "1", Oiartzun = "4"
   )) %>%

### 2021-03-07 analysis/eda.Rmd
eda: 95% for PCA instead of 90%

--- 2021-02-09 analysis/eda.Rmd
@@
 pca_vi <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol_no_response,
-  variance_cap = 0.90,
+  variance_cap = 0.95,
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3, nudge_y = 0.05),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   parallel = TRUE
@@
 pca_vi = pca_vi[[1]] +
-  labs(title = "Target POV: 90%")
+  labs(title = "Target POV: 95%")
 pca_vi[[1]]
@@
 pca_nri <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2_no_response,
-  variance_cap = 0.9,
+  variance_cap = 0.95,
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   parallel = TRUE
@@
 pca_nri = pca_nri[[1]] + 
-  labs(title = "Target POV: 90%")
+  labs(title = "Target POV: 95%")
 pca_nri
@@
 pca_hr = pca_hr[[1]] + 
-  labs(title = "Target POV: 90%")
+  labs(title = "Target POV: 95%")
 pca_hr

### 2021-03-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
merge local changes

--- 2021-02-11 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
+
 It is assumed that the fungi infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
@@
 The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5th 2016.
-The images were taken by an AISAEAGLE-II sensor.
+The images were taken by an AISA EAGLE-II sensor.
 All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.
@@
 \begin{quote}
-The image acquisitions were originally attempted during one day (29) but due to the variable meteorological conditions some stands had to be imaged the next day.
+The image acquisitions were originally attempted during one day (29.10.2016) but due to the variable meteorological conditions some stands had to be imaged the next day.
 \end{quote}

### 2021-03-07 R/01-fun-download.R

@@ -69,7 +69,8 @@ download_hyperspectral <- function(url) {
     curl::curl_download(url,
       destfile = glue::glue(tempdir(), "/hs.zip"), quiet = FALSE
     )
-    unzip(glue::glue(tempdir(), "/hs.zip"), exdir = "data/hyperspectral/")
+    unzip(glue::glue(tempdir(), "/hs.zip"), exdir = "data/hyperspectral/",
+      junkpaths = TRUE)
   }
 
   files <- purrr::map(

### 2021-03-08 docs/00-manuscripts/ieee/ieee-manuscript.tex
Apply suggestions from code review
minor edits (#10)
Co-authored-by: Patrick Schratz <patrick.schratz@gmail.com>

--- 2021-03-04 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
 Defoliation was assessed visually in the field by experienced forest pathologists using an established field protocol.
-Values were recorded at three levels of the tree and combined into an overall percentage. No estimation error was recorded.
+Values were recorded at three levels of the tree and combined into an overall percentage.
+No estimation error was recorded.
 
@@
 \begin{quote}
-The AisaEAGLE II sensor was used for airborne image acquisition with a field of view of 37.7°.
+The AISA EAGLE-II sensor was used for airborne image acquisition with a field of view of 37.7°.
 Its spectral resolution is 2.4 nm in the range from 400 to 1000 nm.

### 2021-03-08 docs/00-manuscripts/ieee/ieee-manuscript.tex
improve in-situ survey descr

@@
 In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
-Field surveys were conducted in September 2016.
-Defoliation was assessed visually in the field by experienced forest pathologists using an established field protocol.
-Values were recorded at three levels of the tree and combined into an overall percentage.
-No estimation error was recorded.
+Field surveys were conducted in September 2016 by experienced forest pathologists.
+Defoliation was measured via visual inspection using 5\% intervals at three height levels (bottom, mid, top) with the help of a dedicated score card.
+Estimating the human observer error when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
+Even though no estimation error was recorded in this study, \cite{maclean1982} estimated human observer errors when assessing defoliation to range between 7\% - 18\%.
 
@@
 
-To account for geometric offsets within the hyperspectral data, which were reported with up to 1 m from \ac{ICGC}, a buffer of two meters around the centroid of each tree was used when extracting the reflectance values.
+To account for geometric offsets within the hyperspectral data, which were reported with up to one meter from \ac{ICGC}, a buffer of one meter around the centroid of each tree was used when extracting the reflectance values.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
@@
 Finding a suitable approach to extract the remote sensing information from each tree was a complex process.
-Due to the reported geometric offset of up to 1 m within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different observation was reasonably high.
-Finally it was decided to apply a buffer around each recorded tree location and not to solely rely on the reported location during value extraction.
-By using a buffer of multiple meters, the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
-However, when using no buffer at all, the difference between single, neighboring pixels might be quite high and involves the risk of including non-tree data.
-It was concluded that using a buffer of two meters can be a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
-Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
-Besides using a buffer to extract the hyperspectral information, a segmentation could be considered.
-However, this method would require more effort for no clear added value in our view and move the focus of this manuscript more to data preprocessing and away from the focus on filter-based feature selection methods.
+Due to the reported geometric offset of up to one meter within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different, possibly non-tree, pixel was reasonably high.
+Also, to represent a tree with a crown diameter of around two meters accordingly, more than one square meter (= one pixel) is needed.
+It was concluded that using a buffer of one meter can be a good compromise between the inclusion of information from too many surrounding trees, mapping a single tree crown accordingly accounting for a possible geometric offset.
+The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius.
+This resulted in almost all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
+Even though no results showing the influence of different buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
+Besides using a buffer to extract the hyperspectral information, a segmentation could have been considered.
+However, this method would have required more effort for no clear added value in our view and would have moved the focus of this manuscript more to data preprocessing and away from the focus on filter-based feature selection methods.
 

### 2021-03-10 docs/00-manuscripts/ieee/ieee-manuscript.tex
correct HS data acquisition descr

@@
 
-\begin{figure} [t!]
-	\centering
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {pca-pov-comparison-1.pdf}
-		\caption{Proportion of variance explained by a Principal Component Analysis for all feature sets.
-		To show at least 2 PC for every feature set, the target POV was set to 98\% for VI and NRI and to 90\% for HR.}
-		\label{fig:pca-pov}
-	\end{center}
-\end{figure}
+% \begin{figure} [t!]
+% 	\centering
+% 	\begin{center}
+% 		\includegraphics[width=0.48\textwidth] {pca-pov-comparison-1.pdf}
+% 		\caption{Proportion of variance explained by a Principal Component Analysis for all feature sets.
+% 		To show at least 2 PC for every feature set, the target POV was set to 98\% for VI and NRI and to 90\% for HR.}
+% 		\label{fig:pca-pov}
+% 	\end{center}
+% \end{figure}
 
@@
 
-\textbf{Sensor Characteristics}
-
-\begin{quote}
 The AISA EAGLE-II sensor was used for airborne image acquisition with a field of view of 37.7°.
-Its spectral resolution is 2.4 nm in the range from 400 to 1000 nm.
-\end{quote}
-
-\textbf{Radiometric Correction}
-
-\begin{quote}
-The conversion of DNs to spectral radiance was made by using a software designed for the instrument.
-Images are originally scaled in 12 bits but they were radiometrically calibrated to 16 bits, reserving the highest value: 65535, for null values.
-The procedure was applied on the 23 previously selected images.
-Thereafter, the geometric and atmospheric corrections were further applied on them.
-\end{quote}
-
-\textbf{Geometric Correction}
-
-\begin{quote}
+Its spectral resolution is 2.4 nm and ranged from 400 to 1000 nm.
+
+The conversion of Digital Numbers (DN) to spectral radiance was made by using a software designed for the instrument.
+Images were originally scaled in 12 bits but were radiometrically calibrated to 16 bits, reserving the highest value (65535) for null values.
+The procedure was applied to the 23 previously selected images.
+Last, the geometric and atmospheric corrections were applied to the images.
+
 The aim of this procedure was to reduce the positional errors of the images.
-The cartography of reference was GeoEuskadi, UTM, zone 30, datum ETRS-89.
-The position was accomplished by coupling an Applanix POS AV 410 system to the sensor, which integrates GPS and IMU systems.
+The cartographic reference system in use was EPSG 25830.
+Positioning was done by coupling an Applanix POS AV 410 system to the sensor, integrating GPS and IMU systems.
 The system provides geographic coordinates of the terrain and relative coordinates of the aircraft (attitude) at each scanned line.
 Additionally a DSM from GeoEuskadi with a spatial resolution of 1 m was used.
-The ortorectified hyperspectral images were compared to orthoimages 1:5000 of GeoEuskadi and the RMSE was calculated, which was below the GSD in the across and along track directions
-\end{quote}
-
-\textbf{Atmospheric Correction}
-
-\begin{quote}
+The ortorectified hyperspectral images were compared to orthoimages (1:5000) from GeoEuskadi.
+This comparison was used as the base to calculate RMSE, which was below the ground sampling distance in the across and along track directions.
+
 The radiance measured by an instrument depends on the illumination geometry and the reflective properties of the observed surface.
-Radiation may be absorbed or scattered (Rayleigh and Mie scattering). Scattering is responsible of the adjacency effect, i.e., radiation coming from neighbors areas to the target pixel.
-The MODTRAN code was used to modelling the effect of the atmosphere on the radiation.
-To represent the aerosols of the study area the rural model was used, and the optical thickness was estimated on pixels with a high vegetation cover.
+Radiation may be absorbed or scattered (Rayleigh and Mie scattering).
+Scattering is responsible for the adjacency effect, i.e., radiation coming from neighbors areas to the target pixel.
+The MODTRAN algorithm was used to model the effect of the atmosphere on the radiation.
+To represent the aerosols of the study area, the rural model was used.
+In addition the optical thickness was estimated on pixels with a high vegetation cover.
 Columnar water vapor was estimated by a linear regression ratio where the spectral radiance of each pixel at the band of the maximum water absorption (~906 nm) is compared to its theoretical value in absence of absorption.
-Nonetheless, this technique is unreliable when a spectral resolution as the one required here is used.
+Nonetheless, this technique is unreliable in presence of a spectral resolution as in this case.
 To resolve this, the water vapor parameter was selected manually according to the smoothness observed on the reflectance peak at 960 nm.
-The mid-latitude summer atmosphere model was also used.
+This was combined with an mid-latitude summer atmosphere model.
 The output of this procedure was reflectance from the target pixel scaled between 0 and 10,000.
-\end{quote}
-
-\textbf{Acquisition days}
-
-\begin{quote}
-The image acquisitions were originally attempted during one day (29.10.2016) but due to the variable meteorological conditions some stands had to be imaged the next day.
-\end{quote}
+
+The image acquisitions were originally attempted during one day (29.10.2016).
+due to the variable meteorological conditions some stands had to be imaged one day later.
 

### 2021-03-11 analysis/eda.Rmd
object name fixes

--- 2021-02-09 analysis/eda.Rmd
@@
 loadd(
-  vi_data_corrected_buffer2, nri_data_corrected_buffer2,
-  bands_data_corrected_buffer2, coords_vi_nri_clean_corrected,
+  vi_data, nri_data,
+  bands_data, coords_vi_nri_clean,
 
@@
 
-  trees_with_bands_corrected_buffer2,
-  trees_with_bands_corrected_no_buffer
+  trees_with_bands,
+  trees_with_bands_no_buffer
 )
@@
 ```{r eda-1 }
-intro <- introduce(vi_data_corrected_buffer2)
+intro <- introduce(vi_data)
 intro_df <- data.frame(
@@
 ```{r eda-2 }
-plot_histogram(vi_data_corrected_buffer2)
+plot_histogram(vi_data)
 ```
@@
 # remove response
-vi_data_corrected_buffer2_no_defol_no_response <- vi_data_corrected_buffer2
-vi_data_corrected_buffer2_no_defol_no_response$defoliation <- NULL
+vi_data_no_defol_no_response <- vi_data
+vi_data_no_defol_no_response$defoliation <- NULL
 
-pca_vi <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol_no_response,
-  variance_cap = 0.90,
+pca_vi <- DataExplorer::plot_prcomp(vi_data_no_defol_no_response,
+  variance_cap = 0.95,
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3, nudge_y = 0.05),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   parallel = TRUE
@@
 ```{r eda-4 }
-plot_correlation(vi_data_corrected_buffer2)
+plot_correlation(vi_data)
 ```
@@
 ```{r eda-5 }
-intro <- introduce(nri_data_corrected_buffer2)
+intro <- introduce(nri_data)
 intro_df <- data.frame(
@@
 # remove response
-nri_data_corrected_buffer2_no_response <- nri_data_corrected_buffer2
-nri_data_corrected_buffer2_no_response$defoliation <- NULL
+nri_data_no_response <- nri_data
+nri_data_no_response$defoliation <- NULL
 
-pca_nri <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2_no_response,
-  variance_cap = 0.9,
+pca_nri <- DataExplorer::plot_prcomp(nri_data_no_response,
+  variance_cap = 0.95,
   ggtheme = ggpubr::theme_pubclean(),
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   parallel = TRUE
@@
 pca_nri = pca_nri[[1]] +
-  labs(title = "Target POV: 90%")
+  labs(title = "Target POV: 95%")
 pca_nri
@@
 ```{r eda-8 }
-intro <- introduce(bands_data_corrected_buffer2)
+intro <- introduce(bands_data)
 intro_df <- data.frame(
@@
 ```{r eda-9 }
-plot_histogram(bands_data_corrected_buffer2)
+plot_histogram(bands_data)
 ```
@@
 # remove response
-bands_data_corrected_buffer2_no_response <- bands_data_corrected_buffer2
-bands_data_corrected_buffer2_no_response$defoliation <- NULL
+bands_data_no_response <- bands_data
+bands_data_no_response$defoliation <- NULL
 
-pca_hr <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2_no_response,
+pca_hr <- DataExplorer::plot_prcomp(bands_data_no_response,
   variance_cap = 0.98,
-  geom_label_args = list(size = 3),
+  geom_label_args = list(size = 2.5, nudge_y = 0.03),
   ggtheme = ggpubr::theme_pubclean()
@@
 pca_hr = pca_hr[[1]] +
-  labs(title = "Target POV: 90%")
+  labs(title = "Target POV: 95%")
 pca_hr
@@
 ```{r eda-11 }
-plot_correlation(bands_data_corrected_buffer2)
+plot_correlation(bands_data)
 ```
@@
 ```{r eda-12 }
-vi_data_plot <- vi_data_corrected_buffer2 %>%
+vi_data_plot <- vi_data %>%
   mutate(plot = factor(rep(
@@
 
-In Meters.
+Mean average distance between all trees in a given plot.
+
+Calculated via `raster::pointDistance(allpairs = TRUE)`.
 
@@
 dist <- map(plots, ~ {
-  coords <- coords_vi_nri_clean_corrected %>%
+  coords <- coords_vi_nri_clean %>%
     mutate(plot = factor(rep(
@@
 
-# Effects of different buffer sizes when extracting values to trees
-
-The following buffer sizes were used to extract the remote sensing information (1 m spatial resolution) to the trees (points):
-
-- 0 m
-- 0.5 m
-- 1 m
-- 1.5 m
-- 2 m
-
-For this EDA only the direct reflectance values are used.
-It is assumed that the VI and NRI datasets will be affected in the same way by the different buffer sizes.
-
-```{r buffer-size-comp}
-compare_buffer <- function(trees_with_buffer, trees_without_buffer, plot_name) {
-  if (plot_name == "laukiz1") {
-    targets_sub_ids <- 1
-  } else if (plot_name == "laukiz2") {
-    targets_sub_ids <- 2
-  } else if (plot_name == "luiando") {
-    targets_sub_ids <- 3
-  } else if (plot_name == "oiartzun") {
-    targets_sub_ids <- 4
-  }
-
-  foo_buffer <- trees_with_buffer[seq(targets_sub_ids, 12, 4)]
-
-  foo_buffer <- foo_buffer %>%
-    map(~ sf::st_set_geometry(., NULL)) %>%
-    map(~ dplyr::summarise_at(., vars(starts_with("B")), mean, na.rm = TRUE)) %>%
-    map(~ tidyr::pivot_longer(., everything(),
-      names_to = "band", values_to = "reflectance"
-    )) %>%
-    dplyr::bind_rows(.id = "Buffer size")
-
-  foo_no_buffer <- trees_without_buffer[[targets_sub_ids]]
-
-  foo_no_buffer <- foo_no_buffer %>%
-    sf::st_set_geometry(NULL) %>%
-    dplyr::summarise_at(vars(starts_with("B")), mean) %>%
-    tidyr::pivot_longer(everything(),
-      names_to = "band", values_to = "reflectance"
-    ) %>%
-    dplyr::bind_rows(.id = "Buffer size") %>%
-    dplyr::mutate(`Buffer size` = "No Buffer")
-
-  all <- bind_rows(foo_buffer, foo_no_buffer) %>%
-    dplyr::mutate(band = as.numeric(stringr::str_replace(band, "B", "")))
-
-  plot <- ggplot(all, aes(x = band, y = reflectance)) +
-    geom_line(aes(color = `Buffer size`), size = 0.6) +
-    labs(title = tools::toTitleCase(plot_name)) +
-    ggsci::scale_color_nejm() +
-    ggpubr::theme_pubr(legend = "right")
-
-  return(plot)
-}
-```
-
-```{r buffer-size-comp-plot}
-map(
-  c("laukiz1", "laukiz2", "luiando", "oiartzun"),
-  ~ compare_buffer(
-    trees_with_bands_corrected_buffer2,
-    trees_with_bands_corrected_no_buffer,
-    .x
-  )
-) %>%
-  patchwork::wrap_plots() +
-  patchwork::guide_area() +
-  patchwork::plot_layout(guides = "collect", ncol = 3, nrow = 2) +
-  patchwork::plot_annotation(
-    title = "Comparing the effect of different buffer sizes when extracting reflectance values to points obsverations (trees)",
-    subtitle = "Mean values of each hyperspectral band for each plot "
-  )
-```
-
 # PCA feature set comparison
@@
 ```
+
+# Buffer extraction scenario sketches
+
+Standard scenario (somewhere within the pixel)
+
+```{r buffer-scenarios-1, echo=FALSE}
+knitr::include_graphics("inst/tree-buffer-vis.png")
+```
+
+Extreme scenario (at the border of one or multiple pixels)
+
+```{r buffer-scenarios-2, echo=FALSE}
+knitr::include_graphics("inst/tree-buffer-vis-max.png")
+```

### 2021-03-11 code/060-modeling/paper/01-eda.R

@@ -26,9 +26,9 @@ filter_eda_plan <- drake_plan(
   ),
 
   task_list_filter = list(
-    task_new_buffer2[[1]],
-    task_new_buffer2[[2]],
-    task_new_buffer2[[3]]
+    task[[1]],
+    task[[2]],
+    task[[3]]
   ),
 
   filter_values = target(

### 2021-03-11 code/060-modeling/paper/02-param-sets.R

@@ -131,15 +131,15 @@ param_sets_plan <- drake_plan(
     # of of the 80% quantile from the min and max of the smallest and largest
     # task, respectively)
 
-    # lambda_ridge_vi_train <- train(lrn_ridge, task_new_buffer2[[2]])
+    # lambda_ridge_vi_train <- train(lrn_ridge, task[[2]])
     lambda_ridge_vi_train <- train(
       makeLearner("regr.cvglmnet"),
-      task_new_buffer2[[2]]
+      task[[2]]
     )
-    # lambda_ridge_hr_nri_vi_train <- train(lrn_ridge, task_new_buffer2[[6]])
+    # lambda_ridge_hr_nri_vi_train <- train(lrn_ridge, task[[6]])
     lambda_ridge_hr_nri_vi_train <- train(
       makeLearner("regr.cvglmnet"),
-      task_new_buffer2[[6]]
+      task[[6]]
     )
 
     lambda_ridge_min_vi <- min(lambda_ridge_vi_train$learner.model$lambda)
@@ -178,11 +178,11 @@ param_sets_plan <- drake_plan(
 
     lambda_lasso_vi_train <- train(
       makeLearner("regr.cvglmnet", alpha = 0),
-      task_new_buffer2[[2]]
+      task[[2]]
     )
     lambda_lasso_hr_nri_vi_train <- train(
       makeLearner("regr.cvglmnet", alpha = 0),
-      task_new_buffer2[[6]]
+      task[[6]]
     )
 
     lambda_lasso_min_vi <- min(lambda_lasso_vi_train$learner.model$lambda)

### 2021-03-11 code/070-benchmark-matrix.R

@@ -12,9 +12,31 @@ benchmark_plan <- drake_plan(
       keep(~ "SVM MBO MRMR" %in% .x$id)
   ),
 
-  tune_wrappers_mbo_xg = tune_wrappers_mbo[7:12],
+  tune_wrappers_mbo_xg = tune_wrappers_mbo[19:21],
 
-  # 174 targets
+  ### 174 targets
+  # Learners:
+  # 1:6   - RF
+  # 7:12  - XGBOOST
+  # 13:18 - SVM
+  # 19    - RF MBO Borda: 19
+  # 20    - XGBOOST MBO Borda: 20
+  # 21    - SVM MBO Borda: 21
+  # 22    - RF MBO PCA: 22
+  # 23    - XGBOOST MBO PCA: 23
+  # 24    - SVM MBO PCA: 24
+  # 25    - RF MBO No Filter
+  # 26    - XGBOOST MBO No Filter
+  # 27    - SVM MBO No Filter
+  # 28    - Lasso MBO
+  # 29    - RIDGE MBO
+  # Tasks:
+  # 1     - hr
+  # 2     - vi
+  # 3     - nri
+  # 4     - hr_nri
+  # 5     - nri_vi
+  # 6     - hr_nri_vi
   benchmark_no_models = target(
     benchmark(
       learners = tune_wrappers_mbo_xg[[1]],
@@ -29,8 +51,7 @@ benchmark_plan <- drake_plan(
         setAggregation(expvar, test.mean)
       )
     ),
-    dynamic = cross(tune_wrappers_mbo_xg, task_reduced_cor[[2]]),
-    resources = list(ncpus = 4, memory = 3500, njobs = 174)
+    dynamic = cross(tune_wrappers_mbo_xg, task_reduced_cor[[2]])
   ),
 
   tune_wrappers_mbo_inspect_tune = tune_wrappers_mbo[c(1, 20, 15)],

### 2021-03-11 analysis/about.Rmd
workflowr::wflow_publish("analysis/about.Rmd")
about.Rmd

@@
 
-Please note that this project was built with R version 4.0.3 on a CentOS 7 operating system.
+Please note that this project was built with R version 4.0.4 on a CentOS 7 operating system.
 Some packages from this project **are not compatible with R versions prior version 3.6.0.**
@@
 
--   `bm_aggregated_new_buffer2`: Aggregated benchmark results of all models using a 2 meter buffer for hyperspectral data extraction.
+-   `bm_aggregated`: Aggregated benchmark results of all models using a 1 meter buffer for hyperspectral data extraction.
 -   `eda_wfr`: Creates the [report which shows Exploratory Data Analysis (EDA) plots and tables](https://pat-s.github.io/2019-feature-selection/eda.html).
@@
 Note that most reports require some/all fitted models.
-Creating these (e.g. target `benchmark_no_models_new_buffer2`) is a costly process and takes several days on a HPC and way longer on a single machine.
+Creating these (e.g. target `benchmark_no_models`) is a costly process and takes several days on a HPC and way longer on a single machine.
 

### 2021-03-20 docs/00-manuscripts/ieee/ieee-manuscript.tex
update xgboost param set

@@
 		\midrule
-		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
-		 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
-		 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
-		 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
-		 & \texttt{gamma}              & int & 0        & 10      & 0   \\
-		 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
-		 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
+		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 70      & -       \\
+		 & \texttt{colsample\_bytree}  & dbl & 0.6      & 1       & 1   \\
+		 & \texttt{subsample}          & dbl & 0.6      & 1       & 1   \\
+		 & \texttt{max\_depth}         & int & 3        & 15      & 6   \\
+		 & \texttt{gamma}              & int & 0.05     & 10      & 0   \\
+		 & \texttt{eta}                & dbl & 0.1      & 1       & 0.3 \\
+		 & \texttt{min\_child\_weight} & int & 1        & 7       & 1   \\
 		\bottomrule

### 2021-03-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

--- 2020-06-21 analysis/eval-performance.Rmd
@@
 loadd(
-  bm_aggregated_new_buffer2,
+  bm_aggregated,
 
@@
 ```{r, warning = FALSE}
-df_perf <- getBMRPerformances(bm_aggregated_new_buffer2, as.df = TRUE) %>%
+df_perf <- getBMRPerformances(bm_aggregated, as.df = TRUE) %>%
   mutate(task.id = recode_factor(task.id,
-    `hr_buffer2` = "HR",
-    `vi_buffer2` = "VI",
-    `nri_buffer2` = "NRI",
-    `nri_vi_buffer2` = "NRI-VI",
-    `hr_nri_buffer2` = "HR-NRI",
-    `hr_vi_buffer2` = "HR-VI",
-    `hr_nri_vi_buffer2` = "HR-NRI-VI",
+    `hr` = "HR",
+    `vi` = "VI",
+    `nri` = "NRI",
+    `nri_vi` = "NRI-VI",
+    `hr_nri` = "HR-NRI",
+    `hr_vi` = "HR-VI",
+    `hr_nri_vi` = "HR-NRI-VI",
   )) %>%
@@
 ```{r}
-df <- mlr::getBMRPerformances(bm_aggregated_new_buffer2,
-  "hr_buffer2", "SVM MBO No Filter",
+df <- mlr::getBMRPerformances(bm_aggregated,
+  "hr", "SVM MBO No Filter",
   as.df = TRUE

### 2021-03-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 
-- Fold 1: Luiando
-- Fold 2: Laukiz1
-- Fold 3: Laukiz2
+- Fold 1: Laukiz1
+- Fold 2: Laukiz2
+- Fold 3: Luiando
 - Fold 4: Oiartzun
@@
   dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
-    Laukiz1 = "2", Laukiz2 = "3",
-    Luiando = "1", Oiartzun = "4"
+    Laukiz1 = "1", Laukiz2 = "2",
+    Luiando = "3", Oiartzun = "4"
   )) %>%
@@
   dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
-    Oiartzun = "2", Luiando = "3",
-    Laukiz1 = "1", Laukiz2 = "4"
+    Oiartzun = "4", Luiando = "3",
+    Laukiz1 = "1", Laukiz2 = "2"
   )) %>%
@@
   dplyr::mutate("\\#" = case_when(
-    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
+    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 479),
     `Test Plot` == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
-    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 301),
-    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
+    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 300),
+    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 529),
   )) %>% 

### 2021-03-29 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 )
-pca_vi = pca_vi[[1]] +
+pca_vi <- pca_vi[[1]] +
   labs(title = "Target POV: 95%")
@@
 )
-pca_nri = pca_nri[[1]] +
+pca_nri <- pca_nri[[1]] +
   labs(title = "Target POV: 95%")
@@
 )
-pca_hr = pca_hr[[1]] +
+pca_hr <- pca_hr[[1]] +
   labs(title = "Target POV: 95%")
@@
     c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-    c(559, 451, 301, 497)
+    c(479, 451, 300, 529)
   )))
@@
       c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-      c(559, 451, 301, 497)
+      c(479, 451, 300, 529)
     ))) %>%

### 2021-03-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 
-```{r performance-results, warning=FALSE, dev = c("png", "pdf")}
+```{r performance-results, warning=FALSE, dev = c("png", "pdf"), out.width = "70%"}
 results_aggr <- df_perf %>%
@@
   #                        stackdir="up") +
-  geom_beeswarm(groupOnX = FALSE, aes(color = learner_group), size = 3) +
+  ggbeeswarm2::geom_beeswarm(groupOnX = FALSE, aes(color = learner_group), size = 2.5) +
   scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
@@
 
-```{r filter-effect-all-vs-no-filter, warning=FALSE, dev = c("png", "pdf")}
+```{r filter-effect-all-vs-no-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%"}
 results_aggr1 <- df_perf %>%
@@
   ggplot(aes(x = perf, y = learner_group)) +
-  geom_beeswarm(
+  ggbeeswarm2::geom_beeswarm(
     data = results_aggr1[results_aggr1$filter != "NF", ], size = 2.2, shape = 3,

### 2021-03-30 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 library("fs")
+library("patchwork")
 ```

### 2021-03-31 analysis/feature-importance.Rmd
line-endings
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 loadd(
-  fi_permut_hr_buffer2,
-  fi_permut_vi_buffer2,
+  # Slurm resources suggestion: at least 20 cores and 2 GB / core = 40 GB pro job
+  fi_permut_hr,
+  fi_permut_vi,
 
-  fi_ale_hr_buffer2,
-  fi_ale_vi_buffer2,
-  fi_ale_hr_buffer2_gs20,
-  fi_ale_vi_buffer2_gs20,
+  fi_ale_hr,
+  fi_ale_vi,
+  fi_ale_hr_gs20,
+  fi_ale_vi_gs20,
 
@@
 ```{r feature-importance-1 }
-fi_ranked_hr <- fi_permut_hr_buffer2$res %>%
+fi_ranked_hr <- fi_permut_hr$res %>%
   tibble::rownames_to_column("measure") %>%
@@
 
-fi_ranked_vi <- fi_permut_vi_buffer2$res %>%
+fi_ranked_vi <- fi_permut_vi$res %>%
   tibble::rownames_to_column("measure") %>%
+  dplyr::select(-tree.number,-decoloration,-canker, -contains("ID..")) %>% 
   tidyr::pivot_longer(
-    cols = starts_with("B"),
+    cols = !measure,
     values_to = "importance", names_to = "feature"
@@
 
-```{r feature-importance-2 }
+```{r feature-importance-2, warning=FALSE}
 spectra_sim <- hsdar::PROSAIL()
@@
 
-data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
-  left_join(spec_sigs, by = c("wavelength")) %>%
-  mutate(class = "HR") %>%
-  mutate(reflectance = as.numeric(reflectance))
+data_hr_merged <- dplyr::left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
+  dplyr::left_join(spec_sigs, by = c("wavelength")) %>%
+  dplyr::mutate(class = "HR") %>%
+  dplyr::mutate(reflectance = as.numeric(reflectance))
 ```
@@
 df_wavelengths_from_indices_imp <- df_wavelengths_from_indices %>%
-  left_join(fi_ranked_vi, by = c("class" = "feature")) %>%
+  dplyr::left_join(fi_ranked_vi, by = c("class" = "feature")) %>%
   na.omit()
@@
   arrange(rank) %>%
-  group_by(class, rank) %>%
+  dplyr::group_by(class, rank) %>%
   dplyr::slice(1) %>%
@@
 df_wavelengths_from_hr <- data_hr_merged %>%
-  arrange(rank) %>%
-  group_by(class, rank) %>%
+  dplyr::arrange(rank) %>%
+  dplyr::group_by(class, rank) %>%
   dplyr::slice(1) %>%
@@
 df_label_all <- df_wavelengths_from_indices_label %>%
-  bind_rows(df_wavelengths_from_hr)
+  dplyr::bind_rows(df_wavelengths_from_hr)
 ```
@@
 ```{r fi-hr-ale, dev = c("png", "pdf"), warning=FALSE}
-ale_p1 = fi_ale_hr_buffer2$plot(
+ale_p1 = fi_ale_hr$plot(
   features = df_wavelengths_from_hr$class,
@@
 
-```{r, warning=FALSE}
-ale_p2 = fi_ale_hr_buffer2_gs20$plot(
+```{r, warning=FALSE, message=FALSE}
+ale_p2 = fi_ale_hr_gs20$plot(
   features = df_wavelengths_from_hr$class,
@@
 ```{r fi-vi-ale, dev = c("png", "pdf"), warning=FALSE}
-ale_p3 = fi_ale_vi_buffer2$plot(
-  features = paste0("bf2_", df_wavelengths_from_indices_label$class),
+ale_p3 = fi_ale_vi$plot(
+  features = df_wavelengths_from_indices_label$class,
   ncol = 2
@@
 ```{r, warning=FALSE}
-ale_p4 = fi_ale_vi_buffer2_gs20$plot(
-  features = paste0("bf2_", df_wavelengths_from_indices_label$class),
+ale_p4 = fi_ale_vi_gs20$plot(
+  features = df_wavelengths_from_indices_label$class,
   ncol = 2

### 2021-04-01 analysis/eda.Rmd
restore old observation counts

@@
     c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-    c(479, 451, 300, 529)
+    c(559, 451, 301, 497)
   )))
@@
       c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
-      c(479, 451, 300, 529)
+      c(559, 451, 301, 497)
     ))) %>%

### 2021-04-01 code/010-download.R

@@ -15,7 +15,7 @@ download_data_plan <- drake_plan(
 
   tree_per_tree = target({
     data_list <- download_trees("https://zenodo.org/record/3630302/files/tree-in-situ-data-corrected.zip") %>%
-      # purrr::map(~ dplyr::rename(., defoliation = DEFOLIATIO)) %>%
+      purrr::map(~ dplyr::rename(., defoliation = DEFOLIATIO)) %>%
       purrr::map(~ sf::st_transform(., 32630))
 
     # data_list[[1]] %<>% dplyr::rename(x = X, y = Y)

### 2021-04-01 code/050-tasks.R

@@ -46,7 +46,7 @@ tasks_plan <- drake_plan(
       data = data[[1]],
       target = "defoliation",
       coordinates = coordinates_list[[1]],
-      blocking = factor(rep(1:4, c(479, 451, 300, 529))),
+      blocking = factor(rep(1:4, c(559, 451, 301, 497))),
     ),
     dynamic = map(
       data,
@@ -61,7 +61,7 @@ tasks_plan <- drake_plan(
       data = data_reduced_cor[[1]],
       target = "defoliation",
       coordinates = coordinates_list[[1]],
-      blocking = factor(rep(1:4, c(479, 451, 300, 529))),
+      blocking = factor(rep(1:4, c(559, 451, 301, 497))),
     ),
     dynamic = map(
       data_reduced_cor,

### 2021-04-05 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   tibble::rownames_to_column("measure") %>%
-  dplyr::select(-tree.number,-decoloration,-canker, -contains("ID..")) %>% 
+  dplyr::select(-contains("ID..")) %>% 
   tidyr::pivot_longer(

### 2021-04-05 analysis/filter-correlation.Rmd
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE, verbose = TRUE)
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE,     verbose = TRUE)

@@
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:89, 7)) %>%
+  mutate(id = rep(1:90, 7)) %>%
   tidyr::spread(filter, value) %>%
@@
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:89, 3)) %>%
+  mutate(id = rep(1:90, 3)) %>%
   tidyr::spread(filter, value) %>%
@@
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:7381, 7)) %>%
+  mutate(id = rep(1:7875, 7)) %>%
   tidyr::spread(filter, value) %>%
@@
   dplyr::select(value, filter) %>%
-  mutate(id = rep(1:7381, 3)) %>%
+  mutate(id = rep(1:7875, 3)) %>%
   tidyr::spread(filter, value) %>%
@@
 ```
-
-```{r filter-correlation-5, echo = FALSE}
-system("sleep 5")
-# file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
-#           "code/98-paper/journal/")
-# file_move("docs/figure/filter-correlation.Rmd/correlation-filter-nri-1.pdf",
-#           "code/98-paper/journal/")
-```

### 2021-04-09 analysis/eval-performance.Rmd
fix beeswarm plots

@@
   ggplot(aes(x = perf, y = learner_group)) +
-  geom_beeswarm(
+  ggbeeswarm2::geom_beeswarm(
     data = results_aggr2[results_aggr2$filter != "Borda", ],

### 2021-04-09 docs/00-manuscripts/ieee/performance-best-per-learner.tex
update plots

@@
 \scalebox{0.9}{
-\begin{tabular}{rlllrrr}
+\begin{tabular}{rlllrrrr}
   \hline
- & Task & Model & Filter & rsq & RMSE & SE \\ 
+ & Task & Model & Filter & rsq & expvar & RMSE & SE \\ 
   \hline
-1 & NRI & SVM & Info Gain & -0.07 & 28.03 & 19.13 \\ 
-  2 & HR-NRI & XGBoost & Car & -0.30 & 31.02 & 17.63 \\ 
-  3 & VI & Lasso-MBO & No Filter & -5.97 & 31.11 & 14.91 \\ 
-  4 & NRI & Ridge-MBO & No Filter & -0.18 & 31.16 & 15.03 \\ 
-  5 & NRI & RF & Car & -5.97 & 32.14 & 15.97 \\ 
+1 & HR-NRI-VI & SVM & Relief & -3.04 & 3.04 & 28.09 & 18.95 \\ 
+  2 & HR & XGBoost & CMIM & -8.50 & 8.51 & 30.88 & 14.47 \\ 
+  3 & NRI & Lasso-MBO & No Filter & -0.18 & 0.18 & 31.16 & 15.03 \\ 
+  4 & NRI & Ridge-MBO & No Filter & -8.92 & 8.92 & 31.16 & 15.03 \\ 
+  5 & HR & RF & Relief & -3.48 & 3.44 & 32.93 & 13.17 \\ 
    \hline

### 2021-04-09 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update plots

@@
-% latex table generated in R 4.0.3 by xtable 1.8-4 package
+% latex table generated in R 4.0.4 by xtable 1.8-4 package
 \begin{table}[ht!]
@@
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.00485 & 3 & 27.77 \\ 
-   & Oiartzun & 0.00002 & 1 & 37.20 \\ 
-   & Luiando & 0.02875 & 9 & 15.57 \\ 
-   & Laukiz2 & 0.04573 & 21 & 38.63 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.50710 & 284 & 38.19 \\ 
-   & Oiartzun & 0.73142 & 364 & 29.85 \\ 
-   & Luiando & 0.94800 & 286 & 36.91 \\ 
-   & Laukiz2 & 0.23877 & 108 & 15.33 \\ 
-  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.19391 & 109 & 14.85 \\ 
-   & Oiartzun & 0.00025 & 1 & 35.31 \\ 
-   & Luiando & 0.04491 & 14 & 36.25 \\ 
-   & Laukiz2 & 0.00015 & 1 & 37.48 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.14212 & 69 & 41.48 \\ 
+   & Laukiz2 & 0.21934 & 99 & 16.47 \\ 
+   & Luiando & 0.02878 & 9 & 46.33 \\ 
+   & Oiartzun & 0.00279 & 2 & 31.82 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.01502 & 8 & 15.48 \\ 
+   & Laukiz2 & 0.66619 & 301 & 44.39 \\ 
+   & Luiando & 0.10428 & 32 & 40.86 \\ 
+   & Oiartzun & 0.66651 & 353 & 34.15 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
+   & Laukiz2 & 0.04106 & 19 & 31.03 \\ 
+   & Luiando & 0.20735 & 63 & 37.19 \\ 
+   & Oiartzun & 0.23928 & 127 & 14.89 \\ 
    \bottomrule

### 2021-04-09 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove PCA correlation analysis figure and only use text

@@
 
-\subsection{Correlation among predictors}
-
-A \ac{PCA} was done to analyse the correlation within the variables of the different feature sets (\autoref{fig:pca-pov}).
-
 \subsection{Hyperspectral data}
@@
 
-The PCA-based correlation analysis in \autoref{fig:pca-pov} shows that, depending on the feature set, between two (HR) and five (VI) \ac{PC} are needed to explain 90\% of variance.
+The PCA-based correlation analysis showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
 The less PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
-Hence, a high intra-feature correlation can be observed for feature set HR which reaches 98\% with only two PCs.
-Feature sets NRI and VI are more diverse and require four (NRI) and five (VI) PCs to explain 90\% of the variance in the data, respectively.
-
-% \begin{figure} [t!]
-% 	\centering
-% 	\begin{center}
-% 		\includegraphics[width=0.48\textwidth] {pca-pov-comparison-1.pdf}
-% 		\caption{Proportion of variance explained by a Principal Component Analysis for all feature sets.
-% 		To show at least 2 PC for every feature set, the target POV was set to 98\% for VI and NRI and to 90\% for HR.}
-% 		\label{fig:pca-pov}
-% 	\end{center}
-% \end{figure}
+Hence, a high intra-feature correlation can be observed for feature set HR which reaches 96\% with only two PCs.
+Feature sets NRI and VI are more diverse and require 42 (NRI) and 12 (VI) PCs to explain 95\% of the variance in the data, respectively.
 

### 2021-04-09 docs/00-manuscripts/ieee/ieee-manuscript.tex
resolve bib issues

@@
 However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
-To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
+To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014a, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
 However, no study that made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
@@
 	MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                & \cite{qi1994}\\
-	MSI             & $\frac{R_{1600}}{R_{817}}$                                                                          & \cite{huntjr1989}\\
+	MSI             & $\frac{R_{1600}}{R_{817}}$                                                                          & \cite{hunt1989}\\
 	mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                           & \cite{sims2002a} \\
@@
 	SAVI           & $\frac{(1+L)\times (R_{800}-R_{670})}{(R_{800}+R_{670}+L)}$                                          & \cite{huete1988} \\
-	SIPI           & $\frac{R_{800}-R_{445}}{R_{800}-R_{680}}$                                                            & \cite{pen-uelas1995} \\
+	SIPI           & $\frac{R_{800}-R_{445}}{R_{800}-R_{680}}$                                                            & \cite{penuelasj.1995} \\
 	\midrule
@@
 	SR8            & $\frac{R_{515}}{R_{550}}$                                                                            & \cite{hernandez-clemente2012}\\
-	SRPI           & $\frac{R_{430}}{R_{680}}$                                                                            & \cite{pen-uelas1995}\\
+	SRPI           & $\frac{R_{430}}{R_{680}}$                                                                            & \cite{penuelasj.1995}\\
 	SRWI           & $\frac{R_{850}}{R_{1240}}$                                                                           & \cite{zarco-tejada2003a}\\

### 2021-04-09 docs/00-manuscripts/ieee/ieee-manuscript.tex
remove ALE plots from manuscript

@@
 
-\subsubsection{ALE Plots}
-
-Most ALE plots show a small absolute change in ALE compared to their respective mean effects (= 0) for all chosen features (less than $\pm 0.001$) (\autoref{fig:fi-hr-ale}).
-Also, a medium-to-high relative dynamic over the respective reflectance value range of each variable was observed.
-
 \section{Discussion}
@@
 
-% ALE plots
-The ALE plots are hard to interpret because their values are interpreted on a relative scale and differ for each band.
-When linking band reflectance values against ALE values of a specific band, a value of 100 has a very different meaning in B67 than in B124 because for the latter, reflectance values of vegetation are naturally higher.
-Hence, a reflectance value of 100 would be considered high for B67 but low for B124 with respect to a spectral curve of vegetation.
-However, no baseline reflectance value for every feature exists since the absolute reflectance values depend on the sensor characteristics.
-ALE plots can be a powerful tool for interpreting feature importance of correlated features but might be limited in their interpretability because only relative comparisons can be conducted.
-
 \subsection{Data Quality}
@@
 
-\section{SVM ALE plots for task HR}
-
-\begin{figure} [ht]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
-		\caption{ALE plots of SVM on dataset HR.
+% \section{SVM ALE plots for task HR}
+
+% \begin{figure} [ht]
+% 	\begin{center}
+% 		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
+% 		\caption{ALE plots of SVM on dataset HR.
 The ten most important features from the permutation-based variable importance estimation were used.
 The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-hr-ale}
-	\end{center}
-\end{figure}
+% 	\end{center}
+% \end{figure}
 

### 2021-04-09 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   slice(which.min(rmse_aggr)) %>%
+  mutate(rsq = round(rsq, 3)) %>%
   dplyr::rename(
@@
     "RMSE" = rmse_aggr,
-    "SE" = se,
+    "SE (RMSE)" = se,
+    "R^2" = rsq,
   ) %>%
@@
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
-  select(-iter, -`Learner ID`) %>%
+  select(-iter, -`Learner ID`, -expvar) %>%
+  relocate(RMSE, .after = Filter) %>%
+  relocate(`SE (RMSE)`, .after = RMSE) %>%
   mutate(RMSE = round(RMSE, 3)) %>%
@@
     caption.placement = "top",
-    timestamp = NULL
+    timestamp = NULL,
+    sanitize.text.function = function(x) {
+      x
+    }
   )
@@
   arrange(rmse_aggr) %>%
+  mutate(rsq = round(rsq, 3)) %>%
   dplyr::rename(
@@
     "RMSE" = rmse_aggr,
-    "SE" = se,
+    "SE (RMSE)" = se,
+    "R^2" = rsq,
   ) %>%
-  select(-iter) %>%
-  select(-`Learner ID`)
+  select(-`Learner ID`) %>%
+  relocate(RMSE, .after = Filter) %>%
+  relocate(`SE (RMSE)`, .after = RMSE) %>%
+  select(-iter, -expvar)
 
@@
     "RMSE" = rmse_aggr,
+    "R^2" = rsq,
     "SE" = se,
+    "SE (RMSE)" = se
   ) %>%
@@
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
-  select(-iter, -`Learner ID`, -SE) %>%
+  select(-iter, -`Learner ID`, -expvar) %>%
+  relocate(RMSE, .after = Filter) %>%
+  relocate(`SE (RMSE)`, .after = RMSE) %>%
   mutate(RMSE = round(RMSE, 3)) %>%
@@
 # remove redundant midrule from first entry
-df_tbl[[1]][first][1] = gsub("\\\\midrule", "", df_tbl[[1]][first][1])
+df_tbl[[1]][first][1] <- gsub("\\\\midrule", "", df_tbl[[1]][first][1])
 

### 2021-04-10 docs/00-manuscripts/ieee/performance-best-per-learner.tex
update tables

--- 2021-02-24 docs/00-manuscripts/ieee/performance-best-per-learner.tex
@@
   \hline
- & Task & Model & Filter & rsq & RMSE & SE \\ 
+ & Task & Model & Filter & RMSE & SE (RMSE) & R^2 \\ 
   \hline
-1 & NRI & SVM & Info Gain & -0.07 & 28.03 & 19.13 \\ 
-  2 & HR-NRI & XGBoost & Car & -0.30 & 31.02 & 17.63 \\ 
-  3 & VI & Lasso-MBO & No Filter & -5.97 & 31.11 & 14.91 \\ 
-  4 & NRI & Ridge-MBO & No Filter & -0.18 & 31.16 & 15.03 \\ 
-  5 & NRI & RF & Car & -5.97 & 32.14 & 15.97 \\ 
+1 & HR-NRI-VI & SVM & Relief & 28.09 & 18.95 & -3.04 \\ 
+  2 & HR & XGBoost & CMIM & 30.88 & 14.47 & -8.50 \\ 
+  3 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 & -0.18 \\ 
+  4 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 & -8.92 \\ 
+  5 & HR & RF & Relief & 32.93 & 13.17 & -3.48 \\ 
    \hline

### 2021-04-10 docs/00-manuscripts/ieee/ieee-manuscript.tex
2m -> 1m buffer adjustments

@@
 
-In this work permutation-based feature importance and \ac{ALE} plots (HR only) were calculated to estimate feature importance / effects \cite{apley2019, molnar2019}.
-With the limitations of both methods in mind when applied to correlated features, the aim was to get a general overview of the feature importance of the hyperspectral bands while trying to avoid an over-interpretation of results.
+In this work permutation-based feature importance was calculated to estimate feature importance / effects \cite{apley2019}.
+With the limitations in mind when applied to correlated features, the aim was to get a general overview of the feature importance of the hyperspectral bands while trying to avoid an over-interpretation of results.
 The best-performing algorithm on the HR task (i.e. SVM) was used for the feature importance calculation.
@@
 Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.).
-SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to seven percentage points (\autoref{fig:perf-result}).
-Ridge faced major issues in four tasks due to one observation which was predicted off the response scale (i.e. $> 100$).
-SVM showed the best overall performance with a mean difference of around three percentage points to the next best model (RF) (\autoref{tab:best-learner-perf}).
+SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five percentage points (\autoref{fig:perf-result}).
+SVM showed the best overall performance with a mean difference of around three percentage points to the next best model (XGBoost) (\autoref{tab:best-learner-perf}).
 Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 p.p.
@@
 The combination of feature sets showed small increases in performance for some learners.
-RF and XGBoost scored slightly better on the combined datasets HR-NRI and NRI-VI, respectively, compared to their standalone variants (HR, NRI, VI) (\autoref{fig:perf-result}).
-Datasets containing derived features only (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR).
-All learners besides SVM showed a substantially worse performance on the VI dataset compared to all others (around five percentage points worse than their respective best performance).
+RF and XGBoost scored slightly better on the combined datasets HR-NRI, NRI-VI and HR-NRI-VI, respectively, compared to their standalone variants (NRI, VI) (\autoref{fig:perf-result}).
+However, the best performances for both RF and XGBoost were scored on HR only.
+Datasets containing derived features only (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR) or combined feature sets.
 
@@
 % - SVM Car is the overall winner
-% - Ridge-CV hit and miss (hit = NRI, miss = VI)
+% - Lasso quite stable (varies within 1 - 4), ridge a bit more variable (2 - 6)
 % - all MBO penalized methods have no difference across tasks
-SVM combined with the \enquote{Carscore} filter achieved the best overall performance (RMSE of 27.98 p.p.) (\autoref{tab:perf-top-10}).
+SVM combined with the \enquote{Relief} filter achieved the best overall performance (RMSE of 28.09 p.p.) (\autoref{tab:perf-top-10}).
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with ridge penalty (L2) showed a high variance when comparing results across tasks: In four out of six tasks (all which VI variables and HR-NRI) the error was enormous (\autoref{tab:perf-worst-10}).
-For NRI and HR RMSE was 31.16 p.p. and 35.45 p.p., respectively.
-In all settings for which ridge showed such a high error, only one observation in one fold was predicted which such a high defoliation value (in the millions).
-The specific observation showed no obvious signs of being anomalous, e.g. extreme values in individual features or in principal components of features.
-This one outlier caused the error estimates of these folds and the average estimate across all folds to be off scale.
+Regression with ridge (L2) and lasso (L1) penalty showed their best performances on the NRI feature set (\autoref{tab:best-learner-perf}).
+Difference to other feature sets were very small for lasso (below one p.p.) and a bit higher for ridge (between two and five p.p. (VI)).
+XGBoost shows bad performances for some feature sets and fills the ten last places of the ranking (\autoref{tab:perf-worst-10}).
+Especially when combined with PCA, the algorithm was not able to achieve adequate performances.
 
@@
 SVM showed no variation in performance across filters (\autoref{fig:filter-effects-no-filter}).
-Using filters for RF showed a substantial increase in performance for all tasks with the exception of VI, for which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
-XGBoost showed a high dependency on filtering the data: In 4 out of 6 tasks using no filter resulted in the worst or second worst performance.
-In contrast, using no filter on dataset NRI resulted in the best performance.
-XGBoost shows the highest overall differences between filters for a single task:
-for feature set HR, the range is up to 14 percentage points (\enquote{Carscore} vs. \enquote{no filter})(\autoref{fig:filter-effects-no-filter}).
+Using filters for RF showed a substantial increase in performance for all tasks with the exception of NRI, for which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
+XGBoost showed a high dependency on filtering the data: In two out of six tasks using no filter resulted in the worst or second worst performance.
+XGBoost shows the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 percentage points (\enquote{CMIM} vs.
+\enquote{no filter}) (\autoref{fig:filter-effects-no-filter}).
 
@@
 % - SVM unaffected
-When comparing the usage of filters against using no filter at all, there was only one instance (XGBoost on the NRI task) where a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
-For SVM, all filters and \enquote{no filter} achieved the same performance on tasks VI and NRI even though \autoref{fig:perf-result} lists "No Filter" as the best option.
+When comparing the usage of filters against using no filter at all, there no instance where a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
+For SVM, all filters and \enquote{no filter} achieved roughly the same performance on all tasks.
 
@@
 % - never the best
-The Borda filter did not achieve the best performance for any learner across any task (\autoref{fig:filter-effects-borda}).
+The Borda filter achieved only in one instance (RF on HR-NRI-VI) the best performance among the used filters (\autoref{fig:filter-effects-borda}).
 For RF and XGBoost it most often ranked within the first 50\% with respect to all filters of a specific task.
-For XGBoost on the VI task, the Borda filter scored the second worst performance.
+For XGBoost on the NRI task, the Borda filter scored the second worst performance.
 
@@
 Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process.
-Most features were selected during optimization when training without Laukiz2 and least when Laukiz1 served as the test set (\autoref{tab:tune-perc-sel-features}).
-RF used only one feature when Luiando and Oiartzun served as the test data (0.00004 \% of 7675) while 34 and 230 features were seen as optimal during optimization when Laukiz1 and Laukiz2 were the test data, respectively.
+While for RF least features were selected when Oiartzun was the test set ($n = 2$), most features were used for this plot during model fitting for XGBoost and SVM (\autoref{tab:tune-perc-sel-features}).
+Most features were selected during optimization when training without Laukiz2 or Oiartzun and least when Laukiz1 served as the test set.
+RF used only two features when Oiartzun served as the test data (0.0028 \% of 7675) while 34 and 230 features were seen as optimal during optimization when Laukiz1 and Laukiz2 were the test data, respectively.
 In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features; test set) more than two-thirds of all available features.
@@
 		\caption{Predictive performance (RMSE) of models across tasks.
-Suffix 'CV' denotes that the learner was optimized using internal 10-fold CV while prefix 'MBO' means that model-based optimization was used for hyperparameter optimization.
+Suffix 'MBO' means that model-based optimization was used for hyperparameter optimization.
 Abbreviations on the vertical axis refer to the combinations of feature sets on which each model was scored on.
-Labels represent the feature selection method (NF = no filter, Car = 'Carscore', Info = 'Information Gain', Borda = 'Borda').
+Labels represent the feature selection method (e.g.
+NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda').
 The second value of each label shows the RMSE value of the respective setting.}\label{fig:perf-result}
@@
 
-The most important features for datasets HR and VI showed an average decrease in RMSE of 1.57 p.p. (HR, B69) and 1.79 p.p (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
-For both datasets most features among the ten most important ones cluster around a wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
+The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+For dataset HR most features cluster around the infrared region (920 nm - 1000 nm) (six out of ten) while for VI eight out of ten concentrate on the wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
 For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around 1 percentage point).
-Overall, most features showed only a small importance with average decreases in RMSE below 0.5 p.p..
+Overall, most features (excluding the top five respectively) showed only a small importance with average decreases in RMSE below 0.5 p.p..
 
@@
 
-The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE 27.99 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE 28.09 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
 Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 percentage points.
@@
 An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
-These cluster around a performance of 31 p.p while SVM is able to score about 3 p.p. better than all other methods.
+These cluster around a performance of 31 p.p while SVM is able to score about three p.p. better than all other methods.
 However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
@@
 Penalized methods showed promising performances, especially when taking runtime into account.
-When removing features with a correlation of nearly 1, lasso is able to score performances around 31 p.p. and shows, somewhat surprisingly, the best performance on the VI task.
-The same general conclusion applies to ridge if one discards the problematic results on four tasks.
-Regarding the problematic predictions of the ridge learner, a careful inspection of the hyperparameter optimization procedure of the fitted models ensured again that everything was correctly implemented.
-One observation turned out to be influential in causing the outlier prediction value even though careful multivariate exploratory data analysis showed no reason for flagging this observation a priori.
-It was therefore left in the model, but it should be acknowledged that this learner seemed particularly sensitive to the input data.
-
-A potential limiting factor in this study could be the upper limit of 600 iterations used for the XGBoost algorithm (hyperparameter \texttt{nrounds}), especially for feature sets including NRIs (\autoref{tab:hyperparameter_limits}).
+When removing features with a correlation of nearly 1, lasso and ridge are able to score performances around 31 p.p..
+
+A potential limiting factor in this study could be the upper limit of 70 iterations used for the XGBoost algorithm (hyperparameter \texttt{nrounds}), especially for feature sets including NRIs (\autoref{tab:hyperparameter_limits}).
 This setting was a compromise between runtime and tuning space extension with the goal to work well for most feature sets.
@@
 With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.
-The performance of RF and XGBoost on the VI feature set was about six percentage points lower than on others.
+The performance of RF and XGBoost on the VI feature set was around four to six percentage points lower than on others.
 One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
@@
 The usefulness of filters with respect to predictive performance in this study varied.
-While the performance of some models (up to 5 p.p.
+While the performance of some models (up to five p.p.
 for RF and XGBoost) was improved by specific filters, some models achieved a poorer performance with filters than without them (\autoref{fig:filter-effects-no-filter}).
-Since these negative cases were not caused by a specific filter method, it is recommended to test multiple filters in a study if filters are going to be used.
+No pattern with respect to lower scores related to a specific filter method could be found.
+Hence, it is recommended to test multiple filters in a study if filters are going to be used.
 While filters can improve the performance of models, they might be more interesting in other aspects than performance: reducing variables can reduce computational efforts in high-dimensional scenarios and might enhance the interpretability of models.
-Filters are a lot cheaper to compute than wrapper methods and the final feature subset selection can be added as an additional hyperparameter to the model optimization stage.
+Filters are a lot cheaper to compute than wrapper methods and the final feature subset selection can be integrated as an additional hyperparameter into the model optimization stage.
 
@@
 Single filter methods might yield better model performances on certain datasets but fail on others.
-The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in almost all cases (besides XGBoost on task HR-NRI-VI) a single filter outperformed the ensemble filter.
+The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in all cases (besides RF on task HR-NRI-VI) a single filter outperformed the ensemble filter.
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
@@
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
-PCA, acting as a filter method in this work, did neither show the best nor worst result compared to other feature-selection approaches.
-It was able to reduce model fitting times substantially.
+PCA, acting as a filter method, more often showed less optimal results, especially for algorithms RF and XGBoost.
+Especially XGBoost had substantial problems when using PCA as a filter method and scored the worst four results (\autoref{tab:perf-worst-10}).
+However, PCA was able to reduce model fitting times substantially across all algorithms.
 Depending on the use case, PCA can be an interesting option to reduce dimensionality while keeping runtime low.
@@
 The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius.
-This resulted in almost all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
+This results in almost all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
 Even though no results showing the influence of different buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
@@
 % What about the general potential of analyzing defol with hyperspectral data?
-Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 27 percentage points scored by the best performing model.
+Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 28 percentage points scored by the best performing model.
 More training data covering a wider range of defoliation values in a larger number of forest plantations is needed to train better models which can create more robust predictions.

### 2021-04-12 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
   fig.height = 6.13,
-  out.width = "100%",
+  out.width = "70%",
   echo = FALSE

### 2021-04-12 analysis/eval-performance.Rmd

@@
 
-flextable::flextable(df_tab)
+DT::datatable(df_tab)
 ```
@@
     "SE (RMSE)" = se,
-    "R^2" = rsq,
+    "$R^2$" = rsq,
   ) %>%
@@
 
-table1 %>%
-  flextable() %>%
-  autofit()
+DT::datatable(table1)
 ```
@@
 
-table2 %>%
-  flextable() %>%
-  autofit()
+DT::datatable(table2)
 ```
@@
     "RMSE" = rmse_aggr,
-    "R^2" = rsq,
+    "$R^2$" = rsq,
     "SE" = se,
@@
 
-table3 %>%
-  ungroup() %>%
-  arrange(desc(RMSE)) %>%
-  slice(1:10) %>%
-  flextable() %>%
-  autofit()
+DT::datatable(table3)
 ```
@@
 
-```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf")}
+```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%"}
 results_aggr2 <- df_perf %>%
@@
   dplyr::mutate(Learner = forcats::fct_recode(Learner,
-    "RF \\\\ Car" = "RF MBO Car", "XGB \\\\ Borda" = "XGBOOST MBO Borda",
+    "RF \\\\ Relief" = "RF MBO Relief", "XGB \\\\ CMIM" = "XGBOOST MBO CMIM",
     "SVM \\\\ Relief" = "SVM MBO Relief"
@@
   dplyr::mutate("\\#" = case_when(
-    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 479),
+    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
     `Test Plot` == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
-    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 300),
-    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 529),
+    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 301),
+    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
   )) %>%
@@
     type = "latex",
-    caption = "Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE",
+    caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted ascending by RMSE.",
     label = "tab:tune-perc-sel-features",
@@
 
-table4 %>%
-  flextable() %>%
-  autofit() %>%
-  fontsize(size = 13)
+DT::datatable(table4)
 ```
@@
   ) %>%
-  flextable() %>%
-  autofit() %>%
-  fontsize(size = 13)
+  DT::datatable()
 ```

### 2021-04-12 analysis/feature-importance.Rmd

@@
   fig.asp = 0.66,
-  out.width = "100%",
+  out.width = "70%",
   echo = FALSE

### 2021-04-12 docs/00-manuscripts/ieee/tune-perc-sel-features.tex

@@
 \centering
-\caption{Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE} 
+\caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted ascending by RMSE.} 
 \label{tab:tune-perc-sel-features}
@@
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.14212 & 69 & 41.48 \\ 
-   & Laukiz2 & 0.21934 & 99 & 16.47 \\ 
-   & Luiando & 0.02878 & 9 & 46.33 \\ 
-   & Oiartzun & 0.00279 & 2 & 31.82 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.01502 & 8 & 15.48 \\ 
-   & Laukiz2 & 0.66619 & 301 & 44.39 \\ 
-   & Luiando & 0.10428 & 32 & 40.86 \\ 
-   & Oiartzun & 0.66651 & 353 & 34.15 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.99973 & 559 & 41.73 \\ 
+   & Laukiz2 & 0.93371 & 422 & 17.12 \\ 
+   & Luiando & 0.00216 & 1 & 33.21 \\ 
+   & Oiartzun & 0.00004 & 1 & 37.20 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 0.30849 & 173 & 33.36 \\ 
+   & Laukiz2 & 0.98122 & 443 & 45.59 \\ 
+   & Luiando & 0.83613 & 252 & 41.01 \\ 
+   & Oiartzun & 0.00515 & 3 & 15.24 \\ 
   \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
@@
    & Luiando & 0.20735 & 63 & 37.19 \\ 
-   & Oiartzun & 0.23928 & 127 & 14.89 \\ 
+   & Oiartzun & 0.23928 & 119 & 14.89 \\ 
    \bottomrule

### 2021-04-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
update "used features during tuning" part

@@
 % Selected percentage of features across plots
-Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process.
-While for RF least features were selected when Oiartzun was the test set ($n = 2$), most features were used for this plot during model fitting for XGBoost and SVM (\autoref{tab:tune-perc-sel-features}).
-Most features were selected during optimization when training without Laukiz2 or Oiartzun and least when Laukiz1 served as the test set.
-RF used only two features when Oiartzun served as the test data (0.0028 \% of 7675) while 34 and 230 features were seen as optimal during optimization when Laukiz1 and Laukiz2 were the test data, respectively.
-In contrast, XGBoost and SVM used in all cases but Laukiz1 (less than 50 features; test set) more than two-thirds of all available features.
+Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process across each learner and plot.
+While for RF least features were selected when Luiando or Oiartzun were the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
+Overall, SVM used the least features across all plots among all learners, with 24\% for test set Oiartzun being the highest single ploot value.
+Every learner behaved quite differently for each plot and no overall pattern could be observed.
+RF and XGBoost used very few features only when Oiartzun was the test set (one and three, respectively) whereas this was the plot for which SVM used most features ($n = 119$).
 

### 2021-04-12 docs/00-manuscripts/ieee/ieee-manuscript.tex
last buffer size adjustments

@@
 While for RF least features were selected when Luiando or Oiartzun were the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
-Overall, SVM used the least features across all plots among all learners, with 24\% for test set Oiartzun being the highest single ploot value.
-Every learner behaved quite differently for each plot and no overall pattern could be observed.
 RF and XGBoost used very few features only when Oiartzun was the test set (one and three, respectively) whereas this was the plot for which SVM used most features ($n = 119$).
+Overall, SVM used the least features across all plots among all learners, with 24\% for test set Oiartzun being the highest single plot value.
+In general every learner behaved quite differently for each plot and no overall pattern could be observed.
 
@@
 Due to the reported geometric offset of up to one meter within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different, possibly non-tree, pixel was reasonably high.
-Also, to represent a tree with a crown diameter of around two meters accordingly, more than one square meter (= one pixel) is needed.
 It was concluded that using a buffer of one meter can be a good compromise between the inclusion of information from too many surrounding trees, mapping a single tree crown accordingly accounting for a possible geometric offset.
-The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius.
+The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e. <= 1m).
 This results in almost all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
@@
 As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of a tree within the pixel grid.
-If a tree observation is located on the border of the plot, some directions of the buffer will contain no values and the subsequent index value will be calculated with fewer pixels than if the tree observation is located in the middle of the plot.
+If a tree observation is located on the border of the plot, some directions of the buffer will contain no values (because the image coverage was cropped to the borders of the plot) and the subsequent index value will be calculated with fewer pixels than if the tree observation is located in the middle of the plot.
+However, in most cases each tree information should be composed out of four hyperspectral pixels.
 

### 2021-04-12 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
 
-```{r filter-effect-all-vs-no-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%"}
+```{r filter-effect-all-vs-no-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%", message=FALSE}
 results_aggr1 <- df_perf %>%
@@
 
-```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%"}
+```{r filter-effect-all-vs-borda-filter, warning=FALSE, dev = c("png", "pdf"), out.width = "70%", message=FALSE}
 results_aggr2 <- df_perf %>%
@@
   dplyr::mutate(Plot = forcats::fct_recode(Plot,
-    Oiartzun = "2", Luiando = "3",
-    Laukiz1 = "1", Laukiz2 = "4"
+    Oiartzun = "4", Luiando = "3",
+    Laukiz1 = "1", Laukiz2 = "2"
   )) %>%
   dplyr::mutate(Learner = forcats::fct_recode(Learner,
-    "RF Car" = "RF MBO Car", "XGBoost Borda" = "XGBOOST MBO Borda",
-    "SVM Relief" = "SVM MBO Relief"
+    "RF \\\\ Relief" = "RF MBO Relief", "XGB \\\\ CMIM" = "XGBOOST MBO CMIM",
+    "SVM \\\\ Relief" = "SVM MBO Relief"
   )) %>%

### 2021-04-23 docs/00-manuscripts/ieee/ieee-manuscript.tex
wording

@@
 The PCA-based correlation analysis showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
-The less PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
-Hence, a high intra-feature correlation can be observed for feature set HR which reaches 96\% with only two PCs.
+The fewer PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
+Hence, a high correlation can be observed for feature set HR which reaches 96\% with only two PCs.
 Feature sets NRI and VI are more diverse and require 42 (NRI) and 12 (VI) PCs to explain 95\% of the variance in the data, respectively.

### 2021-04-29 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   slice(which.min(rmse_aggr)) %>%
-  mutate(rsq = round(rsq, 3)) %>%
   dplyr::rename(
@@
     "RMSE" = rmse_aggr,
-    "SE (RMSE)" = se,
-    "$R^2$" = rsq,
+    "SE" = se
   ) %>%
@@
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
-  select(-iter, -`Learner ID`, -expvar) %>%
+  select(-iter, -`Learner ID`, -expvar, -rsq) %>%
   relocate(RMSE, .after = Filter) %>%
-  relocate(`SE (RMSE)`, .after = RMSE) %>%
+  relocate(`SE`, .after = RMSE) %>%
   mutate(RMSE = round(RMSE, 3)) %>%
@@
   arrange(rmse_aggr) %>%
-  mutate(rsq = round(rsq, 3)) %>%
   dplyr::rename(
@@
     "RMSE" = rmse_aggr,
-    "SE (RMSE)" = se,
-    "$R^2$" = rsq,
+    "SE" = se
   ) %>%
-  select(-`Learner ID`) %>%
+  select(-`Learner ID`, -rsq) %>%
   relocate(RMSE, .after = Filter) %>%
-  relocate(`SE (RMSE)`, .after = RMSE) %>%
+  relocate(`SE`, .after = RMSE) %>%
   select(-iter, -expvar)
@@
     "RMSE" = rmse_aggr,
-    "$R^2$" = rsq,
     "SE" = se,
-    "SE (RMSE)" = se
   ) %>%

### 2021-04-29 docs/00-manuscripts/ieee/performance-best-per-learner.tex
update "inspect tuning" results Rmd

--- 2021-02-24 docs/00-manuscripts/ieee/performance-best-per-learner.tex
@@
 \scalebox{0.9}{
-\begin{tabular}{rlllrrr}
+\begin{tabular}{rlllrr}
   \hline
- & Task & Model & Filter & rsq & RMSE & SE \\ 
+ & Task & Model & Filter & RMSE & SE \\ 
   \hline
-1 & NRI & SVM & Info Gain & -0.07 & 28.03 & 19.13 \\ 
-  2 & HR-NRI & XGBoost & Car & -0.30 & 31.02 & 17.63 \\ 
-  3 & VI & Lasso-MBO & No Filter & -5.97 & 31.11 & 14.91 \\ 
-  4 & NRI & Ridge-MBO & No Filter & -0.18 & 31.16 & 15.03 \\ 
-  5 & NRI & RF & Car & -5.97 & 32.14 & 15.97 \\ 
+1 & HR-NRI-VI & SVM & Relief & 28.09 & 18.95 \\ 
+  2 & HR & XGBoost & CMIM & 30.88 & 14.47 \\ 
+  3 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
+  4 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
+  5 & HR & RF & Relief & 32.93 & 13.17 \\ 
    \hline

### 2021-06-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
machine learning -> ML

@@
 % \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
-\documentclass[peerreview]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
@@
 \begin{document}
-\linenumbers
+% \linenumbers
 %
@@
 % The paper headers
-\markboth{IEEE Transactions on Geoscience and Remote Sensing}%
+\markboth{IEEE JOURNAL OF SELECTED TOPICS IN APPLIED EARTH OBSERVATIONS AND REMOTE SENSING}%
 {Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
@@
 	For some combinations of learners and feature sets, filter methods achieved better predictive performances than the unfiltered feature sets, while ensemble filters did not have a substantial impact on performance.
-
 	Permutation-based feature importance estimated features around the red edge to be most important for the models.
 	However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the best performances.
-
 	More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
@@
 
-Machine-learning methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
+ML methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
 The ability of predicting into unknown space qualifies these tools as a promising toolset for such tasks.
@@
 Most vegetation indices were developed with the aim of being sensitive to changes of specific wavelength regions, serving as a proxy for underlying plant processes.
-However, often enough indices developed for different purposes than the one to be analyzed can help to explain complex relationships.
-This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship.
+However, indices developed/applied for different purposes than the one to be analyzed (i.e.
+defoliation at pine trees) may help to explain complex underlying relationships which are not obvious on the first look.
+This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship \cite{thenkabail2018}.
 A less known index type which can be derived from spectral information is the \ac{NRI}.
-In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands.
+In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands \cite{thenkabail2000}.
 Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
@@
 
-Despite their popularity in environmental modeling, there are no studies so far which used machine-learning algorithms in combination with remote sensing data to analyze defoliation at the tree level.
+Despite their popularity in environmental modeling, there are no studies so far which used ML algorithms in combination with remote sensing data to analyze defoliation at the tree level.
 This study aims to close this gap by analyzing tree defoliation in northern Spain using airborne hyperspectral data.
-The methodology of this study uses machine-learning methods in combination with feature selection and hyperparameter tuning.
+The methodology of this study uses ML methods in combination with feature selection and hyperparameter tuning.
 In addition, feature importance and feature effects are evaluated.
@@
 % clarify the focus of this study
-This study shows how high-dimensional datasets can be handled effectively with machine-learning methods while still being able to interpret the fitted models.
+This study shows how high-dimensional datasets can be handled effectively with ML methods while still being able to interpret the fitted models.
 The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
@@
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in machine learning which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
+Ensemble filters align with the recent rise of the 'ensemble' approach in ML which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
 In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
@@
 More case studies applying ensemble filter methods are needed to verify this finding.
-Nevertheless, ensemble filters can be a promising addition to a machine-learning feature-selection portfolio.
+Nevertheless, ensemble filters can be a promising addition to a ML feature-selection portfolio.
 
@@
 
-This study analyzed defoliation of trees in northern Spain by using hyperspectral data as input for machine-learning models which used hyperparameter tuning and filter-based feature selection.
+This study analyzed defoliation of trees in northern Spain by using hyperspectral data as input for ML models which used hyperparameter tuning and filter-based feature selection.
 Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.

### 2021-06-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
clarify the usefulness of ML in this study
"invasive pathogens" -> "pathogens"

@@
 ML methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
-The ability of predicting into unknown space qualifies these tools as a promising toolset for such tasks.
+The ability of predicting into unknown space qualifies ML algorithms as a helpful tool for such environmental analyses.
 One aspect of this research field is to enhance the understanding of biotic and abiotic stress triggers, for example by analyzing tree defoliation \cite{hawrylo2018}.
@@
 Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata D. Don}) plantations in northern Spain.
-The trees in the plots suffer from infections of invasive pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P.
+The trees in the plots suffer from infections of pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P.
 Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.} and \textit{Dothisthroma septosporum (Dorogin) M.

### 2021-06-05 docs/00-manuscripts/ieee/ieee-manuscript.tex
random forest -> RF

@@
 The last two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
-In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
+In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
 
@@
 
-Random forest and {SVM} are well established algorithms widely used in (environmental) remote sensing.
+RF and SVM are well established algorithms widely used in (environmental) remote sensing.
 Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmark studies in recent years.
@@
 Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process across each learner and plot.
-While for RF least features were selected when Luiando or Oiartzun were the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
+While for RF least features were selected when Luiando or Oiartzun were the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
 RF and XGBoost used very few features only when Oiartzun was the test set (one and three, respectively) whereas this was the plot for which SVM used most features ($n = 119$).
@@
 % write about hyperspectral remote sensing
-The field of (hyperspectral) remote sensing has a strong focus on using random forests for modeling in recent years \cite{belgiu2016}.
+The field of (hyperspectral) remote sensing has a strong focus on using RF for modeling in recent years \cite{belgiu2016}.
 However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
@@
 A grid search was used for hyperparameter tuning and \ac{FFS} for variable selection.
-\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a random forest classifier.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a RF classifier.
 \cite{zandler2015} shows a similar setup: they used hyperspectral vegetation indices, a nested CV approach for performance estimation and estimated variable importance targeting woody biomass as the response.

### 2021-06-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
slighlty rephrase PCA correlation analysis

@@
 % \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[final]{IEEEtran}
 
@@
 %
-\usepackage{cite}
+\usepackage[noadjust]{cite}
 \usepackage[hidelinks]{hyperref}
@@
 In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
-The selected hyperparameter settings are shown in \autoref{tab:hyperparameter_limits}.
+The selected hyperparameter settings are shown in appendix \autoref{tab:hyperparameter_limits}.
 In addition the code which generated the matrix is available in the research compendium of this study (\url{10.5281/zenodo.2635403}).
@@
 The idea of splitting the features into different sets originated from the question whether feature-engineered indices derived from reflectance values have a positive effect on model performance.
-\cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
+Peña et al. 2017 \cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
@@
 
-The PCA-based correlation analysis showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
+PCA was used for a correlation analysis of the feature sets by inspecting the accumulated percentages of the principal components.
 The fewer PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
+Results showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
 Hence, a high correlation can be observed for feature set HR which reaches 96\% with only two PCs.

### 2021-06-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
percentage points -> p.p.

@@
 Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.).
-SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five percentage points (\autoref{fig:perf-result}).
-SVM showed the best overall performance with a mean difference of around three percentage points to the next best model (XGBoost) (\autoref{tab:best-learner-perf}).
+SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p (\autoref{fig:perf-result}).
+SVM showed the best overall performance with a mean difference of around three p.p to the next best model (XGBoost) (\autoref{tab:best-learner-perf}).
 Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 p.p.
@@
 XGBoost showed a high dependency on filtering the data: In two out of six tasks using no filter resulted in the worst or second worst performance.
-XGBoost shows the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 percentage points (\enquote{CMIM} vs.
+XGBoost shows the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p (\enquote{CMIM} vs.
 \enquote{no filter}) (\autoref{fig:filter-effects-no-filter}).
@@
 For dataset HR most features cluster around the infrared region (920 nm - 1000 nm) (six out of ten) while for VI eight out of ten concentrate on the wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
-For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around 1 percentage point).
+For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around 1 p.p.).
 Overall, most features (excluding the top five respectively) showed only a small importance with average decreases in RMSE below 0.5 p.p..
@@
 The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE 28.09 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
-Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 percentage points.
+Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
 However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
@@
 With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.
-The performance of RF and XGBoost on the VI feature set was around four to six percentage points lower than on others.
+The performance of RF and XGBoost on the VI feature set was around four to six p.p lower than on others.
 One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
@@
 % What about the general potential of analyzing defol with hyperspectral data?
-Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 28 percentage points scored by the best performing model.
+Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 28 p.p scored by the best performing model.
 More training data covering a wider range of defoliation values in a larger number of forest plantations is needed to train better models which can create more robust predictions.

### 2021-06-06 docs/00-manuscripts/ieee/ieee-manuscript.tex
typo
remove unnecessary sentence
remove "almost"

@@
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
-In summary, Borda performs no better than a randomly selected filter method in this study.
+In summary, Borda performs not better than a randomly selected filter method in this study.
 More case studies applying ensemble filter methods are needed to verify this finding.
@@
 Environmental datasets always come with some constraints that can have potential influence on the modeling process and its outcome.
-The following paragraph discusses these.
-
 Finding a suitable approach to extract the remote sensing information from each tree was a complex process.
@@
 The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e. <= 1m).
-This results in almost all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
+This results in all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
 Even though no results showing the influence of different buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.

### 2021-06-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
edited response to reviewer 1
fixed affiliations according to corporate style guides

@@
 	\thanks{P.Schratz, J.Muenchow, J.Cortés and A.Brenning are with the Department
-		of Geography, GIScience group, Friedrich-Schiller-University of Jena, Germany.}% <-this % stops a space
-	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-University Munich.}% <-this % stops a space
-	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Arab, Spain.}% <-this % stops a space
+		of Geography, geographic information science group, Friedrich Schiller University Jena, Germany.}% <-this % stops a space
+	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-Universität in Munich, Germany.}% <-this % stops a space
+	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Spain.}% <-this % stops a space
 	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
@@
 However, indices developed/applied for different purposes than the one to be analyzed (i.e.
-defoliation at pine trees) may help to explain complex underlying relationships which are not obvious on the first look.
+defoliation at pine trees) may help to explain complex underlying relationships which are not obvious at first.
 This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship \cite{thenkabail2018}.
@@
 	%	[]{Patrick Schratz}
-	is a Data Scientist working in Zurich, CH.
+	is a Data Scientist working in Zurich, Switzerland.
 	His area of expertise is applied machine learning, more specifically the field of environmental modeling.
-	He is a PhD candidate at the GIScience group, Department of Geography at University of Jena where he acts as a researcher in the environmental modeling field.
+	He is a PhD candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.
 	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.
@@
 	is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
-	He has earned his PhD at the Friedrich-Alexander University Erlangen-Nuremberg (Germany) in 2013.
+	He received his PhD from Friedrich-Alexander University Erlangen-Nürnberg (Germany) in 2013.
 	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia in 2016 (Friedrich Schiller University Jena).
@@
 	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/eiturritxa.jpg}}]{Eugenia Iturritxa}
-	received the Ecology and Ph.D degree in Plant Protection from the Basque country University in 2001.
+	received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.
 	Since 1999 her main research has focused on forest health, on the study of diverse species of native and introduced pathogenic fungi in forests and forest plantations.
-	Her research includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.
+	Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.
 \end{IEEEbiography}
@@
 	received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
-	Currently he is a PhD student at Friedrich-Schiller-University, Jena, Germany.
+	Currently he is a PhD student at Friedrich Schiller University, Jena, Germany.
 	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
@@
 	obtained his Ph.D from Dortmund Technical University in 2013.
-	He is a professor for "Statistical Learning and Data Science" at the Department of Statistics at the Ludwig-Maximilians-University Munich and a director of the "Munich Center for Machine Learning".
+	He is a professor of Statistical Learning and Data Science in the Department of Statistics at Ludwig-Maximilians-Universität in Munich, Germany, and a director of the Munich Center for Machine Learning.
 	His research interests include AutoML, model selection, interpretable ML and XAI.
@@
 	(Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
-	He served as an assistant professor and tenured associate professor in geomatics at the University from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
+	He served as an assistant professor and tenured associate professor in geomatics at the University of Waterloo, Ontario, Canada from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
 	Dr. Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.

### 2021-06-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
edited and commented on response to Reviewer 2

@@
 
-Even though the feature-selection step should be solely left to the filter methods in this study, it was ensured a priori to account for features with a pairwise correlation of 1.
-Having such features within the data can cause undesired effects during model fitting and feature importance calculation.
-Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of $1 - 10^{-10}$, the feature with the largest mean absolute correlation across all variables was removed from the dataset.
-
+Individual features were finally removed from these feature sets if they were numerically equivalent to another feature based on their pairwise correlation being greater than $1 - 10^{-10}$.
 This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).
@@
 
-\subsection{Correlation analysis}
-
-PCA was used for a correlation analysis of the feature sets by inspecting the accumulated percentages of the principal components.
-The fewer PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
-Results showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
-Hence, a high correlation can be observed for feature set HR which reaches 96\% with only two PCs.
-Feature sets NRI and VI are more diverse and require 42 (NRI) and 12 (VI) PCs to explain 95\% of the variance in the data, respectively.
+\subsection{PCA of feature sets}
+
+PCA was used to assess the complexity of the three feature sets.
+Depending on the feature set, 95\% of the variance is explained by two (HR), twelve (VI) and 42 (NRI) \ac{PC}s.
+HR features are therefore highly redundant, while the applied feature transformations enrich the data set, at least from an exploratory linear perspective.
 

### 2021-06-09 docs/00-manuscripts/ieee/ieee-manuscript.tex
clarify pairwise cor feature removal

@@
 
-Individual features were finally removed from these feature sets if they were numerically equivalent to another feature based on their pairwise correlation being greater than $1 - 10^{-10}$.
+Some individual features were finally removed from these feature sets if they were numerically equivalent to another feature based on their pairwise correlation being greater than $1 - 10^{-10}$.
 This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).

### 2021-06-11 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
upd featsel table

--- 2021-04-09 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
@@
 \centering
-\caption{Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE} 
+\caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
+	exttt{#} denotes the absolute number of features selected and 'Features (	exttt{%})' refers to the percentage relative to the overall features available (7675).
+Results were estimated in a separate model tuning step, not within the main cross-validation comparison.}
 \label{tab:tune-perc-sel-features}
@@
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.14212 & 69 & 41.48 \\ 
-   & Laukiz2 & 0.21934 & 99 & 16.47 \\ 
-   & Luiando & 0.02878 & 9 & 46.33 \\ 
-   & Oiartzun & 0.00279 & 2 & 31.82 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.01502 & 8 & 15.48 \\ 
-   & Laukiz2 & 0.66619 & 301 & 44.39 \\ 
-   & Luiando & 0.10428 & 32 & 40.86 \\ 
-   & Oiartzun & 0.66651 & 353 & 34.15 \\ 
-  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
-   & Laukiz2 & 0.04106 & 19 & 31.03 \\ 
-   & Luiando & 0.20735 & 63 & 37.19 \\ 
-   & Oiartzun & 0.23928 & 127 & 14.89 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Relief}} & Luiando & 99.97268 & 301 & 41.73 \\ 
+   & Laukiz2 & 93.37086 & 422 & 17.12 \\ 
+   & Laukiz1 & 0.21573 & 2 & 33.21 \\ 
+   & Oiartzun & 0.00405 & 1 & 37.20 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Luiando & 30.84900 & 93 & 33.36 \\ 
+   & Laukiz1 & 98.12213 & 549 & 45.59 \\ 
+   & Oiartzun & 83.61298 & 416 & 41.01 \\ 
+   & Laukiz2 & 0.51526 & 3 & 15.24 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.06200 & 1 & 35.49 \\ 
+   & Oiartzun & 4.10636 & 21 & 31.03 \\ 
+   & Luiando & 20.73514 & 63 & 37.19 \\ 
+   & Laukiz2 & 23.92820 & 108 & 14.89 \\ 
    \bottomrule

### 2021-06-17 analysis/eval-performance.Rmd

@@
   mutate(learner_group = recode_factor(learner_group, `XG` = "XGBoost")) %>%
+  mutate(learner_group = recode_factor(learner_group, `Ridge-MBO` = "Ridge")) %>%
+  mutate(learner_group = recode_factor(learner_group, `Lasso-MBO` = "Lasso")) %>%
+  # reorder by min RMSE (reviewer request)
   group_by(learner_group, filter, task.id) %>%
@@
   group_by(task.id, learner_group) %>%
-  slice(which.min(rmse_aggr))
+  slice(which.min(rmse_aggr)) %>% 
+  ungroup() %>% 
+  mutate(learner_group = fct_reorder(learner_group, .$rmse_aggr, min))
 
@@
   ggbeeswarm2::geom_beeswarm(groupOnX = FALSE, aes(color = learner_group), size = 2.5) +
-  scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
+  # scale_color_nejm(breaks = sort(levels(results_aggr$learner_group))) +
+  # scale_color_viridis_d(breaks = sort(levels(results_aggr$learner_group))) +
+  scale_color_viridis_d(breaks = results_aggr$learner_group) +
+
   labs(x = "RMSE", y = "Task", color = "Learner") +
@@
 
+Fold IDs are different for each learner, i.e. a specific plot does not always resolve to "fold 1" for each learner.
+See `bmr_inspect_tune[["results"]][["hr_nri_vi"]][["RF MBO Relief"]][["pred"]][["instance"]][["test.inds"]]`.
+
+Thus, we need to manually label the fold IDs to plot names for each learner.
+
+**RF**
+
+Plot 1: Luiando
+Plot 2: Laukiz2
+Plot 3: Laukiz 1
+Plot 4: Oiartzun
+
+**SVM**
+
+Plot 1: Laukiz1
+Plot 2: Oiartzun
+Plot 3: Luiando
+Plot 4: Laukiz2
+
+**XGBoost**
+
+Plot 1: Luiando
+Plot 2: Laukiz1
+Plot 3: Oiartzun
+Plot 4: Laukiz2
+
 ```{r eval-performance-5}
@@
 df_tbl <- getBMRTuneResults(bmr_inspect_tune, as.df = TRUE) %>%
+  # replace iter IDs by individual fold names
+  dplyr::mutate(iter = as.character(iter)) %>% 
+  dplyr::mutate(iter = case_when(learner.id == "XGBOOST MBO CMIM" & iter == 1 ~ "Luiando",
+                                 learner.id == "XGBOOST MBO CMIM" & iter == 2 ~ "Laukiz1",
+                                 learner.id == "XGBOOST MBO CMIM" & iter == 3 ~ "Oiartzun",
+                                 learner.id == "XGBOOST MBO CMIM" & iter == 4 ~ "Laukiz2",
+
+                                 learner.id == "SVM MBO Relief" & iter == 1 ~ "Laukiz1",
+                                 learner.id == "SVM MBO Relief" & iter == 2 ~ "Oiartzun",
+                                 learner.id == "SVM MBO Relief" & iter == 3 ~ "Luiando",
+                                 learner.id == "SVM MBO Relief" & iter == 4 ~ "Laukiz2",
+
+                                 learner.id == "RF MBO Relief" & iter == 1 ~ "Luiando",
+                                 learner.id == "RF MBO Relief" & iter == 2 ~ "Laukiz2",
+                                 learner.id == "RF MBO Relief" & iter == 3 ~ "Laukiz1",
+                                 learner.id == "RF MBO Relief" & iter == 4 ~ "Oiartzun",
+
+                                 TRUE ~ as.character(iter))) %>% 
   dplyr::rename(Learner = learner.id) %>%
@@
   dplyr::rename(RMSE = rmse.test.mean) %>%
+  dplyr::mutate(fw.perc = fw.perc * 100) %>%
   dplyr::rename("Features (\\%)" = fw.perc) %>%
   dplyr::mutate(`Test Plot` = as.character(Plot)) %>%
-  dplyr::mutate(`Test Plot` = forcats::fct_recode(`Test Plot`,
-    Oiartzun = "4", Luiando = "3",
-    Laukiz1 = "1", Laukiz2 = "2"
-  )) %>%
   dplyr::mutate(Learner = forcats::fct_recode(Learner,
@@
   dplyr::mutate("\\#" = case_when(
-    `Test Plot` == "Laukiz1" ~ ceiling(`Features (\\%)` * 559),
-    `Test Plot` == "Laukiz2" ~ ceiling(`Features (\\%)` * 451),
-    `Test Plot` == "Luiando" ~ ceiling(`Features (\\%)` * 301),
-    `Test Plot` == "Oiartzun" ~ ceiling(`Features (\\%)` * 497),
+    `Test Plot` == "Laukiz1" ~ ceiling((`Features (\\%)` * 559) / 100),
+    `Test Plot` == "Laukiz2" ~ ceiling((`Features (\\%)` * 451) / 100),
+    `Test Plot` == "Luiando" ~ ceiling((`Features (\\%)` * 301) / 100),
+    `Test Plot` == "Oiartzun" ~ ceiling((`Features (\\%)` * 497) / 100),
   )) %>%
@@
     type = "latex",
-    caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted ascending by RMSE.",
+    caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
+\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available (7675).
+Results were estimated in a separate model tuning step, not within the main cross-validation comparison.",
     label = "tab:tune-perc-sel-features",

### 2021-06-21 docs/00-manuscripts/ieee/ieee-manuscript.tex
reply to laukiz2 non 5% defol values

@@
 Field surveys were conducted in September 2016 by experienced forest pathologists.
-Defoliation was measured via visual inspection using 5\% intervals at three height levels (bottom, mid, top) with the help of a dedicated score card.
+Defoliation was measured via visual inspection using 5\% intervals with the help of a dedicated score card.
+For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% interval of the other three plots (e.g.
+8.33 \%).
 Estimating the human observer error when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.

### 2021-06-21 analysis/eval-performance.Rmd
upd feature sel table caption

@@
     caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available (7675).
+\\texttt{#} denotes the absolute number of features selected and 'Features (\\texttt{%})' refers to the percentage relative to the overall features available (7675).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.",

### 2021-06-21 docs/00-manuscripts/ieee/tune-perc-sel-features.tex

@@
 \caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-	exttt{#} denotes the absolute number of features selected and 'Features (	exttt{%})' refers to the percentage relative to the overall features available (7675).
+\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available (7675).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.}

### 2021-06-21 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
upd tuning table

--- 2021-06-11 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
@@
 \caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-	exttt{#} denotes the absolute number of features selected and 'Features (	exttt{%})' refers to the percentage relative to the overall features available (7675).
+\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.}
 \label{tab:tune-perc-sel-features}
-\begin{tabular}{llrrr}
+\begin{tabular}{llrlr}
   \toprule
@@
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Relief}} & Luiando & 99.97268 & 301 & 41.73 \\ 
-   & Laukiz2 & 93.37086 & 422 & 17.12 \\ 
-   & Laukiz1 & 0.21573 & 2 & 33.21 \\ 
-   & Oiartzun & 0.00405 & 1 & 37.20 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Luiando & 30.84900 & 93 & 33.36 \\ 
-   & Laukiz1 & 98.12213 & 549 & 45.59 \\ 
-   & Oiartzun & 83.61298 & 416 & 41.01 \\ 
-   & Laukiz2 & 0.51526 & 3 & 15.24 \\ 
-  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.06200 & 1 & 35.49 \\ 
-   & Oiartzun & 4.10636 & 21 & 31.03 \\ 
-   & Luiando & 20.73514 & 63 & 37.19 \\ 
-   & Laukiz2 & 23.92820 & 108 & 14.89 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.21573 & 3/1249 & 33.21 \\ 
+   & Laukiz2 & 93.37086 & 1268/1357 & 17.12 \\ 
+   & Luiando & 99.97268 & 1507/1507 & 41.73 \\ 
+   & Oiartzun & 0.00405 & 1/1311 & 37.20 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.06200 & 1/1249 & 35.49 \\ 
+   & Laukiz2 & 23.92820 & 325/1357 & 14.89 \\ 
+   & Luiando & 20.73514 & 313/1507 & 37.19 \\ 
+   & Oiartzun & 4.10636 & 54/1311 & 31.03 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 98.12213 & 1226/1249 & 45.59 \\ 
+   & Laukiz2 & 0.51526 & 7/1357 & 15.24 \\ 
+   & Luiando & 30.84900 & 465/1507 & 33.36 \\ 
+   & Oiartzun & 83.61298 & 1097/1311 & 41.01 \\ 
    \bottomrule

### 2021-06-21 docs/00-manuscripts/ieee/ieee-manuscript.tex
clarify percentage values in table VII

@@
 The large differences of selected features per fold during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
+Values of selected feature percentages like 99.97268 \% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
+The value for the amount of features to select is initially defined by an internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1.
+
 While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF was able to achieve the best results with a relatively small amount of features.

### 2021-06-21 analysis/eval-performance.Rmd

--- 2021-06-17 analysis/eval-performance.Rmd
@@
 
+Example for RF on fold 1 (Luiando):
+
+- Percentage of selected features during tuning on all plots but Luiando: `bmr_inspect_tune[["results"]][["hr_nri_vi"]][["RF MBO Relief"]][["extract"]][[1]][["mbo.result"]][["x"]][["fw.perc"]]`: 99.972
+- Overall features available in training set: sum(Laukiz1 + Laukiz2 + Oiartzun) = 1507
+- Absolute number of selected features: 1507 * 0.99972 = 1507
+
 **RF**
 
-Plot 1: Luiando
-Plot 2: Laukiz2
-Plot 3: Laukiz 1
-Plot 4: Oiartzun
+- Plot 1: Luiando
+- Plot 2: Laukiz2
+- Plot 3: Laukiz 1
+- Plot 4: Oiartzun
 
@@
 
-Plot 1: Laukiz1
-Plot 2: Oiartzun
-Plot 3: Luiando
-Plot 4: Laukiz2
+- Plot 1: Laukiz1
+- Plot 2: Oiartzun
+- Plot 3: Luiando
+- Plot 4: Laukiz2
 
@@
 
-Plot 1: Luiando
-Plot 2: Laukiz1
-Plot 3: Oiartzun
-Plot 4: Laukiz2
+- Plot 1: Luiando
+- Plot 2: Laukiz1
+- Plot 3: Oiartzun
+- Plot 4: Laukiz2
 
@@
   dplyr::mutate("\\#" = case_when(
-    `Test Plot` == "Laukiz1" ~ ceiling((`Features (\\%)` * 559) / 100),
-    `Test Plot` == "Laukiz2" ~ ceiling((`Features (\\%)` * 451) / 100),
-    `Test Plot` == "Luiando" ~ ceiling((`Features (\\%)` * 301) / 100),
-    `Test Plot` == "Oiartzun" ~ ceiling((`Features (\\%)` * 497) / 100),
+    # numbers here are the respective sums of the training sets (without the mentioned plot acting as the test set)
+    `Test Plot` == "Laukiz1" ~ paste0(as.character(ceiling((`Features (\\%)` * 1249) / 100)), "/1249"),
+    `Test Plot` == "Laukiz2" ~ paste0(as.character(ceiling((`Features (\\%)` * 1357) / 100)),"/1357"),
+    `Test Plot` == "Luiando" ~ paste0(as.character(ceiling((`Features (\\%)` * 1507) / 100)),"/1507"),
+    `Test Plot` == "Oiartzun" ~ paste0(as.character(ceiling((`Features (\\%)` * 1311) / 100)),"/1311"),
   )) %>%
   dplyr::group_by(Learner) %>%
+  dplyr::arrange(Learner, `Test Plot`) %>% 
   dplyr::select(Learner, `Test Plot`, "Features (\\%)", "\\#", RMSE)
@@
     caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available (7675).
+\\texttt{#} denotes the absolute number of features selected and 'Features (\\texttt{%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.",

### 2021-06-21 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update tex of table VII

@@
 \caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
+\texttt{\#} denotes the absolute number of features selected and 'Features (\texttt{\%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.}

### 2021-06-22 analysis/eval-performance.Rmd

@@
   )) %>%
+  dplyr::select(-Plot) %>% 
   dplyr::mutate(RMSE = round(RMSE, 2))
@@
     type = "latex",
-    caption = "Test fold performances for learner SVM on the HR dataset without using a filter.
+    caption = "Test fold performances in RMSE (p.p.) for learner SVM on the HR dataset without using a filter, showcasing performance variance on the fold level.
 For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot.",
@@
     type = "latex",
-    caption = "Best ten results among all combinations, sorted in ascending order of RMSE",
+    caption = "Best ten results among all learner-task-filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE).",
     label = "tab:perf-top-10"
@@
     type = "latex",
-    caption = "Best performance of each learner across any task and filter method, sorted ascending by RMSE",
+    caption = "The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run.",
     label = "tab:best-learner-perf"
@@
     type = "latex",
-    caption = "Worst ten results among all combinations, sorted in decreasing order of RMSE",
+    caption = "Worst ten results among all learner-task-filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE).",
     label = "tab:perf-worst-10"
@@
     caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-\\texttt{#} denotes the absolute number of features selected and 'Features (\\texttt{%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
+\\texttt{\\#} denotes the absolute number of features selected and 'Features (\\texttt{\\%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.",

### 2021-06-22 docs/00-manuscripts/ieee/ieee-manuscript.tex

@@
 		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
-		\caption{Response variable \textit{defoliation} at trees for plots Laukiz1, Laukiz2, Luiando and Oiartzun.
-\texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation.}\label{fig:defol-distr}
+		\caption{Response variable "defoliation at trees" for plots Laukiz1, Laukiz2, Luiando and Oiartzun.
+\texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation, respectively.
+Values for Laukiz1, Luiando and Oiartzun were observed in 5\% intervals; for Laukiz2 defoliation was observed at multiple heights and then averaged, leading to smaller defoliation differences than 5\%.}\label{fig:defol-distr}
 	\end{center}
@@
 		\includegraphics[width=\textwidth] {study-area-hyperspectral.pdf}
-		\caption{Information about location, size and spatial distribution of trees for all plots used in this study.}\label{fig:study_area}
+		\caption{Study area maps showing information about location, size and spatial distribution of trees for all plots (Laukiz1, Laukiz2, Luiando, Oiartzun).
+The background maps used should give a visual impression of the individual plot area but do not necessarily represent the plot state during data acquisition.}\label{fig:study_area}
 	\end{center}
@@
 	\centering
-	\caption{List of filter methods used in this work}
+	\caption{List of filter methods used in this work, their categorization and scientific reference.}
 	\label{tab:filter-methods}
@@
 		\includegraphics[width=0.48\textwidth] {performance-results-1.pdf}
-		\caption{Predictive performance (RMSE) of models across tasks.
-Suffix 'MBO' means that model-based optimization was used for hyperparameter optimization.
-Abbreviations on the vertical axis refer to the combinations of feature sets on which each model was scored on.
-Labels represent the feature selection method (e.g.
+		\caption{Predictive performance in RMSE (p.p.) of models across tasks.
+Different feature sets are shown on the y-axis.
+Labels show the feature selection method (e.g.
 NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda').
-The second value of each label shows the RMSE value of the respective setting.}\label{fig:perf-result}
+The second value of each label shows the RMSE value (p.p.) and the standard error (SE) of the respective setting.}\label{fig:perf-result}
 	\end{center}
@@
 		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-no-filter-1.pdf}
-		\caption{Model performances in RMSE when using no filter method compared to all other filters across all tasks.}\label{fig:filter-effects-no-filter}
+		\caption{Model performances in RMSE across all tasks, split up in facets, when using no filter method (blue dot) compared to any other filter method (red cross) for learners RF, SVM and XGBoost (XG)}\label{fig:filter-effects-no-filter}
 	\end{center}
@@
 \pagebreak
-\section{Correlation among filter methods}
+\section{}
 
@@
 		\includegraphics[width=0.48\textwidth] {correlation-filter-nri-1.pdf}
-		\caption{Spearman correlations of NRI feature rankings obtained with different filters.}\label{fig:correlation-filters}
+		\caption{Spearman correlations of NRI feature rankings obtained with different filters.
+Filter names refer to the nomenclature used by the \texttt{mlr} R package.
+Underscores in names divide the terminology into their upstream R package and the actual filter name.}\label{fig:correlation-filters}
 	\end{center}
@@
 
-\section{Effect of different \texorpdfstring{\(n_{bins}\)}{nbins} values on filter 'information gain'}
+\section{}
 
@@
 		\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
-		\caption{Spearman correlations of rankings obtained with the information gain filter using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.}\label{fig:correlation-nbins}
+		\caption{Spearman correlations of rankings obtained with the information gain filter using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.
+Filter names refer to the nomenclature used by the \texttt{mlr} R package.
+Underscores in names divide the terminology into their upstream R package and the actual filter name.}\label{fig:correlation-nbins}
 	\end{center}
@@
 
-\section{Hyperparameter tuning ranges}
+\section{}
 
@@
 	\caption[t]{Hyperparameter ranges and types for each model.
-		Hyperparameter notations from the respective R packages were used.}
+		Hyperparameter notations from the respective R packages are shown.}
 	\begingroup\scriptsize
@@
 
-\section{Vegetation index formulas}
+\section{}
 

### 2021-06-22 docs/00-manuscripts/ieee/ieee-manuscript.tex
small cosmetic changes

@@
 		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-borda-filter-1.pdf}
-		\caption{Predictive performances in RMSE when using the Borda filter method compared to all other filters for each learner across all tasks.}\label{fig:filter-effects-borda}
+		\caption{Predictive performances in RMSE (p.p.) when using the Borda filter method (blue dot) compared to any other filter (red cross) for each learner across all tasks.}\label{fig:filter-effects-borda}
 	\end{center}
@@
 It was concluded that using a buffer of one meter can be a good compromise between the inclusion of information from too many surrounding trees, mapping a single tree crown accordingly accounting for a possible geometric offset.
-The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e. <= 1m).
+The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e. $<= 1m$).
 This results in all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
@@
 
-Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR}\cite{meng2018, kalin2019}.
+Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
 \cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
@@
 
-In summary, no studies which used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to tree health were found.
+In summary, no studies which used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to tree health could be found.
 This might relate to the fact that most environmental datasets are not high-dimensional.
 In fact, many studies use fewer than ten features and issues related to correlations are often solved manually instead of relying on an automated approach.
-This can be subjective and may limit the reproducibility of results.
+This approach might suffer from subjectivity and may limit the reproducibility of results.
 

### 2021-06-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize abstract

@@
 	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
-	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
-	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
+	Defoliation of trees (\%), derived from in-situ measurements in fall 2016, was modeled as a function of reflectance.
+	Variable importance was assessed using permutation-based feature importance.
+	Overall support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
 	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
-	For some combinations of learners and feature sets, filter methods achieved better predictive performances than the unfiltered feature sets, while ensemble filters did not have a substantial impact on performance.
-	Permutation-based feature importance estimated features around the red edge to be most important for the models.
-	However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the best performances.
+	For some combinations of learners and feature sets, filter methods achieved better predictive performances than using no feature selection.
+	Ensemble filters did not have a substantial impact on performance.
+	Permutation-based feature importance estimated features around the red edge to be most important.
+	However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the overall best performances.
 	More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
@@
 
-Penalized methods showed promising performances, especially when taking runtime into account.
-When removing features with a correlation of nearly 1, lasso and ridge are able to score performances around 31 p.p..
-
 A potential limiting factor in this study could be the upper limit of 70 iterations used for the XGBoost algorithm (hyperparameter \texttt{nrounds}), especially for feature sets including NRIs (\autoref{tab:hyperparameter_limits}).
@@
 
-The large differences of selected features per fold during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
-Values of selected feature percentages like 99.97268 \% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
-The value for the amount of features to select is initially defined by an internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1.
-
-While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF was able to achieve the best results with a relatively small amount of features.
-Realizing early that few features are needed during tuning to reach adequate performances can reduce the overall computational runtime substantially, especially when iterating over parameters such as $m_\textrm{try}$ whose optimum (and range) depends on the number of features.
-Hence, regardless of the potential advantage of using filters for increased predictive performance, it should be noted that these can have a strong positive effect on runtime, at least for RF in this study.
+Arbitrary seeming values of selected feature percentages like 99.97268 \% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
+The specific value to benchmark during tuning is defined by the internal surrogate learner of the MBO tuning method in this study and can be any numeric value between 0 and 1 (interpreted as percentage of available features).
+
+% filters always reduce computation time
+Realizing early during hyperparameter optimization that only few features are needed to reach adequate performances can reduce the overall computational runtime substantially.
+Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially on models which make use of hyperparameters that depend on the available number of features, such as RF $m_\textrm{try}$.
 
@@
 
+% defoliation studies and what data they used
 While most defoliation studies operate on the plot level using coarser-resolution multispectral satellite data \cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
@@
 
-Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
+% defoliation studies and what
+Studies focusing on tree-level defoliation mainly used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
 \cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
@@
 Incorporating such (both spectral and structural) could be a possible enhancement for future works.
+No studies were found modeling defoliation caused by \textit{Diplodia sapinea (Fr.) Fuckel} with remote sensing data and most studies focused on describing the tree conditions based on local sampling \cite{hlebarska2018, kaya2019}.
 
@@
 To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014a, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
-However, no study that made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
-Potential reasons for this gap could be an easier access to wrapper methods and a higher general awareness of such compared to filter methods.
+In general, applying feature selection methods on hyperspectral datasets has shown to be effective, regardless of the method used \cite{pal2010, keller2016}.
+However, no study that made explicit use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
+Potential reasons for this absence could be an easier programmatic access to wrapper methods and a higher general awareness of such compared to filter methods.
 Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.

### 2021-06-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
update intro

@@
 \IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
-Naturally, this coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
+This coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014, which serves as training and prediction data.
 At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from the open-source community.
 Scientists can nowadays process large amounts of (environmental) information with relative ease using various learning algorithms.
-This makes it possible to extend the benchmark comparison matrix of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would never have been explored otherwise \cite{ma2015}.
+This makes it possible to easily extend benchmark comparison matrices of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would not have been explored otherwise \cite{ma2015}.
 
@@
 
+% introduce veg indices
 Vegetation indices have shown the potential to provide valuable information when analyzing forest health \cite{jiang2014, adamczyk2015}.
@@
 In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands \cite{thenkabail2000}.
-Especially when working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
-
-% now that we introduced the importance of remote sensing indices for forest health,
-
+When working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
+
+% now that we introduced the importance of remote sensing indices for forest health, explain what is new in this study
 Despite their popularity in environmental modeling, there are no studies so far which used ML algorithms in combination with remote sensing data to analyze defoliation at the tree level.
@@
 The methodology of this study uses ML methods in combination with feature selection and hyperparameter tuning.
-In addition, feature importance and feature effects are evaluated.
+In addition, feature importance was analyzed.
 Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high dimensionality in environmental modeling \cite{trunk1979, xu2016}.
@@
 Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
+Instead of using wrapper feature selection methods, this study focuses on the use of (ensemble) filter methods which were directly integrated into the hyperparameter optimization step during model creation.
 
 % clarify the focus of this study
-This study shows how high-dimensional datasets can be handled effectively with ML methods while still being able to interpret the fitted models.
+Overall, this study aims to show how high-dimensional datasets can be handled effectively with ML methods while still being able to interpret the fitted models.
 The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
-However, this study clearly focuses on investigating the effects of filter methods and feature set types on predictive performance rather than on interpreting feature effects.
+Yet, this study focuses on investigating the effects of filter methods and feature set types on predictive performance rather than interpreting individual feature effects.
 

### 2021-06-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize data & study area

@@
 Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.} and \textit{Dothisthroma septosporum (Dorogin) M.
-Morelet} causing a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
-The last two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
+Morelet} causing (among others) needle blight, pitch canker and root diseases \cite{mesanza2016, iturritxa2017}.
+The first two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
 In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
@@
 The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-Oiartzun has the most observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
+Oiartzun consists of most tree observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
 All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
@@
 8.33 \%).
-Estimating the human observer error when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
+Estimating the human observer error of such surveys when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
 Even though no estimation error was recorded in this study, \cite{maclean1982} estimated human observer errors when assessing defoliation to range between 7\% - 18\%.
@@
 The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5th 2016.
-The images were taken by an AISA EAGLE-II sensor.
+Images were taken by an AISA EAGLE-II sensor.
 All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.

### 2021-06-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
update methods section

@@
 
-To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI}s were calculated.
-
+To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated.
 NRIs follow the \ac{OMNBR} concept of data-driven information extraction from narrow-band indices of hyperspectral data \cite{thenkabail2000,thenkabail2018}.
 While various index formulations such as band ratios or normalized ratios are available, indices involving the same bands are strongly correlated.
-Since the widely-used NDVI index belongs to the family of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following NDI formula to combine all pairs of reflectances:
+Since the widely-used NDVI index belongs to the family of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following normalized difference index (NDI) formula to combine all pairs of reflectances:
 
@@
 
-To account for geometric offsets within the hyperspectral data, which were reported with up to one meter from \ac{ICGC}, a buffer of one meter around the centroid of each tree was used when extracting the reflectance values.
+To account for geometric offsets within the hyperspectral data, which were reported to be potentially up to one meter from \ac{ICGC}, a buffer of one meter around the centroid of each tree was used when extracting the reflectance values.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
-All of those pixels formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
+The pixel values of such were averaged and formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
 In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated.
@@
 In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
-Due to the focus on filter methods in this manuscript, only this sub-group of feature selection methods will be introduced in greater detail in the following subsections.
+Due to the focus on filter methods in this manuscript, only this sub-group of feature selection methods will be introduced in greater detail in the following sections.
 
@@
 The concept of filters originates from the idea of ranking features following a score calculated by an algorithm \cite{guyon2003}.
-Some filter methods can only deal with specific types of variables (numeric or nominal).
+Some filter methods can only deal with specific types of variables (e.g. numeric or nominal).
 Filters only rank features, they do not decide which covariates to drop or keep \cite{drotar2015}.
-The selection which features to keep for model fitting is usually done within the optimization phase of the model fitting, along with the hyperparameter tuning.
+The selection which features to keep for model fitting is usually done within the optimization phase during model fitting, along with the hyperparameter tuning.
 Essentially, the number of covariates in the model is treated as a additional hyperparameter of the model.
-The goal is to optimize the number of ranked features to the point at which the model achieves the best performance.
+The idea is to optimize the number of ranked features to the point at which the model achieves the best performance.
 
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can possibly enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
 Ensemble filters align with the recent rise of the 'ensemble' approach in ML which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
-In this work the 'Borda' ensemble filter was applied \cite{drotar2017a}.
+In this work the 'Borda' ensemble filter was used \cite{drotar2017a}.
 Its final feature order is determined by the sum of all single filters ranks.
@@
 	\item  Random Forest (RF)
-	\item  Penalized Regression (both L1 (lasso) and L2 (ridge))
+	\item  Penalized Regression (L1 (lasso) and L2 (ridge))
 	\item  Support Vector Machine (SVM, RBF Kernel)
@@
 
-RF and SVM are well established algorithms widely used in (environmental) remote sensing.
+RF and SVM are well established algorithms and widely used in (environmental) remote sensing.
 Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmark studies in recent years.
@@
 The selected hyperparameter settings are shown in appendix \autoref{tab:hyperparameter_limits}.
-In addition the code which generated the matrix is available in the research compendium of this study (\url{10.5281/zenodo.2635403}).
+In addition code and data are available in the research compendium of this study (\url{https://doi.org/10.5281/zenodo.2635403}) to support open and reproducible science.
 
@@
 
-Some individual features were finally removed from these feature sets if they were numerically equivalent to another feature based on their pairwise correlation being greater than $1 - 10^{-10}$.
+Some individual features of such were removed before using the datasets for modelling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
 This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).
@@
 \ac{MBO} may outperform methods that do not use information from previous iterations, such as random search or grid search \cite{bergstra2012}.
-Tuning ranges are shown in \autoref{tab:hyperparameter_limits}.
+Tuning ranges used in this work are shown in \autoref{tab:hyperparameter_limits}.
 
 To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage \cite{binder2020}.
-For \ac{PCA}, the number of principal components was tuned instead.
+For \ac{PCA}, the number of principal components was tuned.
 %The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_{try} = p^x_{try}$.
-The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_\textrm{try} = p_\textrm{sel}^t$ as a function of the number of selected features, $p_\textrm{sel}$.
+The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_\textrm{try} = p_\textrm{sel}^t$, a function of the number of selected features, $p_\textrm{sel}$.
 It was thus tuned on a logarithmic scale by varying $t$ between 0 (i.e. $m_\textrm{try} = 1$) and 0.5 (i.e. $m_\textrm{try}=\sqrt{p_\textrm{sel}}$).
-This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features left after optimizing the feature percentage during tuning.
+This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features that were left after optimizing the feature percentage during tuning.
 
@@
 All tasks of this study were conducted using the open-source statistical programming language R \cite{rcoreteam2019}.
-A complete list of all R packages used in this study can be found in linked repositories.
+A complete list of all R packages used in this study can be found in the linked repositories mentioned in the next paragraph.
 Due to space limitations only the selected packages with high impact on this work will be explicitly cited.
@@
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
-Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://github.com/pat-s/2019-feature-selection}), listing more side-analyses that were carried out during the creation of this study.
+Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://pat-s.github.io/2019-feature-selection}), listing more side-analyses that were carried out during the creation of this study.
 
@@
 
-\subsection{PCA of feature sets}
+\subsection{PCA-based variance analysis of feature sets}
 
@@
 % - MBO tuned penalized methods are substantially worse
-Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.).
-SVM showed no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p (\autoref{fig:perf-result}).
+Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.) (\autoref{fig:perf-result}).
+SVM showed almost no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p (\autoref{fig:perf-result}).
 SVM showed the best overall performance with a mean difference of around three p.p to the next best model (XGBoost) (\autoref{tab:best-learner-perf}).

### 2021-06-28 analysis/eval-performance.Rmd
workflowr::wflow_publish("analysis/eval-performance.Rmd")

@@
   dplyr::mutate(`Test Plot` = as.character(`Test Plot`)) %>%
-  dplyr::mutate("\\#" = case_when(
+  dplyr::mutate("Features (\\#)" = case_when(
     # numbers here are the respective sums of the training sets (without the mentioned plot acting as the test set)
@@
   dplyr::arrange(Learner, `Test Plot`) %>% 
-  dplyr::select(Learner, `Test Plot`, "Features (\\%)", "\\#", RMSE)
+  dplyr::select(Learner, `Test Plot`, "Features (\\%)", "Features (\\#)")
 
@@
     caption = "Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-\\texttt{\\#} denotes the absolute number of features selected and 'Features (\\texttt{\\%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
+'Features (\\texttt{\\#})' denotes the absolute number of features selected and 'Features (\\texttt{\\%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.",
     label = "tab:tune-perc-sel-features",
-    digits = c(0, 0, 0, 5, 0, 2)
+    digits = c(0, 0, 0, 5, 0)
   )

### 2021-06-28 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
update eval-perf

--- 2021-06-21 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
@@
 \caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name.
-\texttt{#} denotes the absolute number of features selected and 'Features (\texttt{%})' refers to the percentage relative to the overall features available (7675).
+'Features (\texttt{\#})' denotes the absolute number of features selected and 'Features (\texttt{\%})' refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311).
 Results were estimated in a separate model tuning step, not within the main cross-validation comparison.}
 \label{tab:tune-perc-sel-features}
-\begin{tabular}{llrrr}
+\begin{tabular}{llrl}
   \toprule
-Learner & Test Plot & Features (\%) & \# & RMSE \\ 
+Learner & Test Plot & Features (\%) & Features (\#) \\ 
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Relief}} & Luiando & 99.97268 & 301 & 41.73 \\ 
-   & Laukiz2 & 93.37086 & 422 & 17.12 \\ 
-   & Laukiz1 & 0.21573 & 2 & 33.21 \\ 
-   & Oiartzun & 0.00405 & 1 & 37.20 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Luiando & 30.84900 & 93 & 33.36 \\ 
-   & Laukiz1 & 98.12213 & 549 & 45.59 \\ 
-   & Oiartzun & 83.61298 & 416 & 41.01 \\ 
-   & Laukiz2 & 0.51526 & 3 & 15.24 \\ 
-  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.06200 & 1 & 35.49 \\ 
-   & Oiartzun & 4.10636 & 21 & 31.03 \\ 
-   & Luiando & 20.73514 & 63 & 37.19 \\ 
-   & Laukiz2 & 23.92820 & 108 & 14.89 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.21573 & 3/1249 \\ 
+   & Laukiz2 & 93.37086 & 1268/1357 \\ 
+   & Luiando & 99.97268 & 1507/1507 \\ 
+   & Oiartzun & 0.00405 & 1/1311 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.06200 & 1/1249 \\ 
+   & Laukiz2 & 23.92820 & 325/1357 \\ 
+   & Luiando & 20.73514 & 313/1507 \\ 
+   & Oiartzun & 4.10636 & 54/1311 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 98.12213 & 1226/1249 \\ 
+   & Laukiz2 & 0.51526 & 7/1357 \\ 
+   & Luiando & 30.84900 & 465/1507 \\ 
+   & Oiartzun & 83.61298 & 1097/1311 \\ 
    \bottomrule

### 2021-06-28 docs/00-manuscripts/ieee/ieee-manuscript.tex
resubmission review 2

@@
 		of Geography, geographic information science group, Friedrich Schiller University Jena, Germany.}% <-this % stops a space
-	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilian-Universität in Munich, Germany.}% <-this % stops a space
+	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilians-Universität in Munich, Germany.}% <-this % stops a space
 	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Spain.}% <-this % stops a space
@@
 % - SVM unaffected
-When comparing the usage of filters against using no filter at all, there no instance where a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
+When comparing the usage of filters against using no filter at all, there were no instances where a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
 For SVM, all filters and \enquote{no filter} achieved roughly the same performance on all tasks.
@@
 % - never the best
-The Borda filter achieved only in one instance (RF on HR-NRI-VI) the best performance among the used filters (\autoref{fig:filter-effects-borda}).
+The Borda filter achieved in only one instance (RF on HR-NRI-VI) the best performance among the used filters (\autoref{fig:filter-effects-borda}).
 For RF and XGBoost it most often ranked within the first 50\% with respect to all filters of a specific task.
@@
 While for RF least features were selected when Luiando or Oiartzun were the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
-RF and XGBoost used very few features only when Oiartzun was the test set (one and three, respectively) whereas this was the plot for which SVM used most features ($n = 119$).
-Overall, SVM used the least features across all plots among all learners, with 24\% for test set Oiartzun being the highest single plot value.
+RF and SVM used very few features only when Laukiz1 was the test set, (three and one, respectively) whereas this was the plot for which XGBoost used almost all features ($n = 1226$).
+Overall, SVM used the least features across all plots among all learners, with 24\% for test set Laukiz2 being the highest single plot value.
 In general every learner behaved quite differently for each plot and no overall pattern could be observed.

### 2021-08-18 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   mutate(Filter = replace(Filter, is.na(Filter), "No Filter")) %>%
-  select(-iter, -`Learner ID`, -expvar, -rsq) %>%
+  select(-iter, -`Learner ID`, -rsq) %>%
   relocate(RMSE, .after = Filter) %>%
@@
   relocate(`SE`, .after = RMSE) %>%
-  select(-iter, -expvar)
+  select(-iter)
 
@@
 results_aggr <- df_perf %>%
+  filter(learner.id != "regr.featureless") %>% 
   mutate(filter = replace(filter, is.na(filter), "NF")) %>%
@@
   filter(learner_group != "Lasso-MBO") %>%
+  filter(learner.id != "regr.featureless") %>%
   # mutate(filter = replace(filter, "No Filter", "NF")) %>%
@@
   ungroup() %>%
-  # we need to reverse the order on porpuse here so that ggplot reverses it
+  # we need to reverse the order on purpose here so that ggplot reverses it
   # again later
-  mutate(learner_group = fct_relevel(learner_group, "XGBoost", "SVM", "RF"))
+  mutate(learner_group = fct_relevel(learner_group, "XG", "SVM", "RF"))
 
@@
   filter(learner_group != "Lasso-MBO") %>%
+  filter(learner.id != "regr.featureless") %>%
   mutate(learner_group = recode_factor(learner_group, `XGBoost` = "XG")) %>%
@@
   # again later
-  mutate(learner_group = fct_relevel(learner_group, "XGBoost", "SVM", "RF"))
+  mutate(learner_group = fct_relevel(learner_group, "XG", "SVM", "RF"))
 

### 2021-08-18 docs/00-manuscripts/ieee/performance-best-per-learner.tex
include regr.featureless in benchmark

@@
 \centering
-\caption{Best performance of each learner across any task and filter method, sorted ascending by RMSE} 
+\caption{The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run.}
 \label{tab:best-learner-perf}
@@
   \hline
-1 & HR-NRI-VI & SVM & Relief & 28.09 & 18.95 \\ 
-  2 & HR & XGBoost & CMIM & 30.88 & 14.47 \\ 
-  3 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
-  4 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
-  5 & HR & RF & Relief & 32.93 & 13.17 \\ 
+1 & HR & SVM & Relief & 28.12 & 19.12 \\ 
+  2 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
+  3 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
+  4 & HR & regr.featureless & No Filter & 31.16 & 15.03 \\ 
+  5 & HR & RF & Relief & 31.25 & 14.96 \\ 
+  6 & HR & XGBoost & Car & 34.07 & 13.57 \\ 
    \hline

### 2021-08-18 code/070-benchmark-matrix.R

@@ -41,7 +41,7 @@ benchmark_plan <- drake_plan(
   # 6     - hr_nri_vi
   benchmark_no_models = target(
     benchmark(
-      learners = tune_wrappers_mbo,
+      learners = tune_wrappers_mbo1[[1]],
       tasks = task_reduced_cor,
       models = FALSE,
       keep.pred = TRUE,
@@ -49,13 +49,12 @@ benchmark_plan <- drake_plan(
       show.info = TRUE,
       measures = list(
         setAggregation(rmse, test.mean),
-        setAggregation(rsq, test.mean),
-        setAggregation(expvar, test.mean)
+        setAggregation(rsq, test.mean)
       )
     ),
     dynamic = cross(
       tune_wrappers_mbo1,
-      task_reduced_cor
+      task_reduced_cor[[1]]
     )
   ),
 
@@ -96,7 +95,6 @@ benchmark_plan <- drake_plan(
       measures = list(
         setAggregation(rmse, test.mean),
         setAggregation(rsq, test.mean),
-        setAggregation(expvar, test.mean)
       )
     ),
     dynamic = cross(learners_keep_models[[1]], task_reduced_cor)

### 2021-08-18 docs/00-manuscripts/ieee/reviews/review2/reviewer2-reply.md

# Reviewer 2 - Iteration 2

> The manuscript deals with the optimization of input data for the detection of different damage classes in trees. The different feature selection approaches should simplify the evaluation of hyperspectral data and lead to better results. The topic and the methods are exciting and deal with current interesting problems. The manuscript is well structured and well written even if some minor shortcomings are noticeable.

## General comments

> The biggest problem I see in the hyperspectral data which were analyzed. The spectral signatures are too noisy and do not correspond to any physical basis. Such ups and downs are not justifiable and therefore the data must be filtered/smoothed beforehand - as usual in other studies.

Image preprocessing is an important step that was handled in our case by an experienced external service provider, which applied any necessary radiometric, geometric and atmospheric corrections.
Additional preprocessing steps in the spectral and spatial domains may potentially be beneficial ([Vaiphasa 2006](https://www.sciencedirect.com/science/article/pii/S0924271605001012?via%3Dihub)), but smoothing techniques also have the potential to introduce new artifacts, for example by propagating errors from one spectral band into adjacent ones.

In this study we opted for a data-driven approach since feature selection techniques are designed to filter spectral characteristics that most reliably predict the response.
The chosen machine learning models are furthermore capable of weighting or averaging the numerous features either explicitly or implicitly, depending on model architecture.
It would be a topic for an interesting follow-up study to determine whether model-based spectral smoothing or data-driven approaches are more effective at using the available information.
We believe that this goes well beyond the scope of our contribution.

We would like to note that all code and data related to this publication is made available on Zenodo, which will facilitate future research related to preprocessing steps or other more specific aspects.

> I would like to see examples of all damage classes from both the field response and the canopy delineated in the EO data.

We attached some pictures taken in the field to our reply.
Note that these should not be generalized as every tree could look different, especially with respect to defoliation caused by Diplodia (which causes brown coloring of the shoots).

> It is not clear how the hyperspectral information was extracted for each tree. I expect that they had GNSS points for each tree. The approach of using a buffer to avoid the use of non-tree pixels is not correct. If the point is outside the tree crown also the have of the four pixels will be outside. On the other hand, if the pixel is on the tree crown the buffer can lead to include pixels from outside. Therefore, the tree crowns should (manual) delineated, and e.g. the mean of each object should be used!

Thanks for discussing the use of the buffer again.
The tree positions were referenced via GPS sensors in the field, correct.
We believe that the use helps to reduce possible positional errors and disagree that this approach is not correct.
Manual delineation is an alternative approach which is fine to use but which we did not do in this study.

> The discussion chapter is mainly a description of the results and doesn’t include a real discussion. I have to be revised and some parts should be shifted to the results. I want to see more comparison with other studies, both dealing with feature selection and hyperspectral data but also some tree health studies should be included.

Thanks for sharing your thoughts related to the discussion section.
We agree with the reviewer that the first part of the discussion deals with the results of this study.
However, rather than describing the results as in the "Results" section, results are discussed and interpreted in the "Discussion" section.
We believe that this is a good structural division and would like to avoid mixing result description and interpretation.
With respect to discussing other studies, there is already a dedicated paragraph titled "Comparison to other studies" which discusses similar studies and includes 18 references.
In a paragraph of this section we also note the following:

"In summary, no studies which used filter methods for FS or made use of NRI indices in their work and had a relation to tree health could be found.
This might relate to the fact that most environmental datasets are not high-dimensional.
In fact, many studies use fewer than ten features and issues related to correlations are often solved manually instead of relying on an automated approach."

The terminology "real discussion" used by the reviewer furthermore indicates that their view of structuring a discussion might differ from ours and that we most likely cannot match the reviewers desired discussion structure perfectly.
However, we believe that the general discussion structure of this manuscript is acceptable.

Nevertheless, to showcase our willingness to incorporate constructive suggestions, we added more references to studies which used feature selection in combination with hyperspectral data ([Pal 2010](https://ieeexplore.ieee.org/abstract/document/5419028?casa_token=MpK3WG8ZqLEAAAAA:E7cICk6rfmT8H8jf4rzyFLPR0ijL5rm_idUOix_WvJyOB6uiZqAwanqkrSd8cIpvHT4lHxMMcd_w), [Keller 2016](https://ieeexplore.ieee.org/abstract/document/8071759)), added two more references showcasing that most studies only describe the loss of defoliation in the field rather than modeling it ([Hlebarska 2018](https://www.cabdirect.org/cabdirect/abstract/20183376613), [Kaya 2019](https://doaj.org/article/3d9b8162586c480caea4dd3e04753868)), and pruned sections "Model differences" and "Performance vs. plot characteristics" to have less potential overlap with the "Results" section.

> Abbreviations introduced should be used consistently afterward.

We harmonized the use of abbreviations.

> The citation style is sometimes strange and I am not sure if it meets the journal's requirements. If there are multiple sources, I would put all numbers in one square bracket. If the source is used in the body text, I think the name of the author should be also given.

We've used the LaTeX citation style as recommended by the author instructions of the Journal.
We are using the following options

```tex
\documentclass[final]{IEEEtran}
\usepackage[noadjust]{cite}
\bibliographystyle{IEEEtran}
```

which should be accepted by the journal and community judging from [this Texexchange post](https://tex.stackexchange.com/a/44086/104937).

> Tables and Figures have to be understandable stand-alone. Therefore, the captions are often not sufficient. Please revise it and provide more details. Also, units are often missing and abbreviations are not explained.

We have updated the captions for the following figures and tables and hope they are more descriptive now and understandable standalone:

- Fig. 1: "Response variable "defoliation at trees" for plots Laukiz1, Laukiz2, Luiando and Oiartzun. n corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation, respectively. Values for Laukiz1, Luiando and Oiartzun were observed in 5\% intervals, for Laukiz2 defoliation was observed at multiple heights and then averaged, leading to smaller defoliation differences than 5\%."
- Fig. 2: "Study area maps showing information about location, size and spatial distribution of trees for all plots (Laukiz1, Laukiz2, Luiando, Oiartzun). The background maps used should give a visual impression of the individual plot area but do not necessarily represent the plot state during data acquisition."
- Fig. 3: "Predictive performance in RMSE (p.p.) of models across tasks. Different feature sets are shown on the y-axis. Labels show the feature selection method (e.g. NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda'). The second value of each label shows the RMSE value (p.p.) and the standard error (SE) of the respective setting."
- Fig. 4: "Model performances in RMSE across all tasks, split up in facets, when using no filter method (blue dot) compared to any other filter method (red cross) for learners RF, SVM and XGBoost (XG)"
- Fig. 5: "Predictive performances in RMSE (p.p.) when using the Borda filter method (blue dot) compared to any other filter (red cross) for each learner across all tasks."
- Table II: "List of filter methods used in this work, their categorization and scientific reference."
- Table III: "Best ten results among all learner-task-filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE)."
- Table IV: "Worst ten results among all learner-task-filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE)."
- Table V: "The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run."
- Table VI: "Test fold performances in RMSE (p.p.) for learner SVM on the HR dataset without using a filter, showcasing performance variance on the fold level. For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot."
- Appendix A: "Spearman correlations of NRI feature rankings obtained with different filters. Filter names refer to the nomenclature used by the mlr R package. Underscores in names divide the terminology into their upstream R package and the actual filter name."
- Appendix B: "Spearman correlations of rankings obtained with the information gain filter using different $n_{bins}$ values for discretization of the numeric response. Filter names refer to the nomenclature used by the mlr R package. Underscores in names divide the terminology into their upstream R package and the actual filter name."

> The quality of the maps should be improved.

We've made some small tweaks to the maps even though we think it was fine before:

- Removed red plot indicators on plot maps
- Referenced Bing Maps as the image source
- Refined legend icon size
- Added Burgos district to overview map (Burgos is not part of the Basque Country)

> Page 2, Column 2, Line 22: Please use the already introduced abbreviation ML

Thanks, we replaced all instances of machine-learning/machine learning with the abbreviation ML.

> P2, C2, L25: citation style is strange

We've used the LaTeX citation style as recommended by the author instructions of the Journal.
We are using the following options

```tex
\documentclass[final]{IEEEtran}
\usepackage[noadjust]{cite}
\bibliographystyle{IEEEtran}
```

which should be accepted by the journal and community judging from [this Texexchange post](https://tex.stackexchange.com/a/44086/104937).

> P2, C2, L25: I didn’t get it, please revise the sentence.

Thanks, we clarified the sentence as follows:
"The ability of predicting into unknown space qualifies ML algorithms as a helpful tool for such environmental analyses."

> P2, C2, L46: the index type is well known, but you are right, often not all possible combinations were tested.
> P3, C1, L3, L17: ML instead of machine-learning

Thanks, we replaced all instances of machine-learning/machine learning with the abbreviation ML.

> P3, C1, L44: Are you sure that all these pathogens are invasive?

We removed the word "invasive" as it is not relevant for this methodological study.

> P3, C1, L54: How was this data obtained? Please add some more details.

We have added "by visual inspection of experts in the field" to make the data collection even more clear.

> P3, C2, Figure 1, L36: The 5% intervals are clear for three test plots. Four Laukiz 2 the intervals between 0 and 15 are smaller. Please clarify. I am also not sure if you can use these values for a regression approach because they are not really metric data.

Thanks for taking such a close look at the data.
The observed values for Laukiz2 are indeed smaller than for all other plots.
The reason is that for plot Laukiz2 defoliation was observed at three different heights (bottom, middle, top) with the final value being composed out of the mathematical average of these three values.
In contrast, the observation data from the other plots consisted of a single value only.
This information was missing so far in the manuscript, we have rephrased the respective part as follows:

"Defoliation was measured via visual inspection using 5\% intervals with the help of a dedicated score card.
For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% interval of the other three plots (e.g. 8.33 \%)."

> P3, C2, L29: Standardize the spelling of the plot names (with or without spaces).

We have not found any instances of Laukiz1 or Laukiz2 with spaces between "Laukiz" and its respective count ID, also not in the sentence you referenced.

> P4, Figure 2: Please revise the maps. Add a detailed view of the tree crowns. What is the background image? Single coordinates are not meaningful.

We do not have custom images showing the plots from a birds-eye perspective; note that the hyperspectral image resolution is 1 m, which does not provide any visual detail at the crown level.
The background images shown are "Bing Aerial" maps queried from within QGIS.
The purpose of these is not to show a high-resolution bird-eye view of the plots but just to give a rough overview of the plot area and its surroundings.

> P4, C1, L44: Spatial resolution instead of geometric resolution?

The difference between these terms is not fully clear, often is assumed that they are the same.
See also [this discussion](https://gis.stackexchange.com/a/170324/43409).

> P5, C1, L15: I traditionally use wrapper methods that obtain sufficient results. Therefore, a comparison of the used filter methods with wrapper methods would be very interesting, both in terms of time and accuracy.

We agree that the comparison of both concepts is an interesting topic, though it is out of scope for this study as the focus is on filter methods.
A literature research for this dedicated question might give more interesting insights.

> P5, C1, L21: I am not a statistician but for me, a filter changes something in the data and does not only rank the original data, but maybe I am wrong.

Filters in the sense of machine learning do not change the data, they simply create a ranking of variables based on their heuristic.
Such rankings can then be used for further variable selection or combined with rankings from other filters (in an ensemble filter scenario).

> P5, C1, L58: Not clear to me. I expect that specific filters work better for specific algorithms. Therefore, I am skeptical that the combination is better if every filter results in different ranks?

The effectiveness of filters is a function of the algorithm, dataset and tuning method used.
Varying one of these parameters might yield substantially different results.
Therefore a benchmarking matrix covering multiple settings and/or combining filter methods is usually used (as also done in this work).
This paragraph refers to the weighting of single filter in the ensemble compilations, which is an important precondition for an unbiased application.

> P5, C2, L41: use the already introduces abbreviation RF

Thanks, we replaced all instances with the acronym.

> P6, C1, L6: The last Table was II, what is the reason to place the next table after five other tables?

Table VIII is placed in the appendix, hence the numbering. We appended the term "appendix" to make this more clear.

> P6, C1, L21: Include Author’s name in the sentence

We manually added it even though it seems not officially supported by the citation style.

> P6, C1, L33-38: I don’t agree. From my experience, the feature selection should be connected to the following method. Therefore, the importance of the features is not the same for all models. Also, high correlated input data are problematic for some algorithms but not for others.

Thank your for this comments, which we believe points to a misunderstanding since this paragraph was not very clear.
The only point we make in this paragraph is that features that are obviously completely redundant were removed.
No filter and no model will be able to retrieve useful information from features that have a correlation of 1 with another feature.
We changed the paragraph to read as follows: 'Some individual features were finally removed from these feature sets if they were numerically equivalent to another feature based on their pairwise correlation being greater than $1 - 10^{-10}$.'"

> P6, C2, L47: not clear to me. Please explain.

For the permutation-based feature importance on the HR dataset, a learner is required, i.e. a ML model.
We used the learner which performed best on this task, i.e. SVM.

> P7, C1, L27: What is PCA-based correlation analysis?

Thanks for pointing out that the wording is imprecise.
We changed the subsection title to 'PCA of feature sets' and shortened the paragraph, which is intended to give a reader a very simplified impression of the complexity of feature space, as requested by a reviewer in the previous round. We changed the text as follows:

"PCA was used to assess the complexity of the three feature sets.
Depending on the feature set, 95\% of the variance is explained by two (HR), twelve (VI) and 42 (NRI) \ac{PC}s.
HR features are therefore highly redundant, while the applied feature transformations enrich the data set, at least from an exploratory linear perspective."

> P7, C1, L44: use the introduced abbreviation p.p.

Thanks, using the abbreviation now.

> P8, C1, Table III, and IV: From my point of view the Tasks with all input data should outperform the others if the feature selections work well. Therefore, I prefer wrapper algorithms and again I think they should be included for comparison (e.g. stepwise feature reduction based on the RF importance measure).

Thanks for suggesting the use/comparison of wrapper methods.
While feature selection wrappers can provide good results and enhance the predictive performance of a model, they also come with certain costs.
They usually take way more time to compute and used with a computational intensive algorithm (e.g. XGBoost or SVM), fitting times become quite large.
In contrast, filter methods are more efficient and can be cached after their first calculation.
Also we wanted to avoid making a subjective pre-selection of feature selection methods in this study for an unbiased comparison and the evaluation of new ensemble filters (Borda).
Comparing different feature selection strategies/categories (i.e. wrapper vs. filter) would be subject of a different study which has access to an unreasonably larger amount computational resources when operating with datasets of 7k or more features.
Besides, such studies already exist, even with a subfocus on specific subareas of modeling ([Talavera 2005](https://link.springer.com/chapter/10.1007/11552253_40)).
However, more recent studies try to make use of hybrid ensemble approaches which combine filter and wrapper methods ([Chen et al. 2020](https://onlinelibrary.wiley.com/doi/full/10.1111/exsy.12553)).

> The order in Table 3 is not clear, 2 to 10 have the same RMSE. Also, the different SE in 8 is not clear.

The results in table VIII are simply the model results.
They are highly similar for the SVM learner, which can happen and is not unusual.
Learner SVM on the HR-NR-VI dataset has a slightly higher standard error.

>Table IV: Why does PCA not work and again it is not clear why PCA based on VI only outperforms PCA based on HR-NRI-VI. Also, the better results (compared to PCA) without filters are not clear.

Applying a filter or using more variables does not guarantee to result in a better performance.
Exploring these differences was one motivation for the extensive benchmark matrix of this study.
Often enough, model results do not follow a pattern one might have from previous modeling experiences.
Explaining why model X is better than model Y can often not be fully determined.
The case mentioned by you here might be one case for which this applies but there might be more to find in this extensive benchmark matrix, with the potential increasing when adding more settings to the matrix.

Judging from the results, it looks like PCA does not go very well with XGBoost - in our view this is everything one can draw from these non-optimal results in Table IV.

> P8, C1, Table VI: Why did you present these results based on the models without filters? I also want to see the best results using filters.

The purpose of this table is only to show the variability of results between the folds without making a deliberate choice which filter to use to estimate such.
We use this table in the first paragraph of Section IV. B. where we want to emphasize that performance differences between the plots are large.
At this point there is no added value to us in showing this information for all learner/filter/dataset combinations, given that it would require a table with four rows for each.

> P8, C1, L50: p.p. instead of percentage points

Thanks, using the abbreviation now.

> P8, C2 Table VII: Please explain Features (%) and values of 0.999, Values are sorted by test plots not ascending RMSE; # means the number of input data? RMSE Values are strange because higher as without filter and also the completely different results for the test plots are not clear to me: Luiando achieved the best result without filter, Laukiz 2 (the worst without filter) the best result with RF, Oiartzun the best with XGB and SVM

Thanks for discussing the results of Table VII.

`#` denotes the absolute number of features used of the respective training set. We've updated the caption.
`features (%)` refer to the relative percentage of features selected during tuning judging from the total amount of features of the respective training set which is the sum of all plots but the test set.
However values were not in percent, we've updated them now. Thanks.
Values were sorted by test plot name which did not match the table caption.
Now the table is sorted by learner and test plot.
Thanks.

In addition we discovered that some results were assigned to the wrong plots names due to an issue in the code generating the table.
We've updated the table.

The results shown here refers only the HR-NRI-VI task and we used the best learner combinations for this feature set.
We assume that by saying "without filter" you mean that almost all features were used for the specific plot and learner combination as there are no "no filter" combinations present in this table.
This table has the sole purpose of showcasing the variance within plots with respect to selected features during tuning and how these might differ among learners.

> P8, C2, Figure 3, include a unit for RMSE, change order of the learners (and also colors) Lasso and Ridge should be next to each other. Why did you include MBO in the name?

Thanks for suggesting possible improvements to Figure 3.
We omitted the unit for RMSE in all other tables/figures.
Adding it only here would make things inconsistent.
We believe that it should be clear to readers of the study that the unit is percentage points as it is mentioned in various places.

Lasso and Ridge were appended with a "MBO" suffix to highlight the use of MBO for internal optimization in contrast to using their respective integrated tuning algorithms.
The added value might not be fully clear, hence we removed the suffix.

We have changed the color palette to "Viridis", a palette with support for colorblind people.

> P9, Figure 4: The filters improved the results mainly for the task HR. For all others, the no Filter results are within the best results. Please explain how can a filter worsen the performance?

Thanks for the detailed look on the results.
There is no guarantee that the use of filters always improves the result, even though the selection of features is optimized via tuning.
As mentioned and discussed in the study, the outcome is a combination of learner+dataset+filter and using a subset of features might not always result in the best performance.
Also the ranking of certain filters might simply not have a positive effect and result in features being removed even though they would have been helpful to the model in the end.

> P9, Figure 5: Why did you highlight Borda Filter?

 We highlighted the Borda filter because the use of ensemble filters (i.e. the Borda filter) is a central evaluation point of this work.

> P10, Figure 6: As I mentioned in the general comments you have to apply smoothing before analyzing the spectral signatures. The peak around 950nm is also very strange.

Please see our previous comment on this topic.

> What does scaled Reflectance or in the caption normalized reflectance mean?

Normalized reflectance between [0, 1] as shown on the axis title.

> P11, C1, L47: not instead of no

Thanks.

> P11, C1, L56-58: PCA obtained the worst results!

Thanks for double-checking the results.
Indeed PCA scored the worst results and this is what we wrote in l.52 - 56.
It also reduced model fitting times, which is fair to mention.

> P11, C2, L10-11: Instead of “not surprisingly” I would prefer citing other studies which obtained similar results.

Using references to support statements is usually a good idea.
In this case however, we believe that this statement does not need to backed by additional references.

> P11, C2, L36-37: Sentence is not necessary.

Thanks, we removed it.

> P11, C2, L48: Why only in almost all cases. I don’t see an option to get less or more pixels. Anyway, this approach is not a solution for your mentioned problem of pixels outside the tree. On the contrary, the method leads to the fact that neighboring pixels are included.

We agree that there are no possibilities to get a different number of pixels than exactly four and removed "almost" from the sentence.

As mentioned in previous comments, we still believe that this approach is a good compromise for the issue at hand.

> P13, C1, 56: Please avoid citations in the conclusion.

We tried to avoid or at least minimize the number of citations in the Conclusion, but we find this one reference particularly pertinent.

> The appendix is often not mentioned in the manuscript except Table VIII.

That's why the resources are listed in the appendix.

> Why do some of the Figures and Tables have regular numbering and some not?

As mentioned we are using the IEEE Tran Latex class and numbering is determined by such.

### 2021-08-24 docs/00-manuscripts/mdpi/manuscript.tex
add mdpi assets

--- 2021-06-28 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
-% \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
-\documentclass[final]{IEEEtran}
-
-\usepackage[switch]{lineno}
-
-% *** CITATION PACKAGES ***
-%
-\usepackage[noadjust]{cite}
-\usepackage[hidelinks]{hyperref}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-%\cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% 'compressed/ranged'. e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-%\cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf{}
+%  LaTeX support: latex@mdpi.com
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
+
+%=================================================================
+\documentclass[remotesensing,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%=================================================================
+% MDPI internal commands
+\firstpage{1}
+\makeatletter
+\setcounter{page}{\@firstpage}
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{2021}
+\copyrightyear{2020}
+%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
+\datereceived{}
+\dateaccepted{}
+\datepublished{}
+\hreflink{https://doi.org/} % If needed use \linebreak
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
+
+% \usepackage[noadjust]{cite}
+% \usepackage[hidelinks]{hyperref}
+\usepackage[nolist]{acronym}
+
+\usepackage{tabularx}
+\usepackage{booktabs}
+\usepackage{multirow}
+\usepackage{adjustbox}
+
+% line breaks in table cells
+\newcommand{\specialcell}[2][l]{%
+  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
+
+\usepackage{xtab}
+\usepackage{csquotes}
+\usepackage[utf8]{inputenc}
+% enables custom font styles which are not available in IEEEtran by default
+% https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran
+
+\renewcommand{\sfdefault}{cmss}
+\renewcommand{\rmdefault}{cmr}
+\renewcommand{\ttdefault}{cmt}
+
+\usepackage[detect-all]{siunitx}
+
 \usepackage[pdftex]{graphicx}
@@
 \DeclareGraphicsExtensions{.pdf,.jpg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
 
@@
 \newcommand*\mean[1]{\bar{#1}}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-
-% *** SPECIALIZED LIST PACKAGES ***
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-\ifCLASSOPTIONcompsoc
- \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\else
- \usepackage[caption=false,font=footnotesize]{subfig}
-\fi
-
-% *** FLOAT PACKAGES ***
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\usepackage[nolist]{acronym}
-
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage{adjustbox}
-
-% line breaks in table cells
-\newcommand{\specialcell}[2][l]{%
-  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
-
-\usepackage{xtab}
-
-\usepackage{csquotes}
-
-\usepackage[utf8]{inputenc}
-% enables custom font styles which are not available in IEEEtran by default
-% https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran
-
-\renewcommand{\sfdefault}{cmss}
-\renewcommand{\rmdefault}{cmr}
-\renewcommand{\ttdefault}{cmt}
-
-\usepackage[detect-all]{siunitx}
-
-\begin{document}
-% \linenumbers
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
-\IEEEpeerreviewmaketitle{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and non-breaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-
-% suppress 'underfull hbox' warnings
-\hbadness=99999
-
-\author{Patrick~Schratz,
-	Jannes~Muenchow,
-	Eugenia~Iturritxa,
-	José~Cortés,
-	Bernd~Bischl,
-	and Alexander~Brenning
-	\thanks{P.Schratz, J.Muenchow, J.Cortés and A.Brenning are with the Department
-		of Geography, geographic information science group, Friedrich Schiller University Jena, Germany.}% <-this % stops a space
-	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilians-Universität in Munich, Germany.}% <-this % stops a space
-	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Spain.}% <-this % stops a space
-	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
+
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
+
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Title}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0003-0748-6624} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-7834-4717} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0003-2567-8689} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0002-6390-5873} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorE}{0000-0001-6002-6980} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorF}{0000-0001-6640-679X} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Patrick Schratz $^{1}$\orcidA{}*, Jannes Muenchow $^{1}$\orcidB{}, José Cortés $^{1}$\orcidC{}, Eugenia Iturritxa $^{2}$\orcidD{}, Bernd Bischl $^{3}$\orcidE{} and Alexander Brenning $^{1,}$\orcidF{}}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{Schratz, P.; Muenchow, J.; Cortés, J.; Iturritxa, E.; Bischl, B.; Brenning, A.}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Friedrich Schiller University Jena, Department of Geography, GIScience group, Germany\\
+$^{2}$ \quad NEIKER Tecnalia, Spain\\
+$^{3}$ \quad Ludwig-Maximilians-Universiy Munich, Department of Statistics, Computational Statistics group, Germany
 }
 
-% note the % following the last \IEEEmembership and also \thanks -
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-%
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those 'LaTeX things'. For
-% instance, '\textbf{A} \textbf{B}' will typeset as 'A B' not 'AB'. To get
-% 'AB' then you have to do: '\textbf{A}\textbf{B}'
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-\markboth{IEEE JOURNAL OF SELECTED TOPICS IN APPLIED EARTH OBSERVATIONS AND REMOTE SENSING}%
-{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-%
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% make the title area
-\maketitle
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
-\begin{abstract}
-	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
+% Contact information of the corresponding author
+\corres{Correspondence: patrick.schratz@uni-jena.de}
+
+% Current address and/or shared authorship
+% \firstnote{Current address: Affiliation 3}
+% \secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+% Abstract (Do not insert blank lines, i.e. \\)
+\abstract{This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
 The effect of filter-based feature-selection methods on predictive performance was compared.
@@
 More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
-	Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.
-
-\end{abstract}
-
-% Note that keywords are not normally used for peer-review papers.
-\begin{IEEEkeywords}
-	hyperspectral imagery, forest health monitoring, machine learning, feature selection, feature effects, model comparison
-\end{IEEEkeywords}
-
-% For peer-review papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle{}
-
-\renewcommand{\IEEEiedlistdecl}{\IEEEsetlabelwidth{SONET}}
+Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.}
+
+% Keywords
+\keyword{hyperspectral imagery; forest health monitoring; machine learning; feature selection; model comparison}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
 \begin{acronym}
@@
 \end{acronym}
-\renewcommand{\IEEEiedlistdecl}{\relax}% reset back
-
+
+\begin{document}
 \section{Introduction}
-% Explain how remote sensing is used in forestry (potential to map forest health)
-
-% Link remote sensing and machine-learning
-
-\IEEEPARstart{T}{he} use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
+
+The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
 This coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014, which serves as training and prediction data.
@@
 
-\section{Data and study area}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Materials and Methods}
+
+\subsection{Data and study area}
 
@@
 
-\subsection{In-situ data}
+\subsubsection{In-situ data}
 
@@
 
-\subsection{Hyperspectral data}
+\subsubsection{Hyperspectral data}
 
@@
 
-\section{Methods}
+Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
+Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
+Please disclose at the submission stage any restrictions on the availability of materials or information.
+New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
+
+Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
+If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
+They must be provided prior to publication.
+
+Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
+\begin{quote}
+This is an example of a quote.
+\end{quote}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
@@
 
-\section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
 
@@
 
-\section{Acknowledgments}
-This work was funded by the EU LIFE Healthy Forest project (LIFE14 ENV/ES/000179) and the German Scholars Organization/Carl Zeiss Foundation.
-
-\appendices{}
-
-% \section{SVM ALE plots for task HR}
-
-% \begin{figure} [ht]
-% 	\begin{center}
-% 		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
-% 		\caption{ALE plots of SVM on dataset HR.
-The ten most important features from the permutation-based variable importance estimation were used.
-The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-hr-ale}
-% 	\end{center}
-% \end{figure}
-
-%\section{SVM ALE plots for task VI}
-
-% \begin{figure} [ht]
-% 	\begin{center}
-% 		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
-% 		\caption{ALE plots on dataset VI of SVM learner.
-Subset showing the ten most important features according to the permutation-based variable importance.
-The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-vi-ale}
-% 	\end{center}
-% \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{Conceptualization, Patrick Schratz and Alexander Brenning; methodology, Patrick Schratz, José Cortés and Jannes Muenchow and Bernd Bischl; software, Patrick Schratz; validation, Patrick Schratz and Alexander Brenning; formal analysis, Patrick Schratz, Alexander Brenning, José Cortés and Bernd Bischl; investigation, Patrick Schratz; resources, Eugenia Iturritxa and Patrick Schratz; data curation, Patrick Schratz and Eugenia Iturritxa; writing---original draft preparation, Patrick Schratz; writing---review and editing, Patrick Schratz and Alexander Brenning; visualization, Patrick Schratz; supervision, Alexander Brenning and Bernd Bischl; project administration, Alexander Brenning; funding acquisition, Alexander Brenning.
+All authors have read and agreed to the published version of the manuscript.}
+
+\funding{This work was funded by the EU LIFE Healthy Forest project (LIFE14 ENV/ES/000179) and the German Scholars Organization/Carl Zeiss Foundation.}
+
+\dataavailability{The data presented in this study are openly available in Zenodo at https://doi.org/10.5281/zenodo.2635403.}
+
+% \acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\conflictsofinterest{The authors declare no conflict of interest.
+The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the~results.}
+
+%% Optional
+% \sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
+
+\noindent
+\begin{tabular}{@{}ll}
+	AGB & above-ground biomass \\
+	ALE & accumulated local effects \\
+	ALS & airborne laser scanning \\
+	ANN & artificial neural network \\
+	AUROC & area under the receiver operating characteristics curve \\
+	BRT & boosted regression trees \\
+	CART & classification and regression trees \\
+	CNN & convolutional neural networks \\
+	CV & cross-validation \\
+	DAP & digital aerial photogrammetry \\
+	ENM & environmental niche modeling \\
+	FFS & forward feature selection \\
+	FPR & false positive rate \\
+	FS & feature selection \\
+	GAM & generalized additive model \\
+	GBM & gradient boosting machine \\
+	GLM & generalized linear model \\
+	ICGC & Institut Cartografic i Geologic de Catalunya \\
+	IQR & interquartile range \\
+	LiDAR & light detection and ranging \\
+	LOWESS & locally weighted scatter plot smoothing \\
+	MARS & multivariate adaptive regression splines \\
+	MBO & model-based optimization \\
+	MEM & maximum entropy model \\
+	ML & machine learning \\
+	NDII & normalized difference infrared index \\
+	NDMI & normalized difference moisture index \\
+	NIR & near-infrared \\
+	NRI & normalized ratio index \\
+	OLS & ordinary least squares \\
+	OMNBR & optimized multiple narrow-band reflectance \\
+	PCA & principal component analysis \\
+	PDP & partial dependence plots \\
+	PISR & potential incoming solar radiation \\
+	PLS & partial least-squares \\
+	POV & proportion of variance explained \\
+	RBF & radial basis function \\
+	RF & random forest \\
+	RMSE & root mean square error \\
+	RR & ridge regression \\
+	RSS & residual sum of squares \\
+	SAR & synthetic aperture radar \\
+	SDM & species distribution modeling \\
+	SMBO & sequential-based model optimization \\
+	SVM & support vector machine \\
+	TPR & true positive rate \\
+	VI & vegetation index \\
+	XGBoost & extreme gradient boosting
+\end{tabular}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
 \pagebreak
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
+\appendix
 \section{}
+\subsection{}
 
@@
 
-\section{}
-
+\pagebreak
+
+\subsection{}
 \begin{figure} [ht]
@@
 
-\section{}
-
-% parameter limits
+\subsection{}
+
 \begin{table}[h!]
@@
 
-\section{}
+\pagebreak
+
+\subsection{}
 
@@
 \setlength\tabcolsep{4pt}  % default value: 6pt
+\begin{table}
+	\caption[t]{List of available vegetation indices in the \texttt{hsdar} package.}
 \begin{xtabular*}{0.3\textwidth}{lll}
@@
 \label{tab:vegindices}
+\end{table}
 \endgroup
@@
 
-\section{Radiometric, Geometric and Atmospheric correction of hyperspectral data}
+\subsection{}
 
@@
 
-\pagebreak
-
-\bibliographystyle{IEEEtran}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{paracol}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% To add notes in main text, please use \endnote{} and un-comment the codes below.
+%\begin{adjustwidth}{-5.0cm}{0cm}
+%\printendnotes[custom]
+%\end{adjustwidth}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\externalbibliography{yes}
 \bibliography{Biblio}
 
-\pagebreak
-
-% The photo area is 1 in wide and 1.25 in long. The IEEE recommends that author photo images should be of 220 dpi (dots per inch) resolution and in gray scale with 8 bits/sample.
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/pschratz.jpeg}}]{Patrick Schratz}
-	%	[]{Patrick Schratz}
-	is a Data Scientist working in Zurich, Switzerland.
+% If authors have biography, please use the format below
+\section*{Short Biography of Authors}
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{pschratz.jpeg}}}
+{\textbf{Patrick Schratz} is a Data Scientist working in Zurich, Switzerland.
 His area of expertise is applied machine learning, more specifically the field of environmental modeling.
 He is a PhD candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.
-	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/jmuenchow.jpg}}]{Jannes Muenchow}
-	is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
+At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{jmuenchow.jpg}}}
+{\textbf{Jannes Muenchow} is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
 He received his PhD from Friedrich-Alexander University Erlangen-Nürnberg (Germany) in 2013.
@@
 Currently, his research focuses on developing open source tools for ecology, geomorphology and qualitative GIS.
-	He is a co-author of the book "Geocomputation with R".
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/eiturritxa.jpg}}]{Eugenia Iturritxa}
-	received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.
+He is a co-author of the book "Geocomputation with R".}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{eiturritxa.jpg}}}
+{\textbf{Eugenia Iturritxa} received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.
 Since 1999 her main research has focused on forest health, on the study of diverse species of native and introduced pathogenic fungi in forests and forest plantations.
-	Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/jcortes.png}}]{José Cortés}
-	received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
+Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{jcortes.png}}}
+{\textbf{José Cortés} received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
 Currently he is a PhD student at Friedrich Schiller University, Jena, Germany.
 He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
-	His research focuses on spatio-temporal trend detection in environmental data.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/bbischl.jpg}}]{Bernd Bischl}
-	obtained his Ph.D from Dortmund Technical University in 2013.
+His research focuses on spatio-temporal trend detection in environmental data.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{bbischl.jpg}}}
+{\textbf{Bernd Bischl} obtained his Ph.D from Dortmund Technical University in 2013.
 He is a professor of Statistical Learning and Data Science in the Department of Statistics at Ludwig-Maximilians-Universität in Munich, Germany, and a director of the Munich Center for Machine Learning.
-	His research interests include AutoML, model selection, interpretable ML and XAI.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/abrenning.jpg}}]{Alexander Brenning}
-	(Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
+His research interests include AutoML, model selection, interpretable ML and XAI.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{abrenning.jpg}}}
+{\textbf{Alexander Brenning} (Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
 He served as an assistant professor and tenured associate professor in geomatics at the University of Waterloo, Ontario, Canada from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
 Dr. Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.
-	He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.
-\end{IEEEbiography}
-
+He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2021-08-24 docs/00-manuscripts/mdpi/template.tex
add mdpi assets

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, appliedchem, applmech, applmicrobiol, applnano, applsci, arts, asi, atmosphere, atoms, audiolres, automation, axioms, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomechanics, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, brainsci, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coatings, colloids, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, crops, cryptography, crystals, curroncol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forensicsci, forests, fractalfract, fuels, futureinternet, futuretransp, futurepharmacol, futurephys, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, land, languages, laws, life, liquids, literature, livers, logistics, lubricants, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, metabolites, metals, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, minerals, mining, modelling, molbank, molecules, mps, mti, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photochem, photonics, physchem, physics, physiolsci, plants, plasma, pollutants, polymers, polysaccharides, proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, systems, taxonomy, technologies, telecom, textiles, thermo, tourismhosp, toxics, toxins, transplantology, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wevj, women, world

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2021}
\copyrightyear{2020}
%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
\datereceived{}
\dateaccepted{}
\datepublished{}
\hreflink{https://doi.org/} % If needed use \linebreak
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3}
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusion: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{Instead of the abstract}
%\entrylink{The Link to this entry published on the encyclopedia platform.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}

The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions on the authors' page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX-related questions please contact latex@mdpi.com.
%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Histories, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item;
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}

% The MDPI table float is called specialtable
\begin{specialtable}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
%%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabular}
\end{specialtable}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{equation}
a = 1,
\end{equation}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\end{paracol}
\nointerlineskip
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}

% Example of a figure that spans the whole page width (the commands \widefigure and \begin{paracol}{2}, \linenumbers, and\switchcolumn need to be present). The same concept works for tables, too.
\begin{figure}[H]
\widefigure
\includegraphics[width=15 cm]{Definitions/logo-mdpi}
\caption{This is a wide figure.\label{fig2}}
\end{figure}
\begin{paracol}{2}
\linenumbers
\switchcolumn

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\institutionalreview{In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals. Please note that the Editorial Office might ask you for further information. Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals. You might also choose to exclude this statement if the study did not involve humans or animals.}

\informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. You might choose to exclude this statement if the study did not report any data.}

\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the~results''.}

%% Optional
\sampleavailability{Samples of the compounds ... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section{}
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

\begin{specialtable}[H]
%\tablesize{\scriptsize}
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab2}}
%\tablesize{} % You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabular}
\end{specialtable}

\section{}
All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{paracol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% To add notes in main text, please use \endnote{} and un-comment the codes below.
%\begin{adjustwidth}{-5.0cm}{0cm}
%\printendnotes[custom]
%\end{adjustwidth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book1}
Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
\bibitem[Author3(year)]{ref-book2}
Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
\bibitem[Author4(year)]{ref-unpublish}
Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
% Reference 5
\bibitem[Author5(year)]{ref-communication}
Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
\bibitem[Author6(year)]{ref-proceeding}
Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of Presentation. In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds. (if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
% Reference 7
\bibitem[Author7(year)]{ref-thesis}
Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
\bibitem[Author8(year)]{ref-url}
Title of Site. Available online: URL (accessed on Day Month Year).
\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% The following MDPI journals use author-date citation: Admsci,  Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Jintelligence, JRFM, Languages, Laws, Literature, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2021-08-24 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   print(
-    file = here::here("docs/00-manuscripts/ieee/performance-svm-single-plot.tex"),
+    file = here::here("docs/00-manuscripts/mdpi/performance-svm-single-plot.tex"),
     include.rownames = TRUE,
@@
   print(
-    file = here("docs/00-manuscripts/ieee/performance-top-10.tex"),
+    file = here("docs/00-manuscripts/mdpi/performance-top-10.tex"),
     include.rownames = TRUE,
@@
   print(
-    file = here("docs/00-manuscripts/ieee/performance-best-per-learner.tex"),
+    file = here("docs/00-manuscripts/mdpi/performance-best-per-learner.tex"),
     include.rownames = TRUE,
@@
   print(
-    file = here("docs/00-manuscripts/ieee/performance-worst-10.tex"),
+    file = here("docs/00-manuscripts/mdpi/performance-worst-10.tex"),
     include.rownames = TRUE,
@@
   print(
-    file = here("docs/00-manuscripts/ieee/tune-perc-sel-features.tex"),
+    file = here("docs/00-manuscripts/mdpi/tune-perc-sel-features.tex"),
     include.rownames = FALSE,

### 2021-08-24 docs/00-manuscripts/mdpi/manuscript.tex
comment on featureless learner

@@
 
-The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE 28.09 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE 28.12 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
 Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
@@
 Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
-However, when inspecting the fold level performances, it can be concluded that the model performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
+However, when inspecting the fold level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
 This applied to all learners of this study.
-The overall performance achieved in this study can be classified as "fair".
+The overall performance of all learners achieved in this study can be classified as "poor" given that only the SVM learner was able to outperform the featureless learner (\autoref{tab:best-learner-perf}).
 

### 2021-08-24 docs/00-manuscripts/mdpi/manuscript.tex
remove nbins correlation figure

@@
 
-In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins and treated as a class variable.
-After feature rank correlations of $> 0.9$ between different bin sizes were observed in a side analysis, \texttt{\(n_{bin}\) = 10} was found to be a reasonable setting to go with.
+In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins \texttt{\(n_{bin}\) = 10} and treated as a class variable.
 
@@
 \subsection{}
-\begin{figure} [ht]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
-		\caption{Spearman correlations of rankings obtained with the information gain filter using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.
-Filter names refer to the nomenclature used by the \texttt{mlr} R package.
-Underscores in names divide the terminology into their upstream R package and the actual filter name.}\label{fig:correlation-nbins}
-	\end{center}
-\end{figure}
-
-\pagebreak
-
-\subsection{}
 

### 2021-08-25 docs/00-manuscripts/mdpi/manuscript.tex
final mdpi conversions

@@
 
+% \usepackage{supertabular}
+\usepackage{supertabular,booktabs}
+
 % line breaks in table cells
@@
 
-\usepackage{xtab}
 \usepackage{csquotes}
@@
 \DeclareGraphicsExtensions{.pdf,.jpg,.png}
+
+% suppress fancyhdr warnings
+\setlength{\headheight}{23.90979pt}
 
@@
 $^{2}$ \quad NEIKER Tecnalia, Spain\\
-$^{3}$ \quad Ludwig-Maximilians-Universiy Munich, Department of Statistics, Computational Statistics group, Germany
+$^{3}$ \quad Ludwig-Maximilians-University Munich, Department of Statistics, Computational Statistics group, Germany
 }
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {defoliation-distribution-plot-1.pdf}
+		\includegraphics[width=0.7\textwidth] {defoliation-distribution-plot-1.pdf}
 		\caption{Response variable "defoliation at trees" for plots Laukiz1, Laukiz2, Luiando and Oiartzun.
@@
 %	\begin{center}
-%		\includegraphics[width=0.48\textwidth] {defol-grid-3000px.jpg}
+%		\includegraphics[width=0.7\textwidth] {defol-grid-3000px.jpg}
 %		\caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
@@
 	\begingroup\footnotesize
-	\begin{adjustbox}{width={0.48\textwidth},totalheight={\textheight},keepaspectratio}
+	\begin{adjustbox}{width={0.7\textwidth},totalheight={\textheight},keepaspectratio}
 		\begin{tabular}{lll}
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {performance-results-1.pdf}
+		\includegraphics[width=0.7\textwidth] {performance-results-1.pdf}
 		\caption{Predictive performance in RMSE (p.p.) of models across tasks.
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-no-filter-1.pdf}
+		\includegraphics[width=0.7\textwidth] {filter-effect-all-vs-no-filter-1.pdf}
 		\caption{Model performances in RMSE across all tasks, split up in facets, when using no filter method (blue dot) compared to any other filter method (red cross) for learners RF, SVM and XGBoost (XG)}\label{fig:filter-effects-no-filter}
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {filter-effect-all-vs-borda-filter-1.pdf}
+		\includegraphics[width=0.7\textwidth] {filter-effect-all-vs-borda-filter-1.pdf}
 		\caption{Predictive performances in RMSE (p.p.) when using the Borda filter method (blue dot) compared to any other filter (red cross) for each learner across all tasks.}\label{fig:filter-effects-borda}
@@
 \appendix
-\section{}
+\section{Appendices}
 \subsection{}
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth] {correlation-filter-nri-1.pdf}
+		\includegraphics[width=0.7\textwidth] {correlation-filter-nri-1.pdf}
 		\caption{Spearman correlations of NRI feature rankings obtained with different filters.
@@
 
-\begingroup\scriptsize
-\setlength\tabcolsep{4pt}  % default value: 6pt
-\begin{table}
-	\caption[t]{List of available vegetation indices in the \texttt{hsdar} package.}
-\begin{xtabular*}{0.3\textwidth}{lll}
+% \begingroup\scriptsize
+% \setlength\tabcolsep{4pt}  % default value: 6pt
+% vertical spacing between rows (default is 1)
+{\def\arraystretch{1.7}
+	\tablecaption{List of available vegetation indices in the \texttt{hsdar} package.}
+	\begin{supertabular}{lll}
+		% \begin{supertabular}{0.3\textwidth}{lll}
 		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference*}         \\
@@
 		mNDVI           & $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$                                         & \cite{sims2002a}              \\
-	MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                           & \cite{hernandez-clemente2011}\\
-	MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                & \cite{qi1994}\\
-	MSI             & $\frac{R_{1600}}{R_{817}}$                                                                          & \cite{hunt1989}\\
+		MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                            & \cite{hernandez-clemente2011} \\
+		MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                 & \cite{qi1994}                 \\
+		MSI             & $\frac{R_{1600}}{R_{817}}$                                                                           & \cite{hunt1989}               \\
 		mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                            & \cite{sims2002a}              \\
-	mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                              & \cite{chen1996}\\
+		mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                               & \cite{chen1996}               \\
 		mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                            & \cite{sims2002a}              \\
-	MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                           & \cite{dash2007}\\
-	\midrule
-	MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}            & \cite{haboudane2002}\\
+		MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                            & \cite{dash2007}               \\
+		\midrule
+		MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}             & \cite{haboudane2002}          \\
 		\midrule
@@
 		PARS            & $\frac{R_{746}}{R_{513}}$                                                                            & \cite{chappelle1992}          \\
-	PRI            & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            & \cite{gamon1992}\\
-	PRI\_norm      & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                & \cite{zarco-tejada2013a}\\
+		PRI             & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            & \cite{gamon1992}              \\
+		PRI\_norm       & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                & \cite{zarco-tejada2013a}      \\
 		PRI*CI2         & $PRI*CI2$                                                                                            & \cite{garrity2011}            \\
@@
 		SR              & $\frac{R_{800}}{R_{680}}$                                                                            & \cite{jordan1969}             \\
-	SR1            & $\frac{R_{750}}{R_{700}}$                                                                            & \cite{gitelson1997}\\
-	SR2            & $\frac{R_{752}}{R_{690}}$                                                                            & \cite{gitelson1997}\\
-	SR3            & $\frac{R_{750}}{R_{550}}$                                                                            & \cite{gitelson1997}\\
+		SR1             & $\frac{R_{750}}{R_{700}}$                                                                            & \cite{gitelson1997}           \\
+		SR2             & $\frac{R_{752}}{R_{690}}$                                                                            & \cite{gitelson1997}           \\
+		SR3             & $\frac{R_{750}}{R_{550}}$                                                                            & \cite{gitelson1997}           \\
 		SR4             & $\frac{R_{700}}{R_{670}}$                                                                            & \cite{mcmurtrey1994}          \\
@@
 		SR7             & $\frac{R_{440}}{R_{690}}$                                                                            & \cite{lichtenthaler1996}      \\
-	SR8            & $\frac{R_{515}}{R_{550}}$                                                                            & \cite{hernandez-clemente2012}\\
-	SRPI           & $\frac{R_{430}}{R_{680}}$                                                                            & \cite{penuelasj.1995}\\
-	SRWI           & $\frac{R_{850}}{R_{1240}}$                                                                           & \cite{zarco-tejada2003a}\\
+		SR8             & $\frac{R_{515}}{R_{550}}$                                                                            & \cite{hernandez-clemente2012} \\
+		SRPI            & $\frac{R_{430}}{R_{680}}$                                                                            & \cite{penuelasj.1995}         \\
+		SRWI            & $\frac{R_{850}}{R_{1240}}$                                                                           & \cite{zarco-tejada2003a}      \\
 		Sum\_Dr1        & $\sum_{i=626}^{795} D1_i$                                                                            & \cite{elvidge1995}            \\
@@
 		\midrule
-	TVI            &\parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}               & \cite{broge2001} \\
+		TVI             & \parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}              & \cite{broge2001}              \\
 		\midrule
@@
 		Vogelmann4      & $\frac{R_{734}-R_{747}}{R_{715}+R_{720}}$                                                            & \cite{vogelmann1993}     \\
-\end{xtabular*}
+	\end{supertabular}
 	\label{tab:vegindices}
-\end{table}
-\endgroup
+}
 

### 2021-08-26 docs/00-manuscripts/mdpi/manuscript.tex
p.p -> p.p.

@@
 Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.) (\autoref{fig:perf-result}).
-SVM showed almost no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p (\autoref{fig:perf-result}).
-SVM showed the best overall performance with a mean difference of around three p.p to the next best model (XGBoost) (\autoref{tab:best-learner-perf}).
+SVM showed almost no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p. (\autoref{fig:perf-result}).
+SVM showed the best overall performance with a mean difference of around three p.p. to the next best model (XGBoost) (\autoref{tab:best-learner-perf}).
 Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 p.p.
@@
 XGBoost showed a high dependency on filtering the data: In two out of six tasks using no filter resulted in the worst or second worst performance.
-XGBoost shows the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p (\enquote{CMIM} vs.
+XGBoost shows the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p.
+(\enquote{CMIM} vs.
 \enquote{no filter}) (\autoref{fig:filter-effects-no-filter}).
@@
 
-The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
 For dataset HR most features cluster around the infrared region (920 nm - 1000 nm) (six out of ten) while for VI eight out of ten concentrate on the wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
@@
 An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
-These cluster around a performance of 31 p.p while SVM is able to score about three p.p. better than all other methods.
+These cluster around a performance of 31 p.p. while SVM is able to score about three p.p. better than all other methods.
 However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
@@
 With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.
-The performance of RF and XGBoost on the VI feature set was around four to six p.p lower than on others.
+The performance of RF and XGBoost on the VI feature set was around four to six p.p. lower than on others.
 One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
@@
 % What about the general potential of analyzing defol with hyperspectral data?
-Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 28 p.p scored by the best performing model.
+Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 28 p.p. scored by the best performing model.
 More training data covering a wider range of defoliation values in a larger number of forest plantations is needed to train better models which can create more robust predictions.

### 2021-08-26 docs/00-manuscripts/mdpi/manuscript.tex
capital D

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization, Patrick Schratz and Alexander Brenning; methodology, Patrick Schratz, José Cortés and Jannes Muenchow and Bernd Bischl; software, Patrick Schratz; validation, Patrick Schratz and Alexander Brenning; formal analysis, Patrick Schratz, Alexander Brenning, José Cortés and Bernd Bischl; investigation, Patrick Schratz; resources, Eugenia Iturritxa and Patrick Schratz; data curation, Patrick Schratz and Eugenia Iturritxa; writing---original draft preparation, Patrick Schratz; writing---review and editing, Patrick Schratz and Alexander Brenning; visualization, Patrick Schratz; supervision, Alexander Brenning and Bernd Bischl; project administration, Alexander Brenning; funding acquisition, Alexander Brenning.
+\authorcontributions{Conceptualization, Patrick Schratz and Alexander Brenning; methodology, Patrick Schratz, José Cortés, Jannes Muenchow and Bernd Bischl; software, Patrick Schratz; validation, Patrick Schratz and Alexander Brenning; formal analysis, Patrick Schratz, Alexander Brenning, José Cortés and Bernd Bischl; investigation, Patrick Schratz; resources, Eugenia Iturritxa and Patrick Schratz; data curation, Patrick Schratz and Eugenia Iturritxa; writing---original draft preparation, Patrick Schratz; writing---review and editing, Patrick Schratz and Alexander Brenning; visualization, Patrick Schratz; supervision, Alexander Brenning and Bernd Bischl; project administration, Alexander Brenning; funding acquisition, Alexander Brenning.
 All authors have read and agreed to the published version of the manuscript.}
@@
 The image acquisitions were originally attempted during one day (29.10.2016).
-due to the variable meteorological conditions some stands had to be imaged one day later.
+Due to the variable meteorological conditions some stands had to be imaged one day later.
 

### 2021-09-02 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
   arrange(rmse_aggr) %>%
+  # remove Task info for featureless.learner
+  dplyr::mutate(task.id = case_when(learner.id == "regr.featureless" ~ "-", TRUE ~ as.character(task.id))) %>%
   dplyr::rename(
@@
   scale_color_viridis_d(breaks = results_aggr$learner_group) +
-
   labs(x = "RMSE", y = "Task", color = "Learner") +
@@
   dplyr::mutate(iter = as.character(iter)) %>%
-  dplyr::mutate(iter = case_when(learner.id == "XGBOOST MBO CMIM" & iter == 1 ~ "Luiando",
+  dplyr::mutate(iter = case_when(
+    learner.id == "XGBOOST MBO CMIM" & iter == 1 ~ "Luiando",
     learner.id == "XGBOOST MBO CMIM" & iter == 2 ~ "Laukiz1",
@@
 
-                                 TRUE ~ as.character(iter))) %>% 
+    TRUE ~ as.character(iter)
+  )) %>%
   dplyr::rename(Learner = learner.id) %>%
@@
     `Test Plot` == "Laukiz1" ~ paste0(as.character(ceiling((`Features (\\%)` * 1249) / 100)), "/1249"),
-    `Test Plot` == "Laukiz2" ~ paste0(as.character(ceiling((`Features (\\%)` * 1357) / 100)),"/1357"),
-    `Test Plot` == "Luiando" ~ paste0(as.character(ceiling((`Features (\\%)` * 1507) / 100)),"/1507"),
-    `Test Plot` == "Oiartzun" ~ paste0(as.character(ceiling((`Features (\\%)` * 1311) / 100)),"/1311"),
+    `Test Plot` == "Laukiz2" ~ paste0(as.character(ceiling((`Features (\\%)` * 1357) / 100)), "/1357"),
+    `Test Plot` == "Luiando" ~ paste0(as.character(ceiling((`Features (\\%)` * 1507) / 100)), "/1507"),
+    `Test Plot` == "Oiartzun" ~ paste0(as.character(ceiling((`Features (\\%)` * 1311) / 100)), "/1311"),
   )) %>%

### 2021-09-02 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
     type = "latex",
-    caption = "The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run.",
+    caption = "The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run.
+For \texttt{regr.featureless} the Task is no applicable and was therefore removed.",
     label = "tab:best-learner-perf"

### 2021-09-02 docs/00-manuscripts/mdpi/performance-best-per-learner.tex
update table caption

--- 2021-04-29 docs/00-manuscripts/ieee/performance-best-per-learner.tex
@@
 \centering
-\caption{Best performance of each learner across any task and filter method, sorted ascending by RMSE} 
+\caption{The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run.
+For 	exttt{regr.featureless} the Task is no applicable and was therefore removed.}
 \label{tab:best-learner-perf}
@@
   \hline
-1 & HR-NRI-VI & SVM & Relief & 28.09 & 18.95 \\ 
-  2 & HR & XGBoost & CMIM & 30.88 & 14.47 \\ 
-  3 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
-  4 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
-  5 & HR & RF & Relief & 32.93 & 13.17 \\ 
+1 & HR & SVM & Relief & 28.12 & 19.12 \\ 
+  2 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
+  3 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
+  4 & - & regr.featureless & No Filter & 31.16 & 15.03 \\ 
+  5 & HR & RF & Relief & 31.25 & 14.96 \\ 
+  6 & HR & XGBoost & Car & 34.07 & 13.57 \\ 
    \hline

### 2021-09-13 docs/00-manuscripts/mdpi/manuscript.tex
add featurless learner to learner list
update author order

@@
 % Authors, for the paper (add full first names)
-\Author{Patrick Schratz $^{1}$\orcidA{}*, Jannes Muenchow $^{1}$\orcidB{}, José Cortés $^{1}$\orcidC{}, Eugenia Iturritxa $^{2}$\orcidD{}, Bernd Bischl $^{3}$\orcidE{} and Alexander Brenning $^{1,}$\orcidF{}}
+\Author{Patrick Schratz $^{1}$\orcidA{}*, Jannes Muenchow $^{1}$\orcidB{}, Eugenia Iturritxa $^{2}$\orcidD{}, José Cortés $^{1}$\orcidC{}, Bernd Bischl $^{3}$\orcidE{} and Alexander Brenning $^{1,}$\orcidF{}}
 
@@
 
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-\begin{quote}
-	This is an example of a quote.
-\end{quote}
-
 \subsection{Derivation of indices}
@@
 	\item  Support Vector Machine (SVM, RBF Kernel)
+	\item  Featureless learner
 \end{itemize}
@@
 Ridge does not remove variables from the model (penalization to zero) but just shrinks them to effectively zero, keeping them in the model.
+A featureless learner was included for a baseline comparison.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Conceptualization, Patrick Schratz and Alexander Brenning; methodology, Patrick Schratz, José Cortés, Jannes Muenchow and Bernd Bischl; software, Patrick Schratz; validation, Patrick Schratz and Alexander Brenning; formal analysis, Patrick Schratz, Alexander Brenning, José Cortés and Bernd Bischl; investigation, Patrick Schratz; resources, Eugenia Iturritxa and Patrick Schratz; data curation, Patrick Schratz and Eugenia Iturritxa; writing---original draft preparation, Patrick Schratz; writing---review and editing, Patrick Schratz and Alexander Brenning; visualization, Patrick Schratz; supervision, Alexander Brenning and Bernd Bischl; project administration, Alexander Brenning; funding acquisition, Alexander Brenning.
-All authors have read and agreed to the published version of the manuscript.}
+\authorcontributions{Conceptualization, Patrick Schratz and Alexander Brenning; Data curation, Patrick Schratz and Eugenia Iturritxa; Formal analysis, Patrick Schratz, José Cortés, Bernd Bischl and Alexander Brenning; Funding acquisition, Alexander Brenning; Investigation, Patrick Schratz; Methodology, Patrick Schratz, Jannes Münchow, José Cortés and Bernd Bischl; Project administration, Alexander Brenning; Resources, Eugenia Iturritxa; Software, Patrick Schratz; Supervision, Bernd Bischl and Alexander Brenning; Validation, Patrick Schratz and Alexander Brenning; Visualization, Patrick Schratz; Writing – original draft, Patrick Schratz; Writing – review \& editing, Jannes Münchow, José Cortés and Alexander Brenning.}
 

### 2021-10-31 docs/00-manuscripts/mdpi/manuscript.tex
Moved l.55 - l.62 to the end of the introduction as suggested by Reviewer2
Add more introductory content for the "tree defoliation" topic

@@
 
-% link to forest health analysis and show exemplary studies
+% link to forest health (defoliation) analysis and show exemplary studies
 
@@
 One aspect of this research field is to enhance the understanding of biotic and abiotic stress triggers, for example by analyzing tree defoliation \cite{hawrylo2018}.
-
+Defoliation is known to be a proxy for pathogen and insect damage \cite{pollastrini2016}.
+While common symptoms are observable across species, some effects and their degree are species-specific \cite{gottardini2020}.
+Defoliation has also been shown to increase predisposition of tree death from secondary biotic factors of up to ten years after the actual defoliation event \cite{oliva2016}.
 Other approaches for analyzing forest health include change detection \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
-In such studies, the defoliation of trees serves as a proxy for forest health by describing the impact of biotic and abiotic pest triggers \cite{townsend2012, goodbody2018}.
 
@@
 
-% now that we introduced the importance of remote sensing indices for forest health, explain what is new in this study
+% mention why FS is used / important
+Even though \ac{ML} algorithms are capable of handling highly correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
+Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
+Instead of using wrapper feature selection methods, this study focuses on the use of (ensemble) filter methods which were directly integrated into the hyperparameter optimization step during model creation.
+
+% clarify the focus of this study
+Overall, this study aims to show how high-dimensional datasets can be handled effectively with ML methods while still being able to interpret the fitted models.
+The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
+Yet, this study focuses on investigating the effects of filter methods and feature set types on predictive performance rather than interpreting individual feature effects.
+
+Considering these opportunities and challenges, the research questions of this study are the following:
+
+\begin{itemize}
+
+	\item Do different (environmental) feature sets show differences in performance when modeling defoliation of trees?
+
+	\item Can predictive performance be substantially improved by combining feature sets?
+
+	\item How do different feature selection methods influence the predictive performance of the models?
+
+	\item Which features are most important and how can these be interpreted in an environmental context?
+
+\end{itemize}
+
+% Explain what is new in this study and link to methodology
 Despite their popularity in environmental modeling, there are no studies so far which used ML algorithms in combination with remote sensing data to analyze defoliation at the tree level.
@@
 Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high dimensionality in environmental modeling \cite{trunk1979, xu2016}.
-
-% mention why FS is used / important
-Even though \ac{ML} algorithms are capable of handling highly correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
-Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
-Instead of using wrapper feature selection methods, this study focuses on the use of (ensemble) filter methods which were directly integrated into the hyperparameter optimization step during model creation.
-
-% clarify the focus of this study
-Overall, this study aims to show how high-dimensional datasets can be handled effectively with ML methods while still being able to interpret the fitted models.
-The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
-Yet, this study focuses on investigating the effects of filter methods and feature set types on predictive performance rather than interpreting individual feature effects.
-
-Considering these opportunities and challenges, the research questions of this study are the following:
-
-\begin{itemize}
-
-	\item Do different (environmental) feature sets show differences in performance when modeling defoliation of trees?
-
-	\item Can predictive performance be substantially improved by combining feature sets?
-
-	\item How do different feature selection methods influence the predictive performance of the models?
-
-	\item Which features are most important and how can these be interpreted in an environmental context?
-
-\end{itemize}
 

### 2021-10-31 docs/00-manuscripts/mdpi/performance-best-per-learner.tex

@@
   \hline
-1 & HR & SVM & Relief & 28.12 & 19.12 \\ 
-  2 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
-  3 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
-  4 & - & regr.featureless & No Filter & 31.16 & 15.03 \\ 
-  5 & HR & RF & Relief & 31.25 & 14.96 \\ 
-  6 & HR & XGBoost & Car & 34.07 & 13.57 \\ 
+1 & HR & SVM & Relief & 28.119 & 19.123 \\ 
+  2 & NRI & Lasso-MBO & No Filter & 31.165 & 15.025 \\ 
+  3 & NRI & Ridge-MBO & No Filter & 31.165 & 15.025 \\ 
+  4 & - & regr.featureless & No Filter & 31.165 & 15.025 \\ 
+  5 & HR & RF & Relief & 31.246 & 14.956 \\ 
+  6 & HR & XGBoost & Car & 34.071 & 13.571 \\ 
    \hline

### 2021-11-02 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

@@
 ```{r buffer-scenarios-1, echo=FALSE}
-knitr::include_graphics("inst/tree-buffer-vis.png")
+knitr::include_graphics("figure/tree-buffer-vis.png")
 ```
@@
 ```{r buffer-scenarios-2, echo=FALSE}
-knitr::include_graphics("inst/tree-buffer-vis-max.png")
+knitr::include_graphics("figure/tree-buffer-vis-max.png")
 ```

### 2021-11-02 docs/00-manuscripts/mdpi/manuscript.tex
Figure 3: Removed erroneous statement about SE values and increased digits from two to three to highlight small differences in the results

@@
 Labels show the feature selection method (e.g.
-NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda').
-The second value of each label shows the RMSE value (p.p.) and the standard error (SE) of the respective setting.}\label{fig:perf-result}
+NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda').}\label{fig:perf-result}
 	\end{center}

### 2021-11-02 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

--- 2021-04-12 analysis/eda.Rmd
@@
 ```{r buffer-scenarios-1, echo=FALSE}
-knitr::include_graphics("inst/tree-buffer-vis.png")
+knitr::include_graphics("docs/figure/tree-buffer-vis.png")
 ```
@@
 ```{r buffer-scenarios-2, echo=FALSE}
-knitr::include_graphics("inst/tree-buffer-vis-max.png")
+knitr::include_graphics("docs/figure/tree-buffer-vis-max.png")
 ```

### 2021-11-02 analysis/eval-performance.Rmd
wflow_publish(knitr_in("analysis/eval-performance.Rmd"), view = FALSE,     verbose = TRUE)

@@
     caption = "Best ten results among all learner-task-filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE).",
-    label = "tab:perf-top-10"
+    label = "tab:perf-top-10",
+    digits = 3
   ) %>%
@@
 For \texttt{regr.featureless} the Task is no applicable and was therefore removed.",
-    label = "tab:best-learner-perf"
+    label = "tab:best-learner-perf",
+    digits = 3
   ) %>%
@@
     caption = "Worst ten results among all learner-task-filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE).",
-    label = "tab:perf-worst-10"
+    label = "tab:perf-worst-10",
+    digits = 3
   ) %>%
@@
     position = position_quasirandom(),
-    aes(label = paste0(filter, ",", round(rmse_aggr, 2))),
+    aes(label = paste0(filter, ",", round(rmse_aggr, 3))),
     size = 4,

### 2021-11-04 docs/00-manuscripts/mdpi/manuscript.tex
Rephrased the introduction and reasoning on why filters methods were chosen over wrapper methods

@@
 Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
-Instead of using wrapper feature selection methods, this study focuses on the use of (ensemble) filter methods which were directly integrated into the hyperparameter optimization step during model creation.
+Instead of using wrapper feature selection methods, which add a substantial overhead to a nested model optimization approach, especially for datasets with many features, this study focuses on the use of (ensemble) filter methods which can be directly integrated into the hyperparameter optimization step during model creation.
 
@@
 In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
-Due to the focus on filter methods in this manuscript, only this sub-group of feature selection methods will be introduced in greater detail in the following sections.
+In addition, filters are less known than wrapper methods and in recent years ensemble filters were introduced which have shown promising results compared to single filter algorithms \cite{drotar2017a}.
+These two points mainly led to the decision to focus on filter methods for this work and evaluate their effectiveness on highly-correlated, high-dimensional datasets.
+Due to this focus, only this sub-group of feature selection methods will be introduced in greater detail in the following sections.
 
@@
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
-However, rather than only looking at these three groups also combinations of such were taken into account:
+However, rather than only looking at these three groups, also the following combinations of such were taken into account:
 

### 2021-11-04 docs/00-manuscripts/mdpi/manuscript.tex
Emphasize feature set differences earlier in the manuscript to underline the benchmarking idea of the study

@@
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
+Each feature set has distinct capabilities which differentiate them from the others.
+This can have both a positive and negative effect on the resulting performance which is one item this study aims to explore.
+For example, feature set VI misses certain parts of the spectral range as the defined indices only cover parts of such.
+Feature set NRI will introduce highly-correlated features which might cause issues for some algorithms.
 However, rather than only looking at these three groups, also the following combinations of such were taken into account:

### 2021-11-06 docs/00-manuscripts/mdpi/manuscript.tex
Explain in more detail why a pairwise-correlation variable selection was applied as a preprocessing step as suggested by Reviewer3

@@
 The pixel values of such were averaged and formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
-In total, \(\frac{121*122}{2} = 7471\) NRIs were calculated.
+In total, \(\frac{121*122}{2} = 7381\) NRIs were calculated.
 
@@
 Some individual features of such were removed before using the datasets for modelling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
-This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).
+This preprocessing step reduced the number of covariates for feature set VI to 86 (from 89).
+This decision was made to prevent issues for the subsequent tuning, filtering and model fitting steps which might occur when passing features with a pairwise correlation of (almost) one.
+Additional variable selection was solely devoted to the respective feature selection methods in this study.
 

### 2021-11-06 analysis/eval-performance.Rmd
chmod 644 -> 755

@@
     position = position_quasirandom(),
-    aes(label = paste0(filter, ",", round(rmse_aggr, 3))),
+    aes(label = paste0(filter, ",", round(rmse_aggr, 2))),
     size = 4,

### 2021-11-06 docs/00-manuscripts/mdpi/performance-svm-single-plot.tex
chmod 644 -> 755

--- 2020-03-05 docs/00-manuscripts/ieee/performance-svm-single-plot.tex
@@
-% latex table generated in R 3.6.1 by xtable 1.8-3 package
+% latex table generated in R 4.0.4 by xtable 1.8-4 package
 \begin{table}[ht!]
 \centering
-\caption{Single fold performances for learner SVM on the HR dataset without using a filter.} 
+\caption{Test fold performances in RMSE (p.p.) for learner SVM on the HR dataset without using a filter, showcasing performance variance on the fold level.
+For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot.}
 \label{tab:svm-single-fold-perf}
-\begin{tabular}{rlr}
+\begin{tabular}{rrl}
   \hline
- & Plot & RMSE \\ 
+ & RMSE & Test Plot \\ 
   \hline
-1 & Luiando & 9.00 \\ 
-  2 & Laukiz1 & 21.17 \\ 
-  3 & Laukiz2 & 54.26 \\ 
-  4 & Oiartzun & 28.05 \\ 
+1 & 54.26 & Laukiz1 \\ 
+  2 & 28.12 & Laukiz2 \\ 
+  3 & 9.00 & Luiando \\ 
+  4 & 21.17 & Oiartzun \\ 
    \hline

### 2021-11-07 docs/00-manuscripts/mdpi/manuscript.tex
minor editorial changes

@@
 $^{2}$ \quad NEIKER Tecnalia, Spain\\
-$^{3}$ \quad Ludwig-Maximilians-University Munich, Department of Statistics, Computational Statistics group, Germany
+$^{3}$ \quad Ludwig-Maximilians-Universität München, Department of Statistics, Computational Statistics group, Germany
 }
@@
 % Abstract (Do not insert blank lines, i.e. \\)
-\abstract{This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
-The effect of filter-based feature-selection methods on predictive performance was compared.
+\abstract{This study analyzed highly correlated, feature-rich datasets from hyperspectral remote sensing data using multiple statistical and machine-learning methods.
+The effect of filter-based feature selection methods on predictive performance was compared.
 Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
-Defoliation of trees (\%), derived from in-situ measurements in fall 2016, was modeled as a function of reflectance.
+Defoliation of trees (\%), derived from in-situ measurements from fall 2016, was modeled as a function of reflectance.
 Variable importance was assessed using permutation-based feature importance.
-Overall support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
+Overall, the support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
 The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
@@
 	\item  Random Forest (RF)
-	\item  Penalized Regression (L1 (lasso) and L2 (ridge))
-	\item  Support Vector Machine (SVM, RBF Kernel)
+	\item  Penalized Regression (with L1 / lasso and L2 / ridge penalties)
+	\item  Support Vector Machine (SVM, radial basis function Kernel)
 	\item  Featureless learner
@@
 
-RF and SVM are well established algorithms and widely used in (environmental) remote sensing.
-Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmark studies in recent years.
+RF and SVM are well-established algorithms and widely used in (environmental) remote sensing.
+Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmarking studies in recent years.
 Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by penalizing the coefficients of the model \cite{hastie2001}.
 Common penalties are 'lasso' (L1) and 'ridge' (L2).
-Ridge does not remove variables from the model (penalization to zero) but just shrinks them to effectively zero, keeping them in the model.
+Ridge regression does not remove variables from the model (penalization to zero) but it just shrinks them towards zero, keeping them in the model.
 A featureless learner was included for a baseline comparison.
 
-In total the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
+In total, the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
 The selected hyperparameter settings are shown in appendix \autoref{tab:hyperparameter_limits}.
-In addition code and data are available in the research compendium of this study (\url{https://doi.org/10.5281/zenodo.2635403}) to support open and reproducible science.
+All code and data are included in the research compendium of this study (\url{https://doi.org/10.5281/zenodo.2635403}).
 
@@
 \begin{itemize}
-	\item The raw hyperspectral band information (HR): no feature engineering) %chktex 13
-	\item Vegetation Indices (\ac{VI}s): expert-based feature engineering)
-	\item Normalized Ratio Indices (\ac{NRI}s): data-driven feature engineering)
+	\item The raw hyperspectral band information (HR): no feature engineering %chktex 13
+	\item Vegetation Indices (\ac{VI}s): expert-based feature engineering
+	\item Normalized Ratio Indices (\ac{NRI}s): data-driven feature engineering
 \end{itemize}
@@
 Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
-Each feature set has distinct capabilities which differentiate them from the others.
-This can have both a positive and negative effect on the resulting performance which is one item this study aims to explore.
-For example, feature set VI misses certain parts of the spectral range as the defined indices only cover parts of such.
-Feature set NRI will introduce highly-correlated features which might cause issues for some algorithms.
-However, rather than only looking at these three groups, also the following combinations of such were taken into account:
+Each feature set has distinct capabilities that differentiate it from the others.
+This can have both a positive and negative effect on the resulting performance, which is one item this study aims to explore.
+For example, feature set VI misses certain parts of the spectral range as the chosen indices only use specific spectral bands.
+Feature set NRI will introduce highly-correlated features, for which some algorithms may be more suitable than others.
+In addition to these feature sets, the following combinations of feature sets were also compared:
 
@@
 
-Some individual features of such were removed before using the datasets for modelling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
+Some individual features were removed before using the datasets for modelling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
 This preprocessing step reduced the number of covariates for feature set VI to 86 (from 89).
 This decision was made to prevent issues for the subsequent tuning, filtering and model fitting steps which might occur when passing features with a pairwise correlation of (almost) one.
-Additional variable selection was solely devoted to the respective feature selection methods in this study.
+Additional variable selection was solely devoted to the respective feature selection methods in this study. <!-- FIXME: I don't understand this sentence. -->
 
@@
 Hyperparameters were tuned using \ac{MBO} within a nested spatial \ac{CV} \cite{mlrmbo, binder2020, schratz2019}.
-In MBO first \textit{n} randomly chosen hyperparameter settings out of a user defined search space are composed.
-After these \textit{n} settings have been evaluated, one new setting, which is going to be evaluated next, is proposed by a fitted surrogate model (by default a kriging method).
+In MBO, first \textit{n} hyperparameter settings are randomly chosen from a user-defined search space.
+After these \textit{n} settings have been evaluated, one new setting, which is going to be evaluated next, is proposed by a fitted surrogate model (by default, a kriging method).
 This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.

### 2021-11-07 docs/00-manuscripts/mdpi/manuscript.tex
numerous minor editorial changes

@@
 	\acro{FFS}{forward feature selection}
-	\acro{FPR}{false positive rate}
+	\acro{FPR}{false-positive rate}
 	\acro{FS}{feature selection}
@@
 	\acro{LiDAR}{light detection and ranging}
-	\acro{LOWESS}{locally weighted scatter plot smoothing}
+	\acro{LOWESS}{locally weighted scatterplot smoothing}
 	\acro{MARS}{multivariate adaptive regression splines}
@@
 	\acro{PISR}{potential incoming solar radiation}
-	\acro{PLS}{partial least-squares}
+	\acro{PLS}{partial least squares}
 	\acro{POV}{proportion of variance explained}
@@
 	\acro{SDM}{species distribution modeling}
-	\acro{SMBO}{sequential-based model optimization}
+	\acro{SMBO}{sequential model-based optimization}
 	\acro{SVM}{support vector machine}
-	\acro{TPR}{true positive rate}
+	\acro{TPR}{true-positive rate}
 	\acro{VI}{vegetation index}
@@
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).
-Besides the availability of code and manuscript sources, a static webpage is available at (\url{https://pat-s.github.io/2019-feature-selection}), listing more side-analyses that were carried out during the creation of this study.
+Apart from the availability of code and manuscript sources, a static webpage is available at (\url{https://pat-s.github.io/2019-feature-selection}), which includes additional side analyses that were carried out during the creation of this study.
 
@@
 
-\subsection{PCA-based variance analysis of feature sets}
+\subsection{Principal component analysis of feature sets}
 
@@
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
-Regression with ridge (L2) and lasso (L1) penalty showed their best performances on the NRI feature set (\autoref{tab:best-learner-perf}).
-Difference to other feature sets were very small for lasso (below one p.p.) and a bit higher for ridge (between two and five p.p. (VI)).
-XGBoost shows bad performances for some feature sets and fills the ten last places of the ranking (\autoref{tab:perf-worst-10}).
+Regression with ridge (L2) and lasso (L1) penalties showed their best performances on the NRI feature set (\autoref{tab:best-learner-perf}).
+Differences to other feature sets were very small for lasso (below one p.p.) and a bit larger for ridge regression (between two and five p.p.).
+XGBoost showed very poor performances for some feature sets and fills the ten last places of the ranking (\autoref{tab:perf-worst-10}).
 Especially when combined with PCA, the algorithm was not able to achieve adequate performances.
@@
 SVM showed no variation in performance across filters (\autoref{fig:filter-effects-no-filter}).
-Using filters for RF showed a substantial increase in performance for all tasks with the exception of NRI, for which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
-XGBoost showed a high dependency on filtering the data: In two out of six tasks using no filter resulted in the worst or second worst performance.
-XGBoost shows the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p.
+The use of filters with RF resulted in a substantial increase in performance in all tasks with the exception of NRI, in which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
+XGBoost's performance depended strongly on feature selection: In two out of six tasks, using no filter resulted in the worst or second worst performance.
+XGBoost showed the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p.
 (\enquote{CMIM} vs.
@@
 The Borda filter achieved in only one instance (RF on HR-NRI-VI) the best performance among the used filters (\autoref{fig:filter-effects-borda}).
-For RF and XGBoost it most often ranked within the first 50\% with respect to all filters of a specific task.
+For RF and XGBoost it most often ranked within the better half among all filters of a specific task.
 For XGBoost on the NRI task, the Borda filter scored the second worst performance.
@@
 % Selected percentage of features across plots
-Large differences were observed between the numbers of features selected during tuning for the subsequent fitting process across each learner and plot.
-While for RF least features were selected when Luiando or Oiartzun were the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
+The number of features selected based on model tuning strongly varied across learners and plots.
+While for RF, the least features were selected when Luiando or Oiartzun was the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
 RF and SVM used very few features only when Laukiz1 was the test set, (three and one, respectively) whereas this was the plot for which XGBoost used almost all features ($n = 1226$).
-Overall, SVM used the least features across all plots among all learners, with 24\% for test set Laukiz2 being the highest single plot value.
-In general every learner behaved quite differently for each plot and no overall pattern could be observed.
+Overall, SVM used the least features across all plots and among all learners, the single highest fraction being 24\% for test set Laukiz2.
+In general, all learners behaved rather differently on the various plots, and therefore no overall pattern could be identified.
 
@@
 Different feature sets are shown on the y-axis.
-Labels show the feature selection method (e.g.
-NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda').}\label{fig:perf-result}
+Labels show the feature selection method (e.g., NF = no filter, Car = 'Carscore', Info Gain = 'Information Gain', Borda = 'Borda').}\label{fig:perf-result}
 	\end{center}
@@
 The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
-For dataset HR most features cluster around the infrared region (920 nm - 1000 nm) (six out of ten) while for VI eight out of ten concentrate on the wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
-For feature set HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important (causing a mean decrease in RMSE of around 1 p.p.).
-Overall, most features (excluding the top five respectively) showed only a small importance with average decreases in RMSE below 0.5 p.p..
+For the HR dataset, most (i.e., six out of ten) relevant features clustered around the infrared region (920 nm - 1000 nm), while for VI, eight out of ten concentrate within the wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
+For HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important, being assciated with a mean decrease in RMSE of around 1 p.p.
+Overall, apart from the top five features, the vast majority of features showed only a small importance with average decreases in RMSE below 0.5 p.p.
 

### 2021-11-09 docs/00-manuscripts/mdpi/manuscript.tex
add brenning2021 reference

@@
 Methods like \ac{PDP} or permutation-based approaches may produce unreliable estimates in such scenarios because unrealistic situations between covariates are created \cite{molnar2019}.
-The development of robust methods which enable an unbiased estimation of feature importance for highly correlated variables are subject to current research.
+The development of robust methods which enable an unbiased estimation of feature importance for highly correlated variables are subject to current research \cite{brenning2021}.
 

### 2021-11-11 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

--- 2021-04-12 analysis/eda.Rmd
@@
 library("patchwork")
+library("here")
 ```
@@
 ```{r buffer-scenarios-1, echo=FALSE}
-knitr::include_graphics("inst/tree-buffer-vis.png")
+knitr::include_graphics(here("docs/figure/tree-buffer-vis.png"))
 ```
@@
 ```{r buffer-scenarios-2, echo=FALSE}
-knitr::include_graphics("inst/tree-buffer-vis-max.png")
+knitr::include_graphics(here("docs/figure/tree-buffer-vis-max.png"))
 ```

### 2021-11-12 docs/00-manuscripts/mdpi/performance-best-per-learner.tex
ensure that only 122 instead uf 126 bands are used for all feature sets

--- 2021-09-02 docs/00-manuscripts/mdpi/performance-best-per-learner.tex
@@
   \hline
-1 & HR & SVM & Relief & 28.12 & 19.12 \\ 
-  2 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
-  3 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
-  4 & - & regr.featureless & No Filter & 31.16 & 15.03 \\ 
-  5 & HR & RF & Relief & 31.25 & 14.96 \\ 
-  6 & HR & XGBoost & Car & 34.07 & 13.57 \\ 
+1 & NRI-VI & SVM & Info Gain & 27.915 & 18.970 \\ 
+  2 & NRI & RF & Relief & 30.842 & 12.028 \\ 
+  3 & HR & XGBoost & Info Gain & 31.165 & 15.025 \\ 
+  4 & NRI & Lasso-MBO & No Filter & 31.165 & 15.025 \\ 
+  5 & NRI & Ridge-MBO & No Filter & 31.165 & 15.025 \\ 
+  6 & - & regr.featureless & No Filter & 31.165 & 15.025 \\ 
    \hline

### 2021-11-12 docs/00-manuscripts/mdpi/tune-perc-sel-features.tex
ensure that only 122 instead uf 126 bands are used for all feature sets

--- 2021-06-28 docs/00-manuscripts/ieee/tune-perc-sel-features.tex
@@
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.21573 & 3/1249 \\ 
-   & Laukiz2 & 93.37086 & 1268/1357 \\ 
-   & Luiando & 99.97268 & 1507/1507 \\ 
-   & Oiartzun & 0.00405 & 1/1311 \\ 
-  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.06200 & 1/1249 \\ 
-   & Laukiz2 & 23.92820 & 325/1357 \\ 
-   & Luiando & 20.73514 & 313/1507 \\ 
-   & Oiartzun & 4.10636 & 54/1311 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 98.12213 & 1226/1249 \\ 
-   & Laukiz2 & 0.51526 & 7/1357 \\ 
-   & Luiando & 30.84900 & 465/1507 \\ 
-   & Oiartzun & 83.61298 & 1097/1311 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.00614 & 1/1249 \\ 
+   & Laukiz2 & 16.58887 & 226/1357 \\ 
+   & Luiando & 0.06681 & 2/1507 \\ 
+   & Oiartzun & 65.06532 & 854/1311 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.50917 & 7/1249 \\ 
+   & Laukiz2 & 91.95721 & 1248/1357 \\ 
+   & Luiando & 0.50917 & 8/1507 \\ 
+   & Oiartzun & 19.70660 & 259/1311 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 93.70844 & 1171/1249 \\ 
+   & Laukiz2 & 3.51986 & 48/1357 \\ 
+   & Luiando & 0.12474 & 2/1507 \\ 
+   & Oiartzun & 85.28181 & 1119/1311 \\ 
    \bottomrule

### 2021-11-12 R/01-fun-download.R

@@ -91,6 +91,24 @@ download_hyperspectral <- function(url, paper = TRUE) {
       ),
       ~ raster::brick(.x)
     )
+    if (!file.exists("data/hyperspectral/paper/122/B103_P1P_A090F04_ATM_S.tif")) {
+      fs::dir_create("data/hyperspectral/paper/122")
+      # we need to write them to disk to create a RasterBrick as follow up operations
+      # only work with bricks
+
+      names <- purrr::map_chr(files, ~ .x@file@name)
+
+      purrr::walk2(files, names, ~ raster::subset(.x, 5:126,
+        filename = sprintf("data/hyperspectral/paper/122/%s", basename(.y)),
+        progress = "text", overwrite = TRUE
+      ))
+    }
+    files <- purrr::map(
+      list.files("data/hyperspectral/paper/122",
+        full.names = TRUE, pattern = ".tif$"
+      ),
+      ~ raster::brick(.x)
+    )
   } else {
     if (!file.exists("data/hyperspectral/all/B101_P1N_A090F03_ATM_S.tif")) {
       curl::curl_download(url,

### 2021-11-12 R/03-fun-hyperspectral.R

@@ -159,7 +159,7 @@ extract_bands_to_plot <- function(plot_name,
                                   hyperspectral_bands) {
   if (!is.null(buffer)) {
     out_bands <- purrr::map(buffer, function(x) {
-      raster::extract(hyperspectral_bands[[plot_name]][[5:126]],
+      raster::extract(hyperspectral_bands[[plot_name]],
         tree_data[[plot_name]],
         buffer = x,
         fun = mean,
@@ -168,7 +168,7 @@ extract_bands_to_plot <- function(plot_name,
       )
     })
   } else {
-    out_bands <- list(raster::extract(hyperspectral_bands[[plot_name]][[5:126]],
+    out_bands <- list(raster::extract(hyperspectral_bands[[plot_name]],
       tree_data[[plot_name]],
       fun = mean,
       df = TRUE,

### 2021-11-12 code/010-download.R

@@ -97,6 +97,28 @@ download_data_plan <- drake_plan(
     )
   ),
 
+  # excluding the first four faulty bands
+  wavelength_reduced = target(
+    c(
+      421.81, 426.27, 430.73, 435.2,
+      439.69, 444.18, 448.68, 453.18, 457.69, 462.22, 466.75, 471.29,
+      475.83, 480.39, 484.95, 489.52, 494.09, 498.68, 503.26, 507.86,
+      512.47, 517.08, 521.7, 526.32, 530.95, 535.58, 540.23, 544.88,
+      549.54, 554.2, 558.86, 563.54, 568.22, 572.9, 577.6, 582.29,
+      586.99, 591.7, 596.41, 601.13, 605.85, 610.58, 615.31, 620.05,
+      624.79, 629.54, 634.29, 639.04, 643.8, 648.56, 653.33, 658.1,
+      662.88, 667.66, 672.44, 677.23, 682.02, 686.81, 691.6, 696.4,
+      701.21, 706.01, 710.82, 715.64, 720.45, 725.27, 730.09, 734.91,
+      739.73, 744.56, 749.39, 754.22, 759.05, 763.89, 768.72, 773.56,
+      778.4, 783.24, 788.08, 792.93, 797.77, 802.62, 807.47, 812.32,
+      817.17, 822.02, 826.87, 831.72, 836.57, 841.42, 846.28, 851.13,
+      855.98, 860.83, 865.69, 870.54, 875.39, 880.24, 885.09, 889.94,
+      894.79, 899.64, 904.49, 909.34, 914.18, 919.03, 923.87, 928.71,
+      933.55, 938.39, 943.23, 948.07, 952.9, 957.73, 962.56, 967.39,
+      972.22, 977.04, 981.87, 986.68, 991.5, 996.31
+    )
+  ),
+
   indices = target(
     c(
       "Boochs", "Boochs2", "CARI", "Carter", "Carter2", "Carter3", "Carter4", "Carter5",

### 2021-11-12 analysis/inspect-glmnet-models.Rmd

--- 2020-03-11 analysis/inspect-glmnet-models.Rmd
@@
   benchmark_models_new_penalized_mbo_buffer2,
-  task_new_buffer2
+  task,
+  task_reduced_cor
 )
@@
 
-obs_train_f4 <- as.matrix(task_new_buffer2[[2]]$env$data[train_inds_fold4, getTaskFeatureNames(task_new_buffer2[[2]])])
-target_f4 <- getTaskTargets(task_new_buffer2[[2]])[train_inds_fold4]
+obs_train_f4 <- as.matrix(task[[2]]$env$data[train_inds_fold4, getTaskFeatureNames(task[[2]])])
+target_f4 <- getTaskTargets(task[[2]])[train_inds_fold4]
 ```
@@
 
-obs_pred_f4 <- as.matrix(task_new_buffer2[[2]]$env$data[pred_inds_fold4, getTaskFeatureNames(task_new_buffer2[[2]])])
+obs_pred_f4 <- as.matrix(task[[2]]$env$data[pred_inds_fold4, getTaskFeatureNames(task[[2]])])
 
@@
 ```{r}
-truth <- task_new_buffer2[[2]]$env$data[pred_inds_fold4, "defoliation"]
+truth <- task[[2]]$env$data[pred_inds_fold4, "defoliation"]
 
@@
 lrn <- makeLearner("regr.cvglmnet", alpha = 0)
-task_f4 <- subsetTask(task_new_buffer2[[2]], train_inds_fold4)
+task_f4 <- subsetTask(task[[2]], train_inds_fold4)
 
@@
 Looks ok - they are all within a normal range with respectv to the PDP estimates.
+
+## Train/predict via {glmnet} manually on dataset with filtered feature correlation
+
+To inspect further, let's refit a {glmnet} model directly on the training data of Fold 4 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:
+
+```{r}
+train_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["train.inds"]][[4]]
+
+obs_train_f4 <- as.matrix(task_reduced_cor[[2]]$env$data[train_inds_fold4, getTaskFeatureNames(task_reduced_cor[[2]])])
+target_f4 <- getTaskTargets(task_reduced_cor[[2]])[train_inds_fold4]
+```
+
+Fit `cv.glmnet`
+
+```{r}
+set.seed(1)
+modf4 <- glmnet::cv.glmnet(obs_train_f4, target_f4, alpha = 0)
+
+modf4$lambda.1se
+```
+
+Predict on Laukiz 2 now.
+
+```{r}
+pred_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["test.inds"]][[4]]
+
+obs_pred_f4 <- as.matrix(task_reduced_cor[[2]]$env$data[pred_inds_fold4, getTaskFeatureNames(task_reduced_cor[[2]])])
+
+pred <- predict(modf4, newx = obs_pred_f4, s = modf4$lambda.1se)
+```
+
+Calculate the error
+
+```{r}
+truth <- task_reduced_cor[[2]]$env$data[pred_inds_fold4, "defoliation"]
+
+mlr:::measureRMSE(truth, pred)
+```

### 2021-11-12 docs/00-manuscripts/mdpi/Definitions/journalnames.tex

%% Updated on 18 June 2021
%,% Website Name,%,%,% Website Name,%%,% Full Name,%%,% Short Name,%%,% DOI abbr,%%,% ISSN,% \continuouspages
%\DeclareOption{}{\gdef\@journal{}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{acoustics}{\gdef\@journal{acoustics}\gdef\@journalfull{Acoustics}\gdef\@journalshort{Acoustics}\gdef\@doiabbr{acoustics}\gdef\@ISSN{2624-599X}\gdef\@continuouspages{yes}}
\DeclareOption{actuators}{\gdef\@journal{actuators}\gdef\@journalfull{Actuators}\gdef\@journalshort{Actuators}\gdef\@doiabbr{act}\gdef\@ISSN{2076-0825}}
\DeclareOption{addictions}{\gdef\@journal{addictions}\gdef\@journalfull{Addictions}\gdef\@journalshort{Addictions}\gdef\@doiabbr{}\gdef\@ISSN{0006-0006}}
\DeclareOption{admsci}{\gdef\@journal{admsci}\gdef\@journalfull{Administrative Sciences}\gdef\@journalshort{Adm. Sci.}\gdef\@doiabbr{admsci}\gdef\@ISSN{2076-3387}}
\DeclareOption{adolescents}{\gdef\@journal{adolescents}\gdef\@journalfull{Adolescents}\gdef\@journalshort{Adolescents}\gdef\@doiabbr{adolescents}\gdef\@ISSN{2673-7051}\gdef\@continuouspages{yes}}
\DeclareOption{aerospace}{\gdef\@journal{aerospace}\gdef\@journalfull{Aerospace}\gdef\@journalshort{Aerospace}\gdef\@doiabbr{aerospace}\gdef\@ISSN{2226-4310}}
\DeclareOption{agriculture}{\gdef\@journal{agriculture}\gdef\@journalfull{Agriculture}\gdef\@journalshort{Agriculture}\gdef\@doiabbr{agriculture}\gdef\@ISSN{2077-0472}}
\DeclareOption{agriengineering}{\gdef\@journal{agriengineering}\gdef\@journalfull{AgriEngineering}\gdef\@journalshort{AgriEngineering}\gdef\@doiabbr{agriengineering}\gdef\@ISSN{2624-7402}\gdef\@continuouspages{yes}}
\DeclareOption{agronomy}{\gdef\@journal{agronomy}\gdef\@journalfull{Agronomy}\gdef\@journalshort{Agronomy}\gdef\@doiabbr{agronomy}\gdef\@ISSN{2073-4395}}
\DeclareOption{ai}{\gdef\@journal{ai}\gdef\@journalfull{AI}\gdef\@journalshort{AI}\gdef\@doiabbr{ai}\gdef\@ISSN{2673-2688}\gdef\@continuouspages{yes}}
\DeclareOption{algorithms}{\gdef\@journal{algorithms}\gdef\@journalfull{Algorithms}\gdef\@journalshort{Algorithms}\gdef\@doiabbr{a}\gdef\@ISSN{1999-4893}}
\DeclareOption{allergies}{\gdef\@journal{allergies}\gdef\@journalfull{Allergies}\gdef\@journalshort{Allergies}\gdef\@doiabbr{allergies}\gdef\@ISSN{2313-5786}}
\DeclareOption{analytica}{\gdef\@journal{analytica}\gdef\@journalfull{Analytica}\gdef\@journalshort{Analytica}\gdef\@doiabbr{analytica}\gdef\@ISSN{2673-4532}\gdef\@continuouspages{yes}}
\DeclareOption{animals}{\gdef\@journal{animals}\gdef\@journalfull{Animals}\gdef\@journalshort{Animals}\gdef\@doiabbr{ani}\gdef\@ISSN{2076-2615}}
\DeclareOption{antibiotics}{\gdef\@journal{antibiotics}\gdef\@journalfull{Antibiotics}\gdef\@journalshort{Antibiotics}\gdef\@doiabbr{antibiotics}\gdef\@ISSN{2079-6382}}
\DeclareOption{antibodies}{\gdef\@journal{antibodies}\gdef\@journalfull{Antibodies}\gdef\@journalshort{Antibodies}\gdef\@doiabbr{antib}\gdef\@ISSN{2073-4468}}
\DeclareOption{antioxidants}{\gdef\@journal{antioxidants}\gdef\@journalfull{Antioxidants}\gdef\@journalshort{Antioxidants}\gdef\@doiabbr{antiox}\gdef\@ISSN{2076-3921}}
\DeclareOption{appliedchem}{\gdef\@journal{appliedchem}\gdef\@journalfull{AppliedChem}\gdef\@journalshort{AppliedChem}\gdef\@doiabbr{appliedchem}\gdef\@ISSN{2673-9623}\gdef\@continuouspages{yes}}
\DeclareOption{appliedmath}{\gdef\@journal{appliedmath}\gdef\@journalfull{AppliedMath}\gdef\@journalshort{AppliedMath}\gdef\@doiabbr{appliedmath}\gdef\@ISSN{2673-9909}\gdef\@continuouspages{yes}}
\DeclareOption{applmech}{\gdef\@journal{applmech}\gdef\@journalfull{Applied Mechanics}\gdef\@journalshort{Appl. Mech.}\gdef\@doiabbr{applmech}\gdef\@ISSN{2673-3161}\gdef\@continuouspages{yes}}
\DeclareOption{applmicrobiol}{\gdef\@journal{applmicrobiol}\gdef\@journalfull{Applied Microbiology}\gdef\@journalshort{Appl. Microbiol.}\gdef\@doiabbr{applmicrobiol}\gdef\@ISSN{2673-8007}\gdef\@continuouspages{yes}}
\DeclareOption{applnano}{\gdef\@journal{applnano}\gdef\@journalfull{Applied Nano}\gdef\@journalshort{Appl. Nano}\gdef\@doiabbr{applnano}\gdef\@ISSN{2673-3501}\gdef\@continuouspages{yes}}
\DeclareOption{applsci}{\gdef\@journal{applsci}\gdef\@journalfull{Applied Sciences}\gdef\@journalshort{Appl. Sci.}\gdef\@doiabbr{app}\gdef\@ISSN{2076-3417}}
\DeclareOption{aquacj}{\gdef\@journal{aquacj}\gdef\@journalfull{Aquaculture Journal}\gdef\@journalshort{Aquac. J.}\gdef\@doiabbr{aj}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{architecture}{\gdef\@journal{architecture}\gdef\@journalfull{Architecture}\gdef\@journalshort{Architecture}\gdef\@doiabbr{architecture}\gdef\@ISSN{2673-8945}\gdef\@continuouspages{yes}}
\DeclareOption{arts}{\gdef\@journal{arts}\gdef\@journalfull{Arts}\gdef\@journalshort{Arts}\gdef\@doiabbr{arts}\gdef\@ISSN{2076-0752}}
\DeclareOption{asc}{\gdef\@journal{asc}\gdef\@journalfull{Automatic Systems and Control}\gdef\@journalshort{Autom. Syst. Control}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{asi}{\gdef\@journal{asi}\gdef\@journalfull{Applied System Innovation}\gdef\@journalshort{Appl. Syst. Innov.}\gdef\@doiabbr{asi}\gdef\@ISSN{2571-5577}}
\DeclareOption{atmosphere}{\gdef\@journal{atmosphere}\gdef\@journalfull{Atmosphere}\gdef\@journalshort{Atmosphere}\gdef\@doiabbr{atmos}\gdef\@ISSN{2073-4433}}
\DeclareOption{atoms}{\gdef\@journal{atoms}\gdef\@journalfull{Atoms}\gdef\@journalshort{Atoms}\gdef\@doiabbr{atoms}\gdef\@ISSN{2218-2004}}
\DeclareOption{audiolres}{\gdef\@journal{audiolres}\gdef\@journalfull{Audiology Research}\gdef\@journalshort{Audiol. Res.}\gdef\@doiabbr{audiolres}\gdef\@ISSN{2039-4349}\gdef\@continuouspages{yes}}
\DeclareOption{automation}{\gdef\@journal{automation}\gdef\@journalfull{Automation}\gdef\@journalshort{Automation}\gdef\@doiabbr{automation}\gdef\@ISSN{2673-4052}\gdef\@continuouspages{yes}}
\DeclareOption{axioms}{\gdef\@journal{axioms}\gdef\@journalfull{Axioms}\gdef\@journalshort{Axioms}\gdef\@doiabbr{axioms}\gdef\@ISSN{2075-1680}}
\DeclareOption{bacteria}{\gdef\@journal{bacteria}\gdef\@journalfull{Bacteria}\gdef\@journalshort{Bacteria}\gdef\@doiabbr{bacteria}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{batteries}{\gdef\@journal{batteries}\gdef\@journalfull{Batteries}\gdef\@journalshort{Batteries}\gdef\@doiabbr{batteries}\gdef\@ISSN{2313-0105}}
\DeclareOption{bdcc}{\gdef\@journal{bdcc}\gdef\@journalfull{Big Data and Cognitive Computing}\gdef\@journalshort{Big Data Cogn. Comput.}\gdef\@doiabbr{bdcc}\gdef\@ISSN{2504-2289}}
\DeclareOption{behavsci}{\gdef\@journal{behavsci}\gdef\@journalfull{Behavioral Sciences}\gdef\@journalshort{Behav. Sci.}\gdef\@doiabbr{bs}\gdef\@ISSN{2076-328X}}
\DeclareOption{beverages}{\gdef\@journal{beverages}\gdef\@journalfull{Beverages}\gdef\@journalshort{Beverages}\gdef\@doiabbr{beverages}\gdef\@ISSN{2306-5710}}
\DeclareOption{biochem}{\gdef\@journal{biochem}\gdef\@journalfull{BioChem}\gdef\@journalshort{BioChem}\gdef\@doiabbr{biochem}\gdef\@ISSN{2673-6411}\gdef\@continuouspages{yes}}
\DeclareOption{bioengineering}{\gdef\@journal{bioengineering}\gdef\@journalfull{Bioengineering}\gdef\@journalshort{Bioengineering}\gdef\@doiabbr{bioengineering}\gdef\@ISSN{2306-5354}}
\DeclareOption{biologics}{\gdef\@journal{biologics}\gdef\@journalfull{Biologics}\gdef\@journalshort{Biologics}\gdef\@doiabbr{biologics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biology}{\gdef\@journal{biology}\gdef\@journalfull{Biology}\gdef\@journalshort{Biology}\gdef\@doiabbr{biology}\gdef\@ISSN{2079-7737}}
\DeclareOption{biomass}{\gdef\@journal{biomass}\gdef\@journalfull{Biomass}\gdef\@journalshort{Biomass}\gdef\@doiabbr{biomass}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biomechanics}{\gdef\@journal{biomechanics}\gdef\@journalfull{Biomechanics}\gdef\@journalshort{Biomechanics}\gdef\@doiabbr{biomechanics}\gdef\@ISSN{2673-7078}\gdef\@continuouspages{yes}}
\DeclareOption{biomed}{\gdef\@journal{biomed}\gdef\@journalfull{BioMed}\gdef\@journalshort{BioMed}\gdef\@doiabbr{biomed}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biomedicines}{\gdef\@journal{biomedicines}\gdef\@journalfull{Biomedicines}\gdef\@journalshort{Biomedicines}\gdef\@doiabbr{biomedicines}\gdef\@ISSN{2227-9059}}
\DeclareOption{biomedinformatics}{\gdef\@journal{biomedinformatics}\gdef\@journalfull{Biomedinformatics}\gdef\@journalshort{Biomedinformatics}\gdef\@doiabbr{biomedinformatics}\gdef\@ISSN{2673-7426}\gdef\@continuouspages{yes}}
\DeclareOption{biomimetics}{\gdef\@journal{biomimetics}\gdef\@journalfull{Biomimetics}\gdef\@journalshort{Biomimetics}\gdef\@doiabbr{biomimetics}\gdef\@ISSN{2313-7673}}
\DeclareOption{biomolecules}{\gdef\@journal{biomolecules}\gdef\@journalfull{Biomolecules}\gdef\@journalshort{Biomolecules}\gdef\@doiabbr{biom}\gdef\@ISSN{2218-273X}}
\DeclareOption{biophysica}{\gdef\@journal{biophysica}\gdef\@journalfull{Biophysica}\gdef\@journalshort{Biophysica}\gdef\@doiabbr{biophysica}\gdef\@ISSN{2673-4125}\gdef\@continuouspages{yes}}
\DeclareOption{biosensors}{\gdef\@journal{biosensors}\gdef\@journalfull{Biosensors}\gdef\@journalshort{Biosensors}\gdef\@doiabbr{bios}\gdef\@ISSN{2079-6374}}
\DeclareOption{biotech}{\gdef\@journal{biotech}\gdef\@journalfull{BioTech}\gdef\@journalshort{BioTech}\gdef\@doiabbr{biotech}\gdef\@ISSN{2673-6284}}
\DeclareOption{birds}{\gdef\@journal{birds}\gdef\@journalfull{Birds}\gdef\@journalshort{Birds}\gdef\@doiabbr{birds}\gdef\@ISSN{2673-6004}\gdef\@continuouspages{yes}}
\DeclareOption{bloods}{\gdef\@journal{bloods}\gdef\@journalfull{Bloods}\gdef\@journalshort{Bloods}\gdef\@doiabbr{bloods}\gdef\@ISSN{2673-415X}}
\DeclareOption{blsf}{\gdef\@journal{	blsf}\gdef\@journalfull{Biology and Life Sciences Forum}\gdef\@journalshort{Biol. Life Sci. Forum}\gdef\@doiabbr{blsf	}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{brainsci}{\gdef\@journal{brainsci}\gdef\@journalfull{Brain Sciences}\gdef\@journalshort{Brain Sci.}\gdef\@doiabbr{brainsci}\gdef\@ISSN{2076-3425}}
\DeclareOption{breath}{\gdef\@journal{breath}\gdef\@journalfull{Breath}\gdef\@journalshort{Breath}\gdef\@doiabbr{breath}\gdef\@ISSN{}}
\DeclareOption{buildings}{\gdef\@journal{buildings}\gdef\@journalfull{Buildings}\gdef\@journalshort{Buildings}\gdef\@doiabbr{buildings}\gdef\@ISSN{2075-5309}}
\DeclareOption{businesses}{\gdef\@journal{businesses}\gdef\@journalfull{Businesses}\gdef\@journalshort{Businesses}\gdef\@doiabbr{businesses}\gdef\@ISSN{2673-7116}\gdef\@continuouspages{yes}}
\DeclareOption{cancers}{\gdef\@journal{cancers}\gdef\@journalfull{Cancers}\gdef\@journalshort{Cancers}\gdef\@doiabbr{cancers}\gdef\@ISSN{2072-6694}}
\DeclareOption{carbon}{\gdef\@journal{carbon}\gdef\@journalfull{C}\gdef\@journalshort{C}\gdef\@doiabbr{c}\gdef\@ISSN{2311-5629}}
\DeclareOption{cardiogenetics}{\gdef\@journal{cardiogenetics}\gdef\@journalfull{Cardiogenetics}\gdef\@journalshort{Cardiogenetics}\gdef\@doiabbr{cardiogenetics}\gdef\@ISSN{2035-8148}\gdef\@continuouspages{yes}}
\DeclareOption{catalysts}{\gdef\@journal{catalysts}\gdef\@journalfull{Catalysts}\gdef\@journalshort{Catalysts}\gdef\@doiabbr{catal}\gdef\@ISSN{2073-4344}}
\DeclareOption{cells}{\gdef\@journal{cells}\gdef\@journalfull{Cells}\gdef\@journalshort{Cells}\gdef\@doiabbr{cells}\gdef\@ISSN{2073-4409}}
\DeclareOption{ceramics}{\gdef\@journal{ceramics}\gdef\@journalfull{Ceramics}\gdef\@journalshort{Ceramics}\gdef\@doiabbr{ceramics}\gdef\@ISSN{2571-6131}\gdef\@continuouspages{yes}}
\DeclareOption{challenges}{\gdef\@journal{challenges}\gdef\@journalfull{Challenges}\gdef\@journalshort{Challenges}\gdef\@doiabbr{challe}\gdef\@ISSN{2078-1547}}
\DeclareOption{chemengineering}{\gdef\@journal{chemengineering}\gdef\@journalfull{ChemEngineering}\gdef\@journalshort{ChemEngineering}\gdef\@doiabbr{chemengineering}\gdef\@ISSN{2305-7084}}
\DeclareOption{chemistry}{\gdef\@journal{chemistry}\gdef\@journalfull{Chemistry}\gdef\@journalshort{Chemistry}\gdef\@doiabbr{chemistry}\gdef\@ISSN{2624-8549}\gdef\@continuouspages{yes}}
\DeclareOption{chemosensors}{\gdef\@journal{chemosensors}\gdef\@journalfull{Chemosensors}\gdef\@journalshort{Chemosensors}\gdef\@doiabbr{chemosensors}\gdef\@ISSN{2227-9040}}
\DeclareOption{chemproc}{\gdef\@journal{chemproc}\gdef\@journalfull{Chemistry Proceedings}\gdef\@journalshort{Chem. Proc.}\gdef\@doiabbr{chemproc}\gdef\@ISSN{2673-4583}}
\DeclareOption{children}{\gdef\@journal{children}\gdef\@journalfull{Children}\gdef\@journalshort{Children}\gdef\@doiabbr{children}\gdef\@ISSN{2227-9067}}
\DeclareOption{cimb}{\gdef\@journal{cimb}\gdef\@journalfull{Current Issues in Molecular Biology}\gdef\@journalshort{Curr. Issues Mol. Biol.}\gdef\@doiabbr{cimb}\gdef\@ISSN{1467-3045}\gdef\@continuouspages{yes}}
\DeclareOption{civileng}{\gdef\@journal{civileng}\gdef\@journalfull{CivilEng}\gdef\@journalshort{CivilEng}\gdef\@doiabbr{civileng}\gdef\@ISSN{2673-4109}\gdef\@continuouspages{yes}}
\DeclareOption{cleantechnol}{\gdef\@journal{cleantechnol}\gdef\@journalfull{Clean Technologies}\gdef\@journalshort{Clean Technol.}\gdef\@doiabbr{cleantechnol}\gdef\@ISSN{2571-8797}\gdef\@continuouspages{yes}}
\DeclareOption{climate}{\gdef\@journal{climate}\gdef\@journalfull{Climate}\gdef\@journalshort{Climate}\gdef\@doiabbr{cli}\gdef\@ISSN{2225-1154}}
\DeclareOption{clinpract}{\gdef\@journal{clinpract}\gdef\@journalfull{Clinics and Practice}\gdef\@journalshort{Clin. Pract.}\gdef\@doiabbr{clinpract}\gdef\@ISSN{2039-7283}\gdef\@continuouspages{yes}}
\DeclareOption{clockssleep}{\gdef\@journal{clockssleep}\gdef\@journalfull{Clocks \& Sleep}\gdef\@journalshort{Clocks\&Sleep}\gdef\@doiabbr{clockssleep}\gdef\@ISSN{2624-5175}\gdef\@continuouspages{yes}}
\DeclareOption{cmd}{\gdef\@journal{cmd}\gdef\@journalfull{Corrosion and Materials Degradation}\gdef\@journalshort{Corros. Mater. Degrad.}\gdef\@doiabbr{cmd}\gdef\@ISSN{2624-5558}\gdef\@continuouspages{yes}}
\DeclareOption{coasts}{\gdef\@journal{coasts}\gdef\@journalfull{Coasts}\gdef\@journalshort{Coasts}\gdef\@doiabbr{coasts}\gdef\@ISSN{2673-964X}\gdef\@continuouspages{yes}}
\DeclareOption{coatings}{\gdef\@journal{coatings}\gdef\@journalfull{Coatings}\gdef\@journalshort{Coatings}\gdef\@doiabbr{coatings}\gdef\@ISSN{2079-6412}}
\DeclareOption{colloids}{\gdef\@journal{colloids}\gdef\@journalfull{Colloids Interfaces}\gdef\@journalshort{Colloids Interfaces}\gdef\@doiabbr{colloids}\gdef\@ISSN{2504-5377}}
\DeclareOption{colorant}{\gdef\@journal{colorant}\gdef\@journalfull{Colorant}\gdef\@journalshort{Colorant}\gdef\@doiabbr{colorant}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{compounds}{\gdef\@journal{compounds}\gdef\@journalfull{Compounds}\gdef\@journalshort{Compounds}\gdef\@doiabbr{compounds}\gdef\@ISSN{2673-6918}\gdef\@continuouspages{yes}}
\DeclareOption{computation}{\gdef\@journal{computation}\gdef\@journalfull{Computation}\gdef\@journalshort{Computation}\gdef\@doiabbr{computation}\gdef\@ISSN{2079-3197}}
\DeclareOption{computers}{\gdef\@journal{computers}\gdef\@journalfull{Computers}\gdef\@journalshort{Computers}\gdef\@doiabbr{computers}\gdef\@ISSN{2073-431X}}
\DeclareOption{condensedmatter}{\gdef\@journal{condensedmatter}\gdef\@journalfull{Condensed Matter}\gdef\@journalshort{Condens. Matter}\gdef\@doiabbr{condmat}\gdef\@ISSN{2410-3896}}
\DeclareOption{conservation}{\gdef\@journal{conservation}\gdef\@journalfull{Conservation}\gdef\@journalshort{Conservation}\gdef\@doiabbr{conservation}\gdef\@ISSN{2673-7159}\gdef\@continuouspages{yes}}
\DeclareOption{constrmater}{\gdef\@journal{constrmater}\gdef\@journalfull{Construction Materials}\gdef\@journalshort{Constr. Mater.}\gdef\@doiabbr{constrmater}\gdef\@ISSN{2673-7108}\gdef\@continuouspages{yes}}
\DeclareOption{cosmetics}{\gdef\@journal{cosmetics}\gdef\@journalfull{Cosmetics}\gdef\@journalshort{Cosmetics}\gdef\@doiabbr{cosmetics}\gdef\@ISSN{2079-9284}}
\DeclareOption{covid}{\gdef\@journal{covid}\gdef\@journalfull{COVID}\gdef\@journalshort{COVID}\gdef\@doiabbr{covid}\gdef\@ISSN{2673-8112}\gdef\@continuouspages{yes}}
\DeclareOption{crops}{\gdef\@journal{crops}\gdef\@journalfull{Crops}\gdef\@journalshort{Crops}\gdef\@doiabbr{crops}\gdef\@ISSN{2673-7655}\gdef\@continuouspages{yes}}
\DeclareOption{cryptography}{\gdef\@journal{cryptography}\gdef\@journalfull{Cryptography}\gdef\@journalshort{Cryptography}\gdef\@doiabbr{cryptography}\gdef\@ISSN{2410-387X}}
\DeclareOption{crystals}{\gdef\@journal{crystals}\gdef\@journalfull{Crystals}\gdef\@journalshort{Crystals}\gdef\@doiabbr{cryst}\gdef\@ISSN{2073-4352}}
\DeclareOption{ctn}{\gdef\@journal{ctn}\gdef\@journalfull{Clinical and Translational Neuroscience}\gdef\@journalshort{CTN}\gdef\@doiabbr{ctn}\gdef\@ISSN{2514-183X}\gdef\@continuouspages{yes}}
\DeclareOption{curroncol}{\gdef\@journal{curroncol}\gdef\@journalfull{Current Oncology}\gdef\@journalshort{Curr. Oncol.}\gdef\@doiabbr{curroncol}\gdef\@ISSN{1718-7729}\gdef\@continuouspages{yes}}
\DeclareOption{cyber}{\gdef\@journal{cyber}\gdef\@journalfull{Cyber}\gdef\@journalshort{Cyber}\gdef\@doiabbr{cyber}\gdef\@ISSN{2624-800X}}
\DeclareOption{dairy}{\gdef\@journal{dairy}\gdef\@journalfull{Dairy}\gdef\@journalshort{Dairy}\gdef\@doiabbr{dairy}\gdef\@ISSN{2624-862X}\gdef\@continuouspages{yes}}
\DeclareOption{data}{\gdef\@journal{data}\gdef\@journalfull{Data}\gdef\@journalshort{Data}\gdef\@doiabbr{data}\gdef\@ISSN{2306-5729}}
\DeclareOption{dentistry}{\gdef\@journal{dentistry}\gdef\@journalfull{Dentistry Journal}\gdef\@journalshort{Dent. J.}\gdef\@doiabbr{dj}\gdef\@ISSN{2304-6767}}
\DeclareOption{dermato}{\gdef\@journal{dermato}\gdef\@journalfull{Dermato}\gdef\@journalshort{Dermato}\gdef\@doiabbr{dermato}\gdef\@ISSN{2673-6179}\gdef\@continuouspages{yes}}
\DeclareOption{dermatopathology}{\gdef\@journal{dermatopathology}\gdef\@journalfull{Dermatopathology}\gdef\@journalshort{Dermatopathology}\gdef\@doiabbr{dermatopathology}\gdef\@ISSN{2296-3529}\gdef\@continuouspages{yes}}
\DeclareOption{designs}{\gdef\@journal{designs}\gdef\@journalfull{Designs}\gdef\@journalshort{Designs}\gdef\@doiabbr{designs}\gdef\@ISSN{2411-9660}}
\DeclareOption{diabetology}{\gdef\@journal{diabetology}\gdef\@journalfull{Diabetology}\gdef\@journalshort{Diabetology}\gdef\@doiabbr{diabetology}\gdef\@ISSN{2673-4540}\gdef\@continuouspages{yes}}
\DeclareOption{diagnostics}{\gdef\@journal{diagnostics}\gdef\@journalfull{Diagnostics}\gdef\@journalshort{Diagnostics}\gdef\@doiabbr{diagnostics}\gdef\@ISSN{2075-4418}}
\DeclareOption{digital}{\gdef\@journal{digital}\gdef\@journalfull{Digital}\gdef\@journalshort{Digital}\gdef\@doiabbr{digital}\gdef\@ISSN{2673-6470}\gdef\@continuouspages{yes}}
\DeclareOption{disabilities}{\gdef\@journal{disabilities}\gdef\@journalfull{Disabilities}\gdef\@journalshort{Disabilities}\gdef\@doiabbr{disabilities}\gdef\@ISSN{2673-7272}\gdef\@continuouspages{yes}}
\DeclareOption{diseases}{\gdef\@journal{diseases}\gdef\@journalfull{Diseases}\gdef\@journalshort{Diseases}\gdef\@doiabbr{diseases}\gdef\@ISSN{2079-9721}}
\DeclareOption{diversity}{\gdef\@journal{diversity}\gdef\@journalfull{Diversity}\gdef\@journalshort{Diversity}\gdef\@doiabbr{d}\gdef\@ISSN{1424-2818}}
\DeclareOption{dna}{\gdef\@journal{dna}\gdef\@journalfull{DNA}\gdef\@journalshort{DNA}\gdef\@doiabbr{dna}\gdef\@ISSN{2673-8856}\gdef\@continuouspages{yes}}
\DeclareOption{drones}{\gdef\@journal{drones}\gdef\@journalfull{Drones}\gdef\@journalshort{Drones}\gdef\@doiabbr{drones}\gdef\@ISSN{2504-446X}}
\DeclareOption{dynamics}{\gdef\@journal{dynamics}\gdef\@journalfull{Dynamics}\gdef\@journalshort{Dynamics}\gdef\@doiabbr{dynamics}\gdef\@ISSN{2673-8716}\gdef\@continuouspages{yes}}
\DeclareOption{earth}{\gdef\@journal{earth}\gdef\@journalfull{Earth}\gdef\@journalshort{Earth}\gdef\@doiabbr{earth}\gdef\@ISSN{2673-4834}\gdef\@continuouspages{yes}}
\DeclareOption{ebj}{\gdef\@journal{ebj}\gdef\@journalfull{European Burn Journal}\gdef\@journalshort{Eur. Burn J.}\gdef\@doiabbr{ebj}\gdef\@ISSN{2673-1991}\gdef\@continuouspages{yes}}
\DeclareOption{ecologies}{\gdef\@journal{ecologies}\gdef\@journalfull{Ecologies}\gdef\@journalshort{Ecologies}\gdef\@doiabbr{ecologies}\gdef\@ISSN{2673-4133}\gdef\@continuouspages{yes}}
\DeclareOption{econometrics}{\gdef\@journal{econometrics}\gdef\@journalfull{Econometrics}\gdef\@journalshort{Econometrics}\gdef\@doiabbr{econometrics}\gdef\@ISSN{2225-1146}}
\DeclareOption{economies}{\gdef\@journal{economies}\gdef\@journalfull{Economies}\gdef\@journalshort{Economies}\gdef\@doiabbr{economies}\gdef\@ISSN{2227-7099}}
\DeclareOption{education}{\gdef\@journal{education}\gdef\@journalfull{Education Sciences}\gdef\@journalshort{Educ. Sci.}\gdef\@doiabbr{educsci}\gdef\@ISSN{2227-7102}}
\DeclareOption{ejihpe}{\gdef\@journal{ejihpe}\gdef\@journalfull{European Journal of Investigation in Health Psychology and Education}\gdef\@journalshort{Eur. J. Investig. Health Psychol. Educ.}\gdef\@doiabbr{ejihpe}\gdef\@ISSN{2254-9625}\gdef\@continuouspages{yes}}
\DeclareOption{electricity}{\gdef\@journal{electricity}\gdef\@journalfull{Electricity}\gdef\@journalshort{Electricity}\gdef\@doiabbr{electricity}\gdef\@ISSN{2673-4826}\gdef\@continuouspages{yes}}
\DeclareOption{electrochem}{\gdef\@journal{electrochem}\gdef\@journalfull{Electrochem}\gdef\@journalshort{Electrochem}\gdef\@doiabbr{electrochem}\gdef\@ISSN{2673-3293}\gdef\@continuouspages{yes}}
\DeclareOption{electronicmat}{\gdef\@journal{electronicmat}\gdef\@journalfull{Electronic Materials}\gdef\@journalshort{Electron. Mater.}\gdef\@doiabbr{electronicmat}\gdef\@ISSN{2673-3978}\gdef\@continuouspages{yes}}
\DeclareOption{electronics}{\gdef\@journal{electronics}\gdef\@journalfull{Electronics}\gdef\@journalshort{Electronics}\gdef\@doiabbr{electronics}\gdef\@ISSN{2079-9292}}
\DeclareOption{encyclopedia}{\gdef\@journal{encyclopedia}\gdef\@journalfull{Encyclopedia}\gdef\@journalshort{Encyclopedia}\gdef\@doiabbr{encyclopedia}\gdef\@ISSN{2673-8392}\gdef\@continuouspages{yes}}
\DeclareOption{endocrines}{\gdef\@journal{endocrines}\gdef\@journalfull{Endocrines}\gdef\@journalshort{Endocrines}\gdef\@doiabbr{endocrines}\gdef\@ISSN{2673-396X}\gdef\@continuouspages{yes}}
\DeclareOption{energies}{\gdef\@journal{energies}\gdef\@journalfull{Energies}\gdef\@journalshort{Energies}\gdef\@doiabbr{en}\gdef\@ISSN{1996-1073}}
\DeclareOption{eng}{\gdef\@journal{eng}\gdef\@journalfull{Eng-Advances in Engineering}\gdef\@journalshort{Eng}\gdef\@doiabbr{eng}\gdef\@ISSN{2673-4117}\gdef\@continuouspages{yes}}
\DeclareOption{engproc}{\gdef\@journal{engproc}\gdef\@journalfull{Engineering Proceedings}\gdef\@journalshort{Eng. Proc.}\gdef\@doiabbr{engproc}\gdef\@ISSN{2673-4591}}
%\DeclareOption{ent}{\gdef\@journal{ent}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{entropy}{\gdef\@journal{entropy}\gdef\@journalfull{Entropy}\gdef\@journalshort{Entropy}\gdef\@doiabbr{e}\gdef\@ISSN{1099-4300}}
\DeclareOption{environments}{\gdef\@journal{environments}\gdef\@journalfull{Environments}\gdef\@journalshort{Environments}\gdef\@doiabbr{environments}\gdef\@ISSN{2076-3298}}
\DeclareOption{environsciproc}{\gdef\@journal{environsciproc}\gdef\@journalfull{Environmental Sciences Proceedings}\gdef\@journalshort{Environ. Sci. Proc.}\gdef\@doiabbr{environsciproc}\gdef\@ISSN{2673-4931}}
\DeclareOption{epidemiologia}{\gdef\@journal{epidemiologia}\gdef\@journalfull{Epidemiologia}\gdef\@journalshort{Epidemiologia}\gdef\@doiabbr{epidemiologia}\gdef\@ISSN{2673-3986}\gdef\@continuouspages{yes}}
\DeclareOption{epigenomes}{\gdef\@journal{epigenomes}\gdef\@journalfull{Epigenomes}\gdef\@journalshort{Epigenomes}\gdef\@doiabbr{epigenomes}\gdef\@ISSN{2075-4655}}
\DeclareOption{est}{\gdef\@journal{est}\gdef\@journalfull{Electrochemical Science and Technology}\gdef\@journalshort{Electrochem. Sci. Technol.}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{fermentation}{\gdef\@journal{fermentation}\gdef\@journalfull{Fermentation}\gdef\@journalshort{Fermentation}\gdef\@doiabbr{fermentation}\gdef\@ISSN{2311-5637}}
\DeclareOption{fibers}{\gdef\@journal{fibers}\gdef\@journalfull{Fibers}\gdef\@journalshort{Fibers}\gdef\@doiabbr{fib}\gdef\@ISSN{2079-6439}}
\DeclareOption{	fintech}{\gdef\@journal{fintech}\gdef\@journalfull{FinTech}\gdef\@journalshort{FinTech}\gdef\@doiabbr{fintech}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{fire}{\gdef\@journal{fire}\gdef\@journalfull{Fire}\gdef\@journalshort{Fire}\gdef\@doiabbr{fire}\gdef\@ISSN{2571-6255}}
\DeclareOption{fishes}{\gdef\@journal{fishes}\gdef\@journalfull{Fishes}\gdef\@journalshort{Fishes}\gdef\@doiabbr{fishes}\gdef\@ISSN{2410-3888}}
\DeclareOption{fluids}{\gdef\@journal{fluids}\gdef\@journalfull{Fluids}\gdef\@journalshort{Fluids}\gdef\@doiabbr{fluids}\gdef\@ISSN{2311-5521}}
\DeclareOption{foods}{\gdef\@journal{foods}\gdef\@journalfull{Foods}\gdef\@journalshort{Foods}\gdef\@doiabbr{foods}\gdef\@ISSN{2304-8158}}
\DeclareOption{forecasting}{\gdef\@journal{forecasting}\gdef\@journalfull{Forecasting}\gdef\@journalshort{Forecasting}\gdef\@doiabbr{forecast}\gdef\@ISSN{2571-9394}\gdef\@continuouspages{yes}}
\DeclareOption{forensicsci}{\gdef\@journal{forensicsci}\gdef\@journalfull{Forensic Sciences}\gdef\@journalshort{Forensic. Sci.}\gdef\@doiabbr{forensicsci}\gdef\@ISSN{2673-6756}\gdef\@continuouspages{yes}}
\DeclareOption{forests}{\gdef\@journal{forests}\gdef\@journalfull{Forests}\gdef\@journalshort{Forests}\gdef\@doiabbr{f}\gdef\@ISSN{1999-4907}}
\DeclareOption{foundations}{\gdef\@journal{foundations}\gdef\@journalfull{Foundations}\gdef\@journalshort{Foundations}\gdef\@doiabbr{foundations}\gdef\@ISSN{2673-9321}\gdef\@continuouspages{yes}}
\DeclareOption{fractalfract}{\gdef\@journal{fractalfract}\gdef\@journalfull{Fractal and Fractional}\gdef\@journalshort{Fractal Fract.}\gdef\@doiabbr{fractalfract}\gdef\@ISSN{2504-3110}}
\DeclareOption{fuels}{\gdef\@journal{fuels}\gdef\@journalfull{Fuels}\gdef\@journalshort{Fuels}\gdef\@doiabbr{fuels}\gdef\@ISSN{2673-3994}\gdef\@continuouspages{yes}}
\DeclareOption{futureinternet}{\gdef\@journal{futureinternet}\gdef\@journalfull{Future Internet}\gdef\@journalshort{Future Internet}\gdef\@doiabbr{fi}\gdef\@ISSN{1999-5903}}
%\DeclareOption{futureparasites}{\gdef\@journal{futureparasites}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{futurepharmacol}{\gdef\@journal{futurepharmacol}\gdef\@journalfull{Future Pharmacology}\gdef\@journalshort{Future Pharm.}\gdef\@doiabbr{futurepharmacol}\gdef\@ISSN{2673-9879}\gdef\@continuouspages{yes}}
\DeclareOption{futurephys}{\gdef\@journal{futurephys}\gdef\@journalfull{Future Physics}\gdef\@journalshort{Future Phys.}\gdef\@doiabbr{futurephys}\gdef\@ISSN{2624-6503}}
\DeclareOption{futuretransp}{\gdef\@journal{futuretransp}\gdef\@journalfull{Future Transportation}\gdef\@journalshort{Future Transp.}\gdef\@doiabbr{futuretransp}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{galaxies}{\gdef\@journal{galaxies}\gdef\@journalfull{Galaxies}\gdef\@journalshort{Galaxies}\gdef\@doiabbr{galaxies}\gdef\@ISSN{2075-4434}}
\DeclareOption{games}{\gdef\@journal{games}\gdef\@journalfull{Games}\gdef\@journalshort{Games}\gdef\@doiabbr{g}\gdef\@ISSN{2073-4336}}
\DeclareOption{gases}{\gdef\@journal{gases}\gdef\@journalfull{Gases}\gdef\@journalshort{Gases}\gdef\@doiabbr{gases}\gdef\@ISSN{2673-5628}\gdef\@continuouspages{yes}}
\DeclareOption{gastroent}{\gdef\@journal{gastroent}\gdef\@journalfull{Gastroenterology Insights}\gdef\@journalshort{Gastroenterol. Insights}\gdef\@doiabbr{gastroent}\gdef\@ISSN{2036-7422}\gdef\@continuouspages{yes}}
\DeclareOption{gastrointestdisord}{\gdef\@journal{gastrointestdisord}\gdef\@journalfull{Gastrointestinal Disorders}\gdef\@journalshort{Gastrointest. Disord.}\gdef\@doiabbr{gidisord}\gdef\@ISSN{2624-5647}\gdef\@continuouspages{yes}}
\DeclareOption{gels}{\gdef\@journal{gels}\gdef\@journalfull{Gels}\gdef\@journalshort{Gels}\gdef\@doiabbr{gels}\gdef\@ISSN{2310-2861}}
\DeclareOption{genealogy}{\gdef\@journal{genealogy}\gdef\@journalfull{Genealogy}\gdef\@journalshort{Genealogy}\gdef\@doiabbr{genealogy}\gdef\@ISSN{2313-5778}}
\DeclareOption{genes}{\gdef\@journal{genes}\gdef\@journalfull{Genes}\gdef\@journalshort{Genes}\gdef\@doiabbr{genes}\gdef\@ISSN{2073-4425}}
\DeclareOption{geographies}{\gdef\@journal{geographies}\gdef\@journalfull{Geographies}\gdef\@journalshort{Geographies}\gdef\@doiabbr{geographies}\gdef\@ISSN{2673-7086}\gdef\@continuouspages{yes}}
\DeclareOption{geohazards}{\gdef\@journal{geohazards}\gdef\@journalfull{GeoHazards}\gdef\@journalshort{GeoHazards}\gdef\@doiabbr{geohazards}\gdef\@ISSN{2624-795X}\gdef\@continuouspages{yes}}
\DeclareOption{geomatics}{\gdef\@journal{geomatics}\gdef\@journalfull{Geomatics}\gdef\@journalshort{Geomatics}\gdef\@doiabbr{geomatics}\gdef\@ISSN{2673-7418}\gdef\@continuouspages{yes}}
\DeclareOption{geosciences}{\gdef\@journal{geosciences}\gdef\@journalfull{Geosciences}\gdef\@journalshort{Geosciences}\gdef\@doiabbr{geosciences}\gdef\@ISSN{2076-3263}}
\DeclareOption{geotechnics}{\gdef\@journal{geotechnics}\gdef\@journalfull{Geotechnics}\gdef\@journalshort{Geotechnics}\gdef\@doiabbr{geotechnics}\gdef\@ISSN{2673-7094}\gdef\@continuouspages{yes}}
\DeclareOption{geriatrics}{\gdef\@journal{geriatrics}\gdef\@journalfull{Geriatrics}\gdef\@journalshort{Geriatrics}\gdef\@doiabbr{geriatrics}\gdef\@ISSN{2308-3417}}
\DeclareOption{hazardousmatters}{\gdef\@journal{hazardousmatters}\gdef\@journalfull{Hazardous Matters}\gdef\@journalshort{Hazard. Matters}\gdef\@doiabbr{}\gdef\@ISSN{0014-0014}}
\DeclareOption{healthcare}{\gdef\@journal{healthcare}\gdef\@journalfull{Healthcare}\gdef\@journalshort{Healthcare}\gdef\@doiabbr{healthcare}\gdef\@ISSN{2227-9032}}
\DeclareOption{hearts}{\gdef\@journal{hearts}\gdef\@journalfull{Hearts}\gdef\@journalshort{Hearts}\gdef\@doiabbr{hearts}\gdef\@ISSN{2673-3846}\gdef\@continuouspages{yes}}
\DeclareOption{hemato}{\gdef\@journal{hemato}\gdef\@journalfull{Hemato}\gdef\@journalshort{Hemato}\gdef\@doiabbr{hemato}\gdef\@ISSN{2673-6357}\gdef\@continuouspages{yes}}
\DeclareOption{heritage}{\gdef\@journal{heritage}\gdef\@journalfull{Heritage}\gdef\@journalshort{Heritage}\gdef\@doiabbr{heritage}\gdef\@ISSN{2571-9408}\gdef\@continuouspages{yes}}
\DeclareOption{highthroughput}{\gdef\@journal{highthroughput}\gdef\@journalfull{High-Throughput}\gdef\@journalshort{High-Throughput}\gdef\@doiabbr{ht}\gdef\@ISSN{2571-5135}}
\DeclareOption{histories}{\gdef\@journal{histories}\gdef\@journalfull{Histories}\gdef\@journalshort{Histories}\gdef\@doiabbr{histories}\gdef\@ISSN{2409-9252}\gdef\@continuouspages{yes}}
\DeclareOption{horticulturae}{\gdef\@journal{horticulturae}\gdef\@journalfull{Horticulturae}\gdef\@journalshort{Horticulturae}\gdef\@doiabbr{horticulturae}\gdef\@ISSN{2311-7524}}
\DeclareOption{humanities}{\gdef\@journal{humanities}\gdef\@journalfull{Humanities}\gdef\@journalshort{Humanities}\gdef\@doiabbr{h}\gdef\@ISSN{2076-0787}}
\DeclareOption{humans}{\gdef\@journal{humans}\gdef\@journalfull{Humans}\gdef\@journalshort{Humans}\gdef\@doiabbr{humans}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{hydrobiology}{\gdef\@journal{hydrobiology}\gdef\@journalfull{Hydrobiology}\gdef\@journalshort{Hydrobiology	}\gdef\@doiabbr{hydrobiology}\gdef\@ISSN{2673-9917}\gdef\@continuouspages{yes}}
\DeclareOption{hydrogen}{\gdef\@journal{hydrogen}\gdef\@journalfull{Hydrogen}\gdef\@journalshort{Hydrogen}\gdef\@doiabbr{hydrogen}\gdef\@ISSN{2673-4141}\gdef\@continuouspages{yes}}
\DeclareOption{hydrology}{\gdef\@journal{hydrology}\gdef\@journalfull{Hydrology}\gdef\@journalshort{Hydrology}\gdef\@doiabbr{hydrology}\gdef\@ISSN{2306-5338}}
\DeclareOption{hygiene}{\gdef\@journal{hygiene}\gdef\@journalfull{Hygiene}\gdef\@journalshort{Hygiene}\gdef\@doiabbr{hygiene}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{idr}{\gdef\@journal{idr}\gdef\@journalfull{Infectious Disease Reports}\gdef\@journalshort{Infect. Dis. Rep.}\gdef\@doiabbr{idr}\gdef\@ISSN{2036-7449}\gdef\@continuouspages{yes}}
\DeclareOption{ijerph}{\gdef\@journal{ijerph}\gdef\@journalfull{International Journal of Environmental Research and Public Health}\gdef\@journalshort{Int. J. Environ. Res. Public Health}\gdef\@doiabbr{ijerph}\gdef\@ISSN{1660-4601}}
\DeclareOption{ijfs}{\gdef\@journal{ijfs}\gdef\@journalfull{International Journal of Financial Studies}\gdef\@journalshort{Int. J. Financial Stud.}\gdef\@doiabbr{ijfs}\gdef\@ISSN{2227-7072}}
\DeclareOption{ijgi}{\gdef\@journal{ijgi}\gdef\@journalfull{ISPRS International Journal of Geo-Information}\gdef\@journalshort{ISPRS Int. J. Geo-Inf.}\gdef\@doiabbr{ijgi}\gdef\@ISSN{2220-9964}}
\DeclareOption{ijms}{\gdef\@journal{ijms}\gdef\@journalfull{International Journal of Molecular Sciences}\gdef\@journalshort{Int. J. Mol. Sci.}\gdef\@doiabbr{ijms}\gdef\@ISSN{1422-0067}}
\DeclareOption{ijns}{\gdef\@journal{ijns}\gdef\@journalfull{International Journal of Neonatal Screening}\gdef\@journalshort{Int. J. Neonatal Screen.}\gdef\@doiabbr{ijns}\gdef\@ISSN{2409-515X}}
\DeclareOption{ijtm}{\gdef\@journal{ijtm}\gdef\@journalfull{International Journal of Translational Medicine}\gdef\@journalshort{Int. J. Transl. Med.}\gdef\@doiabbr{ijtm}\gdef\@ISSN{2673-8937}\gdef\@continuouspages{yes}}
\DeclareOption{ijtpp}{\gdef\@journal{ijtpp}\gdef\@journalfull{International Journal of Turbomachinery Propulsion and Power}\gdef\@journalshort{Int. J. Turbomach. Propuls. Power}\gdef\@doiabbr{ijtpp}\gdef\@ISSN{2504-186X}}
\DeclareOption{immuno}{\gdef\@journal{immuno}\gdef\@journalfull{Immuno}\gdef\@journalshort{Immuno}\gdef\@doiabbr{immuno}\gdef\@ISSN{2673-5601}\gdef\@continuouspages{yes}}
\DeclareOption{informatics}{\gdef\@journal{informatics}\gdef\@journalfull{Informatics}\gdef\@journalshort{Informatics}\gdef\@doiabbr{informatics}\gdef\@ISSN{2227-9709}}
\DeclareOption{information}{\gdef\@journal{information}\gdef\@journalfull{Information}\gdef\@journalshort{Information}\gdef\@doiabbr{info}\gdef\@ISSN{2078-2489}}
\DeclareOption{infrastructures}{\gdef\@journal{infrastructures}\gdef\@journalfull{Infrastructures}\gdef\@journalshort{Infrastructures}\gdef\@doiabbr{infrastructures}\gdef\@ISSN{2412-3811}}
\DeclareOption{inorganics}{\gdef\@journal{inorganics}\gdef\@journalfull{Inorganics}\gdef\@journalshort{Inorganics}\gdef\@doiabbr{inorganics}\gdef\@ISSN{2304-6740}}
\DeclareOption{insects}{\gdef\@journal{insects}\gdef\@journalfull{Insects}\gdef\@journalshort{Insects}\gdef\@doiabbr{insects}\gdef\@ISSN{2075-4450}}
\DeclareOption{instruments}{\gdef\@journal{instruments}\gdef\@journalfull{Instruments}\gdef\@journalshort{Instruments}\gdef\@doiabbr{instruments}\gdef\@ISSN{2410-390X}}
\DeclareOption{inventions}{\gdef\@journal{inventions}\gdef\@journalfull{Inventions}\gdef\@journalshort{Inventions}\gdef\@doiabbr{inventions}\gdef\@ISSN{2411-5134}}
\DeclareOption{iot}{\gdef\@journal{iot}\gdef\@journalfull{IoT}\gdef\@journalshort{IoT}\gdef\@doiabbr{iot}\gdef\@ISSN{2624-831X}\gdef\@continuouspages{yes}}
\DeclareOption{j}{\gdef\@journal{j}\gdef\@journalfull{J}\gdef\@journalshort{J}\gdef\@doiabbr{j}\gdef\@ISSN{2571-8800}\gdef\@continuouspages{yes}}
\DeclareOption{jal}{\gdef\@journal{jal}\gdef\@journalfull{Journal of Ageing and Longevity}\gdef\@journalshort{J. Ageing Longev.}\gdef\@doiabbr{jal}\gdef\@ISSN{2673-9259}\gdef\@continuouspages{yes}}
\DeclareOption{jcdd}{\gdef\@journal{jcdd}\gdef\@journalfull{Journal of Cardiovascular Development and Disease}\gdef\@journalshort{J. Cardiovasc. Dev. Dis.}\gdef\@doiabbr{jcdd}\gdef\@ISSN{2308-3425}}
\DeclareOption{jcm}{\gdef\@journal{jcm}\gdef\@journalfull{Journal of Clinical Medicine}\gdef\@journalshort{J. Clin. Med.}\gdef\@doiabbr{jcm}\gdef\@ISSN{2077-0383}}
\DeclareOption{jcp}{\gdef\@journal{jcp}\gdef\@journalfull{Journal of Cybersecurity and Privacy}\gdef\@journalshort{J. Cybersecur. Priv.}\gdef\@doiabbr{jcp}\gdef\@ISSN{2624-800X}\gdef\@continuouspages{yes}}
\DeclareOption{jcs}{\gdef\@journal{jcs}\gdef\@journalfull{Journal of Composites Science}\gdef\@journalshort{J. Compos. Sci.}\gdef\@doiabbr{jcs}\gdef\@ISSN{2504-477X}}
\DeclareOption{jdb}{\gdef\@journal{jdb}\gdef\@journalfull{Journal of Developmental Biology}\gdef\@journalshort{J. Dev. Biol.}\gdef\@doiabbr{jdb}\gdef\@ISSN{2221-3759}}
\DeclareOption{jfb}{\gdef\@journal{jfb}\gdef\@journalfull{Journal of Functional Biomaterials}\gdef\@journalshort{J. Funct. Biomater.}\gdef\@doiabbr{jfb}\gdef\@ISSN{2079-4983}}
\DeclareOption{jfmk}{\gdef\@journal{jfmk}\gdef\@journalfull{Journal of Functional Morphology and Kinesiology}\gdef\@journalshort{J. Funct. Morphol. Kinesiol.}\gdef\@doiabbr{jfmk}\gdef\@ISSN{2411-5142}}
\DeclareOption{jimaging}{\gdef\@journal{jimaging}\gdef\@journalfull{Journal of Imaging}\gdef\@journalshort{J. Imaging}\gdef\@doiabbr{jimaging}\gdef\@ISSN{2313-433X}}
\DeclareOption{jintelligence}{\gdef\@journal{jintelligence}\gdef\@journalfull{Journal of Intelligence}\gdef\@journalshort{J. Intell.}\gdef\@doiabbr{jintelligence}\gdef\@ISSN{2079-3200}}
\DeclareOption{jlpea}{\gdef\@journal{jlpea}\gdef\@journalfull{Journal of Low Power Electronics and Applications}\gdef\@journalshort{J. Low Power Electron. Appl.}\gdef\@doiabbr{jlpea}\gdef\@ISSN{2079-9268}}
\DeclareOption{jmmp}{\gdef\@journal{jmmp}\gdef\@journalfull{Journal of Manufacturing and Materials Processing}\gdef\@journalshort{J. Manuf. Mater. Process.}\gdef\@doiabbr{jmmp}\gdef\@ISSN{2504-4494}}
\DeclareOption{jmp}{\gdef\@journal{jmp}\gdef\@journalfull{Journal of Molecular Pathology}\gdef\@journalshort{J. Mol. Pathol.}\gdef\@doiabbr{jmp}\gdef\@ISSN{2673-5261}\gdef\@continuouspages{yes}}
\DeclareOption{jmse}{\gdef\@journal{jmse}\gdef\@journalfull{Journal of Marine Science and Engineering}\gdef\@journalshort{J. Mar. Sci. Eng.}\gdef\@doiabbr{jmse}\gdef\@ISSN{2077-1312}}
\DeclareOption{jne}{\gdef\@journal{jne}\gdef\@journalfull{Journal of Nuclear Engineering}\gdef\@journalshort{J. Nucl. Eng.}\gdef\@doiabbr{jne}\gdef\@ISSN{2673-4362}\gdef\@continuouspages{yes}}
\DeclareOption{jnt}{\gdef\@journal{jnt}\gdef\@journalfull{Journal of Nanotheranostics}\gdef\@journalshort{J. Nanotheranostics}\gdef\@doiabbr{jnt}\gdef\@ISSN{2624-845X}\gdef\@continuouspages{yes}}
\DeclareOption{jof}{\gdef\@journal{jof}\gdef\@journalfull{Journal of Fungi}\gdef\@journalshort{J. Fungi}\gdef\@doiabbr{jof}\gdef\@ISSN{2309-608X}}
\DeclareOption{joitmc}{\gdef\@journal{joitmc}\gdef\@journalfull{Journal of Open Innovation: Technology Market and Complexity}\gdef\@journalshort{J. Open Innov. Technol. Mark. Complex.}\gdef\@doiabbr{joitmc}\gdef\@ISSN{2199-8531}}
\DeclareOption{jor}{\gdef\@journal{jor}\gdef\@journalfull{Journal of Respiration}\gdef\@journalshort{J. Respir.}\gdef\@doiabbr{jor}\gdef\@ISSN{2673-527X}\gdef\@continuouspages{yes}}
\DeclareOption{journalmedia}{\gdef\@journal{journalmedia}\gdef\@journalfull{Journalism and Media}\gdef\@journalshort{Journal. Media}\gdef\@doiabbr{journalmedia}\gdef\@ISSN{2673-5172}\gdef\@continuouspages{yes}}
\DeclareOption{jox}{\gdef\@journal{jox}\gdef\@journalfull{Journal of Xenobiotics}\gdef\@journalshort{J. Xenobiot.}\gdef\@doiabbr{jox}\gdef\@ISSN{2039-4713}\gdef\@continuouspages{yes}}
\DeclareOption{jpm}{\gdef\@journal{jpm}\gdef\@journalfull{Journal of Personalized Medicine}\gdef\@journalshort{J. Pers. Med.}\gdef\@doiabbr{jpm}\gdef\@ISSN{2075-4426}}
\DeclareOption{jrfm}{\gdef\@journal{jrfm}\gdef\@journalfull{Journal of Risk and Financial Management}\gdef\@journalshort{J. Risk Financial Manag.}\gdef\@doiabbr{jrfm}\gdef\@ISSN{1911-8074}}
\DeclareOption{jsan}{\gdef\@journal{jsan}\gdef\@journalfull{Journal of Sensor and Actuator Networks}\gdef\@journalshort{J. Sens. Actuator Netw.}\gdef\@doiabbr{jsan}\gdef\@ISSN{2224-2708}}
\DeclareOption{jtaer}{\gdef\@journal{jtaer}\gdef\@journalfull{Journal of Theoretical and Applied Electronic Commerce Research}\gdef\@journalshort{J. Theor. Appl. Electron. Commer. Res.}\gdef\@doiabbr{jtaer}\gdef\@ISSN{0718-1876}\gdef\@continuouspages{yes}}
\DeclareOption{jzbg}{\gdef\@journal{jzbg}\gdef\@journalfull{Journal of Zoological and Botanical Gardens}\gdef\@journalshort{J. Zool. Bot. Gard.}\gdef\@doiabbr{jzbg}\gdef\@ISSN{2673-5636}\gdef\@continuouspages{yes}}
%\DeclareOption{kidney}{\gdef\@journal{kidney}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{kidneydial}{\gdef\@journal{kidneydial}\gdef\@journalfull{Kidney and Dialysis}\gdef\@journalshort{Kidney Dial.}\gdef\@doiabbr{kidneydial}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{knowledge}{\gdef\@journal{knowledge}\gdef\@journalfull{Knowledge}\gdef\@journalshort{Knowledge}\gdef\@doiabbr{knowledge}\gdef\@ISSN{2673-9585}\gdef\@continuouspages{yes}}
\DeclareOption{land}{\gdef\@journal{land}\gdef\@journalfull{Land}\gdef\@journalshort{Land}\gdef\@doiabbr{land}\gdef\@ISSN{2073-445X}}
\DeclareOption{languages}{\gdef\@journal{languages}\gdef\@journalfull{Languages}\gdef\@journalshort{Languages}\gdef\@doiabbr{languages}\gdef\@ISSN{2226-471X}}
\DeclareOption{laws}{\gdef\@journal{laws}\gdef\@journalfull{Laws}\gdef\@journalshort{Laws}\gdef\@doiabbr{laws}\gdef\@ISSN{2075-471X}}
\DeclareOption{life}{\gdef\@journal{life}\gdef\@journalfull{Life}\gdef\@journalshort{Life}\gdef\@doiabbr{life}\gdef\@ISSN{2075-1729}}
\DeclareOption{liquids}{\gdef\@journal{liquids}\gdef\@journalfull{Liquids}\gdef\@journalshort{Liquids}\gdef\@doiabbr{liquids}\gdef\@ISSN{2673-8015}\gdef\@continuouspages{yes}}
\DeclareOption{literature}{\gdef\@journal{literature}\gdef\@journalfull{Literature}\gdef\@journalshort{Literature}\gdef\@doiabbr{literature}\gdef\@ISSN{2410-9789}\gdef\@continuouspages{yes}}
\DeclareOption{livers}{\gdef\@journal{livers}\gdef\@journalfull{Livers}\gdef\@journalshort{Livers}\gdef\@doiabbr{livers}\gdef\@ISSN{2673-4389}\gdef\@continuouspages{yes}}
\DeclareOption{logistics}{\gdef\@journal{logistics}\gdef\@journalfull{Logistics}\gdef\@journalshort{Logistics}\gdef\@doiabbr{logistics}\gdef\@ISSN{2305-6290}}
\DeclareOption{lubricants}{\gdef\@journal{lubricants}\gdef\@journalfull{Lubricants}\gdef\@journalshort{Lubricants}\gdef\@doiabbr{lubricants}\gdef\@ISSN{2075-4442}}
\DeclareOption{machines}{\gdef\@journal{machines}\gdef\@journalfull{Machines}\gdef\@journalshort{Machines}\gdef\@doiabbr{machines}\gdef\@ISSN{2075-1702}}
\DeclareOption{macromol}{\gdef\@journal{macromol}\gdef\@journalfull{Macromol}\gdef\@journalshort{Macromol}\gdef\@doiabbr{macromol}\gdef\@ISSN{2673-6209}\gdef\@continuouspages{yes}}
\DeclareOption{magnetism}{\gdef\@journal{magnetism}\gdef\@journalfull{Magnetism}\gdef\@journalshort{Magnetism}\gdef\@doiabbr{magnetism}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{magnetochemistry}{\gdef\@journal{magnetochemistry}\gdef\@journalfull{Magnetochemistry}\gdef\@journalshort{Magnetochemistry}\gdef\@doiabbr{magnetochemistry}\gdef\@ISSN{2312-7481}}
\DeclareOption{make}{\gdef\@journal{make}\gdef\@journalfull{Machine Learning and Knowledge Extraction}\gdef\@journalshort{Mach. Learn. Knowl. Extr.}\gdef\@doiabbr{make}\gdef\@ISSN{2504-4990}\gdef\@continuouspages{yes}}
\DeclareOption{marinedrugs}{\gdef\@journal{marinedrugs}\gdef\@journalfull{Marine Drugs}\gdef\@journalshort{Mar. Drugs}\gdef\@doiabbr{md}\gdef\@ISSN{1660-3397}}
\DeclareOption{materials}{\gdef\@journal{materials}\gdef\@journalfull{Materials}\gdef\@journalshort{Materials}\gdef\@doiabbr{ma}\gdef\@ISSN{1996-1944}}
\DeclareOption{materproc}{\gdef\@journal{materproc}\gdef\@journalfull{Materials Proceedings}\gdef\@journalshort{Mater. Proc.}\gdef\@doiabbr{materproc}\gdef\@ISSN{2673-4605}}
\DeclareOption{mathematics}{\gdef\@journal{mathematics}\gdef\@journalfull{Mathematics}\gdef\@journalshort{Mathematics}\gdef\@doiabbr{math}\gdef\@ISSN{2227-7390}}
\DeclareOption{mca}{\gdef\@journal{mca}\gdef\@journalfull{Mathematical and Computational Applications}\gdef\@journalshort{Math. Comput. Appl.}\gdef\@doiabbr{mca}\gdef\@ISSN{2297-8747}}
%\DeclareOption{measurements}{\gdef\@journal{measurements}\gdef\@journalfull{Measurements}\gdef\@journalshort{Measurements}\gdef\@doiabbr{measurements}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{medicina}{\gdef\@journal{medicina}\gdef\@journalfull{Medicina}\gdef\@journalshort{Medicina}\gdef\@doiabbr{medicina}\gdef\@ISSN{1648-9144}}
\DeclareOption{medicines}{\gdef\@journal{medicines}\gdef\@journalfull{Medicines}\gdef\@journalshort{Medicines}\gdef\@doiabbr{medicines}\gdef\@ISSN{2305-6320}}
\DeclareOption{medsci}{\gdef\@journal{medsci}\gdef\@journalfull{Medical Sciences}\gdef\@journalshort{Med. Sci.}\gdef\@doiabbr{medsci}\gdef\@ISSN{2076-3271}}
\DeclareOption{membranes}{\gdef\@journal{membranes}\gdef\@journalfull{Membranes}\gdef\@journalshort{Membranes}\gdef\@doiabbr{membranes}\gdef\@ISSN{2077-0375}}
\DeclareOption{merits}{\gdef\@journal{merits}\gdef\@journalfull{Merits}\gdef\@journalshort{Merits}\gdef\@doiabbr{merits}\gdef\@ISSN{2673-8104}\gdef\@continuouspages{yes}}
\DeclareOption{metabolites}{\gdef\@journal{metabolites}\gdef\@journalfull{Metabolites}\gdef\@journalshort{Metabolites}\gdef\@doiabbr{metabo}\gdef\@ISSN{2218-1989}}
\DeclareOption{metals}{\gdef\@journal{metals}\gdef\@journalfull{Metals}\gdef\@journalshort{Metals}\gdef\@doiabbr{met}\gdef\@ISSN{2075-4701}}
\DeclareOption{metrology}{\gdef\@journal{metrology}\gdef\@journalfull{Metrology}\gdef\@journalshort{Metrology}\gdef\@doiabbr{metrology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{micro}{\gdef\@journal{micro}\gdef\@journalfull{Micro}\gdef\@journalshort{Micro}\gdef\@doiabbr{micro}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{microarrays}{\gdef\@journal{microarrays}\gdef\@journalfull{Microarrays}\gdef\@journalshort{Microarrays}\gdef\@doiabbr{microarrays}\gdef\@ISSN{2076-3905}}
\DeclareOption{microbiolres}{\gdef\@journal{microbiolres}\gdef\@journalfull{Microbiology Research}\gdef\@journalshort{Microbiol. Res.}\gdef\@doiabbr{microbiolres}\gdef\@ISSN{2036-7481}\gdef\@continuouspages{yes}}
\DeclareOption{micromachines}{\gdef\@journal{micromachines}\gdef\@journalfull{Micromachines}\gdef\@journalshort{Micromachines}\gdef\@doiabbr{mi}\gdef\@ISSN{2072-666X}}
\DeclareOption{microorganisms}{\gdef\@journal{microorganisms}\gdef\@journalfull{Microorganisms}\gdef\@journalshort{Microorganisms}\gdef\@doiabbr{microorganisms}\gdef\@ISSN{2076-2607}}
\DeclareOption{microplastics}{\gdef\@journal{microplastics}\gdef\@journalfull{Microplastics}\gdef\@journalshort{Microplastics}\gdef\@doiabbr{microplastics}\gdef\@ISSN{2673-8929}\gdef\@continuouspages{yes}}
\DeclareOption{minerals}{\gdef\@journal{minerals}\gdef\@journalfull{Minerals}\gdef\@journalshort{Minerals}\gdef\@doiabbr{min}\gdef\@ISSN{2075-163X}}
\DeclareOption{mining}{\gdef\@journal{mining}\gdef\@journalfull{Mining}\gdef\@journalshort{Mining}\gdef\@doiabbr{mining}\gdef\@ISSN{2673-6489}\gdef\@continuouspages{yes}}
\DeclareOption{modelling}{\gdef\@journal{modelling}\gdef\@journalfull{Modelling}\gdef\@journalshort{Modelling}\gdef\@doiabbr{modelling}\gdef\@ISSN{2673-3951}\gdef\@continuouspages{yes}}
\DeclareOption{molbank}{\gdef\@journal{molbank}\gdef\@journalfull{Molbank}\gdef\@journalshort{Molbank}\gdef\@doiabbr{M}\gdef\@ISSN{1422-8599}}
\DeclareOption{molecules}{\gdef\@journal{molecules}\gdef\@journalfull{Molecules}\gdef\@journalshort{Molecules}\gdef\@doiabbr{molecules}\gdef\@ISSN{1420-3049}}
\DeclareOption{mps}{\gdef\@journal{mps}\gdef\@journalfull{Methods and Protocols}\gdef\@journalshort{Methods Protoc.}\gdef\@doiabbr{mps}\gdef\@ISSN{2409-9279}}
\DeclareOption{msf}{\gdef\@journal{	msf}\gdef\@journalfull{Medical Sciences Forum}\gdef\@journalshort{Med. Sci. Forum}\gdef\@doiabbr{msf}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{mti}{\gdef\@journal{mti}\gdef\@journalfull{Multimodal Technologies and Interaction}\gdef\@journalshort{Multimodal Technol. Interact.}\gdef\@doiabbr{mti}\gdef\@ISSN{2414-4088}}
\DeclareOption{nanoenergyadv}{\gdef\@journal{nanoenergyadv}\gdef\@journalfull{Nanoenergy Advances}\gdef\@journalshort{Nanoenergy Adv.}\gdef\@doiabbr{nanoenergyadv}\gdef\@ISSN{2673-706X}\gdef\@continuouspages{yes}}
\DeclareOption{nanomanufacturing}{\gdef\@journal{nanomanufacturing}\gdef\@journalfull{Nanomanufacturing}\gdef\@journalshort{Nanomanufacturing}\gdef\@doiabbr{nanomanufacturing}\gdef\@ISSN{2673-687X}\gdef\@continuouspages{yes}}
\DeclareOption{nanomaterials}{\gdef\@journal{nanomaterials}\gdef\@journalfull{Nanomaterials}\gdef\@journalshort{Nanomaterials}\gdef\@doiabbr{nano}\gdef\@ISSN{2079-4991}}
\DeclareOption{ncrna}{\gdef\@journal{ncrna}\gdef\@journalfull{Non-coding RNA}\gdef\@journalshort{Non-coding RNA}\gdef\@doiabbr{ncrna}\gdef\@ISSN{2311-553X}}
\DeclareOption{network}{\gdef\@journal{network}\gdef\@journalfull{Network}\gdef\@journalshort{Network}\gdef\@doiabbr{network}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{neuroglia}{\gdef\@journal{neuroglia}\gdef\@journalfull{Neuroglia}\gdef\@journalshort{Neuroglia}\gdef\@doiabbr{neuroglia}\gdef\@ISSN{2571-6980}\gdef\@continuouspages{yes}}
\DeclareOption{neurolint}{\gdef\@journal{neurolint}\gdef\@journalfull{Neurology International}\gdef\@journalshort{Neurol. Int.}\gdef\@doiabbr{neurolint}\gdef\@ISSN{2035-8377}\gdef\@continuouspages{yes}}
\DeclareOption{neurosci}{\gdef\@journal{neurosci}\gdef\@journalfull{NeuroSci}\gdef\@journalshort{NeuroSci}\gdef\@doiabbr{neurosci}\gdef\@ISSN{2673-4087}\gdef\@continuouspages{yes}}
\DeclareOption{nitrogen}{\gdef\@journal{nitrogen}\gdef\@journalfull{Nitrogen}\gdef\@journalshort{Nitrogen}\gdef\@doiabbr{nitrogen}\gdef\@ISSN{2504-3129}\gdef\@continuouspages{yes}}
\DeclareOption{notspecified}{\gdef\@journal{notspecified}\gdef\@journalfull{Journal Not Specified}\gdef\@journalshort{Journal Not Specified}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{nri}{\gdef\@journal{nri}\gdef\@journalfull{Natural Resources Informatics}\gdef\@journalshort{Nat. Resour. Inform.}\gdef\@doiabbr{nri}\gdef\@ISSN{2673-981X}\gdef\@continuouspages{yes}}
\DeclareOption{nursrep}{\gdef\@journal{nursrep}\gdef\@journalfull{Nursing Reports}\gdef\@journalshort{Nurs. Rep.}\gdef\@doiabbr{nursrep}\gdef\@ISSN{2039-4403}\gdef\@continuouspages{yes}}
\DeclareOption{nutraceuticals}{\gdef\@journal{nutraceuticals}\gdef\@journalfull{Nutraceuticals}\gdef\@journalshort{Nutraceuticals}\gdef\@doiabbr{nutraceuticals}\gdef\@ISSN{1661-3821}\gdef\@continuouspages{yes}}
\DeclareOption{nutrients}{\gdef\@journal{nutrients}\gdef\@journalfull{Nutrients}\gdef\@journalshort{Nutrients}\gdef\@doiabbr{nu}\gdef\@ISSN{2072-6643}}
\DeclareOption{obesities}{\gdef\@journal{obesities}\gdef\@journalfull{Obesities}\gdef\@journalshort{Obesities}\gdef\@doiabbr{obesities}\gdef\@ISSN{2673-4168}\gdef\@continuouspages{yes}}
\DeclareOption{oceans}{\gdef\@journal{oceans}\gdef\@journalfull{Oceans}\gdef\@journalshort{Oceans}\gdef\@doiabbr{oceans}\gdef\@ISSN{2673-1924}\gdef\@continuouspages{yes}}
\DeclareOption{ohbm}{\gdef\@journal{ohbm}\gdef\@journalfull{Journal of Otorhinolaryngology Hearing and Balance Medicine}\gdef\@journalshort{J. Otorhinolaryngol. Hear. Balance Med.}\gdef\@doiabbr{ohbm}\gdef\@ISSN{2504-463X}}
\DeclareOption{onco}{\gdef\@journal{onco}\gdef\@journalfull{Onco}\gdef\@journalshort{Onco}\gdef\@doiabbr{onco}\gdef\@ISSN{2673-7523}\gdef\@continuouspages{yes}}
%\DeclareOption{oncopathology}{\gdef\@journal{oncopathology}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{optics}{\gdef\@journal{optics}\gdef\@journalfull{Optics}\gdef\@journalshort{Optics}\gdef\@doiabbr{opt}\gdef\@ISSN{2673-3269}\gdef\@continuouspages{yes}}
\DeclareOption{oral}{\gdef\@journal{oral}\gdef\@journalfull{Oral}\gdef\@journalshort{Oral}\gdef\@doiabbr{oral}\gdef\@ISSN{2673-6373}\gdef\@continuouspages{yes}}
\DeclareOption{organics}{\gdef\@journal{organics}\gdef\@journalfull{Organics}\gdef\@journalshort{Organics}\gdef\@doiabbr{org}\gdef\@ISSN{2673-401X}\gdef\@continuouspages{yes}}
\DeclareOption{osteology}{\gdef\@journal{osteology}\gdef\@journalfull{Osteology}\gdef\@journalshort{Osteology}\gdef\@doiabbr{osteology}\gdef\@ISSN{2673-4036}\gdef\@continuouspages{yes}}
\DeclareOption{oxygen}{\gdef\@journal{oxygen}\gdef\@journalfull{Oxygen}\gdef\@journalshort{Oxygen}\gdef\@doiabbr{oxygen}\gdef\@ISSN{2673-9801}\gdef\@continuouspages{yes}}
\DeclareOption{parasites}{\gdef\@journal{parasites}\gdef\@journalfull{Parasites}\gdef\@journalshort{Parasites}\gdef\@doiabbr{parasites}\gdef\@ISSN{2673-5946}}
\DeclareOption{parasitologia}{\gdef\@journal{parasitologia}\gdef\@journalfull{Parasitologia}\gdef\@journalshort{Parasitologia}\gdef\@doiabbr{parasitologia}\gdef\@ISSN{2673-6772}\gdef\@continuouspages{yes}}
\DeclareOption{particles}{\gdef\@journal{particles}\gdef\@journalfull{Particles}\gdef\@journalshort{Particles}\gdef\@doiabbr{particles}\gdef\@ISSN{2571-712X}\gdef\@continuouspages{yes}}
\DeclareOption{pathogens}{\gdef\@journal{pathogens}\gdef\@journalfull{Pathogens}\gdef\@journalshort{Pathogens}\gdef\@doiabbr{pathogens}\gdef\@ISSN{2076-0817}}
\DeclareOption{pathophysiology}{\gdef\@journal{pathophysiology}\gdef\@journalfull{Pathophysiology}\gdef\@journalshort{Pathophysiology}\gdef\@doiabbr{pathophysiology}\gdef\@ISSN{0928-4680}\gdef\@continuouspages{yes}}
\DeclareOption{pediatrrep}{\gdef\@journal{pediatrrep}\gdef\@journalfull{Pediatric Reports}\gdef\@journalshort{Pediatr. Rep.}\gdef\@doiabbr{pediatric}\gdef\@ISSN{2036-7503}\gdef\@continuouspages{yes}}
\DeclareOption{pharmaceuticals}{\gdef\@journal{pharmaceuticals}\gdef\@journalfull{Pharmaceuticals}\gdef\@journalshort{Pharmaceuticals}\gdef\@doiabbr{ph}\gdef\@ISSN{1424-8247}}
\DeclareOption{pharmaceutics}{\gdef\@journal{pharmaceutics}\gdef\@journalfull{Pharmaceutics}\gdef\@journalshort{Pharmaceutics}\gdef\@doiabbr{pharmaceutics}\gdef\@ISSN{1999-4923}}
\DeclareOption{pharmacy}{\gdef\@journal{pharmacy}\gdef\@journalfull{Pharmacy}\gdef\@journalshort{Pharmacy}\gdef\@doiabbr{pharmacy}\gdef\@ISSN{2226-4787}}
\DeclareOption{philosophies}{\gdef\@journal{philosophies}\gdef\@journalfull{Philosophies}\gdef\@journalshort{Philosophies}\gdef\@doiabbr{philosophies}\gdef\@ISSN{2409-9287}}
\DeclareOption{photochem}{\gdef\@journal{photochem}\gdef\@journalfull{Photochem}\gdef\@journalshort{Photochem}\gdef\@doiabbr{photochem}\gdef\@ISSN{2673-7256}\gdef\@continuouspages{yes}}
\DeclareOption{photonics}{\gdef\@journal{photonics}\gdef\@journalfull{Photonics}\gdef\@journalshort{Photonics}\gdef\@doiabbr{photonics}\gdef\@ISSN{2304-6732}}
\DeclareOption{phycology}{\gdef\@journal{phycology}\gdef\@journalfull{Phycology}\gdef\@journalshort{Phycology}\gdef\@doiabbr{phycology}\gdef\@ISSN{2673-9410}\gdef\@continuouspages{yes}}
\DeclareOption{physchem}{\gdef\@journal{physchem}\gdef\@journalfull{Physchem}\gdef\@journalshort{Physchem}\gdef\@doiabbr{physchem}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{physics}{\gdef\@journal{physics}\gdef\@journalfull{Physics}\gdef\@journalshort{Physics}\gdef\@doiabbr{physics}\gdef\@ISSN{2624-8174}\gdef\@continuouspages{yes}}
\DeclareOption{physiologia}{\gdef\@journal{physiologia}\gdef\@journalfull{Physiologia}\gdef\@journalshort{Physiologia}\gdef\@doiabbr{physiologia}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{plants}{\gdef\@journal{plants}\gdef\@journalfull{Plants}\gdef\@journalshort{Plants}\gdef\@doiabbr{plants}\gdef\@ISSN{2223-7747}}
\DeclareOption{plasma}{\gdef\@journal{plasma}\gdef\@journalfull{Plasma}\gdef\@journalshort{Plasma}\gdef\@doiabbr{plasma}\gdef\@ISSN{2571-6182}\gdef\@continuouspages{yes}}
\DeclareOption{pollutants}{\gdef\@journal{pollutants}\gdef\@journalfull{Pollutants}\gdef\@journalshort{Pollutants}\gdef\@doiabbr{pollutants}\gdef\@ISSN{2673-4672}\gdef\@continuouspages{yes}}
\DeclareOption{polymers}{\gdef\@journal{polymers}\gdef\@journalfull{Polymers}\gdef\@journalshort{Polymers}\gdef\@doiabbr{polym}\gdef\@ISSN{2073-4360}}
\DeclareOption{polysaccharides}{\gdef\@journal{polysaccharides}\gdef\@journalfull{Polysaccharides}\gdef\@journalshort{Polysaccharides}\gdef\@doiabbr{polysaccharides}\gdef\@ISSN{2673-4176}\gdef\@continuouspages{yes}}
\DeclareOption{poultry}{\gdef\@journal{poultry}\gdef\@journalfull{Poultry}\gdef\@journalshort{Poultry}\gdef\@doiabbr{poultry}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{preprints}{\gdef\@journal{preprints}\gdef\@journalfull{Preprints}\gdef\@journalshort{Preprints}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{proceedings}{\gdef\@journal{proceedings}\gdef\@journalfull{Proceedings}\gdef\@journalshort{Proceedings}\gdef\@doiabbr{proceedings}\gdef\@ISSN{2504-3900}}
\DeclareOption{processes}{\gdef\@journal{processes}\gdef\@journalfull{Processes}\gdef\@journalshort{Processes}\gdef\@doiabbr{pr}\gdef\@ISSN{2227-9717}}
\DeclareOption{prosthesis}{\gdef\@journal{prosthesis}\gdef\@journalfull{Prosthesis}\gdef\@journalshort{Prosthesis}\gdef\@doiabbr{prosthesis}\gdef\@ISSN{2673-1592}\gdef\@continuouspages{yes}}
\DeclareOption{proteomes}{\gdef\@journal{proteomes}\gdef\@journalfull{Proteomes}\gdef\@journalshort{Proteomes}\gdef\@doiabbr{proteomes}\gdef\@ISSN{2227-7382}}
\DeclareOption{psf}{\gdef\@journal{psf}\gdef\@journalfull{Physical Sciences Forum}\gdef\@journalshort{Phys. Sci. Forum}\gdef\@doiabbr{psf}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{psych}{\gdef\@journal{psych}\gdef\@journalfull{Psych}\gdef\@journalshort{Psych}\gdef\@doiabbr{psych}\gdef\@ISSN{2624-8611}\gdef\@continuouspages{yes}}
\DeclareOption{psychiatryint}{\gdef\@journal{psychiatryint}\gdef\@journalfull{Psychiatry International}\gdef\@journalshort{Psychiatry Int.}\gdef\@doiabbr{psychiatryint}\gdef\@ISSN{2673-5318}\gdef\@continuouspages{yes}}
\DeclareOption{publications}{\gdef\@journal{publications}\gdef\@journalfull{Publications}\gdef\@journalshort{Publications}\gdef\@doiabbr{publications}\gdef\@ISSN{2304-6775}}
\DeclareOption{quantumrep}{\gdef\@journal{quantumrep}\gdef\@journalfull{Quantum Reports}\gdef\@journalshort{Quantum Rep.}\gdef\@doiabbr{quantum}\gdef\@ISSN{2624-960X}\gdef\@continuouspages{yes}}
\DeclareOption{quaternary}{\gdef\@journal{quaternary}\gdef\@journalfull{Quaternary}\gdef\@journalshort{Quaternary}\gdef\@doiabbr{quat}\gdef\@ISSN{2571-550X}}
\DeclareOption{qubs}{\gdef\@journal{qubs}\gdef\@journalfull{Quantum Beam Science}\gdef\@journalshort{Quantum Beam Sci.}\gdef\@doiabbr{qubs}\gdef\@ISSN{2412-382X}}
\DeclareOption{radiation}{\gdef\@journal{radiation}\gdef\@journalfull{Radiation}\gdef\@journalshort{Radiation}\gdef\@doiabbr{radiation}\gdef\@ISSN{2673-592X}\gdef\@continuouspages{yes}}
\DeclareOption{reactions}{\gdef\@journal{reactions}\gdef\@journalfull{Reactions}\gdef\@journalshort{Reactions}\gdef\@doiabbr{reactions}\gdef\@ISSN{2624-781X}\gdef\@continuouspages{yes}}
\DeclareOption{recycling}{\gdef\@journal{recycling}\gdef\@journalfull{Recycling}\gdef\@journalshort{Recycling}\gdef\@doiabbr{recycling}\gdef\@ISSN{2313-4321}}
\DeclareOption{regeneration}{\gdef\@journal{regeneration}\gdef\@journalfull{Regeneration}\gdef\@journalshort{Regeneration}\gdef\@doiabbr{regeneration}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{religions}{\gdef\@journal{religions}\gdef\@journalfull{Religions}\gdef\@journalshort{Religions}\gdef\@doiabbr{rel}\gdef\@ISSN{2077-1444}}
\DeclareOption{remotesensing}{\gdef\@journal{remotesensing}\gdef\@journalfull{Remote Sensing}\gdef\@journalshort{Remote Sens.}\gdef\@doiabbr{rs}\gdef\@ISSN{2072-4292}}
\DeclareOption{reports}{\gdef\@journal{reports}\gdef\@journalfull{Reports}\gdef\@journalshort{Reports}\gdef\@doiabbr{reports}\gdef\@ISSN{2571-841X}}
\DeclareOption{reprodmed}{\gdef\@journal{reprodmed}\gdef\@journalfull{Reproductive Medicine}\gdef\@journalshort{Reprod. Med.}\gdef\@doiabbr{reprodmed}\gdef\@ISSN{2673-3897}\gdef\@continuouspages{yes}}
\DeclareOption{resources}{\gdef\@journal{resources}\gdef\@journalfull{Resources}\gdef\@journalshort{Resources}\gdef\@doiabbr{resources}\gdef\@ISSN{2079-9276}}
\DeclareOption{rheumato}{\gdef\@journal{rheumato}\gdef\@journalfull{Rheumato}\gdef\@journalshort{Rheumato}\gdef\@doiabbr{rheumato}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{risks}{\gdef\@journal{risks}\gdef\@journalfull{Risks}\gdef\@journalshort{Risks}\gdef\@doiabbr{risks}\gdef\@ISSN{2227-9091}}
\DeclareOption{robotics}{\gdef\@journal{robotics}\gdef\@journalfull{Robotics}\gdef\@journalshort{Robotics}\gdef\@doiabbr{robotics}\gdef\@ISSN{2218-6581}}
\DeclareOption{ruminants}{\gdef\@journal{ruminants}\gdef\@journalfull{Ruminants}\gdef\@journalshort{Ruminants}\gdef\@doiabbr{ruminants}\gdef\@ISSN{2673-933X}\gdef\@continuouspages{yes}}
\DeclareOption{safety}{\gdef\@journal{safety}\gdef\@journalfull{Safety}\gdef\@journalshort{Safety}\gdef\@doiabbr{safety}\gdef\@ISSN{2313-576X}}
\DeclareOption{sci}{\gdef\@journal{sci}\gdef\@journalfull{Sci}\gdef\@journalshort{Sci}\gdef\@doiabbr{sci}\gdef\@ISSN{2413-4155}}
\DeclareOption{scipharm}{\gdef\@journal{scipharm}\gdef\@journalfull{Scientia Pharmaceutica}\gdef\@journalshort{Sci. Pharm.}\gdef\@doiabbr{scipharm}\gdef\@ISSN{2218-0532}}
\DeclareOption{sensors}{\gdef\@journal{sensors}\gdef\@journalfull{Sensors}\gdef\@journalshort{Sensors}\gdef\@doiabbr{s}\gdef\@ISSN{1424-8220}}
\DeclareOption{separations}{\gdef\@journal{separations}\gdef\@journalfull{Separations}\gdef\@journalshort{Separations}\gdef\@doiabbr{separations}\gdef\@ISSN{2297-8739}}
\DeclareOption{sexes}{\gdef\@journal{sexes}\gdef\@journalfull{Sexes}\gdef\@journalshort{Sexes}\gdef\@doiabbr{sexes}\gdef\@ISSN{2411-5118}\gdef\@continuouspages{yes}}
\DeclareOption{signals}{\gdef\@journal{signals}\gdef\@journalfull{Signals}\gdef\@journalshort{Signals}\gdef\@doiabbr{signals}\gdef\@ISSN{2624-6120}\gdef\@continuouspages{yes}}
\DeclareOption{sinusitis}{\gdef\@journal{sinusitis}\gdef\@journalfull{Sinusitis}\gdef\@journalshort{Sinusitis}\gdef\@doiabbr{sinusitis}\gdef\@ISSN{2309-107X}\gdef\@continuouspages{yes}}
\DeclareOption{skins}{\gdef\@journal{skins}\gdef\@journalfull{Skins}\gdef\@journalshort{Skins}\gdef\@doiabbr{skins}\gdef\@ISSN{}}
\DeclareOption{smartcities}{\gdef\@journal{smartcities}\gdef\@journalfull{Smart Cities}\gdef\@journalshort{Smart Cities}\gdef\@doiabbr{smartcities}\gdef\@ISSN{2624-6511}\gdef\@continuouspages{yes}}
\DeclareOption{sna}{\gdef\@journal{sna}\gdef\@journalfull{Sinusitis and Asthma}\gdef\@journalshort{Sinusitis Asthma}\gdef\@doiabbr{sna}\gdef\@ISSN{2624-7003}}
\DeclareOption{societies}{\gdef\@journal{societies}\gdef\@journalfull{Societies}\gdef\@journalshort{Societies}\gdef\@doiabbr{soc}\gdef\@ISSN{2075-4698}}
\DeclareOption{socsci}{\gdef\@journal{socsci}\gdef\@journalfull{Social Sciences}\gdef\@journalshort{Soc. Sci.}\gdef\@doiabbr{socsci}\gdef\@ISSN{2076-0760}}
\DeclareOption{soilsystems}{\gdef\@journal{soilsystems}\gdef\@journalfull{Soil Systems}\gdef\@journalshort{Soil Syst.}\gdef\@doiabbr{soilsystems}\gdef\@ISSN{2571-8789}}
\DeclareOption{solar}{\gdef\@journal{solar}\gdef\@journalfull{Solar}\gdef\@journalshort{Solar}\gdef\@doiabbr{solar}\gdef\@ISSN{2673-9941}\gdef\@continuouspages{yes}}
\DeclareOption{solids}{\gdef\@journal{solids}\gdef\@journalfull{Solids}\gdef\@journalshort{Solids}\gdef\@doiabbr{solids}\gdef\@ISSN{2673-6497}\gdef\@continuouspages{yes}}
\DeclareOption{sports}{\gdef\@journal{sports}\gdef\@journalfull{Sports}\gdef\@journalshort{Sports}\gdef\@doiabbr{sports}\gdef\@ISSN{2075-4663}}
\DeclareOption{standards}{\gdef\@journal{standards}\gdef\@journalfull{Standards}\gdef\@journalshort{Standards}\gdef\@doiabbr{standards}\gdef\@ISSN{2305-6703}\gdef\@continuouspages{yes}}
\DeclareOption{stats}{\gdef\@journal{stats}\gdef\@journalfull{Stats}\gdef\@journalshort{Stats}\gdef\@doiabbr{stats}\gdef\@ISSN{2571-905X}\gdef\@continuouspages{yes}}
\DeclareOption{stresses}{\gdef\@journal{stresses}\gdef\@journalfull{Stresses}\gdef\@journalshort{Stresses}\gdef\@doiabbr{stresses}\gdef\@ISSN{2673-7140}\gdef\@continuouspages{yes}}
\DeclareOption{surfaces}{\gdef\@journal{surfaces}\gdef\@journalfull{Surfaces}\gdef\@journalshort{Surfaces}\gdef\@doiabbr{surfaces}\gdef\@ISSN{2571-9637}\gdef\@continuouspages{yes}}
\DeclareOption{surgeries}{\gdef\@journal{surgeries}\gdef\@journalfull{Surgeries}\gdef\@journalshort{Surgeries}\gdef\@doiabbr{surgeries}\gdef\@ISSN{2017-2017}\gdef\@continuouspages{yes}}
\DeclareOption{suschem}{\gdef\@journal{suschem}\gdef\@journalfull{Sustainable Chemistry}\gdef\@journalshort{Sustain. Chem.}\gdef\@doiabbr{suschem}\gdef\@ISSN{2673-4079}\gdef\@continuouspages{yes}}
\DeclareOption{sustainability}{\gdef\@journal{sustainability}\gdef\@journalfull{Sustainability}\gdef\@journalshort{Sustainability}\gdef\@doiabbr{su}\gdef\@ISSN{2071-1050}}
\DeclareOption{symmetry}{\gdef\@journal{symmetry}\gdef\@journalfull{Symmetry}\gdef\@journalshort{Symmetry}\gdef\@doiabbr{sym}\gdef\@ISSN{2073-8994}}
\DeclareOption{systems}{\gdef\@journal{systems}\gdef\@journalfull{Systems}\gdef\@journalshort{Systems}\gdef\@doiabbr{systems}\gdef\@ISSN{2079-8954}}
\DeclareOption{taxonomy}{\gdef\@journal{taxonomy}\gdef\@journalfull{Taxonomy}\gdef\@journalshort{Taxonomy}\gdef\@doiabbr{taxonomy}\gdef\@ISSN{2673-6500}\gdef\@continuouspages{yes}}
\DeclareOption{technologies}{\gdef\@journal{technologies}\gdef\@journalfull{Technologies}\gdef\@journalshort{Technologies}\gdef\@doiabbr{technologies}\gdef\@ISSN{2227-7080}}
\DeclareOption{telecom}{\gdef\@journal{telecom}\gdef\@journalfull{Telecom}\gdef\@journalshort{Telecom}\gdef\@doiabbr{telecom}\gdef\@ISSN{2673-4001}\gdef\@continuouspages{yes}}
\DeclareOption{test}{\gdef\@journal{test}\gdef\@journalfull{Test}\gdef\@journalshort{Test}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{textiles}{\gdef\@journal{textiles}\gdef\@journalfull{Textiles}\gdef\@journalshort{Textiles}\gdef\@doiabbr{textiles}\gdef\@ISSN{2673-7248}\gdef\@continuouspages{yes}}
\DeclareOption{thermo}{\gdef\@journal{thermo}\gdef\@journalfull{Thermo}\gdef\@journalshort{Thermo}\gdef\@doiabbr{thermo}\gdef\@ISSN{2673-7264}\gdef\@continuouspages{yes}}
\DeclareOption{tomography}{\gdef\@journal{tomography}\gdef\@journalfull{Tomography}\gdef\@journalshort{Tomography}\gdef\@doiabbr{tomography}\gdef\@ISSN{2379-139X}\gdef\@continuouspages{yes}}
\DeclareOption{tourismhosp}{\gdef\@journal{tourismhosp}\gdef\@journalfull{Tourism and Hospitality}\gdef\@journalshort{Tour. Hosp.}\gdef\@doiabbr{tourhosp}\gdef\@ISSN{2673-5768}\gdef\@continuouspages{yes}}
\DeclareOption{toxics}{\gdef\@journal{toxics}\gdef\@journalfull{Toxics}\gdef\@journalshort{Toxics}\gdef\@doiabbr{toxics}\gdef\@ISSN{2305-6304}}
\DeclareOption{toxins}{\gdef\@journal{toxins}\gdef\@journalfull{Toxins}\gdef\@journalshort{Toxins}\gdef\@doiabbr{toxins}\gdef\@ISSN{2072-6651}}
\DeclareOption{transplantology}{\gdef\@journal{transplantology}\gdef\@journalfull{Transplantology}\gdef\@journalshort{Transplantology}\gdef\@doiabbr{transplantology}\gdef\@ISSN{2673-3943}\gdef\@continuouspages{yes}}
\DeclareOption{transportation}{\gdef\@journal{transportation}\gdef\@journalfull{Transportation}\gdef\@journalshort{Transportation}\gdef\@doiabbr{transportation}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{traumacare}{\gdef\@journal{traumacare}\gdef\@journalfull{Trauma Care}\gdef\@journalshort{Trauma Care}\gdef\@doiabbr{traumacare}\gdef\@ISSN{2673-866X}\gdef\@continuouspages{yes}}
\DeclareOption{traumas}{\gdef\@journal{traumas}\gdef\@journalfull{Traumas}\gdef\@journalshort{Traumas}\gdef\@doiabbr{traumas}\gdef\@ISSN{2673-7477}\gdef\@continuouspages{yes}}
\DeclareOption{tropicalmed}{\gdef\@journal{tropicalmed}\gdef\@journalfull{Tropical Medicine and Infectious Disease}\gdef\@journalshort{Trop. Med. Infect. Dis.}\gdef\@doiabbr{tropicalmed}\gdef\@ISSN{2414-6366}}
\DeclareOption{universe}{\gdef\@journal{universe}\gdef\@journalfull{Universe}\gdef\@journalshort{Universe}\gdef\@doiabbr{universe}\gdef\@ISSN{2218-1997}}
\DeclareOption{urbansci}{\gdef\@journal{urbansci}\gdef\@journalfull{Urban Science}\gdef\@journalshort{Urban Sci.}\gdef\@doiabbr{urbansci}\gdef\@ISSN{2413-8851}}
\DeclareOption{uro}{\gdef\@journal{uro}\gdef\@journalfull{Uro}\gdef\@journalshort{Uro}\gdef\@doiabbr{uro}\gdef\@ISSN{2673-4397}\gdef\@continuouspages{yes}}
\DeclareOption{vaccines}{\gdef\@journal{vaccines}\gdef\@journalfull{Vaccines}\gdef\@journalshort{Vaccines}\gdef\@doiabbr{vaccines}\gdef\@ISSN{2076-393X}}
\DeclareOption{vehicles}{\gdef\@journal{vehicles}\gdef\@journalfull{Vehicles}\gdef\@journalshort{Vehicles}\gdef\@doiabbr{vehicles}\gdef\@ISSN{2624-8921}\gdef\@continuouspages{yes}}
\DeclareOption{venereology}{\gdef\@journal{venereology}\gdef\@journalfull{Venereology}\gdef\@journalshort{Venereology}\gdef\@doiabbr{venereology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{vetsci}{\gdef\@journal{vetsci}\gdef\@journalfull{Veterinary Sciences}\gdef\@journalshort{Vet. Sci.}\gdef\@doiabbr{vetsci}\gdef\@ISSN{2306-7381}}
\DeclareOption{vibration}{\gdef\@journal{vibration}\gdef\@journalfull{Vibration}\gdef\@journalshort{Vibration}\gdef\@doiabbr{vibration}\gdef\@ISSN{2571-631X}\gdef\@continuouspages{yes}}
\DeclareOption{viruses}{\gdef\@journal{viruses}\gdef\@journalfull{Viruses}\gdef\@journalshort{Viruses}\gdef\@doiabbr{v}\gdef\@ISSN{1999-4915}}
\DeclareOption{vision}{\gdef\@journal{vision}\gdef\@journalfull{Vision}\gdef\@journalshort{Vision}\gdef\@doiabbr{vision}\gdef\@ISSN{2411-5150}}
\DeclareOption{water}{\gdef\@journal{water}\gdef\@journalfull{Water}\gdef\@journalshort{Water}\gdef\@doiabbr{w}\gdef\@ISSN{2073-4441}}
\DeclareOption{wem}{\gdef\@journal{wem}\gdef\@journalfull{Wildlife Ecology and Management}\gdef\@journalshort{Wildl. Ecol. Manag.}\gdef\@doiabbr{}\gdef\@ISSN{1234-4321}}
\DeclareOption{wevj}{\gdef\@journal{wevj}\gdef\@journalfull{World Electric Vehicle Journal}\gdef\@journalshort{World Electr. Veh. J.}\gdef\@doiabbr{wevj}\gdef\@ISSN{2032-6653}}
\DeclareOption{wind}{\gdef\@journal{wind}\gdef\@journalfull{Wind}\gdef\@journalshort{Wind}\gdef\@doiabbr{wind}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{women}{\gdef\@journal{women}\gdef\@journalfull{Women}\gdef\@journalshort{Women}\gdef\@doiabbr{women}\gdef\@ISSN{2673-4184}\gdef\@continuouspages{yes}}
\DeclareOption{world}{\gdef\@journal{world}\gdef\@journalfull{World}\gdef\@journalshort{World}\gdef\@doiabbr{world}\gdef\@ISSN{2673-4060}\gdef\@continuouspages{yes}}
\DeclareOption{youth}{\gdef\@journal{youth}\gdef\@journalfull{Youth}\gdef\@journalshort{Youth}\gdef\@doiabbr{youth}\gdef\@ISSN{2673-995X}\gdef\@continuouspages{yes}}
\DeclareOption{zoonoses}{\gdef\@journal{zoonoses}\gdef\@journalfull{Zoonoses}\gdef\@journalshort{Zoonoses}\gdef\@doiabbr{zoonoses}\gdef\@ISSN{2673-9968}\gdef\@continuouspages{yes}}

### 2021-11-12 docs/00-manuscripts/mdpi/manuscript.tex

@@
 % Explain what is new in this study and link to methodology
-Despite their popularity in environmental modeling, there are no studies so far which used ML algorithms in combination with remote sensing data to analyze defoliation at the tree level.
-This study aims to close this gap by analyzing tree defoliation in northern Spain using airborne hyperspectral data.
+In recent years, various studies which have utilized hyperspectral data to analyze pest/fungi infections on trees were published:
+Pinto et al.
+\citep{pinto2020} successfully used hyperspectral imagery to characterize pest infections on peanut leaves using random forest while Yu et.
+al \cite{yu2021} aimed to detect Pine wilt disease in pine plots in China using vegetation indices derived from hyperspectral data.
+Other studies which applied hyperspectral data for forest health monitoring are \cite{lin2014,kayet2019,dash2017}.
+Building upon these successful applications of hyperspectral remote sensing usage in the field of leaf and tree health monitoring, this work analyses tree defoliation in northern Spain using airborne hyperspectral data.
 The methodology of this study uses ML methods in combination with feature selection and hyperparameter tuning.
@@
 The first two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
-In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).
+In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0--100 (in \%) (\autoref{fig:defol-distr}).
 
@@
 		\includegraphics[width=0.7\textwidth] {defoliation-distribution-plot-1.pdf}
-		\caption{Response variable "defoliation at trees" for plots Laukiz1, Laukiz2, Luiando and Oiartzun.
+		\caption{Response variable ``defoliation at trees'' for plots~Laukiz1, Laukiz2, Luiando and Oiartzun.
 \texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation, respectively.
@@
 The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-Oiartzun consists of most tree observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
-All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
-In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
+Oiartzun consists of most tree observations (n = 559) while Laukiz2 shows the largest area size (1.44~ha).
+All plots besides Luiando are located within 100~km from the coast (\autoref{fig:study_area}).
+A total of 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
 Field surveys were conducted in September 2016 by experienced forest pathologists.
 Defoliation was measured via visual inspection using 5\% intervals with the help of a dedicated score card.
-For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% interval of the other three plots (e.g.
-8.33 \%).
+For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% intervals of the other three plots (e.g., 8.33\%).
 Estimating the human observer error of such surveys when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
@@
 
-The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5th 2016.
+The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5, 2016.
 Images were taken by an AISA EAGLE-II sensor.
@@
 		\toprule
-		Geometric resolution   & 1 m                                 \\
+		Geometric resolution   & 1~m                                 \\
 		Radiometric resolution & 12 bit                              \\
-		Spectral resolution    & 126 bands (404.08 nm --- 996.31 nm) \\
+		Spectral resolution    & 126 bands (404.08~nm~--- 996.31~nm) \\
 		Correction:            & Radiometric, geometric, atmospheric
@@
 While various index formulations such as band ratios or normalized ratios are available, indices involving the same bands are strongly correlated.
-Since the widely-used NDVI index belongs to the family of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following normalized difference index (NDI) formula to combine all pairs of reflectances:
+Since the widely\ used NDVI index belongs to the family of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following normalized difference index (NDI) formula to combine all pairs of reflectances:
 
@@
 In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
-In addition, filters are less known than wrapper methods and in recent years ensemble filters were introduced which have shown promising results compared to single filter algorithms \cite{drotar2017a}.
-These two points mainly led to the decision to focus on filter methods for this work and evaluate their effectiveness on highly-correlated, high-dimensional datasets.
-Due to this focus, only this sub-group of feature selection methods will be introduced in greater detail in the following sections.
+In addition, filters are less known than wrapper methods and in recent years ensemble filters were introduced which have shown promising results compared to single filter algorithms \cite{drotar2017}.
+These two points mainly led to the decision to focus on filter methods for this work and evaluate their effectiveness on highly\ correlated, high-dimensional datasets.
+Due to this focus, only this subgroup of feature selection methods will be introduced in greater detail in the following sections.
 
@@
 The concept of filters originates from the idea of ranking features following a score calculated by an algorithm \cite{guyon2003}.
-Some filter methods can only deal with specific types of variables (e.g. numeric or nominal).
+Some filter methods can only deal with specific types of variables (e.g., numeric or nominal).
 Filters only rank features, they do not decide which covariates to drop or keep \cite{drotar2015}.
 The selection which features to keep for model fitting is usually done within the optimization phase during model fitting, along with the hyperparameter tuning.
-Essentially, the number of covariates in the model is treated as a additional hyperparameter of the model.
+Essentially, the number of covariates in the model is treated as an additional hyperparameter of the model.
 The idea is to optimize the number of ranked features to the point at which the model achieves the best performance.
@@
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can possibly enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017a}.
+Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can possibly enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017}.
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
 Ensemble filters align with the recent rise of the 'ensemble' approach in ML which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
-In this work the 'Borda' ensemble filter was used \cite{drotar2017a}.
+In this work the 'Borda' ensemble filter was used \cite{drotar2017}.
 Its final feature order is determined by the sum of all single filters ranks.
@@
 \begin{itemize}
-	\item Univariate/multivariate (scoring based on a single variable / multiple variables).
+	\item Univariate/multivariate (scoring based on a single variable/multiple variables).
 	\item Linear/non-linear (usage of linear/non-linear calculations).
@@
 	\item  Random Forest (RF)
-	\item  Penalized Regression (with L1 / lasso and L2 / ridge penalties)
+	\item  Penalized Regression (with L1/lasso and L2/ridge penalties)
 	\item  Support Vector Machine (SVM, radial basis function Kernel)
@@
 Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmarking studies in recent years.
-Penalized regression is a statistical modeling technique capable of dealing with highly-correlated covariates by penalizing the coefficients of the model \cite{hastie2001}.
+Penalized regression is a statistical modeling technique capable of dealing with highly\ correlated covariates by penalizing the coefficients of the model \cite{hastie2001}.
 Common penalties are 'lasso' (L1) and 'ridge' (L2).
@@
 
-In total, the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3 ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
+In total, the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3~ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets $\times$ 2 models.
 The selected hyperparameter settings are shown in appendix \autoref{tab:hyperparameter_limits}.
@@
 Peña et al. 2017 \cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
-Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method constant makes it possible to draw conclusions on their individual impact.
+Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method steady makes it possible to draw conclusions on their individual impact.
 Each feature set has distinct capabilities that differentiate it from the others.
@@
 For example, feature set VI misses certain parts of the spectral range as the chosen indices only use specific spectral bands.
-Feature set NRI will introduce highly-correlated features, for which some algorithms may be more suitable than others.
+Feature set NRI will introduce highly\ correlated features, for which some algorithms may be more suitable than others.
 In addition to these feature sets, the following combinations of feature sets were also compared:
@@
 
-Some individual features were removed before using the datasets for modelling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
-This preprocessing step reduced the number of covariates for feature set VI to 86 (from 89).
+Some individual features were removed before using the datasets for modeling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
+This preprocessing step reduced the number of covariates for feature set~VI to 86 (from 89).
 This decision was made to prevent issues for the subsequent tuning, filtering and model fitting steps which might occur when passing features with a pairwise correlation of (almost) one.
@@
 The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_\textrm{try} = p_\textrm{sel}^t$, a function of the number of selected features, $p_\textrm{sel}$.
-It was thus tuned on a logarithmic scale by varying $t$ between 0 (i.e. $m_\textrm{try} = 1$) and 0.5 (i.e. $m_\textrm{try}=\sqrt{p_\textrm{sel}}$).
+It was thus tuned on a logarithmic scale by varying $t$ between 0 (i.e., $m_\textrm{try} = 1$) and 0.5 (i.e., $m_\textrm{try}=\sqrt{p_\textrm{sel}}$).
 This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features that were left after optimizing the feature percentage during tuning.
@@
 
-In this work permutation-based feature importance was calculated to estimate feature importance / effects \cite{apley2019}.
+In this work permutation-based feature importance was calculated to estimate feature importance/effects \cite{apley2019}.
 With the limitations in mind when applied to correlated features, the aim was to get a general overview of the feature importance of the hyperspectral bands while trying to avoid an over-interpretation of results.
-The best-performing algorithm on the HR task (i.e. SVM) was used for the feature importance calculation.
+The best-performing algorithm on the HR task (i.e., SVM) was used for the feature importance calculation.
 
@@
 
-For environmental interpretation purposes the ten most important indices of the best performing models of feature sets HR and VI were linked to the spectral regions of the hyperspectral data.
+For environmental interpretation purposes, the ten most important indices of the best performing models of feature sets HR and VI were linked to the spectral regions of the hyperspectral data.
 The aim was to visualize the most important features along the spectral curve of the plots to better understand which spectral regions were most important for the model.
@@
 A complete list of all R packages used in this study can be found in the linked repositories mentioned in the next paragraph.
-Due to space limitations only the selected packages with high impact on this work will be explicitly cited.
+Due to space limitations, only the selected packages with high impact on this work will be explicitly cited.
 
@@
 Differences to other feature sets were very small for lasso (below one p.p.) and a bit larger for ridge regression (between two and five p.p.).
-XGBoost showed very poor performances for some feature sets and fills the ten last places of the ranking (\autoref{tab:perf-worst-10}).
+XGBoost showed very poor performances for some feature sets and fills the last\ ten places of the ranking (\autoref{tab:perf-worst-10}).
 Especially when combined with PCA, the algorithm was not able to achieve adequate performances.
@@
 SVM showed no variation in performance across filters (\autoref{fig:filter-effects-no-filter}).
-The use of filters with RF resulted in a substantial increase in performance in all tasks with the exception of NRI, in which the difference among all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
-XGBoost's performance depended strongly on feature selection: In two out of six tasks, using no filter resulted in the worst or second worst performance.
+The use of filters with RF resulted in a substantial increase in performance in all tasks with the exception of NRI, in which the difference between all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
+XGBoost's performance depended strongly on feature selection: In two out of six tasks, using no filter resulted in the worst or second\-/worst performance.
 XGBoost showed the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p.
@@
 % - never the best
-The Borda filter achieved in only one instance (RF on HR-NRI-VI) the best performance among the used filters (\autoref{fig:filter-effects-borda}).
-For RF and XGBoost it most often ranked within the better half among all filters of a specific task.
-For XGBoost on the NRI task, the Borda filter scored the second worst performance.
+The Borda filter achieved in only one instance (RF on HR-NRI-VI) the best performance of\ all filters (\autoref{fig:filter-effects-borda}).
+For RF and XGBoost, it most often ranked within the better half among all filters of a specific task.
+For XGBoost on the NRI task, the Borda filter scored the second\-/worst performance.
 
@@
 The number of features selected based on model tuning strongly varied across learners and plots.
-While for RF, the least features were selected when Luiando or Oiartzun was the test set ($n = 1$), more than 90\% of all features were used for test sets Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
-RF and SVM used very few features only when Laukiz1 was the test set, (three and one, respectively) whereas this was the plot for which XGBoost used almost all features ($n = 1226$).
-Overall, SVM used the least features across all plots and among all learners, the single highest fraction being 24\% for test set Laukiz2.
+While for RF, the least features were selected when Luiando or Oiartzun was the test set ($n = 1$), more than 90\% of all features were used for test sets~Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
+RF and SVM used very few features only when Laukiz1 was the test set (three and one, respectively) whereas this was the plot for which XGBoost used almost all features ($n = 1226$).
+Overall, SVM used the least features across all plots and among all learners, the highest\ single fraction being 24\% for test set~Laukiz2.
 In general, all learners behaved rather differently on the various plots, and therefore no overall pattern could be identified.
@@
 The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
-For the HR dataset, most (i.e., six out of ten) relevant features clustered around the infrared region (920 nm - 1000 nm), while for VI, eight out of ten concentrate within the wavelength range of 700 nm - 750 nm (the so called \enquote{red edge}).
-For HR, four features in the infrared region (920 nm - 1000 nm) were identified by the model to be most important, being assciated with a mean decrease in RMSE of around 1 p.p.
+For the HR dataset, most (i.e., six out of ten) relevant features clustered around the infrared region (920~nm - 1000~nm), while for VI, eight out of ten concentrate within the wavelength range of 700~nm - 750~nm (the so\-/called \enquote{red edge}).
+For HR, four features in the infrared region (920~nm - 1000~nm) were identified by the model to be most important, being associated with a mean decrease in RMSE of around 1 p.p.
 Overall, apart from the top five features, the vast majority of features showed only a small importance with average decreases in RMSE below 0.5 p.p.
@@
 		\includegraphics[width=0.9\textwidth] {fi-permut-vi-hr-1.pdf}
-		\caption{Variable importance for feature sets HR and VI: Mean decrease in RMSE for one-hundred feature permutations using the SVM learner.
-The wavelength range on the x-axis matches the range of the hyperspectral sensor (400 nm - 1000 nm).
+		\caption{Variable importance for feature sets HR and VI: Mean decrease in RMSE for one\ hundred feature permutations using the SVM learner.
+The wavelength range on the x-axis matches the range of the hyperspectral sensor (400~nm - 1000~nm).
 For each dataset, the ten most important features were highlighted as black dots and labeled by name.
-Grey dots represent features from importance rank 11 to last.
+Gray dots represent features from importance rank~11 to last.
 The spectral signature (mean) of each plot was added as a reference on a normalized reflectance scale [0, 1] (secondary y-axis).
@@
 
-The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE 28.12 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE~28.12 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
 Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
 However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
-Due to the apparent presence of model overfitting in this study it can be postulated that more training data representing a greater variety of situations is needed.
+Due to the apparent presence of model overfitting in this study, it can be postulated that more training data representing a greater variety of situations is needed.
 A model can only make robust predictions if it has learned relationships across the whole range of the response.
 Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
-However, when inspecting the fold level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
+However, when inspecting the fold\-/level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
 This applied to all learners of this study.
-The overall performance of all learners achieved in this study can be classified as "poor" given that only the SVM learner was able to outperform the featureless learner (\autoref{tab:best-learner-perf}).
+The overall performance of all learners achieved in this study can be classified as ``poor'' given that only the SVM learner was able to outperform the featureless learner (\autoref{tab:best-learner-perf}).
 
 In addition, data quality issues (\autoref{subsec:data-quality}) might have an influence on model performances.
-These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g.
-soil reflectance) and the possible positional offset of measured GPS coordinates of trees.
+These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g., soil reflectance) and the possible positional offset of measured GPS coordinates of trees.
 
@@
 The performance of RF and XGBoost on the VI feature set was around four to six p.p. lower than on others.
-One reason could be the lack of coverage in the wavelength area between 810 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
-In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e. NRI-VI, HR-NRI, HR-NRI-VI).
+One reason could be the lack of coverage in the wavelength area between 810~nm and 1000~nm (\autoref{fig:fi-permut-vi-hr}).
+In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e., NRI-VI, HR-NRI, HR-NRI-VI).
 
@@
 
-Arbitrary seeming values of selected feature percentages like 99.97268 \% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
-The specific value to benchmark during tuning is defined by the internal surrogate learner of the MBO tuning method in this study and can be any numeric value between 0 and 1 (interpreted as percentage of available features).
+Arbitrary seeming values of selected feature percentages like 99.97268\% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
+The specific value to benchmark during tuning is defined by the internal surrogate learner of the MBO tuning method in this study and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features).
 
@@
 Ultimately, the results of \autoref{tab:tune-perc-sel-features} should be taken with care as they rely on single model-filter combinations and are subject to random variation.
-More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criteria optimization problem.
+More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criterion optimization problem.
 
@@
 While filters can improve the performance of models, they might be more interesting in other aspects than performance: reducing variables can reduce computational efforts in high-dimensional scenarios and might enhance the interpretability of models.
-Filters are a lot cheaper to compute than wrapper methods and the final feature subset selection can be integrated as an additional hyperparameter into the model optimization stage.
+Filters are a lot cheaper to compute than wrapper methods, and the final feature subset selection can be integrated as an additional hyperparameter into the model optimization stage.
 
@@
 The models which used the Borda ensemble method in this study did not score better on average than models which used a single filter or no filter at all.
-Ensemble methods have higher stability and robustness than single ones and have shown promising results in \cite{drotar2017a}.
-Hence, their main advantage are stable performances across datasets with varying characteristics.
+Ensemble methods have higher stability and robustness than single ones and have shown promising results in \cite{drotar2017}.
+Hence, their main advantage is stable performances across datasets with varying characteristics.
 Single filter methods might yield better model performances on certain datasets but fail on others.
@@
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
-In summary, Borda performs not better than a randomly selected filter method in this study.
+In summary, Borda does\ not\ perform better than a randomly selected filter method in this study.
 More case studies applying ensemble filter methods are needed to verify this finding.
@@
 PCA, acting as a filter method, more often showed less optimal results, especially for algorithms RF and XGBoost.
-Especially XGBoost had substantial problems when using PCA as a filter method and scored the worst four results (\autoref{tab:perf-worst-10}).
+Especially XGBoost had substantial problems when using PCA as a filter method and scored the four\ worst results (\autoref{tab:perf-worst-10}).
 However, PCA was able to reduce model fitting times substantially across all algorithms.
@@
 
-Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 680 nm to 750 nm.
-
-This area has the highest ability to distinguish between reflectances related to a high density / high foliage density und thus the health status of vegetation and its respective counterpart \cite{horler1983}.
-However, four out of ten of the most important features of dataset HR are located between 920 nm and 1000 nm.
-Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area - especially for plot Oiartzun - which might explain why these features were considered important by the model.
-
-A possible explanation for the worse performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850 nm and 1000 nm (\autoref{fig:fi-permut-vi-hr}).
-The majority of VI features covers the range between 550 nm - 800 nm.
-Only one index (PWI) covers information in the range beyond 900 nm.
-
-\subsection{Data Quality}
+Not surprisingly the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 680~nm to 750~nm.
+
+This area has the highest ability to distinguish between reflectances related to a high density/high foliage density and thus the health status of vegetation and its respective counterpart \cite{horler1983}.
+However, four out of ten of the most important features of dataset HR are located between 920~nm and 1000~nm.
+Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area---especially for plot Oiartzun---which might explain why these features were considered important by the model.
+
+A possible explanation for the poor performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850~nm and 1000~nm (\autoref{fig:fi-permut-vi-hr}).
+The majority of VI features covers the range between 550~nm - 800~nm.
+Only one index (PWI) covers information in the range beyond 900~nm.
+
+\subsection{Data quality}
 \label{subsec:data-quality}
@@
 It was concluded that using a buffer of one meter can be a good compromise between the inclusion of information from too many surrounding trees, mapping a single tree crown accordingly accounting for a possible geometric offset.
-The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e. $<= 1m$).
+The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e., $<= 1m$).
 This results in all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
 Even though no results showing the influence of different buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
-Besides using a buffer to extract the hyperspectral information, a segmentation could have been considered.
+Besides using a buffer to extract the hyperspectral information, segmentation could have been considered.
 However, this method would have required more effort for no clear added value in our view and would have moved the focus of this manuscript more to data preprocessing and away from the focus on filter-based feature selection methods.
@@
 
-The available hyperspectral data covered a wavelength between 400 nm and 1000 nm.
+The available hyperspectral data covered a wavelength between 400~nm and 1000~nm.
 Hence, the wavelength range of the shortwave infrared (SWIR) region is not covered in this study.
-Given that this wavelength range is often used in forest health studies \cite{hais2019}, e.g. when calculating the \ac{NDMI} index \cite{gao1996}, this marks a clear limitation of the dataset at hand.
+Given that this wavelength range is often used in forest health studies \cite{hais2019}, e.g., when calculating the \ac{NDMI} index \cite{gao1996}, this marks a clear limitation of the dataset at hand.
 
@@
 The R package hsdar was used for the calculation of vegetation indices \cite{lehnert2016}.
-All indices that could be calculated with the given spectral range of the data (400 nm - 1000 nm) were used.
+All indices that could be calculated with the given spectral range of the data (400~nm - 1000~nm) were used.
 This means even though \autoref{tab:vegindices} lists all available indices of the package, not all listed indices were used in this study.
@@
 Such limitations and uncertainties apply to most environmental studies and cannot be completely avoided.
+
+\subsection{Practical implications on defoliation and tree health mapping}
+Even though this work has a strong methodological focus by comparing different benchmark settings on highly\ correlated feature sets, implications on tree health should be briefly discussed in the following.
+Due to the outlined dataset issues in \autoref{subsec:data-quality}, which are mainly responsible for the resulting poor model performances, the trained models are not suited for practical use, e.g., to predict defoliation in unknown areas, due to the high mapping uncertainty.
+Yet the general approach of utilizing hyperspectral data to classify the health status of trees partly led to promising results:
+For example, due to the narrow bandwidth of the hyperspectral sensor small parts of the spectra (mainly in the infrared region) were of higher importance to the models (e.g., see \autoref{fig:fi-permut-vi-hr}), meaning they helped the models to distinguish between low and high tree defoliation.
+If spatial offset errors of the image data and possible background noise can be reduced (possibly by making use of image segmentation), we believe that model performances could be substantially enhanced.
+Such improved models, starting around an RMSE of 20\% and less, should be able to provide added value to support the long\-/term monitoring of forest health and early detection of fungi affected tree plots.
+Though overall the usage of defoliation as a proxy to forest health should be critically questioned as it comes with various practical issues, starting from potential offsets during data collection, varying leaf density due to tree age and differing effects between tree species, to just name a few.
 
@@
 % defoliation studies and what data they used
-While most defoliation studies operate on the plot level using coarser-resolution multispectral satellite data \cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
+While most defoliation studies operate on the plot level using coarser\ resolution multispectral satellite data \cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
 Among these, \cite{meng2018, kalin2019} used ground-level methods such as \ac{ALS} or \ac{LiDAR}.
@@
 However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
-To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014a, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
+To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
 In general, applying feature selection methods on hyperspectral datasets has shown to be effective, regardless of the method used \cite{pal2010, keller2016}.
 However, no study that made explicit use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
-Potential reasons for this absence could be an easier programmatic access to wrapper methods and a higher general awareness of such compared to filter methods.
+Potential reasons for this absence could be an easier programmatic access to wrapper methods, and a higher general awareness of such compared to filter methods.
 Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.
@@
 A grid search was used for hyperparameter tuning and \ac{FFS} for variable selection.
-\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with a RF classifier.
+\cite{ludwig2019} analyzed woody cover in South Africa using spatial \ac{CV} and \ac{FS} approach \cite{meyer2018} with an RF classifier.
 \cite{zandler2015} shows a similar setup: they used hyperspectral vegetation indices, a nested CV approach for performance estimation and estimated variable importance targeting woody biomass as the response.
 In the results, lasso showed the best performance among the chosen methods.
-However, the authors did not optimize the hyperparameters of RF which makes a fair comparison problematic since the other models used perform internal optimization.
+However, the authors did not optimize the hyperparameters of RF which makes a fair comparison problematic since the other models used\ perform internal optimization.
 The discussion section of \cite{zandler2015} lists additional studies that made use of shrinkage models for high dimensional remote sensing modeling.
@@
 
-Other fields (e.g. bioinformatics) face high-dimensional datasets more often.
+Other fields (e.g., bioinformatics) face high-dimensional datasets more often.
 Hence more studies using (filter-based) feature-selection approaches can be found for this field \cite{guo2019, radovic2017}.
 Yet bioinformatics differs conceptually in many ways from environmental modeling and therefore no greater focus was put into comparing studies of this field.
-The availability of high dimensional feature sets will increase in the future due to higher temporal and spectral resolutions of sensors.
+The availability of high\-/dimensional feature sets will increase in the future due to higher temporal and spectral resolutions of sensors.
 In addition, a high spatial resolution comes with the possibility to calculate many textural features.
-Hence, the ability to deal with high dimensional datasets becomes more important and unbiased robust approaches are needed.
+Hence, the ability to deal with high\-/dimensional datasets becomes more important and unbiased robust approaches are needed.
 We hope that this work and its methodology raises awareness about the application of filter methods to tackle high-dimensional problems in the environmental modeling field.
@@
 Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
-SVM showed the most robust behavior across all highly-correlated datasets and was able to predict the response variable of this study substantially better than other methods.
+SVM showed the most robust behavior across all highly\ correlated datasets and was able to predict the response variable of this study substantially better than other methods.
 
@@
 The addition of derived feature sets was in most cases able to improve predictive performance.
-In contrast, feature sets which focused on only a small fraction of the available spectral range (i.e.
-dataset VI) showed a worse performance than the ones which covered wider range (400 nm - 1000 nm; HR, NRI).
+In contrast, feature sets which focused on only a small fraction of the available spectral range (i.e., dataset~VI) showed a worse performance than the ones which covered a\ wider range (400~nm - 1000~nm; HR, NRI).
 NRIs can be seen as a valuable addition for optimizing predictive performance in remote sensing of vegetation.
@@
 		Carter6         & $R_{550}$                                                                                            & \cite{carter1994}             \\
-		CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                    & \cite{zarco-tejada2003a}      \\
-		CI2             & $R_{760}/R_{700}-1$                                                                                  & \cite{gitelson2003}           \\
+		CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                    & \cite{zarco-tejada2003}      \\
+		CI~2             & $R_{760}/R_{700}-1$                                                                                  & \cite{gitelson2003}           \\
 		ClAInt          & $\int_{600 nm}^{735 nm} R$                                                                           & \cite{oppelt2004}             \\
@@
 		CRI4            & $1/R_{515}-1/R_{700}\times R_{770}$                                                                  & \cite{gitelson2003}           \\
-		D1              & $D_{730}/D_{706}$                                                                                    & \cite{zarco-tejada2003a}      \\
-		D2              & $D_{705}/D_{722}$                                                                                    & \cite{zarco-tejada2003a}      \\
+		D1              & $D_{730}/D_{706}$                                                                                    & \cite{zarco-tejada2003}      \\
+		D2              & $D_{705}/D_{722}$                                                                                    & \cite{zarco-tejada2003}      \\
 		Datt            & $(R_{850}-R_{710})/(R_{850}-R_{680})$                                                                & \cite{datt1999}               \\
@@
 		DDn             & $2\times (R_{710}-R_{660}-R_{760})$                                                                  & \cite{lemaire2008}            \\
-		DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                        & \cite{zarco-tejada2003a}      \\
+		DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                        & \cite{zarco-tejada2003}      \\
 		DWSI1           & $R_{80}/R_{1660}$                                                                                    & \cite{apan2004}               \\
@@
 		EGFR            & $\max(D_{650:750})/\max(D_{500:550})$                                                                & \cite{penuelas1994}           \\
-		EVI             & $\frac{2.5\times ((R_{800}-R_{670}) }{ (R_{800}-(6\times R_{670})-(7.5\times R_{475})+1)}$           & \cite{huete1997a}             \\
+		EVI             & $\frac{2.5\times ((R_{800}-R_{670}) }{ (R_{800}-(6\times R_{670})-(7.5\times R_{475})+1)}$           & \cite{huete1997}             \\
 		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$**                                                    & \cite{wu2014}                 \\
@@
 		\midrule
-		MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}          & \cite{wu2008a}                \\
-		\midrule
-		mND705          & $\frac{(R_{750}-R_{705})}{R_{750}+R_{705}-2\times R_{445}}$                                          & \cite{sims2002a}              \\
-		mNDVI           & $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$                                         & \cite{sims2002a}              \\
+		MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}          & \cite{wu2008}                \\
+		\midrule
+		mND705          & $\frac{(R_{750}-R_{705})}{R_{750}+R_{705}-2\times R_{445}}$                                          & \cite{sims2002}              \\
+		mNDVI           & $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$                                         & \cite{sims2002}              \\
 		MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                            & \cite{hernandez-clemente2011} \\
@@
 		MSI             & $\frac{R_{1600}}{R_{817}}$                                                                           & \cite{hunt1989}               \\
-		mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                            & \cite{sims2002a}              \\
+		mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                            & \cite{sims2002}              \\
 		mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                               & \cite{chen1996}               \\
-		mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                            & \cite{sims2002a}              \\
+		mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                            & \cite{sims2002}              \\
 		MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                            & \cite{dash2007}               \\
@@
 		OSAVI           & $\frac{(1+0.16) \times (R_{800}-R_{670})}{R_{800}+R_{670}+0.16 }$                                    & \cite{rondeaux1996}           \\
-		OSAVI2          & $\frac{(1+0.16)\times (R_{750}-R_{705})}{R_{750}+R_{705}+0.16) }$                                    & \cite{wu2008a}                \\
+		OSAVI2          & $\frac{(1+0.16)\times (R_{750}-R_{705})}{R_{750}+R_{705}+0.16) }$                                    & \cite{wu2008}                \\
 		PARS            & $\frac{R_{746}}{R_{513}}$                                                                            & \cite{chappelle1992}          \\
 		PRI             & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            & \cite{gamon1992}              \\
-		PRI\_norm       & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                & \cite{zarco-tejada2013a}      \\
+		PRI\_norm       & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                & \cite{zarco-tejada2013}      \\
 		PRI*CI2         & $PRI*CI2$                                                                                            & \cite{garrity2011}            \\
@@
 		SRPI            & $\frac{R_{430}}{R_{680}}$                                                                            & \cite{penuelasj.1995}         \\
-		SRWI            & $\frac{R_{850}}{R_{1240}}$                                                                           & \cite{zarco-tejada2003a}      \\
+		SRWI            & $\frac{R_{850}}{R_{1240}}$                                                                           & \cite{zarco-tejada2003}      \\
 		Sum\_Dr1        & $\sum_{i=626}^{795} D1_i$                                                                            & \cite{elvidge1995}            \\
@@
 		\midrule
-		TCARI2          & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} & \cite{wu2008a}                \\
-		\midrule
-		TCARI2/OSAVI2   & TCARI2/OSAVI2                                                                                        & \cite{wu2008a}                \\
+		TCARI2          & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} & \cite{wu2008}                \\
+		\midrule
+		TCARI2/OSAVI2   & TCARI2/OSAVI2                                                                                        & \cite{wu2008}                \\
 		\midrule
@@
 
-The AISA EAGLE-II sensor was used for airborne image acquisition with a field of view of 37.7°.
-Its spectral resolution is 2.4 nm and ranged from 400 to 1000 nm.
-
-The conversion of Digital Numbers (DN) to spectral radiance was made by using a software designed for the instrument.
+The AISA EAGLE-II sensor was used for airborne image acquisition with a field of view of 37.7~°.
+Its spectral resolution is 2.4~nm and ranged from 400~nm\ to 1000~nm.
+
+The conversion of Digital Numbers (DN) to spectral radiance was made by using software designed for the instrument.
 Images were originally scaled in 12 bits but were radiometrically calibrated to 16 bits, reserving the highest value (65535) for null values.
@@
 The aim of this procedure was to reduce the positional errors of the images.
-The cartographic reference system in use was EPSG 25830.
+The cartographic reference system in use was EPSG~25830.
 Positioning was done by coupling an Applanix POS AV 410 system to the sensor, integrating GPS and IMU systems.
 The system provides geographic coordinates of the terrain and relative coordinates of the aircraft (attitude) at each scanned line.
-Additionally a DSM from GeoEuskadi with a spatial resolution of 1 m was used.
+Additionally a DSM from GeoEuskadi with a spatial resolution of 1~m was used.
 The ortorectified hyperspectral images were compared to orthoimages (1:5000) from GeoEuskadi.
@@
 Radiation may be absorbed or scattered (Rayleigh and Mie scattering).
-Scattering is responsible for the adjacency effect, i.e., radiation coming from neighbors areas to the target pixel.
+Scattering is responsible for the adjacency effect, i.e., radiation coming from neighbors' areas to the target pixel.
 The MODTRAN algorithm was used to model the effect of the atmosphere on the radiation.
@@
 In addition the optical thickness was estimated on pixels with a high vegetation cover.
-Columnar water vapor was estimated by a linear regression ratio where the spectral radiance of each pixel at the band of the maximum water absorption (~906 nm) is compared to its theoretical value in absence of absorption.
-Nonetheless, this technique is unreliable in presence of a spectral resolution as in this case.
-To resolve this, the water vapor parameter was selected manually according to the smoothness observed on the reflectance peak at 960 nm.
-This was combined with an mid-latitude summer atmosphere model.
+Columnar water vapor was estimated by a linear regression ratio where the spectral radiance of each pixel at the band of the maximum water absorption (~906~nm) is compared to its theoretical value in the\ absence of absorption.
+Nonetheless, this technique is unreliable in the\ presence of a spectral resolution as in this case.
+To resolve this, the water vapor parameter was selected manually according to the smoothness observed on the reflectance peak at 960~nm.
+This was combined with a mid-latitude summer atmosphere model.
 The output of this procedure was reflectance from the target pixel scaled between 0 and 10,000.
@@
 The image acquisitions were originally attempted during one day (29.10.2016).
-Due to the variable meteorological conditions some stands had to be imaged one day later.
+Due to the variable meteorological conditions, some stands had to be imaged one day later.
 
@@
 	His area of expertise is applied machine learning, more specifically the field of environmental modeling.
-	He is a PhD candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.
-	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.}
+	He is a Ph.D. candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.
+	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatiotemporal data handling.}
 
@@
 {\textbf{Jannes Muenchow} is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
-	He received his PhD from Friedrich-Alexander University Erlangen-Nürnberg (Germany) in 2013.
-	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia in 2016 (Friedrich Schiller University Jena).
+	He received his Ph.D. from Friedrich-Alexander University of\ Erlangen-Nürnberg (Germany) in 2013.
+	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia in 2016 (Friedrich Schiller University of\ Jena).
 	Currently, his research focuses on developing open source tools for ecology, geomorphology and qualitative GIS.
-	He is a co-author of the book "Geocomputation with R".}
+	He is a co-author of the book ``Geocomputation with R''.}
 
@@
 {\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{eiturritxa.jpg}}}
-{\textbf{Eugenia Iturritxa} received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.
+{\textbf{Eugenia Iturritxa} received the Ecology and Ph.D. degree in Plant Protection from the University of the Basque Country in 2001.
 	Since 1999 her main research has focused on forest health, on the study of diverse species of native and introduced pathogenic fungi in forests and forest plantations.
@@
 {\textbf{José Cortés} received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
-	Currently he is a PhD student at Friedrich Schiller University, Jena, Germany.
+	Currently he is a Ph.D. student at Friedrich Schiller University, Jena, Germany.
 	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
-	His research focuses on spatio-temporal trend detection in environmental data.}
+	His research focuses on spatiotemporal trend detection in environmental data.}
 
@@
 {\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{abrenning.jpg}}}
-{\textbf{Alexander Brenning} (Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
+{\textbf{Alexander Brenning} (Ph.D.~2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu~Berlin.
 	He served as an assistant professor and tenured associate professor in geomatics at the University of Waterloo, Ontario, Canada from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
-	Dr. Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.
+	Dr.~Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.
 	He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.}

### 2021-11-13 docs/00-manuscripts/mdpi/manuscript.tex
clarify confusing sentence

@@
 This decision was made to prevent issues for the subsequent tuning, filtering and model fitting steps which might occur when passing features with a pairwise correlation of (almost) one.
-Additional variable selection was solely devoted to the respective feature selection methods in this study. <!-- FIXME: I don't understand this sentence. -->
+The remaining features were then used as input for the filter-based feature selection within the CV.
 
@@
 The combination of feature sets showed small increases in performance for some learners.
-RF and XGBoost scored slightly better on the combined datasets HR-NRI, NRI-VI and HR-NRI-VI, respectively, compared to their standalone variants (NRI, VI) (\autoref{fig:perf-result}).
-However, the best performances for both RF and XGBoost were scored on HR only.
-Datasets containing derived features only (VI, NRI) showed no improvement in performance compared to the raw hyperspectral band information (HR) or combined feature sets.
+XGBoost scored slightly better on the combined datasets HR-NRI, NRI-VI and HR-NRI-VI, respectively, compared to their standalone variants (NRI, VI) (\autoref{fig:perf-result}).
+However, the best performances for RF and XGBoost were scored on NRI and HR, respectively.
+RF showed a substantial performance increase when using only NRI compared to all other feature sets whereas for XGBoost the worst performances were associated with the VI and NRI only feature sets (\autoref{fig:perf-result}).
 
 % learner
-% - SVM Car is the overall winner
-% - Lasso quite stable (varies within 1 - 4), ridge a bit more variable (2 - 6)
-% - all MBO penalized methods have no difference across tasks
-SVM combined with the \enquote{Relief} filter achieved the best overall performance (RMSE of 28.09 p.p.) (\autoref{tab:perf-top-10}).
+% - SVM Info Gain is the overall winner
+% - Combined sets do not help for XBoost and Ridge
+% - RF NRI quite strong
+SVM combined with the \enquote{Information Gain} filter achieved the best overall performance (RMSE of 27.915 p.p.) (\autoref{tab:perf-top-10}).
 %L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
 Regression with ridge (L2) and lasso (L1) penalties showed their best performances on the NRI feature set (\autoref{tab:best-learner-perf}).
-Differences to other feature sets were very small for lasso (below one p.p.) and a bit larger for ridge regression (between two and five p.p.).
+Combining feature sets for lasso and ridge did not help to increase performance and while there was no subatantial difference for lasso, ridge scored around two percentage points on such.
 XGBoost showed very poor performances for some feature sets and fills the last\ ten places of the ranking (\autoref{tab:perf-worst-10}).
@@
 SVM showed no variation in performance across filters (\autoref{fig:filter-effects-no-filter}).
-The use of filters with RF resulted in a substantial increase in performance in all tasks with the exception of NRI, in which the difference between all filters was also the smallest (\autoref{fig:filter-effects-no-filter}).
-XGBoost's performance depended strongly on feature selection: In two out of six tasks, using no filter resulted in the worst or second\-/worst performance.
-XGBoost showed the highest overall differences between filters for a single task: for feature set HR, the range is up to 15 p.p.
+The use of filters for RF resulted in a substantial increase in performance in all tasks, especially on the HR feature set where all filters showed an improved score compared to using no filter (\autoref{fig:filter-effects-no-filter}).
+XGBoost's performance depended strongly on feature selection: In two out of six tasks (HR, VI), using no filter resulted in the worst performance.
+XGBoost showed the highest overall differences between filters for a single task: for feature set HR, the range is up to 13 p.p.
 (\enquote{CMIM} vs.
@@
 % - SVM unaffected
-When comparing the usage of filters against using no filter at all, there were no instances where a model without filtering scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
+When comparing the usage of filters against using no filter at all, there were no instances in which a non-filtered model scored a better performance than the best filtered one (\autoref{fig:filter-effects-no-filter}).
 For SVM, all filters and \enquote{no filter} achieved roughly the same performance on all tasks.
@@
 % - never the best
-The Borda filter achieved in only one instance (RF on HR-NRI-VI) the best performance of\ all filters (\autoref{fig:filter-effects-borda}).
+The Borda filter was not able to score the best performance in any learner/filter setting (\autoref{fig:filter-effects-borda}).
 For RF and XGBoost, it most often ranked within the better half among all filters of a specific task.
-For XGBoost on the NRI task, the Borda filter scored the second\-/worst performance.
 
 % Selected percentage of features across plots
+FIXME:
 The number of features selected based on model tuning strongly varied across learners and plots.
@@
 
+FIXME:
 The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
@@
 
-The best aggregated performance of this study (SVM + \enquote{Relief} filter, RMSE~28.12 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE~27.915 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
 Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
@@
 
+FIXME:
 Arbitrary seeming values of selected feature percentages like 99.97268\% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
@@
 Single filter methods might yield better model performances on certain datasets but fail on others.
-The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in all cases (besides RF on task HR-NRI-VI) a single filter outperformed the ensemble filter.
+The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in all cases a single filter outperformed the ensemble filter.
 However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
@@
 PCA, acting as a filter method, more often showed less optimal results, especially for algorithms RF and XGBoost.
-Especially XGBoost had substantial problems when using PCA as a filter method and scored the four\ worst results (\autoref{tab:perf-worst-10}).
+Especially XGBoost had substantial problems when using PCA as a filter method and scored four of the six worst results (\autoref{tab:perf-worst-10}).
 However, PCA was able to reduce model fitting times substantially across all algorithms.

### 2021-11-14 docs/00-manuscripts/mdpi/tune-perc-sel-features.tex
update tune results table

@@
   \midrule
-\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.00614 & 1/1249 \\ 
-   & Laukiz2 & 16.58887 & 226/1357 \\ 
-   & Luiando & 0.06681 & 2/1507 \\ 
-   & Oiartzun & 65.06532 & 854/1311 \\ 
-  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.50917 & 7/1249 \\ 
-   & Laukiz2 & 91.95721 & 1248/1357 \\ 
-   & Luiando & 0.50917 & 8/1507 \\ 
-   & Oiartzun & 19.70660 & 259/1311 \\ 
-  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 93.70844 & 1171/1249 \\ 
-   & Laukiz2 & 3.51986 & 48/1357 \\ 
-   & Luiando & 0.12474 & 2/1507 \\ 
-   & Oiartzun & 85.28181 & 1119/1311 \\ 
+\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.00245 & 1/1249 \\ 
+   & Laukiz2 & 0.00359 & 1/1357 \\ 
+   & Luiando & 0.12448 & 2/1507 \\ 
+   & Oiartzun & 2.80356 & 37/1311 \\ 
+  \midrule\multirow{4}{*}{\specialcell{SVM \\ Car}} & Laukiz1 & 16.76686 & 210/1249 \\ 
+   & Laukiz2 & 40.77700 & 554/1357 \\ 
+   & Luiando & 43.80604 & 661/1507 \\ 
+   & Oiartzun & 81.23205 & 1065/1311 \\ 
+  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 79.54091 & 994/1249 \\ 
+   & Laukiz2 & 0.96545 & 14/1357 \\ 
+   & Luiando & 66.27871 & 999/1507 \\ 
+   & Oiartzun & 41.89759 & 550/1311 \\ 
    \bottomrule

### 2021-11-14 analysis/eval-performance.Rmd

--- 2021-11-02 analysis/eval-performance.Rmd
@@
   dplyr::mutate(iter = case_when(
-    learner.id == "XGBOOST MBO CMIM" & iter == 1 ~ "Luiando",
-    learner.id == "XGBOOST MBO CMIM" & iter == 2 ~ "Laukiz1",
-    learner.id == "XGBOOST MBO CMIM" & iter == 3 ~ "Oiartzun",
-    learner.id == "XGBOOST MBO CMIM" & iter == 4 ~ "Laukiz2",
+    learner.id == "XGBOOST MBO Borda" & iter == 1 ~ "Luiando",
+    learner.id == "XGBOOST MBO Borda" & iter == 2 ~ "Laukiz1",
+    learner.id == "XGBOOST MBO Borda" & iter == 3 ~ "Oiartzun",
+    learner.id == "XGBOOST MBO Borda" & iter == 4 ~ "Laukiz2",
 
-    learner.id == "SVM MBO Relief" & iter == 1 ~ "Laukiz1",
-    learner.id == "SVM MBO Relief" & iter == 2 ~ "Oiartzun",
-    learner.id == "SVM MBO Relief" & iter == 3 ~ "Luiando",
-    learner.id == "SVM MBO Relief" & iter == 4 ~ "Laukiz2",
+    learner.id == "SVM MBO Car" & iter == 1 ~ "Laukiz1",
+    learner.id == "SVM MBO Car" & iter == 2 ~ "Oiartzun",
+    learner.id == "SVM MBO Car" & iter == 3 ~ "Luiando",
+    learner.id == "SVM MBO Car" & iter == 4 ~ "Laukiz2",
 
-    learner.id == "RF MBO Relief" & iter == 1 ~ "Luiando",
-    learner.id == "RF MBO Relief" & iter == 2 ~ "Laukiz2",
-    learner.id == "RF MBO Relief" & iter == 3 ~ "Laukiz1",
-    learner.id == "RF MBO Relief" & iter == 4 ~ "Oiartzun",
+    learner.id == "RF MBO Car" & iter == 1 ~ "Luiando",
+    learner.id == "RF MBO Car" & iter == 2 ~ "Laukiz2",
+    learner.id == "RF MBO Car" & iter == 3 ~ "Laukiz1",
+    learner.id == "RF MBO Car" & iter == 4 ~ "Oiartzun",
 
@@
   dplyr::mutate(Learner = forcats::fct_recode(Learner,
-    "RF \\\\ Relief" = "RF MBO Relief", "XGB \\\\ CMIM" = "XGBOOST MBO CMIM",
-    "SVM \\\\ Relief" = "SVM MBO Relief"
+    "RF \\\\ Car" = "RF MBO Car", "XGB \\\\ Borda" = "XGBOOST MBO Borda",
+    "SVM \\\\ Car" = "SVM MBO Car"
   )) %>%

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
update selected features part WRT to changes of results

@@
 % Selected percentage of features across plots
-FIXME:
-The number of features selected based on model tuning strongly varied across learners and plots.
-While for RF, the least features were selected when Luiando or Oiartzun was the test set ($n = 1$), more than 90\% of all features were used for test sets~Laukiz1 and Laukiz2 (\autoref{tab:tune-perc-sel-features}).
-RF and SVM used very few features only when Laukiz1 was the test set (three and one, respectively) whereas this was the plot for which XGBoost used almost all features ($n = 1226$).
-Overall, SVM used the least features across all plots and among all learners, the highest\ single fraction being 24\% for test set~Laukiz2.
-In general, all learners behaved rather differently on the various plots, and therefore no overall pattern could be identified.
+The number of features selected during model optimization strongly varied across learners and plots.
+RF selected the least features of all three learners, and with the expection of Oiartzun, only one or two variables got selected.
+SVM used 210 features or more in all instances and varied between 16\% (Laukiz1) - 81\% (Oiartzun) (\autoref{tab:tune-perc-sel-features}).
+XGBoost also favored using several hundred features with the exeption of Laukiz2 for which only 14 (0.96\%) were selected.
 
@@
 
-FIXME:
-The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B69) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B65) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
 For the HR dataset, most (i.e., six out of ten) relevant features clustered around the infrared region (920~nm - 1000~nm), while for VI, eight out of ten concentrate within the wavelength range of 700~nm - 750~nm (the so\-/called \enquote{red edge}).
@@
 
-FIXME:
-Arbitrary seeming values of selected feature percentages like 99.97268\% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
-The specific value to benchmark during tuning is defined by the internal surrogate learner of the MBO tuning method in this study and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features).
+Arbitrary seeming values of selected feature percentages in \autoref{tab:tune-perc-sel-features} emerge from the specific hyperparameter setting which got elected as the best one during the tuning step.
+The specific performance value to benchmark against during tuning is defined by the internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features) for the hyperparameter that controls the number of features.
 

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
affiliations, details in bios

@@
 {\textbf{Jannes Muenchow} is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
-	He received his Ph.D. from Friedrich-Alexander University of\ Erlangen-Nürnberg (Germany) in 2013.
-	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia in 2016 (Friedrich Schiller University of\ Jena).
-	Currently, his research focuses on developing open source tools for ecology, geomorphology and qualitative GIS.
+	He received his Ph.D. from Friedrich-Alexander-Universität Erlangen-Nürnberg (Germany) in 2013.
+	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia from 2016 to 2019 (Friedrich Schiller University Jena).
+	His research focuses on developing open-source tools for ecology, geomorphology and qualitative GIS.
 	He is a co-author of the book ``Geocomputation with R''.}
@@
 {\textbf{José Cortés} received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
-	Currently he is a Ph.D. student at Friedrich Schiller University, Jena, Germany.
+	Currently he is a Ph.D. student at Friedrich Schiller University Jena, Germany.
 	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
@@
 {\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{bbischl.jpg}}}
-{\textbf{Bernd Bischl} obtained his Ph.D from Dortmund Technical University in 2013.
+{\textbf{Bernd Bischl} obtained his Ph.D. from TU Dortmund University in 2013.
 	He is a professor of Statistical Learning and Data Science in the Department of Statistics at Ludwig-Maximilians-Universität in Munich, Germany, and a director of the Munich Center for Machine Learning.

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
revised abstract

@@
 Variable importance was assessed using permutation-based feature importance.
-Overall, the support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
+Overall, the support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), and lasso (L1) and ridge (L2) regressions by at least three percentage points.
 The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
@@
 Ensemble filters did not have a substantial impact on performance.
-Permutation-based feature importance estimated features around the red edge to be most important.
-However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the overall best performances.
-More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
-Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.}
+The most important features were located around the red edge.
+Additional features in the near-infrared region (800 nm - 1000 nm) were also essential to achieve the overall best performances.
+Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.
+Nevertheless, more training data and replication in similar benchmarking studies are needed to be able to generalize the results.}
 

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
revised conclusion

@@
 
-This study analyzed defoliation of trees in northern Spain by using hyperspectral data as input for ML models which used hyperparameter tuning and filter-based feature selection.
-Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
+This study analyzed the effectiveness of filter-based feature selection in improving various machine-learning models of defoliation of trees in northern Spain based on hyperspectral remote-sensing data.
+Substantial differences in performance occurred depending on which feature-selection and machine-learning methods were combined.
 SVM showed the most robust behavior across all highly\ correlated datasets and was able to predict the response variable of this study substantially better than other methods.

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
revised intro, chaged order of research questions

@@
 The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade \cite{lary2016}.
-This coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014, which serves as training and prediction data.
+This coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
 At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from the open-source community.
 Scientists can nowadays process large amounts of (environmental) information with relative ease using various learning algorithms.
-This makes it possible to easily extend benchmark comparison matrices of studies in a semi-automated way, possibly stumbling across unexpected findings of process settings that would not have been explored otherwise \cite{ma2015}.
+This makes it possible to easily extend benchmark comparison matrices of studies in a semi-automated way, possibly stumbling upon unexpected findings such as process settings that would not have been explored otherwise \cite{ma2015}.
 
@@
 ML methods in combination with remote sensing data are used in many environmental fields such as vegetation cover analysis or forest carbon storage mapping \cite{mascaro2014, urban2018}.
-The ability of predicting into unknown space qualifies ML algorithms as a helpful tool for such environmental analyses.
+The ability to predict these environmental variables in unknown regions qualifies ML algorithms as a helpful tool for such environmental analyses.
 One aspect of this research field is to enhance the understanding of biotic and abiotic stress triggers, for example by analyzing tree defoliation \cite{hawrylo2018}.
 Defoliation is known to be a proxy for pathogen and insect damage \cite{pollastrini2016}.
-While common symptoms are observable across species, some effects and their degree are species-specific \cite{gottardini2020}.
-Defoliation has also been shown to increase predisposition of tree death from secondary biotic factors of up to ten years after the actual defoliation event \cite{oliva2016}.
-Other approaches for analyzing forest health include change detection \cite{zhang2016} or describing the current health status of forests on a stand level \cite{townsend2012}.
+While common symptoms are observable across species, some effects and their degree of severity are species-specific \cite{gottardini2020}.
+Defoliation has also been shown to increase predisposition of tree death from secondary biotic factors up to ten years after the actual defoliation event \cite{oliva2016}.
+Other approaches for analyzing forest health include change detection \cite{zhang2016}, or describing the current health status of forests on a stand level \cite{townsend2012}.
 
@@
 Vegetation indices have shown the potential to provide valuable information when analyzing forest health \cite{jiang2014, adamczyk2015}.
-Most vegetation indices were developed with the aim of being sensitive to changes of specific wavelength regions, serving as a proxy for underlying plant processes.
-However, indices developed/applied for different purposes than the one to be analyzed (i.e.
-defoliation at pine trees) may help to explain complex underlying relationships which are not obvious at first.
+Most vegetation indices were developed with the aim of being sensitive to changes in specific wavelength regions, serving proxies for underlying plant physiological processes.
+In some cases, indices developed for different purposes than the one to be analyzed (e.g., defoliation of pine trees) may help to explain complex underlying relationships that are not obvious at first.
 This emphasizes the need to extract as much information as possible from the available input data to generate promising features which can help to understand the modeled relationship \cite{thenkabail2018}.
-A less known index type which can be derived from spectral information is the \ac{NRI}.
-In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics but instead makes use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands \cite{thenkabail2000}.
+A less known index type that can be derived from spectral information is the \ac{NRI}.
+In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics; instead, they make use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands \cite{thenkabail2000}.
 When working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.
@@
 % mention why FS is used / important
-Even though \ac{ML} algorithms are capable of handling highly correlated input variables, model fitting becomes computationally more demanding, and model interpretation more complex.
-Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplify model interpretability and possibly enhance predictive performance \cite{cai2018}.
-Instead of using wrapper feature selection methods, which add a substantial overhead to a nested model optimization approach, especially for datasets with many features, this study focuses on the use of (ensemble) filter methods which can be directly integrated into the hyperparameter optimization step during model creation.
+Even though \ac{ML} algorithms are capable of handling highly correlated input variables, model fitting becomes computationally more demanding, and model interpretation more challenging.
+Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplifing model interpretability, and possibly enhancing predictive performance \cite{cai2018}.
+Instead of using wrapper feature selection methods, which add a substantial overhead to a nested model optimization approach, especially for datasets with many features, this study focuses on (ensemble) filter methods, which can be directly integrated into the hyperparameter optimization step during model construction.
 
 % clarify the focus of this study
-Overall, this study aims to show how high-dimensional datasets can be handled effectively with ML methods while still being able to interpret the fitted models.
+Overall, this study aims to show how high-dimensional datasets can be handled effectively with ML methods while still allowing the interpretation of the fitted models.
 The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
-Yet, this study focuses on investigating the effects of filter methods and feature set types on predictive performance rather than interpreting individual feature effects.
+Yet, this study focuses mainly on investigating the effects of filter methods and feature set types on predictive performance rather than interpreting individual feature effects.
 
@@
 
-	\item Do different (environmental) feature sets show differences in performance when modeling defoliation of trees?
+	\item How do different feature selection methods influence the predictive performance of ML models of the defoliation of trees?
+
+	\item Do different (environmental) feature sets show differences in performance?
 
@@
 
-	\item How do different feature selection methods influence the predictive performance of the models?
-
-	\item Which features are most important and how can these be interpreted in an environmental context?
+	\item Which features are most important and how can these be interpreted in this context?
 
@@
 % Explain what is new in this study and link to methodology
-In recent years, various studies which have utilized hyperspectral data to analyze pest/fungi infections on trees were published:
+In recent years, various studies that have used hyperspectral data to analyze pest/fungi infections on trees have been published:
 Pinto et al.
 \citep{pinto2020} successfully used hyperspectral imagery to characterize pest infections on peanut leaves using random forest while Yu et.
-al \cite{yu2021} aimed to detect Pine wilt disease in pine plots in China using vegetation indices derived from hyperspectral data.
+al \cite{yu2021} aimed to detect pine wilt disease in pine plots in China using vegetation indices derived from hyperspectral data.
 Other studies which applied hyperspectral data for forest health monitoring are \cite{lin2014,kayet2019,dash2017}.

### 2021-11-14 docs/00-manuscripts/mdpi/manuscript.tex
revised data and methods section

@@
 Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata D. Don}) plantations in northern Spain.
-The trees in the plots suffer from infections of pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P.
+The trees in the plots suffer from infections with pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P.
 Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.} and \textit{Dothisthroma septosporum (Dorogin) M.
-Morelet} causing (among others) needle blight, pitch canker and root diseases \cite{mesanza2016, iturritxa2017}.
+Morelet} causing (among others) needle blight, pitch canker, and root diseases \cite{mesanza2016, iturritxa2017}.
 The first two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
-In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0--100 (in \%) (\autoref{fig:defol-distr}).
+In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection by experts.
+\textit{Defoliation} in percent was used as the response variable (\autoref{fig:defol-distr}).
 
 It is assumed that the fungi infect the trees through open wounds, possibly caused by previous hail damage \cite{iturritxa2014}.
-The dieback of these trees, which are mainly used as timber, causes high economic damages \cite{ganley2009}.
+The dieback of these trees, which are mainly used as timber, causes substantial economic losses \cite{ganley2009}.
 
@@
 		\includegraphics[width=0.7\textwidth] {defoliation-distribution-plot-1.pdf}
-		\caption{Response variable ``defoliation at trees'' for plots~Laukiz1, Laukiz2, Luiando and Oiartzun.
-\texttt{n} corresponds to the total number of trees in the plot, $\bar{x}$ refers to the mean defoliation, respectively.
+		\caption{Response variable ``defoliation of trees'' for plots~Laukiz1, Laukiz2, Luiando and Oiartzun.
+\texttt{n} is the total number of trees in each plot, and $\bar{x}$ the mean defoliation.
 Values for Laukiz1, Luiando and Oiartzun were observed in 5\% intervals; for Laukiz2 defoliation was observed at multiple heights and then averaged, leading to smaller defoliation differences than 5\%.}\label{fig:defol-distr}
@@
 The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
-Oiartzun consists of most tree observations (n = 559) while Laukiz2 shows the largest area size (1.44~ha).
+Oiartzun has the largest number of observations ($n = 559$ trees) while Laukiz2 is the largest in area (1.44~ha).
 All plots besides Luiando are located within 100~km from the coast (\autoref{fig:study_area}).
-A total of 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
+A total of 1808 observations are available (Laukiz1: 559, Laukiz2: 451, Luiando: 301, Oiartzun: 497).
 Field surveys were conducted in September 2016 by experienced forest pathologists.
-Defoliation was measured via visual inspection using 5\% intervals with the help of a dedicated score card.
-For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% intervals of the other three plots (e.g., 8.33\%).
-Estimating the human observer error of such surveys when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
-Even though no estimation error was recorded in this study, \cite{maclean1982} estimated human observer errors when assessing defoliation to range between 7\% - 18\%.
+Defoliation was measured in 5\% steps through visual inspection with the help of a score card.
+For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, resulting in values that are not multiples of 5\% (e.g., 8.33\%).
+The magnitude of human observer errors in such surveys, including the present one, is not precisely known and has being discussed for many years \cite{innes1993}.
+\cite{maclean1982} estimated human observer errors in defoliation surveys to range between 7\% and 18\%.
 
@@
 		\caption{Study area maps showing information about location, size and spatial distribution of trees for all plots (Laukiz1, Laukiz2, Luiando, Oiartzun).
-The background maps used should give a visual impression of the individual plot area but do not necessarily represent the plot state during data acquisition.}\label{fig:study_area}
+The background maps give a visual impression of the individual plot area but do not necessarily represent the plot's state during data acquisition.}\label{fig:study_area}
 	\end{center}
@@
 
-The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5, 2016.
-Images were taken by an AISA EAGLE-II sensor.
+The airborne hyperspectral data was acquired by an AISA EAGLE-II sensor during two flight campaigns on September 28th and October 5, 2016 at noon.
 All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.
@@
 
-To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated.
+To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package \texttt{hsdar} (89 in total) as well as all possible \ac{NRI} combinations were calculated.
 NRIs follow the \ac{OMNBR} concept of data-driven information extraction from narrow-band indices of hyperspectral data \cite{thenkabail2000,thenkabail2018}.
 While various index formulations such as band ratios or normalized ratios are available, indices involving the same bands are strongly correlated.
-Since the widely\ used NDVI index belongs to the family of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following normalized difference index (NDI) formula to combine all pairs of reflectances:
+Since the widely\ used NDVI index belongs to the group of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following normalized difference index (NDI) formula to combine all pairs of reflectances:
 
@@
 
-To account for geometric offsets within the hyperspectral data, which were reported to be potentially up to one meter from \ac{ICGC}, a buffer of one meter around the centroid of each tree was used when extracting the reflectance values.
+To account for geometric offsets within the hyperspectral data, which were reported by \ac{ICGC} to be potentially up to one meter, a buffer of one meter radius around the centroid of each tree was used when extracting the reflectance values.
 A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
-The pixel values of such were averaged and formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
+The pixel values within a buffer zone were averaged and formed the final reflectance value of a single tree and were used as the base information to derive all additional feature sets.
 In total, \(\frac{121*122}{2} = 7381\) NRIs were calculated.
@@
 
-To reduce the feature space of a dataset, conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge) or \ac{PCA} \cite{bommert2020, das2001, guyon2003, jolliffe2016}.
-In contrast to wrapper methods, filters can be added to the hyperparameter optimization step and have a lower computational footprint.
-In addition, filters are less known than wrapper methods and in recent years ensemble filters were introduced which have shown promising results compared to single filter algorithms \cite{drotar2017}.
+To reduce the feature space of a dataset, several conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge), and \ac{PCA} \cite{bommert2020, das2001, guyon2003, jolliffe2016}.
+In contrast to wrapper methods, filters have a much lower computational cost, and their tuning can be added to the hyperparameter optimization step.
+In addition, filters are less known than wrapper methods and in recent years, ensemble filters were introduced which have shown promising results compared to single filter algorithms \cite{drotar2017}.
 These two points mainly led to the decision to focus on filter methods for this work and evaluate their effectiveness on highly\ correlated, high-dimensional datasets.
@@
 Filters only rank features, they do not decide which covariates to drop or keep \cite{drotar2015}.
-The selection which features to keep for model fitting is usually done within the optimization phase during model fitting, along with the hyperparameter tuning.
-Essentially, the number of covariates in the model is treated as an additional hyperparameter of the model.
-The idea is to optimize the number of ranked features to the point at which the model achieves the best performance.
+The decision which features to keep for model fitting can be integrated into the optimization phase during model fitting, along with hyperparameter tuning.
+Thus, the number of top-ranked features to be included in the model is treated as an additional hyperparameter of the model.
+This hyperparameter is tuned to optimize the model's performance.
 
 % Ensemble filter methods
-Besides the concept of choosing a specific filter method to rank variables, studies showed that combining several filters using statistical operations such as 'minimum' or 'mean' can possibly enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017}.
+Beyond the use of individual filter methods to rank and select features, recent studies have shown that combining several filters by using statistical operations such as 'minimum' or 'mean' may enhance the predictive performance of the resulting models, especially when applied to multiple datasets \cite{abeel2010, drotar2017}.
 This approach is referred to as 'ensemble filtering' \cite{dietterich2000}.
-Ensemble filters align with the recent rise of the 'ensemble' approach in ML which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
-In this work the 'Borda' ensemble filter was used \cite{drotar2017}.
-Its final feature order is determined by the sum of all single filters ranks.
+Ensemble filters align with the recent rise of the ensemble approach in ML, which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance \cite{polikar2012, feurer2015, bolon-canedo2019}.
+In this work the Borda ensemble filter was used \cite{drotar2017}.
+Its overall feature order is determined by the sum of filter ranks of all individual filters in the ensemble.
 
 % Ensuring a fair weighting in the ensemble
-Filter methods can be grouped into groups which are formed out of three binary classes: multivariate or univariate feature use, correlation or entropy-based importance weighting and linear and non-linear filter methodology.
-Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the final ranking result will be biased.
+Filter methods can be categorized based on three binary criteria: multivariate or univariate feature use, correlation or entropy-based importance weighting, and linear and non-linear filter methodology.
+Care needs to be taken to not weigh certain classes more than others in the ensemble as otherwise the ranking will be biased.
 In this study this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
-If filter pairs showed a correlation of 0.9 or higher, only one of the two was included into the ensemble filter, selected at random.
-This ensured that the ensemble filter composition was not biased towards a certain group of filter methods.
+If filter pairs showed a correlation of 0.9 or higher, only one of the two was included in the ensemble filter, selecting it at random.
 
@@
 
-The filter 'Information Gain' is only defined for nominal response variables:
+The filter 'Information Gain' in its original form is only defined for nominal response variables:
 
@@
 
-where \(H\) is the conditional entropy of the response variable (class or Y) or the feature (attribute or X), respectively.
+where \(H\) is the conditional entropy of the response variable (class or Y) and the feature (attribute or X), respectively.
 $H(X)$ is Shannon's Entropy \cite{shannon1948} for a variable X and $H(X, Y)$ is a joint Shannon's Entropy for a variable X with a condition to Y.
@@
 
-where $b$ is the base of the logarithm used (with $b$ commonly being set to 2).
-
-In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins \texttt{\(n_{bin}\) = 10} and treated as a class variable.
+where $b$ is the base of the logarithm used, most commonly 2.
+
+In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins \texttt{\(n_{bin}\) = 10} and treated as a categorical variable.
 
@@
 
-RF and SVM are well-established algorithms and widely used in (environmental) remote sensing.
+RF and SVM are well-established algorithms and widely used in environmental remote sensing.
 Extreme Gradient Boosting (commonly abbreviated as XGBoost) has shown promising results in benchmarking studies in recent years.
-Penalized regression is a statistical modeling technique capable of dealing with highly\ correlated covariates by penalizing the coefficients of the model \cite{hastie2001}.
+Penalized regression is a statistical modeling technique capable of dealing with highly\ correlated covariates by penalizing the model coefficients \cite{hastie2001}.
 Common penalties are 'lasso' (L1) and 'ridge' (L2).
-Ridge regression does not remove variables from the model (penalization to zero) but it just shrinks them towards zero, keeping them in the model.
+Ridge regression does not remove variables from the model (penalization to zero), but it shrinks them towards zero, keeping them in the model.
 A featureless learner was included for a baseline comparison.
@@
 Peña et al. 2017 \cite{pena2017} is an exemplary study which used this approach in a spectro-temporal setting.
-Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method steady makes it possible to draw conclusions on their individual impact.
+Benchmarking learners on these feature sets while keeping all other variables such as model type, tuning strategy and partitioning method fixed makes it possible to draw conclusions on their individual impact.
 Each feature set has distinct capabilities that differentiate it from the others.
-This can have both a positive and negative effect on the resulting performance, which is one item this study aims to explore.
+This can have both a positive and negative effect on the resulting performance, which is one question this study aims to explore.
 For example, feature set VI misses certain parts of the spectral range as the chosen indices only use specific spectral bands.
 Feature set NRI will introduce highly\ correlated features, for which some algorithms may be more suitable than others.
-In addition to these feature sets, the following combinations of feature sets were also compared:
+
+In addition to these individual feature sets, the following combinations of feature sets were also compared:
 
@@
 This preprocessing step reduced the number of covariates for feature set~VI to 86 (from 89).
-This decision was made to prevent issues for the subsequent tuning, filtering and model fitting steps which might occur when passing features with a pairwise correlation of (almost) one.
+This decision was made to prevent numerical issues that may occur in the subsequent tuning, filtering and model fitting steps when offering features with a pairwise correlation of (almost) one.
 The remaining features were then used as input for the filter-based feature selection within the CV.
 
-\subsubsection{Hyperparameter Optimization}
+\subsubsection{Hyperparameter optimization}
 
@@
 After these \textit{n} settings have been evaluated, one new setting, which is going to be evaluated next, is proposed by a fitted surrogate model (by default, a kriging method).
-This strategy continues until a termination criterion, defined by the user, is reached \cite{hutter2011, jones1998}.
-
-In this work, an initial design of 30 randomly composed hyperparameter settings in combination with a termination criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
+This strategy continues until a user-defined stopping criterion is satisfied \cite{hutter2011, jones1998}.
+
+In this work, an initial design of 30 randomly drawn hyperparameter settings in combination with a stopping criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
 The advantage of this tuning approach is a substantial reduction of the tuning budget that is required to find a setting close to the global optimization minimum.
@@
 It was thus tuned on a logarithmic scale by varying $t$ between 0 (i.e., $m_\textrm{try} = 1$) and 0.5 (i.e., $m_\textrm{try}=\sqrt{p_\textrm{sel}}$).
-This was necessary to ensure that \texttt{\(m_{try}\)} was not chosen higher than the available number of features that were left after optimizing the feature percentage during tuning.
+This was necessary to ensure that \texttt{\(m_{try}\)} did not exceed the number of features available after optimizing the feature percentage during tuning.
 
@@
 
-A spatial nested cross-validation on the plot level was chosen to reduce the influence of spatial autocorrelation as much as possible \cite{schratz2019, sperrorest}.
+A spatial nested cross-validation on the plot level was chosen to account for spatial autocorrelation within the plots and assess model transferability to different plots \cite{schratz2019, sperrorest}.
 The \ac{RMSE} was chosen as the error measure.
-Each plot served as one fold within the cross-validation setting, resulting in four iterations in total.
-For the inner level (hyperparameter tuning), \(k - 1\) folds were used with \(k\) being the number of plots.
+Each plot served as one cross-validation fold, resulting in four iterations in total.
+The inner level of cross-validation for hyperparameter tuning also used plot-level cross-validation.
 
@@
 
-Estimating feature importance for datasets with highly correlated features is a complicated task for which many different approaches, model-specific and agnostic, exist \cite{friedman2001, hastie2001, greenwell2018}.
-The correlation between covariates makes it challenging to calculate an unbiased estimate for single features \cite{molnar2019}.
-Methods like \ac{PDP} or permutation-based approaches may produce unreliable estimates in such scenarios because unrealistic situations between covariates are created \cite{molnar2019}.
-The development of robust methods which enable an unbiased estimation of feature importance for highly correlated variables are subject to current research \cite{brenning2021}.
-
-In this work permutation-based feature importance was calculated to estimate feature importance/effects \cite{apley2019}.
+Estimating feature importance for datasets with highly correlated features is a challenging task for which numerous model-specific and model-agnostic approaches exist \cite{friedman2001, hastie2001, greenwell2018}.
+The strong correlations among features make it challenging to calculate an unbiased estimate for single features \cite{molnar2019}.
+Methods like \ac{PDP} or permutation-based approaches may produce unreliable estimates in such scenarios because unrealistic combinations of feature values are created \cite{molnar2019}.
+The development of robust methods that enable an unbiased estimation of feature importance for highly correlated variables are subject to current research \cite{brenning2021}.
+
+In this work, permutation-based feature importance was calculated to estimate feature importance or effects \cite{apley2019}.
 With the limitations in mind when applied to correlated features, the aim was to get a general overview of the feature importance of the hyperspectral bands while trying to avoid an over-interpretation of results.
@@
 
-\subsection{Linking feature importance to wavelength regions}
-
-For environmental interpretation purposes, the ten most important indices of the best performing models of feature sets HR and VI were linked to the spectral regions of the hyperspectral data.
+%\subsection{Linking feature importance to wavelength regions}
+
+To facilitate interpretation, the ten most important indices of the best performing models using feature sets HR and VI were linked to the spectral regions of the hyperspectral data that went into their calculation.
 The aim was to visualize the most important features along the spectral curve of the plots to better understand which spectral regions were most important for the model.
@@
 A complete list of all R packages used in this study can be found in the linked repositories mentioned in the next paragraph.
-Due to space limitations, only the selected packages with high impact on this work will be explicitly cited.
+Due to space limitations, only selected packages with high impact on this work will be explicitly cited.
 
@@
 Package mlr \cite{mlr} was used for all modeling related steps.
-drake \cite{drake} was used for structuring the work and reproducibility.
+drake \cite{drake} was used for structuring the work and for reproducibility.
 This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}).

### 2021-11-15 docs/00-manuscripts/mdpi/manuscript.tex

@@
 Regression with ridge (L2) and lasso (L1) penalties showed their best performances on the NRI feature set (\autoref{tab:best-learner-perf}).
-Combining feature sets for lasso and ridge did not help to increase performance and while there was no subatantial difference for lasso, ridge scored around two percentage points on such.
+Combining feature sets for lasso and ridge did not help to increase performance, and while there was no substantial difference for lasso, the performance of ridge improved by around two percentage points.
 XGBoost showed very poor performances for some feature sets and fills the last\ ten places of the ranking (\autoref{tab:perf-worst-10}).
@@
 % - never the best
-The Borda filter was not able to score the best performance in any learner/filter setting (\autoref{fig:filter-effects-borda}).
+The Borda ensemble filter was not able to score the best performance in any learner/filter setting (\autoref{fig:filter-effects-borda}).
 For RF and XGBoost, it most often ranked within the better half among all filters of a specific task.
@@
 The number of features selected during model optimization strongly varied across learners and plots.
-RF selected the least features of all three learners, and with the expection of Oiartzun, only one or two variables got selected.
-SVM used 210 features or more in all instances and varied between 16\% (Laukiz1) - 81\% (Oiartzun) (\autoref{tab:tune-perc-sel-features}).
-XGBoost also favored using several hundred features with the exeption of Laukiz2 for which only 14 (0.96\%) were selected.
+RF selected the least features of all three learners, and with the exception of Oiartzun, only one or two variables were selected.
+SVM used 210 features or more in all instances and selected between 16\% (Laukiz1) and 81\% (Oiartzun) of the features (\autoref{tab:tune-perc-sel-features}).
+XGBoost also favored using several hundred features with the exception of Laukiz2 for which only 14 (0.96\%) were selected.
 
@@
 
-The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B65) and 1.93 p.p. (VI, Vogelmann2) (\autoref{fig:fi-permut-vi-hr}).
+The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B65) and 1.93 p.p. (VI, Vogelmann2) when permuted (\autoref{fig:fi-permut-vi-hr}).
 For the HR dataset, most (i.e., six out of ten) relevant features clustered around the infrared region (920~nm - 1000~nm), while for VI, eight out of ten concentrate within the wavelength range of 700~nm - 750~nm (the so\-/called \enquote{red edge}).

### 2021-11-15 docs/00-manuscripts/mdpi/manuscript.tex
revising discussion (1)

@@
 
-The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE~27.915 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
+The best overall performance achieved in this study (SVM with the \enquote{Info Gain} filter, RMSE~27.915 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
 Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
 However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
-Due to the apparent presence of model overfitting in this study, it can be postulated that more training data representing a greater variety of situations is needed.
+Due to the apparent presence of model overfitting in this study, it is suggested that more training data representing a greater variety of situations is needed.
 A model can only make robust predictions if it has learned relationships across the whole range of the response.
-Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
-However, when inspecting the fold\-/level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
+Hence, care should be taken when predicting on the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
+However, when inspecting the fold\-/level performances, it can be concluded that the models performed reasonably well predicting defoliation greater than 50\%, but they failed for lower levels.
 This applied to all learners of this study.
-The overall performance of all learners achieved in this study can be classified as ``poor'' given that only the SVM learner was able to outperform the featureless learner (\autoref{tab:best-learner-perf}).
-
-In addition, data quality issues (\autoref{subsec:data-quality}) might have an influence on model performances.
-These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g., soil reflectance) and the possible positional offset of measured GPS coordinates of trees.
+In this study, the overall performance across all learners can be classified as ``poor'' given that only the SVM learner was able to substantially outperform the featureless learner (\autoref{tab:best-learner-perf}).
+It is worth noting that data quality issues may have affecting model performances, as discussed below in detail  (\autoref{subsec:data-quality}).
 
@@
 An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (\autoref{tab:best-learner-perf}).
-These cluster around a performance of 31 p.p. while SVM is able to score about three p.p. better than all other methods.
+These cluster around a performance of 31 p.p. while SVM scored about three p.p. better than all other methods.
 However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).
@@
 
-One objective of this study was whether expert-based or data-driven feature engineering has a positive influence on model performance.
-With respect to \autoref{fig:perf-result}, no overall positive or negative trend was found for all models that related to specific feature sets.
+One objective of this study was to determine whether expert-based and data-driven feature engineering have a positive influence on model performance.
+With respect to \autoref{fig:perf-result}, no overall positive or negative pattern related to specific feature sets was found that would be valid across all models.
 The performance of RF and XGBoost on the VI feature set was around four to six p.p. lower than on others.
-One reason could be the lack of coverage in the wavelength area between 810~nm and 1000~nm (\autoref{fig:fi-permut-vi-hr}).
+One reason could be the lack of coverage in the wavelength range between 810~nm and 1000~nm (\autoref{fig:fi-permut-vi-hr}).
 In addition, for all learners but SVM a better performance was observed when NRI indices were included in the feature set (i.e., NRI-VI, HR-NRI, HR-NRI-VI).
@@
 
-Arbitrary seeming values of selected feature percentages in \autoref{tab:tune-perc-sel-features} emerge from the specific hyperparameter setting which got elected as the best one during the tuning step.
+There was also no clear pattern in the percentage of features selected based on hyperparameter tuning (\autoref{tab:tune-perc-sel-features}).
 The specific performance value to benchmark against during tuning is defined by the internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features) for the hyperparameter that controls the number of features.
+% Nevertheless, considering SVM's relative success, the use of at least a few hundred features from the combined feature set appears to be beneficial, or at least not harmful when the model's built-in regularization is capable of dealing with the resulting high-dimensional situation.
 
@@
 Realizing early during hyperparameter optimization that only few features are needed to reach adequate performances can reduce the overall computational runtime substantially.
-Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially on models which make use of hyperparameters that depend on the available number of features, such as RF $m_\textrm{try}$.
+Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially of models making use of hyperparameters that depend on the available number of features, such as RF with $m_\textrm{try}$.
 
 Ultimately, the results of \autoref{tab:tune-perc-sel-features} should be taken with care as they rely on single model-filter combinations and are subject to random variation.
-More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criterion optimization problem.
+More in-depth research is needed to investigate the effect of filters on other criteria than performance (such as runtime), leading to a multi-criteria optimization problem.
 

### 2021-11-15 docs/00-manuscripts/mdpi/manuscript.tex
revised discussion (part 2/2)

@@
 for RF and XGBoost) was improved by specific filters, some models achieved a poorer performance with filters than without them (\autoref{fig:filter-effects-no-filter}).
-No pattern with respect to lower scores related to a specific filter method could be found.
-Hence, it is recommended to test multiple filters in a study if filters are going to be used.
-While filters can improve the performance of models, they might be more interesting in other aspects than performance: reducing variables can reduce computational efforts in high-dimensional scenarios and might enhance the interpretability of models.
+There was no pattern of specific filters consistently resulting in better scores.
+Hence, it is recommended to test multiple filters in a study if it is intended to use filters.
+While filters can improve the performance of models, they may be more appealing based on other aspects than performance: Reducing variables can reduce computational efforts in high-dimensional scenarios and may enhance the interpretability of models.
 Filters are a lot cheaper to compute than wrapper methods, and the final feature subset selection can be integrated as an additional hyperparameter into the model optimization stage.
@@
 % - Ensemble FS shows no advantage to simple methods -> always one simple filter which is better. Ensemble filters are good across many different datasets
-The models which used the Borda ensemble method in this study did not score better on average than models which used a single filter or no filter at all.
+Models that used the Borda ensemble method in this study did not perform better on average than models which used a single filter or no filter at all.
 Ensemble methods have higher stability and robustness than single ones and have shown promising results in \cite{drotar2017}.
-Hence, their main advantage is stable performances across datasets with varying characteristics.
+Hence, their expected main advantage is stable performances across datasets with varying characteristics.
 Single filter methods might yield better model performances on certain datasets but fail on others.
 The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation why in all cases a single filter outperformed the ensemble filter.
-However, studies which used ensemble filters are still rare and usually these are not compared against single filters \cite{ghosh2019}.
-In summary, Borda does\ not\ perform better than a randomly selected filter method in this study.
+However, studies which used ensemble filters are still rare, and these are usually not compared against single filters \cite{ghosh2019}.
+In summary, in this study Borda did\ not\ perform better than a randomly selected filter method.
 More case studies applying ensemble filter methods are needed to verify this finding.
@@
 % - PCA shows similar performance but probably requires tuning of main components -> check runtime!
-PCA, acting as a filter method, more often showed less optimal results, especially for algorithms RF and XGBoost.
+PCA, acting as a filter method, more often showed less than optimal results, especially for algorithms RF and XGBoost.
 Especially XGBoost had substantial problems when using PCA as a filter method and scored four of the six worst results (\autoref{tab:perf-worst-10}).
@@
 However, four out of ten of the most important features of dataset HR are located between 920~nm and 1000~nm.
-Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral area---especially for plot Oiartzun---which might explain why these features were considered important by the model.
+Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral range---especially for plot Oiartzun---, which might explain why these features were considered important by the model.
 
@@
 Finding a suitable approach to extract the remote sensing information from each tree was a complex process.
-Due to the reported geometric offset of up to one meter within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different, possibly non-tree, pixel was reasonably high.
-It was concluded that using a buffer of one meter can be a good compromise between the inclusion of information from too many surrounding trees, mapping a single tree crown accordingly accounting for a possible geometric offset.
-The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e., $<= 1m$).
+Due to the reported geometric offset of up to one meter within the hyperspectral data, the risk of assigning a value to an observation that would actually refer to a different, possibly non-tree, pixel was reasonably high.
+It was concluded that using a buffer radius of one meter can be a good compromise between the inclusion of information from too many surrounding trees and an under-coverage of the tree crown.
+With the chosen radius, we are confident that we were able to map individual tree crowns while accounting for a possible geometric offset.
+% The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e., $<= 1m$).
 This results in all cases in four contributing pixels (= four square meters) for the extraction of hyperspectral information for a given tree.
 Even though no results showing the influence of different buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
-Besides using a buffer to extract the hyperspectral information, segmentation could have been considered.
-However, this method would have required more effort for no clear added value in our view and would have moved the focus of this manuscript more to data preprocessing and away from the focus on filter-based feature selection methods.
-
-Another point worth discussing is that the exact number of contributing pixels to the final index value of an observation cannot be determined precisely: it depends on the location of the tree within the pixel grid.
-According to the extract function of the raster package, a pixel is included if its centroid (and not just any part of the grid cell) falls inside the buffer.
-As the buffer is circular, the total number of contributing pixels of each tree depends on the exact location of a tree within the pixel grid.
-If a tree observation is located on the border of the plot, some directions of the buffer will contain no values (because the image coverage was cropped to the borders of the plot) and the subsequent index value will be calculated with fewer pixels than if the tree observation is located in the middle of the plot.
-However, in most cases each tree information should be composed out of four hyperspectral pixels.
+Instead of using a buffer to extract the hyperspectral information, segmentation could have been considered.
+However, this method would have required more effort for no clear added value in our view and would have moved the focus of this manuscript more to data preprocessing and away from feature selection methods.
+
+Trees located within grid cells on the border of a plot are a notable exception where the exact number of pixels contributing to the observation's feature value may be reduced since the image was cropped to the plot's extent.
+Cropping was applied to avoid the accidental inclusion of background data such as forest roads.
+However, this effect was deemed to be of negligible importance.
 
 The available hyperspectral data covered a wavelength between 400~nm and 1000~nm.
-Hence, the wavelength range of the shortwave infrared (SWIR) region is not covered in this study.
-Given that this wavelength range is often used in forest health studies \cite{hais2019}, e.g., when calculating the \ac{NDMI} index \cite{gao1996}, this marks a clear limitation of the dataset at hand.
-
-The dataset consists of in-situ data collected within September 2016 matched against remote sensing data acquired at the end of September 2016.
+Hence, the spectral range of the shortwave infrared (SWIR) region is not covered in this study.
+Given that this range is often used in forest health studies \cite{hais2019}, e.g., when calculating the \ac{NDMI} index \cite{gao1996}, this marks a clear limitation of the dataset at hand.
+
+The dataset consists of in-situ data collected during September 2016, which was matched against remote sensing data acquired at the end of September 2016.
 A multi-temporal dataset consisting of in-situ data from different phenology stages would possibly improve the achieved model performances.
-However, this would also require matching hyperspectral data of these additional timestamps.
-
-The R package hsdar was used for the calculation of vegetation indices \cite{lehnert2016}.
+However, this would also require the costly acquisition of hyperspectral data of these additional timestamps.
+
+The R package \texttt{hsdar} was used for the calculation of vegetation indices \cite{lehnert2016}.
 All indices that could be calculated with the given spectral range of the data (400~nm - 1000~nm) were used.
-This means even though \autoref{tab:vegindices} lists all available indices of the package, not all listed indices were used in this study.
-Even though this selection included a large number of indices, some possibly helpful indices might have been missed due to the restriction of the hyperspectral data.
+This means even though \autoref{tab:vegindices} lists all indices available in the package, not all listed indices were used in this study.
+Although this selection included a large number of indices, some possibly helpful indices might have been missed due to the restriction of the hyperspectral data.
 
@@
 Yet the general approach of utilizing hyperspectral data to classify the health status of trees partly led to promising results:
-For example, due to the narrow bandwidth of the hyperspectral sensor small parts of the spectra (mainly in the infrared region) were of higher importance to the models (e.g., see \autoref{fig:fi-permut-vi-hr}), meaning they helped the models to distinguish between low and high tree defoliation.
+For example, due to the narrow bandwidth of the hyperspectral sensor, small parts of the spectrum (mainly in the infrared region) were of higher importance to the models (e.g., see \autoref{fig:fi-permut-vi-hr}), meaning that they helped the models to distinguish between low and high tree defoliation.
 If spatial offset errors of the image data and possible background noise can be reduced (possibly by making use of image segmentation), we believe that model performances could be substantially enhanced.
-Such improved models, starting around an RMSE of 20\% and less, should be able to provide added value to support the long\-/term monitoring of forest health and early detection of fungi affected tree plots.
-Though overall the usage of defoliation as a proxy to forest health should be critically questioned as it comes with various practical issues, starting from potential offsets during data collection, varying leaf density due to tree age and differing effects between tree species, to just name a few.
+Such improved models, starting around an RMSE of 20\% and less, should be able to provide added value to support the long\-/term monitoring of forest health and early detection of fungi-affected tree plots.
+Nevertheless, overall the use of defoliation as a proxy for forest health should be critically reconsidered as it comes with various practical issues, starting from potential offsets during data collection, varying leaf density due to tree age and differing effects between tree species, to just name a few.
 
@@
 \cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
-However, both did not use spatial \ac{CV} and \cite{kalin2019} no \ac{FS}.
+However, neither of them used spatial \ac{CV} for model assessment, and \cite{kalin2019} did not perform \ac{FS}.
 \cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
 Study results indicated that spectral features were found to be most helpful for the model.
-Incorporating such (both spectral and structural) could be a possible enhancement for future works.
-No studies were found modeling defoliation caused by \textit{Diplodia sapinea (Fr.) Fuckel} with remote sensing data and most studies focused on describing the tree conditions based on local sampling \cite{hlebarska2018, kaya2019}.
+Incorporating such features (both spectral and structural) could be a possible enhancement for future works.
+No studies were found modeling defoliation caused by \textit{Diplodia sapinea (Fr.) Fuckel} with remote sensing data, and most studies focused on describing the tree conditions based on local sampling \cite{hlebarska2018, kaya2019}.
 
@@
 The field of (hyperspectral) remote sensing has a strong focus on using RF for modeling in recent years \cite{belgiu2016}.
-However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
+However, in high-dimensional scenarios, tuning the parameter \texttt{\(m_{try}\)} becomes computationally expensive.
 To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
 In general, applying feature selection methods on hyperspectral datasets has shown to be effective, regardless of the method used \cite{pal2010, keller2016}.
-However, no study that made explicit use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
-Potential reasons for this absence could be an easier programmatic access to wrapper methods, and a higher general awareness of such compared to filter methods.
+However, no studies were found that made explicit use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
+Potential reasons for this absence could be an easier programmatic access to wrapper methods, and a higher general awareness of these compared to filter methods.
 Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.
@@
 \cite{shendryk2016, ludwig2019} are more similar in their methodology but focus on a different response variable (woody cover).
-\cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees for eucalyptus forests.
+\cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees in eucalyptus forests.
 A grid search was used for hyperparameter tuning and \ac{FFS} for variable selection.
@@
 
-In summary, no studies which used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to tree health could be found.
+In summary, no studies could be found that used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to tree health.
 This might relate to the fact that most environmental datasets are not high-dimensional.
 In fact, many studies use fewer than ten features and issues related to correlations are often solved manually instead of relying on an automated approach.
-This approach might suffer from subjectivity and may limit the reproducibility of results.
+These manual approaches might suffer from subjectivity and may limit the reproducibility of results.
 

### 2021-12-06 analysis/eda.Rmd
[pre-commit.ci] auto fixes from pre-commit.com hooks
for more information, see https://pre-commit.ci

@@
   bands_data, coords_vi_nri_clean,
-
   filter_values,
   filter_info_gain_nbins,
-
   trees_with_bands,

### 2021-12-06 analysis/report-defoliation.Rmd
[pre-commit.ci] auto fixes from pre-commit.com hooks
for more information, see https://pre-commit.ci

@@
   defoliation_df_relative,
-
   tune_xgboost_plan_project
@@
 ```{r echo = FALSE}
-#system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/")
+# system("rsync -rlptDivzog --chown=*:www-data --chmod=g+r,o+r /home/patrick/papers/2019-feature-selection/docs/figure/report-defoliation.Rmd/defol* patrick@jupiter.geogr.uni-jena.de:/home/www/life-healthy-forest/action-B1-spatial-mapping/defoliation-maps/")
 ```

### 2021-12-06 R/04-fun-sentinel-processing.R

@@ -20,7 +20,8 @@ get_records <- function(aoi, date, processing_level) {
   login_CopHub("be-marc", password = "ISQiwQDl")
 
   # Get records from Sentinel-2 in time range
-  records <- getSpatialData::get_records(time_range = date,
+  records <- getSpatialData::get_records(
+    time_range = date,
     products = "Sentinel-2", as_sf = FALSE
   )
 

### 2021-12-06 _drake.R

@@ -35,10 +35,8 @@ R.utils::sourceDirectory("code")
 plan_paper <- bind_plans(
   download_data_plan,
   hyperspectral_processing_plan,
-
   data_preprocessing_plan,
   tasks_plan,
-
   filter_eda_plan,
   param_sets_plan,
   learners_plan,
@@ -46,10 +44,8 @@ plan_paper <- bind_plans(
   resampling_plan,
   tune_ctrl_plan,
   tune_wrapper_plan,
-
   benchmark_plan,
   bm_aggregated_plan,
-
   feature_imp_plan,
   reports_plan_paper
 )
@@ -58,11 +54,11 @@ plan_paper <- bind_plans(
 
 # config for long running tasks
 drake_config(plan_paper,
-   targets = c(
-     # replace this with the target name that should be computed
+  targets = c(
+    # replace this with the target name that should be computed
     #  c("eda_wfr")
     "benchmark_no_models"
-   ),
+  ),
   verbose = 1,
   lazy_load = "eager",
   packages = NULL,

### 2021-12-06 code/020-hyperspectral-processing.R

@@ -6,7 +6,6 @@ hyperspectral_processing_plan <- drake_plan(
       id = name_id_paper
     )
   ),
-
   ndvi_rasters = target(
     map(data_hs_preprocessed_paper, ~ HyperSpecRaster(.x, wavelength_reduced)) %>%
       future_imap(~ vegindex(.x, "NDVI2",
@@ -14,12 +13,10 @@ hyperspectral_processing_plan <- drake_plan(
         bnames = "NDVI"
       ))
   ),
-
   veg_indices = target(
     map(data_hs_preprocessed_paper, ~ HyperSpecRaster(.x, wavelength_reduced)) %>%
       calc_veg_indices(indices)
   ),
-
   nri_indices = target(
     map(data_hs_preprocessed_paper, ~ HyperSpecRaster(.x, wavelength_reduced)) %>%
       calc_nri_indices(indices)
@@ -28,7 +25,6 @@ hyperspectral_processing_plan <- drake_plan(
   # corrected data from Jan 2020 -----------------------------------------------
 
   plot_names = c("laukiz1", "laukiz2", "luiando", "oiartzun"),
-
   trees_with_indices = target(
     list(
       extract_indices_to_plot(plot_names,
@@ -42,7 +38,6 @@ hyperspectral_processing_plan <- drake_plan(
       plot_names
     )
   ),
-
   trees_with_bands = target(
     list(
       extract_bands_to_plot(plot_names,

### 2021-12-06 code/040-data-preprocessing.R

@@ -46,7 +46,6 @@ data_preprocessing_plan <- drake_plan(
   coords_vi_nri_clean = target(
     as_tibble(rbindlist(data_clean_single_plots_coords[[1]]))
   ),
-
   data_bands_clean = target(
     as_tibble(rbindlist(data_clean_single_plots[[2]], fill = TRUE)) %>%
       Filter(function(x) !any(is.na(x)), .)
@@ -73,30 +72,24 @@ data_preprocessing_plan <- drake_plan(
   nri_data = target(
     split_into_feature_sets(data_trim_defoliation, "nri")
   ),
-
   vi_data = target(
     split_into_feature_sets(data_trim_defoliation, "vi")
   ),
-
   bands_data = target(
     split_into_feature_sets(data_trim_defoliation, "bands")
   ),
-
   nri_vi_data = target(
     cbind(nri_data, vi_data) %>%
       subset(select = which(!duplicated(names(.)))) # remove duplicate "defoliation" column
   ),
-
   hr_nri_data = target(
     cbind(bands_data, nri_data) %>%
       subset(select = which(!duplicated(names(.)))) # remove duplicate "defoliation" column
   ),
-
   hr_vi_data = target(
     cbind(bands_data, vi_data) %>%
       subset(select = which(!duplicated(names(.)))) # remove duplicate "defoliation" column
   ),
-
   hr_nri_vi_data = target(
     cbind(
       bands_data, nri_data,
@@ -121,7 +114,6 @@ data_preprocessing_plan <- drake_plan(
       nri_data
     }
   }),
-
   vi_data_trim_cor = target({
     inds <- caret::findCorrelation(cor(vi_data[, -1]),
       cutoff = 0.9999999999, exact = TRUE
@@ -133,7 +125,6 @@ data_preprocessing_plan <- drake_plan(
       vi_data
     }
   }),
-
   bands_data_trim_cor = target({
     inds <- caret::findCorrelation(cor(bands_data[, -1]),
       cutoff = 0.9999999999, exact = TRUE
@@ -145,7 +136,6 @@ data_preprocessing_plan <- drake_plan(
       bands_data
     }
   }),
-
   nri_vi_data_trim_cor = target({
     inds <- caret::findCorrelation(cor(nri_vi_data[, -1]),
       cutoff = 0.9999999999, exact = TRUE
[…]

### 2021-12-06 code/050-tasks.R

@@ -1,5 +1,4 @@
 tasks_plan <- drake_plan(
-
   coordinates_list = list(
     coords_bands_clean,
     coords_vi_nri_clean,
@@ -8,7 +7,6 @@ tasks_plan <- drake_plan(
     coords_vi_nri_clean,
     coords_vi_nri_clean
   ),
-
   data =
     list(
       bands_data,
@@ -18,7 +16,6 @@ tasks_plan <- drake_plan(
       nri_vi_data,
       hr_nri_vi_data
     ),
-
   data_reduced_cor =
     list(
       bands_data_trim_cor,
@@ -39,7 +36,6 @@ tasks_plan <- drake_plan(
     "nri_vi",
     "hr_nri_vi"
   ),
-
   task = target(
     makeRegrTask(
       id = task_names,
@@ -54,7 +50,6 @@ tasks_plan <- drake_plan(
       coordinates_list
     )
   ),
-
   task_reduced_cor = target(
     makeRegrTask(
       id = task_names,

### 2021-12-06 code/060-modeling/paper/01-eda.R

@@ -14,7 +14,6 @@ filter_eda_plan <- drake_plan(
     "FSelectorRcpp_gain.ratio",
     "FSelectorRcpp_symmetrical.uncertainty"
   ),
-
   names_filter_values = c(
     "carscore",
     "FSelectorRcpp_relief",
@@ -24,13 +23,11 @@ filter_eda_plan <- drake_plan(
     "praznik_CMIM",
     "variance"
   ),
-
   task_list_filter = list(
     task[[1]],
     task[[2]],
     task[[3]]
   ),
-
   filter_values = target(
     generateFilterValuesData(
       task = task_list_filter[[1]],
@@ -38,7 +35,6 @@ filter_eda_plan <- drake_plan(
     ),
     dynamic = cross(task_list_filter, names_filter_values)
   ),
-
   filter_values_fselectorrcpp = target(
     generateFilterValuesData(
       task = task_list_filter[[1]],
@@ -47,9 +43,7 @@ filter_eda_plan <- drake_plan(
     ),
     dynamic = cross(task_list_filter, filters_FSelectorRcpp)
   ),
-
   nbins = seq(5, 30, 5),
-
   filter_info_gain_nbins = target(
     generateFilterValuesData(
       task = task_list_filter[[1]],

### 2021-12-06 code/060-modeling/paper/02-param-sets.R

@@ -17,7 +17,6 @@ param_sets_plan <- drake_plan(
       makeNumericParam("fw.perc", lower = 0, upper = 1)
     )
   ),
-
   ps_xgboost = target(
     makeParamSet(
       makeNumericParam("eta", lower = 0.1, upper = 1),
@@ -32,7 +31,6 @@ param_sets_plan <- drake_plan(
       makeIntegerParam("nrounds", lower = 10, upper = 70)
     )
   ),
-
   ps_xgboost_pca = target(
     makeParamSet(
       makeNumericParam("eta", lower = 0.1, upper = 1),
@@ -63,7 +61,6 @@ param_sets_plan <- drake_plan(
       makeNumericParam("fw.perc", lower = 0, upper = 1)
     )
   ),
-
   ps_svm = target(
     makeParamSet(
       makeNumericParam("C",
@@ -76,7 +73,6 @@ param_sets_plan <- drake_plan(
       )
     )
   ),
-
   ps_svm_pca = target(
     makeParamSet(
       makeNumericParam("C",
@@ -101,7 +97,6 @@ param_sets_plan <- drake_plan(
       makeNumericParam("fw.perc", lower = 0, upper = 1)
     )
   ),
-
   ps_rf = target(
     makeParamSet(
       makeNumericParam("mtry.power", lower = 0, upper = 0.5),
@@ -109,7 +104,6 @@ param_sets_plan <- drake_plan(
       makeNumericParam("sample.fraction", lower = 0.2, upper = 0.9)
     )
   ),
-
   ps_rf_pca = target(
     makeParamSet(
       makeNumericParam("mtry.power", lower = 0, upper = 0.5),
@@ -152,7 +146,6 @@ param_sets_plan <- drake_plan(
       probs = c(0.10, 0.90)
     )
   }),
-
   ps_ridge = target(
     makeParamSet(
       makeNumericParam("s",
@@ -195,7 +188,6 @@ param_sets_plan <- drake_plan(
       probs = c(0.10, 0.90)
     )
   }),
-
   ps_lasso = target(
     makeParamSet(makeNumericParam("s",
       # lower = quan_lasso[1],

### 2021-12-06 code/060-modeling/paper/08-train.R

@@ -2,12 +2,10 @@ train_plan <- drake_plan(
   train_hr_xgboost = target(
     train(xgboost_borda, task[[1]])
   ),
-
   iml_new_hr = target(Predictor$new(
     train_hr_xgboost,
     task[[1]]$env$data
   )),
-
   iml_ale_hr = target({
     cl <- makePSOCKcluster(4)
     registerDoParallel(cl)

### 2021-12-06 code/060-modeling/paper/09-feature-importance.R

@@ -14,7 +14,6 @@ feature_imp_plan <- drake_plan(
       setAggregation(rmse, test.mean)
     )
   )),
-
   df_wavelengths_from_indices = target({
 
     # The table is a cleaned version from https://raw.githubusercontent.com/cran/hsdar/master/man/vegindex.Rd
@@ -59,7 +58,6 @@ feature_imp_plan <- drake_plan(
 
     return(df_wide_formatted)
   }),
-
   fi_ale_hr = target({
     mod_hr <- mlr::train(
       tune_wrappers_mbo[[13]], # SVM MBO Car
@@ -75,8 +73,7 @@ feature_imp_plan <- drake_plan(
     )
     return(fe)
   }),
-
-    fi_ale_hr_gs20 = target({
+  fi_ale_hr_gs20 = target({
     mod_hr <- mlr::train(
       tune_wrappers_mbo[[13]], # SVM MBO Car
       task_reduced_cor[[1]]
@@ -91,7 +88,6 @@ feature_imp_plan <- drake_plan(
     )
     return(fe)
   }),
-
   fi_ale_vi = target({
     mod_hr <- mlr::train(
       tune_wrappers_mbo[[13]], # SVM MBO Car
@@ -107,8 +103,7 @@ feature_imp_plan <- drake_plan(
     )
     return(fe)
   }),
-
-    fi_ale_vi_gs20 = target({
+  fi_ale_vi_gs20 = target({
     mod_hr <- mlr::train(
       tune_wrappers_mbo[[13]], # SVM MBO Car
       task_reduced_cor[[2]]

### 2021-12-06 code/090-reports.R

@@ -36,11 +36,10 @@ reports_plan_paper <- drake_plan(
     )
     copy_figures()
   }),
-
   inspect_glmnet_wfr = target({
     wflow_publish(knitr_in("analysis/inspect-glmnet-models.Rmd"),
-                  view = FALSE,
-                  verbose = TRUE
+      view = FALSE,
+      verbose = TRUE
     )
     copy_figures()
   }),

### 2021-12-06 code/099-packages.R

@@ -1,4 +1,4 @@
-load_packages = function() {
+load_packages <- function() {
   suppressPackageStartupMessages(library(drake))
   suppressPackageStartupMessages(library(rgdal))
   suppressPackageStartupMessages(library(hsdar))

### 2021-12-06 inst/HS-filter-borda.R

@@ -7,30 +7,34 @@ library(magrittr)
 
 # Read data
 
-data = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
-coords = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
+data <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
+coords <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
 
 # Merge into one dataset
 
-data = tibble::as_tibble(data.table::rbindlist(data, fill = TRUE))
-data = Filter(function(x) !any(is.na(x)), data)
-coords = tibble::as_tibble(data.table::rbindlist(coords))
+data <- tibble::as_tibble(data.table::rbindlist(data, fill = TRUE))
+data <- Filter(function(x) !any(is.na(x)), data)
+coords <- tibble::as_tibble(data.table::rbindlist(coords))
 
-task = makeRegrTask(id = "defoliation-all-plots", data = data,
-                    target = "defoliation", coordinates = coords,
-                    blocking = factor(rep(1:4, c(479, 451, 291, 529))))
+task <- makeRegrTask(
+  id = "defoliation-all-plots", data = data,
+  target = "defoliation", coordinates = coords,
+  blocking = factor(rep(1:4, c(479, 451, 291, 529)))
+)
 
 # Learners ---------------------
 
 lrn_xgboost <- makeLearner("regr.xgboost",
-                           par.vals = list(
-                             objective = "reg:squarederror",
-                             eval_metric = "error"
-                           )) %>%
+  par.vals = list(
+    objective = "reg:squarederror",
+    eval_metric = "error"
+  )
+) %>%
   makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
 
 lrn_ksvm <- makeLearner("regr.ksvm",
-                        kernel = "rbfdot") %>%
+  kernel = "rbfdot"
+) %>%
   makeFilterWrapper(fw.method = "E-Borda", cache = TRUE)
 
 # Resampling ----------------
@@ -48,67 +52,94 @@ ps_xgboost <- makeParamSet(
   makeNumericParam("gamma", lower = 0, upper = 10),
   makeNumericParam("eta", lower = 0.001, upper = 0.6),
[…]

### 2021-12-06 analysis/eval-performance.Rmd

@@
     learner.id == "XGBOOST MBO Borda" & iter == 4 ~ "Laukiz2",
-
     learner.id == "SVM MBO Car" & iter == 1 ~ "Laukiz1",
@@
     learner.id == "SVM MBO Car" & iter == 4 ~ "Laukiz2",
-
     learner.id == "RF MBO Car" & iter == 1 ~ "Luiando",
@@
     learner.id == "RF MBO Car" & iter == 4 ~ "Oiartzun",
-
     TRUE ~ as.character(iter)

### 2021-12-06 analysis/feature-importance.Rmd

@@
   fi_permut_vi,
-
   fi_ale_hr,
@@
   fi_ale_vi_gs20,
-
   df_wavelengths_from_indices,
@@
     ) +
-
     scale_color_manual(values = c(
@@
 ```{r fi-hr-ale, dev = c("png", "pdf"), warning=FALSE}
-ale_p1 = fi_ale_hr$plot(
+ale_p1 <- fi_ale_hr$plot(
   features = df_wavelengths_from_hr$class,
@@
 # Custom y axis breaks for subplots
-ale_p1[[2]] = ale_p1[[2]] + scale_y_continuous(
+ale_p1[[2]] <- ale_p1[[2]] + scale_y_continuous(
   n.breaks = 3
@@
 
-ale_p1[[3]] = ale_p1[[3]] + scale_y_continuous(
+ale_p1[[3]] <- ale_p1[[3]] + scale_y_continuous(
   n.breaks = 4
@@
 
-ale_p1[[7]] = ale_p1[[7]] + scale_y_continuous(
+ale_p1[[7]] <- ale_p1[[7]] + scale_y_continuous(
   n.breaks = 4
@@
 
-ale_p1[[8]] = ale_p1[[8]] + scale_y_continuous(
+ale_p1[[8]] <- ale_p1[[8]] + scale_y_continuous(
   n.breaks = 4
@@
 
-ale_p1[[9]] = ale_p1[[9]] + scale_y_continuous(
+ale_p1[[9]] <- ale_p1[[9]] + scale_y_continuous(
   n.breaks = 4
@@
 
-ale_p1[[10]] = ale_p1[[10]] + scale_y_continuous(
+ale_p1[[10]] <- ale_p1[[10]] + scale_y_continuous(
   n.breaks = 4
@@
 ```{r, warning=FALSE, message=FALSE}
-ale_p2 = fi_ale_hr_gs20$plot(
+ale_p2 <- fi_ale_hr_gs20$plot(
   features = df_wavelengths_from_hr$class,
@@
 # Custom y axis breaks for subplots
-ale_p2[[1]] = ale_p2[[1]] + scale_y_continuous(
+ale_p2[[1]] <- ale_p2[[1]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p2[[2]] = ale_p2[[2]] + scale_y_continuous(
+ale_p2[[2]] <- ale_p2[[2]] + scale_y_continuous(
   n.breaks = 5
 )
-ale_p2[[3]] = ale_p2[[3]] + scale_y_continuous(
+ale_p2[[3]] <- ale_p2[[3]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p2[[4]] = ale_p2[[4]] + scale_y_continuous(
+ale_p2[[4]] <- ale_p2[[4]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p2[[6]] = ale_p2[[6]] + scale_y_continuous(
+ale_p2[[6]] <- ale_p2[[6]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p2[[7]] = ale_p2[[7]] + scale_y_continuous(
+ale_p2[[7]] <- ale_p2[[7]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p2[[10]] = ale_p2[[10]] + scale_y_continuous(
+ale_p2[[10]] <- ale_p2[[10]] + scale_y_continuous(
   n.breaks = 3
@@
 ```{r fi-vi-ale, dev = c("png", "pdf"), warning=FALSE}
-ale_p3 = fi_ale_vi$plot(
+ale_p3 <- fi_ale_vi$plot(
   features = df_wavelengths_from_indices_label$class,
@@
 # Custom y axis breaks for subplots
-ale_p3[[1]] = ale_p3[[1]] + scale_y_continuous(
+ale_p3[[1]] <- ale_p3[[1]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p3[[2]] = ale_p3[[2]] + scale_y_continuous(
+ale_p3[[2]] <- ale_p3[[2]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p3[[3]] = ale_p3[[3]] + scale_y_continuous(
+ale_p3[[3]] <- ale_p3[[3]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p3[[4]] = ale_p3[[4]] + scale_y_continuous(
+ale_p3[[4]] <- ale_p3[[4]] + scale_y_continuous(
   n.breaks = 4
@@
 ```{r, warning=FALSE}
-ale_p4 = fi_ale_vi_gs20$plot(
+ale_p4 <- fi_ale_vi_gs20$plot(
   features = df_wavelengths_from_indices_label$class,
@@
 # Custom y axis breaks for subplots
-ale_p4[[2]] = ale_p3[[2]] + scale_y_continuous(
+ale_p4[[2]] <- ale_p3[[2]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p4[[3]] = ale_p3[[3]] + scale_y_continuous(
+ale_p4[[3]] <- ale_p3[[3]] + scale_y_continuous(
   n.breaks = 4
 )
-ale_p4[[4]] = ale_p3[[4]] + scale_y_continuous(
+ale_p4[[4]] <- ale_p3[[4]] + scale_y_continuous(
   n.breaks = 4

### 2021-12-06 analysis/test/cloud-cover.Rmd

@@
 ```{r setup, include=FALSE}
-knitr::opts_knit$set(base.dir = 'analysis/rmd/')
-knitr::opts_chunk$set(fig.retina = 3, fig.align = 'center', fig.width = 6.93, 
-                      fig.height = 6.13, out.width = "100%")
+knitr::opts_knit$set(base.dir = "analysis/rmd/")
+knitr::opts_chunk$set(
+  fig.retina = 3, fig.align = "center", fig.width = 6.93,
+  fig.height = 6.13, out.width = "100%"
+)
 source(here::here("code/99-packages.R"))
@@
 # Unique dates
-date_scenes =  
+date_scenes <-
   records %>%
@@
 # Scenes per mosaic/ date
-count_scenes =
+count_scenes <-
   records %>%
@@
 # Row number of mosaic
-row_scenes =
+row_scenes <-
   records %>%
@@
 # Calculate area of aoi
-aoi_area = st_area(aoi) %>%
+aoi_area <- st_area(aoi) %>%
   sum()
@@
 # Calculate cloud cover
-cloud_scenes =
+cloud_scenes <-
   cloud_mosaic %>%
-  map(~ st_read(., quiet=TRUE)) %>%
+  map(~ st_read(., quiet = TRUE)) %>%
   map(~ st_set_crs(., 25830)) %>%
@@
   map_dbl(possibly(~ sum(.), NA)) %>%
-  map_dbl(possibly(~ ./aoi_area, NA))
+  map_dbl(possibly(~ . / aoi_area, NA))
 
-ext_scenes =
+ext_scenes <-
   records %>%

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/final-submission/remotesensing-1398065-final.tex
add assets of all submissions

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[remotesensing,article,accept,moreauthors,pdftex]{Definitions/mdpi}

%--------------------
% Class Options:
%--------------------

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%=================================================================
% MDPI internal commands
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2021}
\copyrightyear{2021}
\externaleditor{Academic Editor: {Hanna Meyer}
} % For journal Automation, please change Academic Editor to "Communicated by"
\datereceived{12 September 2021}
\dateaccepted{23 November 2021}
\datepublished{}
\hreflink{https://doi.org/} % If needed use \linebreak
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

% \usepackage[noadjust]{cite}
% \usepackage[hidelinks]{hyperref}
\usepackage[nolist]{acronym}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{adjustbox}

% \usepackage{supertabular}
\usepackage{supertabular,booktabs}

% line breaks in table cells
\newcommand{\specialcell}[2][l]{%
  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

\usepackage{csquotes}
\usepackage[utf8]{inputenc}
% enables custom font styles which are not available in IEEEtran by default
% https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran

%\renewcommand{\sfdefault}{cmss}
%\renewcommand{\rmdefault}{cmr}
%\renewcommand{\ttdefault}{cmt}

\usepackage[detect-all]{siunitx}

\usepackage[pdftex]{graphicx}
% declare the path(s) where your graphic files are
\graphicspath{{./pdf/}{./jpg/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

% suppress fancyhdr warnings
\setlength{\headheight}{23.90979pt}

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
\newcommand*\mean[1]{\bar{#1}}

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Monitoring Forest Health Using Hyperspectral Imagery: Does Feature Selection Improve the Performance of Machine-\linebreak Learning Techniques?}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Monitoring Forest Health Using Hyperspectral Imagery: Does Feature Selection Improve the Performance of Machine-Learning Techniques? }

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0003-0748-6624} % Add \orcidA{} behind the author's name
\newcommand{\orcidauthorB}{0000-0001-7834-4717} % Add \orcidB{} behind the author's name
\newcommand{\orcidauthorC}{0000-0003-2567-8689} % Add \orcidC{} behind the author's name
\newcommand{\orcidauthorD}{0000-0002-6390-5873} % Add \orcidD{} behind the author's name
\newcommand{\orcidauthorE}{0000-0001-6002-6980} % Add \orcidE{} behind the author's name
\newcommand{\orcidauthorF}{0000-0001-6640-679X} % Add \orcidF{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Patrick Schratz $^{1,}$*\orcidA{}, Jannes Muenchow $^{1}$\orcidB{}, Eugenia Iturritxa $^{2}$\orcidD{}, José Cortés $^{1}$\orcidC{}, Bernd Bischl $^{3}$\orcidE{} and~Alexander~Brenning $^{1}$\orcidF{}}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Patrick Schratz, Jannes Muenchow, Eugenia Iturritxa, José Cortés, Bernd Bischl and Alexander Brenning}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Schratz, P.; Muenchow, J.; Iturritxa, E.; Cortés, J.; Bischl, B.; Brenning, A.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad GIScience Group, Department of Geography, Friedrich Schiller University Jena, Germany;
 malnamalja@gmx.de~(J.M.); jose.cortes@uni-jena.de~(J.C.); alexander.brenning@uni-jena.de~(A.B.)\\
$^{2}$ \quad NEIKER Tecnalia, Spain; eiturritxa@neiker.eus\\
$^{3}$ \quad Department of Statistics, Ludwig-Maximilians-Universität München, Germany; bernd.bischl@stat.uni-muenchen.de
}

% Contact information of the corresponding author
\corres{Correspondence: patrick.schratz@uni-jena.de}

% Current address and/or shared authorship
% \firstnote{Current address: Affiliation 3}
% \secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{This study analyzed highly correlated, feature-rich datasets from hyperspectral remote sensing data using multiple statistical and machine-learning methods.
The effect of filter-based feature selection methods on predictive performance was compared.
In addition, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
Defoliation of trees (\%), derived from in situ measurements from fall 2016, was modeled as a function of reflectance.
Variable importance was assessed using permutation-based feature importance.
Overall, the support vector machine (SVM) outperformed other algorithms, such as random forest (RF), extreme gradient boosting (XGBoost), and lasso (L1) and ridge (L2) regressions by at least three percentage points.
The combination of certain feature sets showed small increases in predictive performance, while no substantial differences between individual feature sets were observed.
For some combinations of learners and feature sets, filter methods achieved better predictive performances than using no feature selection.
Ensemble filters did not have a substantial impact on performance.
The most important features were located around the red edge.
Additional features in the near-infrared region (800 nm--1000 nm) were also essential to achieve the overall best performances.
Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.
Nevertheless, more training data and replication in similar benchmarking studies are needed to be able to generalize the results.}

% Keywords
\keyword{hyperspectral imagery; forest health monitoring; machine learning; feature selection; model comparison}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

\begin{acronym}

	% geringerer Zeilenabstand

	%\setlength{\itemsep}{-\parsep}
	\acro{AGB}{above-ground biomass}
	\acro{ALE}{accumulated local effects}
	\acro{ALS}{airborne laser scanning}
	\acro{ANN}{artificial neural network}
	\acro{AUROC}{area under the receiver operating characteristics curve}
	\acro{BRT}{boosted regression trees}
	\acro{CART}{classification and regression trees}
	\acro{CNN}{convolutional neural networks}
	\acro{CV}{cross-validation}
	\acro{DAP}{digital aerial photogrammetry}
	\acro{ENM}{environmental niche modeling}
	\acro{FFS}{forward feature selection}
	\acro{FPR}{false-positive rate}
	\acro{FS}{feature selection}
	\acro{GAM}{generalized additive model}
	\acro{GBM}{gradient boosting machine}
	\acro{GLM}{generalized linear model}
	\acro{ICGC}{Institut Cartografic i Geologic de Catalunya}
	\acro{IQR}{interquartile range}
	\acro{LiDAR}{light detection and ranging}
	\acro{LOWESS}{locally weighted scatterplot smoothing}
	\acro{MARS}{multivariate adaptive regression splines}
	\acro{MBO}{model-based optimization}
	\acro{MEM}{maximum entropy model}
	\acro{ML}{machine learning}
	\acro{NDII}{normalized difference infrared index}
	\acro{NDMI}{normalized difference moisture index}
	\acro{NIR}{near-infrared}
	\acro{NRI}{normalized ratio index}
	\acro{OLS}{ordinary least squares}
	\acro{OMNBR}{optimized multiple narrow-band reflectance}
	\acro{PCA}{principal component analysis}
	\acro{PDP}{partial dependence plots}
	\acro{PISR}{potential incoming solar radiation}
	\acro{PLS}{partial least squares}
	\acro{POV}{proportion of variance explained}
	\acro{RBF}{radial basis function}
	\acro{RF}{random forest}
	\acro{RMSE}{root mean square error}
	\acro{RR}{ridge regression}
	\acro{RSS}{residual sum of squares}
	\acro{SAR}{synthetic aperture radar}
	\acro{SDM}{species distribution modeling}
	\acro{SMBO}{sequential model-based optimization}
	\acro{SVM}{support vector machine}
	\acro{TPR}{true-positive rate}
	\acro{VI}{vegetation index}
	\acro{XGBoost}{extreme gradient boosting}
	\acro{PC}{principal component}
\end{acronym}

\begin{document}
\section{Introduction}

The use of \ac{ML} algorithms for analyzing remote sensing data has seen a huge increase in the last decade~\cite{lary2016}.
This coincided with the increased availability of remote sensing imagery, especially since the launch of the first Sentinel satellite in the year 2014.
At the same time, the implementation and usability of learning algorithms has been greatly simplified with many contributions from the open-source community.
Scientists can nowadays process large amounts of (environmental) information with relative ease using various learning algorithms.
This makes it possible to easily extend benchmark comparison matrices of studies in a semi-automated way, possibly stumbling upon unexpected findings, such as process settings, that would not have been explored otherwise~\cite{ma2015}.

% link to forest health (defoliation) analysis and show exemplary studies

ML methods in combination with remote sensing data are used in many environmental fields, such as vegetation cover analysis and forest carbon storage mapping~\cite{mascaro2014, urban2018}.
The ability to predict these environmental variables in unknown regions qualifies ML algorithms as a helpful tool for such environmental analyses.
One aspect of this research field is to enhance the understanding of biotic and abiotic stress triggers, for example, by analyzing tree defoliation~\cite{hawrylo2018}.
Defoliation is known to be a proxy for pathogen and insect damage~\cite{pollastrini2016}.
While common symptoms are observable across species, some effects and their degree of severity are species-specific~\cite{gottardini2020}.
Defoliation has also been shown to increase predisposition of tree death from secondary biotic factors up to ten years after the actual defoliation event~\cite{oliva2016}.
Other approaches for analyzing forest health include change detection~\cite{zhang2016} or describing the current health status of forests on a stand level~\cite{townsend2012}.

% introduce veg indices
Vegetation indices have shown the potential to provide valuable information when analyzing forest health~\cite{jiang2014, adamczyk2015}.
Most vegetation indices were developed with the aim of being sensitive to changes in specific wavelength regions, serving proxies for underlying plant physiological processes.
In some cases, indices developed for different purposes than the one to be analyzed (e.g., defoliation of pine trees) may help to explain complex underlying relationships that are not obvious at first.
This emphasizes the need to extract as much information as possible from the available input data to generate promising features that can help in improving our understanding of the modeled relationship~\cite{thenkabail2018}.
A less known index type that can be derived from spectral information is the \ac{NRI}.
In contrast to most vegetation indices, \ac{NRI}s do not use an expert-based formula following environmental heuristics; instead, they make use of a data-driven feature engineering approach by combining (all possible) combinations of spectral bands~\cite{thenkabail2000}.
When working with hyperspectral data, thousands of \ac{NRI} features can be derived this way.

% mention why FS is used / important
Even though \ac{ML} algorithms are capable of handling highly correlated input variables, model fitting becomes computationally more demanding and model interpretation more challenging.
Feature selection approaches can help to address this issue, reducing possible noise in the feature space, simplifying model interpretability, and possibly enhancing predictive performance~\cite{cai2018}.
Instead of using wrapper feature selection methods, which add a substantial overhead to a nested model optimization approach, especially for datasets with many features, this study focuses on (ensemble) filter methods, which can be directly integrated into the hyperparameter optimization step during model construction.

% clarify the focus of this study
Overall, this study aims to show how high-dimensional datasets can be handled effectively with ML methods while still allowing the interpretation of the fitted models.
The predictive power of non-linear methods and their ability to handle highly correlated predictors is combined with common and new approaches for assessing feature importance and feature effects.
However, this study focuses mainly on investigating the effects of filter methods and feature set types on predictive performance rather than interpreting individual feature effects.

Considering these opportunities and challenges, the research questions of this study are as follows:

\begin{itemize}

	\item How do different feature selection methods influence the predictive performance of ML models of the defoliation of trees?

	\item Do different (environmental) feature sets show differences in performance?

	\item Can predictive performance be substantially improved by combining feature sets?

	\item Which features are most important and how can these be interpreted in this context?

\end{itemize}

% Explain what is new in this study and link to methodology
In recent years, various studies that have used hyperspectral data to analyze pest/fungi infections on trees have been published.
Pinto et al. \citep{pinto2020} successfully used hyperspectral imagery to characterize pest infections on peanut leaves using random forest, while Yu~et~al.~\cite{yu2021} aimed to detect pine wilt disease in pine plots in China using vegetation indices derived from hyperspectral data.
Other studies which applied hyperspectral data for forest health monitoring are~\cite{lin2014,kayet2019,dash2017}.
Building upon these successful applications of hyperspectral remote sensing usage in the field of leaf and tree health monitoring, this work analyzes tree defoliation in northern Spain using airborne hyperspectral data.
The methodology of this study uses ML methods in combination with feature selection and hyperparameter tuning.
In addition, feature importance was analyzed.
Incorporating the idea of creating data-driven \ac{NRI}s, this study also discusses the practical problems of high dimensionality in environmental modeling~\cite{trunk1979, xu2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

\subsection{Data and Study Area}

Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands were available for four Monterey Pine (\textit{Pinus radiata D. Don}) plantations in northern Spain.
The trees in the plots suffer from infections with pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P. Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.}, and \textit{Dothistroma septosporum (Dorogin) M. Morelet} causing (among others) needle blight, pitch canker, and root diseases~\cite{mesanza2016, iturritxa2017}.
The first two fungi are mainly responsible for the foliation loss of the trees analyzed in this study~\cite{iturritxa2014}.
In situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection by experts. \textit{Defoliation} in percent was used as the response variable (Figure~\ref{fig:defol-distr}).

It is assumed that the fungi infect the trees through open wounds, possibly caused by previous hail damage~\cite{iturritxa2014}.
The dieback of these trees, which are mainly used as timber, causes substantial economic losses~\cite{ganley2009}.

\begin{figure}[H]
	%\centering
	%\begin{center}
		\includegraphics[width=0.7\textwidth] {defoliation-distribution-plot-1.pdf}
		\caption{Response variable ``defoliation of trees'' for plots~Laukiz1, Laukiz2, Luiando, and Oiartzun. \texttt{n} is the total number of trees in each plot, and $\bar{x}$ the mean defoliation. Values for Laukiz1, Luiando, and Oiartzun were observed in 5\% intervals; for Laukiz2, defoliation was observed at multiple heights and then averaged, leading to smaller defoliation differences than 5\%.}\label{fig:defol-distr}
	%\end{center}
\end{figure}

\subsubsection{In Situ Data}

The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando, and Oiartzun, are located in the northern part of the Basque Country (Figure~\ref{fig:study_area}).
Oiartzun has the largest number of observations ($n = 559$ trees), while Laukiz2 is the largest in area (1.44~ha).
All plots besides Luiando are located within 100~km from the coast (Figure~\ref{fig:study_area}).
A total of 1808~observations are available (Laukiz1: 559, Laukiz2: 451, Luiando: 301, Oiartzun: 497).
Field surveys were conducted in September 2016 by experienced forest pathologists.
Defoliation was measured in 5\% steps through visual inspection with the help of a score card.
For Laukiz2, values at three height levels (bottom, mid, and top) were available and averaged into an overall defoliation value, resulting in values that are not multiples of 5\% (e.g., 8.33\%).
The magnitude of human observer errors in such surveys, including the present one, is not precisely known and has being discussed for many years~\cite{innes1993}.
Ref.~\cite{maclean1982} estimated human observer errors in defoliation surveys to range between 7\% and 18\%.

% start a new page without indent 4.6cm
\vspace{-12pt}
\end{paracol}
\nointerlineskip
\begin{figure} [H]
\widefigure
	%\begin{center}
		%\centering
		\includegraphics[width=\textwidth] {study-area-hyperspectral.pdf}
		\caption{Study area maps showing information about location, size, and spatial distribution of trees for all plots (Laukiz1, Laukiz2, Luiando, and Oiartzun). The background maps give a visual impression of the individual plot area but do not necessarily represent the plot's state during data acquisition.}\label{fig:study_area}
	%\end{center}
\end{figure}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

%\begin{figure}[H]
%	%\begin{center}
%		\includegraphics[width=0.7\textwidth] {defol-grid-3000px.jpg}
%		\caption{Example trees showing different levels of defoliation as interpreted by the surveyor: 10 \% (top-left), 20 \% (top-right), 40 \% (bottom-left), 60-70 \% (bottom-right).}
%		\label{fig:defol-trees}
%	%\end{center}
%\end{figure}

\vspace{-12pt}
\subsubsection{Hyperspectral Data}

The airborne hyperspectral data were acquired by an AISA EAGLE-II sensor during two flight campaigns on 28 September and 5 October 2016 at noon.
All preprocessing steps (geometric, radiometric, and atmospheric) were conducted by the \ac{ICGC}.
The first four bands were corrupted, leaving 122 bands with valid information.
Additional metadata information is available in Table~\ref{tab:hyperparameter_metadata}.

% parameter limits

\begin{specialtable}[H]
\setlength{\tabcolsep}{11.3mm}
	%\centering
	\caption[t]{Specifications of hyperspectral data.}
	\begingroup
	\begin{tabular}{ll}
	\toprule
		\textbf{Characteristic}         & \textbf{Value}                               \\\midrule

		Geometric resolution   & 1~m                                 \\
		Radiometric resolution & 12 bit                              \\
		Spectral resolution    & 126 bands (404.08~nm--996.31~nm) \\
		Correction:            & Radiometric, geometric, atmospheric\\\bottomrule
	\end{tabular}
	\endgroup\label{tab:hyperparameter_metadata}
\end{specialtable}

\subsection{Derivation of Indices}

To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package \texttt{hsdar} (89 in total) as well as all possible \ac{NRI} combinations were calculated.
NRIs follow the \ac{OMNBR} concept of data-driven information extraction from narrow-band indices of hyperspectral data~\cite{thenkabail2000,thenkabail2018}.
While various index formulations, such as band ratios or normalized ratios, are available, indices involving the same bands are strongly correlated.
Since the widely\ used NDVI index belongs to the group of normalized ratio indices (NRIs), which are implemented in the \texttt{hsdar} R package, we used the following normalized difference index (NDI) formula to combine all pairs of reflectances:
\begin{equation}
	NRI_{i,j} = \frac{band_{i} - band_{j}}{band_{i} + band_{j}}
\end{equation}
where \(i\) and \(j\) are the respective band numbers.

%\bigbreak{}

To account for geometric offsets within the hyperspectral data, which were reported by \ac{ICGC} to be potentially up to one meter, a buffer of one meter radius around the centroid of each tree was used when extracting the reflectance values.
A pixel was considered to fall into a tree's buffer zone if the centroid of the respective pixel was touched by the buffer.
The pixel values within a buffer zone were averaged and formed the final reflectance value of a single tree, and they were used as the base information to derive all additional feature sets.
In total, \(\frac{121*122}{2} = 7381\) NRIs were calculated.

\subsection{Feature Selection}

High-dimensional, feature-rich datasets come with several challenges for both model fitting and evaluation:

\begin{itemize}
	\item Model fitting times increase.
	\item Noise is possibly introduced into models by highly correlated variables~\cite{johnstoneiainm.2009}.
	\item Model interpretation and prediction become more challenging~\cite{johnstoneiainm.2009}.
\end{itemize}

To reduce the feature space of a dataset, several conceptually differing approaches exist: wrapper methods, filters, penalization methods (lasso and ridge), and \ac{PCA}~\cite{bommert2020, das2001, guyon2003, jolliffe2016}.
In contrast to wrapper methods, filters have a much lower computational cost, and their tuning can be added to the hyperparameter optimization step.
In addition, filters are less known than wrapper methods, and, in recent years, ensemble filters, which have shown promising results compared to single filter algorithms, were introduced~\cite{drotar2017}.
These two points mainly led to the decision to focus on filter methods for this work and evaluate their effectiveness on highly\ correlated, high-dimensional datasets.
Due to this focus, only this subgroup of feature selection methods is be introduced in greater detail in the following sections.

\subsubsection{Filter Methods}

% Filter methods
The concept of filters originates from the idea of ranking features following a score calculated by an algorithm~\cite{guyon2003}.
Some filter methods can only deal with specific types of variables (e.g., numeric or nominal).
Filters only rank features; they do not decide which covariates to drop or keep~\cite{drotar2015}.
The decision of which features to keep for model fitting can be integrated into the optimization phase during model fitting, along with hyperparameter tuning.
Thus, the number of top-ranked features to be included in the model is treated as an additional hyperparameter of the model.
This hyperparameter is tuned to optimize the model's performance.

% Ensemble filter methods
Beyond the use of individual filter methods to rank and select features, recent studies have shown that combining several filters by using statistical operations such as "minimum" or "mean" may enhance the predictive performance of the resulting models, especially when applied to multiple datasets~\cite{abeel2010, drotar2017}.
This approach is referred to as "ensemble filtering"~\cite{dietterich2000}.
Ensemble filters align with the recent rise of the ensemble approach in ML, which uses the idea of stacking to combine the predictions of multiple models, aiming to enhance predictive performance~\cite{polikar2012, feurer2015, bolon-canedo2019}.
In this work, the Borda ensemble filter was used~\cite{drotar2017}.
Its overall feature order is determined by the sum of filter ranks of all individual filters in the~ensemble.

% Ensuring a fair weighting in the ensemble
Filter methods can be categorized based on three binary criteria: multivariate or univariate feature use, correlation or entropy-based importance weighting, and linear and non-linear filter methodology.
Care needs to be taken to not weigh certain classes more than others in the ensemble, as, otherwise, the ranking will be biased.
In this study, this was taken care of by checking the rank correlations (Spearman's correlation) of the generated feature rankings of all methods against each other.
If filter pairs showed a correlation of 0.9 or higher, only one of the two was included in the ensemble filter, selecting it at random.

\subsubsection{Description of Used Filter Methods}

Filter methods can be classified as follows (Table~\ref{tab:filter-methods}):

\begin{itemize}
	\item Univariate/multivariate (scoring based on a single variable/multiple variables).
	\item Linear/non-linear (usage of linear/non-linear calculations).
	\item Entropy/correlation (scoring based on derivations of entropy or correlation-based approaches).
\end{itemize}
\vspace{-6pt}

% filter methods
\begin{specialtable}[H]
\setlength{\tabcolsep}{5.2mm}
	%\centering
	\caption{List of filter methods used in this work, their categorization, and scientific reference.}
	\label{tab:filter-methods}
	\begingroup\footnotesize
%	\begin{adjustbox}{width={0.7\textwidth},totalheight={\textheight},keepaspectratio}
		\begin{tabular}{lll}
				\toprule
			\textbf{Name}                                         & \textbf{Group} & \textbf{Ref. }              \\\midrule

			Linear correlation (Pearson)                 & univariate, linear, correlation   &~\cite{pearson1901} \\
			Information gain                             & univariate, non-linear, entropy   &~\cite{quinlan1986} \\
			Minimum redundancy, maximum relevance        & multivariate, non-linear, entropy &~\cite{zhao2013}    \\
			Carscore                                     & multivariate, linear, correlation &~\cite{zuber2011}   \\
			Relief                                       & multivariate, linear, entropy     &~\cite{kira1992}    \\
			Conditional minimal information maximization & multivariate, linear, entropy     &~\cite{fleuret2004}\\\bottomrule
		\end{tabular}
%	\end{adjustbox}
	\endgroup
\end{specialtable}

The filter "Information Gain" in its original form is only defined for nominal response variables:
\begin{equation}
	H(Class) + H(Attribute) - H(Class, Attribute)
\end{equation}
where \(H\) is the conditional entropy of the response variable (class or Y) and the feature (attribute or X).
$H(X)$ is Shannon's entropy~\cite{shannon1948} for a variable X, and $H(X, Y)$ is a joint Shannon's entropy for a variable X with a condition to Y.
$H(X)$ itself is defined as
\begin{equation}
	H(X) = - \sum_{i=1}^{n} P(x_i)\log_bP(x_i)
\end{equation}
where $b$ is the base of the logarithm used, most commonly 2.

In order to use this method with a numeric response (percentage defoliation of trees), the variable was discretized into equal bins \texttt{\(n_{bin}\) = 10} and treated as a categorical variable.

\subsection{Benchmarking Design}

\subsubsection{Algorithms}

The following learners were used in this work:

\begin{itemize}
	\item  Extreme gradient boosting (XGBoost);
	\item  Random forest (RF);
	\item  Penalized regression (with L1/lasso and L2/ridge penalties);
	\item  Support vector machine (SVM, radial basis function Kernel);
	\item  Featureless learner.
\end{itemize}

RF and SVM are well-established algorithms and widely used in environmental remote sensing.
Extreme gradient boosting (commonly abbreviated as XGBoost) has shown promising results in benchmarking studies in recent years.
Penalized regression is a statistical modeling technique capable of dealing with highly\ correlated covariates by penalizing the model coefficients~\cite{hastie2001}.
Common penalties are "lasso" (L1) and "ridge" (L2).
Ridge regression does not remove variables from the model (penalization to zero), but it shrinks them towards zero, keeping them in the model.
A featureless learner was included for a baseline comparison.

In total, the benchmarking grid consisted of 156 experiments (6 feature sets $\times$ 3~ML algorithms $\times$ 8 feature-selection methods and for the L1/L2 models, 6 feature sets~$\times$~2~models.
The selected hyperparameter settings are shown in Appendix~\ref{appendixA} Table~\ref{tab:hyperparameter_limits}.
All code and data are included in the research compendium of this study (\url{https://doi.org/10.5281/zenodo.2635403} (accessed on 27 November 2021).

\subsubsection{Feature Sets}

Three feature sets were used in this study, each representing a different approach to feature engineering:

\begin{itemize}
	\item The raw hyperspectral band information (HR): no feature engineering %chktex 13;
	\item Vegetation indices (\ac{VI}s): expert-based feature engineering;
	\item Normalized ratio indices (\ac{NRI}s): data-driven feature engineering.
\end{itemize}

The idea of splitting the features into different sets originated from the question of whether feature-engineered indices derived \textls[-5]{from reflectance values have a positive effect on model performance.
Peña et al. 2017~\cite{pena2017} is an exemplary study that used this approach in a spectro-temporal setting.}
Benchmarking learners on these feature sets while keeping all other variables, such as model type, tuning strategy, and a partitioning method, fixed makes it possible to draw conclusions on their individual impact.
Each feature set has distinct capabilities that differentiate it from the others.
This can have both a positive and negative effect on the resulting performance, which is one question this study aims to explore.
For example, feature set VI misses certain parts of the spectral range, as the chosen indices only use specific spectral bands.
Feature set NRI will introduce highly\ correlated features, for which some algorithms may be more suitable than others.

In addition to these individual feature sets, the following combinations of feature sets were also compared:

\begin{itemize}
	\item HR + VI %chktex 13;
	\item HR + NRI;
	\item HR + VI + NRI.
\end{itemize}

Some individual features were removed before using the datasets for modeling when being numerically equivalent to another feature based on the pairwise correlation being greater than $1 - 10^{-10}$.
This preprocessing step reduced the number of covariates for feature set~VI to 86 (from 89).
This decision was made to prevent numerical issues that may occur in the subsequent tuning, filtering, and model fitting steps when offering features with a pairwise correlation of (almost) one.
The remaining features were then used as input for the filter-based feature selection within the CV.

\subsubsection{Hyperparameter Optimization}

Hyperparameters were tuned using \ac{MBO} within a nested spatial \ac{CV}~\cite{mlrmbo, binder2020, schratz2019}.
In MBO, first, \textit{n} hyperparameter settings are randomly chosen from a user-defined search space.
After these \textit{n} settings have been evaluated, one new setting, which is evaluated next, is proposed by a fitted surrogate model (by default, a kriging method).
This strategy continues until a user-defined stopping criterion is satisfied~\cite{hutter2011, jones1998}.

In this work, an initial design of 30 randomly drawn hyperparameter settings in combination with a stopping criterion of 70 iterations was used, resulting in a total budget of 100 evaluated hyperparameter settings per fold.
The advantage of this tuning approach is a substantial reduction of the tuning budget that is required to find a setting close to the global optimization minimum.
\ac{MBO} may outperform methods that do not use information from previous iterations, such as random search or grid search~\cite{bergstra2012}.
Tuning ranges used in this work are shown in Table~\ref{tab:hyperparameter_limits}.

To optimize the number of features used for model fitting, the percentage of features was added as a hyperparameter during the optimization stage~\cite{binder2020}.
For \ac{PCA}, the number of principal components was tuned.
%The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_{try} = p^x_{try}$.
The RF hyperparameter \texttt{\(m_{try}\)} was re-expressed as $m_\textrm{try} = p_\textrm{sel}^t$, a function of the number of selected features, $p_\textrm{sel}$.
It was thus tuned on a logarithmic scale by varying $t$ between 0 (i.e., $m_\textrm{try} = 1$) and 0.5 (i.e., $m_\textrm{try}=\sqrt{p_\textrm{sel}}$).
This was necessary to ensure that \texttt{\(m_{try}\)} did not exceed the number of features available after optimizing the feature percentage during tuning.

\subsubsection{Spatial Resampling}

A spatial nested cross-validation on the plot level was chosen to account for spatial autocorrelation within the plots and assess model transferability to different plots~\cite{schratz2019, sperrorest}.
The \ac{RMSE} was chosen as the error measure.
Each plot served as one cross-validation fold, resulting in four iterations in total.
The inner level of cross-validation for hyperparameter tuning also used plot-level cross-validation.

\subsection{Feature Importance and Feature Effects}

Estimating feature importance for datasets with highly correlated features is a challenging task for which numerous model-specific and model-agnostic approaches exist~\cite{friedman2001, hastie2001, greenwell2018}.
The strong correlations among features make it challenging to calculate an unbiased estimate for single features~\cite{molnar2019}.
Methods such as \ac{PDP} or permutation-based approaches may produce unreliable estimates in such scenarios because unrealistic combinations of feature values are created~\cite{molnar2019}.
The development of robust methods that enable an unbiased estimation of feature importance for highly correlated variables is subject to current research~\cite{brenning2021}.

In this work, permutation-based feature importance was calculated to estimate feature importance or effects~\cite{apley2019}.
With the limitations in mind when applied to correlated features, the aim was to get a general overview of the feature importance of the hyperspectral bands while trying to avoid an over-interpretation of results.
The best-performing algorithm on the HR task (i.e., SVM) was used for the feature importance calculation.

%\subsection{Linking feature importance to wavelength regions}

To facilitate interpretation, the ten most important indices of the best performing models using feature sets HR and VI were linked to the spectral regions of the hyperspectral data that went into their calculation.
The aim was to visualize the most important features along the spectral curve of the plots to better understand which spectral regions were most important for the model.

\subsection{Research Compendium}

All tasks of this study were conducted using the open-source statistical programming language R~\cite{rcoreteam2019}.
A complete list of all R packages used in this study can be found in the linked repositories mentioned in the next paragraph.
Due to space limitations, only selected packages with high impact on this work are explicitly cited.

The algorithm implementations of the following packages were used: xgboost~\cite{chen2016} (extreme gradient boosting), kernlab~\cite{kernlab} (support vector machine) and glmnet~\cite{glmnet} (penalized regression).
The filter implementations of the following packages were used: praznik~\cite{praznik} and FSelectorRcpp~\cite{fselectorrcpp}.
Package mlr~\cite{mlr} was used for all modeling related steps, and drake~\cite{drake} was used for structuring the work and for reproducibility.
This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}, (accessed on 27 November 2021).
Apart from the availability of code and manuscript sources, a static webpage is available at \url{https://pat-s.github.io/2019-feature-selection} (accessed on 27 November 2021) which includes additional side analyses that were carried out during the creation of this study.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\subsection{Principal Component Analysis of Feature Sets}

PCA was used to assess the complexity of the three feature sets.
Depending on the feature set, 95\% of the variance is explained by two (HR), twelve (VI), and 42 (NRI) \ac{PC}s.
HR features are therefore highly redundant, while the applied feature transformations enrich the data set, at least from an exploratory linear perspective.

\subsection{Predictive Performance}

% - RMSE around 27
% - MBO tuned penalized methods are substantially worse
Overall, the response variable \enquote{tree defoliation} could be modeled with an \ac{RMSE} of 28 percentage points (p.p.) (Figure~\ref{fig:perf-result}).
SVM showed almost no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p. (Figure~\ref{fig:perf-result}).
SVM showed the best overall performance with a mean difference of around three p.p. to the next best model (XGBoost) (Table~\ref{tab:best-learner-perf}).
Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 p.p. for learner SVM (without filter) but up to 54.3 p.p. when testing on Laukiz2 (Table~\ref{tab:svm-single-fold-perf}).

% feature sets
% - no substantial difference between feature sets
% - combining feature sets has only small impact
% - expert based feature sets show no better result than data-driven ones
% - penalized methods seem to have problems with VI feature set
The combination of feature sets showed small increases in performance for some learners.
XGBoost scored slightly better on the combined datasets HR-NRI, NRI-VI, and HR-NRI-VI compared to their standalone variants (NRI and VI) (Figure~\ref{fig:perf-result}).
However, the best performances for RF and XGBoost were scored on NRI and HR, respectively.
RF showed a substantial performance increase when using only NRI compared to all other feature sets, whereas for XGBoost, the worst performances were associated with the VI- and NRI-only feature sets (Figure~\ref{fig:perf-result}).

% learner
% - SVM Info Gain is the overall winner
% - Combined sets do not help for XBoost and Ridge
% - RF NRI quite strong
SVM combined with the \enquote{Information Gain} filter achieved the best overall performance (RMSE of 27.915 p.p.) (Table~\ref{tab:perf-top-10}).
%L1 penalized regression showed slightly better results in three out of six tasks when being optimized using the package internal 10-fold CV compared to MBO (RMSE of 50 vs 58).
Regressions with ridge (L2) and lasso (L1) penalties showed their best performances on the NRI feature set (Table~\ref{tab:best-learner-perf}).
Combining feature sets for lasso and ridge did not help to increase performance, and while there was no substantial difference for lasso, the performance of ridge improved by around two percentage points.
XGBoost showed very poor performances for some feature sets and fills the last\ ten places of the ranking (Table~\ref{tab:perf-worst-10}).
Especially when combined with PCA, the algorithm was not able to achieve adequate performances.

% filters
% - influence varies based on algorithm and task (SVM no influence, RF medium, XGB medium)
% - no substantial decrease in performance when using filters
% - no advantage of ensemble borda filter compared to others
The effects of filter methods on performance differed greatly between the algorithms:
SVM showed no variation in performance across filters (Figure~\ref{fig:filter-effects-no-filter}).
The use of filters for RF resulted in a substantial increase in performance in all tasks, especially on the HR feature set where all filters showed an improved score compared to using no filter (Figure~\ref{fig:filter-effects-no-filter}).
XGBoost's performance depended strongly on feature selection. In two out of six tasks (HR, VI), using no filter resulted in the worst performance.
XGBoost showed the highest overall differences between filters for a single task---for feature set HR, the range is up to 13~p.p. (\enquote{CMIM} vs. \enquote{no filter}) (Figure~\ref{fig:filter-effects-no-filter}).

% filter vs. no filter
% - Substantial effect for XGBoost and RF
% - SVM unaffected
When comparing the usage of filters against using no filter at all, there were no instances in which a non-filtered model scored a better performance than the best filtered one (Figure~\ref{fig:filter-effects-no-filter}).
For SVM, all filters and \enquote{no filter} achieved roughly the same performance on all tasks.

% Borda vs. other filters
% - most often among the first best filters for each task
% - never the best
The Borda ensemble filter was not able to score the best performance in any learner/\linebreak filter setting (Figure~\ref{fig:filter-effects-borda}).
For RF and XGBoost, it most often ranked within the better half among all filters of a specific task.

% Selected percentage of features across plots
The number of features selected during model optimization strongly varied across learners and plots.
RF selected the least features of all three learners, and with the exception of Oiartzun, only one or two variables were selected.
SVM used 210 features or more in all instances and selected between 16\% (Laukiz1) and 81\% (Oiartzun) of the features (Table~\ref{tab:tune-perc-sel-features}).
XGBoost also favored using several hundred features with the exception of Laukiz2, for which only 14 (0.96\%) were selected.

% latex table best performance per learner
%\input{performance-best-per-learner.tex}
% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{specialtable}[H]
\setlength{\tabcolsep}{6.8mm}
%\centering
\caption{The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, lasso, and ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run. For exttt{regr.featureless}, the task is not applicable and was therefore removed.}
\label{tab:best-learner-perf}
\scalebox{0.9}{
\begin{tabular}{rlllrr}
\toprule
 & \textbf{Task} & \textbf{Model} & \textbf{Filter} & \textbf{RMSE} & \textbf{SE} \\
  \midrule
1 & NRI-VI & SVM & Info Gain & 27.915 & 18.970 \\
  2 & NRI & RF & Relief & 30.842 & 12.028 \\
  3 & HR & XGBoost & Info Gain & 31.165 & 15.025 \\
  4 & NRI & Lasso-MBO & No Filter & 31.165 & 15.025 \\
  5 & NRI & Ridge-MBO & No Filter & 31.165 & 15.025 \\
  6 & - & regr.featureless & No Filter & 31.165 & 15.025 \\
   \bottomrule
\end{tabular}
}
\end{specialtable}
\vspace{-12pt}

% latex table top 20 absolute performances
%\input{performance-svm-single-plot.tex}
% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{specialtable}[H]
\setlength{\tabcolsep}{18.7mm}
%\centering
\caption{Test fold performances in RMSE (p.p.) for learner SVM on the HR dataset without using a filter, showcasing performance variance on the fold level. For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot.}
\label{tab:svm-single-fold-perf}
\begin{tabular}{rrl}
\toprule
 & \textbf{RMSE} & \textbf{Test Plot} \\
  \midrule
1 & 28.12 & Laukiz1 \\
  2 & 54.26 & Laukiz2 \\
  3 & 9.00 & Luiando \\
  4 & 21.17 & Oiartzun \\
   \bottomrule
\end{tabular}
\end{specialtable}
\vspace{-12pt}

% latex table top 10 absolute performances
%\input{performance-top-10}
% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{specialtable}[H]
\setlength{\tabcolsep}{6mm}
%\centering
\caption{Best ten results among all learner--task--filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE).}
\label{tab:perf-top-10}
\begin{tabular}{rlllrr}
\toprule
 & \textbf{Task} & \textbf{Model} & \textbf{Filter} & \textbf{RMSE} & \textbf{SE} \\
  \midrule
1 & NRI-VI & SVM & Info Gain & 27.915 & 18.970 \\
  2 & NRI & SVM & CMIM & 28.044 & 19.101 \\
  3 & VI & SVM & Relief & 28.082 & 19.140 \\
  4 & NRI-VI & SVM & Borda & 28.102 & 19.128 \\
  5 & HR & SVM & CMIM & 28.119 & 19.123 \\
  6 & HR & SVM & MRMR & 28.119 & 19.123 \\
  7 & VI & SVM & Info Gain & 28.121 & 19.123 \\
  8 & NRI & SVM & PCA & 28.121 & 19.123 \\
  9 & HR-NRI & SVM & PCA & 28.121 & 19.123 \\
  10 & HR-NRI-VI & SVM & PCA & 28.121 & 19.123 \\
   \bottomrule
\end{tabular}
\end{specialtable}
\vspace{-12pt}

% latex table worst 10 absolute performances
%\input{performance-worst-10}
% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{specialtable}[H]
\setlength{\tabcolsep}{5.7mm}
%\centering
\caption{Worst ten results among all learner--task--filter combinations, sorted in decreasing order of RMSE (p.p.) and their respective standard error (SE).}
\label{tab:perf-worst-10}
\begin{tabular}{rlllrr}
\toprule
 & \textbf{Task} & \textbf{Model} & \textbf{Filter} & \textbf{RMSE} & \textbf{SE} \\
  \midrule
1 & VI & XGBoost & No Filter & 45.366 & 6.672 \\
  2 & HR & XGBoost & No Filter & 44.982 & 5.378 \\
  3 & VI & XGBoost & PCA & 44.539 & 8.187 \\
  4 & HR & XGBoost & PCA & 44.032 & 6.183 \\
  5 & NRI & XGBoost & PCA & 43.433 & 9.543 \\
  6 & HR-NRI & XGBoost & PCA & 43.220 & 2.557 \\
  7 & HR-NRI-VI & XGBoost & PCA & 41.076 & 9.862 \\
  8 & VI & RF & CMIM & 39.980 & 10.144 \\
  9 & VI & RF & Info Gain & 39.623 & 10.616 \\
  10 & NRI & XGBoost & Pearson & 39.492 & 11.548 \\
   \bottomrule
\end{tabular}
\end{specialtable}

% latex table percentage of selected features during tuning
%\input{tune-perc-sel-features.tex}
% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{specialtable}[H]
\setlength{\tabcolsep}{9.2mm}
%\centering
\caption{Selected feature portions during tuning for the best performing learner--filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted by plot name. "Features ({\#})" denotes the absolute number of features selected, and "Features ({\%})" refers to the percentage relative to the overall features available in the training sets for each plot (Laukiz1 = 1249, Laukiz2 = 1357, Luiando = 1507, Oiartzun = 1311). Results were estimated in a separate model tuning step, not within the main cross-validation comparison.}
\label{tab:tune-perc-sel-features}
\begin{tabular}{llrl}
  \toprule
\textbf{Learner} & \textbf{Test Plot} & \textbf{Features (\%)} & \textbf{Features (\#)} \\
  \midrule
\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.00245 & 1/1249 \\
   & Laukiz2 & 0.00359 & 1/1357 \\
   & Luiando & 0.12448 & 2/1507 \\
   & Oiartzun & 2.80356 & 37/1311 \\
  \midrule\multirow{4}{*}{\specialcell{SVM \\ Car}} & Laukiz1 & 16.76686 & 210/1249 \\
   & Laukiz2 & 40.77700 & 554/1357 \\
   & Luiando & 43.80604 & 661/1507 \\
   & Oiartzun & 81.23205 & 1065/1311 \\
  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 79.54091 & 994/1249 \\
   & Laukiz2 & 0.96545 & 14/1357 \\
   & Luiando & 66.27871 & 999/1507 \\
   & Oiartzun & 41.89759 & 550/1311 \\
   \bottomrule
\end{tabular}
\end{specialtable}

\vspace{-18pt}
% plot performance results
\begin{figure}[H]
	%\centering
	%\begin{center}
		\includegraphics[width=0.7\textwidth] {performance-results-1.pdf}
		\caption{Predictive performance in RMSE (p.p.) of models across tasks. Different feature sets are shown on the y-axis. Labels show the feature selection method (e.g., NF = no filter, Car = "Carscore", Info Gain = "Information Gain", Borda = "Borda").}\label{fig:perf-result}
	%\end{center}
\end{figure}
\vspace{-15pt}

% plot no filter vs all other filters for each model and task
\begin{figure}[H]
	%\centering
	%\begin{center}
		\includegraphics[width=0.67\textwidth] {filter-effect-all-vs-no-filter-1.pdf}
		\caption{Model performances in RMSE across all tasks, split up in facets, when using no filter method (blue dot) compared to any other filter method (red cross) for learners RF, SVM, and XGBoost~(XG).}\label{fig:filter-effects-no-filter}
	%\end{center}
\end{figure}
\vspace{-18pt}

% plot Borda vs all other filters for each model and task
\begin{figure}[H]
	%\centering
	%\begin{center}
		\includegraphics[width=0.67\textwidth] {filter-effect-all-vs-borda-filter-1.pdf}
		\caption{Predictive performances in RMSE (p.p.) when using the Borda filter method (blue dot) compared to any other filter (red cross) for each learner across all tasks.}\label{fig:filter-effects-borda}
	%\end{center}
\end{figure}

\subsection{Variable Importance}

\subsubsection*{Permutation-Based Variable Importance}

% Variable Importance
% - VI: Vogelmann indices are best
% - All: Clustering around the red edge (700 - 750 nm)
% - Overall highest: 1.7 RMSE decrease - not that much

\textls[-15]{The most important features for datasets HR and VI showed an average decrease in RMSE of 1.06 p.p. (HR, B65) and 1.93 p.p. (VI, Vogelmann2) when permuted (Figure~\ref{fig:fi-permut-vi-hr}).}
For the HR dataset, most (i.e., six out of ten) relevant features clustered around the infrared region (920~nm--1000~nm), while for VI, eight out of ten concentrate within the wavelength range of 700~nm--750~nm (the so\-/called \enquote{red edge}).
For HR, four features in the infrared region (920~nm--1000~nm) were identified by the model to be most important, being associated with a mean decrease in RMSE of around 1 p.p.
Overall, apart from the top five features, the vast majority of features showed only a small importance with average decreases in RMSE below 0.5 p.p.
\vspace{-10pt}
% start a new page without indent 4.6cm

\end{paracol}
\nointerlineskip
% permutation based var imp for datasets HR and VI
\begin{figure} [H]
\widefigure
	%\centering
	%\begin{center}
		\includegraphics[width=0.75\textwidth] {fi-permut-vi-hr-1.pdf}
		\caption{Variable importance for feature sets HR and VI: Mean decrease in RMSE for one\ hundred feature permutations using the SVM learner. The wavelength range on the x-axis matches the range of the hyperspectral sensor (400~nm--1000~nm). For each dataset, the ten most important features are highlighted as black dots and labeled by name. Gray dots represent features from importance rank~11 to last. The spectral signature (mean) of each plot was added as a reference on a normalized reflectance scale [0, 1] (secondary y-axis). VI features were decomposed into their individual formula parts, all instances being connected via dashed lines. Each VI feature is composed out of at least two instances.}\label{fig:fi-permut-vi-hr}
	%\end{center}
\end{figure}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\subsection{Predictive Performance}

The best overall performance achieved in this study (SVM with the \enquote{Info Gain} filter, RMSE~27.915 p.p.) has to be seen in the light of model overfitting (see Section~\ref{subsec:perf-plot-char}).
Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 p.p.
However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
Due to the apparent presence of model overfitting in this study, it is suggested that more training data representing a greater variety of situations are needed.
A model can only make robust predictions if it has learned relationships across the whole range of the response.
Hence, care should be taken when predicting on the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
However, when inspecting the fold\-/level performances, it can be concluded that the models performed reasonably well, predicting defoliation greater than 50\%, but they failed for lower levels.
This applied to all learners of this study.
In this study, the overall performance across all learners can be classified as ``poor'' given that only the SVM learner was able to substantially outperform the featureless learner \linebreak (Table~\ref{tab:best-learner-perf}).
It is worth noting that data quality issues may have affected model performances, as discussed below in detail  (Section~\ref{subsec:data-quality}).

\subsubsection{Model Differences}

An interesting finding is the strength of the SVM algorithm when comparing its predictive performance to its competitors (Table~\ref{tab:best-learner-perf}).
These cluster around a performance of 31 p.p., while SVM scored about 3 p.p. better than all other methods.
However, we refrain from comparing these results (both relatively and absolute) to other studies since many study design points have an influence on the final result (optimization strategy, data characteristics, feature selection methods, etc.).

A potential limiting factor in this study could be the upper limit of 70 iterations used for the XGBoost algorithm (hyperparameter \texttt{nrounds}), especially for feature sets including NRIs (Table~\ref{tab:hyperparameter_limits}).
This setting was a compromise between runtime and tuning space extension with the goal to work well for most feature sets.
It may be recommendable to increase this upper limit to a value closer to the number of features in the dataset in order to be able to exploit the full potential of this hyperparameter.

\subsubsection{Feature Set Differences}

One objective of this study was to determine whether expert-based and data-driven feature engineering have a positive influence on model performance.
With respect to Figure~\ref{fig:perf-result}, no overall positive or negative pattern related to specific feature sets was found that would be valid across all models.
The performance of RF and XGBoost on the VI feature set was around 4 to 6 p.p. lower than on others.
One reason could be the lack of coverage in the wavelength range between 810~nm and 1000~nm (Figure~\ref{fig:fi-permut-vi-hr}).
In addition, for all learners but SVM, a better performance was observed when NRI indices were included in the feature set (i.e., NRI-VI, HR-NRI, and HR-NRI-VI).

\subsection{Performance vs.\ Plot Characteristics}
\label{subsec:perf-plot-char}

The large differences in RMSE obtained on different test folds can be attributed to model overfitting (Table~\ref{tab:svm-single-fold-perf}).
An RMSE of 54.26 p.p. reveals the model's inability to predict tree defoliation on this plot (Laukiz2).
Laukiz2 differs highly in the distribution of the response variable defoliation compared to all other plots (Figure~\ref{fig:defol-distr}).
In the prediction scenario for Laukiz2, the model was trained on data containing mostly medium-to-high defoliation values and only few low ones.
This caused overfitting on the medium-to-high values, degrading the model's predictive performance in other scenarios.
When Laukiz2 was in the training set, the overall mean RMSE was reduced by up to 50\% with single fold performances as good as 9 p.p. RMSE (with Luiando as test set).

There was also no clear pattern in the percentage of features selected based on hyperparameter tuning (Table~\ref{tab:tune-perc-sel-features}).
The optimal value for the number of features (interpreted as a\ percentage of available features), which are selected from the ranked filter results, is determined by the internal surrogate learner of the MBO tuning method using the results from the previous tuning iterations.
Due to this iterative approach, MBO is in some ways able to evaluate how well a learning algorithm plays together with a certain amount of selected features and is subsequently able to adjust the number of variables to an optimal value.
In general, considering SVM's relative success, the use of at least a few hundred features from the combined feature set appears to be beneficial, or at least not harmful when the model's built-in regularization is capable of dealing with the resulting high-dimensional situation.

% filters always reduce computation time
Realizing early during hyperparameter optimization that only a few features are needed to reach adequate performances can reduce the overall computational runtime substantially.
Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially of models making use of hyperparameters that depend on the available number of features, such as RF with $m_\textrm{try}$.

Ultimately, the results of Table~\ref{tab:tune-perc-sel-features} should be taken with care, as they rely on single model--filter combinations and are subject to random variation.
More in-depth research is needed to investigate the effect of filters on criteria other than performance (such as runtime), leading to a multi-criteria optimization problem.

\subsection{Feature Selection Methods}

% - Effect varies across learners -> suggest to use FS even on feature sets with p < 100?
% - Filters can also have a negative impact
% - Filters are interesting for reducing runtime and improve interpretability, not only for performance
The usefulness of filters with respect to predictive performance in this study varied.
While the performance of some models (up to 5 p.p. for RF and XGBoost) was improved by specific filters, some models achieved a poorer performance with filters than without them (Figure~\ref{fig:filter-effects-no-filter}).
There was no pattern of specific filters consistently resulting in better scores.
Hence, it is recommended to test multiple filters in a study if it is intended to use filters.
While filters can improve the performance of models, they may be more appealing based on other aspects than performance. Reducing variables can reduce computational efforts in high-dimensional scenarios and may enhance the interpretability of models.
Filters are a lot cheaper to compute than wrapper methods, and the final feature subset selection can be integrated as an additional hyperparameter into the model optimization~stage.

% - Ensemble FS shows no advantage to simple methods -> always one simple filter which is better. Ensemble filters are good across many different datasets
Models that used the Borda ensemble method in this study did not perform better on average than models that used a single filter or no filter at all.
Ensemble methods have higher stability and robustness than single ones and have shown promising results in~\cite{drotar2017}.
Hence, their expected main advantage is stable performances across datasets with varying characteristics.
Single filter methods might yield better model performances on certain datasets but fail on others.
The fact that this study used multiple feature sets but only one dataset and tested many single filters could be a potential explanation of why, in all cases, a single filter outperformed the ensemble filter.
However, studies that use ensemble filters are still rare, and these are usually not compared against single filters~\cite{ghosh2019}.
In summary, in this study, Borda did\ not\ perform better than a randomly selected filter method.
More case studies applying ensemble filter methods are needed to verify this finding.
Nevertheless, ensemble filters can be a promising addition to an ML feature selection portfolio.

% - PCA shows similar performance but probably requires tuning of main components -> check runtime!
PCA, acting as a filter method, more often showed less than optimal results, especially for algorithms RF and XGBoost.
XGBoost in particular had substantial problems when using PCA as a filter method and accounted for four of the six worst results (Table~\ref{tab:perf-worst-10}).
However, PCA was able to reduce model fitting times substantially across all algorithms.
Depending on the use case, PCA can be an interesting option to reduce dimensionality while keeping runtime low.
However, information about the total number of features used by the model is lost when applying this technique.
Since filter scores only need to be calculated once for a given dataset in a benchmark setting, the runtime advantage of a PCA vs. filter methods might in fact be negligible in practice.

\subsection{Linking Feature Importance to Spectral Characteristics}

Unsurprisingly, the most important features for both HR and VI datasets were identified around the red edge of the spectra, specifically in the range of 680~nm to 750~nm.

This area has the highest ability to distinguish between reflectances related to a high density/high foliage density and thus the health status of vegetation and its respective counterpart~\cite{horler1983}.
However, four out of ten of the most important features of dataset HR are located between 920~nm and 1000~nm.
Looking at the spectral curves of the plots, apparent reflectance differences can be observed in this spectral range---especially for plot Oiartzun---which might explain why these features were considered important by the~model.

A possible explanation for the poor performances of most models scored on the VI dataset compared to all other feature sets could be the lack of features covering the area between 850~nm and 1000~nm (Figure~\ref{fig:fi-permut-vi-hr}).
The majority of VI features covers the range between 550~nm--800~nm.
Only one index (PWI) covers information in the range beyond 900~nm.

\subsection{Data Quality}
\label{subsec:data-quality}

Environmental datasets always come with some constraints that can have potential influence on the modeling process and its outcome.
Finding a suitable approach to extract the remote sensing information from each tree was a complex process.
Due to the reported geometric offset of up to one meter within the hyperspectral data, the risk of assigning a value to an observation that would actually refer to a different, possibly non-tree, pixel was reasonably high.
It was concluded that using a buffer radius of one meter can be a good compromise between the inclusion of information from too many surrounding trees and an under-coverage of the tree crown.
With the chosen radius, we are confident that we were able to map individual tree crowns while accounting for a possible geometric offset.
% The applied buffer only included a pixel value if the distance to the centroid of a pixel was smaller than the buffer radius (i.e., $<= 1m$).
This results in all cases in four contributing pixels (=four square meters) for the extraction of hyperspectral information for a given tree.
Even though no results showing the influence of different buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
Instead of using a buffer to extract the hyperspectral information, segmentation could have been considered.
However, this method would have required more effort for no clear added value in our view and would have moved the focus of this manuscript more to data preprocessing and away from feature selection methods.

Trees located within grid cells on the border of a plot are a notable exception where the exact number of pixels contributing to the observation's feature value may be reduced since the image was cropped to the plot's extent.
Cropping was applied to avoid the accidental inclusion of background data such as forest roads.
However, this effect was deemed to be of negligible importance.

The available hyperspectral data covered a wavelength between 400~nm and 1000~nm.
Hence, the spectral range of the shortwave infrared (SWIR) region is not covered in this study.
Given that this range is often used in forest health studies~\cite{hais2019}, e.g., when calculating the \ac{NDMI} index~\cite{gao1996}, this marks a clear limitation of the dataset at hand.

The dataset consists of in situ data collected during September 2016, which was matched against remote sensing data acquired at the end of September 2016.
A multi-temporal dataset consisting of in situ data from different phenology stages would possibly improve the achieved model performances.
However, this would also require the costly acquisition of hyperspectral data of these additional timestamps.

The R package \texttt{hsdar} was used for the calculation of vegetation indices~\cite{lehnert2016}.
All indices that could be calculated with the given spectral range of the data (400~nm--1000~nm) were used.
This means that even though Table~\ref{tab:vegindices} lists all indices available in the package, not all listed indices were used in this study.
Although this selection included a large number of indices, some possibly helpful indices might have been missed due to the restriction of the hyperspectral data.

Overall, the magnitude of uncertainty introduced by the mentioned effects during index derivation cannot be quantified.
Such limitations and uncertainties apply to most environmental studies and cannot be completely avoided.

\subsection{Practical Implications on Defoliation and Tree Health Mapping}
Even though this work has a strong methodological focus by comparing different benchmark settings on highly\ correlated feature sets, implications on tree health should be briefly discussed in the following.
Due to the outlined dataset issues in Section~\ref{subsec:data-quality}, which are mainly responsible for the resulting poor model performances, the trained models are not suited for practical use, e.g., to predict defoliation in unknown areas, due to the high mapping uncertainty.
However, the general approach of utilizing hyperspectral data to classify the health status of trees partly led to promising results.
For example, due to the narrow bandwidth of the hyperspectral sensor, small parts of the spectrum (mainly in the infrared region) were of higher importance to the models (e.g., see Figure~\ref{fig:fi-permut-vi-hr}), meaning that they helped the models to distinguish between low and high tree defoliation.
If spatial offset errors of the image data and possible background noise can be reduced (possibly by making use of image segmentation), we believe that model performances could be substantially enhanced.
Such improved models, starting around an RMSE of 20\% and less, should be able to provide added value to support the long-term monitoring of forest health and early detection of fungi-affected tree plots.
Nevertheless, overall the use of defoliation as a proxy for forest health should be critically reconsidered as it comes with various practical issues, starting from potential offsets during data collection, varying leaf density due to tree age, and differing effects between tree species, to  name just a few.

\subsection{Comparison to Other Studies}
% - no environmental studies use filter methods, some use FFS
% - only few studies analyze defoliation
% - most FS is done in bioinformatics

% defoliation studies and what data they used
While most defoliation studies operate on the plot level using coarser\ resolution multispectral satellite data~\cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
Among these, refs.~\cite{meng2018, kalin2019} used ground-level methods, such as \ac{ALS} and \ac{LiDAR}.

% defoliation studies and what
Studies focusing on tree-level defoliation mainly used ground-level methods, such as \ac{ALS} or \ac{LiDAR}~\cite{meng2018, kalin2019}.
Ref.~\cite{meng2018} used \ac{OLS} regression methods while~\cite{kalin2019} retrieving information from ground-level RGB photos using \ac{CNN}.
However, neither of them used spatial \ac{CV} for model assessment, and~\cite{kalin2019} did not perform \ac{FS}.
The authors of \cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
Study results indicated that spectral features were found to be most helpful for the model.
Incorporating such features (both spectral and structural) could be a possible enhancement for future works.
No studies were found to model defoliation caused by \textit{Diplodia sapinea (Fr.) Fuckel} with remote sensing data, and most studies focused on describing the tree conditions based on local sampling~\cite{hlebarska2018, kaya2019}.

% write about hyperspectral remote sensing
The field of (hyperspectral) remote sensing has had a strong focus on using RF for modeling in recent years~\cite{belgiu2016}.
However, in high-dimensional scenarios, tuning the parameter \texttt{\(m_{try}\)} becomes computationally expensive.
To account for this and the high dimensionality in general, studies used feature selection approaches, such as  semi-supervised feature extraction~\cite{xia2015}, wrapper methods~\cite{fassnacht2014, feng2016, georganos2018}, PCA, and adjusted feature selection~\cite{rochac2016}.
In general, applying feature selection methods on hyperspectral datasets has shown to be effective, regardless of the method used~\cite{pal2010, keller2016}.
However, no studies were found that made explicit use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing.
Potential reasons for this absence could be an easier programmatic access to wrapper methods and a higher general awareness of these compared to filter methods.
Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.

% RS studies with more/other algorithms than RF
When looking for remote sensing studies that compare multiple models, it turned out that these often operate in a low-dimensional predictor space~\cite{xu2019} or use wrapper methods explicitly~\cite{georganos2018}.

Refs.~\cite{shendryk2016, ludwig2019} are more similar in their methodology but focus on a different response variable (woody cover).
Ref.~\cite{shendryk2016} used machine learning with \ac{ALS} data to study dieback of trees in eucalyptus forests.
A grid search was used for hyperparameter tuning and \ac{FFS} for variable selection.
Ref.~\cite{ludwig2019} analyzed woody cover in South Africa using a spatial \ac{CV} and \ac{FS} approach~\cite{meyer2018} with an RF classifier.
Ref.~\cite{zandler2015} shows a similar setup; they used hyperspectral vegetation indices and a nested CV approach for performance estimation, and they estimated variable importance targeting woody biomass as the response.
In the results, lasso showed the best performance among the chosen methods.
However, the authors did not optimize the hyperparameters of RF, which makes a fair comparison problematic since the other models used internal optimization.
The discussion section of~\cite{zandler2015} lists additional studies that made use of shrinkage models for high-dimensional remote sensing~modeling.

In summary, no studies could be found that used filter methods for \ac{FS} or made use of \ac{NRI} indices in their work and had a relation to tree health.
This might relate to the fact that most environmental datasets are not high dimensional.
In fact, many studies use fewer than ten features, and issues related to correlations are often solved manually instead of relying on an automated approach.
These manual approaches might suffer from subjectivity and may limit the reproducibility of results.

Other fields (e.g., bioinformatics) encounter high-dimensional datasets more often.
Hence, more studies using (filter-based) feature selection approaches can be found in this \linebreak field~\cite{guo2019, radovic2017}.
However, bioinformatics differs conceptually in many ways from environmental modeling, and, therefore, no greater focus was put into comparing studies of this field.
The availability of high-dimensional feature sets will increase in the future due to higher temporal and spectral resolutions of sensors.
In addition, a high-spatial resolution comes with the possibility of calculating many textural features.
Hence, the ability to deal with high-dimensional datasets becomes more important, and unbiased robust approaches are needed.
We hope that this work and its methodology raise awareness about the application of filter methods to tackle high-dimensional problems in the environmental modeling~field.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This study analyzed the effectiveness of filter-based feature selection in improving various machine-learning models of defoliation of trees in northern Spain based on hyperspectral remote-sensing data.
Substantial differences in performance occurred depending on which feature selection and machine learning methods were combined.
SVM showed the most robust behavior across all highly\ correlated datasets and was able to predict the response variable of this study substantially better than other methods.

% RQ3: How are feature-selection methods influencing the predictive performance of the models?
Filter methods were able to improve the predictive performance on datasets in some instances, although there was no clear and systematic pattern.
Their effectiveness depends on the algorithm and the dataset characteristics.
Ensemble filter methods did not show a substantial improvement over individual filter methods in this study.

% RQ2: Does combining feature sets have an substantial effect on predictive performance?
% RQ1: Do different environmental feature sets show differences in performance when modeling defoliation at trees?
The addition of derived feature sets was, in most cases, able to improve predictive performance.
In contrast, feature sets that focused on only a small fraction of the available spectral range (i.e., dataset~VI) showed a worse performance than the ones that covered a\ wider range (400~nm--1000~nm; HR, NRI).
NRIs can be seen as a valuable addition to the optimization of predictive performance in the remote sensing of vegetation.

% RQ4: Which features are most important for the models and how can these be interpreted in an ecological context?
Features along the red-edge wavelength region were most important for models during prediction.
With respect to dedicated vegetation indices, all versions of the Vogelmann index were seen as the most important indices for the best performing SVM model.
This matches well with the actual purpose of these indices---they were invented to detect defoliation on sugar maple trees (\textit{Acer saccharum Marsh.}) caused by pear thrips (\textit{Taeniothrips inconsequens Uzel})~\cite{vogelmann1993}.
However, assessing feature importance for highly correlated features remains a challenging task.
Results might be biased and should be taken with care to avoid overgeneralizing from individual studies.

% What about the general potential of analyzing defol with hyperspectral data?
Finally, the potential of predicting defoliation with the given study design was rather limited with respect to the average RMSE of 28 p.p. scored by the best performing model.
More training data covering a wider range of defoliation values in a larger number of forest plantations are needed to train better models that can create more robust predictions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{Conceptualization, P.S. and A.B.; data curation, P.S. and E.I.; formal analysis, P.S., J.C., B.B., and A.B.; funding acquisition, A.B.; investigation, P.S.; methodology, P.S., J.M., J.C., and B.B.; project administration, A.B.; resources, E.I.; software, P.S.; supervision, B.B. and A.B.; validation, P.S. and A.B.; visualization, P.S.; writing---original draft, P.S.; Writing---review and editing, J.M., J.C., and A.B. All authors have read and agreed to the published version of the manuscript.}

\funding{This work was funded by the EU LIFE Healthy Forest project (LIFE14 ENV/ES/000179) and the German Scholars Organization/Carl Zeiss Foundation.}

\dataavailability{The data presented in this study are openly available in Zenodo at \url{https://doi.org/10.5281/zenodo.2635403} (accessed on 27 November 2021).}

\acknowledgments{We acknowledge support by the German Research Foundation and the Open Access Publication Fund of the Thueringer Universitaets- und Landesbibliothek Jena Projekt-Nr. 433052568.}

\conflictsofinterest{The authors declare no conflict of interest. The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results.}

%% Optional
% \sampleavailability{Samples of the compounds ... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\abbreviations{Abbreviations}{
	The following abbreviations are used in this manuscript:\\

	\noindent
	\begin{tabular}{@{}ll}
		AGB     & above-ground biomass                                    \\
		ALE     & accumulated local effects                               \\
		ALS     & airborne laser scanning                                 \\
		ANN     & artificial neural network                               \\
		AUROC   & area under the receiver operating characteristics curve \\
		BRT     & boosted regression trees                                \\
		CART    & classification and regression trees                     \\
		CNN     & convolutional neural networks                           \\
		CV      & cross-validation                                        \\
		DAP     & digital aerial photogrammetry                           \\
		ENM     & environmental niche modeling                            \\
		FFS     & forward feature selection                               \\
		FPR     & false positive rate                                     \\
		FS      & feature selection                                       \\
		GAM     & generalized additive model                              \\
		GBM     & gradient boosting machine                               \\
		GLM     & generalized linear model                                \\
		ICGC    & Institut Cartografic i Geologic de Catalunya            \\
		IQR     & interquartile range                                     \\
		LiDAR   & light detection and ranging                             \\
		LOWESS  & locally weighted scatter plot smoothing                 \\
		MARS    & multivariate adaptive regression splines                \\
		MBO     & model-based optimization                                \\
		MEM     & maximum entropy model                                   \\
		ML      & machine learning                                        \\
		NDII    & normalized difference infrared index                    \\
%		NDMI    & normalized difference moisture index                    \\
%		NIR     & near-infrared                                           \\
%		NRI     & normalized ratio index                                  \\
%		OLS     & ordinary least squares                                  \\
%		OMNBR   & optimized multiple narrow-band reflectance              \\
%		PCA     & principal component analysis                            \\
%		PDP     & partial dependence plots                                \\
%		PISR    & potential incoming solar radiation                      \\
%		PLS     & partial least-squares                                   \\
%		POV     & proportion of variance explained                        \\
%		RBF     & radial basis function                                   \\
%		RF      & random forest                                           \\
%		RMSE    & root mean square error                                  \\
%		RR      & ridge regression                                        \\
%		RSS     & residual sum of squares                                 \\
%		SAR     & synthetic aperture radar                                \\
%		SDM     & species distribution modeling                           \\
%		SMBO    & sequential-based model optimization                     \\
%		SVM     & support vector machine                                  \\
%		TPR     & true positive rate                                      \\
%		VI      & vegetation index                                        \\
%		XGBoost & extreme gradient boosting
	\end{tabular}

	\noindent
	\begin{tabular}{@{}ll}
%		AGB     & above-ground biomass                                    \\
%		ALE     & accumulated local effects                               \\
%		ALS     & airborne laser scanning                                 \\
%		ANN     & artificial neural network                               \\
%		AUROC   & area under the receiver operating characteristics curve \\
%		BRT     & boosted regression trees                                \\
%		CART    & classification and regression trees                     \\
%		CNN     & convolutional neural networks                           \\
%		CV      & cross-validation                                        \\
%		DAP     & digital aerial photogrammetry                           \\
%		ENM     & environmental niche modeling                            \\
%		FFS     & forward feature selection                               \\
%		FPR     & false positive rate                                     \\
%		FS      & feature selection                                       \\
%		GAM     & generalized additive model                              \\
%		GBM     & gradient boosting machine                               \\
%		GLM     & generalized linear model                                \\
%		ICGC    & Institut Cartografic i Geologic de Catalunya            \\
%		IQR     & interquartile range                                     \\
%		LiDAR   & light detection and ranging                             \\
%		LOWESS  & locally weighted scatter plot smoothing                 \\
%		MARS    & multivariate adaptive regression splines                \\
%		MBO     & model-based optimization                                \\
%%		MEM     & maximum entropy model                                   \\
%		ML      & machine learning                                        \\
%		NDII    & normalized difference infrared index                    \\
		NDMI    & normalized difference moisture index                    \\
		NIR     & near-infrared                                           \\
		NRI     & normalized ratio index                                  \\
		OLS     & ordinary least squares                                  \\
		OMNBR   & optimized multiple narrow-band reflectance              \\
		PCA     & principal component analysis                            \\
		PDP     & partial dependence plots                                \\
		PISR    & potential incoming solar radiation                      \\
		PLS     & partial least-squares                                   \\
		POV     & proportion of variance explained                        \\
		RBF     & radial basis function                                   \\
		RF      & random forest                                           \\
		RMSE    & root mean square error                                  \\
		RR      & ridge regression                                        \\
		RSS     & residual sum of squares                                 \\
		SAR     & synthetic aperture radar                                \\
		SDM     & species distribution modeling                           \\
		SMBO    & sequential-based model optimization                     \\
		SVM     & support vector machine                                  \\
		TPR     & true positive rate                                      \\
		VI      & vegetation index                                        \\
		XGBoost & extreme gradient boosting
	\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%\pagebreak
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section{}\label{appendixA}
\subsection{}

\vspace{-12pt}
\begin{figure} [H]
	%\begin{center}
		\includegraphics[width=0.7\textwidth] {correlation-filter-nri-1.pdf}
		\caption{Spearman correlations of NRI feature rankings obtained with different filters. Filter names refer to the nomenclature used by the \texttt{mlr} R package. Underscores in names divide the terminology into their upstream R package and the actual filter name.}\label{fig:correlation-filters}
	%\end{center}
\end{figure}

%\pagebreak

\subsection{}
\vspace{-12pt}
\begin{specialtable}[H]
\setlength{\tabcolsep}{5mm}
	%\centering
	\caption[H]{Hyperparameter ranges and types for each model.
		Hyperparameter notations from the respective R packages are shown.}
	\begingroup\scriptsize
	\begin{tabularx}{.75\textwidth}{llXXrXX}
		\toprule
		\textbf{\specialcell{Model                                              \\ (Package)}}                & \textbf{Hyperparameter}              & \textbf{Type}    & \textbf{Start}     & \textbf{End}      & \textbf{Default} \\
	\midrule
		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_{try}$}    & dbl & 0         & 0.5      & -       \\
		 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
		 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
		\midrule
		\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
		 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
		\midrule
		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 70      & -       \\
		 & \texttt{colsample\_bytree}  & dbl & 0.6      & 1       & 1   \\
		 & \texttt{subsample}          & dbl & 0.6      & 1       & 1   \\
		 & \texttt{max\_depth}         & int & 3        & 15      & 6   \\
		 & \texttt{gamma}              & int & 0.05     & 10      & 0   \\
		 & \texttt{eta}                & dbl & 0.1      & 1       & 0.3 \\
		 & \texttt{min\_child\_weight} & int & 1        & 7       & 1   \\
		\bottomrule
	\end{tabularx}
	\endgroup
	\label{tab:hyperparameter_limits}
\end{specialtable}
\vspace{-6pt}
%\pagebreak

\subsection{}
\vspace{-12pt}
% \begingroup\scriptsize
% \setlength\tabcolsep{4pt}  % default value: 6pt
% vertical spacing between rows (default is 1)
\renewcommand{\thespecialtable}{A\arabic{specialtable}}
\begin{specialtable}[H]
\setlength{\tabcolsep}{10.4mm}
{
	\caption{List of available vegetation indices in the \texttt{hsdar} package.}
	\begin{tabular}{lll}\toprule
		% \begin{supertabular}{0.3\textwidth}{lll}
		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference~}
         \\		\midrule
		Boochs          & $D_{703}$                                                                                            &~\cite{boochs1990}             \\
		Boochs2         & $D_{720}$                                                                                            &~\cite{boochs1990}             \\
		CAI             & $0.5 \times (R_{2000} + R_{2200}) -R_{2100}$                                                         &~\cite{nagler2003}             \\
		\midrule
		CARI            & $a = (R_{700}-R_{550}) / 150$                                                                        &~\cite{walthall1994}           \\
		& $b = R_{550}-(a\times 550)$                                                                          &                               \\
		& $\frac{R_{700}\,\times\, | (a\,\times\,670\,+\,R_{670}\,+\,b)}{R_{670}\,\times\,(a^2+1)| ^{0.5}}$                       &                               \\
		\midrule
		Carter          & $R_{695}/R_{420}$                                                                                    &~\cite{carter1994}             \\
		Carter2         & $R_{695}/R_{760}$                                                                                    &~\cite{carter1994}             \\
		Carter3         & $R_{605}/R_{760}$                                                                                    &~\cite{carter1994}             \\
		Carter4         & $R_{710}/R_{760}$                                                                                    &~\cite{carter1994}             \\
		Carter5         & $R_{695}/R_{670}$                                                                                    &~\cite{carter1994}             \\
		Carter6         & $R_{550}$                                                                                            &~\cite{carter1994}             \\
		CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                    &~\cite{zarco-tejada2003}      \\
		CI~2             & $R_{760}/R_{700}-1$                                                                                  &~\cite{gitelson2003}           \\
		ClAInt          & $\int_{600 nm}^{735 nm} R$                                                                           &~\cite{oppelt2004}             \\
		CRI1            & $1/R_{515}-1/R_{550}$                                                                                &~\cite{gitelson2003}           \\
		CRI2            & $1/R_{515}-1/R_{770}$                                                                                &~\cite{gitelson2003}           \\
		CRI3            & $1/R_{515}-1/R_{550}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
		CRI4            & $1/R_{515}-1/R_{700}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
		D1              & $D_{730}/D_{706}$                                                                                    &~\cite{zarco-tejada2003}      \\
		D2              & $D_{705}/D_{722}$                                                                                    &~\cite{zarco-tejada2003}      \\
		Datt            & $(R_{850}-R_{710})/(R_{850}-R_{680})$                                                                &~\cite{datt1999}               \\
		Datt2           & $R_{850}/R_{710}$                                                                                    &~\cite{datt1999}               \\
		Datt3           & $D_{754}/D_{704}$                                                                                    &~\cite{datt1999}               \\
		Datt4           & $R_{672}/(R_{550} \times R_{708})$                                                                   &~\cite{datt1998}               \\
		Datt5           & $R_{672}/R_{550}$                                                                                    &~\cite{datt1998}               \\
		Datt6           & $(R_{860})/(R_{550}\times R_{708})$                                                                  &~\cite{datt1998}               \\
%		Datt7           & $(R_{860} - R_{2218})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
%		Datt8           & $(R_{860} - R_{1788})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
%		DD              & $(R_{749}-R_{720})-(R_{701}-R_{672})$                                                                &~\cite{maire2004}              \\
%		DDn             & $2\times (R_{710}-R_{660}-R_{760})$                                                                  &~\cite{lemaire2008}            \\
%		DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                        &~\cite{zarco-tejada2003}      \\
%		DWSI1           & $R_{80}/R_{1660}$                                                                                    &~\cite{apan2004}               \\
%		DWSI2           & $R_{1660}/R_{550}$                                                                                   &~\cite{apan2004}               \\
%		DWSI3           & $R_{1660}/R_{680}$                                                                                   &~\cite{apan2004}               \\
%		DWSI4           & $R_{550}/R_{680}$                                                                                    &~\cite{apan2004}               \\
%		DWSI5           & $(R_{800} + R_{550})/(R_{1660} + R_{680})$                                                           &~\cite{apan2004}               \\
%		EGFN            & $\frac{(\max(D_{650:750})-\max(D_{500:550}))}{(\max(D_{650:750})+\max(D_{500:550}))}$                &~\cite{penuelas1994}           \\
%		EGFR            & $\max(D_{650:750})/\max(D_{500:550})$                                                                &~\cite{penuelas1994}           \\
%		EVI             & $\frac{2.5\times ((R_{800}-R_{670}) }{ (R_{800}-(6\times R_{670})-(7.5\times R_{475})+1)}$           &~\cite{huete1997}             \\
%		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$**                                                    &~\cite{wu2014}                 \\
%		GI              & $R_{554}/R_{677}$                                                                                    &~\cite{smith1995}              \\
%		Gitelson        & $1/R_{700}$                                                                                          &~\cite{gitelson1999}           \\
%		Gitelson2       & $(R_{750}-R_{800}/R_{695}-R_{740})-1$                                                                &~\cite{gitelson2003}           \\
%		GMI1            & $R_{750}/R_{550}$                                                                                    &~\cite{gitelson2003}           \\
%		GMI2            & $R_{750}/R_{700}$                                                                                    &~\cite{gitelson2003}           \\
%		Green NDVI      & $\frac{R_{800}-R_{550}}{R_{800}+R_{550}}$                                                            &~\cite{gitelson1996}           \\
%		LWVI\_1         & $\frac{(R_{1094}-R_{983})}{(R_{1094}+R_{983})}$                                                      &~\cite{galvao2005}             \\
%		LWVI\_2         & $\frac{R_{1094}-R_{1205}}{R_{1094}+R_{1205}}$                                                        &~\cite{galvao2005}             \\
%		Maccioni        & $\frac{R_{780}-R_{710})}{R_{780}-R_{680}}$                                                           &~\cite{maccioni2001}           \\
%		\midrule
%		MCARI           & \parbox{5.5cm}{$((R_{700}-R_{670})-0.2\times (R_{700}-R_{550})) \times (R_{700}/R_{670})$}           &~\cite{daughtry2000}           \\
%		\midrule
%		MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}          &~\cite{wu2008}                \\
%		\midrule
%		mND705          & $\frac{(R_{750}-R_{705})}{R_{750}+R_{705}-2\times R_{445}}$                                          &~\cite{sims2002}              \\
%		mNDVI           & $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$                                         &~\cite{sims2002}              \\
%		MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                            &~\cite{hernandez-clemente2011} \\
%		MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                 &~\cite{qi1994}                 \\
%		MSI             & $\frac{R_{1600}}{R_{817}}$                                                                           &~\cite{hunt1989}               \\
%		mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                            &~\cite{sims2002}              \\
%		mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                               &~\cite{chen1996}               \\
%		mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                            &~\cite{sims2002}              \\
%		MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                            &~\cite{dash2007}               \\
%		\midrule
%		MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}             &~\cite{haboudane2002}          \\
%		\midrule
%		NDLI            & $\frac{log(1/R_{1754}) - log(1/R_{1680})}{log(1/R_{1754}) + log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
%		NDNI            & $\frac{log(1/R_{1510}) - log(1/R_{1680})}{log(1/R_{1510}) + log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
%		NDVI            & $\frac{R_{800}-R_{680}}{R_{800}+R_{680}}$                                                            &~\cite{tucker1979}             \\
%		NDVI2           & $\frac{R_{750}-R_{705}}{R_{750}+R_{705}}$                                                            &~\cite{gitelson1994}           \\
%		NDVI3           & $\frac{R_{682}-R_{553}}{R_{682}+R_{553}}$                                                            &~\cite{guanter2005}            \\
%		NDWI            & $\frac{R_{860}-R_{1240}}{R_{860}+R_{1240}}$                                                          &~\cite{gao1996}                \\
%		NPCI            & $\frac{R_{680}-R_{430}}{R_{680}+R_{430}}$                                                            &~\cite{penuelas1994}           \\
%		OSAVI           & $\frac{(1+0.16) \times (R_{800}-R_{670})}{R_{800}+R_{670}+0.16 }$                                    &~\cite{rondeaux1996}           \\
%		OSAVI2          & $\frac{(1+0.16)\times (R_{750}-R_{705})}{R_{750}+R_{705}+0.16) }$                                    &~\cite{wu2008}                \\
%		PARS            & $\frac{R_{746}}{R_{513}}$                                                                            &~\cite{chappelle1992}          \\
%		PRI             & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            &~\cite{gamon1992}              \\
%		PRI\_norm       & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                &~\cite{zarco-tejada2013}      \\
%		PRI*CI2         & $PRI*CI2$                                                                                            &~\cite{garrity2011}            \\
%		PSRI            & $\frac{R_{678}-R_{500}}{R_{750}}$                                                                    &~\cite{merzlyak1999}           \\
%		PSSR            & $\frac{R_{800}}{R_{635}}$                                                                            &~\cite{blackburn1998}          \\
%		PSND            & $\frac{R_{800}-R_{470}}{R_{800}-R_{470})}$                                                           &~\cite{blackburn1998}          \\
%		PWI             & $\frac{R_{900}}{R_{970}}$                                                                            &~\cite{penuelas1997}           \\
%		RDVI            & $\frac{R_{800}-R_{670}}{ \sqrt{R_{800}+R_{670}}}$                                                    &~\cite{roujean1995}            \\
%		\midrule
%		REP\_LE         & \parbox{3.8cm}{Red-edge position through linear extrapolation}                                       &~\cite{cho2006}                \\
%		\midrule
%		REP\_Li         & $R_{re}=\frac{R_{670}+R_{780})}{2}$                                                                  &~\cite{guyot1988}              \\
%		& $\frac{700 + 40 \times ((R_{re} -R_{700})}{(R_{740}-R_{700}))}$                                      &                               \\
%		\midrule
%		SAVI            & $\frac{(1+L)\times (R_{800}-R_{670})}{(R_{800}+R_{670}+L)}$                                          &~\cite{huete1988}              \\
%		SIPI            & $\frac{R_{800}-R_{445}}{R_{800}-R_{680}}$                                                            &~\cite{penuelasj.1995}         \\
%		\midrule
%		SPVI            & \parbox{3.8cm}{$0.4 \times 3.7 \times (R_{800}-R_{670})-1.2 \times ((R_{530}-R_{670})^2)^{0.5}$}     &~\cite{vincini2006}            \\
%		\midrule
%		SR              & $\frac{R_{800}}{R_{680}}$                                                                            &~\cite{jordan1969}             \\
%		SR1             & $\frac{R_{750}}{R_{700}}$                                                                            &~\cite{gitelson1997}           \\
%		SR2             & $\frac{R_{752}}{R_{690}}$                                                                            &~\cite{gitelson1997}           \\
%		SR3             & $\frac{R_{750}}{R_{550}}$                                                                            &~\cite{gitelson1997}           \\
%		SR4             & $\frac{R_{700}}{R_{670}}$                                                                            &~\cite{mcmurtrey1994}          \\
%		SR5             & $\frac{R_{675}}{R_{700}}$                                                                            &~\cite{chappelle1992}          \\
%		SR6             & $\frac{R_{750}}{R_{710}}$                                                                            &~\cite{zarco-tejada1999}       \\
%		SR7             & $\frac{R_{440}}{R_{690}}$                                                                            &~\cite{lichtenthaler1996}      \\
%		SR8             & $\frac{R_{515}}{R_{550}}$                                                                            &~\cite{hernandez-clemente2012} \\
%		SRPI            & $\frac{R_{430}}{R_{680}}$                                                                            &~\cite{penuelasj.1995}         \\
%		SRWI            & $\frac{R_{850}}{R_{1240}}$                                                                           &~\cite{zarco-tejada2003}      \\
%		Sum\_Dr1        & $\sum_{i=626}^{795} D1_i$                                                                            &~\cite{elvidge1995}            \\
%		Sum\_Dr2        & $\sum_{i=680}^{780} D1_i$                                                                            &~\cite{filella1994}            \\
%		SWIR FI         & $\frac{R_{2133}^2}{R_{2225} \times R_{2209}^3}$                                                      &~\cite{levin2007}              \\
%		\midrule
%		SWIR LI         & \parbox{3.8cm}{$3.87  \times (R_{2210} - R_{2090}) - 27.51 \times (R_{2280} - R_{2090}) - 0.2$}      &~\cite{lobell2001}             \\
%		\midrule
%		SWIR SI         & \parbox{3.8cm}{$-41.59 \times (R_{2210} - R_{2090}) + 1.24 \times (R_{2280} - R_{2090}) + 0.64 $}    &~\cite{lobell2001}             \\
%		\midrule
%		SWIR VI         & \parbox{3.8cm}{$37.72  \times (R_{2210} - R_{2090}) + 6.27 \times (R_{2280} - R_{2090}) + 0.57$}     &~\cite{lobell2001}             \\
%		\midrule
%		TCARI           & \parbox{3.8cm}{$3*((R_{700}-R_{670})-0.2\times R_{700}-R_{550})\times (R_{700}/R_{670}))$}           &~\cite{haboudane2002}          \\
%		\midrule
%		TCARI/OSAVI     & TCARI/OSAVI                                                                                          &~\cite{haboudane2002}          \\
%		\midrule
%		TCARI2          & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} &~\cite{wu2008}                \\
%		\midrule
%		TCARI2/OSAVI2   & TCARI2/OSAVI2                                                                                        &~\cite{wu2008}                \\
%		\midrule
%		TGI             & \parbox{3.8cm}{$-0.5 (190 (R_{670} - R_{550} ) - 120 (R_{670} - R_{480}))$}                          &~\cite{hunt2013}               \\
%		\midrule
%		TVI             & \parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}              &~\cite{broge2001}              \\
%		\midrule
%		Vogelmann       & $\frac{R_{740}}{R_{720}}$                                                                            &~\cite{vogelmann1993}          \\
%		Vogelmann2      & $\frac{R_{734}-R_{747}}{R_{715}+R_{726})}$                                                           &~\cite{vogelmann1993}          \\
%		Vogelmann3      & $\frac{D_{715}}{D_{705}}$                                                                            &~\cite{vogelmann1993}          \\
%		Vogelmann4      & $\frac{R_{734}-R_{747}}{R_{715}+R_{720}}$                                                            &~\cite{vogelmann1993}     \\
\bottomrule
	\end{tabular}
	\label{tab:vegindices}}
	\end{specialtable}

\begin{specialtable}[H]\ContinuedFloat
\setlength{\tabcolsep}{7.5mm}
{
	\caption{{\em Cont.}}
	\begin{tabular}{lll}\toprule
		% \begin{supertabular}{0.3\textwidth}{lll}
		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference~*}         \\		\midrule
%		Boochs          & $D_{703}$                                                                                            &~\cite{boochs1990}             \\
%		Boochs2         & $D_{720}$                                                                                            &~\cite{boochs1990}             \\
%		CAI             & $0.5 \times (R_{2000} + R_{2200}) -R_{2100}$                                                         &~\cite{nagler2003}             \\
%		\midrule
%		CARI            & $a = (R_{700}-R_{550}) / 150$                                                                        &~\cite{walthall1994}           \\
%		& $b = R_{550}-(a\times 550)$                                                                          &                               \\
%		& $\frac{R_{700}\times | (a\times 670+R_{670}+b)}{R_{670}\times(a^2+1)| ^{0.5}}$                       &                               \\
%		\midrule
%		Carter          & $R_{695}/R_{420}$                                                                                    &~\cite{carter1994}             \\
%		Carter2         & $R_{695}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter3         & $R_{605}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter4         & $R_{710}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter5         & $R_{695}/R_{670}$                                                                                    &~\cite{carter1994}             \\
%		Carter6         & $R_{550}$                                                                                            &~\cite{carter1994}             \\
%		CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                    &~\cite{zarco-tejada2003}      \\
%		CI~2             & $R_{760}/R_{700}-1$                                                                                  &~\cite{gitelson2003}           \\
%		ClAInt          & $\int_{600 nm}^{735 nm} R$                                                                           &~\cite{oppelt2004}             \\
%		CRI1            & $1/R_{515}-1/R_{550}$                                                                                &~\cite{gitelson2003}           \\
%		CRI2            & $1/R_{515}-1/R_{770}$                                                                                &~\cite{gitelson2003}           \\
%		CRI3            & $1/R_{515}-1/R_{550}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
%		CRI4            & $1/R_{515}-1/R_{700}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
%		D1              & $D_{730}/D_{706}$                                                                                    &~\cite{zarco-tejada2003}      \\
%		D2              & $D_{705}/D_{722}$                                                                                    &~\cite{zarco-tejada2003}      \\
%		Datt            & $(R_{850}-R_{710})/(R_{850}-R_{680})$                                                                &~\cite{datt1999}               \\
%		Datt2           & $R_{850}/R_{710}$                                                                                    &~\cite{datt1999}               \\
%		Datt3           & $D_{754}/D_{704}$                                                                                    &~\cite{datt1999}               \\
%		Datt4           & $R_{672}/(R_{550} \times R_{708})$                                                                   &~\cite{datt1998}               \\
%		Datt5           & $R_{672}/R_{550}$                                                                                    &~\cite{datt1998}               \\
%		Datt6           & $(R_{860})/(R_{550}\times R_{708})$                                                                  &~\cite{datt1998}               \\
		Datt7           & $(R_{860} - R_{2218})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
		Datt8           & $(R_{860} - R_{1788})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
		DD              & $(R_{749}-R_{720})-(R_{701}-R_{672})$                                                                &~\cite{maire2004}              \\
		DDn             & $2\times (R_{710}-R_{660}-R_{760})$                                                                  &~\cite{lemaire2008}            \\
		DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                        &~\cite{zarco-tejada2003}      \\
		DWSI1           & $R_{80}/R_{1660}$                                                                                    &~\cite{apan2004}               \\
		DWSI2           & $R_{1660}/R_{550}$                                                                                   &~\cite{apan2004}               \\
		DWSI3           & $R_{1660}/R_{680}$                                                                                   &~\cite{apan2004}               \\
		DWSI4           & $R_{550}/R_{680}$                                                                                    &~\cite{apan2004}               \\
		DWSI5           & $(R_{800} + R_{550})/(R_{1660} + R_{680})$                                                           &~\cite{apan2004}               \\
		EGFN            & $\frac{(\max(D_{650:750})\,-\,\max(D_{500:550}))}{(\max(D_{650:750})\,+\,\max(D_{500:550}))}$                &~\cite{penuelas1994}           \\
		EGFR            & $\max(D_{650:750})/\max(D_{500:550})$                                                                &~\cite{penuelas1994}           \\
		EVI             & $\frac{2.5\,\times\, (R_{800}\,-\,R_{670}) }{ (R_{800}\,-\,(6\,\times\, R_{670})\,-\,(7.5\,\times\, R_{475})\,+\,1)}$           &~\cite{huete1997}             \\
		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$  &~\cite{wu2014}                 \\
		GI              & $R_{554}/R_{677}$                                                                                    &~\cite{smith1995}              \\
		Gitelson        & $1/R_{700}$                                                                                          &~\cite{gitelson1999}           \\
		Gitelson2       & $(R_{750}-R_{800}/R_{695}-R_{740})-1$                                                                &~\cite{gitelson2003}           \\
		GMI1            & $R_{750}/R_{550}$                                                                                    &~\cite{gitelson2003}           \\
		GMI2            & $R_{750}/R_{700}$                                                                                    &~\cite{gitelson2003}           \\
		Green NDVI      & $\frac{R_{800}\,-\,R_{550}}{R_{800}\,+\,R_{550}}$                                                            &~\cite{gitelson1996}           \\
		LWVI\_1         & $\frac{(R_{1094}\,-\,R_{983})}{(R_{1094}\,+\,R_{983})}$                                                      &~\cite{galvao2005}             \\
		LWVI\_2         & $\frac{R_{1094}\,-\,R_{1205}}{R_{1094}\,+\,R_{1205}}$                                                        &~\cite{galvao2005}             \\
		Maccioni        & $\frac{R_{780}\,-\,R_{710})}{R_{780}\,-\,R_{680}}$                                                           &~\cite{maccioni2001}           \\
		\midrule
		MCARI           & \parbox{5.5cm}{$((R_{700}-R_{670})-0.2\times (R_{700}-R_{550})) \times (R_{700}/R_{670})$}           &~\cite{daughtry2000}           \\
		\midrule
		MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}          &~\cite{wu2008}                \\
		\midrule
		mND705          & $\frac{(R_{750}\,-\,R_{705})}{R_{750}\,+\,R_{705}\,-\,2\,\times\, R_{445}}$                                          &~\cite{sims2002}              \\
		mNDVI           & $\frac{(R_{800}\,-\,R_{680})}{R_{800}\,+\,R_{680}\,-\,2 \,\times\, R_{445}}$                                         &~\cite{sims2002}              \\
		MPRI            & $\frac{R_{515}\,-\,R_{530}}{R_{515}\,+\,R_{530}}$                                                            &~\cite{hernandez-clemente2011} \\
		MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                 &~\cite{qi1994}                 \\
		MSI             & $\frac{R_{1600}}{R_{817}}$                                                                           &~\cite{hunt1989}               \\
		mSR             & $\frac{R_{800}\,-\,R_{445}}{R_{680}\,-\,R_{445}}$                                                            &~\cite{sims2002}              \\
		mSR2            & $\frac{(R_{750}/R_{705})\,-\,1}{R_{750}/R_{705}\,+\,1)^{0.5}}$                                               &~\cite{chen1996}               \\
		mSR705          & $\frac{R_{750}\,-\,R_{445}}{R_{705}\,-\,R_{445}}$                                                            &~\cite{sims2002}              \\
		MTCI            & $\frac{R_{754}\,-\,R_{709}}{R_{709}\,-\,R_{681}}$                                                            &~\cite{dash2007}               \\
		\midrule
		MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}             &~\cite{haboudane2002}          \\
		\midrule
		NDLI            & $\frac{log(1/R_{1754}) \,-\, log(1/R_{1680})}{log(1/R_{1754}) \,+\, log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
		NDNI            & $\frac{log(1/R_{1510}) \,-\, log(1/R_{1680})}{log(1/R_{1510}) \,+\, log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
		NDVI            & $\frac{R_{800}\,-\,R_{680}}{R_{800}\,+\,R_{680}}$                                                            &~\cite{tucker1979}             \\
%		NDVI2           & $\frac{R_{750}-R_{705}}{R_{750}+R_{705}}$                                                            &~\cite{gitelson1994}           \\
%		NDVI3           & $\frac{R_{682}-R_{553}}{R_{682}+R_{553}}$                                                            &~\cite{guanter2005}            \\
%		NDWI            & $\frac{R_{860}-R_{1240}}{R_{860}+R_{1240}}$                                                          &~\cite{gao1996}                \\
%		NPCI            & $\frac{R_{680}-R_{430}}{R_{680}+R_{430}}$                                                            &~\cite{penuelas1994}           \\
%		OSAVI           & $\frac{(1+0.16) \times (R_{800}-R_{670})}{R_{800}+R_{670}+0.16 }$                                    &~\cite{rondeaux1996}           \\
%		OSAVI2          & $\frac{(1+0.16)\times (R_{750}-R_{705})}{R_{750}+R_{705}+0.16) }$                                    &~\cite{wu2008}                \\
%		PARS            & $\frac{R_{746}}{R_{513}}$                                                                            &~\cite{chappelle1992}          \\
%		PRI             & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            &~\cite{gamon1992}              \\
%		PRI\_norm       & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                &~\cite{zarco-tejada2013}      \\
%		PRI*CI2         & $PRI*CI2$                                                                                            &~\cite{garrity2011}            \\
%		PSRI            & $\frac{R_{678}-R_{500}}{R_{750}}$                                                                    &~\cite{merzlyak1999}           \\
%		PSSR            & $\frac{R_{800}}{R_{635}}$                                                                            &~\cite{blackburn1998}          \\
%		PSND            & $\frac{R_{800}-R_{470}}{R_{800}-R_{470})}$                                                           &~\cite{blackburn1998}          \\
%		PWI             & $\frac{R_{900}}{R_{970}}$                                                                            &~\cite{penuelas1997}           \\
%		RDVI            & $\frac{R_{800}-R_{670}}{ \sqrt{R_{800}+R_{670}}}$                                                    &~\cite{roujean1995}            \\
%		\midrule
%		REP\_LE         & \parbox{3.8cm}{Red-edge position through linear extrapolation}                                       &~\cite{cho2006}                \\
%		\midrule
%		REP\_Li         & $R_{re}=\frac{R_{670}+R_{780})}{2}$                                                                  &~\cite{guyot1988}              \\
%		& $\frac{700 + 40 \times ((R_{re} -R_{700})}{(R_{740}-R_{700}))}$                                      &                               \\
%		\midrule
%		SAVI            & $\frac{(1+L)\times (R_{800}-R_{670})}{(R_{800}+R_{670}+L)}$                                          &~\cite{huete1988}              \\
%		SIPI            & $\frac{R_{800}-R_{445}}{R_{800}-R_{680}}$                                                            &~\cite{penuelasj.1995}         \\
%		\midrule
%		SPVI            & \parbox{3.8cm}{$0.4 \times 3.7 \times (R_{800}-R_{670})-1.2 \times ((R_{530}-R_{670})^2)^{0.5}$}     &~\cite{vincini2006}            \\
%		\midrule
%		SR              & $\frac{R_{800}}{R_{680}}$                                                                            &~\cite{jordan1969}             \\
%		SR1             & $\frac{R_{750}}{R_{700}}$                                                                            &~\cite{gitelson1997}           \\
%		SR2             & $\frac{R_{752}}{R_{690}}$                                                                            &~\cite{gitelson1997}           \\
%		SR3             & $\frac{R_{750}}{R_{550}}$                                                                            &~\cite{gitelson1997}           \\
%		SR4             & $\frac{R_{700}}{R_{670}}$                                                                            &~\cite{mcmurtrey1994}          \\
%		SR5             & $\frac{R_{675}}{R_{700}}$                                                                            &~\cite{chappelle1992}          \\
%		SR6             & $\frac{R_{750}}{R_{710}}$                                                                            &~\cite{zarco-tejada1999}       \\
%		SR7             & $\frac{R_{440}}{R_{690}}$                                                                            &~\cite{lichtenthaler1996}      \\
%		SR8             & $\frac{R_{515}}{R_{550}}$                                                                            &~\cite{hernandez-clemente2012} \\
%		SRPI            & $\frac{R_{430}}{R_{680}}$                                                                            &~\cite{penuelasj.1995}         \\
%		SRWI            & $\frac{R_{850}}{R_{1240}}$                                                                           &~\cite{zarco-tejada2003}      \\
%		Sum\_Dr1        & $\sum_{i=626}^{795} D1_i$                                                                            &~\cite{elvidge1995}            \\
%		Sum\_Dr2        & $\sum_{i=680}^{780} D1_i$                                                                            &~\cite{filella1994}            \\
%		SWIR FI         & $\frac{R_{2133}^2}{R_{2225} \times R_{2209}^3}$                                                      &~\cite{levin2007}              \\
%		\midrule
%		SWIR LI         & \parbox{3.8cm}{$3.87  \times (R_{2210} - R_{2090}) - 27.51 \times (R_{2280} - R_{2090}) - 0.2$}      &~\cite{lobell2001}             \\
%		\midrule
%		SWIR SI         & \parbox{3.8cm}{$-41.59 \times (R_{2210} - R_{2090}) + 1.24 \times (R_{2280} - R_{2090}) + 0.64 $}    &~\cite{lobell2001}             \\
%		\midrule
%		SWIR VI         & \parbox{3.8cm}{$37.72  \times (R_{2210} - R_{2090}) + 6.27 \times (R_{2280} - R_{2090}) + 0.57$}     &~\cite{lobell2001}             \\
%		\midrule
%		TCARI           & \parbox{3.8cm}{$3*((R_{700}-R_{670})-0.2\times R_{700}-R_{550})\times (R_{700}/R_{670}))$}           &~\cite{haboudane2002}          \\
%		\midrule
%		TCARI/OSAVI     & TCARI/OSAVI                                                                                          &~\cite{haboudane2002}          \\
%		\midrule
%		TCARI2          & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} &~\cite{wu2008}                \\
%		\midrule
%		TCARI2/OSAVI2   & TCARI2/OSAVI2                                                                                        &~\cite{wu2008}                \\
%		\midrule
%		TGI             & \parbox{3.8cm}{$-0.5 (190 (R_{670} - R_{550} ) - 120 (R_{670} - R_{480}))$}                          &~\cite{hunt2013}               \\
%		\midrule
%		TVI             & \parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}              &~\cite{broge2001}              \\
%		\midrule
%		Vogelmann       & $\frac{R_{740}}{R_{720}}$                                                                            &~\cite{vogelmann1993}          \\
%		Vogelmann2      & $\frac{R_{734}-R_{747}}{R_{715}+R_{726})}$                                                           &~\cite{vogelmann1993}          \\
%		Vogelmann3      & $\frac{D_{715}}{D_{705}}$                                                                            &~\cite{vogelmann1993}          \\
%		Vogelmann4      & $\frac{R_{734}-R_{747}}{R_{715}+R_{720}}$                                                            &~\cite{vogelmann1993}     \\
		\bottomrule
	\end{tabular}
	\label{tab:vegindices}}
	\end{specialtable}

\begin{specialtable}[H]\ContinuedFloat
\setlength{\tabcolsep}{11mm}
{
	\caption{{\em Cont.}}
	\begin{tabular}{lll}\toprule
		% \begin{supertabular}{0.3\textwidth}{lll}
		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference~*}         \\		\midrule
%		Boochs          & $D_{703}$                                                                                            &~\cite{boochs1990}             \\
%		Boochs2         & $D_{720}$                                                                                            &~\cite{boochs1990}             \\
%		CAI             & $0.5 \times (R_{2000} + R_{2200}) -R_{2100}$                                                         &~\cite{nagler2003}             \\
%		\midrule
%		CARI            & $a = (R_{700}-R_{550}) / 150$                                                                        &~\cite{walthall1994}           \\
%		& $b = R_{550}-(a\times 550)$                                                                          &                               \\
%		& $\frac{R_{700}\times | (a\times 670+R_{670}+b)}{R_{670}\times(a^2+1)| ^{0.5}}$                       &                               \\
%		\midrule
%		Carter          & $R_{695}/R_{420}$                                                                                    &~\cite{carter1994}             \\
%		Carter2         & $R_{695}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter3         & $R_{605}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter4         & $R_{710}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter5         & $R_{695}/R_{670}$                                                                                    &~\cite{carter1994}             \\
%		Carter6         & $R_{550}$                                                                                            &~\cite{carter1994}             \\
%		CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                    &~\cite{zarco-tejada2003}      \\
%		CI~2             & $R_{760}/R_{700}-1$                                                                                  &~\cite{gitelson2003}           \\
%		ClAInt          & $\int_{600 nm}^{735 nm} R$                                                                           &~\cite{oppelt2004}             \\
%		CRI1            & $1/R_{515}-1/R_{550}$                                                                                &~\cite{gitelson2003}           \\
%		CRI2            & $1/R_{515}-1/R_{770}$                                                                                &~\cite{gitelson2003}           \\
%		CRI3            & $1/R_{515}-1/R_{550}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
%		CRI4            & $1/R_{515}-1/R_{700}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
%		D1              & $D_{730}/D_{706}$                                                                                    &~\cite{zarco-tejada2003}      \\
%		D2              & $D_{705}/D_{722}$                                                                                    &~\cite{zarco-tejada2003}      \\
%		Datt            & $(R_{850}-R_{710})/(R_{850}-R_{680})$                                                                &~\cite{datt1999}               \\
%		Datt2           & $R_{850}/R_{710}$                                                                                    &~\cite{datt1999}               \\
%		Datt3           & $D_{754}/D_{704}$                                                                                    &~\cite{datt1999}               \\
%		Datt4           & $R_{672}/(R_{550} \times R_{708})$                                                                   &~\cite{datt1998}               \\
%		Datt5           & $R_{672}/R_{550}$                                                                                    &~\cite{datt1998}               \\
%		Datt6           & $(R_{860})/(R_{550}\times R_{708})$                                                                  &~\cite{datt1998}               \\
%		Datt7           & $(R_{860} - R_{2218})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
%		Datt8           & $(R_{860} - R_{1788})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
%		DD              & $(R_{749}-R_{720})-(R_{701}-R_{672})$                                                                &~\cite{maire2004}              \\
%		DDn             & $2\times (R_{710}-R_{660}-R_{760})$                                                                  &~\cite{lemaire2008}            \\
%		DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                        &~\cite{zarco-tejada2003}      \\
%		DWSI1           & $R_{80}/R_{1660}$                                                                                    &~\cite{apan2004}               \\
%		DWSI2           & $R_{1660}/R_{550}$                                                                                   &~\cite{apan2004}               \\
%		DWSI3           & $R_{1660}/R_{680}$                                                                                   &~\cite{apan2004}               \\
%		DWSI4           & $R_{550}/R_{680}$                                                                                    &~\cite{apan2004}               \\
%		DWSI5           & $(R_{800} + R_{550})/(R_{1660} + R_{680})$                                                           &~\cite{apan2004}               \\
%		EGFN            & $\frac{(\max(D_{650:750})-\max(D_{500:550}))}{(\max(D_{650:750})+\max(D_{500:550}))}$                &~\cite{penuelas1994}           \\
%		EGFR            & $\max(D_{650:750})/\max(D_{500:550})$                                                                &~\cite{penuelas1994}           \\
%		EVI             & $\frac{2.5\times ((R_{800}-R_{670}) }{ (R_{800}-(6\times R_{670})-(7.5\times R_{475})+1)}$           &~\cite{huete1997}             \\
%		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$**                                                    &~\cite{wu2014}                 \\
%		GI              & $R_{554}/R_{677}$                                                                                    &~\cite{smith1995}              \\
%		Gitelson        & $1/R_{700}$                                                                                          &~\cite{gitelson1999}           \\
%		Gitelson2       & $(R_{750}-R_{800}/R_{695}-R_{740})-1$                                                                &~\cite{gitelson2003}           \\
%		GMI1            & $R_{750}/R_{550}$                                                                                    &~\cite{gitelson2003}           \\
%		GMI2            & $R_{750}/R_{700}$                                                                                    &~\cite{gitelson2003}           \\
%		Green NDVI      & $\frac{R_{800}-R_{550}}{R_{800}+R_{550}}$                                                            &~\cite{gitelson1996}           \\
%		LWVI\_1         & $\frac{(R_{1094}-R_{983})}{(R_{1094}+R_{983})}$                                                      &~\cite{galvao2005}             \\
%		LWVI\_2         & $\frac{R_{1094}-R_{1205}}{R_{1094}+R_{1205}}$                                                        &~\cite{galvao2005}             \\
%		Maccioni        & $\frac{R_{780}-R_{710})}{R_{780}-R_{680}}$                                                           &~\cite{maccioni2001}           \\
%		\midrule
%		MCARI           & \parbox{5.5cm}{$((R_{700}-R_{670})-0.2\times (R_{700}-R_{550})) \times (R_{700}/R_{670})$}           &~\cite{daughtry2000}           \\
%		\midrule
%		MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}          &~\cite{wu2008}                \\
%		\midrule
%		mND705          & $\frac{(R_{750}-R_{705})}{R_{750}+R_{705}-2\times R_{445}}$                                          &~\cite{sims2002}              \\
%		mNDVI           & $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$                                         &~\cite{sims2002}              \\
%		MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                            &~\cite{hernandez-clemente2011} \\
%		MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                 &~\cite{qi1994}                 \\
%		MSI             & $\frac{R_{1600}}{R_{817}}$                                                                           &~\cite{hunt1989}               \\
%		mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                            &~\cite{sims2002}              \\
%		mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                               &~\cite{chen1996}               \\
%		mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                            &~\cite{sims2002}              \\
%		MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                            &~\cite{dash2007}               \\
%		\midrule
%		MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}             &~\cite{haboudane2002}          \\
%		\midrule
%		NDLI            & $\frac{log(1/R_{1754}) - log(1/R_{1680})}{log(1/R_{1754}) + log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
%		NDNI            & $\frac{log(1/R_{1510}) - log(1/R_{1680})}{log(1/R_{1510}) + log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
%		NDVI            & $\frac{R_{800}-R_{680}}{R_{800}+R_{680}}$                                                            &~\cite{tucker1979}             \\
		NDVI2           & $\frac{R_{750}\,-\,R_{705}}{R_{750}\,+\,R_{705}}$                                                            &~\cite{gitelson1994}           \\
		NDVI3           & $\frac{R_{682}\,-\,R_{553}}{R_{682}\,+\,R_{553}}$                                                            &~\cite{guanter2005}            \\
		NDWI            & $\frac{R_{860}\,-\,R_{1240}}{R_{860}\,+\,R_{1240}}$                                                          &~\cite{gao1996}                \\
		NPCI            & $\frac{R_{680}\,-\,R_{430}}{R_{680}\,+\,R_{430}}$                                                            &~\cite{penuelas1994}           \\
		OSAVI           & $\frac{(1\,+\,0.16) \,\times\, (R_{800}\,-\,R_{670})}{R_{800}\,+\,R_{670}\,+\,0.16 }$                                    &~\cite{rondeaux1996}           \\
		OSAVI2          & $\frac{(1\,+\,0.16)\,\times\, (R_{750}\,-\,R_{705})}{R_{750}\,+\,R_{705}\,+\,0.16) }$                                    &~\cite{wu2008}                \\
		PARS            & $\frac{R_{746}}{R_{513}}$                                                                            &~\cite{chappelle1992}          \\
		PRI             & $\frac{R_{531}\,-\,R_{570}}{R_{531}\,+\,R_{570}}$                                                            &~\cite{gamon1992}              \\
		PRI\_norm       & $\frac{PRI \,\times\, (-1) }{RDVI\,\times\, R_{700}/R_{670}}$                                                &~\cite{zarco-tejada2013}      \\
		PRI\,*\,CI2         & $PRI*CI2$                                                                                            &~\cite{garrity2011}            \\
		PSRI            & $\frac{R_{678}\,-\,R_{500}}{R_{750}}$                                                                    &~\cite{merzlyak1999}           \\
		PSSR            & $\frac{R_{800}}{R_{635}}$                                                                            &~\cite{blackburn1998}          \\
		PSND            & $\frac{R_{800}\,-\,R_{470}}{R_{800}-R_{470}}$                                                           &~\cite{blackburn1998}          \\
		PWI             & $\frac{R_{900}}{R_{970}}$                                                                            &~\cite{penuelas1997}           \\
		RDVI            & $\frac{R_{800}\,-\,R_{670}}{ \sqrt{R_{800}\,+\,R_{670}}}$                                                    &~\cite{roujean1995}            \\
		\midrule
		REP\_LE         & \parbox{3.8cm}{Red-edge position through linear extrapolation}                                       &~\cite{cho2006}                \\
		\midrule
		REP\_Li         & $R_{re}=\frac{R_{670}\,+\,R_{780}}{2}$                                                                  &~\cite{guyot1988}              \\
		& $\frac{700 \,+\, 40 \,\times\, (R_{re} \,-\,R_{700})}{(R_{740}\,-\,R_{700}))}$                                      &                               \\
		\midrule
		SAVI            & $\frac{(1\,+\,L)\,\times\, (R_{800}\,-\,R_{670})}{(R_{800}\,+\,R_{670}\,+\,L)}$                                          &~\cite{huete1988}              \\
		SIPI            & $\frac{R_{800}\,-\,R_{445}}{R_{800}\,-\,R_{680}}$                                                            &~\cite{penuelasj.1995}         \\
		\midrule
		SPVI            & \parbox{3.8cm}{$0.4 \,\times\, 3.7 \,\times\, (R_{800}\,-\,R_{670})\,-\,1.2 \,\times\, ((R_{530}-R_{670})^2)^{0.5}$}     &~\cite{vincini2006}            \\
		\midrule
		SR              & $\frac{R_{800}}{R_{680}}$                                                                            &~\cite{jordan1969}             \\
		SR1             & $\frac{R_{750}}{R_{700}}$                                                                            &~\cite{gitelson1997}           \\
		SR2             & $\frac{R_{752}}{R_{690}}$                                                                            &~\cite{gitelson1997}           \\
		SR3             & $\frac{R_{750}}{R_{550}}$                                                                            &~\cite{gitelson1997}           \\
		SR4             & $\frac{R_{700}}{R_{670}}$                                                                            &~\cite{mcmurtrey1994}          \\
		SR5             & $\frac{R_{675}}{R_{700}}$                                                                            &~\cite{chappelle1992}          \\
		SR6             & $\frac{R_{750}}{R_{710}}$                                                                            &~\cite{zarco-tejada1999}       \\
		SR7             & $\frac{R_{440}}{R_{690}}$                                                                            &~\cite{lichtenthaler1996}      \\
		SR8             & $\frac{R_{515}}{R_{550}}$                                                                            &~\cite{hernandez-clemente2012} \\
		SRPI            & $\frac{R_{430}}{R_{680}}$                                                                            &~\cite{penuelasj.1995}         \\
		SRWI            & $\frac{R_{850}}{R_{1240}}$\vspace{6pt}                                                                                                                                                      &~\cite{zarco-tejada2003}      \\
		Sum\_Dr1        & $\sum_{i=626}^{795} D1_i$\vspace{6pt}                                                                            &~\cite{elvidge1995}            \\
		Sum\_Dr2        & $\sum_{i=680}^{780} D1_i$                                                                            &~\cite{filella1994}            \\
		SWIR FI         & $\frac{R_{2133}^2}{R_{2225} \,\times\, R_{2209}^3}$                                                      &~\cite{levin2007}              \\
%		\midrule
%		SWIR LI         & \parbox{3.8cm}{$3.87  \times (R_{2210} - R_{2090}) - 27.51 \times (R_{2280} - R_{2090}) - 0.2$}      &~\cite{lobell2001}             \\
%		\midrule
%		SWIR SI         & \parbox{3.8cm}{$-41.59 \times (R_{2210} - R_{2090}) + 1.24 \times (R_{2280} - R_{2090}) + 0.64 $}    &~\cite{lobell2001}             \\
%		\midrule
%		SWIR VI         & \parbox{3.8cm}{$37.72  \times (R_{2210} - R_{2090}) + 6.27 \times (R_{2280} - R_{2090}) + 0.57$}     &~\cite{lobell2001}             \\
%		\midrule
%		TCARI           & \parbox{3.8cm}{$3*((R_{700}-R_{670})-0.2\times R_{700}-R_{550})\times (R_{700}/R_{670}))$}           &~\cite{haboudane2002}          \\
%		\midrule
%		TCARI/OSAVI     & TCARI/OSAVI                                                                                          &~\cite{haboudane2002}          \\
%		\midrule
%		TCARI2          & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} &~\cite{wu2008}                \\
%		\midrule
%		TCARI2/OSAVI2   & TCARI2/OSAVI2                                                                                        &~\cite{wu2008}                \\
%		\midrule
%		TGI             & \parbox{3.8cm}{$-0.5 (190 (R_{670} - R_{550} ) - 120 (R_{670} - R_{480}))$}                          &~\cite{hunt2013}               \\
%		\midrule
%		TVI             & \parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}              &~\cite{broge2001}              \\
%		\midrule
%		Vogelmann       & $\frac{R_{740}}{R_{720}}$                                                                            &~\cite{vogelmann1993}          \\
%		Vogelmann2      & $\frac{R_{734}-R_{747}}{R_{715}+R_{726})}$                                                           &~\cite{vogelmann1993}          \\
%		Vogelmann3      & $\frac{D_{715}}{D_{705}}$                                                                            &~\cite{vogelmann1993}          \\
%		Vogelmann4      & $\frac{R_{734}-R_{747}}{R_{715}+R_{720}}$                                                            &~\cite{vogelmann1993}     \\
		\bottomrule
	\end{tabular}
	\label{tab:vegindices}}
	\end{specialtable}

\begin{specialtable}[H]\ContinuedFloat
\setlength{\tabcolsep}{9mm}
{
	\caption{{\em Cont.}}
	\begin{tabular}{lll}\toprule
		% \begin{supertabular}{0.3\textwidth}{lll}
		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference~*}         \\		\midrule
%		Boochs          & $D_{703}$                                                                                            &~\cite{boochs1990}             \\
%		Boochs2         & $D_{720}$                                                                                            &~\cite{boochs1990}             \\
%		CAI             & $0.5 \times (R_{2000} + R_{2200}) -R_{2100}$                                                         &~\cite{nagler2003}             \\
%		\midrule
%		CARI            & $a = (R_{700}-R_{550}) / 150$                                                                        &~\cite{walthall1994}           \\
%		& $b = R_{550}-(a\times 550)$                                                                          &                               \\
%		& $\frac{R_{700}\times | (a\times 670+R_{670}+b)}{R_{670}\times(a^2+1)| ^{0.5}}$                       &                               \\
%		\midrule
%		Carter          & $R_{695}/R_{420}$                                                                                    &~\cite{carter1994}             \\
%		Carter2         & $R_{695}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter3         & $R_{605}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter4         & $R_{710}/R_{760}$                                                                                    &~\cite{carter1994}             \\
%		Carter5         & $R_{695}/R_{670}$                                                                                    &~\cite{carter1994}             \\
%		Carter6         & $R_{550}$                                                                                            &~\cite{carter1994}             \\
%		CI              & $R_{675}\times R_{690}/R_{683}^2$                                                                    &~\cite{zarco-tejada2003}      \\
%		CI~2             & $R_{760}/R_{700}-1$                                                                                  &~\cite{gitelson2003}           \\
%		ClAInt          & $\int_{600 nm}^{735 nm} R$                                                                           &~\cite{oppelt2004}             \\
%		CRI1            & $1/R_{515}-1/R_{550}$                                                                                &~\cite{gitelson2003}           \\
%		CRI2            & $1/R_{515}-1/R_{770}$                                                                                &~\cite{gitelson2003}           \\
%		CRI3            & $1/R_{515}-1/R_{550}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
%		CRI4            & $1/R_{515}-1/R_{700}\times R_{770}$                                                                  &~\cite{gitelson2003}           \\
%		D1              & $D_{730}/D_{706}$                                                                                    &~\cite{zarco-tejada2003}      \\
%		D2              & $D_{705}/D_{722}$                                                                                    &~\cite{zarco-tejada2003}      \\
%		Datt            & $(R_{850}-R_{710})/(R_{850}-R_{680})$                                                                &~\cite{datt1999}               \\
%		Datt2           & $R_{850}/R_{710}$                                                                                    &~\cite{datt1999}               \\
%		Datt3           & $D_{754}/D_{704}$                                                                                    &~\cite{datt1999}               \\
%		Datt4           & $R_{672}/(R_{550} \times R_{708})$                                                                   &~\cite{datt1998}               \\
%		Datt5           & $R_{672}/R_{550}$                                                                                    &~\cite{datt1998}               \\
%		Datt6           & $(R_{860})/(R_{550}\times R_{708})$                                                                  &~\cite{datt1998}               \\
%		Datt7           & $(R_{860} - R_{2218})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
%		Datt8           & $(R_{860} - R_{1788})/(R_{860} - R_{1928})$                                                          &~\cite{datt1999a}              \\
%		DD              & $(R_{749}-R_{720})-(R_{701}-R_{672})$                                                                &~\cite{maire2004}              \\
%		DDn             & $2\times (R_{710}-R_{660}-R_{760})$                                                                  &~\cite{lemaire2008}            \\
%		DPI             & $(D_{688}*D_{710})/D_{697}^2$                                                                        &~\cite{zarco-tejada2003}      \\
%		DWSI1           & $R_{80}/R_{1660}$                                                                                    &~\cite{apan2004}               \\
%		DWSI2           & $R_{1660}/R_{550}$                                                                                   &~\cite{apan2004}               \\
%		DWSI3           & $R_{1660}/R_{680}$                                                                                   &~\cite{apan2004}               \\
%		DWSI4           & $R_{550}/R_{680}$                                                                                    &~\cite{apan2004}               \\
%		DWSI5           & $(R_{800} + R_{550})/(R_{1660} + R_{680})$                                                           &~\cite{apan2004}               \\
%		EGFN            & $\frac{(\max(D_{650:750})-\max(D_{500:550}))}{(\max(D_{650:750})+\max(D_{500:550}))}$                &~\cite{penuelas1994}           \\
%		EGFR            & $\max(D_{650:750})/\max(D_{500:550})$                                                                &~\cite{penuelas1994}           \\
%		EVI             & $\frac{2.5\times ((R_{800}-R_{670}) }{ (R_{800}-(6\times R_{670})-(7.5\times R_{475})+1)}$           &~\cite{huete1997}             \\
%		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$**                                                    &~\cite{wu2014}                 \\
%		GI              & $R_{554}/R_{677}$                                                                                    &~\cite{smith1995}              \\
%		Gitelson        & $1/R_{700}$                                                                                          &~\cite{gitelson1999}           \\
%		Gitelson2       & $(R_{750}-R_{800}/R_{695}-R_{740})-1$                                                                &~\cite{gitelson2003}           \\
%		GMI1            & $R_{750}/R_{550}$                                                                                    &~\cite{gitelson2003}           \\
%		GMI2            & $R_{750}/R_{700}$                                                                                    &~\cite{gitelson2003}           \\
%		Green NDVI      & $\frac{R_{800}-R_{550}}{R_{800}+R_{550}}$                                                            &~\cite{gitelson1996}           \\
%		LWVI\_1         & $\frac{(R_{1094}-R_{983})}{(R_{1094}+R_{983})}$                                                      &~\cite{galvao2005}             \\
%		LWVI\_2         & $\frac{R_{1094}-R_{1205}}{R_{1094}+R_{1205}}$                                                        &~\cite{galvao2005}             \\
%		Maccioni        & $\frac{R_{780}-R_{710})}{R_{780}-R_{680}}$                                                           &~\cite{maccioni2001}           \\
%		\midrule
%		MCARI           & \parbox{5.5cm}{$((R_{700}-R_{670})-0.2\times (R_{700}-R_{550})) \times (R_{700}/R_{670})$}           &~\cite{daughtry2000}           \\
%		\midrule
%		MCARI2          & \parbox{5.5cm}{$((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550})) \times (R_{750}/R_{705})$}          &~\cite{wu2008}                \\
%		\midrule
%		mND705          & $\frac{(R_{750}-R_{705})}{R_{750}+R_{705}-2\times R_{445}}$                                          &~\cite{sims2002}              \\
%		mNDVI           & $\frac{(R_{800}-R_{680})}{R_{800}+R_{680}-2 \times R_{445}}$                                         &~\cite{sims2002}              \\
%		MPRI            & $\frac{R_{515}-R_{530}}{R_{515}+R_{530}}$                                                            &~\cite{hernandez-clemente2011} \\
%		MSAVI           & \parbox{5.5cm}{$0.5 \times ((2\times R_{800}+1)^2-8\times (R_{800}-R_{670}))^{0.5}$}                 &~\cite{qi1994}                 \\
%		MSI             & $\frac{R_{1600}}{R_{817}}$                                                                           &~\cite{hunt1989}               \\
%		mSR             & $\frac{R_{800}-R_{445}}{R_{680}-R_{445}}$                                                            &~\cite{sims2002}              \\
%		mSR2            & $\frac{(R_{750}/R_{705})-1}{R_{750}/R_{705}+1)^{0.5}}$                                               &~\cite{chen1996}               \\
%		mSR705          & $\frac{R_{750}-R_{445}}{R_{705}-R_{445}}$                                                            &~\cite{sims2002}              \\
%		MTCI            & $\frac{R_{754}-R_{709}}{R_{709}-R_{681}}$                                                            &~\cite{dash2007}               \\
%		\midrule
%		MTVI            & \parbox{3.8cm}{$1.2 \times (1.2 \times (R_{800}-R_{550})-2.5 \times (R_{670}-R_{550}))$}             &~\cite{haboudane2002}          \\
%		\midrule
%		NDLI            & $\frac{log(1/R_{1754}) - log(1/R_{1680})}{log(1/R_{1754}) + log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
%		NDNI            & $\frac{log(1/R_{1510}) - log(1/R_{1680})}{log(1/R_{1510}) + log(1/R_{1680})}$                        &~\cite{serrano2002}            \\
%		NDVI            & $\frac{R_{800}-R_{680}}{R_{800}+R_{680}}$                                                            &~\cite{tucker1979}             \\
%		NDVI2           & $\frac{R_{750}-R_{705}}{R_{750}+R_{705}}$                                                            &~\cite{gitelson1994}           \\
%		NDVI3           & $\frac{R_{682}-R_{553}}{R_{682}+R_{553}}$                                                            &~\cite{guanter2005}            \\
%		NDWI            & $\frac{R_{860}-R_{1240}}{R_{860}+R_{1240}}$                                                          &~\cite{gao1996}                \\
%		NPCI            & $\frac{R_{680}-R_{430}}{R_{680}+R_{430}}$                                                            &~\cite{penuelas1994}           \\
%		OSAVI           & $\frac{(1+0.16) \times (R_{800}-R_{670})}{R_{800}+R_{670}+0.16 }$                                    &~\cite{rondeaux1996}           \\
%		OSAVI2          & $\frac{(1+0.16)\times (R_{750}-R_{705})}{R_{750}+R_{705}+0.16) }$                                    &~\cite{wu2008}                \\
%		PARS            & $\frac{R_{746}}{R_{513}}$                                                                            &~\cite{chappelle1992}          \\
%		PRI             & $\frac{R_{531}-R_{570}}{R_{531}+R_{570}}$                                                            &~\cite{gamon1992}              \\
%		PRI\_norm       & $\frac{PRI \times (-1) }{RDVI\times R_{700}/R_{670}}$                                                &~\cite{zarco-tejada2013}      \\
%		PRI*CI2         & $PRI*CI2$                                                                                            &~\cite{garrity2011}            \\
%		PSRI            & $\frac{R_{678}-R_{500}}{R_{750}}$                                                                    &~\cite{merzlyak1999}           \\
%		PSSR            & $\frac{R_{800}}{R_{635}}$                                                                            &~\cite{blackburn1998}          \\
%		PSND            & $\frac{R_{800}-R_{470}}{R_{800}-R_{470})}$                                                           &~\cite{blackburn1998}          \\
%		PWI             & $\frac{R_{900}}{R_{970}}$                                                                            &~\cite{penuelas1997}           \\
%		RDVI            & $\frac{R_{800}-R_{670}}{ \sqrt{R_{800}+R_{670}}}$                                                    &~\cite{roujean1995}            \\
%		\midrule
%		REP\_LE         & \parbox{3.8cm}{Red-edge position through linear extrapolation}                                       &~\cite{cho2006}                \\
%		\midrule
%		REP\_Li         & $R_{re}=\frac{R_{670}+R_{780})}{2}$                                                                  &~\cite{guyot1988}              \\
%		& $\frac{700 + 40 \times ((R_{re} -R_{700})}{(R_{740}-R_{700}))}$                                      &                               \\
%		\midrule
%		SAVI            & $\frac{(1+L)\times (R_{800}-R_{670})}{(R_{800}+R_{670}+L)}$                                          &~\cite{huete1988}              \\
%		SIPI            & $\frac{R_{800}-R_{445}}{R_{800}-R_{680}}$                                                            &~\cite{penuelasj.1995}         \\
%		\midrule
%		SPVI            & \parbox{3.8cm}{$0.4 \times 3.7 \times (R_{800}-R_{670})-1.2 \times ((R_{530}-R_{670})^2)^{0.5}$}     &~\cite{vincini2006}            \\
%		\midrule
%		SR              & $\frac{R_{800}}{R_{680}}$                                                                            &~\cite{jordan1969}             \\
%		SR1             & $\frac{R_{750}}{R_{700}}$                                                                            &~\cite{gitelson1997}           \\
%		SR2             & $\frac{R_{752}}{R_{690}}$                                                                            &~\cite{gitelson1997}           \\
%		SR3             & $\frac{R_{750}}{R_{550}}$                                                                            &~\cite{gitelson1997}           \\
%		SR4             & $\frac{R_{700}}{R_{670}}$                                                                            &~\cite{mcmurtrey1994}          \\
%		SR5             & $\frac{R_{675}}{R_{700}}$                                                                            &~\cite{chappelle1992}          \\
%		SR6             & $\frac{R_{750}}{R_{710}}$                                                                            &~\cite{zarco-tejada1999}       \\
%		SR7             & $\frac{R_{440}}{R_{690}}$                                                                            &~\cite{lichtenthaler1996}      \\
%		SR8             & $\frac{R_{515}}{R_{550}}$                                                                            &~\cite{hernandez-clemente2012} \\
%		SRPI            & $\frac{R_{430}}{R_{680}}$                                                                            &~\cite{penuelasj.1995}         \\
%		SRWI            & $\frac{R_{850}}{R_{1240}}$                                                                           &~\cite{zarco-tejada2003}      \\
%		Sum\_Dr1        & $\sum_{i=626}^{795} D1_i$                                                                            &~\cite{elvidge1995}            \\
%		Sum\_Dr2        & $\sum_{i=680}^{780} D1_i$                                                                            &~\cite{filella1994}            \\
%		SWIR FI         & $\frac{R_{2133}^2}{R_{2225} \times R_{2209}^3}$                                                      &~\cite{levin2007}              \\
%		\midrule
		SWIR LI         & \parbox{3.8cm}{$3.87  \times (R_{2210} - R_{2090}) - 27.51 \times (R_{2280} - R_{2090}) - 0.2$}      &~\cite{lobell2001}             \\
		\midrule
		SWIR SI         & \parbox{3.8cm}{$-41.59 \times (R_{2210} - R_{2090}) + 1.24 \times (R_{2280} - R_{2090}) + 0.64 $}    &~\cite{lobell2001}             \\
		\midrule
		SWIR VI         & \parbox{3.8cm}{$37.72  \times (R_{2210} - R_{2090}) + 6.27 \times (R_{2280} - R_{2090}) + 0.57$}     &~\cite{lobell2001}             \\
		\midrule
		TCARI           & \parbox{3.8cm}{$3*((R_{700}-R_{670})-0.2\times R_{700}-R_{550})\times (R_{700}/R_{670}))$}           &~\cite{haboudane2002}          \\
		\midrule
		TCARI/OSAVI     & TCARI/OSAVI                                                                                          &~\cite{haboudane2002}          \\
		\midrule
		TCARI2          & \parbox{3.8cm}{$3 \times ((R_{750}-R_{705})-0.2 \times (R_{750}-R_{550}) \times (R_{750}/R_{705}))$} &~\cite{wu2008}                \\
		\midrule
		TCARI2/OSAVI2   & TCARI2/OSAVI2                                                                                        &~\cite{wu2008}                \\
		\midrule
		TGI             & \parbox{3.8cm}{$-0.5 (190 (R_{670} - R_{550} ) - 120 (R_{670} - R_{480}))$}                          &~\cite{hunt2013}               \\
		\midrule
		TVI             & \parbox{3.8cm}{$0.5\times (120 \times (R_{750}-R_{550})-200 \times (R_{670}-R_{550}))$}              &~\cite{broge2001}              \\
		\midrule
		Vogelmann       & $\frac{R_{740}}{R_{720}}$                                                                            &~\cite{vogelmann1993}          \\
		Vogelmann2      & $\frac{R_{734}\,-\,R_{747}}{R_{715}\,+\,R_{726})}$                                                           &~\cite{vogelmann1993}          \\
		Vogelmann3      & $\frac{D_{715}}{D_{705}}$                                                                            &~\cite{vogelmann1993}          \\
		Vogelmann4      & $\frac{R_{734}\,-\,R_{747}}{R_{715}\,+\,R_{720}}$                                                            &~\cite{vogelmann1993}     \\
		\bottomrule
	\end{tabular}
	\label{tab:vegindices}}
	\end{specialtable}
%\pagebreak

\subsection{}

The following information was provided by the Institut Carogràfic i Geològic de Catalunya, which was in charge of image acquisition and data preprocessing.

The AISA EAGLE-II sensor was used for airborne image acquisition with a field of view of 37.7~°.
Its spectral resolution is 2.4~nm and ranges from 400~nm\ to 1000~nm.

The conversion of digital numbers (DN) to spectral radiance was made  using software designed for the instrument.
Images were originally scaled in 12 bits but were radiometrically calibrated to 16 bits, reserving the highest value (65,535) for null values.
The procedure was applied to the 23 previously selected images.
Finally, the geometric and atmospheric corrections were applied to the images.

The aim of this procedure was to reduce the positional errors of the images.
The cartographic reference system in use was EPSG~25830.
Positioning was achieved by coupling an Applanix POS AV 410 system to the sensor, integrating GPS and IMU systems.
The system provides geographic coordinates of the terrain and relative coordinates of the aircraft (attitude) at each scanned line.
Additionally a DSM from GeoEuskadi with a spatial resolution of 1~m was used.
The orthorectified hyperspectral images were compared to orthoimages (1:5000) from GeoEuskadi.
This comparison was used as the base to calculate RMSE, which was below the ground sampling distance in the across and along track~directions.

The radiance measured by an instrument depends on the illumination geometry and the reflective properties of the observed surface.
Radiation may be absorbed or scattered (Rayleigh and Mie scattering).
Scattering is responsible for the adjacency effect, i.e., radiation coming from neighbors' areas to the target pixel.
The MODTRAN algorithm was used to model the effect of the atmosphere on the radiation.
To represent the aerosols of the study area, the rural model was used.
In addition, optical thickness was estimated on pixels with a high vegetation cover.
Columnar water vapor was estimated by a linear regression ratio where the spectral radiance of each pixel at the band of the maximum water absorption (906~nm) is compared to its theoretical value in the absence of absorption.
Nonetheless, this technique is unreliable in the presence of a spectral resolution as in this case.
To resolve this, the water vapor parameter was selected manually according to the smoothness observed on the reflectance peak at 960~nm.
This was combined with a mid-latitude summer atmosphere model.
The output of this procedure was reflectance from the target pixel scaled between 0 and 10,000.

The image acquisitions were originally attempted during one day (29 October 2016).
Due to the variable meteorological conditions, some stands had to be imaged one day later.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{paracol}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% To add notes in main text, please use \endnote{} and un-comment the codes below.
%\begin{adjustwidth}{-5.0cm}{0cm}
%\printendnotes[custom]
%\end{adjustwidth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{Biblio}
\begin{thebibliography}{999}

\bibitem[Lary \em{et~al.}(2016)Lary, Alavi, Gandomi, and Walker]{lary2016}
Lary, D.J.; Alavi, A.H.; Gandomi, A.H.; Walker, A.L.
\newblock Machine Learning in Geosciences and Remote Sensing.
\newblock {\em Geosci. Front.} {\bf 2016}, {\em 7},~3--10, doi:10.1016/j.gsf.2015.07.003.

\bibitem[Ma \em{et~al.}(2015)Ma, Wu, Wang, Huang, Ranjan, Zomaya, and
  Jie]{ma2015}
Ma, Y.; Wu, H.; Wang, L.; Huang, B.; Ranjan, R.; Zomaya, A.; Jie, W.
\newblock Remote Sensing Big Data Computing: Challenges and Opportunities.
\newblock {\em Future Gener. Comput. Syst.} {\bf 2015}, {\em
  51},~47--60.

\bibitem[Mascaro \em{et~al.}(2014)Mascaro, Asner, Knapp, {Kennedy-Bowdoin},
  Martin, Anderson, Higgins, and Chadwick]{mascaro2014}
Mascaro, J.; Asner, G.P.; Knapp, D.E.; {Kennedy-Bowdoin}, T.; Martin, R.E.;
  Anderson, C.; Higgins, M.; Chadwick, K.D.
\newblock A {{Tale}} of {{Two}} ``{{Forests}}'': Random {{Forest Machine
  Learning Aids Tropical Forest Carbon Mapping}}.
\newblock {\em PLoS ONE} {\bf 2014}, {\em 9},~e85993.

\bibitem[Urban \em{et~al.}(2018)Urban, Berger, Mudau, Heckel, Truckenbrodt,
  Onyango~Odipo, Smit, and Schmullius]{urban2018}
Urban, M.; Berger, C.; Mudau, T.E.; Heckel, K.; Truckenbrodt, J.;
  Onyango~Odipo, V.; Smit, I.P.J.; Schmullius, C.
\newblock Surface {{Moisture}} and {{Vegetation Cover Analysis}} for {{Drought
  Monitoring}} in the {{Southern Kruger National Park Using Sentinel}}-1,
  {{Sentinel}}-2, and {{Landsat}}-8.
\newblock {\em Remote Sens.} {\bf 2018}, {\em 10},~1482.

\bibitem[Hawry{\l}o \em{et~al.}(2018)Hawry{\l}o, Bednarz, Wezyk, and
  Szostak]{hawrylo2018}
Hawry{\l}o, P.; Bednarz, B.; Wezyk, P.; Szostak, M.
\newblock Estimating Defoliation of {{Scots}} Pine Stands Using Machine
  Learning Methods and Vegetation Indices of {{Sentinel}}-2.
\newblock {\em Eur. J. Remote Sens.} {\bf 2018}, {\em
  51},~194--204.

\bibitem[Pollastrini \em{et~al.}(2016)Pollastrini, Feducci, Bonal, Fotelli,
  Gessler, Grossiord, Guyot, Jactel, Nguyen, Radoglou, and
  Bussotti]{pollastrini2016}
Pollastrini, M.; Feducci, M.; Bonal, D.; Fotelli, M.; Gessler, A.; Grossiord,
  C.; Guyot, V.; Jactel, H.; Nguyen, D.; Radoglou, K.; et~al.
\newblock Physiological Significance of Forest Tree Defoliation: Results from a
  Survey in a Mixed Forest in {{Tuscany}} (Central {{Italy}}).
\newblock {\em For. Ecol. Manag.} {\bf 2016}, {\em 361},~170--178, doi:10.1016/j.foreco.2015.11.018.

\bibitem[Gottardini \em{et~al.}(2020)Gottardini, Cristofolini, Cristofori,
  Pollastrini, Camin, and Ferretti]{gottardini2020}
Gottardini, E.; Cristofolini, F.; Cristofori, A.; Pollastrini, M.; Camin, F.;
  Ferretti, M.
\newblock A Multi-Proxy Approach Reveals Common and Species-Specific Features
  Associated with Tree Defoliation in Broadleaved Species.
\newblock {\em For. Ecol. Manag.} {\bf 2020}, {\em 467},~118151, doi:10.1016/j.foreco.2020.118151.

\bibitem[Oliva \em{et~al.}(2016)Oliva, Stenlid, {Gr{\"o}nkvist-Wichmann},
  Wahlstr{\"o}m, Jonsson, Drobyshev, and Stenstr{\"o}m]{oliva2016}
Oliva, J.; Stenlid, J.; {Gr{\"o}nkvist-Wichmann}, L.; Wahlstr{\"o}m, K.;
  Jonsson, M.; Drobyshev, I.; Stenstr{\"o}m, E.
\newblock Pathogen-Induced Defoliation of {{Pinus}} Sylvestris Leads to Tree
  Decline and Death from Secondary Biotic Factors.
\newblock {\em For. Ecol. Manag.} {\bf 2016}, {\em 379},~273--280, doi:10.1016/j.foreco.2016.08.011.

\bibitem[Zhang \em{et~al.}(2016)Zhang, Thapa, Ross, and Gann]{zhang2016}
Zhang, K.; Thapa, B.; Ross, M.; Gann, D.
\newblock Remote Sensing of Seasonal Changes and Disturbances in Mangrove
  Forest: A Case Study from {{South Florida}}.
\newblock {\em Ecosphere} {\bf 2016}, \emph{7}, e01366, doi:10.1002/ecs2.1366.

\bibitem[Townsend \em{et~al.}(2012)Townsend, Singh, Foster, Rehberg, Kingdon,
  Eshleman, and Seagle]{townsend2012}
Townsend, P.A.; Singh, A.; Foster, J.R.; Rehberg, N.J.; Kingdon, C.C.;
  Eshleman, K.N.; Seagle, S.W.
\newblock A General {{Landsat}} Model to Predict Canopy Defoliation in
  Broadleaf Deciduous Forests.
\newblock {\em Remote Sens. Environ.} {\bf 2012}, {\em 119},~255--265, doi:10.1016/j.rse.2011.12.023.

\bibitem[Jiang \em{et~al.}(2014)Jiang, Wang, {de Bie}, Skidmore, Liu, Song,
  Zhang, Wang, and Shao]{jiang2014}
Jiang, Y.; Wang, T.; {de Bie}, C.A.J.M.; Skidmore, A.K.; Liu, X.; Song, S.;
  Zhang, L.; Wang, J.; Shao, X.
\newblock Satellite-Derived Vegetation Indices Contribute Significantly to the
  Prediction of Epiphyllous Liverworts.
\newblock {\em Ecol. Indic.} {\bf 2014}, {\em 38},~72--80, doi:10.1016/j.ecolind.2013.10.024.

\bibitem[Adamczyk and Osberger(2015)]{adamczyk2015}
Adamczyk, J.; Osberger, A.
\newblock Red-Edge Vegetation Indices for Detecting and Assessing Disturbances
  in {{Norway}} Spruce Dominated Mountain Forests.
\newblock {\em Int. J. Appl. Earth Obs. Geoinf.} {\bf 2015}, {\em 37},~90--99, doi:10.1016/j.jag.2014.10.013.

\bibitem[Thenkabail \em{et~al.}(2018)Thenkabail, Lyon, and
  Huete]{thenkabail2018}
Thenkabail, P.S.; Lyon, J.G.; Huete, A. (Eds.)
\newblock {\em Hyperspectral Indices and Image Classifications for Agriculture
  and Vegetation}; {CRC Press}:  Boca Raton, FL, USA, 2018, doi:10.1201/9781315159331.

\bibitem[Thenkabail \em{et~al.}(2000)Thenkabail, Smith, and
  De~Pauw]{thenkabail2000}
Thenkabail, P.S.; Smith, R.B.; De~Pauw, E.
\newblock Hyperspectral Vegetation Indices and Their Relationships with
  Agricultural Crop Characteristics.
\newblock {\em Remote Sens. Environ.} {\bf 2000}, {\em 71},~158--182.

\bibitem[Cai \em{et~al.}(2018)Cai, Luo, Wang, and Yang]{cai2018}
Cai, J.; Luo, J.; Wang, S.; Yang, S.
\newblock Feature Selection in Machine Learning: A New Perspective.
\newblock {\em Neurocomputing} {\bf 2018}, {\em 300},~70--79, doi:10.1016/j.neucom.2017.11.077.

\bibitem[Pinto \em{et~al.}(2020)Pinto, Powell, Peterson, Rosalen, and
  Fernandes]{pinto2020}
Pinto, J.; Powell, S.; Peterson, R.; Rosalen, D.; Fernandes, O.
\newblock Detection of {{Defoliation Injury}} in {{Peanut}} with
  {{Hyperspectral Proximal Remote Sensing}}.
\newblock {\em Remote Sens.} {\bf 2020}, {\em 12},~3828, doi:10.3390/rs12223828.

\bibitem[Yu \em{et~al.}(2021)Yu, Ren, and Luo]{yu2021}
Yu, R.; Ren, L.; Luo, Y.
\newblock Early Detection of Pine Wilt Disease in {{Pinus}} Tabuliformis in
  {{North China}} Using a Field Portable Spectrometer and {{UAV}}-Based
  Hyperspectral Imagery.
\newblock {\em For. Ecosyst.} {\bf 2021}, {\em 8},~44, doi:10.1186/s40663-021-00328-6.

\bibitem[Lin \em{et~al.}(2014)Lin, Yan, Wang, and Song]{lin2014}
Lin, H.; Yan, E.; Wang, G.; Song, R.
\newblock Analysis of Hyperspectral Bands for the Health Diagnosis of Tree
  Species.
\newblock  In Proceedings of the 2014 {{Third International Workshop}} on {{Earth Observation}} and
  {{Remote Sensing Applications}} ({{EORSA}}), Changsha, China, 11--14 June 2014; pp. 448--451, doi:10.1109/EORSA.2014.6927931.

\bibitem[Kayet \em{et~al.}(2019)Kayet, Pathak, Chakrabarty, Singh, Chowdary,
  Kumar, and Sahoo]{kayet2019}
Kayet, N.; Pathak, K.; Chakrabarty, A.; Singh, C.P.; Chowdary, V.M.; Kumar, S.;
  Sahoo, S.
\newblock Forest Health Assessment for Geo-Environmental Planning and
  Management in Hilltop Mining Areas Using {{Hyperion}} and {{Landsat}} Data.
\newblock {\em Ecol. Indic.} {\bf 2019}, {\em 106},~105471, doi:10.1016/j.ecolind.2019.105471.

\bibitem[Dash \em{et~al.}(2017)Dash, Watt, Pearse, Heaphy, and
  Dungey]{dash2017}
Dash, J.P.; Watt, M.S.; Pearse, G.D.; Heaphy, M.; Dungey, H.S.
\newblock Assessing Very High Resolution {{UAV}} Imagery for Monitoring Forest
  Health during a Simulated Disease Outbreak.
\newblock {\em ISPRS J. Photogramm. Remote Sens.} {\bf 2017},
  {\em 131},~1--14, doi:10.1016/j.isprsjprs.2017.07.007.

\bibitem[Trunk(1979)]{trunk1979}
Trunk, G.V.
\newblock A {{Problem}} of {{Dimensionality}}: A {{Simple Example}}.
\newblock {\em IEEE Trans. Pattern Anal. Mach. Intell.}
  {\bf 1979}, {\em PAMI-1}, 306--307.

\bibitem[Xu \em{et~al.}(2016)Xu, Caramanis, and Mannor]{xu2016}
Xu, H.; Caramanis, C.; Mannor, S.
\newblock Statistical {{Optimization}} in {{High Dimensions}}.
\newblock {\em Oper. Res.} {\bf 2016}, {\em 64},~958--979.

\bibitem[Mesanza \em{et~al.}(2016)Mesanza, Iturritxa, and Patten]{mesanza2016}
Mesanza, N.; Iturritxa, E.; Patten, C.L.
\newblock Native Rhizobacteria as Biocontrol Agents of {{Heterobasidion}}
  Annosum s.s. and {{Armillaria}} Mellea Infection of {{Pinus}} Radiata.
\newblock {\em Biol. Control} {\bf 2016}, {\em 101},~8--16, doi:10.1016/j.biocontrol.2016.06.003.

\bibitem[Iturritxa \em{et~al.}(2017)Iturritxa, Trask, Mesanza, Raposo,
  {Elvira-Recuenco}, and Patten]{iturritxa2017}
Iturritxa, E.; Trask, T.; Mesanza, N.; Raposo, R.; {Elvira-Recuenco}, M.;
  Patten, C.L.
\newblock Biocontrol of {{Fusarium}} Circinatum Infection of Young {{Pinus}}
  Radiata Trees.
\newblock {\em Forests} {\bf 2017}, {\em 8},~32, doi:10.3390/f8020032.

\bibitem[Iturritxa \em{et~al.}(2014)Iturritxa, Mesanza, and
  Brenning]{iturritxa2014}
Iturritxa, E.; Mesanza, N.; Brenning, A.
\newblock Spatial Analysis of the Risk of Major Forest Diseases in {{Monterey}}
  Pine Plantations.
\newblock {\em Plant Pathol.} {\bf 2014}, {\em 64},~880--889, doi:10.1111/ppa.12328.

\bibitem[Ganley \em{et~al.}(2009)Ganley, Watt, Manning, and
  Iturritxa]{ganley2009}
Ganley, R.J.; Watt, M.S.; Manning, L.; Iturritxa, E.
\newblock A Global Climatic Risk Assessment of Pitch Canker Disease.
\newblock {\em Can. J. For. Res.} {\bf 2009}, {\em
  39},~2246--2256, doi:10.1139/x09-131.

\bibitem[Innes(1993)]{innes1993}
Innes, J.
\newblock Methods to Estimate Forest Health.
\newblock {\em Silva Fenn.} {\bf 1993}, {\em 27}, doi:10.14214/sf.a15668.

\bibitem[MacLean and Lidstone(1982)]{maclean1982}
MacLean, D.A.; Lidstone, R.G.
\newblock Defoliation by Spruce Budworm: Estimation by Ocular and Shoot-Count
  Methods and Variability among Branches, Trees, and Stands.
\newblock {\em Can. J. For. Res.} {\bf 1982}, {\em
  12},~582--594, doi:10.1139/x82-090.

\bibitem[{Johnstone Iain M.} and {Titterington D.
  Michael}(2009)]{johnstoneiainm.2009}
{Johnstone Iain M.}.; {Titterington D. Michael}.
\newblock Statistical Challenges of High-Dimensional Data.
\newblock {\em Philos. Trans. R. Soc. A Math. Phys. Eng. Sci.} {\bf 2009}, {\em 367},~4237--4253.

\bibitem[Bommert \em{et~al.}(2020)Bommert, Sun, Bischl, Rahnenf{\"u}hrer, and
  Lang]{bommert2020}
Bommert, A.; Sun, X.; Bischl, B.; Rahnenf{\"u}hrer, J.; Lang, M.
\newblock Benchmark for Filter Methods for Feature Selection in
  High-Dimensional Classification Data.
\newblock {\em Comput. Stat. Data Anal.} {\bf 2020}, {\em
  143},~106839, doi:10.1016/j.csda.2019.106839.

\bibitem[Das(2001)]{das2001}
Das, S.
\newblock \emph{Filters, {{Wrappers}} and a {{Boosting}}-{{Based Hybrid}} for
  {{Feature Selection}}};
\newblock  {{ICML}}, Citeseer, 2001.

\bibitem[Guyon and Elisseeff(2003)]{guyon2003}
Guyon, I.; Elisseeff, A.
\newblock An Introduction to Variable and Feature Selection.
\newblock {\em J. Mach. Learn. Res.} {\bf 2003}, {\em
  3},~1157--1182.

\bibitem[Jolliffe and Cadima(2016)]{jolliffe2016}
Jolliffe, I.; Cadima, J.
\newblock Principal Component Analysis: A Review and Recent Developments.
\newblock {\em Philos. Trans. R. Soc. A Math. Phys. Eng. Sci.} {\bf 2016}, {\em 374},~20150202.

\bibitem[Drot{\'a}r \em{et~al.}(2017)Drot{\'a}r, {\v S}imo{\v n}{\'a}k,
  Pietrikov{\'a}, Chovanec, Chovancov{\'a}, {\'A}d{\'a}m, Szab{\'o},
  Bal{\'a}{\v z}, and Bi{\v n}as]{drotar2017}
Drot{\'a}r, P.; {\v S}imo{\v n}{\'a}k, S.; Pietrikov{\'a}, E.; Chovanec, M.;
  Chovancov{\'a}, E.; {\'A}d{\'a}m, N.; Szab{\'o}, C.; Bal{\'a}{\v z}, A.;
  Bi{\v n}as, M.
\newblock Comparison of {{Filter Techniques}} for {{Two}}-{{Step Feature
  Selection}}.
\newblock {\em Comput. Inform.} {\bf 2017}, {\em 36},~597--617, doi:10.4149/cai\_2017\_3\_597.

\bibitem[Drot{\'a}r \em{et~al.}(2015)Drot{\'a}r, Gazda, and
  Sm{\'e}kal]{drotar2015}
Drot{\'a}r, P.; Gazda, J.; Sm{\'e}kal, Z.
\newblock An Experimental Comparison of Feature Selection Methods on Two-Class
  Biomedical Datasets.
\newblock {\em Comput. Biol. Med.} {\bf 2015}, {\em 66},~1--10, doi:10.1016/j.compbiomed.2015.08.010.

\bibitem[Abeel \em{et~al.}(2010)Abeel, Helleputte, {Van de Peer}, Dupont, and
  Saeys]{abeel2010}
Abeel, T.; Helleputte, T.; {Van de Peer}, Y.; Dupont, P.; Saeys, Y.
\newblock Robust Biomarker Identification for Cancer Diagnosis with Ensemble
  Feature Selection Methods.
\newblock {\em Bioinformatics} {\bf 2010}, {\em 26},~392--398.

\bibitem[Dietterich(2000)]{dietterich2000}
Dietterich, T.G.
\newblock Ensemble {{Methods}} in {{Machine Learning}}.
\newblock  In Proceedings of the {{First International Workshop}} on {{Multiple
  Classifier Systems}}, Cagliari, Italy, 21--23 June 2000; {Springer}:  Berlin/Heidelberg, Germany,
 2000; pp. 1--15.

\bibitem[Polikar(2012)]{polikar2012}
Polikar, R.
\newblock Ensemble {{Learning}}. In {\em Ensemble {{Machine Learning}}: Methods
  and {{Applications}}}; Zhang, C., Ma, Y., Eds.; {Springer}: {Boston, MA, USA},
  2012; pp. 1--34, doi:10.1007/978-1-4419-9326-7\_1.

\bibitem[Feurer \em{et~al.}(2015)Feurer, Klein, Eggensperger, Springenberg,
  Blum, and Hutter]{feurer2015}
Feurer, M.; Klein, A.; Eggensperger, K.; Springenberg, J.; Blum, M.; Hutter, F.
\newblock Efficient and {{Robust Automated Machine Learning}}. In {\em Advances
  in {{Neural Information Processing Systems}} 28}; Cortes, C., Lawrence, N.D.,
  Lee, D.D., Sugiyama, M., Garnett, R., Eds.; {Curran Associates, Inc.}:  2015;
  pp. 2962--2970.

\bibitem[{Bol{\'o}n-Canedo} and {Alonso-Betanzos}(2019)]{bolon-canedo2019}
{Bol{\'o}n-Canedo}, V.; {Alonso-Betanzos}, A.
\newblock Ensembles for Feature Selection: A Review and Future Trends.
\newblock {\em Inf. Fusion} {\bf 2019}, {\em 52},~1--12, doi:10.1016/j.inffus.2018.11.008.

\bibitem[Pearson(1901)]{pearson1901}
Pearson, K.
\newblock {{LIII}}. {{On}} Lines and Planes of Closest Fit to Systems of Points
  in Space.
\newblock {\em  Lond. Edinb. Dublin Philos. Mag. J. Sci.} {\bf 1901}, {\em 2},~559--572.

\bibitem[Quinlan(1986)]{quinlan1986}
Quinlan, J.R.
\newblock Induction of Decision Trees.
\newblock {\em Mach. Learn.} {\bf 1986}, {\em 1},~81--106.

\bibitem[Zhao(2013)]{zhao2013}
Zhao, X.M.
\newblock Maximum {{Relevance}}/{{Minimum Redundancy}} ({{MRMR}}). In {\em
  Encyclopedia of {{Systems Biology}}}; Dubitzky, W., Wolkenhauer, O., Cho,
  K.H.,Yokota, H., Eds.; {Springer}: {New York, NY, USA},  2013; pp.
  1191--1192, doi:10.1007/978-1-4419-9863-7\_432.

\bibitem[Zuber and Strimmer(2011)]{zuber2011}
Zuber, V.; Strimmer, K.
\newblock High-{{Dimensional Regression}} and {{Variable Selection Using CAR
  Scores}}.
\newblock {\em Stat. Appl. Genet. Mol. Biol.} {\bf
  2011}, {\em 10}, 1-27, doi:10/c62xhp.

\bibitem[Kira and Rendell(1992)]{kira1992}
Kira, K.; Rendell, L.A.
\newblock The Feature Selection Problem: Traditional Methods and a New
  Algorithm.
\newblock  In Proceedings of the Tenth National Conference on {{Artificial}}
  Intelligence, San Jose, CA, USA, 12--16 July 1992; {AAAI Press}:  1992; pp. 129--134.

\bibitem[Fleuret(2004)]{fleuret2004}
Fleuret, F.
\newblock Fast {{Binary Feature Selection}} with {{Conditional Mutual
  Information}}.
\newblock {\em  J. Mach. Learn. Res.} {\bf 2004}, {\em
  5},~1531--1555.

\bibitem[Shannon(1948)]{shannon1948}
Shannon, C.E.
\newblock A Mathematical Theory of Communication.
\newblock {\em  Bell Syst. Tech. J.} {\bf 1948}, {\em
  27},~379--423, doi:10.1002/j.1538-7305.1948.tb01338.x.

\bibitem[Hastie \em{et~al.}(2001)Hastie, Friedman, and Tibshirani]{hastie2001}
Hastie, T.; Friedman, J.; Tibshirani, R.
\newblock {\em The Elements of Statistical Learning}; {Springer}: New York, NY, USA,
  2001, doi:10.1007/978-0-387-21606-5.

\bibitem[Pe{\~n}a \em{et~al.}(2017)Pe{\~n}a, Liao, and Brenning]{pena2017}
Pe{\~n}a, M.; Liao, R.; Brenning, A.
\newblock Using Spectrotemporal Indices to Improve the Fruit-Tree Crop
  Classification Accuracy.
\newblock {\em ISPRS J. Photogramm. Remote Sens.} {\bf 2017},
  {\em 128},~158--169.

\bibitem[Bischl \em{et~al.}(2017)Bischl, Richter, Bossek, Horn, Thomas, and
  Lang]{mlrmbo}
Bischl, B.; Richter, J.; Bossek, J.; Horn, D.; Thomas, J.; Lang, M.
\newblock {{mlrMBO}}: A {{Modular Framework}} for {{Model}}-{{Based
  Optimization}} of {{Expensive Black}}-{{Box Functions}}.
\newblock {\em arXiv} {\bf 2017}, arXiv:1703.03373.

\bibitem[Binder \em{et~al.}(2020)Binder, Moosbauer, Thomas, and
  Bischl]{binder2020}
Binder, M.; Moosbauer, J.; Thomas, J.; Bischl, B.
\newblock Multi-{{Objective Hyperparameter Tuning}} and {{Feature Selection}}
  Using {{Filter Ensembles}}.
\newblock {\em arXiv} {\bf 2020},
  arXiv:1912.12912

\bibitem[Schratz \em{et~al.}(2019)Schratz, Muenchow, Iturritxa, Richter, and
  Brenning]{schratz2019}
Schratz, P.; Muenchow, J.; Iturritxa, E.; Richter, J.; Brenning, A.
\newblock Hyperparameter Tuning and Performance Assessment of Statistical and
  Machine-Learning Algorithms Using Spatial Data.
\newblock {\em Ecol. Model.} {\bf 2019}, {\em 406},~109--120, doi:10.1016/j.ecolmodel.2019.06.002.

\bibitem[Hutter \em{et~al.}(2011)Hutter, Hoos, and {Leyton-Brown}]{hutter2011}
Hutter, F.; Hoos, H.H.; {Leyton-Brown}, K.
\newblock \emph{Sequential Model-Based Optimization for General Algorithm
  Configuration}; Lecture {{Notes}} in {{Computer Science}}; {Springer:
  Berlin/Heidelberg, Germany},  2011; pp. 507--523, doi:10.1007/978-3-642-25566-3\_40.

\bibitem[Jones \em{et~al.}(1998)Jones, Schonlau, and Welch]{jones1998}
Jones, D.R.; Schonlau, M.; Welch, W.J.
\newblock Efficient Global Optimization of Expensive Black-Box Functions.
\newblock {\em J. Glob. Optim.} {\bf 1998}, {\em 13},~455--492, doi:10.1023/a:1008306431147.

\bibitem[Bergstra and Bengio(2012)]{bergstra2012}
Bergstra, J.; Bengio, Y.
\newblock Random {{Search}} for {{Hyper}}-Parameter {{Optimization}}.
\newblock {\em J. Mach. Learn. Res.} {\bf 2012}, {\em 13},~281--305.

\bibitem[Brenning(2012)]{sperrorest}
Brenning, A.
\newblock Spatial Cross-Validation and Bootstrap for the Assessment of
  Prediction Rules in Remote Sensing: The {{R}} Package Sperrorest.
\newblock  In Proceedings of the 2012 {{IEEE International Geoscience}} and {{Remote Sensing
  Symposium}}, Munich, Germany, 22--27 July 2012; {R} Package Version 2.1.0; {IEEE}:  2012,
\newblock
  doi:{\changeurlcolor{black}\href{https://doi.org/10.1109/igarss.2012.6352393}{\detokenize{10.1109/igarss.2012.6352393}}}.

\bibitem[Friedman(2001)]{friedman2001}
Friedman, J.H.
\newblock Greedy Function Approximation: A Gradient Boosting Machine.
\newblock {\em Ann. Stat.} {\bf 2001}, {\em 29},~1189--1232, doi:10.1214/aos/1013203451.

\bibitem[Greenwell \em{et~al.}(2018)Greenwell, Boehmke, and
  McCarthy]{greenwell2018}
Greenwell, B.M.; Boehmke, B.C.; McCarthy, A.J.
\newblock A {{Simple}} and {{Effective Model}}-{{Based Variable Importance
  Measure}}.
\newblock {\em arXiv} {\bf 2018}, arXiv:1805.04755

\bibitem[Molnar(2019)]{molnar2019}
Molnar, C.
\newblock {\em Interpretable Machine Learning---{{A Guide}} for {{Making Black
  Box Models Explainable}}}; self-published: Munich, Germany, 2019.

\bibitem[Brenning(2021)]{brenning2021}
Brenning, A.
\newblock Transforming {{Feature Space}} to {{Interpret Machine Learning
  Models}}.
\newblock {\em arXiv} {\bf 2021}, arXiv:2104.04295

\bibitem[Apley and Zhu(2019)]{apley2019}
Apley, D.W.; Zhu, J.
\newblock Visualizing the {{Effects}} of {{Predictor Variables}} in {{Black Box
  Supervised Learning Models}}.
\newblock {\em arXiv} {\bf 2019}, arXiv:1612.08468

\bibitem[{R Core Team}(2019)]{rcoreteam2019}
{R Core Team}.
\newblock {\em R: A {{Language}} and {{Environment}} for {{Statistical
  Computing}}};  R Core Team: 2019, Vienna, Austria.

\bibitem[Chen and Guestrin(2016)]{chen2016}
Chen, T.; Guestrin, C.
\newblock {{XGBoost}}: A {{Scalable Tree Boosting System}}.
\newblock  In Proceedings of the {{22nd ACM SIGKDD International Conference}} on
  {{Knowledge Discovery}} and {{Data Mining}}, {{KDD}} '16, San Francisco, CA, USA, 13--17 August 2016; {ACM}: {New York, NY, USA},
  2016; pp. 785--794, doi:10.1145/2939672.2939785.

\bibitem[Karatzoglou \em{et~al.}(2004)Karatzoglou, Smola, Hornik, and
  Zeileis]{kernlab}
Karatzoglou, A.; Smola, A.; Hornik, K.; Zeileis, A.
\newblock Kernlab---{{An S4 Package}} for {{Kernel Methods}} in {{R}}.
\newblock {\em J. Stat. Softw.} {\bf 2004}, {\em 11},~1--20, doi:10.18637/jss.v011.i09.

\bibitem[Friedman \em{et~al.}(2010)Friedman, Hastie, and Tibshirani]{glmnet}
Friedman, J.; Hastie, T.; Tibshirani, R.
\newblock Regularization Paths for Generalized Linear Models via Coordinate
  Descent.
\newblock {\em J. Stat. Softw.} {\bf 2010}, {\em 33},~1--22, doi:10.18637/jss.v033.i01.

\bibitem[Kursa(2018)]{praznik}
Kursa, M.B.
\newblock {\em Praznik: Collection of {{Information}}-{{Based Feature Selection
  Filters}}};  2018.

\bibitem[Zawadzki and Kosinski(2019)]{fselectorrcpp}
Zawadzki, Z.; Kosinski, M.
\newblock {\em {{FSelectorRcpp}}: '{{Rcpp}}' {{Implementation}} of
  '{{FSelector}}' {{Entropy}}-{{Based Feature Selection Algorithms}} with a
  {{Sparse Matrix Support}}};  2019.

\bibitem[Bischl \em{et~al.}(2016)Bischl, Lang, Kotthoff, Schiffner, Richter,
  Studerus, Casalicchio, and Jones]{mlr}
Bischl, B.; Lang, M.; Kotthoff, L.; Schiffner, J.; Richter, J.; Studerus, E.;
  Casalicchio, G.; Jones, Z.M.
\newblock mlr: Machine learning in {R}.
\newblock {\em J. Mach. Learn. Res.} {\bf 2016}, {\em
  17},~1--5.

\bibitem[Landau(2018)]{drake}
Landau, W.M.
\newblock The {drake} {{R}} Package: A Pipeline Toolkit for Reproducibility and
  High-Performance Computing.
\newblock {\em J. Open Source Softw.} {\bf 2018}, {\em 3}.

\bibitem[Ghosh \em{et~al.}(2019)Ghosh, Adhikary, Ghosh, Sardar, Begum, and
  Sarkar]{ghosh2019}
Ghosh, M.; Adhikary, S.; Ghosh, K.K.; Sardar, A.; Begum, S.; Sarkar, R.
\newblock Genetic Algorithm Based Cancerous Gene Identification from Microarray
  Data Using Ensemble of Filter Methods.
\newblock {\em Med. Biol. Eng. Comput.} {\bf 2019}, {\em
  57},~159--176, doi:10.1007/s11517-018-1874-4.

\bibitem[Horler \em{et~al.}(1983)Horler, Dockray, and Barber]{horler1983}
Horler, D.N.H.; Dockray, M.; Barber, J.
\newblock The Red Edge of Plant Leaf Reflectance.
\newblock {\em Int. J. Remote Sens.} {\bf 1983}, {\em
  4},~273--288, doi:10.1080/01431168308948546.

\bibitem[Hais \em{et~al.}(2019)Hais, Hellebrandov{\'a}, and {\v
  S}r{\'a}mek]{hais2019}
Hais, M.; Hellebrandov{\'a}, K.N.; {\v S}r{\'a}mek, V.
\newblock Potential of {{Landsat}} Spectral Indices in Regard to the Detection
  of Forest Health Changes Due to Drought Effects.
\newblock {\em J. For. Sci.} {\bf 2019}, {\em 65},~70--78, doi:10.17221/137/2018-JFS.

\bibitem[Gao(1996)]{gao1996}
Gao, B.C.
\newblock {{NDWI}}\textemdash{{A}} Normalized Difference Water Index for Remote
  Sensing of Vegetation Liquid Water from Space.
\newblock {\em Remote Sens. Environ.} {\bf 1996}, {\em 58},~257--266.

\bibitem[Lehnert \em{et~al.}(2016)Lehnert, Meyer, and Bendix]{lehnert2016}
Lehnert, L.W.; Meyer, H.; Bendix, J.
\newblock {\em Hsdar: Manage, Analyse and Simulate Hyperspectral Data in
  {{R}}};  2016.

\bibitem[{de Beurs} and Townsend(2008)]{debeurs2008}
{de Beurs}, K.M.; Townsend, P.A.
\newblock Estimating the Effect of Gypsy Moth Defoliation Using {{MODIS}}.
\newblock {\em Remote Sens. Environ.} {\bf 2008}, {\em
  112},~3983--3990.

\bibitem[Rengarajan and Schott(2016)]{rengarajan2016}
Rengarajan, R.; Schott, J.R.
\newblock Modeling Forest Defoliation Using Simulated {{BRDF}} and Assessing
  Its Effect on Reflectance and Sensor Reaching Radiance.
\newblock  In \emph{Remote {{Sensing}} and {{Modeling}} of {{Ecosystems}} for
  {{Sustainability XIII}}}; {International Society for Optics and Photonics}:
  2016; San Diego, California, United States, Volume 9975, p. 997503, doi:10.1117/12.2235391.

\bibitem[Meng \em{et~al.}(2018)Meng, Dennison, Zhao, Shendryk, Rickert,
  Hanavan, Cook, and Serbin]{meng2018}
Meng, R.; Dennison, P.E.; Zhao, F.; Shendryk, I.; Rickert, A.; Hanavan, R.P.;
  Cook, B.D.; Serbin, S.P.
\newblock Mapping Canopy Defoliation by Herbivorous Insects at the Individual
  Tree Level Using Bi-Temporal Airborne Imaging Spectroscopy and {{LiDAR}}
  Measurements.
\newblock {\em Remote Sens. Environ.} {\bf 2018}, {\em 215},~170--183.

\bibitem[K{\"a}lin \em{et~al.}(2019)K{\"a}lin, Lang, Hug, Gessler, and
  Wegner]{kalin2019}
K{\"a}lin, U.; Lang, N.; Hug, C.; Gessler, A.; Wegner, J.D.
\newblock Defoliation Estimation of Forest Trees from Ground-Level Images.
\newblock {\em Remote Sens. Environ.} {\bf 2019}, {\em 223},~143--153.

\bibitem[Goodbody \em{et~al.}(2018)Goodbody, Coops, Hermosilla, Tompalski,
  McCartney, and MacLean]{goodbody2018}
Goodbody, T.R.H.; Coops, N.C.; Hermosilla, T.; Tompalski, P.; McCartney, G.;
  MacLean, D.A.
\newblock Digital Aerial Photogrammetry for Assessing Cumulative Spruce Budworm
  Defoliation and Enhancing Forest Inventories at a Landscape-Level.
\newblock {\em ISPRS J. Photogramm. Remote Sens.} {\bf 2018},
  {\em 142},~1--11, doi:10.1016/j.isprsjprs.2018.05.012.

\bibitem[Hlebarska and Georgieva(2018)]{hlebarska2018}
Hlebarska, S.; Georgieva, M.
\newblock {Distribution of the invasive pathogen Diplodia sapinea on \emph{Pinus} spp.
  in Bulgaria.}
\newblock In Proceedings of the 90 Years Forest Research Institute---For
  the Society and Nature, Sofia, Bulgaria, 24--26 October 2018;   pp.
  61--70.

\bibitem[Kaya \em{et~al.}(2019)Kaya, Yeltekin, Lehtijarvi, Lehtijarvi, and
  Woodward]{kaya2019}
Kaya, A.G.A.; Yeltekin, {\c S}.; Lehtijarvi, T.D.; Lehtijarvi, A.; Woodward, S.
\newblock Severity of {{Diplodia}} Shoot Blight (Caused by {{Diplodia}}
  Sapinea) Was Greatest on {{Pinus}} Sylvestris and {{Pinus}} Nigra in a
  Plantation Containing Five Pine Species.
\newblock {\em Phytopathol. Mediterr.} {\bf 2019}, {\em 58},~249--259.

\bibitem[Belgiu and Dr{\u a}gu{\c t}(2016)]{belgiu2016}
Belgiu, M.; Dr{\u a}gu{\c t}, L.
\newblock Random Forest in Remote Sensing: A Review of Applications and Future
  Directions.
\newblock {\em ISPRS J. Photogramm. Remote Sens.} {\bf 2016},
  {\em 114},~24--31, doi:10.1016/j.isprsjprs.2016.01.011.

\bibitem[Xia \em{et~al.}(2015)Xia, Liao, Chanussot, Du, Song, and
  Philips]{xia2015}
Xia, J.; Liao, W.; Chanussot, J.; Du, P.; Song, G.; Philips, W.
\newblock Improving {{Random Forest with Ensemble}} of {{Features}} and
  {{Semisupervised Feature Extraction}}.
\newblock {\em IEEE Geosci. Remote Sens. Lett.} {\bf 2015}, {\em
  12},~1471--1475, doi:10.1109/LGRS.2015.2409112.

\bibitem[Fassnacht \em{et~al.}(2014)Fassnacht, Neumann, F{\"o}rster,
  Buddenbaum, Ghosh, Clasen, Joshi, and Koch]{fassnacht2014}
Fassnacht, F.E.; Neumann, C.; F{\"o}rster, M.; Buddenbaum, H.; Ghosh, A.;
  Clasen, A.; Joshi, P.K.; Koch, B.
\newblock Comparison of {{Feature Reduction Algorithms}} for {{Classifying Tree
  Species With Hyperspectral Data}} on {{Three Central European Test Sites}}.
\newblock {\em IEEE J. Sel. Top. Appl. Earth Obs. Remote Sens.} {\bf 2014}, {\em 7},~2547--2561, doi:10.1109/JSTARS.2014.2329390.

\bibitem[Feng \em{et~al.}(2016)Feng, Jiao, Liu, Sun, and Zhang]{feng2016}
Feng, J.; Jiao, L.; Liu, F.; Sun, T.; Zhang, X.
\newblock Unsupervised Feature Selection Based on Maximum Information and
  Minimum Redundancy for Hyperspectral Images.
\newblock {\em Pattern Recognit.} {\bf 2016}, {\em 51},~295--309, doi:10.1016/j.patcog.2015.08.018.

\bibitem[Georganos \em{et~al.}(2018)Georganos, Grippa, Vanhuysse, Lennert,
  Shimoni, Kalogirou, and Wolff]{georganos2018}
Georganos, S.; Grippa, T.; Vanhuysse, S.; Lennert, M.; Shimoni, M.; Kalogirou,
  S.; Wolff, E.
\newblock Less Is More: Optimizing Classification Performance through Feature
  Selection in a Very-High-Resolution Remote Sensing Object-Based Urban
  Application.
\newblock {\em GISci. Remote Sens.} {\bf 2018}, {\em 55},~221--242, doi:10.1080/15481603.2017.1408892.

\bibitem[Rochac and Zhang(2016)]{rochac2016}
Rochac, J.F.R.; Zhang, N.
\newblock Feature Extraction in Hyperspectral Imaging Using Adaptive Feature
  Selection Approach.
\newblock  In Proceedings of the 2016 {{Eighth International Conference}} on {{Advanced Computational
  Intelligence}} ({{ICACI}}),  Chiang Mai, Thailand, 14--16 February 2016; pp. 36--40, doi:10.1109/ICACI.2016.7449799.

\bibitem[Pal and Foody(2010)]{pal2010}
Pal, M.; Foody, G.M.
\newblock Feature {{Selection}} for {{Classification}} of {{Hyperspectral
  Data}} by {{SVM}}.
\newblock {\em IEEE Trans. Geosci. Remote Sens.} {\bf 2010},
  {\em 48},~2297--2307, doi:10.1109/TGRS.2009.2039484.

\bibitem[Keller \em{et~al.}(2016)Keller, Braun, Hinz, and Weinmann]{keller2016}
Keller, S.; Braun, A.C.; Hinz, S.; Weinmann, M.
\newblock Investigation of the Impact of Dimensionality Reduction and Feature
  Selection on the Classification of Hyperspectral {{EnMAP}} Data.
\newblock  In Proceedings of the 2016 8th {{Workshop}} on {{Hyperspectral Image}} and {{Signal
  Processing}}: Evolution in {{Remote Sensing}} ({{WHISPERS}}),  Los Angeles, CA, USA, 21--24 August 2016; pp. 1--5, doi:10.1109/WHISPERS.2016.8071759.

\bibitem[Xu \em{et~al.}(2019)Xu, Zhao, Yin, Zhang, Liu, and Yang]{xu2019}
Xu, S.; Zhao, Q.; Yin, K.; Zhang, F.; Liu, D.; Yang, G.
\newblock Combining Random Forest and Support Vector Machines for Object-Based
  Rural-Land-Cover Classification Using High Spatial Resolution Imagery.
\newblock {\em J. Appl. Remote Sens.} {\bf 2019}, {\em 13},~014521, doi:10.1117/1.JRS.13.014521.

\bibitem[Shendryk \em{et~al.}(2016)Shendryk, Broich, Tulbure, McGrath, Keith,
  and Alexandrov]{shendryk2016}
Shendryk, I.; Broich, M.; Tulbure, M.G.; McGrath, A.; Keith, D.; Alexandrov,
  S.V.
\newblock Mapping Individual Tree Health Using Full-Waveform Airborne Laser
  Scans and Imaging Spectroscopy: A Case Study for a Floodplain Eucalypt
  Forest.
\newblock {\em Remote Sens. Environ.} {\bf 2016}, {\em 187},~202--217.

\bibitem[Ludwig \em{et~al.}(2019)Ludwig, Morgenthal, Detsch, Higginbottom,
  Lezama~Valdes, Nau{\ss}, and Meyer]{ludwig2019}
Ludwig, M.; Morgenthal, T.; Detsch, F.; Higginbottom, T.P.; Lezama~Valdes, M.;
  Nau{\ss}, T.; Meyer, H.
\newblock Machine Learning and Multi-Sensor Based Modelling of Woody Vegetation
  in the {{Molopo Area}}, {{South Africa}}.
\newblock {\em Remote Sens. Environ.} {\bf 2019}, {\em 222},~195--203.

\bibitem[Meyer \em{et~al.}(2018)Meyer, Reudenbach, Hengl, Katurji, and
  Nauss]{meyer2018}
Meyer, H.; Reudenbach, C.; Hengl, T.; Katurji, M.; Nauss, T.
\newblock Improving Performance of Spatio-Temporal Machine Learning Models
  Using Forward Feature Selection and Target-Oriented Validation.
\newblock {\em Environ. Model. Softw.} {\bf 2018}, {\em
  101},~1--9, doi:10.1016/j.envsoft.2017.12.001.

\bibitem[Zandler \em{et~al.}(2015)Zandler, Brenning, and Samimi]{zandler2015}
Zandler, H.; Brenning, A.; Samimi, C.
\newblock Quantifying Dwarf Shrub Biomass in an Arid Environment: Comparing
  Empirical Methods in a High Dimensional Setting.
\newblock {\em Remote Sens. Environ.} {\bf 2015}, {\em 158},~140--155, doi:10.1016/j.rse.2014.11.007.

\bibitem[Guo \em{et~al.}(2019)Guo, Chung, Li, and Zhang]{guo2019}
Guo, Y.; Chung, F.L.; Li, G.; Zhang, L.
\newblock Multi-{{Label Bioinformatics Data Classification with Ensemble
  Embedded Feature Selection}}.
\newblock {\em IEEE Access} {\bf 2019}, {\em 7},~103863--103875, doi:10.1109/ACCESS.2019.2931035.

\bibitem[Radovic \em{et~al.}(2017)Radovic, Ghalwash, Filipovic, and
  Obradovic]{radovic2017}
Radovic, M.; Ghalwash, M.; Filipovic, N.; Obradovic, Z.
\newblock Minimum Redundancy Maximum Relevance Feature Selection Approach for
  Temporal Gene Expression Data.
\newblock {\em BMC Bioinform.} {\bf 2017}, {\em 18},~9, doi:10.1186/s12859-016-1423-9.

\bibitem[Vogelmann \em{et~al.}(1993)Vogelmann, Rock, and Moss]{vogelmann1993}
Vogelmann, J.E.; Rock, B.N.; Moss, D.M.
\newblock Red Edge Spectral Measurements from Sugar Maple Leaves.
\newblock {\em Int. J. Remote Sens.} {\bf 1993}, {\em
  14},~1563--1575, doi:10.1080/01431169308953986.

\bibitem[Boochs \em{et~al.}(1990)Boochs, Kupfer, Dockter, and
  K{\"u}hbauch]{boochs1990}
Boochs, F.; Kupfer, G.; Dockter, K.; K{\"u}hbauch, W.
\newblock Shape of the Red Edge as Vitality Indicator for Plants.
\newblock {\em Int. J. Remote Sens.} {\bf 1990}, {\em
  11},~1741--1753, doi:10.1080/01431169008955127.

\bibitem[Nagler \em{et~al.}(2003)Nagler, Inoue, Glenn, Russ, and
  Daughtry]{nagler2003}
Nagler, P.L.; Inoue, Y.; Glenn, E.P.; Russ, A.L.; Daughtry, C.S.T.
\newblock Cellulose Absorption Index ({{CAI}}) to Quantify Mixed
  Soil\textendash Plant Litter Scenes.
\newblock {\em Remote Sens. Environ.} {\bf 2003}, {\em 87},~310--325, doi:10.1016/j.rse.2003.06.001.

\bibitem[Walthall \em{et~al.}(1994)Walthall, Daughtry, Chappelle, Mcmurtrey,
  and Kim]{walthall1994}
Walthall, C.L.; Daughtry, C.S.T.; Chappelle, E.W.; Mcmurtrey, J.E.; Kim, M.S.
\newblock \emph{The {{Use}} of {{High Spectral Resolution Bands}} for {{Estimating
  Absorbed Photosynthetically Active Radiation}} ({{A Par}})}; NASA Gov, Washington, { 1994}.

\bibitem[Carter(1994)]{carter1994}
Carter, G.A.
\newblock Ratios of Leaf Reflectances in Narrow Wavebands as Indicators of
  Plant Stress.
\newblock {\em Int. J. Remote Sens.} {\bf 1994}, {\em
  15},~697--703, doi:10.1080/01431169408954109.

\bibitem[{Zarco-Tejada} \em{et~al.}(2003){Zarco-Tejada}, Pushnik, Dobrowski,
  and Ustin]{zarco-tejada2003}
{Zarco-Tejada}, P.J.; Pushnik, J.C.; Dobrowski, S.; Ustin, S.L.
\newblock Steady-State Chlorophyll a Fluorescence Detection from Canopy
  Derivative Reflectance and Double-Peak Red-Edge Effects.
\newblock {\em Remote Sens. Environ.} {\bf 2003}, {\em 84},~283--294, doi:10.1016/s0034-4257(02)00113-x.

\bibitem[Gitelson \em{et~al.}(2003)Gitelson, Gritz~{\textdagger}, and
  Merzlyak]{gitelson2003}
Gitelson, A.A.; Gritz, Y.; Merzlyak, M.N.
\newblock Relationships between Leaf Chlorophyll Content and Spectral
  Reflectance and Algorithms for Non-Destructive Chlorophyll Assessment in
  Higher Plant Leaves.
\newblock {\em J. Plant Physiol.} {\bf 2003}, {\em 160},~271--282, doi:10.1078/0176-1617-00887.

\bibitem[Oppelt and Mauser(2004)]{oppelt2004}
Oppelt, N.; Mauser, W.
\newblock Hyperspectral Monitoring of Physiological Parameters of Wheat during
  a Vegetation Period Using {{AVIS}} Data.
\newblock {\em Int. J. Remote Sens.} {\bf 2004}, {\em
  25},~145--159, doi:10.1080/0143116031000115300.

\bibitem[Datt(1999)]{datt1999}
Datt, B.
\newblock Visible/near Infrared Reflectance and Chlorophyll Content in
  {{Eucalyptus}} Leaves.
\newblock {\em Int. J. Remote Sens.} {\bf 1999}, {\em
  20},~2741--2759, doi:10.1080/014311699211778.

\bibitem[Datt(1998)]{datt1998}
Datt, B.
\newblock Remote {{Sensing}} of {{Chlorophyll}} a, {{Chlorophyll}} b,
  {{Chlorophyll}} A+b, and {{Total Carotenoid
  Content}} in {{Eucalyptus Leaves}}.
\newblock {\em Remote Sens. Environ.} {\bf 1998}, {\em 66},~111--121, doi:10.1016/s0034-4257(98)00046-7.

\bibitem[Datt(1999)]{datt1999a}
Datt, B.
\newblock Remote {{Sensing}} of {{Water Content}} in {{Eucalyptus Leaves}}.
\newblock {\em Aust. J. Bot.} {\bf 1999}, {\em 47},~909, doi:10.1071/bt98042.

\bibitem[le~Maire \em{et~al.}(2004)le~Maire, Fran{\c c}ois, and
  Dufr{\^e}ne]{maire2004}
le~Maire, G.; Fran{\c c}ois, C.; Dufr{\^e}ne, E.
\newblock Towards Universal Broad Leaf Chlorophyll Indices Using {{PROSPECT}}
  Simulated Database and Hyperspectral Reflectance Measurements.
\newblock {\em Remote Sens. Environ.} {\bf 2004}, {\em 89},~1--28, doi:10.1016/j.rse.2003.09.004.

\bibitem[Lemaire \em{et~al.}(2008)Lemaire, Francois, Soudani, Berveiller,
  Pontailler, Breda, Genet, Davi, and Dufrene]{lemaire2008}
Lemaire, G.; Francois, C.; Soudani, K.; Berveiller, D.; Pontailler, J.; Breda,
  N.; Genet, H.; Davi, H.; Dufrene, E.
\newblock Calibration and Validation of Hyperspectral Indices for the
  Estimation of Broadleaved Forest Leaf Chlorophyll Content, Leaf Mass per
  Area, Leaf Area Index and Leaf Canopy Biomass.
\newblock {\em Remote Sens. Environ.} {\bf 2008}, {\em
  112},~3846--3864, doi:10.1016/j.rse.2008.06.005.

\bibitem[Apan \em{et~al.}(2004)Apan, Held, Phinn, and Markley]{apan2004}
Apan, A.; Held, A.; Phinn, S.; Markley, J.
\newblock Detecting Sugarcane `orange Rust' Disease Using {{EO}}-1 {{Hyperion}}
  Hyperspectral Imagery.
\newblock {\em Int. J. Remote Sens.} {\bf 2004}, {\em
  25},~489--498, doi:10.1080/01431160310001618031.

\bibitem[Pe{\~n}uelas \em{et~al.}(1994)Pe{\~n}uelas, Gamon, Fredeen, Merino,
  and Field]{penuelas1994}
Pe{\~n}uelas, J.; Gamon, J.A.; Fredeen, A.L.; Merino, J.; Field, C.B.
\newblock Reflectance Indices Associated with Physiological Changes in
  Nitrogen- and Water-Limited Sunflower Leaves.
\newblock {\em Remote Sens. Environ.} {\bf 1994}, {\em 48},~135--146, doi:10.1016/0034-4257(94)90136-8.

\bibitem[Huete \em{et~al.}(1997)Huete, Liu, Batchily, and {van
  Leeuwen}]{huete1997}
Huete, A.R.; Liu, H.Q.; Batchily, K.; {van Leeuwen}, W.
\newblock A Comparison of Vegetation Indices over a Global Set of {{TM}} Images
  for {{EOS}}-{{MODIS}}.
\newblock {\em Remote Sens. Environ.} {\bf 1997}, {\em 59},~440--451, doi:10.1016/s0034-4257(96)00112-5.

\bibitem[Wu(2014)]{wu2014}
Wu, W.
\newblock The {{Generalized Difference Vegetation Index}} ({{GDVI}}) for
  {{Dryland Characterization}}.
\newblock {\em Remote Sens.} {\bf 2014}, {\em 6},~1211--1233, doi:10.3390/rs6021211.

\bibitem[Smith \em{et~al.}(1995)Smith, Adams, Stephens, and Hick]{smith1995}
Smith, R.C.G.; Adams, J.; Stephens, D.J.; Hick, P.T.
\newblock Forecasting Wheat Yield in a {{Mediterranean}}-Type Environment from
  the {{NOAA}} Satellite.
\newblock {\em Aust. J. Agric. Res.} {\bf 1995}, {\em
  46},~113, doi:10.1071/ar9950113.

\bibitem[Gitelson \em{et~al.}(1999)Gitelson, Buschmann, and
  Lichtenthaler]{gitelson1999}
Gitelson, A.A.; Buschmann, C.; Lichtenthaler, H.K.
\newblock The {{Chlorophyll Fluorescence Ratio F735}}/{{F700}} as an {{Accurate
  Measure}} of the {{Chlorophyll Content}} in {{Plants}}.
\newblock {\em Remote Sens. Environ.} {\bf 1999}, {\em 69},~296--302, doi:10.1016/s0034-4257(99)00023-1.

\bibitem[Gitelson \em{et~al.}(1996)Gitelson, Kaufman, and
  Merzlyak]{gitelson1996}
Gitelson, A.A.; Kaufman, Y.J.; Merzlyak, M.N.
\newblock Use of a Green Channel in Remote Sensing of Global Vegetation from
  {{EOS}}-{{MODIS}}.
\newblock {\em Remote Sens. Environ.} {\bf 1996}, {\em 58},~289--298, doi:10.1016/s0034-4257(96)00072-7.

\bibitem[Galv{\~a}o \em{et~al.}(2005)Galv{\~a}o, Formaggio, and
  Tisot]{galvao2005}
Galv{\~a}o, L.S.; Formaggio, A.R.; Tisot, D.A.
\newblock Discrimination of Sugarcane Varieties in {{Southeastern Brazil}} with
  {{EO}}-1 {{Hyperion}} Data.
\newblock {\em Remote Sens. Environ.} {\bf 2005}, {\em 94},~523--534, doi:10.1016/j.rse.2004.11.012.

\bibitem[Maccioni \em{et~al.}(2001)Maccioni, Agati, and
  Mazzinghi]{maccioni2001}
Maccioni, A.; Agati, G.; Mazzinghi, P.
\newblock New Vegetation Indices for Remote Measurement of Chlorophylls Based
  on Leaf Directional Reflectance Spectra.
\newblock {\em J. Photochem. Photobiol. B Biol.} {\bf
  2001}, {\em 61},~52--61, doi:10.1016/s1011-1344(01)00145-2.

\bibitem[Daughtry(2000)]{daughtry2000}
Daughtry, C.
\newblock Estimating {{Corn Leaf Chlorophyll Concentration}} from {{Leaf}} and
  {{Canopy Reflectance}}.
\newblock {\em Remote Sens. Environ.} {\bf 2000}, {\em 74},~229--239, doi:10.1016/s0034-4257(00)00113-9.

\bibitem[Wu \em{et~al.}(2008)Wu, Niu, Tang, and Huang]{wu2008}
Wu, C.; Niu, Z.; Tang, Q.; Huang, W.
\newblock Estimating Chlorophyll Content from Hyperspectral Vegetation Indices:
  Modeling and Validation.
\newblock {\em Agric. For. Meteorol.} {\bf 2008}, {\em
  148},~1230--1241, doi:10.1016/j.agrformet.2008.03.005.

\bibitem[Sims and Gamon(2002)]{sims2002}
Sims, D.A.; Gamon, J.A.
\newblock Relationships between Leaf Pigment Content and Spectral Reflectance
  across a Wide Range of Species, Leaf Structures and Developmental Stages.
\newblock {\em Remote Sens. Environ.} {\bf 2002}, {\em 81},~337--354, doi:10.1016/s0034-4257(02)00010-x.

\bibitem[{Hern{\'a}ndez-Clemente} \em{et~al.}(2011){Hern{\'a}ndez-Clemente},
  {Navarro-Cerrillo}, Su{\'a}rez, Morales, and
  {Zarco-Tejada}]{hernandez-clemente2011}
{Hern{\'a}ndez-Clemente}, R.; {Navarro-Cerrillo}, R.M.; Su{\'a}rez, L.;
  Morales, F.; {Zarco-Tejada}, P.J.
\newblock Assessing Structural Effects on {{PRI}} for Stress Detection in
  Conifer Forests.
\newblock {\em Remote Sens. Environ.} {\bf 2011}, {\em
  115},~2360--2375, doi:10.1016/j.rse.2011.04.036.

\bibitem[Qi \em{et~al.}(1994)Qi, Chehbouni, Huete, Kerr, and
  Sorooshian]{qi1994}
Qi, J.; Chehbouni, A.; Huete, A.R.; Kerr, Y.H.; Sorooshian, S.
\newblock A Modified Soil Adjusted Vegetation Index.
\newblock {\em Remote Sens. Environ.} {\bf 1994}, {\em 48},~119--126, doi:10.1016/0034-4257(94)90134-1.

\bibitem[Hunt and Rock(1989)]{hunt1989}
Hunt, E.; Rock, B.
\newblock Detection of Changes in Leaf Water Content Using {{Near}}- and
  {{Middle}}-{{Infrared}} Reflectances.
\newblock {\em Remote Sens. Environ.} {\bf 1989}, {\em 30},~43--54, doi:10.1016/0034-4257(89)90046-1.

\bibitem[Chen(1996)]{chen1996}
Chen, J.M.
\newblock Evaluation of Vegetation Indices and a {{Modified Simple Ratio}} for
  Boreal Applications;
\newblock {\emph{Canadian Journal of Remote Sensing}}:  1996, \emph{22}, 229-242, doi:10.4095/218303.

\bibitem[Dash and Curran(2007)]{dash2007}
Dash, J.; Curran, P.
\newblock Evaluation of the {{MERIS}} Terrestrial Chlorophyll Index ({{MTCI}}).
\newblock {\em Adv. Space Res.} {\bf 2007}, {\em 39},~100--104, doi:10.1016/j.asr.2006.02.034.

\bibitem[Haboudane \em{et~al.}(2002)Haboudane, Miller, Tremblay,
  {Zarco-Tejada}, and Dextraze]{haboudane2002}
Haboudane, D.; Miller, J.R.; Tremblay, N.; {Zarco-Tejada}, P.J.; Dextraze, L.
\newblock Integrated Narrow-Band Vegetation Indices for Prediction of Crop
  Chlorophyll Content for Application to Precision Agriculture.
\newblock {\em Remote Sens. Environ.} {\bf 2002}, {\em 81},~416--426, doi:10.1016/s0034-4257(02)00018-4.

\bibitem[Serrano \em{et~al.}(2002)Serrano, Pe{\~n}uelas, and
  Ustin]{serrano2002}
Serrano, L.; Pe{\~n}uelas, J.; Ustin, S.L.
\newblock Remote Sensing of Nitrogen and Lignin in {{Mediterranean}} Vegetation
  from {{AVIRIS}} Data.
\newblock {\em Remote Sens. Environ.} {\bf 2002}, {\em 81},~355--364, doi:10.1016/s0034-4257(02)00011-1.

\bibitem[Tucker(1979)]{tucker1979}
Tucker, C.J.
\newblock Red and Photographic Infrared Linear Combinations for Monitoring
  Vegetation.
\newblock {\em Remote Sens. Environ.} {\bf 1979}, {\em 8},~127--150, doi:10.1016/0034-4257(79)90013-0.

\bibitem[Gitelson and Merzlyak(1994)]{gitelson1994}
Gitelson, A.; Merzlyak, M.N.
\newblock Quantitative Estimation of Chlorophyll-a Using Reflectance Spectra:
  Experiments with Autumn Chestnut and Maple Leaves.
\newblock {\em J. Photochem. Photobiol. B Biol.} {\bf
  1994}, {\em 22},~247--252, doi:10.1016/1011-1344(93)06963-4.

\bibitem[Guanter \em{et~al.}(2005)Guanter, Alonso, and Moreno]{guanter2005}
Guanter, L.; Alonso, L.; Moreno, J.
\newblock A Method for the Surface Reflectance Retrieval from
  {{PROBA}}/{{CHRIS}} Data over Land: Application to {{ESA SPARC}} Campaigns.
\newblock {\em IEEE Trans. Geosci. Remote Sens.} {\bf 2005},
  {\em 43},~2908--2917, doi:10.1109/tgrs.2005.857915.

\bibitem[Rondeaux \em{et~al.}(1996)Rondeaux, Steven, and Baret]{rondeaux1996}
Rondeaux, G.; Steven, M.; Baret, F.
\newblock Optimization of Soil-Adjusted Vegetation Indices.
\newblock {\em Remote Sens. Environ.} {\bf 1996}, {\em 55},~95--107, doi:10.1016/0034-4257(95)00186-7.

\bibitem[Chappelle \em{et~al.}(1992)Chappelle, Kim, and
  McMurtrey]{chappelle1992}
Chappelle, E.W.; Kim, M.S.; McMurtrey, J.E.
\newblock Ratio Analysis of Reflectance Spectra ({{RARS}}): An Algorithm for
  the Remote Estimation of the Concentrations of Chlorophyll {{A}}, Chlorophyll
  {{B}}, and Carotenoids in Soybean Leaves.
\newblock {\em Remote Sens. Environ.} {\bf 1992}, {\em 39},~239--247, doi:10.1016/0034-4257(92)90089-3.

\bibitem[Gamon \em{et~al.}(1992)Gamon, Pe{\~n}uelas, and Field]{gamon1992}
Gamon, J.; Pe{\~n}uelas, J.; Field, C.
\newblock A Narrow-Waveband Spectral Index That Tracks Diurnal Changes in
  Photosynthetic Efficiency.
\newblock {\em Remote Sens. Environ.} {\bf 1992}, {\em 41},~35--44, doi:10.1016/0034-4257(92)90059-s.

\bibitem[{Zarco-Tejada} \em{et~al.}(2013){Zarco-Tejada}, {Gonz{\'a}lez-Dugo},
  Williams, Su{\'a}rez, Berni, Goldhamer, and Fereres]{zarco-tejada2013}
{Zarco-Tejada}, P.J.; {Gonz{\'a}lez-Dugo}, V.; Williams, L.E.; Su{\'a}rez, L.;
  Berni, J.A.J.; Goldhamer, D.; Fereres, E.
\newblock A {{PRI}}-Based Water Stress Index Combining Structural and
  Chlorophyll Effects: Assessment Using Diurnal Narrow-Band Airborne Imagery
  and the {{CWSI}} Thermal Index.
\newblock {\em Remote Sens. Environ.} {\bf 2013}, {\em 138},~38--50, doi:10.1016/j.rse.2013.07.024.

\bibitem[Garrity \em{et~al.}(2011)Garrity, Eitel, and Vierling]{garrity2011}
Garrity, S.R.; Eitel, J.U.; Vierling, L.A.
\newblock Disentangling the Relationships between Plant Pigments and the
  Photochemical Reflectance Index Reveals a New Approach for Remote Estimation
  of Carotenoid Content.
\newblock {\em Remote Sens. Environ.} {\bf 2011}, {\em 115},~628--635, doi:10.1016/j.rse.2010.10.007.

\bibitem[Merzlyak \em{et~al.}(1999)Merzlyak, Gitelson, Chivkunova, and
  Rakitin]{merzlyak1999}
Merzlyak, M.N.; Gitelson, A.A.; Chivkunova, O.B.; Rakitin, V.Y.
\newblock Non-Destructive Optical Detection of Pigment Changes during Leaf
  Senescence and Fruit Ripening.
\newblock {\em Physiol. Plant.} {\bf 1999}, {\em 106},~135--141, doi:10.1034/j.1399-3054.1999.106119.x.

\bibitem[Blackburn(1998)]{blackburn1998}
Blackburn, G.A.
\newblock Quantifying Chlorophylls and Caroteniods at Leaf and Canopy Scales.
\newblock {\em Remote Sens. Environ.} {\bf 1998}, {\em 66},~273--285, doi:10.1016/s0034-4257(98)00059-5.

\bibitem[Penuelas \em{et~al.}(1997)Penuelas, Pinol, Ogaya, and
  Filella]{penuelas1997}
Penuelas, J.; Pinol, J.; Ogaya, R.; Filella, I.
\newblock Estimation of Plant Water Concentration by the Reflectance {{Water
  Index WI}} ({{R900}}/{{R970}}).
\newblock {\em Int. J. Remote Sens.} {\bf 1997}, {\em
  18},~2869--2875, doi:10.1080/014311697217396.

\bibitem[Roujean and Breon(1995)]{roujean1995}
Roujean, J.L.; Breon, F.M.
\newblock Estimating {{PAR}} Absorbed by Vegetation from Bidirectional
  Reflectance Measurements.
\newblock {\em Remote Sens. Environ.} {\bf 1995}, {\em 51},~375--384, doi:10.1016/0034-4257(94)00114-3.

\bibitem[Cho and Skidmore(2006)]{cho2006}
Cho, M.A.; Skidmore, A.K.
\newblock A New Technique for Extracting the Red Edge Position from
  Hyperspectral Data: The Linear Extrapolation Method.
\newblock {\em Remote Sens. Environ.} {\bf 2006}, {\em 101},~181--193, doi:10.1016/j.rse.2005.12.011.

\bibitem[Guyot and Baret(1988)]{guyot1988}
Guyot, G.; Baret, F.
\newblock Utilisation de La Haute Resolution Spectrale Pour Suivre l'etat Des
  Couverts Vegetaux.
\newblock  In \emph{Spectral Signatures of Objects in Remote Sensing}; Guyenne, T.D.,
  Hunt, J.J., Eds.;  {ESA Special Publication}: 1988; Aussois, France, Volume 287, p. 279.

\bibitem[Huete(1988)]{huete1988}
Huete, A.
\newblock A Soil-Adjusted Vegetation Index ({{SAVI}}).
\newblock {\em Remote Sens. Environ.} {\bf 1988}, {\em 25},~295--309, doi:10.1016/0034-4257(88)90106-x.

\bibitem[Penuelas \em{et~al.}(1995)Penuelas, Filella, Lloret, Munoz, and
  Vilaleliu]{penuelasj.1995}
Penuelas, J.; Filella, I.; Lloret, P.; Munoz, F.; Vilaleliu, M.
\newblock Reflectance Assessment of Mite Effects on Apple Trees.
\newblock {\em Int. J. Remote Sens.} {\bf 1995}, {\em
  16},~2727--2733, doi:10.1080/01431169508954588.

\bibitem[Vincini \em{et~al.}(2006)Vincini, Frazzi, and D'Alessio]{vincini2006}
Vincini, M.; Frazzi, E.; D'Alessio, P.
\newblock Angular Dependence of Maize and Sugar Beet {{VIs}} from Directional
  {{CHRIS}}/{{Proba}} Data.
\newblock  In Proceedings of the 4th {{ESA CHRIS PROBA}} Workshop,  Frascati, Italy, 19--21 September 2006; Volume 2006, pp.
  19--21.

\bibitem[Jordan(1969)]{jordan1969}
Jordan, C.F.
\newblock Derivation of Leaf-Area Index from Quality of Light on the Forest
  Floor.
\newblock {\em Ecology} {\bf 1969}, {\em 50},~663--666, doi:10.2307/1936256.

\bibitem[Gitelson and Merzlyak(1997)]{gitelson1997}
Gitelson, A.A.; Merzlyak, M.N.
\newblock Remote Estimation of Chlorophyll Content in Higher Plant Leaves.
\newblock {\em Int. J. Remote Sens.} {\bf 1997}, {\em
  18},~2691--2697, doi:10.1080/014311697217558.

\bibitem[McMurtrey \em{et~al.}(1994)McMurtrey, Chappelle, Kim, Meisinger, and
  Corp]{mcmurtrey1994}
McMurtrey, J.; Chappelle, E.; Kim, M.; Meisinger, J.; Corp, L.
\newblock Distinguishing Nitrogen Fertilization Levels in Field Corn (\emph{{{Zea}}
  mays} {{L}}.) with Actively Induced Fluorescence and Passive Reflectance
  Measurements.
\newblock {\em Remote Sens. Environ.} {\bf 1994}, {\em 47},~36--44, doi:10.1016/0034-4257(94)90125-2.

\bibitem[{Zarco-Tejada} and Miller(1999)]{zarco-tejada1999}
{Zarco-Tejada}, P.J.; Miller, J.R.
\newblock Land Cover Mapping at {{BOREAS}} Using Red Edge Spectral Parameters
  from {{CASI}} Imagery.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 1999}, {\em
  104},~27921--27933, doi:10.1029/1999jd900161.

\bibitem[Lichtenthaler \em{et~al.}(1996)Lichtenthaler, Lang, Sowinska, Heisel,
  and Mieh{\'e}]{lichtenthaler1996}
Lichtenthaler, H.; Lang, M.; Sowinska, M.; Heisel, F.; Mieh{\'e}, J.
\newblock Detection of Vegetation Stress via a New High Resolution Fluorescence
  Imaging System.
\newblock {\em J. Plant Physiol.} {\bf 1996}, {\em 148},~599--612, doi:10.1016/s0176-1617(96)80081-2.

\bibitem[{Hern{\'a}ndez-Clemente} \em{et~al.}(2012){Hern{\'a}ndez-Clemente},
  {Navarro-Cerrillo}, and {Zarco-Tejada}]{hernandez-clemente2012}
{Hern{\'a}ndez-Clemente}, R.; {Navarro-Cerrillo}, R.M.; {Zarco-Tejada}, P.J.
\newblock Carotenoid Content Estimation in a Heterogeneous Conifer Forest Using
  Narrow-Band Indices and {{PROSPECTDART}} Simulations.
\newblock {\em Remote Sens. Environ.} {\bf 2012}, {\em 127},~298--315, doi:10.1016/j.rse.2012.09.014.

\bibitem[Elvidge and Chen(1995)]{elvidge1995}
Elvidge, C.D.; Chen, Z.
\newblock Comparison of Broad-Band and Narrow-Band Red and near-Infrared
  Vegetation Indices.
\newblock {\em Remote Sens. Environ.} {\bf 1995}, {\em 54},~38--48, doi:10.1016/0034-4257(95)00132-k.

\bibitem[Filella and Penuelas(1994)]{filella1994}
Filella, I.; Penuelas, J.
\newblock The Red Edge Position and Shape as Indicators of Plant Chlorophyll
  Content, Biomass and Hydric Status.
\newblock {\em Int. J. Remote Sens.} {\bf 1994}, {\em
  15},~1459--1470, doi:10.1080/01431169408954177.

\bibitem[Levin \em{et~al.}(2007)Levin, Kidron, and {Ben-Dor}]{levin2007}
Levin, N.; Kidron, G.J.; {Ben-Dor}, E.
\newblock Surface Properties of Stabilizing Coastal Dunes: Combining Spectral
  and Field Analyses.
\newblock {\em Sedimentology} {\bf 2007}, {\em 54},~771--788, doi:10.1111/j.1365-3091.2007.00859.x.

\bibitem[Lobell \em{et~al.}(2001)Lobell, Asner, Law, and Treuhaft]{lobell2001}
Lobell, D.B.; Asner, G.P.; Law, B.E.; Treuhaft, R.N.
\newblock Subpixel Canopy Cover Estimation of Coniferous Forests in {{Oregon}}
  Using {{SWIR}} Imaging Spectrometry.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 2001}, {\em
  106},~5151--5160, doi:10.1029/2000jd900739.

\bibitem[Hunt \em{et~al.}(2013)Hunt, Doraiswamy, McMurtrey, Daughtry, Perry,
  and Akhmedov]{hunt2013}
Hunt, E.R.; Doraiswamy, P.C.; McMurtrey, J.E.; Daughtry, C.S.; Perry, E.M.;
  Akhmedov, B.
\newblock A Visible Band Index for Remote Sensing Leaf Chlorophyll Content at
  the Canopy Scale.
\newblock {\em Int. J. Appl. Earth Obs. Geoinf.} {\bf 2013}, {\em 21},~103--112, doi:10.1016/j.jag.2012.07.020.

\bibitem[Broge and Leblanc(2001)]{broge2001}
Broge, N.; Leblanc, E.
\newblock Comparing Prediction Power and Stability of Broadband and
  Hyperspectral Vegetation Indices for Estimation of Green Leaf Area Index and
  Canopy Chlorophyll Density.
\newblock {\em Remote Sens. Environ.} {\bf 2001}, {\em 76},~156--172, doi:10.1016/s0034-4257(00)00197-8.

\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{pschratz.jpeg}}}
%{\textbf{Patrick Schratz} is a Data Scientist working in Zurich, Switzerland.
%	His area of expertise is applied machine learning, more specifically the field of environmental modeling.
%	He is a Ph.D. candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.
%	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatiotemporal data handling.}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{jmuenchow.jpg}}}
%{\textbf{Jannes Muenchow} is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
%	He received his Ph.D. from Friedrich-Alexander-Universität Erlangen-Nürnberg (Germany) in 2013.
%	He joined the business location department of a large consulting company as a geo-data scientist for more than two years until the prediction of a strong El Niño event brought him back to academia from 2016 to 2019 (Friedrich Schiller University Jena).
%	His research focuses on developing open-source tools for ecology, geomorphology and qualitative GIS.
%	He is a co-author of the book ``Geocomputation with R''.}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{eiturritxa.jpg}}}
%{\textbf{Eugenia Iturritxa} received the Ecology and Ph.D. degree in Plant Protection from the University of the Basque Country in 2001.
%	Since 1999 her main research has focused on forest health, on the study of diverse species of native and introduced pathogenic fungi in forests and forest plantations.
%	Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{jcortes.png}}}
%{\textbf{José Cortés} received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
%	Currently he is a Ph.D. student at Friedrich Schiller University Jena, Germany.
%	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
%	His research focuses on spatiotemporal trend detection in environmental data.}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{bbischl.jpg}}}
%{\textbf{Bernd Bischl} obtained his Ph.D. from TU Dortmund University in 2013.
%	He is a professor of Statistical Learning and Data Science in the Department of Statistics at Ludwig-Maximilians-Universität in Munich, Germany, and a director of the Munich Center for Machine Learning.
%	His research interests include AutoML, model selection, interpretable ML and XAI.}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{abrenning.jpg}}}
%{\textbf{Alexander Brenning} (Ph.D.~2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu~Berlin.
%	He served as an assistant professor and tenured associate professor in geomatics at the University of Waterloo, Ontario, Canada from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
%	Dr.~Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.
%	He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/proofreading-before-pub/Definitions/journalnames.tex
add assets of all submissions

--- 2021-11-12 docs/00-manuscripts/mdpi/Definitions/journalnames.tex
@@
-%% Updated on 18 June 2021
+%%Updated on 12 October 2021
 %,% Website Name,%,%,% Website Name,%%,% Full Name,%%,% Short Name,%%,% DOI abbr,%%,% ISSN,% \continuouspages
@@
 \DeclareOption{allergies}{\gdef\@journal{allergies}\gdef\@journalfull{Allergies}\gdef\@journalshort{Allergies}\gdef\@doiabbr{allergies}\gdef\@ISSN{2313-5786}}
+\DeclareOption{alloys}{\gdef\@journal{alloys}\gdef\@journalfull{Alloys}\gdef\@journalshort{Alloys}\gdef\@doiabbr{alloys}\gdef\@ISSN{2674-063X}\gdef\@continuouspages{yes}}
 \DeclareOption{analytica}{\gdef\@journal{analytica}\gdef\@journalfull{Analytica}\gdef\@journalshort{Analytica}\gdef\@doiabbr{analytica}\gdef\@ISSN{2673-4532}\gdef\@continuouspages{yes}}
@@
 \DeclareOption{antioxidants}{\gdef\@journal{antioxidants}\gdef\@journalfull{Antioxidants}\gdef\@journalshort{Antioxidants}\gdef\@doiabbr{antiox}\gdef\@ISSN{2076-3921}}
+\DeclareOption{applbiosci}{\gdef\@journal{applbiosci}\gdef\@journalfull{Applied Biosciences}\gdef\@journalshort{Appl. Biosci.}\gdef\@doiabbr{applbiosci}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{appliedchem}{\gdef\@journal{appliedchem}\gdef\@journalfull{AppliedChem}\gdef\@journalshort{AppliedChem}\gdef\@doiabbr{appliedchem}\gdef\@ISSN{2673-9623}\gdef\@continuouspages{yes}}
@@
 \DeclareOption{asi}{\gdef\@journal{asi}\gdef\@journalfull{Applied System Innovation}\gdef\@journalshort{Appl. Syst. Innov.}\gdef\@doiabbr{asi}\gdef\@ISSN{2571-5577}}
+\DeclareOption{astronomy}{\gdef\@journal{astronomy}\gdef\@journalfull{Astronomy}\gdef\@journalshort{Astronomy}\gdef\@doiabbr{astronomy}\gdef\@ISSN{2674-0346}\gdef\@continuouspages{yes}}
 \DeclareOption{atmosphere}{\gdef\@journal{atmosphere}\gdef\@journalfull{Atmosphere}\gdef\@journalshort{Atmosphere}\gdef\@doiabbr{atmos}\gdef\@ISSN{2073-4433}}
@@
 \DeclareOption{bloods}{\gdef\@journal{bloods}\gdef\@journalfull{Bloods}\gdef\@journalshort{Bloods}\gdef\@doiabbr{bloods}\gdef\@ISSN{2673-415X}}
-\DeclareOption{blsf}{\gdef\@journal{	blsf}\gdef\@journalfull{Biology and Life Sciences Forum}\gdef\@journalshort{Biol. Life Sci. Forum}\gdef\@doiabbr{blsf	}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{blsf}{\gdef\@journal{blsf}\gdef\@journalfull{Biology and Life Sciences Forum}\gdef\@journalshort{Biol. Life Sci. Forum}\gdef\@doiabbr{blsf	}\gdef\@ISSN{2673-9976}}
 \DeclareOption{brainsci}{\gdef\@journal{brainsci}\gdef\@journalfull{Brain Sciences}\gdef\@journalshort{Brain Sci.}\gdef\@doiabbr{brainsci}\gdef\@ISSN{2076-3425}}
@@
 \DeclareOption{children}{\gdef\@journal{children}\gdef\@journalfull{Children}\gdef\@journalshort{Children}\gdef\@doiabbr{children}\gdef\@ISSN{2227-9067}}
+\DeclareOption{chips}{\gdef\@journal{chips}\gdef\@journalfull{Chips}\gdef\@journalshort{Chips}\gdef\@doiabbr{chips}\gdef\@ISSN{2674-0710}\gdef\@continuouspages{yes}}
 \DeclareOption{cimb}{\gdef\@journal{cimb}\gdef\@journalfull{Current Issues in Molecular Biology}\gdef\@journalshort{Curr.
@@
 \DeclareOption{colloids}{\gdef\@journal{colloids}\gdef\@journalfull{Colloids Interfaces}\gdef\@journalshort{Colloids Interfaces}\gdef\@doiabbr{colloids}\gdef\@ISSN{2504-5377}}
-\DeclareOption{colorant}{\gdef\@journal{colorant}\gdef\@journalfull{Colorant}\gdef\@journalshort{Colorant}\gdef\@doiabbr{colorant}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{colorants}{\gdef\@journal{colorants}\gdef\@journalfull{Colorants}\gdef\@journalshort{Colorants}\gdef\@doiabbr{colorants}\gdef\@ISSN{2079-6447}\gdef\@continuouspages{yes}}
+\DeclareOption{commodities}{\gdef\@journal{commodities}\gdef\@journalfull{Commodities}\gdef\@journalshort{Commodities}\gdef\@doiabbr{commodities}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{compounds}{\gdef\@journal{compounds}\gdef\@journalfull{Compounds}\gdef\@journalshort{Compounds}\gdef\@doiabbr{compounds}\gdef\@ISSN{2673-6918}\gdef\@continuouspages{yes}}
@@
 \DeclareOption{crystals}{\gdef\@journal{crystals}\gdef\@journalfull{Crystals}\gdef\@journalshort{Crystals}\gdef\@doiabbr{cryst}\gdef\@ISSN{2073-4352}}
-\DeclareOption{ctn}{\gdef\@journal{ctn}\gdef\@journalfull{Clinical and Translational Neuroscience}\gdef\@journalshort{CTN}\gdef\@doiabbr{ctn}\gdef\@ISSN{2514-183X}\gdef\@continuouspages{yes}}
+\DeclareOption{ctn}{\gdef\@journal{ctn}\gdef\@journalfull{Clinical and Translational Neuroscience}\gdef\@journalshort{CTN}\gdef\@doiabbr{Clin. Transl. Neurosci.}\gdef\@ISSN{2514-183X}}
 \DeclareOption{curroncol}{\gdef\@journal{curroncol}\gdef\@journalfull{Current Oncology}\gdef\@journalshort{Curr. Oncol.}\gdef\@doiabbr{curroncol}\gdef\@ISSN{1718-7729}\gdef\@continuouspages{yes}}
@@
 \DeclareOption{diagnostics}{\gdef\@journal{diagnostics}\gdef\@journalfull{Diagnostics}\gdef\@journalshort{Diagnostics}\gdef\@doiabbr{diagnostics}\gdef\@ISSN{2075-4418}}
+\DeclareOption{dietetics}{\gdef\@journal{dietetics}\gdef\@journalfull{Dietetics}\gdef\@journalshort{Dietetics}\gdef\@doiabbr{dietetics}\gdef\@ISSN{2674-0311}\gdef\@continuouspages{yes}}
 \DeclareOption{digital}{\gdef\@journal{digital}\gdef\@journalfull{Digital}\gdef\@journalshort{Digital}\gdef\@doiabbr{digital}\gdef\@ISSN{2673-6470}\gdef\@continuouspages{yes}}
@@
 %\DeclareOption{ent}{\gdef\@journal{ent}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{entomology}{\gdef\@journal{entomology}\gdef\@journalfull{Entomology}\gdef\@journalshort{Entomology}\gdef\@doiabbr{entomology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}	
 \DeclareOption{entropy}{\gdef\@journal{entropy}\gdef\@journalfull{Entropy}\gdef\@journalshort{Entropy}\gdef\@doiabbr{e}\gdef\@ISSN{1099-4300}}
@@
 \DeclareOption{fibers}{\gdef\@journal{fibers}\gdef\@journalfull{Fibers}\gdef\@journalshort{Fibers}\gdef\@doiabbr{fib}\gdef\@ISSN{2079-6439}}
-\DeclareOption{	fintech}{\gdef\@journal{fintech}\gdef\@journalfull{FinTech}\gdef\@journalshort{FinTech}\gdef\@doiabbr{fintech}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{fintech}{\gdef\@journal{fintech}\gdef\@journalfull{FinTech}\gdef\@journalshort{FinTech}\gdef\@doiabbr{fintech}\gdef\@ISSN{2674-1032}\gdef\@continuouspages{yes}}
 \DeclareOption{fire}{\gdef\@journal{fire}\gdef\@journalfull{Fire}\gdef\@journalshort{Fire}\gdef\@doiabbr{fire}\gdef\@ISSN{2571-6255}}
@@
 \DeclareOption{livers}{\gdef\@journal{livers}\gdef\@journalfull{Livers}\gdef\@journalshort{Livers}\gdef\@doiabbr{livers}\gdef\@ISSN{2673-4389}\gdef\@continuouspages{yes}}
+\DeclareOption{logics}{\gdef\@journal{logics}\gdef\@journalfull{Logics}\gdef\@journalshort{Logics}\gdef\@doiabbr{logics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{logistics}{\gdef\@journal{logistics}\gdef\@journalfull{Logistics}\gdef\@journalshort{Logistics}\gdef\@doiabbr{logistics}\gdef\@ISSN{2305-6290}}
 \DeclareOption{lubricants}{\gdef\@journal{lubricants}\gdef\@journalfull{Lubricants}\gdef\@journalshort{Lubricants}\gdef\@doiabbr{lubricants}\gdef\@ISSN{2075-4442}}
+\DeclareOption{lymphatics}{\gdef\@journal{lymphatics}\gdef\@journalfull{Lymphatics}\gdef\@journalshort{Lymphatics}\gdef\@doiabbr{lymphatics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}	
 \DeclareOption{machines}{\gdef\@journal{machines}\gdef\@journalfull{Machines}\gdef\@journalshort{Machines}\gdef\@doiabbr{machines}\gdef\@ISSN{2075-1702}}
@@
 \DeclareOption{mca}{\gdef\@journal{mca}\gdef\@journalfull{Mathematical and Computational Applications}\gdef\@journalshort{Math. Comput. Appl.}\gdef\@doiabbr{mca}\gdef\@ISSN{2297-8747}}
-%\DeclareOption{measurements}{\gdef\@journal{measurements}\gdef\@journalfull{Measurements}\gdef\@journalshort{Measurements}\gdef\@doiabbr{measurements}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{measurements}{\gdef\@journal{measurements}\gdef\@journalfull{Measurements}\gdef\@journalshort{Measurements}\gdef\@doiabbr{measurements}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{medicina}{\gdef\@journal{medicina}\gdef\@journalfull{Medicina}\gdef\@journalshort{Medicina}\gdef\@doiabbr{medicina}\gdef\@ISSN{1648-9144}}
@@
 \DeclareOption{metals}{\gdef\@journal{metals}\gdef\@journalfull{Metals}\gdef\@journalshort{Metals}\gdef\@doiabbr{met}\gdef\@ISSN{2075-4701}}
+\DeclareOption{meteorology}{\gdef\@journal{meteorology}\gdef\@journalfull{Meteorology}\gdef\@journalshort{Meteorology}\gdef\@doiabbr{meteorology}\gdef\@ISSN{2674-0494}\gdef\@continuouspages{yes}}
+\DeclareOption{methane}{\gdef\@journal{methane}\gdef\@journalfull{Methane}\gdef\@journalshort{Methane}\gdef\@doiabbr{methane}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{metrology}{\gdef\@journal{metrology}\gdef\@journalfull{Metrology}\gdef\@journalshort{Metrology}\gdef\@doiabbr{metrology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
@@
 \DeclareOption{mps}{\gdef\@journal{mps}\gdef\@journalfull{Methods and Protocols}\gdef\@journalshort{Methods Protoc.}\gdef\@doiabbr{mps}\gdef\@ISSN{2409-9279}}
-\DeclareOption{msf}{\gdef\@journal{	msf}\gdef\@journalfull{Medical Sciences Forum}\gdef\@journalshort{Med. Sci. Forum}\gdef\@doiabbr{msf}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{msf}{\gdef\@journal{msf}\gdef\@journalfull{Medical Sciences Forum}\gdef\@journalshort{Med. Sci. Forum}\gdef\@doiabbr{msf}\gdef\@ISSN{2673-9992}}
 \DeclareOption{mti}{\gdef\@journal{mti}\gdef\@journalfull{Multimodal Technologies and Interaction}\gdef\@journalshort{Multimodal Technol. Interact.}\gdef\@doiabbr{mti}\gdef\@ISSN{2414-4088}}
@@
 \DeclareOption{organics}{\gdef\@journal{organics}\gdef\@journalfull{Organics}\gdef\@journalshort{Organics}\gdef\@doiabbr{org}\gdef\@ISSN{2673-401X}\gdef\@continuouspages{yes}}
+\DeclareOption{organoids}{\gdef\@journal{organoids}\gdef\@journalfull{Organoids}\gdef\@journalshort{Organoids}\gdef\@doiabbr{organoids}\gdef\@ISSN{2674-1172}\gdef\@continuouspages{yes}}
 \DeclareOption{osteology}{\gdef\@journal{osteology}\gdef\@journalfull{Osteology}\gdef\@journalshort{Osteology}\gdef\@doiabbr{osteology}\gdef\@ISSN{2673-4036}\gdef\@continuouspages{yes}}
@@
 \DeclareOption{polysaccharides}{\gdef\@journal{polysaccharides}\gdef\@journalfull{Polysaccharides}\gdef\@journalshort{Polysaccharides}\gdef\@doiabbr{polysaccharides}\gdef\@ISSN{2673-4176}\gdef\@continuouspages{yes}}
-\DeclareOption{poultry}{\gdef\@journal{poultry}\gdef\@journalfull{Poultry}\gdef\@journalshort{Poultry}\gdef\@doiabbr{poultry}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{poultry}{\gdef\@journal{poultry}\gdef\@journalfull{Poultry}\gdef\@journalshort{Poultry}\gdef\@doiabbr{poultry}\gdef\@ISSN{2674-1164}\gdef\@continuouspages{yes}}
+\DeclareOption{powders}{\gdef\@journal{powders}\gdef\@journalfull{Powders}\gdef\@journalshort{Powders}\gdef\@doiabbr{powders}\gdef\@ISSN{2674-0516}\gdef\@continuouspages{yes}}
 \DeclareOption{preprints}{\gdef\@journal{preprints}\gdef\@journalfull{Preprints}\gdef\@journalshort{Preprints}\gdef\@doiabbr{}\gdef\@ISSN{}}
@@
 \DeclareOption{proteomes}{\gdef\@journal{proteomes}\gdef\@journalfull{Proteomes}\gdef\@journalshort{Proteomes}\gdef\@doiabbr{proteomes}\gdef\@ISSN{2227-7382}}
-\DeclareOption{psf}{\gdef\@journal{psf}\gdef\@journalfull{Physical Sciences Forum}\gdef\@journalshort{Phys. Sci. Forum}\gdef\@doiabbr{psf}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{psf}{\gdef\@journal{psf}\gdef\@journalfull{Physical Sciences Forum}\gdef\@journalshort{Phys. Sci. Forum}\gdef\@doiabbr{psf}\gdef\@ISSN{2673-9984}}
 \DeclareOption{psych}{\gdef\@journal{psych}\gdef\@journalfull{Psych}\gdef\@journalshort{Psych}\gdef\@doiabbr{psych}\gdef\@ISSN{2624-8611}\gdef\@continuouspages{yes}}
 \DeclareOption{psychiatryint}{\gdef\@journal{psychiatryint}\gdef\@journalfull{Psychiatry International}\gdef\@journalshort{Psychiatry Int.}\gdef\@doiabbr{psychiatryint}\gdef\@ISSN{2673-5318}\gdef\@continuouspages{yes}}
+\DeclareOption{psychoactives}{\gdef\@journal{psychoactives}\gdef\@journalfull{Psychoactives}\gdef\@journalshort{Psychoactives}\gdef\@doiabbr{psychoactives}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{publications}{\gdef\@journal{publications}\gdef\@journalfull{Publications}\gdef\@journalshort{Publications}\gdef\@doiabbr{publications}\gdef\@ISSN{2304-6775}}
@@
 \DeclareOption{resources}{\gdef\@journal{resources}\gdef\@journalfull{Resources}\gdef\@journalshort{Resources}\gdef\@doiabbr{resources}\gdef\@ISSN{2079-9276}}
-\DeclareOption{rheumato}{\gdef\@journal{rheumato}\gdef\@journalfull{Rheumato}\gdef\@journalshort{Rheumato}\gdef\@doiabbr{rheumato}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{rheumato}{\gdef\@journal{rheumato}\gdef\@journalfull{Rheumato}\gdef\@journalshort{Rheumato}\gdef\@doiabbr{rheumato}\gdef\@ISSN{2674-0621}\gdef\@continuouspages{yes}}
 \DeclareOption{risks}{\gdef\@journal{risks}\gdef\@journalfull{Risks}\gdef\@journalshort{Risks}\gdef\@doiabbr{risks}\gdef\@ISSN{2227-9091}}
@@
 \DeclareOption{scipharm}{\gdef\@journal{scipharm}\gdef\@journalfull{Scientia Pharmaceutica}\gdef\@journalshort{Sci. Pharm.}\gdef\@doiabbr{scipharm}\gdef\@ISSN{2218-0532}}
+\DeclareOption{seeds}{\gdef\@journal{seeds}\gdef\@journalfull{Seeds}\gdef\@journalshort{Seeds}\gdef\@doiabbr{seeds}\gdef\@ISSN{2674-1024}\gdef\@continuouspages{yes}}
 \DeclareOption{sensors}{\gdef\@journal{sensors}\gdef\@journalfull{Sensors}\gdef\@journalshort{Sensors}\gdef\@doiabbr{s}\gdef\@ISSN{1424-8220}}
@@
 \DeclareOption{socsci}{\gdef\@journal{socsci}\gdef\@journalfull{Social Sciences}\gdef\@journalshort{Soc. Sci.}\gdef\@doiabbr{socsci}\gdef\@ISSN{2076-0760}}
+\DeclareOption{software}{\gdef\@journal{software}\gdef\@journalfull{Software}\gdef\@journalshort{Software}\gdef\@doiabbr{software}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{soilsystems}{\gdef\@journal{soilsystems}\gdef\@journalfull{Soil Systems}\gdef\@journalshort{Soil Syst.}\gdef\@doiabbr{soilsystems}\gdef\@ISSN{2571-8789}}
@@
 \DeclareOption{symmetry}{\gdef\@journal{symmetry}\gdef\@journalfull{Symmetry}\gdef\@journalshort{Symmetry}\gdef\@doiabbr{sym}\gdef\@ISSN{2073-8994}}
+\DeclareOption{synbio}{\gdef\@journal{synbio}\gdef\@journalfull{Synbio}\gdef\@journalshort{Synbio}\gdef\@doiabbr{synbio}\gdef\@ISSN{2674-0583}\gdef\@continuouspages{yes}}
 \DeclareOption{systems}{\gdef\@journal{systems}\gdef\@journalfull{Systems}\gdef\@journalshort{Systems}\gdef\@doiabbr{systems}\gdef\@ISSN{2079-8954}}
@@
 \DeclareOption{vehicles}{\gdef\@journal{vehicles}\gdef\@journalfull{Vehicles}\gdef\@journalshort{Vehicles}\gdef\@doiabbr{vehicles}\gdef\@ISSN{2624-8921}\gdef\@continuouspages{yes}}
-\DeclareOption{venereology}{\gdef\@journal{venereology}\gdef\@journalfull{Venereology}\gdef\@journalshort{Venereology}\gdef\@doiabbr{venereology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{venereology}{\gdef\@journal{venereology}\gdef\@journalfull{Venereology}\gdef\@journalshort{Venereology}\gdef\@doiabbr{venereology}\gdef\@ISSN{2674-0710}\gdef\@continuouspages{yes}}
 \DeclareOption{vetsci}{\gdef\@journal{vetsci}\gdef\@journalfull{Veterinary Sciences}\gdef\@journalshort{Vet. Sci.}\gdef\@doiabbr{vetsci}\gdef\@ISSN{2306-7381}}
@@
 \DeclareOption{vision}{\gdef\@journal{vision}\gdef\@journalfull{Vision}\gdef\@journalshort{Vision}\gdef\@doiabbr{vision}\gdef\@ISSN{2411-5150}}
+\DeclareOption{waste}{\gdef\@journal{waste}\gdef\@journalfull{Waste}\gdef\@journalshort{Waste}\gdef\@doiabbr{waste}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
 \DeclareOption{water}{\gdef\@journal{water}\gdef\@journalfull{Water}\gdef\@journalshort{Water}\gdef\@doiabbr{w}\gdef\@ISSN{2073-4441}}
@@
 \DeclareOption{wevj}{\gdef\@journal{wevj}\gdef\@journalfull{World Electric Vehicle Journal}\gdef\@journalshort{World Electr. Veh. J.}\gdef\@doiabbr{wevj}\gdef\@ISSN{2032-6653}}
-\DeclareOption{wind}{\gdef\@journal{wind}\gdef\@journalfull{Wind}\gdef\@journalshort{Wind}\gdef\@doiabbr{wind}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
+\DeclareOption{wind}{\gdef\@journal{wind}\gdef\@journalfull{Wind}\gdef\@journalshort{Wind}\gdef\@doiabbr{wind}\gdef\@ISSN{2674-032X}\gdef\@continuouspages{yes}}
 \DeclareOption{women}{\gdef\@journal{women}\gdef\@journalfull{Women}\gdef\@journalshort{Women}\gdef\@doiabbr{women}\gdef\@ISSN{2673-4184}\gdef\@continuouspages{yes}}

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/proofreading-before-pub/remotesensing-1398065 english editing done.tex
add assets of all submissions

--- 2021-12-17 docs/00-manuscripts/mdpi/submissions/final-submission/remotesensing-1398065-final.tex
@@
 \documentclass[remotesensing,article,accept,moreauthors,pdftex]{Definitions/mdpi}
+
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
 
@@
 \newcommand{\orcidauthorB}{0000-0001-7834-4717} % Add \orcidB{} behind the author's name
-\newcommand{\orcidauthorC}{0000-0003-2567-8689} % Add \orcidC{} behind the author's name
-\newcommand{\orcidauthorD}{0000-0002-6390-5873} % Add \orcidD{} behind the author's name
-\newcommand{\orcidauthorE}{0000-0001-6002-6980} % Add \orcidE{} behind the author's name
-\newcommand{\orcidauthorF}{0000-0001-6640-679X} % Add \orcidF{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0003-2567-8689} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0002-6390-5873} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorE}{0000-0001-6002-6980} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorF}{0000-0001-6640-679X} % Add \orcidB{} behind the author's name
 
 % Authors, for the paper (add full first names)
-\Author{Patrick Schratz $^{1,}$*\orcidA{}, Jannes Muenchow $^{1}$\orcidB{}, Eugenia Iturritxa $^{2}$\orcidD{}, José Cortés $^{1}$\orcidC{}, Bernd Bischl $^{3}$\orcidE{} and~Alexander~Brenning $^{1}$\orcidF{}}
+\Author{\hl{Patrick Schratz} %MDPI: Please carefully check the accuracy of names and affiliations.
+ $^{1,}$*\orcidA{}, \hl{Jannes Muenchow} %MDPI: This author is different from redmine, please confirm.
+ $^{1}$\orcidB{}, Eugenia Iturritxa $^{2}$\orcidD{}, José Cortés $^{1}$\orcidC{}, Bernd Bischl $^{3}$\orcidE{} and~Alexander~Brenning $^{1}$\orcidF{}}
 
@@
 \address{%
-$^{1}$ \quad GIScience Group, Department of Geography, Friedrich Schiller University Jena, Germany;
+$^{1}$ \quad \hl{GIScience Group,} %MDPI: Please check if the addresses are arranged from small to large. The same to aff. 3.
+ Department of Geography, Friedrich Schiller University Jena, \hl{Germany;} %MDPI: please add city and post code. The same to affs. 2 and 3.
  malnamalja@gmx.de~(J.M.); jose.cortes@uni-jena.de~(J.C.); alexander.brenning@uni-jena.de~(A.B.)\\
-$^{2}$ \quad NEIKER Tecnalia, Spain; eiturritxa@neiker.eus\\
-$^{3}$ \quad Department of Statistics, Ludwig-Maximilians-Universität München, Germany; bernd.bischl@stat.uni-muenchen.de
+$^{2}$ \quad \hl{NEIKER} %MDPI: please check if this address is correct.
+ Tecnalia, \hl{Spain;} eiturritxa@neiker.eus\\
+$^{3}$ \quad \hl{Computational Statistics Group,} Department of Statistics, Ludwig-Maximilians-Universität München, \hl{Germany;} bernd.bischl@stat.uni-muenchen.de
 }
@@
 The magnitude of human observer errors in such surveys, including the present one, is not precisely known and has being discussed for many years~\cite{innes1993}.
-Ref.~\cite{maclean1982} estimated human observer errors in defoliation surveys to range between 7\% and 18\%.
+\hl{Ref.}%MDPI: Reference citation can't be in the begining of sentence, we add Ref., please confirm.
+~\cite{maclean1982} estimated human observer errors in defoliation surveys to range between 7\% and 18\%.
 
@@
 The selected hyperparameter settings are shown in Appendix~\ref{appendixA} Table~\ref{tab:hyperparameter_limits}.
-All code and data are included in the research compendium of this study (\url{https://doi.org/10.5281/zenodo.2635403} (accessed on 27 November 2021).
+All code and data are included in the research compendium of this study (\url{https://doi.org/10.5281/zenodo.2635403}, \hl{accessed on).} %MDPI: Please add access date after the URL; like “http://xxxxxxx (access on 8 January 2021)”, please check all URL in manuscript and add the specific date.
 
@@
 
-The algorithm implementations of the following packages were used: xgboost~\cite{chen2016} (extreme gradient boosting), kernlab~\cite{kernlab} (support vector machine) and glmnet~\cite{glmnet} (penalized regression).
+The algorithm implementations of the following packages were used: xgboost~\cite{chen2016} (\textit{\hl{Extreme Gradient Boosting}}), %MDPI: Is the italic necessary?
+ kernlab~\cite{kernlab} (support vector machine)m and glmnet~\cite{glmnet} (penalized regression).
 The filter implementations of the following packages were used: praznik~\cite{praznik} and FSelectorRcpp~\cite{fselectorrcpp}.
-Package mlr~\cite{mlr} was used for all modeling related steps, and drake~\cite{drake} was used for structuring the work and for reproducibility.
-This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}, (accessed on 27 November 2021).
-Apart from the availability of code and manuscript sources, a static webpage is available at \url{https://pat-s.github.io/2019-feature-selection} (accessed on 27 November 2021) which includes additional side analyses that were carried out during the creation of this study.
+Package mlr~\cite{mlr} was used for all modeling related steps, and
+drake~\cite{drake} was used for structuring the work and for reproducibility.
+This study is available as a research compendium on Zenodo (\url{10.5281/zenodo.2635403}, \hl{accessed on).} %MDPI: Please add access date after the URL; like “http://xxxxxxx (access on 8 January 2021)”, please check all URL in manuscript and add the specific date.
+Apart from the availability of code and manuscript sources, a static webpage is available at (\url{https://pat-s.github.io/2019-feature-selection}, \hl{accessed on),} %MDPI: Please add access date after the URL; like “http://xxxxxxx (access on 8 January 2021)”, please check all URL in manuscript and add the specific date.
+which includes additional side analyses that were carried out during the creation of this study.
 
@@
 SVM showed almost no differences in RMSE across feature sets whereas other learners (RF, SVM, XGBoost, lasso and ridge) differed up to five p.p. (Figure~\ref{fig:perf-result}).
-SVM showed the best overall performance with a mean difference of around three p.p. to the next best model (XGBoost) (Table~\ref{tab:best-learner-perf}).
+SVM showed the best overall performance with a mean difference of around three p.p.
+to the next best model (XGBoost) (\hl{Table}%MDPI: we adjust the order of Tables 3, 4 and Tables 5, 6, please confirm.
+~\ref{tab:best-learner-perf}).
 Performance differences between test folds were large: Predicting on Luiando resulted in an RMSE of 9.0 p.p.
@@
 
-\dataavailability{The data presented in this study are openly available in Zenodo at \url{https://doi.org/10.5281/zenodo.2635403} (accessed on 27 November 2021).}
-
-\acknowledgments{We acknowledge support by the German Research Foundation and the Open Access Publication Fund of the Thueringer Universitaets- und Landesbibliothek Jena Projekt-Nr. 433052568.}
+\institutionalreview{\hl{ } %MDPI: In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals.
+Please note that the Editorial Office might ask you for further information.
+Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals.
+You might also choose to exclude this statement if the study did not involve humans or animals.
+}
+
+\informedconsent{\hl{ } %MDPI: Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
+}
+
+\dataavailability{The data presented in this study are openly available in Zenodo at \url{https://doi.org/10.5281/zenodo.2635403} \hl{(accessed on).} %MDPI: Please add access date after the URL; like “http://xxxxxxx (access on 8 January 2021)”, please check all URL in manuscript and add the specific date.
+}
+
+% \acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
 
@@
 		% \begin{supertabular}{0.3\textwidth}{lll}
-		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference~}
+		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference~\hl{*}} %MDPI: there is no explanation for *.
          \\		\midrule
@@
 		EVI             & $\frac{2.5\,\times\, (R_{800}\,-\,R_{670}) }{ (R_{800}\,-\,(6\,\times\, R_{670})\,-\,(7.5\,\times\, R_{475})\,+\,1)}$           &~\cite{huete1997}             \\
-		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$  &~\cite{wu2014}                 \\
+		GDVI            & $(R_{800}^n-R_{680}^n) / (R_{800}^n+R_{680}^n)$ \hl{**                                                   } %MDPI: there is no explanation for **.
+ &~\cite{wu2014}                 \\
 		GI              & $R_{554}/R_{677}$                                                                                    &~\cite{smith1995}              \\
@@
 \newblock {\em Hyperspectral Indices and Image Classifications for Agriculture
-  and Vegetation}; {CRC Press}:  Boca Raton, FL, USA, 2018, doi:10.1201/9781315159331.
+  and Vegetation}; {CRC Press}:  \hl{Boca Raton, FL, USA,} %newly added information, please confirm
+ 2018, doi:10.1201/9781315159331.
 
@@
 \newblock  In Proceedings of the 2014 {{Third International Workshop}} on {{Earth Observation}} and
-  {{Remote Sensing Applications}} ({{EORSA}}), Changsha, China, 11--14 June 2014; pp. 448--451, doi:10.1109/EORSA.2014.6927931.
+  {{Remote Sensing Applications}} ({{EORSA}}),  \hl{Changsha, China, 11--14 June} %MDPI: newly added, please confirm. The same to refs. 37, 45, 56, 63, 87, 89, 145
+ 2014; pp. 448--451, doi:10.1109/EORSA.2014.6927931.
 
@@
 \newblock {\em IEEE Trans. Pattern Anal. Mach. Intell.}
-  {\bf 1979}, {\em PAMI-1}, 306--307.
+  {\bf 1979}, {\em \hl{PAMI-1}}, %MDPI: please confirm if this volume is correct.
+ 306--307.
 
@@
 \newblock Methods to Estimate Forest Health.
-\newblock {\em Silva Fenn.} {\bf 1993}, {\em 27}, doi:10.14214/sf.a15668.
+\newblock {\em Silva Fenn.} {\bf 1993}, {\em \hl{27}}, %MDPI: please add page.
+ doi:10.14214/sf.a15668.
 
@@
   {{Feature Selection}}};
-\newblock  {{ICML}}, Citeseer, 2001.
+\newblock  \hl{{{ICML}}:} %MDPI: please add location of publisher. The same to refs. 39, 45, 56, 62, 76, 125, 142
+  2001.
 
@@
 \newblock  In Proceedings of the {{First International Workshop}} on {{Multiple
-  Classifier Systems}}, Cagliari, Italy, 21--23 June 2000; {Springer}:  Berlin/Heidelberg, Germany,
+  Classifier Systems}}, \hl{Cagliari, Italy, 21--23 June} 2000; {Springer}:  \hl{Berlin/Heidelberg, Germany,} %newly added information, please confirm
  2000; pp. 1--15.
@@
   in {{Neural Information Processing Systems}} 28}; Cortes, C., Lawrence, N.D.,
-  Lee, D.D., Sugiyama, M., Garnett, R., Eds.; {Curran Associates, Inc.}:  2015;
+  Lee, D.D., Sugiyama, M., Garnett, R., Eds.; {\hl{Curran Associates, Inc.}}:  2015;
   pp. 2962--2970.
@@
 \newblock {\em Stat. Appl. Genet. Mol. Biol.} {\bf
-  2011}, {\em 10}, 1-27, doi:10/c62xhp.
+  2011}, {\em \hl{10}}. %MDPI: please add page. The same to ref. 69
 
@@
 \newblock  In Proceedings of the Tenth National Conference on {{Artificial}}
-  Intelligence, San Jose, CA, USA, 12--16 July 1992; {AAAI Press}:  1992; pp. 129--134.
+  Intelligence, \hl{San Jose, CA, USA, 12--16 July} 1992; {\hl{AAAI Press}}:  1992; pp. 129--134.
 
@@
 \newblock  In Proceedings of the 2012 {{IEEE International Geoscience}} and {{Remote Sensing
-  Symposium}}, Munich, Germany, 22--27 July 2012; {R} Package Version 2.1.0; {IEEE}:  2012,
+  Symposium}}, \hl{Munich, Germany, 22--27 July} 2012; {R} Package Version 2.1.0; {\hl{IEEE}}:  2012,
 \newblock
@@
 \newblock {\em Interpretable Machine Learning---{{A Guide}} for {{Making Black
-  Box Models Explainable}}}; self-published: Munich, Germany, 2019.
+  Box Models Explainable}}};  \hl{2019.} %MDPI: please add publisher and location. The same to ref. 66, 67, 74, 100
 
@@
 \newblock {\em R: A {{Language}} and {{Environment}} for {{Statistical
-  Computing}}};  R Core Team: 2019, Vienna, Austria.
+  Computing}}};  \hl{R Core Team:} 2019.
 
@@
 \newblock  In Proceedings of the {{22nd ACM SIGKDD International Conference}} on
-  {{Knowledge Discovery}} and {{Data Mining}}, {{KDD}} '16, San Francisco, CA, USA, 13--17 August 2016; {ACM}: {New York, NY, USA},
+  {{Knowledge Discovery}} and {{Data Mining}}, {{KDD}} '16, \hl{San Francisco, CA, USA, 13--17 August} 2016; {ACM}: {New York, NY, USA},
   2016; pp. 785--794, doi:10.1145/2939672.2939785.
@@
 \newblock {\em Praznik: Collection of {{Information}}-{{Based Feature Selection
-  Filters}}};  2018.
+  Filters}}};  \hl{2018.}
 
@@
   '{{FSelector}}' {{Entropy}}-{{Based Feature Selection Algorithms}} with a
-  {{Sparse Matrix Support}}};  2019.
+  {{Sparse Matrix Support}}};  \hl{2019.}
 
@@
 Landau, W.M.
-\newblock The {drake} {{R}} Package: A Pipeline Toolkit for Reproducibility and
+\newblock The Drake {{R}} Package: A Pipeline Toolkit for Reproducibility and
   High-Performance Computing.
-\newblock {\em J. Open Source Softw.} {\bf 2018}, {\em 3}.
+\newblock {\em J. Open Source Softw.} {\bf 2018}, {\em \hl{3}}.
 
@@
 \newblock {\em Hsdar: Manage, Analyse and Simulate Hyperspectral Data in
-  {{R}}};  2016.
+  {{R}}};  \hl{2016.}
 
@@
 \newblock  In \emph{Remote {{Sensing}} and {{Modeling}} of {{Ecosystems}} for
-  {{Sustainability XIII}}}; {International Society for Optics and Photonics}:
-  2016; San Diego, California, United States, Volume 9975, p. 997503, doi:10.1117/12.2235391.
+  {{Sustainability XIII}}}; {\hl{International Society for Optics and Photonics}}:
+  2016; Volume 9975, p. 997503, doi:10.1117/12.2235391.
 
@@
 \newblock  In Proceedings of the 2016 {{Eighth International Conference}} on {{Advanced Computational
-  Intelligence}} ({{ICACI}}),  Chiang Mai, Thailand, 14--16 February 2016; pp. 36--40, doi:10.1109/ICACI.2016.7449799.
+  Intelligence}} ({{ICACI}}),  \hl{Chiang Mai, Thailand, 14--16 February} 2016; pp. 36--40, doi:10.1109/ICACI.2016.7449799.
 
@@
 \newblock  In Proceedings of the 2016 8th {{Workshop}} on {{Hyperspectral Image}} and {{Signal
-  Processing}}: Evolution in {{Remote Sensing}} ({{WHISPERS}}),  Los Angeles, CA, USA, 21--24 August 2016; pp. 1--5, doi:10.1109/WHISPERS.2016.8071759.
+  Processing}}: Evolution in {{Remote Sensing}} ({{WHISPERS}}),  \hl{Los Angeles, CA, USA, 21--24 August} 2016; pp.
+  1--5, doi:10.1109/WHISPERS.2016.8071759.
 
@@
 \newblock \emph{The {{Use}} of {{High Spectral Resolution Bands}} for {{Estimating
-  Absorbed Photosynthetically Active Radiation}} ({{A Par}})}; NASA Gov, Washington, { 1994}.
+  Absorbed Photosynthetically Active Radiation}} ({{A Par}})}; { \hl{1994}}.
 
@@
 Chen, J.M.
-\newblock Evaluation of Vegetation Indices and a {{Modified Simple Ratio}} for
-  Boreal Applications;
-\newblock {\emph{Canadian Journal of Remote Sensing}}:  1996, \emph{22}, 229-242, doi:10.4095/218303.
+\newblock \emph{Evaluation of Vegetation Indices and a {{Modified Simple Ratio}} for
+  Boreal Applications};
+\newblock Technical Report; {\hl{Natural Resources Canada/ESS/Scientific and
+  Technical Publishing Services}}:  1996, doi:10.4095/218303.
 
@@
 \newblock  In \emph{Spectral Signatures of Objects in Remote Sensing}; Guyenne, T.D.,
-  Hunt, J.J., Eds.;  {ESA Special Publication}: 1988; Aussois, France, Volume 287, p. 279.
+  Hunt, J.J., Eds.;  {\hl{ESA Special Publication}}: 1988; Volume 287, p. 279.
 
@@
   {{CHRIS}}/{{Proba}} Data.
-\newblock  In Proceedings of the 4th {{ESA CHRIS PROBA}} Workshop,  Frascati, Italy, 19--21 September 2006; Volume 2006, pp.
+\newblock  In Proceedings of the 4th {{ESA CHRIS PROBA}} Workshop,  \hl{Frascati, Italy, 19--21 September} 2006; Volume 2006, pp.
   19--21.

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/submission1/submission-1/manuscript.tex
add assets of all submissions

--- 2021-09-02 docs/00-manuscripts/ieee/ieee-manuscript.tex
@@
-% \documentclass[letterpaper, onecolumn, peerreview]{IEEEtran}
-\documentclass[final]{IEEEtran}
-
-\usepackage[switch]{lineno}
-
-% *** CITATION PACKAGES ***
-%
-\usepackage[noadjust]{cite}
-\usepackage[hidelinks]{hyperref}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-%\cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% 'compressed/ranged'. e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-%\cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf{}
+%  LaTeX support: latex@mdpi.com
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
+
+%=================================================================
+\documentclass[remotesensing,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%=================================================================
+% MDPI internal commands
+\firstpage{1}
+\makeatletter
+\setcounter{page}{\@firstpage}
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{2021}
+\copyrightyear{2020}
+%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
+\datereceived{}
+\dateaccepted{}
+\datepublished{}
+\hreflink{https://doi.org/} % If needed use \linebreak
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
+
+% \usepackage[noadjust]{cite}
+% \usepackage[hidelinks]{hyperref}
+\usepackage[nolist]{acronym}
+
+\usepackage{tabularx}
+\usepackage{booktabs}
+\usepackage{multirow}
+\usepackage{adjustbox}
+
+% \usepackage{supertabular}
+\usepackage{supertabular,booktabs}
+
+% line breaks in table cells
+\newcommand{\specialcell}[2][l]{%
+  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
+
+\usepackage{csquotes}
+\usepackage[utf8]{inputenc}
+% enables custom font styles which are not available in IEEEtran by default
+% https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran
+
+\renewcommand{\sfdefault}{cmss}
+\renewcommand{\rmdefault}{cmr}
+\renewcommand{\ttdefault}{cmt}
+
+\usepackage[detect-all]{siunitx}
+
 \usepackage[pdftex]{graphicx}
@@
 \DeclareGraphicsExtensions{.pdf,.jpg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
+
+% suppress fancyhdr warnings
+\setlength{\headheight}{23.90979pt}
 
@@
 \newcommand*\mean[1]{\bar{#1}}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-
-% *** SPECIALIZED LIST PACKAGES ***
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-\ifCLASSOPTIONcompsoc
- \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\else
- \usepackage[caption=false,font=footnotesize]{subfig}
-\fi
-
-% *** FLOAT PACKAGES ***
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\usepackage[nolist]{acronym}
-
-\usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multirow}
-\usepackage{adjustbox}
-
-% line breaks in table cells
-\newcommand{\specialcell}[2][l]{%
-  \begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
-
-\usepackage{xtab}
-
-\usepackage{csquotes}
-
-\usepackage[utf8]{inputenc}
-% enables custom font styles which are not available in IEEEtran by default
-% https://tex.stackexchange.com/questions/331228/there-is-no-bold-text-italics-in-ieeetran
-
-\renewcommand{\sfdefault}{cmss}
-\renewcommand{\rmdefault}{cmr}
-\renewcommand{\ttdefault}{cmt}
-
-\usepackage[detect-all]{siunitx}
-
-\begin{document}
-% \linenumbers
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
-\title{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
-\IEEEpeerreviewmaketitle{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and non-breaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-
-% suppress 'underfull hbox' warnings
-\hbadness=99999
-
-\author{Patrick~Schratz,
-	Jannes~Muenchow,
-	Eugenia~Iturritxa,
-	José~Cortés,
-	Bernd~Bischl,
-	and Alexander~Brenning
-	\thanks{P.Schratz, J.Muenchow, J.Cortés and A.Brenning are with the Department
-		of Geography, geographic information science group, Friedrich Schiller University Jena, Germany.}% <-this % stops a space
-	\thanks{B.Bischl is head of the computational statistics group at the Department of Statistics, Ludwig-Maximilians-Universität in Munich, Germany.}% <-this % stops a space
-	\thanks{E.Iturritxa is with NEIKER Tecnalia, Vitoria-Gasteiz, Spain.}% <-this % stops a space
-	%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}
+
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?}
+
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Title}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0003-0748-6624} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-7834-4717} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0003-2567-8689} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0002-6390-5873} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorE}{0000-0001-6002-6980} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorF}{0000-0001-6640-679X} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Patrick Schratz $^{1}$\orcidA{}*, Jannes Muenchow $^{1}$\orcidB{}, José Cortés $^{1}$\orcidC{}, Eugenia Iturritxa $^{2}$\orcidD{}, Bernd Bischl $^{3}$\orcidE{} and Alexander Brenning $^{1,}$\orcidF{}}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{Schratz, P.; Muenchow, J.; Cortés, J.; Iturritxa, E.; Bischl, B.; Brenning, A.}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Friedrich Schiller University Jena, Department of Geography, GIScience group, Germany\\
+$^{2}$ \quad NEIKER Tecnalia, Spain\\
+$^{3}$ \quad Ludwig-Maximilians-University Munich, Department of Statistics, Computational Statistics group, Germany
 }
 
-% note the % following the last \IEEEmembership and also \thanks -
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-%
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those 'LaTeX things'. For
-% instance, '\textbf{A} \textbf{B}' will typeset as 'A B' not 'AB'. To get
-% 'AB' then you have to do: '\textbf{A}\textbf{B}'
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-\markboth{IEEE JOURNAL OF SELECTED TOPICS IN APPLIED EARTH OBSERVATIONS AND REMOTE SENSING}%
-{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-%
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% make the title area
-\maketitle
-
-% As a general rule, do not put math, special symbols or citations
-% in the abstract or keywords.
-\begin{abstract}
-	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
+% Contact information of the corresponding author
+\corres{Correspondence: patrick.schratz@uni-jena.de}
+
+% Current address and/or shared authorship
+% \firstnote{Current address: Affiliation 3}
+% \secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+% Abstract (Do not insert blank lines, i.e. \\)
+\abstract{This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
 The effect of filter-based feature-selection methods on predictive performance was compared.
@@
 
-\section{Methods}
-
 \subsection{Derivation of indices}
@@
 
-\section{Acknowledgments}
-This work was funded by the EU LIFE Healthy Forest project (LIFE14 ENV/ES/000179) and the German Scholars Organization/Carl Zeiss Foundation.
-
-\appendices{}
-
-% \section{SVM ALE plots for task HR}
-
-% \begin{figure} [ht]
-% 	\begin{center}
-% 		\includegraphics[width=0.48\textwidth] {fi-hr-ale-1.pdf}
-% 		\caption{ALE plots of SVM on dataset HR.
-The ten most important features from the permutation-based variable importance estimation were used.
-The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-hr-ale}
-% 	\end{center}
-% \end{figure}
-
-%\section{SVM ALE plots for task VI}
-
-% \begin{figure} [ht]
-% 	\begin{center}
-% 		\includegraphics[width=0.48\textwidth] {fi-vi-ale-1.pdf}
-% 		\caption{ALE plots on dataset VI of SVM learner.
-Subset showing the ten most important features according to the permutation-based variable importance.
-The y-axis shows the deviation to the mean prediction for each feature, with the mean prediction being centered at zero.}\label{fig:fi-vi-ale}
-% 	\end{center}
-% \end{figure}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{Conceptualization, Patrick Schratz and Alexander Brenning; Data curation, Patrick Schratz and Eugenia Iturritxa; Formal analysis, Patrick Schratz, José Cortés, Bernd Bischl and Alexander Brenning; Funding acquisition, Alexander Brenning; Investigation, Patrick Schratz; Methodology, Patrick Schratz, Jannes Münchow, José Cortés and Bernd Bischl; Project administration, Alexander Brenning; Resources, Eugenia Iturritxa; Software, Patrick Schratz; Supervision, Bernd Bischl and Alexander Brenning; Validation, Patrick Schratz and Alexander Brenning; Visualization, Patrick Schratz; Writing – original draft, Patrick Schratz; Writing – review \& editing, Jannes Münchow, José Cortés and Alexander Brenning.}
+
+\funding{This work was funded by the EU LIFE Healthy Forest project (LIFE14 ENV/ES/000179) and the German Scholars Organization/Carl Zeiss Foundation.}
+
+\dataavailability{The data presented in this study are openly available in Zenodo at https://doi.org/10.5281/zenodo.2635403.}
+
+% \acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\conflictsofinterest{The authors declare no conflict of interest.
+The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the~results.}
+
+%% Optional
+% \sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\abbreviations{Abbreviations}{
+	The following abbreviations are used in this manuscript:\\
+
+	\noindent
+	\begin{tabular}{@{}ll}
+		AGB     & above-ground biomass                                    \\
+		ALE     & accumulated local effects                               \\
+		ALS     & airborne laser scanning                                 \\
+		ANN     & artificial neural network                               \\
+		AUROC   & area under the receiver operating characteristics curve \\
+		BRT     & boosted regression trees                                \\
+		CART    & classification and regression trees                     \\
+		CNN     & convolutional neural networks                           \\
+		CV      & cross-validation                                        \\
+		DAP     & digital aerial photogrammetry                           \\
+		ENM     & environmental niche modeling                            \\
+		FFS     & forward feature selection                               \\
+		FPR     & false positive rate                                     \\
+		FS      & feature selection                                       \\
+		GAM     & generalized additive model                              \\
+		GBM     & gradient boosting machine                               \\
+		GLM     & generalized linear model                                \\
+		ICGC    & Institut Cartografic i Geologic de Catalunya            \\
+		IQR     & interquartile range                                     \\
+		LiDAR   & light detection and ranging                             \\
+		LOWESS  & locally weighted scatter plot smoothing                 \\
+		MARS    & multivariate adaptive regression splines                \\
+		MBO     & model-based optimization                                \\
+		MEM     & maximum entropy model                                   \\
+		ML      & machine learning                                        \\
+		NDII    & normalized difference infrared index                    \\
+		NDMI    & normalized difference moisture index                    \\
+		NIR     & near-infrared                                           \\
+		NRI     & normalized ratio index                                  \\
+		OLS     & ordinary least squares                                  \\
+		OMNBR   & optimized multiple narrow-band reflectance              \\
+		PCA     & principal component analysis                            \\
+		PDP     & partial dependence plots                                \\
+		PISR    & potential incoming solar radiation                      \\
+		PLS     & partial least-squares                                   \\
+		POV     & proportion of variance explained                        \\
+		RBF     & radial basis function                                   \\
+		RF      & random forest                                           \\
+		RMSE    & root mean square error                                  \\
+		RR      & ridge regression                                        \\
+		RSS     & residual sum of squares                                 \\
+		SAR     & synthetic aperture radar                                \\
+		SDM     & species distribution modeling                           \\
+		SMBO    & sequential-based model optimization                     \\
+		SVM     & support vector machine                                  \\
+		TPR     & true positive rate                                      \\
+		VI      & vegetation index                                        \\
+		XGBoost & extreme gradient boosting
+	\end{tabular}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
 \pagebreak
-\section{}
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
+\appendix
+\section{Appendices}
+\subsection{}
 
@@
 
-\section{}
-
-\begin{figure} [ht]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth] {correlation-nbins-1.pdf}
-		\caption{Spearman correlations of rankings obtained with the information gain filter using different \texttt{\(n_{bins}\)} values for discretization of the numeric response.
-Filter names refer to the nomenclature used by the \texttt{mlr} R package.
-Underscores in names divide the terminology into their upstream R package and the actual filter name.}\label{fig:correlation-nbins}
-	\end{center}
-\end{figure}
-
 \pagebreak
 
-\section{}
-
-% parameter limits
+\subsection{}
+
 \begin{table}[h!]
@@
 
-\section{}
-
-\begingroup\scriptsize
-\setlength\tabcolsep{4pt}  % default value: 6pt
-\begin{xtabular*}{0.3\textwidth}{lll}
+\pagebreak
+
+\subsection{}
+
+% \begingroup\scriptsize
+% \setlength\tabcolsep{4pt}  % default value: 6pt
+% vertical spacing between rows (default is 1)
+{\def\arraystretch{1.7}
+	\tablecaption{List of available vegetation indices in the \texttt{hsdar} package.}
+	\begin{supertabular}{lll}
+		% \begin{supertabular}{0.3\textwidth}{lll}
 		\bfseries{Name} & \bfseries{Formula}                                                                                   & \bfseries{Reference*}         \\
@@
 		Vogelmann4      & $\frac{R_{734}-R_{747}}{R_{715}+R_{720}}$                                                            & \cite{vogelmann1993}     \\
-\end{xtabular*}
+	\end{supertabular}
 	\label{tab:vegindices}
-\endgroup
+}
 
@@
 The image acquisitions were originally attempted during one day (29.10.2016).
-due to the variable meteorological conditions some stands had to be imaged one day later.
-
-\pagebreak
-
-\bibliographystyle{IEEEtran}
+Due to the variable meteorological conditions some stands had to be imaged one day later.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{paracol}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% To add notes in main text, please use \endnote{} and un-comment the codes below.
+%\begin{adjustwidth}{-5.0cm}{0cm}
+%\printendnotes[custom]
+%\end{adjustwidth}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\externalbibliography{yes}
 \bibliography{Biblio}
 
-\pagebreak
-
-% The photo area is 1 in wide and 1.25 in long. The IEEE recommends that author photo images should be of 220 dpi (dots per inch) resolution and in gray scale with 8 bits/sample.
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/pschratz.jpeg}}]{Patrick Schratz}
-	%	[]{Patrick Schratz}
-	is a Data Scientist working in Zurich, Switzerland.
+% If authors have biography, please use the format below
+\section*{Short Biography of Authors}
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{pschratz.jpeg}}}
+{\textbf{Patrick Schratz} is a Data Scientist working in Zurich, Switzerland.
 	His area of expertise is applied machine learning, more specifically the field of environmental modeling.
 	He is a PhD candidate in geographic information science in the Department of Geography at Friedrich Schiller University Jena where he conducts research in environmental modeling.
-	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/jmuenchow.jpg}}]{Jannes Muenchow}
-	is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
+	At cynkra GmbH in Zurich, Patrick is an R consultant with many years of experience in the following fields: CI/CD, package development, DevOps tasks, machine learning and spatio-temporal data handling.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{jmuenchow.jpg}}}
+{\textbf{Jannes Muenchow} is a GIScientist working in tropical ecology since 2007 with a special interest in ENSO, biodiversity, species distribution modeling and predictive mapping.
 	He received his PhD from Friedrich-Alexander University Erlangen-Nürnberg (Germany) in 2013.
@@
 	Currently, his research focuses on developing open source tools for ecology, geomorphology and qualitative GIS.
-	He is a co-author of the book "Geocomputation with R".
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/eiturritxa.jpg}}]{Eugenia Iturritxa}
-	received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.
+	He is a co-author of the book "Geocomputation with R".}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{eiturritxa.jpg}}}
+{\textbf{Eugenia Iturritxa} received the Ecology and Ph.D degree in Plant Protection from the University of the Basque Country in 2001.
 	Since 1999 her main research has focused on forest health, on the study of diverse species of native and introduced pathogenic fungi in forests and forest plantations.
-	Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/jcortes.png}}]{José Cortés}
-	received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
+	Her research the NEIKER research center includes the distribution of diseases, analysis of predisposing factors for them, genetic and phenotypic studies of populations and their epidemiology.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{jcortes.png}}}
+{\textbf{José Cortés} received his B.S. in Mathematics and M.S. in Statistics from Arizona State University, Arizona, USA, in 2014 and 2016, respectively.
 	Currently he is a PhD student at Friedrich Schiller University, Jena, Germany.
 	He is a member of the International Max Planck Research School on Global Biogeochemical Cycles (IMPRS-gBGC), a joint program with the Max Planck Institute for Biogeochemistry.
-	His research focuses on spatio-temporal trend detection in environmental data.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/bbischl.jpg}}]{Bernd Bischl}
-	obtained his Ph.D from Dortmund Technical University in 2013.
+	His research focuses on spatio-temporal trend detection in environmental data.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{bbischl.jpg}}}
+{\textbf{Bernd Bischl} obtained his Ph.D from Dortmund Technical University in 2013.
 	He is a professor of Statistical Learning and Data Science in the Department of Statistics at Ludwig-Maximilians-Universität in Munich, Germany, and a director of the Munich Center for Machine Learning.
-	His research interests include AutoML, model selection, interpretable ML and XAI.
-\end{IEEEbiography}
-\begin{IEEEbiography}
-	[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{./jpg/abrenning.jpg}}]{Alexander Brenning}
-	(Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
+	His research interests include AutoML, model selection, interpretable ML and XAI.}
+
+\bio
+{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{abrenning.jpg}}}
+{\textbf{Alexander Brenning} (Ph.D. 2005) graduated in mathematics at Technical University of Freiberg, Germany, and received his Ph.D. in geography from Humboldt-Universität zu Berlin.
 	He served as an assistant professor and tenured associate professor in geomatics at the University of Waterloo, Ontario, Canada from 2007 until 2015, when he was appointed as a full professor in geographic information science at Friedrich Schiller University Jena, Germany.
 	Dr. Brenning's research interests include statistical and machine-learning techniques for environmental modeling and remote sensing.
-	He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.
-\end{IEEEbiography}
-
+	He has also contributed to open-source geocomputation by developing R packages for spatial cross-validation and GIS coupling.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/submission1/submission-1/performance-best-per-learner.tex
add assets of all submissions

--- 2021-08-18 docs/00-manuscripts/ieee/performance-best-per-learner.tex
@@
   3 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
-  4 & HR & regr.featureless & No Filter & 31.16 & 15.03 \\ 
+  4 & - & regr.featureless & No Filter & 31.16 & 15.03 \\ 
   5 & HR & RF & Relief & 31.25 & 14.96 \\ 

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/submission2/submission-2.tex
add assets of all submissions

--- 2021-11-15 docs/00-manuscripts/mdpi/manuscript.tex
@@
 There was also no clear pattern in the percentage of features selected based on hyperparameter tuning (\autoref{tab:tune-perc-sel-features}).
-The specific performance value to benchmark against during tuning is defined by the internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1 (interpreted as a\ percentage of available features) for the hyperparameter that controls the number of features.
-% Nevertheless, considering SVM's relative success, the use of at least a few hundred features from the combined feature set appears to be beneficial, or at least not harmful when the model's built-in regularization is capable of dealing with the resulting high-dimensional situation.
+The optimal value for the number of features (interpreted as a\ percentage of available features), which are selected from the ranked filter results, is determined by the internal surrogate learner of the MBO tuning method using the results from the previous tuning iterations.
+Due to this iterative approach MBO is in some ways able to evaluate how well a learning algorithm plays together with a certain amount of selected features and is subsequently able to adjust the number of variables to an optimal value.
+All in all, considering SVM's relative success, the use of at least a few hundred features from the combined feature set appears to be beneficial, or at least not harmful when the model's built-in regularization is capable of dealing with the resulting high-dimensional situation.
 

### 2021-12-17 docs/00-manuscripts/mdpi/submissions/submission1/submission-1/jpg/create-img-grid.R

@@ -1,0 +1,31 @@
+# using imagemagick system library
+
+# paper ------------------------------------------------------------------------
+
+system2("montage", args = c("./code/98-paper/ieee/jpg/10-defol.jpg",
+                            "./code/98-paper/ieee/jpg/20-defol.jpg",
+                            "./code/98-paper/ieee/jpg/40-defol.jpg",
+                            "./code/98-paper/ieee/jpg/60-70-defol.jpg",
+                            "-geometry +20+20",
+                            "-tile 2x2",
+                            "./code/98-paper/ieee/jpg/defol-grid-2000px.jpg"))
+
+# resize
+system2("convert", args = c("./code/98-paper/ieee/jpg/defol-grid-2000px.jpg",
+                           "-resize 2000x2000",
+                           "./code/98-paper/ieee/jpg/defol-grid-2000px.jpg"))
+
+# presentation -----------------------------------------------------------------
+
+system2("montage", args = c("./code/98-paper/ieee/jpg/10-defol.jpg",
+                            "./code/98-paper/ieee/jpg/20-defol.jpg",
+                            "./code/98-paper/ieee/jpg/40-defol.jpg",
+                            "./code/98-paper/ieee/jpg/60-70-defol.jpg",
+                            "-geometry +20+20",
+                            "-tile 4x1",
+                            "./code/98-paper/ieee/jpg/defol-grid-pres-1500px.jpg"))
+
+# resize
+system2("convert", args = c("./code/98-paper/ieee/jpg/defol-grid-pres-1500px.jpg",
+                            "-resize 2000x1500",
+                            "./code/98-paper/ieee/jpg/defol-grid-pres-1500px.jpg"))
